### 2021-08-10 Notes/notes.tex
created notes skeleton

\documentclass[onehalfspacing]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{ wasysym }

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}

\author{Alexander Pluska}
\title{Proofs as Programs in Classical Logic\\Notes}

\begin{document}

\maketitle

\bibliographystyle{acm}
\bibliography{references}

\end{document}

### 2021-08-22 Notes/notes.tex
First draft. Calculus defined. Simple example.

@@
 \usepackage{ wasysym }
+\usepackage{ stmaryrd }
+\usepackage{ mathpartir }
+\usepackage{xcolor}
 
@@
 \newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+
+\newcommand{\llb}{\llbracket}
+\newcommand{\rrb}{\rrbracket}
+\newcommand{\blue}[1]{{\color{blue}#1}}
+\newcommand{\var}{\text{var}}
 
@@
 
+\section{Proof translation for saturation-based theorem proving with induction}
+
+Most modern theorem provers for first order logic utilize resolution and saturation based techniques.
+The standard intuitionistic proofs-as-programs correspondence fails as even the very method itself crucially depends on double negation elimination.
+As a first step we work out a way to convert proofs of $\forall\exists$-proofs based on the calculus presented in~\cite{Echenim_2019} into programs.
+An advantage of the calculus as presented in~\cite{Echenim_2019} is that it can (at least in theory) handle general induction for inductive datatype.
+Many advances have recently been made for incorporating induction in existing theorem provers~\cite{hajdu2021automating}\cite{10.1007/978-3-030-53518-6_8} however all of the calculi fail to support induction in the general case, in particular sentences involving existential quantifiers and requiring induction depth greater than $1$ are problematic.
+
+The approach in~\cite{Echenim_2019} seems promising as, despite being able to handle general induction, there is no explosion in the number of clauses due to a clever use of so-called \textit{constrained clauses}, special symbols $\textbf{T}_\alpha^\prec$ for induction and a special activation mechanism for potential inductive invariants, although how it performs in practice is yet to be seen.
+
+Program translation for classical proofs has been outlined in~\cite{Griffin_1990}.
+Based on this~\cite{de_Groote_1995} and~\cite{Barbanera_1996} have been developed.
+Our a approach is a combination of both.
+At the moment it is quite crude and is mostly based on the exception method from~\cite{de_Groote_1995}.
+We aim to later translate this to the CPS approach presented in the same paper to better see how it relates to the normal proofs-as-program correspondence known from intuitionistic logic.
+
+\subsection{The Calculus}
+
+As explained above our calculus is based on~\cite{Echenim_2019}. We take the core rules, which are not mentioned in~\cite{Echenim_2019}, from~\cite{Kov_cs_2013}.
+
+The first five inferences are standard in resolution based theorem proving. We assume some simplification ordering $\succeq^*$.\\
+\textbf{Resolution.}
+$$\inferrule{{A}\vee C_1\\{\neg B}\vee C_2}{(C_1\vee C_2)\theta}$$
+where $\theta$ is a mgu of $A$ and $\neg B$.\\
+\textbf{Factoring.}
+$$\inferrule{{A}\vee{\neg B}\vee C}{(A\vee C)\theta}$$
+where theta is a mgu of $A$ and $B$.\\
+\textbf{Superposition.}
+$$\inferrule{{l=r}\vee C_1\\{L[s]}\vee C_2}{(L[r]\vee C_1\vee C_2)\theta}\:
+\inferrule{{l=r}\vee C_1\\ {t[s]=t'}\vee C_2}{(t[r]=t'\vee C_1\vee C_2)\theta}\:
+\inferrule{{l=r}\vee C_1\\{t[s]\neq t'}\vee C_2}{(t[r]\neq t'\vee C_1\vee C_2)\theta}$$
+where $\theta$ is a mgu of $l$ and $s$, $s$ is not a variable, $r\theta\not\succeq^* l\theta$, $L[s]$ is not an equality literal, and $t'\theta\succeq^* t[s]\theta$.\\
+\textbf{Equality Resolution.}
+$$\inferrule{{s\neq t}\vee C}{C\theta}$$
+where $\theta$ is a mgu of $s$ and $t$.\\
+\textbf{Equality Factoring.}
+$$\inferrule{{s=t}\vee{s'=t'}\vee C}{(s=t\vee t=t'\vee C)\theta}$$
+where $\theta$ is an mgu of $s$ and $s'$, $t\theta\not\succeq^* s\theta$, and $t'\theta\succeq^* t\theta$.
+
+In addition to regular clauses and inference rules we introduce c-clauses and special rules for them, i.e. (all of this will be made more precise later)
+\begin{definition}
+	A \textit{constrained clause} (c-clause) is an expression of  the form $\llb C\:|\:\mathcal X \rrb$, where
+	\begin{itemize}
+		\item C is a clause,
+		\item $\mathcal{X}$ is a conjunction of the form $\bigwedge_{i=1}^ n f_i(\mathbf{u_i})\simeq v_i\wedge\bigwedge_{i=1}^m \mathbf{T}_{\alpha_i}^\prec(z_i,\textbf{w}_i)$, where $n, m\geq 0$ and $\forall i\in\{1\dots n\}, f_i$ are uninterpreted function symbols (possibly nullary).
+	\end{itemize}
+\end{definition}
+The semantics of $\llb C\:|\:\mathcal X \rrb$ is more or less ``if all constraints in $\mathcal X$ hold, then $C$ holds" and of $\mathbf{T}_{\alpha_i}^\prec(z,\mathbf{w})$ ``for all $z'\prec z$ $\alpha(z', \mathbf{w})$ holds" where $\prec$ is the standard wfo of an inductive type.
+
+All of the above inference rules extend naturally to constrained clauses, i.e.
+if $\tiny{\inferrule{H_1,\dots,H_n}{C\theta}}$ is an inference rule for regular clauses, then $$\inferrule{\llb H_1\:|\:\mathcal X_1\rrb,\dots,\llb H_1\:|\:\mathcal X_1\rrb}{\llb C\:|\:\mathcal X_1\wedge\dots\wedge\mathcal X_n\rrb\theta}$$ is the corresponding inference rule for c-clauses.
+
+Furthermore we have some special rules for c-clauses:\\
+\textbf{Constraint Factorization.}
+$$\inferrule{\llb C\:|\:l_1\wedge l_2\wedge \mathcal{X}\rrb}{\llb C\:|\:l_1\wedge \mathcal{X}\rrb\theta}$$where $\theta$ is an mgu of $l_1$ and $l_2$.\\
+\textbf{Abstraction.}
+$$\inferrule{\llb C[f(\mathbf{t})]_p\:|\:\mathcal{X}\rrb}{\llb C[x]_p\:|\:f(\mathbf{t})\simeq x\wedge\mathcal{X}\rrb}$$ where $f$ is a uninterpreted function symbol, $x$ is a fresh variable.\\
+\textbf{Instantiation.}
+$$\inferrule{C\:|\:\mathcal{X}\wedge t\simeq s\rrb}{\llb C\:|\:\mathcal{X}\rrb\theta}$$
+where $\theta$ is a mgu of $t$ and $s$.
+
+In addition there are some rules that generate candidates for an inductive invariant in~\cite{Echenim_2019}.
+We will leave those out for now an choose our variant ``magically" by hand.
+For every formula $\alpha[z,\textbf{w}]$ we than have axiom $$\Gamma_\mathbf{T_\alpha^\prec}:= \llb z' \not\prec z\vee \alpha(z', \mathbf{w})\:|\:\mathbf{T}_\alpha^\prec(z,\mathbf{w})\rrb$$ and $\prec $ is axiomatized by $$\Gamma_\prec^s:=\llb x_i\prec s(x_1,\dots, x_n)\:|\:\top\rrb\indent \Gamma_\prec:=\llb x\not\prec y\vee y\not\prec z\vee x\prec z\:|\:\top\rrb$$for every n-ary inductive constructor $s$, $i\in\{1\dots n\}$.
+
+Finally we have additional rules for inductive types.
+Since we are (for now) not interested in how to generate inductive invariants it is possible to present just one combination rule for Domain Decomposition and Induction.\\
+\textbf{Induction.}
+$$\inferrule{\llb C_1\:|\:\beta_1(t_1)\wedge \mathcal{X}_1\rrb\dots\llb C_n\:|\:\beta_n(t_n)\wedge \mathcal{X}_n\rrb}{\llb C_1\vee\dots\vee C_n\vee \alpha(x,\mathbf{w}_1)\:|\:\mathcal{X}_1\setminus\{\mathbf{T}_{\alpha}^\prec(t_1,\mathbf w_1)\}\wedge\dots\wedge\mathcal{X}_n\setminus\{\mathbf{T}_{\alpha}^\prec(t_n,\mathbf w_n)\}\wedge\beta_1(x)\rrb\theta}$$where $\theta$ is the most general idempotent substitution such that $\beta_j\theta\subseteq\beta_1\theta$ for $n\in\{2\dots n\}$ and a mgu for $\mathbf{w}_1,\dots,\mathbf{w}_n$, the variables occurring in $t_1,\dots,t_n$ do not occur in $C_i,\beta_i,\mathbf{w}_i$ or $\mathcal{X}_i$.
+$\{t_1\dots t_n\}$ is \textit{covering}, i.e.
+the whole type can be constructed using $\{t_1\dots t_n\}$, for instance for $nat$ both $\{z, s(x)\}$ and $\{z, s(z), s(s(x))\}$ are covering and $x$ is some fresh variable.
+
+Note that setting $\alpha=\bot$ and $\mathbf{w_i}=\emptyset$ we obtain the regular Domain Decomposition Rule from~\cite{Echenim_2019}.
+
+\subsection{Examples and Translations}
+
+For now we will consider two minimal examples.
+The first on doesn't use induction.
+There are two constant symbols $a, b$ and a binary proposition variable $p(-, -)$.
+The only axiom is $\forall x\:p(x,a)\vee p(x, b)$.
+We seek to prove $\forall x\exists y p(x, y)$.
+It is immediately clear that unless the axiom has computational content, our extracted program cannot yield a value.
+On the other hand we will see that we can still create a well-typed program.
+And if the axiom contains computational content it will indeed yield a result.
+
+\begin{example}\hfill
+	\begin{align}
+		& p(x_1,a)\vee p(x_1, b) & \text{Axiom}\\
+		& \llb \neg p(x_2, y_2) | x\simeq x_2\rrb & \text{negated Hypothesis}\\
+		& \llb p(x_1, b) | x\simeq x_1\rrb & res.\:(1) (2) [x_1/x_2, a/y_2]\\
+		& \llb \square | x\simeq x_1\rrb & res.\:(3) (2) [x_1/x_2, b/y_2]\\
+		& \square &\text{Instantiation }(4)
+	\end{align}
+\end{example}
+
+Now for our computational translation we introduce a special exception terms $e_\alpha(t_1,\dots,t_n)$ and $\overline e_\alpha(t_1,\dots,t_n)$ for every atom $\alpha[t_1,\dots,t_n]$ with $n$ term-variables, which intuitively indicate that $\alpha[t_1,\dots,t_n]$ is not false.
+They are usually introduced in the manner of the last derivation of definition 3.4 in~\cite{de_Groote_1995}.
+If one were to assign types we would have $e_\alpha(t_1.\dots,t_n) : (\alpha(t_1,\dots,t_n)\to\bot)\to\bot$ and $\overline e_\alpha(t_1.\dots,t_n) : ((\alpha(t_1,\dots,t_n)\to\bot)\to\bot)\to\bot$.
+However as we have the (classical) identification, we can simplify $\overline e_\alpha(t_1.\dots,t_n) : \alpha(t_1,\dots,t_n)\to\bot$.
+Having this in mind both $e_\alpha(t_1.\dots,t_n) \overline e_\alpha(t_1.\dots,t_n) : \bot$ and $\overline e_\alpha(t_1.\dots,t_n) e_\alpha(t_1.\dots,t_n) : \bot$ make ``type-sense" in the exception sense of~\cite{de_Groote_1995}.
+We will use this in our program.
+Let us now translate the above proof.
+Note that the terms for axioms are given.
+For clarity we will use braces for function application.
+For convenience we will write $\alpha_i$ for the $i$-the generated term.
+\setcounter{equation}{0}
+\begin{example}\hfill
+	\begin{align}
+		& \alpha_1 & (p(x_1,a)\to\bot)\to(p(x_1, b)\to\bot)\to\bot\\
+		& \overline e_p(x_2, y_2) & p(x_2, y_2)\to\bot\\
+		& \lambda t. \mathbf{let}\:v : \neg p(x_1, a)&(p(x_1, b)\to\bot)\to\bot\\
+		& \phantom{\lambda t.}\mathbf{in}\:\alpha_1(v, t) &\\
+		& \phantom{\lambda t.}\mathbf{handle}\:v(w)\Rightarrow \overline e_p(x_1, a)(w)\indent [= \alpha_2[x_1/x_2, a/y_2](w)]\hspace*{-4cm}\nonumber&\\
+		& \mathbf{let}\:v : \neg p(x_1, b)\:&\bot\\
+		& \mathbf{in}\:\alpha_3(v) &\nonumber\\
+		& \mathbf{handle}\:v(w)\Rightarrow \overline e_p(x_1, b)(w)\indent[= \alpha_2[x_1/x_2, b/y_2](w)]\hspace*{-4cm}\nonumber&\\
+		& \lambda x_1.\alpha_4 &\forall x\bot
+	\end{align}
+\end{example}
+
+Now to retrieve a value we just need to handle the final exception, i.e.
+$$\lambda x_1.\alpha_4\:\mathbf{handle}\: \overline{e}_p(x_2, y_2)(w)\Rightarrow y_2$$
+Of course this will not yield an actual value unless $\alpha$ yields.\\
+We will now consider a minimal inductive example.\pagebreak
+
+Consider the inductive type $nat$ with constructors $z : nat$ and $s: nat \to nat$ and some predicate $p(x, y)\subseteq nat^2$.
+We are going to prove $\forall x\exists y\: p(x,y)$ from $p(z, z)$ and $\forall n,m:nat(p(n, m)\Rightarrow p(s(n), s(s(m))))$.
+
+\begin{example}\hfill
+	\begin{align}
+		& p(z, z) & \text{Axiom}\\
+		& \neg p(n_2, m_2)\vee p(s(n_2), s(s(m_2)))& \text{Axiom}\\
+		& \llb \neg p(n_3, m_3) | x\simeq n_3\rrb & \text{negated Hypothesis}\\
+		& \llb \square | x\simeq z\rrb & res.\:(1) (3) [z/n_3, z/m_3]\\
+		& \llb \neg p(n_2, m_2) | x\simeq s(n_2)\rrb &res. (2) (3) [s(n_2)/n_3], s(s(m_2))/m_3]\\
+		& \text{Activate $\alpha := p(a, b)[a, b]$}&\text{Trigger  (4)}\\
+		& \llb n_7\not\prec n_7'\vee p(n_7, m_7) | \mathbf{T}_\alpha^\prec(n_7', m_7)\rrb& \Gamma_{\mathbf{T}_\alpha^\prec}\\
+		& \llb n_7\not\prec n_7'| x\simeq s(n_7)\wedge \mathbf{T}_\alpha^\prec(n_7', m_7)\rrb & res. (5) (7) [n_7/n_2, m_7/ m_2] \\
+		& \llb n_7\prec s(n_7)|\top\rrb &\Gamma_\prec^s\\
+		& \llb \square | x\simeq s(n_7)\wedge\mathbf{T}_\alpha^\prec(s(n_7), m_7)\rrb&res. (8)(9)[s(n_7)/n_7']\\
+		& \llb \square | x\simeq n_{10}\rrb &\text{Induction }(4)(9)\\
+		& \square &\text{Instantiation }(10)
+	\end{align}
+\end{example}
+
 \bibliographystyle{acm}

### 2021-09-06 Notes/notes.tex
Introduced Agda

@@
 
+\subsection{Type system}
+
+Since we want to apply our method to first-order resolution proofs we will require a dependant type system. For this purpose we shall use the practical type system of agda as described in~\cite{agda}.
+
 \subsection{Examples and Translations}
@@
 For now we will consider two minimal examples.
-The first on doesn't use induction.
+The first doesn't use induction.
 There are two constant symbols $a, b$ and a binary proposition variable $p(-, -)$.
@@
 		& \lambda t. \mathbf{let}\:v : \neg p(x_1, a)&(p(x_1, b)\to\bot)\to\bot\\
-		& \phantom{\lambda t.}\mathbf{in}\:\alpha_1(v, t) &\\
+		& \phantom{\lambda t.}\mathbf{in}\:\alpha_1(v, t) &\nonumber\\
 		& \phantom{\lambda t.}\mathbf{handle}\:v(w)\Rightarrow \overline e_p(x_1, a)(w)\indent [= \alpha_2[x_1/x_2, a/y_2](w)]\hspace*{-4cm}\nonumber&\\
@@
 		& \mathbf{handle}\:v(w)\Rightarrow \overline e_p(x_1, b)(w)\indent[= \alpha_2[x_1/x_2, b/y_2](w)]\hspace*{-4cm}\nonumber&\\
-		& \lambda x_1.\alpha_4 &\forall x\bot
+		& \lambda x_1.\alpha_4 &(x : X) \to \bot
 	\end{align}
@@
 $$\lambda x_1.\alpha_4\:\mathbf{handle}\: \overline{e}_p(x_2, y_2)(w)\Rightarrow y_2$$
-Of course this will not yield an actual value unless $\alpha$ yields.\\
+Of course this will not yield an actual value unless $\alpha_1$ yields.\\
 We will now consider a minimal inductive example.\pagebreak

### 2021-09-28 Notes/notes.tex
Added intuitionistic example for simple Curry Howard

@@
 \usepackage{xcolor}
+\usepackage{bussproofs}
 
@@
 Of course this will not yield an actual value unless $\alpha_1$ yields.\\
-We will now consider a minimal inductive example.\pagebreak
+We will now consider a minimal inductive example.
 
@@
 \end{example}
+\pagebreak
+\subsection{Intuitionistic Example}
 
+For simplicity we will use a single sort (0-level type).\\
+Function symbols: $f(-)$\\
+Predicate symbols: $P(-), Q(-, -)$\\
+Axioms: $\forall x P(x)$, $\forall x\exists y x=f(y)$, $\forall x(P(x)\supset Q(f(x), x))$\\
+To Show: $\forall x \exists y Q(x, y)$\\
+
+\noindent Below is a ND-style proof and translation based on the calculus and correspondence presented in~\cite{pfenning2000constructive}.\\
+
+\small{
+\AxiomC{}
+\RightLabel{\scriptsize Axiom}
+\UnaryInfC{$\vdash\forall x\exists y f(y)=x$}
+\RightLabel{\scriptsize $\forall E$}
+\UnaryInfC{$\vdash\exists y f(y)=x$}
+\AxiomC{}
+\RightLabel{\scriptsize Axiom}
+\UnaryInfC{$\vdash \forall x(P(x)\supset Q(f(x), x))$}
+\RightLabel{\scriptsize $\forall E$}
+\UnaryInfC{$\vdash P(y)\supset Q(f(y), y)$}
+\AxiomC{}
+\RightLabel{\scriptsize Axiom}
+\UnaryInfC{$\vdash\forall x P(x)$}
+\RightLabel{\scriptsize $\forall E$}
+\UnaryInfC{$\vdash P(y)$}
+\RightLabel{\scriptsize $\supset\hspace{-2pt}E$}
+\BinaryInfC{$\vdash Q(f(y), y)$}
+\AxiomC{}
+\RightLabel{\scriptsize u}
+\UnaryInfC{[$\vdash f(y) = x$]}
+\RightLabel{\scriptsize subst}
+\BinaryInfC{$\vdash Q(x, y)$}
+\RightLabel{\scriptsize $\supset\hspace{-2pt}I^u$}
+\UnaryInfC{$f(y)=x\vdash Q(x, y)$}
+\RightLabel{\scriptsize$\exists I$}
+\UnaryInfC{$f(y)=x\vdash \exists y Q(x, y)$}
+\RightLabel{\scriptsize $\exists E$}
+\BinaryInfC{$\vdash \exists Q(x, y)$}
+\RightLabel{\scriptsize $\forall I$}
+\UnaryInfC{$\vdash\forall x \exists yQ(x, y)$}
+
+\hspace*{-1cm}
+\DisplayProof
+}
+\vspace{3cm}
+\small{
+	\AxiomC{}
+	\RightLabel{\scriptsize Axiom}
+	\UnaryInfC{$\vdash \alpha : \Pi_x\Sigma_y f(y)=x$}
+	\RightLabel{\scriptsize $\forall E$}
+	\UnaryInfC{$\vdash \alpha(x): \Sigma_y f(y)=x$}
+	\AxiomC{}
+	\RightLabel{\scriptsize Axiom}
+	\UnaryInfC{$\vdash \beta : \Pi_x(P(x)\supset Q(f(x), x))$}
+	\RightLabel{\scriptsize $\forall E$}
+	\UnaryInfC{$\vdash \beta(y) : P(y)\supset Q(f(y), y)$}
+	\AxiomC{}
+	\RightLabel{\scriptsize Axiom}
+	\UnaryInfC{$\vdash\Gamma\Pi_x P(x)$}
+	\RightLabel{\scriptsize $\forall E$}
+	\UnaryInfC{$\vdash\gamma(y) : P(y)$}
+	\RightLabel{\scriptsize $\supset\hspace{-2pt}E$}
+	\BinaryInfC{$\vdash \beta(y)(\gamma(y)) : Q(f(y), y)$}
+	\AxiomC{}
+	\RightLabel{\scriptsize u}
+	\UnaryInfC{[$\vdash\delta : f(y) = x$]}
+	\RightLabel{\scriptsize subst}
+	\BinaryInfC{$\vdash \beta(y)(\gamma(y))_\delta : Q(x, y)$}
+	\RightLabel{\scriptsize $\supset\hspace{-2pt}I^u$}
+	\UnaryInfC{$\delta : f(y)=x\vdash\beta(y)(\gamma(y))_\delta : Q(x, y)$}
+	\RightLabel{\scriptsize$\exists I$}
+	\UnaryInfC{$\delta : f(y)=x\vdash \langle y, \beta(y)(\gamma(y))_\delta\rangle :\Sigma_y Q(x, y)$}
+	\RightLabel{\scriptsize $\exists E$}
+	\BinaryInfC{$\vdash \textbf{let }\langle y, \delta\rangle = \alpha(x)\textbf{ in }\langle y,\beta(y)(\gamma(y)\delta)\rangle: \Sigma_y Q(x, y)$}
+	\RightLabel{\scriptsize $\forall I$}
+	\UnaryInfC{$\lambda x.\textbf{let }\langle y, \delta\rangle = \alpha(x)\textbf{ in }\langle y,\beta(y)(\gamma(y))_\delta\rangle:\Pi_x \Sigma_y Q(x, y)$}
+
+	\hspace*{-1cm}
+	\DisplayProof
+}
+\hfill
+\\\\\hspace{2cm}We can reduce the last term to $\lambda x.\langle \alpha(x)_1, \beta(\alpha(x)_1)(\gamma(\alpha(x)_1)_{\alpha(x)_2}\rangle$.
+
+\noindent Finally replacing the proof terms by units we obtain $\lambda x.\langle \alpha(x)_1, ()\rangle : \Pi_x\Sigma_y \top$ from which we may further extract $\lambda x. \alpha(x)_1$.
+
+\pagebreak
 \bibliographystyle{acm}

### 2021-10-20 Notes/notes.tex
equality first thoughts

@@
 \maketitle
+
+\section{Type theory}
+
+As we are interested in a proof-as-programs correspondence for first order logic our underlying type theory has to be of dependant nature, meaning that types can and will depend on terms.
+
+One interesting aspect of choosing a type theory is how to handle equality.
+There are many different ways to do this but for our purposes the relatively simple model in agda~\cite{agda} is sufficient (?).
+Namely the only equality inherent to the type theory is definitional equality, i.e.
+equality induced by equivalence under $\beta$-reductions and $\eta$-equivalences whereas propositional equality is itself an inductive type (or rather inductive schema) with a single constructor $$\text{refl}_x : x =_A x$$ (the type is parametrized by a type $A$ and $x:A$), i.e.
+for all types $A$ and $x:A$ we have the rule
+\begin{center}
+\AxiomC{}
+\RightLabel{\scriptsize $=$-intro}
+\UnaryInfC{$\text{refl}_x : x =_A x$}
+\DisplayProof
+\end{center}
+Many of the expected properties of equality hold. For instance we can derive the term
+\begin{align*}
+	\text{subst} &: \Pi_{A}\Pi_{x, y:A}\Pi_{P:A\to Type}\Pi_{e : x=_Ay}\Pi_{p:P x}P y\\
+	\text{subst} &\: A\:x\:y\:P\: \text{refl}_x\: px = px 
+\end{align*}
+by pattern matching, i.e. matching refl the type checker can unify $x$ and $y$.
+
+However there are some issues with this handling of equality.
+For example it is impossible to define quotient structures: Let's say we prove prime decomposition for PIDs and want to apply it to some ring.
+Then the ring will be most likely defined as a setoid-like structure with the equality being a more coarse relation on the set of terms and in particular not equal to the equality above.
+Our proof-as-programs translation will have to take this into account.
 
@@
 	\RightLabel{\scriptsize Axiom}
-	\UnaryInfC{$\vdash\Gamma\Pi_x P(x)$}
+	\UnaryInfC{$\vdash\gamma : \Pi_x P(x)$}
 	\RightLabel{\scriptsize $\forall E$}
@@
 	\RightLabel{\scriptsize subst}
-	\BinaryInfC{$\vdash \beta(y)(\gamma(y))_\delta : Q(x, y)$}
+	\BinaryInfC{$\vdash \text{subst}(f(y),x,\Pi_tQ(t, y),\delta,\beta(y)(\gamma(y)) : Q(x, y)$}
 	\RightLabel{\scriptsize $\supset\hspace{-2pt}I^u$}
-	\UnaryInfC{$\delta : f(y)=x\vdash\beta(y)(\gamma(y))_\delta : Q(x, y)$}
+	\UnaryInfC{$\delta : f(y)=x\vdash\text{subst}(f(y),x,\Pi_tQ(t, y),\delta,\beta(y)(\gamma(y)) : Q(x, y)$}
 	\RightLabel{\scriptsize$\exists I$}
-	\UnaryInfC{$\delta : f(y)=x\vdash \langle y, \beta(y)(\gamma(y))_\delta\rangle :\Sigma_y Q(x, y)$}
+	\UnaryInfC{$\delta : f(y)=x\vdash \langle y, \text{subst}(f(y),x,\Pi_tQ(t, y),\delta,\beta(y)(\gamma(y))\rangle :\Sigma_y Q(x, y)$}
 	\RightLabel{\scriptsize $\exists E$}
-	\BinaryInfC{$\vdash \textbf{let }\langle y, \delta\rangle = \alpha(x)\textbf{ in }\langle y,\beta(y)(\gamma(y)\delta)\rangle: \Sigma_y Q(x, y)$}
+	\BinaryInfC{$\vdash \textbf{let }\langle y, \delta\rangle = \alpha(x)\textbf{ in }\langle y,\text{subst}(f(y),x,\Pi_tQ(t, y),\delta,\beta(y)(\gamma(y))\rangle: \Sigma_y Q(x, y)$}
 	\RightLabel{\scriptsize $\forall I$}
-	\UnaryInfC{$\lambda x.\textbf{let }\langle y, \delta\rangle = \alpha(x)\textbf{ in }\langle y,\beta(y)(\gamma(y))_\delta\rangle:\Pi_x \Sigma_y Q(x, y)$}
+	\UnaryInfC{$\lambda x.\textbf{let }\langle y, \delta\rangle = \alpha(x)\textbf{ in }\langle y,\text{subst}(f(y),x,\Pi_tQ(t, y),\delta,\beta(y)(\gamma(y))\rangle:\Pi_x \Sigma_y Q(x, y)$}
 
@@
 \hfill
-\\\\\hspace{2cm}We can reduce the last term to $\lambda x.\langle \alpha(x)_1, \beta(\alpha(x)_1)(\gamma(\alpha(x)_1)_{\alpha(x)_2}\rangle$.
+\\\\\hspace{2cm}We can reduce the last term to $\lambda x.\langle \alpha(x)_1, \text{subst}(f(\alpha(x)_1),x,\Pi_tQ(t, \alpha(x)_1),\alpha(x)_2,\beta(\alpha(x)_1)(\gamma(\alpha(x)_1))\rangle$.
 

### 2021-11-08 Notes/notes.tex
added vampire example

@@
 \subsection{Examples and Translations}
+
+\begin{example}\hfill
+	\begin{align*}
+		1. &\forall x_0 : \text{Nat}(z + x_0 = x_0) & \text{Axiom}\\
+		2.&\forall x_0x_1: \text{Nat}(s(x_0 + x_1) = s(x_0) + x_1) & \text{Axiom}\\
+		3.&\neg\forall x_0 : \text{Nat}\exists x_1:\text{Nat}(x_1 + z = x_0) & \text{negated claim}\\
+		4.&(\forall x_0:\text{Nat}\exists x_1:\text{Nat}, x_1 + z = x_0)\vee(\forall x_0:\text{Nat},z\neq x_0 + z)\vee&\\
+		&\exists x_0:\text{Nat}(\forall x_1:\text{Nat}, x_1 + z\neq s(x_0)\wedge \exists x_1:\text{Nat}, x_1+z=x_0)&\text{Induction axiom (ENNF)}\\
+		5.&\exists x_0: \text{Nat}\forall x_1: \text{Nat}(x_1 + z \neq x_0) & \text{ENNF transformation 3}\\
+		6.&\forall x_1: \text{Nat}(x_1 + z \neq sk_0) & \text{skolemization 5}\\
+		7.&(\forall x_0:\text{Nat}, sk_1(x_0) + z = x_0)\vee(\forall x_0:\text{Nat},z\neq x_0 + z)\vee&\\
+		&(\forall x_1:\text{Nat}, x_1 + z\neq s(sk_2)\wedge sk_3+z=sk_2)&\text{skolemization 4}\\
+		8.&x_1 + z\neq sk_0&\text{cnf 6}\\
+		9.&sk_1(x_0) + z = x_0\vee z\neq x_1 + z\vee sk_2 = sk_3 + z&\text{cnf 7}\\
+		10.&sk_1(x_0) + z = x_0\vee z\neq x_1 + z\vee x_2 + z\neq s(sk_2)&\text{cnf 7}\\
+		11.&s(x_0 + x_1) = s(x_0) + x_1&\text{cnf 2}\\
+		12.&z + x_0 = x_0&\text{cnf 1}\\
+		13.&sk_1(x_0) + z = x_0\vee z\neq z\vee sk_2 = sk_3 + z&\text{superposition 9, 12}\\
+		14.&sk_1(x_0) + z = x_0\vee sk_2 = sk_3 + z&\text{trivial inequality removal 13}\\
+		15.&sk_1(x_0) + z = x_0\vee z\neq x_1 + z\vee s(x_3 + z)\neq s(sk_2)&\text{supserposition 10, 11}\\
+		16.&sk_0\neq x_0\vee sk_2 = sk_3 + z&\text{superposition 8, 14}\\
+		17.&sk_2 = sk_3 + z&\text{equality resolution 16}\\
+		18.&sk_1(x_0) + z = x_0\vee z\neq x_1 + z\vee s(sk_2)\neq s(sk_2)&\text{superposition 15, 17}\\
+		19.&sk_1(x_0) + z = x_0\vee z\neq x_1 + z&\text{trivial inequality removal 18}\\
+		20.&sk_1(x_0) + z = x_0\vee z\neq z&\text{superposition 12, 19}\\
+		21.&sk_1(x_0) + z = x_0&\text{trivial inequality removal 20}\\
+		22.&sk_0 \neq x_0&\text{superoisition 21, 8}\\
+		23.&\bot&\text{equality resolution 22}
+	\end{align*}
+\end{example}
+
+\begin{example}\hfill
+	\begin{align*}
+		1. &\vdash \lambda f. f(\lambda x.\text{id}_{x}): \text{Cont }\Pi_{x_0 : \text{Nat}}(z + x_0 = x_0) & \text{Axiom}\\
+		2.&\vdash \lambda f. f(\lambda x_0x_1.\text{id}_{s(x_0 + x_1)}): \text{Cont }\Pi_{x_0:\text{Nat}}\Pi_{x_1: \text{Nat}}(s(x_0 + x_1) = s(x_0) + x_1) & \text{Axiom}\\
+		3.&\vdash f: (\Pi_{x_0 : \text{Nat}}\Sigma_{x_1:\text{Nat}}x_1 + z = x_0)\to\bot & \text{negated claim}\\
+		4.&\vdash \lambda f.:\text{Cont }(\Pi_{x_0 : \text{Nat}}\Sigma_{x_1:\text{Nat}}x_1 + z = x_0)\vee(\Pi_{x_0 : \text{Nat}}z\neq x_0 + z)\vee&\\
+		&\Sigma_{x_0:\text{Nat}}(\Pi_{x_1:\text{Nat}}x_1 + z\neq s(x_0)\wedge \Sigma_{x_1:\text{Nat}} x_1+z=x_0)&\text{Induction axiom (ENNF)}\\
+		5.&\exists x_0: \text{Nat}\forall x_1: \text{Nat}(x_1 + z \neq x_0) & \text{ENNF transformation 3}\\
+		6.&\forall x_1: \text{Nat}(x_1 + z \neq sk_0) & \text{skolemization 5}\\
+		7.&(\forall x_0:\text{Nat}, sk_1(x_0) + z = x_0)\vee(\forall x_0:\text{Nat},z\neq x_0 + z)\vee&\\
+		&(\forall x_1:\text{Nat}, x_1 + z\neq s(sk_2)\wedge sk_3+z=sk_2)&\text{skolemization 4}\\
+		8.&x_1 + z\neq sk_0&\text{cnf 6}\\
+		9.&sk_1(x_0) + z = x_0\vee z\neq x_1 + z\vee sk_2 = sk_3 + z&\text{cnf 7}\\
+		10.&sk_1(x_0) + z = x_0\vee z\neq x_1 + z\vee x_2 + z\neq s(sk_2)&\text{cnf 7}\\
+		11.&s(x_0 + x_1) = s(x_0) + x_1&\text{cnf 2}\\
+		12.&z + x_0 = x_0&\text{cnf 1}\\
+		13.&sk_1(x_0) + z = x_0\vee z\neq z\vee sk_2 = sk_3 + z&\text{superposition 9, 12}\\
+		14.&sk_1(x_0) + z = x_0\vee sk_2 = sk_3 + z&\text{trivial inequality removal 13}\\
+		15.&sk_1(x_0) + z = x_0\vee z\neq x_1 + z\vee s(x_3 + z)\neq s(sk_2)&\text{supserposition 10, 11}\\
+		16.&sk_0\neq x_0\vee sk_2 = sk_3 + z&\text{superposition 8, 14}\\
+		17.&sk_2 = sk_3 + z&\text{equality resolution 16}\\
+		18.&sk_1(x_0) + z = x_0\vee z\neq x_1 + z\vee s(sk_2)\neq s(sk_2)&\text{superposition 15, 17}\\
+		19.&sk_1(x_0) + z = x_0\vee z\neq x_1 + z&\text{trivial inequality removal 18}\\
+		20.&sk_1(x_0) + z = x_0\vee z\neq z&\text{superposition 12, 19}\\
+		21.&sk_1(x_0) + z = x_0&\text{trivial inequality removal 20}\\
+		22.&sk_0 \neq x_0&\text{superoisition 21, 8}\\
+		23.&\bot&\text{equality resolution 22}
+	\end{align*}
+\end{example}
 

### 2021-11-16 Notes/notes.tex
Plan

@@
 \maketitle
+
+\section{Plan}
+
+Goal:
+\begin{itemize}
+	\item Extract program from resolution proofs of $\forall\exists$-sentences over inductive datatypes.
+\end{itemize}
+
+\noindent Steps:
+
+\begin{itemize}
+	\item Give extensions of G\"odel's System \textbf{T} and HAS + EM$_1$ + SK$_1$ to arbitrary inductive datatypes (possibly GADTs) and prove (or rather check) properties, i.e.
+strong normalization, cut-elimination, uniqueness of normal forms.
+	\item Adapt the results of~\cite{aschieri2014interactive} to exhibit realizers in $\mathcal{F} + \Phi$ for cut-free proofs in the extended version of HAS + EM$_1$ + SK$_1$ of formulas $\forall x\exists y Pxy$ where $P$ is a predicate in the extended version of \textbf{T}.
+	\item Adapt the iterative learning from~\cite{aschieri2014interactive} to extract $\lambda$-terms from realizers.
+	\item Give a proof translation from resolution proofs to cut-free sequent calculus proofs (already done?).
+\end{itemize}
+
+\noindent Questions:
+\begin{itemize}
+	\item When does the translated proof require more than EM$_1$?
+	\item Using the methods in~\cite{aschieri2014interactive} the extracted term will be in simple $\lambda$-calculus. Is is possible to obtain a term in $\mathcal{F}$? (Talk with Federico Aschieri)
+	\item Can the predicates be defined in $\mathcal{F}$ instead of $\mathcal{T}$? (probably yes)
+	\item What happens if we add non-inductive theories?
+
+\end{itemize}
+
+\pagebreak
 

### 2022-01-11 Notes/notes.tex
Explained mechanism + examples

--- 2021-10-20 Notes/notes.tex
@@
 \usepackage{bussproofs}
+\usepackage[left=3.00cm, right=3.00cm, top=3.00cm,bottom=3.0cm]{geometry}
 
@@
 
-\section{Type theory}
+\section{Plan}
 
-As we are interested in a proof-as-programs correspondence for first order logic our underlying type theory has to be of dependant nature, meaning that types can and will depend on terms.
+Goal:
+\begin{itemize}
+	\item Extract program from resolution proofs of $\forall\exists$-sentences over inductive datatypes.
+\end{itemize}
 
-One interesting aspect of choosing a type theory is how to handle equality.
-There are many different ways to do this but for our purposes the relatively simple model in agda~\cite{agda} is sufficient (?).
-Namely the only equality inherent to the type theory is definitional equality, i.e.
-equality induced by equivalence under $\beta$-reductions and $\eta$-equivalences whereas propositional equality is itself an inductive type (or rather inductive schema) with a single constructor $$\text{refl}_x : x =_A x$$ (the type is parametrized by a type $A$ and $x:A$), i.e.
-for all types $A$ and $x:A$ we have the rule
+\noindent Steps:
+
+\begin{itemize}
+	\item Give extensions of G\"odel's System \textbf{T} and HAS + EM$_1$ + SK$_1$ to arbitrary inductive datatypes (possibly GADTs) and prove (or rather check) properties, i.e.
+strong normalization, cut-elimination, uniqueness of normal forms.
+	\item Adapt the results of~\cite{aschieri2014interactive} to exhibit realizers in $\mathcal{F} + \Phi$ for cut-free proofs in the extended version of HAS + EM$_1$ + SK$_1$ of formulas $\forall x\exists y Pxy$ where $P$ is a predicate in the extended version of \textbf{T}.
+	\item Adapt the iterative learning from~\cite{aschieri2014interactive} to extract $\lambda$-terms from realizers.
+	\item Give a proof translation from resolution proofs to cut-free sequent calculus proofs (already done?).
+\end{itemize}
+
+\noindent Questions:
+\begin{itemize}
+	\item When does the translated proof require more than EM$_1$?
+	\item Using the methods in~\cite{aschieri2014interactive} the extracted term will be in simple $\lambda$-calculus. Is is possible to obtain a term in $\mathcal{F}$? (Talk with Federico Aschieri)
+	\item Can the predicates be defined in $\mathcal{F}$ instead of $\mathcal{T}$? (probably yes)
+	\item What happens if we add non-inductive theories?
+
+\end{itemize}
+
+\pagebreak
+
+\section{Extracting constructive content from resolution proofs}
+\subsection{The superposition calculus}
+First let us define the calculus from which we wish to extract programs.
+It will comprise a core set of rules used in the vampire theorem prover taken from~\cite{Kov_cs_2013}.
+Note that we pay special attention to the usually neglected (and highly non-constructive) part of CNF transformation and Skolemization.
+Also note that we neglect the simplification ordering which is necessary to formulate a strategy for proof search but not for our proof transformation.
+
+\noindent\textbf{Resolution.}
 \begin{center}
-\AxiomC{}
-\RightLabel{\scriptsize $=$-intro}
-\UnaryInfC{$\text{refl}_x : x =_A x$}
+\AxiomC{$A\vee B$}
+\AxiomC{$\neg A'\vee C$}
+\BinaryInfC{$(B\vee C)\theta$}
 \DisplayProof
 \end{center}
-Many of the expected properties of equality hold. For instance we can derive the term
-\begin{align*}
-	\text{subst} &: \Pi_{A}\Pi_{x, y:A}\Pi_{P:A\to Type}\Pi_{e : x=_Ay}\Pi_{p:P x}P y\\
-	\text{subst} &\: A\:x\:y\:P\: \text{refl}_x\: px = px 
-\end{align*}
-by pattern matching, i.e. matching refl the type checker can unify $x$ and $y$.
+where $\theta$ is an mgu of $A$ and $A'$.
 
-However there are some issues with this handling of equality.
-For example it is impossible to define quotient structures: Let's say we prove prime decomposition for PIDs and want to apply it to some ring.
-Then the ring will be most likely defined as a setoid-like structure with the equality being a more coarse relation on the set of terms and in particular not equal to the equality above.
-Our proof-as-programs translation will have to take this into account.
+\noindent\textbf{Factoring.}
+\begin{center}
+	\AxiomC{$A\vee A'\vee B$}
+	\UnaryInfC{$(A\vee B)\theta$}
+	\DisplayProof
+\end{center}
+where $\theta$ is an mgu of $A$ and $A'$.
 
-\section{Proof translation for saturation-based theorem proving with induction}
+\noindent\textbf{Superposition.}
+\begin{center}
+	\AxiomC{$l=r\vee B$}
+	\AxiomC{$L[s]\vee C$}
+	\BinaryInfC{$(L[r]\vee B\vee C)\theta$}
+	\DisplayProof
+	\AxiomC{$l=r\vee B$}
+	\AxiomC{$t[s] = t'\vee C$}
+	\BinaryInfC{$(t[r] = t'\vee B\vee C)\theta$}
+	\DisplayProof
+	\AxiomC{$l=r\vee B$}
+	\AxiomC{$t[s]\neq t'\vee C$}
+	\BinaryInfC{$(t[r]\neq t'\vee B\vee C)\theta$}
+	\DisplayProof
+\end{center}
+where $\theta$ is an mgu of $l$ and $s$, $s$ is not a variable, $L[s]$ is not an equality literal.
 
-Most modern theorem provers for first order logic utilize resolution and saturation based techniques.
-The standard intuitionistic proofs-as-programs correspondence fails as even the very method itself crucially depends on double negation elimination.
-As a first step we work out a way to convert proofs of $\forall\exists$-proofs based on the calculus presented in~\cite{Echenim_2019} into programs.
-An advantage of the calculus as presented in~\cite{Echenim_2019} is that it can (at least in theory) handle general induction for inductive datatype.
-Many advances have recently been made for incorporating induction in existing theorem provers~\cite{hajdu2021automating}\cite{10.1007/978-3-030-53518-6_8} however all of the calculi fail to support induction in the general case, in particular sentences involving existential quantifiers and requiring induction depth greater than $1$ are problematic.
+\noindent\textbf{Equality Resolution.}
+\begin{center}
+	\AxiomC{$s\neq t\vee C$}
+	\UnaryInfC{$C\theta$}
+	\DisplayProof
+\end{center}
+where $\theta$ is an mgu of $s$ and $t$.
 
-The approach in~\cite{Echenim_2019} seems promising as, despite being able to handle general induction, there is no explosion in the number of clauses due to a clever use of so-called \textit{constrained clauses}, special symbols $\textbf{T}_\alpha^\prec$ for induction and a special activation mechanism for potential inductive invariants, although how it performs in practice is yet to be seen.
+\noindent\textbf{Equality Factoring.}
+\begin{center}
+	\AxiomC{$s=t\vee s'=t'\vee C$}
+	\UnaryInfC{$(s=t\vee t\neq t'\vee C)\theta$}
+	\DisplayProof
+\end{center}
+where $\theta$ is an mgu of $s$ and $s'$.
 
-Program translation for classical proofs has been outlined in~\cite{Griffin_1990}.
-Based on this~\cite{de_Groote_1995} and~\cite{Barbanera_1996} have been developed.
-Our a approach is a combination of both.
-At the moment it is quite crude and is mostly based on the exception method from~\cite{de_Groote_1995}.
-We aim to later translate this to the CPS approach presented in the same paper to better see how it relates to the normal proofs-as-program correspondence known from intuitionistic logic.
+First let us look at an example of the transformation we going to do.
+We shall look at a proof of $\forall x\exists y: f(y) = g(x)$ from $\forall x: f(x) = g(h(x))$ and $\forall x: h(h'(x)) = x$.
+The program we extract hopefully is $h'$.
 
-\subsection{The Calculus}
+\begin{align}
+	[\text{axiom}]&&\forall x: f(x) = g(h(x))\\
+	[\text{cnf 1}]&&f(x_0) = g(h(x_0))\\
+	[\text{axiom}]&&\forall x: h(h'(x)) = x\\
+	[\text{cnf 3}]&&h(h'(x_0)) = x_0\\
+	[\text{negated conjecture}]&&\neg\forall x\exists y: f(y) = g(x)\\
+	[\text{ennf 5}]&&\exists x\forall y: f(y)\neq g(x)\\
+	[\text{choice axiom}]&&\exists x\forall y: f(y)\neq g(x)\rightarrow \forall x : f(x) \neq g(s_0)\\
+	[\text{skolemization 6, 7}]&&\forall x: f(x)\neq g(s_0)\\
+	[\text{cnf 8}]&&f(x_1)\neq g(s_0)\\
+	[\text{superposition 9, 2}]&\{\}&g(s_0)\neq g(h(x_0))\\
+	[\text{superposition 10, 4}]&\{x_0\mapsto h'(x_0)\}&g(x_0)\neq g(s_0)\\
+	[\text{equality resolution 11}]&&\bot
+\end{align}
 
-As explained above our calculus is based on~\cite{Echenim_2019}. We take the core rules, which are not mentioned in~\cite{Echenim_2019}, from~\cite{Kov_cs_2013}.
+We first use the well known trick of adding the original conjecture to the negated one from~\cite{Luckham_1971} to transform this into a classical proof of $\forall x\exists y: f(x) = g(y)$ rather than a refutation of its negation:
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}]&&\forall x: f(x) = g(h(x))\\
+	[\text{cnf 1}]&&f(x_0) = g(h(x_0))\\
+	[\text{axiom}]&&\forall x: h(h'(x)) = x\\
+	[\text{cnf 3}]&&h(h'(x_0)) = x_0\\
+	[\text{\color{red}{tautology}}]&&\forall x\exists y: f(y) = g(x)\vee\neg\forall x\exists y: f(y) = g(x)\\
+	[\text{ennf 5}]&&\forall x\exists y: f(y) = g(x)\vee\exists x\forall y: f(y)\neq g(x)\\
+	[\text{choice axiom}]&&\exists x\forall y: f(y)\neq g(x)\rightarrow \forall x : f(x) \neq g(s_0)\\
+	[\text{choice axiom}]&&\forall x(\exists y: f(y)= g(x)\rightarrow f(s_1(x)) = g(x))\\
+	[\text{skolemization 6, 7}]&&\forall x\exists y: f(y) = g(x)\vee\forall x: f(x)\neq g(s_0)\\
+	[\text{skolemization 9, 8}]&&\forall x: f(s_1(x)) = g(x)\vee\forall x: f(x)\neq g(s_0)\\
+	[\text{cnf 8}]&&f(s_1(x_1)) = g(x_1)\vee f(x_0)\neq g(s_0)\\
+	[\text{superposition 11, 2}]&\{\}&f(s_1(x_1)) = g(x_1)\vee g(s_0)\neq g(h(x_0))\\
+	[\text{superposition 12, 4}]&\{x_0\mapsto h'(x_0)\}&f(s_1(x_1)) = g(x_1)\vee g(x_0)\neq g(s_0)\\
+	[\text{equality resolution 13}]&\{x_0\to s_0\}&f(s_1(x_1)) = g(x_1)
+\end{align}
 
-The first five inferences are standard in resolution based theorem proving. We assume some simplification ordering $\succeq^*$.\\
-\textbf{Resolution.}
-$$\inferrule{{A}\vee C_1\\{\neg B}\vee C_2}{(C_1\vee C_2)\theta}$$
-where $\theta$ is a mgu of $A$ and $\neg B$.\\
-\textbf{Factoring.}
-$$\inferrule{{A}\vee{\neg B}\vee C}{(A\vee C)\theta}$$
-where theta is a mgu of $A$ and $B$.\\
-\textbf{Superposition.}
-$$\inferrule{{l=r}\vee C_1\\{L[s]}\vee C_2}{(L[r]\vee C_1\vee C_2)\theta}\:
-\inferrule{{l=r}\vee C_1\\ {t[s]=t'}\vee C_2}{(t[r]=t'\vee C_1\vee C_2)\theta}\:
-\inferrule{{l=r}\vee C_1\\{t[s]\neq t'}\vee C_2}{(t[r]\neq t'\vee C_1\vee C_2)\theta}$$
-where $\theta$ is a mgu of $l$ and $s$, $s$ is not a variable, $r\theta\not\succeq^* l\theta$, $L[s]$ is not an equality literal, and $t'\theta\succeq^* t[s]\theta$.\\
-\textbf{Equality Resolution.}
-$$\inferrule{{s\neq t}\vee C}{C\theta}$$
-where $\theta$ is a mgu of $s$ and $t$.\\
-\textbf{Equality Factoring.}
-$$\inferrule{{s=t}\vee{s'=t'}\vee C}{(s=t\vee t=t'\vee C)\theta}$$
-where $\theta$ is an mgu of $s$ and $s'$, $t\theta\not\succeq^* s\theta$, and $t'\theta\succeq^* t\theta$.
+Next we eliminate free variables by propagating substitutions:
 
-In addition to regular clauses and inference rules we introduce c-clauses and special rules for them, i.e. (all of this will be made more precise later)
-\begin{definition}
-	A \textit{constrained clause} (c-clause) is an expression of  the form $\llb C\:|\:\mathcal X \rrb$, where
-	\begin{itemize}
-		\item C is a clause,
-		\item $\mathcal{X}$ is a conjunction of the form $\bigwedge_{i=1}^ n f_i(\mathbf{u_i})\simeq v_i\wedge\bigwedge_{i=1}^m \mathbf{T}_{\alpha_i}^\prec(z_i,\textbf{w}_i)$, where $n, m\geq 0$ and $\forall i\in\{1\dots n\}, f_i$ are uninterpreted function symbols (possibly nullary).
-	\end{itemize}
-\end{definition}
-The semantics of $\llb C\:|\:\mathcal X \rrb$ is more or less ``if all constraints in $\mathcal X$ hold, then $C$ holds" and of $\mathbf{T}_{\alpha_i}^\prec(z,\mathbf{w})$ ``for all $z'\prec z$ $\alpha(z', \mathbf{w})$ holds" where $\prec$ is the standard wfo of an inductive type.
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}]&&\forall x: f(x) = g(h(x))\\
+	[\text{cnf 1}]&&f(h'(x_0)) = g(h(h'(x_0)))\\
+	[\text{axiom}]&&\forall x: h(h'(x)) = x\\
+	[\text{cnf 3}]&&h(h'(x_0)) = x_0\\
+	[\text{tautology}]&&\forall x\exists y: f(y) = g(x)\vee\neg\forall x\exists y: f(y) = g(x)\\
+	[\text{ennf 5}]&&\forall x\exists y: f(y) = g(x)\vee\exists x\forall y: f(y)\neq g(x)\\
+	[\text{choice axiom}]&&\exists x\forall y: f(y)\neq g(x)\rightarrow \forall x : f(x) \neq g(s_0)\\
+	[\text{choice axiom}]&&\forall x(\exists y: f(y)= g(x)\rightarrow f(s_1(x)) = g(x))\\
+	[\text{skolemization 6, 7}]&&\forall x\exists y: f(y) = g(x)\vee\forall x: f(x)\neq g(s_0)\\
+	[\text{skolemization 9, 8}]&&\forall x: f(s_1(x)) = g(x)\vee\forall x: f(x)\neq g(s_0)\\
+	[\text{cnf 8}]&&f(s_1(x_1)) = g(x_1)\vee f(h'(x_0))\neq g(s_0)\\
+	[\text{superposition 11, 2}]&\{\}&f(s_1(x_1)) = g(x_1)\vee g(s_0)\neq g(h(h'(x_0)))\\
+	[\text{superposition 12, 4}]&\{\}&f(s_1(x_1)) = g(x_1)\vee g(x_0)\neq g(s_0)\\
+	[\text{equality resolution 13}]&\{x_0\to s_0\}&f(s_1(x_1)) = g(x_1)
+\end{align}
 
-All of the above inference rules extend naturally to constrained clauses, i.e.
-if $\tiny{\inferrule{H_1,\dots,H_n}{C\theta}}$ is an inference rule for regular clauses, then $$\inferrule{\llb H_1\:|\:\mathcal X_1\rrb,\dots,\llb H_1\:|\:\mathcal X_1\rrb}{\llb C\:|\:\mathcal X_1\wedge\dots\wedge\mathcal X_n\rrb\theta}$$ is the corresponding inference rule for c-clauses.
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}]&&\forall x: f(x) = g(h(x))\\
+	[\text{cnf 1}]&&f(h'(s_0)) = g(h(h'(s_0)))\\
+	[\text{axiom}]&&\forall x: h(h'(x)) = x\\
+	[\text{cnf 3}]&&h(h'(s_0)) = s_0\\
+	[\text{tautology}]&&\forall x\exists y: f(y) = g(x)\vee\neg\forall x\exists y: f(y) = g(x)\\
+	[\text{ennf 5}]&&\forall x\exists y: f(y) = g(x)\vee\exists x\forall y: f(y)\neq g(x)\\
+	[\text{choice axiom}]&&\exists x\forall y: f(y)\neq g(x)\rightarrow \forall x : f(x) \neq g(s_0)\\
+	[\text{choice axiom}]&&\forall x(\exists y: f(y)= g(x)\rightarrow f(s_1(x)) = g(x))\\
+	[\text{skolemization 6, 7}]&&\forall x\exists y: f(y) = g(x)\vee\forall x: f(x)\neq g(s_0)\\
+	[\text{skolemization 9, 8}]&&\forall x: f(s_1(x)) = g(x)\vee\forall x: f(x)\neq g(s_0)\\
+	[\text{cnf 8}]&&f(s_1(x_1)) = g(x_1)\vee f(h'(s_0))\neq g(s_0)\\
+	[\text{superposition 11, 2}]&\{\}&f(s_1(x_1)) = g(x_1)\vee g(s_0)\neq g(h(h'(s_0)))\\
+	[\text{superposition 12, 4}]&\{\}&f(s_1(x_1)) = g(x_1)\vee g(s_0)\neq g(s_0)\\
+	[\text{equality resolution 13}]&\{\}&f(s_1(x_1)) = g(x_1)
+\end{align}
 
-Furthermore we have some special rules for c-clauses:\\
-\textbf{Constraint Factorization.}
-$$\inferrule{\llb C\:|\:l_1\wedge l_2\wedge \mathcal{X}\rrb}{\llb C\:|\:l_1\wedge \mathcal{X}\rrb\theta}$$where $\theta$ is an mgu of $l_1$ and $l_2$.\\
-\textbf{Abstraction.}
-$$\inferrule{\llb C[f(\mathbf{t})]_p\:|\:\mathcal{X}\rrb}{\llb C[x]_p\:|\:f(\mathbf{t})\simeq x\wedge\mathcal{X}\rrb}$$ where $f$ is a uninterpreted function symbol, $x$ is a fresh variable.\\
-\textbf{Instantiation.}
-$$\inferrule{C\:|\:\mathcal{X}\wedge t\simeq s\rrb}{\llb C\:|\:\mathcal{X}\rrb\theta}$$
-where $\theta$ is a mgu of $t$ and $s$.
+Next we remove all the skolem constants in the conjecture-tautology by unification, propagate this change, and finally reinterpet superposition as composition to yield a valid intuitionistic proof:
 
-In addition there are some rules that generate candidates for an inductive invariant in~\cite{Echenim_2019}.
-We will leave those out for now an choose our variant ``magically" by hand.
-For every formula $\alpha[z,\textbf{w}]$ we than have axiom $$\Gamma_\mathbf{T_\alpha^\prec}:= \llb z' \not\prec z\vee \alpha(z', \mathbf{w})\:|\:\mathbf{T}_\alpha^\prec(z,\mathbf{w})\rrb$$ and $\prec $ is axiomatized by $$\Gamma_\prec^s:=\llb x_i\prec s(x_1,\dots, x_n)\:|\:\top\rrb\indent \Gamma_\prec:=\llb x\not\prec y\vee y\not\prec z\vee x\prec z\:|\:\top\rrb$$for every n-ary inductive constructor $s$, $i\in\{1\dots n\}$.
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}]&&\forall x: f(x) = g(h(x))\\
+	[\text{instantiation 1}]&&f(h'(x_1)) = g(h(h'(x_1)))\\
+	[\text{axiom}]&&\forall x: h(h'(x)) = x\\
+	[\text{instantiation 3}]&&h(h'(x_1)) = x_1\\
+	[\text{tautology}]&&f(h'(x_1)) = g(x_1)\Rightarrow f(h'(x_1)) = g(x_1)\\
+	[\text{equality 5, 2}]&\{\}&g(x_1)= g(h(h'(x_1)))\Rightarrow f(h'(x_1)) = g(x_1)\\
+	[\text{equality 6, 4}]&\{\}&g(x_1)= g(x_1)\Rightarrow f(h'(x_1)) = g(x_1)\\
+	[\text{equality 13}]&\{\}&f(h'(x_1)) = g(x_1)
+\end{align}
 
-Finally we have additional rules for inductive types.
-Since we are (for now) not interested in how to generate inductive invariants it is possible to present just one combination rule for Domain Decomposition and Induction.\\
-\textbf{Induction.}
-$$\inferrule{\llb C_1\:|\:\beta_1(t_1)\wedge \mathcal{X}_1\rrb\dots\llb C_n\:|\:\beta_n(t_n)\wedge \mathcal{X}_n\rrb}{\llb C_1\vee\dots\vee C_n\vee \alpha(x,\mathbf{w}_1)\:|\:\mathcal{X}_1\setminus\{\mathbf{T}_{\alpha}^\prec(t_1,\mathbf w_1)\}\wedge\dots\wedge\mathcal{X}_n\setminus\{\mathbf{T}_{\alpha}^\prec(t_n,\mathbf w_n)\}\wedge\beta_1(x)\rrb\theta}$$where $\theta$ is the most general idempotent substitution such that $\beta_j\theta\subseteq\beta_1\theta$ for $n\in\{2\dots n\}$ and a mgu for $\mathbf{w}_1,\dots,\mathbf{w}_n$, the variables occurring in $t_1,\dots,t_n$ do not occur in $C_i,\beta_i,\mathbf{w}_i$ or $\mathcal{X}_i$.
-$\{t_1\dots t_n\}$ is \textit{covering}, i.e.
-the whole type can be constructed using $\{t_1\dots t_n\}$, for instance for $nat$ both $\{z, s(x)\}$ and $\{z, s(z), s(s(x))\}$ are covering and $x$ is some fresh variable.
+Let us look at a second example: Consider the sentences $\forall x : p(x_0, f(x_0))$, $\forall x_0, x_1: (p(x_0, x_1)\Rightarrow q(x_0, g(x_1))$, $\forall x_0, x_1: (q(x_0, x_1)\Rightarrow r(x_0, h(x_1)))$.
+We take a look at a proof of $\forall x_0\exists x_1: r(x_0, x_1)$ which hopefully gives us $h\circ g\circ f$.
+First consider the output of vampire
 
-Note that setting $\alpha=\bot$ and $\mathbf{w_i}=\emptyset$ we obtain the regular Domain Decomposition Rule from~\cite{Echenim_2019}.
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{negated conjecture}] &&\neg\forall  x_0 : \exists x_1 : r(x_0,x_1)\\
+	[\text{ennf transformation 2}] && \forall x_0, x_1 : (q(x_0,g(x_1)) \vee \neg p(x_0,x_1)) \\
+	[\text{ennf transformation 3}] &&\forall x_0, x_1 : (r(x_0,h(x_1)) \vee \neg q(x_0, x_1)) \\
+	[\text{ennf transformation 4}] && \exists x_0 \forall x_1 : \neg r(x_0, x_1)\\
+	[\text{choice axiom}] && \exists x_0 \forall x_1 : \neg r(x_0, x_1) \Rightarrow \forall x_1 : \neg r(s_0,x_1)\\
+	[\text{skolemisation 7,8}] && \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{cnf transformation 1}] && p(x_0,f(x_0))\\
+	[\text{cnf transformation 5}] && q(x_1,g(x_2)) \vee \neg p(x_1, x_2)\\
+	[\text{cnf transformation 6}] && r(x_3,h(x_4)) \vee \neg q(x_3, x_4)\\
+	[\text{cnf transformation 9}] &&. \neg r(s_0,x_5)\\
+	[\text{resolution 12, 13}] &\{x_3\mapsto s_0, x_5\mapsto h(x_4)\}& \neg q(s_0,x_4)\\
+	[\text{resolution 14, 11}] &\{x_1\mapsto s_0, x_4\mapsto g(x_2)\}&\neg p(s_0,x_2)\\
+	[\text{resolution 15, 10}] &\{x_0\mapsto s_0, x_2\mapsto f(s_0)\}& \bot
+\end{align}
 
-\subsection{Type system}
+Again replace negated conjecture by tautology
 
-Since we want to apply our method to first-order resolution proofs we will require a dependant type system. For this purpose we shall use the practical type system of agda as described in~\cite{agda}.
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{tautology}] &&\forall  x_0 \exists x_1 : r(x_0,x_1)\vee \neg\forall  x_0 \exists x_1 : r(x_0,x_1)\\
+	[\text{ennf transformation 2}] && \forall x_0, x_1 : (q(x_0,g(x_1)) \vee \neg p(x_0,x_1)) \\
+	[\text{ennf transformation 3}] &&\forall x_0, x_1 : (r(x_0,h(x_1)) \vee \neg q(x_0, x_1)) \\
+	[\text{ennf transformation 4}] && \forall  x_0 : \exists x_1 : r(x_0,x_1)\vee \exists x_0 \forall x_1 : \neg r(x_0, x_1)\\
+	[\text{choice axiom}] && \exists x_0 : \forall x_1 : \neg r(x_0, x_1) \Rightarrow \forall x_1 : \neg r(s_0,x_1)\\
+	[\text{choice axiom}] &&\forall x_0(\exists x_1: r(x_0, x_1)\Rightarrow r(x_0, s_1(x_0)))\\
+	[\text{skolemisation 7,8}] && \forall  x_0 \exists x_1 : r(x_0,x_1)\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{skolemisation 9,10}] && \forall  x_0 :  r(x_0, s_1(x_0))\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{cnf transformation 1}] && p(x_0,f(x_0))\\
+	[\text{cnf transformation 5}] && q(x_1,g(x_2)) \vee \neg p(x_1, x_2)\\
+	[\text{cnf transformation 6}] && r(x_3,h(x_4)) \vee \neg q(x_3, x_4)\\
+	[\text{cnf transformation 11}] && r(x_6, s_1(x_6))\vee\neg r(s_0, x_5)\\
+	[\text{resolution 14, 15}] &\{x_3\mapsto s_0, x_5\mapsto h(x_4)\}&r(x_6, s_1(x_6))\vee \neg q(s_0,x_4)\\
+	[\text{resolution 16, 13}] &\{x_1\mapsto s_0, x_4\mapsto g(x_2)\}&r(x_6, s_1(x_6))\vee\neg p(s_0,x_2)\\
+	[\text{resolution 17, 12}] &\{x_0\mapsto s_0, x_2\mapsto f(s_0)\}&r(x_6, s_1(x_6))
+\end{align}
 
-\subsection{Examples and Translations}
+We propagate the substitutions
 
-For now we will consider two minimal examples.
-The first doesn't use induction.
-There are two constant symbols $a, b$ and a binary proposition variable $p(-, -)$.
-The only axiom is $\forall x\:p(x,a)\vee p(x, b)$.
-We seek to prove $\forall x\exists y p(x, y)$.
-It is immediately clear that unless the axiom has computational content, our extracted program cannot yield a value.
-On the other hand we will see that we can still create a well-typed program.
-And if the axiom contains computational content it will indeed yield a result.
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{tautology}] &&\forall  x_0 \exists x_1 : r(x_0,x_1)\vee \neg\forall  x_0 \exists x_1 : r(x_0,x_1)\\
+	[\text{ennf transformation 2}] && \forall x_0, x_1 : (q(x_0,g(x_1)) \vee \neg p(x_0,x_1)) \\
+	[\text{ennf transformation 3}] &&\forall x_0, x_1 : (r(x_0,h(x_1)) \vee \neg q(x_0, x_1)) \\
+	[\text{ennf transformation 4}] && \forall  x_0 : \exists x_1 : r(x_0,x_1)\vee \exists x_0 \forall x_1 : \neg r(x_0, x_1)\\
+	[\text{choice axiom}] && \exists x_0 : \forall x_1 : \neg r(x_0, x_1) \Rightarrow \forall x_1 : \neg r(s_0,x_1)\\
+	[\text{choice axiom}] &&\forall x_0(\exists x_1: r(x_0, x_1)\Rightarrow r(x_0, s_1(x_0)))\\
+	[\text{skolemisation 7,8}] && \forall  x_0 \exists x_1 : r(x_0,x_1)\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{skolemisation 9,10}] && \forall  x_0 :  r(x_0, s_1(x_0))\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{cnf transformation 1}] && p(x_0,f(x_0))\\
+	[\text{cnf transformation 5}] && q(x_1,g(x_2)) \vee \neg p(x_1, x_2)\\
+	[\text{cnf transformation 6}] && r(s_0,h(x_4)) \vee \neg q(s_0, x_4)\\
+	[\text{cnf transformation 11}] && r(x_6, s_1(x_6))\vee\neg r(s_0,h(x_4))\\
+	[\text{resolution 14, 15}] &\{\}&r(x_6, s_1(x_6))\vee\neg q(s_0,x_4)\\
+	[\text{resolution 16, 13}] &\{x_1\mapsto s_0, x_4\mapsto g(x_2)\}&r(x_6, s_1(x_6))\vee\neg p(s_0,x_2)\\
+	[\text{resolution 17, 12}] &\{x_0\mapsto s_0, x_2\mapsto f(s_0)\}&r(x_6, s_1(x_6))
+\end{align}
 
-\begin{example}\hfill
+\setcounter{equation}{0}
 \begin{align}
-		& p(x_1,a)\vee p(x_1, b) & \text{Axiom}\\
-		& \llb \neg p(x_2, y_2) | x\simeq x_2\rrb & \text{negated Hypothesis}\\
-		& \llb p(x_1, b) | x\simeq x_1\rrb & res.\:(1) (2) [x_1/x_2, a/y_2]\\
-		& \llb \square | x\simeq x_1\rrb & res.\:(3) (2) [x_1/x_2, b/y_2]\\
-		& \square &\text{Instantiation }(4)
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{tautology}] &&\forall  x_0 \exists x_1 : r(x_0,x_1)\vee \neg\forall  x_0 \exists x_1 : r(x_0,x_1)\\
+	[\text{ennf transformation 2}] && \forall x_0, x_1 : (q(x_0,g(x_1)) \vee \neg p(x_0,x_1)) \\
+	[\text{ennf transformation 3}] &&\forall x_0, x_1 : (r(x_0,h(x_1)) \vee \neg q(x_0, x_1)) \\
+	[\text{ennf transformation 4}] && \forall  x_0 : \exists x_1 : r(x_0,x_1)\vee \exists x_0 \forall x_1 : \neg r(x_0, x_1)\\
+	[\text{choice axiom}] && \exists x_0 : \forall x_1 : \neg r(x_0, x_1) \Rightarrow \forall x_1 : \neg r(s_0,x_1)\\
+	[\text{choice axiom}] &&\forall x_0(\exists x_1: r(x_0, x_1)\Rightarrow r(x_0, s_1(x_0)))\\
+	[\text{skolemisation 7,8}] && \forall  x_0 \exists x_1 : r(x_0,x_1)\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{skolemisation 9,10}] && \forall  x_0 :  r(x_0, s_1(x_0))\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{cnf transformation 1}] && p(x_0,f(x_0))\\
+	[\text{cnf transformation 5}] && q(s_0,g(x_2)) \vee \neg p(s_0, x_2)\\
+	[\text{cnf transformation 6}] && r(s_0,h(g(x_2))) \vee \neg q(s_0, g(x_2))\\
+	[\text{cnf transformation 11}] && r(x_6, s_1(x_6))\vee\neg r(s_0,h(g(x_2)))\\
+	[\text{resolution 14, 15}] &\{\}&r(x_6, s_1(x_6))\vee\neg q(s_0,g(x_2))\\
+	[\text{resolution 16, 13}] &\{\}&r(x_6, s_1(x_6))\vee\neg p(s_0,x_2)\\
+	[\text{resolution 17, 12}] &\{x_0\mapsto s_0, x_2\mapsto f(s_0)\}&r(x_6, s_1(x_6))
 \end{align}
-\end{example}
 
-Now for our computational translation we introduce a special exception terms $e_\alpha(t_1,\dots,t_n)$ and $\overline e_\alpha(t_1,\dots,t_n)$ for every atom $\alpha[t_1,\dots,t_n]$ with $n$ term-variables, which intuitively indicate that $\alpha[t_1,\dots,t_n]$ is not false.
-They are usually introduced in the manner of the last derivation of definition 3.4 in~\cite{de_Groote_1995}.
-If one were to assign types we would have $e_\alpha(t_1.\dots,t_n) : (\alpha(t_1,\dots,t_n)\to\bot)\to\bot$ and $\overline e_\alpha(t_1.\dots,t_n) : ((\alpha(t_1,\dots,t_n)\to\bot)\to\bot)\to\bot$.
-However as we have the (classical) identification, we can simplify $\overline e_\alpha(t_1.\dots,t_n) : \alpha(t_1,\dots,t_n)\to\bot$.
-Having this in mind both $e_\alpha(t_1.\dots,t_n) \overline e_\alpha(t_1.\dots,t_n) : \bot$ and $\overline e_\alpha(t_1.\dots,t_n) e_\alpha(t_1.\dots,t_n) : \bot$ make ``type-sense" in the exception sense of~\cite{de_Groote_1995}.
-We will use this in our program.
-Let us now translate the above proof.
-Note that the terms for axioms are given.
-For clarity we will use braces for function application.
-For convenience we will write $\alpha_i$ for the $i$-the generated term.
 \setcounter{equation}{0}
-\begin{example}\hfill
 \begin{align}
-		& \alpha_1 & (p(x_1,a)\to\bot)\to(p(x_1, b)\to\bot)\to\bot\\
-		& \overline e_p(x_2, y_2) & p(x_2, y_2)\to\bot\\
-		& \lambda t. \mathbf{let}\:v : \neg p(x_1, a)&(p(x_1, b)\to\bot)\to\bot\\
-		& \phantom{\lambda t.}\mathbf{in}\:\alpha_1(v, t) &\nonumber\\
-		& \phantom{\lambda t.}\mathbf{handle}\:v(w)\Rightarrow \overline e_p(x_1, a)(w)\indent [= \alpha_2[x_1/x_2, a/y_2](w)]\hspace*{-4cm}\nonumber&\\
-		& \mathbf{let}\:v : \neg p(x_1, b)\:&\bot\\
-		& \mathbf{in}\:\alpha_3(v) &\nonumber\\
-		& \mathbf{handle}\:v(w)\Rightarrow \overline e_p(x_1, b)(w)\indent[= \alpha_2[x_1/x_2, b/y_2](w)]\hspace*{-4cm}\nonumber&\\
-		& \lambda x_1.\alpha_4 &(x : X) \to \bot
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{tautology}] &&\forall  x_0 \exists x_1 : r(x_0,x_1)\vee \neg\forall  x_0 \exists x_1 : r(x_0,x_1)\\
+	[\text{ennf transformation 2}] && \forall x_0, x_1 : (q(x_0,g(x_1)) \vee \neg p(x_0,x_1)) \\
+	[\text{ennf transformation 3}] &&\forall x_0, x_1 : (r(x_0,h(x_1)) \vee \neg q(x_0, x_1)) \\
+	[\text{ennf transformation 4}] && \forall  x_0 : \exists x_1 : r(x_0,x_1)\vee \exists x_0 \forall x_1 : \neg r(x_0, x_1)\\
+	[\text{choice axiom}] && \exists x_0 : \forall x_1 : \neg r(x_0, x_1) \Rightarrow \forall x_1 : \neg r(s_0,x_1)\\
+	[\text{choice axiom}] &&\forall x_0(\exists x_1: r(x_0, x_1)\Rightarrow r(x_0, s_1(x_0)))\\
+	[\text{skolemisation 7,8}] && \forall  x_0 \exists x_1 : r(x_0,x_1)\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{skolemisation 9,10}] && \forall  x_0 :  r(x_0, s_1(x_0))\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{cnf transformation 1}] && p(s_0, f(s_0))\\
+	[\text{cnf transformation 5}] && q(s_0,g(f(s_0))) \vee \neg p(s_0, f(s_0))\\
+	[\text{cnf transformation 6}] && r(s_0,h(g(f(s_0)))) \vee \neg q(s_0, g(f(s_0)))\\
+	[\text{cnf transformation 11}] && r(x_6, s_1(x_6))\vee\neg r(s_0,h(g(f(s_0))))\\
+	[\text{resolution 14, 15}] &\{\}&r(x_6, s_1(x_6))\vee\neg q(s_0,g(f(s_0)))\\
+	[\text{resolution 16, 13}] &\{\}&r(x_6, s_1(x_6))\vee\neg p(s_0,f(s_0))\\
+	[\text{resolution 17, 12}] &\{\}&r(x_6, s_1(x_6))
 \end{align}
-\end{example}
 
-Now to retrieve a value we just need to handle the final exception, i.e.
-$$\lambda x_1.\alpha_4\:\mathbf{handle}\: \overline{e}_p(x_2, y_2)(w)\Rightarrow y_2$$
-Of course this will not yield an actual value unless $\alpha_1$ yields.\\
-We will now consider a minimal inductive example.
+Then unification at 15 gives $s_1 = h\circ g\circ f$ and $s_0 = x_6$ and the final transformation gives:
 
-Consider the inductive type $nat$ with constructors $z : nat$ and $s: nat \to nat$ and some predicate $p(x, y)\subseteq nat^2$.
-We are going to prove $\forall x\exists y\: p(x,y)$ from $p(z, z)$ and $\forall n,m:nat(p(n, m)\Rightarrow p(s(n), s(s(m))))$.
-
-\begin{example}\hfill
+\setcounter{equation}{0}
 \begin{align}
-		& p(z, z) & \text{Axiom}\\
-		& \neg p(n_2, m_2)\vee p(s(n_2), s(s(m_2)))& \text{Axiom}\\
-		& \llb \neg p(n_3, m_3) | x\simeq n_3\rrb & \text{negated Hypothesis}\\
-		& \llb \square | x\simeq z\rrb & res.\:(1) (3) [z/n_3, z/m_3]\\
-		& \llb \neg p(n_2, m_2) | x\simeq s(n_2)\rrb &res. (2) (3) [s(n_2)/n_3], s(s(m_2))/m_3]\\
-		& \text{Activate $\alpha := p(a, b)[a, b]$}&\text{Trigger  (4)}\\
-		& \llb n_7\not\prec n_7'\vee p(n_7, m_7) | \mathbf{T}_\alpha^\prec(n_7', m_7)\rrb& \Gamma_{\mathbf{T}_\alpha^\prec}\\
-		& \llb n_7\not\prec n_7'| x\simeq s(n_7)\wedge \mathbf{T}_\alpha^\prec(n_7', m_7)\rrb & res. (5) (7) [n_7/n_2, m_7/ m_2] \\
-		& \llb n_7\prec s(n_7)|\top\rrb &\Gamma_\prec^s\\
-		& \llb \square | x\simeq s(n_7)\wedge\mathbf{T}_\alpha^\prec(s(n_7), m_7)\rrb&res. (8)(9)[s(n_7)/n_7']\\
-		& \llb \square | x\simeq n_{10}\rrb &\text{Induction }(4)(9)\\
-		& \square &\text{Instantiation }(10)
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{instantiation 1}] && p(x_6, f(x_6))\\
+	[\text{instantiation 2}] && p(x_6, f(x_6))\Rightarrow q(x_6,g(f(x_6))) \\
+	[\text{instantiation 3}] && q(x_6, g(f(x_6)))\Rightarrow r(x_6,h(g(f(x_6)))) \\
+	[\text{tautology}] && r(x_6, h(g(f(x_6))))\Rightarrow r(x_6,h(g(f(x_6))))\\
+	[\text{composition 6, 7}] &\{\}&q(x_6, g(f(x_6)))\Rightarrow r(x_6,h(g(f(x_6))))\\
+	[\text{composition 5, 8}] &\{\}&p(x_6,f(x_6))\Rightarrow r(x_6, h(g(f(x_6))))\\
+	[\text{composition 4, 9}] &\{\}&r(x_6, h(g(f(x_6))))
 \end{align}
-\end{example}
-\pagebreak
-\subsection{Intuitionistic Example}
-
-For simplicity we will use a single sort (0-level type).\\
-Function symbols: $f(-)$\\
-Predicate symbols: $P(-), Q(-, -)$\\
-Axioms: $\forall x P(x)$, $\forall x\exists y x=f(y)$, $\forall x(P(x)\supset Q(f(x), x))$\\
-To Show: $\forall x \exists y Q(x, y)$\\
-
-\noindent Below is a ND-style proof and translation based on the calculus and correspondence presented in~\cite{pfenning2000constructive}.\\
-
-\small{
-\AxiomC{}
-\RightLabel{\scriptsize Axiom}
-\UnaryInfC{$\vdash\forall x\exists y f(y)=x$}
-\RightLabel{\scriptsize $\forall E$}
-\UnaryInfC{$\vdash\exists y f(y)=x$}
-\AxiomC{}
-\RightLabel{\scriptsize Axiom}
-\UnaryInfC{$\vdash \forall x(P(x)\supset Q(f(x), x))$}
-\RightLabel{\scriptsize $\forall E$}
-\UnaryInfC{$\vdash P(y)\supset Q(f(y), y)$}
-\AxiomC{}
-\RightLabel{\scriptsize Axiom}
-\UnaryInfC{$\vdash\forall x P(x)$}
-\RightLabel{\scriptsize $\forall E$}
-\UnaryInfC{$\vdash P(y)$}
-\RightLabel{\scriptsize $\supset\hspace{-2pt}E$}
-\BinaryInfC{$\vdash Q(f(y), y)$}
-\AxiomC{}
-\RightLabel{\scriptsize u}
-\UnaryInfC{[$\vdash f(y) = x$]}
-\RightLabel{\scriptsize subst}
-\BinaryInfC{$\vdash Q(x, y)$}
-\RightLabel{\scriptsize $\supset\hspace{-2pt}I^u$}
-\UnaryInfC{$f(y)=x\vdash Q(x, y)$}
-\RightLabel{\scriptsize$\exists I$}
-\UnaryInfC{$f(y)=x\vdash \exists y Q(x, y)$}
-\RightLabel{\scriptsize $\exists E$}
-\BinaryInfC{$\vdash \exists Q(x, y)$}
-\RightLabel{\scriptsize $\forall I$}
-\UnaryInfC{$\vdash\forall x \exists yQ(x, y)$}
-
-\hspace*{-1cm}
-\DisplayProof
-}
-\vspace{3cm}
-\small{
-	\AxiomC{}
-	\RightLabel{\scriptsize Axiom}
-	\UnaryInfC{$\vdash \alpha : \Pi_x\Sigma_y f(y)=x$}
-	\RightLabel{\scriptsize $\forall E$}
-	\UnaryInfC{$\vdash \alpha(x): \Sigma_y f(y)=x$}
-	\AxiomC{}
-	\RightLabel{\scriptsize Axiom}
-	\UnaryInfC{$\vdash \beta : \Pi_x(P(x)\supset Q(f(x), x))$}
-	\RightLabel{\scriptsize $\forall E$}
-	\UnaryInfC{$\vdash \beta(y) : P(y)\supset Q(f(y), y)$}
-	\AxiomC{}
-	\RightLabel{\scriptsize Axiom}
-	\UnaryInfC{$\vdash\gamma : \Pi_x P(x)$}
-	\RightLabel{\scriptsize $\forall E$}
-	\UnaryInfC{$\vdash\gamma(y) : P(y)$}
-	\RightLabel{\scriptsize $\supset\hspace{-2pt}E$}
-	\BinaryInfC{$\vdash \beta(y)(\gamma(y)) : Q(f(y), y)$}
-	\AxiomC{}
-	\RightLabel{\scriptsize u}
-	\UnaryInfC{[$\vdash\delta : f(y) = x$]}
-	\RightLabel{\scriptsize subst}
-	\BinaryInfC{$\vdash \text{subst}(f(y),x,\Pi_tQ(t, y),\delta,\beta(y)(\gamma(y)) : Q(x, y)$}
-	\RightLabel{\scriptsize $\supset\hspace{-2pt}I^u$}
-	\UnaryInfC{$\delta : f(y)=x\vdash\text{subst}(f(y),x,\Pi_tQ(t, y),\delta,\beta(y)(\gamma(y)) : Q(x, y)$}
-	\RightLabel{\scriptsize$\exists I$}
-	\UnaryInfC{$\delta : f(y)=x\vdash \langle y, \text{subst}(f(y),x,\Pi_tQ(t, y),\delta,\beta(y)(\gamma(y))\rangle :\Sigma_y Q(x, y)$}
-	\RightLabel{\scriptsize $\exists E$}
-	\BinaryInfC{$\vdash \textbf{let }\langle y, \delta\rangle = \alpha(x)\textbf{ in }\langle y,\text{subst}(f(y),x,\Pi_tQ(t, y),\delta,\beta(y)(\gamma(y))\rangle: \Sigma_y Q(x, y)$}
-	\RightLabel{\scriptsize $\forall I$}
-	\UnaryInfC{$\lambda x.\textbf{let }\langle y, \delta\rangle = \alpha(x)\textbf{ in }\langle y,\text{subst}(f(y),x,\Pi_tQ(t, y),\delta,\beta(y)(\gamma(y))\rangle:\Pi_x \Sigma_y Q(x, y)$}
-
-	\hspace*{-1cm}
-	\DisplayProof
-}
-\hfill
-\\\\\hspace{2cm}We can reduce the last term to $\lambda x.\langle \alpha(x)_1, \text{subst}(f(\alpha(x)_1),x,\Pi_tQ(t, \alpha(x)_1),\alpha(x)_2,\beta(\alpha(x)_1)(\gamma(\alpha(x)_1))\rangle$.
-
-\noindent Finally replacing the proof terms by units we obtain $\lambda x.\langle \alpha(x)_1, ()\rangle : \Pi_x\Sigma_y \top$ from which we may further extract $\lambda x. \alpha(x)_1$.
 

### 2022-02-02 Implementation/formula.py
Code

@@ -1,0 +1,107 @@
+from dataclasses import dataclass
+from typing import Sequence
+from term import Term, Variable
+
+class Formula:  
+    def variables(self):
+        if type(self) == Atom:
+            vars = {}
+            for arg in self.arguments:
+                if vars:
+                    vars.update(arg.variables())
+                else:
+                    vars = arg.variables()
+            return vars
+        if type(self) in [Conjunction, Disjunction, Implication, Equivalence]:
+            vars = {}
+            for arg in self.arguments:
+                if vars:
+                    vars.update(arg.variables())
+                else:
+                    vars = arg.variables()
+            return vars
+        if type(self) == Negation:
+            return self.argument.variables()
+        if type(self) in [Forall, Exists]:
+            return self.formula.variables()
+        else:
+            return {}
+    
+    def substitute(self, d):
+        if type(self) == Atom:
+            return type(self)(name=self.name, arguments=[arg.substitute(d) for arg in self.arguments])
+        if type(self) in [Conjunction, Disjunction, Implication, Equivalence]:
+            return type(self)(arguments=[arg.substitute(d) for arg in self.arguments])
+        if type(self) == Negation:
+            return Negation(argument=self.argument.substitute(d))
+        if type(self) in [BTrue, BFalse]:
+            return self
+        if type(self) in [Forall, Exists]:
+            return type(self)(bound_variables=self.bound_variables, formula=self.formula.substitute({k : d[k] for k in d.keys() if k not in self.bound_variables}))
+
+    def __str__(self):
+        if type(self) == Atom:
+            return self.name + '(' + ','.join(list(map(str, self.arguments))) +')'
+        if type(self) == Conjunction:
+            return '(' + ') & ('.join(list(map(str, self.arguments))) + ')'
+        if type(self) == Disjunction:
+            return '(' + ') | ('.join(list(map(str, self.arguments))) + ')'
+        if type(self) == Implication:
+            return '(' + ') => ('.join(list(map(str, self.arguments))) + ')'
+        if type(self) == Equivalence:
+            return '(' + ') <=> ('.join(list(map(str, self.arguments))) + ')'
+        if type(self) == Negation:
+            return '~(' + str(self.argument) + ')'
+        if type(self) == BTrue:
+            return 'True'
[…]

### 2022-02-02 Implementation/parse.py

@@ -1,0 +1,118 @@
+from parsimonious.grammar import Grammar, NodeVisitor
+from formula import Disjunction, Conjunction, Implication, Equivalence, Forall, Exists, Negation, Atom, BTrue, BFalse
+from term import Term, Variable, FunctionApplication, Constant
+
+
+proof = Grammar(
+    """
+    proof = "\\n"? (rule "\\n"?)*
+    rule = id ". " formula " [" rulename id* "]"
+    id = ~"[0-9]+"","?
+    rulename = ~"[a-z\s]+"
+    formula = disjunction / conjunction / implication / equivalence / parensFormula / quantifiedFormula / negation / atom
+    disjunction = (quantifiedFormula / negation / atom / parensFormula) (" | " (quantifiedFormula / negation / atom / parensFormula))+
+    conjunction = (quantifiedFormula / negation / atom / parensFormula) (" & " (quantifiedFormula / negation / atom / parensFormula))+
+    implication = (quantifiedFormula / negation / atom / parensFormula) (" => " (quantifiedFormula / negation / atom / parensFormula))+
+    equivalence = (quantifiedFormula / negation / atom / parensFormula) (" <=> " (quantifiedFormula / negation / atom / parensFormula))+
+    parensFormula = "(" formula ")"
+    quantifiedFormula = forall / exists
+    forall = "! [" variable+ "] : " (parensFormula / quantifiedFormula / negation / atom)
+    exists = "? [" variable+ "] : " (parensFormula / quantifiedFormula / negation / atom)
+    negation = "~" (quantifiedFormula / atom / parensFormula)
+    atom = equality / inequality / predicate / bool
+    equality = term " = " term 
+    inequality = term " != " term
+    predicate = (name "(" term+ ")")
+    bool = "$true" / "$false"
+    term = (functionApplication / variable / constant)","?
+    functionApplication = (skolemName / name) "(" term+ ")"
+    name = ~"[a-z A-Z]([a-z A-Z 0-9]*)"
+    variable = ~"X([0-9]+)"","?
+    constant = skolemName / name
+    skolemName = ~"sK([0-9]+)"
+    """
+)
+
+class ProofVisitor(NodeVisitor):
+    def __init__(self) -> None:
+        self.result = {}
+
+    def visit_rule(self, node, visited_children):
+        id, _, formula , _, rulename, ids, _ = visited_children
+        self.result[id] = (rulename, [] if ids == '' else ids, formula)
+
[…]

### 2022-02-02 Implementation/proof.py

@@ -1,0 +1,217 @@
+from ast import arguments
+from dataclasses import dataclass
+from formula import BFalse, BTrue, Formula, Disjunction, Negation, Atom
+from typing import Dict, Sequence
+from term import Term, Variable
+
+class Proof:
+    def parse(d):
+        head_id = max(d.keys())
+        if d[head_id][0] == 'input':
+            return Input(d[head_id][2])
+        if d[head_id][0] == 'negated conjecture ':
+            return NegatedConjecture(d[head_id][2])
+        if d[head_id][0] == 'rectify ':
+            return Rectify(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
+        if d[head_id][0] == 'ennf transformation ':
+            return EnnfTransformation(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
+        if d[head_id][0] == 'nnf transformation ':
+            return NnfTransformation(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
+        if d[head_id][0] == 'cnf transformation ':
+            return CnfTransformation(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
+        if d[head_id][0] == 'flattening ':
+            return Flattening(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
+        if d[head_id][0] == 'choice axiom':
+            return ChoiceAxiom(d[head_id][2])
+        if d[head_id][0] == 'equality resolution ':
+            return EqualityResolution(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
+        if d[head_id][0] == 'skolemisation ':
+            return Skoemisation([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], d[head_id][2])
+        if d[head_id][0] == 'superposition ':
+            return Superposition([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
+        if d[head_id][0] == 'forward demodulation ':
+            return ForwardDemodulation([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
[…]

### 2022-02-02 Implementation/term.py

@@ -1,0 +1,72 @@
+from dataclasses import dataclass
+from typing import Sequence
+
+
+class Term:
+    def variables(self):
+        if type(self) == FunctionApplication:
+            vars = None
+            for arg in self.arguments:
+                if vars:
+                    vars.update(arg.variables())
+                else:
+                    vars = arg.variables()
+            return vars
+        if type(self) == Constant:
+            return {}
+        if type(self) == Variable:
+            return {self}
+
+    def substitute(self, d):
+        if not d:
+            return self
+        if type(self) == FunctionApplication:
+            return FunctionApplication(self.name, arguments=[arg.substitute(d) for arg in self.arguments])
+        if type(self) == Constant:
+            return Constant(self.name)
+        if type(self) == Variable:
+            if self in d.keys():
+                return(d[self])
+            else:
+                return(Variable(self.name))
+
+    def mgu(self, f):
+        if type(self) == Variable:
+            return ({self : f}, {})
+        if type(f) == Variable:
+            return ({}, {f : self})
+        if type(self) == FunctionApplication and type(f) == FunctionApplication:
+            s = [{}, {}]
+            for arg1, arg2 in zip(self.arguments, f.arguments):
+                arg1, arg2 = arg1.substitute(s[0]), arg2.substitute(s[1])
+                si = arg1.mgu(arg2)
+                for i in [0, 1]:
+                    if s[i]:
+                        s[i].update(si[i])
+                    else:
+                        s[i] = si[i]
+            return s
+        return ({}, {})
+
+    def __str__(self):
+        if type(self) in [Variable, Constant]:
+            return self.name
+        if type(self) == FunctionApplication:
+            return self.name + '(' + ','.join(list(map(str, self.arguments))) + ')'
+
+
+@dataclass
+class Variable(Term):
+    name : str
+
+    def __hash__(self):
+        return hash(self.name)
+
+@dataclass
+class Constant(Term):
+    name : str
+
+@dataclass
+class FunctionApplication(Term):
+    name : str
+    arguments : Sequence[Term]

### 2022-02-13 Notes/notes.tex
Code
implementation

@@
 \usepackage{bussproofs}
-\usepackage[left=3.00cm, right=3.00cm, top=3.00cm,bottom=3.0cm]{geometry}
+\usepackage[left=2.00cm, right=2.00cm, top=2.00cm,bottom=2.0cm]{geometry}
+\usepackage{pdflscape}
 
@@
 \pagebreak
+
+\pagebreak
+
+To explain our approach let us first recall how resolution works and why it is effective for classical logic.
+In a nutshell the principle of resolution works as follows: From a set of formulas $A, B, C\dots$ obtain a set of sequents $A'_1\dots  A'_n\vdash A_1\dots A_n$, $B'_1\dots B'_n\vdash B_1\dots B_n$, $C'_1\dots C'_n\vdash C_1\dots C_n$ in which all occurring formulas are somehow simple.
+Then iteratively apply the resolution rule (and some other rules, but this is the crucial one), which is an analogon to cut, i.e.
+\begin{center}
+	\AxiomC{$A, \Gamma\vdash \Delta$}
+	\AxiomC{$\Gamma'\vdash A', \Delta'$}
+	\BinaryInfC{$(\Gamma\cup \Gamma')\theta\vdash (\Delta\cup\Delta')\theta$}
+	\DisplayProof
+\end{center}
+where $\theta$ is the mgu of $A$ and $A'$, and by process of saturation (attempt to) obtain the empty Sequent.
+
+Now for classical logic we can simply interpret $A'_1,\dots  A'_n\vdash A_1,\dots, A_n$ as $\neg A'_1\vee\dots\vee\neg A'_n\vee A_1\vee\dots\vee A_n$.
+Furthermore we can normalize every formula such that each of the $A'_i, A_i$ is atomic.
+The resolution calculus is then immediately refutationally complete, i.e.
+if $A, B, C\dots $ are inconsistent there exists a successful resolution, giving a refutation of $A\wedge B\wedge C\dots$, which is classically equivalent to a proof of $B\wedge C\dots\rightarrow\neg A$.
+I.e.
+if we want to prove $B\wedge C\dots\rightarrow A$ we can simply preform resolution on $\neg A\wedge B\wedge C\dots$
+
+Now there are a number of hurdles in applying this strategy to intuitionistic logic, or to extract intuitionistic proof from classical resolution proofs.
+First and foremost is the non-existence of normal forms.
+In particular with the (intuitionally speaking) strict interpretation of $A'_1,\dots  A'_n\vdash A_1,\dots, A_n$ as $\neg A'_1\vee\dots\vee\neg A'_n\vee A_1\vee\dots\vee A_n$ we can hardly translate any formulas into such a form where the $A_i$ are atomic, or even literals.
+This can be remedied of course by interpreting $A'_1,\dots  A'_n\vdash A_1,\dots, A_n$ as $\neg(A'_1\wedge\dots\wedge A'_n\wedge(\neg A_1\vee\dots\vee \neg A_n))$, i.e.
+a double negation translation.
+But this brings us to the second problem: Even if our calculus is refutationally complete a refutation of $\neg A\wedge B\wedge C\dots$ does not give us a proof of $B\wedge C\dots\to A$ but rather $B\wedge C\dots\to \neg\neg A$.
+This problem is especially pronounced in the second interpretation of $\vdash$ where we are not even using $A$ but $\neg \neg A$.
+
+Now there are a few saving graces: First of all, if $A$ is decidable, i.e.
+$\neg\neg A\to A$, we immediately obtain a proof of $A$.
+Furthermore we may attempt Friedman Translation~\cite{Friedman_1978} of our proof and indeed~\cite{Berger_2002} gives us some quite liberal criteria when this is possible in principle, e.g.
+when $B, C\dots$ are theorems of PA and $A$ is a $\Pi_0^2$ formula.
+And indeed interpreting $A'_1,\dots  A'_n\vdash A_1,\dots, A_n$ as $A'_1\to\dots\to A'_n\to A_1\vee\dots\vee A_n$, we are able to obtain an intuitionistic proof if their conditions are met and the input formulas can be transformed to such normal forms.
 
@@
 \pagebreak
+
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{negated conjecture}] &&\neg\forall  x_0 : \exists x_1 : r(x_0,x_1)\\
+	[\text{ennf transformation 2}] && \forall x_0, x_1 : (q(x_0,g(x_1)) \vee \neg p(x_0,x_1)) \\
+	[\text{ennf transformation 3}] &&\forall x_0, x_1 : (r(x_0,h(x_1)) \vee \neg q(x_0, x_1)) \\
+	[\text{ennf transformation 4}] && \exists x_0 \forall x_1 : \neg r(x_0, x_1)\\
+	[\text{choice axiom}] && \exists x_0 \forall x_1 : \neg r(x_0, x_1) \Rightarrow \forall x_1 : \neg r(s_0,x_1)\\
+	[\text{skolemisation 7,8}] && \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{cnf transformation 1}] && p(x_0,f(x_0))\\
+	[\text{cnf transformation 5}] && q(x_1,g(x_2)) \vee \neg p(x_1, x_2)\\
+	[\text{cnf transformation 6}] && r(x_3,h(x_4)) \vee \neg q(x_3, x_4)\\
+	[\text{cnf transformation 9}] &&. \neg r(s_0,x_5)\\
+	[\text{resolution 12, 13}] &\{x_3\mapsto s_0, x_5\mapsto h(x_4)\}& \neg q(s_0,x_4)\\
+	[\text{resolution 14, 11}] &\{x_1\mapsto s_0, x_4\mapsto g(x_2)\}&\neg p(s_0,x_2)\\
+	[\text{resolution 15, 10}] &\{x_0\mapsto s_0, x_2\mapsto f(s_0)\}& \bot
+\end{align}
+
+Again replace negated conjecture by tautology
+
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{tautology}] &&\forall  x_0 \exists x_1 : r(x_0,x_1)\vee \neg\forall  x_0 \exists x_1 : r(x_0,x_1)\\
+	[\text{ennf transformation 2}] && \forall x_0, x_1 : (q(x_0,g(x_1)) \vee \neg p(x_0,x_1)) \\
+	[\text{ennf transformation 3}] &&\forall x_0, x_1 : (r(x_0,h(x_1)) \vee \neg q(x_0, x_1)) \\
+	[\text{ennf transformation 4}] && \forall  x_0 : \exists x_1 : r(x_0,x_1)\vee \exists x_0 \forall x_1 : \neg r(x_0, x_1)\\
+	[\text{choice axiom}] && \exists x_0 : \forall x_1 : \neg r(x_0, x_1) \Rightarrow \forall x_1 : \neg r(s_0,x_1)\\
+	[\text{choice axiom}] &&\forall x_0(\exists x_1: r(x_0, x_1)\Rightarrow r(x_0, s_1(x_0)))\\
+	[\text{skolemisation 7,8}] && \forall  x_0 \exists x_1 : r(x_0,x_1)\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{skolemisation 9,10}] && \forall  x_0 :  r(x_0, s_1(x_0))\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{cnf transformation 1}] && p(x_0,f(x_0))\\
+	[\text{cnf transformation 5}] && q(x_1,g(x_2)) \vee \neg p(x_1, x_2)\\
+	[\text{cnf transformation 6}] && r(x_3,h(x_4)) \vee \neg q(x_3, x_4)\\
+	[\text{cnf transformation 11}] && r(x_6, s_1(x_6))\vee\neg r(s_0, x_5)\\
+	[\text{resolution 14, 15}] &\{x_3\mapsto s_0, x_5\mapsto h(x_4)\}&r(x_6, s_1(x_6))\vee \neg q(s_0,x_4)\\
+	[\text{resolution 16, 13}] &\{x_1\mapsto s_0, x_4\mapsto g(x_2)\}&r(x_6, s_1(x_6))\vee\neg p(s_0,x_2)\\
+	[\text{resolution 17, 12}] &\{x_0\mapsto s_0, x_2\mapsto f(s_0)\}&r(x_6, s_1(x_6))
+\end{align}
+
+We propagate the substitutions
+
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{tautology}] &&\forall  x_0 \exists x_1 : r(x_0,x_1)\vee \neg\forall  x_0 \exists x_1 : r(x_0,x_1)\\
+	[\text{ennf transformation 2}] && \forall x_0, x_1 : (q(x_0,g(x_1)) \vee \neg p(x_0,x_1)) \\
+	[\text{ennf transformation 3}] &&\forall x_0, x_1 : (r(x_0,h(x_1)) \vee \neg q(x_0, x_1)) \\
+	[\text{ennf transformation 4}] && \forall  x_0 : \exists x_1 : r(x_0,x_1)\vee \exists x_0 \forall x_1 : \neg r(x_0, x_1)\\
+	[\text{choice axiom}] && \exists x_0 : \forall x_1 : \neg r(x_0, x_1) \Rightarrow \forall x_1 : \neg r(s_0,x_1)\\
+	[\text{choice axiom}] &&\forall x_0(\exists x_1: r(x_0, x_1)\Rightarrow r(x_0, s_1(x_0)))\\
+	[\text{skolemisation 7,8}] && \forall  x_0 \exists x_1 : r(x_0,x_1)\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{skolemisation 9,10}] && \forall  x_0 :  r(x_0, s_1(x_0))\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{cnf transformation 1}] && p(x_0,f(x_0))\\
+	[\text{cnf transformation 5}] && q(x_1,g(x_2)) \vee \neg p(x_1, x_2)\\
+	[\text{cnf transformation 6}] && r(s_0,h(x_4)) \vee \neg q(s_0, x_4)\\
+	[\text{cnf transformation 11}] && r(x_6, s_1(x_6))\vee\neg r(s_0,h(x_4))\\
+	[\text{resolution 14, 15}] &\{\}&r(x_6, s_1(x_6))\vee\neg q(s_0,x_4)\\
+	[\text{resolution 16, 13}] &\{x_1\mapsto s_0, x_4\mapsto g(x_2)\}&r(x_6, s_1(x_6))\vee\neg p(s_0,x_2)\\
+	[\text{resolution 17, 12}] &\{x_0\mapsto s_0, x_2\mapsto f(s_0)\}&r(x_6, s_1(x_6))
+\end{align}
+
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{tautology}] &&\forall  x_0 \exists x_1 : r(x_0,x_1)\vee \neg\forall  x_0 \exists x_1 : r(x_0,x_1)\\
+	[\text{ennf transformation 2}] && \forall x_0, x_1 : (q(x_0,g(x_1)) \vee \neg p(x_0,x_1)) \\
+	[\text{ennf transformation 3}] &&\forall x_0, x_1 : (r(x_0,h(x_1)) \vee \neg q(x_0, x_1)) \\
+	[\text{ennf transformation 4}] && \forall  x_0 : \exists x_1 : r(x_0,x_1)\vee \exists x_0 \forall x_1 : \neg r(x_0, x_1)\\
+	[\text{choice axiom}] && \exists x_0 : \forall x_1 : \neg r(x_0, x_1) \Rightarrow \forall x_1 : \neg r(s_0,x_1)\\
+	[\text{choice axiom}] &&\forall x_0(\exists x_1: r(x_0, x_1)\Rightarrow r(x_0, s_1(x_0)))\\
+	[\text{skolemisation 7,8}] && \forall  x_0 \exists x_1 : r(x_0,x_1)\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{skolemisation 9,10}] && \forall  x_0 :  r(x_0, s_1(x_0))\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{cnf transformation 1}] && p(x_0,f(x_0))\\
+	[\text{cnf transformation 5}] && q(s_0,g(x_2)) \vee \neg p(s_0, x_2)\\
+	[\text{cnf transformation 6}] && r(s_0,h(g(x_2))) \vee \neg q(s_0, g(x_2))\\
+	[\text{cnf transformation 11}] && r(x_6, s_1(x_6))\vee\neg r(s_0,h(g(x_2)))\\
+	[\text{resolution 14, 15}] &\{\}&r(x_6, s_1(x_6))\vee\neg q(s_0,g(x_2))\\
+	[\text{resolution 16, 13}] &\{\}&r(x_6, s_1(x_6))\vee\neg p(s_0,x_2)\\
+	[\text{resolution 17, 12}] &\{x_0\mapsto s_0, x_2\mapsto f(s_0)\}&r(x_6, s_1(x_6))
+\end{align}
+
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{tautology}] &&\forall  x_0 \exists x_1 : r(x_0,x_1)\vee \neg\forall  x_0 \exists x_1 : r(x_0,x_1)\\
+	[\text{ennf transformation 2}] && \forall x_0, x_1 : (q(x_0,g(x_1)) \vee \neg p(x_0,x_1)) \\
+	[\text{ennf transformation 3}] &&\forall x_0, x_1 : (r(x_0,h(x_1)) \vee \neg q(x_0, x_1)) \\
+	[\text{ennf transformation 4}] && \forall  x_0 : \exists x_1 : r(x_0,x_1)\vee \exists x_0 \forall x_1 : \neg r(x_0, x_1)\\
+	[\text{choice axiom}] && \exists x_0 : \forall x_1 : \neg r(x_0, x_1) \Rightarrow \forall x_1 : \neg r(s_0,x_1)\\
+	[\text{choice axiom}] &&\forall x_0(\exists x_1: r(x_0, x_1)\Rightarrow r(x_0, s_1(x_0)))\\
+	[\text{skolemisation 7,8}] && \forall  x_0 \exists x_1 : r(x_0,x_1)\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{skolemisation 9,10}] && \forall  x_0 :  r(x_0, s_1(x_0))\vee \forall x_1 : \neg r(s_0, x_1)\\
+	[\text{cnf transformation 1}] && p(s_0, f(s_0))\\
+	[\text{cnf transformation 5}] && q(s_0,g(f(s_0))) \vee \neg p(s_0, f(s_0))\\
+	[\text{cnf transformation 6}] && r(s_0,h(g(f(s_0)))) \vee \neg q(s_0, g(f(s_0)))\\
+	[\text{cnf transformation 11}] && r(x_6, s_1(x_6))\vee\neg r(s_0,h(g(f(s_0))))\\
+	[\text{resolution 14, 15}] &\{\}&r(x_6, s_1(x_6))\vee\neg q(s_0,g(f(s_0)))\\
+	[\text{resolution 16, 13}] &\{\}&r(x_6, s_1(x_6))\vee\neg p(s_0,f(s_0))\\
+	[\text{resolution 17, 12}] &\{\}&r(x_6, s_1(x_6))
+\end{align}
+
+Then unification at 15 gives $s_1 = h\circ g\circ f$ and $s_0 = x_6$ and the final transformation gives:
+
+\setcounter{equation}{0}
+\begin{align}
+	[\text{axiom}] && \forall x_0: p(x_0,f(x_0))\\
+	[\text{axiom}] && \forall x_0, x_1 : (p(x_0,x_1) \Rightarrow q(x_0,g(x_1)))\\
+	[\text{axiom}] && \forall x_0, x_1 : (q(x_0,x_1) \Rightarrow r(x_0,h(x_1)))\\
+	[\text{instantiation 1}] && p(x_6, f(x_6))\\
+	[\text{instantiation 2}] && p(x_6, f(x_6))\Rightarrow q(x_6,g(f(x_6))) \\
+	[\text{instantiation 3}] && q(x_6, g(f(x_6)))\Rightarrow r(x_6,h(g(f(x_6)))) \\
+	[\text{tautology}] && r(x_6, h(g(f(x_6))))\Rightarrow r(x_6,h(g(f(x_6))))\\
+	[\text{composition 6, 7}] &\{\}&q(x_6, g(f(x_6)))\Rightarrow r(x_6,h(g(f(x_6))))\\
+	[\text{composition 5, 8}] &\{\}&p(x_6,f(x_6))\Rightarrow r(x_6, h(g(f(x_6))))\\
+	[\text{composition 4, 9}] &\{\}&r(x_6, h(g(f(x_6))))
+\end{align}
+
+\pagebreak
 \bibliographystyle{acm}

### 2022-02-13 Implementation/formula.py

@@ -1,107 +1,97 @@
-from dataclasses import dataclass
-from typing import Sequence
-from term import Term, Variable
+from ast import arguments
+from term import Term, Signature
 
-class Formula:  
-    def variables(self):
-        if type(self) == Atom:
-            vars = {}
-            for arg in self.arguments:
-                if vars:
-                    vars.update(arg.variables())
+from functools import reduce
+from typing import Dict, Set
+
+
+class Signature:
+    def __init__(self, atoms: Dict[str, int], connectors: Dict[str, int], quantifiers: Set[str], termType: type):
+        self._d = {**atoms, **connectors}
+        self.termType = termType
+        self.atoms = set(atoms.keys()) | {c for c, a in connectors.items() if a == 0}
+        self.connectors = {c for c, a in connectors.items() if a > 0}
+        self.quantifiers = quantifiers
+        self.symbols = self.connectors | self.atoms | self.quantifiers
+        self.arity = lambda x: self._d[x]
+
+    def __repr__(self):
+        return f"""
+        Atoms: {", ".join([f"{s}/{self.termArity(s)}" for s in self.atoms])}
+        Connectors: {", ".join([f"{s}/{self.formulaArity(s)}" for s in self.connectors])}
+        Quantifiers: {", ".join([f"{s}" for s in self.quantifiers])}
+        """
+
+    def __eq__(self, other) -> bool:
+        return self._d == other._d
+
+def Formula(sig: Signature):
+    class _Formula:
+        signature = sig
+        
+        def __init__(self, name: str, arguments = [], boundVariables = []):
+            if name in _Formula.signature.quantifiers:
+                self.name = name
+                self.boundVariables = boundVariables
+                self.arguments = arguments
+            else:
+                if name in _Formula.signature.atoms | _Formula.signature.connectors:
+                    assert len(arguments) == _Formula.signature.arity(name)
+                    assert all(type(arg) == _Formula.signature.termType if name in _Formula.signature.atoms else type(arg) == type(self) for arg in arguments) 
+                    self.name = name
+                    self.arguments = arguments
                 else:
-                    vars = arg.variables()
-            return vars
-        if type(self) in [Conjunction, Disjunction, Implication, Equivalence]:
-            vars = {}
-            for arg in self.arguments:
-                if vars:
-                    vars.update(arg.variables())
-                else:
-                    vars = arg.variables()
[…]

### 2022-02-13 Implementation/proof.py

@@ -1,217 +1,31 @@
-from ast import arguments
-from dataclasses import dataclass
-from formula import BFalse, BTrue, Formula, Disjunction, Negation, Atom
-from typing import Dict, Sequence
-from term import Term, Variable
+from typing import Any, Dict
+from formula import Signature, Formula
 
-class Proof:
-    def parse(d):
-        head_id = max(d.keys())
-        if d[head_id][0] == 'input':
-            return Input(d[head_id][2])
-        if d[head_id][0] == 'negated conjecture ':
-            return NegatedConjecture(d[head_id][2])
-        if d[head_id][0] == 'rectify ':
-            return Rectify(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
-        if d[head_id][0] == 'ennf transformation ':
-            return EnnfTransformation(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
-        if d[head_id][0] == 'nnf transformation ':
-            return NnfTransformation(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
-        if d[head_id][0] == 'cnf transformation ':
-            return CnfTransformation(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
-        if d[head_id][0] == 'flattening ':
-            return Flattening(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
-        if d[head_id][0] == 'choice axiom':
-            return ChoiceAxiom(d[head_id][2])
-        if d[head_id][0] == 'equality resolution ':
-            return EqualityResolution(Proof.parse({k : d[k] for k in d.keys() if k <= d[head_id][1][0]}), d[head_id][2])
-        if d[head_id][0] == 'skolemisation ':
-            return Skoemisation([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], d[head_id][2])
-        if d[head_id][0] == 'superposition ':
-            return Superposition([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
-        if d[head_id][0] == 'forward demodulation ':
[…]

### 2022-02-13 Implementation/term.py

@@ -1,72 +1,83 @@
 from dataclasses import dataclass
-from typing import Sequence
+from functools import reduce
+from typing import Dict
 
 
-class Term:
-    def variables(self):
-        if type(self) == FunctionApplication:
-            vars = None
-            for arg in self.arguments:
-                if vars:
-                    vars.update(arg.variables())
-                else:
-                    vars = arg.variables()
-            return vars
-        if type(self) == Constant:
-            return {}
-        if type(self) == Variable:
-            return {self}
+class Signature:
+    def __init__(self, d: Dict[str, int]):
+        self._d = d
+        self.constants = {c for c, a in self._d.items() if a == 0}
+        self.functions = {c for c, a in self._d.items() if a > 0}
+        self.symbols = self.constants | self.functions
+        self.arity = lambda x: self._d[x]
 
-    def substitute(self, d):
-        if not d:
-            return self
-        if type(self) == FunctionApplication:
-            return FunctionApplication(self.name, arguments=[arg.substitute(d) for arg in self.arguments])
-        if type(self) == Constant:
-            return Constant(self.name)
-        if type(self) == Variable:
-            if self in d.keys():
-                return(d[self])
+    def __repr__(self):
+        return "[" + ", ".join([f"{s}/{self.arity(s)}" for s in self.symbols]) + "]"
+
+    def __eq__(self, other) -> bool:
+        return self._d == other._d
+
+
+def Term(sig: Signature):
+    class _Term():
+        signature = sig
+
+        def __init__(self, name: str, arguments = []):
+            assert all(type(self).signature == type(arg).signature for arg in arguments)
+            if name in _Term.signature.symbols:
+                assert len(arguments) == _Term.signature.arity(name)
+                self.name = name
+                self.arguments = arguments
             else:
-                return(Variable(self.name))
+                assert len(arguments) == 0
+                self.name = name
+                self.arguments = []
 
-    def mgu(self, f):
-        if type(self) == Variable:
-            return ({self : f}, {})
-        if type(f) == Variable:
-            return ({}, {f : self})
-        if type(self) == FunctionApplication and type(f) == FunctionApplication:
-            s = [{}, {}]
-            for arg1, arg2 in zip(self.arguments, f.arguments):
-                arg1, arg2 = arg1.substitute(s[0]), arg2.substitute(s[1])
-                si = arg1.mgu(arg2)
-                for i in [0, 1]:
-                    if s[i]:
-                        s[i].update(si[i])
-                    else:
[…]

### 2022-02-13 Implementation_old/formula.py

@@ -1,0 +1,97 @@
+from dataclasses import dataclass, field
+from typing import Sequence
+from term import Term, Variable
+
+class Formula:  
+    def variables(self):
+        vars = {}
+        for arg in self.arguments:
+            if vars:
+                vars.update(arg.variables())
+            else:
+                vars = arg.variables()
+        return vars
+    
+    def substitute(self, d):
+        if type(self) == Atom:
+            return type(self)(name=self.name, arguments=[arg.substitute(d) for arg in self.arguments])
+        if type(self) in [Conjunction, Disjunction, Implication, Equivalence, Negation]:
+            return type(self)(arguments=[arg.substitute(d) for arg in self.arguments])
+        if type(self) in [Forall, Exists]:
+            return type(self)(bound_variables=self.bound_variables, arguments=[self.arguments[0].substitute({k : d[k] for k in d.keys() if k not in self.bound_variables})])
+        else:
+            return self
+
+    def __str__(self):
+        if type(self) == Atom:
+            return self.name + '(' + ','.join(list(map(str, self.arguments))) +')'
+        if type(self) == Conjunction:
+            return '(' + ') & ('.join(list(map(str, self.arguments))) + ')'
+        if type(self) == Disjunction:
+            return '(' + ') | ('.join(list(map(str, self.arguments))) + ')'
+        if type(self) == Implication:
+            return '(' + ') => ('.join(list(map(str, self.arguments))) + ')'
+        if type(self) == Equivalence:
+            return '(' + ') <=> ('.join(list(map(str, self.arguments))) + ')'
+        if type(self) == Negation:
+            return '~(' + str(self.arguments[0]) + ')'
+        if type(self) == BTrue:
+            return 'True'
+        if type(self) == BFalse:
+            return 'False'
+        if type(self) == Forall:
+            return '! [' + ','.join(list(map(str, self.bound_variables))) + ']: ' + str(self.arguments[0])
+        if type(self) == Exists:
+            return '? [' + ','.join(list(map(str, self.bound_variables))) + ']: ' + str(self.arguments[0])
+
+    def normalize(self):
+        if type(self) == Forall:
+            return self.arguments[0].substitute({x : Variable("input" + x.name[1:], []) for x in self.bound_variables}).normalize()
+        if type(self) == Exists:
[…]

### 2022-02-13 Implementation_old/parse.py

@@ -1,6 +1,6 @@
 from parsimonious.grammar import Grammar, NodeVisitor
 from formula import Disjunction, Conjunction, Implication, Equivalence, Forall, Exists, Negation, Atom, BTrue, BFalse
-from term import Term, Variable, FunctionApplication, Constant
+from term import Variable, FunctionApplication, Constant
 
 
 proof = Grammar(
@@ -72,14 +72,14 @@ class ProofVisitor(NodeVisitor):
 
     def visit_forall(self, node, visited_children):
         _, vars, _, f = visited_children
-        return Forall(vars, f[0])
+        return Forall(vars, f)
 
     def visit_exists(self, node, visited_children):
         _, vars, _, f = visited_children
-        return Exists(vars, f[0])
+        return Exists(vars, f)
 
     def visit_negation(self, node, visited_children):
-        return Negation(visited_children[1][0])
+        return Negation(visited_children[1])
     
     def visit_atom(self, node, visited_children):
         return visited_children[0]

### 2022-02-13 Implementation_old/proof.py

@@ -1,0 +1,248 @@
+from dataclasses import dataclass
+from formula import BFalse, BTrue, Formula, Disjunction, Negation, Atom, Forall, Exists
+from typing import Dict, Sequence
+from term import Term, Variable
+
+class Proof:
+    def parse(d):
+        head_id = max(d.keys())
+        if d[head_id][0] == 'input':
+            return Input([], None, d[head_id][2])
+        if d[head_id][0] == 'negated conjecture ':
+            return NegatedConjecture([], None, d[head_id][2])
+        if d[head_id][0] == 'rectify ':
+            return Rectify([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
+        if d[head_id][0] == 'ennf transformation ':
+            return EnnfTransformation([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
+        if d[head_id][0] == 'nnf transformation ':
+            return NnfTransformation([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
+        if d[head_id][0] == 'cnf transformation ':
+            return CnfTransformation([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
+        if d[head_id][0] == 'flattening ':
+            return Flattening([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
+        if d[head_id][0] == 'choice axiom':
+            return ChoiceAxiom([], None, d[head_id][2])
+        if d[head_id][0] == 'equality resolution ':
+            return EqualityResolution([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
+        if d[head_id][0] == 'skolemisation ':
+            return Skoemisation([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
+        if d[head_id][0] == 'superposition ':
+            return Superposition([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
+        if d[head_id][0] == 'forward demodulation ':
[…]

### 2022-02-13 Implementation_old/term.py

@@ -1,0 +1,75 @@
+from ast import arguments
+from dataclasses import dataclass, field
+from typing import Sequence
+
+class Term:
+    def variables(self):
+        if type(self) == FunctionApplication:
+            vars = None
+            for arg in self.arguments:
+                if vars:
+                    vars.update(arg.variables())
+                else:
+                    vars = arg.variables()
+            return vars
+        if type(self) == Constant:
+            return {}
+        if type(self) == Variable:
+            return {self}
+
+    def substitute(self, d):
+        if not d:
+            return self
+        if type(self) == FunctionApplication:
+            return FunctionApplication(self.name, arguments=[arg.substitute(d) for arg in self.arguments])
+        if type(self) == Constant:
+            return Constant(self.name, [])
+        if type(self) == Variable:
+            if self in d.keys():
+                return(d[self])
+            else:
+                return(Variable(self.name, []))
+
+    def mgu(self, f):
+        if type(self) == Variable:
+            return ({self : f}, {})
+        if type(f) == Variable:
+            return ({}, {f : self})
+        if type(self) == FunctionApplication and type(f) == FunctionApplication:
+            s = [{}, {}]
+            for arg1, arg2 in zip(self.arguments, f.arguments):
+                arg1, arg2 = arg1.substitute(s[0]), arg2.substitute(s[1])
+                si = arg1.mgu(arg2)
+                for i in [0, 1]:
+                    if s[i]:
+                        s[i].update(si[i])
+                    else:
+                        s[i] = si[i]
+            return s
+        return ({}, {})
+
+    def __str__(self):
+        if type(self) in [Variable, Constant]:
+            return self.name
+        if type(self) == FunctionApplication:
+            return self.name + '(' + ','.join(list(map(str, self.arguments))) + ')'
+
+
+
+@dataclass
+class Variable(Term):
+    name : str
+    arguments : Sequence[Term] = field(default_factory = list)
+
+    def __hash__(self):
+        return hash(self.name)
+
+@dataclass
+class Constant(Term):
+    name : str
+    arguments : Sequence[Term] = field(default_factory = list)
+
+@dataclass
+class FunctionApplication(Term):
+    name : str
+    arguments : Sequence[Term]

### 2022-03-03 Notes/notes.tex
Added algorithm to pdf

@@
 \usepackage{pdflscape}
+\usepackage{xspace}
+\usepackage{todonotes}
+\usepackage{stackengine}
 
+\setlength\parindent{0pt}
+
+\theoremstyle{definition}
 \newtheorem{theorem}{Theorem}[section]
+\theoremstyle{definition}
 \newtheorem{corollary}[theorem]{Corollary}
+\theoremstyle{definition}
 \newtheorem{lemma}[theorem]{Lemma}
+\theoremstyle{definition}
 \newtheorem{proposition}[theorem]{Proposition}
+\theoremstyle{definition}
 \newtheorem{definition}[theorem]{Definition}
+\theoremstyle{definition}
 \newtheorem{example}[theorem]{Example}
@@
 \newcommand{\var}{\text{var}}
+\newcommand{\Gc}{\textbf{Gc}\xspace}
+\newcommand{\Gi}{\textbf{Gi}\xspace}
+\newcommand{\Res}{\textbf{Res}\xspace}
+\newcommand{\I}{\mathcal{I}}
 
@@
 
-To explain our approach let us first recall how resolution works and why it is effective for classical logic.
+To exdefinition our approach let us first recall how resolution works and why it is effective for classical logic.
 In a nutshell the principle of resolution works as follows: From a set of formulas $A, B, C\dots$ obtain a set of sequents $A'_1\dots  A'_n\vdash A_1\dots A_n$, $B'_1\dots B'_n\vdash B_1\dots B_n$, $C'_1\dots C'_n\vdash C_1\dots C_n$ in which all occurring formulas are somehow simple.
@@
 I.e.
-if we want to prove $B\wedge C\dots\rightarrow A$ we can simply preform resolution on $\neg A\wedge B\wedge C\dots$
+if we want to prove $B\wedge C\dots\rightarrow A$ we can simply perform resolution on $\neg A\wedge B\wedge C\dots$
 
@@
 \pagebreak
+\newgeometry{left=3.00cm, right=3.00cm, top=3.00cm,bottom=3.0cm}
+\section{Preliminaries}
+
+\subsection{First-Order Logic}
+Since the details of syntax are a delicate concern for proof theory we will collect some important definitions and basic facts in this chapter.
+We define terms inductively for the sequent calculus as usual.
+
+\begin{definition}
+	A \textit{Signature} consists of
+	\begin{enumerate}
+		\item a finite set $S_f$ of function symbols $f_1, \dots, f_n$
+		\item a finite set $S_R$ of relation symbols $R_1,\dots, R_m$
+		\item  a function ar$: S_f\cup S_R\to \mathbb{N}$ assigning to each symbol its arity (possibly 0).
+	\end{enumerate}
+\end{definition}
+
+\noindent We will sometimes denote a Signature by $\{f_1/\text{ar}(f_1),\dots,f_n/\text{ar}(f_n), R_1/\text{ar}(R_1).\dots,R_m/\text{ar}(R_m)\}$.
+Fix a countably infinite collection of free variables $a, b, c\dots$ and bound variables $x, y, z\dots$. 
+Note that the following definitions could be simplified by not distinguishing between bound and free variables.
+However this has some other drawbacks and in particular when defining a sequent calculus for first-order logic having disjoint sets of bound and free variables is convenient.
+
+\begin{definition}
+	A \textit{Semiterm} is one of the following
+	\begin{enumerate}
+		\item A bound or free variable
+		\item $f(t_1, \dots, t_n)$ given semiterms $t_1,\dots,t_n$ and an $n$-ary function symbol $f$.
+	\end{enumerate}
+	A \textit{Term} is defined as above but without allowing bound variables.
+We denote the set of bound variables as $\mathcal{V}_b$, of free variables as $\mathcal{V}_f$ and the set of terms as $\mathcal{T}$.
+\end{definition}
+
+\begin{definition}
+	A \textit{Semiformula} is one of the following
+	\begin{enumerate}
+		\item $R(t_1, \dots, t_n), t_1 = t_2, \bot, \top$ given semiterms $t_1,\dots,t_n$ and an $n$-ary function symbol $R$.
+		\item $\forall x A, \exists x A, A\wedge B, A\vee B, A\to B$ given Semiformulas $A, B$ and a bound variable $x$.
+	\end{enumerate}
+	A \textit{Formula} is semiformula in which every occurrence of a bound variable is guarded by a quantifier. A formula created by the first rule is called \textit{atomic}.
+\end{definition}
+
+\begin{definition}
+	A \textit{Substitution} is a function $\sigma: \mathcal{V}_\text{f}\to\mathcal{T}$. For a semiterm $t$ we define $t\sigma$ as follows:
+	\begin{itemize}
+		\item $x\sigma = x$ for $x\in \mathcal{V}_b$
+		\item $a\sigma = \sigma(a)$ for $a\in \mathcal{V}_f$
+		\item $f(t_1,\dots,t_n)\sigma = f(t_1\sigma,\dots,t_n\sigma)$ for an $n$-ary function symbol $f$ and terms $t_1,\dots,t_n$.
+	\end{itemize}
+	Analogously we define $F\sigma$ for formulas $F$.
+Given free variables $a_1,\dots,a_n$ and terms $t_1,\dots,t_n$ we define a special substitution $[t_1/a_1,\dots,t_n/a_n]$ as $$\sigma(a) = \begin{cases}
+		t_i & \text{if } a = a_i \text{ for } i \in \{1\dots n\}\\
+		a & \text{else}
+	\end{cases}$$this notion can be naturally extended where $a_i$ are arbitrary terms.
+\end{definition}
+
+\subsubsection{Semantics}
+\begin{definition}
+	A $\Sigma$-\textit{Structure} $\mathcal{A}$ is a pair $(A, I)$ where
+	\begin{itemize}
+		\item $A$ is a non-empty set, called the carrier of $\mathcal{A}$
+		\item $I$ is a function that assigns
+		\begin{itemize}
+			\item to each $n$-ary $f\in S_f$ a function $I(f): A^n\to A$
+			\item to each $n$-ary $R\in S_R$ a relation $I(R)\subseteq A^n$.
+		\end{itemize}
+		we will usually write $f^\mathcal{A}, R^\mathcal{A}$ instead of $I(f), I(R)$.
+	\end{itemize}
+\end{definition}
+
+\begin{definition}
+	An \textit{Interpretation} is a pair $(\mathcal{A}, \beta)$ where $\mathcal{A}$ is a $\Sigma$-structure and $\beta :\mathcal{V}_f\to A$ is called the \textit{Valuation}.
+As usual we can extend $\beta$ to quantifier-free semi-terms.
+We denote this extension by $\I_\beta$ and write $\I$ for $\I_\beta$ where it is not confusing.
+
+	For $a\in A, v\in \mathcal{V}_f$ and valuation $\beta$ we define $\beta[a/v]$ as
+	$$\beta[a/v](w) = \begin{cases}
+		v & \text{ if } v = w\\
+		w & \text{ else }
+	\end{cases}$$
+	and similarly $\I[a/v] = \langle \I, \beta[a/v]\rangle$.
+\end{definition}
+
+\begin{definition}We can now define the model relation $\I\models F$ for $\Sigma$-Formulas $F$:\\
+	\begin{tabular}{lcll}
+		$\I\models\top, \mathcal I\not\models\bot$ &&& always\\
+		$\I\models R(t_1,\dots t_n)$ & iff & $(\I(t_1),\dots,\I(t_n))\in R^\mathcal A$ & for $n$-ary $R\in S_R$ and terms $t_1\dots t_n$\\
+		$\I\models t_1 = t_2$ & iff & $\I(t_1)=\I(t_2)$ & for terms $t_1, t_2$\\
+		$\I\models \varphi\wedge\psi$ & iff & $\I\models \varphi$ and $\I\models \psi$ & \\
+		$\I\models \varphi\vee\psi$ & iff & $\I\models \varphi$ or $\I\models \psi$ & \\
+		$\I\models \varphi\to\psi$ & iff & $\I\models \varphi$ implies $\I\models \psi$ & \\
+		$\I\models \forall x\varphi$ & iff & $\I[a/v]\models\varphi[v/x]$ for all $a\in A$ & where $v$ is a new free variable\\
+		$\I\models \forall x\varphi$ & iff & $\I[a/v]\models\varphi[v/x]$ for some $a\in A$ & where $v$ is a new free variable\\
+		\end{tabular}
+	As usual for sets of formulas $\Phi$ we write $\I\models\Phi$ iff $\I\models\varphi$ for all $\varphi\in\Phi$.
+We say $\I$ \textit{models} $\Phi$ and $\Phi$ \textit{satisfies} $\I$.
+A set $\Phi$ is \textit{valid} iff it satisfies every interpretation.
+
+	Furthermore we write $\Phi\models\varphi$ if every interpretation that satisfies $\Phi$ also satisfies $\varphi$ and say that $\varphi$ \textit{semantically follows} from $\Phi$.
+\end{definition}
+
+\begin{example}We now introduce our running example: Consider $\Sigma = \{f/1, P/2, Q/2\}$.
+We show that from $\forall x \exists y P(x, y)\wedge\forall x \forall y(P(x, y)\to Q(x, f(y)))$ follows $\forall x\exists yQ(x, y)$.
+Consider the following semantic argument: Suppose $\mathcal{I}\models \forall x \exists y P(x, y)\wedge\forall x \forall y(P(x, y)\to Q(x, f(y)))$.
+We want to show $\I\models\forall x\exists y(Q)$.
+Let $a\in A$.
+Then $\mathcal{I}\models \exists yP(a, y)$, i.e.
+there exists $b\in A$ such that $\I\models P(a, b)$.
+Furthermore $\I\models P(a, b)\to Q(a, f(b))$ and thus $\I\models Q(a, f(b))$ and finally $\I\models \exists y Q(a, y)$.
+Since $a$ was chosen arbitrarily at the start $\I\models\forall x\exists yQ(x,y)$.
+Therefore $\forall x \exists y P(x, y)\wedge\forall x \forall y(P(x, y)\to Q(x, f(y)))\models\forall x\exists yQ(x, y)$.
+\end{example}
+
+The following is a central theorem resulting from the above definitions
+
+\begin{theorem}[Deduction Theorem]
+$\{\varphi_1,\dots,\varphi_n\}\models\varphi$ iff $\models \varphi_1\to\dots\to\varphi_n\to\varphi$.
+\end{theorem}
+
+\subsubsection{Proof Theory}
+\begin{definition}
+   Pairs of multisets of formulas $A = \{A_1,\dots, A_n\}, B = \{B_1,\dots, B_m\}$ form a \textit{Sequent} $A\Rightarrow B$.
+Usually we will write $A_1,\dots, A_n\Rightarrow B_1,\dots, B_m$, note however that $A$ and $B$ are \textbf{not} to be interpreted as sequences but as multisets.
+For a substitution $\theta$ define $$(A_1,\dots, A_n\Rightarrow B_1,\dots, B_m)\theta = A_1\theta,\dots, A_n\theta\Rightarrow B_1\theta,\dots, B_m\theta.$$
+\end{definition}
+
+We define a sequent calculus very similar to that in~\cite[p.77]{basicprooftheory}. We have following inferences\\
+
+\begin{tabular}{lll}
+	\AxiomC{\hphantom{x}}
+	\RightLabel{Ax ($P$ atomic)}
+	\UnaryInfC{$P,\Gamma\Rightarrow \Delta, P$}
+	\DisplayProof&
+	\AxiomC{\hphantom{x}}
+	\RightLabel{L$\bot$}
+	\UnaryInfC{$\bot,\Gamma\Rightarrow\Delta$}
+	\DisplayProof&
+\\&&\\
+	\AxiomC{$A, B,\Gamma\Rightarrow\Delta$}
+	\RightLabel{L$\wedge$}
+	\UnaryInfC{$A\wedge B, \Gamma\Rightarrow \Delta$}
+	\DisplayProof&
+	\AxiomC{$\Gamma\Rightarrow\Delta, A$}
+	\AxiomC{$\Gamma\Rightarrow\Delta, B$}
+	\RightLabel{R$\wedge$}
+	\BinaryInfC{$\Gamma\Rightarrow \Delta, A\wedge B$}
+	\DisplayProof&
+\\&&\\
+	\AxiomC{$A, \Gamma\Rightarrow\Delta$}
+	\AxiomC{$B, \Gamma\Rightarrow\Delta$}
+	\RightLabel{L$\vee$}
+	\BinaryInfC{$A\vee B, \Gamma\Rightarrow \Delta$}
+	\DisplayProof&
+	\AxiomC{$\Gamma\Rightarrow\Delta, A, B$}
+	\RightLabel{R$\vee$}
+	\UnaryInfC{$\Gamma\Rightarrow \Delta, A\vee B$}
+	\DisplayProof&
+\\&&\\
+	\AxiomC{$A\to B, \Gamma\Rightarrow\Delta, A$}
+	\AxiomC{$B, \Gamma\Rightarrow\Delta$}
+	\RightLabel{L$\to$}
+	\BinaryInfC{$A\to B, \Gamma\Rightarrow \Delta$}
+	\DisplayProof&
+	\AxiomC{$A,\Gamma\Rightarrow\Delta, B$}
+	\RightLabel{R$\to$}
+	\UnaryInfC{$\Gamma\Rightarrow \Delta, A\to B$}
+	\DisplayProof&
+\\&&\\
+	\AxiomC{$\forall xA, A[t/x], \Gamma\Rightarrow\Delta$}
+	\RightLabel{L$\forall$}
+	\UnaryInfC{$\forall xA, \Gamma\Rightarrow \Delta$}
+	\DisplayProof&
+	\AxiomC{$\Gamma\Rightarrow\Delta, A[a/x]$}
+	\RightLabel{R$\forall$}
+	\UnaryInfC{$\Gamma\Rightarrow \Delta, \forall xA$}
+	\DisplayProof&
+\\&&\\
+	\AxiomC{$A[a/x], \Gamma\Rightarrow\Delta$}
+	\RightLabel{L$\exists$}
+	\UnaryInfC{$\exists xA, \Gamma\Rightarrow \Delta$}
+	\DisplayProof&
+	\AxiomC{$\Gamma\Rightarrow\Delta, A[t/x], \exists xA$}
+	\RightLabel{R$\exists$}
+	\UnaryInfC{$\Gamma\Rightarrow \Delta, \exists xA$}
+	\DisplayProof&
+\\&&\\
+	\AxiomC{$t = t, \Gamma\Rightarrow \Delta$}
+	\RightLabel{Ref}
+	\UnaryInfC{$\Gamma\Rightarrow \Delta$}
+	\DisplayProof&
+	\AxiomC{$s = t, P[s/x], P[t/x], \Gamma\Rightarrow \Delta$}
+	\RightLabel{Ref}
+	\UnaryInfC{$s = t, P[s/x],\Gamma\Rightarrow \Delta$}
+	\DisplayProof&
+\\&&\\
+\end{tabular}
+
+where in R$\forall$ and L$\exists$ $a$ is a free variable not occurring in the conclusion.
+
+The above rules make up the calculus called \textbf{Gc}. It corresponds to \textbf{G3c}$^=$ from~\cite{basicprooftheory}.
+
+\begin{definition}
+	A \textit{Derivation} is a rooted finite labelled tree, in which each leaf is labelled with one of the two axiom rules and each other label is derived from the labels of the successor nodes in accordance to the rules above.
+The label of the root is called \textit{Head}.
+A formula that is the head of a derivation is said to be \textit{derivable}.
+\end{definition}
+
+\begin{example} We return to our running example and give a Derivation:\\
+
+	\hspace{-1cm}
+	\AxiomC{}
+	\RightLabel{Ax}
+	\UnaryInfC{\scriptsize{\stackanchor{$P(a, b)\to Q(a, f(b)), \forall x\forall y(P(x, y)\to Q(x, f(y)))$}{$\forall x\exists yP(x, y), P(a, b), \forall y(P(a, y)\to Q(a, f(y)))$}$\Rightarrow$\stackanchor{$P(a, b), Q(a, f(b))$}{$\exists yQ(a, f(y))$}}}
+	\RightLabel{Ax}
+	\AxiomC{}
+	\UnaryInfC{\scriptsize{\stackanchor{$Q(a, f(b)), P(a, b), \forall y(P(a, y)\to Q(a, f(y)))$}{$\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))$}$\Rightarrow$\stackanchor{$Q(a, f(b))$}{$\exists yQ(a, f(y))$}}}
+	\RightLabel{L$\to$}
+	\BinaryInfC{\stackanchor{$\forall y(P(a, y)\to Q(a, f(y))), P(a, b)\to Q(a, f(b))$}{$ \forall x\exists yP(x, y), P(a, b), \forall x\forall y(P(x, y)\to Q(x, f(y))) $}$\Rightarrow Q(a, f(b)), \exists yQ(a, f(y))$}
+	\RightLabel{L$\forall$}
+	\UnaryInfC{\stackanchor{$\forall x\forall y(P(x, y)\to Q(x, f(y))), P(a, b)$}{$\forall y(P(a, y)\to Q(a, f(y))), \forall x\exists yP(x, y)$}$\Rightarrow Q(a, f(b)), \exists yQ(a, f(y))$}
+	\RightLabel{L$\forall$}
+	\UnaryInfC{$P(a, b),\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow Q(a, f(b)), \exists yQ(a, f(y))$}
+	\RightLabel{R$\exists$}
+	\UnaryInfC{$P(a, b),\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow \exists yQ(a, f(y))$}
+	\RightLabel{L$\exists$}
+	\UnaryInfC{$\exists yP(a, y),\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow \exists yQ(a, f(y))$}
+	\RightLabel{L$\forall$}
+	\UnaryInfC{$\forall x\exists yP(x, y),\forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow \exists yQ(a, f(y))$}
+	\DisplayProof
+\end{example}
+
+The following is one of the fundamental results of first-order Logic:
+
+\begin{theorem}[Completeness theorem]
+	A sequent $\Phi\Rightarrow \varphi$ is derivable iff $\Phi\models \varphi$.
+\end{theorem}
+
+Since we are ultimately interested in program extraction we shall also define a calculus for intuitionistic logic.
+Usually this is done by restricting all sequents to a single Formula in the succedent and adjusting R$\vee$.
+Instead we remain with a multi-succedent system by adjusting R$\to$ and R$\forall$:\\
+
+\begin{center}
+	\AxiomC{$A, \Gamma\Rightarrow B$}
+	\RightLabel{R$\to$}
+	\UnaryInfC{$\Gamma\Rightarrow\Delta, A\to B$}
+	\DisplayProof\hspace{2cm}
+	\AxiomC{$\Gamma\Rightarrow A[a/x]$}
+	\RightLabel{R$\forall$}
+	\UnaryInfC{$\Gamma\Rightarrow\Delta, \forall xA$}
+	\DisplayProof
+\end{center}
+
+We call this modified calculus \Gi, it corresponds to \textbf{m-G3i}$^=$ from~\cite{basicprooftheory}.
+Note that any proof in \Gc is also a proof in \Gi and many proofs in \Gc are also valid in \Gi, such as the above example.
+
+\section{Resolution}
+
+Resolution systems usually consist of two parts: First formulas are transformed into a normal form, usually a disjunction of ground atoms, and then the obtained clause set is extended by process of saturation until the empty clause is found.
+The first step is usually part of the preprocessing and treated in passing.
+However from the perspective of constructive content normal from translation is crucial, when taken naively many of the steps are intuitionistically invalid, e.g.
+going from $A\to B$ to $\neg A\vee B$.
+Therefore we must give a comprehensive account of the normal form translation.
+
+\subsection{Normal Form Translation}
+
+\begin{tabular}{lll}
+	\AxiomC{$A\vee B, \Delta\Rightarrow \Gamma$}
+	\RightLabel{L$\vee$1}
+	\UnaryInfC{$A, \Delta\Rightarrow \Gamma$}
+	\DisplayProof&
+	\AxiomC{$A\vee B, \Delta\Rightarrow \Gamma$}
+	\RightLabel{L$\vee$2}
+	\UnaryInfC{$B, \Delta\Rightarrow \Gamma$}
+	\DisplayProof&
+	\AxiomC{$\Delta\Rightarrow \Gamma, A\vee B$}
+	\RightLabel{R$\vee$}
+	\UnaryInfC{$\Delta\Rightarrow \Gamma, A, B$}
+	\DisplayProof
+	\\&\\
+	\AxiomC{$A\wedge B, \Delta\Rightarrow \Gamma$}
+	\RightLabel{L$\wedge$}
+	\UnaryInfC{$A, B, \Delta\Rightarrow \Gamma$}
+	\DisplayProof&
+	\AxiomC{$\Delta\Rightarrow \Gamma, A\wedge B$}
+	\RightLabel{R$\vee$1}
+	\UnaryInfC{$\Delta\Rightarrow\Gamma, A$}
+	\DisplayProof&
+	\AxiomC{$\Delta\Rightarrow \Gamma, A\wedge B$}
+	\RightLabel{R$\vee$1}
+	\UnaryInfC{$\Delta\Rightarrow\Gamma, B$}
+	\DisplayProof
+	\\&\\
+	\AxiomC{$A\to B, \Delta\Rightarrow \Gamma$}
+	\RightLabel{L$\to$1}
+	\UnaryInfC{$B, \Delta\Rightarrow \Gamma$}
+	\DisplayProof&
+	\AxiomC{$A\to B,\Delta\Rightarrow \Gamma$}
+	\RightLabel{L$\to$2}
+	\UnaryInfC{$\Delta\Rightarrow\Gamma, A$}
+	\DisplayProof&
+	\AxiomC{$\Delta\Rightarrow \Gamma, A\to B$}
+	\RightLabel{R$\to$}
+	\UnaryInfC{$A, \Delta\Rightarrow\Gamma, B$}
+	\DisplayProof
+	\\&\\
+	\AxiomC{$\top, \Delta\Rightarrow \Gamma$}
+	\RightLabel{L$\top$}
+	\UnaryInfC{$\Delta\Rightarrow \Gamma$}
+	\DisplayProof&
+	\AxiomC{$\Delta\Rightarrow \Gamma, \bot$}
+	\RightLabel{R$\bot$}
+	\UnaryInfC{$\Delta\Rightarrow\Gamma$}
+	\DisplayProof&
+	\\&\\
+	\AxiomC{$\forall xA, \Delta\Rightarrow \Gamma$}
+	\RightLabel{L$\forall$}
+	\UnaryInfC{$A[s(a_1,\dots,a_n)/x], \Delta\Rightarrow \Gamma$}
+	\DisplayProof&
+	\AxiomC{$\Delta\Rightarrow \Gamma, \forall xA$}
+	\RightLabel{R$\forall$}
+	\UnaryInfC{$\Delta\Rightarrow\Gamma, A[a/x]$}
+	\DisplayProof&
+	\\&\\
+	\AxiomC{$\exists xA, \Delta\Rightarrow \Gamma$}
+	\RightLabel{L$\exists$}
+	\UnaryInfC{$A[a/x], \Delta\Rightarrow \Gamma$}
+	\DisplayProof&
+	\AxiomC{$\Delta\Rightarrow \Gamma, \exists xA$}
+	\RightLabel{R$\exists$}
+	\UnaryInfC{$\Delta\Rightarrow\Gamma, A[s(a_1,\dots,a_n)/x]$}
+	\DisplayProof&
+	\\&\\
+\end{tabular}
+
+where $a_1,\dots,a_n$ are the free variables occurring in $A$ and $s$ is a new function symbol.
+
+\subsection{Unification}
+
+\begin{definition}
+	Let $A, B$ be formulas with disjoint sets of free variables occurring in them.
+A \textit{Unifier} of $A$ and $B$ is a substitution $\theta$ such that $A\theta = B\theta$, if a such a unifier exists $A, B$ are said to be \textit{unifiable}.
+A \textit{mgu} of most general unifier of $A$ and $B$ is a unifier $\theta$ such that for every other unifier $\theta'$ there exists a substitution $\tau$ such that $\tau\circ \theta = \theta'$.
+\end{definition}
+
+\begin{lemma}
+	If $A, B$ are unifiable there exists an mgu.
+\end{lemma}
+
+\subsection{Ground Resolution}
+The following rules are only to be applied to ground sequents.
+
+\begin{center}
+	\AxiomC{$\Delta\to \Gamma, A$}
+	\AxiomC{$B, \Delta'\to \Gamma'$}
+	\RightLabel{Res}
+	\BinaryInfC{$(\Delta, \Delta'\to \Gamma\cup\Gamma')\theta$}
+	\DisplayProof
+\end{center}
+where $\theta$ is an mgu of $A$ and $B$.
+
+\begin{center}
+	\AxiomC{$A, B,\Delta\to \Gamma$}
+	\RightLabel{LFac}
+	\UnaryInfC{$(A, \Delta\to\Gamma)\theta$}
+	\DisplayProof\hspace*{1cm}
+	\AxiomC{$\Delta\to \Gamma, A, B$}
+	\RightLabel{RFac}
+	\UnaryInfC{$(\Delta\to\Gamma, A)\theta$}
+	\DisplayProof
+\end{center}
+where $\theta$ is an mgu of $A$ and $B$.
+
+\begin{center}
+	\AxiomC{$\Delta\to\Gamma, l=r$}
+	\AxiomC{$L[s], \Delta'\to \Gamma'$}
+	\RightLabel{LSup}
+	\BinaryInfC{$(L[r],\Delta, \Delta'\to \Gamma,\Gamma')\theta$}
+	\DisplayProof\hspace*{1cm}
+	\AxiomC{$\Delta\to\Gamma, l=r$}
+	\AxiomC{$\Delta'\to \Gamma',L[s]$}
+	\RightLabel{RSup}
+	\BinaryInfC{$(\Delta, \Delta'\to \Gamma,\Gamma', L[r])\theta$}
+	\DisplayProof
+\end{center}
+where $\theta$ is an mgu of $l$ and $s$, $s$ is not a variable.
+
+\begin{center}
+	\AxiomC{$s= t,\Delta\to\Gamma$}
+	\RightLabel{eqRes}
+	\UnaryInfC{$(\Delta\to\Gamma)\theta$}
+	\DisplayProof
+\end{center}
+where $\theta$ is an mgu of $s$ and $t$.
+
+\begin{center}
+	\AxiomC{$\Gamma\to\Delta, s=t, s'=t'$}
+	\RightLabel{eqFac}
+	\UnaryInfC{$(t=t', \Delta\to\Gamma, s=t)\theta$}
+	\DisplayProof
+\end{center}
+where $\theta$ is an mgu of $s$ and $s'$.
+
+\begin{definition}
+	A \textit{Resolution Refutation} of a set of Sequents $\{A_1,\dots, A_n\}$ is a finite rooted tree whose roots are labelled with some $A_i$, the label of each other node follows from the labels of its children nodes by one of the normalization rules or the rules above, and the root is labelled with the empty sequent.
+A \textit{Resolution Proof} of $A_1,\dots,A_n\Rightarrow B_1,\dots,B_n$ is a resolution refutation of $\{\Rightarrow A_i\}\cup\{B_i\Rightarrow\}$.
+\end{definition}
+
+\begin{example}
+	Let's prove our running example.\\\\
+	\AxiomC{$\Rightarrow \forall x\exists y P(x, y)$}
+	\RightLabel{R$\forall$}
+	\UnaryInfC{$\Rightarrow\exists yP(a, y)$}
+	\RightLabel{R$\exists$}
+	\UnaryInfC{$\Rightarrow P(a, s_1(a))$}
+	\AxiomC{$\Rightarrow\forall x\forall y(P(x, y) \to Q(x, f(y)))$}
+	\RightLabel{R$\forall$}
+	\UnaryInfC{$\Rightarrow\forall y(P(b, y)\to Q(b, f(y)))$}
+	\RightLabel{R$\forall$}
+	\UnaryInfC{$\Rightarrow P(b, c)\to Q(b, f(c))$}
+	\RightLabel{R$\to$}
+	\UnaryInfC{$P(b, c)\Rightarrow Q(b, f(c))$}
+	\RightLabel{Res}
+	\BinaryInfC{$\Rightarrow Q(a, f(s_1(a)))$}
+	\AxiomC{$\forall x\exists y Q(x, y)\Rightarrow$}
+	\RightLabel{L$\forall$}
+	\UnaryInfC{$\exists y Q(s_2, y)\Rightarrow$}
+	\RightLabel{L$\exists$}
+	\UnaryInfC{$Q(s_2, b)\Rightarrow$}
+	\RightLabel{Res}
+	\BinaryInfC{$\vphantom{h}\Rightarrow$}
+	\DisplayProof
+\end{example}
+
+\begin{theorem}
+	\Res is refutationally complete, i.e. $\Phi\Rightarrow\varphi$ has a resolution proof iff $\Phi\models \varphi$.
+\end{theorem}
+
+\section{From \Res to \Gc}
+
+We now detail a transformation algorithm that takes Resolution proofs and transforms the into sequent calculus proofs.
+Note that at most of the intermediate steps the finite labeled tree will constitute a valid proof in neither \Res nor \Gc.
+
+At the bottom of this algorithm lies the observation that the normalization rules are (almost) dual to the rules of \Gc, and indeed we will construct the \Gc proof "upside down".
+A notable difference of \Res and \Gc is that in \Res it is possible that already assigned free variables are substituted with other terms.
+This does not happen in \Gc, so let us first eliminate all substitutions.
+To that end let us define substitution for (sub)trees via
+
+$$\left\lbrack
+	\AxiomC{$\mathcal{T}_1$}
+	\AxiomC{$\dots$}
+	\AxiomC{$\mathcal T_n$}
+	\TrinaryInfC{$S$}
+	\DisplayProof
+\right\rbrack\theta = 
+	\AxiomC{$\mathcal{T}_1\theta$}
+	\AxiomC{$\dots$}
+	\AxiomC{$\mathcal T_n\theta$}
+	\TrinaryInfC{$S\theta$}
+	\DisplayProof$$
+
+We can then eliminate all substitutions from a \Res tree by defining a function $f$ on finite labeled rooted trees\\
+
+\begin{tabular}{llr}
+$\mathfrak f_1\left(\AxiomC{$\mathcal T_1$}
+\AxiomC{$\mathcal T_2$}
+\RightLabel{Res}
+\BinaryInfC{$(\Delta, \Delta'\to \Gamma\cup\Gamma')\theta$}
+\DisplayProof\right)$&$=$&\AxiomC{$\mathfrak f_1(\mathcal T_1)\theta$}
+\AxiomC{$\mathfrak f_1(\mathcal T_2)\theta$}
+\RightLabel{Cut}
+\BinaryInfC{$(\Delta, \Delta'\to \Gamma\cup\Gamma')\theta$}
+\DisplayProof\\&&\\
+$\mathfrak f_1\left(\AxiomC{$\mathcal T$}
+\RightLabel{LFac}
+\UnaryInfC{$(A, \Delta\to\Gamma)\theta$}
+\DisplayProof\right)$&$=$&\AxiomC{$\mathfrak f_1(\mathcal T)\theta$}
+\RightLabel{LWeak}
+\UnaryInfC{$(A, \Delta\to\Gamma)\theta$}
+\DisplayProof\\&&\\
+$\mathfrak f_1\left(\AxiomC{$\mathcal T$}
+\RightLabel{RFac}
+\UnaryInfC{$(\Delta\to\Gamma, A)\theta$}
+\DisplayProof\right)$&$=$&\AxiomC{$\mathfrak f_1(\mathcal T)\theta$}
+\RightLabel{RWeak}
+\UnaryInfC{$(\Delta\to\Gamma, A)\theta$}
+\DisplayProof\\&&\\
+$\mathfrak f_1\left(\AxiomC{$\mathcal T$}
+\RightLabel{eqRes}
+\UnaryInfC{$(\Delta\to\Gamma)\theta$}
+\DisplayProof\right)$&$=$&$\mathfrak f_1(\mathcal T')\theta\hspace*{2.1cm}$
+\end{tabular}
+
+where in $\mathcal{T}'$ we remove $s=t$ in the antecedent of every sequent.
+
+\begin{lemma}
+	All applications of Cut, LWeak and RWeak in a such transformed Resolution tree are valid.
+\end{lemma}
+
+\begin{example} Let's perform this transformation on our example
+	\\\\
+	\AxiomC{$\Rightarrow \forall x\exists y P(x, y)$}
+	\RightLabel{R$\forall$}
+	\UnaryInfC{$\Rightarrow\exists yP(s_2, y)$}
+	\RightLabel{R$\exists$}
+	\UnaryInfC{$\Rightarrow P(s_2, s_1(s_2))$}
+	\AxiomC{$\Rightarrow\forall x\forall y(P(x, y) \to Q(x, f(y)))$}
+	\RightLabel{R$\forall$}
+	\UnaryInfC{$\Rightarrow\forall y(P(s_2, y)\to Q(s_2, f(y)))$}
+	\RightLabel{R$\forall$}
+	\UnaryInfC{$\Rightarrow P(s_2, s_1(s_2))\to Q(s_2, f(s_1(s_2)))$}
+	\RightLabel{R$\to$}
+	\UnaryInfC{$P(s_2, s_1(s_2))\Rightarrow Q(s_2, f(s_1(s_2)))$}
+	\RightLabel{Cut}
+	\BinaryInfC{$\Rightarrow Q(s_2, f(s_1(s_2)))$}
+	\AxiomC{$\forall x\exists y Q(x, y)\Rightarrow$}
+	\RightLabel{L$\forall$}
+	\UnaryInfC{$\exists y Q(s_2, y)\Rightarrow$}
+	\RightLabel{L$\exists$}
+	\UnaryInfC{$Q(s_2, f(s_1(s_2)))\Rightarrow$}
+	\RightLabel{Cut}
+	\BinaryInfC{$\vphantom{h}\Rightarrow$}
+	\DisplayProof
+\end{example}
+
+We observe of course that the applications of the qunatifier rules have become invalid in doing this.
+Now we are ready to give a definition of the transformation\\
+
+\newsavebox\InnerProofa
+\sbox\InnerProofa{
+	\AxiomC{$A_1\Rightarrow B_1$}\UnaryInfC{$\ddots$}
+	\AxiomC{$A_n\Rightarrow B_n$}\UnaryInfC{\reflectbox{$\ddots$}}
+	\noLine \BinaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+
+\newsavebox\InnerProofba
+\sbox\InnerProofba{
+	\AxiomC{$\forall xA, \Delta\Rightarrow\Gamma$}\RightLabel{L$\forall$}\UnaryInfC{$A[t/x], \Delta\Rightarrow\Gamma$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \UnaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+\newsavebox\InnerProofbb
+\sbox\InnerProofbb{
+	\AxiomC{$A[t/x],\Delta\Rightarrow\Gamma$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \UnaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+\newsavebox\InnerProofca
+\sbox\InnerProofca{
+	\AxiomC{$\Delta\Rightarrow\Gamma, \forall xA$}\RightLabel{R$\forall$}\UnaryInfC{$\Delta\Rightarrow\Gamma, A[t/x]$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \UnaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+\newsavebox\InnerProofcb
+\sbox\InnerProofcb{
+	\AxiomC{$\Delta\Rightarrow\Gamma, A[t/x]$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \UnaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+\newsavebox\InnerProofda
+\sbox\InnerProofda{
+	\AxiomC{$\exists xA, \Delta\Rightarrow\Gamma$}\RightLabel{L$\exists$}\UnaryInfC{$A[t/x], \Delta\Rightarrow\Gamma$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \UnaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+\newsavebox\InnerProofdb
+\sbox\InnerProofdb{
+	\AxiomC{$A[t/x],\Delta\Rightarrow\Gamma$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \UnaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+\newsavebox\InnerProofea
+\sbox\InnerProofea{
+	\AxiomC{$\Delta\Rightarrow\Gamma, \exists xA$}\RightLabel{R$\exists$}\UnaryInfC{$\Delta\Rightarrow\Gamma, A[t/x]$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \UnaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+\newsavebox\InnerProofeb
+\sbox\InnerProofeb{
+	\AxiomC{$\Delta\Rightarrow\Gamma, A[t/x]$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \UnaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+\newsavebox\InnerProoffa
+\sbox\InnerProoffa{
+	\AxiomC{$A\to B,\Delta\Rightarrow\Gamma$}\RightLabel{L$\to$1}\UnaryInfC{$B, \Delta\Rightarrow\Gamma$}\noLine\UnaryInfC{$\vdots$}\AxiomC{$A\to B, \Delta\Rightarrow\Gamma$}\RightLabel{L$\to$2}\UnaryInfC{$\Delta\Rightarrow\Gamma, A$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \BinaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+\newsavebox\InnerProoffb
+\sbox\InnerProoffb{
+	\AxiomC{$B, \Delta\Rightarrow\Gamma$}\noLine\UnaryInfC{$\vdots$}
+	\AxiomC{$\Delta\Rightarrow\Gamma, A$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \BinaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+
+\newsavebox\InnerProofga
+\sbox\InnerProofga{
+	\AxiomC{$\Delta\Rightarrow\Gamma, A\to B$}\RightLabel{R$\to$}\UnaryInfC{$A, \Delta\Rightarrow\Gamma, B$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \UnaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+\newsavebox\InnerProofgb
+\sbox\InnerProofgb{
+	\AxiomC{$A, \Delta\Rightarrow\Gamma, B$}\noLine\UnaryInfC{$\vdots$}
+	\noLine \UnaryInfC{\vphantom{T}}\UnaryInfC{\vphantom{T}$\Rightarrow$}\DisplayProof
+}
+
+{\hspace*{-1cm}}
+\begin{tabular}{ccc}
+	$\mathfrak f_2\left(\usebox\InnerProofa\right)$& = &
+	$\mathfrak f_3\left(\usebox\InnerProofa, B_1\dots B_n\Rightarrow A_1\dots A_n\right)$\\&&\\
+
+	$\mathfrak f_3\left(\usebox\InnerProofba, \Gamma'\Rightarrow\Delta'\right)$& = &
+	\AxiomC{$\mathfrak f_3\left(\left(\usebox\InnerProofbb\right)[a/t],(\Gamma'\Rightarrow \Delta'\setminus\{\forall xA\}, A[a/x])\right)$}\RightLabel{R$\forall$}\UnaryInfC{$\Gamma'\Rightarrow \Delta'$}\DisplayProof\\\multicolumn{3}{l}{where $a$ is a new free variable, unless $t$ is a free variable, then $t = a$.}\\&&\\
+
+	$\mathfrak f_3\left(\usebox\InnerProofca, \Gamma'\Rightarrow\Delta'\right)$& = &
+	\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofcb, A[t/x], \Gamma'\Rightarrow \Delta' \right)$}\RightLabel{L$\forall$}\UnaryInfC{$\Gamma'\Rightarrow \Delta'$}\DisplayProof\\&&\\
+
+	$\mathfrak f_3\left(\usebox\InnerProofda, \Gamma'\Rightarrow\Delta'\right)$& = &
+	\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofdb, \Gamma'\Rightarrow \Delta', A[t/x]\right)$}\RightLabel{R$\exists$}\UnaryInfC{$\Gamma'\Rightarrow \Delta'$}\DisplayProof\\&&\\
+
+	$\mathfrak f_3\left(\usebox\InnerProofea, \Gamma'\Rightarrow\Delta'\right)$& = &
+	\AxiomC{$\mathfrak f_3\left(\left(\usebox\InnerProofeb\right)[a/t], (A[a/x], \Gamma'\setminus\{\exists xA\}\Rightarrow \Delta')\right)$}\RightLabel{L$\exists$}\UnaryInfC{$\Gamma'\Rightarrow \Delta'$}\DisplayProof\\\multicolumn{3}{l}{where $a$ is a new free variable, unless $t$ is a free variable, then $t = a$.}\\&&\\
+\end{tabular}
+
+	$${\hspace*{-2.5cm}}\mathfrak f_3\left(\usebox\InnerProoffa, \Gamma'\Rightarrow\Delta'\right) = 
+	\AxiomC{$\mathfrak f_3\left(\usebox\InnerProoffb, A, \Gamma'\Rightarrow \Delta'\setminus\{A\to B\}, B\right)$}\RightLabel{R$\to$}\UnaryInfC{$\Gamma'\Rightarrow \Delta'$}\DisplayProof$$
+
+	where the occurrence of one of the rules on the left is optional.
+
+	$${\hspace*{-2.5cm}}\mathfrak f_3\left(\usebox\InnerProofga, \Gamma'\Rightarrow\Delta'\right)= 
+	\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofgb, \Gamma'\Rightarrow \Delta', A\right)$}\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofgb, B, \Gamma'\setminus\{A\to B\}\Rightarrow \Delta'\right)$}\RightLabel{L$\to$}\BinaryInfC{$\Gamma'\Rightarrow \Delta'$}\DisplayProof$$
+
+	And finally $\mathfrak f_3\left(T, \Gamma\Rightarrow\Delta\right) = \Gamma\Rightarrow\Delta$ if no more noramlization rules are left in $T$.
+(Note that so far $\mathfrak{f}_3$ does not technically determine a function since we have not specified which leaf to match on)
+
+Let's evaluate $\mathfrak f_2$ on our example:
+
+$$\mathfrak{f}_2\left(\AxiomC{$\Rightarrow \forall x\exists y P(x, y)$}
+\RightLabel{R$\forall$}
+\UnaryInfC{$\Rightarrow\exists yP(s_2, y)$}
+\RightLabel{R$\exists$}
+\UnaryInfC{$\Rightarrow P(s_2, s_1(s_2))$}
+\AxiomC{$\Rightarrow\forall x\forall y(P(x, y) \to Q(x, f(y)))$}
+\RightLabel{R$\forall$}
+\UnaryInfC{$\Rightarrow\forall y(P(s_2, y)\to Q(s_2, f(y)))$}
+\RightLabel{R$\forall$}
+\UnaryInfC{$\Rightarrow P(s_2, s_1(s_2))\to Q(s_2, f(s_1(s_2)))$}
+\RightLabel{R$\to$}
+\UnaryInfC{$P(s_2, s_1(s_2))\Rightarrow Q(s_2, f(s_1(s_2)))$}
+\RightLabel{Cut}
+\BinaryInfC{$\Rightarrow Q(s_2, f(s_1(s_2)))$}
+\AxiomC{$\forall x\exists y Q(x, y)\Rightarrow$}
+\RightLabel{L$\forall$}
+\UnaryInfC{$\exists y Q(s_2, y)\Rightarrow$}
+\RightLabel{L$\exists$}
+\UnaryInfC{$Q(s_2, f(s_1(s_2)))\Rightarrow$}
+\RightLabel{Cut}
+\BinaryInfC{$\vphantom{h}\Rightarrow$}
+\DisplayProof\right)=$$
+
+$$\mathfrak{f}_3\left(\begin{array}{l}\AxiomC{$\Rightarrow \forall x\exists y P(x, y)$}
+\RightLabel{R$\forall$}
+\UnaryInfC{$\Rightarrow\exists yP(a, y)$}
+\RightLabel{R$\exists$}
+\UnaryInfC{$\Rightarrow P(a, s_1(a))$}
+\AxiomC{$\Rightarrow\forall x\forall y(P(x, y) \to Q(x, f(y)))$}
+\RightLabel{R$\forall$}
+\UnaryInfC{$\Rightarrow\forall y(P(a, y)\to Q(a, f(y)))$}
+\RightLabel{R$\forall$}
+\UnaryInfC{$\Rightarrow P(a, s_1(a))\to Q(a, f(s_1(a)))$}
+\RightLabel{R$\to$}
+\UnaryInfC{$P(a, s_1(a))\Rightarrow Q(a, f(s_1(a)))$}
+\RightLabel{Cut}
+\BinaryInfC{$\Rightarrow Q(a, f(s_1(a)))$}
+\AxiomC{$\forall x\exists y Q(x, y)\Rightarrow$}
+\RightLabel{L$\forall$}
+\UnaryInfC{$\exists y Q(a, y)\Rightarrow$}
+\RightLabel{L$\exists$}
+\UnaryInfC{$Q(a, f(s_1(a)))\Rightarrow$}
+\RightLabel{Cut}
+\BinaryInfC{$\vphantom{h}\Rightarrow$}
+\DisplayProof\\\\\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)\end{array}\right)=$$
+\newsavebox\InnerProof
+\sbox\InnerProof{$\mathfrak{f}_3\left(\begin{array}{l}
+		\AxiomC{$\Rightarrow\exists yP(a, y)$}
+		\RightLabel{R$\exists$}
+		\UnaryInfC{$\Rightarrow P(a, s_1(a))$}
+		\AxiomC{$\Rightarrow\forall x\forall y(P(x, y) \to Q(x, f(y)))$}
+		\RightLabel{R$\forall$}
+		\UnaryInfC{$\Rightarrow\forall y(P(a, y)\to Q(a, f(y)))$}
+		\RightLabel{R$\forall$}
+		\UnaryInfC{$\Rightarrow P(a, s_1(a))\to Q(a, f(s_1(a)))$}
+		\RightLabel{R$\to$}
+		\UnaryInfC{$P(a, s_1(a))\Rightarrow Q(a, f(s_1(a)))$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\Rightarrow Q(a, f(s_1(a)))$}
+		\AxiomC{$\forall x\exists y Q(x, y)\Rightarrow$}
+		\RightLabel{L$\forall$}
+		\UnaryInfC{$\exists y Q(a, y)\Rightarrow$}
+		\RightLabel{L$\exists$}
+		\UnaryInfC{$Q(a, f(s_1(a)))\Rightarrow$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\vphantom{h}\Rightarrow$}
+		\DisplayProof\\\\\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)\end{array}\right)$
+}
+$$\AxiomC{\usebox\InnerProof}\RightLabel{L$\forall$}\UnaryInfC{$\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\DisplayProof=$$
+
+\newsavebox\InnerProoff
+\sbox\InnerProoff{$\mathfrak{f}_3\left(\begin{array}{l}
+		\AxiomC{$\Rightarrow P(a, s_1(a))$}
+		\AxiomC{$\Rightarrow\forall x\forall y(P(x, y) \to Q(x, f(y)))$}
+		\RightLabel{R$\forall$}
+		\UnaryInfC{$\Rightarrow\forall y(P(a, y)\to Q(a, f(y)))$}
+		\RightLabel{R$\forall$}
+		\UnaryInfC{$\Rightarrow P(a, s_1(a))\to Q(a, f(s_1(a)))$}
+		\RightLabel{R$\to$}
+		\UnaryInfC{$P(a, s_1(a))\Rightarrow Q(a, f(s_1(a)))$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\Rightarrow Q(a, f(s_1(a)))$}
+		\AxiomC{$\forall x\exists y Q(x, y)\Rightarrow$}
+		\RightLabel{L$\forall$}
+		\UnaryInfC{$\exists y Q(a, y)\Rightarrow$}
+		\RightLabel{L$\exists$}
+		\UnaryInfC{$Q(a, f(s_1(a)))\Rightarrow$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\vphantom{h}\Rightarrow$}
+		\DisplayProof\\\\P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)\end{array}\right)$
+}
+$$\AxiomC{\usebox\InnerProoff}\RightLabel{L$\exists$}\UnaryInfC{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\DisplayProof=$$
+
+\newsavebox\InnerProofff
+\sbox\InnerProofff{$\mathfrak{f}_3\left(\begin{array}{l}
+		\AxiomC{$\Rightarrow P(a, s_1(a))$}
+		\AxiomC{$\Rightarrow\forall y(P(a, y)\to Q(a, f(y)))$}
+		\RightLabel{R$\forall$}
+		\UnaryInfC{$\Rightarrow P(a, s_1(a))\to Q(a, f(s_1(a)))$}
+		\RightLabel{R$\to$}
+		\UnaryInfC{$P(a, s_1(a))\Rightarrow Q(a, f(s_1(a)))$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\Rightarrow Q(a, f(s_1(a)))$}
+		\AxiomC{$\forall x\exists y Q(x, y)\Rightarrow$}
+		\RightLabel{L$\forall$}
+		\UnaryInfC{$\exists y Q(a, y)\Rightarrow$}
+		\RightLabel{L$\exists$}
+		\UnaryInfC{$Q(a, f(s_1(a)))\Rightarrow$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\vphantom{h}\Rightarrow$}
+		\DisplayProof\\\\\stackanchor{$\forall y(P(a, y)\to Q(a, f(y))), P(a, s_1(a))$}{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)\end{array}\right)$
+}
+$$\AxiomC{\usebox\InnerProofff}\RightLabel{L$\forall$}\UnaryInfC{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\exists$}\UnaryInfC{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\DisplayProof=$$
+
+\newsavebox\InnerProoffff
+\sbox\InnerProoffff{$\mathfrak{f}_3\left(\begin{array}{l}
+		\AxiomC{$\Rightarrow P(a, s_1(a))$}
+		\AxiomC{$\Rightarrow P(a, s_1(a))\to Q(a, f(s_1(a)))$}
+		\RightLabel{R$\to$}
+		\UnaryInfC{$P(a, s_1(a))\Rightarrow Q(a, f(s_1(a)))$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\Rightarrow Q(a, f(s_1(a)))$}
+		\AxiomC{$\forall x\exists y Q(x, y)\Rightarrow$}
+		\RightLabel{L$\forall$}
+		\UnaryInfC{$\exists y Q(a, y)\Rightarrow$}
+		\RightLabel{L$\exists$}
+		\UnaryInfC{$Q(a, f(s_1(a)))\Rightarrow$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\vphantom{h}\Rightarrow$}
+		\DisplayProof\\\\\stackanchor{$P(a, s_1(a))\to Q(a, f(s_1(a))),\forall y(P(a, y)\to Q(a, f(y))), P(a, s_1(a))$}{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)\end{array}\right)$
+}
+$$\AxiomC{\usebox\InnerProoffff}\RightLabel{L$\forall$}\UnaryInfC{$\stackanchor{$P(a, s_1(a))\to Q(a, f(s_1(a))),\forall y(P(a, y)\to Q(a, f(y)))$}{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\exists$}\UnaryInfC{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\DisplayProof=$$
+
+\newsavebox\InnerProofffff
+\sbox\InnerProofffff{$\mathfrak{f}_3\left(\begin{array}{l}
+		\AxiomC{$\Rightarrow P(a, s_1(a))$}
+		\AxiomC{$P(a, s_1(a))\Rightarrow Q(a, f(s_1(a)))$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\Rightarrow Q(a, f(s_1(a)))$}
+		\AxiomC{$\forall x\exists y Q(x, y)\Rightarrow$}
+		\RightLabel{L$\forall$}
+		\UnaryInfC{$\exists y Q(a, y)\Rightarrow$}
+		\RightLabel{L$\exists$}
+		\UnaryInfC{$Q(a, f(s_1(a)))\Rightarrow$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\vphantom{h}\Rightarrow$}
+		\DisplayProof\\\\\stackanchor{$Q(a, f(s_1(a))), \forall y(P(a, y)\to Q(a, f(y))), P(a, s_1(a))$}{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)\end{array}\right)$
+}
+$$\hspace*{-1.5cm}\AxiomC{$\mathfrak f_3(\dots)$}\AxiomC{\usebox\InnerProofffff}\RightLabel{L$\to$}\BinaryInfC{$\stackanchor{$P(a, s_1(a))\to Q(a, f(s_1(a))),\forall y(P(a, y)\to Q(a, f(y)))$}{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$\stackanchor{$\forall y(P(a, y)\to Q(a, f(y)))$}{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\exists$}\UnaryInfC{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\DisplayProof=$$
+
+\newsavebox\InnerProoffffff
+\sbox\InnerProoffffff{$\mathfrak{f}_3\left(\begin{array}{l}
+		\AxiomC{$\Rightarrow P(a, s_1(a))$}
+		\AxiomC{$P(a, s_1(a))\Rightarrow Q(a, f(s_1(a)))$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\Rightarrow Q(a, f(s_1(a)))$}
+		\AxiomC{$\exists y Q(a, y)\Rightarrow$}
+		\RightLabel{L$\exists$}
+		\UnaryInfC{$Q(a, f(s_1(a)))\Rightarrow$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\vphantom{h}\Rightarrow$}
+		\DisplayProof\\\\\stackanchor{$Q(a, f(s_1(a))), \forall y(P(a, y)\to Q(a, f(y))), P(a, s_1(a))$}{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \exists y Q(a, y)\end{array}\right)$
+}
+$$\hspace*{-1.5cm}\AxiomC{$\mathfrak f_3(\dots)$}\AxiomC{\usebox\InnerProoffffff}\RightLabel{R$\forall$}\UnaryInfC{$\stackanchor{$Q(a, f(s_1(a))), \forall y(P(a, y)\to Q(a, f(y))), P(a, s_1(a))$}{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\to$}\BinaryInfC{$\stackanchor{$P(a, s_1(a))\to Q(a, f(s_1(a))),\forall y(P(a, y)\to Q(a, f(y)))$}{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$\stackanchor{$\forall y(P(a, y)\to Q(a, f(y)))$}{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\exists$}\UnaryInfC{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\DisplayProof=$$
+
+\newsavebox\InnerProofffffff
+\sbox\InnerProofffffff{$\mathfrak{f}_3\left(\begin{array}{l}
+		\AxiomC{$\Rightarrow P(a, s_1(a))$}
+		\AxiomC{$P(a, s_1(a))\Rightarrow Q(a, f(s_1(a)))$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\Rightarrow Q(a, f(s_1(a)))$}
+		\AxiomC{$Q(a, f(s_1(a)))\Rightarrow$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\vphantom{h}\Rightarrow$}
+		\DisplayProof\\\\\stackanchor{$Q(a, f(s_1(a))), \forall y(P(a, y)\to Q(a, f(y))), P(a, s_1(a))$}{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \exists y Q(a, y), Q(a, f(s_1(a)))\end{array}\right)$
+}
+$$\hspace*{-1.5cm}\AxiomC{$\mathfrak f_3(\dots)$}\AxiomC{\usebox\InnerProofffffff}\RightLabel{R$\exists$}\UnaryInfC{$\stackanchor{$Q(a, f(s_1(a))), \forall y(P(a, y)\to Q(a, f(y))), P(a, s_1(a))$}{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \exists y Q(a, y)$}\RightLabel{R$\forall$}\UnaryInfC{$\stackanchor{$Q(a, f(s_1(a))), \forall y(P(a, y)\to Q(a, f(y))), P(a, s_1(a))$}{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\to$}\BinaryInfC{$\stackanchor{$P(a, s_1(a))\to Q(a, f(s_1(a))),\forall y(P(a, y)\to Q(a, f(y)))$}{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$\stackanchor{$\forall y(P(a, y)\to Q(a, f(y)))$}{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\exists$}\UnaryInfC{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\DisplayProof=$$
+
+$$\AxiomC{$\mathfrak f_3(\dots)$}\AxiomC{$\stackanchor{$Q(a, f(s_1(a))), \forall y(P(a, y)\to Q(a, f(y))), P(a, s_1(a))$}{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \exists y Q(a, y), Q(a, f(s_1(a)))$}\RightLabel{R$\exists$}\UnaryInfC{$\stackanchor{$Q(a, f(s_1(a))), \forall y(P(a, y)\to Q(a, f(y))), P(a, s_1(a))$}{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \exists y Q(a, y)$}\RightLabel{R$\forall$}\UnaryInfC{$\stackanchor{$Q(a, f(s_1(a))), \forall y(P(a, y)\to Q(a, f(y))), P(a, s_1(a))$}{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\to$}\BinaryInfC{$\stackanchor{$P(a, s_1(a))\to Q(a, f(s_1(a))),\forall y(P(a, y)\to Q(a, f(y)))$}{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$\stackanchor{$\forall y(P(a, y)\to Q(a, f(y)))$}{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))$}\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$P(a, s_1(a)), \exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\exists$}\UnaryInfC{$\exists yP(a, y),\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\RightLabel{L$\forall$}\UnaryInfC{$\forall x\exists y P(x, y),\forall x\forall y(P(x, y) \to Q(x, f(y)))\Rightarrow \forall x\exists y Q(x, y)$}\DisplayProof$$
+
+The left subtree is evaluated analogously giving us indeed a valid \Gc and even \Gi proof.
+
+\begin{lemma}[Correctness]
+	If $T$ is a valid Resolution proof then $\mathfrak{f}_2(\mathfrak{f}_1(T))$ is a valid \Gc proof.
+\end{lemma}
+
+Now the extracted proof is intuitionistically valid if for $\mathfrak{f}_2$ in the cases L$\to$ we have $\Delta' = A\to B$ and in L$\forall$ we have $\Delta' = \forall xA$.
+Note that this is however not even deterministic, since we have not specified which leaf to match on!
+Consider the resolution proof of the trivial $A\to B\Rightarrow A\to B$.
+$$
+\AxiomC{$\Rightarrow A\to B$}
+\RightLabel{R$\to$}
+\UnaryInfC{$A\Rightarrow B$}
+\AxiomC{$A\to B\Rightarrow$}
+\RightLabel{L$\to$1}
+\UnaryInfC{$B\Rightarrow$}
+\RightLabel{Res}
+\BinaryInfC{A$\Rightarrow$}
+\AxiomC{$A\to B\Rightarrow$}
+\RightLabel{L$\to$1}
+\UnaryInfC{$\Rightarrow A$}
+\RightLabel{Res}
+\BinaryInfC{$\Rightarrow$}
+\DisplayProof
+$$
+
+Then we can have
+
+$$\mathfrak{f_2}\left(\AxiomC{$\Rightarrow A\to B$}
+\RightLabel{R$\to$}
+\UnaryInfC{$A\Rightarrow B$}
+\AxiomC{$A\to B\Rightarrow$}
+\RightLabel{L$\to$1}
+\UnaryInfC{$B\Rightarrow$}
+\RightLabel{Res}
+\BinaryInfC{A$\Rightarrow$}
+\AxiomC{$A\to B\Rightarrow$}
+\RightLabel{L$\to$1}
+\UnaryInfC{$\Rightarrow A$}
+\RightLabel{Res}
+\BinaryInfC{$\Rightarrow$}
+\DisplayProof\right)$$ can be, if we match on L$\to$ first, the intuitionistically valid $$\AxiomC{$A\to B, A\Rightarrow B, A$}\AxiomC{$A, B\Rightarrow B$}\RightLabel{R$\to$}\BinaryInfC{$A\to B, A\Rightarrow B$}\RightLabel{L$\to$}\UnaryInfC{$A\to B\Rightarrow A\to B$}\DisplayProof$$
+or, if we match on R$\to$ first, the not intuitionistically valid $$\AxiomC{$A\to B, A\Rightarrow B, A$}\RightLabel{R$\to$}\UnaryInfC{$A\to B\Rightarrow A\to B, A$}\AxiomC{$A, B\Rightarrow B$}\RightLabel{R$\to$}\UnaryInfC{$B\Rightarrow A\to B$}\RightLabel{L$\to$}\BinaryInfC{$A\to B\Rightarrow A\to B$}\DisplayProof$$
+a
+\\
+
+Now towards the main result:
+\begin{definition}
+	A formula is \textit{relevant} if it can constructed by the following steps:
+	\begin{itemize}
+		\item $\bot$ is a relevant
+		\item If $B$ is relevant and $A$ is arbitrary $A\to B$ is relevant.
+		\item if $A$ is relevant then $\forall x A[x/a]$ is relevant.
+	\end{itemize}
+	it is \textit{irrelevant} if it is not relevant.
+\end{definition}
+
+Intuitively a relevant formula is a formula that ends in $\bot$.
+
+We can now by simultaneous induction define the notions of goal formulas and definite formulas
+
+\begin{definition}
+	A formula is a \textit{Goal Formula} if it can be constructed by the following steps:
+	\begin{itemize}
+		\item Any atomic formula is a goal formula
+		\item For a definite formula $D$, goal formula $G$, if $D$ is relevant or quantifier-free, $D\to G$ is a goal formula.
+		\item If $G$ is a goal formula then $\forall xG[x/a]$ is a goal formula.
+	\end{itemize}
+	A formula is a \textit{Definite Formula} if it can be constructed by the following steps:
+	\begin{itemize}
+		\item Any atomic formula is a definite formula
+		\item For a definite formula $D$, goal formula $G$, if $D$ is relevant or $G$ irrelevant, $G\to D$ is a definite formula.
+		\item If $D$ is a definite formula then $\forall xD[x/a]$ is a definite formula.
+	\end{itemize}
+\end{definition}
+
+\begin{theorem}
+	Let $D_1,\dots, D_n$ be definite formulas and $G$ a goal formula. If $D_1,\dots,D_n\Rightarrow G$ has a resolution proof then $D_1,\dots,D_n\Rightarrow G$ is provable in \Gi.
+\end{theorem}
+
+\begin{lemma}
+	Let $A_1,\dots, A_n\Rightarrow B_1, \dots, B_m$ be the label in the node of a resolution refutation of $\{\Rightarrow D_i\}\cup\{G\}$. Then each $A_i$ is a goal forumla and each $B_i$ is definite.
+\end{lemma}
+\begin{proof}
+	Induction.
+\end{proof}
+\pagebreak
+
 \bibliographystyle{acm}

### 2022-03-21 Grant Proposal/Draft Alexander/Proposal.tex
Grant Proposal First Draft

\documentclass[onehalfspacing]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{ wasysym }
\usepackage{ stmaryrd }
\usepackage{ mathpartir }
\usepackage{xcolor}
\usepackage{bussproofs}
\usepackage[left=3.00cm, right=3.00cm, top=3.00cm,bottom=3.0cm]{geometry}
\usepackage{pdflscape}
\usepackage{xspace}

\title{WWTF Grant Proposal\\Leveraging Resolution for Verified Program Synthesis}

\begin{document}

	\maketitle

	\section{Project summary}

	Program synthesis is a long-standing problem crossing many areas of computer science and mathematics including programming languages, artificial intelligence, formal methods and proof theory. There have been major advances in NLP (natural language processing) based program synthesis recently, most prominently DeeepMind's AlphaCode~\cite{AlphaCode1}, OpenAI's Codex~\cite{Codex1} and Github's Copilot~\cite{Copilot1}.

	However a major drawback of such an approach is that there is little control over the generated code leaving us with no guarantees for reliability or security of the resulting programs.

	Another route to program synthesis is offered through the Curry-Howard correspondence. Exploiting the equivalence of proofs in intuitionistic logic and functional programs allows us to reduce the problem of synthesising a program satisfying a specification into proving its validity. The added benefit of this approach is that a program generated from such a proof is guaranteed to be correct. From the perspective of formal methods there are benefits as well. At the moment creating a verified program requires both writing specification and code. Being able to automate the latter step would save considerable overhead.

	However automated theorem proving in intuitionistic logic is notoriously difficult in particular due to the absence of convenient normal forms. Most state of the art theorem provers like Vampire~\cite{Vampire1} and E~\cite{E1} use resolution which is a great technique for automation but inherently non-intuitionistic in that it requires the formulas to be transformed into conjunctive normal form and works through refutation of the negated formula, which is not  intuitionistically sufficient for a proof of validity.

	The goal of this project is to establish when and to what degree theorem provers based on classical logic can still be utilized, allowing us to leverage the simplicity of resolution to synthesize programs, and to then implement a functionality to this end on top of the Vampire theorem prover.

	\section{Introduction, background and state of the art}

	Program synthesis in the sense of this project is concerned with generating a program that satisfies a specification given in FOL (first order logic). Instead of directly generating a program however we obtain it via the Curry-Howard correspondence from a proof of the validity of the specification. Such a program will always be valid.

	Cutting edge theorem provers like Vampire~\cite{Vampire1} and E~\cite{E1} have advanced to a point where they can already fully automatically prove many specifications~\cite{CASC} and are ready to be used in practical systems, in particular when provided with hints. With recent development to inductive reasoning~\cite{Vampire2}\cite{Vampire3} we can expect that more and more specifications will be provable in the future.

	However for the Curry-Howard correspondence to apply the initial proof has to be valid in intuitionistic logic but state-of-the-art theorem provers utilize resolution which is an inherently classical techniques. There are 2 approaches  present in literature that can help to remedy this fact.

	On the one hand the Curry-Howard correspondence can be extended to classical logic using control operators~\cite{Control1}, Parigot's $\lambda\mu$-Calculus~\cite{Parigot1} or CPS-translation~\cite{CPS1}. However in this process we must in general give up on some desirable properties, at the very least realizability of $\vee$ and $\exists$, i.e. a proof of $\exists xP(x)$ will not necessarily give us an $x$ such that $P(x)$ and a proof of $A\vee B$ might give us neither a proof of $A$ nor a proof of $B$.

	On the other hand in certain regimes it is possible to transform a classical proof into an intuitionistic one, e.g. proofs of $\Pi_2$ sentences in Peano Arithmetic~\cite{HAPA}, via Friedman's Translation or similar methods.~\cite{Friedman} This approach has been examined in different small-scale settings and would have to be adapted for a general setting.

	Finally we suggest a third approach which has not been covered in literature yet, that is to pass a modified formula to the classical prover which can give us additional information about what an intuitionistic proof can look like.

	By combining the above approaches we hope to examine to what extent we can extract intuitionistic information from a specification using a classical theorem prover and implement a real-world system for this on top of the vampire theorem prover.

	\section{Research questions, objectives and hypotheses}

	The final goal of our project is to leverage state-of-the-art first-order automated theorem provers, in particular Vampire, for program synthesis. This is motivated by the recent and prospective advancements in their capabilities.

	Before beginning an implementation a lot of theoretical groundwork will have to be laid, this will be (WP1). An optimal procedure for synthesis would consist of
	\begin{itemize}
		\item transforming the specification into an alternate form that would guarantee the use of as few non-intuitionistic rules as possible and applicability of Friedman translation in as many places as possible
		\item applying the prover to this modified specification obtaining a classical proof
		\item apply Friedman translation or similar translation procedures in as many places as possible
		\item extracting a program via Curry-Howard correspondence, applying control operators to non-intuitionistic remnants as necessary.
	\end{itemize}
	All of these elements except the initial modification of the specification already exist separately and in model settings such as arithmetic. Extending these results and combining them will be the first major undertaking of this project. This is also a necessary precondition for determining what initial modifications should be undertaken. We expect that for most real-world synthesis problems our procedure will be sufficient to transform classical proofs into usable and verified programs.

	Finally we want to implement a real system on top of Vampire realizing the synthesis problem (WP2) which will target a suitable functional programming language.

	\section{Expected results, novelty, and relevance}

	As outlined we expect to obtain the necessary theoretical foundations as well as a practical system to leverage state-of-the-art classical theorem provers for program extraction. At the current time no such system exists. Much of the theoretical groundwork for implementing such a system is already present, but is scattered and mostly applied in  theoretical model settings.

	A successful completion of this project would lay the foundation for verified program synthesis via automated theorem provers.

	\section{Methods and feasibility}

	\section{Potential to span across disciplines and to build bridges to application fields}

	\bibliographystyle{acm}
	\bibliography{references}

\end{document}

### 2022-03-21 Notes/notes.tex
asdf
Grant Proposal First Draft

@@
 \usepackage{stackengine}
+\usepackage{tikz}
+\usepackage{tikz-qtree}
 
@@
 \\&&\\
-	\AxiomC{$A\to B, \Gamma\Rightarrow\Delta, A$}
+	\AxiomC{$\Gamma\Rightarrow\Delta, A$}
 	\AxiomC{$B, \Gamma\Rightarrow\Delta$}
@@
 \\&&\\
-	\AxiomC{$\forall xA, A[t/x], \Gamma\Rightarrow\Delta$}
+	\AxiomC{$A[t/x], \Gamma\Rightarrow\Delta$}
 	\RightLabel{L$\forall$}
@@
 	\DisplayProof&
-	\AxiomC{$\Gamma\Rightarrow\Delta, A[t/x], \exists xA$}
+	\AxiomC{$\Gamma\Rightarrow\Delta, A[t/x]$}
 	\RightLabel{R$\exists$}
@@
 
-where in R$\forall$ and L$\exists$ $a$ is a free variable not occurring in the conclusion.
+where in R$\forall$ and L$\exists$ $a$ is a free variable not occurring in the sequent otherwise.
 
@@
 
-	\hspace{-1cm}
+	\hspace{-.5cm}
 	\AxiomC{}
 	\RightLabel{Ax}
-	\UnaryInfC{\scriptsize{\stackanchor{$P(a, b)\to Q(a, f(b)), \forall x\forall y(P(x, y)\to Q(x, f(y)))$}{$\forall x\exists yP(x, y), P(a, b), \forall y(P(a, y)\to Q(a, f(y)))$}$\Rightarrow$\stackanchor{$P(a, b), Q(a, f(b))$}{$\exists yQ(a, f(y))$}}}
+	\UnaryInfC{\scriptsize{\stackanchor{$\forall x\forall y(P(x, y)\to Q(x, f(y))), P(a, b)$}{$\forall x\exists yP(x, y), \forall y(P(a, y)\to Q(a, f(y)))$}$\Rightarrow$\stackanchor{$P(a, b), Q(a, f(b))$}{$\exists yQ(a, f(y))$}}}
 	\RightLabel{Ax}
@@
 	\AxiomC{$A, \Gamma\Rightarrow B$}
-	\RightLabel{R$\to$}
+	\RightLabel{R$\to_1$}
 	\UnaryInfC{$\Gamma\Rightarrow\Delta, A\to B$}
@@
 	\AxiomC{$\Gamma\Rightarrow A[a/x]$}
-	\RightLabel{R$\forall$}
+	\RightLabel{R$\forall_1$}
 	\UnaryInfC{$\Gamma\Rightarrow\Delta, \forall xA$}
@@
 
-	$${\hspace*{-2.5cm}}\mathfrak f_3\left(\usebox\InnerProofga, \Gamma'\Rightarrow\Delta'\right)= 
-	\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofgb, \Gamma'\Rightarrow \Delta', A\right)$}\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofgb, B, \Gamma'\setminus\{A\to B\}\Rightarrow \Delta'\right)$}\RightLabel{L$\to$}\BinaryInfC{$\Gamma'\Rightarrow \Delta'$}\DisplayProof$$
+	$${\hspace*{-2.95cm}}\mathfrak f_3\left(\usebox\InnerProofga, \Gamma'\Rightarrow\Delta'\right)= 
+	\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofgb, \Gamma'\setminus\{A\to B\}\Rightarrow \Delta', A\right)$}\AxiomC{$\mathfrak f_3\left(\usebox\InnerProofgb, B, \Gamma'\setminus\{A\to B\}\Rightarrow \Delta'\right)$}\RightLabel{L$\to$}\BinaryInfC{$\Gamma'\Rightarrow \Delta'$}\DisplayProof$$
 
@@
 or, if we match on R$\to$ first, the not intuitionistically valid $$\AxiomC{$A\to B, A\Rightarrow B, A$}\RightLabel{R$\to$}\UnaryInfC{$A\to B\Rightarrow A\to B, A$}\AxiomC{$A, B\Rightarrow B$}\RightLabel{R$\to$}\UnaryInfC{$B\Rightarrow A\to B$}\RightLabel{L$\to$}\BinaryInfC{$A\to B\Rightarrow A\to B$}\DisplayProof$$
-a
 \\
 
-Now towards the main result:
+Now we may ask what structural conditions we have to impose on a sequent so that we can always extract an intuitionistic proof, i.e.
+every instance of R$\to$ and R$\to$ is a valid instance of R$\to_1$ and R$\forall_1$.
+
+As usual we can associate with every sequent and formula a tree:
+
 \begin{definition}
-	A formula is \textit{relevant} if it can constructed by the following steps:
+	A \textit{Formula Tree} is a labelled ordered rooted tree in which every node is labelled with a connector, quantifier or atom.
+We can define a function $\iota$ that transforms sequents and formulas into formula trees as follows:
 	\begin{itemize}
-		\item $\bot$ is a relevant
-		\item If $B$ is relevant and $A$ is arbitrary $A\to B$ is relevant.
-		\item if $A$ is relevant then $\forall x A[x/a]$ is relevant.
-	\end{itemize}
-	it is \textit{irrelevant} if it is not relevant.
-\end{definition}
-
-Intuitively a relevant formula is a formula that ends in $\bot$.
-
-We can now by simultaneous induction define the notions of goal formulas and definite formulas
-
-\begin{definition}
-	A formula is a \textit{Goal Formula} if it can be constructed by the following steps:
-	\begin{itemize}
-		\item Any atomic formula is a goal formula
-		\item For a definite formula $D$, goal formula $G$, if $D$ is relevant or quantifier-free, $D\to G$ is a goal formula.
-		\item If $G$ is a goal formula then $\forall xG[x/a]$ is a goal formula.
-	\end{itemize}
-	A formula is a \textit{Definite Formula} if it can be constructed by the following steps:
-	\begin{itemize}
-		\item Any atomic formula is a definite formula
-		\item For a definite formula $D$, goal formula $G$, if $D$ is relevant or $G$ irrelevant, $G\to D$ is a definite formula.
-		\item If $D$ is a definite formula then $\forall xD[x/a]$ is a definite formula.
+		\item $\iota(A_1,\dots, A_n\Rightarrow B) = \iota(A_1,\dots, A_{n-1}\Rightarrow A_n\to B)$, $\iota(\Rightarrow B) = \iota(B)$
+		\item $\iota(A\square B) = $\tikz[scale=0.9, baseline=-5mm, thick]{\Tree [.$\square$ [.$\iota(A)$  ] [.$\iota(B)$  ] ]} for $\square\in\{\wedge,\vee,\to\}$.
+		\item $\iota(\square xA) = $\tikz[scale=0.9, baseline=-5mm, thick]{\Tree [.$\square x$ [.$\iota(A)$  ]]} for $\square\in\{\forall, \exists\}$.
+		\item $\iota(A) = A$ for $A$ atomic.
 	\end{itemize}
@@
 
+\begin{definition}
+	To each node of a formula tree we assign a \textit{Polarity} as follows:
+	The polarity of the root node is $1$.
+If the label of a node with polarity $p$ is in $\{\wedge,\vee,\forall,\exists\}$ then the polarity of its children is also $p$.
+If the label of a node with polarity $p$ is $\to$ then then polarity of its left child is $-p$ and of its right child $p$.
+\end{definition}
+ We annotate the polarity of a node with a superscript $+$ or $-$.
+
+\begin{example}
+	For our running example $\forall x\exists yP(x, y), \forall x\forall y(P(x, y)\to Q(x, f(y)))\Rightarrow\forall x\exists Q(x, y)$ the associated formula tree is
+	\begin{center}
+		\begin{tikzpicture}
+			\Tree [.$\to^+$ [.$\to^-$  [.$\forall x^+$ [.$\exists y^+$ $P(x,y)^+$ ] ] [.$\exists x^-$ [.$\forall y^-$ [.$\to^-$ $P(x,y)^+$ $Q(x,f(x))^-$ ] ] ] ]  [.$\forall x^+$  [.$\exists y^+$ $Q(x,y)^+$ ] ] ]
+		\end{tikzpicture}
+	\end{center}
+\end{example}
+
+We can now give sufficient conditions for when an extracted proof will always be intuitionstic:
+
 \begin{theorem}
-	Let $D_1,\dots, D_n$ be definite formulas and $G$ a goal formula. If $D_1,\dots,D_n\Rightarrow G$ has a resolution proof then $D_1,\dots,D_n\Rightarrow G$ is provable in \Gi.
+	Suppose in the formula tree of $\varphi$ there is no occurrence of $\to^+$ or $\forall^+$ below $\to^-$ or $\vee^+$.
+Then for any resolution proof $\mathcal{T}$ of $\varphi$ the extracted \Gc proof $\mathfrak{f}_2(\mathcal{T})$ is a valid \Gi proof.\\
 \end{theorem}
 
-\begin{lemma}
-	Let $A_1,\dots, A_n\Rightarrow B_1, \dots, B_m$ be the label in the node of a resolution refutation of $\{\Rightarrow D_i\}\cup\{G\}$. Then each $A_i$ is a goal forumla and each $B_i$ is definite.
-\end{lemma}
-\begin{proof}
-	Induction.
-\end{proof}
+This is however quite restrictive.
+Even quite simple, intuitionistically valid sequents like $$(A\to A)\to A\Rightarrow A$$ or $$(A\to B)\to C, A\to B\Rightarrow C$$ are not covered by the above theorem.\\
+However by examining alternatives to rules R$\to_1$, R$\forall_1$ we can find more broadly applicable results.
+
+\subsection{$\Pi_2$ sentences in PA}
+
+\subsection{A refined algorithm}
+
+Consider the calculus \Gi' obtained from \Gi by replacing R$\to_1$ and R$\forall_1$ with the rules 
+\begin{center}
+	\AxiomC{$A, \Gamma\Rightarrow \Delta, B$}
+	\RightLabel{R$\to_2$}
+	\UnaryInfC{$\Gamma\Rightarrow\Delta, A\to B$}
+	\DisplayProof\hspace{2cm}
+	\AxiomC{$\Gamma\Rightarrow \Delta,A[a/x]$}
+	\RightLabel{R$\forall_2$}
+	\UnaryInfC{$\Gamma\Rightarrow\Delta, \forall xA$}
+	\DisplayProof
+\end{center}
+where the introductory axiom in the branch is not between atoms from $\Delta$ and $A$.
+
+In we analyse our reasoning for theorem 5.7 under this calculus we identify for each formula precisely the atoms we may not resolve with each other.
+
+Intuitively this corresponds to the fact that doubly negated atoms cannot resolve request.
+
 \pagebreak
+
+We now propose a translation procedure for formulas $\varphi$ such that $\varphi^\circ$ is provable by resolution iff $\varphi$ is valid in \Gi'.
+The key insight is that atoms that occur left below $\to^-$ may not be resolved with atoms that occur below only positive nodes and atoms that occur below an initial $\vee^+$ may not be resolved with atoms from the other branch of $\vee^+$:
+
+First rename every atom $P$ to $P^0$.
+Rename all atoms that do not occur below a $\to^-$ to $P^1$ and all positive atoms that occurr below only positive nodes to $P^{-1}$, these atoms are not affected by any further relabelling.
+If a node is labelled $\vee^+$ and every node above is positive relabel all atoms $P^i$ in the left branch to $P^{i+1}$ and in the right branch to $P^{i+2}$.
+The thus obtained formula we call $\varphi^\circ$ and for each atom $P$ we denote the maximum $i$ such that $P^i$ occurs in $\varphi^\circ$ with $\chi_P(\varphi)$.
+\begin{example}The sequent $(A\to A)\to A\Rightarrow A$ gets transformed to $(A^0\to A^0)\to A^1\Rightarrow A^1$, the sequent $(A\to \bot)\to \bot\Rightarrow A$ gets transformed to $(A^0\to \bot)\to\bot\Rightarrow A^1$.
+
+\begin{theorem}
+	$\{P^i\to P^0\:|\:P \text{ atom in $\varphi$}, i\leq \chi_{P}(\varphi)\}\Rightarrow \varphi^\circ$ is provable by resolution iff $\varphi$ provable in \Gi'.
+\end{theorem}
+
+\end{example}
 

### 2022-03-21 WWTF/main.tex

\documentclass[a4paper,12pt]{article}
\begin{document}
\title{WWTF ICT 2022}
\author{Florian Zuleger and Alexander Pluska}
\date{\today}
\maketitle

\begin{abstract}
  %Project Summary/Abstract (max. 1500 characters)
  %
  %Please provide an overall project summary. Clearly state the research question and the aims of the team. Describe the most innovative aspects of your proposal, and its potential relevance and impact.
  tba
\end{abstract}

\section{Introduction, background and state of the art (max. 2500 characters)}
%What is the topic and broad background of this research project? What is the state of the art in this field?

``WORKSHOP ON SYNTHESIS''
https://www.cs.technion.ac.il/~shaull/SYNT2022/index.html

workshop on ``Synthesis of Models and Systems'' 
https://simons.berkeley.edu/workshops/tfcs2021-1

\section{Research questions, hypotheses and objectives (max. 2000 characters)}
%What are the fundamental research questions/hypotheses that the project seeks to address? What are the goals the project seeks to achieve? Provide an outline of the scientific approach that will be used to address the questions and reach the objectives. The objectives should be achievable within the duration of your project. Please provide preliminary data and any relevant research experience.

\section{Expected results, novelty and relevance (max. 2000 characters)}
%What kind of advancements are expected to be gained from the research project? Which aspects of the proposed project are especially innovative? Describe the scientific relevance and the timeliness of the research project. Please note that the development of source code, systems or protocols should not be the main goal of the project; however, they are encouraged in the context of demonstrating the practical applicability of new scientific results.

\section{Methods and feasibility (max. 2000 characters)}
%Specify the methodology intended to be used in order to answer the research question(s) and objective(s).
%Describe the basic working principles and concepts, and why the chosen approach/ specific mix of approaches is the most suitable for the research question. Include a short assessment on the feasibility of the approach.

\section{Role of team members and collaborative aspects (max. 1500 characters)}
%Each team member should describe his/her role in accomplishing the goal of the team. If applicable, which different disciplines are represented in the project and which partnerships across research groups and institutions will result from the project? Which collaborative elements are essential for the project to succeed; what makes the team more than the sum of its individual contributions? A team member’s contribution should be integrated into the overall plan and should not appear merely as a resource.

\section{Potential for interdisciplinarity and application to other research fields (max. 1000 characters)}
%How will the scientific advancements gained through this project benefit other research fields that are beyond the immediate scope of the project? Describe how the knowledge, methods or technology developed in this project could further understanding or contribute to applications in other research disciplines.

\section{Potential medium-term economic / societal benefits (max. 500 characters)}
%Provide a brief statement how the results from this research project could contribute towards solving a recognised problem that affects broader society or the economy.

Key references (max. 10 citations, max. 1250 characters)
Cite a maximum of 10 of the most relevant background publications for the proposal. There is no required citation format, nevertheless publications should be cited in a way that allows the reader to easily retrieve the key information.

Schedule/ Project overview (1 page Gantt chart, PDF upload)
Upload a Gantt chart to give an overview of the milestones to be achieved during the project period. Specify the time periods and the respective PIs responsible for each milestone.

\section{Potential ethical aspects (max. 500 characters)}
%If required, information should be given with respect to the ethical approval. Is an ethical approval available that covers the prospective use of data and the planned analysis? If not, state whether an additional approval is required.

\bibliographystyle{plain} % We choose the "plain" reference style
\bibliography{refs} % Entries are in the refs.bib file

\end{document}

### 2022-03-21 WWTF/Draft Alexander/Proposal.tex
Initial Draft Idea

--- 2022-03-21 Grant Proposal/Draft Alexander/Proposal.tex
@@
 	Program synthesis is a long-standing problem crossing many areas of computer science and mathematics including programming languages, artificial intelligence, formal methods and proof theory.
-There have been major advances in NLP (natural language processing) based program synthesis recently, most prominently DeeepMind's AlphaCode~\cite{AlphaCode1}, OpenAI's Codex~\cite{Codex1} and Github's Copilot~\cite{Copilot1}.
+There have been major advances in NLP (natural language processing) based program synthesis recently, most prominently DeeepMind's AlphaCode, OpenAI's Codex and Github's Copilot.
 
@@
 	Another route to program synthesis is offered through the Curry-Howard correspondence.
-Exploiting the equivalence of proofs in intuitionistic logic and functional programs allows us to reduce the problem of synthesising a program satisfying a specification into proving its validity.
+Exploiting the equivalence of proofs in intuitionistic logic and functional programs allows us to reduce the problem of synthesising a program satisfying a specification into proving the validity of said specification.
 The added benefit of this approach is that a program generated from such a proof is guaranteed to be correct.
-From the perspective of formal methods there are benefits as well.
-At the moment creating a verified program requires both writing specification and code.
-Being able to automate the latter step would save considerable overhead.
 
 	However automated theorem proving in intuitionistic logic is notoriously difficult in particular due to the absence of convenient normal forms.
-Most state of the art theorem provers like Vampire~\cite{Vampire1} and E~\cite{E1} use resolution which is a great technique for automation but inherently non-intuitionistic in that it requires the formulas to be transformed into conjunctive normal form and works through refutation of the negated formula, which is not  intuitionistically sufficient for a proof of validity.
+Most state of the art provers use techniques which are inherently non-intuitionistic.
 
@@
 Instead of directly generating a program however we obtain it via the Curry-Howard correspondence from a proof of the validity of the specification.
-Such a program will always be valid.
+Such a program will always satisfy the specification.
 
@@
 
-	However for the Curry-Howard correspondence to apply the initial proof has to be valid in intuitionistic logic but state-of-the-art theorem provers utilize resolution which is an inherently classical techniques.
-There are 2 approaches  present in literature that can help to remedy this fact.
+	However for the Curry-Howard correspondence to apply the initial proof has to be valid in intuitionistic logic but state-of-the-art theorem provers utilize resolution which is an inherently classical technique.
+There are 2 approaches  present in literature that can help to remedy this.
 
@@
 	On the other hand in certain regimes it is possible to transform a classical proof into an intuitionistic one, e.g.
-proofs of $\Pi_2$ sentences in Peano Arithmetic~\cite{HAPA}, via Friedman's Translation or similar methods.~\cite{Friedman} This approach has been examined in different small-scale settings and would have to be adapted for a general setting.
+proofs of $\Pi_2$ sentences in Peano Arithmetic~\cite{HAPA}, via Friedman's Translation or similar methods.~\cite{Friedman} This approach has been examined in different model settings and would have to be adapted and extended for our general problem.
 
@@
 
-	\bibliographystyle{acm}
+	\bibliographystyle{plain}
 	\bibliography{references}

### 2022-03-21 WWTF/main.tex
summary

@@
 \begin{document}
-\title{WWTF ICT 2022}
+\title{Leveraging The Progress in Automated Theorem Proving for Automated Program Synthesis\\\\
+\small{WWTF ICT 2022}}
 \author{Florian Zuleger and Alexander Pluska}
@@
 Describe the most innovative aspects of your proposal, and its potential relevance and impact.
-  tba
+
 \end{abstract}
@@
 %What is the topic and broad background of this research project? What is the state of the art in this field?
-
-``WORKSHOP ON SYNTHESIS''
-https://www.cs.technion.ac.il/~shaull/SYNT2022/index.html
-
-workshop on ``Synthesis of Models and Systems'' 
-https://simons.berkeley.edu/workshops/tfcs2021-1
+Programs are usually first written and then tested/verified.
+In contrast, program synthesis allows a system developer to derive a program in one step.
+At the same time, program synthesis holds the promise of freeing the programmer from low-level details and of enabling the programmer to focus on the high-level intent.
+Program synthesis is a long-standing problem that crosses many areas of computer science and mathematics including programming languages, AI, formal methods and proof theory.
+Traditionally, program synthesis has been formalized as a problem in deductive theorem proving~\cite{conf/ijcai/MannaW79}:
+Given a logical formula that states that for all inputs there exists an output such that the required specification holds, one searches for a constructive proof of this formula and then extracts a program from this proof.
+Progress on the deductive synthesis problem was slow for many years because of a lack of automated and scalable provers.
+However, a recent variant of the program synthesis problem has enabled dramatic progress (we refer to the special issue~\cite{fisman2022special} and the synthesis workshop\footnote{https://simons.berkeley.edu/workshops/tfcs2021-1} for an overview):
+%\footnote{https://www.cs.technion.ac.il/~shaull/SYNT2022/index.html}
+The logical specification is accompanied by a syntactic template that constrains the space of possible implementations.
+The synthesis procedure then searches for/enumerates candidate implementations and either rejects these candidates by finding violating input/output pairs or accept the generated program by verifying their correctness.
+The syntactic approach is very versatile and synthesis approaches have been explored that solely start from input/output examples or that mine possible program fragments from public code repositories.
+While the reported progress is impressive, we believe that the traditional deductive approach to synthesis has been unjustly neglected.
+In this project we aim at connecting classical ideas of proofs-as-programs~\cite{?} with modern automated theorem provers, in particular first-order solvers (and SMT solvers???), which have made impressive progress over the recent years (as evidenced by yearly competitions).
+We believe this approach offers the following advantages:
+1) First-order solvers are particularly suitable to solving \emph{quantified} formulas as needed for synthesis.
+2) Proofs-as-programs represent a longstanding approach to program extraction for which a large body of theory has been developed.
+3) We are able to take advantage of the highly sophisticated and optimized proof search strategies of first-order solvers instead of needing to define and implement \emph{ad-hoc} search strategies.
 
@@
 Please provide preliminary data and any relevant research experience.
+
+gap in the literature
+
+input/output pairs, existing program fragments
 
@@
 \bibliographystyle{plain} % We choose the "plain" reference style
-\bibliography{refs} % Entries are in the refs.bib file
+\bibliography{literature} % Entries are in the refs.bib file
 

### 2022-03-21 WWTF/main.tex
asdf

@@
 Progress on the deductive synthesis problem was slow for many years because of a lack of automated and scalable provers.
-However, a recent variant of the program synthesis problem has enabled dramatic progress (we refer to the special issue~\cite{fisman2022special} and the synthesis workshop\footnote{https://simons.berkeley.edu/workshops/tfcs2021-1} for an overview):
+However, a recent variant of the program synthesis problem has enabled dramatic progress (we refer to the special issue~\cite{fisman2022special} and the workshop\footnote{https://simons.berkeley.edu/workshops/tfcs2021-1} for an overview):
 %\footnote{https://www.cs.technion.ac.il/~shaull/SYNT2022/index.html}

### 2022-03-22 Notes/A reverse double negation translation/RDNT.tex
Added formula modification for propositional calculus

\documentclass[a4paper,12pt]{article}

\title{A reverse double negation translation}
\author{Alexander Pluska}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{ wasysym }
\usepackage{ stmaryrd }
\usepackage{ mathpartir }
\usepackage{bussproofs}

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}[section]
\theoremstyle{definition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{definition}
\newtheorem{example}[theorem]{Example}

\begin{document}

	\maketitle

	\begin{abstract}
		The aim of this paper is for a propositional sequent to give a transformed sequent that is valid in classical logic if and only if the original one is valid in intuitionistic logic. Since intuitionistic propositional logic is decidable this is of course trivially possible, however our procedure will be local. This means that our transformation will also be sound for first-order logic, allowing us to reduce the non-intuitionistic content of such sequents which yields benefits for program extraction in a situation where only a classical prover is available. 
	\end{abstract}

	\section{Porpositional logic}

	A sequent $A\Rightarrow B$ consists of two sets of formulas $A, B$. We will be using the following rules for our classical calculus:\\\\
	\begin{center}
	\begin{tabular}{lll}
		\AxiomC{\hphantom{x}}
		\RightLabel{Ax ($P$ atomic)}
		\UnaryInfC{$P,\Gamma\Rightarrow \Delta, P$}
		\DisplayProof&
		\AxiomC{\hphantom{x}}
		\RightLabel{L$\bot$}
		\UnaryInfC{$\bot,\Gamma\Rightarrow\Delta$}
		\DisplayProof&
		\\&&\\
		\AxiomC{$A, B,\Gamma\Rightarrow\Delta$}
		\RightLabel{L$\wedge$}
		\UnaryInfC{$A\wedge B, \Gamma\Rightarrow \Delta$}
		\DisplayProof&
		\AxiomC{$\Gamma\Rightarrow\Delta, A$}
		\AxiomC{$\Gamma\Rightarrow\Delta, B$}
		\RightLabel{R$\wedge$}
		\BinaryInfC{$\Gamma\Rightarrow \Delta, A\wedge B$}
		\DisplayProof&
		\\&&\\
		\AxiomC{$A, \Gamma\Rightarrow\Delta$}
		\AxiomC{$B, \Gamma\Rightarrow\Delta$}
		\RightLabel{L$\vee$}
		\BinaryInfC{$A\vee B, \Gamma\Rightarrow \Delta$}
		\DisplayProof&
		\AxiomC{$\Gamma\Rightarrow\Delta, A, B$}
		\RightLabel{R$\vee$}
		\UnaryInfC{$\Gamma\Rightarrow \Delta, A\vee B$}
		\DisplayProof&
		\\&&\\
		\AxiomC{$\Gamma\Rightarrow\Delta, A$}
		\AxiomC{$B, \Gamma\Rightarrow\Delta$}
		\RightLabel{L$\to$}
		\BinaryInfC{$A\to B, \Gamma\Rightarrow \Delta$}
		\DisplayProof&
		\AxiomC{$A,\Gamma\Rightarrow\Delta, B$}
		\RightLabel{R$\to$}
		\UnaryInfC{$\Gamma\Rightarrow \Delta, A\to B$}
		\DisplayProof&
		\\&&\\
	\end{tabular}
	\end{center}

	For the intuitionistic calculus we just need to modify R$\to$:\\
	\begin{center}
		\AxiomC{$A, \Gamma\Rightarrow B$}
		\RightLabel{R$\to'$}
		\UnaryInfC{$\Gamma\Rightarrow\Delta, A\to B$}
		\DisplayProof
	\end{center}

	\section{The Translation Procedure}

	\begin{definition}
		We say that two sequents $S, S'$ are \textit{equi-valid} if $S$ is provable in intuitionistic logic iff $S'$ is provable.
	\end{definition}

	\begin{example}
		The sequents $\Rightarrow A\vee B$ and $A\to C, B\to C\Rightarrow C$ are equi-valid:  By examining how a proof of the latter must end it is easy to argue that if there is a proof of one there must be a proof of the other, an alternative argument of course is that neither is provable.
	\end{example}

		Note that they are not equivalent in the typical sense however, due to the absence of $C$ in the former. The reason that equivalence is insufficient for our procedure is that new atoms may be introduced. In trying to reverse double negation translation this is to be expected however: In double negation translation we collapse the veracity of $A$ and $\neg\neg A$, which are not equivalent before the translation. In our case we do the opposite, i.e. we must modifiy our formula such that $A$ and $\neg\neg A$ are no longer classically equivalent. For that we require new atoms.

	\begin{definition}
		A sequent $A_1,\dots,A_n\Rightarrow B$ is \textit{implicational} if $B$ is atomic and each $A_i$ is of the form either $C\to (D\to E)$, $(C\to D)\to E$, $C\to D$ or $C$ for some atomic $C, D, E$.
	\end{definition}

	The first result is that for every sequent there is some equi-valid implicational sequent. This transformation is not strictly necessary for what follows, but it makes the arguments and procedure considerably simpler.

	Consider the following rules, where $X$ is not atomic, $N$ is a new atom:
	\begin{center}
		\begin{tabular}{lcl}
			$\Delta\Rightarrow \Gamma, A, B$&$\rightsquigarrow$&$\Delta, A\to N, B\to N\Rightarrow\Gamma, N$\\
			$\Delta\Rightarrow A\vee B$&$\rightsquigarrow$&$\Delta, A\to N, B\to N\Rightarrow N$\\
			$\Delta\Rightarrow A\wedge B$&$\rightsquigarrow$&$\Delta, A\to (B\to N)\Rightarrow N$\\
			$\Delta\Rightarrow A\to B$&$\rightsquigarrow$&$\Delta, A\Rightarrow B$\\

			$\Delta, A\wedge B\Rightarrow C$&$\rightsquigarrow$&$\Delta, A, B\Rightarrow C$\\
			$\Delta, A\vee B\Rightarrow C$&$\rightsquigarrow$&$\Delta, (A\to C)\to (B\to C)\to C\Rightarrow  C$\\
			$\Delta, (A\wedge B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to (B\to C)\Rightarrow D$\\
			$\Delta, (A\vee B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to C, B\to C\Rightarrow D$\\
			$\Delta, A\to (B\vee C)\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to (B\to D)\to (C\to D)\to D\Rightarrow D$\\
			$\Delta, A\to (B\wedge C)\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to B, A\to C\Rightarrow D$\\
			$\Delta, X\to (A\to B)\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, N\to (A\to B)\Rightarrow C$\\
			$\Delta, A\to (X\to B)\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, A\to (N\to B)\Rightarrow C$\\
			$\Delta, A\to (B\to X)\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, A\to (B\to N)\Rightarrow C$\\
			$\Delta, (X\to A)\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, (N\to A)\to B\Rightarrow C$\\
			$\Delta, (A\to X)\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, (A\to N)\to B\Rightarrow C$\\
			$\Delta, (A\to B)\to X\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, (A\to B)\to N\Rightarrow C$\\
		\end{tabular}
	\end{center}
	\pagebreak
	\begin{example}\hphantom{x}
	\begin{center}
		\begin{tabular}{ll}
			&$(A\to B)\to (C\vee D)\to E\Rightarrow F\vee G$\\
			$\rightsquigarrow$&$(A\to B)\to (C\vee D)\to E, F\to N, G\to N\Rightarrow N$\\
			$\rightsquigarrow$&$(A\to B)\to M, M\to (C\vee D)\to E, F\to N, G\to N\Rightarrow N$\\
			$\rightsquigarrow$&$(A\to B)\to M, M\to O\to E, (C\vee D)\to O, F\to N, G\to N\Rightarrow N$\\
			$\rightsquigarrow$&$(A\to B)\to M, M\to O\to E, C\to O, D\to O, F\to N, G\to N\Rightarrow N$\\
		\end{tabular}
	\end{center}
	\end{example}

	\begin{lemma}
		Iteratively applying the above transformations we can obtain from every sequent $S$ an equi-valid implicational sequent $S_\to$.
	\end{lemma}

	\begin{proof}[Proofsketch]

		Check that for any formula that is not implicational there is an applicable rule.

		Check that the sequents on both sides of each rule are equi-valid.

		Termination can be shown by well-foundedness: First the number of formulas in the succedent is reduced to $1$, then the succedent is reduced to an atom. Then in each step the number of disjunctions, conjunctions or the depth of nested implications is reduced. 
	\end{proof}

	For each atom $A$ consider a new atom $A'$ and a special atom $\omega = \bot'$. Have $\top' = \top$. Fix some implicational sequent $S := \Delta\Rightarrow C$ for now. Let $\mathcal{A}$ be the set of atoms occurring in $S$. Then define 
	\begin{align*}
		\Delta' = &\{A\to A', \omega\to A'\:|\:A\in\mathcal{A}\}
		\\&\cup \{(A'\to B')\to C\:|\: (A\to B)\to C\in\Delta\}
		\\&\cup \{A'\to (B'\to C'), A\to(B\to C)\:|\: A\to (B\to C)\in\Delta\}
		\\&\cup \{A'\to B', A\to B\:|\: A\to (B\to C)\in\Delta\}
	\end{align*}
and have $$S' := \Delta' \Rightarrow C$$

	\begin{lemma}
		An implicational sequent $S$ is intuitionistically provable if and only if $S'$ is classically provable
	\end{lemma}

	\begin{proof}[Proofsketch]

		$\Rightarrow$ the classical proof can be obtained by appropriate renaming of atoms and weakening. It will still be intuitionistically sound.

		$\Leftarrow$ Rename all variables $A'$ to $A$ and rewrite the R$\to$ rules. At the end apply a number of cuts to remove the $A\to A,\bot\to A$.

	\end{proof}

	\begin{theorem}
		A sequent $S$ in intuitionistically provable if and only if $S^\circ_\to$ is classically provable.
	\end{theorem}

	\begin{example}
		Consider Pierce's law $P := (A\to B)\to A\Rightarrow A$. This is a classic non-intuitionistic tautology. Note that it is already implicational. $P'$ is $$(A'\to B')\to A, A\to A', B\to B', \omega\to A', \omega\to B'\Rightarrow A$$ As proven this sequent is not classically valid, $A = B = B' = \omega = 0$, $A' = 1$  provides a countermodel.

		If we add $A\to B$ to the hypothesis set $P'$ is $$(A'\to B')\to A, A'\to B', A\to B, A\to A', B\to B', \omega\to A', \omega\to B'\Rightarrow A$$ which is classically and intuitionistically valid. Argueing in BHK terms, we can provide a terms of type $A'\to B'$ to $(A'\to B')\to A$ in order to obtain a term of type $A$, which is what is needed.
	\end{example}

	Intuitively the primed atoms may only be used in implications, i.e. as arguments to formulas of type $(A\to B)\to C$ but never as proper atoms, i.e. as arguments to $A\to B$.

	\bibliographystyle{plain}
	\bibliography{references}

\end{document}

### 2022-03-22 WWTF/main.tex
aasdf

@@
 \documentclass[a4paper,12pt]{article}
+
+\usepackage{url}
+
 \begin{document}
-\title{Leveraging The Progress in Automated Theorem Proving for Automated Program Synthesis\\\\
+\title{Leveraging The Progress in Automated Theorem Proving for Program Synthesis\\
 \small{WWTF ICT 2022}}
@@
 Given a logical formula that states that for all inputs there exists an output such that the required specification holds, one searches for a constructive proof of this formula and then extracts a program from this proof.
-Progress on the deductive synthesis problem was slow for many years because of a lack of automated and scalable provers.
-However, a recent variant of the program synthesis problem has enabled dramatic progress (we refer to the special issue~\cite{fisman2022special} and the workshop\footnote{https://simons.berkeley.edu/workshops/tfcs2021-1} for an overview):
+A classic route to program extraction is the \emph{Curry-Howard correspondence}~\cite{?} (also known as \emph{proofs-as-programs}) where one exploits the equivalence between proofs in intuitionist logic and functional programs in order to directly extract a functional program from an intuitionistic proof.
+The Curry-Howard correspondence denotes the fundamental relationship between two seemingly unrelated formalisms, namely proofs systems on one hand and computational models on the other hand.
+The Curry-Howard correspondence has been a major driving force for research spanning functional programming languages, type theory, proof systems and constructive logic.
+However, progress on the deductive synthesis problem has been slow for many years because of a lack of automated and scalable provers.
+Instead, a recent variant of the program synthesis problem has enabled dramatic progress (we refer to the special issue~\cite{fisman2022special} and the workshop\footnote{\url{https://simons.berkeley.edu/workshops/tfcs2021-1}} for an overview):
 %\footnote{https://www.cs.technion.ac.il/~shaull/SYNT2022/index.html}
@@
 While the reported progress is impressive, we believe that the traditional deductive approach to synthesis has been unjustly neglected.
-In this project we aim at connecting classical ideas of proofs-as-programs~\cite{?} with modern automated theorem provers, in particular first-order solvers (and SMT solvers???), which have made impressive progress over the recent years (as evidenced by yearly competitions).
-We believe this approach offers the following advantages:
-1) First-order solvers are particularly suitable to solving \emph{quantified} formulas as needed for synthesis.
-2) Proofs-as-programs represent a longstanding approach to program extraction for which a large body of theory has been developed.
-3) We are able to take advantage of the highly sophisticated and optimized proof search strategies of first-order solvers instead of needing to define and implement \emph{ad-hoc} search strategies.
 
@@
 
-gap in the literature
+The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{?}--, with modern automated theorem provers, in particular first-order solvers (and SMT solvers???), which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).
+
+We believe this approach offers the following advantages:
+1) First-order solvers are particularly suitable to solving \emph{quantified} formulas as needed for synthesis.
+2) Proofs-as-programs represents a longstanding approach to program extraction for which a large body of theory has been developed.
+That is to say, the groundwork for program extraction from proofs has been laid, we just need to adapt this theory to the specifics of modern automated theorem provers.
+3) We are able to take advantage of the highly sophisticated and optimized proof search strategies of first-order solvers instead of needing to define and implement \emph{ad-hoc} search strategies.
+Moreover, we will profit from any future advancements in automated theorem proving.
+
+We believe this research goal is very timely, because the improved hardware and the progress in automated theorem proving have finally made automated proof search feasible.
+Thus the proposed research promises to fill a longstanding gap in the literature.
+
+The main challenge consists in the fact that the proofs-as-programs approach to program synthesis requires proofs that are valid in intuitionistic logic, while state-of-the-art theorem provers
+utilize resolution which is an inherently classical technique.
+While the literature has identified
+We plan to address this challenge by adapting techniques from the literature as well as directions identified in our own preliminary research:
+1) The Curry-Howard correspondence can be extended to classical logic using control operators~\cite{Control1}, Parigot's $\lambda\mu$-Calculus~\cite{Parigot1} or CPS-translation~\cite{CPS1}.
+% However in this process we must in general give up on some desirable properties, at the very least realizability of $\vee$ and $\exists$, i.e.
+a proof of $\exists xP(x)$ will not necessarily give us an $x$ such that $P(x)$ and a proof of $A\vee B$ might give us neither a proof of $A$ nor a proof of $B$.
+2) Under certain conditions, classically valid proofs can be transformed into an intuitionistic proofs, e.g.
+proofs of $\Pi_2$ sentences in Peano Arithmetic~\cite{HAPA}, via Friedman's Translation~\cite{Friedman} or similar methods.
+%This approach has been examined in different model settings and would have to be adapted and extended for our general problem.
+
+3) Finally we suggest a third approach which has not been covered in literature yet, that is to pass a modified formula to the classical prover which can give us additional information about what an intuitionistic proof can look like.
+(Florian: unclear what is meant here)
+
+We plan to examine the following
 
@@
 Please note that the development of source code, systems or protocols should not be the main goal of the project; however, they are encouraged in the context of demonstrating the practical applicability of new scientific results.
+
+The final goal of our project is to leverage state-of-the-art first-order automated
+theorem provers, in particular Vampire, for program synthesis. 
+This is motivated by the recent and prospective advancements in their capabilities.
+As outlined we expect to obtain the necessary theoretical foundations as well
+as a practical system to leverage state-of-the-art classical theorem provers for
+program extraction. 
+At the current time no such system exists. 
+Much of the theoretical groundwork for implementing such a system is already present,
+but is scattered and mostly applied in theoretical model settings.
 

### 2022-03-22 WWTF/main.tex
asdf

@@
 A classic route to program extraction is the \emph{Curry-Howard correspondence}~\cite{?} (also known as \emph{proofs-as-programs}) where one exploits the equivalence between proofs in intuitionist logic and functional programs in order to directly extract a functional program from an intuitionistic proof.
-The Curry-Howard correspondence denotes the fundamental relationship between two seemingly unrelated formalisms, namely proofs systems on one hand and computational models on the other hand.
+The Curry-Howard correspondence denotes a fundamental relationship between two seemingly unrelated formalisms, namely proofs systems on one hand and computational models on the other hand.
 The Curry-Howard correspondence has been a major driving force for research spanning functional programming languages, type theory, proof systems and constructive logic.
@@
 %\footnote{https://www.cs.technion.ac.il/~shaull/SYNT2022/index.html}
-The logical specification is accompanied by a syntactic template that constrains the space of possible implementations.
+The logical specification is accompanied by a \emph{syntactic template} that constrains the space of possible implementations.
 The synthesis procedure then searches for/enumerates candidate implementations and either rejects these candidates by finding violating input/output pairs or accept the generated program by verifying their correctness.
@@
 1) First-order solvers are particularly suitable to solving \emph{quantified} formulas as needed for synthesis.
-2) Proofs-as-programs represents a longstanding approach to program extraction for which a large body of theory has been developed.
+2) Proofs-as-programs represents a longstanding approach to program extraction.
 That is to say, the groundwork for program extraction from proofs has been laid, we just need to adapt this theory to the specifics of modern automated theorem provers.
 3) We are able to take advantage of the highly sophisticated and optimized proof search strategies of first-order solvers instead of needing to define and implement \emph{ad-hoc} search strategies.
-Moreover, we will profit from any future advancements in automated theorem proving.
+%Moreover, we will profit from any future advancements in automated theorem proving.
 
-We believe this research goal is very timely, because the improved hardware and the progress in automated theorem proving have finally made automated proof search feasible.
-Thus the proposed research promises to fill a longstanding gap in the literature.
+We believe the proposed research is very timely, because improved hardware and the progress in automated theorem proving have finally made automated proof search feasible.
+Thus the proposed research promises to fill a longstanding gap in the state-of-the-art.
 
-The main challenge consists in the fact that the proofs-as-programs approach to program synthesis requires proofs that are valid in intuitionistic logic, while state-of-the-art theorem provers
+We will advance the above main goal by two work packages (WPs):
+
+(WP1) We will lay the theoretical groundwork for program extraction with automated theorem provers.
+The challenge here is that the proofs-as-programs approach requires proofs in intuitionistic logic, while state-of-the-art theorem provers
 utilize resolution which is an inherently classical technique.
-While the literature has identified
-We plan to address this challenge by adapting techniques from the literature as well as directions identified in our own preliminary research:
-1) The Curry-Howard correspondence can be extended to classical logic using control operators~\cite{Control1}, Parigot's $\lambda\mu$-Calculus~\cite{Parigot1} or CPS-translation~\cite{CPS1}.
-% However in this process we must in general give up on some desirable properties, at the very least realizability of $\vee$ and $\exists$, i.e.
-a proof of $\exists xP(x)$ will not necessarily give us an $x$ such that $P(x)$ and a proof of $A\vee B$ might give us neither a proof of $A$ nor a proof of $B$.
-2) Under certain conditions, classically valid proofs can be transformed into an intuitionistic proofs, e.g.
-proofs of $\Pi_2$ sentences in Peano Arithmetic~\cite{HAPA}, via Friedman's Translation~\cite{Friedman} or similar methods.
-%This approach has been examined in different model settings and would have to be adapted and extended for our general problem.
+For this, we will adapt and extend techniques from the literature that connect the Curry-Howard correspondence to classical logic.
 
-3) Finally we suggest a third approach which has not been covered in literature yet, that is to pass a modified formula to the classical prover which can give us additional information about what an intuitionistic proof can look like.
-(Florian: unclear what is meant here)
-
-We plan to examine the following
-
-input/output pairs, existing program fragments
+(WP2) We will implement a proof-of-concept tool that demonstrates the feasibility of our approach.
+This tool will take proofs produced by automated theorem provers as an input, and will output programs that are correct by construction.
+We will further experiment with techniques that relate to modern syntax-guided approaches to synthesis.
 
@@
 This is motivated by the recent and prospective advancements in their capabilities.
-As outlined we expect to obtain the necessary theoretical foundations as well
-as a practical system to leverage state-of-the-art classical theorem provers for
-program extraction. 
+We will lay the necessary theoretical foundations as well as implement a prototype system to leverage state-of-the-art classical theorem provers for program extraction. 
 At the current time no such system exists.
+
 Much of the theoretical groundwork for implementing such a system is already present,
-but is scattered and mostly applied in theoretical model settings.
+but is scattered and mostly limited to theoretical settings.
+We sketch here the two main directions that connect the Curry-Howard correspondence to classical logic:
+1) The Curry-Howard correspondence can be extended to classical logic using control operators~\cite{Control1}, Parigot's $\lambda\mu$-Calculus~\cite{Parigot1} or CPS-translation~\cite{CPS1}.
+However in this process we must in general give up on some desirable properties, at the very least the realizability of $\vee$ and $\exists$.
+2) Under certain conditions, classically valid proofs can be transformed into intuitionistic proofs, e.g.
+proofs of $\Pi_2$ sentences in Peano Arithmetic~\cite{HAPA}, via Friedman's Translation~\cite{Friedman} or similar methods.
+We will adapt and specialise these techniques from the literature to the proofs produced by automated theorem provers such as Vampire:
+A) We will develop the first\emph{ program extraction calculi} for \emph{resolution resp. superposition proofs}.
+In particular, we need to pay careful attentiont to the normal-form transformation --- a pre-processing step --- in automated theorem provers, which also involve equivalences that are only classically valid.
+C) We will identify assumptions under which our adaptions of the techniques 1) and 2) guarantee that such a program extraction is possible.
+We are not aware of related results in the literature.
+
+%This approach has been examined in different model settings and would have to be adapted and extended for our general problem.
 
@@
 Include a short assessment on the feasibility of the approach.
+
+We now give the details on (WP1):
+An optimal procedure for synthesis would consist of
+\begin{itemize}
+\item transforming the specification into an alternate form that would guarantee the use of as few non-intuitionistic rules as possible and applicability of Friedman translation in as many places as possible
+\item applying the prover to this modified specification obtaining a classical proof
+\item apply Friedman translation or similar translation procedures in as many places as possible
+\item extracting a program via Curry-Howard correspondence, applying control operators to non-intuitionistic remnants as necessary.
+\end{itemize}
+All of these elements except the initial modification of the specification already exist separately and in model settings such as arithmetic. 
+Extending these results and combining them will be the first major undertaking of this project. 
+This is also a necessary precondition for determining what initial modifications should be undertaken. 
+We expect that for most real-world synthesis problems our procedure will be sufficient to transform classical proofs into usable and verified programs.
+
+Finally we want to implement a real system on top of Vampire realizing the synthesis problem (WP2) which will target a suitable functional programming language.
+
+Finally we suggest a third approach which has not been covered in literature yet, that is to pass a modified formula to the classical prover which can give us additional information about what an intuitionistic proof can look like.
+
+We will investigate how to restrict proof search in order to steer the search towards programs from certain syntactic fragments, we will experiment with different axioms that can be given as input to the automated theorem provers and we will examine input/output pairs to the tool.
 

### 2022-03-27 WWTF/Draft Alexander/Proposal.tex
Formatting and packages in Proposal.tex
expected results, novelty and relevance

@@
-\documentclass[onehalfspacing]{article}
-
-\usepackage{amsmath,amssymb,amsthm}
-\usepackage{ wasysym }
-\usepackage{ stmaryrd }
-\usepackage{ mathpartir }
-\usepackage{xcolor}
-\usepackage{bussproofs}
-\usepackage[left=3.00cm, right=3.00cm, top=3.00cm,bottom=3.0cm]{geometry}
-\usepackage{pdflscape}
-\usepackage{xspace}
+\documentclass[a4paper,12pt]{article}
 
@@
 
-	As outlined we expect to obtain the necessary theoretical foundations as well as a practical system to leverage state-of-the-art classical theorem provers for program extraction.
-At the current time no such system exists.
-Much of the theoretical groundwork for implementing such a system is already present, but is scattered and mostly applied in  theoretical model settings.
+	We now give details on (WP1).
+As mentioned in the previous section there are currently two main approaches in the literature for obtaining programs from classical proofs: 1) Parigot's $\lambda\mu$-calculus~\cite{Parigot1} gives a direct translation of classical proofs with datatypes and is essentially ready for implementation.
+However, since non-intuitionistic tautologies like $\neg\forall xP(x)\to \exists x\neg P(x)$ or $A\vee \neg A$ are assigned a program, we lose the realizability of $\exists$ and $\vee$.
+As discussed in~\cite{practical}, in practice it is often possible to eliminate the non-intuitionistic fragments.
+We aim to establish stronger theoretical guarantees on when this is possible but also more practical benchmarks on the feasibility of this approach, even when realizability is not ensured by a theorem.
+2) Sometimes classical validity guarantees intuitionistic validity, most famously for $\Pi_2$-formulas in Arithmetic~\cite{Friedman}.
+There have been various generalizations of this result, most notably by Schwichtenberg's group~\cite{schwichtenberg}.
+We will adapt their work to our more general setting.
 
-	A successful completion of this project would lay the foundation for verified program synthesis via automated theorem provers.
+	We expect that in practice these approaches will be sufficient in many situations.
+However we also suggest a completely new direction: We aim for a translation between provers rather than proofs, so we examine what can be learned by attempting proofs of modified versions of the target formula.
+A hopeful sign is the reverse direction: given an intuitionistic prover we can establish classical provability via double negation translation.
+A first approximation has been outlined in~\cite{RDNT}.
+
+	Using this theoretical groundwork we will implement a synthesis system on top of Vampire for (WP2) targeting a suitable functional programming language.
+We will investigate how syntactic context and redundant
+	axioms can be used to restrict and guide proof search and we will examine the merit of our approach when only input/output pairs are given.
 

### 2022-03-27 WWTF/main.tex
expected results, novelty and relevance

@@
 \usepackage{url}
+\usepackage{xcolor}
 
@@
 Given a logical formula that states that for all inputs there exists an output such that the required specification holds, one searches for a constructive proof of this formula and then extracts a program from this proof.
-A classic route to program extraction is the \emph{Curry-Howard correspondence}~\cite{?} (also known as \emph{proofs-as-programs}) where one exploits the equivalence between proofs in intuitionist logic and functional programs in order to directly extract a functional program from an intuitionistic proof.
+A classic route to program extraction is the \emph{Curry-Howard correspondence}~\cite{?} {\color{red}Standardreferenz Proofs and Types~\cite{girard1989proofs} oder Original Curry~\cite{curry1934functionality}} (also known as \emph{proofs-as-programs}) where one exploits the equivalence between proofs in intuitionist logic and functional programs in order to directly extract a functional program from an intuitionistic proof.
 The Curry-Howard correspondence denotes a fundamental relationship between two seemingly unrelated formalisms, namely proofs systems on one hand and computational models on the other hand.
@@
 
-The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{?}--, with modern automated theorem provers, in particular first-order solvers (and SMT solvers???), which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).
+The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{?}{ \color{red}s.o.}--, with modern automated theorem provers, in particular first-order solvers (and SMT solvers???), which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).
 
@@
 That is to say, the groundwork for program extraction from proofs has been laid, we just need to adapt this theory to the specifics of modern automated theorem provers.
+{\color{red}in the classical case not everything has been said, perhaps "much groundwork..." and "...adapt and extend..."}
 3) We are able to take advantage of the highly sophisticated and optimized proof search strategies of first-order solvers instead of needing to define and implement \emph{ad-hoc} search strategies.
@@
 We sketch here the two main directions that connect the Curry-Howard correspondence to classical logic:
-1) The Curry-Howard correspondence can be extended to classical logic using control operators~\cite{Control1}, Parigot's $\lambda\mu$-Calculus~\cite{Parigot1} or CPS-translation~\cite{CPS1}.
+1) The Curry-Howard correspondence can be extended to classical logic using control operators~\cite{Control1}, Parigot's $\lambda\mu$-Calculus~\cite{Parigot1} or CPS-translation.
 However in this process we must in general give up on some desirable properties, at the very least the realizability of $\vee$ and $\exists$.
-2) Under certain conditions, classically valid proofs can be transformed into intuitionistic proofs, e.g.
-proofs of $\Pi_2$ sentences in Peano Arithmetic~\cite{HAPA}, via Friedman's Translation~\cite{Friedman} or similar methods.
+2) Under certain conditions, classically valid proofs can be transformed into intuitionistic proofs, e.g. proofs of $\Pi_2$-sentences in Peano Arithmetic~\cite{HAPA}.
 We will adapt and specialise these techniques from the literature to the proofs produced by automated theorem provers such as Vampire:
@@
 In particular, we need to pay careful attentiont to the normal-form transformation --- a pre-processing step --- in automated theorem provers, which also involve equivalences that are only classically valid.
-C) We will identify assumptions under which our adaptions of the techniques 1) and 2) guarantee that such a program extraction is possible.
+B) We will identify assumptions under which our adaptions of the techniques 1) and 2) guarantee that such a program extraction is possible.
 We are not aware of related results in the literature.

### 2022-03-28 WWTF/Draft Alexander/Proposal.tex
Reorganized sections
wording

@@
 
+	\section{Methods and feasibility}
+
 	We now give details on (WP1).
-As mentioned in the previous section there are currently two main approaches in the literature for obtaining programs from classical proofs: 1) Parigot's $\lambda\mu$-calculus~\cite{Parigot1} gives a direct translation of classical proofs with datatypes and is essentially ready for implementation.
-However, since non-intuitionistic tautologies like $\neg\forall xP(x)\to \exists x\neg P(x)$ or $A\vee \neg A$ are assigned a program, we lose the realizability of $\exists$ and $\vee$.
+We will consider two approaches from the literature for obtaining programs from classical proofs: 1) Parigot's $\lambda\mu$-calculus~\cite{Parigot1} gives a direct translation of classical proofs with datatypes and is essentially ready for implementation.
+However, since non-intuitionistic tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ are assigned a program, we lose the realizability of $\exists$ and $\vee$.
 As discussed in~\cite{practical}, in practice it is often possible to eliminate the non-intuitionistic fragments.
@@
 2) Sometimes classical validity guarantees intuitionistic validity, most famously for $\Pi_2$-formulas in Arithmetic~\cite{Friedman}.
-There have been various generalizations of this result, most notably by Schwichtenberg's group~\cite{schwichtenberg}.
+There have been various variations of this result, most notably by Schwichtenberg's group~\cite{schwichtenberg}.
 We will adapt their work to our more general setting.
@@
 
-	\section{Methods and feasibility}
-
 	\section{Potential to span across disciplines and to build bridges to application fields}

### 2022-03-28 WWTF/main.tex
asdf

@@
 Given a logical formula that states that for all inputs there exists an output such that the required specification holds, one searches for a constructive proof of this formula and then extracts a program from this proof.
-A classic route to program extraction is the \emph{Curry-Howard correspondence}~\cite{?} {\color{red}Standardreferenz Proofs and Types~\cite{girard1989proofs} oder Original Curry~\cite{curry1934functionality}} (also known as \emph{proofs-as-programs}) where one exploits the equivalence between proofs in intuitionist logic and functional programs in order to directly extract a functional program from an intuitionistic proof.
+A classic route to program extraction is the \emph{Curry-Howard correspondence}~\cite{girard1989proofs} (also known as \emph{proofs-as-programs}) where one exploits the equivalence between proofs in intuitionist logic and functional programs in order to directly extract a functional program from an intuitionistic proof.
 The Curry-Howard correspondence denotes a fundamental relationship between two seemingly unrelated formalisms, namely proofs systems on one hand and computational models on the other hand.
@@
 
-The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{?}{ \color{red}s.o.}--, with modern automated theorem provers, in particular first-order solvers (and SMT solvers???), which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).
+The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{girard1989proofs} --, with modern automated theorem provers, in particular first-order solvers (and SMT solvers???), which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).
 
@@
 2) Proofs-as-programs represents a longstanding approach to program extraction.
-That is to say, the groundwork for program extraction from proofs has been laid, we just need to adapt this theory to the specifics of modern automated theorem provers.
-{\color{red}in the classical case not everything has been said, perhaps "much groundwork..." and "...adapt and extend..."}
+That is to say, much groundwork for program extraction from proofs has been laid, we just need to adapt and extend this theory to the specifics of modern automated theorem provers.
 3) We are able to take advantage of the highly sophisticated and optimized proof search strategies of first-order solvers instead of needing to define and implement \emph{ad-hoc} search strategies.
@@
 
-We now give the details on (WP1):
-An optimal procedure for synthesis would consist of
-\begin{itemize}
-\item transforming the specification into an alternate form that would guarantee the use of as few non-intuitionistic rules as possible and applicability of Friedman translation in as many places as possible
-\item applying the prover to this modified specification obtaining a classical proof
-\item apply Friedman translation or similar translation procedures in as many places as possible
-\item extracting a program via Curry-Howard correspondence, applying control operators to non-intuitionistic remnants as necessary.
-\end{itemize}
-All of these elements except the initial modification of the specification already exist separately and in model settings such as arithmetic. 
-Extending these results and combining them will be the first major undertaking of this project. 
-This is also a necessary precondition for determining what initial modifications should be undertaken. 
-We expect that for most real-world synthesis problems our procedure will be sufficient to transform classical proofs into usable and verified programs.
+We now give details on (WP1).
+We will consider two approaches from the literature:
 
-Finally we want to implement a real system on top of Vampire realizing the synthesis problem (WP2) which will target a suitable functional programming language.
+1) Parigot's $\lambda\mu$-calculus~\cite{Parigot1} gives a direct translation of classical proofs with datatypes and is essentially ready for implementation.
+However, since non-intuitionistic tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ are assigned a program, we lose the realizability of $\exists$ and $\vee$.
+{\color{red} unclear: ``are assigned a program''}
+As discussed in~\cite{practical}, in practice it is often possible to eliminate the non-intuitionistic fragments.
+{\color{red} unclear: ``fragments''}
+We aim to establish stronger theoretical guarantees on when this is possible but also more practical benchmarks on the feasibility of this approach, even when realizability is not ensured by a theorem.
+{\color{red} unclear: ``more practical benchmarks''}
 
-Finally we suggest a third approach which has not been covered in literature yet, that is to pass a modified formula to the classical prover which can give us additional information about what an intuitionistic proof can look like.
+2) Sometimes classical validity guarantees intuitionistic validity, most famously for $\Pi_2$-formulas in Arithmetic~\cite{Friedman}.
+There have been various variations of this result, most notably by Schwichtenberg's group~\cite{schwichtenberg}.
+We will adapt their work to our more general setting.
+{\color{red} Can the previous sentence be made more concrete? Also why we this promising and feasiable?}
+%	
+%We expect that in practice these approaches will be sufficient in many situations.
 
-We will investigate how to restrict proof search in order to steer the search towards programs from certain syntactic fragments, we will experiment with different axioms that can be given as input to the automated theorem provers and we will examine input/output pairs to the tool.
+Third, we also investigate a completely new direction:
+We aim for a translation between provers rather than proofs. 
+E.g., we study under what conditions classic proofs of a modified input formula allow to extract intuitionistic proofs of the actual input.
+A promising example is actually the reverse direction:
+given an intuitionistic prover one can establish classical provability via double negation translation.
+%A first approximation has been outlined in~\cite{RDNT}.
+
+Using the results of (WP1) we will implement a synthesis system on top of Vampire (WP2), targeting a suitable functional programming language.
+We will experiment with various benchmark problems (e.g., median computation, sorting algorithms, arithmetic functions such as sqrt) and compare our results with the SyGuS-competition.
+We will use ideas from syntax-guided synthesis (sketched in 1) in order to restrict and guide proof search:
+We will experiment with different formulations of arithmetic axioms, induction axioms and redundant axioms.
+We will further examine if proof search is possible when only input/output pairs are given.
 
@@
 
+Programs that are correct by construction represent a longstanding goal of computer science.
+Correct Their is holds particularly true for safet
+
+\section{Potential ethical aspects (max. 500 characters)}
+%If required, information should be given with respect to the ethical approval.
+Is an ethical approval available that covers the prospective use of data and the planned analysis?
+If not, state whether an additional approval is required.
+
+Due to the focus on basic research and the mathematical and formal methodology, the proposed research does not entail ethical aspects beyond the usual mandate of good scientific practice.
+Similarly, the proposed research does not entail sex-specific or gender-related aspects.
+
+\section{Further Needed Stuff}
+
 Key references (max. 10 citations, max. 1250 characters)
@@
 
-\section{Potential ethical aspects (max. 500 characters)}
-%If required, information should be given with respect to the ethical approval.
-Is an ethical approval available that covers the prospective use of data and the planned analysis?
-If not, state whether an additional approval is required.
-
 \bibliographystyle{plain} % We choose the "plain" reference style

### 2022-03-28 WWTF/main.tex
abstract

@@
 
+%DIESEL: Deductive Synthesis Reloaded
+%DITTO: Deductive Synthesis with Automated Theorem Provers
+%LEGION: Leveraging The Progress in Automated Theorem Proving for Program Synthesis
+
 \begin{abstract}
@@
 
+  Program synthesis is a long-standing problem that crosses many areas of computer science and mathematics.
+  Program synthesis holds the promise to  increase the reliability of software systems (because the synthesized program are correct by construction) and for increasing programmer productivity by enabling programmers to focus on high-level ideas instead of low-level details.
+
+  %
+  Traditionally, program synthesis has been formalized as a problem in deductive theorem proving:
+  %
+  Given a logical formula that states that for all inputs there exists an output such that the required specification holds, one searches for a constructive proof of this formula and then extracts a program from this proof.
+
+  However, progress on deductive synthesis has been slow because of a lack of automated and scalable provers.
+  We believe that the impressive progress in automated theorem proving finally has made automated proof search feasible.
+
+  Thus, the main goal of our project is to leverage state-of-the-art first-order automated theorem provers (such as Vampire) for program synthesis, demonstrating the feasibility of deductive synthesis via automated theorem proving.
+
+  Specifically, we plan to 1) adapt and extend the theory on program extraction to the specifics of modern automated theorem provers, and to 2) implement a proof-of-concept tool that demonstrates the feasibility of our approach.
 \end{abstract}
@@
 
+%The project promises to make advancements in the design of correct of systems, automated theorem proving and proof theory.
+
+%The complexity of modern software systems, including in particular, \emph{safety-critical systems}, continues to grow.
+Traditionally quality assurance techniques rely on first building the desired system and then testing/verifying the system.
+%, which may become costly or even impossible for sufficiently complex systems.
+In contrast, \emph{correct-by-construction} approaches seek to develop a product that continues to be correct, from the initial design to the final product.
+Advances in deductive synthesis promise to increase the applicability of the correct-by-construction approach, leading to higher reliability in software development.
+
+Automated theorem proving has made impressive progress in recent years, demonstrated in yearly solver competitions. 
+We will submit our benchmarks to these competitions in order to interest solver developers in improving their tools wrt to the specific needs of deductive synthesis.
+%(formulas with quantifier alternation, support for induction, etc.).
+
+We expect that a successful prototype for program extraction with automated theorem provers will lead to a revival of proof-as-programs for automated deductive synthesis, inspiring more members of the community to investigate this approach.
+
 \section{Potential medium-term economic / societal benefits (max. 500 characters)}
@@
 
-Programs that are correct by construction represent a longstanding goal of computer science.
-Correct Their is holds particularly true for safet
+- As a correct-by-construction methodology, program synthesis will increase the reliability of important safety-critical systems such as avionics, cars, medical devices, etc.
+
+\noindent - Enabling programmers to focus on high-level ideas instead of low-level details will increase programmer productivity.
+
+\noindent - Non-expert/ end-user programmers are also expected to promise from automated synthesis by only needing to communicate their intent to the synthesis engine (instead of writing a whole program).
 

### 2022-03-28 WWTF/main.tex
incorporated remarks

@@
 \usepackage{xcolor}
+\usepackage[normalem]{ulem}
 
@@
 1) Parigot's $\lambda\mu$-calculus~\cite{Parigot1} gives a direct translation of classical proofs with datatypes and is essentially ready for implementation.
-However, since non-intuitionistic tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ are assigned a program, we lose the realizability of $\exists$ and $\vee$.
+However, since \sout{non-intuitionistic} {\color{blue}non-constructive} tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ \sout{are assigned a program} {\color{blue} are classically provable}, we lose the realizability of $\exists$ and $\vee$.
 {\color{red} unclear: ``are assigned a program''}
-As discussed in~\cite{practical}, in practice it is often possible to eliminate the non-intuitionistic fragments.
+As discussed in~\cite{practical}, in practice it is \sout{often} {\color{blue} sometimes} possible to eliminate the \sout{non-intuitionistic fragments} {\color{blue} non-constructive reasoning steps from proofs}.
 {\color{red} unclear: ``fragments''}
-We aim to establish stronger theoretical guarantees on when this is possible but also more practical benchmarks on the feasibility of this approach, even when realizability is not ensured by a theorem.
+We aim to establish stronger theoretical guarantees on when this is possible but also \sout{more practical benchmarks on the feasibility of} {\color{blue}evaluate} this approach, even when realizability is not ensured by a theorem.
 {\color{red} unclear: ``more practical benchmarks''}
 
-2) Sometimes classical validity guarantees intuitionistic validity, most famously for $\Pi_2$-formulas in Arithmetic~\cite{Friedman}.
-There have been various variations of this result, most notably by Schwichtenberg's group~\cite{schwichtenberg}.
-We will adapt their work to our more general setting.
+2) {\color{blue}Under certain conditions on the specification} \sout{Sometimes} classical validity guarantees intuitionistic validity, most famously for $\Pi_2$-formulas in Arithmetic~\cite{Friedman}.
+There have been \sout{various} {\color{blue} many} variations of this result, most notably by Schwichtenberg's group~\cite{schwichtenberg}.
+\sout{We will adapt their work to our more general setting.} {\color{blue} We will collect, adapt and extend them with regard to our concrete setting.
+This approach is particularly promising when combined with the following, which allows us to consider altered specifications.}
 {\color{red} Can the previous sentence be made more concrete? Also why we this promising and feasiable?}

### 2022-03-29 WWTF/main.tex
asdf

--- 2022-03-28 WWTF/main.tex
@@
 1) Parigot's $\lambda\mu$-calculus~\cite{Parigot1} gives a direct translation of classical proofs with datatypes and is essentially ready for implementation.
-However, since non-intuitionistic tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ are assigned a program, we lose the realizability of $\exists$ and $\vee$.
-{\color{red} unclear: ``are assigned a program''}
-As discussed in~\cite{practical}, in practice it is often possible to eliminate the non-intuitionistic fragments.
-{\color{red} unclear: ``fragments''}
-We aim to establish stronger theoretical guarantees on when this is possible but also more practical benchmarks on the feasibility of this approach, even when realizability is not ensured by a theorem.
-{\color{red} unclear: ``more practical benchmarks''}
+However, since non-constructive tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ {\color{blue} are classically provable}, we lose the realizability of $\exists$ and $\vee$.
+As discussed in~\cite{practical}, in practice it is sometimes possible to eliminate the non-constructive reasoning steps from proofs.
+We aim to establish stronger theoretical guarantees on when this is possible but also evaluate this approach when realizability is not ensured by a theorem.
 
 2) Sometimes classical validity guarantees intuitionistic validity, most famously for $\Pi_2$-formulas in Arithmetic~\cite{Friedman}.
-There have been various variations of this result, most notably by Schwichtenberg's group~\cite{schwichtenberg}.
-We will adapt their work to our more general setting.
-{\color{red} Can the previous sentence be made more concrete? Also why we this promising and feasiable?}
+There have been many variations of this result, most notably by Schwichtenberg's group~\cite{schwichtenberg}.
+We will collect, adapt and extend these results with regard to our concrete setting. 
+This is particularly promising when considering altered specifications (see next point).
 %	
@@
 
-Third, we also investigate a completely new direction:
+Third, we investigate a new direction:
 We aim for a translation between provers rather than proofs.
-E.g., we study under what conditions classic proofs of a modified input formula allow to extract intuitionistic proofs of the actual input.
-A promising example is actually the reverse direction:
+E.g., we study under what conditions classical proofs of a modified input formula allow to extract intuitionistic proofs of the actual input.
+A motivating example is actually the reverse direction:
 given an intuitionistic prover one can establish classical provability via double negation translation.

### 2022-03-30 Notes/A reverse double negation translation/RDNT.tex
added detail RDNT

@@
 
-	A sequent $A\Rightarrow B$ consists of two sets of formulas $A, B$. We will be using the following rules for our classical calculus:\\\\
+	A sequent $A\Rightarrow B$ consists of two multi-sets of formulas $A, B$. We will be using the following rules for our classical calculus:\\
 	\begin{center}
@@
 		\AxiomC{\hphantom{x}}
-		\RightLabel{Ax ($P$ atomic)}
-		\UnaryInfC{$P,\Gamma\Rightarrow \Delta, P$}
+		\RightLabel{Ax}
+		\UnaryInfC{$A,\Gamma\Rightarrow \Delta, A$}
 		\DisplayProof&
@@
 		\\&&\\
-		\AxiomC{$\Gamma\Rightarrow\Delta, A$}
+		\AxiomC{$A\to B, \Gamma\Rightarrow\Delta, A$}
 		\AxiomC{$B, \Gamma\Rightarrow\Delta$}
@@
 
-	For the intuitionistic calculus we just need to modify R$\to$:\\
+	For the intuitionistic calculus we just need to modify R$\to$:
 	\begin{center}
@@
 
+	These calculi have the following nice property:
+
+	\begin{theorem}[subformula property]
+		Any formula that appears in a derivation is a subformula of a forumla that occurs in the root sequent.
+	\end{theorem}
+
+	We consider the following additional rules, which are not a priori part of our calculus:
+		\begin{center}
+		\begin{tabular}{lll}
+			\AxiomC{$\Gamma\Rightarrow\Delta$}
+			\RightLabel{Lweak}
+			\UnaryInfC{$A,\Gamma\Rightarrow \Delta$}
+			\DisplayProof&
+			\AxiomC{$\Gamma\Rightarrow\Delta$}
+			\RightLabel{Rweak}
+			\UnaryInfC{$\Gamma\Rightarrow \Delta, A$}
+			\DisplayProof&
+			\\&&\\
+			\AxiomC{$A, A,\Gamma\Rightarrow\Delta$}
+			\RightLabel{Lcontr}
+			\UnaryInfC{$A, \Gamma\Rightarrow \Delta$}
+			\DisplayProof&
+			\AxiomC{$\Gamma\Rightarrow\Delta, A, A$}
+			\RightLabel{Rcontr}
+			\UnaryInfC{$\Gamma\Rightarrow \Delta, A$}
+			\DisplayProof&
+			\\&&\\
+		\end{tabular}
+
+		\AxiomC{$\Gamma\Rightarrow A, \Delta$}
+		\AxiomC{$\Gamma', A\Rightarrow \Delta'$}
+		\RightLabel{Cut}
+		\BinaryInfC{$\Gamma, \Gamma'\Rightarrow\Delta, \Delta'$}
+		\DisplayProof
+	\end{center}
+
+	The following central result of proof theory will be useful to us in many places:
+	\begin{theorem}[admissibility of cut (and structural rules)]
+		The 5 above rules are admissible in our calculi for classical and intuitionistic logic, i.e. adding them we are not able to prove any additional theorems.
+	\end{theorem}
+
+	This means that we may assume that all proofs are free of these rules, but in constructing new proofs we may apply them.
+We exclude them from our calculus to keep it simple for proofs and because cut destroys the subformula property.
+
 	\section{The Translation Procedure}
@@
 	\begin{definition}
-		We say that two sequents $S, S'$ are \textit{equi-valid} if $S$ is provable in intuitionistic logic iff $S'$ is provable.
+		We say that two sequents $S, S'$ are \textit{equi-provable} if $S$ is provable iff $S'$ is provable in intuitionistic logic.
 	\end{definition}
 
+	Clearly equi-provability is an equivalence relation.
+
 	\begin{example}
-		The sequents $\Rightarrow A\vee B$ and $A\to C, B\to C\Rightarrow C$ are equi-valid:  By examining how a proof of the latter must end it is easy to argue that if there is a proof of one there must be a proof of the other, an alternative argument of course is that neither is provable.
+		The sequents $\Rightarrow A\vee B$ and $A\to C, B\to C\Rightarrow C$ are equi-provable. Note that they are not equivalent in the typical sense however, due to the absence of $C$ in the former. 
 	\end{example}
 
-		Note that they are not equivalent in the typical sense however, due to the absence of $C$ in the former.
 		The reason that equivalence is insufficient for our procedure is that new atoms may be introduced.
-In trying to reverse double negation translation this is to be expected however: In double negation translation we collapse the veracity of $A$ and $\neg\neg A$, which are not equivalent before the translation.
+In trying to reverse double negation translation this is to be expected however: In double negation translation we collapse the veracity of $A$ and $\neg\neg A$, which are not intuitionistically equivalent.
 In our case we do the opposite, i.e.
@@
 For that we require new atoms.
+
+		However there is clearly some collection between the two notions:
+
+		\begin{lemma}
+			Let $A$ and $B$ be equivalent formulas, i.e.
+$A\Rightarrow B$ and $B\Rightarrow A$ are provable, then for any $\Gamma, \Delta$ the sequents $A,\Gamma\Rightarrow\Delta$ and $B, \Gamma\Rightarrow\Delta$ as well as $\Gamma\Rightarrow\Delta, A$ and $\Gamma\Rightarrow\Delta, B$ are equi-provable.
+		\end{lemma}
+
+		\begin{proof}
+			This is a direct consequence of the admissibility of cut, e.g.
+			\begin{center}
+				\AxiomC{$\vdots$}
+				\noLine
+				\UnaryInfC{$A,\Gamma\Rightarrow\Delta$}
+				\AxiomC{$\vdots$}
+				\noLine
+				\UnaryInfC{$B\Rightarrow A$}
+				\RightLabel{Cut}
+				\BinaryInfC{$B, \Gamma\Rightarrow\Delta$}
+				\DisplayProof
+			\end{center}
+		\end{proof}
+
+		We observe some other basic properties of equiprovability.
+
+	\begin{lemma}
+		Let $N$ be a atomic proposition that does not occur in $\Gamma$ or $\Delta$, $A, B$ arbitrary formulas that do not contain $N$, then:
+		\begin{enumerate}
+			\item $\Gamma\Rightarrow\Delta$ and $\Gamma\Rightarrow\Delta, N$ are equi-provable.
+			\item $\Gamma\Rightarrow\Delta, A$ and $\Gamma, A\to N\Rightarrow \Delta, N$ are equi-provable.
+		\end{enumerate} 
+	\end{lemma}
+
+	\begin{proof}
+		1.
+If $\Gamma\Rightarrow\Delta$ is provable then we obtain a proof of $\Gamma\Rightarrow\Delta, N$ through a simple application of weakening.
+Suppose on the other hand we have a proof of  $\Gamma\Rightarrow\Delta, N$.
+Then by the subformula property $N$ may only occur as a pure atom.
+In particular $N$ can never appear in the antecedent as it would occur in $\Gamma$ then.
+Since we may assume that our proof does not contain weakenings, $N$ has to be introduced as part of $\Delta$ in the axioms.
+Therefore removing $N$ from all succedents produces a valid proof of $\Gamma\Rightarrow\Delta$.
+
+		2. If $\Gamma\Rightarrow\Delta, A$ is provable we obtain $\Gamma, A\to N\Rightarrow \Delta, N$ through a simple application of cut:
+		\begin{center}
+			\AxiomC{$\vdots$}
+			\noLine
+			\UnaryInfC{$\Gamma\Rightarrow\Delta, A$}
+			\AxiomC{\hphantom{X}}
+			\RightLabel{Ax}
+			\UnaryInfC{$A, A\to N\Rightarrow P, A$}
+			\AxiomC{\hphantom{X}}
+			\RightLabel{Ax}
+			\UnaryInfC{$N, A\Rightarrow N$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$A, A\to N\Rightarrow N$}
+			\RightLabel{Cut}
+			\BinaryInfC{$\Gamma, A\to N\Rightarrow\Delta, N$}
+			\DisplayProof
+		\end{center}
+		Suppose on the other hand we have a proof of $\Gamma, A\to N\Rightarrow \Delta, N$.
+Consider the proof obtained by replacing every occurrence of $N$ with $A$.
+By induction on the proof height it can be checked that this constitutes a valid proof of $\Gamma, A\to A\Rightarrow\Delta, N$.
+We obtain $\Gamma\Rightarrow\Delta, A$ through an application fo Cut:
+		\begin{center}
+			\AxiomC{\hphantom{X}}
+			\RightLabel{Ax}
+			\UnaryInfC{$A, \Rightarrow A$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$\Rightarrow A\to A$}
+			\AxiomC{$\vdots$}
+			\noLine
+			\UnaryInfC{$\Gamma, A\to A\Rightarrow\Delta, A$}
+			\RightLabel{Cut}
+			\BinaryInfC{$\Gamma\Rightarrow\Delta, A$}
+			\DisplayProof
+		\end{center}
+	\end{proof}
 
@@
 
-	The first result is that for every sequent there is some equi-valid implicational sequent.
+	The first result is that for every sequent there is some equi-provable implicational sequent.
 This transformation is not strictly necessary for what follows, but it makes the arguments and procedure considerably simpler.
+It is reminiscent of the definitions of connectors in second order logic.
 
@@
 		\begin{tabular}{lcl}
-			$\Delta\Rightarrow \Gamma, A, B$&$\rightsquigarrow$&$\Delta, A\to N, B\to N\Rightarrow\Gamma, N$\\
-			$\Delta\Rightarrow A\vee B$&$\rightsquigarrow$&$\Delta, A\to N, B\to N\Rightarrow N$\\
-			$\Delta\Rightarrow A\wedge B$&$\rightsquigarrow$&$\Delta, A\to (B\to N)\Rightarrow N$\\
+			$\Delta\Rightarrow \Gamma, A, B$&$\rightsquigarrow$&$\Delta\Rightarrow\Gamma, A\vee B$\\
+			$\Delta\Rightarrow A\vee B$&$\rightsquigarrow$&$\Delta, (A\vee B)\to N\Rightarrow N$\\
+			$\Delta\Rightarrow A\wedge B$&$\rightsquigarrow$&$\Delta,(A\wedge B)\to N\Rightarrow N$\\
 			$\Delta\Rightarrow A\to B$&$\rightsquigarrow$&$\Delta, A\Rightarrow B$\\
@@
 	\end{center}
-	\pagebreak
+
 	\begin{example}\hphantom{x}
@@
 	\begin{lemma}
-		Iteratively applying the above transformations we can obtain from every sequent $S$ an equi-valid implicational sequent $S_\to$.
+		Iteratively applying the above transformations we can obtain from every sequent $S$ an equi-provable implicational sequent $S_\to$.
 	\end{lemma}
@@
 
-		Check that the sequents on both sides of each rule are equi-valid.
+		Check that the sequents on both sides of each rule are equi-provable.
 

### 2022-03-31 WWTF/main.tex
if life gives you lemons
Incorporated SMT

@@
 \begin{document}
-\title{Leveraging The Progress in Automated Theorem Proving for Program Synthesis\\
-\small{WWTF ICT 2022}}
+\title{LEMONADE: Leveraging Modern Automated Deduction for Program Synthesis\\
+\vspace{0.5cm}
+\large{WWTF ICT 2022}}
 \author{Florian Zuleger and Alexander Pluska}
@@
 %LEGION: Leveraging The Progress in Automated Theorem Proving for Program Synthesis
+%PROGRESS: Program Synthesis via (oder despite? :)) Resolution
+%LEMONADE: Leveraging Modern Automated Deduction for Program Synthesis
 
@@
 
-The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{girard1989proofs} --, with modern automated theorem provers, in particular first-order solvers (and SMT solvers???), which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).
+The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{girard1989proofs} --, with modern automated theorem provers, in particular first-order solvers \sout{(}and SMT solvers\sout{???)}, which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).
 
@@
 However in this process we must in general give up on some desirable properties, at the very least the realizability of $\vee$ and $\exists$.
-2) Under certain conditions, classically valid proofs can be transformed into intuitionistic proofs, e.g. proofs of $\Pi_2$-sentences in Peano Arithmetic~\cite{HAPA}.
+2) Under certain conditions, classically valid proofs can be transformed into intuitionistic proofs, e.g.
+proofs of $\Pi_2$-sentences in Peano Arithmetic~\cite{HAPA}, {\color{blue}this point is of particular interest for SMT}.
 We will adapt and specialise these techniques from the literature to the proofs produced by automated theorem provers such as Vampire:
@@
 1) Parigot's $\lambda\mu$-calculus~\cite{Parigot1} gives a direct translation of classical proofs with datatypes and is essentially ready for implementation.
-However, since non-constructive tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ {\color{blue} are classically provable}, we lose the realizability of $\exists$ and $\vee$.
+However, since non-constructive tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ are classically provable, we lose the realizability of $\exists$ and $\vee$.
 As discussed in~\cite{practical}, in practice it is sometimes possible to eliminate the non-constructive reasoning steps from proofs.
@@
 We will collect, adapt and extend these results with regard to our concrete setting. 
-This is particularly promising when considering altered specifications (see next point).
-%	
+This is particularly promising when  {\color{blue} applied to SMT, which can be interpreted as a restricted subproblem of the full first-order case~\cite{reynolds2015counterexample}, or when} considering altered specifications (see next point).
 %We expect that in practice these approaches will be sufficient in many situations.

### 2022-03-31 WWTF/main.tex
asdf

@@
 
-The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{girard1989proofs} --, with modern automated theorem provers, in particular first-order solvers \sout{(}and SMT solvers\sout{???)}, which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).
+The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{girard1989proofs} --, with modern automated theorem provers, in particular first-order solvers (including SMT solvers), which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).
 
@@
 2) Under certain conditions, classically valid proofs can be transformed into intuitionistic proofs, e.g.
-proofs of $\Pi_2$-sentences in Peano Arithmetic~\cite{HAPA}, {\color{blue}this point is of particular interest for SMT}.
+proofs of $\Pi_2$-sentences in Peano Arithmetic~\cite{HAPA} (this point is of particular interest for SMT).
 We will adapt and specialise these techniques from the literature to the proofs produced by automated theorem provers such as Vampire:
@@
 We will collect, adapt and extend these results with regard to our concrete setting.
-This is particularly promising when  {\color{blue} applied to SMT, which can be interpreted as a restricted subproblem of the full first-order case~\cite{reynolds2015counterexample}, or when} considering altered specifications (see next point).
+This is particularly promising when applied to SMT, which can be interpreted as a restricted subproblem of the full first-order case~\cite{reynolds2015counterexample}, or when considering altered specifications (see next point).
 %We expect that in practice these approaches will be sufficient in many situations.

### 2022-04-03 WWTF/main.tex
asdf

@@
 1) Parigot's $\lambda\mu$-calculus~\cite{Parigot1} gives a direct translation of classical proofs with datatypes and is essentially ready for implementation.
-However, since non-constructive tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ are classically provable, we lose the realizability of $\exists$ and $\vee$.
+Unfortunately, we loose the realizability of $\exists$ and $\vee$, since non-constructive tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ are classically provable.
 As discussed in~\cite{practical}, in practice it is sometimes possible to eliminate the non-constructive reasoning steps from proofs.

### 2022-04-04 WWTF/main.tex
asdf

@@
 
-The main goal of this project is to connect the classical approach to deductive synthesis --  proofs-as-programs~\cite{girard1989proofs} --, with modern automated theorem provers, in particular first-order solvers (including SMT solvers), which have made impressive progress over the recent years (as evidenced by yearly competitions\footnote{\url{http://www.tptp.org/CASC/}}).
+The main goal of this project is to connect the classical approach to deductive synthesis --- proofs-as-programs~\cite{girard1989proofs} ---, with modern automated theorem provers, in particular first-order solvers (including SMT solvers), which have made impressive progress over the recent years (as evidenced by yearly competitions).
+%\footnote{\url{http://www.tptp.org/CASC/}})
 
-We believe this approach offers the following advantages:
+We believe that this approach offers the following advantages:
 1) First-order solvers are particularly suitable to solving \emph{quantified} formulas as needed for synthesis.
@@
 We believe the proposed research is very timely, because improved hardware and the progress in automated theorem proving have finally made automated proof search feasible.
-Thus the proposed research promises to fill a longstanding gap in the state-of-the-art.
-
 We will advance the above main goal by two work packages (WPs):
@@
 At the current time no such system exists.
+Thus the proposed research promises to fill a longstanding gap in the state-of-the-art.
 
@@
 
-We now give details on (WP1).
-We will consider two approaches from the literature:
+We now give details on (WP1):
+%We will consider two approaches from the literature:
 
@@
 Unfortunately, we loose the realizability of $\exists$ and $\vee$, since non-constructive tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ are classically provable.
-As discussed in~\cite{practical}, in practice it is sometimes possible to eliminate the non-constructive reasoning steps from proofs.
+In practice it is sometimes possible to eliminate the non-constructive reasoning steps from proofs~\cite{practical}.
 We aim to establish stronger theoretical guarantees on when this is possible but also evaluate this approach when realizability is not ensured by a theorem.
@@
 We will collect, adapt and extend these results with regard to our concrete setting.
-This is particularly promising when applied to SMT, which can be interpreted as a restricted subproblem of the full first-order case~\cite{reynolds2015counterexample}, or when considering altered specifications (see next point).
+This is particularly promising when applied to SMT, interpreted as a restricted subproblem of the full first-order case~\cite{reynolds2015counterexample}, or when considering altered specifications (see next point).
 %We expect that in practice these approaches will be sufficient in many situations.
 
-Third, we investigate a new direction:
+3) We investigate a new direction:
 We aim for a translation between provers rather than proofs.
@@
 Using the results of (WP1) we will implement a synthesis system on top of Vampire (WP2), targeting a suitable functional programming language.
-We will experiment with various benchmark problems (e.g., median computation, sorting algorithms, arithmetic functions such as sqrt) and compare our results with the SyGuS-competition.
-We will use ideas from syntax-guided synthesis (sketched in 1) in order to restrict and guide proof search:
+We will experiment with various benchmark problems %(e.g., median computation, sorting algorithms, arithmetic functions such as sqrt) 
+and compare our results with the SyGuS-competition.
+We will use ideas from syntax-guided synthesis in order to restrict and guide proof search:
 We will experiment with different formulations of arithmetic axioms, induction axioms and redundant axioms.
@@
 
+%[1] Ulrich Berger, Wilfried Buchholz, and Helmut Schwichtenberg. Refined
+%program extraction from classical proofs. 2002.
+%
+%[2] Dana Fisman, Rishabh Singh, and Armando Solar-Lezama. Special issue
+%on syntax-guided synthesis preface. 2022.
+%
+%[3] Harvey Friedman. Classically and intuitionistically provably recursive
+%functions. 1978.
+%
+%[4] Jean-Yves Girard, Paul Taylor, and Yves Lafont. Proofs and types. 1989.
+%
+%[5] Timothy G Griffin. A formulae-as-type notion of control. 1989.
+%
+%[6] Ulrich Kohlenbach. The Friedman A-translation. 2008.
+%
+%[7] Yevgeniy Makarov. Practical program extraction from classical proofs.  2006.
+%
+%[8] Zohar Manna and Richard Waldinger. A deductive approach to program
+%synthesis. 1979.
+%
+%[9] Michel Parigot. λμ-calculus: an algorithmic interpretation of classical
+%natural deduction. 1992.
+%
+%[10] Andrew Reynolds, Morgan Deters, Viktor Kuncak, Cesare Tinelli, and
+%Clark Barrett. Counterexample-guided quantifier instantiation for synthesis
+%in SMT. 2015.
+
 \end{document}

### 2022-04-05 WWTF/main.tex

@@
 Given a logical formula that states that for all inputs there exists an output such that the required specification holds, one searches for a constructive proof of this formula and then extracts a program from this proof.
-A classic route to program extraction is the \emph{Curry-Howard correspondence}~\cite{girard1989proofs} (also known as \emph{proofs-as-programs}) where one exploits the equivalence between proofs in intuitionist logic and functional programs in order to directly extract a functional program from an intuitionistic proof.
+A classic route to program extraction is the \emph{Curry-Howard correspondence}~\cite{girard1989proofs} (also known as \emph{proofs-as-programs}) where one exploits the equivalence between proofs in intuitionistic logic and functional programs in order to directly extract a functional program from an intuitionistic proof.
 The Curry-Howard correspondence denotes a fundamental relationship between two seemingly unrelated formalisms, namely proofs systems on one hand and computational models on the other hand.
@@
 The logical specification is accompanied by a \emph{syntactic template} that constrains the space of possible implementations.
-The synthesis procedure then searches for/enumerates candidate implementations and either rejects these candidates by finding violating input/output pairs or accept the generated program by verifying their correctness.
+The synthesis procedure then searches for/enumerates candidate implementations and either rejects these candidates by finding violating input/output pairs or accepts the generated program by verifying their correctness.
 The syntactic approach is very versatile and synthesis approaches have been explored that solely start from input/output examples or that mine possible program fragments from public code repositories.
@@
 (WP2) We will implement a proof-of-concept tool that demonstrates the feasibility of our approach.
-This tool will take proofs produced by automated theorem provers as an input, and will output programs that are correct by construction.
+This tool will take first-order specifications as an input, and leverage automated theorem provers to output programs that are correct by construction.
 We will further experiment with techniques that relate to modern syntax-guided approaches to synthesis.
@@
 
-The final goal of our project is to leverage state-of-the-art first-order automated
-theorem provers, in particular Vampire, for program synthesis.
+The final goal of our project is to leverage state-of-the-art first-order automated theorem provers, in particular Vampire~\cite{conf/cav/KovacsV13}, for program synthesis.
 This is motivated by the recent and prospective advancements in their capabilities.
@@
 1) Parigot's $\lambda\mu$-calculus~\cite{Parigot1} gives a direct translation of classical proofs with datatypes and is essentially ready for implementation.
-Unfortunately, we loose the realizability of $\exists$ and $\vee$, since non-constructive tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ are classically provable.
+Unfortunately, we lose the realizability of $\exists$ and $\vee$ since non-constructive tautologies like $\neg\forall xA(x)\to \exists x\neg A(x)$ or $A\vee \neg A$ are classically provable.
 In practice it is sometimes possible to eliminate the non-constructive reasoning steps from proofs~\cite{practical}.

### 2022-04-07 Notes/A reverse double negation translation/RDNT.tex
Full Proof Lemma 2.7

@@
 \usepackage{bussproofs}
+
+\usepackage{hyperref}% http://ctan.org/pkg/hyperref
+\usepackage{cleveref}% http://ctan.org/pkg/cleveref
 
@@
 
+\DeclareMathOperator*{\argmax}{arg\,max}
+
 \begin{document}
@@
 
-	A sequent $A\Rightarrow B$ consists of two multi-sets of formulas $A, B$. We will be using the following rules for our classical calculus:\\
+	A sequent $A\Rightarrow B$ consists of two finite multi-sets of formulas $A, B$. We will be using the following rules for our classical calculus:\\
 	\begin{center}
@@
 	\begin{definition}
-		We say that two sequents $S, S'$ are \textit{equi-provable} if $S$ is provable iff $S'$ is provable in intuitionistic logic.
+		We say that two sequents $S, S'$ are \textit{equiprovable} if $S$ is intuitionistically provable iff $S'$ is intuitionistically provable and write $$S\sim_p S'.$$
 	\end{definition}
 
-	Clearly equi-provability is an equivalence relation.
+	Clearly equiprovability is an equivalence relation.
+Since every sequent is either provable or not the above notion is trivial in a sense.
+However we can often argue that two sequents are equiprovable even when we cannot determine their individual provability.
+This is in particular useful when discussing transformations of sequents that are to preserve intuitionistic provability.
 
 	\begin{example}
-		The sequents $\Rightarrow A\vee B$ and $A\to C, B\to C\Rightarrow C$ are equi-provable. Note that they are not equivalent in the typical sense however, due to the absence of $C$ in the former. 
+		The sequents $\Rightarrow A\vee B$ and $A\to C, B\to C\Rightarrow C$ are equiprovable. Note that they are not equivalent in the typical sense however, due to the absence of $C$ in the former. 
 	\end{example}
@@
 
-		However there is clearly some collection between the two notions:
+		However there is clearly some connection between the two notions:
 
-		\begin{lemma}
-			Let $A$ and $B$ be equivalent formulas, i.e.
-$A\Rightarrow B$ and $B\Rightarrow A$ are provable, then for any $\Gamma, \Delta$ the sequents $A,\Gamma\Rightarrow\Delta$ and $B, \Gamma\Rightarrow\Delta$ as well as $\Gamma\Rightarrow\Delta, A$ and $\Gamma\Rightarrow\Delta, B$ are equi-provable.
+		\begin{lemma}\label{ep1}
+			Let $A$ and $B$ be equivalent formulas, i.e. $A\Rightarrow B$ and $B\Rightarrow A$ are provable. Then \begin{enumerate}
+				\item $A,\Gamma\Rightarrow\Delta \sim_p B, \Gamma\Rightarrow\Delta$
+				\item $\Gamma\Rightarrow\Delta, A\sim_p\Gamma\Rightarrow\Delta, B$
+			\end{enumerate}
 		\end{lemma}
@@
 
-	\begin{lemma}
+	\begin{lemma}\label{ep2}
 		Let $N$ be a atomic proposition that does not occur in $\Gamma$ or $\Delta$, $A, B$ arbitrary formulas that do not contain $N$, then:
 		\begin{enumerate}
-			\item $\Gamma\Rightarrow\Delta$ and $\Gamma\Rightarrow\Delta, N$ are equi-provable.
-			\item $\Gamma\Rightarrow\Delta, A$ and $\Gamma, A\to N\Rightarrow \Delta, N$ are equi-provable.
+			\item $\Gamma\Rightarrow\Delta\sim_p\Gamma\Rightarrow\Delta, N$
+			\item $\Gamma\Rightarrow\Delta, A\sim_p\Gamma, A\to N\Rightarrow \Delta, N$
+			\item $\Gamma, A, B\Rightarrow\Delta\sim_p\Gamma, A\wedge B\Rightarrow\Delta$
+			\item $\Gamma\Rightarrow A\vee B,\Delta\sim_p\Gamma\Rightarrow A, B,\Delta$
 		\end{enumerate} 
@@
 			\RightLabel{Ax}
-			\UnaryInfC{$A, \Rightarrow A$}
+			\UnaryInfC{$A \Rightarrow A$}
 			\RightLabel{R$\to$}
@@
 		\end{center}
+
+		3. Going from $\Gamma, A, B\Rightarrow\Delta$ to $\Gamma, A\wedge B\Rightarrow\Delta$ is a single $L\wedge$ step. The other direction is a simple application of Cut:
+
+		\begin{center}
+			\AxiomC{\hphantom{X}}
+			\RightLabel{Ax}
+			\UnaryInfC{$A, B \Rightarrow A$}
+			\AxiomC{\hphantom{X}}
+			\RightLabel{Ax}
+			\UnaryInfC{$A, B \Rightarrow B$}
+			\RightLabel{R$\wedge$}
+			\BinaryInfC{$A, B \Rightarrow A\wedge B$}
+			\AxiomC{$\vdots$}
+			\noLine
+			\UnaryInfC{$\Gamma, A\wedge B\Rightarrow\Delta$}
+			\RightLabel{Cut}
+			\BinaryInfC{$\Gamma, A, B\Rightarrow\Delta$}
+			\DisplayProof
+		\end{center}
+
+		4. Analogous to 3.
 	\end{proof}
@@
 
-	The first result is that for every sequent there is some equi-provable implicational sequent.
+	The first result is that for every sequent there is some equiprovable implicational sequent.
 This transformation is not strictly necessary for what follows, but it makes the arguments and procedure considerably simpler.
@@
 			$\Delta\Rightarrow \Gamma, A, B$&$\rightsquigarrow$&$\Delta\Rightarrow\Gamma, A\vee B$\\
+			$\Delta\Rightarrow A\to B$&$\rightsquigarrow$&$\Delta, A\Rightarrow B$\\
 			$\Delta\Rightarrow A\vee B$&$\rightsquigarrow$&$\Delta, (A\vee B)\to N\Rightarrow N$\\
 			$\Delta\Rightarrow A\wedge B$&$\rightsquigarrow$&$\Delta,(A\wedge B)\to N\Rightarrow N$\\
-			$\Delta\Rightarrow A\to B$&$\rightsquigarrow$&$\Delta, A\Rightarrow B$\\
 
 			$\Delta, A\wedge B\Rightarrow C$&$\rightsquigarrow$&$\Delta, A, B\Rightarrow C$\\
-			$\Delta, A\vee B\Rightarrow C$&$\rightsquigarrow$&$\Delta, (A\to C)\to (B\to C)\to C\Rightarrow  C$\\
+			$\Delta, A\vee B\Rightarrow C$&$\rightsquigarrow$&$\Delta, (A\to C)\to (B\to C)\to N\Rightarrow  N$\\
 			$\Delta, (A\wedge B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to (B\to C)\Rightarrow D$\\
 			$\Delta, (A\vee B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to C, B\to C\Rightarrow D$\\
-			$\Delta, A\to (B\vee C)\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to (B\to D)\to (C\to D)\to D\Rightarrow D$\\
+			$\Delta, A\to (B\vee C)\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to (B\to D)\to (C\to D)\to N\Rightarrow N$\\
 			$\Delta, A\to (B\wedge C)\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to B, A\to C\Rightarrow D$\\
-			$\Delta, X\to (A\to B)\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, N\to (A\to B)\Rightarrow C$\\
-			$\Delta, A\to (X\to B)\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, A\to (N\to B)\Rightarrow C$\\
-			$\Delta, A\to (B\to X)\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, A\to (B\to N)\Rightarrow C$\\
+			$\Delta, X\to A\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, N\to (A\to B)\Rightarrow C$\\
+			$\Delta, A\to X\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, A\to (N\to B)\Rightarrow C$\\
+			$\Delta, A\to B\to X\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, A\to (B\to N)\Rightarrow C$\\
 			$\Delta, (X\to A)\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, (N\to A)\to B\Rightarrow C$\\
@@
 	\begin{lemma}
-		Iteratively applying the above transformations we can obtain from every sequent $S$ an equi-provable implicational sequent $S_\to$.
+		Iteratively applying the above transformations we can obtain from every sequent $S$ an equiprovable implicational sequent $S_\to$.
 	\end{lemma}
 
-	\begin{proof}[Proofsketch]
+	\begin{proof}[Proof]
 
-		Check that for any formula that is not implicational there is an applicable rule.
+		There are three things to check: 1.
+There is an applicable transformation for every non-implicational sequent.
+2.
+The translation procedure terminates on every input.
+3.
+The sequents involved in every transformation are equiprovable.
 
-		Check that the sequents on both sides of each rule are equi-provable.
+		1.
+Suppose $\Gamma\Rightarrow\Delta$ is not implicational.
+Then the succedent is not a single atom, or there is a formula in the antecedent that is not of the form $(C\to D)\to E, C\to (D\to E), C\to D$ or $D$ for some atoms $C, D, E$.
+In the former case one of transformations 1-4 is applicable.
+In the latter case one of the other rules is.
 
-		Termination can be shown by well-foundedness: First the number of formulas in the succedent is reduced to $1$, then the succedent is reduced to an atom.
-Then in each step the number of disjunctions, conjunctions or the depth of nested implications is reduced.
+		2.
+Define the complexity $c(\varphi)$ of a formula as $c(A) = 0$ for atoms $A$, $c(A\circ B) = c(A) + c(B) + 1$ for $\circ\in\{\wedge,\vee,\to\}$.
+We can use $c$ to define a partial order $\succ$ on finite multi-sets of formulas, i.e.
+have $\emptyset\not \succ\Delta$ and $\Gamma\succ\emptyset$ for $\Gamma$ non-empty and for both $\Gamma,\Delta$ non-empty have\begin{align*}
+			\Gamma \succ\Delta:\Longleftrightarrow&\max_\Gamma c > \max_\Delta c\text{ or }\\&\max_\Delta c =\max_\Gamma c\text{ and }\Gamma - \argmax_\Gamma c \succ\Delta - \argmax_\Delta c.
+		\end{align*} $\succ$ is well-founded: Suppose towards contradiction that there is an infinite chain $$\Gamma_0^{(0)} \succ \Gamma_1^{(0)} \succ\dots\succ\Gamma_n^{(0)}\succ\dots$$Then every $\Gamma_i^{(0)}$ must be non-empty and $$\max_{\Gamma_0^{(0)}}c \geq \max_{\Gamma_1^{(0)}}c \geq\dots \geq \max_{\Gamma_n^{(0)}}c \geq\dots$$ and since $\mathbb{N}$ is well-ordered this chain must become stationary, i.e.
+there is $N_0$ such that $\max_{\Gamma_i^{(0)}}c = \max_{\Gamma_{N_0}^{(0)}}c$ for $i\geq N_0$ and $\max_{\Gamma_i^{(0)}}c>\max_{\Gamma_{N_0}^{(0)}}c$ for $i<N_0$.
+In particular for $i\geq N_0$ we have $\Gamma_i^{(0)}-\argmax_{\Gamma_i^{(0)}}c >_c \Gamma_{i+1}^{(0)}-\argmax_{\Gamma_{i+1}^{(0)}}c$, thus we obtain a second chain $$\Gamma_0^{(1)} \succ\Gamma_1^{(1)} \succ\dots\succ\Gamma_n^{(1)}\succ\dots$$ with $$\Gamma_i^{(1)} = \Gamma_{N+i}^{(0)} - \argmax_{\Gamma_{N+i}^{(0)}}c.$$We can repeat this construction to obtain an sequence $N_0, N_1,\dots$ Note that $$ \max_{\Gamma_{N_i}^{(i)}}c\geq \max_{\Gamma_{0}^{(i+1)}}c\geq \max_{\Gamma_{N_{i+1}}^{(i+1)}}c$$where the last is even a strict inequality if $N_{i+1} > 0$.
+Observe that $\Gamma_{N_0+\dots+N_i}^{(0)}$ by construction must contain at least $i+1$ formulas, i.e.
+$N_0,N_1,\dots$ cannot become constant 0.
+Consider a subsequence $N_{\alpha(0)},N_{\alpha(1)},\dots$ that does not contain a zero.
+Then we have $$\max_{\Gamma_{N_{\alpha(0)}}^{({\alpha(0)})}}c > \max_{\Gamma_{N_{\alpha(1)}}^{({\alpha(1)})}}c>\dots >\max_{\Gamma_{N_{\alpha(n)}}^{({\alpha(n)})}}c>\dots$$ which is an infinite decreasing chain, but $\mathbb{N}$ is a well-order.
+$\lightning$
+
+	 	We can now use $\succ$ to define a partial order $\gg$ on the set of sequents as follows: $$\Gamma\Rightarrow\Delta \gg \Gamma'\Rightarrow\Delta'$$ iff
+		\begin{enumerate}
+			\item $|\Delta| > |\Delta'|$ 
+			\item not 1 and $\Delta \succ \Delta'$
+			\item not 1, 2 and the number of occurrences of $\wedge$ in $\Gamma$ is higher than in $\Gamma' $.
+			\item not 1 - 3 and the number of occurrences of $\vee$ in $\Gamma$ is higher than in $\Gamma' $.
+			\item not 1 - 4 and $\Gamma \succ\Gamma' $.
+		\end{enumerate}
+		Since each of 1 - 5 is well-founded, $\gg$ is well-founded, i.e. for each set we can obtain a minimal element by considering first a minimal subset with regard to 1, then to 2 and so on.
+
+		Now every transformation is reductive with regard to $\gg$.
+Therefore iteratively applying transformations creates a descending chain.
+Since every descending chain in a well-founded order is finite this proves termination.
+
+		3.
+For most cases equiprovability follows from Lemmas \ref{ep1} and \ref{ep2}, e.g.
+consider $$\Delta, A\to (B\wedge C)\Rightarrow D\indent\rightsquigarrow\indent\Delta, A\to B, A\to C\Rightarrow D$$Then we have
+		\begin{align*}
+			\Delta, A\to (B\wedge C)\Rightarrow D&\overset{\ref{ep1}.1}{\sim_p}\Delta, (A\to B)\wedge (A\to C)\Rightarrow D\\&\overset{\ref{ep2}.3}{\sim_p}\Delta, A\to B, A\to C\Rightarrow D
+		\end{align*}
+
+		Some cases require simple applications of cut using intuitionistic tautologies, e.g. for
+		$$\Delta, A\vee B\Rightarrow C\indent\rightsquigarrow\indent\Delta, (A\to C)\to (B\to C)\to N\Rightarrow  N$$
+		we have 
+		\begin{align*}
+			\Delta, A\vee B\Rightarrow C&\overset{\star}{\sim_p}\Delta\Rightarrow (A\vee B)\to C\\&\overset{\ref{ep2}.3}{\sim_p}\Delta\Rightarrow (A\to C)\wedge(B\to C)\\&\overset{\ref{ep2}.2}{\sim_p} \Delta, ((A\to C)\wedge (B\to C))\to N\Rightarrow N\\&\overset{\ref{ep1}.1}{\sim_p}\Delta, (A\to C)\to(B\to C)\to N\Rightarrow N.
+		\end{align*}
+		Where one direction of $\star$ is trivial and the other is obtained via cut with $$A\vee B, (A\vee B)\to C\Rightarrow C$$
+
+		For the last 6 rules one direction in obtained via cut and the other one by substituting all occurrences of $N$ in a proof with $X$.
 	\end{proof}

### 2022-04-11 WWTF/main.tex
asdf

@@
   Program synthesis is a long-standing problem that crosses many areas of computer science and mathematics.
-  Program synthesis holds the promise to  increase the reliability of software systems (because the synthesized program are correct by construction) and for increasing programmer productivity by enabling programmers to focus on high-level ideas instead of low-level details.
+  Program synthesis holds the promise to  increase the reliability of software systems (because the synthesized program are correct by construction) and to increase programmer productivity by enabling programmers to focus on high-level ideas instead of low-level details.
 
@@
 %, which may become costly or even impossible for sufficiently complex systems.
-In contrast, \emph{correct-by-construction} approaches seek to develop a product that continues to be correct, from the initial design to the final product.
+In contrast, \emph{correct-by-construction} approaches seek to develop a system that continues to be correct, from the initial design to the final product.
 Advances in deductive synthesis promise to increase the applicability of the correct-by-construction approach, leading to higher reliability in software development.
@@
 Automated theorem proving has made impressive progress in recent years, demonstrated in yearly solver competitions.
-We will submit our benchmarks to these competitions in order to interest solver developers in improving their tools wrt to the specific needs of deductive synthesis.
+We will submit our benchmarks to these competitions in order to interest solver developers in improving their tools wrt the specific needs of deductive synthesis.
 %(formulas with quantifier alternation, support for induction, etc.).

### 2022-04-29 Notes/A reverse double negation translation/RDNT.tex
Completed proofs of propositional part

@@
 
-\title{A reverse double negation translation}
+\title{Embedding intuitionistic into classical logic}
 \author{Alexander Pluska}
@@
 \usepackage{bussproofs}
+\usepackage{enumerate}
+\usepackage{tikz-qtree}
+\usepackage{stackengine}
 
@@
 
-	A sequent $A\Rightarrow B$ consists of two finite multi-sets of formulas $A, B$. We will be using the following rules for our classical calculus:\\
+	Since the details of syntax are a delicate concern for proof theory we will first collect some important definitions and facts.
+Fix a countably infinite number of atomic propositions $A, B, C,\dots$ as well as a special atomic proposition $\bot$.
+
+	\begin{definition}
+		A \textit{Fomrula} can be constructed only in the following ways:
+		\begin{itemize}
+			\item every atomic proposition is a formula.
+			\item given formulas $A, B$ we have formulas $A\wedge B, A\vee B, A\to B$.
+		\end{itemize}
+	 	With every formula $\varphi$ we can associate a unique labelled rooted binary tree, the \textit{syntax tree} $T(\varphi)$:
+		 \begin{itemize}
+		 	\item If $\varphi$ is atomic then $T(\varphi)$ consists of a single node with label $\varphi$: \tikz[scale=0.9, baseline=0mm, thick]{\node[draw] {$ \varphi$}}
+		 	\item If $\varphi = \varphi_1\circ\varphi_2$ then the root node is again labelled with gamma and has two child nodes, at which the trees $T(\varphi_1)$ and $T(\varphi_2)$ are attached:
+		 	\begin{center}
+		 	\tikz[scale=1, baseline=-5mm, thick]{\Tree [.\node[draw]{$\varphi$}; [.$T(\varphi_1)$  ] [.$T(\varphi_2)$  ] ]}
+	 		\end{center}
+		 \end{itemize}
+	 	$\psi$ is a \textit{subformula} of $\varphi$ if it is the label of some node of $T(\varphi)$.
+	 \end{definition}
+
+	 \begin{example}
+	 	Consider the formula $((A\wedge B)\to A) \to (B\vee A)$. Then its syntax tree is
+	 	\begin{center}
+	 		\tikz[scale=1, baseline=-5mm, thick]{\Tree [.\node[draw]{$((A\to B)\to A)\to (B\vee A)$}; [.\node[draw]{$(A\to B)\to A$}; [.\node[draw]{$A\to B$}; [.\node[draw]{$A$}; ] [.\node[draw]{$B$}; ] ] [.\node[draw]{$A$}; ] ] [.\node[draw]{$B\vee A$}; [.\node[draw]{$B$}; ] [.\node[draw]{$A$};] ] ]}
+	 	\end{center}
+	 \end{example}
+
+ 	We are now ready to introduce our calculi for intuitionistic and classical logic:
+
+	\begin{definition}
+			A \textit{sequent} $A\Rightarrow B$ consists of two multisets of formulas $A, B$.
+The position of $A_i$ in $A_1,\dots, A_n\Rightarrow B_1,\dots, B_m$ is $0i$ and that of $B_i$ is $1i$.
+A formula $\varphi$ occurs at $p = (ji, p')$ if $\varphi$ occurs at $p'$ in $A_i$ if $j=0$ and in $B_i$ if $j=1$.
+The following rules make up our classical calculus $\mathbf{GC}$, corresponding roughly to the propositional fragment of \textbf{G3c} from~\cite{basicprooftheory}:\\
 		\begin{center}
@@
 
-	For the intuitionistic calculus we just need to modify R$\to$:
+		To obtain our intuitionistic calculus \textbf{GI} (corresponding to \textbf{m-G3i} from~\cite{basicprooftheory}) substitute R$\to$ with
 		\begin{center}
@@
 
-	These calculi have the following nice property:
+		A \textit{proof} of a formula $\varphi$ in a calculus $\mathbf{G}$ is a labelled binary rooted tree in which every node is labelled with a pair of the name of a rule in $\mathbf{G}$, its rule, and a sequent, its sequent, such that
+		\begin{itemize}
+			\item Every leaf is labelled with (Ax, $A,\Gamma\Rightarrow\Delta, A$) or (L$\bot$, $\bot, \Gamma\Rightarrow\Delta$) for some formula $A$, multi-sets of formulas $\Gamma,\Delta$.
+			\item The sequent of an interior node can be derived from sequents of its child nodes according to its rule.
+		\end{itemize}
+		We usually present a proof directly as a sequence of applications of rules rather than in a more usual tree presentation.
+Note that we implicitly permute the antecedent and succedent, i.e.
+we implicitly may apply a number of exchange rules at any point in the proof.
+	\end{definition}
+
+	\begin{example}
+		The following is a proof of $(A\to B)\to A\Rightarrow B$:
+		\begin{center}
+			\AxiomC{}
+			\RightLabel{Ax}
+			\UnaryInfC{$A\to B, A\Rightarrow B, A$}
+			\AxiomC{}
+			\RightLabel{Ax}
+			\UnaryInfC{$A, B\Rightarrow B$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$A\to B, A\Rightarrow B$}
+			\DisplayProof
+		\end{center}
+	\end{example}
+
+	\noindent We now recall some important properties of the given systems:
 
@@
 	\begin{definition}
-		A sequent $A_1,\dots,A_n\Rightarrow B$ is \textit{implicational} if $B$ is atomic and each $A_i$ is of the form either $C\to (D\to E)$, $(C\to D)\to E$, $C\to D$ or $C$ for some atomic $C, D, E$.
+		A sequent $\Gamma\Rightarrow \Delta$ is \textit{implicational} if each formula $A$ in $\Gamma$ is of the form $C_A\to (D_A\to E_A)$, $(C_A\to D_A)\to E_A$, $C_A\to D_A$ or $C_A$ for some atomic $C_A, D_A, E_A$ and each formula $B$ in $\Delta$ is of the form $C_B$ or $C_B\to D_B$ for some atomic $C_B, D_B$.
 	\end{definition}
 
-	The first result is that for every sequent there is some equiprovable implicational sequent.
-This transformation is not strictly necessary for what follows, but it makes the arguments and procedure considerably simpler.
-It is reminiscent of the definitions of connectors in second order logic.
+	While at first this might seems quite restrictive we can give a translation procedure that transforms every sequent into an equiprovable implicational sequent - even one containing a single atom in the succedent - so in some sense this is a normal form for intuitionistic sequents.
+One nice property of implicational sequents is that their proofs consist of implicational sequents and contain only the rules Ax, L$\bot$, L$\to$ and R$\to$
 
-	Consider the following rules, where $X$ is not atomic, $N$ is a new atom:
+	The translation steps are reminiscent of the definitions of connectors in second order logic. Consider the following rules, where $X$ is not atomic, $N$ is a new atom:
 	\begin{center}
-		\begin{tabular}{lcl}
-			$\Delta\Rightarrow \Gamma, A, B$&$\rightsquigarrow$&$\Delta\Rightarrow\Gamma, A\vee B$\\
-			$\Delta\Rightarrow A\to B$&$\rightsquigarrow$&$\Delta, A\Rightarrow B$\\
-			$\Delta\Rightarrow A\vee B$&$\rightsquigarrow$&$\Delta, (A\vee B)\to N\Rightarrow N$\\
-			$\Delta\Rightarrow A\wedge B$&$\rightsquigarrow$&$\Delta,(A\wedge B)\to N\Rightarrow N$\\
+		\begin{tabular}{clcl}
+			(1)&$\Gamma\Rightarrow \Gamma, A, B$&$\rightsquigarrow$&$\Gamma\Rightarrow\Gamma, A\vee B$\\
+			(2)&$\Gamma\Rightarrow A\to B$&$\rightsquigarrow$&$\Gamma, A\Rightarrow B$\\
+			(3)&$\Gamma\Rightarrow A\vee B$&$\rightsquigarrow$&$\Gamma, (A\vee B)\to N\Rightarrow N$\\
+			(4)&$\Gamma\Rightarrow A\wedge B$&$\rightsquigarrow$&$\Gamma,(A\wedge B)\to N\Rightarrow N$\\
 
-			$\Delta, A\wedge B\Rightarrow C$&$\rightsquigarrow$&$\Delta, A, B\Rightarrow C$\\
-			$\Delta, A\vee B\Rightarrow C$&$\rightsquigarrow$&$\Delta, (A\to C)\to (B\to C)\to N\Rightarrow  N$\\
-			$\Delta, (A\wedge B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to (B\to C)\Rightarrow D$\\
-			$\Delta, (A\vee B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to C, B\to C\Rightarrow D$\\
-			$\Delta, A\to (B\vee C)\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to (B\to D)\to (C\to D)\to N\Rightarrow N$\\
-			$\Delta, A\to (B\wedge C)\Rightarrow D$&$\rightsquigarrow$&$\Delta, A\to B, A\to C\Rightarrow D$\\
-			$\Delta, X\to A\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, N\to (A\to B)\Rightarrow C$\\
-			$\Delta, A\to X\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, A\to (N\to B)\Rightarrow C$\\
-			$\Delta, A\to B\to X\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, A\to (B\to N)\Rightarrow C$\\
-			$\Delta, (X\to A)\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, (N\to A)\to B\Rightarrow C$\\
-			$\Delta, (A\to X)\to B\Rightarrow C$&$\rightsquigarrow$&$\Delta, X\to N, (A\to N)\to B\Rightarrow C$\\
-			$\Delta, (A\to B)\to X\Rightarrow C$&$\rightsquigarrow$&$\Delta, N\to X, (A\to B)\to N\Rightarrow C$\\
+			(5)&$\Gamma, A\wedge B\Rightarrow C$&$\rightsquigarrow$&$\Gamma, A, B\Rightarrow C$\\
+			(6)&$\Gamma, A\vee B\Rightarrow C$&$\rightsquigarrow$&$\Gamma, (A\to C)\to (B\to C)\to N\Rightarrow  N$\\
+			(7)&$\Gamma, (A\wedge B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Gamma, A\to (B\to C)\Rightarrow D$\\
+			(8)&$\Gamma, (A\vee B)\to C\Rightarrow D$&$\rightsquigarrow$&$\Gamma, A\to C, B\to C\Rightarrow D$\\
+			(9)&$\Gamma, A\to (B\wedge C)\Rightarrow D$&$\rightsquigarrow$&$\Gamma, A\to B, A\to C\Rightarrow D$\\
+			(10)&$\Gamma, A\to (B\vee C)\Rightarrow D$&$\rightsquigarrow$&$\Gamma, A\to (B\to D)\to (C\to D)\to N\Rightarrow N$\\
+			(11)&$\Gamma, X\to A\to B\Rightarrow C$&$\rightsquigarrow$&$\Gamma, X\to N, N\to (A\to B)\Rightarrow C$\\
+			(12)&$\Gamma, A\to X\to B\Rightarrow C$&$\rightsquigarrow$&$\Gamma, X\to N, A\to (N\to B)\Rightarrow C$\\
+			(13)&$\Gamma, A\to B\to X\Rightarrow C$&$\rightsquigarrow$&$\Gamma, N\to X, A\to (B\to N)\Rightarrow C$\\
+			(14)&$\Gamma, (X\to A)\to B\Rightarrow C$&$\rightsquigarrow$&$\Gamma, N\to X, (N\to A)\to B\Rightarrow C$\\
+			(15)&$\Gamma, (A\to X)\to B\Rightarrow C$&$\rightsquigarrow$&$\Gamma, X\to N, (A\to N)\to B\Rightarrow C$\\
+			(16)&$\Gamma, (A\to B)\to X\Rightarrow C$&$\rightsquigarrow$&$\Gamma, N\to X, (A\to B)\to N\Rightarrow C$\\
 		\end{tabular}
@@
 Then the succedent is not a single atom, or there is a formula in the antecedent that is not of the form $(C\to D)\to E, C\to (D\to E), C\to D$ or $D$ for some atoms $C, D, E$.
-In the former case one of transformations 1-4 is applicable.
-In the latter case one of the other rules is.
+If the succedent is not a single atom one of transformations (1)-(4) is applicable.
+If it is but the antecedent contains a formula which is not an implication (5) or (6) are.
+If that formula is an implication but one of its arguments is not then (7)-(10) are applicable.
+Finally if it is an implication of implications but the involved formulas are not atomic one of the rules (11)-(16) applies.
 
 		2.
-Define the complexity $c(\varphi)$ of a formula as $c(A) = 0$ for atoms $A$, $c(A\circ B) = c(A) + c(B) + 1$ for $\circ\in\{\wedge,\vee,\to\}$.
-We can use $c$ to define a partial order $\succ$ on finite multi-sets of formulas, i.e.
-have $\emptyset\not \succ\Delta$ and $\Gamma\succ\emptyset$ for $\Gamma$ non-empty and for both $\Gamma,\Delta$ non-empty have\begin{align*}
-			\Gamma \succ\Delta:\Longleftrightarrow&\max_\Gamma c > \max_\Delta c\text{ or }\\&\max_\Delta c =\max_\Gamma c\text{ and }\Gamma - \argmax_\Gamma c \succ\Delta - \argmax_\Delta c.
-		\end{align*} $\succ$ is well-founded: Suppose towards contradiction that there is an infinite chain $$\Gamma_0^{(0)} \succ \Gamma_1^{(0)} \succ\dots\succ\Gamma_n^{(0)}\succ\dots$$Then every $\Gamma_i^{(0)}$ must be non-empty and $$\max_{\Gamma_0^{(0)}}c \geq \max_{\Gamma_1^{(0)}}c \geq\dots \geq \max_{\Gamma_n^{(0)}}c \geq\dots$$ and since $\mathbb{N}$ is well-ordered this chain must become stationary, i.e.
-there is $N_0$ such that $\max_{\Gamma_i^{(0)}}c = \max_{\Gamma_{N_0}^{(0)}}c$ for $i\geq N_0$ and $\max_{\Gamma_i^{(0)}}c>\max_{\Gamma_{N_0}^{(0)}}c$ for $i<N_0$.
-In particular for $i\geq N_0$ we have $\Gamma_i^{(0)}-\argmax_{\Gamma_i^{(0)}}c >_c \Gamma_{i+1}^{(0)}-\argmax_{\Gamma_{i+1}^{(0)}}c$, thus we obtain a second chain $$\Gamma_0^{(1)} \succ\Gamma_1^{(1)} \succ\dots\succ\Gamma_n^{(1)}\succ\dots$$ with $$\Gamma_i^{(1)} = \Gamma_{N+i}^{(0)} - \argmax_{\Gamma_{N+i}^{(0)}}c.$$We can repeat this construction to obtain an sequence $N_0, N_1,\dots$ Note that $$ \max_{\Gamma_{N_i}^{(i)}}c\geq \max_{\Gamma_{0}^{(i+1)}}c\geq \max_{\Gamma_{N_{i+1}}^{(i+1)}}c$$where the last is even a strict inequality if $N_{i+1} > 0$.
-Observe that $\Gamma_{N_0+\dots+N_i}^{(0)}$ by construction must contain at least $i+1$ formulas, i.e.
-$N_0,N_1,\dots$ cannot become constant 0.
-Consider a subsequence $N_{\alpha(0)},N_{\alpha(1)},\dots$ that does not contain a zero.
-Then we have $$\max_{\Gamma_{N_{\alpha(0)}}^{({\alpha(0)})}}c > \max_{\Gamma_{N_{\alpha(1)}}^{({\alpha(1)})}}c>\dots >\max_{\Gamma_{N_{\alpha(n)}}^{({\alpha(n)})}}c>\dots$$ which is an infinite decreasing chain, but $\mathbb{N}$ is a well-order.
-$\lightning$
+We proceed by defining a well-founded order on the set of sequents $\mathcal S$ and showing that each of the transformations is reductive.
+Define the complexity $c(\varphi)$ of a formula as usual as $c(A) = 0$ for atoms $A$ and $c(A\circ B) = c(A) + c(B) + 1$ for $\circ\in\{\wedge,\vee,\to\}$.
+$c$ induces a partial order $<_c$ on formulas.
+We can extend this to a ordering $<_{DM}$ on finite multisets of formulas $\mathcal M$, also knows as the Dershowitz-Manna ordering as follows:
 
-	 	We can now use $\succ$ to define a partial order $\gg$ on the set of sequents as follows: $$\Gamma\Rightarrow\Delta \gg \Gamma'\Rightarrow\Delta'$$ iff
-		\begin{enumerate}
-			\item $|\Delta| > |\Delta'|$ 
-			\item not 1 and $\Delta \succ \Delta'$
-			\item not 1, 2 and the number of occurrences of $\wedge$ in $\Gamma$ is higher than in $\Gamma' $.
-			\item not 1 - 3 and the number of occurrences of $\vee$ in $\Gamma$ is higher than in $\Gamma' $.
-			\item not 1 - 4 and $\Gamma \succ\Gamma' $.
+		$M <_{DM}N$ iff there exist $X, Y\in\mathcal{M}$ such that
+		\begin{itemize}
+			\item $X\subseteq N$ is non-empty,
+			\item $M = (N \setminus X) \cup Y$,
+			\item for all $y\in Y$ there exists $x\in X$ such that $y <_S x$.
+		\end{itemize}
+		Since $<_c$ is well-founded so is $<_{DM}$~\cite{Dershowitz-Manna_1979}. Now consider the lexicographic order $\ll$ on $\mathcal{S}^5$ induced by the following orders:
+		\begin{enumerate}[(i)]
+			\item $(\Gamma\Rightarrow\Delta)\ll(\Gamma'\Rightarrow\Delta')$ iff $|\Delta| < |\Delta'|$ 
+			\item $(\Gamma\Rightarrow\Delta)\ll(\Gamma'\Rightarrow\Delta')$ iff $\Delta <_{DM} \Delta'$
+			\item $(\Gamma\Rightarrow\Delta)\ll(\Gamma'\Rightarrow\Delta')$ iff the number of occurrences of $\wedge$ in $\Gamma$ is lower than in $\Gamma' $.
+			\item $(\Gamma\Rightarrow\Delta)\ll(\Gamma'\Rightarrow\Delta')$ iff the number of occurrences of $\vee$ in $\Gamma$ is lower than in $\Gamma' $.
+			\item $(\Gamma\Rightarrow\Delta)\ll(\Gamma'\Rightarrow\Delta')$ iff $\Gamma <_{DM}\Gamma $.
 		\end{enumerate}
-		Since each of 1 - 5 is well-founded, $\gg$ is well-founded, i.e. for each set we can obtain a minimal element by considering first a minimal subset with regard to 1, then to 2 and so on.
-
-		Now every transformation is reductive with regard to $\gg$.
-Therefore iteratively applying transformations creates a descending chain.
-Since every descending chain in a well-founded order is finite this proves termination.
+		Then the diagonal of $\ll$ is our desired order $<$.
+Since each involved order is well-founded so is $\ll$ and therefore also $<$.
+Now every transformation is reductive with regard to $<$, i.e.
+(1) is reductive w.r.t (i), (2)-(4) are reductive w.r.t.
+(ii), (5), (7), (9) are reductive w.r.t (iii), (6), (8), (10) w.r.t 4 (iv) and (11) - (16) w.r.t.
+(v) and non are increasing with any of the higher orders.
+This proves termination.
 
@@
 For most cases equiprovability follows from Lemmas \ref{ep1} and \ref{ep2}, e.g.
-consider $$\Delta, A\to (B\wedge C)\Rightarrow D\indent\rightsquigarrow\indent\Delta, A\to B, A\to C\Rightarrow D$$Then we have
+consider $$\Gamma, A\to (B\wedge C)\Rightarrow D\indent\rightsquigarrow\indent\Gamma, A\to B, A\to C\Rightarrow D$$Then we have
 		\begin{align*}
-			\Delta, A\to (B\wedge C)\Rightarrow D&\overset{\ref{ep1}.1}{\sim_p}\Delta, (A\to B)\wedge (A\to C)\Rightarrow D\\&\overset{\ref{ep2}.3}{\sim_p}\Delta, A\to B, A\to C\Rightarrow D
+			\Gamma, A\to (B\wedge C)\Rightarrow D&\overset{\ref{ep1}.1}{\sim_p}\Gamma, (A\to B)\wedge (A\to C)\Rightarrow D\\&\overset{\ref{ep2}.3}{\sim_p}\Gamma, A\to B, A\to C\Rightarrow D
 		\end{align*}
@@
 		Some cases require simple applications of cut using intuitionistic tautologies, e.g. for
-		$$\Delta, A\vee B\Rightarrow C\indent\rightsquigarrow\indent\Delta, (A\to C)\to (B\to C)\to N\Rightarrow  N$$
+		$$\Gamma, A\vee B\Rightarrow C\indent\rightsquigarrow\indent\Gamma, (A\to C)\to (B\to C)\to N\Rightarrow  N$$
 		we have 
 		\begin{align*}
-			\Delta, A\vee B\Rightarrow C&\overset{\star}{\sim_p}\Delta\Rightarrow (A\vee B)\to C\\&\overset{\ref{ep2}.3}{\sim_p}\Delta\Rightarrow (A\to C)\wedge(B\to C)\\&\overset{\ref{ep2}.2}{\sim_p} \Delta, ((A\to C)\wedge (B\to C))\to N\Rightarrow N\\&\overset{\ref{ep1}.1}{\sim_p}\Delta, (A\to C)\to(B\to C)\to N\Rightarrow N.
+			\Gamma, A\vee B\Rightarrow C&\overset{\star}{\sim_p}\Gamma\Rightarrow (A\vee B)\to C\\&\overset{\ref{ep2}.3}{\sim_p}\Gamma\Rightarrow (A\to C)\wedge(B\to C)\\&\overset{\ref{ep2}.2}{\sim_p} \Gamma, ((A\to C)\wedge (B\to C))\to N\Rightarrow N\\&\overset{\ref{ep1}.1}{\sim_p}\Gamma, (A\to C)\to(B\to C)\to N\Rightarrow N.
 		\end{align*}
@@
 
-	For each atom $A$ consider a new atom $A'$ and a special atom $\omega = \bot'$.
-Have $\top' = \top$.
-Fix some implicational sequent $S := \Delta\Rightarrow C$ for now.
-Let $\mathcal{A}$ be the set of atoms occurring in $S$.
+	This reduces our task to embedding just the implicational sequents from intuitionistic into classical logic.
+And we shall immediately present this embedding, even though showing its correctness will bee a bit involved:
+
+	 Consider some implicational sequent $S := \Gamma\Rightarrow \Delta$.
+Let $\{(C_i\to D_i)\to E_i\}_{i\in I}$ be an enumeration of sequents of that form occurring in $\Gamma$ and $\{C_j\to D_j\}_{j\in J}$ of such sequents in $\Delta$, and let $i = j$ iff.
+For each atom $A\neq\bot$ occurring in $S$ and $i\in I\cup J$ define a new atom $A^{(i)}$ with the stipulation that $A^{(i)} = A^{(j)}$ if $C^{(i)} = C^{(j)}$ and $D^{(i)} = D^{(j)}$.
+Furthermore have $\bot^{(i)} = \bot$.
+We call all such $A^{(i)}$ primed.
 Then define
 	\begin{align*}
-		\Delta' = &\{A\to A', \omega\to A'\:|\:A\in\mathcal{A}\}
-		\\&\cup \{(A'\to B')\to C\:|\: (A\to B)\to C\in\Delta\}
-		\\&\cup \{A'\to (B'\to C'), A\to(B\to C)\:|\: A\to (B\to C)\in\Delta\}
-		\\&\cup \{A'\to B', A\to B\:|\: A\to (B\to C)\in\Delta\}
+		\Gamma' = &\: (\Gamma\setminus\{(C_{i}\to D_i)\to E_i\:|\:i\in I\})
+		\\&\cup \{(C_{i}^{(i)}\to D_i^{(i)})\to E_i\:|\:i\in I\}
+		\\&\cup \{A_i^{(i)}\to B_i^{(i)}\:|\: A, B\text{ atomic}, A\to B\in\Gamma, i\in I\cup J\}
+		\\&\cup \{C^{(i)}\to (D^{(i)}\to E^{(i)})\:|\: C, D, E\text{ atomic}, C\to (D\to E)\in\Gamma, i\in I\cup J\}
+		\\&\cup \{D_i\to D_i^{(i)}\:|\:i\in I\cup J\}
 	\end{align*}
-and have $$S' := \Delta' \Rightarrow C$$
+as well as
+\begin{align*}
+	\Delta' = &\:(\Delta\setminus\{C_{j}\to D_j\:|\:j\in J\})
+	\\&\cup \{C_j^{(j)}\to D_j^{(j)}\:|\: C, D\text{ atomic}, C\to D\in\Delta, j\in J\}
+\end{align*}
+and have $$S' := \Gamma' \Rightarrow \Delta' $$
+
+	We prove that $S$ is intuitionistically valid iff $S'$ is classically valid by first showing that $S$ and $S'$ are equiprovable and then that $S'$ is provable intuitionistically iff it is classically provable.
 
 	\begin{lemma}
-		An implicational sequent $S$ is intuitionistically provable if and only if $S'$ is classically provable
+		Any implicational sequent $S$ is equiprovable with $S'$.
 	\end{lemma}
 
-	\begin{proof}[Proofsketch]
+	\begin{proof}
 
-		$\Rightarrow$ the classical proof can be obtained by appropriate renaming of atoms and weakening. It will still be intuitionistically sound.
+		Suppose $S'$ is provable.
+Consider the proof obtained by replacing every occurrence of primed atom with its regular version, then a number of contractions and cuts with sequents of the form $\Rightarrow A\to A$ yields a proof of $S$.
+The proof of the other direction is much more involved.
 
-		$\Leftarrow$ Rename all variables $A'$ to $A$ and rewrite the R$\to$ rules. At the end apply a number of cuts to remove the $A\to A,\bot\to A$.
+		Suppose $S$ is provable.
+We show that $S'$ is provable by induction on the height of the proof of $S$.
+If the proof of $S$ consists of a single Ax/L$\bot$ then $S'$ can also be proven by a single Ax/L$\bot$.
+Suppose that last rule in the proof of $S$ is $L\to$, i.e.
+it ends in
 
+		\begin{center}
+			\AxiomC{$A\to B, \Gamma\Rightarrow\Delta, A$}
+			\AxiomC{$B, \Gamma\Rightarrow\Delta$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$A\to B, \Gamma\Rightarrow \Delta$}
+			\DisplayProof
+		\end{center}
+
+		There are three cases:
+
+		1.
+$A, B$ are atoms.
+By induction hypothesis $A\to B, (A^{(i)}\to B^{(i)})_{i\in I}, \Gamma' \Rightarrow \Delta', A$ and $B,\Gamma' \Rightarrow \Delta'$ are provable and thus by weakening also $B, (A^{(i)}\to B^{(i)})_{i\in I}, \Gamma' \Rightarrow \Delta'$ and we have
+		\begin{center}
+			\AxiomC{$A\to B, \{A^{(i)}\to B^{(i)}\}_{i\in I},\Gamma' \Rightarrow\Delta' , A$}
+			\AxiomC{$B, \{A^{(i)}\to B^{(i)}\}_{i\in I}, \Gamma' \Rightarrow\Delta' $}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$A\to B, \{A^{(i)}\to B^{(i)}\}_{i\in I}, \Gamma' \Rightarrow \Delta' $}
+			\DisplayProof
+		\end{center}
+		so $S'$ is indeed provable.
+
+		2.
+$A$ atomic, $B = C\to D$ with $C, D$ atomic.
+Again by induction hypothesis we obtain that $A\to (C\to D), \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}, \Gamma'\Rightarrow \Delta', A$ and $C\to D, \{C^{(i)}\to D^{(i)}\}_{i\in I}, \Gamma'\Rightarrow\Delta'$ are provable.
+Since $\Gamma'$ contains $A\to A^{(i)}$ a cut followed by a contraction gives that $A\to (C\to D), \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}, \Gamma'\Rightarrow\Delta', A^{(i)}$ is provable for each $i$ and a cut with $C\to D\Rightarrow A\to (C\to D)$ gives that $C\to D, \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}, \Gamma'\Rightarrow\Delta', A^{(i)}$ is provable.
+Then a number of cuts with $C^{(i)}\to D^{(i)}\Rightarrow A^{(i)}\to (C^{(i)}\to D^{(i)})$ gives for each $j\in I=\{1,\dots,n\}$ (w.l.o.g.
+since $I$ was just chosen as just some arbitrary finite index set) $$C\to D, \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\leq j}, \{C^{(i)}\to D^{(i)}\}_{i> j}, \Gamma'\Rightarrow\Delta', A^{(i)}$$ Note that for each $j$ we have
+
+		\begin{center}
+			\hspace*{-.7cm}
+			\AxiomC{$\stackanchor{$\{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\leq j},$}{$C\to D,\{C^{(i)}\to D^{(i)}\}_{i> j},$} \Gamma'\Rightarrow\Delta', A^{(j)}$}
+			\AxiomC{$\stackanchor{$\{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i< j},$}{$C\to D,\{C^{(i)}\to D^{(i)}\}_{i\geq j}$}\Gamma'\Rightarrow\Delta'$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$\{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i< j+1},C\to D,\{C^{(i)}\to D^{(i)}\}_{i\geq j+1},\Gamma'\Rightarrow\Delta'$}
+			\DisplayProof
+		\end{center}
+		and we have already established the right argument for $j=0$ and the left argument for every $j\in I$.
+Therefore applying the above iteratively $n$ times we obtain $C\to D, \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}\Gamma'\Rightarrow\Delta'$ and finally
+		\begin{center}
+			\AxiomC{$\stackanchor{$A\to (C\to D),\Gamma',$}{$\{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I},$} \Rightarrow \Delta', A$}
+			\AxiomC{$\stackanchor{$C\to D, \Gamma',$}{$\{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}$}\Rightarrow \Delta' $}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$A\to (C\to D), \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}, \Gamma' \Rightarrow \Delta' $}
+			\DisplayProof
+		\end{center}
+		which shows that $S'$ is indeed provable.
+
+		3.
+$A = C\to D$  with $C, D$ atomic, $B$ atomic.
+Then by induction hypothesis $(C^{(i)}\to D^{(i)})\to B, \Gamma'\Rightarrow\Delta', C^{(j)}\to D^{(j)}$ and $B, \Gamma'\Rightarrow\Delta'$ are provable, where $C^{(i)} = C^{(j)}$ and $D^{(i)} = D^{(j)}$ so a single application of L$\to$ gives us $S'$.
+
+		Finally suppose the last rule of the proof is R$\to$, i.e.
+
+		\begin{center}
+			\AxiomC{$A,\Gamma\Rightarrow B$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$\Gamma\Rightarrow \Delta, A\to B$}
+			\DisplayProof
+		\end{center}
+		Then $A, B$ are atoms.
+Let $i$ be some new index.
+Let $\Gamma''$ be $\Gamma$ where every non-primed atom $X$ has been renamed to $X^{(i)}$.
+Then $A^{(i)}, \Gamma''\Rightarrow B^{(i)}$ is provable and thus so is $\bigwedge\Gamma''\Rightarrow A^{(i)}\to B^{(i)}$.
+Since $\Gamma'\Rightarrow \bigwedge\Gamma''$ is a tautology a single cut yields that $S'$ is provable.
+	\end{proof}
+
+	\begin{lemma}
+		Let $S$ be an implicational sequent such that $S'$ is provable classically. Then $S'$ is provable intuitionistically.
+	\end{lemma}
+
+	\begin{proof}
+		We proceed by induction on the proof height of the classical proof of $S'$.
+If the last rule is Ax or L$\bot$ (or L$\to$) the proof is already intuitionistically valid (or it follows directly from the induction hypothesis).
+Otherwise the proof ends in $n$ applications of R$\to$ where $n$ is chosen maximally, i.e.
+		\begin{center}
+			\AxiomC{}
+			\RightLabel{K}
+			\UnaryInfC{$A_1^{(1)},\dots, A_n^{(n)},\Gamma'\Rightarrow\Delta', B_1^{(1)},\dots B_n^{(n)}$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{\vdots}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$\Gamma'\Rightarrow\Delta', A_1^{(1)}\to B_1^{(1)},\dots, A_n^{(n)}\to B_n^{(n)}$}
+			\DisplayProof
+		\end{center}
+		with $\text{K }\neq\text{ R}\to$. If K $=$ L$\bot$ we have $\bot\in\Delta'$ and therefore we immediately have
+		\begin{center}
+			\AxiomC{}
+			\RightLabel{L$\bot$}
+			\UnaryInfC{$\Gamma'\Rightarrow\Delta', A_1^{(1)}\to B_1^{(1)},\dots, A_n^{(n)}\to B_n^{(n)}$}
+			\DisplayProof
+		\end{center}
+	For K $=$ Ax observe that the atomic formulas in $\Delta'$ and $\Gamma'$ are not primed.
+Therefore the tautology in Ax must be between some $A_k^{(k)}$ and $B_l^{(l)}$ or some formulas in $\Gamma'$ and $\Delta'$.
+In the latter case we just need a single application of $S'$.
+In the former if $k = l$ we get the intuitionistic proof
+	\begin{center}
+		\AxiomC{}
+		\RightLabel{Ax}
+		\UnaryInfC{$A_k^{(k)},\Gamma'\Rightarrow\Delta', B_k^{(k)}$}
+		\RightLabel{R$\to$}
+		\UnaryInfC{$\Gamma'\Rightarrow\Delta', A_1^{(1)}\to B_1^{(1)},\dots, A_n^{(n)}\to B_n^{(n)}$}
+		\DisplayProof
+	\end{center}
+	otherwise $k\neq l$ but $A^{(k)} = B^{(l)}$ can only be the case if $A^{(k)} = A^{(l)} = B^{(k)} = B^{(l)}$ and the above proof works thee same.
 	\end{proof}

### 2022-05-02 Notes/A reverse double negation translation/RDNT.tex
updated 2.8 - 2.12

@@
 
-	This reduces our task to embedding just the implicational sequents from intuitionistic into classical logic.
-And we shall immediately present this embedding, even though showing its correctness will bee a bit involved:
-
-	 Consider some implicational sequent $S := \Gamma\Rightarrow \Delta$.
-Let $\{(C_i\to D_i)\to E_i\}_{i\in I}$ be an enumeration of sequents of that form occurring in $\Gamma$ and $\{C_j\to D_j\}_{j\in J}$ of such sequents in $\Delta$, and let $i = j$ iff.
-For each atom $A\neq\bot$ occurring in $S$ and $i\in I\cup J$ define a new atom $A^{(i)}$ with the stipulation that $A^{(i)} = A^{(j)}$ if $C^{(i)} = C^{(j)}$ and $D^{(i)} = D^{(j)}$.
-Furthermore have $\bot^{(i)} = \bot$.
-We call all such $A^{(i)}$ primed.
-Then define
-	\begin{align*}
-		\Gamma' = &\: (\Gamma\setminus\{(C_{i}\to D_i)\to E_i\:|\:i\in I\})
-		\\&\cup \{(C_{i}^{(i)}\to D_i^{(i)})\to E_i\:|\:i\in I\}
-		\\&\cup \{A_i^{(i)}\to B_i^{(i)}\:|\: A, B\text{ atomic}, A\to B\in\Gamma, i\in I\cup J\}
-		\\&\cup \{C^{(i)}\to (D^{(i)}\to E^{(i)})\:|\: C, D, E\text{ atomic}, C\to (D\to E)\in\Gamma, i\in I\cup J\}
-		\\&\cup \{D_i\to D_i^{(i)}\:|\:i\in I\cup J\}
-	\end{align*}
-as well as
-\begin{align*}
-	\Delta' = &\:(\Delta\setminus\{C_{j}\to D_j\:|\:j\in J\})
-	\\&\cup \{C_j^{(j)}\to D_j^{(j)}\:|\: C, D\text{ atomic}, C\to D\in\Delta, j\in J\}
-\end{align*}
-and have $$S' := \Gamma' \Rightarrow \Delta' $$
-
-	We prove that $S$ is intuitionistically valid iff $S'$ is classically valid by first showing that $S$ and $S'$ are equiprovable and then that $S'$ is provable intuitionistically iff it is classically provable.
-
-	\begin{lemma}
-		Any implicational sequent $S$ is equiprovable with $S'$.
-	\end{lemma}
-
-	\begin{proof}
-
-		Suppose $S'$ is provable.
-Consider the proof obtained by replacing every occurrence of primed atom with its regular version, then a number of contractions and cuts with sequents of the form $\Rightarrow A\to A$ yields a proof of $S$.
-The proof of the other direction is much more involved.
-
-		Suppose $S$ is provable.
-We show that $S'$ is provable by induction on the height of the proof of $S$.
-If the proof of $S$ consists of a single Ax/L$\bot$ then $S'$ can also be proven by a single Ax/L$\bot$.
-Suppose that last rule in the proof of $S$ is $L\to$, i.e.
-it ends in
-
-		\begin{center}
-			\AxiomC{$A\to B, \Gamma\Rightarrow\Delta, A$}
-			\AxiomC{$B, \Gamma\Rightarrow\Delta$}
-			\RightLabel{L$\to$}
-			\BinaryInfC{$A\to B, \Gamma\Rightarrow \Delta$}
-			\DisplayProof
-		\end{center}
-
-		There are three cases:
-
-		1.
-$A, B$ are atoms.
-By induction hypothesis $A\to B, (A^{(i)}\to B^{(i)})_{i\in I}, \Gamma' \Rightarrow \Delta', A$ and $B,\Gamma' \Rightarrow \Delta'$ are provable and thus by weakening also $B, (A^{(i)}\to B^{(i)})_{i\in I}, \Gamma' \Rightarrow \Delta'$ and we have
-		\begin{center}
-			\AxiomC{$A\to B, \{A^{(i)}\to B^{(i)}\}_{i\in I},\Gamma' \Rightarrow\Delta' , A$}
-			\AxiomC{$B, \{A^{(i)}\to B^{(i)}\}_{i\in I}, \Gamma' \Rightarrow\Delta' $}
-			\RightLabel{L$\to$}
-			\BinaryInfC{$A\to B, \{A^{(i)}\to B^{(i)}\}_{i\in I}, \Gamma' \Rightarrow \Delta' $}
-			\DisplayProof
-		\end{center}
-		so $S'$ is indeed provable.
-
-		2.
-$A$ atomic, $B = C\to D$ with $C, D$ atomic.
-Again by induction hypothesis we obtain that $A\to (C\to D), \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}, \Gamma'\Rightarrow \Delta', A$ and $C\to D, \{C^{(i)}\to D^{(i)}\}_{i\in I}, \Gamma'\Rightarrow\Delta'$ are provable.
-Since $\Gamma'$ contains $A\to A^{(i)}$ a cut followed by a contraction gives that $A\to (C\to D), \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}, \Gamma'\Rightarrow\Delta', A^{(i)}$ is provable for each $i$ and a cut with $C\to D\Rightarrow A\to (C\to D)$ gives that $C\to D, \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}, \Gamma'\Rightarrow\Delta', A^{(i)}$ is provable.
-Then a number of cuts with $C^{(i)}\to D^{(i)}\Rightarrow A^{(i)}\to (C^{(i)}\to D^{(i)})$ gives for each $j\in I=\{1,\dots,n\}$ (w.l.o.g.
-since $I$ was just chosen as just some arbitrary finite index set) $$C\to D, \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\leq j}, \{C^{(i)}\to D^{(i)}\}_{i> j}, \Gamma'\Rightarrow\Delta', A^{(i)}$$ Note that for each $j$ we have
-
-		\begin{center}
-			\hspace*{-.7cm}
-			\AxiomC{$\stackanchor{$\{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\leq j},$}{$C\to D,\{C^{(i)}\to D^{(i)}\}_{i> j},$} \Gamma'\Rightarrow\Delta', A^{(j)}$}
-			\AxiomC{$\stackanchor{$\{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i< j},$}{$C\to D,\{C^{(i)}\to D^{(i)}\}_{i\geq j}$}\Gamma'\Rightarrow\Delta'$}
-			\RightLabel{L$\to$}
-			\BinaryInfC{$\{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i< j+1},C\to D,\{C^{(i)}\to D^{(i)}\}_{i\geq j+1},\Gamma'\Rightarrow\Delta'$}
-			\DisplayProof
-		\end{center}
-		and we have already established the right argument for $j=0$ and the left argument for every $j\in I$.
-Therefore applying the above iteratively $n$ times we obtain $C\to D, \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}\Gamma'\Rightarrow\Delta'$ and finally
-		\begin{center}
-			\AxiomC{$\stackanchor{$A\to (C\to D),\Gamma',$}{$\{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I},$} \Rightarrow \Delta', A$}
-			\AxiomC{$\stackanchor{$C\to D, \Gamma',$}{$\{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}$}\Rightarrow \Delta' $}
-			\RightLabel{L$\to$}
-			\BinaryInfC{$A\to (C\to D), \{A^{(i)}\to(C^{(i)}\to D^{(i)})\}_{i\in I}, \Gamma' \Rightarrow \Delta' $}
-			\DisplayProof
-		\end{center}
-		which shows that $S'$ is indeed provable.
-
-		3.
-$A = C\to D$  with $C, D$ atomic, $B$ atomic.
-Then by induction hypothesis $(C^{(i)}\to D^{(i)})\to B, \Gamma'\Rightarrow\Delta', C^{(j)}\to D^{(j)}$ and $B, \Gamma'\Rightarrow\Delta'$ are provable, where $C^{(i)} = C^{(j)}$ and $D^{(i)} = D^{(j)}$ so a single application of L$\to$ gives us $S'$.
-
-		Finally suppose the last rule of the proof is R$\to$, i.e.
+	This reduces our task to embedding just the implicational sequents from intuitionistic into classical logic. The basic idea is as follows: When intuitionistically proving a sequent via R$\to$
 
@@
 	\end{center}
-		Then $A, B$ are atoms.
-Let $i$ be some new index.
-Let $\Gamma''$ be $\Gamma$ where every non-primed atom $X$ has been renamed to $X^{(i)}$.
-Then $A^{(i)}, \Gamma''\Rightarrow B^{(i)}$ is provable and thus so is $\bigwedge\Gamma''\Rightarrow A^{(i)}\to B^{(i)}$.
-Since $\Gamma'\Rightarrow \bigwedge\Gamma''$ is a tautology a single cut yields that $S'$ is provable.
+
+	the formulas in $\Delta$ may not play a role in the derivation of $A, \Gamma\Rightarrow B$ whereas in classical logic they can
+
+	\begin{center}
+		\AxiomC{$A,\Gamma\Rightarrow \Delta, B$}
+		\RightLabel{R$\to$}
+		\UnaryInfC{$\Gamma\Rightarrow \Delta, A\to B$}
+		\DisplayProof
+	\end{center}
+
+	We resolve this dichotomy by using a modified formula $A'\to B'$ and a modified antecedent $\Gamma'$ such that the formulas in $\Delta$ are useless for proving $A', \Gamma'\Rightarrow B'$.
+Furthermore for an implicational formula we know all possible ways R$\to$ can be applied, that is in deriving formulas of type $C\to D$ in the succedent and more indirectly $(C\to D)\to E$ in the antecedent.
+An additional complication arises due to the possibility that multiple R$\to$ are nested.
+Taking all these points into consideration we define the following:
+
+	\begin{definition}\label{translation}
+		Suppose we are given an implicational sequent $\Gamma\to\Delta$ of the form $$\stackanchor{$\{(C_i\to D_i)\to E_i\}_{i\in I}, \{C_j\}_{j\in J},$}{$\{C_k\to D_k\}_{k\in K},\{C_l\to(D_l\to E_l)\}_{l\in L}$}\Rightarrow \{C_m\}_{m\in M}, \{C_n\to D_n\}_{n\in N}$$
+
+		Let $\mathcal{A}$ denote the atoms in the sequent.
+For each $A\in\mathcal{A}-\{\bot\}$ and $S\subseteq I\cup N$ introduce a new atom $A^{S}$, which is a variant of $A$ for the subproof before the applications of R$\to$ associated with $S$.
+Have $\bot^S = \bot$ in the following:
+		\begin{align*}
+			\Gamma' = &\:\{(C_i^{S\cup \{i\}}\to D_i^{S\cup\{i\}})\to E_i^{S}\:|\: S\subset I\cup N - i\}
+			\\&\cup \{C_{j}^S\:|\:j\in J, S\subset I\cup N\}
+			\\&\cup \{C_{k}^S\to D_k^S\:|\:k\in K, S\subset I\cup N\}
+			\\&\cup \{C_l^S\to(D_l^S\to E_l^S)\:|\: l\in L, S\subset I\cup N\}
+			\\&\cup \{A^S\to A^{S\cup \{i\}}\:|\:A\in\mathcal{A}, S\subsetneq I\cup N, i\in I\cup N - S\}\\
+			\Delta' = &\:\{C_m^\emptyset\:|\:m\in M\}\cup \{C_n^{\{n\}}\to D_n^{\{n\}}\:|\: n\in N\}
+		\end{align*}
+	\end{definition}
+
+	We claim that the sequent $\Gamma'\to\Delta'$ is classically provable iff $\Gamma\to\Delta$ is intuitionistically provable.
+Note that the blowup of the sequent can be quite large, i.e.
+$O(2^{|I| + |N|})$.
+We will see that this will be far more manageable in the first-order case where we can do encodings via terms.
+Before proving our claim we shall examine some examples.
+	\begin{example}
+		Consider \begin{enumerate}
+			\item $(A\to \bot)\to \bot\Rightarrow A$
+			\item $\Rightarrow A\to B, B\to A$
+			\item $(A\to B)\to C, A\to B\Rightarrow C$
+			\item $N\to (A\to B), (A\to B)\to C, (N\to C)\to D\Rightarrow D$ 
+		\end{enumerate}
+	which have the translated versions, where the sets $S$ have been encoded as natural numbers,
+	\begin{enumerate}
+		\item $(A^1\to \bot)\to \bot, A^0\to A^1\Rightarrow A^0$
+		\item $A^0\to A^1, A^1\to A^3, B^0\to B^1, B^1\to B^3\Rightarrow A^1\to B^1, B^2\to A^2$
+		\item $\stackanchor{$(A^1\to B^1)\to C^0, A^0\to B^0, A^1\to B^1$}{$A^0\to A^1, B^0\to B^1, C^0\to C^1$}\Rightarrow C^0$
+		\item $\stackanchor{$\{N^i\to (A^i\to B^i)\:|\:i\},(A^1\to B^1)\to C^0, (A^3\to B^3)\to C^2$}{$(N^2\to C^2)\to D^0, (N^3\to C^3)\to D^1, \{X^0\to X^i, X^i\to X^3\:|\:i, X\}$}\Rightarrow D^0$ 
+	\end{enumerate}
+	We can see that the sequents have become quite a lot more complex.
+As we would expect the first two, which stem from non-intuitionistic tautologies, are not provable.
+In the first case it boils down to not being able to prove $A_0$ from $A_1$.
+In the second to not being able to prove $A^2$ from $A^1$ or $B^2$ from $B^1$ ($A^3/B^3$ would be provable!).
+
+	The third example features a minimal intuitionistically valid sequent which requires an application or R$\to$.
+It remains valid,  as $$(A^1\to B^1)\to C^0, A^1\to B^1\Rightarrow C^0$$ is provable.
+The final sequent is more involved requiring applications of R$\to$.
+Let's look at the proof of the modified version (with some implicit weakenings):
+	\begin{center}
+		\hspace*{-2cm}
+		\AxiomC{}
+		\RightLabel{Ax}
+		\UnaryInfC{$N^3\to N^3$}
+		\AxiomC{}
+		\RightLabel{Ax}
+		\UnaryInfC{$N^2\to N^2$}
+		\RightLabel{L$\to$}
+		\BinaryInfC{$N^2\to N^3, N^2\Rightarrow N^3$}
+		\AxiomC{}
+		\RightLabel{Ax}
+		\UnaryInfC{$A^3\to B^3\Rightarrow A^3\to B^3$}
+		\RightLabel{L$\to$}
+		\BinaryInfC{$N^2\to N^3, N^3\to (A^3\to B^3), N^2\Rightarrow A^3\to B^3$}
+		\AxiomC{}
+		\RightLabel{Ax}
+		\UnaryInfC{$C^2\Rightarrow C^2$}
+		\RightLabel{L$\to$}
+		\BinaryInfC{$N^2\to N^3, N^3\to (A^3\to B^3), (A^3\to B^3)\to C^2, N^2\Rightarrow C^2$}
+		\RightLabel{R$\to$}
+		\UnaryInfC{$N^2\to N^3, N^3\to (A^3\to B^3), (A^3\to B^3)\to C^2\Rightarrow N^2\to C^2$}
+		\AxiomC{}
+		\RightLabel{Ax}
+		\UnaryInfC{$D^0\Rightarrow D^0$}
+		\RightLabel{L$\to$}
+		\BinaryInfC{$N^2\to N^3, N^3\to (A^3\to B^3), (A^3\to B^3)\to C^2, (N^2\to C^2)\to D^0\Rightarrow D^0$}
+		\DisplayProof
+	\end{center}
+	We can see that above R$\to$ as applied to $N^2\to C^2$ we only see atoms indexed with $2$ and $3$.
+	\end{example}
+
+	We now proceed to show that this is indeed a valid embedding of implicational sequents into classical logic.
+To that end we first show that $\Gamma\to\Delta$ and $\Gamma'\to\Delta'$ are equiprovable and then that the existence of a classical proof implies the existence of an intuitionistic proof (the other direction is trivial).
+
+	\begin{lemma}
+		Let $\Gamma\to\Delta$ be an implicational sequent. Then $\Gamma\to\Delta$ and $\Gamma'\to\Delta'$ are equiprovable.
+	\end{lemma}
+
+	\begin{proof}
+		We adopt the notation from~\ref{translation}.
+Suppose $\Gamma'\to\Delta'$ are provable.
+Consider some proof and rename each atom $A^S$ to $A$.
+Each formula in the root sequent is then either part of $\Gamma\to\Delta$ or an intuitionistic tautology.
+Thus modulo some contractions and simple cuts we obtain a valid proof of $\Gamma\to\Delta$.
+The other direction is much more involved.
+
+		Suppose we have a valid proof of $\Gamma\to\Delta$.
+We proceed by induction on the proof height.
+If the proof consists of a single Ax/L$\bot$ we can prove $\Gamma'\to\Delta'$ by Ax/L$\bot$ as well.
+Suppose the last rule is L$\to$:
+		\begin{center}
+			\AxiomC{$\Gamma\Rightarrow\Delta, A$}
+			\AxiomC{$B, \Gamma\setminus\{A\to B\}\Rightarrow\Delta$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$\Gamma\Rightarrow\Delta$}
+			\DisplayProof
+		\end{center}
+		with $A\to B\in\Gamma$. There are 3 cases:
+
+		1.
+$A$ and $B$ are both atoms.
+Then by induction hypothesis $\Gamma'\Rightarrow\Delta', A^\emptyset$ and $\{B^S|S\subset I\cup N\}\cup \Gamma'\setminus\{A^S\to B^S|S\}\Rightarrow\Delta'$, i.e.
+$\{B^S|S\}\cup \Gamma'\Rightarrow\Delta'$ after weakening, are provable.
+Observe that for any $S\subset I\cup N$ $$\{A^T\to A^{T\cup \{i\}}\:|\:T\subsetneq S, i\in S - T\}, A^\emptyset, A^S\to B^S\Rightarrow B^S$$ is valid and $\{A^T\to A^{T\cup \{i\}}\:|\:T\subsetneq S, i\in S - T\}\subset\Gamma'$, $A^S\to B^S\in\Gamma$, i.e.
+we can prove $\Gamma'\Rightarrow\Delta', B^S$ via cut and weakening.
+From this and $\{B^S|S\}\cup \Gamma'\Rightarrow\Delta'$ we can obtain $\Gamma'\to\Delta'$ via $2^{|I| + |N|}$ cuts.
+(We don't really need cuts, but this is conceptually simpler and generalizes better to 2.
+and 3.)
+
+		2.
+$B = C\to D$ where $A, C, D$ are atoms.
+By induction hypothesis $\Gamma'\Rightarrow\Delta', A^\emptyset$ and $\{C^S\to D^S|S\}\cup \Gamma'\Rightarrow\Delta'$ are provable.
+The argument is then analogous to 1 but instead of $A^S\to B^S$ we use $A^S\to(C^S\to D^S)$.
+
+		3.
+$A = C\to D$ where $B, C, D$ are atoms.
+Let $i\in I$ be the index of $A\to B$, i.e.
+$A\to B = (C_i\to D_i)\to E_i$.
+Then by induction hypothesis $\Gamma'\to\Delta', C^{\{i\}}\to D^{\{i\}}$ and $\{B^S|S\}\cup \Gamma'\Rightarrow\Delta'$ are provable.
+Similarly to 1 observe that for any $S\subset I\cup N$ $$\{B^T\to B^{T\cup \{j\}}\:|\:T\subsetneq S, j\in S - T\}, C^{\{i\}}\to D^{\{i\}}, (C^{\{i\}}\to D^{\{i\}})\to B^\emptyset\Rightarrow B^S$$
+		is valid and we have $\{B^T\to B^{T\cup \{j\}}\:|\:T\subsetneq S, j\in S - T\}\subseteq \Gamma'$ and also $(C^{\{i\}}\to D^{\{i\}})\to B^\emptyset\in\Gamma'$. The rest is analogous.
+
+		Finally suppose the last rule of the proof of $\Gamma\to\Delta$ is R$\to$:
+		\begin{center}
+			\AxiomC{$\Gamma, A\Rightarrow B$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$\Gamma\Rightarrow\Delta$}
+			\DisplayProof
+		\end{center}
+		with $A\to B\in\Delta$.
+Let $i$ be the index of $A\to B$.
+Let $\Gamma''$ be $\Gamma'$ in which each formula containing some $X^S$ with $i\notin S$ is deleted evey remaining atom $X^S$ is replaced by $X^{S - i}$.
+Then by induction hypothesis $$\Gamma'', \{A^S\:|\:S\subseteq I\cup N - i\}\to B^\emptyset$$ is provable.
+Now rename every atom $X^S$ in the proof of this sequent with $X^{S\cup\{i\}}$ and perform left weakening using the formulas in $\Gamma'$ that contain some $X^S$ where $i\notin S$.
+This gives us a proof of $\Gamma', \{A^{S\cup\{i\}}S\subset I\cup N - i\}\to B^{\{i\}}$.
+Finally observe that for all $S\subset I\cup N - i$ $$\{A^{T\cup\{i\}}\to A^{T\cup \{i, j\}}\:|\:T\subsetneq S, j\in S - T\}, A^{\{i\}}\Rightarrow A^{S\cup\{i\}}$$ is valid and $\{A^{T\cup\{i\}}\to A^{T\cup \{i, j\}}\:|\:T\subsetneq S, j\in S - T\}\subset\Gamma'$.
+Therefore using cut and weakening we obtain $\Gamma', A^{\{i\}}\Rightarrow B^{\{i\}}$ and via R$\to$ and weakening $\Gamma'\Rightarrow\Delta'$.
+	\end{proof}
+
+	For the following lemma we first need a technical result:
+	\begin{lemma}
+		Let $\Gamma\Rightarrow\Delta$ be a provable implicational sequent. Then there exists $B\in\Delta\cup\{\bot\}$ such that $\Gamma\Rightarrow B$ is provable.
+	\end{lemma}
+
+	\begin{proof}
+		By induction on the proof height. If it consists of a single Ax, L$\bot$ or the last rule is R$\to$ the claim is clear. Suppose it ends in L$\to$
+		\begin{center}
+			\AxiomC{$\Gamma\Rightarrow\Delta, A$}
+			\AxiomC{$B, \Gamma\setminus\{A\to B\}\Rightarrow\Delta$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$\Gamma\Rightarrow\Delta$}
+			\DisplayProof
+		\end{center}
+
+		We apply the induction hypothesis to the first argument.
+If $\Gamma\Rightarrow B_\Delta$ is valid for some $B_\Delta\in\Delta\cup\{\bot\}$ we are done so suppose $\Gamma\Rightarrow A$.
+But then applying the induction hypothesis to the second argument we get
+		\begin{center}
+			\AxiomC{$\Gamma\Rightarrow A$}
+			\AxiomC{$B, \Gamma\setminus\{A\to B\}\Rightarrow B_\Delta$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$\Gamma\Rightarrow B_\Delta$}
+			\DisplayProof
+		\end{center}
 	\end{proof}
@@
 	\begin{lemma}
-		Let $S$ be an implicational sequent such that $S'$ is provable classically. Then $S'$ is provable intuitionistically.
+		Let $\Gamma\to\Delta$ be an implicational sequent. Suppose $\Gamma'\to\Delta'$ is classically provable. Then $\Gamma'\to\Delta'$ is intuitionistically provable.
 	\end{lemma}
@@
 	\begin{proof}
-		We proceed by induction on the proof height of the classical proof of $S'$.
-If the last rule is Ax or L$\bot$ (or L$\to$) the proof is already intuitionistically valid (or it follows directly from the induction hypothesis).
-Otherwise the proof ends in $n$ applications of R$\to$ where $n$ is chosen maximally, i.e.
+		Again we use the terminology from ~\ref{translation} and proceed by the induction on the number of R$\to$ in the proof of $\Gamma'\to\Delta'$ that have non-empty set of side formulas $\Delta$.
+If there are no such R$\to$ the proof is already intuitionistically valid.
+Suppose otherwise, i.e.
+the proof contains
 		\begin{center}
-			\AxiomC{}
-			\RightLabel{K}
-			\UnaryInfC{$A_1^{(1)},\dots, A_n^{(n)},\Gamma'\Rightarrow\Delta', B_1^{(1)},\dots B_n^{(n)}$}
+			\AxiomC{$\Gamma', A^{\{i\}}\Rightarrow \Delta'-(A^{\{i\}}\to B^{\{i\}}), B^{\{i\}}$}
 			\RightLabel{R$\to$}
-			\UnaryInfC{\vdots}
-			\RightLabel{R$\to$}
-			\UnaryInfC{$\Gamma'\Rightarrow\Delta', A_1^{(1)}\to B_1^{(1)},\dots, A_n^{(n)}\to B_n^{(n)}$}
+			\UnaryInfC{$\Gamma'\to\Delta'$}
 			\DisplayProof
 		\end{center}
-		with $\text{K }\neq\text{ R}\to$. If K $=$ L$\bot$ we have $\bot\in\Delta'$ and therefore we immediately have
-		\begin{center}
-			\AxiomC{}
-			\RightLabel{L$\bot$}
-			\UnaryInfC{$\Gamma'\Rightarrow\Delta', A_1^{(1)}\to B_1^{(1)},\dots, A_n^{(n)}\to B_n^{(n)}$}
-			\DisplayProof
-		\end{center}
-	For K $=$ Ax observe that the atomic formulas in $\Delta'$ and $\Gamma'$ are not primed.
-Therefore the tautology in Ax must be between some $A_k^{(k)}$ and $B_l^{(l)}$ or some formulas in $\Gamma'$ and $\Delta'$.
-In the latter case we just need a single application of $S'$.
-In the former if $k = l$ we get the intuitionistic proof
-	\begin{center}
-		\AxiomC{}
-		\RightLabel{Ax}
-		\UnaryInfC{$A_k^{(k)},\Gamma'\Rightarrow\Delta', B_k^{(k)}$}
-		\RightLabel{R$\to$}
-		\UnaryInfC{$\Gamma'\Rightarrow\Delta', A_1^{(1)}\to B_1^{(1)},\dots, A_n^{(n)}\to B_n^{(n)}$}
-		\DisplayProof
-	\end{center}
-	otherwise $k\neq l$ but $A^{(k)} = B^{(l)}$ can only be the case if $A^{(k)} = A^{(l)} = B^{(k)} = B^{(l)}$ and the above proof works thee same.
+		for some atomic $A, B$ and $i\in N$ with $\Delta - (A^{\{i\}}\to B^{\{i\}}) \neq\emptyset$.
+Consider $\overline\Gamma', \overline\Delta'$ obtained from $\Gamma', \Delta'$ by renaming every atom $X^S$ to $X^{S - i}$.
+Then this renaming applied to the proof produces one of $\overline\Gamma', A^\emptyset\Rightarrow\overline\Delta' - (A^\emptyset\to B^\emptyset), B^\emptyset$, i.e.
+after weakening  $$(\Gamma\cup\{A\})'\Rightarrow(\Delta\cup\{B\} - (A\to B))'$$ which by inductions hypothesis is intuitionistically valid.
 	\end{proof}
@@
 
-	\begin{example}
-		Consider Pierce's law $P := (A\to B)\to A\Rightarrow A$.
-This is a classic non-intuitionistic tautology.
-Note that it is already implicational.
-$P'$ is $$(A'\to B')\to A, A\to A', B\to B', \omega\to A', \omega\to B'\Rightarrow A$$ As proven this sequent is not classically valid, $A = B = B' = \omega = 0$, $A' = 1$  provides a countermodel.
-
-		If we add $A\to B$ to the hypothesis set $P'$ is $$(A'\to B')\to A, A'\to B', A\to B, A\to A', B\to B', \omega\to A', \omega\to B'\Rightarrow A$$ which is classically and intuitionistically valid.
-Argueing in BHK terms, we can provide a terms of type $A'\to B'$ to $(A'\to B')\to A$ in order to obtain a term of type $A$, which is what is needed.
-	\end{example}
-
-	Intuitively the primed atoms may only be used in implications, i.e. as arguments to formulas of type $(A\to B)\to C$ but never as proper atoms, i.e. as arguments to $A\to B$.
-
 	\bibliographystyle{plain}

### 2022-05-03 Notes/A reverse double negation translation/RDNT.tex
Lemma 2.11 and 2.12

@@
 
-	\begin{lemma}
+	\begin{lemma}\label{equi}
 		Let $\Gamma\to\Delta$ be an implicational sequent. Then $\Gamma\to\Delta$ and $\Gamma'\to\Delta'$ are equiprovable.
@@
 	\begin{proof}
+
 		By induction on the proof height. If it consists of a single Ax, L$\bot$ or the last rule is R$\to$ the claim is clear. Suppose it ends in L$\to$
@@
 
-	\begin{lemma}
+	\begin{lemma}\label{classical}
 		Let $\Gamma\to\Delta$ be an implicational sequent. Suppose $\Gamma'\to\Delta'$ is classically provable. Then $\Gamma'\to\Delta'$ is intuitionistically provable.
@@
 	\begin{proof}
-		Again we use the terminology from ~\ref{translation} and proceed by the induction on the number of R$\to$ in the proof of $\Gamma'\to\Delta'$ that have non-empty set of side formulas $\Delta$.
-If there are no such R$\to$ the proof is already intuitionistically valid.
-Suppose otherwise, i.e.
-the proof contains
+		We adopt again the notation from~\ref{translation}. Suppose we have a classical proof of $\Gamma'\to\Delta'$. We give a transformation into an intuitionistic proof.
+
+		First observe that we can move applications of R$\to$ downward the proof tree, as long as the resulting right implication is not an active formula, i.e.
+$\Gamma\to\Delta, A\to B$ is replaced with $A, \Gamma\to \Delta, B$:
 		\begin{center}
-			\AxiomC{$\Gamma', A^{\{i\}}\Rightarrow \Delta'-(A^{\{i\}}\to B^{\{i\}}), B^{\{i\}}$}
+			\AxiomC{$A, C,\Gamma\Rightarrow\Delta, B, D$}
 			\RightLabel{R$\to$}
-			\UnaryInfC{$\Gamma'\to\Delta'$}
+			\UnaryInfC{$C,\Gamma\Rightarrow\Delta, A\to B, D$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$\Gamma\Rightarrow\Delta, A\to B, C\to D$}
+			\DisplayProof
+
+			$\downarrow$
+
+			\AxiomC{$A, C,\Gamma\Rightarrow\Delta, B, D$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$A,\Gamma\Rightarrow\Delta, B, C\to D$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$\Gamma\Rightarrow\Delta, A\to B, C\to D$}
+			\DisplayProof
+
+			\AxiomC{$C\to D, A,\Gamma\Rightarrow\Delta, B, D$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$ C\to D,\Gamma\Rightarrow\Delta, A\to B, D$}
+			\AxiomC{$A, D,\Gamma\Rightarrow\Delta, B$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$D,\Gamma\Rightarrow\Delta, A\to B$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$ C\to D,\Gamma\Rightarrow\Delta, A\to B$}
+			\DisplayProof
+
+			$\downarrow$
+
+			\AxiomC{$C\to D, A, \Gamma\Rightarrow\Delta, B, D$}
+			\AxiomC{$A, D,\Gamma\Rightarrow\Delta, B$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$ C\to D, A,\Gamma\Rightarrow\Delta, B$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$C\to D,\Gamma\Rightarrow\Delta, A\to B$}
 			\DisplayProof
 		\end{center}
-		for some atomic $A, B$ and $i\in N$ with $\Delta - (A^{\{i\}}\to B^{\{i\}}) \neq\emptyset$.
-Consider $\overline\Gamma', \overline\Delta'$ obtained from $\Gamma', \Delta'$ by renaming every atom $X^S$ to $X^{S - i}$.
-Then this renaming applied to the proof produces one of $\overline\Gamma', A^\emptyset\Rightarrow\overline\Delta' - (A^\emptyset\to B^\emptyset), B^\emptyset$, i.e.
-after weakening  $$(\Gamma\cup\{A\})'\Rightarrow(\Delta\cup\{B\} - (A\to B))'$$ which by inductions hypothesis is intuitionistically valid.
+		Note that in the L$\to$ case it is not a problem if $A\to B$ was not introduced via R$\to$ in one of the branches but via Ax, we then just have to modify Ax accordingly.
+
+		Next consider the case where R$\to$ is followed by L$\to$ where the derived sequent is active, in that case we can move both rules downward: 
+		\begin{center}
+			\AxiomC{$(A\to B)\to C, A, D, \Gamma\Rightarrow\Delta, B, E$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$(A\to B)\to C, D, \Gamma\Rightarrow\Delta, A\to B, E$}
+			\AxiomC{$C, D,\Gamma\Rightarrow\Delta, E$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$(A\to B)\to C, D, \Gamma\Rightarrow\Delta, E$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$(A\to B)\to C, \Gamma\Rightarrow\Delta, D\to E$}
+			\DisplayProof
+
+			$\downarrow$
+
+			\AxiomC{$(A\to B)\to C, A, D, \Gamma\Rightarrow\Delta, B, E$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$(A\to B)\to C, A, \Gamma\Rightarrow\Delta, B, D\to E$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$(A\to B)\to C, \Gamma\Rightarrow\Delta, A\to B, D\to E$}
+			\AxiomC{$C, D,\Gamma\Rightarrow\Delta, E$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$C, \Gamma\Rightarrow\Delta, D\to E$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$(A\to B)\to C, \Gamma\Rightarrow\Delta, D\to E$}
+			\DisplayProof			
+		\end{center}
+		and analogously for L$\to$.
+		\begin{center}
+			\hspace*{-4cm}
+			{\footnotesize
+			\AxiomC{$(A\to B)\to C, A, D\to E, \Gamma\Rightarrow\Delta, B, D$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$(A\to B)\to C, D\to E, \Gamma\Rightarrow\Delta, A\to B, D$}
+			\AxiomC{$C, D\to E,\Gamma\Rightarrow\Delta, D$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$(A\to B)\to C, D\to E, \Gamma\Rightarrow\Delta, D$}
+			\AxiomC{$(A\to B)\to C, A, E, \Gamma\Rightarrow\Delta, B$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$(A\to B)\to C, E, \Gamma\Rightarrow\Delta, A\to B$}
+			\AxiomC{$C, E,\Gamma\Rightarrow\Delta$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$(A\to B)\to C, E, \Gamma\Rightarrow\Delta$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$(A\to B)\to C, D\to E, \Gamma\Rightarrow\Delta$}
+			\DisplayProof
+		}
+
+			$\downarrow$
+
+			\hspace*{-3cm}
+			{\footnotesize
+				\AxiomC{$(A\to B)\to C, A, D\to E, \Gamma\Rightarrow\Delta, B, D$}
+				\AxiomC{$(A\to B)\to C, A, E, \Gamma\Rightarrow\Delta, B$}
+				\RightLabel{L$\to$}
+				\BinaryInfC{$(A\to B)\to C, A, D\to E, \Gamma\Rightarrow\Delta, B$}
+				\RightLabel{R$\to$}
+				\UnaryInfC{$(A\to B)\to C, D\to E, \Gamma\Rightarrow\Delta, A\to B$}
+				\AxiomC{$C, D\to E,\Gamma\Rightarrow\Delta, D$}
+				\AxiomC{$C, E,\Gamma\Rightarrow\Delta$}
+				\RightLabel{L$\to$}
+				\BinaryInfC{$C, D\to E, \Gamma\Rightarrow\Delta$}
+				\RightLabel{L$\to$}
+				\BinaryInfC{$(A\to B)\to C, D\to E, \Gamma\Rightarrow\Delta$}
+				\DisplayProof
+			}	
+		\end{center}
+		where the same remark as before applies. This allows us to move the applications of R$\to$
+		\begin{center}
+			\AxiomC{$C^S, \Gamma\Rightarrow \Delta, D^S$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$\Gamma\Rightarrow \Delta, C^S\to D^S$}
+			\DisplayProof
+		\end{center}
+		such that
+		\begin{enumerate}[i.]
+			\item no application of R$\to$ below this one with corresponding $S'$ has $S'\supsetneq S$.
+			\item if $S\neq\{i\}$ with $i\in N$ then R$\to$ is directly followed by L$\to$ where the derived application is active, i.e.
+			\begin{center}
+				\AxiomC{$(C^{S\cup\{i\}}\to D^{S\cup\{i\}})\to E^S, C^{S\cup\{i\}}, \Gamma\Rightarrow \Delta, D^{S\cup\{i\}}$}
+				\RightLabel{R$\to$}
+				\UnaryInfC{$(C^{S\cup\{i\}}\to D^{S\cup\{i\}})\to E^S,\Gamma\Rightarrow \Delta, C^{S\cup\{i\}}\to D^{S\cup\{i\}}$}
+				\AxiomC{$\Gamma, E^S\Rightarrow \Delta$}
+				\RightLabel{L$\to$}
+				\BinaryInfC{$(C^{S\cup\{i\}}\to D^{S\cup\{i\}})\to E^S,\Gamma\Rightarrow \Delta$}
+				\DisplayProof
+			\end{center}
+		\end{enumerate}
+
+		We now proceed by induction on the number of R$\to$ in this transformed proof. There are three cases
+
+		1. There are no R$\to$ left, then we are done as the proof is already intuitionistically valid.
+
+		2. The last rule is R$\to$, i.e. of the form 
+		\begin{center}
+			\AxiomC{$C^{\{i\}}, \Gamma\Rightarrow \Delta, D^{\{i\}}$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$\Gamma\Rightarrow \Delta, C^{\{i\}}\to D^{\{i\}}$}
+			\DisplayProof
+		\end{center}
+		for some $i\in N$.
+If $C^{\{i\}}, \Gamma\Rightarrow D^{\{i\}}$ perform a renaming in the above proof to one of $\{C^S|S\}, \Gamma\Rightarrow D$.
+Then by induction hypothesis we obtain an intuitionistic proof of that and reverting the transformation we obtain one of $C^{\{i\}}, \Gamma\Rightarrow D^{\{i\}}$.
+But then we can apply R$\to$ to obtain $\Gamma\Rightarrow \Delta, C^{\{i\}}\to D^{\{i\}}$.
+Otherwise $\Gamma\Rightarrow\Delta$ and we are done by weakening.
+
+		3. The last two rules are of the form
+		\begin{center}
+			\AxiomC{$(C^{S\cup\{i\}}\to D^{S\cup\{i\}})\to E^S, C^{S\cup\{i\}}, \Gamma\Rightarrow \Delta, D^{S\cup\{i\}}$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$(C^{S\cup\{i\}}\to D^{S\cup\{i\}})\to E^S,\Gamma\Rightarrow \Delta, C^{S\cup\{i\}}\to D^{S\cup\{i\}}$}
+			\AxiomC{$\Gamma, E^S\Rightarrow \Delta$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$(C^{S\cup\{i\}}\to D^{S\cup\{i\}})\to E^S,\Gamma\Rightarrow \Delta$}
+			\DisplayProof
+		\end{center}
+		then again if $C^{S\cup\{i\}}, \Gamma\Rightarrow D^{S\cup\{i\}}$ is provable perform a renaming to obtain one of $\{C^S|S\}, \Gamma\Rightarrow D$, apply the induction hypothesis, and revert the renaming to obtain an intuitionistically valid proof of $(C^{S\cup\{i\}}\to D^{S\cup\{i\}})\to E^S, C^{S\cup\{i\}}, \Gamma\Rightarrow D^{S\cup\{i\}}$ which we can then use in the above.
+Otherwise we again obtain that $\Gamma\to\Delta$ is provable on which we just need to perform weakening.
 	\end{proof}
 
+	From~\ref{equi} and~\ref{classical} we finally obtain the result
+
 	\begin{theorem}
-		A sequent $S$ in intuitionistically provable if and only if $S^\circ_\to$ is classically provable.
+		A sequent $S$ in intuitionistically provable if and only if $S'_\to$ is classically provable.
 	\end{theorem}

### 2022-06-23 Notes/A reverse double negation translation/RDNT.tex
QBF

@@
 
-	 \begin{example}
-	 	Consider the formula $((A\wedge B)\to A) \to (B\vee A)$. Then its syntax tree is
-	 	\begin{center}
-	 		\tikz[scale=1, baseline=-5mm, thick]{\Tree [.\node[draw]{$((A\to B)\to A)\to (B\vee A)$}; [.\node[draw]{$(A\to B)\to A$}; [.\node[draw]{$A\to B$}; [.\node[draw]{$A$}; ] [.\node[draw]{$B$}; ] ] [.\node[draw]{$A$}; ] ] [.\node[draw]{$B\vee A$}; [.\node[draw]{$B$}; ] [.\node[draw]{$A$};] ] ]}
-	 	\end{center}
-	 \end{example}
-
  	We are now ready to introduce our calculi for intuitionistic and classical logic:
@@
 	\begin{definition}
-			A \textit{sequent} $A\Rightarrow B$ consists of two multisets of formulas $A, B$.
-The position of $A_i$ in $A_1,\dots, A_n\Rightarrow B_1,\dots, B_m$ is $0i$ and that of $B_i$ is $1i$.
-A formula $\varphi$ occurs at $p = (ji, p')$ if $\varphi$ occurs at $p'$ in $A_i$ if $j=0$ and in $B_i$ if $j=1$.
-The following rules make up our classical calculus $\mathbf{GC}$, corresponding roughly to the propositional fragment of \textbf{G3c} from~\cite{basicprooftheory}:\\
+			A \textit{sequent} $\Gamma\Rightarrow \Delta$ consists of two multisets of formulas $\Gamma, \Delta$.
+The following rules make up our classical calculus $\mathbf{GCp}$, corresponding roughly to the propositional fragment of \textbf{G3c} from~\cite{basicprooftheory}:\\
 		\begin{center}
@@
 
-		To obtain our intuitionistic calculus \textbf{GI} (corresponding to \textbf{m-G3i} from~\cite{basicprooftheory}) substitute R$\to$ with
+		To obtain our intuitionistic calculus \textbf{GIp} (corresponding to \textbf{m-G3i} from~\cite{basicprooftheory}) substitute R$\to$ with
 		\begin{center}
 			\AxiomC{$A, \Gamma\Rightarrow B$}
-			\RightLabel{R$\to'$}
+			\RightLabel{R$\to$}
 			\UnaryInfC{$\Gamma\Rightarrow\Delta, A\to B$}
@@
 
-		A \textit{proof} of a formula $\varphi$ in a calculus $\mathbf{G}$ is a labelled binary rooted tree in which every node is labelled with a pair of the name of a rule in $\mathbf{G}$, its rule, and a sequent, its sequent, such that
+		A \textit{proof} of a formula $\varphi$ in a calculus $\mathbf{G}$ is a labelled binary rooted tree in which every node is labelled with a sequent such that
 		\begin{itemize}
-			\item Every leaf is labelled with (Ax, $A,\Gamma\Rightarrow\Delta, A$) or (L$\bot$, $\bot, \Gamma\Rightarrow\Delta$) for some formula $A$, multi-sets of formulas $\Gamma,\Delta$.
-			\item The sequent of an interior node can be derived from sequents of its child nodes according to its rule.
+			\item Every leaf is labelled in accordance to Ax or L$\bot$.
+			\item The sequent of an interior node can be derived from sequents of its child nodes according to one of the rules.
 		\end{itemize}
 		We usually present a proof directly as a sequence of applications of rules rather than in a more usual tree presentation.
-Note that we implicitly permute the antecedent and succedent, i.e.
-we implicitly may apply a number of exchange rules at any point in the proof.
 	\end{definition}
@@
 	We now proceed to show that this is indeed a valid embedding of implicational sequents into classical logic.
-To that end we first show that $\Gamma\to\Delta$ and $\Gamma'\to\Delta'$ are equiprovable and then that the existence of a classical proof implies the existence of an intuitionistic proof (the other direction is trivial).
+To that end we first show that $\Gamma\to\Delta$ and $\Gamma'\to\Delta'$ are equiprovable and then that the existence of a classical proof of $\Gamma'\to\Delta'$ implies the existence of an intuitionistic proof of $\Gamma'\to\Delta'$ (the other direction is trivial).
 
@@
 		for some $i\in N$.
-If $C^{\{i\}}, \Gamma\Rightarrow D^{\{i\}}$ perform a renaming in the above proof to one of $\{C^S|S\}, \Gamma\Rightarrow D$.
+If $C^{\{i\}}, \Gamma\Rightarrow D^{\{i\}}$ is provable perform a renaming in its proof to one of $\{C^S|S\}, \Gamma\Rightarrow D$.
 Then by induction hypothesis we obtain an intuitionistic proof of that and reverting the transformation we obtain one of $C^{\{i\}}, \Gamma\Rightarrow D^{\{i\}}$.
@@
 
+	\section{First-order logic}
+
+	Let us first introduce the calculi and recapitulate some facts. First we need to introduce our term language
+	\begin{definition}
+		A \textit{Signature} consists of
+		\begin{enumerate}
+			\item a finite set $S_f$ of function symbols $f_1, \dots, f_n$
+			\item a finite set $S_R$ of relation symbols $R_1,\dots, R_m$
+			\item  a function ar$: S_f\cup S_R\to \mathbb{N}$ assigning to each symbol its arity (possibly 0).
+		\end{enumerate}
+	\end{definition}
+
+	\noindent Usually the signature we be implicit and contain a binary relation symbol $=$ for equality.
+If explicitly stated we denote it by $$\{f_1/\text{ar}(f_1),\dots,f_n/\text{ar}(f_n), R_1/\text{ar}(R_1).\dots,R_m/\text{ar}(R_m)\}.$$
+	Fix a countably infinite collection of free variables $a, b, c\dots$ and bound variables $x, y, z\dots$.
+	Note that the following definitions could be simplified by not distinguishing between bound and free variables.
+However this has some other drawbacks and in particular when defining a sequent calculus for first-order logic having disjoint sets of bound and free variables is convenient.
+
+	\begin{definition}
+		A \textit{Semiterm} is one of the following
+		\begin{enumerate}
+			\item A bound or free variable
+			\item $f(t_1, \dots, t_n)$ given semiterms $t_1,\dots,t_n$ and an $n$-ary function symbol $f$.
+		\end{enumerate}
+		A \textit{Term} is defined as above but without allowing bound variables.
+We denote the set of bound variables as $\mathcal{V}_b$, of free variables as $\mathcal{V}_f$ and the set of terms as $\mathcal{T}$.
+	\end{definition}
+
+	\begin{definition}
+		A \textit{Semiformula} is one of the following
+		\begin{enumerate}
+			\item $R(t_1, \dots, t_n), t_1 = t_2, \bot, \top$ given semiterms $t_1,\dots,t_n$ and an $n$-ary function symbol $R$.
+			\item $\forall x A, \exists x A, A\wedge B, A\vee B, A\to B$ given Semiformulas $A, B$ and a bound variable $x$.
+		\end{enumerate}
+		A \textit{Formula} is semiformula in which every occurrence of a bound variable is guarded by a quantifier. A formula created by the first rule is called \textit{atomic}.
+	\end{definition}
+
+	\begin{definition}
+		A \textit{Substitution} is a function $\sigma: \mathcal{V}_\text{f}\to\mathcal{T}$. For a semiterm $t$ we define $t\sigma$ as follows:
+		\begin{itemize}
+			\item $x\sigma = x$ for $x\in \mathcal{V}_b$
+			\item $a\sigma = \sigma(a)$ for $a\in \mathcal{V}_f$
+			\item $f(t_1,\dots,t_n)\sigma = f(t_1\sigma,\dots,t_n\sigma)$ for an $n$-ary function symbol $f$ and terms $t_1,\dots,t_n$.
+		\end{itemize}
+		Analogously we define $F\sigma$ for formulas $F$.
+Given free variables $a_1,\dots,a_n$ and terms $t_1,\dots,t_n$ we define a special substitution $[t_1/a_1,\dots,t_n/a_n]$ as $$\sigma(a) = \begin{cases}
+			t_i & \text{if } a = a_i \text{ for } i \in \{1\dots n\}\\
+			a & \text{else}
+		\end{cases}$$this notion can be naturally extended where $a_i$ are arbitrary terms.
+	\end{definition}
+
+	We now extend our calculi from the previous section to be able to reason about terms. We obtain roughly calculi G3c and m-G3i from~\cite{basicprooftheory}. 
+	\begin{definition}
+	Obtain \textbf{GC} by extending \textbf{GCp} by following quantifier and equality rules:\\
+
+	\begin{tabular}{lll}
+		\AxiomC{$A[t/x], \Gamma\Rightarrow\Delta$}
+		\RightLabel{L$\forall$}
+		\UnaryInfC{$\forall xA, \Gamma\Rightarrow \Delta$}
+		\DisplayProof&
+		\AxiomC{$\Gamma\Rightarrow\Delta, A[a/x]$}
+		\RightLabel{R$\forall$}
+		\UnaryInfC{$\Gamma\Rightarrow \Delta, \forall xA$}
+		\DisplayProof&
+		\\&&\\
+		\AxiomC{$A[a/x], \Gamma\Rightarrow\Delta$}
+		\RightLabel{L$\exists$}
+		\UnaryInfC{$\exists xA, \Gamma\Rightarrow \Delta$}
+		\DisplayProof&
+		\AxiomC{$\Gamma\Rightarrow\Delta, A[t/x]$}
+		\RightLabel{R$\exists$}
+		\UnaryInfC{$\Gamma\Rightarrow \Delta, \exists xA$}
+		\DisplayProof&
+		\\&&\\
+		\AxiomC{$t = t, \Gamma\Rightarrow \Delta$}
+		\RightLabel{Ref}
+		\UnaryInfC{$\Gamma\Rightarrow \Delta$}
+		\DisplayProof&
+		\AxiomC{$s = t, P[s/x], P[t/x], \Gamma\Rightarrow \Delta$}
+		\RightLabel{Ref}
+		\UnaryInfC{$s = t, P[s/x],\Gamma\Rightarrow \Delta$}
+		\DisplayProof&
+		\\&&\\
+	\end{tabular}
+
+	where in R$\forall$ and L$\exists$ $a$ is a free variable not occurring in the sequent otherwise. Again for \textbf{GI} we modify R$\to$ as before and R$\forall$ to
+	\begin{center}
+		\AxiomC{$\Gamma\Rightarrow A[a/x]$}
+		\RightLabel{R$\forall_1$}
+		\UnaryInfC{$\Gamma\Rightarrow\Delta, \forall xA$}
+		\DisplayProof
+	\end{center}
+	\end{definition}
+
+	All theorems for cut elimination etc. from the previous section apply.
+
 	\bibliographystyle{plain}

### 2022-06-23 Notes/Embedding intuitionistic into classical logic/EIICL.tex
QBF

\documentclass[a4paper,12pt]{report}

\title{Embedding intuitionistic into classical logic}
\author{Alexander Pluska}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{ wasysym }
\usepackage{ stmaryrd }
\usepackage{ mathpartir }
\usepackage{bussproofs}
\usepackage{enumerate}
\usepackage{tikz-qtree}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usepackage{stackengine}
\usepackage{bbold}
\usepackage{footmisc}

\usepackage{hyperref}% http://ctan.org/pkg/hyperref
\usepackage{cleveref}% http://ctan.org/pkg/cleveref

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}[section]
\theoremstyle{definition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{definition}
\newtheorem{remark}[theorem]{Remark}

\DeclareMathOperator*{\argmax}{arg\,max}

\begin{document}

	\maketitle

	\begin{abstract}
		We present an embedding of intuitionistic into classical logic, a quasi-inverse to the famous double-negation translation, i.e. a procedure that for each formula $\varphi$ gives a formula $\varphi_c$ such that $\varphi$ is intuitionistically valid if and only if $\varphi_c$ is classically valid. We consider both the propositional and predicate case. Since many of the more involved results for the more interesting predicate scenario involve similar arguments as in the propositional case we shall develop them simultaneously.
	\end{abstract}

	\tableofcontents

	\chapter{Preliminaries}

	First let us recapitulate the most important definitions and some facts about them. This will serve to fix our notation as well.

	\section{Syntax}

	\subsection{Propositional logic}

	Let us fix some countably infinite set of propositional variables $A, B, C\dots$.

	\begin{definition}
		A \textit{formula} and \textit{subformula} is defined inductively via the following rules:
		\begin{itemize}
			\item every propositional variable as well as a special symbol $\bot$ is a formula, such a formula is called \textit{atomic} and $A,B,C,\dots$ and $\bot$ are called \textit{atoms}.
			\item if $\varphi$ and $\psi$ are formulas then $\varphi\circ\psi$ is a formula for $\circ\in\{\wedge,\vee,\to\}$.
			\item every formula is a subformula of itself.
			\item for a formula $\chi = \varphi\circ\psi$ every subformula of $\varphi$ and $\psi$ is also a subformula of $\chi$.
		\end{itemize}
	We write $\neg \varphi$ for $\varphi\to \bot$.
	\end{definition}

	\subsection{Predicate logic}

	The analog for predicate logic is much more involved as we need to define the term language.

	\begin{definition}
		A \textit{signature} consists of
		\begin{itemize}
			\item a finite set $S_f$ of function symbols $f_1, \dots, f_n$
			\item a finite set $S_R$ of relation symbols $R_1,\dots, R_m$
			\item  a function ar$: S_f\cup S_R\to \mathbb{N}$ assigning to each symbol its arity (possibly 0).
		\end{itemize}
	\end{definition}

	\noindent Usually the signature we be left implicit. If explicitly stated we denote it by $$\{f_1/\text{ar}(f_1),\dots,f_n/\text{ar}(f_n), R_1/\text{ar}(R_1).\dots,R_m/\text{ar}(R_m)\}.$$
	Fix a countably infinite collection of free variables $a, b, c\dots$ and bound variables $x, y, z\dots$.
	Note that the following definitions could be simplified by not distinguishing between bound and free variables. However this has some other drawbacks and in particular when defining a sequent calculus for first-order logic having disjoint sets of bound and free variables is convenient.

	\begin{definition}
		A \textit{term} is defined inductively via the following rules
		\begin{itemize}
			\item each free variable is a term
			\item if $t_1,\dots,t_n$ are terms and $f$ an $n$-ary function symbol then $f(t_1, \dots, t_n)$ is a term.
		\end{itemize}
		A \textit{semiterm} is defined in the same way with the difference that bound variables are allowed in 1. Clearly every term is a semiterm. Analogously to before we can define \textit{subterms} and \textit{subsemiterms}.
	\end{definition}

	\begin{definition}
		A \textit{Substitution} is a function $\sigma$ from free variables to semiterms. For a semiterm $t$ we define $t\sigma$ as follows:
		\begin{itemize}
			\item $x\sigma = x$ if $x$ is bound.
			\item $a\sigma = \sigma(a)$ if $a$ is free.
			\item $f(t_1,\dots,t_n)\sigma = f(t_1\sigma,\dots,t_n\sigma)$.
		\end{itemize}
	\end{definition}

	\begin{definition}
		A \textit{formula} is defined inductively via the following rules
		\begin{itemize}
			\item $\bot$ is a formula.
			\item if $t_1,\dots,t_n$ are terms and $R$ and an $n$-ary relation symbol then $R(t_1, \dots, t_n)$ is a formula.
			\item if $s, t$ are terms then $s = t$ is a formula.
		\end{itemize}	
		Formulas formed by the above rules are called \textit{atomic}.
		\begin{itemize}
			\item If $\varphi$ and $\psi$ are formulas then $\varphi\circ\psi$ is a formula for $\circ\in\{\wedge,\vee,\to\}$.
			\item If $\varphi$ is a formula and $a$ is a free variable occurring in $\varphi$ and $x$ is a bound variable not occurring in $\varphi$ then $Qx(\varphi[x/a])$ is a formula for $Q\in\{\exists,\forall\}$ where the \textit{substitution} $\varphi\sigma$ is defined inductively via
			\begin{itemize}
				\item $\bot\sigma = \bot$
				\item $R(t_1,\dots,t_n)\sigma = R(t_1\sigma,\dots, t_n\sigma)$
				\item $(s = t)\sigma = (s\sigma = t\sigma)$
				\item $(\varphi\circ\psi)\sigma = \varphi\sigma\circ\psi\sigma$ for $\circ\in\{\wedge,\vee,\to\}$.
				\item $(Qy\varphi)\sigma = Qy(\varphi\sigma)$ for $Q\in\{\exists,\forall\}$.
			\end{itemize}
			\item A formula with no free variables is a \textit{sentence}.
		\end{itemize}
	\end{definition}

	\section{Semantics}

	We now give meaning to formulas in classical and intuitionistic logic.

	\subsection{Propositional Logic}
	\subsubsection{Classical Semantics}

	\begin{definition}
		A \textit{valuation} $v$ is a function that assigns to each propositional variable $A$ a \textit{true value} $v(A)\in\{0, 1\}$. We extend $v$ to a function $v^*$ on formulas:
		\begin{itemize}
			\item $v^*(\bot) = 0$
			\item $v^*(A) = v(A)$ for each propositional variable $A$.
			\item $v^*(\varphi\wedge\psi) = \min(v^*(\varphi), v^*(\psi))$
			\item $v^*(\varphi\vee\psi) = \max(v^*(\varphi), v^*(\psi))$
			\item $v^*(\varphi\to \psi) = \max(1 - v^*(\varphi), v^*(\psi))$
		\end{itemize}
		$v$ is a \textit{model} for $\varphi$ if $v(\varphi) = 1$, we write $v\models\varphi$. $\varphi$ is \textit{satisfiable} if there exists a valuation which is a model. It is \textit{valid} if every valuation is a model and we write $\models \varphi$. We denote the set of valid formulas with CPC. We can furthermore define an entailment relation between sets of clauses, that is $\mathcal S\models_C \mathcal T$ if for every valuation $v$ with $v\models\bigwedge S$ we have $v\models\bigvee \mathcal T$.
	\end{definition}

	One notable property of classical logic is that a formula $\varphi$ is valid if and only if its negation, i.e. $\neg\varphi$ is not satisfiable. The same does not hold true for intuitionistic logic as we shall see.

	\subsubsection{Intuitionistic Semantics}

	\begin{definition}
		A \textit{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of nodes an a family of valuations $(v_w)_{w\in W}$ such that for $u\leq w$ we have $v_u(A)\leq v_u(A)$ for every propositional variable $A$, this is called the \textit{persistency condition}. We extend the valuations $v_u$ slightly differently to $v_u^\#$ than before, i.e. have
		\begin{itemize}
			\item $v_u^\#(\bot) = 0$
			\item $v_u^\#(\varphi) = v_u(A)$ if $\varphi$ is atomic.
			\item $v_u^\#(\varphi\wedge\psi) = \min(v_u^\#(\varphi), v_u^\#(\psi))$
			\item $v_u^\#(\varphi\vee\psi) = \max(v_u^\#(\varphi), v_u^\#(\psi))$
			\item $v_u^\#(\varphi\to \psi) = \min\{\max(1 - v_w^\#(\varphi), v_w^\#(\psi))\:|\:w\geq u\}$
		\end{itemize}
		That is $v_u^\#(\varphi\to \psi) = 1$ if at every world $w\geq u$ we have $v_w^\#(\varphi) = 0$ or $v_w^\#(\psi) = 1$. We say that $\varphi$ is satisfied at a node $u$ if $v_u^\#(\varphi) = 1$ and write $u\models\varphi$. If a formula is satisfied at every node $K$ is a \textit{model} for $\varphi$ and we write $K\models \varphi$. $\varphi$ is satisfiable if there exists a Kripke structure which is a model. It is \textit{valid} if every Kripke structure is a model and we write $\models \varphi$. We denote the set of valid formulas with IPC. As before we define an entailment relation $\models_I$ between sets of formulas.
	\end{definition}
	There are many classical tautologies which are not intuitionistically valid
	\begin{example}\label{LEMcounterexample}
		Consider the famous law of excluded middle $A\vee\neg A$. As a counter model consider a Kripke structure with two nodes $u\leq w$ and $v_u(A) = 0, v_w(A) = 1$. Then $u\not\models A$ and $u\not\models \neg A$, i.e. $u\not\models A\vee\neg A$. 
	\end{example}
	The following is a famous and important result due to Glivenko~\cite{glivenko1929}

	\begin{theorem}[Glivenko's theorem]
		A propositional formula $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
	\end{theorem}
	\noindent Note however that this difference between classical and intuitionistic logic is not reflected in satisfiability, only in validity:
	\begin{lemma}\label{propsat}
		A propositional formula $\varphi$ is intuitionistically satisfiable if and only if it is classically satisfiable.
	\end{lemma}

	\begin{proof}
		Every classical model can be viewed as a Kripke model with a single node. Therefore classical satisfiability trivially implies intuitionistic satisfiability.	On the other hand in every Kripke structure that is a model every node is a classical model so intuitionistic satisfiability implies intuitionistic satisfiability.
	\end{proof}

	In classical logic a proven strategy for showing that a formula valid is to consider its negation and show that it is not satisfiable. Due to the previous lemma this is insufficient for intuitionistic logic. There are formulas like $A\vee\neg A$ which are not a intuitionistic tautology while their negation is not satisfiable either. This serves to motivate the consideration of equivalidity in the following chapters. In the context of classical logic it can be fully explained via equisatisfiability and is therefore neglected but for intuitionistic logic it is more fundamental.

	Finally let us note the complexities of these various decision problems, which play a fundamental role in all of theoretical computer science.

	\begin{theorem}
		Satisfiability for classical and intuitionistic propositional logic is NP-complete.
	\end{theorem}

	For classical logic this is the famous Cook-levin theorem first published in Cook's seminal paper~\cite{cook1971complexity}. Of course we have seen that the intuitionistic case is the same decision problem.

	\begin{corollary}
		Validity for classical logic is co-NP-complete
	\end{corollary}

	This is clear since validity is just the complementary problem to satisfiability of the negated formula. However in intuitionistic logic unfortunately validity is really harder than satisfiability as shown by Statman~\cite{statman1979intuitionistic}.

	\begin{theorem}
		Validity for intuitionistic logic in PSPACE-complete.
	\end{theorem}

	One thing to note is that in contrast to predicate logic all these problems are decidable.

	\subsection{Predicate Logic}

	We now give account of the more involved semantics of first-order logic

	\subsubsection{Classical Semantics}

	\begin{definition}
		Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the domain of $\mathcal{M}$, and a function $I$ that assigns
		\begin{itemize}
			\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
			\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
		\end{itemize}
		A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
			m&\text{if $b=a$}\\
			v(b)&\text{otherwise}
		\end{cases}$$
		Then terms are interpreted as follows:
		\begin{itemize}
			\item $a^{I, v} = v(a)$ for each free variable $a$.
			\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$.
		\end{itemize}
		We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
		\begin{itemize}
			\item $\mathcal M, v\not\models\bot$.
			\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$.
			\item $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
			\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$.
			\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
			\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
			\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
			\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\end{itemize}
		$\varphi$ is satisfiable if for some $\mathcal{M}, v$ we have  $\mathcal M, v\models\varphi$. It is valid if for all  $\mathcal M, v$ we have  $\mathcal M, v\models\varphi$. We write $\mathcal{M}\models\varphi$ if for every $v$ we have  $\mathcal M, v\models\varphi$. We denote the set of valid formulas with CQC. As before define an entailment relation $\models_C$ between sets of formulas.
	\end{definition}

	\subsubsection{Intuitionistic Semantics}

	\begin{definition}
		A $\Sigma$-\textit{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$ and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$
		\begin{itemize}
			\item $M_u\subseteq M_w$.
			\item $f^{I_w}|_{M_u} = f^{I_u}$.
			\item $R^{I_w}|_{M_u} = f^{I_u}$.
		\end{itemize}
		Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
		\begin{itemize}
			\item $u, v\not\models\bot$
			\item $u, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I_u, v},\dots,t_n^{I_u, v})\in R^{I_u}$.
			\item $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
			\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$.
			\item $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
			\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
			\item $u, v\models\exists x\varphi$ iff there exists $m\in M_u$ such that $u, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
			\item $u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\end{itemize}
		We say that $\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. As always a formula is satisfiable if it is satisfies by some Kripke structure and it is valid if it is satisfied in every Kripke structure. We denote the set of valid formulas with IQC.  As before define an entailment relation $\models_I$ between sets of formulas.
	\end{definition}
	In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically.
	Consider for instance the classical tautology $\varphi$ $$\neg\forall x A(x)\to \exists x \neg A(x)$$We exhibit an intuitionistic counter model. Consider the Kripke structure with two nodes $u\leq w$ with $M_u = \{0\}, M_w = \{0, 1\}$ and $A^{I_u} = A^{I_w} = \{0\}$. Then $u, v\not\models \varphi$: First note that $w, v\not\models\forall xA(x)$ and $u, v\not\models\forall xA(x)$ for any valuation $v$ since $w, v[1, a]\not\models A(a)$, i.e. $u, v\models\neg\forall xA(x)$ and also $w, v\models\neg\forall xA(x)$. Furthermore $u, v[0/a]\not\models \neg A(a)$ and $w, v[0/a]\not\models \neg A(a)$ so $u, v\not\models \exists x\neg A(x)$ for any valuation $v$ (note however that this is not true at $w$!). Therefore $u, v\not\models\varphi$.

	Intuitively the problem is that the existential quantifier at $u$ does not see the elements at $w$, i.e. it must "choose" between the currently available elements. As in the propositional case there are well known negative translation of classical logic into intuitionistic logic going back to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit}.

	\begin{definition}
		For a predicate formula $\varphi$ define inductively $\varphi^N$ as follows:
		\begin{itemize}
			\item $\varphi^N = \varphi$ if $\varphi$ is atomic
			\item $(\varphi\wedge\psi)^N = \varphi^N\wedge\psi^N$
			\item $(\varphi\vee\psi)^N = \neg(\neg\varphi^N\wedge\neg\varphi^N)$
			\item $(\varphi\to\psi)^N = \varphi^N\to\psi^N$
			\item $(\forall x\varphi)^N = \forall x\varphi^N$
			\item $(\exists x\varphi)^N = \neg\forall x\neg\varphi^N$
		\end{itemize}
	\end{definition}
	\begin{theorem}
		A $\Sigma$-formula $\varphi$ is classically valid if and only if $\varphi^N$ is intuitionistically valid.
	\end{theorem}

	But again this is not reflected in satisfiability, i.e. with an completely analogous argument to~\ref{propsat} we obtain

	\begin{lemma}
		Any $\Sigma$-formula $\varphi$ is classically satisfiable if and only if it is intuitionistically satisfiable.
	\end{lemma}

	\section{Proof theory}

	Logical truth can also be approached via syntax rather than semantics. This gives rise to proof theory which will be especially important in the third chapter.

	\begin{definition}
		A \textit{sequent} $\Gamma\vdash \Delta$ consists of two multisets of formulas $\Gamma, \Delta$. The following rules make up our classical calculus $\mathbf{GC}$, corresponding roughly to \textbf{G3c} from~\cite{basicprooftheory}:\\
		\begin{center}
			\begin{tabular}{lll}
				\AxiomC{\hphantom{x}}
				\RightLabel{Ax ($P$ atomic)}
				\UnaryInfC{$P,\Gamma\vdash \Delta, P$}
				\DisplayProof&
				\AxiomC{\hphantom{x}}
				\RightLabel{L$\bot$}
				\UnaryInfC{$\bot,\Gamma\vdash\Delta$}
				\DisplayProof&
				\\&&\\
				\AxiomC{$A, B,\Gamma\vdash\Delta$}
				\RightLabel{L$\wedge$}
				\UnaryInfC{$A\wedge B, \Gamma\vdash \Delta$}
				\DisplayProof&
				\AxiomC{$\Gamma\vdash\Delta, A$}
				\AxiomC{$\Gamma\vdash\Delta, B$}
				\RightLabel{R$\wedge$}
				\BinaryInfC{$\Gamma\vdash \Delta, A\wedge B$}
				\DisplayProof&
				\\&&\\
				\AxiomC{$A, \Gamma\vdash\Delta$}
				\AxiomC{$B, \Gamma\vdash\Delta$}
				\RightLabel{L$\vee$}
				\BinaryInfC{$A\vee B, \Gamma\vdash \Delta$}
				\DisplayProof&
				\AxiomC{$\Gamma\vdash\Delta, A, B$}
				\RightLabel{R$\vee$}
				\UnaryInfC{$\Gamma\vdash \Delta, A\vee B$}
				\DisplayProof&
				\\&&\\
				\AxiomC{$\Gamma\vdash\Delta, A$}
				\AxiomC{$B, \Gamma\vdash\Delta$}
				\RightLabel{L$\to$}
				\BinaryInfC{$A\to B, \Gamma\vdash \Delta$}
				\DisplayProof&
				\AxiomC{$A,\Gamma\vdash\Delta, B$}
				\RightLabel{R$\to$}
				\UnaryInfC{$\Gamma\vdash \Delta, A\to B$}
				\DisplayProof&
				\\&&\\
				\AxiomC{$A[t/x], \Gamma\vdash\Delta$}
				\RightLabel{L$\forall$}
				\UnaryInfC{$\forall xA, \Gamma\vdash \Delta$}
				\DisplayProof&
				\AxiomC{$\Gamma\vdash\Delta, A[a/x]$}
				\RightLabel{R$\forall$}
				\UnaryInfC{$\Gamma\vdash \Delta, \forall xA$}
				\DisplayProof&
				\\&&\\
				\AxiomC{$A[a/x], \Gamma\vdash\Delta$}
				\RightLabel{L$\exists$}
				\UnaryInfC{$\exists xA, \Gamma\vdash \Delta$}
				\DisplayProof&
				\AxiomC{$\Gamma\vdash\Delta, A[t/x]$}
				\RightLabel{R$\exists$}
				\UnaryInfC{$\Gamma\vdash \Delta, \exists xA$}
				\DisplayProof&
				\\&&\\
				\AxiomC{$t = t, \Gamma\vdash \Delta$}
				\RightLabel{Ref}
				\UnaryInfC{$\Gamma\vdash \Delta$}
				\DisplayProof&
				\AxiomC{$s = t, P[s/x], P[t/x], \Gamma\vdash \Delta$}
				\RightLabel{Ref}
				\UnaryInfC{$s = t, P[s/x],\Gamma\vdash \Delta$}
				\DisplayProof&
				\\&&\\
			\end{tabular}
		\end{center}

		where in R$\forall$ and L$\exists$ $a$ is a free variable not occurring in the sequent otherwise.
		The above rules make up the calculus called \textbf{Gc}. It corresponds to \textbf{G3c}$^=$ from~\cite{basicprooftheory}.

		For intuitionistic logic we remain with a multi-succedent system by adjusting R$\to$ and R$\forall$:\\

		\begin{center}
			\AxiomC{$A, \Gamma\vdash B$}
			\RightLabel{R$\to$}
			\UnaryInfC{$\Gamma\vdash\Delta, A\to B$}
			\DisplayProof\hspace{2cm}
			\AxiomC{$\Gamma\vdash A[a/x]$}
			\RightLabel{R$\forall$}
			\UnaryInfC{$\Gamma\vdash\Delta, \forall xA$}
			\DisplayProof
		\end{center}

		We will call the calculus containing this formulas $\mathbf{GI}$ roughly corresponding to $\mathbf{G3i}$ from~\cite{basicprooftheory}. Usually for intuitionistic logic a single-formula succedent system is considered, but we are interested in reducing the difference between the calculi as much as possible.

		A \textit{proof} of a formula $\varphi$ in a calculus $\mathbf{G}$ is a labelled binary rooted tree in which every node is labelled with a sequent such that
		\begin{itemize}
			\item Every leaf is labelled in accordance to Ax or L$\bot$.
			\item The sequent of an interior node can be derived from sequents of its child nodes according to one of the rules.
		\end{itemize}
		We usually present a proof directly as a sequence of applications of rules rather than in a more usual tree presentation. If a sequent $\mathcal S\vdash\mathcal T$ is provable in $\mathbf{GC}$ we write $\mathcal S\vdash_C\mathcal T$, if it is in $\mathbf{GI}$ we write $\mathcal S\vdash_I\mathcal T$.
	\end{definition}

	\begin{example}
		The following is a proof of $(A\to B)\to A\vdash B$:
		\begin{center}
			\AxiomC{}
			\RightLabel{Ax}
			\UnaryInfC{$A\to B, A\vdash B, A$}
			\AxiomC{}
			\RightLabel{Ax}
			\UnaryInfC{$A, B\vdash B$}
			\RightLabel{L$\to$}
			\BinaryInfC{$A\to B, A\vdash B$}
			\DisplayProof
		\end{center}
	\end{example}

	\noindent We now recall some important properties of the given systems:

	\begin{theorem}[subformula property]
		Any formula that appears in a derivation is a subformula of a forumla that occurs in the root sequent.
	\end{theorem}

	We consider the following additional rules, which are not a priori part of our calculus:
	\begin{center}
		\begin{tabular}{lll}
			\AxiomC{$\Gamma\vdash\Delta$}
			\RightLabel{Lweak}
			\UnaryInfC{$A,\Gamma\vdash \Delta$}
			\DisplayProof&
			\AxiomC{$\Gamma\vdash\Delta$}
			\RightLabel{Rweak}
			\UnaryInfC{$\Gamma\vdash \Delta, A$}
			\DisplayProof&
			\\&&\\
			\AxiomC{$A, A,\Gamma\vdash\Delta$}
			\RightLabel{Lcontr}
			\UnaryInfC{$A, \Gamma\vdash \Delta$}
			\DisplayProof&
			\AxiomC{$\Gamma\vdash\Delta, A, A$}
			\RightLabel{Rcontr}
			\UnaryInfC{$\Gamma\vdash \Delta, A$}
			\DisplayProof&
			\\&&\\
		\end{tabular}

		\AxiomC{$\Gamma\vdash A, \Delta$}
		\AxiomC{$\Gamma', A\vdash \Delta'$}
		\RightLabel{Cut}
		\BinaryInfC{$\Gamma, \Gamma'\vdash\Delta, \Delta'$}
		\DisplayProof
	\end{center}

	The following central result of proof theory will be useful to us in many places:
	\begin{theorem}[admissibility of cut (and structural rules)]
		The 5 above rules are admissible in our calculi for classical and intuitionistic logic, i.e. adding them we are not able to prove any additional theorems.
	\end{theorem}

	This means that we may assume that all proofs are free of these rules, but in constructing new proofs we may apply them. We exclude them from our calculus to keep it simple for proofs and because cut destroys the subformula property.

	The semantic and syntactic approach to logic are of course equivalent.

	\begin{theorem}\label{completeness}[Completeness and Soundness]
		For sets of formulas $\mathcal S,\mathcal T$ both classically and intuitionistically
		$$\mathcal S\models\mathcal T \text{ iff }\mathcal S\vdash\mathcal T$$
	\end{theorem}

	\chapter{Embedding intuitionistic into classical logic}

	The goal of this section is to give a complete embedding of intuitionistic into classical logic. We shall start with the simpler case of propositional formulas. The process will be divided into three steps. The first is a first-order encoding of intuitionistic semantics, which gives an obvious but unsatisfactory embedding from propositional intuitionistic logic. We then discuss how some of the quantifiers can be eliminated via Herbrandization. As a third part we give a certain normal form that then finally allows us to realize that all quantifiers can be indeed eliminated from the translated formula and we end up with a propositional formula. As a quick note we also give an embedding form IPC to QBF.

	For predicate formulas the process is a bit more involved in that we also need to respect the constructive content of quantifiers. For this we use orderization as proposed in~\cite{iemhoff2010eskolemization}. This allows us to utilize a quite similar process as in the first section, even though it is a bit more involved.

	\section{Propositional logic}

	The idea is to encode the Kripke semantics of intuitionistic logic classically and is quite close to what was done in~\cite{iemhoff2010eskolemization}. We first present an embedding from IPC to CQC which is conceptually easier to understand, but will then reduce it to CPC.

	\subsection{IPC to CQC}

	The most obvious approach to embedding intuitionistic into classical logic is to express intuitionistic semantics, i.e. Kripke frames, as a first order theory. For every atom $A$ consider a unary predicate $A$ of the same name where $A(u)$ expresses that $A$ is true at some world $u$. We can then naively encode formulas as follows:

	\begin{definition}
		Let $\varphi$ be a propositional formula. Define $\varphi^{k}$ inductively as follows:
		\begin{itemize}
			\item $A^{k} = A(k)$ for every atom $A$.
			\item $(\varphi\wedge\psi)^k = \varphi^k\wedge\psi^k$.
			\item $(\varphi\vee\psi)^k = \varphi^k\vee\psi^k$.
			\item $(\varphi\to \psi)^k = \forall u(k\preceq u\to\varphi^{u}\to\psi^{u})$ where $u$ is some new free variable.
		\end{itemize}
		Let $K(\varphi)$ encode that $\preceq$ is a partial order, the persistency condition for all propositional variables occurring in $\varphi$ and ex falso quodlibet for each world, i.e.
		\begin{align*}
			K(\varphi) = &\forall k(k\preceq k)\wedge \forall k\forall l(k\preceq l\to l\preceq k\to l = k) \wedge \\&\forall k\forall l\forall m(k\preceq l \to l\preceq m\to k\preceq m)\wedge\\&\bigwedge \{\forall k\forall l(k\preceq l\to A(k)\to A(l))\:|\: \text{ $A$ occurs in $\varphi$}\}\wedge\\&\bigwedge\{\forall k(\bot(k)\to A(k))\:|\: A \text{ occurs in }\varphi\}
		\end{align*}
		Then define
		$$\varphi^{C} = K(\varphi)\to \varphi^{u}$$
		where $u$ is some free variable.
	\end{definition}

	All that we have done is model Kripke frames with a finite number of propositional variables as a first-order theory. The only attention must be paid with regard to $\bot$, which we have encoded because it will be useful for translation to propositional logic later. Consider the formula $\varphi^C_\bot$, which is exactly as above except that $\bot^k = \bot$. If we have a counter model to $\varphi^C_\bot$ we obtain a counter model for $\varphi^C$ by setting $\bot(k)$ false for all $k$. On the other hand if we have a counter model for $\varphi^C$ then we obtain one for $\varphi^C_\bot$ by restricting our domain to all $u$ for which $\bot(u)$ does not hold. Therefore this modification is unproblematic and we have:

	\begin{lemma}
		$\varphi$ is intuitionistically valid iff $\varphi^C$ is classically valid.
	\end{lemma}

	\begin{example}
		Consider the law of excluded middle $$\varphi = (A\to \bot)\vee A$$which is not intuitionistically valid. Then
		\begin{align*}
			\varphi^{C} &= K(\varphi)\to ((A\to \bot)\vee A)^u\\
						 &= K(\varphi)\to (A\to \bot)^u\vee A^u\\
						 &= K(\varphi)\to (\forall u'( u\preceq u' \to A^{u'}\to \bot^{u'})\vee A^\#(u))\\
						 &= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot(u'))\vee A(u))
		\end{align*}
		and indeed $\varphi^{C}$ admits a counter model with domain $\{u, u'\}$ where $u \preceq u'$, $A(u) = \bot(u) = \bot(u') = 0, A(u') = 1$.
	\end{example}

	Note how this counter model corresponds exactly to the intuitionistic counter model presented in example~\ref{LEMcounterexample}. We can make this correspondence explicit.

	\begin{definition}
		Let $\mathcal K = (W, (v_w)_{v\in W})$ be a propositional Kripke structure. Define the $\mathcal M(\mathcal K) = (M, I)$ as follows:
		\begin{itemize}
			\item As a domain take $M = W$.
			\item Have $u\in A^{I}$ iff $u, v_u\models A$.
			\item Interpret $\preceq$ as the partial order on $W$.
		\end{itemize}
	\end{definition}

	\noindent By definition we then have
	\begin{lemma}
		$\mathcal K\models \varphi$ iff $\mathcal M(\mathcal K)\models \varphi^{C}$.
	\end{lemma}

	\indent Similarly for every $\Sigma$-structure $\mathcal M$ with $\mathcal M\models K(\varphi)$ we could construct a Kripke Frame $\mathcal K(\mathcal M)$ such that $\mathcal M\models \varphi^C$ iff $\mathcal K(\mathcal M)\models \varphi$.

	\subsection{Quantifiers in Classical Logic}

	Of course translating a propositional formula to a first-order formula is somewhat unsatisfying. As a first step to obtaining a propositional formula we wish to eliminate the quantifiers. Since we are interested in preserving validity we use Herbrandization. In this process we introduce free variables and also additional function symbols to the signature. Whenever we speak of a new free variable what is meant is a free variable that does not occur in any of the considered formulas. Whenever a new function symbol is introduced we implicitly extend the signature by some not previously contained symbol.

	Now usually in classical logic before removing quantifiers prenexification is performed, however since we wish to apply a similar procedure to intuitionistic logic later, where prenexification is not valid, we will do without.

	\begin{definition}
		For formulas $\varphi$ we define the Skolemization $\varphi^S_T$ and Herbrandization $\varphi^H_T$ with respect to $T$ by simultaneous induction as follows:
		\begin{itemize}
			\item $A^S_T = A^H_T = A$ for each atomic $A$.
			\item $(\varphi\circ\psi)^X_T = \varphi^X_T\circ\psi^X_T$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
			\item $(\varphi\to\psi)^S_T = \varphi^H_T\to \psi^S_T$\\$(\varphi\to\psi)^H_T = \varphi^S_T\to\psi^H_T$.
			\item $(\forall x\varphi)^S_T = \forall x(\varphi[a/x]^S_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_T = \varphi[s(t_1,\dots,t_n)/x]^H_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$.
			\item $(\exists x\varphi)^S_T = \varphi[s(t_1,\dots,t_n)/x]^S_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$\\$(\exists x\varphi)^H_T = \exists x(\varphi[a/x]^H_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable.
		\end{itemize}
		Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
	\end{definition}

	\begin{theorem}For every formula $\varphi$
		\begin{itemize}
			\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
			\item $\varphi$ and $\varphi^H$ are classically equivalid.
		\end{itemize}
	\end{theorem}

	which follows from the following two Lemmata:

	\begin{lemma}
		Let $\chi$ be a formula and $\{t_1\dots t_n\} = T$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
		\begin{itemize}
			\item if $\mathcal M, v\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
			\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
		\end{itemize}
	\end{lemma}

	\begin{proof}
		We proceed by simultaneous induction on the formula height.

		For atomic $\chi$ the claims are clear.

		Otherwise there are $3$ cases:

		1. $\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$. By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v}), \mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as in the lemma. Then we can define $I_{\chi, v}$ as follows:

		For every symbol $p$ that occurs in $\varphi^S$ or $\varphi^H$ but not $\psi^S, \psi^H$ have $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$. For every symbol $p$ that occurs in $\psi^S$ or $\psi^H$ but not $\varphi^S, \varphi^H$ have $p^{I_{\chi, v}} = p^{I_{\psi, v}}$. Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.

		Let $\circ=\to$. Suppose $\mathcal M, v\models\chi$. Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_T$ or $\mathcal M_\chi\models\psi^H_T$, i.e. $\mathcal M_\chi\models\chi^S_T$. On the other hand suppose $\mathcal M, v\not\models\chi$. Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_T$ and $\mathcal M_\chi, v\not\models\psi^S_T$, i.e. $\mathcal M_\chi\not\models\chi^H_T$. Analogous arguments work for $\circ\in\{\wedge, \vee\}$.

		2. $\chi = \forall x\varphi$. Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise. Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.

		By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma. Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.

		Suppose $\mathcal M, v\models\chi$. Then for all $m\in M$ $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{M\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{M\cup\{a\}}[x/a])$, i.e. $\mathcal M_\chi,v\models \chi^S$. On the other hand suppose $\mathcal M, v\not\models\chi$. Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{T\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(t_1,\dots t_n)/x]^H_T$, i.e. $\mathcal M_\chi, v\not\models(\forall x\varphi)^H_T$.

		3. $\chi = \exists x\varphi$. The argument runs dually to 2. Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise. Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.

		By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma. Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.

		Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_T$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_T$.
	\end{proof}

	\begin{lemma}
		For every structure $\mathcal M$ and $\{t_1\dots t_n\} = T$ that contains all free variables in $\chi$ and variable assignment $v$
		\begin{itemize}
			\item if $\mathcal M, v\models\varphi^S_T$ then $\mathcal M, v\models \varphi$
			\item if $\mathcal M, v\not\models\varphi^H_T$ then $\mathcal M, v\not\models\varphi$.
		\end{itemize}
	\end{lemma}

	\begin{proof}
		Again we proceed by simultaneous induction on the formula height.

		For atoms the claims are clear. We distinguish 5 cases.

		1. $\chi = \varphi\to\psi$. Suppose $\mathcal M, v\models\chi^S_T$, i.e. $\mathcal M, v\not\models\varphi^H_T$ or $\mathcal M, v\models\psi^S_T$. By induction hypothesis $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e. $\mathcal M, v\models \chi$. On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e. $\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_T$. Again by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.

		2. + 3. Conjunctions and Disjunctions are dealt with analogously.

		4. $\chi = \forall x\varphi$.  Suppose $\mathcal M, v\models\chi^S_T$, i.e. for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{T\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$. Then it follows that $\mathcal M, v\models\varphi$. On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e. there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{T\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e. $\mathcal M, v\not\models\forall x\chi$.

		5. $\chi = \exists x\varphi$.  The argument runs dually to 4.
	\end{proof}

	\subsection{Equivalid normalization}

	Now trying to apply Herbrandization to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated. We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form. A similar translation was proposed in~\cite{statman1979intuitionistic}.

	\begin{definition}
		Let $\varphi$ be some formula. For each sub(semi)formula $\psi$ of $\varphi$ consider some new $n$-ary relation symbol $P_\psi$ where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$. We denote those variables with $\vec t_\psi$ in some arbitrary but fixed order. We inductively define clause sets $\mathcal S^+(\varphi)$ and $\mathcal S^-(\varphi)$ as follows:
		\begin{itemize}
			\item For every atomic formula $A$\\$\mathcal S^+(A) = \{P_A(\vec t_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec t_A)\}$
			\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$\\$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec t_{\varphi\circ\psi})\to P_{\varphi}(\vec t_\varphi)\circ P_{\psi}(\vec t_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec t_\varphi)\circ P_{\psi}(\vec t_\psi)\to P_{\varphi\circ\psi}(\vec t_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
			\item For every implication $\varphi \to\psi$\\$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec t_{\varphi\to\psi})\wedge P_{\varphi}(\vec t_\varphi))\to P_{\psi}(\vec t_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec t_\varphi)\to P_{\psi}(\vec t_\psi))\to P_{\varphi\to\psi}(\vec t_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\varphi)$
			\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$\\$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec t_{Qx\varphi})\to QxP_{\varphi}(\vec t_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec t_{\varphi})\to P_{Qx\varphi}(\vec t_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
		\end{itemize}
		and have $\mathcal S(\varphi) = \mathcal S^+(\varphi)\cup\mathcal S^-(\varphi)$.
	\end{definition}

	\begin{lemma}The following are true for both classical and intuitionistic semantics:
		\begin{itemize}
			\item $\bigwedge S^+(\varphi)\wedge P_\varphi\models\varphi$. 
			\item $\bigwedge S^-(\varphi)\wedge \varphi\models P_\varphi$. 
		\end{itemize}
	\end{lemma}

	\begin{proof}
		The claims are verified in a straightforward manner by induction on the height of $\varphi$.
	\end{proof}

	\begin{example}
		Consider $\varphi = \neg\forall x A(x)\to \exists x\neg A(x)$. Then
		\begin{align*}
			\mathcal S^+(\varphi) = &\{(P_\varphi\wedge P_{\neg\forall x A(x)})\to P_{\exists x\neg A(x)}, (P_{\forall xA(x)}\to\bot)\to P_{\neg \forall xA(x)}\}\\&\cup\{\forall xP_{A(x)}(x)\to P_{\forall xA(x)}, \forall x(P_{A(x)}(x)\to A(x))\}\\&\cup\{P_{\exists x\neg A(x)}\to \exists xP_{\neg A(x)}(x), \forall x(P_{\neg A(x)}(x)\to P_{A(x)}(x)\to \bot)\}\\&\cup\{\forall x(A(x)\to P_{A(a)}(x))\}\\
			\mathcal S^-(\varphi) = &\{(P_{\neg\forall x A(x)}\to P_{\exists x\neg A(x)})\to P_\varphi,  (P_{\neg \forall xA(x)}\wedge P_{\forall xA(x)})\to\bot\}\\&\cup\{P_{\forall xA(x)}\to \forall xP_{A(x)}(x), \forall x(A(x)\to P_{A(x)}(x))\}\\&\cup\{\exists xP_{\neg A(x)}(x)\to P_{\exists x\neg A(x)}, \forall x((P_{A(x)}(x)\to \bot)\to P_{\neg A(x)}(x))\}\\&\cup\{\forall x(P_{A(a)}(x)\to A(x))\}
		\end{align*}
	\end{example}

	We can give a transformation of structures that corresponds to the above transformation of sentences.

	\begin{definition}
		For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms $P_\psi$. By slight abuse of notation we mean with $\vec t_\psi$ elements of the domain instead of variables. In the classical case have	
		$$P_{\varphi\wedge\psi}^I(\vec t_{\varphi\wedge\psi}) \Leftrightarrow P_{\varphi}^I(\vec t_\varphi)\wedge P_{\psi}^I(\vec t_\psi)\indent P_{\varphi\vee\psi}^I(\vec t_{\varphi\vee\psi}) \Leftrightarrow P_{\varphi}^I(\vec t_\varphi)\vee P_{\psi}^I((\vec t_\psi))$$$$ P_{\varphi\to\psi}^I(\vec t_{\varphi\to\psi}) \Leftrightarrow (\neg P_{\varphi}^I(\vec t_{\varphi}))\vee P_{\psi}^I(\vec t_{\psi})$$
		$$P_{\forall x\varphi}^I(\vec t_{\forall x\varphi}) \Leftrightarrow \{\vec t\:|\:\:P_{\varphi}^I(\vec t_\varphi) \text{ for all $x\in M$}\}$$$$ P_{\exists x\varphi}^I(\vec t_{\exists x\varphi}) \Leftrightarrow \{\vec t\:|\:\:P_{\varphi}^I(\vec t_\varphi) \text{ for some $x\in M$}\}$$
		\\

		and for intuitionistic logic and some world $u$ 		
		$$P_{\varphi\wedge\psi}^{I_u}(\vec t_{\varphi\wedge\psi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi)\wedge P_{\psi}^{I_u}(\vec t_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec t_{\varphi\vee\psi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi)\vee P_{\psi}^{I_u}((\vec t_\psi))$$$$ P_{\varphi\to\psi}^{I_u}(\vec t_{\varphi\to\psi}) \Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec t_{\varphi}))\vee P_{\psi}^{I_w}(\vec t_{\psi})\text{ for all $w\geq u$}$$
		$$P_{\forall x\varphi}^{I_u}(\vec t_{\forall x\varphi}) \Leftrightarrow P_{\varphi}^{I_w}(\vec t_\varphi) \text{ for all $w\geq u$, $x\in M_w$}$$$$ P_{\exists x\varphi}^{I_u}(\vec t_{\exists x\varphi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi) \text{ for some $x\in M_u$}$$
		\\
	\end{definition}

	The above definitions are motivated by the following results which follow directly from the involved definitions and are true both for classical and intuitionistic semantics:

	\begin{lemma}
		For every formula $\varphi$ and structure $\mathcal M$
		$$\mathcal S(\mathcal M, \varphi)\models\mathcal S(\varphi)$$
	\end{lemma}

	\begin{lemma}
		For every structure $\mathcal M$
		$$\mathcal M\models \varphi\text{ iff }S(\mathcal M, \varphi)\models P_\varphi$$
	\end{lemma}

	Then from the previous three Lemmas we get

	\begin{corollary}
		In both intuitionistic and classical logic
		\begin{itemize}
			\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec t_\varphi)$ is. 
			\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec t_\varphi)$ is.
		\end{itemize}
	\end{corollary}

	\subsection{IPC to CPC}

	We are now ready to put everything together. Recall that our goal is for a formula to give a modified formula that is valid classically iff the original formula is valid intuitionistically. As a first step we perform the normalization from the last section, i.e. instead of $\varphi$ we consider $$\varphi' = \bigwedge \mathcal S^-(\varphi)\to P_\varphi$$

	Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
	$$\hspace{-0.8cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$

	where we can treat $A\to B$ as a special case $A\to (B\wedge B)$. We then apply the transformation to CQC, leading to a formula
	$$\varphi'' = K(\varphi')\to\bigwedge  \mathcal S\to P_\varphi(u)$$
	where $u$ is some new free variable and each $\psi\in\mathcal S$ is of one of the forms
	$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\wedge (B_\psi(k))\to C_\psi(k)))$$
	$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\vee (B_\psi(k))\to C_\psi(k)))$$
	$$\forall k(u\preceq k\to (\forall l(k\preceq l \to A_\psi(l)\to B_\psi(l))\to C_\psi(k)))$$

	Applying Herbrandization we obtain a formula
	$$\varphi''' = K(\varphi')\to \bigwedge \mathcal S'\to P_\varphi(s)$$
	where $s$ is some new constant term and each $\psi\in\mathcal S'$ is of one of the forms
	$$\hspace*{-2mm}\forall k(s\preceq k\to A_\psi(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k(s\preceq k\to (A_\psi(k)\wedge (B_\psi(k))\to C_\psi(k)))$$
	$$\hspace*{-2mm}\forall k(s\preceq k\to A_\psi(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(s\preceq k\to (A_\psi(k)\vee (B_\psi(k))\to C_\psi(k)))$$
	$$\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A_\psi(f_\psi(k))\to B_\psi(f_\psi(k)))\to C_\psi(k))$$
	and $f$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model then we have a counter-model $(M',I')$ where $$M' = \{m\in M|s\preceq m\}$$ and 
	$$f_\psi^{I'}(u) = \begin{cases}
		f^I_\psi(u)&\text{ iff $u\preceq f^{I}_\psi(u)$}\\
		u&\text{ else}
	\end{cases}$$ that is where $s$ is a least element and the $f_\psi$ increase their argument. Therefore instead of $\varphi'''$ we may consider a formula $$\varphi^\circ = \forall k(s\preceq k)\to \bigwedge\{\forall k(k\preceq f_\psi(k))\:|\:\psi\in\mathcal F_\to\}\to K(\varphi')\to \bigwedge \mathcal S^\circ\to P_\varphi(s)$$ where each formula $\psi\in\mathcal S^\circ$ is of one of the forms
	$$\forall k(A_\psi(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k((A_\psi(k)\wedge B_\psi(k))\to C_\psi(k))$$
	$$\forall k(A_\psi(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k((A_\psi(k)\vee B_\psi(k))\to C_\psi(k))$$
	$$\forall k((A_\psi(f_\psi(k))\to B_\psi(f_\psi(k)))\to C_\psi(k))$$
	and $\mathcal F_\to$ denotes the set of formulas of the last kind.

	We now argue that if a counter-model exists for $\varphi^\circ$ then a bounded counter-model exists, allowing us to return to the propositional case.

	\begin{definition}
		Suppose we are given a counter-model $(M, I)$ for $\varphi^\circ$. For $\psi\in\mathcal F_\to$ we say that is it \textit{fulfilled} at a world $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true. For $\psi\in\mathcal F_\to$ define $$g_\psi : M\to M, u\mapsto\begin{cases}
			u&\text{ if $\psi$ is fulfilled in $u$}\\
			f^I_\psi(u)&\text{ else}		
		\end{cases}$$Define a $(M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ is the set of sequences without repetition on $\mathcal F_\to$.
		\item Interpret $\preceq$ as the prefix-order.
		\item Have $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
			\psi_1\dots\psi_n&\text{if $\psi$ occurs in $\psi_1\dots\psi_n$}\\
			\psi_1\dots\psi_n\psi&\text{else}			
		\end{cases}$$
		\item For propositional variables $P$ have $$P^{I_T}(x_1\dots x_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots))$$
	\end{itemize}
	\end{definition}

	The following easy technical Lemma is quite useful.
	\begin{lemma}
		Let $(M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
		\begin{enumerate}
			\item $\psi$ is fulfilled at $f_\psi^I(u)$ for all $u\in M, \psi\in\mathcal F_\to$.
			\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\geq u$.
		\end{enumerate}

	\end{lemma}

	\begin{proof}
		1. If $C_\psi^I(u)$ is true then we are done due to persistency. Otherwise we know that $(A_\psi^I(f_\psi^I(u))\to B_\psi^I(f_\psi^I(u)))\to C_\psi^I(f_\psi^I(u))$ holds, so $A_\psi^I(f_\psi^I(u))\to B_\psi^I(f_\psi^I(u))$ must be true.

		2. If $C_\psi^I(u)$ is true then $C_\psi^I(w)$ is also true and we are done. So suppose $A_\psi^I(u)\to B_\psi^I(u)$ is false, i.e. $A_\psi^I(u)$ true and $B_\psi^I(u)$ false. Then due to the persistency condition $A_\psi^I(w)$ is also true. Again if $C_\psi^I(w)$ is true we are done. Otherwise we know that $B^I(f^I_\psi(w))$ must be false, since $(A^I_\psi(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I_\psi(w)$ and $A^I_\psi(f^I_\psi(w))$ hold. But then due to persistency so is $B^I_\psi(w)$ and we have that $\psi$ is fulfilled at $w$.
	\end{proof}

	Essentially what fulfilment of $\psi$ at $u$ tells us is that having $f_\psi(w) = w$ for all $w\geq u$ works. 
	With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.

	\begin{corollary}
		If $(M, I)$ is a counter-model to $\varphi^\circ$ then so is $(M_T, I_T)$.
	\end{corollary}

	The size of this model is $|\mathcal F_\to|!$ in the theoretical worst case. While this would already allow us to return to the propositional world we can do better than that. In fact we can merge all branches into a model of size $2^{|\mathcal F_\to|}$.

	\begin{definition}
		Suppose we are given a counter-example $(M_T, I_T)$ to $\varphi^\circ$ as above. Consider $(M_F, I_F)$ as follows:
		\begin{itemize}
			\item $M_F = 2^{F_\to}$, i.e. the set of subsets of $F_\to$.
			\item $\preceq$ is interpreted as the subset relation.
			\item $s^{I_M} = \emptyset$, $f_\psi^{I_F}(u) = u\cup\{\psi\}$.
			\item For every atom $P$ have $P^{I_F}(u)$ iff for every $w\in M_T$ where each $\psi\in u$ is fulfilled we have $$P^{I_T}(w)$$
		\end{itemize}
	\end{definition}

	Note that in this definition for the first time it is of crucial importance that we have also transformed $\bot$ into a unary predicate that we may set to true, as this could be forced in certain cases, in particular if there is no $w$ at which each $\psi\in u$ is fulfilled.

	Of course we have to prove that this model transformation is correct:

	\begin{definition}
		If $(M_T, I_T)$ is a counter-model for $\varphi^\circ$ then so is $(M_F, I_F)$.
	\end{definition}
	\begin{proof}
		We have to check that all 5 types of formulas are still satisfied. The proofs for the first 4 are very similar, so we shall just present one.

		Suppose $\chi$ is of  the form $\forall k(A_\chi(k)\to (B_\chi(k)\vee C_\chi(k))$. Let $u\subseteq \mathcal F_\to$.
	\end{proof}

	Therefore we know that $\varphi^\circ$ is valid iff it is valid for structures with domain $2^{F_\to}$, i.e. the power-set of $F_\to$, ordered by the subset relation and $s^I = \emptyset, f_\psi^I(x) = x\cup\{\psi\}$. Thus we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula. This gives us the following translation:

	\begin{definition}
		Suppose $\mathcal S$ is a set of formulas of the form
		$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C),(A\vee B)\to C, (A\to B)\to C$$
		Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$. For each atom $A$ occurring in $\mathcal S$ and $M\subseteq \mathcal F_\to$ consider a new atom $A^{M}$, which will correspond to $A^{I_F}(M)$ from the previous definition. Obtain $\mathcal S^\#$ by including the formulas
		\begin{itemize}
			\item $\bot^N\to A^N, A^N\to A^M$ for atoms $A$ occurring in $\mathcal S$ and $N\subseteq M\subseteq \mathcal F_\to$.
			\item $A^M\to (B^M\circ C^M)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $M\subseteq \mathcal F_\to$.
			\item $(A^M\circ B^M)\to C^M$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $M\subseteq \mathcal F_\to$.
			\item $(A^{M\cup \{\psi\}}\to B^{M\cup \{\psi\}})\to C^M$ for $\psi = (A\to B)\to C\in\mathcal S$, $M\subseteq \mathcal F_\to$.
		\end{itemize}
	\end{definition}

	\begin{theorem}
		$$\mathcal S\models_C P\text{ iff }\mathcal S^\#\models_I P^\emptyset$$
	\end{theorem}

	\begin{example}
		Let us consider the law of excluded middle $\varphi = A\vee\neg A$. Then we get
		$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\}$$
		and we know that $\varphi$ is valid iff $\bigwedge \mathcal S(\varphi)\to P_\varphi$ is valid. Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
		\begin{align*}
			\mathcal S^\# =&\{P_\varphi^\emptyset\to P_\varphi^{\{\psi\}}, P_A^\emptyset\to P_A^{\{\psi\}},P_{\neg A}^\emptyset\to P_{\neg A}^{\{\psi\}},A^\emptyset\to A^{\{\psi\}}\}\\ &\cup\{(P_A^\emptyset\vee P_{\neg A}^\emptyset)\to P_\varphi^\emptyset, (P_A^{\{\psi\}}\vee P_{\neg A}^{\{\psi\}})\to P_\varphi^{\{\psi\}},  A^\emptyset\to P_A^\emptyset\}\\ &\cup \{P_A^{\{\psi\}}\to P_A^{\{\psi\}}, (A^{\{\psi\}}\to \bot)\to P_{\neg A}^\emptyset, (A^{\{\psi\}}\to \bot)\to P_{\neg A}^{\{\psi\}}\}
		\end{align*}
		It is easy to construct a counter example:
		$$0 = P_\varphi^\emptyset  = P_A^\emptyset = P_{\neg A}^\emptyset = A^\emptyset$$
		$$1=  P_{\varphi}^{\{\psi\}} = P_A^{\{\psi\}} =  P_{\neg A}^{\{\psi\}} = A^{\{\psi\}} $$
	\end{example}

	\subsection{IPC to QBF}

	It is well known that a polynomial time translation between QBF and IPC must exists, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}. However giving an explicit translation from IPC to QBF can still be useful as good solvers exists for QBF but for IPC not so much. Furthermore we will be able to give fragments of IPC for which validity is complete for other stages of the polynomial hierarchy. The translation from the last section will serve as a useful tool for this. We will not explain the semantics of QBF in depth. Intuitively just think that the interpretation of existentially quantified variables may depend on variables quantified above them.

	 Fix some set $\mathcal S$ of formulas $\psi$ of the form $$\hspace*{-.7cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$and a formula $$\varphi = \bigwedge \mathcal S\to P$$
	and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a linear-size formula is instead of expressing validity at every node of $\mathcal M_F$ - which was done in the CPC translation - we express that in each path in $\mathcal M_F$ from $\emptyset$ to $F_\to$ the worlds on the way satisfy all conditions for $\mathcal M_F$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition will express which path is considered.

	\begin{definition}
		For every atom $A$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $A^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$. Define
		\begin{itemize}
			\item $K^n = \{\bot^i\to A^i, A^i\to A^j\:|\:\text{ for all atoms $A$ and $i < j\leq n$}\}$
			\item $K_\to^n = \bigvee \{F_\psi^{n}\wedge\bigwedge\{\neg F^i_\psi\:|\:i<n\}\:|\:\psi\in\mathcal F_\to\}$, i.e. encoding that at each stage $n$ an $F_\psi^{n}$ that was not true at a previous stage is true, and thereby via the pigeon-hole principle also that exactly one of the $F_\psi^{n}$ is true at each of the $|\mathcal F_\to|$ stages.
			\item $\mathcal S^n$ as $\mathcal S\setminus\mathcal F_\to$ where each atom $A$ has been replaced by $A^n$.
			\item $\mathcal F_\to^n:=\left\{\bigvee\{F_\psi^i\:|\:i \leq n\}\to (A^{n+1}_\psi\to B^{n+1}_\psi)\to C^n_\psi\:|\:\psi\in\mathcal F_\to\right\}$
		\end{itemize}
		and let $k = |\mathcal F_\to|$. Have
		$$\hspace*{-1cm}\varphi^Q = \exists \vec X^0\forall\vec F^0\dots\exists \vec X^k\left(\bigwedge \{K^{i}_\to\:|\:i < k\}\to\left(\neg P^0\wedge K^k\wedge\bigwedge\{\mathcal S^i\:|\:i\leq k\}\cup\{\mathcal F^i_\to\:|\: i< k\}\right)\right)$$
	\end{definition}

	\begin{example}
		For double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
		$$\hspace*{-.3cm}
			\varphi^Q = \exists A^0,\bot^0\forall F^1\exists A^1,\bot^1\left(F^1\to\left(\neg A^0\wedge\bigwedge\left\{\begin{matrix}
				\bot^0\to A^0,\bot^1\to A^1,\bot^0\to \bot^1,\\
				 A^0\to A^1, (A^1\to \bot^1)\to \bot^0 
			\end{matrix}\right\}\right)\right)
		$$
		which is not a satisfiable QBF since $A^0 = \bot^0 = \bot^1 = 0, A^1 = 1$ falsifies it for any $F^1$.
	\end{example}

	\begin{lemma}
		$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
	\end{lemma}
	\begin{proof}
		Suppose $\varphi$ is not intuitionistically valid, i.e. there exists a counter-model $\mathcal M$ for $\mathcal S^\#\to P^\emptyset$. For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\emptyset$. Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and there exists a sequence $\emptyset = M_0\subseteq M_1\dots\subseteq M_{n-1}$ of subsets of $\mathcal F_\to$ such that $M_i$ contains exactly the $F_\psi$ for which $F_\psi^j$ is true for some $j\leq i$ and $A^i$ is interpreted as $A^{M_i}$ is in $\mathcal M$. Consider arbitrary interpretations for $F_\to^{n}$. If none is interpreted as true then $K^n_\to$ fails, if more than one is then $K^m_\to$ will fail for some $m\ > n$, in any case $\varphi^Q$ would be automatically satisfied. If exactly one is satisfied, say $\psi_n$, let $M_n = M_{n-1}\cup\{\psi_n\}$. Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{M_n}$. Continue this construction until $M_{|\mathcal F_\to|}  = \mathcal F_\to$ is reached. Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.

		On the other hand suppose $\varphi^Q$ is satisfiable. We construct a Kripke structure that is a counter-example to $\varphi$.
		\begin{itemize}
			\item As a set of potential worlds take non-repeating sequences of elements in $\mathcal F_\to$, some of these worlds will be removed later.
			\item We define the valuations at each world inductively. At $\epsilon$ (i.e. the empty sequence) interpret $A$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$. Given a sequence $\chi_0\dots \chi_n$ such that for $i<n$ interpreting $A^i$ as it is interpreted in $\chi_0\dots\chi_i$ is part of a satisfying interpretation of $\varphi^Q$ where we have chosen $F^i_\psi$ true iff $\psi = \chi_i$, let $\chi_0\dots \chi_n$ interpret $A$ as $A^n$ in that interpretation of $\varphi^Q$ where we have chosen $F^n_\psi$ true iff $\psi = \chi_n$.
			\item Defined as above this might require us to interpret $\bot$ as true at some worlds. We simply remove those worlds from the considered set.
			\item The order on the remaining worlds is the usual prefix order.
		\end{itemize}
		From the definitions it directly follows that this constitutes a counter-example to $\varphi$.
	\end{proof}

	Since we have not discussed QBF in detail the above argument might be a bit difficult to follow, in particular the second part, but is simple in essence. We shall demonstrate with help of a example.

	\begin{example}
		Consider the formula $\varphi = (((A\to B)\to C)\wedge((B\to A)\to C))\to C$, which is a non-intuitionistic tautology. Its associated QBF is $$\varphi^Q = \exists A^0B^0C^0\bot^0\forall F_{\psi_1}^1F_{\psi_2}^1 \exists A^1B^1C^1\bot^1\forall F_{\psi_1}^2F_{\psi_2}^2 \exists A^2B^2C^2\bot^2\psi'$$where $\psi_1 := (A\to B)\to C, \psi_2 = (B\to A)\to C$ and $$\psi' = \bigwedge\left\{\begin{matrix}
			F^1_{\psi_1}\vee F^1_{\psi_2}, (F^2_{\psi_1}\wedge\neg F^1_{\psi_1})\vee (F^2_{\psi_2}\wedge\neg F^1_{\psi_2}),\\\bot^0\to A^0, \bot^0\to B^0,\bot^0\to C^0,\bot^1\to A^1,\bot^1\to B^1, \bot^1\to C^1,\\
			\bot^2\to A^2, \bot^2\to B^2,\bot^2\to C^2,\bot^0\to \bot^1, \bot^1\to\bot^2, A^0\to A^1,
			\\A^1\to A^2, B^0\to B^1, B^1\to B^2, C^0\to C^1, C^1\to C^2,\\
			  \neg C^0, F^1_{\psi_1}\to (A^1\to B^1)\to C^0, F^1_{\psi_2}\to(B^1\to A^1)\to C^0\\
			 (F^1_{\psi_1}\vee F^2_{\psi_1})\to (A^2\to B^2)\to C^1, (F^1_{\psi_2}\vee F^2_{\psi_2})\to(B^2\to A^2)\to C^1
		\end{matrix}\right\}$$
	This QBF is satisfiable: Have $P^0 = 0$ for all atoms $P$, always $P^2 = 1$ for all atoms $P$ and $C^1 = 1$, $\bot^1 = 0$. Furthermore have $A^1 = 1$ iff $F^1_{\psi_1} = 1$ and $B^1 = 1$ iff $F^1_{\psi_2} = 1$. The construction of an intuitionistic counter-model then proceeds as follows:
	\begin{itemize}
		\item As a set of potential worlds we have $\{\epsilon, \varphi_1, \varphi_2, \varphi_1\varphi_2, \varphi_2\varphi_1\}$.
		\item Since $\bot^2$ is always true irrespective of how the $F^i_\psi$ are chosen, the worlds $\varphi_1\varphi_2, \varphi_2\varphi_1$ are deleted.
		\item For the remaining three worlds we have
		\begin{itemize}
			\item$\epsilon\not\models A, B, C$
			\item$\psi_1\models A, C, \psi_1\not\models B$ (corresponding to $F^1_{\psi_1} = 1, F^1_{\psi_2} = 0$)
			\item$\psi_2\models B, C, \psi_2\not\models A$ (corresponding to $F^1_{\psi_1} = 0, F^1_{\psi_2} = 1$)
		\end{itemize}
		according to the above satisfying interpretation.
	\end{itemize}
	\end{example}

	Note that the last set of existentially quantified variables is superfluous because there is only one assignment which even has the chance to falsify $\varphi^Q$, namely that which assigns $1$ to the only $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$. That is in our final formula we can replace $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^i$. With that we get the following:

	\begin{corollary}
		Let $N$-INTINVALID be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid. Then $N$-INTINVALID is in $\Sigma_{2N-1}^P$. The dual problem $N$-INTVALID is in $\Pi_{2N-1}^P$.
	\end{corollary}

	As a final result we would like to prove that the above problems are complete for their respective complexity classes. It is sufficient to provide a polynomial reduction from QBF with bounded number of quantifier alternation to them. Such translations exists~\cite{statman1979intuitionistic,vsvejdar2003polynomial}, however they are insufficient in that they generate more formulas of form $(A\to B)\to C$ than there are quantifier alternations. We therefore would need a slightly new translation.

	\section{Predicate logic}

	We now give an analogous transformation for first order logic. However including quantifiers provides us with a new source of non-constructivism, i.e. $\neg\forall x A(x)\to \exists x\neg A(x)$ is a classical tautology that is not automatically eliminated by our above considerations. This makes skolemization tricky.

	Analogously to the previous section we define a clause sets $S_\psi$ as follows:
	\begin{itemize}
		\item $S_\psi = \{P_\psi(\vec s)\to\psi(\vec t), \psi(\vec t)\to P_\psi(\vec t)\}$ if $\psi(\vec t)$ is atomic.
		\item $S_\psi = \{P_\psi(\vec t)\to P_{\psi_1}(\vec t), P_\psi(\vec t)\to P_{\psi_2}(\vec t),P_{\psi_1}(\vec t)\to P_{\psi_2}(\vec t)\to P_\psi(\vec t)\}$ if $\psi(\vec t) = \psi_1(\vec t)\wedge\psi_2(\vec t)$.
		\item $S_\psi = \{P_{\psi_1}(\vec t)\to P_\psi(\vec t), P_{\psi_2}(\vec t)\to P_\psi(\vec t)\}\cup $ $$\hspace{-1.5cm}\{(P_{\psi_1}(\vec t)\to P_\chi(\vec a))\to (P_{\psi_2}(\vec t)\to P_\chi(\vec a))\to P_\psi(\vec t)\to P_\chi(\vec a)\:|\:\chi \text{ is a subformula of $\varphi$}\}$$ if $\psi(\vec t) = \psi_1(\vec t)\vee\psi_2(\vec t)$ for appropriate vectors of new free variables $\vec a$.
		\item $S_\psi = \{(P_{\psi_1}(\vec t)\to P_{\psi_2}(\vec t))\to P_\psi(\vec t), P_\psi(\vec t)\to P_{\psi_1}(\vec t)\to P_{\psi_2}(\vec t)\}$ if $\psi(\vec t) = \psi_1(\vec t)\to\psi_2(\vec t)$.
		\item $S_\psi = \{\forall xP_{\psi_1}(\vec t, x) \to P_\psi(\vec t), P_\psi(\vec t)\to \forall xP_{\psi_1}(\vec t, x)\}$ if $\psi(\vec t) = \forall x(\psi_1(\vec t, x))$.
		\item $S_\psi = \{\exists xP_{\psi_1}(\vec t, x) \to P_\psi(\vec t), P_\psi(\vec t)\to \exists xP_{\psi_1}(\vec t, x)\}$ if $\psi(\vec t) = \exists x(\psi_1(\vec t, x))$.
	\end{itemize}
	Again let $\mathcal S = (P_\bot \to\bot)\cup\bigcup\{S_\psi\:|\:\psi\text{ is a subformula of $\varphi$}\}$ and define
	$$\varphi^\to = \bigwedge{\mathcal S} \to P_\varphi(\vec t)$$
	Analogously to the previous section we have
	\begin{lemma}
		$\varphi$ and $\varphi^\to$ are intuitionistically equivalid.
	\end{lemma}
	The proof works exactly as in the propositional case. Similarly to before every formula in $\mathcal S$ is of one of the following forms $$A\to B, A\to B\to C, (A\to B)\to C, (A\to D)\to(B\to D)\to C\to D$$$$\forall xA\to B, A\to\forall xB, \exists xA\to B, A\to\exists xB$$

	\subsection{Eskolemization}

	Before being able to apply the procedure from the previous section we need to eliminate the quantifiers from $\varphi^\to$. The key idea for this stems from~\cite{baaz2006skolemization} and was expanded in~\cite{iemhoff2010eskolemization}. We will be able to use a condensed version. The basic principle is to introduce a new unary predicate $E(x)$ which encodes roughly that $x$ exists in a constructive sense. Note that besides adding $E$ to the signature step we need to add new function symbols, as is usual in skolemization. We then obtain $\mathcal S^E$ from $\mathcal S$ by including formulas as follows:

	\begin{itemize}
		\item every quantifier-free $\varphi\in\mathcal S$
		\item $\bigwedge_i E a_i \to Ef(a_1,\dots a_n)$ for every $n$-ary function symbol $f$ and new free variables $a_1,\dots, a_n$.
		\item $Ea$ for every free variable $a$ occurring in $\varphi^\to$.
		\item $(Es\to A[s/x])\to B$ for each $\forall xA\to B\in\mathcal{S}$ where $s$ is a new constant symbol.
		\item $A\to Ea\to B[a/x]$ for each $A\to \forall xB\in\mathcal S$  where $a$ is some new free variable.
		\item $Ea\to A[a/x]\to B$ for each $\exists A\to B$ where $a$ is some new free variable..
		\item $A\to B[s(\vec a)/x]$ and $Es(\vec a)$ for each $A\to \exists xB$ where $s$ is some new function symbol and $\vec a$ are the free variables occurring in $B$.
	\end{itemize}
	As usual we define $$\varphi^E = \bigwedge \mathcal S^E\to P_\varphi(\vec t)$$
	and have 
	\begin{lemma}
		$\varphi^\to$ and $\varphi^E$ are equivalid.
	\end{lemma}
	\begin{proof}
		As usual we do a translation of counter models.

		Suppose we have a Kripke structure $\mathcal K = (W, (M_u)_{u\in W}, (I_u)_{u\in W})$ with world $u$ and variable assignment $v$ such that $u, v\models \bigwedge \mathcal S^\to$ and $u, v\not\models P_\varphi(\vec t)$.

		First we will show that there must exists a linear substructure $\mathcal K^L = (W^L, (M^L_u)_{u\in W}, (I^L_u)_{u\in W})$ (i.e. $W^L$ is a total order) that is also a counter example. We will argue indirectly, i.e. suppose towards contradiction that every linear substructure is a model of $\varphi^\to$. In particular in every linear substructure we must have $u, v\not\models\mathcal\bigwedge S^\to$.

		 Consider the following Kripke structure $\mathcal K^E = (W^E, (M^E_u)_{u\in W}, (I^E_u)_{u\in W})$:
		\begin{itemize}
			\item Take $W^E = W$. For clarity we will denote worlds in $W^E$ with $u^E$ instead of $u$.
			\item Let the domain $M^E_u$ be constant across all worlds with $M^E_u = \underset{u\in U}{\bigcup} M_u$.
			\item \vspace{-3mm}Interpret $E$ at $u$ as $M_u$, i.e. $E^{I_u^E}(t)$ iff $t\in M_u$.
			\item Interpret each $n$-ary function symbol $f$ at $u$ via $$f^{I_u^E}(\vec t) = \begin{cases}f^{I_u}(\vec t) &\text{if $\vec t\in M_u^n$}\\m_u&\text{otherwise}\end{cases}\\$$ for with some arbitrary $m_u\in M^E_u\setminus M_u$.\footnote{Note that if $M^E_u\setminus M_u = \emptyset$ the second case does not occur in the first place.}
			\item Interpret each relation symbol as in $\mathcal K$, i.e. $R^{I_u^E} = R^{I_u}$.
		\end{itemize}
		We claim that $u^E, v\not\models \varphi^E$. Clearly $u^E, v\not\models P_\varphi(\vec t)$. Furthermore it is easy to see that if $\psi\in\mathcal S^E$ was not newly introduced in the the above construction then $u^E, v\not\models \psi$ follows directly from $u, v\not\models\psi$. We then have $6$ remaining cases to consider.

		1. $\psi$ is of the form $\bigwedge_i E a_i \to Ef(a_1,\dots a_n)$. Which is satisfied by the interpretation of functions and $E$ we gave.

		2. $\psi$ is of the form $Ea$. Since we are using the same variable assignment this is also satisfied.

		3. $\psi$ is of the form $(Es\to A[s/x])\to B$. Let $w^E\geq u^E$. If $w^E, v\models B$ we are done, so suppose otherwise. We know that $u, v\models \forall xA\to B$, i.e. if $w, v\not\models B$ then $w, v\not\models \forall xA$ so there exists $w' \geq w$ and $m_{w'}\in M_{w'}$ such that $m_{w'}\notin A^{I_{w'}}$. But then $m_{w'}\in E^{I_{w'^E}^E}$ and $m_{w'}\notin A^{I_{w'^E}^E}$, i.e. $w'^E, u\models Es$ and $w'^E, u\not\models As$
	\end{proof}

	\subsection{Classical Translation}

	At this point we proceed similarly to the previous section. But instead of introducing new atoms $A^M$, we will encode the used hypotheses via terms, i.e. for each $n$-ary relation symbol $A$ we use a new $n+1$-ary relation symbol $A^\#$. Intuitively the additional argument ranges over worlds and encodes if $A$ is true at that world. The idea is analogous to the previous section in that this allows us to force certain hypotheses. We then obtain $\mathcal S^\#$ from $\mathcal S$ by including formulas as follows:
	\begin{itemize}
		\item $A^\#(\vec t, k)\to B^\#(\vec t, k)$ for some new free variable $k$ for every formula $A(\vec t)\to B(\vec t)\in\mathcal S$.
		\item $A^\#(\vec t, k)\to B^\#(\vec t, k)\to C^\#(\vec t, k)$ for some new free variable $k$ for every $A(\vec t)\to B(\vec t)\to C(\vec t)\in\mathcal S$.
		\item $B^\#(\vec t, f(k))\to C^\#(\vec t, k)$ and $A^\#(\vec t, f(l))$ where $f$ is some new function symbol and $k, l$ are new free variables for every $(A(\vec t)\to B(\vec t))\to C(\vec t)\in\mathcal S$.
	\end{itemize}
	Let $F_\to$ denote the set of function symbols $f$ introduced in this steps.
	\begin{itemize}
		\item $D^\#(\vec t, f(k))\to D^\#(\vec t, g(k))\to C^\#(t, k)\to D^\#(t, k)$, $A^\#(\vec t, f(l))$ and $B^\#(\vec t, g(m))$ where $f, g$ are new function symbols $k,l,m$ are new free variables for every $(A(\vec t)\to D(\vec t))\to (B(\vec t)\to D(\vec t))\to C(\vec t)\to D(\vec t)\in\mathcal S$.
	\end{itemize}
	Let $F'_\to$ denote the set of function symbols $f, g$ introduced in this steps.
	\begin{itemize}
		\item $A(\vec t, s(k), f(k))\to B(\vec t, k)$ and $E(s(l), f(l))$ where $k,l$ are new free variables $s, f$ are new function symbols for every $\forall xA(\vec t, x)\to B(\vec t)\in\mathcal S$.
	\end{itemize}
	Let $F_\forall$ denote the set of function symbols $f$ introduced in this step and $s_f$ the corresponding functions $s$.
	\begin{itemize}
		\item For every $f, g\in F_\to\cup F'_\to\cup F_\forall$ include $f(f(a)) = f(a)$, $g(f(a)) = f(g(a))$ and $A(\vec t, k)\to A(\vec t, f(k))$ for every relation symbol $A$ (including $E$).
		\item $E(a, k)\to A^\#(\vec t, k)\to B^\#(\vec t, a, k)$ where $a, k$ are some new free variables for all $A(\vec t_1)\to \forall xB(\vec t_2, x)\in\mathcal S$.
		\item $E(a, k)\to A^\#(\vec t, a, k)\to B^\#(\vec t, k)$ where $a, k$ are some new free variables for all $\exists xA(\vec t, x)\to B(\vec t)\in\mathcal S$.
		\item $A^\#(\vec t, k)\to B^\#(\vec t, s(\vec t, k), k)$ and $A^\#(\vec t, k)\to E(s(\vec t, k), k)$ where $k$ is a new free variables and $s$ a new function symbol for all $A(\vec t)\to \exists xB(\vec t, x)\in\mathcal S$.
	\end{itemize}
		Let $S_\exists$ denote the set of function symbols $s$ introduced in this step.
	\begin{itemize}
		\item $\bot^\#(k)\to A^\#(\vec a, k)$ for each relation $A$ and new free variables $\vec a, k$.
	\end{itemize}
	Define $$\varphi^\# = \bigwedge S^\#\to P_\varphi^\#(\vec t, s_\varphi)$$ where $s_\varphi$ is some new constant (i.e. $0$-ary function symbol).
	\begin{lemma}
		$\varphi^\to$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
	\end{lemma}

	\begin{proof}
		We proceed analogously to the previous chapter.

		Suppose $\varphi^\to$ is not valid intuitionistically, i.e. there is a Kripke model $\mathcal K$ with world $u_0\in W$ and variable assignment $v$ such that $u_0, v\not\models P_\varphi(\vec t)$ and $u_0, v\models \psi(\vec t)$ for each $\psi(\vec t)\in\mathcal{S}$. For each world $u$ let $M_{u}$ be the domain at $u$ and $I_u$ the interpretation.
		W.l.o.g. for $u_1, u_2\in W$ and $m_1\in M_{u_1}, m_2\in M_{u_2}$ have $m_1 = m_2$ only if there exists a world $w\in W$ that is comparable with both $u_1$ and $u_2$ and with $m_1 = m_2\in M_w$ (if this does not exists we can make them distinct by indexing $m$ with the set of comparable worlds it occurs at).

		We then construct a classical counter model $\mathcal M$ to $\varphi^\#$ as follows:
		\begin{itemize}
			\item As a domain $M$ take the union of
			\begin{itemize}
				\item the union of domains $M_u$ at worlds $u\geq u_0$.
				\item the set of worlds $u\in W$ with $u\geq u_0$.
				\item an additional element $m_\bot$, that will be the result of applying a function to invalid arguments.
			\end{itemize}
			These three sets are also pairwise disjoint w.l.o.g.
			\item As a variable assignment take $v$.
			\item Define the interpretation $I$ as follows:
			\begin{itemize}
				\item $s_\varphi^I = u_0$.
				\item For each function symbol $f\in F_\to$ recall the corresponding formula $(A(\vec t)\to B(\vec t))\to C(\vec t)$ with which it was introduced. Let $u\in W$ with $u\geq u_0$. We know that $u, v\models (A(\vec t)\to B(\vec t))\to C(\vec t)$. If $u, v\models C(\vec t)$ let $f^I(u) = u$. Otherwise choose $w\geq u$ such that $w, v\models A(\vec t)$ and $w, v\not\models B(\vec t)$ and let $f^I(u) = w$. For every other $m\in M$ let $f^I(m) = m_\bot$.
				\item Analogously for $f\in F'_\to$ recall the corresponding formula $(A(\vec t)\to D(\vec t))\to (B(\vec t)\to D(\vec t))\to C(\vec t)\to D(\vec t)$. Let $u\in W$ with $u\geq u_0$. We know that $u, v\models (A(\vec t)\to D(\vec t))\to (B(\vec t)\to D(\vec t))\to C(\vec t)\to D(\vec t)$. If $u, v\models D(\vec t)$ let $f^I(u) = u$. Otherwise choose $w\geq u$ such that $w, v\not\models D(\vec t)$ and one of $w, v\not\models A(\vec t)$, $w, v\models B(\vec t)$, $w, v\models C(\vec t)$
				and let $f^I(u) = w$. For every other $m\in M$ let $f^I(m) = m_\bot$.
				\item Similarly for each $f\in F_\forall$ recall its formula $\forall xA(\vec t, x)\to B(\vec t)$. Let $u\in W$ with $u\geq u_0$. If $u, v\models B(\vec t)$ again have $f^I(u) = u$ and let $c_f^I(u)\in M_u$ be arbitrary. Otherwise choose a world $w\geq u$ and $m\in M_w$ such that $w, v[w/a]\not\models A(\vec t, a)$ and let $f^I(u) = w$ and $c_f^I(u) = m$. For every other $m\in M$ let $f^I(m) = c_f^I(m) = m_\bot$.
				\item Finally for each $s\in S_\exists$ recall the corresponding formula $A(\vec t)\to \exists xB(\vec t, x)$. Let $u\in W$ with $u\geq u_0$. If $u, v\not\models A(\vec t)$ have $s^I(\vec t, u) = m_\bot$. Otherwise choose $m\in M_u$ such that $u, v[m/a]\models B(\vec t, a)$ and let $s^I(\vec t^{I_u, v}, u) = m$. For every other $m\in M$ let $s^I(m, u) = m_\bot$.
				\item Every other $n$-ary function symbol $f$ is part of the original signature and we interpret it accordingly, i.e. if there exists a world $u$ with $\vec m\in M_u^n$ then $f^I(\vec m) = f^{I_u}(\vec m)$ and otherwise  $f^I(\vec m) = m_\bot$. Our distinctness condition at the start makes this unique and well-defined, i.e. there are no two worlds $u_1, u_2$ with $\vec m\in M_{u_1}^n\cap M_{u_2}^n$ but $ f^{I_{u_1}}(\vec m)\neq  f^{I_{u_2}}(\vec m)$.
				\item We interpret $E$ via $(m, u)\in E^{I}$ iff $u\in W$, $m\in M_u$.
				\item Finally $A$ is part of the original signature for every other $n+1$-ary relation symbol $A^\#$ and we interpret it accordingly, i.e. $(\vec m, u)\in A^I$ iff $u\in W$ and $\vec m\in A^{I_u}$. 
			\end{itemize}
		\end{itemize}
		We claim that $\mathcal M, v\not\models P^\#(\vec t, s_\varphi)$ and $\mathcal M, v\models \psi$ for every $\psi\in \mathcal S^\#$. The former is clear since $s_\varphi^I = u$ and $u, v\models P(\vec t)$. For the latter there are 12 cases.
	\end{proof}

	\chapter{Proof translation}

	\section{Translation of counter-models}

	There have been various results on when classical provability implies intuitionistic provability, most famously for $\Pi_2$-formulas in Arithmetic~\cite{Friedman_1978}, but most recently also in some other cases~\cite{schwichtenberg, practical}. The constructibility of a bounded counter-model as in~\ref{finitecountermodel} can be used as an inspiration to give a similar result, i.e. roughly when such a construction is possible then classical validity implies intuitionistic validity. Note that the results in this section will applicable to predicate logic already.

	For this section we fix some set $\mathcal S$ of (predicate) formulas where each formula $\varphi$ is of the forms $\forall \vec x\psi$ for some $\psi$ of the form
	$$A_\psi\to(B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi)$$$$(A_\psi\vee B_\psi)\to C_\psi, A_\psi\to B_\psi\to C_\psi, (A_\psi\to B_\psi)\to  C_\psi$$$$\forall xA_\psi\to B_\psi, A_\psi\to \forall xB_\psi, \exists xA_\psi\to B_\psi, A_\psi\to\exists xB_\psi$$
	for some atomic relations $A_\psi, B_\psi, C_\psi$. We write $\varphi(\vec t)$ for $\psi[\vec t/\vec x]$. Note that we can interpret $A\to B$ as e.g. $A\to (B\wedge B)$.

	\subsection{Dependence}

	\begin{definition}
		A \textit{dependence order} $\precsim$ (w.r.t. $\mathcal S$) is a pre-order on a subset of ground predicates such that for all $\varphi\in\mathcal S$, ground predicates $A, B, C$ and ground terms $\vec t$
		\begin{itemize}
			\item If $\varphi(\vec t) = A\to (B\wedge C)$ then $A\precsim B$ and $A\precsim C$.
			\item If $\varphi(\vec t) = (A\wedge B)\to C$ then for each $D$ with $D\precsim B$ and $D\precsim C$ we have $D\precsim A$.
			\item If $\varphi(\vec t) = A\to (B\vee C)$ then for each $D$ with $B\precsim D$ and $C\precsim D$ we have $A\precsim D$.
			\item If $\varphi(\vec t) = (A\vee B)\to C$ then $A\precsim C$ and $B\precsim C$.
			\item If $\varphi(\vec t) = A\to \forall xB$ then $A\precsim B[t/x]$ for every ground term $t$.
			\item If $\varphi(\vec t) = \exists xA\to B$ then $A[t/x]\precsim B$ for every ground term $t$.
			\item If $\varphi(\vec t) = A\to \exists xB$ then for each $D$ with $B[t/x]\precsim D$ for each $t$ we have $A\precsim D$.
		\end{itemize}
		We say that a dependence order is \textit{sufficient} for some ground predicate $P$ if for each $\varphi\in\mathcal S$, ground predicates $A, B, C$ and ground terms $\vec t$
		\begin{itemize}
			\item If $\varphi(\vec t) = (A\to B)\to C$ and $C\precsim P$ then $A\not\precsim P$ and $\precsim$ is sufficient for $B$.
			\item If $\varphi(\vec t) = \forall xA\to B$ and $B\precsim P$ then $\precsim$ is sufficient for $A[t/x]$ for every ground term $t$.
		\end{itemize}
	\end{definition}

	\begin{theorem}~\label{dependencetheorem}
		Consider some formula $\varphi = \bigwedge\mathcal S\to P$. If there is a dependence order w.r.t. $\mathcal S$ that is sufficient for $P(s)$ and $\varphi$ is classically valid then $\varphi$ is intuitionistically valid.
	\end{theorem}

	\begin{example}
		Consider the formula $(A\to B)\vee (B\to A)$. We know from the previous section that this is intuitionistically equivalid to $$\bigwedge\mathcal S\to P_\varphi(s)$$
		where
		\begin{align*}
			\mathcal S = &\{\forall u((P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u))\}\\
			&\cup \{\forall u((P_A(f(u))\to P_B(f(u)))\to P_{A\to B}(u))\}\\
			&\cup \{\forall u((P_B(f(u))\to P_A(f(u)))\to P_{B\to A}(u))\}\\
			&\cup \{\forall u(P_A(u)\to A(u)), \forall u(A(u)\to P_A(u))\}\\
			&\cup \forall u(P_B(u)\to B(u)), \forall u(B(u)\to P_B(u))\}
		\end{align*}
		Consider the following dependence order

		\begin{center}
			\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
				\node[squarednode]      (upper)                          {$P_\varphi(s), P_{A\to B}(s), P_{B\to A}(s)$};
				\node[squarednode]      (lower)       	[below=of upper] {$P_A(f(s)), P_B(f(s)), A(f(s)), B(f(s))$};
				\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north) -- (upper.south);
			\end{tikzpicture}
		\end{center}

		i.e. for all $P, Q$ in the same rectangle have $P\precsim Q$ and $P\precsim Q$, and for $P$ in the lower $Q$ in the upper $P\prec Q$.
	\end{example}

	\chapter{Implementation}

	\bibliographystyle{plain}
	\bibliography{EIICL}

\end{document}

### 2022-07-07 CLS/lipics-v2021-EIICL.tex
CLS paper

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

\usepackage{tikz-qtree}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{arrows,shapes,quotes}

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Embedding intuitionistic into classical logic} %TODO Please add

%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line

\author{Alexander {Pluska}}{Faculty of Mathematics, Universtät Wien, Austria}{e11941874@student.tuwien.ac.at}{https://orcid.org/my-orcid?orcid=0000-0002-7709-3335}{}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Florian Zuleger}{Institute of Logic and Computation, Technische Universität Wien, Austria}{florian.zuleger@tuwien.ac.at}{https://orcid.org/0000-0003-1468-8398}{}

\authorrunning{A. Pluska and F. Zuleger} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Jane Open Access and Joan R. Public} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[100]{\textcolor{red}{Theory of Computation Logic Constructive Mathematics}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Intuitionistic Logic, Automated Reasoning, Proof Theory} %TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
The famous double negation translation~\cite{glivenko1929quelques, godel1933intuitionistischen} establishes an embedding from classical into intuitionistic logic. Curiously the reverse direction has not been covered in literature, perhaps because going from the more expressive logic to the less expressive one necessitates a much more involved encoding. We present an embedding from intuitionistic to classical logic, both in the propositional and first-order case, as well as an embedding of intuitionistic propositional logic into quantified boolean formulas. 
Furthermore in the propositional case we establish a parameter that bounds the proof complexity of intuitionistic formulas. In general validity of intuitionistic propositional logic is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} but this gives us a finer-grained result. In this process we also present a Tseytin-like translation that preserves intuitionistic validity and reduces formulas to a normal form.
Finally in the first order-case we hope that our translation allows the use state-of-the-art theorem provers for classical first-order logic such as Vampire~\cite{kovacs2013first} for checking intuitionistic validity and program synthesis. 
\end{abstract}

\section{Introduction}

Most state-of-the-art first-order theorem provers leverage classical logic~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}. Besides historic reasons a main advantage of classical logic over intuitionistic logic is the existence of convenient normal forms which enable reasoning techniques such as superposition and resolution. However there are also some drawbacks of this approach, in particular the non-constructiveness of classical logic poses a problem in synthesis-related tasks. On the other hand the Curry-Howard correspondence makes program extraction from intuitionistic proofs trivial~\cite{howard1980formulae} but automated deduction for intuitionistic logic is notoriously difficult.

There are different approaches for program extraction using classical provers. The first is simply to examine under which circumstances classical validity entails intuitionistic validity. Most famously this holds for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}. There are also other circumstances under which this entailment holds~\cite{schwichtenberg}, however this approach cannot encapsulate the general case. Another popular approach is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}. In this case we necessarily lose realizability at least of $\exists$ and $\vee$, which - one could argue - is why we want a program in the first place.

We examine a new direction, i.e. for each formula $\varphi$ giving a translation $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid. We shall start with the simpler case of propositional formulas. The process will be divided into three steps. The first is a first-order encoding of intuitionistic Kripke semantics, which gives an obvious but unsatisfactory embedding. We then discuss how some of the quantifiers can be eliminated via Herbrandization and give a Tseytin-like transformation that exposes a parameter bounding the proof complexity of the propositional formula. This finally allows us to eliminate all remaining quantifiers and we end up with a propositional formula. Since validity of intuitionistic logic is \verb+PSPACE+-complete while for classical logic it only \verb+coNP+-complete as one could expect there is a blow-up which is exponential in the parameter that we previously isolated.
Instead of targetting classical propositional logic at the last step we give an alternate translation to QBF, which is linear in the size of our complexity parameter. For predicate formulas the process is a bit more involved in that we also need to respect the constructive content of quantifiers. In particular the different semantics of the $\forall$-quantifier proves challenging. After dealing with this, we are able to utilize a quite similar process as in the propositional case.

\section{IPC to CQC}

The most obvious approach to embedding intuitionistic into classical logic is to express intuitionistic semantics, i.e. Kripke frames, as a first order theory. For every propositional variable $A$ consider a unary predicate $A$ of the same name where $A(u)$ expresses that $A$ is true at some world $u$. We can then naively encode formulas as follows:

\begin{definition}
	Let $\varphi$ be a propositional formula. Define $\varphi^{k}$ inductively as follows:
	\begin{itemize}
		\item $A^{k} = A(k)$ for every propositional variable $A$.
		\item $(\varphi\wedge\psi)^k = \varphi^k\wedge\psi^k$.
		\item $(\varphi\vee\psi)^k = \varphi^k\vee\psi^k$.
		\item $(\varphi\to \psi)^k = \forall u(k\preceq u\to\varphi^{u}\to\psi^{u})$ where $u$ is some new free variable.
	\end{itemize}
	Let $K(\varphi)$ encode that $\preceq$ is a partial order, the persistency condition for all propositional variables occurring in $\varphi$
	\begin{align*}
		K(\varphi) = &\forall k(k\preceq k)\wedge \forall k\forall l(k\preceq l\to l\preceq k\to l = k) \wedge \\&\forall k\forall l\forall m(k\preceq l \to l\preceq m\to k\preceq m)\wedge\\&\bigwedge \{\forall k\forall l(k\preceq l\to A(k)\to A(l))\:|\: \text{ $A$ occurs in $\varphi$}\}
	\end{align*}
	Then define
	$$\varphi^{C} = K(\varphi)\to \varphi^{u}$$
	where $u$ is some free variable.
\end{definition}

All that we have done is model Kripke frames with a finite number of propositional variables as a first-order theory and by definition we have

\begin{lemma}
	$\varphi$ is intuitionistically valid iff $\varphi^C$ is classically valid.
\end{lemma}

\begin{example}
	Consider the law of excluded middle $$\varphi = (A\to \bot)\vee A$$ which is not intuitionistically valid. Then
	\begin{align*}
		\varphi^{C} &= K(\varphi)\to ((A\to \bot)\vee A)^u\\
		&= K(\varphi)\to (A\to \bot)^u\vee A^u\\
		&= K(\varphi)\to (\forall u'( u\preceq u' \to A^{u'}\to \bot^{u'})\vee A(u))\\
		&= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot(u'))\vee A(u))
	\end{align*}
	and indeed $\varphi^{C}$ admits a counter model with domain $\{u, u'\}$ where $u \preceq u'$, $A(u) = 0, A(u') = 1$.
\end{example}

\section{Quantifiers in Classical Logic}

Of course translating a propositional formula to a first-order formula is somewhat unsatisfying. As a first step to obtaining a propositional formula we wish to eliminate the quantifiers. Since we are interested in preserving validity we use Herbrandization. In this process we introduce free variables and also additional function symbols to the signature. Whenever we speak of a new free variable what is meant is a free variable that does not occur in any of the considered formulas. Whenever a new function symbol is introduced we implicitly extend the signature by some not previously contained symbol.

Now usually in classical logic before removing quantifiers prenexification is performed, however since we wish to apply a similar procedure to intuitionistic logic later, where prenexification is not valid, we will do without.

\begin{definition}
	For formulas $\varphi$ we define the Skolemization $\varphi^S_T$ and Herbrandization $\varphi^H_T$ with respect to $T$ by simultaneous induction as follows:
	\begin{itemize}
		\item $A^S_T = A^H_T = A$ for each atomic $A$.
		\item $(\varphi\circ\psi)^X_T = \varphi^X_T\circ\psi^X_T$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
		\item $(\varphi\to\psi)^S_T = \varphi^H_T\to \psi^S_T$\\$(\varphi\to\psi)^H_T = \varphi^S_T\to\psi^H_T$.
		\item $(\forall x\varphi)^S_T = \forall x(\varphi[a/x]^S_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_T = \varphi[s(t_1,\dots,t_n)/x]^H_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$.
		\item $(\exists x\varphi)^S_T = \varphi[s(t_1,\dots,t_n)/x]^S_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$\\$(\exists x\varphi)^H_T = \exists x(\varphi[a/x]^H_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable.
	\end{itemize}
	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
\end{definition}

\begin{theorem}For every formula $\varphi$
	\begin{itemize}
		\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
		\item $\varphi$ and $\varphi^H$ are classically equivalid.
	\end{itemize}
\end{theorem}

which follows from the following two Lemmata, proofs of which can be found in the appendix.

\begin{lemma}\label{ap1}
	Let $\chi$ be a formula and $\{t_1\dots t_n\} = T$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
	\begin{itemize}
		\item if $\mathcal M, v\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
	\end{itemize}
\end{lemma}

\begin{lemma}\label{ap2}
	For every structure $\mathcal M$ and $\{t_1\dots t_n\} = T$ that contains all free variables in $\chi$ and variable assignment $v$
	\begin{itemize}
		\item if $\mathcal M, v\models\varphi^S_T$ then $\mathcal M, v\models \varphi$
		\item if $\mathcal M, v\not\models\varphi^H_T$ then $\mathcal M, v\not\models\varphi$.
	\end{itemize}
\end{lemma}

\section{Equivalid normalization}

Now trying to apply Herbrandization to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated. We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form. A similar translation was proposed in~\cite{statman1979intuitionistic}.

\begin{definition}
	Let $\varphi$ be some formula. For each sub(semi)formula $\psi$ of $\varphi$ consider some new $n$-ary relation symbol $P_\psi$ where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$. We denote those variables with $\vec t_\psi$ in some arbitrary but fixed order. We inductively define clause sets $\mathcal S^+(\varphi)$ and $\mathcal S^-(\varphi)$ as follows:
	\begin{itemize}
		\item For every atomic formula $A$\\$\mathcal S^+(A) = \{P_A(\vec t_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec t_A)\}$
		\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$\\$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec t_{\varphi\circ\psi})\to P_{\varphi}(\vec t_\varphi)\circ P_{\psi}(\vec t_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec t_\varphi)\circ P_{\psi}(\vec t_\psi)\to P_{\varphi\circ\psi}(\vec t_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
		\item For every implication $\varphi \to\psi$\\$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec t_{\varphi\to\psi})\wedge P_{\varphi}(\vec t_\varphi))\to P_{\psi}(\vec t_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec t_\varphi)\to P_{\psi}(\vec t_\psi))\to P_{\varphi\to\psi}(\vec t_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\varphi)$
		\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$\\$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec t_{Qx\varphi})\to QxP_{\varphi}(\vec t_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec t_{\varphi})\to P_{Qx\varphi}(\vec t_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
	\end{itemize}
	and have $\mathcal S(\varphi) = \mathcal S^+(\varphi)\cup\mathcal S^-(\varphi)$.
\end{definition}

\begin{lemma}The following are true for both classical and intuitionistic semantics:
	\begin{itemize}
		\item $\bigwedge S^+(\varphi)\wedge P_\varphi\models\varphi$. 
		\item $\bigwedge S^-(\varphi)\wedge \varphi\models P_\varphi$. 
	\end{itemize}
\end{lemma}

\begin{proof}
	The claims are verified in a straightforward manner by induction on the height of $\varphi$.
\end{proof}

\begin{example}
	Consider $\varphi = \neg\forall x A(x)\to \exists x\neg A(x)$. Then
	\begin{align*}
		\mathcal S^+(\varphi) = &\{(P_\varphi\wedge P_{\neg\forall x A(x)})\to P_{\exists x\neg A(x)}, (P_{\forall xA(x)}\to\bot)\to P_{\neg \forall xA(x)}\}\\&\cup\{\forall xP_{A(x)}(x)\to P_{\forall xA(x)}, \forall x(P_{A(x)}(x)\to A(x))\}\\&\cup\{P_{\exists x\neg A(x)}\to \exists xP_{\neg A(x)}(x), \forall x(P_{\neg A(x)}(x)\to P_{A(x)}(x)\to \bot)\}\\&\cup\{\forall x(A(x)\to P_{A(a)}(x))\}\\
		\mathcal S^-(\varphi) = &\{(P_{\neg\forall x A(x)}\to P_{\exists x\neg A(x)})\to P_\varphi,  (P_{\neg \forall xA(x)}\wedge P_{\forall xA(x)})\to\bot\}\\&\cup\{P_{\forall xA(x)}\to \forall xP_{A(x)}(x), \forall x(A(x)\to P_{A(x)}(x))\}\\&\cup\{\exists xP_{\neg A(x)}(x)\to P_{\exists x\neg A(x)}, \forall x((P_{A(x)}(x)\to \bot)\to P_{\neg A(x)}(x))\}\\&\cup\{\forall x(P_{A(a)}(x)\to A(x))\}
	\end{align*}
\end{example}

We can give a transformation of structures that corresponds to the above transformation of sentences.

\begin{definition}
	For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms $P_\psi$. By slight abuse of notation we denote with $\vec t_\psi$ elements of the domain instead of variables. In the classical case have	
	$$P_{\varphi\wedge\psi}^I(\vec t_{\varphi\wedge\psi}) \Leftrightarrow P_{\varphi}^I(\vec t_\varphi)\wedge P_{\psi}^I(\vec t_\psi)\indent P_{\varphi\vee\psi}^I(\vec t_{\varphi\vee\psi}) \Leftrightarrow P_{\varphi}^I(\vec t_\varphi)\vee P_{\psi}^I((\vec t_\psi))$$$$ P_{\varphi\to\psi}^I(\vec t_{\varphi\to\psi}) \Leftrightarrow (\neg P_{\varphi}^I(\vec t_{\varphi}))\vee P_{\psi}^I(\vec t_{\psi})$$
	$$P_{\forall x\varphi}^I(\vec t_{\forall x\varphi}) \Leftrightarrow \{\vec t\:|\:\:P_{\varphi}^I(\vec t_\varphi) \text{ for all $x\in M$}\}$$$$ P_{\exists x\varphi}^I(\vec t_{\exists x\varphi}) \Leftrightarrow \{\vec t\:|\:\:P_{\varphi}^I(\vec t_\varphi) \text{ for some $x\in M$}\}$$
	\\

	and for intuitionistic logic and some world $u$ 		
	$$P_{\varphi\wedge\psi}^{I_u}(\vec t_{\varphi\wedge\psi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi)\wedge P_{\psi}^{I_u}(\vec t_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec t_{\varphi\vee\psi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi)\vee P_{\psi}^{I_u}((\vec t_\psi))$$$$ P_{\varphi\to\psi}^{I_u}(\vec t_{\varphi\to\psi}) \Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec t_{\varphi}))\vee P_{\psi}^{I_w}(\vec t_{\psi})\text{ for all $w\geq u$}$$
	$$P_{\forall x\varphi}^{I_u}(\vec t_{\forall x\varphi}) \Leftrightarrow P_{\varphi}^{I_w}(\vec t_\varphi) \text{ for all $w\geq u$, $x\in M_w$}$$$$ P_{\exists x\varphi}^{I_u}(\vec t_{\exists x\varphi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi) \text{ for some $x\in M_u$}$$
	\\
\end{definition}

The above definitions are motivated by the following results which follow directly from the involved definitions and are true both for classical and intuitionistic semantics:

\begin{lemma}
	For every formula $\varphi$ and structure $\mathcal M$
	$$\mathcal S(\mathcal M, \varphi)\models\mathcal S(\varphi)$$
\end{lemma}

\begin{lemma}
	For every structure $\mathcal M$
	$$\mathcal M\models \varphi\text{ iff }S(\mathcal M, \varphi)\models P_\varphi$$
\end{lemma}

Then from the previous three Lemmas we get

\begin{corollary}\label{equivalid}
	In both intuitionistic and classical logic
	\begin{itemize}
		\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec t_\varphi)$ is. 
		\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec t_\varphi)$ is.
	\end{itemize}
\end{corollary}

\section{IPC to CPC}

We are now ready to put everything together. Recall that our goal is for a formula to give a modified formula that is valid classically iff the original formula is valid intuitionistically. As a first step we perform the normalization from the last section, i.e. instead of $\varphi$ we consider $$\varphi' = \bigwedge \mathcal S^-(\varphi)\to P_\varphi$$

Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
$$\hspace{-0.8cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$

where we can treat $A\to B$ as a special case $A\to (B\wedge B)$. We then apply the transformation to CQC, leading to a formula
$$\varphi'' = K(\varphi')\to\bigwedge  \mathcal S\to P_\varphi(u)$$
where $u$ is some new free variable and each $\psi\in\mathcal S$ is of one of the forms
$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\wedge (B_\psi(k))\to C_\psi(k)))$$
$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\vee B_\psi(k))\to C_\psi(k))$$
$$\forall k(u\preceq k\to (\forall l(k\preceq l \to A_\psi(l)\to B_\psi(l))\to C_\psi(k)))$$
with each $P(x)$ possibly being $\bot$. Applying Herbrandization we obtain a formula
$$\varphi''' = K(\varphi')\to \bigwedge \mathcal S'\to P_\varphi(s)$$
where $s$ is some new constant term and each $\psi\in\mathcal S'$ is of one of the forms
$$\hspace*{-2mm}\forall k(s\preceq k\to A_\psi(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k(s\preceq k\to (A_\psi(k)\wedge (B_\psi(k))\to C_\psi(k)))$$
$$\hspace*{-2mm}\forall k(s\preceq k\to A_\psi(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(s\preceq k\to (A_\psi(k)\vee (B_\psi(k))\to C_\psi(k)))$$
$$\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A_\psi(f_\psi(k))\to B_\psi(f_\psi(k)))\to C_\psi(k))$$
and $f$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model then we have a counter-model $(M',I')$ where $$M' = \{m\in M|s\preceq m\}$$ and 
$$f_\psi^{I'}(u) = \begin{cases}
	f^I_\psi(u)&\text{ iff $u\preceq f^{I}_\psi(u)$}\\
	u&\text{ else}
\end{cases}$$ that is where $s$ is a least element and the $f_\psi$ increase their argument. Therefore instead of $\varphi'''$ we may consider a formula $$\varphi^\circ = \forall k(s\preceq k)\to \bigwedge\{\forall k(k\preceq f_\psi(k))\:|\:\psi\in\mathcal F_\to\}\to K(\varphi')\to \bigwedge \mathcal S^\circ\to P_\varphi(s)$$ where each formula $\psi\in\mathcal S^\circ$ is of one of the forms
$$\forall k(A_\psi(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k((A_\psi(k)\wedge B_\psi(k))\to C_\psi(k))$$
$$\forall k(A_\psi(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k((A_\psi(k)\vee B_\psi(k))\to C_\psi(k))$$
$$\forall k((A_\psi(f_\psi(k))\to B_\psi(f_\psi(k)))\to C_\psi(k))$$
and $\mathcal F_\to$ denotes the set of formulas of the last kind.

We now argue that if a counter-model exists for $\varphi^\circ$ then a bounded counter-model exists, allowing us to return to the propositional case.

\begin{definition}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. For $\psi\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true. For $\psi\in\mathcal F_\to$ define $$g_\psi : M\to M, u\mapsto\begin{cases}
		u&\text{ if $\psi$ is fulfilled in $u$}\\
		f^I_\psi(u)&\text{ else}		
	\end{cases}$$Define a $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ is the set of sequences without repetition on $\mathcal F_\to$.
		\item Interpret $\preceq$ as the prefix-order.
		\item Have $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
			\psi_1\dots\psi_n&\text{if $\psi$ occurs in $\psi_1\dots\psi_n$}\\
			\psi_1\dots\psi_n\psi&\text{else}			
		\end{cases}$$
		\item For propositional variables $P$ have $$P^{I_T}(\psi_1\dots \psi_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots))$$
	\end{itemize}
\end{definition}
\begin{lemma}
	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
	\begin{enumerate}
		\item $\psi$ is fulfilled at $f_\psi^I(u)$ for all $u\in M, \psi\in\mathcal F_\to$.
		\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\geq u$.
	\end{enumerate}

\end{lemma}

\begin{proof}
	1. If $C_\psi^I(u)$ is true then we are done due to persistency. Otherwise we know that $(A_\psi^I(f_\psi^I(u))\to B_\psi^I(f_\psi^I(u)))\to C_\psi^I(f_\psi^I(u))$ holds, so $A_\psi^I(f_\psi^I(u))\to B_\psi^I(f_\psi^I(u))$ must be true.

	2. If $C_\psi^I(u)$ is true then $C_\psi^I(w)$ is also true and we are done. So suppose $A_\psi^I(u)\to B_\psi^I(u)$ is false, i.e. $A_\psi^I(u)$ true and $B_\psi^I(u)$ false. Then due to the persistency condition $A_\psi^I(w)$ is also true. Again if $C_\psi^I(w)$ is true we are done. Otherwise we know that $B^I(f^I_\psi(w))$ must be false, since $(A^I_\psi(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I_\psi(w)$ and $A^I_\psi(f^I_\psi(w))$ hold. But then due to persistency so is $B^I_\psi(w)$ and we have that $\psi$ is fulfilled at $w$.
\end{proof}

Essentially what fulfilment of $\psi$ at $u$ tells us is that having $f_\psi(w) = w$ for all $w\geq u$ works. 
With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.

\begin{corollary}
	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
\end{corollary}

\begin{example}
	Consider
	$$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model
	\begin{center}
		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
			\node (center) {};
			\node[squarednode]      (right)   [right=of center] {$B, C$};
			\node[squarednode]      (left)    [left=of center] {$A, D$};
			\node[squarednode]      (lower)       	[below=of center] {};
			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left.south east);
			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right.south west);			
		\end{tikzpicture}
	\end{center}
	\vspace*{-.3cm}where at each node the true propositions are indicated.\pagebreak

	There is a corresponding classical counter-model $\mathcal M = (M, I)$ to $\varphi^\circ$ with $M = \{u, u_l, u_r\}$, $A^{I}(l) = D^I(l) = B^I(r) = C^I(r) = 1$ and $0$ else.

	Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (A\to B)\to C$ this corresponds to a counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below
	\begin{center}
		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
			\node (center) {};
			\node[squarednode]      (right1)   [right=of center] {$B, C$};
			\node[squarednode]      (right2)   [above=of right1] {$B, C$};
			\node[squarednode]      (left1)    [left=of center] {$A, D$};
			\node[squarednode]      (left2)    [above=of left1] {$A, D$};
			\node[squarednode]      (lower)       	[below=of center] {};
			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left1.south east) node [midway, below left] {$f_{\psi_1}$};
			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right1.south west) node [midway, below right] {$f_{\psi_2}$};
			\draw[-{Stealth[length=3mm, width=3mm]}] (left1.north) -- (left2.south) node [midway, right] {$f_{\psi_2}$};
			\draw[-{Stealth[length=3mm, width=3mm]}] (right1.north) -- (right2.south) node [midway, left] {$f_{\psi_1}$};
			\draw[-{Stealth[length=3mm, width=3mm]}] (left1.south west) to [out=225, in=135, loop, looseness=3] node [left] {$f_{\psi_1}$} (left1.north west);		
			\draw[-{Stealth[length=3mm, width=3mm]}] (right1.south east) to [out=315, in=45, loop, looseness=3] node [right] {$f_{\psi_2}$} (right1.north east);
			\draw[-{Stealth[length=3mm, width=3mm]}] (left2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (left2.north east);
			\draw[-{Stealth[length=3mm, width=3mm]}] (right2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (right2.north east);		
		\end{tikzpicture}
	\end{center}
\end{example}

So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
	x&\text{ if $\psi$ occurs in $x$}\\
	x\psi&\text{ otherwise}
\end{cases}$$ Now we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula. This gives us the following translation:

\begin{definition}
	Suppose $\mathcal S$ is a set of formulas of the form
	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C),(A\vee B)\to C, (A\to B)\to C$$
	Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$. For each atom $A$ occurring in $\mathcal S$ and $k\in \mathcal F_\to$ consider a new atom $A^{k}$, which will correspond to $A^{I_F}(k)$ from the previous definition. Obtain $\mathcal S^\#$ by including the formulas
	\begin{itemize}
		\item $A^k\to A^{k\psi}$ for $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
		\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
	\end{itemize}
\end{definition}

\begin{theorem}
	$$\mathcal S\models_C P\text{ iff }\mathcal S^\#\models_I P^\emptyset$$
\end{theorem}

\begin{example}
	Let us consider the law of excluded middle $\varphi = A\vee\neg A$. Then we get
	$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\}$$
	and we know that $\varphi$ is valid iff $\bigwedge \mathcal S(\varphi)\to P_\varphi$ is valid. Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
	\begin{align*}
		\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi}\}\\ &\cup\{(P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi},  A^\epsilon\to P_A^\epsilon\}\\ &\cup \{P_A^{\psi}\to P_A^{\psi}, (A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}
	\end{align*}
	It is easy to construct a counter example:
	$$P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$$
	$$P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$$
\end{example}

\section{IPC to QBF}

It is well known that a polynomial time translation between QBF and IPC must exists, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}. However giving an explicit translation from IPC to QBF can still be useful as good solvers exists for QBF but for IPC not so much. Furthermore we will be able to give fragments of IPC for which validity is complete for other stages of the polynomial hierarchy. The translation from the last section will serve as a useful tool for this. We will not explain the semantics of QBF in depth. Intuitively just think that the interpretation of existentially quantified variables may depend on variables quantified above them.

Fix some set $\mathcal S$ of formulas $\psi$ of the form $$\hspace*{-.7cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$and a formula $$\varphi = \bigwedge \mathcal S\to P$$
and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e. express which path is considered.

\begin{definition}
	For every propositional variable $X$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $X^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$. Let $\vec X^n$ range over all propositional variables $X^n$ and $\vec F^n$ range over all $F_\psi^n$. Define
	\begin{itemize}
		\item valid$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e. if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
		\item per($n$) which encodes that the persistency condition holds.
		\item sat$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
		\item sat$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
	\end{itemize}
	Have $k = |\mathcal F_\to|$ and define
	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{per}(k)\wedge \text{sat}_{\mathcal S}(k)\wedge \text{sat}_{\mathcal F_\to}(k)\right)$$
	for $0 < i < k$
	$$\hspace*{-.8cm}\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{per}(i)\wedge \text{sat}_{\mathcal S}(i)\wedge \text{sat}_{\mathcal F_\to}(i)\wedge\left(\text{valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
	and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{sat}_{\mathcal S}(0)\wedge(\text{valid}(0)\to \varphi_1^Q)\right)$$
	Example encodings of the above formulas would be
	\begin{align*}
		\text{valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\wedge\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
		\text{per}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
		\text{sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\to C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
		\text{sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
	\end{align*}
\end{definition}

\begin{example}
	With the above encodings for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
	$$\hspace*{-.3cm}
	\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right)
	$$
	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $F^0$.
\end{example}

\begin{lemma}
	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
\end{lemma}
\begin{proof}
	Suppose $\varphi$ is not intuitionistically valid, i.e. there exists a counter-model $\mathcal M$ for $\mathcal S^\#\to P^\epsilon$. For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$. Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.

	Let the $F^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified). If not exactly one is interpreted as true then valid$(n)$ fails, i.e. have $\psi_n\in\mathcal F_\to$ the only element with $F^n_{\psi_n} = 1$. If $F^i_{\psi_n} = 1$ for some $i < n$ then valid$(n)$ also fails, so we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions. Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{\psi_1\dots\psi_n}$. Continue this construction until $n  = |\mathcal F_\to|$. Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.

	On the other hand suppose $\varphi^Q$ is satisfiable. We construct a counter-example to $\varphi^\#$.
	Again we proceed iteratively. Interpret $A^\epsilon$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$. Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$ in which $F^i_\psi$ is chosen true iff $\psi = \psi_i$. Let $\psi_n\in\mathcal F_\to$. Consider some interpretation of the $A^n$ that are part of a satisfying assignment where $F^n_\psi$ is true iff $\psi = \psi_n$ and all variables quantified above are chosen as before. Have $A^{\psi_0\dots\psi_n} = A^n$ for each propositional variable $A$. From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.
\end{proof}

Since we have not discussed QBF in detail the above argument might be a bit difficult to follow, in particular the second part, but is simple in essence. We shall demonstrate with help of a example.

\begin{example}
	Consider the formula $\varphi = (((A\to B)\to C)\wedge((B\to A)\to C))\to C$, which is a non-intuitionistic tautology. Its associated QBF is $$\varphi^Q = \exists A^0B^0C^0\forall F_{\psi_1}^1F_{\psi_2}^1 \exists A^1B^1C^1\forall F_{\psi_1}^2F_{\psi_2}^2 \exists A^2B^2C^2\psi'$$where $\psi_1 := (A\to B)\to C, \psi_2 = (B\to A)\to C$ and $$\psi' = \bigwedge\left\{\begin{matrix}
		F^1_{\psi_1}\vee F^1_{\psi_2}, \neg(F^1_{\psi_1}\wedge F^1_{\psi_2}), F^2_{\psi_1}\vee F^2_{\psi_2},\neg(F^2_{\psi_1}\wedge F^2_{\psi_2}),\\F^1_{\psi_1}\to\neg F^2_{\psi_1},F^1_{\psi_2}\to\neg F^2_{\psi_2}, A^0\to A^1,A^1\to A^2, 
		\\B^0\to B^1, B^1\to B^2, C^0\to C^1, C^1\to C^2,\neg C^0, \\
		F^1_{\psi_1}\to (A^1\to B^1)\to C^0, F^1_{\psi_2}\to(B^1\to A^1)\to C^0\\
		F^2_{\psi_1}\to (A^2\to B^2)\to C^1, F^2_{\psi_2}\to(B^2\to A^2)\to C^1
	\end{matrix}\right\}$$
	This QBF is satisfiable: Have $P^0 = 0$ for all atoms $P$, always $P^2 = 1$ for all atoms $P$ and $C^1 = 1$. Furthermore have $A^1 = 1$ iff $F^1_{\psi_1} = 1$ and $B^1 = 1$ iff $F^1_{\psi_2} = 1$. The construction of an counter-model to $\varphi^\#$ then proceeds as follows:
	\begin{itemize}
		\item First have $P^\epsilon = 0$ for each propositional variable $P$.
		\item For $P^{\psi_1}$ we consider what happens when we choose $F^1_{\psi_1}$ to be true, i.e. we have $A^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
		\item For $P^{\psi_2}$ we consider what happens when we choose $F^1_{\psi_2}$ to be true, i.e. we have $B^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
		\item Finally for $P^{\psi_1\psi_2} = P^{\psi_2\psi_1} = 1$ for all propositional variables $P$ as $P^2$ is true regardless of choice for the $F^i_\psi$. 
	\end{itemize}
\end{example}

Note that the last set of existentially quantified variables is superfluous because there is only one assignment which even has the chance to falsify $\varphi^Q$, namely that which assigns $1$ to the only $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$. That is in our final formula we can replace $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^i$. With that we get the following:

\begin{corollary}
	Let $N$-INTINVALID be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid. Then $N$-INTINVALID is in $\Sigma_{2N-1}^P$. The dual problem $N$-INTVALID is in $\Pi_{2N-1}^P$.
\end{corollary}

As a final result we would like to prove that the above problems are complete for their respective complexity classes. It is sufficient to provide a polynomial reduction from QBF with bounded number of quantifier alternation to them. Such translations exists~\cite{statman1979intuitionistic,vsvejdar2003polynomial}, however they are insufficient in that they generate more formulas of form $(A\to B)\to C$ than there are quantifier alternations. We therefore would need a slightly new translation.

\section{CQC to IQC}

We now give an analogous transformation for first order logic. However our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic. However we do it nonetheless! The domain of our classical model will feature on the one hand elements representing worlds in a Kripke frame and on the other hand the domain-elements from the Kripke model. We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$. In addition we use a unary predicate $W$ that encodes that an element represents a world.

First of all recall that due to~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all $\psi\in\mathcal S$ are of one of the forms

$$\forall \vec t(A(\vec t_{A})\to (B(\vec t_B)\wedge C(\vec t_C))), \forall \vec t((A(\vec t_{A})\wedge B(\vec t_B))\to C(\vec t_C)),$$$$ \forall \vec t(A(\vec t_{A})\to (B(\vec t_B)\vee C(\vec t_C))),
\forall \vec t((A(\vec t_{A})\vee B(\vec t_B))\to C(\vec t_C)),$$$$ \forall \vec t((A(\vec t_A)\to B(\vec t_B))\to C(\vec t_C)),\forall \vec t(\forall xA(\vec t_A)\to B(\vec t_B)),$$$$ \forall \vec t(A(\vec t_A)\to\forall xB(t_B)), \forall \vec t(\exists xA(\vec t_A)\to B(\vec t_B)), \forall \vec t(A(\vec t_A)\to\exists xB(t_B))$$	
First let us note how we axiomatize $E$ and $W$. For a vector $\vec x = x_1\dots x_n$ we define $\vec E(\vec x, u) = \bigwedge\{E(x_i, u)\:|\:i\leq n\}$.
\begin{align*}
	E(\varphi) = & \forall u\forall x(E(x, u)\to(\neg W(x)\wedge W(u)))\wedge\\
	& \forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))\wedge\\
	& \bigwedge\{\forall\vec t\forall u(\vec E(\vec t, u)\to E(f(\vec t), u))\:|\:\text{$f$ is an $n$-ary function symbol}\}
\end{align*}
The one possible problem in this encoding is that we have a global interpretation of function symbols, i.e. in an intuitionistic model we could have worlds $u, v$ such that $f^{I_u}\neq f^{I_v}$. However this is not a problem: We can take as the domain of our classical model equivalence classes $\{[(x, u)]\:|\: x\in M_u\}$ under the relation $(x, u)\sim (y, v)$ iff $x = y$ and there exists $w$ comparable with $u, v$ such that $x\in w$ and interpret $f^I([x, u]) = [f^{I_u}(x), u]$. Then our classical translation only forces functions to be equal on arguments on which they must be equal anyway. So the Kripke model does in fact not afford us any additional generality.

As before we need to axiomatize $\preceq$. For each $n$-ary relation symbol consider $P$ a new $n+1$ ary relation symbol $P^\#$.  The only difference in that for persistence we may require additional quantification. 
\begin{align*}
	K(\varphi) = & \forall k(k\preceq k)\wedge \forall k\forall l((k\preceq l\wedge l\preceq k)\to l = k)\wedge\\
	& \forall k\forall l\forall m((k\preceq l\wedge l\preceq m)\to k\preceq m)\wedge\\
	& \bigwedge\{\forall k\forall l(k\preceq l\to \forall \vec x(A^\#(\vec x, k)\to A^\#(\vec x, l)))\:|\:\text{$A$ occurs in $\varphi$}\}
\end{align*}
Using $E$ and $W$ we obtain a set $\mathcal S^\circ$ by including formulas as follows:

\begin{itemize}
	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to (A_\psi^\#(t_A, u)\to (B_\psi^\#(\vec t_B, u)\circ C_\psi^\#(\vec t_C, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec t(A_\psi(\vec t_A)\to (B_\psi(\vec t_B)\circ C_\psi(\vec t_C)))\in\mathcal S$
	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec t_A, w)\to B_\psi^\#(\vec t_B, w))\to C_\psi^\#(\vec t_C, u))$ for each $\psi = \forall \vec t((A_\psi(\vec t_A)\to B_\psi(\vec t_B))\to C_\psi(\vec t_C))\in\mathcal S$
	\item  $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec t_A, w)))\to B^\#(\vec t_B, u))$ for each $\forall \vec t(\forall xA_\psi(\vec t_A)\to B_\psi(\vec t_B))\in\mathcal S$.
	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to A^\#_\psi(\vec t_A, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec t_B, u)))$\\for each $\forall \vec t(A(\vec t_A)\to \forall xB(\vec t_B))\in\mathcal S$.
	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec t_A, u))\to B^\#_\psi(\vec t_B))$\\for each $\forall \vec t(\exists xA_\psi(\vec t_A)\to B_\psi(\vec t_B))\in\mathcal S$
	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to A^\#_\psi(\vec t_B)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec t_B, u)))$\\for each $\forall \vec t(A_\psi(\vec t_A)\to \exists xB_\psi(\vec t_B))\in\mathcal S$
\end{itemize}
Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid iff
$$\varphi^\circ= E(\varphi)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
is classically valid.

We proceed now as in the previous section. Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec t((A_\psi(\vec t_A)\to B_\psi(\vec t_B))\to C_\psi(\vec t_C))$ with $\mathcal F_\to$ and of the form $\forall \vec t(\forall xA_\psi(\vec t_A)\to B_\psi(\vec t_B))$ with $\mathcal F_\forall$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$. As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
$$\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec t_A, w)\to B_\psi^\#(\vec t_B, w))\to C_\psi^\#(\vec t_C, u))$$ with
$$\forall \vec t\forall u(\vec E(\vec t, u)\to (A_\psi^\#(\vec t_A, f_\psi(\vec t, u))\to B_\psi^\#(\vec t_B, f_\psi(\vec t, u)))\to C_\psi^\#(\vec t_C, u))$$ and
$$\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec t_A, w)))\to B^\#(\vec t_B, u))$$ with $$\forall \vec t\forall u(\vec E(\vec t, u)\to (\forall x(E(x, f_\psi(\vec t, u))\to A_\psi^\#(\vec t_A, f_\psi(\vec t, u))))\to B^\#(\vec t_B, u))$$
and in each other formula in $\mathcal S^\circ$ remove the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S$. Then the previous formulas are equivalid to
$$\hspace*{-.5cm}\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f_\psi(\vec t, u))\:|\:\psi\in\mathcal F\}\to E(\varphi)\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
As in the previous section we can define a tree-counter model for this.
\begin{definition}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. For $\psi = \forall\vec t\psi'\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff $(A_\psi^I(\vec t_A, u)\to B_\psi^I(\vec t_B, u))[\vec a/\vec t]$ is false or $C_\psi^I(\vec t_C, u)[\vec a/\vec t]$ is true.
	For $\psi = \forall\vec t\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff there is $a\in M_u$ such that $A^{I[a/x]}_\psi(\vec t_A)[\vec a/\vec t]$ is false or $B_\psi^I(\vec t_B)[\vec a/\vec t]$ is true. For $\psi\in\mathcal F$ define $$g_{\vec a, \psi} : M\to M, u\mapsto\begin{cases}
		u&\text{ if $\psi$ is fulfilled in $u$ with $\vec a$}\\
		f^I_\psi(\vec a, u)&\text{ else}		
	\end{cases}$$		
	Define a $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ are the sequences on $\{ \vec t, \psi\:|\:\psi\in \mathcal F, \vec t\in M^n\}$ without repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\to$ and without immediate repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\forall$.
		\item Interpret $\preceq$ as the prefix-order.
		\item Have $$f_\psi^{I_T}(x_1\dots x_n) = \begin{cases}
			x_1\dots x_n&\text{if $x=x_n$}\\
			x_1\dots x_nx&\text{else}			
		\end{cases}$$
		\item For propositional variables $P$ have $$P^{I_T}(x_1\dots x_n) = P^I(g_{x_n}(\dots(g_{x_1}(s^I))\dots))$$
	\end{itemize}
\end{definition}

\begin{lemma}
	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
	\begin{enumerate}
		\item $\psi$ is fulfilled at $f_\psi^I(\vec a, u)$ with $\vec a$ for all $u\in M, \psi\in\mathcal F$, $\vec a\in M^n$.
		\item If $\psi\in\mathcal F_\to$ is fulfilled at $u\in M$ with $\vec a$ then $\psi$ is fulfilled at all $w\geq u$ with $\vec a$.
	\end{enumerate}	
\end{lemma}

\begin{proof}
	For $\psi\in\mathcal F_\to$ the proof is analogous to the previous section.

	1. Let $\psi\in\mathcal F_\forall$. If $B^I_\psi(\vec t_B, u)[\vec a/\vec t]$ holds then we are done due to persistency. Otherwise there is $a$ with $E^I(a, f^I_\psi(\vec a, u))$ such that $A^{I[a/x]}_\psi(\vec t_A, f^I_\psi(\vec a, u))$ is false. So $\psi$ is fulfilled at $f^I_\psi(\vec a, u)$ with $\vec a$.
\end{proof}

That the second point does not work for $\psi\in\mathcal F_\forall$ is why we must switch to immediate repetition. It does not work because even if for all $x$ with $E(x, w$) we have $A(x, w)$  there could be some term $t$ with $\neg E(t, w)$ and $v$ with $w\preceq v$ such that $E(t, v)$ and $\neg A(t, w)$.

\begin{example}
	Consider the $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model
	\begin{center}
		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
			\node[squarednode]      (0)   [] {$ $};
			\node[squarednode]      (1r)   [above right=of 0] {{\color{blue}$1$}};
			\node[squarednode]      (1l)   [above left=of 0] {{\color{red}$1$}};
			\node[squarednode]      (2r)   [above=of 1r] {{1, \color{blue}$2$}};
			\node[squarednode]      (2l)   [above=of 1l] {{1, \color{red}$2$}};
			\node[squarednode]      (3r)   [above=of 2r] {{1, 2, \color{blue}$3$}};
			\node[squarednode]      (3l)   [above=of 2l] {{1, 2, \color{red}$3$}};
			\node      (4r)   [above=of 3r] {{$\vdots$}};
			\node      (4l)   [above=of 3l] {{$\vdots$}};
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (0.north east) -- (1r.south west);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (0.north west) -- (1l.south east);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1r.north) -- (2r.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1l.north) -- (2l.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1l.north east) -- (2r.south west);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1r.north west) -- (2l.south east);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2r.north) -- (3r.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2l.north) -- (3l.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2l.north east) -- (3r.south west);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2r.north west) -- (3l.south east);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3r.north) -- (4r.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3l.north) -- (4l.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3l.north east) -- (4r.south west);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3r.north west) -- (4l.south east);
		\end{tikzpicture}
	\end{center}
	where at each node the elements that exist at it are indicated. For black ones both $A$ and $B$ hold, for red ones only $B$, for blue ones only $A$. Furthermore red arrows indicate $f_{\neg \forall xA(x)}$ and blue arrows $f_{\neg \forall xB(x)}$. Note that applying $f_{\neg \forall xA(x)}$ we must always reach the left column and $f_{\neg \forall xB(x)}$ we must always reach the right column. Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we can never remain stationary. This is in contrast to the propositional case.
\end{example}

\begin{corollary}
	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
\end{corollary}

Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$ is sufficient. In particular this allows us to completely eliminate $\preceq$ from the formula, that is we replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec t\forall u(E(x, u)\to E(x, f_\psi(\vec t, u)))\:|\:\psi\in\mathcal F_\to\}$ in $E(\varphi)$ to get $E^\#(\varphi)$ and obtain a formula

$$\varphi^{\#} = E^\#(\varphi)\to \bigwedge S^\#\to P^\#$$

\begin{theorem}
	$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
\end{theorem}

\begin{example}
	Consider the double negation shift (DNS)
	$$\varphi = \forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$$
	which is a famous non-intuitionistic tautology. We get
	\begin{align*}
		\mathcal S = & \{(P_{\forall x\neg\neg A(x)}\to P_{\neg\neg\forall xA(x)})\to P_\varphi, P_{\forall x\neg\neg A(x)}\to \forall xP_{\neg\neg A(x)}(x)\}\cup\\
		& \{\forall t((P_{\neg\neg A(x)}(t)\wedge P_{\neg A(x)}(t))\to \bot), \forall t((A(t)\to \bot)\to P_{\neg A(x)})(t)\}\cup\\
		& \{(P_{\neg\forall xA(x)}\to \bot)\to P_{\neg\neg\forall xA(x)}, (P_{\neg\forall xA(x)}\wedge P_{\forall xA(x)})\to \bot\}\cup\\
		&\{\forall xA(x)\to P_{\forall xA(x)}\}
	\end{align*}
	and thus, denoting $\psi_1 = (P_{\neg\forall xA(x)}\to \bot)\to P_{\neg\neg\forall xA(x)}$ and $\psi_2 = \forall xA(x)\to P_{\forall xA(x)}$, we have
	\begin{align*}
		\mathcal S^\# = & \{\forall u(P_{\forall x\neg\neg A(x)}(u)\to P_{\neg\neg\forall xA(x)}(u))\to P_\varphi(u)\}\cup\\& \{\forall u(P_{\forall x\neg\neg A(x)}(u)\to \forall xP_{\neg\neg A(x)}(x, u)\}\cup\\
		& \{\forall t, u((P_{\neg\neg A(x)}(t, u)\wedge P_{\neg A(x)}(t, u))\to \bot)\}\\& \{\forall t, u((A(t, u)\to \bot)\to P_{\neg A(x)}(t, u))\}\cup\\
		& \{\forall u((P_{\neg\forall xA(x)}(f_{\psi_1}(u))\to \bot)\to P_{\neg\neg\forall xA(x)}(u))\}\\& \{\forall u((P_{\neg\forall xA(x)}(u)\wedge P_{\forall xA(x)}(u))\to \bot)\}\cup\\
		&\{\forall u(\forall xA(x, f_{\psi_2}(u))\to P_{\forall xA(x)}(u))\}
	\end{align*}
\end{example}

\bibliography{lipics-v2021-EIICL}

\appendix

\begin{proof}[Lemma \ref{ap1}]
	We proceed by simultaneous induction on the formula height.

	For atomic $\chi$ the claims are clear.

	Otherwise there are $3$ cases:

	1. $\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$. By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v}), \mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as in the lemma. Then we can define $I_{\chi, v}$ as follows:

	For every symbol $p$ that occurs in $\varphi^S$ or $\varphi^H$ but not $\psi^S, \psi^H$ have $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$. For every symbol $p$ that occurs in $\psi^S$ or $\psi^H$ but not $\varphi^S, \varphi^H$ have $p^{I_{\chi, v}} = p^{I_{\psi, v}}$. Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.

	Let $\circ=\to$. Suppose $\mathcal M, v\models\chi$. Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_T$ or $\mathcal M_\chi\models\psi^H_T$, i.e. $\mathcal M_\chi\models\chi^S_T$. On the other hand suppose $\mathcal M, v\not\models\chi$. Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_T$ and $\mathcal M_\chi, v\not\models\psi^S_T$, i.e. $\mathcal M_\chi\not\models\chi^H_T$. Analogous arguments work for $\circ\in\{\wedge, \vee\}$.

	2. $\chi = \forall x\varphi$. Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise. Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.

	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma. Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.

	Suppose $\mathcal M, v\models\chi$. Then for all $m\in M$ $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{M\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{M\cup\{a\}}[x/a])$, i.e. $\mathcal M_\chi,v\models \chi^S$. On the other hand suppose $\mathcal M, v\not\models\chi$. Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{T\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(t_1,\dots t_n)/x]^H_T$, i.e. $\mathcal M_\chi, v\not\models(\forall x\varphi)^H_T$.

	3. $\chi = \exists x\varphi$. The argument runs dually to 2. Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise. Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.

	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma. Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.

	Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_T$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_T$.
\end{proof}

\begin{proof}[Lemma \ref{ap2}]
	Again we proceed by simultaneous induction on the formula height.

	For atoms the claims are clear. We distinguish 5 cases.

	1. $\chi = \varphi\to\psi$. Suppose $\mathcal M, v\models\chi^S_T$, i.e. $\mathcal M, v\not\models\varphi^H_T$ or $\mathcal M, v\models\psi^S_T$. By induction hypothesis $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e. $\mathcal M, v\models \chi$. On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e. $\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_T$. Again by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.

	2. + 3. Conjunctions and Disjunctions are dealt with analogously.

	4. $\chi = \forall x\varphi$.  Suppose $\mathcal M, v\models\chi^S_T$, i.e. for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{T\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$. Then it follows that $\mathcal M, v\models\varphi$. On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e. there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{T\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e. $\mathcal M, v\not\models\forall x\chi$.

	5. $\chi = \exists x\varphi$.  The argument runs dually to 4.
\end{proof}

\end{document}

### 2022-07-07 Notes/Embedding intuitionistic into classical logic/EIICL.tex
CLS paper

@@
 \usetikzlibrary{arrows.meta}
+\usetikzlibrary{arrows,shapes,quotes}
 \usepackage{stackengine}
@@
 			\item $f^{I_w}|_{M_u} = f^{I_u}$.
-			\item $R^{I_w}|_{M_u} = f^{I_u}$.
+			\item $R^{I_w}|_{M_u} = R^{I_u}$.
 		\end{itemize}
@@
 Kripke frames, as a first order theory.
-For every atom $A$ consider a unary predicate $A$ of the same name where $A(u)$ expresses that $A$ is true at some world $u$.
+For every propositional variable $A$ consider a unary predicate $A$ of the same name where $A(u)$ expresses that $A$ is true at some world $u$.
 We can then naively encode formulas as follows:
@@
 		\begin{itemize}
-			\item $A^{k} = A(k)$ for every atom $A$.
+			\item $A^{k} = A(k)$ for every propositional variable $A$.
 			\item $(\varphi\wedge\psi)^k = \varphi^k\wedge\psi^k$.
@@
 		\end{itemize}
-		Let $K(\varphi)$ encode that $\preceq$ is a partial order, the persistency condition for all propositional variables occurring in $\varphi$ and ex falso quodlibet for each world, i.e.
+		Let $K(\varphi)$ encode that $\preceq$ is a partial order, the persistency condition for all propositional variables occurring in $\varphi$
 		\begin{align*}
-			K(\varphi) = &\forall k(k\preceq k)\wedge \forall k\forall l(k\preceq l\to l\preceq k\to l = k) \wedge \\&\forall k\forall l\forall m(k\preceq l \to l\preceq m\to k\preceq m)\wedge\\&\bigwedge \{\forall k\forall l(k\preceq l\to A(k)\to A(l))\:|\: \text{ $A$ occurs in $\varphi$}\}\wedge\\&\bigwedge\{\forall k(\bot(k)\to A(k))\:|\: A \text{ occurs in }\varphi\}
+			K(\varphi) = &\forall k(k\preceq k)\wedge \forall k\forall l(k\preceq l\to l\preceq k\to l = k) \wedge \\&\forall k\forall l\forall m(k\preceq l \to l\preceq m\to k\preceq m)\wedge\\&\bigwedge \{\forall k\forall l(k\preceq l\to A(k)\to A(l))\:|\: \text{ $A$ occurs in $\varphi$}\}
 		\end{align*}
@@
 
-	All that we have done is model Kripke frames with a finite number of propositional variables as a first-order theory.
-The only attention must be paid with regard to $\bot$, which we have encoded because it will be useful for translation to propositional logic later.
-Consider the formula $\varphi^C_\bot$, which is exactly as above except that $\bot^k = \bot$.
-If we have a counter model to $\varphi^C_\bot$ we obtain a counter model for $\varphi^C$ by setting $\bot(k)$ false for all $k$.
-On the other hand if we have a counter model for $\varphi^C$ then we obtain one for $\varphi^C_\bot$ by restricting our domain to all $u$ for which $\bot(u)$ does not hold.
-Therefore this modification is unproblematic and we have:
+	All that we have done is model Kripke frames with a finite number of propositional variables as a first-order theory and by definition we have
 
@@
 	\begin{example}
-		Consider the law of excluded middle $$\varphi = (A\to \bot)\vee A$$which is not intuitionistically valid. Then
+		Consider the law of excluded middle $$\varphi = (A\to \bot)\vee A$$ which is not intuitionistically valid. Then
 		\begin{align*}
@@
 						 &= K(\varphi)\to (A\to \bot)^u\vee A^u\\
-						 &= K(\varphi)\to (\forall u'( u\preceq u' \to A^{u'}\to \bot^{u'})\vee A^\#(u))\\
+						 &= K(\varphi)\to (\forall u'( u\preceq u' \to A^{u'}\to \bot^{u'})\vee A(u))\\
 						 &= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot(u'))\vee A(u))
 		\end{align*}
-		and indeed $\varphi^{C}$ admits a counter model with domain $\{u, u'\}$ where $u \preceq u'$, $A(u) = \bot(u) = \bot(u') = 0, A(u') = 1$.
+		and indeed $\varphi^{C}$ admits a counter model with domain $\{u, u'\}$ where $u \preceq u'$, $A(u) = 0, A(u') = 1$.
 	\end{example}
@@
 
-	\begin{corollary}
+	\begin{corollary}\label{equivalid}
 		In both intuitionistic and classical logic
@@
 	$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\wedge (B_\psi(k))\to C_\psi(k)))$$
-	$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\vee (B_\psi(k))\to C_\psi(k)))$$
+	$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\vee B_\psi(k))\to C_\psi(k))$$
 	$$\forall k(u\preceq k\to (\forall l(k\preceq l \to A_\psi(l)\to B_\psi(l))\to C_\psi(k)))$$
-
-	Applying Herbrandization we obtain a formula
+	with each $P(x)$ possibly being $\bot$. Applying Herbrandization we obtain a formula
 	$$\varphi''' = K(\varphi')\to \bigwedge \mathcal S'\to P_\varphi(s)$$
@@
 	\begin{definition}
-		Suppose we are given a counter-model $(M, I)$ for $\varphi^\circ$.
-For $\psi\in\mathcal F_\to$ we say that is it \textit{fulfilled} at a world $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true.
+		Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
+For $\psi\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true.
 For $\psi\in\mathcal F_\to$ define $$g_\psi : M\to M, u\mapsto\begin{cases}
@@
 			f^I_\psi(u)&\text{ else}		
-		\end{cases}$$Define a $(M_T, I_T)$ as follows:
+		\end{cases}$$Define a $\mathcal M_T = (M_T, I_T)$ as follows:
 	\begin{itemize}
@@
 		\end{cases}$$
-		\item For propositional variables $P$ have $$P^{I_T}(x_1\dots x_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots))$$
-	\end{itemize}
-	\end{definition}
-
-	The following easy technical Lemma is quite useful.
+		\item For propositional variables $P$ have $$P^{I_T}(\psi_1\dots \psi_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots))$$
+	\end{itemize}
+	\end{definition}
 	\begin{lemma}
-		Let $(M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
+		Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
 		\begin{enumerate}
@@
 	\begin{corollary}
-		If $(M, I)$ is a counter-model to $\varphi^\circ$ then so is $(M_T, I_T)$.
+		If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
 	\end{corollary}
 
-	The size of this model is $|\mathcal F_\to|!$ in the theoretical worst case.
-While this would already allow us to return to the propositional world we can do better than that.
-In fact we can merge all branches into a model of size $2^{|\mathcal F_\to|}$.
-
-	\begin{definition}
-		Suppose we are given a counter-example $(M_T, I_T)$ to $\varphi^\circ$ as above. Consider $(M_F, I_F)$ as follows:
-		\begin{itemize}
-			\item $M_F = 2^{F_\to}$, i.e. the set of subsets of $F_\to$.
-			\item $\preceq$ is interpreted as the subset relation.
-			\item $s^{I_M} = \emptyset$, $f_\psi^{I_F}(u) = u\cup\{\psi\}$.
-			\item For every atom $P$ have $P^{I_F}(u)$ iff for every $w\in M_T$ where each $\psi\in u$ is fulfilled we have $$P^{I_T}(w)$$
-		\end{itemize}
-	\end{definition}
-
-	Note that in this definition for the first time it is of crucial importance that we have also transformed $\bot$ into a unary predicate that we may set to true, as this could be forced in certain cases, in particular if there is no $w$ at which each $\psi\in u$ is fulfilled.
-
-	Of course we have to prove that this model transformation is correct:
-
-	\begin{definition}
-		If $(M_T, I_T)$ is a counter-model for $\varphi^\circ$ then so is $(M_F, I_F)$.
-	\end{definition}
-	\begin{proof}
-		We have to check that all 5 types of formulas are still satisfied. The proofs for the first 4 are very similar, so we shall just present one.
-
-		Suppose $\chi$ is of  the form $\forall k(A_\chi(k)\to (B_\chi(k)\vee C_\chi(k))$. Let $u\subseteq \mathcal F_\to$.
-	\end{proof}
-
-	Therefore we know that $\varphi^\circ$ is valid iff it is valid for structures with domain $2^{F_\to}$, i.e.
-the power-set of $F_\to$, ordered by the subset relation and $s^I = \emptyset, f_\psi^I(x) = x\cup\{\psi\}$.
-Thus we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula.
+	\begin{example}
+		Consider
+		$$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model
+		\begin{center}
+			\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
+				\node (center) {};
+				\node[squarednode]      (right)   [right=of center] {$B, C$};
+				\node[squarednode]      (left)    [left=of center] {$A, D$};
+				\node[squarednode]      (lower)       	[below=of center] {};
+				\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left.south east);
+				\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right.south west);			
+			\end{tikzpicture}
+		\end{center}
+		\vspace*{-.3cm}where at each node the true propositions are indicated.\pagebreak
+
+		There is a corresponding classical counter-model $\mathcal M = (M, I)$ to $\varphi^\circ$ with $M = \{u, u_l, u_r\}$, $A^{I}(l) = D^I(l) = B^I(r) = C^I(r) = 1$ and $0$ else.
+
+		 Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (A\to B)\to C$ this corresponds to a counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below
+		\begin{center}
+			\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
+				\node (center) {};
+				\node[squarednode]      (right1)   [right=of center] {$B, C$};
+				\node[squarednode]      (right2)   [above=of right1] {$B, C$};
+				\node[squarednode]      (left1)    [left=of center] {$A, D$};
+				\node[squarednode]      (left2)    [above=of left1] {$A, D$};
+				\node[squarednode]      (lower)       	[below=of center] {};
+				\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left1.south east) node [midway, below left] {$f_{\psi_1}$};
+				\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right1.south west) node [midway, below right] {$f_{\psi_2}$};
+				\draw[-{Stealth[length=3mm, width=3mm]}] (left1.north) -- (left2.south) node [midway, right] {$f_{\psi_2}$};
+				\draw[-{Stealth[length=3mm, width=3mm]}] (right1.north) -- (right2.south) node [midway, left] {$f_{\psi_1}$};
+				\draw[-{Stealth[length=3mm, width=3mm]}] (left1.south west) to [out=225, in=135, loop, looseness=3] node [left] {$f_{\psi_1}$} (left1.north west);		
+				\draw[-{Stealth[length=3mm, width=3mm]}] (right1.south east) to [out=315, in=45, loop, looseness=3] node [right] {$f_{\psi_2}$} (right1.north east);
+				\draw[-{Stealth[length=3mm, width=3mm]}] (left2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (left2.north east);
+				\draw[-{Stealth[length=3mm, width=3mm]}] (right2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (right2.north east);		
+			\end{tikzpicture}
+		\end{center}
+	\end{example}
+
+	So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
+		x&\text{ if $\psi$ occurs in $x$}\\
+		x\psi&\text{ otherwise}
+	\end{cases}$$ Now we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula.
 This gives us the following translation:
@@
 		$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C),(A\vee B)\to C, (A\to B)\to C$$
-		Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$.
-For each atom $A$ occurring in $\mathcal S$ and $M\subseteq \mathcal F_\to$ consider a new atom $A^{M}$, which will correspond to $A^{I_F}(M)$ from the previous definition.
+		Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$.
+For each atom $A$ occurring in $\mathcal S$ and $k\in \mathcal F_\to$ consider a new atom $A^{k}$, which will correspond to $A^{I_F}(k)$ from the previous definition.
 Obtain $\mathcal S^\#$ by including the formulas
 		\begin{itemize}
-			\item $\bot^N\to A^N, A^N\to A^M$ for atoms $A$ occurring in $\mathcal S$ and $N\subseteq M\subseteq \mathcal F_\to$.
-			\item $A^M\to (B^M\circ C^M)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $M\subseteq \mathcal F_\to$.
-			\item $(A^M\circ B^M)\to C^M$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $M\subseteq \mathcal F_\to$.
-			\item $(A^{M\cup \{\psi\}}\to B^{M\cup \{\psi\}})\to C^M$ for $\psi = (A\to B)\to C\in\mathcal S$, $M\subseteq \mathcal F_\to$.
+			\item $A^k\to A^{k\psi}$ for $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
+			\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+			\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+			\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
 		\end{itemize}
@@
 		\begin{align*}
-			\mathcal S^\# =&\{P_\varphi^\emptyset\to P_\varphi^{\{\psi\}}, P_A^\emptyset\to P_A^{\{\psi\}},P_{\neg A}^\emptyset\to P_{\neg A}^{\{\psi\}},A^\emptyset\to A^{\{\psi\}}\}\\ &\cup\{(P_A^\emptyset\vee P_{\neg A}^\emptyset)\to P_\varphi^\emptyset, (P_A^{\{\psi\}}\vee P_{\neg A}^{\{\psi\}})\to P_\varphi^{\{\psi\}},  A^\emptyset\to P_A^\emptyset\}\\ &\cup \{P_A^{\{\psi\}}\to P_A^{\{\psi\}}, (A^{\{\psi\}}\to \bot)\to P_{\neg A}^\emptyset, (A^{\{\psi\}}\to \bot)\to P_{\neg A}^{\{\psi\}}\}
+			\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi}\}\\ &\cup\{(P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi},  A^\epsilon\to P_A^\epsilon\}\\ &\cup \{P_A^{\psi}\to P_A^{\psi}, (A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}
 		\end{align*}
 		It is easy to construct a counter example:
-		$$0 = P_\varphi^\emptyset  = P_A^\emptyset = P_{\neg A}^\emptyset = A^\emptyset$$
-		$$1=  P_{\varphi}^{\{\psi\}} = P_A^{\{\psi\}} =  P_{\neg A}^{\{\psi\}} = A^{\{\psi\}} $$
+		$$P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$$
+		$$P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$$
 	\end{example}
@@
 We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model.
-The idea how to get a linear-size formula is instead of expressing validity at every node of $\mathcal M_F$ - which was done in the CPC translation - we express that in each path in $\mathcal M_F$ from $\emptyset$ to $F_\to$ the worlds on the way satisfy all conditions for $\mathcal M_F$ being a counter-model.
-The universally quantified variables $F_\psi$ in the next definition will express which path is considered.
-
-	\begin{definition}
-		For every atom $A$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $A^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$. Define
-		\begin{itemize}
-			\item $K^n = \{\bot^i\to A^i, A^i\to A^j\:|\:\text{ for all atoms $A$ and $i < j\leq n$}\}$
-			\item $K_\to^n = \bigvee \{F_\psi^{n}\wedge\bigwedge\{\neg F^i_\psi\:|\:i<n\}\:|\:\psi\in\mathcal F_\to\}$, i.e.
-encoding that at each stage $n$ an $F_\psi^{n}$ that was not true at a previous stage is true, and thereby via the pigeon-hole principle also that exactly one of the $F_\psi^{n}$ is true at each of the $|\mathcal F_\to|$ stages.
-			\item $\mathcal S^n$ as $\mathcal S\setminus\mathcal F_\to$ where each atom $A$ has been replaced by $A^n$.
-			\item $\mathcal F_\to^n:=\left\{\bigvee\{F_\psi^i\:|\:i \leq n\}\to (A^{n+1}_\psi\to B^{n+1}_\psi)\to C^n_\psi\:|\:\psi\in\mathcal F_\to\right\}$
-		\end{itemize}
-		and let $k = |\mathcal F_\to|$. Have
-		$$\hspace*{-1cm}\varphi^Q = \exists \vec X^0\forall\vec F^0\dots\exists \vec X^k\left(\bigwedge \{K^{i}_\to\:|\:i < k\}\to\left(\neg P^0\wedge K^k\wedge\bigwedge\{\mathcal S^i\:|\:i\leq k\}\cup\{\mathcal F^i_\to\:|\: i< k\}\right)\right)$$
+The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
+The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e.
+express which path is considered.
+
+	\begin{definition}
+		For every propositional variable $X$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $X^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$.
+Let $\vec X^n$ range over all propositional variables $X^n$ and $\vec F^n$ range over all $F_\psi^n$.
+Define
+		\begin{itemize}
+			\item valid$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
+if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
+			\item per($n$) which encodes that the persistency condition holds.
+			\item sat$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
+			\item sat$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
+		\end{itemize}
+		Have $k = |\mathcal F_\to|$ and define
+		$$\varphi^Q_{k} = \exists \vec X^k\left(\text{per}(k)\wedge \text{sat}_{\mathcal S}(k)\wedge \text{sat}_{\mathcal F_\to}(k)\right)$$
+		for $0 < i < k$
+		$$\hspace*{-.8cm}\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{per}(i)\wedge \text{sat}_{\mathcal S}(i)\wedge \text{sat}_{\mathcal F_\to}(i)\wedge\left(\text{valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
+		and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{sat}_{\mathcal S}(0)\wedge(\text{valid}(0)\to \varphi_1^Q)\right)$$
+		Example encodings of the above formulas would be
+		\begin{align*}
+			\text{valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\wedge\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
+			\text{per}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
+			\text{sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
+			&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\to C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
+			\text{sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
+		\end{align*}
 	\end{definition}
@@
 	\begin{example}
-		For double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
+		With the above encodings for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
 		$$\hspace*{-.3cm}
-			\varphi^Q = \exists A^0,\bot^0\forall F^1\exists A^1,\bot^1\left(F^1\to\left(\neg A^0\wedge\bigwedge\left\{\begin{matrix}
-				\bot^0\to A^0,\bot^1\to A^1,\bot^0\to \bot^1,\\
-				 A^0\to A^1, (A^1\to \bot^1)\to \bot^0 
-			\end{matrix}\right\}\right)\right)
+			\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right)
 		$$
-		which is not a satisfiable QBF since $A^0 = \bot^0 = \bot^1 = 0, A^1 = 1$ falsifies it for any $F^1$.
+		which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $F^0$.
 	\end{example}
@@
 		Suppose $\varphi$ is not intuitionistically valid, i.e.
-there exists a counter-model $\mathcal M$ for $\mathcal S^\#\to P^\emptyset$.
-For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\emptyset$.
-Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and there exists a sequence $\emptyset = M_0\subseteq M_1\dots\subseteq M_{n-1}$ of subsets of $\mathcal F_\to$ such that $M_i$ contains exactly the $F_\psi$ for which $F_\psi^j$ is true for some $j\leq i$ and $A^i$ is interpreted as $A^{M_i}$ is in $\mathcal M$.
-Consider arbitrary interpretations for $F_\to^{n}$.
-If none is interpreted as true then $K^n_\to$ fails, if more than one is then $K^m_\to$ will fail for some $m\ > n$, in any case $\varphi^Q$ would be automatically satisfied.
-If exactly one is satisfied, say $\psi_n$, let $M_n = M_{n-1}\cup\{\psi_n\}$.
-Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{M_n}$.
-Continue this construction until $M_{|\mathcal F_\to|}  = \mathcal F_\to$ is reached.
+there exists a counter-model $\mathcal M$ for $\mathcal S^\#\to P^\epsilon$.
+For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$.
+Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.
+
+		Let the $F^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified).
+If not exactly one is interpreted as true then valid$(n)$ fails, i.e.
+have $\psi_n\in\mathcal F_\to$ the only element with $F^n_{\psi_n} = 1$.
+If $F^i_{\psi_n} = 1$ for some $i < n$ then valid$(n)$ also fails, so we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions.
+Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{\psi_1\dots\psi_n}$.
+Continue this construction until $n  = |\mathcal F_\to|$.
 Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.
 
-		On the other hand suppose $\varphi^Q$ is satisfiable. We construct a Kripke structure that is a counter-example to $\varphi$.
-		\begin{itemize}
-			\item As a set of potential worlds take non-repeating sequences of elements in $\mathcal F_\to$, some of these worlds will be removed later.
-			\item We define the valuations at each world inductively.
-At $\epsilon$ (i.e.
-the empty sequence) interpret $A$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$.
-Given a sequence $\chi_0\dots \chi_n$ such that for $i<n$ interpreting $A^i$ as it is interpreted in $\chi_0\dots\chi_i$ is part of a satisfying interpretation of $\varphi^Q$ where we have chosen $F^i_\psi$ true iff $\psi = \chi_i$, let $\chi_0\dots \chi_n$ interpret $A$ as $A^n$ in that interpretation of $\varphi^Q$ where we have chosen $F^n_\psi$ true iff $\psi = \chi_n$.
-			\item Defined as above this might require us to interpret $\bot$ as true at some worlds. We simply remove those worlds from the considered set.
-			\item The order on the remaining worlds is the usual prefix order.
-		\end{itemize}
-		From the definitions it directly follows that this constitutes a counter-example to $\varphi$.
+		On the other hand suppose $\varphi^Q$ is satisfiable. We construct a counter-example to $\varphi^\#$.
+		Again we proceed iteratively.
+Interpret $A^\epsilon$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$.
+Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$ in which $F^i_\psi$ is chosen true iff $\psi = \psi_i$.
+Let $\psi_n\in\mathcal F_\to$.
+Consider some interpretation of the $A^n$ that are part of a satisfying assignment where $F^n_\psi$ is true iff $\psi = \psi_n$ and all variables quantified above are chosen as before.
+Have $A^{\psi_0\dots\psi_n} = A^n$ for each propositional variable $A$.
+From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.
 	\end{proof}
@@
 		Consider the formula $\varphi = (((A\to B)\to C)\wedge((B\to A)\to C))\to C$, which is a non-intuitionistic tautology.
-Its associated QBF is $$\varphi^Q = \exists A^0B^0C^0\bot^0\forall F_{\psi_1}^1F_{\psi_2}^1 \exists A^1B^1C^1\bot^1\forall F_{\psi_1}^2F_{\psi_2}^2 \exists A^2B^2C^2\bot^2\psi'$$where $\psi_1 := (A\to B)\to C, \psi_2 = (B\to A)\to C$ and $$\psi' = \bigwedge\left\{\begin{matrix}
-			F^1_{\psi_1}\vee F^1_{\psi_2}, (F^2_{\psi_1}\wedge\neg F^1_{\psi_1})\vee (F^2_{\psi_2}\wedge\neg F^1_{\psi_2}),\\\bot^0\to A^0, \bot^0\to B^0,\bot^0\to C^0,\bot^1\to A^1,\bot^1\to B^1, \bot^1\to C^1,\\
-			\bot^2\to A^2, \bot^2\to B^2,\bot^2\to C^2,\bot^0\to \bot^1, \bot^1\to\bot^2, A^0\to A^1,
-			\\A^1\to A^2, B^0\to B^1, B^1\to B^2, C^0\to C^1, C^1\to C^2,\\
-			  \neg C^0, F^1_{\psi_1}\to (A^1\to B^1)\to C^0, F^1_{\psi_2}\to(B^1\to A^1)\to C^0\\
-			 (F^1_{\psi_1}\vee F^2_{\psi_1})\to (A^2\to B^2)\to C^1, (F^1_{\psi_2}\vee F^2_{\psi_2})\to(B^2\to A^2)\to C^1
+Its associated QBF is $$\varphi^Q = \exists A^0B^0C^0\forall F_{\psi_1}^1F_{\psi_2}^1 \exists A^1B^1C^1\forall F_{\psi_1}^2F_{\psi_2}^2 \exists A^2B^2C^2\psi'$$where $\psi_1 := (A\to B)\to C, \psi_2 = (B\to A)\to C$ and $$\psi' = \bigwedge\left\{\begin{matrix}
+			F^1_{\psi_1}\vee F^1_{\psi_2}, \neg(F^1_{\psi_1}\wedge F^1_{\psi_2}), F^2_{\psi_1}\vee F^2_{\psi_2},\neg(F^2_{\psi_1}\wedge F^2_{\psi_2}),\\F^1_{\psi_1}\to\neg F^2_{\psi_1},F^1_{\psi_2}\to\neg F^2_{\psi_2}, A^0\to A^1,A^1\to A^2,
+			\\B^0\to B^1, B^1\to B^2, C^0\to C^1, C^1\to C^2,\neg C^0, \\
+			  F^1_{\psi_1}\to (A^1\to B^1)\to C^0, F^1_{\psi_2}\to(B^1\to A^1)\to C^0\\
+			 F^2_{\psi_1}\to (A^2\to B^2)\to C^1, F^2_{\psi_2}\to(B^2\to A^2)\to C^1
 		\end{matrix}\right\}$$
-	This QBF is satisfiable: Have $P^0 = 0$ for all atoms $P$, always $P^2 = 1$ for all atoms $P$ and $C^1 = 1$, $\bot^1 = 0$.
+	This QBF is satisfiable: Have $P^0 = 0$ for all atoms $P$, always $P^2 = 1$ for all atoms $P$ and $C^1 = 1$.
 Furthermore have $A^1 = 1$ iff $F^1_{\psi_1} = 1$ and $B^1 = 1$ iff $F^1_{\psi_2} = 1$.
-The construction of an intuitionistic counter-model then proceeds as follows:
-	\begin{itemize}
-		\item As a set of potential worlds we have $\{\epsilon, \varphi_1, \varphi_2, \varphi_1\varphi_2, \varphi_2\varphi_1\}$.
-		\item Since $\bot^2$ is always true irrespective of how the $F^i_\psi$ are chosen, the worlds $\varphi_1\varphi_2, \varphi_2\varphi_1$ are deleted.
-		\item For the remaining three worlds we have
-		\begin{itemize}
-			\item$\epsilon\not\models A, B, C$
-			\item$\psi_1\models A, C, \psi_1\not\models B$ (corresponding to $F^1_{\psi_1} = 1, F^1_{\psi_2} = 0$)
-			\item$\psi_2\models B, C, \psi_2\not\models A$ (corresponding to $F^1_{\psi_1} = 0, F^1_{\psi_2} = 1$)
-		\end{itemize}
-		according to the above satisfying interpretation.
+The construction of an counter-model to $\varphi^\#$ then proceeds as follows:
+	\begin{itemize}
+		\item First have $P^\epsilon = 0$ for each propositional variable $P$.
+		\item For $P^{\psi_1}$ we consider what happens when we choose $F^1_{\psi_1}$ to be true, i.e. we have $A^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
+		\item For $P^{\psi_2}$ we consider what happens when we choose $F^1_{\psi_2}$ to be true, i.e. we have $B^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
+		\item Finally for $P^{\psi_1\psi_2} = P^{\psi_2\psi_1} = 1$ for all propositional variables $P$ as $P^2$ is true regardless of choice for the $F^i_\psi$. 
 	\end{itemize}
@@
 	We now give an analogous transformation for first order logic.
-However including quantifiers provides us with a new source of non-constructivism, i.e.
-$\neg\forall x A(x)\to \exists x\neg A(x)$ is a classical tautology that is not automatically eliminated by our above considerations.
-This makes skolemization tricky.
-
-	Analogously to the previous section we define a clause sets $S_\psi$ as follows:
-	\begin{itemize}
-		\item $S_\psi = \{P_\psi(\vec s)\to\psi(\vec t), \psi(\vec t)\to P_\psi(\vec t)\}$ if $\psi(\vec t)$ is atomic.
-		\item $S_\psi = \{P_\psi(\vec t)\to P_{\psi_1}(\vec t), P_\psi(\vec t)\to P_{\psi_2}(\vec t),P_{\psi_1}(\vec t)\to P_{\psi_2}(\vec t)\to P_\psi(\vec t)\}$ if $\psi(\vec t) = \psi_1(\vec t)\wedge\psi_2(\vec t)$.
-		\item $S_\psi = \{P_{\psi_1}(\vec t)\to P_\psi(\vec t), P_{\psi_2}(\vec t)\to P_\psi(\vec t)\}\cup $ $$\hspace{-1.5cm}\{(P_{\psi_1}(\vec t)\to P_\chi(\vec a))\to (P_{\psi_2}(\vec t)\to P_\chi(\vec a))\to P_\psi(\vec t)\to P_\chi(\vec a)\:|\:\chi \text{ is a subformula of $\varphi$}\}$$ if $\psi(\vec t) = \psi_1(\vec t)\vee\psi_2(\vec t)$ for appropriate vectors of new free variables $\vec a$.
-		\item $S_\psi = \{(P_{\psi_1}(\vec t)\to P_{\psi_2}(\vec t))\to P_\psi(\vec t), P_\psi(\vec t)\to P_{\psi_1}(\vec t)\to P_{\psi_2}(\vec t)\}$ if $\psi(\vec t) = \psi_1(\vec t)\to\psi_2(\vec t)$.
-		\item $S_\psi = \{\forall xP_{\psi_1}(\vec t, x) \to P_\psi(\vec t), P_\psi(\vec t)\to \forall xP_{\psi_1}(\vec t, x)\}$ if $\psi(\vec t) = \forall x(\psi_1(\vec t, x))$.
-		\item $S_\psi = \{\exists xP_{\psi_1}(\vec t, x) \to P_\psi(\vec t), P_\psi(\vec t)\to \exists xP_{\psi_1}(\vec t, x)\}$ if $\psi(\vec t) = \exists x(\psi_1(\vec t, x))$.
-	\end{itemize}
-	Again let $\mathcal S = (P_\bot \to\bot)\cup\bigcup\{S_\psi\:|\:\psi\text{ is a subformula of $\varphi$}\}$ and define
-	$$\varphi^\to = \bigwedge{\mathcal S} \to P_\varphi(\vec t)$$
-	Analogously to the previous section we have
+However our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic.
+However we do it nonetheless!
+The domain of our classical model will feature on the one hand elements representing worlds in a Kripke frame and on the other hand the domain-elements from the Kripke model.
+We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$.
+In addition we use a unary predicate $W$ that encodes that an element represents a world.
+
+	\subsection{CQC to IQC}
+
+	First of all recall that due to~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all $\psi\in\mathcal S$ are of one of the forms
+
+	$$\forall \vec t(A(\vec t_{A})\to (B(\vec t_B)\wedge C(\vec t_C))), \forall \vec t((A(\vec t_{A})\wedge B(\vec t_B))\to C(\vec t_C)),$$$$ \forall \vec t(A(\vec t_{A})\to (B(\vec t_B)\vee C(\vec t_C))),
+	\forall \vec t((A(\vec t_{A})\vee B(\vec t_B))\to C(\vec t_C)),$$$$ \forall \vec t((A(\vec t_A)\to B(\vec t_B))\to C(\vec t_C)),\forall \vec t(\forall xA(\vec t_A)\to B(\vec t_B)),$$$$ \forall \vec t(A(\vec t_A)\to\forall xB(t_B)), \forall \vec t(\exists xA(\vec t_A)\to B(\vec t_B)), \forall \vec t(A(\vec t_A)\to\exists xB(t_B))$$
+	First let us note how we axiomatize $E$ and $W$. For a vector $\vec x = x_1\dots x_n$ we define $\vec E(\vec x, u) = \bigwedge\{E(x_i, u)\:|\:i\leq n\}$.
+	\begin{align*}
+		E(\varphi) = & \forall u\forall x(E(x, u)\to(\neg W(x)\wedge W(u)))\wedge\\
+		 & \forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))\wedge\\
+		 & \bigwedge\{\forall\vec t\forall u(\vec E(\vec t, u)\to E(f(\vec t), u))\:|\:\text{$f$ is an $n$-ary function symbol}\}
+	\end{align*}
+	The one possible problem in this encoding is that we have a global interpretation of function symbols, i.e.
+in an intuitionistic model we could have worlds $u, v$ such that $f^{I_u}\neq f^{I_v}$.
+However this is not a problem: We can take as the domain of our classical model equivalence classes $\{[(x, u)]\:|\: x\in M_u\}$ under the relation $(x, u)\sim (y, v)$ iff $x = y$ and there exists $w$ comparable with $u, v$ such that $x\in w$ and interpret $f^I([x, u]) = [f^{I_u}(x), u]$.
+Then our classical translation only forces functions to be equal on arguments on which they must be equal anyway.
+So the Kripke model does in fact not afford us any additional generality.
+
+	As before we need to axiomatize $\preceq$.
+For each $n$-ary relation symbol consider $P$ a new $n+1$ ary relation symbol $P^\#$.
+The only difference in that for persistence we may require additional quantification.
+	\begin{align*}
+		K(\varphi) = & \forall k(k\preceq k)\wedge \forall k\forall l((k\preceq l\wedge l\preceq k)\to l = k)\wedge\\
+		& \forall k\forall l\forall m((k\preceq l\wedge l\preceq m)\to k\preceq m)\wedge\\
+		& \bigwedge\{\forall k\forall l(k\preceq l\to \forall \vec x(A^\#(\vec x, k)\to A^\#(\vec x, l)))\:|\:\text{$A$ occurs in $\varphi$}\}
+	\end{align*}
+	Using $E$ and $W$ we obtain a set $\mathcal S^\circ$ by including formulas as follows:
+
+	\begin{itemize}
+		\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to (A_\psi^\#(t_A, u)\to (B_\psi^\#(\vec t_B, u)\circ C_\psi^\#(\vec t_C, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec t(A_\psi(\vec t_A)\to (B_\psi(\vec t_B)\circ C_\psi(\vec t_C)))\in\mathcal S$
+		\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec t_A, w)\to B_\psi^\#(\vec t_B, w))\to C_\psi^\#(\vec t_C, u))$ for each $\psi = \forall \vec t((A_\psi(\vec t_A)\to B_\psi(\vec t_B))\to C_\psi(\vec t_C))\in\mathcal S$
+		\item  $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec t_A, w)))\to B^\#(\vec t_B, u))$ for each $\forall \vec t(\forall xA_\psi(\vec t_A)\to B_\psi(\vec t_B))\in\mathcal S$.
+		\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to A^\#_\psi(\vec t_A, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec t_B, u)))$\\for each $\forall \vec t(A(\vec t_A)\to \forall xB(\vec t_B))\in\mathcal S$.
+		\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec t_A, u))\to B^\#_\psi(\vec t_B))$\\for each $\forall \vec t(\exists xA_\psi(\vec t_A)\to B_\psi(\vec t_B))\in\mathcal S$
+		\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to A^\#_\psi(\vec t_B)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec t_B, u)))$\\for each $\forall \vec t(A_\psi(\vec t_A)\to \exists xB_\psi(\vec t_B))\in\mathcal S$
+	\end{itemize}
+	Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid iff
+	$$\varphi^\circ= E(\varphi)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
+	is classically valid.
+
+	We proceed now as in the previous section.
+Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec t((A_\psi(\vec t_A)\to B_\psi(\vec t_B))\to C_\psi(\vec t_C))$ with $\mathcal F_\to$ and of the form $\forall \vec t(\forall xA_\psi(\vec t_A)\to B_\psi(\vec t_B))$ with $\mathcal F_\forall$.
+Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
+For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$.
+As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
+	$$\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec t_A, w)\to B_\psi^\#(\vec t_B, w))\to C_\psi^\#(\vec t_C, u))$$ with
+		$$\forall \vec t\forall u(\vec E(\vec t, u)\to (A_\psi^\#(\vec t_A, f_\psi(\vec t, u))\to B_\psi^\#(\vec t_B, f_\psi(\vec t, u)))\to C_\psi^\#(\vec t_C, u))$$ and
+		$$\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec t_A, w)))\to B^\#(\vec t_B, u))$$ with $$\forall \vec t\forall u(\vec E(\vec t, u)\to (\forall x(E(x, f_\psi(\vec t, u))\to A_\psi^\#(\vec t_A, f_\psi(\vec t, u))))\to B^\#(\vec t_B, u))$$
+	and in each other formula in $\mathcal S^\circ$ remove the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S$. Then the previous formulas are equivalid to
+	$$\hspace*{-.5cm}\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f_\psi(\vec t, u))\:|\:\psi\in\mathcal F\}\to E(\varphi)\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
+	As in the previous section we can define a tree-counter model for this.
+	\begin{definition}
+		Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
+For $\psi = \forall\vec t\psi'\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff $(A_\psi^I(\vec t_A, u)\to B_\psi^I(\vec t_B, u))[\vec a/\vec t]$ is false or $C_\psi^I(\vec t_C, u)[\vec a/\vec t]$ is true.
+		For $\psi = \forall\vec t\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff there is $a\in M_u$ such that $A^{I[a/x]}_\psi(\vec t_A)[\vec a/\vec t]$ is false or $B_\psi^I(\vec t_B)[\vec a/\vec t]$ is true.
+For $\psi\in\mathcal F$ define $$g_{\vec a, \psi} : M\to M, u\mapsto\begin{cases}
+			u&\text{ if $\psi$ is fulfilled in $u$ with $\vec a$}\\
+			f^I_\psi(\vec a, u)&\text{ else}		
+		\end{cases}$$		
+		Define a $\mathcal M_T = (M_T, I_T)$ as follows:
+		\begin{itemize}
+			\item $M_T$ are the sequences on $\{ \vec t, \psi\:|\:\psi\in \mathcal F, \vec t\in M^n\}$ without repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\to$ and without immediate repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\forall$.
+			\item Interpret $\preceq$ as the prefix-order.
+			\item Have $$f_\psi^{I_T}(x_1\dots x_n) = \begin{cases}
+				x_1\dots x_n&\text{if $x=x_n$}\\
+				x_1\dots x_nx&\text{else}			
+			\end{cases}$$
+			\item For propositional variables $P$ have $$P^{I_T}(x_1\dots x_n) = P^I(g_{x_n}(\dots(g_{x_1}(s^I))\dots))$$
+		\end{itemize}
+	\end{definition}
+
 	\begin{lemma}
-		$\varphi$ and $\varphi^\to$ are intuitionistically equivalid.
+		Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
+		\begin{enumerate}
+			\item $\psi$ is fulfilled at $f_\psi^I(\vec a, u)$ with $\vec a$ for all $u\in M, \psi\in\mathcal F$, $\vec a\in M^n$.
+			\item If $\psi\in\mathcal F_\to$ is fulfilled at $u\in M$ with $\vec a$ then $\psi$ is fulfilled at all $w\geq u$ with $\vec a$.
+		\end{enumerate}	
 	\end{lemma}
-	The proof works exactly as in the propositional case.
-Similarly to before every formula in $\mathcal S$ is of one of the following forms $$A\to B, A\to B\to C, (A\to B)\to C, (A\to D)\to(B\to D)\to C\to D$$$$\forall xA\to B, A\to\forall xB, \exists xA\to B, A\to\exists xB$$
-
-	\subsection{Eskolemization}
-
-	Before being able to apply the procedure from the previous section we need to eliminate the quantifiers from $\varphi^\to$.
-The key idea for this stems from~\cite{baaz2006skolemization} and was expanded in~\cite{iemhoff2010eskolemization}.
-We will be able to use a condensed version.
-The basic principle is to introduce a new unary predicate $E(x)$ which encodes roughly that $x$ exists in a constructive sense.
-Note that besides adding $E$ to the signature step we need to add new function symbols, as is usual in skolemization.
-We then obtain $\mathcal S^E$ from $\mathcal S$ by including formulas as follows:
-
-	\begin{itemize}
-		\item every quantifier-free $\varphi\in\mathcal S$
-		\item $\bigwedge_i E a_i \to Ef(a_1,\dots a_n)$ for every $n$-ary function symbol $f$ and new free variables $a_1,\dots, a_n$.
-		\item $Ea$ for every free variable $a$ occurring in $\varphi^\to$.
-		\item $(Es\to A[s/x])\to B$ for each $\forall xA\to B\in\mathcal{S}$ where $s$ is a new constant symbol.
-		\item $A\to Ea\to B[a/x]$ for each $A\to \forall xB\in\mathcal S$  where $a$ is some new free variable.
-		\item $Ea\to A[a/x]\to B$ for each $\exists A\to B$ where $a$ is some new free variable..
-		\item $A\to B[s(\vec a)/x]$ and $Es(\vec a)$ for each $A\to \exists xB$ where $s$ is some new function symbol and $\vec a$ are the free variables occurring in $B$.
-	\end{itemize}
-	As usual we define $$\varphi^E = \bigwedge \mathcal S^E\to P_\varphi(\vec t)$$
-	and have 
-	\begin{lemma}
-		$\varphi^\to$ and $\varphi^E$ are equivalid.
-	\end{lemma}
+
 	\begin{proof}
-		As usual we do a translation of counter models.
-
-		Suppose we have a Kripke structure $\mathcal K = (W, (M_u)_{u\in W}, (I_u)_{u\in W})$ with world $u$ and variable assignment $v$ such that $u, v\models \bigwedge \mathcal S^\to$ and $u, v\not\models P_\varphi(\vec t)$.
-
-		First we will show that there must exists a linear substructure $\mathcal K^L = (W^L, (M^L_u)_{u\in W}, (I^L_u)_{u\in W})$ (i.e.
-$W^L$ is a total order) that is also a counter example.
-We will argue indirectly, i.e.
-suppose towards contradiction that every linear substructure is a model of $\varphi^\to$.
-In particular in every linear substructure we must have $u, v\not\models\mathcal\bigwedge S^\to$.
-
-		 Consider the following Kripke structure $\mathcal K^E = (W^E, (M^E_u)_{u\in W}, (I^E_u)_{u\in W})$:
-		\begin{itemize}
-			\item Take $W^E = W$. For clarity we will denote worlds in $W^E$ with $u^E$ instead of $u$.
-			\item Let the domain $M^E_u$ be constant across all worlds with $M^E_u = \underset{u\in U}{\bigcup} M_u$.
-			\item \vspace{-3mm}Interpret $E$ at $u$ as $M_u$, i.e. $E^{I_u^E}(t)$ iff $t\in M_u$.
-			\item Interpret each $n$-ary function symbol $f$ at $u$ via $$f^{I_u^E}(\vec t) = \begin{cases}f^{I_u}(\vec t) &\text{if $\vec t\in M_u^n$}\\m_u&\text{otherwise}\end{cases}\\$$ for with some arbitrary $m_u\in M^E_u\setminus M_u$.\footnote{Note that if $M^E_u\setminus M_u = \emptyset$ the second case does not occur in the first place.}
-			\item Interpret each relation symbol as in $\mathcal K$, i.e. $R^{I_u^E} = R^{I_u}$.
-		\end{itemize}
-		We claim that $u^E, v\not\models \varphi^E$.
-Clearly $u^E, v\not\models P_\varphi(\vec t)$.
-Furthermore it is easy to see that if $\psi\in\mathcal S^E$ was not newly introduced in the the above construction then $u^E, v\not\models \psi$ follows directly from $u, v\not\models\psi$.
-We then have $6$ remaining cases to consider.
-
-		1. $\psi$ is of the form $\bigwedge_i E a_i \to Ef(a_1,\dots a_n)$. Which is satisfied by the interpretation of functions and $E$ we gave.
-
-		2. $\psi$ is of the form $Ea$. Since we are using the same variable assignment this is also satisfied.
-
-		3.
-$\psi$ is of the form $(Es\to A[s/x])\to B$.
-Let $w^E\geq u^E$.
-If $w^E, v\models B$ we are done, so suppose otherwise.
-We know that $u, v\models \forall xA\to B$, i.e.
-if $w, v\not\models B$ then $w, v\not\models \forall xA$ so there exists $w' \geq w$ and $m_{w'}\in M_{w'}$ such that $m_{w'}\notin A^{I_{w'}}$.
-But then $m_{w'}\in E^{I_{w'^E}^E}$ and $m_{w'}\notin A^{I_{w'^E}^E}$, i.e.
-$w'^E, u\models Es$ and $w'^E, u\not\models As$
+		For $\psi\in\mathcal F_\to$ the proof is analogous to the previous section.
+
+		1.
+Let $\psi\in\mathcal F_\forall$.
+If $B^I_\psi(\vec t_B, u)[\vec a/\vec t]$ holds then we are done due to persistency.
+Otherwise there is $a$ with $E^I(a, f^I_\psi(\vec a, u))$ such that $A^{I[a/x]}_\psi(\vec t_A, f^I_\psi(\vec a, u))$ is false.
+So $\psi$ is fulfilled at $f^I_\psi(\vec a, u)$ with $\vec a$.
 	\end{proof}
 
-	\subsection{Classical Translation}
-
-	At this point we proceed similarly to the previous section.
-But instead of introducing new atoms $A^M$, we will encode the used hypotheses via terms, i.e.
-for each $n$-ary relation symbol $A$ we use a new $n+1$-ary relation symbol $A^\#$.
-Intuitively the additional argument ranges over worlds and encodes if $A$ is true at that world.
-The idea is analogous to the previous section in that this allows us to force certain hypotheses.
-We then obtain $\mathcal S^\#$ from $\mathcal S$ by including formulas as follows:
-	\begin{itemize}
-		\item $A^\#(\vec t, k)\to B^\#(\vec t, k)$ for some new free variable $k$ for every formula $A(\vec t)\to B(\vec t)\in\mathcal S$.
-		\item $A^\#(\vec t, k)\to B^\#(\vec t, k)\to C^\#(\vec t, k)$ for some new free variable $k$ for every $A(\vec t)\to B(\vec t)\to C(\vec t)\in\mathcal S$.
-		\item $B^\#(\vec t, f(k))\to C^\#(\vec t, k)$ and $A^\#(\vec t, f(l))$ where $f$ is some new function symbol and $k, l$ are new free variables for every $(A(\vec t)\to B(\vec t))\to C(\vec t)\in\mathcal S$.
-	\end{itemize}
-	Let $F_\to$ denote the set of function symbols $f$ introduced in this steps.
-	\begin{itemize}
-		\item $D^\#(\vec t, f(k))\to D^\#(\vec t, g(k))\to C^\#(t, k)\to D^\#(t, k)$, $A^\#(\vec t, f(l))$ and $B^\#(\vec t, g(m))$ where $f, g$ are new function symbols $k,l,m$ are new free variables for every $(A(\vec t)\to D(\vec t))\to (B(\vec t)\to D(\vec t))\to C(\vec t)\to D(\vec t)\in\mathcal S$.
-	\end{itemize}
-	Let $F'_\to$ denote the set of function symbols $f, g$ introduced in this steps.
-	\begin{itemize}
-		\item $A(\vec t, s(k), f(k))\to B(\vec t, k)$ and $E(s(l), f(l))$ where $k,l$ are new free variables $s, f$ are new function symbols for every $\forall xA(\vec t, x)\to B(\vec t)\in\mathcal S$.
-	\end{itemize}
-	Let $F_\forall$ denote the set of function symbols $f$ introduced in this step and $s_f$ the corresponding functions $s$.
-	\begin{itemize}
-		\item For every $f, g\in F_\to\cup F'_\to\cup F_\forall$ include $f(f(a)) = f(a)$, $g(f(a)) = f(g(a))$ and $A(\vec t, k)\to A(\vec t, f(k))$ for every relation symbol $A$ (including $E$).
-		\item $E(a, k)\to A^\#(\vec t, k)\to B^\#(\vec t, a, k)$ where $a, k$ are some new free variables for all $A(\vec t_1)\to \forall xB(\vec t_2, x)\in\mathcal S$.
-		\item $E(a, k)\to A^\#(\vec t, a, k)\to B^\#(\vec t, k)$ where $a, k$ are some new free variables for all $\exists xA(\vec t, x)\to B(\vec t)\in\mathcal S$.
-		\item $A^\#(\vec t, k)\to B^\#(\vec t, s(\vec t, k), k)$ and $A^\#(\vec t, k)\to E(s(\vec t, k), k)$ where $k$ is a new free variables and $s$ a new function symbol for all $A(\vec t)\to \exists xB(\vec t, x)\in\mathcal S$.
-	\end{itemize}
-		Let $S_\exists$ denote the set of function symbols $s$ introduced in this step.
-	\begin{itemize}
-		\item $\bot^\#(k)\to A^\#(\vec a, k)$ for each relation $A$ and new free variables $\vec a, k$.
-	\end{itemize}
-	Define $$\varphi^\# = \bigwedge S^\#\to P_\varphi^\#(\vec t, s_\varphi)$$ where $s_\varphi$ is some new constant (i.e. $0$-ary function symbol).
-	\begin{lemma}
-		$\varphi^\to$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
-	\end{lemma}
-
-	\begin{proof}
-		We proceed analogously to the previous chapter.
-
-		Suppose $\varphi^\to$ is not valid intuitionistically, i.e.
-there is a Kripke model $\mathcal K$ with world $u_0\in W$ and variable assignment $v$ such that $u_0, v\not\models P_\varphi(\vec t)$ and $u_0, v\models \psi(\vec t)$ for each $\psi(\vec t)\in\mathcal{S}$.
-For each world $u$ let $M_{u}$ be the domain at $u$ and $I_u$ the interpretation.
-		W.l.o.g.
-for $u_1, u_2\in W$ and $m_1\in M_{u_1}, m_2\in M_{u_2}$ have $m_1 = m_2$ only if there exists a world $w\in W$ that is comparable with both $u_1$ and $u_2$ and with $m_1 = m_2\in M_w$ (if this does not exists we can make them distinct by indexing $m$ with the set of comparable worlds it occurs at).
-
-		We then construct a classical counter model $\mathcal M$ to $\varphi^\#$ as follows:
-		\begin{itemize}
-			\item As a domain $M$ take the union of
-			\begin{itemize}
-				\item the union of domains $M_u$ at worlds $u\geq u_0$.
-				\item the set of worlds $u\in W$ with $u\geq u_0$.
-				\item an additional element $m_\bot$, that will be the result of applying a function to invalid arguments.
-			\end{itemize}
-			These three sets are also pairwise disjoint w.l.o.g.
-			\item As a variable assignment take $v$.
-			\item Define the interpretation $I$ as follows:
-			\begin{itemize}
-				\item $s_\varphi^I = u_0$.
-				\item For each function symbol $f\in F_\to$ recall the corresponding formula $(A(\vec t)\to B(\vec t))\to C(\vec t)$ with which it was introduced.
-Let $u\in W$ with $u\geq u_0$.
-We know that $u, v\models (A(\vec t)\to B(\vec t))\to C(\vec t)$.
-If $u, v\models C(\vec t)$ let $f^I(u) = u$.
-Otherwise choose $w\geq u$ such that $w, v\models A(\vec t)$ and $w, v\not\models B(\vec t)$ and let $f^I(u) = w$.
-For every other $m\in M$ let $f^I(m) = m_\bot$.
-				\item Analogously for $f\in F'_\to$ recall the corresponding formula $(A(\vec t)\to D(\vec t))\to (B(\vec t)\to D(\vec t))\to C(\vec t)\to D(\vec t)$.
-Let $u\in W$ with $u\geq u_0$.
-We know that $u, v\models (A(\vec t)\to D(\vec t))\to (B(\vec t)\to D(\vec t))\to C(\vec t)\to D(\vec t)$.
-If $u, v\models D(\vec t)$ let $f^I(u) = u$.
-Otherwise choose $w\geq u$ such that $w, v\not\models D(\vec t)$ and one of $w, v\not\models A(\vec t)$, $w, v\models B(\vec t)$, $w, v\models C(\vec t)$
-				and let $f^I(u) = w$. For every other $m\in M$ let $f^I(m) = m_\bot$.
-				\item Similarly for each $f\in F_\forall$ recall its formula $\forall xA(\vec t, x)\to B(\vec t)$.
-Let $u\in W$ with $u\geq u_0$.
-If $u, v\models B(\vec t)$ again have $f^I(u) = u$ and let $c_f^I(u)\in M_u$ be arbitrary.
-Otherwise choose a world $w\geq u$ and $m\in M_w$ such that $w, v[w/a]\not\models A(\vec t, a)$ and let $f^I(u) = w$ and $c_f^I(u) = m$.
-For every other $m\in M$ let $f^I(m) = c_f^I(m) = m_\bot$.
-				\item Finally for each $s\in S_\exists$ recall the corresponding formula $A(\vec t)\to \exists xB(\vec t, x)$.
-Let $u\in W$ with $u\geq u_0$.
-If $u, v\not\models A(\vec t)$ have $s^I(\vec t, u) = m_\bot$.
-Otherwise choose $m\in M_u$ such that $u, v[m/a]\models B(\vec t, a)$ and let $s^I(\vec t^{I_u, v}, u) = m$.
-For every other $m\in M$ let $s^I(m, u) = m_\bot$.
-				\item Every other $n$-ary function symbol $f$ is part of the original signature and we interpret it accordingly, i.e.
-if there exists a world $u$ with $\vec m\in M_u^n$ then $f^I(\vec m) = f^{I_u}(\vec m)$ and otherwise  $f^I(\vec m) = m_\bot$.
-Our distinctness condition at the start makes this unique and well-defined, i.e.
-there are no two worlds $u_1, u_2$ with $\vec m\in M_{u_1}^n\cap M_{u_2}^n$ but $ f^{I_{u_1}}(\vec m)\neq  f^{I_{u_2}}(\vec m)$.
-				\item We interpret $E$ via $(m, u)\in E^{I}$ iff $u\in W$, $m\in M_u$.
-				\item Finally $A$ is part of the original signature for every other $n+1$-ary relation symbol $A^\#$ and we interpret it accordingly, i.e.
-$(\vec m, u)\in A^I$ iff $u\in W$ and $\vec m\in A^{I_u}$.
-			\end{itemize}
-		\end{itemize}
-		We claim that $\mathcal M, v\not\models P^\#(\vec t, s_\varphi)$ and $\mathcal M, v\models \psi$ for every $\psi\in \mathcal S^\#$.
-The former is clear since $s_\varphi^I = u$ and $u, v\models P(\vec t)$.
-For the latter there are 12 cases.
-	\end{proof}
+	That the second point does not work for $\psi\in\mathcal F_\forall$ is why we must switch to immediate repetition.
+It does not work because even if for all $x$ with $E(x, w$) we have $A(x, w)$  there could be some term $t$ with $\neg E(t, w)$ and $v$ with $w\preceq v$ such that $E(t, v)$ and $\neg A(t, w)$.
+
+	\begin{example}
+		Consider the $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model
+		\begin{center}
+			\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
+					\node[squarednode]      (0)   [] {$ $};
+					\node[squarednode]      (1r)   [above right=of 0] {{\color{blue}$1$}};
+					\node[squarednode]      (1l)   [above left=of 0] {{\color{red}$1$}};
+					\node[squarednode]      (2r)   [above=of 1r] {{1, \color{blue}$2$}};
+					\node[squarednode]      (2l)   [above=of 1l] {{1, \color{red}$2$}};
+					\node[squarednode]      (3r)   [above=of 2r] {{1, 2, \color{blue}$3$}};
+					\node[squarednode]      (3l)   [above=of 2l] {{1, 2, \color{red}$3$}};
+					\node      (4r)   [above=of 3r] {{$\vdots$}};
+					\node      (4l)   [above=of 3l] {{$\vdots$}};
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (0.north east) -- (1r.south west);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (0.north west) -- (1l.south east);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1r.north) -- (2r.south);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1l.north) -- (2l.south);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1l.north east) -- (2r.south west);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1r.north west) -- (2l.south east);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2r.north) -- (3r.south);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2l.north) -- (3l.south);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2l.north east) -- (3r.south west);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2r.north west) -- (3l.south east);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3r.north) -- (4r.south);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3l.north) -- (4l.south);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3l.north east) -- (4r.south west);
+					\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3r.north west) -- (4l.south east);
+			\end{tikzpicture}
+		\end{center}
+		where at each node the elements that exist at it are indicated.
+For black ones both $A$ and $B$ hold, for red ones only $B$, for blue ones only $A$.
+Furthermore red arrows indicate $f_{\neg \forall xA(x)}$ and blue arrows $f_{\neg \forall xB(x)}$.
+Note that applying $f_{\neg \forall xA(x)}$ we must always reach the left column and $f_{\neg \forall xB(x)}$ we must always reach the right column.
+Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we can never remain stationary.
+This is in contrast to the propositional case.
+	\end{example}
+
+	\begin{corollary}
+		If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
+	\end{corollary}
+
+	Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$ is sufficient.
+In particular this allows us to completely eliminate $\preceq$ from the formula, that is we replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec t\forall u(E(x, u)\to E(x, f_\psi(\vec t, u)))\:|\:\psi\in\mathcal F_\to\}$ in $E(\varphi)$ to get $E^\#(\varphi)$ and obtain a formula
+
+	$$\varphi^{\#} = E^\#(\varphi)\to \bigwedge S^\#\to P^\#$$
+
+	\begin{theorem}
+		$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
+	\end{theorem}
+
+	\begin{example}
+		Consider the double negation shift (DNS)
+		$$\varphi = \forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$$
+		which is a famous non-intuitionistic tautology. We get
+		\begin{align*}
+			\mathcal S = & \{(P_{\forall x\neg\neg A(x)}\to P_{\neg\neg\forall xA(x)})\to P_\varphi, P_{\forall x\neg\neg A(x)}\to \forall xP_{\neg\neg A(x)}(x)\}\cup\\
+			             & \{\forall t((P_{\neg\neg A(x)}(t)\wedge P_{\neg A(x)}(t))\to \bot), \forall t((A(t)\to \bot)\to P_{\neg A(x)})(t)\}\cup\\
+			             & \{(P_{\neg\forall xA(x)}\to \bot)\to P_{\neg\neg\forall xA(x)}, (P_{\neg\forall xA(x)}\wedge P_{\forall xA(x)})\to \bot\}\cup\\
+			             &\{\forall xA(x)\to P_{\forall xA(x)}\}
+		\end{align*}
+		and thus, denoting $\psi_1 = (P_{\neg\forall xA(x)}\to \bot)\to P_{\neg\neg\forall xA(x)}$ and $\psi_2 = \forall xA(x)\to P_{\forall xA(x)}$, we have
+			\begin{align*}
+				\mathcal S^\# = & \{\forall u(P_{\forall x\neg\neg A(x)}(u)\to P_{\neg\neg\forall xA(x)}(u))\to P_\varphi(u)\}\cup\\& \{\forall u(P_{\forall x\neg\neg A(x)}(u)\to \forall xP_{\neg\neg A(x)}(x, u)\}\cup\\
+				& \{\forall t, u((P_{\neg\neg A(x)}(t, u)\wedge P_{\neg A(x)}(t, u))\to \bot)\}\\& \{\forall t, u((A(t, u)\to \bot)\to P_{\neg A(x)}(t, u))\}\cup\\
+				& \{\forall u((P_{\neg\forall xA(x)}(f_{\psi_1}(u))\to \bot)\to P_{\neg\neg\forall xA(x)}(u))\}\\& \{\forall u((P_{\neg\forall xA(x)}(u)\wedge P_{\forall xA(x)}(u))\to \bot)\}\cup\\
+				&\{\forall u(\forall xA(x, f_{\psi_2}(u))\to P_{\forall xA(x)}(u))\}
+			\end{align*}
+	\end{example}
 

### 2022-07-09 CSL/lipics-v2021-EIICL.tex
CSL Abstract, Intro, Overview

--- 2022-07-07 CLS/lipics-v2021-EIICL.tex
@@
 
-\Copyright{Jane Open Access and Joan R. Public} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+\Copyright{Alexander Pluska and Florian Zuleger} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
 
@@
 The famous double negation translation~\cite{glivenko1929quelques, godel1933intuitionistischen} establishes an embedding from classical into intuitionistic logic.
-Curiously the reverse direction has not been covered in literature, perhaps because going from the more expressive logic to the less expressive one necessitates a much more involved encoding.
-We present an embedding from intuitionistic to classical logic, both in the propositional and first-order case, as well as an embedding of intuitionistic propositional logic into quantified boolean formulas.
+Curiously the reverse direction has not been covered in literature.
+We present an effective embedding from intuitionistic to classical logic, both in the propositional and first-order case, as well as an effective embedding of intuitionistic propositional logic into quantified boolean formulas.
 Furthermore in the propositional case we establish a parameter that bounds the proof complexity of intuitionistic formulas.
-In general validity of intuitionistic propositional logic is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} but this gives us a finer-grained result.
 In this process we also present a Tseytin-like translation that preserves intuitionistic validity and reduces formulas to a normal form.
-Finally in the first order-case we hope that our translation allows the use state-of-the-art theorem provers for classical first-order logic such as Vampire~\cite{kovacs2013first} for checking intuitionistic validity and program synthesis.
+The key notion used in the proofs is the existence of counter-models - an analogon to the satisfiability of the negation in classical logic, which is not equivalent to invalidity in the intuitionistic case.
+We hope that this will allow leveraging the tremendous recent progress in automated reasoning in classical logic for checking intuitionistic validity, for which there has not been much progress in recent decades.
 \end{abstract}
@@
 
-Most state-of-the-art first-order theorem provers leverage classical logic~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}.
+Most state-of-the-art first-order theorem provers use classical logic~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}.
 Besides historic reasons a main advantage of classical logic over intuitionistic logic is the existence of convenient normal forms which enable reasoning techniques such as superposition and resolution.
-However there are also some drawbacks of this approach, in particular the non-constructiveness of classical logic poses a problem in synthesis-related tasks.
-On the other hand the Curry-Howard correspondence makes program extraction from intuitionistic proofs trivial~\cite{howard1980formulae} but automated deduction for intuitionistic logic is notoriously difficult.
+However there are also some drawbacks to this approach, in particular the non-constructiveness of classical logic poses a problem in synthesis-related tasks.
 
-There are different approaches for program extraction using classical provers.
-The first is simply to examine under which circumstances classical validity entails intuitionistic validity.
+Several approaches for program extraction from classical proofs exist in literature.
+The first is simply to examine under which circumstances classical validity entails intuitionistic validity and therefore constructiveness.
 Most famously this holds for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}.
@@
 
-We examine a new direction, i.e.
-for each formula $\varphi$ giving a translation $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
-We shall start with the simpler case of propositional formulas.
-The process will be divided into three steps.
-The first is a first-order encoding of intuitionistic Kripke semantics, which gives an obvious but unsatisfactory embedding.
-We then discuss how some of the quantifiers can be eliminated via Herbrandization and give a Tseytin-like transformation that exposes a parameter bounding the proof complexity of the propositional formula.
-This finally allows us to eliminate all remaining quantifiers and we end up with a propositional formula.
-Since validity of intuitionistic logic is \verb+PSPACE+-complete while for classical logic it only \verb+coNP+-complete as one could expect there is a blow-up which is exponential in the parameter that we previously isolated.
-Instead of targetting classical propositional logic at the last step we give an alternate translation to QBF, which is linear in the size of our complexity parameter.
-For predicate formulas the process is a bit more involved in that we also need to respect the constructive content of quantifiers.
-In particular the different semantics of the $\forall$-quantifier proves challenging.
-After dealing with this, we are able to utilize a quite similar process as in the propositional case.
+Intuitionistic logic~\cite{sep-logic-intuitionistic} is a non-classical logic in which the law of excluded middle ($A\vee \neg A$) is not a theorem.
+A key motivation for studying intuitionistic logic is the Curry-Howard correspondence~\cite{howard1980formulae} which allows an effective bi-interpretation of proofs and programs.
+
+One aspect that makes intuitionistic validity more difficult than classical validity is that it is not dual to satisfiability, i.e.
+unsatisfiability of $\neg\varphi$ is not sufficient for validity of $\varphi$.
+In fact intuitionistic satisfiability is equivalent to classical satisfiability and thus \verb+NP+-complete, while validity for intuitionistic logic is \verb+PSPACE+-complete~\cite{statman1979intuitionistic}.
+In classical logic this duality is the foundation of methods like resolution, tableaux~\cite{fitting2012first} and forms the backbone of most automated theorem provers.
+There is however a obvious dual notion to intuitionistic validity in the existence of Kripke counter-models, a standard semantics for intuitionistic logic~\cite{sep-logic-intuitionistic}.
+We will show that for each formula $\varphi$, if there exists a counter-model then there is one with fixed geometry completely determined by the structure of $\varphi$.
+This key observation will allow us to give a complete and effective embedding of intuitionistic logic into classical logic, i.e.
+for each formula $\varphi$ we give a formula $\varphi^\#$ that is classically valid if and only if $\varphi$ is intuitionistically valid, as well as an effective procedure to translate counter-models of the formulas.
+Intuitionistic validity can then be established using a classical prover by
+\begin{itemize}
+	\item performing our translation on the input formula $\varphi$ to obtain $\varphi^\#$
+	\item checking the classical validity of $\varphi^\#$ with the prover resulting in a proof or counter-model
+	\item translating the proof/counter-model of $\varphi^\#$ to a (intuitionistic) proof/counter-model of $\varphi$.
+\end{itemize}
+
+\section{Overview}
+
+We shall now give an overview of the main results and highlight the key arguments.
+Recall that our goal is to give an effective translation procedure that, for a given formula $\varphi$, yields a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
+We start with the predicate case.
+The key arguments are the same as in the first-order case while it is technically simpler and less cluttered and therefore more clear.
+
+Before performing the main transformation we propose a preprocessing to a simpler form.
+A popular pre-processing step in classical automated reasoning is the Tseytin transformation~\cite{tseitin1983complexity}.
+It gives an equisatisfiable sentence in conjunctive normal form.
+Eliminating all implications is not possible in intuitionistic logic since $A\to B$ is not equivalent to $\neg A\vee B$.
+Still we propose a similar transformation.
+A nice feature will be that the non-classical content will be encapsulated in formulas of type $(A\to B)\to C$, i.e.
+if there are no such formulas in the transformed formula then the classical validity of the original formula immediately implies intuitionistic validity.
+
+\begin{theorem}
+	For every propositional formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set of $\mathcal S$ of formulas $\psi$ of one of the forms
+	$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$for atomic $A_\psi, B_\psi, C_\psi$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
+The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
+\end{theorem}
+
+The main transformation then proceeds in three steps.
+
+First we encode as a first-order sentence that the considered formula holds in every Kripke frame.
+Since Kripke frames for propositional logic over a fixed alphabet of propostional variables form a first-order theory this is rather straightforward.
+
+Next we eliminate some quantifiers via Herbrandization. The only quantifiers eliminated occur in formulas derived from formulas of type $(A\to B)\to C$.
+
+This leaves us with a formula with only weak $\forall$-quantifiers.
+We can then argue that if there is a counter-model for the resulting formula then there is a counter-model with a certain Kripke frame completely determined by the number of formulas of type $(A\to B)\to C$.
+This allows us to eliminate the remaining quantifiers by simply enumerating the worlds in that Kripke model.
+
+All in all this leaves us with the following:
+
+\begin{theorem}
+	Let $\mathcal S$ be as above.
+Let $\mathcal F_\to\subseteq\mathcal S$ denote formulas $\psi$ of the form $(A_\psi\to B_\psi)\to C_\psi$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
+For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
+Obtain $\mathcal S^\#$ by including formulas as follows
+	\begin{itemize}
+		\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
+		\item $A_\psi^k\to (B_\psi^k\circ C_\psi^k)$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
+		\item $(A_\psi^k\circ B_\psi^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
+		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A_\psi\to B_\psi)\to C_\psi\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
+	\end{itemize}
+	Then $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
+The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$.
+Futhermore there is an effective procedure for translating counter-models between the sentences.
+\end{theorem}
+
+Examining the proof of the above theorem we will be able to notice that due to its tree form the constructed Kripke frame lends itself as an interpretation for a certain quantified boolean formula.
+We can show that this QBF is satisfiable if and only if there exists a counter-model for our original formula.
+
+\begin{theorem}
+	Let $\mathcal S$ be as above.
+There is an effective procedure that produces a QBF $\varphi^Q$ of size $\mathcal O(|\mathcal S|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$ with $2\cdot |\mathcal F_\to|-1$ quantifier alternations such that $\varphi$ is intuitionistically if and only if $\varphi^Q$ is not satisfiable.
+Fixing $N\in\mathbb N$ deciding intuitionistic validity for formulas $\bigwedge \mathcal S\to P$ where $\mathcal S$ is as above and $|\mathcal F_\to| = N$ is therefore in $\Sigma_{2N-1}$.
+\end{theorem}
+
+We then move to the first-order case.
+Here things are a bit more involved but the underlying approach is the same.
+First we again perform a Tseytin-like transformation.
+Here the non-classical content in encapsulated by formulas of type $(A\to B)\to C$ and $\forall xA\to B$.
+
+\begin{theorem}
+	For every predicate formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set $\mathcal S$ of formulas $\psi$ of the forms
+	$$\forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\wedge C_\psi(\vec c_\psi))), \forall \vec z_\psi((A_\psi(\vec a_\psi)\wedge B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi)),$$$$\forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\vee C_\psi(\vec c_\psi))),
+	\forall \vec z_\psi((A_\psi(\vec a_\psi)\vee B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi)),$$$$ \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi)),\forall \vec z_\psi(\forall xA(\vec a_\psi)\to B_\psi(\vec b_\psi)),$$$$ \forall \vec z_\psi(A_\psi(\vec a_\psi)\to\forall xB_\psi(b_\psi)), \forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi)), \forall \vec z_\psi(A_\psi(\vec a_\psi)\to\exists xB_\psi(b_\psi))$$for atomic $A_\psi, B_\psi, C_\psi$ and vectors of variables $\vec z_\psi, \vec a_\psi, \vec b_\psi, \vec c_\psi$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
+The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
+\end{theorem}
+
+We then proceed quite similarly to the propositional case.
+Note however that one difficulty arises from the fact that our (classical) domain will now contain on the one hand worlds in the Kripke frame but on the other hand also proper domain elements.
+We resolve this apparent conflict by introducing a special binary predicate $E$, inspired by~\cite{iemhoff2010eskolemization}, encoding which domain elements exists at which world.
+In particular $E(x, u)$ entails that $x$ represents a domain element and $u$ a world.
+
+The Herbrandization is completely analogous the the propositional case.
+
+The real difference arises in the final step: While we can also establish a universal Kripke Frame that is parametrized only by the formulas of type $(A\to B)\to C$ and $\forall xA\to B$, it is possibly countably infinite.
+This is not surprising since certain intuitionistically invalid formulas like the double negation shift $\forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$ don't have finite counter-models.
+Therefore we are not able to eliminate the weak $\forall$-quantifiers associated with the Kripke semantics.
+However since we started in first-order logic to begin with this is perhaps not a big deal.
+
+We end up with the following translation:
+
+\begin{theorem}
+	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$.
+Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
+For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$.
+For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$.
+Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
+Obtain $\mathcal S^\#$ by including formulas as follows:
+	\begin{itemize}
+		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\to (B_\psi^\#(\vec b_\psi, u)\circ C_\psi^\#(\vec c_\psi, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\circ C_\psi(\vec c_\psi)))\in\mathcal S$
+		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\circ B_\psi^\#(\vec b_\psi, u))\to C_\psi^\#(\vec c_\psi, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\circ B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
+		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to(A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u))\to B_\psi^\#(\vec b_\psi, f_\psi(\vec z_\psi, u)))\to C_\psi^\#(\vec c_\psi, u))$ for each $\psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
+		\item  $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \forall x(E(x, f_\psi(\vec z_\psi, u))\to A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u)))\to B^\#(\vec b_\psi, u))$\\for each $\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec t_B))\in\mathcal S$.
+		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec a_\psi, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec t(A_\psi(\vec a_\psi)\to \forall xB_\psi(\vec b_\psi))\in\mathcal S$.
+		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec z_\psi, u))\to B^\#_\psi(\vec b_\psi))$\\for each $\forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\in\mathcal S$
+		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec b_\psi)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec z_\psi(A_\psi(\vec a_\psi)\to \exists xB_\psi(\vec b_\psi))\in\mathcal S$
+	\end{itemize}
+	Then $\bigwedge S\to P$ is intuitionistically iff $\mathcal E\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol and
+	\begin{align*}
+		\mathcal E := &\exists xE(x, s)\wedge \forall x\forall y\forall z(\neg(E(x, z)\wedge E(z, y)))\wedge\\
+		& \bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
+		& \bigwedge\{\forall\vec z\forall u(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is an $n$-ary function symbol in original signature}\}
+	\end{align*}
+	The size of the obtained formula is linear in the input.
+However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
+\end{theorem}
+
+These results provide an effective framework for using classical provers to check intuitionistic validity. First experiments with the Vampire theorem prover show promise.
 

### 2022-07-09 CSL/lipics-v2021-EIICL.tex
typos

@@
 	Let $\mathcal S$ be as above.
-There is an effective procedure that produces a QBF $\varphi^Q$ of size $\mathcal O(|\mathcal S|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$ with $2\cdot |\mathcal F_\to|-1$ quantifier alternations such that $\varphi$ is intuitionistically if and only if $\varphi^Q$ is not satisfiable.
+There is an effective procedure that produces a QBF $\varphi^Q$ of size $\mathcal O(|\mathcal S|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$ with $2\cdot |\mathcal F_\to|-1$ quantifier alternations such that $\varphi$ is intuitionistically valid if and only if $\varphi^Q$ is not satisfiable.
 Fixing $N\in\mathbb N$ deciding intuitionistic validity for formulas $\bigwedge \mathcal S\to P$ where $\mathcal S$ is as above and $|\mathcal F_\to| = N$ is therefore in $\Sigma_{2N-1}$.
@@
 First we again perform a Tseytin-like transformation.
-Here the non-classical content in encapsulated by formulas of type $(A\to B)\to C$ and $\forall xA\to B$.
+Here the non-classical content is encapsulated by formulas of type $(A\to B)\to C$ and $\forall xA\to B$.
 

### 2022-07-09 CSL/lipics-v2021-EIICL.tex
modificaitons to abstract

@@
 
-\title{Embedding intuitionistic into classical logic} %TODO Please add
+\title{Embedding Intuitionistic into Classical Logic} %TODO Please add
 
@@
 Curiously the reverse direction has not been covered in literature.
-We present an effective embedding from intuitionistic to classical logic, both in the propositional and first-order case, as well as an effective embedding of intuitionistic propositional logic into quantified boolean formulas.
-Furthermore in the propositional case we establish a parameter that bounds the proof complexity of intuitionistic formulas.
-In this process we also present a Tseytin-like translation that preserves intuitionistic validity and reduces formulas to a normal form.
-The key notion used in the proofs is the existence of counter-models - an analogon to the satisfiability of the negation in classical logic, which is not equivalent to invalidity in the intuitionistic case.
+We present an effective embedding from intuitionistic into classical logic, both in the propositional and first-order case, as well as an effective embedding of intuitionistic propositional logic into quantified boolean formulas.
+Furthermore, in the propositional case we establish a parameter that bounds the proof complexity of intuitionistic formulas.
+As part of our construction, we also present a Tseytin-like translation that preserves intuitionistic validity and reduces formulas to a normal form.
+
+%The key notion used in the proofs is the existence of counter-models - an analogon to the satisfiability of the negation in classical logic, which is not equivalent to invalidity in the intuitionistic case.
 We hope that this will allow leveraging the tremendous recent progress in automated reasoning in classical logic for checking intuitionistic validity, for which there has not been much progress in recent decades.
+
+We hope that our embeddings of classical into intuitionistic logic will allow leveraging the tremendous progress in automated reasoning in classical logic for checking intuitionistic validity, for which there has not been much progress in recent decades.
+Crucially, our constructions support the transfer of counter-models to validity, which is important to model checking applications, where the output of counter-examples is an appreciated feature.
 \end{abstract}

### 2022-07-12 CSL/lipics-v2021-EIICL.tex
keywords
IQC to CQC, Conclusion

@@
 
-\keywords{Intuitionistic Logic, Automated Reasoning, Proof Theory} %TODO mandatory; please add comma-separated list of keywords
+\keywords{Intuitionistic Logic, Automated Reasoning, Propositional Logic, First-order Logic, QBF} %TODO mandatory; please add comma-separated list of keywords
 
@@
 
-\begin{theorem}
+\begin{theorem}\label{fullFOtranslation}
 	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$.
@@
 	\end{itemize}
-	Then $\bigwedge S\to P$ is intuitionistically iff $\mathcal E\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol and
+	Then $\bigwedge S\to P$ is intuitionistically iff $K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol and
 	\begin{align*}
-		\mathcal E := &\exists xE(x, s)\wedge \forall x\forall y\forall z(\neg(E(x, z)\wedge E(z, y)))\wedge\\
+		K := &\exists xE(x, s)\wedge \forall x\forall y\forall z(\neg(E(x, z)\wedge E(z, y)))\wedge\\
 		& \bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
@@
 
+\section{Preliminaries}
+
+Before proceeding towards the results let us quickly fix our notation and recapitulate the semantics for intuitionistic logic.
+
+\subsection{Propositional Semantics}
+
+\begin{definition}
+	A \textbf{valuation} $v$ is a function that assigns to each propositional variable $A$ a \textbf{truth value} $v(A)\in\{0, 1\}$. We extend $v$ to a function $v^*$ on formulas:
+	\begin{itemize}
+		\item $v^*(\bot) = 0$
+		\item $v^*(A) = v(A)$ for each propositional variable $A$.
+		\item $v^*(\varphi\wedge\psi) = \min(v^*(\varphi), v^*(\psi))$
+		\item $v^*(\varphi\vee\psi) = \max(v^*(\varphi), v^*(\psi))$
+		\item $v^*(\varphi\to \psi) = \max(1 - v^*(\varphi), v^*(\psi))$
+	\end{itemize}
+	$v$ is a \textbf{model} for $\varphi$ if $v(\varphi) = 1$. We denote the set of valid formulas with CPC.
+\end{definition}
+
+One notable property of classical logic is that a formula $\varphi$ is valid if and only if its negation, i.e.
+$\neg\varphi$ is not satisfiable.
+The same does not hold true for intuitionistic logic as we shall see.
+
+\begin{definition}
+	A \textbf{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of nodes an a family of valuations $(v_w)_{w\in W}$ such that for $u\leq w$ we have $v_u(A)\leq v_u(A)$ for every propositional variable $A$, this is called the \textbf{persistency condition}.
+We extend the valuations $v_u$ slightly differently to $v_u^\#$ than before, i.e.
+have
+	\begin{itemize}
+		\item $v_u^\#(\bot) = 0$
+		\item $v_u^\#(\varphi) = v_u(A)$ if $\varphi$ is atomic.
+		\item $v_u^\#(\varphi\wedge\psi) = \min(v_u^\#(\varphi), v_u^\#(\psi))$
+		\item $v_u^\#(\varphi\vee\psi) = \max(v_u^\#(\varphi), v_u^\#(\psi))$
+		\item $v_u^\#(\varphi\to \psi) = \min\{\max(1 - v_w^\#(\varphi), v_w^\#(\psi))\:|\:w\geq u\}$
+	\end{itemize}
+	That is $v_u^\#(\varphi\to \psi) = 1$ if at every world $w\geq u$ we have $v_w^\#(\varphi) = 0$ or $v_w^\#(\psi) = 1$.
+We say that $\varphi$ is satisfied at a node $u$ if $v_u^\#(\varphi) = 1$ and write $u\models\varphi$.
+If a formula is satisfied at every node then $\mathcal K$ is a \textbf{model} for $\varphi$.
+We denote the set of valid formulas with IPC.
+\end{definition}
+There are many classical tautologies which are not intuitionistically valid
+\begin{example}\label{LEMcounterexample}
+	Consider the famous law of excluded middle $A\vee\neg A$.
+Recall that $\neg A := A\to \bot$.
+As a counter-model consider a Kripke structure with two nodes $u\leq w$ and $v_u(A) = 0, v_w(A) = 1$.
+Then $u\not\models A$ and $u\not\models \neg A$, i.e.
+$u\not\models A\vee\neg A$.
+\end{example}
+In classical logic a proven strategy for showing that a formula valid is to consider its negation and show that it is not satisfiable.
+This is insufficient for intuitionistic logic, e.g.
+$A\vee\neg A$ is not a intuitionistic tautology but its negation is not satisfiable either.
+
+\subsection{Predicate Semantics}
+
+We now give account of the more involved semantics of first-order logic
+
+\begin{definition}
+	Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the domain of $\mathcal{M}$, and a function $I$ that assigns
+	\begin{itemize}
+		\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.
+		\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
+	\end{itemize}
+	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
+		m&\text{if $b=a$}\\
+		v(b)&\text{otherwise}
+	\end{cases}$$
+	Then terms are interpreted as follows:
+	\begin{itemize}
+		\item $a^{I, v} = v(a)$ for each free variable $a$.
+		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$.
+	\end{itemize}
+	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
+	\begin{itemize}
+		\item $\mathcal M, v\not\models\bot$.
+		\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$.
+		\item $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
+		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$.
+		\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
+		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
+		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
+		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
+	\end{itemize}
+	$\mathcal{M}$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$. We denote the set of valid formulas with CQC.
+\end{definition}
+
+\begin{definition}
+	A $\Sigma$-\textit{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$ and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$
+	\begin{itemize}
+		\item $M_u\subseteq M_w$.
+		\item $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
+		\item $R^{I_w}|_{M_u} = R^{I_u}$.
+	\end{itemize}
+	Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
+	\begin{itemize}
+		\item $u, v\not\models\bot$
+		\item $u, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I_u, v},\dots,t_n^{I_u, v})\in R^{I_u}$.
+		\item $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
+		\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$.
+		\item $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
+		\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
+		\item $u, v\models\exists x\varphi$ iff there exists $m\in M_u$ such that $u, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
+		\item $u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
+	\end{itemize}
+	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. We denote the set of valid formulas with IQC.
+\end{definition}
+In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically.
+Consider for instance the classical tautology $\varphi$ $$\neg\forall x A(x)\to \exists x \neg A(x)$$Analogous remarks with regard to validity and satisfiability hold as in the propositional case.
+
+\subsection{Skolemization and Herbrandization}
+An important step in the embedding will be the elimination of quantifiers via Herbrandization.
+In this process we introduce free variables and also additional function symbols to the signature.
+Whenever we speak of a new free variable what is meant is a free variable that does not occur in any of the considered formulas.
+Whenever a new function symbol is introduced we implicitly extend the signature by some not previously contained symbol.
+
+\begin{definition}
+	For formulas $\varphi$ we define the Skolemization $\varphi^S_T$ and Herbrandization $\varphi^H_T$ with respect to $T$ by simultaneous induction as follows:
+	\begin{itemize}
+		\item $A^S_T = A^H_T = A$ for each atomic $A$.
+		\item $(\varphi\circ\psi)^X_T = \varphi^X_T\circ\psi^X_T$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
+		\item $(\varphi\to\psi)^S_T = \varphi^H_T\to \psi^S_T$\\$(\varphi\to\psi)^H_T = \varphi^S_T\to\psi^H_T$.
+		\item $(\forall x\varphi)^S_T = \forall x(\varphi[a/x]^S_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_T = \varphi[s(t_1,\dots,t_n)/x]^H_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$.
+		\item $(\exists x\varphi)^S_T = \varphi[s(t_1,\dots,t_n)/x]^S_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$\\$(\exists x\varphi)^H_T = \exists x(\varphi[a/x]^H_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable.
+	\end{itemize}
+	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
+\end{definition}
+
+\begin{theorem}For every formula $\varphi$
+	\begin{itemize}
+		\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
+		\item $\varphi$ and $\varphi^H$ are classically equivalid.
+	\end{itemize}
+\end{theorem}
+
+which follows from Lemmata~\ref{ap1} and~\ref{ap2} in the appendix. We are now ready to present our embedding of intuitionistic into classical logic.
+
 \section{IPC to CQC}
@@
 \begin{definition}
-	Let $\varphi$ be a propositional formula. Define $\varphi^{k}$ inductively as follows:
-	\begin{itemize}
-		\item $A^{k} = A(k)$ for every propositional variable $A$.
-		\item $(\varphi\wedge\psi)^k = \varphi^k\wedge\psi^k$.
-		\item $(\varphi\vee\psi)^k = \varphi^k\vee\psi^k$.
-		\item $(\varphi\to \psi)^k = \forall u(k\preceq u\to\varphi^{u}\to\psi^{u})$ where $u$ is some new free variable.
-	\end{itemize}
-	Let $K(\varphi)$ encode that $\preceq$ is a partial order, the persistency condition for all propositional variables occurring in $\varphi$
+	Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively as follows:
+	\begin{itemize}
+		\item $A^{u} = A(u)$ for every propositional variable $A$.
+		\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
+		\item $(\varphi\to \psi)^u = \forall w(u\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new variable.
+	\end{itemize}
+	Let $K(\varphi)$ encode the Kripke semantics, i.e.
+	$$K(\varphi) = \text{PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{Persistent}(u, w))$$
+	with e.g.
 	\begin{align*}
-		K(\varphi) = &\forall k(k\preceq k)\wedge \forall k\forall l(k\preceq l\to l\preceq k\to l = k) \wedge \\&\forall k\forall l\forall m(k\preceq l \to l\preceq m\to k\preceq m)\wedge\\&\bigwedge \{\forall k\forall l(k\preceq l\to A(k)\to A(l))\:|\: \text{ $A$ occurs in $\varphi$}\}
+		\text{PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
+		Persistent(u, w)=&\:\bigwedge \{A(u)\to A(w))\:|\: \text{ $A$ is a propositional variable that occurs in $\varphi$}\}
 	\end{align*}
@@
 
-\section{Quantifiers in Classical Logic}
-
-Of course translating a propositional formula to a first-order formula is somewhat unsatisfying.
-As a first step to obtaining a propositional formula we wish to eliminate the quantifiers.
-Since we are interested in preserving validity we use Herbrandization.
-In this process we introduce free variables and also additional function symbols to the signature.
-Whenever we speak of a new free variable what is meant is a free variable that does not occur in any of the considered formulas.
-Whenever a new function symbol is introduced we implicitly extend the signature by some not previously contained symbol.
-
-Now usually in classical logic before removing quantifiers prenexification is performed, however since we wish to apply a similar procedure to intuitionistic logic later, where prenexification is not valid, we will do without.
-
-\begin{definition}
-	For formulas $\varphi$ we define the Skolemization $\varphi^S_T$ and Herbrandization $\varphi^H_T$ with respect to $T$ by simultaneous induction as follows:
-	\begin{itemize}
-		\item $A^S_T = A^H_T = A$ for each atomic $A$.
-		\item $(\varphi\circ\psi)^X_T = \varphi^X_T\circ\psi^X_T$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
-		\item $(\varphi\to\psi)^S_T = \varphi^H_T\to \psi^S_T$\\$(\varphi\to\psi)^H_T = \varphi^S_T\to\psi^H_T$.
-		\item $(\forall x\varphi)^S_T = \forall x(\varphi[a/x]^S_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_T = \varphi[s(t_1,\dots,t_n)/x]^H_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$.
-		\item $(\exists x\varphi)^S_T = \varphi[s(t_1,\dots,t_n)/x]^S_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$\\$(\exists x\varphi)^H_T = \exists x(\varphi[a/x]^H_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable.
-	\end{itemize}
-	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
-\end{definition}
-
-\begin{theorem}For every formula $\varphi$
-	\begin{itemize}
-		\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
-		\item $\varphi$ and $\varphi^H$ are classically equivalid.
-	\end{itemize}
-\end{theorem}
-
-which follows from the following two Lemmata, proofs of which can be found in the appendix.
-
-\begin{lemma}\label{ap1}
-	Let $\chi$ be a formula and $\{t_1\dots t_n\} = T$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
-	\begin{itemize}
-		\item if $\mathcal M, v\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
-		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
-	\end{itemize}
-\end{lemma}
-
-\begin{lemma}\label{ap2}
-	For every structure $\mathcal M$ and $\{t_1\dots t_n\} = T$ that contains all free variables in $\chi$ and variable assignment $v$
-	\begin{itemize}
-		\item if $\mathcal M, v\models\varphi^S_T$ then $\mathcal M, v\models \varphi$
-		\item if $\mathcal M, v\not\models\varphi^H_T$ then $\mathcal M, v\not\models\varphi$.
-	\end{itemize}
-\end{lemma}
-
 \section{Equivalid normalization}
 
-Now trying to apply Herbrandization to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated.
+Of course transforming a propositional formula into a predicate formula is somewhat unsatisfying.
+As a first step we want to apply Herbrandization.
+Now trying to apply it to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated.
 We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form.
@@
 Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
-$$\hspace{-0.8cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$
+$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$
 
@@
 \begin{definition}
-	Suppose $\mathcal S$ is a set of formulas of the form
-	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C),(A\vee B)\to C, (A\to B)\to C$$
+	Suppose $\mathcal S$ is a set of formulas $\psi$ of the form
+	$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi),(A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$
 	Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$.
@@
 	\begin{itemize}
-		\item $A^k\to A^{k\psi}$ for $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
-		\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
+		\item $A^k\to A^{k\psi}$ for each $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
+		\item $A^k_\psi\to (B^k_\psi\circ C^k_\psi)$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
+		\item $(A^k_\psi\circ B^k_\psi)\to C^k_\psi$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
+		\item $(A^{k\psi}_\psi\to B^{k\psi}_\psi)\to C^k_\psi$ for $(A_\psi\to B_\psi)\to C_\psi\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
 	\end{itemize}
@@
 However giving an explicit translation from IPC to QBF can still be useful as good solvers exists for QBF but for IPC not so much.
-Furthermore we will be able to give fragments of IPC for which validity is complete for other stages of the polynomial hierarchy.
 The translation from the last section will serve as a useful tool for this.
@@
 
-Fix some set $\mathcal S$ of formulas $\psi$ of the form $$\hspace*{-.7cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$and a formula $$\varphi = \bigwedge \mathcal S\to P$$
+Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$and$$\varphi = \bigwedge \mathcal S\to P$$
 and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$.
@@
 	\begin{itemize}
-		\item valid$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
+		\item {\normalfont Valid}$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
 if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
-		\item per($n$) which encodes that the persistency condition holds.
-		\item sat$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
-		\item sat$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
-	\end{itemize}
-	Have $k = |\mathcal F_\to|$ and define
-	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{per}(k)\wedge \text{sat}_{\mathcal S}(k)\wedge \text{sat}_{\mathcal F_\to}(k)\right)$$
-	for $0 < i < k$
-	$$\hspace*{-.8cm}\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{per}(i)\wedge \text{sat}_{\mathcal S}(i)\wedge \text{sat}_{\mathcal F_\to}(i)\wedge\left(\text{valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
-	and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{sat}_{\mathcal S}(0)\wedge(\text{valid}(0)\to \varphi_1^Q)\right)$$
+		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
+		\item {\normalfont Sat}$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
+		\item {\normalfont Sat}$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
+	\end{itemize}
+	Have $k = |\mathcal F_\to|$ and define$$\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal S}(i)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
+	for $0 < i < k$ as well as the special cases
+	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
+	for leafs and
+	and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
 	Example encodings of the above formulas would be
 	\begin{align*}
-		\text{valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\wedge\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
-		\text{per}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
-		\text{sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
+		\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\wedge\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
+		\text{\normalfont Persistent}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
+		\text{\normalfont Sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
 		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\to C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
-		\text{sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
+		\text{\normalfont Sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
 	\end{align*}
@@
 
-As a final result we would like to prove that the above problems are complete for their respective complexity classes.
-It is sufficient to provide a polynomial reduction from QBF with bounded number of quantifier alternation to them.
-Such translations exists~\cite{statman1979intuitionistic,vsvejdar2003polynomial}, however they are insufficient in that they generate more formulas of form $(A\to B)\to C$ than there are quantifier alternations.
-We therefore would need a slightly new translation.
-
 \section{CQC to IQC}
@@
 We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$.
-In addition we use a unary predicate $W$ that encodes that an element represents a world.
-
-First of all recall that due to~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all $\psi\in\mathcal S$ are of one of the forms
-
-$$\forall \vec t(A(\vec t_{A})\to (B(\vec t_B)\wedge C(\vec t_C))), \forall \vec t((A(\vec t_{A})\wedge B(\vec t_B))\to C(\vec t_C)),$$$$ \forall \vec t(A(\vec t_{A})\to (B(\vec t_B)\vee C(\vec t_C))),
-\forall \vec t((A(\vec t_{A})\vee B(\vec t_B))\to C(\vec t_C)),$$$$ \forall \vec t((A(\vec t_A)\to B(\vec t_B))\to C(\vec t_C)),\forall \vec t(\forall xA(\vec t_A)\to B(\vec t_B)),$$$$ \forall \vec t(A(\vec t_A)\to\forall xB(t_B)), \forall \vec t(\exists xA(\vec t_A)\to B(\vec t_B)), \forall \vec t(A(\vec t_A)\to\exists xB(t_B))$$
-First let us note how we axiomatize $E$ and $W$. For a vector $\vec x = x_1\dots x_n$ we define $\vec E(\vec x, u) = \bigwedge\{E(x_i, u)\:|\:i\leq n\}$.
+We fix some signature $\Sigma$ for this section.
+To encode that some $n$-ary relation $A$ holds at a world $u$ we replace each $n$-ary relation symbol $A$ with a $n+1$-ary relation symbol $A^\#$, interpreting $A^\#(\vec x, u)$ as "$A(\vec x)$ holds at u".
+Furthermore we write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$.
+
+First of all recall that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all $\psi\in\mathcal S$ are of one of the forms
+
+$$\forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\wedge C_\psi(\vec c_\psi))), \forall \vec z_\psi((A_\psi(\vec a_\psi)\wedge B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi)),$$$$ \forall \vec z_\psi(A_\psi(\vec a)\to (B_\psi(\vec b)\vee C_\psi(\vec c))),
+\forall \vec z_\psi((A_\psi(\vec a)\vee B_\psi(\vec b))\to C_\psi(\vec c)),$$$$ \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi)),\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi)),$$$$ \forall \vec z_\psi(A_\psi(\vec a_\psi)\to\forall xB_\psi(b_\psi)), \forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi)), \forall \vec z_\psi(A_\psi(\vec a_\psi)\to\exists xB_\psi(b_\psi))$$
+As in the propositional case we will now encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
+\begin{itemize}
+	\item PartialOrder($\preceq$) encoding that $\preceq$ is a partial order.
+	\item DomainSubset$(u, w)$ encoding that the domain at $u$ is a subset of that at $w$.
+	\item World($u$) encoding that $u$ represents a world.
+	\item DomainClosed($u$) which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
+	\item Persistent$(u, w)$ which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$ at $u$ we have $A(x)$ at $w$.
+\end{itemize}
+Then analogously to the predicate case define
 \begin{align*}
-	E(\varphi) = & \forall u\forall x(E(x, u)\to(\neg W(x)\wedge W(u)))\wedge\\
-	& \forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))\wedge\\
-	& \bigwedge\{\forall\vec t\forall u(\vec E(\vec t, u)\to E(f(\vec t), u))\:|\:\text{$f$ is an $n$-ary function symbol}\}
+	K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \\
+				 & \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
+				 & \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge\\
+				 & \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
 \end{align*}
-The one possible problem in this encoding is that we have a global interpretation of function symbols, i.e.
-in an intuitionistic model we could have worlds $u, v$ such that $f^{I_u}\neq f^{I_v}$.
-However this is not a problem: We can take as the domain of our classical model equivalence classes $\{[(x, u)]\:|\: x\in M_u\}$ under the relation $(x, u)\sim (y, v)$ iff $x = y$ and there exists $w$ comparable with $u, v$ such that $x\in w$ and interpret $f^I([x, u]) = [f^{I_u}(x), u]$.
-Then our classical translation only forces functions to be equal on arguments on which they must be equal anyway.
-So the Kripke model does in fact not afford us any additional generality.
-
-As before we need to axiomatize $\preceq$.
-For each $n$-ary relation symbol consider $P$ a new $n+1$ ary relation symbol $P^\#$.
-The only difference in that for persistence we may require additional quantification.
+Example encodings could be
 \begin{align*}
-	K(\varphi) = & \forall k(k\preceq k)\wedge \forall k\forall l((k\preceq l\wedge l\preceq k)\to l = k)\wedge\\
-	& \forall k\forall l\forall m((k\preceq l\wedge l\preceq m)\to k\preceq m)\wedge\\
-	& \bigwedge\{\forall k\forall l(k\preceq l\to \forall \vec x(A^\#(\vec x, k)\to A^\#(\vec x, l)))\:|\:\text{$A$ occurs in $\varphi$}\}
+	\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
+	\text{DomainSubset}(u, w) = &\:\forall z(E(z, u)\to E(z, w))\\
+	\text{World}(u) = &\:\exists xE(x, u)\\
+	\text{DomainClosed}(u) = &\:\bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f\in\Sigma$ is a function symbol}\}\\
+	\text{Pesistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
 \end{align*}
-Using $E$ and $W$ we obtain a set $\mathcal S^\circ$ by including formulas as follows:
-
-\begin{itemize}
-	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to (A_\psi^\#(t_A, u)\to (B_\psi^\#(\vec t_B, u)\circ C_\psi^\#(\vec t_C, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec t(A_\psi(\vec t_A)\to (B_\psi(\vec t_B)\circ C_\psi(\vec t_C)))\in\mathcal S$
-	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec t_A, w)\to B_\psi^\#(\vec t_B, w))\to C_\psi^\#(\vec t_C, u))$ for each $\psi = \forall \vec t((A_\psi(\vec t_A)\to B_\psi(\vec t_B))\to C_\psi(\vec t_C))\in\mathcal S$
-	\item  $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec t_A, w)))\to B^\#(\vec t_B, u))$ for each $\forall \vec t(\forall xA_\psi(\vec t_A)\to B_\psi(\vec t_B))\in\mathcal S$.
-	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to A^\#_\psi(\vec t_A, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec t_B, u)))$\\for each $\forall \vec t(A(\vec t_A)\to \forall xB(\vec t_B))\in\mathcal S$.
-	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec t_A, u))\to B^\#_\psi(\vec t_B))$\\for each $\forall \vec t(\exists xA_\psi(\vec t_A)\to B_\psi(\vec t_B))\in\mathcal S$
-	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to A^\#_\psi(\vec t_B)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec t_B, u)))$\\for each $\forall \vec t(A_\psi(\vec t_A)\to \exists xB_\psi(\vec t_B))\in\mathcal S$
-\end{itemize}
+
+Using $E$ we then obtain $\mathcal S^\circ$ analogously to the predicate case by including formulas:
+
+\begin{itemize}
+	\item $\forall \vec z_\psi\forall u(s\preceq u\to \vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\to (B_\psi^\#(\vec b_\psi, u)\circ C_\psi^\#(\vec c_\psi, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\circ C_\psi(\vec c_\psi)))\in\mathcal S$
+	\item $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\circ B_\psi^\#(\vec b_\psi, u))\to C_\psi^\#(\vec c_\psi, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\circ B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
+	\item $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to\forall w(u\preceq w\to A_\psi^\#(\vec a_\psi, w)\to B_\psi^\#(\vec b_\psi, w))\to C_\psi^\#(\vec c_\psi, u))$\\ for each $\psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
+	\item  $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec a_\psi, w)))\to B^\#(\vec b_\psi, u))$\\for each $\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec t_B))\in\mathcal S$.
+	\item $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec a_\psi, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec t(A_\psi(\vec a_\psi)\to \forall xB_\psi(\vec b_\psi))\in\mathcal S$.
+	\item $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec z_\psi, u))\to B^\#_\psi(\vec b_\psi))$\\for each $\forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\in\mathcal S$
+	\item $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec b_\psi)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec z_\psi(A_\psi(\vec a_\psi)\to \exists xB_\psi(\vec b_\psi))\in\mathcal S$
+\end{itemize}
+
 Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid iff
-$$\varphi^\circ= E(\varphi)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
-is classically valid.
+$$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
+is classically valid where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced, so let's present a proof.
+
+\begin{proof}
+	We proceed by translation of counter-examples.
+Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi^\circ$.
+As a Kripke frame $(W, \leq)$ take all $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ let $\leq$ be $\preceq^I$ restricted to $W$.
+Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A^\#(\vec x, u)$.
+It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
+
+	The other direction is a bit more involved.
+Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
+In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
+Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
+Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$ and extend $\sim$ by having $[w] = \{w\}$ for $w\in W_0$.
+Now have
+	\begin{itemize}
+		\item $s^I = w_0$.
+		\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
+		\item $f^I(m_1\dots m_n) =\begin{cases}
+			 f^{I_u}(x_1\dots x_n) &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$}\\
+			 w_0 & \text{otherwise}
+		\end{cases}$
+		\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
+			A^{I_u}(x_1\dots x_n) &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$}\\
+			\top & \text{otherwise}
+		\end{cases}$
+	\end{itemize}
+	One easily verifies that these are well-defined.
+Let us now briefly check that this indeed defines a counter-model.
+First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is.
+Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified.
+All that remains to show is $M, I\models\bigwedge\mathcal S^\circ$.
+Consider e.g.
+the case where $\psi\in\mathcal S$ is of the form
+	$$ \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))$$
+	We have to show that
+	$$M, I\models \forall \vec z_\psi\forall u(\vec E(t_\psi, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec a_\psi, w)\to B_\psi^\#(\vec b_\psi, w))\to C_\psi^\#(\vec c_\psi, u))$$
+	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec t_\psi\in M$ such that for all $w\in M$
+	$$\vec E^I_\psi(t_\psi, u)\to (u\preceq w\to {A^\#}^I_\psi(\vec a_\psi, w)\to {B^\#}^I_\psi(\vec b_\psi, w))\to {C^\#}^I_\psi(\vec c_\psi, u)$$ is false.
+For that ${C^\#}^I_\psi(\vec c_\psi, u)$ must be false and in particular $u\in W_0$.
+Futhermore $\vec E^I_\psi(z_\psi, u)$ must be true and we can write $\vec z_\psi = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$.
+Note that the above is false in particular for $w\in M_0$ with $u\preceq w$.
+But this would imply
+	$$((A_\psi^{I_u}(\vec a_\psi)\to B_\psi^{I_u}(\vec b_\psi))\to C_\psi^{I_u}(\vec c_\psi))[\vec z_\psi/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
+\end{proof}
 
 We proceed now as in the previous section.
-Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec t((A_\psi(\vec t_A)\to B_\psi(\vec t_B))\to C_\psi(\vec t_C))$ with $\mathcal F_\to$ and of the form $\forall \vec t(\forall xA_\psi(\vec t_A)\to B_\psi(\vec t_B))$ with $\mathcal F_\forall$.
+Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))$ with $\mathcal F_\to$ and of the form $\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))$ with $\mathcal F_\forall$.
 Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
@@
 As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
-$$\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec t_A, w)\to B_\psi^\#(\vec t_B, w))\to C_\psi^\#(\vec t_C, u))$$ with
-$$\forall \vec t\forall u(\vec E(\vec t, u)\to (A_\psi^\#(\vec t_A, f_\psi(\vec t, u))\to B_\psi^\#(\vec t_B, f_\psi(\vec t, u)))\to C_\psi^\#(\vec t_C, u))$$ and
-$$\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec t_A, w)))\to B^\#(\vec t_B, u))$$ with $$\forall \vec t\forall u(\vec E(\vec t, u)\to (\forall x(E(x, f_\psi(\vec t, u))\to A_\psi^\#(\vec t_A, f_\psi(\vec t, u))))\to B^\#(\vec t_B, u))$$
-and in each other formula in $\mathcal S^\circ$ remove the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S$. Then the previous formulas are equivalid to
-$$\hspace*{-.5cm}\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f_\psi(\vec t, u))\:|\:\psi\in\mathcal F\}\to E(\varphi)\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
+$$\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec a_\psi, w)\to B_\psi^\#(\vec b_\psi, w))\to C_\psi^\#(\vec c_\psi, u))$$ with
+$$\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, f_\psi(\vec t, u))\to B_\psi^\#(\vec b_\psi, f_\psi(\vec z_\psi, u)))\to C_\psi^\#(\vec c_\psi, u))$$ and
+$$\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec a_\psi, w)))\to B^\#(\vec b_\psi, u))$$ with $$\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (\forall x(E(x, f_\psi(\vec z_\psi, u))\to A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u))))\to B^\#(\vec b_\psi, u))$$
+and in each other formula in $\mathcal S^\circ$ remove the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
+$$\hspace*{-.5cm}\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f_\psi(\vec t, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
 As in the previous section we can define a tree-counter model for this.
@@
 
-That the second point does not work for $\psi\in\mathcal F_\forall$ is why we must switch to immediate repetition.
-It does not work because even if for all $x$ with $E(x, w$) we have $A(x, w)$  there could be some term $t$ with $\neg E(t, w)$ and $v$ with $w\preceq v$ such that $E(t, v)$ and $\neg A(t, w)$.
+Part 2 of this Lemma does not hold for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, w$) we have $A(x, w)$ there could be some term $t$ with $\neg E(t, w)$ and $v$ with $w\preceq v$ such that $E(t, v)$ and $\neg A(t, w)$.
+This is why we must consider sequences that have repetitions of such formulas:
 
@@
 
-Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$ is sufficient.
-In particular this allows us to completely eliminate $\preceq$ from the formula, that is we replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec t\forall u(E(x, u)\to E(x, f_\psi(\vec t, u)))\:|\:\psi\in\mathcal F_\to\}$ in $E(\varphi)$ to get $E^\#(\varphi)$ and obtain a formula
-
-$$\varphi^{\#} = E^\#(\varphi)\to \bigwedge S^\#\to P^\#$$
-
+Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$$ is sufficient.
+In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PO-encoding from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
+Define
+
+$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#$$
+then we have 
 \begin{theorem}
@@
 \end{theorem}
+The complete translation is summarised in Theorem~\ref{fullFOtranslation}.
 
@@
 
+\section{Conclusion}
+
+We have presented a complete embedding of intuitionistic into classical logic.
+The transformation saw an exponential blow-up parametrized by $|\mathcal F_\to|$ in the propositional case and a complexity increase reflected by arity-increase of all relations and introduction of new function symbols in the predicate case.
+A key motivation for our work is the possibility of leveraging classical provers for showing intuitionistic validity.
+The practical feasibility of our translation is yet to be tested.
+In particular in the first-order case it is difficult to estimate how the increased complexity of the output formula impacts proof search.
+Initial tests show promise but we plan on conducting a more thorough evaluation.
+A key factor that is not yet clear is how large the parameters $|\mathcal F_\to|$ in the propositional case and $|\mathcal F|$ in the predicate case respectively are in practice.
+
+The complexity increase fell directly out of our very straightforward construction.
+We plan to establish better bounds in future work by more intelligently utilizing structural properties of the formula, in particular by using that certain atoms are independent from each other it is possible to construct a smaller counter-example $\mathcal M_T$ in many cases.
+Besides improving the practical feasibility of our approach we hope that this will allow us to also give a new translation from QBF to IPC respecting complexity considerations and shine more light on the link between intuitionistic logic and the polynomial hierarchy.
+
+We hope that this work will provide a foundation for a new practical approach to automated reasoning for intuitionistic logic, a problem that has not seen much progress in recent times.
+
 \bibliography{lipics-v2021-EIICL}
@@
 
-\begin{proof}[Lemma \ref{ap1}]
+\section{Omitted Proofs}
+
+\begin{lemma}\label{ap1}
+	Let $\chi$ be a formula and $\{t_1\dots t_n\} = T$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
+	\begin{itemize}
+		\item if $\mathcal M, v\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
+		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
+	\end{itemize}
+\end{lemma}
+
+\begin{proof}
 	We proceed by simultaneous induction on the formula height.
@@
 
-\begin{proof}[Lemma \ref{ap2}]
+\begin{lemma}\label{ap2}
+	For every structure $\mathcal M$ and $\{t_1\dots t_n\} = T$ that contains all free variables in $\chi$ and variable assignment $v$
+	\begin{itemize}
+		\item if $\mathcal M, v\models\varphi^S_T$ then $\mathcal M, v\models \varphi$
+		\item if $\mathcal M, v\not\models\varphi^H_T$ then $\mathcal M, v\not\models\varphi$.
+	\end{itemize}
+\end{lemma}
+
+\begin{proof}
 	Again we proceed by simultaneous induction on the formula height.

### 2022-07-12 CSL/lipics-v2021-EIICL.tex
typos

@@
 	\end{itemize}
-	Then $\bigwedge S\to P$ is intuitionistically iff $K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol and
+	Then $\bigwedge S\to P$ is intuitionistically iff $\exists xE(x, s)\to K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol, $\Sigma$ is the original signature and
 	\begin{align*}
-		K := &\exists xE(x, s)\wedge \forall x\forall y\forall z(\neg(E(x, z)\wedge E(z, y)))\wedge\\
-		& \bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
-		& \bigwedge\{\forall\vec z\forall u(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is an $n$-ary function symbol in original signature}\}
+		K := &\:\bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
+		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
+		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}
 	\end{align*}

### 2022-07-12 CSL/lipics-v2021-EIICL.tex
Fixed minor errors, removed subsripts

@@
 \begin{theorem}
-	For every propositional formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set of $\mathcal S$ of formulas $\psi$ of one of the forms
-	$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$for atomic $A_\psi, B_\psi, C_\psi$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
+	For every propositional formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set of $\mathcal S$ of formulas  of one of the forms
+	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$for atomic $A, B, C$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
 The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
@@
 		\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
-		\item $A_\psi^k\to (B_\psi^k\circ C_\psi^k)$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A_\psi^k\circ B_\psi^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A_\psi\to B_\psi)\to C_\psi\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
+		\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
 	\end{itemize}
@@
 \begin{theorem}
-	For every predicate formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set $\mathcal S$ of formulas $\psi$ of the forms
-	$$\forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\wedge C_\psi(\vec c_\psi))), \forall \vec z_\psi((A_\psi(\vec a_\psi)\wedge B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi)),$$$$\forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\vee C_\psi(\vec c_\psi))),
-	\forall \vec z_\psi((A_\psi(\vec a_\psi)\vee B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi)),$$$$ \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi)),\forall \vec z_\psi(\forall xA(\vec a_\psi)\to B_\psi(\vec b_\psi)),$$$$ \forall \vec z_\psi(A_\psi(\vec a_\psi)\to\forall xB_\psi(b_\psi)), \forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi)), \forall \vec z_\psi(A_\psi(\vec a_\psi)\to\exists xB_\psi(b_\psi))$$for atomic $A_\psi, B_\psi, C_\psi$ and vectors of variables $\vec z_\psi, \vec a_\psi, \vec b_\psi, \vec c_\psi$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
+	For every predicate formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set $\mathcal S$ of formulas of the forms
+	$$\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),$$$$\forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),
+	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)),$$$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),$$$$ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))$$for atomic $A, B, C$ and vectors of variables $\vec z, \vec a, \vec b, \vec c$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
 The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
@@
 	\begin{itemize}
-		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\to (B_\psi^\#(\vec b_\psi, u)\circ C_\psi^\#(\vec c_\psi, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\circ C_\psi(\vec c_\psi)))\in\mathcal S$
-		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\circ B_\psi^\#(\vec b_\psi, u))\to C_\psi^\#(\vec c_\psi, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\circ B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
-		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to(A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u))\to B_\psi^\#(\vec b_\psi, f_\psi(\vec z_\psi, u)))\to C_\psi^\#(\vec c_\psi, u))$ for each $\psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
-		\item  $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \forall x(E(x, f_\psi(\vec z_\psi, u))\to A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u)))\to B^\#(\vec b_\psi, u))$\\for each $\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec t_B))\in\mathcal S$.
-		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec a_\psi, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec t(A_\psi(\vec a_\psi)\to \forall xB_\psi(\vec b_\psi))\in\mathcal S$.
-		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec z_\psi, u))\to B^\#_\psi(\vec b_\psi))$\\for each $\forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\in\mathcal S$
-		\item $\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec b_\psi)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec z_\psi(A_\psi(\vec a_\psi)\to \exists xB_\psi(\vec b_\psi))\in\mathcal S$
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
+		\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
 	\end{itemize}
@@
 \begin{definition}
-	For formulas $\varphi$ we define the Skolemization $\varphi^S_T$ and Herbrandization $\varphi^H_T$ with respect to $T$ by simultaneous induction as follows:
-	\begin{itemize}
-		\item $A^S_T = A^H_T = A$ for each atomic $A$.
-		\item $(\varphi\circ\psi)^X_T = \varphi^X_T\circ\psi^X_T$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
-		\item $(\varphi\to\psi)^S_T = \varphi^H_T\to \psi^S_T$\\$(\varphi\to\psi)^H_T = \varphi^S_T\to\psi^H_T$.
-		\item $(\forall x\varphi)^S_T = \forall x(\varphi[a/x]^S_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_T = \varphi[s(t_1,\dots,t_n)/x]^H_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$.
-		\item $(\exists x\varphi)^S_T = \varphi[s(t_1,\dots,t_n)/x]^S_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$\\$(\exists x\varphi)^H_T = \exists x(\varphi[a/x]^H_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable.
-	\end{itemize}
-	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
+	For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to $Z$ by simultaneous induction as follows:
+	\begin{itemize}
+		\item $A^S_T = A^H_Z = A$ for each atomic $A$.
+		\item $(\varphi\circ\psi)^X_Z = \varphi^X_Z\circ\psi^X_Z$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
+		\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$\\$(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
+		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
+		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.
+	\end{itemize}
+	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
 \end{definition}
@@
 		\item $A^{u} = A(u)$ for every propositional variable $A$.
+		\item $\bot^u = \bot$.
 		\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
@@
 	Let $K(\varphi)$ encode the Kripke semantics, i.e.
-	$$K(\varphi) = \text{PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{Persistent}(u, w))$$
+	$$K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))$$
 	with e.g.
 	\begin{align*}
-		\text{PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
-		Persistent(u, w)=&\:\bigwedge \{A(u)\to A(w))\:|\: \text{ $A$ is a propositional variable that occurs in $\varphi$}\}
+		\text{\normalfont PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
+		\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w))\:|\: \text{ $A$ is a propositional variable that occurs in $\varphi$}\}
 	\end{align*}
@@
 		&= K(\varphi)\to (\forall u'( u\preceq u' \to A^{u'}\to \bot^{u'})\vee A(u))\\
-		&= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot(u'))\vee A(u))
+		&= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot)\vee A(u))
 	\end{align*}
@@
 For each sub(semi)formula $\psi$ of $\varphi$ consider some new $n$-ary relation symbol $P_\psi$ where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$.
-We denote those variables with $\vec t_\psi$ in some arbitrary but fixed order.
+We denote those variables with $\vec z_\psi$ in some arbitrary but fixed order.
 We inductively define clause sets $\mathcal S^+(\varphi)$ and $\mathcal S^-(\varphi)$ as follows:
 	\begin{itemize}
-		\item For every atomic formula $A$\\$\mathcal S^+(A) = \{P_A(\vec t_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec t_A)\}$
-		\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$\\$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec t_{\varphi\circ\psi})\to P_{\varphi}(\vec t_\varphi)\circ P_{\psi}(\vec t_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec t_\varphi)\circ P_{\psi}(\vec t_\psi)\to P_{\varphi\circ\psi}(\vec t_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
-		\item For every implication $\varphi \to\psi$\\$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec t_{\varphi\to\psi})\wedge P_{\varphi}(\vec t_\varphi))\to P_{\psi}(\vec t_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec t_\varphi)\to P_{\psi}(\vec t_\psi))\to P_{\varphi\to\psi}(\vec t_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\varphi)$
-		\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$\\$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec t_{Qx\varphi})\to QxP_{\varphi}(\vec t_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec t_{\varphi})\to P_{Qx\varphi}(\vec t_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
+		\item For every atomic formula $A$\\$\mathcal S^+(A) = \{P_A(\vec z_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec z_A)\}$
+		\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$\\$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\to P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\to P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
+		\item For every implication $\varphi \to\psi$\\$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\wedge P_{\varphi}(\vec z_\varphi))\to P_{\psi}(\vec z_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec z_\varphi)\to P_{\psi}(\vec z_\psi))\to P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\psi)$
+		\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$\\$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec z_{Qx\varphi})\to QxP_{\varphi}(\vec z_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec z_{\varphi})\to P_{Qx\varphi}(\vec z_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
 	\end{itemize}
@@
 \begin{definition}
-	For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms $P_\psi$.
-By slight abuse of notation we denote with $\vec t_\psi$ elements of the domain instead of variables.
+	For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms of the form $P_\psi$.
+By slight abuse of notation we denote with $\vec z_\psi$ elements of the domain instead of variables.
 In the classical case have
-	$$P_{\varphi\wedge\psi}^I(\vec t_{\varphi\wedge\psi}) \Leftrightarrow P_{\varphi}^I(\vec t_\varphi)\wedge P_{\psi}^I(\vec t_\psi)\indent P_{\varphi\vee\psi}^I(\vec t_{\varphi\vee\psi}) \Leftrightarrow P_{\varphi}^I(\vec t_\varphi)\vee P_{\psi}^I((\vec t_\psi))$$$$ P_{\varphi\to\psi}^I(\vec t_{\varphi\to\psi}) \Leftrightarrow (\neg P_{\varphi}^I(\vec t_{\varphi}))\vee P_{\psi}^I(\vec t_{\psi})$$
-	$$P_{\forall x\varphi}^I(\vec t_{\forall x\varphi}) \Leftrightarrow \{\vec t\:|\:\:P_{\varphi}^I(\vec t_\varphi) \text{ for all $x\in M$}\}$$$$ P_{\exists x\varphi}^I(\vec t_{\exists x\varphi}) \Leftrightarrow \{\vec t\:|\:\:P_{\varphi}^I(\vec t_\varphi) \text{ for some $x\in M$}\}$$
+	$$P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)$$
+	and
+	$$P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))$$$$ P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})$$
+	$$P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}$$$$ P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\}$$
 	\\
-
 	and for intuitionistic logic and some world $u$
-	$$P_{\varphi\wedge\psi}^{I_u}(\vec t_{\varphi\wedge\psi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi)\wedge P_{\psi}^{I_u}(\vec t_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec t_{\varphi\vee\psi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi)\vee P_{\psi}^{I_u}((\vec t_\psi))$$$$ P_{\varphi\to\psi}^{I_u}(\vec t_{\varphi\to\psi}) \Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec t_{\varphi}))\vee P_{\psi}^{I_w}(\vec t_{\psi})\text{ for all $w\geq u$}$$
-	$$P_{\forall x\varphi}^{I_u}(\vec t_{\forall x\varphi}) \Leftrightarrow P_{\varphi}^{I_w}(\vec t_\varphi) \text{ for all $w\geq u$, $x\in M_w$}$$$$ P_{\exists x\varphi}^{I_u}(\vec t_{\exists x\varphi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi) \text{ for some $x\in M_u$}$$
+	$$P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)$$
+	and
+	$$P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))$$$$ P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}$$
+	$$P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}$$$$ P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}$$
 	\\
@@
 	For every formula $\varphi$ and structure $\mathcal M$
-	$$\mathcal S(\mathcal M, \varphi)\models\mathcal S(\varphi)$$
+	$$\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$$
 \end{lemma}
@@
 	\begin{itemize}
-		\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec t_\varphi)$ is.
-		\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec t_\varphi)$ is.
+		\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)$ is.
+		\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec z_\varphi)$ is.
 	\end{itemize}
@@
 
-First of all recall that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all $\psi\in\mathcal S$ are of one of the forms
-
-$$\forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\wedge C_\psi(\vec c_\psi))), \forall \vec z_\psi((A_\psi(\vec a_\psi)\wedge B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi)),$$$$ \forall \vec z_\psi(A_\psi(\vec a)\to (B_\psi(\vec b)\vee C_\psi(\vec c))),
-\forall \vec z_\psi((A_\psi(\vec a)\vee B_\psi(\vec b))\to C_\psi(\vec c)),$$$$ \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi)),\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi)),$$$$ \forall \vec z_\psi(A_\psi(\vec a_\psi)\to\forall xB_\psi(b_\psi)), \forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi)), \forall \vec z_\psi(A_\psi(\vec a_\psi)\to\exists xB_\psi(b_\psi))$$
+First of all recall that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
+
+$$\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),$$$$ \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),
+\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)),$$$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),$$$$ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))$$
 As in the propositional case we will now encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
@@
 \begin{itemize}
-	\item $\forall \vec z_\psi\forall u(s\preceq u\to \vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\to (B_\psi^\#(\vec b_\psi, u)\circ C_\psi^\#(\vec c_\psi, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi(A_\psi(\vec a_\psi)\to (B_\psi(\vec b_\psi)\circ C_\psi(\vec c_\psi)))\in\mathcal S$
-	\item $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, u)\circ B_\psi^\#(\vec b_\psi, u))\to C_\psi^\#(\vec c_\psi, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\circ B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
-	\item $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to\forall w(u\preceq w\to A_\psi^\#(\vec a_\psi, w)\to B_\psi^\#(\vec b_\psi, w))\to C_\psi^\#(\vec c_\psi, u))$\\ for each $\psi = \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))\in\mathcal S$
-	\item  $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec a_\psi, w)))\to B^\#(\vec b_\psi, u))$\\for each $\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec t_B))\in\mathcal S$.
-	\item $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec a_\psi, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec t(A_\psi(\vec a_\psi)\to \forall xB_\psi(\vec b_\psi))\in\mathcal S$.
-	\item $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec z_\psi, u))\to B^\#_\psi(\vec b_\psi))$\\for each $\forall \vec z_\psi(\exists xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\in\mathcal S$
-	\item $\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to A^\#_\psi(\vec b_\psi)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec b_\psi, u)))$\\for each $\forall \vec z_\psi(A_\psi(\vec a_\psi)\to \exists xB_\psi(\vec b_\psi))\in\mathcal S$
+	\item $\forall \vec z\forall u(s\preceq u\to \vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
+	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
+	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to\forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$\\ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
+	\item  $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
+	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
+	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
+	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
 \end{itemize}
@@
 the case where $\psi\in\mathcal S$ is of the form
-	$$ \forall \vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))$$
+	$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
 	We have to show that
-	$$M, I\models \forall \vec z_\psi\forall u(\vec E(t_\psi, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec a_\psi, w)\to B_\psi^\#(\vec b_\psi, w))\to C_\psi^\#(\vec c_\psi, u))$$
-	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec t_\psi\in M$ such that for all $w\in M$
-	$$\vec E^I_\psi(t_\psi, u)\to (u\preceq w\to {A^\#}^I_\psi(\vec a_\psi, w)\to {B^\#}^I_\psi(\vec b_\psi, w))\to {C^\#}^I_\psi(\vec c_\psi, u)$$ is false.
-For that ${C^\#}^I_\psi(\vec c_\psi, u)$ must be false and in particular $u\in W_0$.
-Futhermore $\vec E^I_\psi(z_\psi, u)$ must be true and we can write $\vec z_\psi = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$.
+	$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
+	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
+	$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false.
+For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$.
+Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$.
 Note that the above is false in particular for $w\in M_0$ with $u\preceq w$.
 But this would imply
-	$$((A_\psi^{I_u}(\vec a_\psi)\to B_\psi^{I_u}(\vec b_\psi))\to C_\psi^{I_u}(\vec c_\psi))[\vec z_\psi/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
+	$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
 \end{proof}
@@
 We proceed now as in the previous section.
-Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z_\psi((A_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))\to C_\psi(\vec c_\psi))$ with $\mathcal F_\to$ and of the form $\forall \vec z_\psi(\forall xA_\psi(\vec a_\psi)\to B_\psi(\vec b_\psi))$ with $\mathcal F_\forall$.
+Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$.
 Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
-For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$.
+For each  $\psi\in\mathcal F$ consider a new function symbol $f$.
 As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
-$$\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec a_\psi, w)\to B_\psi^\#(\vec b_\psi, w))\to C_\psi^\#(\vec c_\psi, u))$$ with
-$$\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (A_\psi^\#(\vec a_\psi, f_\psi(\vec t, u))\to B_\psi^\#(\vec b_\psi, f_\psi(\vec z_\psi, u)))\to C_\psi^\#(\vec c_\psi, u))$$ and
-$$\forall \vec z_\psi\forall u(s\preceq u\to\vec E(\vec z_\psi, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec a_\psi, w)))\to B^\#(\vec b_\psi, u))$$ with $$\forall \vec z_\psi\forall u(\vec E(\vec z_\psi, u)\to (\forall x(E(x, f_\psi(\vec z_\psi, u))\to A_\psi^\#(\vec a_\psi, f_\psi(\vec z_\psi, u))))\to B^\#(\vec b_\psi, u))$$
+$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$ with
+$$\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$$ and
+$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$$ with $$\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u))))\to B^\#(\vec b, u))$$
 and in each other formula in $\mathcal S^\circ$ remove the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
-$$\hspace*{-.5cm}\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f_\psi(\vec t, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
+$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
 As in the previous section we can define a tree-counter model for this.
@@
 Let $\psi\in\mathcal F_\forall$.
-If $B^I_\psi(\vec t_B, u)[\vec a/\vec t]$ holds then we are done due to persistency.
+If $B^I_\psi(\vec b, u)[\vec a/\vec t]$ holds then we are done due to persistency.
 Otherwise there is $a$ with $E^I(a, f^I_\psi(\vec a, u))$ such that $A^{I[a/x]}_\psi(\vec t_A, f^I_\psi(\vec a, u))$ is false.
@@
 
-Part 2 of this Lemma does not hold for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, w$) we have $A(x, w)$ there could be some term $t$ with $\neg E(t, w)$ and $v$ with $w\preceq v$ such that $E(t, v)$ and $\neg A(t, w)$.
+Part 2 of this Lemma does not hold for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$) we have $A(x, v)$ there could be some term $t$ with $\neg E(t, v)$ and $w$ with $v\preceq w$ such that $E(t, w)$ and $\neg A(t, w)$.
 This is why we must consider sequences that have repetitions of such formulas:

### 2022-07-13 Notes/Embedding intuitionistic into classical logic/EIICL.tex
Removed indices

@@
 
-	\subsection{Equivalid normalization}
-
-	Now trying to apply Herbrandization to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated.
+	\section{Equivalid normalization}
+
+	Of course transforming a propositional formula into a predicate formula is somewhat unsatisfying.
+As a first step we want to apply Herbrandization.
+Now trying to apply it to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated.
 We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form.
@@
 For each sub(semi)formula $\psi$ of $\varphi$ consider some new $n$-ary relation symbol $P_\psi$ where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$.
-We denote those variables with $\vec t_\psi$ in some arbitrary but fixed order.
+We denote those variables with $\vec z_\psi$ in some arbitrary but fixed order.
 We inductively define clause sets $\mathcal S^+(\varphi)$ and $\mathcal S^-(\varphi)$ as follows:
 		\begin{itemize}
-			\item For every atomic formula $A$\\$\mathcal S^+(A) = \{P_A(\vec t_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec t_A)\}$
-			\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$\\$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec t_{\varphi\circ\psi})\to P_{\varphi}(\vec t_\varphi)\circ P_{\psi}(\vec t_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec t_\varphi)\circ P_{\psi}(\vec t_\psi)\to P_{\varphi\circ\psi}(\vec t_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
-			\item For every implication $\varphi \to\psi$\\$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec t_{\varphi\to\psi})\wedge P_{\varphi}(\vec t_\varphi))\to P_{\psi}(\vec t_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec t_\varphi)\to P_{\psi}(\vec t_\psi))\to P_{\varphi\to\psi}(\vec t_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\varphi)$
-			\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$\\$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec t_{Qx\varphi})\to QxP_{\varphi}(\vec t_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec t_{\varphi})\to P_{Qx\varphi}(\vec t_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
+			\item For every atomic formula $A$\\$\mathcal S^+(A) = \{P_A(\vec z_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec z_A)\}$
+			\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$\\$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\to P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\to P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
+			\item For every implication $\varphi \to\psi$\\$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\wedge P_{\varphi}(\vec z_\varphi))\to P_{\psi}(\vec z_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec z_\varphi)\to P_{\psi}(\vec z_\psi))\to P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\psi)$
+			\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$\\$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec z_{Qx\varphi})\to QxP_{\varphi}(\vec z_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec z_{\varphi})\to P_{Qx\varphi}(\vec z_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
 		\end{itemize}
@@
 	\begin{definition}
-		For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms $P_\psi$.
-By slight abuse of notation we mean with $\vec t_\psi$ elements of the domain instead of variables.
+		For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms of the form $P_\psi$.
+By slight abuse of notation we denote with $\vec z_\psi$ elements of the domain instead of variables.
 In the classical case have
-		$$P_{\varphi\wedge\psi}^I(\vec t_{\varphi\wedge\psi}) \Leftrightarrow P_{\varphi}^I(\vec t_\varphi)\wedge P_{\psi}^I(\vec t_\psi)\indent P_{\varphi\vee\psi}^I(\vec t_{\varphi\vee\psi}) \Leftrightarrow P_{\varphi}^I(\vec t_\varphi)\vee P_{\psi}^I((\vec t_\psi))$$$$ P_{\varphi\to\psi}^I(\vec t_{\varphi\to\psi}) \Leftrightarrow (\neg P_{\varphi}^I(\vec t_{\varphi}))\vee P_{\psi}^I(\vec t_{\psi})$$
-		$$P_{\forall x\varphi}^I(\vec t_{\forall x\varphi}) \Leftrightarrow \{\vec t\:|\:\:P_{\varphi}^I(\vec t_\varphi) \text{ for all $x\in M$}\}$$$$ P_{\exists x\varphi}^I(\vec t_{\exists x\varphi}) \Leftrightarrow \{\vec t\:|\:\:P_{\varphi}^I(\vec t_\varphi) \text{ for some $x\in M$}\}$$
+		$$P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)$$
+		and
+		$$P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))$$$$ P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})$$
+		$$P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}$$$$ P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\}$$
 		\\
-
 		and for intuitionistic logic and some world $u$
-		$$P_{\varphi\wedge\psi}^{I_u}(\vec t_{\varphi\wedge\psi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi)\wedge P_{\psi}^{I_u}(\vec t_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec t_{\varphi\vee\psi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi)\vee P_{\psi}^{I_u}((\vec t_\psi))$$$$ P_{\varphi\to\psi}^{I_u}(\vec t_{\varphi\to\psi}) \Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec t_{\varphi}))\vee P_{\psi}^{I_w}(\vec t_{\psi})\text{ for all $w\geq u$}$$
-		$$P_{\forall x\varphi}^{I_u}(\vec t_{\forall x\varphi}) \Leftrightarrow P_{\varphi}^{I_w}(\vec t_\varphi) \text{ for all $w\geq u$, $x\in M_w$}$$$$ P_{\exists x\varphi}^{I_u}(\vec t_{\exists x\varphi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi) \text{ for some $x\in M_u$}$$
+		$$P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)$$
+		and
+		$$P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))$$$$ P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}$$
+		$$P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}$$$$ P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}$$
 		\\
@@
 		For every formula $\varphi$ and structure $\mathcal M$
-		$$\mathcal S(\mathcal M, \varphi)\models\mathcal S(\varphi)$$
+		$$\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$$
 	\end{lemma}
@@
 		\begin{itemize}
-			\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec t_\varphi)$ is. 
-			\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec t_\varphi)$ is.
+			\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)$ is.
+			\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec z_\varphi)$ is.
 		\end{itemize}
@@
 
-	\subsection{IPC to CPC}
+	\section{IPC to CPC}
 
@@
 	Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
-	$$\hspace{-0.8cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$
+	$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$
 
@@
 	\begin{definition}
-		Suppose $\mathcal S$ is a set of formulas of the form
-		$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C),(A\vee B)\to C, (A\to B)\to C$$
+		Suppose $\mathcal S$ is a set of formulas $\psi$ of the form
+		$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi),(A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$
 		Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$.
@@
 		\begin{itemize}
-			\item $A^k\to A^{k\psi}$ for $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
-			\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-			\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-			\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
+			\item $A^k\to A^{k\psi}$ for each $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
+			\item $A^k_\psi\to (B^k_\psi\circ C^k_\psi)$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
+			\item $(A^k_\psi\circ B^k_\psi)\to C^k_\psi$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
+			\item $(A^{k\psi}_\psi\to B^{k\psi}_\psi)\to C^k_\psi$ for $(A_\psi\to B_\psi)\to C_\psi\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
 		\end{itemize}
@@
 
-	\subsection{IPC to QBF}
+	\section{IPC to QBF}
 
@@
 
-	 Fix some set $\mathcal S$ of formulas $\psi$ of the form $$\hspace*{-.7cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$and a formula $$\varphi = \bigwedge \mathcal S\to P$$
+	Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$and$$\varphi = \bigwedge \mathcal S\to P$$
 	and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$.
@@
 		\begin{itemize}
-			\item valid$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
+			\item {\normalfont Valid}$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
 if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
-			\item per($n$) which encodes that the persistency condition holds.
-			\item sat$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
-			\item sat$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
-		\end{itemize}
-		Have $k = |\mathcal F_\to|$ and define
-		$$\varphi^Q_{k} = \exists \vec X^k\left(\text{per}(k)\wedge \text{sat}_{\mathcal S}(k)\wedge \text{sat}_{\mathcal F_\to}(k)\right)$$
-		for $0 < i < k$
-		$$\hspace*{-.8cm}\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{per}(i)\wedge \text{sat}_{\mathcal S}(i)\wedge \text{sat}_{\mathcal F_\to}(i)\wedge\left(\text{valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
-		and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{sat}_{\mathcal S}(0)\wedge(\text{valid}(0)\to \varphi_1^Q)\right)$$
+			\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
+			\item {\normalfont Sat}$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
+			\item {\normalfont Sat}$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
+		\end{itemize}
+		Have $k = |\mathcal F_\to|$ and define$$\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal S}(i)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
+		for $0 < i < k$ as well as the special cases
+		$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
+		for leafs and
+		and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
 		Example encodings of the above formulas would be
 		\begin{align*}
-			\text{valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\wedge\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
-			\text{per}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
-			\text{sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
+			\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\wedge\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
+			\text{\normalfont Persistent}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
+			\text{\normalfont Sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
 			&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\to C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
-			\text{sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
+			\text{\normalfont Sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
 		\end{align*}
@@
 
-	As a final result we would like to prove that the above problems are complete for their respective complexity classes.
-It is sufficient to provide a polynomial reduction from QBF with bounded number of quantifier alternation to them.
-Such translations exists~\cite{statman1979intuitionistic,vsvejdar2003polynomial}, however they are insufficient in that they generate more formulas of form $(A\to B)\to C$ than there are quantifier alternations.
-We therefore would need a slightly new translation.
-
-	\section{Predicate logic}
+	\section{CQC to IQC}
 
@@
 We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$.
-In addition we use a unary predicate $W$ that encodes that an element represents a world.
-
-	\subsection{CQC to IQC}
-
-	First of all recall that due to~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all $\psi\in\mathcal S$ are of one of the forms
-
-	$$\forall \vec t(A(\vec t_{A})\to (B(\vec t_B)\wedge C(\vec t_C))), \forall \vec t((A(\vec t_{A})\wedge B(\vec t_B))\to C(\vec t_C)),$$$$ \forall \vec t(A(\vec t_{A})\to (B(\vec t_B)\vee C(\vec t_C))),
-	\forall \vec t((A(\vec t_{A})\vee B(\vec t_B))\to C(\vec t_C)),$$$$ \forall \vec t((A(\vec t_A)\to B(\vec t_B))\to C(\vec t_C)),\forall \vec t(\forall xA(\vec t_A)\to B(\vec t_B)),$$$$ \forall \vec t(A(\vec t_A)\to\forall xB(t_B)), \forall \vec t(\exists xA(\vec t_A)\to B(\vec t_B)), \forall \vec t(A(\vec t_A)\to\exists xB(t_B))$$
-	First let us note how we axiomatize $E$ and $W$. For a vector $\vec x = x_1\dots x_n$ we define $\vec E(\vec x, u) = \bigwedge\{E(x_i, u)\:|\:i\leq n\}$.
+We fix some signature $\Sigma$ for this section.
+To encode that some $n$-ary relation $A$ holds at a world $u$ we replace each $n$-ary relation symbol $A$ with a $n+1$-ary relation symbol $A^\#$, interpreting $A^\#(\vec x, u)$ as "$A(\vec x)$ holds at u".
+Furthermore we write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$.
+
+	First of all recall that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
+
+	$$\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),$$$$ \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),
+	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)),$$$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),$$$$ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))$$
+	As in the propositional case we will now encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
+	\begin{itemize}
+		\item PartialOrder($\preceq$) encoding that $\preceq$ is a partial order.
+		\item DomainSubset$(u, w)$ encoding that the domain at $u$ is a subset of that at $w$.
+		\item World($u$) encoding that $u$ represents a world.
+		\item DomainClosed($u$) which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
+		\item Persistent$(u, w)$ which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$ at $u$ we have $A(x)$ at $w$.
+	\end{itemize}
+	Then analogously to the predicate case define
 	\begin{align*}
-		E(\varphi) = & \forall u\forall x(E(x, u)\to(\neg W(x)\wedge W(u)))\wedge\\
-		 & \forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))\wedge\\
-		 & \bigwedge\{\forall\vec t\forall u(\vec E(\vec t, u)\to E(f(\vec t), u))\:|\:\text{$f$ is an $n$-ary function symbol}\}
+		K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \\
+		& \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
+		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge\\
+		& \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
 	\end{align*}
-	The one possible problem in this encoding is that we have a global interpretation of function symbols, i.e.
-in an intuitionistic model we could have worlds $u, v$ such that $f^{I_u}\neq f^{I_v}$.
-However this is not a problem: We can take as the domain of our classical model equivalence classes $\{[(x, u)]\:|\: x\in M_u\}$ under the relation $(x, u)\sim (y, v)$ iff $x = y$ and there exists $w$ comparable with $u, v$ such that $x\in w$ and interpret $f^I([x, u]) = [f^{I_u}(x), u]$.
-Then our classical translation only forces functions to be equal on arguments on which they must be equal anyway.
-So the Kripke model does in fact not afford us any additional generality.
-
-	As before we need to axiomatize $\preceq$.
-For each $n$-ary relation symbol consider $P$ a new $n+1$ ary relation symbol $P^\#$.
-The only difference in that for persistence we may require additional quantification.
+	Example encodings could be
 	\begin{align*}
-		K(\varphi) = & \forall k(k\preceq k)\wedge \forall k\forall l((k\preceq l\wedge l\preceq k)\to l = k)\wedge\\
-		& \forall k\forall l\forall m((k\preceq l\wedge l\preceq m)\to k\preceq m)\wedge\\
-		& \bigwedge\{\forall k\forall l(k\preceq l\to \forall \vec x(A^\#(\vec x, k)\to A^\#(\vec x, l)))\:|\:\text{$A$ occurs in $\varphi$}\}
+		\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
+		\text{DomainSubset}(u, w) = &\:\forall z(E(z, u)\to E(z, w))\\
+		\text{World}(u) = &\:\exists xE(x, u)\\
+		\text{DomainClosed}(u) = &\:\bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f\in\Sigma$ is a function symbol}\}\\
+		\text{Pesistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
 	\end{align*}
-	Using $E$ and $W$ we obtain a set $\mathcal S^\circ$ by including formulas as follows:
-
-	\begin{itemize}
-		\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to (A_\psi^\#(t_A, u)\to (B_\psi^\#(\vec t_B, u)\circ C_\psi^\#(\vec t_C, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec t(A_\psi(\vec t_A)\to (B_\psi(\vec t_B)\circ C_\psi(\vec t_C)))\in\mathcal S$
-		\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec t_A, w)\to B_\psi^\#(\vec t_B, w))\to C_\psi^\#(\vec t_C, u))$ for each $\psi = \forall \vec t((A_\psi(\vec t_A)\to B_\psi(\vec t_B))\to C_\psi(\vec t_C))\in\mathcal S$
-		\item  $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec t_A, w)))\to B^\#(\vec t_B, u))$ for each $\forall \vec t(\forall xA_\psi(\vec t_A)\to B_\psi(\vec t_B))\in\mathcal S$.
-		\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to A^\#_\psi(\vec t_A, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec t_B, u)))$\\for each $\forall \vec t(A(\vec t_A)\to \forall xB(\vec t_B))\in\mathcal S$.
-		\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec t_A, u))\to B^\#_\psi(\vec t_B))$\\for each $\forall \vec t(\exists xA_\psi(\vec t_A)\to B_\psi(\vec t_B))\in\mathcal S$
-		\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to A^\#_\psi(\vec t_B)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec t_B, u)))$\\for each $\forall \vec t(A_\psi(\vec t_A)\to \exists xB_\psi(\vec t_B))\in\mathcal S$
-	\end{itemize}
+
+	Using $E$ we then obtain $\mathcal S^\circ$ analogously to the predicate case by including formulas:
+
+	\begin{itemize}
+		\item $\forall \vec z\forall u(s\preceq u\to \vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
+		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
+		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to\forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$\\ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
+		\item  $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
+		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
+		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
+		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
+	\end{itemize}
+
 	Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid iff
-	$$\varphi^\circ= E(\varphi)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
-	is classically valid.
+	$$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
+	is classically valid where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced, so let's present a proof.
+
+	\begin{proof}
+		We proceed by translation of counter-examples.
+Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi^\circ$.
+As a Kripke frame $(W, \leq)$ take all $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ let $\leq$ be $\preceq^I$ restricted to $W$.
+Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A^\#(\vec x, u)$.
+It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
+
+		The other direction is a bit more involved.
+Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
+In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
+Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
+Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$ and extend $\sim$ by having $[w] = \{w\}$ for $w\in W_0$.
+Now have
+		\begin{itemize}
+			\item $s^I = w_0$.
+			\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
+			\item $f^I(m_1\dots m_n) =\begin{cases}
+				f^{I_u}(x_1\dots x_n) &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$}\\
+				w_0 & \text{otherwise}
+			\end{cases}$
+			\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
+				A^{I_u}(x_1\dots x_n) &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$}\\
+				\top & \text{otherwise}
+			\end{cases}$
+		\end{itemize}
+		One easily verifies that these are well-defined.
+Let us now briefly check that this indeed defines a counter-model.
+First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is.
+Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified.
+All that remains to show is $M, I\models\bigwedge\mathcal S^\circ$.
+Consider e.g.
+the case where $\psi\in\mathcal S$ is of the form
+		$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
+		We have to show that
+		$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
+		holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
+		$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false.
+For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$.
+Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$.
+Note that the above is false in particular for $w\in M_0$ with $u\preceq w$.
+But this would imply
+		$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
+	\end{proof}
 
 	We proceed now as in the previous section.
-Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec t((A_\psi(\vec t_A)\to B_\psi(\vec t_B))\to C_\psi(\vec t_C))$ with $\mathcal F_\to$ and of the form $\forall \vec t(\forall xA_\psi(\vec t_A)\to B_\psi(\vec t_B))$ with $\mathcal F_\forall$.
+Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$.
 Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
-For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$.
+For each  $\psi\in\mathcal F$ consider a new function symbol $f$.
 As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
-	$$\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec t_A, w)\to B_\psi^\#(\vec t_B, w))\to C_\psi^\#(\vec t_C, u))$$ with
-		$$\forall \vec t\forall u(\vec E(\vec t, u)\to (A_\psi^\#(\vec t_A, f_\psi(\vec t, u))\to B_\psi^\#(\vec t_B, f_\psi(\vec t, u)))\to C_\psi^\#(\vec t_C, u))$$ and
-		$$\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec t_A, w)))\to B^\#(\vec t_B, u))$$ with $$\forall \vec t\forall u(\vec E(\vec t, u)\to (\forall x(E(x, f_\psi(\vec t, u))\to A_\psi^\#(\vec t_A, f_\psi(\vec t, u))))\to B^\#(\vec t_B, u))$$
-	and in each other formula in $\mathcal S^\circ$ remove the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S$. Then the previous formulas are equivalid to
-	$$\hspace*{-.5cm}\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f_\psi(\vec t, u))\:|\:\psi\in\mathcal F\}\to E(\varphi)\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
+	$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$ with
+	$$\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$$ and
+	$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$$ with $$\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u))))\to B^\#(\vec b, u))$$
+	and in each other formula in $\mathcal S^\circ$ remove the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
+	$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
 	As in the previous section we can define a tree-counter model for this.
@@
 
-	That the second point does not work for $\psi\in\mathcal F_\forall$ is why we must switch to immediate repetition.
-It does not work because even if for all $x$ with $E(x, w$) we have $A(x, w)$  there could be some term $t$ with $\neg E(t, w)$ and $v$ with $w\preceq v$ such that $E(t, v)$ and $\neg A(t, w)$.
+	Part 2 of this Lemma does not hold for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$) we have $A(x, v)$ there could be some term $t$ with $\neg E(t, v)$ and $w$ with $v\preceq w$ such that $E(t, w)$ and $\neg A(t, w)$.
+This is why we must consider sequences that have repetitions of such formulas:
 
@@
 
-	Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$ is sufficient.
-In particular this allows us to completely eliminate $\preceq$ from the formula, that is we replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec t\forall u(E(x, u)\to E(x, f_\psi(\vec t, u)))\:|\:\psi\in\mathcal F_\to\}$ in $E(\varphi)$ to get $E^\#(\varphi)$ and obtain a formula
-
-	$$\varphi^{\#} = E^\#(\varphi)\to \bigwedge S^\#\to P^\#$$
-
+	Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$$ is sufficient.
+In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PO-encoding from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
+Define
+
+	$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#$$
+	then we have 
 	\begin{theorem}
@@
 	\end{theorem}
+	The complete translation is summarised in Theorem~\ref{fullFOtranslation}.
 

### 2022-07-14 CSL/lipics-v2021-EIICL.tex
Overview

@@
 
-Most state-of-the-art first-order theorem provers use classical logic~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}.
-Besides historic reasons a main advantage of classical logic over intuitionistic logic is the existence of convenient normal forms which enable reasoning techniques such as superposition and resolution.
-However there are also some drawbacks to this approach, in particular the non-constructiveness of classical logic poses a problem in synthesis-related tasks.
-
-Several approaches for program extraction from classical proofs exist in literature.
+Constructive mathematics refers to a flavour of mathematics in which existence of an object can only be established by explicit construction, as opposed to classical mathematics where existence can be shown implicitly, e.g.
+by assuming non-existence and deriving a contradiction.
+Besides the philosophical issues that one may take with classical mathematics, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation when looking from the position of computer science in that intuitionistic proofs correspond directly to computer programs, which is expressed in the Curry-Howard-Isomorphism~\cite{howard1980formulae}.
+The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the facts that the law of excluded middle $A\vee\neg A$ and double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
+
+However comparing benchmarks for classical and intuitionistic automated theorem proving, e.g.
+TPTP~\cite{casc} and ILTP~\cite{iltp}, we see that classical provers currently have a far better success rate.
+Furthermore all the most popular and used first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover} utilize classical logic.
+This difference can be partially explained by fundamental differences between the logics.
+First of all in the propositional case determining intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas the analogous decision problem for classical logic is only \verb+coNP+-complete~\cite{cook1971complexity}.
+A further advantage of classical logic is the existence of calculi that are particularly suitable for automation like superposition~\cite{bachmair2001resolution} which rely on the existence of convenient normal forms such as CNF and the duality between validity and satisfiability, i.e.
+to show validity they show unsatisfiability of the negation.
+While some kind of normal form would of course obtainable in the intuitionistic case, crucially the duality does not hold, e.g.
+$\neg(A\vee \neg A)$ is not satisfiable while $A\vee\neg A$ is not a theorem.
+Therefore most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient, tammet1996resolution} use the naive inverse method, i.e.
+directly constructing a cut-free proof, enhanced by search strategies such as focussing and polarization.
+This strategy generally leads to a much more complex search and is difficult to implement efficiently.
+Finally in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, this is particularly true for the propositional case, i.e.
+SAT-solvers.
+
+With this work we want to open the path for leveraging the progress that has been made in classical theorem proving for intuitionistic logic.
+To that end we give an embedding from intuitionistic logic into classical logic.
+That is we give an effective procedure that gives for each formula $\varphi$ we a modified formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
+Interestingly the reverse direction, the famous double negation translation, is a long established translation that goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case and G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
+In the propositional case it is particularly simple: $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
+Intuitively the translation collapses for each subformula $\psi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent.
+This gives us a first idea why the reverse direction is perhaps more difficult: We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e.
+if they both occur in $\varphi$ we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$.
+In particular this necessitates the introduction of new propositional variables, which already marks a big difference to the double negation translation.
+
+In addition to the translation of formulas we give an effective translation of counter-models, i.e.
+for each intuitionistic counter-model for $\varphi$ we give a classical counter-model for $\varphi^\#$ and vice versa.
+In fact the existence of counter-models is a key notion.
+For classical logic it just corresponds to the satisfiability of the negation.
+However for intuitionistic logic it forms a proper dual to validity, while satisfiability of the negation is not necessary for invalidity and would therefore not form a proper dual.
+Transforming and reducing counter-models is what ultimately enables our translation.
+
+Finally we should mention that approaches to obtain constructive content from classical proofs exist in literature, however they are insufficient in our view.
 The first is simply to examine under which circumstances classical validity entails intuitionistic validity and therefore constructiveness.
@@
 
-Intuitionistic logic~\cite{sep-logic-intuitionistic} is a non-classical logic in which the law of excluded middle ($A\vee \neg A$) is not a theorem.
-A key motivation for studying intuitionistic logic is the Curry-Howard correspondence~\cite{howard1980formulae} which allows an effective bi-interpretation of proofs and programs.
-
-One aspect that makes intuitionistic validity more difficult than classical validity is that it is not dual to satisfiability, i.e.
-unsatisfiability of $\neg\varphi$ is not sufficient for validity of $\varphi$.
-In fact intuitionistic satisfiability is equivalent to classical satisfiability and thus \verb+NP+-complete, while validity for intuitionistic logic is \verb+PSPACE+-complete~\cite{statman1979intuitionistic}.
-In classical logic this duality is the foundation of methods like resolution, tableaux~\cite{fitting2012first} and forms the backbone of most automated theorem provers.
-There is however a obvious dual notion to intuitionistic validity in the existence of Kripke counter-models, a standard semantics for intuitionistic logic~\cite{sep-logic-intuitionistic}.
-We will show that for each formula $\varphi$, if there exists a counter-model then there is one with fixed geometry completely determined by the structure of $\varphi$.
-This key observation will allow us to give a complete and effective embedding of intuitionistic logic into classical logic, i.e.
-for each formula $\varphi$ we give a formula $\varphi^\#$ that is classically valid if and only if $\varphi$ is intuitionistically valid, as well as an effective procedure to translate counter-models of the formulas.
-Intuitionistic validity can then be established using a classical prover by
-\begin{itemize}
-	\item performing our translation on the input formula $\varphi$ to obtain $\varphi^\#$
-	\item checking the classical validity of $\varphi^\#$ with the prover resulting in a proof or counter-model
-	\item translating the proof/counter-model of $\varphi^\#$ to a (intuitionistic) proof/counter-model of $\varphi$.
+We suggest a new approach for obtaining a constructive proof using a classical prover:
+\begin{itemize}
+	\item Performing our translation on the input formula $\varphi$ to obtain $\varphi^\#$.
+	\item Checking the classical validity of $\varphi^\#$ with the prover resulting in a proof or counter-model.
+	\item Translating the proof/counter-model of $\varphi^\#$ to a (intuitionistic) proof/counter-model of $\varphi$.
 \end{itemize}

### 2022-07-14 CSL/lipics-v2021-EIICL.tex
Typos

@@
 by assuming non-existence and deriving a contradiction.
-Besides the philosophical issues that one may take with classical mathematics, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation when looking from the position of computer science in that intuitionistic proofs correspond directly to computer programs, which is expressed in the Curry-Howard-Isomorphism~\cite{howard1980formulae}.
-The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the facts that the law of excluded middle $A\vee\neg A$ and double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
-
-However comparing benchmarks for classical and intuitionistic automated theorem proving, e.g.
+Besides the philosophical issues that one may take with classical mathematics, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that intuitionistic proofs correspond directly to computer programs, which is expressed in the Curry-Howard-Isomorphism~\cite{howard1980formulae}.
+The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the fact that the law of excluded middle $A\vee\neg A$ and the double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
+
+Comparing benchmarks for classical and intuitionistic automated theorem proving, e.g.
 TPTP~\cite{casc} and ILTP~\cite{iltp}, we see that classical provers currently have a far better success rate.
-Furthermore all the most popular and used first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover} utilize classical logic.
+Furthermore all the most popular first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover} utilize classical logic.
 This difference can be partially explained by fundamental differences between the logics.
-First of all in the propositional case determining intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas the analogous decision problem for classical logic is only \verb+coNP+-complete~\cite{cook1971complexity}.
+First of all determining intuitionistic validity is computationally harder, i.e.
+in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
 A further advantage of classical logic is the existence of calculi that are particularly suitable for automation like superposition~\cite{bachmair2001resolution} which rely on the existence of convenient normal forms such as CNF and the duality between validity and satisfiability, i.e.
 to show validity they show unsatisfiability of the negation.
-While some kind of normal form would of course obtainable in the intuitionistic case, crucially the duality does not hold, e.g.
-$\neg(A\vee \neg A)$ is not satisfiable while $A\vee\neg A$ is not a theorem.
+While some kind of normal form is of course be obtainable for intuitionistic logic, albeit more complex, crucially the duality does not hold, e.g.
+$\neg(A\vee \neg A)$ is not satisfiable while $A\vee\neg A$ is also not a theorem.
 Therefore most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient, tammet1996resolution} use the naive inverse method, i.e.
-directly constructing a cut-free proof, enhanced by search strategies such as focussing and polarization.
-This strategy generally leads to a much more complex search and is difficult to implement efficiently.
+directly constructing a cut-free proof from the theorem by applying the inverse of rules from some proof calculus, enhanced by search strategies such as focussing and polarization.
+This strategy generally leads to a much more complex search and is therefore difficult to implement efficiently.
 Finally in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, this is particularly true for the propositional case, i.e.
@@
 In fact the existence of counter-models is a key notion.
-For classical logic it just corresponds to the satisfiability of the negation.
-However for intuitionistic logic it forms a proper dual to validity, while satisfiability of the negation is not necessary for invalidity and would therefore not form a proper dual.
-Transforming and reducing counter-models is what ultimately enables our translation.
+For classical logic it just corresponds to the satisfiability of the negation but for intuitionistic logic it forms a proper dual to validity, while satisfiability of the negation is not necessary for invalidity and would therefore not form a proper dual.
+Transforming and reducing counter-models to a normal form is also what ultimately enables our translation.
 
@@
 
-We suggest a new approach for obtaining a constructive proof using a classical prover:
-\begin{itemize}
-	\item Performing our translation on the input formula $\varphi$ to obtain $\varphi^\#$.
-	\item Checking the classical validity of $\varphi^\#$ with the prover resulting in a proof or counter-model.
-	\item Translating the proof/counter-model of $\varphi^\#$ to a (intuitionistic) proof/counter-model of $\varphi$.
+Our work suggests a new approach for obtaining a constructive proof using a classical prover, that is
+\begin{itemize}
+	\item performing our translation on the input formula $\varphi$ to obtain $\varphi^\#$,
+	\item checking the classical validity of $\varphi^\#$ with the prover resulting in a proof or counter-model,
+	\item translating the proof/counter-model of $\varphi^\#$ to a (intuitionistic) proof/counter-model of $\varphi$.
 \end{itemize}

### 2022-07-14 CSL/lipics-v2021-EIICL.tex
Unified notation for prop sem. and pred sem. definition

@@
 directly constructing a cut-free proof from the theorem by applying the inverse of rules from some proof calculus, enhanced by search strategies such as focussing and polarization.
-This strategy generally leads to a much more complex search and is therefore difficult to implement efficiently.
+This approach generally leads to a much more complex search and is therefore difficult to implement efficiently.
 Finally in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, this is particularly true for the propositional case, i.e.
@@
 
-Before proceeding towards the results let us quickly fix our notation and recapitulate the semantics for intuitionistic logic.
+Before proceeding towards the results let us quickly fix our notation and recapitulate the semantics for classical and more importantly intuitionistic logic.
 
@@
 \begin{definition}
-	A \textbf{valuation} $v$ is a function that assigns to each propositional variable $A$ a \textbf{truth value} $v(A)\in\{0, 1\}$. We extend $v$ to a function $v^*$ on formulas:
-	\begin{itemize}
-		\item $v^*(\bot) = 0$
-		\item $v^*(A) = v(A)$ for each propositional variable $A$.
-		\item $v^*(\varphi\wedge\psi) = \min(v^*(\varphi), v^*(\psi))$
-		\item $v^*(\varphi\vee\psi) = \max(v^*(\varphi), v^*(\psi))$
-		\item $v^*(\varphi\to \psi) = \max(1 - v^*(\varphi), v^*(\psi))$
-	\end{itemize}
-	$v$ is a \textbf{model} for $\varphi$ if $v(\varphi) = 1$. We denote the set of valid formulas with CPC.
+	A \emph{valuation} $v$ is a function that assigns to each propositional variable $A$ a \textbf{truth value} $v(A)\in\{0, 1\}$.
+We inductively define the model relation $\models$ between $v$ and formulas:
+	\begin{itemize}
+		\item $v\not\models \bot$
+		\item $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
+		\item $v\models \varphi\wedge\psi$ iff $v\models\varphi$ and $v\models\psi$.
+		\item $v\models\varphi\vee\psi$ iff $v\models\varphi$ or $v\models\psi$.
+		\item $v\models\varphi\to \psi$ iff $v\not\models\varphi$ and $v\models\psi$.
+	\end{itemize}
+	$v$ is a \textbf{model} for $\varphi$ if $v\models\varphi$. If every valuation is a model for $\varphi$ then we say $\varphi$ is \emph{valid}. We denote the set of valid formulas with \emph{CPC}.
 \end{definition}
 
-One notable property of classical logic is that a formula $\varphi$ is valid if and only if its negation, i.e.
-$\neg\varphi$ is not satisfiable.
+One notable property of classical logic is that a formula $\varphi$ is valid if and only if its negation $\neg\varphi := \varphi\to\bot$ is not satisfiable, i.e.
+there does not exist a model for it.
 The same does not hold true for intuitionistic logic as we shall see.
@@
 \begin{definition}
-	A \textbf{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of nodes an a family of valuations $(v_w)_{w\in W}$ such that for $u\leq w$ we have $v_u(A)\leq v_u(A)$ for every propositional variable $A$, this is called the \textbf{persistency condition}.
-We extend the valuations $v_u$ slightly differently to $v_u^\#$ than before, i.e.
+	A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of nodes an a family of valuations $(v_w)_{w\in W}$ such that for $u\leq w$ we have $v_u(A)\leq v_u(A)$ for every propositional variable $A$, this is called the \emph{persistency condition}.
+The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, i.e.
 have
 	\begin{itemize}
-		\item $v_u^\#(\bot) = 0$
-		\item $v_u^\#(\varphi) = v_u(A)$ if $\varphi$ is atomic.
-		\item $v_u^\#(\varphi\wedge\psi) = \min(v_u^\#(\varphi), v_u^\#(\psi))$
-		\item $v_u^\#(\varphi\vee\psi) = \max(v_u^\#(\varphi), v_u^\#(\psi))$
-		\item $v_u^\#(\varphi\to \psi) = \min\{\max(1 - v_w^\#(\varphi), v_w^\#(\psi))\:|\:w\geq u\}$
-	\end{itemize}
-	That is $v_u^\#(\varphi\to \psi) = 1$ if at every world $w\geq u$ we have $v_w^\#(\varphi) = 0$ or $v_w^\#(\psi) = 1$.
-We say that $\varphi$ is satisfied at a node $u$ if $v_u^\#(\varphi) = 1$ and write $u\models\varphi$.
+		\item $v_u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
+	\end{itemize}
+	We say that $\varphi$ is satisfied at a node $u$ if $v_u\models\varphi$ and write $u\models\varphi$.
 If a formula is satisfied at every node then $\mathcal K$ is a \textbf{model} for $\varphi$.
+A formula is valid if every Kripke structure is a model of it.
 We denote the set of valid formulas with IPC.
@@
 \begin{definition}
-	A $\Sigma$-\textit{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$ and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$
+	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$ and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$
 	\begin{itemize}
@@
 	\end{center}
-	\vspace*{-.3cm}where at each node the true propositions are indicated.\pagebreak
+	\vspace*{-.3cm}where at each node the true propositions are indicated.
 

### 2022-07-14 CSL/lipics-v2021-EIICL.tex
Restructuring

@@
 
-We end up with the following translation:
-
-\begin{theorem}\label{fullFOtranslation}
-	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$.
-Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
-For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$.
-For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$.
-Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
-Obtain $\mathcal S^\#$ by including formulas as follows:
-	\begin{itemize}
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
-		\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
-	\end{itemize}
-	Then $\bigwedge S\to P$ is intuitionistically iff $\exists xE(x, s)\to K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol, $\Sigma$ is the original signature and
-	\begin{align*}
-		K := &\:\bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
-		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
-		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}
-	\end{align*}
-	The size of the obtained formula is linear in the input.
-However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
+\begin{theorem}
+	There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
+The size of $\varphi^\#$ is in linear in the size of $|\varphi|$, however for each $n$-ary relation symbol in $\varphi$ there is a corresponding $n+1$-ary relation symbol in $\varphi^\#$ and furthermore $\varphi$ contains a new binary predicate $E$ as well as a number of new function symbols.
+Furthermore there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
 \end{theorem}
+The complete translation is stated in Theorem~\ref{fullFOtranslation}.
 
@@
 		\item $v\models\varphi\vee\psi$ iff $v\models\varphi$ or $v\models\psi$.
-		\item $v\models\varphi\to \psi$ iff $v\not\models\varphi$ and $v\models\psi$.
+		\item $v\models\varphi\to \psi$ iff $v\not\models\varphi$ or $v\models\psi$.
 	\end{itemize}
@@
 
-\section{IPC to CPC}
-
-We are now ready to put everything together.
-Recall that our goal is for a formula to give a modified formula that is valid classically iff the original formula is valid intuitionistically.
+\section{IPC to simplified CQC}
+
+We are now ready to put these two reduction together with our embedding f.
+Let $\varphi$ be some propositional formula.
 As a first step we perform the normalization from the last section, i.e.
@@
 Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
-$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$
+$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$
 
@@
 where $u$ is some new free variable and each $\psi\in\mathcal S$ is of one of the forms
-$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\wedge (B_\psi(k))\to C_\psi(k)))$$
-$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\vee B_\psi(k))\to C_\psi(k))$$
-$$\forall k(u\preceq k\to (\forall l(k\preceq l \to A_\psi(l)\to B_\psi(l))\to C_\psi(k)))$$
+$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B(k)\wedge C(k))), \forall k(u\preceq k\to (A(k)\wedge (B(k))\to C(k)))$$
+$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B(k)\vee C(k))), \forall k(u\preceq k\to (A(k)\vee B(k))\to C(k))$$
+$$\forall k(u\preceq k\to (\forall l(k\preceq l \to A(l)\to B(l))\to C(k)))$$
 with each $P(x)$ possibly being $\bot$. Applying Herbrandization we obtain a formula
@@
 where $s$ is some new constant term and each $\psi\in\mathcal S'$ is of one of the forms
-$$\hspace*{-2mm}\forall k(s\preceq k\to A_\psi(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k(s\preceq k\to (A_\psi(k)\wedge (B_\psi(k))\to C_\psi(k)))$$
-$$\hspace*{-2mm}\forall k(s\preceq k\to A_\psi(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(s\preceq k\to (A_\psi(k)\vee (B_\psi(k))\to C_\psi(k)))$$
-$$\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A_\psi(f_\psi(k))\to B_\psi(f_\psi(k)))\to C_\psi(k))$$
-and $f$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model then we have a counter-model $(M',I')$ where $$M' = \{m\in M|s\preceq m\}$$ and
+$$\hspace*{-2mm}\forall k(s\preceq k\to A(k)\to (B(k)\wedge C(k))), \forall k(s\preceq k\to (A(k)\wedge (B(k))\to C(k)))$$
+$$\hspace*{-2mm}\forall k(s\preceq k\to A(k)\to (B(k)\vee C(k))), \forall k(s\preceq k\to (A(k)\vee (B(k))\to C(k)))$$
+$$\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A(f_\psi(k))\to B(f_\psi(k)))\to C(k))$$
+and $f_\psi$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model then we have a counter-model $(M',I')$ where $$M' = \{m\in M|s\preceq m\}$$ and
 $$f_\psi^{I'}(u) = \begin{cases}
@@
 
+This is as far as the first-order encoding can be simplified. What remains is the elimination of the remaining weak $\forall$-quantifiers.
+
+\section{IPC to CPC}
+
 We now argue that if a counter-model exists for $\varphi^\circ$ then a bounded counter-model exists, allowing us to return to the propositional case.
@@
 	1.
-If $C_\psi^I(u)$ is true then we are done due to persistency.
-Otherwise we know that $(A_\psi^I(f_\psi^I(u))\to B_\psi^I(f_\psi^I(u)))\to C_\psi^I(f_\psi^I(u))$ holds, so $A_\psi^I(f_\psi^I(u))\to B_\psi^I(f_\psi^I(u))$ must be true.
+If $C^I(u)$ is true then we are done due to persistency.
+Otherwise we know that $(A^I(f_\psi^I(u))\to B^I(f_\psi^I(u)))\to C^I(f_\psi^I(u))$ holds, so $A^I(f_\psi^I(u))\to B^I(f_\psi^I(u))$ must be true.
 
 	2.
-If $C_\psi^I(u)$ is true then $C_\psi^I(w)$ is also true and we are done.
-So suppose $A_\psi^I(u)\to B_\psi^I(u)$ is false, i.e.
-$A_\psi^I(u)$ true and $B_\psi^I(u)$ false.
-Then due to the persistency condition $A_\psi^I(w)$ is also true.
-Again if $C_\psi^I(w)$ is true we are done.
-Otherwise we know that $B^I(f^I_\psi(w))$ must be false, since $(A^I_\psi(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I_\psi(w)$ and $A^I_\psi(f^I_\psi(w))$ hold.
-But then due to persistency so is $B^I_\psi(w)$ and we have that $\psi$ is fulfilled at $w$.
+If $C^I(u)$ is true then $C^I(w)$ is also true and we are done.
+So suppose $A^I(u)\to B^I(u)$ is false, i.e.
+$A^I(u)$ true and $B^I(u)$ false.
+Then due to the persistency condition $A^I(w)$ is also true.
+Again if $C^I(w)$ is true we are done.
+Otherwise we know that $B^I(f^I(w))$ must be false, since $(A^I(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I(w)$ and $A^I(f^I_\psi(w))$ hold.
+But then due to persistency so is $B^I(w)$ and we have that $\psi$ is fulfilled at $w$.
 \end{proof}
@@
 	Suppose $\mathcal S$ is a set of formulas $\psi$ of the form
-	$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi),(A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$
+	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C),(A\vee B)\to C, (A\to B)\to C$$
 	Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$.
@@
 		\item $A^k\to A^{k\psi}$ for each $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
-		\item $A^k_\psi\to (B^k_\psi\circ C^k_\psi)$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A^k_\psi\circ B^k_\psi)\to C^k_\psi$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A^{k\psi}_\psi\to B^{k\psi}_\psi)\to C^k_\psi$ for $(A_\psi\to B_\psi)\to C_\psi\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
+		\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $(A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
 	\end{itemize}
@@
 
-Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$and$$\varphi = \bigwedge \mathcal S\to P$$
+Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P$$
 and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$.
@@
 \forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)),$$$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),$$$$ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))$$
-As in the propositional case we will now encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
+
+The complete result that we will show is as follows:
+\begin{theorem}\label{fullFOtranslation}
+	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain the formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$.
+Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
+For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$.
+For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$.
+Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
+Obtain $\mathcal S^\#$ by including formulas as follows:
+	\begin{itemize}
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
+		\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
+	\end{itemize}
+	Then $\bigwedge S\to P$ is intuitionistically iff $\exists xE(x, s)\to K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol, $\Sigma$ is the original signature and
+	\begin{align*}
+		K := &\:\bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
+		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
+		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}
+	\end{align*}
+	The size of the obtained formula is linear in the input.
+However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
+\end{theorem}
+
+It is not possible to completely eliminate worlds from the domain as in the propositional case, since counter-models possibly must be infinite. We will see why that is in more detail later.
+
+\subsection{Encoding Kripke Semantics}
+
+As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
 \begin{itemize}
@@
 
-We proceed now as in the previous section.
+\subsection*{Reducing the encoding}
+
+As in the predicate case we now show the existence of a canonical counter-model, if a counter-model exists.
 Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$.
 Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
-For each  $\psi\in\mathcal F$ consider a new function symbol $f$.
+For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$.
 As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
 $$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$ with
-$$\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$$ and
-$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$$ with $$\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u))))\to B^\#(\vec b, u))$$
-and in each other formula in $\mathcal S^\circ$ remove the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
+$$\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$$ and
+$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$$ with $$\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$$
+and in each other formula in $\mathcal S^\circ$ removing the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
 $$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
@@
 
-Part 2 of this Lemma does not hold for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$) we have $A(x, v)$ there could be some term $t$ with $\neg E(t, v)$ and $w$ with $v\preceq w$ such that $E(t, w)$ and $\neg A(t, w)$.
+However Part 2 of this Lemma does not hold for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$) we have $A(x, v)$ there could be some term $t$ with $\neg E(t, v)$ and $w$ with $v\preceq w$ such that $E(t, w)$ and $\neg A(t, w)$.
 This is why we must consider sequences that have repetitions of such formulas:

### 2022-07-14 CSL/lipics-v2021-EIICL.tex
typo
added some comments

@@
 \usetikzlibrary{arrows,shapes,quotes}
+
+\usepackage{todonotes}
 
@@
 
-Before proceeding towards the results let us quickly fix our notation and recapitulate the semantics for classical and more importantly intuitionistic logic.
+In this section we fix notation and recall the semantics for classical and intuitionistic logic.
 
@@
 \begin{definition}
-	A \emph{valuation} $v$ is a function that assigns to each propositional variable $A$ a \textbf{truth value} $v(A)\in\{0, 1\}$.
-We inductively define the model relation $\models$ between $v$ and formulas:
+A \emph{valuation} $v$ maps each propositional variable $A$ to a \emph{truth value} $v(A)\in\{0, 1\}$. We inductively define the model relation $\models$ between $v$ and formulas:
 	\begin{itemize}
@@
 	\end{itemize}
-	$v$ is a \textbf{model} for $\varphi$ if $v\models\varphi$. If every valuation is a model for $\varphi$ then we say $\varphi$ is \emph{valid}. We denote the set of valid formulas with \emph{CPC}.
+A valuation $v$ is a \emph{model} for $\varphi$ if $v\models\varphi$.
+If every valuation is a model for $\varphi$ then we say $\varphi$ is \emph{valid}.
+We denote the set of valid formulas with \emph{CPC}.
+\todo{What does CPC stand for? Introduce abbreviation somewhere like classical predicate ... (CPC).}
 \end{definition}
@@
 \begin{definition}
-	A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of nodes an a family of valuations $(v_w)_{w\in W}$ such that for $u\leq w$ we have $v_u(A)\leq v_u(A)$ for every propositional variable $A$, this is called the \emph{persistency condition}.
-The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, i.e.
-have
+\todo{We later also use the notion `Kripke frame', we  might introduce it here.}
+\todo{Note that I have replaced `nodes' by `worlds' here.}
+A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds} and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
+The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, where we set
 	\begin{itemize}
@@
 	\end{itemize}
-	We say that $\varphi$ is satisfied at a node $u$ if $v_u\models\varphi$ and write $u\models\varphi$.
-If a formula is satisfied at every node then $\mathcal K$ is a \textbf{model} for $\varphi$.
+We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$.
+If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$.
 A formula is valid if every Kripke structure is a model of it.
 We denote the set of valid formulas with IPC.
+\todo{introduce abbreviation}
 \end{definition}
-There are many classical tautologies which are not intuitionistically valid
+There are many classical tautologies which are not intuitionistically valid:
 \begin{example}\label{LEMcounterexample}
@@
 This is insufficient for intuitionistic logic, e.g.
-$A\vee\neg A$ is not a intuitionistic tautology but its negation is not satisfiable either.
+$A\vee\neg A$ is not an intuitionistic tautology but its negation is not satisfiable either.
 
@@
 
-We now give account of the more involved semantics of first-order logic
+We now recall the semantics of first-order logic:
 
 \begin{definition}
-	Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the domain of $\mathcal{M}$, and a function $I$ that assigns
+Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
 	\begin{itemize}
@@
 \subsection{Skolemization and Herbrandization}
+
 An important step in the embedding will be the elimination of quantifiers via Herbrandization.
@@
 
-\begin{theorem}For every formula $\varphi$
+\begin{theorem}
+\label{thm:herbrand-skolem}
+For every formula $\varphi$
 	\begin{itemize}
@@
 
-which follows from Lemmata~\ref{ap1} and~\ref{ap2} in the appendix. We are now ready to present our embedding of intuitionistic into classical logic.
+Theorem~\ref{thm:herbrand-skolem} follows from Lemmata~\ref{ap1} and~\ref{ap2} in the appendix.
 
@@
 
-\subsection*{Reducing the encoding}
+\subsection{Reducing the encoding}
 

### 2022-07-14 CSL/lipics-v2021-EIICL.tex
comments for propositional case

@@
 	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
-		m&\text{if $b=a$}\\
-		v(b)&\text{otherwise}
+		m, &\text{if $b=a$,}\\
+		v(b), &\text{otherwise.}
 	\end{cases}$$
@@
 \end{definition}
-In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically.
-Consider for instance the classical tautology $\varphi$ $$\neg\forall x A(x)\to \exists x \neg A(x)$$Analogous remarks with regard to validity and satisfiability hold as in the propositional case.
+In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
+Analogous remarks with regard to validity and satisfiability hold as in the propositional case.
 
@@
 An important step in the embedding will be the elimination of quantifiers via Herbrandization. 
-In this process we introduce free variables and also additional function symbols to the signature.
-Whenever we speak of a new free variable what is meant is a free variable that does not occur in any of the considered formulas.
-Whenever a new function symbol is introduced we implicitly extend the signature by some not previously contained symbol.
+In this process we introduce fresh variables and add additional function symbols to the signature. 
+A fresh variable is any variable that does not occur in any of the considered formulas. 
+Whenever we add function symbol we implicitly extend the signature by some not previously contained symbol.
 
@@
 
-Theorem~\ref{thm:herbrand-skolem} follows from Lemmata~\ref{ap1} and~\ref{ap2} in the appendix.
+Theorem~\ref{thm:herbrand-skolem} follows from Lemmata~\ref{ap1} and~\ref{ap2} stated in the appendix.
 
@@
 
-All that we have done is model Kripke frames with a finite number of propositional variables as a first-order theory and by definition we have
+\noindent As all we have done is modelling Kripke frames as a first-order theory we directly obtain:
 
@@
 \begin{example}
-	Consider the law of excluded middle $$\varphi = (A\to \bot)\vee A$$ which is not intuitionistically valid. Then
+	Consider the law of excluded middle $$\varphi = (A\to \bot)\vee A$$ which is not intuitionistically valid. Then,
 	\begin{align*}
@@
 		&= K(\varphi)\to (\forall u'( u\preceq u' \to A^{u'}\to \bot^{u'})\vee A(u))\\
-		&= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot)\vee A(u))
+		&= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot)\vee A(u)),
 	\end{align*}
@@
 A similar translation was proposed in~\cite{statman1979intuitionistic}.
+\todo{what has been done in the previous work?
+just the propositional case?
+just classical logic?
+did you find a citation for the full Tseytin-like transformation somewhere?
+If this is new we might add it to the intro.}
 
@@
 \begin{proof}
-	The claims are verified in a straightforward manner by induction on the height of $\varphi$.
+The claims are verified in a straightforward manner by induction on $\varphi$.
 \end{proof}
@@
 Then from the previous three Lemmas we get
+\todo{as a general rule, I like to finish sentences with a proper grammatical ending, here `:'}
 
@@
 instead of $\varphi$ we consider $$\varphi' = \bigwedge \mathcal S^-(\varphi)\to P_\varphi$$
+\todo{here the previous sentence did not finish with a proper grammatical ending as well, I would add `.' after $P_\varphi$}
 
 Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
-$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$
-
+$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$$
 where we can treat $A\to B$ as a special case $A\to (B\wedge B)$. We then apply the transformation to CQC, leading to a formula
-$$\varphi'' = K(\varphi')\to\bigwedge  \mathcal S\to P_\varphi(u)$$
+$$\varphi'' = K(\varphi')\to\bigwedge  \mathcal S\to P_\varphi(u),$$
 where $u$ is some new free variable and each $\psi\in\mathcal S$ is of one of the forms
@@
 $$\hspace*{-2mm}\forall k(s\preceq k\to A(k)\to (B(k)\vee C(k))), \forall k(s\preceq k\to (A(k)\vee (B(k))\to C(k)))$$
-$$\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A(f_\psi(k))\to B(f_\psi(k)))\to C(k))$$
+$$\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A(f_\psi(k))\to B(f_\psi(k)))\to C(k)),$$
 and $f_\psi$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model then we have a counter-model $(M',I')$ where $$M' = \{m\in M|s\preceq m\}$$ and
 $$f_\psi^{I'}(u) = \begin{cases}
-	f^I_\psi(u)&\text{ iff $u\preceq f^{I}_\psi(u)$}\\
-	u&\text{ else}
+	f^I_\psi(u), &\text{ iff $u\preceq f^{I}_\psi(u)$,}\\
+	u,&\text{ else.}
 \end{cases}$$ that is where $s$ is a least element and the $f_\psi$ increase their argument.
 Therefore instead of $\varphi'''$ we may consider a formula $$\varphi^\circ = \forall k(s\preceq k)\to \bigwedge\{\forall k(k\preceq f_\psi(k))\:|\:\psi\in\mathcal F_\to\}\to K(\varphi')\to \bigwedge \mathcal S^\circ\to P_\varphi(s)$$ where each formula $\psi\in\mathcal S^\circ$ is of one of the forms
-$$\forall k(A_\psi(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k((A_\psi(k)\wedge B_\psi(k))\to C_\psi(k))$$
-$$\forall k(A_\psi(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k((A_\psi(k)\vee B_\psi(k))\to C_\psi(k))$$
-$$\forall k((A_\psi(f_\psi(k))\to B_\psi(f_\psi(k)))\to C_\psi(k))$$
+$$\forall k(A(k)\to (B(k)\wedge C(k))), \forall k((A(k)\wedge B(k))\to C(k))$$
+$$\forall k(A(k)\to (B(k)\vee C(k))), \forall k((A(k)\vee B_\psi(k))\to C(k))$$
+$$\forall k((A(f_\psi(k))\to B(f_\psi(k)))\to C(k))$$
 and $\mathcal F_\to$ denotes the set of formulas of the last kind.
@@
 This is as far as the first-order encoding can be simplified. What remains is the elimination of the remaining weak $\forall$-quantifiers.
+\todo{delete the last two sentences?}
 
@@
 Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
-For $\psi\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true.
+We say $\psi\in\mathcal F_\to$ is \emph{fulfilled} at $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true.
 For $\psi\in\mathcal F_\to$ define $$g_\psi : M\to M, u\mapsto\begin{cases}
-		u&\text{ if $\psi$ is fulfilled in $u$}\\
-		f^I_\psi(u)&\text{ else}		
-	\end{cases}$$Define a $\mathcal M_T = (M_T, I_T)$ as follows:
+		u,&\text{ if $\psi$ is fulfilled in $u$,}\\
+		f^I_\psi(u),&\text{ else.}		
+	\end{cases}$$Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
 	\begin{itemize}
@@
 		\item Interpret $\preceq$ as the prefix-order.
-		\item Have $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
-			\psi_1\dots\psi_n&\text{if $\psi$ occurs in $\psi_1\dots\psi_n$}\\
-			\psi_1\dots\psi_n\psi&\text{else}			
+		\item We set $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
+			\psi_1\dots\psi_n, &\text{if $\psi$ occurs in $\psi_1\dots\psi_n$,}\\
+			\psi_1\dots\psi_n\psi, &\text{else.}			
 		\end{cases}$$
-		\item For propositional variables $P$ have $$P^{I_T}(\psi_1\dots \psi_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots))$$
+		\item For propositional variables $P$, we set $$P^{I_T}(\psi_1\dots \psi_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots)).$$
 	\end{itemize}
@@
 \begin{proof}
-	1.
-If $C^I(u)$ is true then we are done due to persistency.
-Otherwise we know that $(A^I(f_\psi^I(u))\to B^I(f_\psi^I(u)))\to C^I(f_\psi^I(u))$ holds, so $A^I(f_\psi^I(u))\to B^I(f_\psi^I(u))$ must be true.
-
-	2.
-If $C^I(u)$ is true then $C^I(w)$ is also true and we are done.
-So suppose $A^I(u)\to B^I(u)$ is false, i.e.
-$A^I(u)$ true and $B^I(u)$ false.
-Then due to the persistency condition $A^I(w)$ is also true.
-Again if $C^I(w)$ is true we are done.
-Otherwise we know that $B^I(f^I(w))$ must be false, since $(A^I(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I(w)$ and $A^I(f^I_\psi(w))$ hold.
+\todo{I corrected the proof, please check again.}
+
+	1. If $C^I(u)$ is true then we are done due to persistency. Otherwise $C^I(u)$ is false.
+Because $(A^I(f_\psi^I(u))\to B^I(f_\psi^I(u)))\to C^I(u)$ holds, then $A^I(f_\psi^I(u))\to B^I(f_\psi^I(u))$ must be false.
+
+	2. Let $w$ be some element with $w\geq u$.
+If $C^I(w)$ is true, we are done. 
+Otherwise $C^I(w)$ is false.
+Due to persistency $C^I(u)$ is also false.
+Because $(A^I(f_\psi^I(u))\to B^I(f_\psi^I(u)))\to C^I(u)$ holds, then $A^I(f_\psi^I(u))\to B^I(f_\psi^I(u))$ must be false, i.e.
+$A^I(u)$ is true and $B^I(u)$ is false.
+Then $A^I(w)$ and $A^I(f^I_\psi(w))$ is also true due to persistency.
+Because $(A^I(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I(w)$ holds, we now get that  $B^I(f^I(w))$ must be false.
 But then due to persistency so is $B^I(w)$ and we have that $\psi$ is fulfilled at $w$.
@@
 Essentially what fulfilment of $\psi$ at $u$ tells us is that having $f_\psi(w) = w$ for all $w\geq u$ works.
+\todo{what is meant by `having $f_\psi(w) = w$ for all $w\geq u$ works`'? Reformulate the sentence.}
 With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.
@@
 	Consider
-	$$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model
+	$$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model, depicted below,
 	\begin{center}
@@
 	\vspace*{-.3cm}where at each node the true propositions are indicated.
-
+%	
 There is a corresponding classical counter-model $\mathcal M = (M, I)$ to $\varphi^\circ$ with $M = \{u, u_l, u_r\}$, $A^{I}(l) = D^I(l) = B^I(r) = C^I(r) = 1$ and $0$ else.
-
-	Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (A\to B)\to C$ this corresponds to a counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below
+%
+Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (A\to B)\to C$ \todo{I guess this should be $\psi_2 := (B\to A)\to D$}, this corresponds to a transformed counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below:
 	\begin{center}
@@
 So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
-	x&\text{ if $\psi$ occurs in $x$}\\
-	x\psi&\text{ otherwise}
+	x, &\text{ if $\psi$ occurs in $x$,}\\
+	x, \psi&\text{ otherwise.}
 \end{cases}$$ Now we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula.
@@
 	Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$.
-For each atom $A$ occurring in $\mathcal S$ and $k\in \mathcal F_\to$ consider a new atom $A^{k}$, which will correspond to $A^{I_F}(k)$ from the previous definition.
+For each atom $A$ occurring in $\mathcal S$ and $k\in \mathcal F_\to$, consider a new atom $A^{k}$, which will correspond to $A^{I_F}(k)$ from the previous definition.
+\todo{$A^{I_F}(k)$ does not exist anymore. I guess it is from a previous version?}
+\todo{As a general rule, I would remove explanations from definitions, and do remarks after the definition}
 Obtain $\mathcal S^\#$ by including the formulas
@@
 		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $(A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
+		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $(A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occur in $k$.
 	\end{itemize}
@@
 	and we know that $\varphi$ is valid iff $\bigwedge \mathcal S(\varphi)\to P_\varphi$ is valid.
-Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
+Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically valid, where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
 	\begin{align*}
@@
 	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
-	for leafs and
-	and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
+	for leafs and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
 	Example encodings of the above formulas would be
@@
 		\item $f^I(m_1\dots m_n) =\begin{cases}
-			 f^{I_u}(x_1\dots x_n) &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$}\\
-			 w_0 & \text{otherwise}
+			 f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
+			 w_0, & \text{otherwise.}
 		\end{cases}$
 		\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
-			A^{I_u}(x_1\dots x_n) &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$}\\
-			\top & \text{otherwise}
+			A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
+			\top, & \text{otherwise.}
 		\end{cases}$
@@
 For $\psi\in\mathcal F$ define $$g_{\vec a, \psi} : M\to M, u\mapsto\begin{cases}
-		u&\text{ if $\psi$ is fulfilled in $u$ with $\vec a$}\\
-		f^I_\psi(\vec a, u)&\text{ else}		
+		u, &\text{ if $\psi$ is fulfilled in $u$ with $\vec a$,}\\
+		f^I_\psi(\vec a, u), &\text{ else.}		
 	\end{cases}$$		
@@
 		\item Have $$f_\psi^{I_T}(x_1\dots x_n) = \begin{cases}
-			x_1\dots x_n&\text{if $x=x_n$}\\
-			x_1\dots x_nx&\text{else}			
+			x_1\dots x_n,&\text{if $x=x_n$,}\\
+			x_1\dots x_nx,&\text{else.}			
 		\end{cases}$$
@@
 
+\todo{the lemmata still use the set `T', so the notation is probably not consistent anymore}
+
 \begin{lemma}\label{ap1}

### 2022-07-14 CSL/lipics-v2021-EIICL.tex
all comments for the technical sections added

@@
 
-It is well known that a polynomial time translation between QBF and IPC must exists, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}.
-However giving an explicit translation from IPC to QBF can still be useful as good solvers exists for QBF but for IPC not so much.
-The translation from the last section will serve as a useful tool for this.
-We will not explain the semantics of QBF in depth.
-Intuitively just think that the interpretation of existentially quantified variables may depend on variables quantified above them.
-
-Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P$$
+It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}. 
+However, giving an explicit translation from IPC to QBF still promises to be useful, allowing to leverage the progress in QBF solving.
+The translation from the last section will serve as starting point for developing the translation in this section.
+For lack of space, we do not define the semantics of QBF in this paper and refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
+
+Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
 and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$.
@@
 \begin{proof}
-	Suppose $\varphi$ is not intuitionistically valid, i.e.
-there exists a counter-model $\mathcal M$ for $\mathcal S^\#\to P^\epsilon$.
+\todo{The proof contains some old notation at several places.}
+\todo{Also the proof could be beautified a bit.}
+Suppose $\varphi$ is not intuitionistically valid, i.e. there exists a counter-model $\mathcal M$ for $\mathcal S^\#\to P^\epsilon$.
 For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$.
@@
 
-Since we have not discussed QBF in detail the above argument might be a bit difficult to follow, in particular the second part, but is simple in essence. We shall demonstrate with help of a example.
-
 \begin{example}
+\todo{I think this example is a candidate for the appendix.}
 	Consider the formula $\varphi = (((A\to B)\to C)\wedge((B\to A)\to C))\to C$, which is a non-intuitionistic tautology.
@@
 
-Note that the last set of existentially quantified variables is superfluous because there is only one assignment which even has the chance to falsify $\varphi^Q$, namely that which assigns $1$ to the only $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$.
-That is in our final formula we can replace $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^i$.
+Note that the last set of existentially\todo{universally?} quantified variables can be avoided because there is only one assignment that has the chance to falsify $\varphi^Q$, namely the assignment that assigns $1$ to the single $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$.
+Hence, in our final formula we can replace every $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^n$.
 With that we get the following:
@@
 \begin{corollary}
-	Let $N$-INTINVALID be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid.
-Then $N$-INTINVALID is in $\Sigma_{2N-1}^P$.
-The dual problem $N$-INTVALID is in $\Pi_{2N-1}^P$.
+	Let $N$-\textbf{IntInvalid} be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid.
+Then $N$-\textbf{IntInvalid} is in $\Sigma_{2N-1}^P$.
+The dual problem $N$-\textbf{IntValid} is in $\Pi_{2N-1}^P$.
 \end{corollary}
+
+\todo{The size analysis of the QBF reduction should be stated in some theorem in this section}
 
@@
 	The other direction is a bit more involved.
-Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
+Suppose we have a Kripke counter-model to $\varphi$ with Frame \todo{frame?
+Also make sure frame is defined} $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
 In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
 Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
-Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$ and extend $\sim$ by having $[w] = \{w\}$ for $w\in W_0$.
+Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$ and extend $\sim$ by having $[w] = \{w\}$ for $w\in W_0$ \todo{why is the later needed?
+I don't think it is used}.
 Now have
@@
 As in the previous section we can define a tree-counter model for this.
+\todo[inline]{Important: Why can't we have Theorem~\ref{thm:fo-reduction} directly, without Def.~\ref{def:fo-fulfilled} and Lemma~\ref{thm:fo-countermodel-reduction}? They don't seem to be needed.
+However, Lemma~\ref{thm:fo-countermodel-reduction} could still be used as in the propositional case to eliminate the function symbols for the implications at the cost of introducing additional predicates.
+One could also add the idempotence as axioms in order to aid the solver. 
+We could mentioned that as a justification for Lemma~\ref{thm:fo-countermodel-reduction}.
+}
+
 \begin{definition}
+\label{def:fo-fulfilled}
+\todo{this definition still uses variables $t$, please use a different letter}
 	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
@@
 		\item $M_T$ are the sequences on $\{ \vec t, \psi\:|\:\psi\in \mathcal F, \vec t\in M^n\}$ without repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\to$ and without immediate repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\forall$.
+\todo{do you mean sequences of pairs here?
+Then it is probably better to use brackets as in $(\vec t, \psi)$}
 		\item Interpret $\preceq$ as the prefix-order.
@@
 \begin{lemma}
+\label{thm:fo-countermodel-reduction}
 	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
@@
 Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$$ is sufficient.
-In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PO-encoding from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
+In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PO-encoding \todo{PO-encoding used for the first time, say PartialOrder?}
+from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
 Define
@@
 \begin{theorem}
+\label{thm:fo-reduction}
 	$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.

### 2022-07-15 CSL/lipics-v2021-EIICL.tex
Corrections

@@
 If every valuation is a model for $\varphi$ then we say $\varphi$ is \emph{valid}.
-We denote the set of valid formulas with \emph{CPC}.
-\todo{What does CPC stand for? Introduce abbreviation somewhere like classical predicate ... (CPC).}
+We denote the set of valid formulas with \emph{CPC} (Classical Propositional Calculus).
 \end{definition}
@@
 \begin{definition}
-\todo{We later also use the notion `Kripke frame', we  might introduce it here.}
-\todo{Note that I have replaced `nodes' by `worlds' here.}
 A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds} and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
@@
 A formula is valid if every Kripke structure is a model of it.
-We denote the set of valid formulas with IPC.
-\todo{introduce abbreviation}
+We denote the set of valid formulas with \emph{IPC} (Intuitionistic Propositional Calculus).
 \end{definition}
@@
 	\end{itemize}
-	$\mathcal{M}$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$. We denote the set of valid formulas with CQC.
-\end{definition}
-
-\begin{definition}
-	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$ and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$
+	$\mathcal{M}$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$.
+A formula is valid if every $\Sigma$-structure satisfies it.
+We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).
+\end{definition}
+
+\begin{definition}
+	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, its (Kripke) \emph{Frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$
 	\begin{itemize}
@@
 		\item $u, v\not\models\bot$
-		\item $u, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I_u, v},\dots,t_n^{I_u, v})\in R^{I_u}$.
+		\item $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$.
 		\item $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
@@
 	\end{itemize}
-	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. We denote the set of valid formulas with IQC.
+	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$.
+A formula is valid if it is satisfied in eveery $\Sigma$-Kripke structure, it is valid for some Frame (write $W\models\varphi$) if it is satisfied in every Krike structure with that Frame.
+We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
 \end{definition}
@@
 Of course transforming a propositional formula into a predicate formula is somewhat unsatisfying.
-As a first step we want to apply Herbrandization.
+As a first step towards obtaining a propositional formula we want to apply Herbrandization.
 Now trying to apply it to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated.
 We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form.
-A similar translation was proposed in~\cite{statman1979intuitionistic}.
-\todo{what has been done in the previous work?
-just the propositional case?
-just classical logic?
-did you find a citation for the full Tseytin-like transformation somewhere?
-If this is new we might add it to the intro.}
+In fact our translation is often given as a first step in the Tseytin transformation after which the individual formulas are then converted to CNF, so it is well known, but we are not aware of work that uses just this step as a stand-alone.
+A similar translation for intuitionistic logic was also proposed in~\cite{statman1979intuitionistic}.
 
@@
 
-Then from the previous three Lemmas we get
-\todo{as a general rule, I like to finish sentences with a proper grammatical ending, here `:'}
+Then from the previous three Lemmas we get:
 
@@
 As a first step we perform the normalization from the last section, i.e.
-instead of $\varphi$ we consider $$\varphi' = \bigwedge \mathcal S^-(\varphi)\to P_\varphi$$
-\todo{here the previous sentence did not finish with a proper grammatical ending as well, I would add `.' after $P_\varphi$}
+instead of $\varphi$ we consider $$\varphi' = \bigwedge \mathcal S^-(\varphi)\to P_\varphi.$$
 
@@
 
-This is as far as the first-order encoding can be simplified. What remains is the elimination of the remaining weak $\forall$-quantifiers.
-\todo{delete the last two sentences?}
-
 \section{IPC to CPC}
@@
 \begin{proof}
-\todo{I corrected the proof, please check again.}
 
@@
 Due to persistency $C^I(u)$ is also false.
-Because $(A^I(f_\psi^I(u))\to B^I(f_\psi^I(u)))\to C^I(u)$ holds, then $A^I(f_\psi^I(u))\to B^I(f_\psi^I(u))$ must be false, i.e.
-$A^I(u)$ is true and $B^I(u)$ is false.
-Then $A^I(w)$ and $A^I(f^I_\psi(w))$ is also true due to persistency.
+Because $\psi$ is fulfilled at $u$, $A^I(u)\to B^I(u)$ must be false, i.e. $A^I(u)$ is true and $B^I(u)$ is false. Then $A^I(w)$ and $A^I(f^I_\psi(w))$ are also true due to persistency.
 Because $(A^I(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I(w)$ holds, we now get that  $B^I(f^I(w))$ must be false.
@@
 
-Essentially what fulfilment of $\psi$ at $u$ tells us is that having $f_\psi(w) = w$ for all $w\geq u$ works.
-\todo{what is meant by `having $f_\psi(w) = w$ for all $w\geq u$ works`'? Reformulate the sentence.}
+Essentially what this tells us is that for every $u$ at which $\psi$ is fulfilled, in particular each $u$ in the image of $f_\psi$, we may add the assumption that $f_\psi(w) = w$ for all $w\geq u$ without changing the validity of $\psi$.
 With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.
@@
 \end{corollary}
+
+\begin{remark}
+	It is well known that Kripke Frames which are trees are complete with regard to intuitionistic logic.
+Our contribution is giving an explicit geometry that depends only on the structure of the considered formula.
+\end{remark}
 
@@
 %
-Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (A\to B)\to C$ \todo{I guess this should be $\psi_2 := (B\to A)\to D$}, this corresponds to a transformed counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below:
+Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (A\to B)\to D$, this corresponds to a transformed counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below:
 	\begin{center}
@@
 	Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$.
-For each atom $A$ occurring in $\mathcal S$ and $k\in \mathcal F_\to$, consider a new atom $A^{k}$, which will correspond to $A^{I_F}(k)$ from the previous definition.
-\todo{$A^{I_F}(k)$ does not exist anymore. I guess it is from a previous version?}
-\todo{As a general rule, I would remove explanations from definitions, and do remarks after the definition}
+For each atom $A$ occurring in $\mathcal S$ and $k\in \mathcal F_\to$, consider a new atom $A^{k}$.
 Obtain $\mathcal S^\#$ by including the formulas
@@
 \begin{theorem}
-	$$\mathcal S\models_C P\text{ iff }\mathcal S^\#\models_I P^\emptyset$$
+	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
 \end{theorem}
@@
 	\begin{align*}
-		\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\wedge\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
+		\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
 		\text{\normalfont Persistent}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
 		\text{\normalfont Sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
-		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\to C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
+		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
 		\text{\normalfont Sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
@@
 \todo{Also the proof could be beautified a bit.}
-Suppose $\varphi$ is not intuitionistically valid, i.e. there exists a counter-model $\mathcal M$ for $\mathcal S^\#\to P^\epsilon$.
+Suppose $\varphi$ is not intuitionistically valid, i.e. there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
 For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$.
@@
 \begin{example}
-\todo{I think this example is a candidate for the appendix.}
 	Consider the formula $\varphi = (((A\to B)\to C)\wedge((B\to A)\to C))\to C$, which is a non-intuitionistic tautology.
@@
 
-Note that the last set of existentially\todo{universally?} quantified variables can be avoided because there is only one assignment that has the chance to falsify $\varphi^Q$, namely the assignment that assigns $1$ to the single $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$.
+A simple counting argument shows that for each $n < |\mathcal F_\to|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal F_\to|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal S}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal F_\to}(n)|\in \mathcal O(|\mathcal F_\to|)$ so overall we get
+\begin{lemma}
+	The size of $\varphi^Q$ is in $\mathcal O(|\varphi|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$.
+\end{lemma}
+
+Note that the last set of universally quantified variables can be avoided because there is only one assignment that has the chance to falsify $\varphi^Q$, namely the assignment that assigns $1$ to the single $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$.
 Hence, in our final formula we can replace every $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^n$.
@@
 \end{corollary}
-
-\todo{The size analysis of the QBF reduction should be stated in some theorem in this section}
 
@@
 	The other direction is a bit more involved.
-Suppose we have a Kripke counter-model to $\varphi$ with Frame \todo{frame?
-Also make sure frame is defined} $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
+Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
 In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
 Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
-Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$ and extend $\sim$ by having $[w] = \{w\}$ for $w\in W_0$ \todo{why is the later needed?
-I don't think it is used}.
+Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
 Now have
@@
 and in each other formula in $\mathcal S^\circ$ removing the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
-$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
-As in the previous section we can define a tree-counter model for this.
-\todo[inline]{Important: Why can't we have Theorem~\ref{thm:fo-reduction} directly, without Def.~\ref{def:fo-fulfilled} and Lemma~\ref{thm:fo-countermodel-reduction}? They don't seem to be needed.
-However, Lemma~\ref{thm:fo-countermodel-reduction} could still be used as in the propositional case to eliminate the function symbols for the implications at the cost of introducing additional predicates.
-One could also add the idempotence as axioms in order to aid the solver. 
-We could mentioned that as a justification for Lemma~\ref{thm:fo-countermodel-reduction}.
-}
+$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#(s)$$
+As in the previous section we can define a tree counter-model for this.
+
+\begin{definition}
+	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
+	\begin{itemize}
+		\item $M_T$ are the sequences on $\{(\vec z, \psi)\:|\:\psi\in \mathcal F, \vec z\in M^n\}$.
+		\item Interpret $\preceq$ as the prefix-order.
+		\item Have $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi)$$
+		\item For propositional variables $P$ have $$P^{I_T}((\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = P^I(f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots))$$
+	\end{itemize}
+\end{definition}
+
+\begin{lemma}
+	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
+\end{lemma}
+
+Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec z_1, f_{\psi_2}(\vec z_2, \dots f_{\psi_n}(\vec z_n, s)\dots))$$ is sufficient.
+In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
+from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
+Define
+
+$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#$$
+then we have
+\begin{theorem}
+	\label{thm:fo-reduction}
+	$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
+\end{theorem}
+The complete translation is summarised in Theorem~\ref{fullFOtranslation}. Let us understand why a further reduction as in the propositional case is impossible.
 
@@
 \label{def:fo-fulfilled}
-\todo{this definition still uses variables $t$, please use a different letter}
 	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
-For $\psi = \forall\vec t\psi'\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff $(A_\psi^I(\vec t_A, u)\to B_\psi^I(\vec t_B, u))[\vec a/\vec t]$ is false or $C_\psi^I(\vec t_C, u)[\vec a/\vec t]$ is true.
-	For $\psi = \forall\vec t\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff there is $a\in M_u$ such that $A^{I[a/x]}_\psi(\vec t_A)[\vec a/\vec t]$ is false or $B_\psi^I(\vec t_B)[\vec a/\vec t]$ is true.
-For $\psi\in\mathcal F$ define $$g_{\vec a, \psi} : M\to M, u\mapsto\begin{cases}
-		u, &\text{ if $\psi$ is fulfilled in $u$ with $\vec a$,}\\
-		f^I_\psi(\vec a, u), &\text{ else.}		
-	\end{cases}$$		
-	Define a $\mathcal M_T = (M_T, I_T)$ as follows:
-	\begin{itemize}
-		\item $M_T$ are the sequences on $\{ \vec t, \psi\:|\:\psi\in \mathcal F, \vec t\in M^n\}$ without repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\to$ and without immediate repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\forall$.
-\todo{do you mean sequences of pairs here?
-Then it is probably better to use brackets as in $(\vec t, \psi)$}
-		\item Interpret $\preceq$ as the prefix-order.
-		\item Have $$f_\psi^{I_T}(x_1\dots x_n) = \begin{cases}
-			x_1\dots x_n,&\text{if $x=x_n$,}\\
-			x_1\dots x_nx,&\text{else.}			
-		\end{cases}$$
-		\item For propositional variables $P$ have $$P^{I_T}(x_1\dots x_n) = P^I(g_{x_n}(\dots(g_{x_1}(s^I))\dots))$$
-	\end{itemize}
+For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
+	For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
 \end{definition}
@@
 
-However Part 2 of this Lemma does not hold for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$) we have $A(x, v)$ there could be some term $t$ with $\neg E(t, v)$ and $w$ with $v\preceq w$ such that $E(t, w)$ and $\neg A(t, w)$.
-This is why we must consider sequences that have repetitions of such formulas:
+However Part 2 of this Lemma fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$) we have $A(x, v)$ there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
+This is why we must consider sequences that have repetitions:
 
@@
 
-\begin{corollary}
-	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
-\end{corollary}
-
-Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$$ is sufficient.
-In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PO-encoding \todo{PO-encoding used for the first time, say PartialOrder?}
-from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
-Define
-
-$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#$$
-then we have
-\begin{theorem}
-\label{thm:fo-reduction}
-	$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
-\end{theorem}
-The complete translation is summarised in Theorem~\ref{fullFOtranslation}.
-
-\begin{example}
-	Consider the double negation shift (DNS)
-	$$\varphi = \forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$$
-	which is a famous non-intuitionistic tautology. We get
-	\begin{align*}
-		\mathcal S = & \{(P_{\forall x\neg\neg A(x)}\to P_{\neg\neg\forall xA(x)})\to P_\varphi, P_{\forall x\neg\neg A(x)}\to \forall xP_{\neg\neg A(x)}(x)\}\cup\\
-		& \{\forall t((P_{\neg\neg A(x)}(t)\wedge P_{\neg A(x)}(t))\to \bot), \forall t((A(t)\to \bot)\to P_{\neg A(x)})(t)\}\cup\\
-		& \{(P_{\neg\forall xA(x)}\to \bot)\to P_{\neg\neg\forall xA(x)}, (P_{\neg\forall xA(x)}\wedge P_{\forall xA(x)})\to \bot\}\cup\\
-		&\{\forall xA(x)\to P_{\forall xA(x)}\}
-	\end{align*}
-	and thus, denoting $\psi_1 = (P_{\neg\forall xA(x)}\to \bot)\to P_{\neg\neg\forall xA(x)}$ and $\psi_2 = \forall xA(x)\to P_{\forall xA(x)}$, we have
-	\begin{align*}
-		\mathcal S^\# = & \{\forall u(P_{\forall x\neg\neg A(x)}(u)\to P_{\neg\neg\forall xA(x)}(u))\to P_\varphi(u)\}\cup\\& \{\forall u(P_{\forall x\neg\neg A(x)}(u)\to \forall xP_{\neg\neg A(x)}(x, u)\}\cup\\
-		& \{\forall t, u((P_{\neg\neg A(x)}(t, u)\wedge P_{\neg A(x)}(t, u))\to \bot)\}\\& \{\forall t, u((A(t, u)\to \bot)\to P_{\neg A(x)}(t, u))\}\cup\\
-		& \{\forall u((P_{\neg\forall xA(x)}(f_{\psi_1}(u))\to \bot)\to P_{\neg\neg\forall xA(x)}(u))\}\\& \{\forall u((P_{\neg\forall xA(x)}(u)\wedge P_{\forall xA(x)}(u))\to \bot)\}\cup\\
-		&\{\forall u(\forall xA(x, f_{\psi_2}(u))\to P_{\forall xA(x)}(u))\}
-	\end{align*}
-\end{example}
+However Lemma~\ref{thm:fo-countermodel-reduction} still enables us to add some additional axioms which could help guide proof search.
+In particular we may assume that $f_\psi(w) = w$ if $\exists u(w = f_\psi(u))$ for $\psi\in\mathcal F_\forall$, i.e.
+$f_\psi$ is idempotent, and  $f_\psi(w) = w$ if $\exists u(w \geq f_\psi(u))$ for each $\psi\in\mathcal F_\to$, which is even stronger than idempotence.
+How adding these properties as assumption helps with proof search has to be examined.
+We have left them out of our translation for now for the sake of simplicity.
 

### 2022-07-15 CSL/lipics-v2021-EIICL.tex
pass over intro finished

@@
 
-Constructive mathematics refers to a flavour of mathematics in which existence of an object can only be established by explicit construction, as opposed to classical mathematics where existence can be shown implicitly, e.g.
+Constructive mathematics refers to a flavour of mathematics in which the existence of an object can only be established by explicit construction, as opposed to classical mathematics where existence can be shown implicitly, e.g.
 by assuming non-existence and deriving a contradiction.
-Besides the philosophical issues that one may take with classical mathematics, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that intuitionistic proofs correspond directly to computer programs, which is expressed in the Curry-Howard-Isomorphism~\cite{howard1980formulae}.
 The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the fact that the law of excluded middle $A\vee\neg A$ and the double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
-
-Comparing benchmarks for classical and intuitionistic automated theorem proving, e.g.
-TPTP~\cite{casc} and ILTP~\cite{iltp}, we see that classical provers currently have a far better success rate.
-Furthermore all the most popular first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover} utilize classical logic.
+Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in famous Curry-Howard correspondence~\cite{howard1980formulae}.
+
+The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmarks problems (see e.g.
+the
+ILTP library website~\cite{iltp}).
+However, progress in automated theorem proving for intuitionistic logic has been slow whereas reasoners for classical logic have made tremendous progress, see e.g.
+the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
+%, e.g., the first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}
+%Comparing benchmarks for classical and intuitionistic automated theorem proving, e.g.
+TPTP~\cite{casc} and , we see that classical provers currently have a far better success rate.
+Furthermore all the most popular first-order utilize classical logic.
 This difference can be partially explained by fundamental differences between the logics.
-First of all determining intuitionistic validity is computationally harder, i.e.
+First of all, determining intuitionistic validity is computationally harder, i.e.
 in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
-A further advantage of classical logic is the existence of calculi that are particularly suitable for automation like superposition~\cite{bachmair2001resolution} which rely on the existence of convenient normal forms such as CNF and the duality between validity and satisfiability, i.e.
-to show validity they show unsatisfiability of the negation.
-While some kind of normal form is of course be obtainable for intuitionistic logic, albeit more complex, crucially the duality does not hold, e.g.
-$\neg(A\vee \neg A)$ is not satisfiable while $A\vee\neg A$ is also not a theorem.
+A further advantage of classical logic is the existence of calculi that are particularly suitable for automation such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability, i.e.
+to show the validity of a formula it suffices to show the unsatisfiability of the negated formula.
+While some (albeit more complex) normal forms also exist  for intuitionistic logic, crucially the duality between validity and satisfiability of the negation does not hold.
 Therefore most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient, tammet1996resolution} use the naive inverse method, i.e.
-directly constructing a cut-free proof from the theorem by applying the inverse of rules from some proof calculus, enhanced by search strategies such as focussing and polarization.
+directly search for a cut-free proof by applying the rules from some proof calculus inversely, enhanced by search strategies such as focussing and polarization.
 This approach generally leads to a much more complex search and is therefore difficult to implement efficiently.
-Finally in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, this is particularly true for the propositional case, i.e.
+Finally we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, in particular for the propositional case, i.e.
 SAT-solvers.
 
-With this work we want to open the path for leveraging the progress that has been made in classical theorem proving for intuitionistic logic.
-To that end we give an embedding from intuitionistic logic into classical logic.
-That is we give an effective procedure that gives for each formula $\varphi$ we a modified formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
-Interestingly the reverse direction, the famous double negation translation, is a long established translation that goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case and G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
+With this work we want to open a path for leveraging the progress in classical theorem proving for intuitionistic logic.
+To this end we give an embedding from intuitionistic logic into classical logic.
+That is, we give an effective procedure that returns for each formula $\varphi$ a modified formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
+Interestingly, the reverse direction, the famous double-negation translation, has long been established and goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
 In the propositional case it is particularly simple: $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
-Intuitively the translation collapses for each subformula $\psi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent.
+Intuitively the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent.
 This gives us a first idea why the reverse direction is perhaps more difficult: We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e.
 if they both occur in $\varphi$ we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$.
-In particular this necessitates the introduction of new propositional variables, which already marks a big difference to the double negation translation.
-
+In particular, this necessitates the introduction of new propositional variables in our procedure, which already marks a big difference to the double negation translation.
+%
 In addition to the translation of formulas we give an effective translation of counter-models, i.e.
-for each intuitionistic counter-model for $\varphi$ we give a classical counter-model for $\varphi^\#$ and vice versa.
-In fact the existence of counter-models is a key notion.
-For classical logic it just corresponds to the satisfiability of the negation but for intuitionistic logic it forms a proper dual to validity, while satisfiability of the negation is not necessary for invalidity and would therefore not form a proper dual.
+for each intuitionistic counter-model of $\varphi$ we can effectively construct a classical counter-model of $\varphi^\#$ and vice versa.
+We note that the existence of counter-models is a key notion, which forms a proper dual to validity --- whereas the satisfiability of the negation is not necessary for invalidity (in contrast to classical logic, where it is a proper dual to validity).
 Transforming and reducing counter-models to a normal form is also what ultimately enables our translation.
 
-Finally we should mention that approaches to obtain constructive content from classical proofs exist in literature, however they are insufficient in our view.
-The first is simply to examine under which circumstances classical validity entails intuitionistic validity and therefore constructiveness.
-Most famously this holds for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}.
-There are also other circumstances under which this entailment holds~\cite{schwichtenberg}, however this approach cannot encapsulate the general case.
-Another popular approach is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}.
-In this case we necessarily lose realizability at least of $\exists$ and $\vee$, which - one could argue - is why we want a program in the first place.
-
-Our work suggests a new approach for obtaining a constructive proof using a classical prover, that is
+In summary, our work suggests a new approach for obtaining a constructive proof using a classical prover, that is
 \begin{itemize}
@@
 \end{itemize}
+
+\textbf{Related Work.}
+Several approaches to obtain constructive content from classical proofs have been studied in the literature.
+An important direction is to examine under which circumstances classical validity entails intuitionistic validity and therefore constructiveness.
+This entailment has most famously been established for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}.
+Further circumstances under which this entailment holds have been identified in~\cite{schwichtenberg}, however this approach cannot encapsulate the general case.\todo{can you elaborate a few words on ``further circumstances'' and ``the general case''}
+Another popular approach is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}.
+In this case we necessarily lose realizability at least of $\exists$ and $\vee$, which --- one could argue --- is why we want a program in the first place.\todo{as the curry-howard correspondence is only mentioned in passing now, we should adapt the last part of sentence}
 

### 2022-07-15 CSL/lipics-v2021-EIICL.tex
pass over overview finished

@@
 
-We shall now give an overview of the main results and highlight the key arguments.
+We shall now give an overview of our main results and highlight the key arguments.
 Recall that our goal is to give an effective translation procedure that, for a given formula $\varphi$, yields a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
-We start with the predicate case.
-The key arguments are the same as in the first-order case while it is technically simpler and less cluttered and therefore more clear.
-
-Before performing the main transformation we propose a preprocessing to a simpler form.
-A popular pre-processing step in classical automated reasoning is the Tseytin transformation~\cite{tseitin1983complexity}.
-It gives an equisatisfiable sentence in conjunctive normal form.
-Eliminating all implications is not possible in intuitionistic logic since $A\to B$ is not equivalent to $\neg A\vee B$.
+We start with the propositional case.
+The key arguments are the same as in the first-order case, while it is technically simpler and less cluttered.
+
+Before our main transformation we employ a preprocessing step akin to the Tseytin transformation~\cite{tseitin1983complexity}, which is a popular pre-processing step in classical automated reasoning: 
+It gives an equisatisfiable sentence (over an extended set of propositions) in conjunctive normal form.
+Eliminating all implications, however, is not possible in intuitionistic logic since $A\to B$ is not equivalent to $\neg A\vee B$. 
 Still we propose a similar transformation. 
-A nice feature will be that the non-classical content will be encapsulated in formulas of type $(A\to B)\to C$, i.e.
-if there are no such formulas in the transformed formula then the classical validity of the original formula immediately implies intuitionistic validity.
+A notable feature of our transformation is that all non-classical content is encapsulated in formulas of type $(A\to B)\to C$, i.e.
+if there are no such formulas in the transformed formula then the classical validity of the transformed formula immediately implies intuitionistic validity.
 
 \begin{theorem}
-	For every propositional formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set of $\mathcal S$ of formulas  of one of the forms
-	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$for atomic $A, B, C$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
-The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
+For every propositional formula $\varphi$ there effectively are an atom $P$ and a set of $\mathcal S$ of formulas (over an extended set of propositions) of one of the forms\\
+$\text{ \hspace{1cm}} A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$\\
+for atomic $A, B, C$, such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
+The time complexity of the transformation is linear in the input size.
 \end{theorem}
 
-The main transformation then proceeds in three steps.
-
-First we encode as a first-order sentence that the considered formula holds in every Kripke frame.
-Since Kripke frames for propositional logic over a fixed alphabet of propostional variables form a first-order theory this is rather straightforward.
-
-Next we eliminate some quantifiers via Herbrandization. The only quantifiers eliminated occur in formulas derived from formulas of type $(A\to B)\to C$.
-
-This leaves us with a formula with only weak $\forall$-quantifiers.
-We can then argue that if there is a counter-model for the resulting formula then there is a counter-model with a certain Kripke frame completely determined by the number of formulas of type $(A\to B)\to C$.
+The main transformation then proceeds in three steps:
+%
+1) We encode as a first-order sentence that the considered formula holds for every Kripke frame. 
+Since Kripke frames for propositional logic over a fixed alphabet of propositional variables form a first-order theory this step is rather straightforward.
+%
+2) Next we eliminate some quantifiers via Herbrandization. The only quantifiers eliminated occur in formulas derived from formulas of type $(A\to B)\to C$.
+%
+3) We can then argue that if there is a counter-model for the resulting formula then there is a counter-model with a certain Kripke frame of bounded size completely determined by the number of formulas of type $(A\to B)\to C$.
 This allows us to eliminate the remaining quantifiers by simply enumerating the worlds in that Kripke model.
 
-All in all this leaves us with the following:
+Summarizing, we obtain the following result:
 
 \begin{theorem}
+\label{thm:reduction-propositional}
 	Let $\mathcal S$ be as above.
-Let $\mathcal F_\to\subseteq\mathcal S$ denote formulas $\psi$ of the form $(A_\psi\to B_\psi)\to C_\psi$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
+Let $\mathcal F_\to\subseteq\mathcal S$ denote formulas $\psi$ of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
 For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
-Obtain $\mathcal S^\#$ by including formulas as follows
+Obtain $\mathcal S^\#$ by including the following formulas:
 	\begin{itemize}
@@
 	\end{itemize}
-	Then $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
+Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
 The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$.
-Futhermore there is an effective procedure for translating counter-models between the sentences.
+Futhermore, there is an effective procedure for translating counter-models between the sentences.
 \end{theorem}
 
-Examining the proof of the above theorem we will be able to notice that due to its tree form the constructed Kripke frame lends itself as an interpretation for a certain quantified boolean formula.
-We can show that this QBF is satisfiable if and only if there exists a counter-model for our original formula.
+Instead of explicitly enumerating the worlds of a counter-example of bounded size (as in the reduction stated in the above theorem), we can also do this enumeration implicitly with a quantified boolean formula (QBF).
+We can show that this QBF is satisfiable if and only if there exists a counter-model for our original formula:
 
@@
 
-We then move to the first-order case.
-Here things are a bit more involved but the underlying approach is the same.
-First we again perform a Tseytin-like transformation.
+%\begin{theorem}
+%	For every predicate formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set $\mathcal S$ of formulas of the forms
+%	$$\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),$$$$\forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),
+%	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)),$$$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),$$$$ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))$$for atomic $A, B, C$ and vectors of variables $\vec z, \vec a, \vec b, \vec c$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
+The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
+%\end{theorem}
+
+We then move to the first-order case. Here the construction is a bit more involved but the underlying approach is the same.
+Again we first perform a Tseytin-like transformation.
 Here the non-classical content is encapsulated by formulas of type $(A\to B)\to C$ and $\forall xA\to B$.
-
-\begin{theorem}
-	For every predicate formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set $\mathcal S$ of formulas of the forms
-	$$\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),$$$$\forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),
-	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)),$$$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),$$$$ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))$$for atomic $A, B, C$ and vectors of variables $\vec z, \vec a, \vec b, \vec c$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
-The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
-\end{theorem}
-
-We then proceed quite similarly to the propositional case.
+The details of this translation are given in Section~\ref{section:tseytin}.
+We then proceed similarly to the propositional case. 
 Note however that one difficulty arises from the fact that our (classical) domain will now contain on the one hand worlds in the Kripke frame but on the other hand also proper domain elements. 
@@
 In particular $E(x, u)$ entails that $x$ represents a domain element and $u$ a world.
-
-The Herbrandization is completely analogous the the propositional case.
-
-The real difference arises in the final step: While we can also establish a universal Kripke Frame that is parametrized only by the formulas of type $(A\to B)\to C$ and $\forall xA\to B$, it is possibly countably infinite.
+The main transformation then again proceeds in three steps:
+Step 1) and 2) are analogous to the propositional case.
+The real difference arises in step 3): 
+While we can also establish the existence of a canonic counter-model whose size only depends on formulas of type $(A\to B)\to C$ and $\forall xA\to B$, the size of this counter-model is countably infinite in general.
 This is not surprising since certain intuitionistically invalid formulas like the double negation shift $\forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$ don't have finite counter-models. 
-Therefore we are not able to eliminate the weak $\forall$-quantifiers associated with the Kripke semantics.
-However since we started in first-order logic to begin with this is perhaps not a big deal.
+Therefore we are not able to eliminate the $\forall$-quantifiers associated with the Kripke semantics. 
+However since our translation targets first-order logic, we believe that the introduction of these quantifiers is acceptable. 
+%
+Summarizing, we obtain the following result:
 
 \begin{theorem}
+\label{thm:reduction-first-order-short}
 There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
@@
 \end{theorem}
-The complete translation is stated in Theorem~\ref{fullFOtranslation}.
+A more detailed account of Theorem~\ref{thm:reduction-first-order-short} is given in Theorem~\ref{fullFOtranslation}.
 
@@
 \section{Equivalid normalization}
+\label{section:tseytin}
+
+\todo{find better name for section, as the normalization works for equisatsifiability as well as for equivalidity}
 

### 2022-07-15 CSL/lipics-v2021-EIICL.tex
pass over conclusion finished

@@
 
-We have presented a complete embedding of intuitionistic into classical logic.
-The transformation saw an exponential blow-up parametrized by $|\mathcal F_\to|$ in the propositional case and a complexity increase reflected by arity-increase of all relations and introduction of new function symbols in the predicate case.
-A key motivation for our work is the possibility of leveraging classical provers for showing intuitionistic validity.
-The practical feasibility of our translation is yet to be tested.
-In particular in the first-order case it is difficult to estimate how the increased complexity of the output formula impacts proof search.
+We have presented embeddings of intuitionistic into classical logic for the propositional and the predicate case.
+The transformation saw an exponential blow-up parameterized by $|\mathcal F_\to|$ in the propositional case and a complexity increase reflected by an arity-increase of one for all relations and the introduction of new function symbols in the predicate case.
+A key motivation for our work is the potential of leveraging classical provers for showing intuitionistic validity.
+%The practical feasibility of our translation is yet to be tested. 
+%In particular in the first-order case it is difficult to estimate how the increased complexity of the output formula impacts proof search.
 Initial tests show promise but we plan on conducting a more thorough evaluation. 
-A key factor that is not yet clear is how large the parameters $|\mathcal F_\to|$ in the propositional case and $|\mathcal F|$ in the predicate case respectively are in practice.
-
-The complexity increase fell directly out of our very straightforward construction.
-We plan to establish better bounds in future work by more intelligently utilizing structural properties of the formula, in particular by using that certain atoms are independent from each other it is possible to construct a smaller counter-example $\mathcal M_T$ in many cases.
-Besides improving the practical feasibility of our approach we hope that this will allow us to also give a new translation from QBF to IPC respecting complexity considerations and shine more light on the link between intuitionistic logic and the polynomial hierarchy.
-
-We hope that this work will provide a foundation for a new practical approach to automated reasoning for intuitionistic logic, a problem that has not seen much progress in recent times.
+A key factor yet to be determined is how large the parameters $|\mathcal F_\to|$ in the propositional case and $|\mathcal F|$ in the predicate case are in practice.
+
+The above mentioned complexity considerations are direct consequences of our straightforward constructions.
+We plan to establish better bounds in future work by utilizing structural properties of the input formula, in particular by constructing smaller counter-examples $\mathcal M_T$ in case certain atoms are independent from each other.
+On the theoretical side, we also hope to give a new translation from QBF to IPC that improves our understanding of the relationship between intuitionistic propositional logic and the polynomial hierarchy.
+
+%We hope that this work will provide a foundation for a new practical approach to automated reasoning for intuitionistic logic, a problem that has not seen much progress in recent times.
+\todo{can you add a few sentences on proof translations, which is an important goal because it allows program extraction using the curry-howard correspondence?}
 

### 2022-07-15 CSL/lipics-v2021-EIICL.tex
Space

--- 2022-07-15 CSL/lipics-v2021-EIICL.tex
@@
 
+\usepackage{wrapfig}
 \usepackage{tikz-qtree}
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:prop-translation}
 	Let $\mathcal S$ be as above.
@@
 	For every predicate formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set $\mathcal S$ of formulas of the forms
-	$$\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),$$$$\forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),
-	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)),$$$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),$$$$ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))$$for atomic $A, B, C$ and vectors of variables $\vec z, \vec a, \vec b, \vec c$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
+	$$\begin{matrix}
+		\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),\forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),\\
+		\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)), \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),\\ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))
+	\end{matrix}$$
+	for atomic $A, B, C$ and vectors of variables $\vec z, \vec a, \vec b, \vec c$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
 The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
@@
 	\begin{itemize}
-		\item $v\not\models \bot$
-		\item $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
+		\item $v\not\models \bot$ and $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
 		\item $v\models \varphi\wedge\psi$ iff $v\models\varphi$ and $v\models\psi$.
@@
 	\begin{itemize}
-		\item $\mathcal M, v\not\models\bot$.
-		\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$.
+		\item $\mathcal M, v\not\models\bot$ and  $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$.
 		\item $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
@@
 	\begin{itemize}
-		\item $u, v\not\models\bot$
-		\item $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$.
+		\item $u, v\not\models\bot$ and $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$.
 		\item $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
@@
 
-\begin{example}
-	Consider the law of excluded middle $$\varphi = (A\to \bot)\vee A$$ which is not intuitionistically valid. Then,
-	\begin{align*}
-		\varphi^{C} &= K(\varphi)\to ((A\to \bot)\vee A)^u\\
-		&= K(\varphi)\to (A\to \bot)^u\vee A^u\\
-		&= K(\varphi)\to (\forall u'( u\preceq u' \to A^{u'}\to \bot^{u'})\vee A(u))\\
-		&= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot)\vee A(u)),
-	\end{align*}
-and indeed $\varphi^{C}$ admits a counter model with domain $\{u, u'\}$ where $u \preceq u'$, $A(u) = 0, A(u') = 1$.
-\end{example}
-
 \section{Equivalid normalization}
@@
 
-\begin{lemma}The following are true for both classical and intuitionistic semantics:
-	\begin{itemize}
-		\item $\bigwedge S^+(\varphi)\wedge P_\varphi\models\varphi$.
-		\item $\bigwedge S^-(\varphi)\wedge \varphi\models P_\varphi$.
-	\end{itemize}
+\begin{lemma}
+	$\bigwedge S^+(\varphi)\wedge P_\varphi\to\varphi$, $\bigwedge S^-(\varphi)\wedge \varphi\to P_\varphi$ hold classically and intuitionistically.
 \end{lemma}
 
-\begin{proof}
-The claims are verified in a straightforward manner by induction on $\varphi$.
-\end{proof}
-
-\begin{example}
-	Consider $\varphi = \neg\forall x A(x)\to \exists x\neg A(x)$. Then
-	\begin{align*}
-		\mathcal S^+(\varphi) = &\{(P_\varphi\wedge P_{\neg\forall x A(x)})\to P_{\exists x\neg A(x)}, (P_{\forall xA(x)}\to\bot)\to P_{\neg \forall xA(x)}\}\\&\cup\{\forall xP_{A(x)}(x)\to P_{\forall xA(x)}, \forall x(P_{A(x)}(x)\to A(x))\}\\&\cup\{P_{\exists x\neg A(x)}\to \exists xP_{\neg A(x)}(x), \forall x(P_{\neg A(x)}(x)\to P_{A(x)}(x)\to \bot)\}\\&\cup\{\forall x(A(x)\to P_{A(a)}(x))\}\\
-		\mathcal S^-(\varphi) = &\{(P_{\neg\forall x A(x)}\to P_{\exists x\neg A(x)})\to P_\varphi,  (P_{\neg \forall xA(x)}\wedge P_{\forall xA(x)})\to\bot\}\\&\cup\{P_{\forall xA(x)}\to \forall xP_{A(x)}(x), \forall x(A(x)\to P_{A(x)}(x))\}\\&\cup\{\exists xP_{\neg A(x)}(x)\to P_{\exists x\neg A(x)}, \forall x((P_{A(x)}(x)\to \bot)\to P_{\neg A(x)}(x))\}\\&\cup\{\forall x(P_{A(a)}(x)\to A(x))\}
-	\end{align*}
-\end{example}
-
-We can give a transformation of structures that corresponds to the above transformation of sentences.
+We give a transformation of structures that corresponds to the above.
 
@@
 In the classical case have
-	$$P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)$$
-	and
-	$$P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))$$$$ P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})$$
-	$$P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}$$$$ P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\}$$
-	\\
-	and for intuitionistic logic and some world $u$
-	$$P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)$$
-	and
-	$$P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))$$$$ P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}$$
-	$$P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}$$$$ P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}$$
-	\\
+	$$\begin{matrix}
+		P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)\indent P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))\\P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})\\P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}\indent P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\}
+	\end{matrix}$$
+	and for intuitionistic logic and each world $u$
+	$$\begin{matrix}
+		P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)\indent P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))\\
+		 P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}\\
+		 P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}\indent P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}
+	\end{matrix}$$
 \end{definition}
@@
 	For every formula $\varphi$ and structure $\mathcal M$
-	$$\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$$
+	$\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$
 \end{lemma}
@@
 	For every structure $\mathcal M$
-	$$\mathcal M\models \varphi\text{ iff }S(\mathcal M, \varphi)\models P_\varphi$$
+	$\mathcal M\models \varphi$ if and only if $S(\mathcal M, \varphi)\models P_\varphi$
 \end{lemma}
@@
 where $u$ is some new free variable and each $\psi\in\mathcal S$ is of one of the forms
-$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B(k)\wedge C(k))), \forall k(u\preceq k\to (A(k)\wedge (B(k))\to C(k)))$$
-$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B(k)\vee C(k))), \forall k(u\preceq k\to (A(k)\vee B(k))\to C(k))$$
-$$\forall k(u\preceq k\to (\forall l(k\preceq l \to A(l)\to B(l))\to C(k)))$$
+$$\begin{matrix}
+	\forall k(u\preceq k\to A(k)\to (B(k)\wedge C(k)))\indent  \forall k(u\preceq k\to (A(k)\wedge (B(k))\to C(k)))\\
+	\forall k(u\preceq k\to A(k)\to (B(k)\vee C(k)))\indent \forall k(u\preceq k\to (A(k)\vee B(k))\to C(k))\\
+	\forall k(u\preceq k\to (\forall l(k\preceq l \to A(l)\to B(l))\to C(k)))
+\end{matrix}$$
 with each $P(x)$ possibly being $\bot$. Applying Herbrandization we obtain a formula
@@
 where $s$ is some new constant term and each $\psi\in\mathcal S'$ is of one of the forms
-$$\hspace*{-2mm}\forall k(s\preceq k\to A(k)\to (B(k)\wedge C(k))), \forall k(s\preceq k\to (A(k)\wedge (B(k))\to C(k)))$$
-$$\hspace*{-2mm}\forall k(s\preceq k\to A(k)\to (B(k)\vee C(k))), \forall k(s\preceq k\to (A(k)\vee (B(k))\to C(k)))$$
-$$\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A(f_\psi(k))\to B(f_\psi(k)))\to C(k)),$$
-and $f_\psi$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model then we have a counter-model $(M',I')$ where $$M' = \{m\in M|s\preceq m\}$$ and
-$$f_\psi^{I'}(u) = \begin{cases}
-	f^I_\psi(u), &\text{ iff $u\preceq f^{I}_\psi(u)$,}\\
-	u,&\text{ else.}
-\end{cases}$$ that is where $s$ is a least element and the $f_\psi$ increase their argument.
+$$\begin{matrix}
+	\forall k(s\preceq k\to A(k)\to (B(k)\wedge C(k)))\indent \forall k(s\preceq k\to (A(k)\wedge (B(k))\to C(k)))\\
+	\forall k(s\preceq k\to A(k)\to (B(k)\vee C(k)))\indent \forall k(s\preceq k\to (A(k)\vee (B(k))\to C(k)))\\
+	\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A(f_\psi(k))\to B(f_\psi(k)))\to C(k))\\
+\end{matrix}$$
+and $f_\psi$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model then we have a counter-model $(M',I')$ where $M' = \{m\in M|s\preceq m\}$ and
+$f_\psi^{I'}(u) = f^I_\psi(u) \text{ iff $u\preceq f^{I}_\psi(u)$ and $f_\psi^{I'}(u) = u$ else.}$ that is where $s$ is a least element and the $f_\psi$ increase their argument.
 Therefore instead of $\varphi'''$ we may consider a formula $$\varphi^\circ = \forall k(s\preceq k)\to \bigwedge\{\forall k(k\preceq f_\psi(k))\:|\:\psi\in\mathcal F_\to\}\to K(\varphi')\to \bigwedge \mathcal S^\circ\to P_\varphi(s)$$ where each formula $\psi\in\mathcal S^\circ$ is of one of the forms
-$$\forall k(A(k)\to (B(k)\wedge C(k))), \forall k((A(k)\wedge B(k))\to C(k))$$
-$$\forall k(A(k)\to (B(k)\vee C(k))), \forall k((A(k)\vee B_\psi(k))\to C(k))$$
-$$\forall k((A(f_\psi(k))\to B(f_\psi(k)))\to C(k))$$
+$$\begin{matrix}
+	\forall k(A(k)\to (B(k)\wedge C(k)))\indent\forall k((A(k)\wedge B(k))\to C(k))\\
+	\forall k(A(k)\to (B(k)\vee C(k)))\indent\forall k((A(k)\vee B_\psi(k))\to C(k))\\
+	\forall k((A(f_\psi(k))\to B(f_\psi(k)))\to C(k))
+\end{matrix}$$
 and $\mathcal F_\to$ denotes the set of formulas of the last kind.
@@
 
-\begin{corollary}
+\begin{corollary}\label{cor:prop-tree-model}
 	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
 \end{corollary}
+
+An examples of this translation can be found in the appendix.
 
@@
 
+So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
+	x, &\text{ if $\psi$ occurs in $x$,}\\
+	x, \psi&\text{ otherwise.}
+\end{cases}$$ Now we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula.
+This gives us the translation stated in Theorem~\ref{thm:prop-translation}.
+\begin{theorem}
+	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
+\end{theorem}
+
 \begin{example}
-	Consider
+	Let us consider the law of excluded middle $\varphi = A\vee\neg A$. Then we get
+	$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\}$$
+	and we know that $\varphi$ is valid iff $\bigwedge \mathcal S(\varphi)\to P_\varphi$ is valid.
+Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically valid, where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
+	\begin{align*}
+		\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi}, (P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi}\}\cup\\ &\{A^\epsilon\to P_A^\epsilon, P_A^{\psi}\to P_A^{\psi}, (A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}
+	\end{align*}
+	and $P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$, $P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$ defines a counter-model.
+\end{example}
+
+\section{IPC to QBF}
+
+It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}.
+However, giving an explicit translation from IPC to QBF still promises to be useful, allowing to leverage the progress in QBF solving.
+The translation from the last section will serve as starting point for developing the translation in this section.
+For lack of space, we do not define the semantics of QBF in this paper and refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
+
+Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
+and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$.
+We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model.
+The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
+The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e.
+express which path is considered.
+
+\begin{definition}
+	For every propositional variable $X$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $X^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$.
+Let $\vec X^n$ range over all propositional variables $X^n$ and $\vec F^n$ range over all $F_\psi^n$.
+Define
+	\begin{itemize}
+		\item {\normalfont Valid}$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
+if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
+		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
+		\item {\normalfont Sat}$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
+		\item {\normalfont Sat}$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
+	\end{itemize}
+	Have $k = |\mathcal F_\to|$ and define$$\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal S}(i)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
+	for $0 < i < k$ as well as the special cases
+	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
+	for leafs and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
+	Example encodings of the above formulas would be
+	\begin{align*}
+		\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
+		\text{\normalfont Persistent}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
+		\text{\normalfont Sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
+		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
+		\text{\normalfont Sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
+	\end{align*}
+\end{definition}
+
+\begin{example}
+	With the above encodings for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
+	$$\hspace*{-.3cm}
+	\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right)
+	$$
+	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $F^0$.
+\end{example}
+
+\begin{lemma}
+	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
+\end{lemma}
+\begin{proof}
+Suppose $\varphi$ is not intuitionistically valid i.e.
+it $\bigwedge\mathcal S\to P$ has an intuitionistic counter-model.
+Then by the precious section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
+For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$.
+Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.
+
+	Let the $F^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified).
+If not exactly one is interpreted as true then valid$(n)$ fails, i.e.
+have $\psi_n\in\mathcal F_\to$ the only element with $F^n_{\psi_n} = 1$.
+If $F^i_{\psi_n} = 1$ for some $i < n$ then valid$(n)$ also fails, so we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions.
+Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{\psi_1\dots\psi_n}$.
+Continue this construction until $n  = |\mathcal F_\to|$.
+Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.
+
+	On the other hand suppose $\varphi^Q$ is satisfiable. We construct a counter-example to $\varphi^\#$.
+	Again we proceed iteratively.
+Interpret $A^\epsilon$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$.
+Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$ in which $F^i_\psi$ is chosen true iff $\psi = \psi_i$.
+Let $\psi_n\in\mathcal F_\to$.
+Consider some interpretation of the $A^n$ that are part of a satisfying assignment where $F^n_\psi$ is true iff $\psi = \psi_n$ and all variables quantified above are chosen as before.
+Have $A^{\psi_0\dots\psi_n} = A^n$ for each propositional variable $A$.
+From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.
+\end{proof}
+
+A simple counting argument shows that for each $n < |\mathcal F_\to|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal F_\to|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal S}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal F_\to}(n)|\in \mathcal O(|\mathcal F_\to|)$ so overall we get
+\begin{lemma}
+	The size of $\varphi^Q$ is in $\mathcal O(|\varphi|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$.
+\end{lemma}
+
+Note that the last set of universally quantified variables can be avoided because there is only one assignment that has the chance to falsify $\varphi^Q$, namely the assignment that assigns $1$ to the single $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$.
+Hence, in our final formula we can replace every $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^n$.
+With that we get the following:
+
+\begin{corollary}
+	Let $N$-\textbf{IntInvalid} be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid.
+Then $N$-\textbf{IntInvalid} is in $\Sigma_{2N-1}^P$.
+The dual problem $N$-\textbf{IntValid} is in $\Pi_{2N-1}^P$.
+\end{corollary}
+
+\section{CQC to IQC}
+
+We now give an analogous transformation for first order logic.
+However our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic.
+However we do it nonetheless!
+The domain of our classical model will feature on the one hand elements representing worlds in a Kripke frame and on the other hand the domain-elements from the Kripke model.
+We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$.
+We fix some signature $\Sigma$ for this section.
+To encode that some $n$-ary relation $A$ holds at a world $u$ we replace each $n$-ary relation symbol $A$ with a $n+1$-ary relation symbol $A^\#$, interpreting $A^\#(\vec x, u)$ as "$A(\vec x)$ holds at u".
+Furthermore we write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$.
+
+First of all recall that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
+\[\begin{matrix}
+	\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c)))\indent \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c))\indent \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c)))\\
+	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c))\indent \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\indent\forall \vec z(\forall xA(\vec a)\to B(\vec b))\\ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))
+\end{matrix}\]
+The complete result that we will show is as follows:
+\begin{theorem}\label{fullFOtranslation}
+	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain the formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$.
+Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
+For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$.
+For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$.
+Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
+Obtain $\mathcal S^\#$ by including formulas as follows:
+	\begin{itemize}
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
+		\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
+	\end{itemize}
+	Then $\bigwedge S\to P$ is intuitionistically iff $\exists xE(x, s)\to K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol, $\Sigma$ is the original signature and
+	\begin{align*}
+		K := &\:\bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
+		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
+		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}
+	\end{align*}
+	The size of the obtained formula is linear in the input.
+However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
+\end{theorem}
+
+It is not possible to completely eliminate worlds from the domain as in the propositional case, since counter-models possibly must be infinite. We will see why that is in more detail later.
+
+\subsection{Encoding Kripke Semantics}
+
+As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
+\begin{itemize}
+	\item PartialOrder($\preceq$) encoding that $\preceq$ is a partial order.
+	\item DomainSubset$(u, w)$ encoding that the domain at $u$ is a subset of that at $w$.
+	\item World($u$) encoding that $u$ represents a world.
+	\item DomainClosed($u$) which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
+	\item Persistent$(u, w)$ which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$ at $u$ we have $A(x)$ at $w$.
+\end{itemize}
+Then analogously to the predicate case define
+\begin{align*}
+	K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \\
+				 & \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
+				 & \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge\\
+				 & \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
+\end{align*}
+Example encodings could be
+\begin{align*}
+	\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
+	\text{DomainSubset}(u, w) = &\:\forall z(E(z, u)\to E(z, w))\\
+	\text{World}(u) = &\:\exists xE(x, u)\\
+	\text{DomainClosed}(u) = &\:\bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f\in\Sigma$ is a function symbol}\}\\
+	\text{Pesistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
+\end{align*}
+
+Using $E$ we then obtain $\mathcal S^\circ$ analogously to the predicate case by including formulas:
+
+\begin{itemize}
+	\item $\forall \vec z\forall u(s\preceq u\to \vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
+	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
+	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to\forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$\\ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
+	\item  $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
+	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
+	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
+	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
+\end{itemize}
+
+Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid iff
+$$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
+is classically valid where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced, so let's present a proof.
+
+\begin{proof}
+	We proceed by translation of counter-examples.
+Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi^\circ$.
+As a Kripke frame $(W, \leq)$ take all $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ let $\leq$ be $\preceq^I$ restricted to $W$.
+Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A^\#(\vec x, u)$.
+It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
+
+	The other direction is a bit more involved.
+Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
+In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
+Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
+Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
+Now have
+	\begin{itemize}
+		\item $s^I = w_0$.
+		\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
+		\item $f^I(m_1\dots m_n) =\begin{cases}
+			 f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
+			 w_0, & \text{otherwise.}
+		\end{cases}$
+		\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
+			A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
+			\top, & \text{otherwise.}
+		\end{cases}$
+	\end{itemize}
+	One easily verifies that these are well-defined.
+Let us now briefly check that this indeed defines a counter-model.
+First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is.
+Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified.
+All that remains to show is $M, I\models\bigwedge\mathcal S^\circ$.
+Consider e.g.
+the case where $\psi\in\mathcal S$ is of the form
+	$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
+	We have to show that
+	$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
+	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
+	$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false.
+For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$.
+Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$.
+Note that the above is false in particular for $w\in M_0$ with $u\preceq w$.
+But this would imply
+	$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
+\end{proof}
+
+\subsection{Reducing the encoding}
+
+As in the predicate case we now show the existence of a canonical counter-model, if a counter-model exists.
+Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$.
+Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
+For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$.
+As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
+$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$ with
+$$\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$$ and
+$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$$ with $$\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$$
+and in each other formula in $\mathcal S^\circ$ removing the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
+$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#(s)$$
+As in the previous section we can define a tree counter-model for this.
+
+\begin{definition}
+	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
+	\begin{itemize}
+		\item $M_T$ are the sequences on $\{(\vec z, \psi)\:|\:\psi\in \mathcal F, \vec z\in M^n\}$.
+		\item Interpret $\preceq$ as the prefix-order.
+		\item Have $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi)$$
+		\item For propositional variables $P$ have $$P^{I_T}((\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = P^I(f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots))$$
+	\end{itemize}
+\end{definition}
+
+\begin{lemma}
+	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
+\end{lemma}
+
+Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec z_1, f_{\psi_2}(\vec z_2, \dots f_{\psi_n}(\vec z_n, s)\dots))$$ is sufficient.
+In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
+from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
+Define
+
+$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#$$
+then we have
+\begin{theorem}
+	\label{thm:fo-reduction}
+	$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
+\end{theorem}
+The complete translation is summarised in Theorem~\ref{fullFOtranslation}. Let us understand why a further reduction as in the propositional case is impossible.
+
+\begin{definition}
+\label{def:fo-fulfilled}
+	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
+For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
+	For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
+\end{definition}
+
+\begin{lemma}
+\label{thm:fo-countermodel-reduction}
+	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
+	\begin{enumerate}
+		\item $\psi$ is fulfilled at $f_\psi^I(\vec a, u)$ with $\vec a$ for all $u\in M, \psi\in\mathcal F$, $\vec a\in M^n$.
+		\item If $\psi\in\mathcal F_\to$ is fulfilled at $u\in M$ with $\vec a$ then $\psi$ is fulfilled at all $w\geq u$ with $\vec a$.
+	\end{enumerate}	
+\end{lemma}
+
+\begin{proof}
+	For $\psi\in\mathcal F_\to$ the proof is analogous to the previous section.
+
+	1.
+Let $\psi\in\mathcal F_\forall$.
+If $B^I_\psi(\vec b, u)[\vec a/\vec t]$ holds then we are done due to persistency.
+Otherwise there is $a$ with $E^I(a, f^I_\psi(\vec a, u))$ such that $A^{I[a/x]}_\psi(\vec t_A, f^I_\psi(\vec a, u))$ is false.
+So $\psi$ is fulfilled at $f^I_\psi(\vec a, u)$ with $\vec a$.
+\end{proof}
+
+However Part 2 of this Lemma fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$) we have $A(x, v)$ there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
+This is why we must consider sequences that have repetitions:
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\begin{center}
+		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
+			\node[squarednode]      (0)   [] {$ $};
+			\node[squarednode]      (1r)   [above right=of 0] {{\color{blue}$1$}};
+			\node[squarednode]      (1l)   [above left=of 0] {{\color{red}$1$}};
+			\node[squarednode]      (2r)   [above=of 1r] {{1, \color{blue}$2$}};
+			\node[squarednode]      (2l)   [above=of 1l] {{1, \color{red}$2$}};
+			\node[squarednode]      (3r)   [above=of 2r] {{1, 2, \color{blue}$3$}};
+			\node[squarednode]      (3l)   [above=of 2l] {{1, 2, \color{red}$3$}};
+			\node      (4r)   [above=of 3r] {{$\vdots$}};
+			\node      (4l)   [above=of 3l] {{$\vdots$}};
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (0.north east) -- (1r.south west);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (0.north west) -- (1l.south east);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1r.north) -- (2r.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1l.north) -- (2l.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1l.north east) -- (2r.south west);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1r.north west) -- (2l.south east);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2r.north) -- (3r.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2l.north) -- (3l.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2l.north east) -- (3r.south west);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2r.north west) -- (3l.south east);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3r.north) -- (4r.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3l.north) -- (4l.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3l.north east) -- (4r.south west);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3r.north west) -- (4l.south east);
+		\end{tikzpicture}
+	\end{center}
+\end{wrapfigure}
+
+\begin{example}
+	Consider the $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model
+
+	where at each node the elements that exist at it are indicated.
+For black ones both $A$ and $B$ hold, for red ones only $B$, for blue ones only $A$.
+Furthermore red arrows indicate $f_{\neg \forall xA(x)}$ and blue arrows $f_{\neg \forall xB(x)}$.
+Note that applying $f_{\neg \forall xA(x)}$ we must always reach the left column and $f_{\neg \forall xB(x)}$ we must always reach the right column.
+Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we can never remain stationary.
+This is in contrast to the propositional case.
+\end{example}
+
+However Lemma~\ref{thm:fo-countermodel-reduction} still enables us to add some additional axioms which could help guide proof search.
+In particular we may assume that $f_\psi(w) = w$ if $\exists u(w = f_\psi(u))$ for $\psi\in\mathcal F_\forall$, i.e.
+$f_\psi$ is idempotent, and  $f_\psi(w) = w$ if $\exists u(w \geq f_\psi(u))$ for each $\psi\in\mathcal F_\to$, which is even stronger than idempotence.
+How adding these properties as assumption helps with proof search has to be examined.
+We have left them out of our translation for now for the sake of simplicity.
+
+\section{Conclusion}
+
+We have presented a complete embedding of intuitionistic into classical logic.
+The transformation saw an exponential blow-up parametrized by $|\mathcal F_\to|$ in the propositional case and a complexity increase reflected by arity-increase of all relations and introduction of new function symbols in the predicate case.
+A key motivation for our work is the possibility of leveraging classical provers for showing intuitionistic validity.
+The practical feasibility of our translation is yet to be tested.
+In particular in the first-order case it is difficult to estimate how the increased complexity of the output formula impacts proof search.
+Initial tests show promise but we plan on conducting a more thorough evaluation.
+A key factor that is not yet clear is how large the parameters $|\mathcal F_\to|$ in the propositional case and $|\mathcal F|$ in the predicate case respectively are in practice.
+
+The complexity increase fell directly out of our very straightforward construction.
+We plan to establish better bounds in future work by more intelligently utilizing structural properties of the formula, in particular by using that certain atoms are independent from each other it is possible to construct a smaller counter-example $\mathcal M_T$ in many cases.
+Besides improving the practical feasibility of our approach we hope that this will allow us to also give a new translation from QBF to IPC respecting complexity considerations and shine more light on the link between intuitionistic logic and the polynomial hierarchy.
+
+We hope that this work will provide a foundation for a new practical approach to automated reasoning for intuitionistic logic, a problem that has not seen much progress in recent times.
+
+\bibliography{lipics-v2021-EIICL}
+
+\appendix
+
+\section{Omitted Proofs}
+
+\todo{the lemmata still use the set `T', so the notation is probably not consistent anymore}
+
+\begin{lemma}\label{ap1}
+	Let $\chi$ be a formula and $\{t_1\dots t_n\} = T$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
+	\begin{itemize}
+		\item if $\mathcal M, v\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
+		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
+	\end{itemize}
+\end{lemma}
+
+\begin{proof}
+	We proceed by simultaneous induction on the formula height.
+
+	For atomic $\chi$ the claims are clear.
+
+	Otherwise there are $3$ cases:
+
+	1.
+$\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$.
+By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v}), \mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as in the lemma.
+Then we can define $I_{\chi, v}$ as follows:
+
+	For every symbol $p$ that occurs in $\varphi^S$ or $\varphi^H$ but not $\psi^S, \psi^H$ have $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$.
+For every symbol $p$ that occurs in $\psi^S$ or $\psi^H$ but not $\varphi^S, \varphi^H$ have $p^{I_{\chi, v}} = p^{I_{\psi, v}}$.
+Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.
+
+	Let $\circ=\to$.
+Suppose $\mathcal M, v\models\chi$.
+Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_T$ or $\mathcal M_\chi\models\psi^H_T$, i.e.
+$\mathcal M_\chi\models\chi^S_T$.
+On the other hand suppose $\mathcal M, v\not\models\chi$.
+Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_T$ and $\mathcal M_\chi, v\not\models\psi^S_T$, i.e.
+$\mathcal M_\chi\not\models\chi^H_T$.
+Analogous arguments work for $\circ\in\{\wedge, \vee\}$.
+
+	2.
+$\chi = \forall x\varphi$.
+Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise.
+Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.
+
+	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
+Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
+
+	Suppose $\mathcal M, v\models\chi$.
+Then for all $m\in M$ $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{M\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{M\cup\{a\}}[x/a])$, i.e.
+$\mathcal M_\chi,v\models \chi^S$.
+On the other hand suppose $\mathcal M, v\not\models\chi$.
+Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{T\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(t_1,\dots t_n)/x]^H_T$, i.e.
+$\mathcal M_\chi, v\not\models(\forall x\varphi)^H_T$.
+
+	3.
+$\chi = \exists x\varphi$.
+The argument runs dually to 2.
+Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise.
+Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.
+
+	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
+Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
+
+	Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_T$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_T$.
+\end{proof}
+
+\begin{lemma}\label{ap2}
+	For every structure $\mathcal M$ and $\{t_1\dots t_n\} = T$ that contains all free variables in $\chi$ and variable assignment $v$
+	\begin{itemize}
+		\item if $\mathcal M, v\models\varphi^S_T$ then $\mathcal M, v\models \varphi$
+		\item if $\mathcal M, v\not\models\varphi^H_T$ then $\mathcal M, v\not\models\varphi$.
+	\end{itemize}
+\end{lemma}
+
+\begin{proof}
+	Again we proceed by simultaneous induction on the formula height.
+
+	For atoms the claims are clear. We distinguish 5 cases.
+
+	1.
+$\chi = \varphi\to\psi$.
+Suppose $\mathcal M, v\models\chi^S_T$, i.e.
+$\mathcal M, v\not\models\varphi^H_T$ or $\mathcal M, v\models\psi^S_T$.
+By induction hypothesis $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e.
+$\mathcal M, v\models \chi$.
+On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e.
+$\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_T$.
+Again by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.
+
+	2. + 3. Conjunctions and Disjunctions are dealt with analogously.
+
+	4.
+$\chi = \forall x\varphi$.
+Suppose $\mathcal M, v\models\chi^S_T$, i.e.
+for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{T\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$.
+Then it follows that $\mathcal M, v\models\varphi$.
+On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e.
+there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{T\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e.
+$\mathcal M, v\not\models\forall x\chi$.
+
+	5. $\chi = \exists x\varphi$.  The argument runs dually to 4.
+\end{proof}
+
+\begin{example}
+	This is an example for the model translation from Corollary~\ref{cor:prop-tree-model}. Consider
 	$$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model, depicted below,
@@
 
-So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
-	x, &\text{ if $\psi$ occurs in $x$,}\\
-	x, \psi&\text{ otherwise.}
-\end{cases}$$ Now we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula.
-This gives us the following translation:
-
-\begin{definition}
-	Suppose $\mathcal S$ is a set of formulas $\psi$ of the form
-	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C),(A\vee B)\to C, (A\to B)\to C$$
-	Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$.
-For each atom $A$ occurring in $\mathcal S$ and $k\in \mathcal F_\to$, consider a new atom $A^{k}$.
-Obtain $\mathcal S^\#$ by including the formulas
-	\begin{itemize}
-		\item $A^k\to A^{k\psi}$ for each $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
-		\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $(A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occur in $k$.
-	\end{itemize}
-\end{definition}
-
-\begin{theorem}
-	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
-\end{theorem}
-
-\begin{example}
-	Let us consider the law of excluded middle $\varphi = A\vee\neg A$. Then we get
-	$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\}$$
-	and we know that $\varphi$ is valid iff $\bigwedge \mathcal S(\varphi)\to P_\varphi$ is valid.
-Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically valid, where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
-	\begin{align*}
-		\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi}\}\\ &\cup\{(P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi},  A^\epsilon\to P_A^\epsilon\}\\ &\cup \{P_A^{\psi}\to P_A^{\psi}, (A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}
-	\end{align*}
-	It is easy to construct a counter example:
-	$$P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$$
-	$$P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$$
-\end{example}
-
-\section{IPC to QBF}
-
-It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}.
-However, giving an explicit translation from IPC to QBF still promises to be useful, allowing to leverage the progress in QBF solving.
-The translation from the last section will serve as starting point for developing the translation in this section.
-For lack of space, we do not define the semantics of QBF in this paper and refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
-
-Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
-and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$.
-We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model.
-The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
-The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e.
-express which path is considered.
-
-\begin{definition}
-	For every propositional variable $X$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $X^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$.
-Let $\vec X^n$ range over all propositional variables $X^n$ and $\vec F^n$ range over all $F_\psi^n$.
-Define
-	\begin{itemize}
-		\item {\normalfont Valid}$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
-if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
-		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
-		\item {\normalfont Sat}$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
-		\item {\normalfont Sat}$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
-	\end{itemize}
-	Have $k = |\mathcal F_\to|$ and define$$\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal S}(i)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
-	for $0 < i < k$ as well as the special cases
-	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
-	for leafs and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
-	Example encodings of the above formulas would be
-	\begin{align*}
-		\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
-		\text{\normalfont Persistent}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
-		\text{\normalfont Sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
-		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
-		\text{\normalfont Sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
-	\end{align*}
-\end{definition}
-
-\begin{example}
-	With the above encodings for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
-	$$\hspace*{-.3cm}
-	\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right)
-	$$
-	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $F^0$.
-\end{example}
-
-\begin{lemma}
-	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
-\end{lemma}
-\begin{proof}
-\todo{The proof contains some old notation at several places.}
-\todo{Also the proof could be beautified a bit.}
-Suppose $\varphi$ is not intuitionistically valid, i.e. there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
-For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$.
-Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.
-
-	Let the $F^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified).
-If not exactly one is interpreted as true then valid$(n)$ fails, i.e.
-have $\psi_n\in\mathcal F_\to$ the only element with $F^n_{\psi_n} = 1$.
-If $F^i_{\psi_n} = 1$ for some $i < n$ then valid$(n)$ also fails, so we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions.
-Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{\psi_1\dots\psi_n}$.
-Continue this construction until $n  = |\mathcal F_\to|$.
-Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.
-
-	On the other hand suppose $\varphi^Q$ is satisfiable. We construct a counter-example to $\varphi^\#$.
-	Again we proceed iteratively.
-Interpret $A^\epsilon$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$.
-Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$ in which $F^i_\psi$ is chosen true iff $\psi = \psi_i$.
-Let $\psi_n\in\mathcal F_\to$.
-Consider some interpretation of the $A^n$ that are part of a satisfying assignment where $F^n_\psi$ is true iff $\psi = \psi_n$ and all variables quantified above are chosen as before.
-Have $A^{\psi_0\dots\psi_n} = A^n$ for each propositional variable $A$.
-From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.
-\end{proof}
-
-\begin{example}
-	Consider the formula $\varphi = (((A\to B)\to C)\wedge((B\to A)\to C))\to C$, which is a non-intuitionistic tautology.
-Its associated QBF is $$\varphi^Q = \exists A^0B^0C^0\forall F_{\psi_1}^1F_{\psi_2}^1 \exists A^1B^1C^1\forall F_{\psi_1}^2F_{\psi_2}^2 \exists A^2B^2C^2\psi'$$where $\psi_1 := (A\to B)\to C, \psi_2 = (B\to A)\to C$ and $$\psi' = \bigwedge\left\{\begin{matrix}
-		F^1_{\psi_1}\vee F^1_{\psi_2}, \neg(F^1_{\psi_1}\wedge F^1_{\psi_2}), F^2_{\psi_1}\vee F^2_{\psi_2},\neg(F^2_{\psi_1}\wedge F^2_{\psi_2}),\\F^1_{\psi_1}\to\neg F^2_{\psi_1},F^1_{\psi_2}\to\neg F^2_{\psi_2}, A^0\to A^1,A^1\to A^2,
-		\\B^0\to B^1, B^1\to B^2, C^0\to C^1, C^1\to C^2,\neg C^0, \\
-		F^1_{\psi_1}\to (A^1\to B^1)\to C^0, F^1_{\psi_2}\to(B^1\to A^1)\to C^0\\
-		F^2_{\psi_1}\to (A^2\to B^2)\to C^1, F^2_{\psi_2}\to(B^2\to A^2)\to C^1
-	\end{matrix}\right\}$$
-	This QBF is satisfiable: Have $P^0 = 0$ for all atoms $P$, always $P^2 = 1$ for all atoms $P$ and $C^1 = 1$.
-Furthermore have $A^1 = 1$ iff $F^1_{\psi_1} = 1$ and $B^1 = 1$ iff $F^1_{\psi_2} = 1$.
-The construction of an counter-model to $\varphi^\#$ then proceeds as follows:
-	\begin{itemize}
-		\item First have $P^\epsilon = 0$ for each propositional variable $P$.
-		\item For $P^{\psi_1}$ we consider what happens when we choose $F^1_{\psi_1}$ to be true, i.e. we have $A^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
-		\item For $P^{\psi_2}$ we consider what happens when we choose $F^1_{\psi_2}$ to be true, i.e. we have $B^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
-		\item Finally for $P^{\psi_1\psi_2} = P^{\psi_2\psi_1} = 1$ for all propositional variables $P$ as $P^2$ is true regardless of choice for the $F^i_\psi$.
-	\end{itemize}
-\end{example}
-
-A simple counting argument shows that for each $n < |\mathcal F_\to|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal F_\to|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal S}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal F_\to}(n)|\in \mathcal O(|\mathcal F_\to|)$ so overall we get
-\begin{lemma}
-	The size of $\varphi^Q$ is in $\mathcal O(|\varphi|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$.
-\end{lemma}
-
-Note that the last set of universally quantified variables can be avoided because there is only one assignment that has the chance to falsify $\varphi^Q$, namely the assignment that assigns $1$ to the single $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$.
-Hence, in our final formula we can replace every $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^n$.
-With that we get the following:
-
-\begin{corollary}
-	Let $N$-\textbf{IntInvalid} be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid.
-Then $N$-\textbf{IntInvalid} is in $\Sigma_{2N-1}^P$.
-The dual problem $N$-\textbf{IntValid} is in $\Pi_{2N-1}^P$.
-\end{corollary}
-
-\section{CQC to IQC}
-
-We now give an analogous transformation for first order logic.
-However our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic.
-However we do it nonetheless!
-The domain of our classical model will feature on the one hand elements representing worlds in a Kripke frame and on the other hand the domain-elements from the Kripke model.
-We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$.
-We fix some signature $\Sigma$ for this section.
-To encode that some $n$-ary relation $A$ holds at a world $u$ we replace each $n$-ary relation symbol $A$ with a $n+1$-ary relation symbol $A^\#$, interpreting $A^\#(\vec x, u)$ as "$A(\vec x)$ holds at u".
-Furthermore we write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$.
-
-First of all recall that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
-
-$$\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),$$$$ \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),
-\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)),$$$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),$$$$ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))$$
-
-The complete result that we will show is as follows:
-\begin{theorem}\label{fullFOtranslation}
-	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain the formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$.
-Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
-For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$.
-For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$.
-Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
-Obtain $\mathcal S^\#$ by including formulas as follows:
-	\begin{itemize}
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
-		\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
-	\end{itemize}
-	Then $\bigwedge S\to P$ is intuitionistically iff $\exists xE(x, s)\to K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol, $\Sigma$ is the original signature and
-	\begin{align*}
-		K := &\:\bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
-		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
-		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}
-	\end{align*}
-	The size of the obtained formula is linear in the input.
-However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
-\end{theorem}
-
-It is not possible to completely eliminate worlds from the domain as in the propositional case, since counter-models possibly must be infinite. We will see why that is in more detail later.
-
-\subsection{Encoding Kripke Semantics}
-
-As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
-\begin{itemize}
-	\item PartialOrder($\preceq$) encoding that $\preceq$ is a partial order.
-	\item DomainSubset$(u, w)$ encoding that the domain at $u$ is a subset of that at $w$.
-	\item World($u$) encoding that $u$ represents a world.
-	\item DomainClosed($u$) which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
-	\item Persistent$(u, w)$ which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$ at $u$ we have $A(x)$ at $w$.
-\end{itemize}
-Then analogously to the predicate case define
-\begin{align*}
-	K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \\
-				 & \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
-				 & \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge\\
-				 & \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
-\end{align*}
-Example encodings could be
-\begin{align*}
-	\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
-	\text{DomainSubset}(u, w) = &\:\forall z(E(z, u)\to E(z, w))\\
-	\text{World}(u) = &\:\exists xE(x, u)\\
-	\text{DomainClosed}(u) = &\:\bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f\in\Sigma$ is a function symbol}\}\\
-	\text{Pesistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
-\end{align*}
-
-Using $E$ we then obtain $\mathcal S^\circ$ analogously to the predicate case by including formulas:
-
-\begin{itemize}
-	\item $\forall \vec z\forall u(s\preceq u\to \vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
-	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
-	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to\forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$\\ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
-	\item  $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
-	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
-	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
-	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
-\end{itemize}
-
-Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid iff
-$$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
-is classically valid where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced, so let's present a proof.
-
-\begin{proof}
-	We proceed by translation of counter-examples.
-Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi^\circ$.
-As a Kripke frame $(W, \leq)$ take all $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ let $\leq$ be $\preceq^I$ restricted to $W$.
-Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A^\#(\vec x, u)$.
-It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
-
-	The other direction is a bit more involved.
-Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
-In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
-Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
-Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
-Now have
-	\begin{itemize}
-		\item $s^I = w_0$.
-		\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
-		\item $f^I(m_1\dots m_n) =\begin{cases}
-			 f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
-			 w_0, & \text{otherwise.}
-		\end{cases}$
-		\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
-			A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
-			\top, & \text{otherwise.}
-		\end{cases}$
-	\end{itemize}
-	One easily verifies that these are well-defined.
-Let us now briefly check that this indeed defines a counter-model.
-First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is.
-Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified.
-All that remains to show is $M, I\models\bigwedge\mathcal S^\circ$.
-Consider e.g.
-the case where $\psi\in\mathcal S$ is of the form
-	$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
-	We have to show that
-	$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
-	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
-	$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false.
-For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$.
-Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$.
-Note that the above is false in particular for $w\in M_0$ with $u\preceq w$.
-But this would imply
-	$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
-\end{proof}
-
-\subsection{Reducing the encoding}
-
-As in the predicate case we now show the existence of a canonical counter-model, if a counter-model exists.
-Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$.
-Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
-For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$.
-As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
-$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$ with
-$$\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$$ and
-$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$$ with $$\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$$
-and in each other formula in $\mathcal S^\circ$ removing the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
-$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#(s)$$
-As in the previous section we can define a tree counter-model for this.
-
-\begin{definition}
-	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
-	\begin{itemize}
-		\item $M_T$ are the sequences on $\{(\vec z, \psi)\:|\:\psi\in \mathcal F, \vec z\in M^n\}$.
-		\item Interpret $\preceq$ as the prefix-order.
-		\item Have $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi)$$
-		\item For propositional variables $P$ have $$P^{I_T}((\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = P^I(f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots))$$
-	\end{itemize}
-\end{definition}
-
-\begin{lemma}
-	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
-\end{lemma}
-
-Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec z_1, f_{\psi_2}(\vec z_2, \dots f_{\psi_n}(\vec z_n, s)\dots))$$ is sufficient.
-In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
-from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
-Define
-
-$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#$$
-then we have
-\begin{theorem}
-	\label{thm:fo-reduction}
-	$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
-\end{theorem}
-The complete translation is summarised in Theorem~\ref{fullFOtranslation}. Let us understand why a further reduction as in the propositional case is impossible.
-
-\begin{definition}
-\label{def:fo-fulfilled}
-	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
-For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
-	For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
-\end{definition}
-
-\begin{lemma}
-\label{thm:fo-countermodel-reduction}
-	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
-	\begin{enumerate}
-		\item $\psi$ is fulfilled at $f_\psi^I(\vec a, u)$ with $\vec a$ for all $u\in M, \psi\in\mathcal F$, $\vec a\in M^n$.
-		\item If $\psi\in\mathcal F_\to$ is fulfilled at $u\in M$ with $\vec a$ then $\psi$ is fulfilled at all $w\geq u$ with $\vec a$.
-	\end{enumerate}	
-\end{lemma}
-
-\begin{proof}
-	For $\psi\in\mathcal F_\to$ the proof is analogous to the previous section.
-
-	1.
-Let $\psi\in\mathcal F_\forall$.
-If $B^I_\psi(\vec b, u)[\vec a/\vec t]$ holds then we are done due to persistency.
-Otherwise there is $a$ with $E^I(a, f^I_\psi(\vec a, u))$ such that $A^{I[a/x]}_\psi(\vec t_A, f^I_\psi(\vec a, u))$ is false.
-So $\psi$ is fulfilled at $f^I_\psi(\vec a, u)$ with $\vec a$.
-\end{proof}
-
-However Part 2 of this Lemma fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$) we have $A(x, v)$ there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
-This is why we must consider sequences that have repetitions:
-
-\begin{example}
-	Consider the $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model
-	\begin{center}
-		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
-			\node[squarednode]      (0)   [] {$ $};
-			\node[squarednode]      (1r)   [above right=of 0] {{\color{blue}$1$}};
-			\node[squarednode]      (1l)   [above left=of 0] {{\color{red}$1$}};
-			\node[squarednode]      (2r)   [above=of 1r] {{1, \color{blue}$2$}};
-			\node[squarednode]      (2l)   [above=of 1l] {{1, \color{red}$2$}};
-			\node[squarednode]      (3r)   [above=of 2r] {{1, 2, \color{blue}$3$}};
-			\node[squarednode]      (3l)   [above=of 2l] {{1, 2, \color{red}$3$}};
-			\node      (4r)   [above=of 3r] {{$\vdots$}};
-			\node      (4l)   [above=of 3l] {{$\vdots$}};
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (0.north east) -- (1r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (0.north west) -- (1l.south east);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1r.north) -- (2r.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1l.north) -- (2l.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1l.north east) -- (2r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1r.north west) -- (2l.south east);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2r.north) -- (3r.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2l.north) -- (3l.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2l.north east) -- (3r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2r.north west) -- (3l.south east);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3r.north) -- (4r.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3l.north) -- (4l.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3l.north east) -- (4r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3r.north west) -- (4l.south east);
-		\end{tikzpicture}
-	\end{center}
-	where at each node the elements that exist at it are indicated.
-For black ones both $A$ and $B$ hold, for red ones only $B$, for blue ones only $A$.
-Furthermore red arrows indicate $f_{\neg \forall xA(x)}$ and blue arrows $f_{\neg \forall xB(x)}$.
-Note that applying $f_{\neg \forall xA(x)}$ we must always reach the left column and $f_{\neg \forall xB(x)}$ we must always reach the right column.
-Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we can never remain stationary.
-This is in contrast to the propositional case.
-\end{example}
-
-However Lemma~\ref{thm:fo-countermodel-reduction} still enables us to add some additional axioms which could help guide proof search.
-In particular we may assume that $f_\psi(w) = w$ if $\exists u(w = f_\psi(u))$ for $\psi\in\mathcal F_\forall$, i.e.
-$f_\psi$ is idempotent, and  $f_\psi(w) = w$ if $\exists u(w \geq f_\psi(u))$ for each $\psi\in\mathcal F_\to$, which is even stronger than idempotence.
-How adding these properties as assumption helps with proof search has to be examined.
-We have left them out of our translation for now for the sake of simplicity.
-
-\section{Conclusion}
-
-We have presented a complete embedding of intuitionistic into classical logic.
-The transformation saw an exponential blow-up parametrized by $|\mathcal F_\to|$ in the propositional case and a complexity increase reflected by arity-increase of all relations and introduction of new function symbols in the predicate case.
-A key motivation for our work is the possibility of leveraging classical provers for showing intuitionistic validity.
-The practical feasibility of our translation is yet to be tested.
-In particular in the first-order case it is difficult to estimate how the increased complexity of the output formula impacts proof search.
-Initial tests show promise but we plan on conducting a more thorough evaluation.
-A key factor that is not yet clear is how large the parameters $|\mathcal F_\to|$ in the propositional case and $|\mathcal F|$ in the predicate case respectively are in practice.
-
-The complexity increase fell directly out of our very straightforward construction.
-We plan to establish better bounds in future work by more intelligently utilizing structural properties of the formula, in particular by using that certain atoms are independent from each other it is possible to construct a smaller counter-example $\mathcal M_T$ in many cases.
-Besides improving the practical feasibility of our approach we hope that this will allow us to also give a new translation from QBF to IPC respecting complexity considerations and shine more light on the link between intuitionistic logic and the polynomial hierarchy.
-
-We hope that this work will provide a foundation for a new practical approach to automated reasoning for intuitionistic logic, a problem that has not seen much progress in recent times.
-
-\bibliography{lipics-v2021-EIICL}
-
-\appendix
-
-\section{Omitted Proofs}
-
-\todo{the lemmata still use the set `T', so the notation is probably not consistent anymore}
-
-\begin{lemma}\label{ap1}
-	Let $\chi$ be a formula and $\{t_1\dots t_n\} = T$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
-	\begin{itemize}
-		\item if $\mathcal M, v\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
-		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
-	\end{itemize}
-\end{lemma}
-
-\begin{proof}
-	We proceed by simultaneous induction on the formula height.
-
-	For atomic $\chi$ the claims are clear.
-
-	Otherwise there are $3$ cases:
-
-	1.
-$\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$.
-By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v}), \mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as in the lemma.
-Then we can define $I_{\chi, v}$ as follows:
-
-	For every symbol $p$ that occurs in $\varphi^S$ or $\varphi^H$ but not $\psi^S, \psi^H$ have $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$.
-For every symbol $p$ that occurs in $\psi^S$ or $\psi^H$ but not $\varphi^S, \varphi^H$ have $p^{I_{\chi, v}} = p^{I_{\psi, v}}$.
-Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.
-
-	Let $\circ=\to$.
-Suppose $\mathcal M, v\models\chi$.
-Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_T$ or $\mathcal M_\chi\models\psi^H_T$, i.e.
-$\mathcal M_\chi\models\chi^S_T$.
-On the other hand suppose $\mathcal M, v\not\models\chi$.
-Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_T$ and $\mathcal M_\chi, v\not\models\psi^S_T$, i.e.
-$\mathcal M_\chi\not\models\chi^H_T$.
-Analogous arguments work for $\circ\in\{\wedge, \vee\}$.
-
-	2.
-$\chi = \forall x\varphi$.
-Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise.
-Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.
-
-	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
-Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
-
-	Suppose $\mathcal M, v\models\chi$.
-Then for all $m\in M$ $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{M\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{M\cup\{a\}}[x/a])$, i.e.
-$\mathcal M_\chi,v\models \chi^S$.
-On the other hand suppose $\mathcal M, v\not\models\chi$.
-Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{T\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(t_1,\dots t_n)/x]^H_T$, i.e.
-$\mathcal M_\chi, v\not\models(\forall x\varphi)^H_T$.
-
-	3.
-$\chi = \exists x\varphi$.
-The argument runs dually to 2.
-Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise.
-Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.
-
-	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
-Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
-
-	Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_T$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_T$.
-\end{proof}
-
-\begin{lemma}\label{ap2}
-	For every structure $\mathcal M$ and $\{t_1\dots t_n\} = T$ that contains all free variables in $\chi$ and variable assignment $v$
-	\begin{itemize}
-		\item if $\mathcal M, v\models\varphi^S_T$ then $\mathcal M, v\models \varphi$
-		\item if $\mathcal M, v\not\models\varphi^H_T$ then $\mathcal M, v\not\models\varphi$.
-	\end{itemize}
-\end{lemma}
-
-\begin{proof}
-	Again we proceed by simultaneous induction on the formula height.
-
-	For atoms the claims are clear. We distinguish 5 cases.
-
-	1.
-$\chi = \varphi\to\psi$.
-Suppose $\mathcal M, v\models\chi^S_T$, i.e.
-$\mathcal M, v\not\models\varphi^H_T$ or $\mathcal M, v\models\psi^S_T$.
-By induction hypothesis $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e.
-$\mathcal M, v\models \chi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e.
-$\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_T$.
-Again by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.
-
-	2. + 3. Conjunctions and Disjunctions are dealt with analogously.
-
-	4.
-$\chi = \forall x\varphi$.
-Suppose $\mathcal M, v\models\chi^S_T$, i.e.
-for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{T\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$.
-Then it follows that $\mathcal M, v\models\varphi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e.
-there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{T\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e.
-$\mathcal M, v\not\models\forall x\chi$.
-
-	5. $\chi = \exists x\varphi$.  The argument runs dually to 4.
-\end{proof}
-
 \end{document}

### 2022-07-15 CSL/lipics-v2021-EIICL.tex

@@
 
-We shall now give an overview of the main results and highlight the key arguments.
+We shall now give an overview of our main results and highlight the key arguments.
 Recall that our goal is to give an effective translation procedure that, for a given formula $\varphi$, yields a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
-We start with the predicate case.
-The key arguments are the same as in the first-order case while it is technically simpler and less cluttered and therefore more clear.
-
-Before performing the main transformation we propose a preprocessing to a simpler form.
-A popular pre-processing step in classical automated reasoning is the Tseytin transformation~\cite{tseitin1983complexity}.
-It gives an equisatisfiable sentence in conjunctive normal form.
-Eliminating all implications is not possible in intuitionistic logic since $A\to B$ is not equivalent to $\neg A\vee B$.
+We start with the propositional case.
+The key arguments are the same as in the first-order case, while it is technically simpler and less cluttered.
+
+Before our main transformation we employ a preprocessing step akin to the Tseytin transformation~\cite{tseitin1983complexity}, which is a popular pre-processing step in classical automated reasoning: 
+It gives an equisatisfiable sentence (over an extended set of propositions) in conjunctive normal form.
+Eliminating all implications, however, is not possible in intuitionistic logic since $A\to B$ is not equivalent to $\neg A\vee B$. 
 Still we propose a similar transformation. 
-A nice feature will be that the non-classical content will be encapsulated in formulas of type $(A\to B)\to C$, i.e.
-if there are no such formulas in the transformed formula then the classical validity of the original formula immediately implies intuitionistic validity.
+A notable feature of our transformation is that all non-classical content is encapsulated in formulas of type $(A\to B)\to C$, i.e.
+if there are no such formulas in the transformed formula then the classical validity of the transformed formula immediately implies intuitionistic validity.
 
 \begin{theorem}
-	For every propositional formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set of $\mathcal S$ of formulas  of one of the forms
-	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$for atomic $A, B, C$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
-The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
+For every propositional formula $\varphi$ there effectively are an atom $P$ and a set of $\mathcal S$ of formulas (over an extended set of propositions) of one of the forms\\
+$\text{ \hspace{1cm}} A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$\\
+for atomic $A, B, C$, such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
+The time complexity of the transformation is linear in the input size.
 \end{theorem}
 
-The main transformation then proceeds in three steps.
-
-First we encode as a first-order sentence that the considered formula holds in every Kripke frame.
-Since Kripke frames for propositional logic over a fixed alphabet of propostional variables form a first-order theory this is rather straightforward.
-
-Next we eliminate some quantifiers via Herbrandization. The only quantifiers eliminated occur in formulas derived from formulas of type $(A\to B)\to C$.
-
-This leaves us with a formula with only weak $\forall$-quantifiers.
-We can then argue that if there is a counter-model for the resulting formula then there is a counter-model with a certain Kripke frame completely determined by the number of formulas of type $(A\to B)\to C$.
+The main transformation then proceeds in three steps:
+%
+1) We encode as a first-order sentence that the considered formula holds for every Kripke frame. 
+Since Kripke frames for propositional logic over a fixed alphabet of propositional variables form a first-order theory this step is rather straightforward.
+%
+2) Next we eliminate some quantifiers via Herbrandization. The only quantifiers eliminated occur in formulas derived from formulas of type $(A\to B)\to C$.
+%
+3) We can then argue that if there is a counter-model for the resulting formula then there is a counter-model with a certain Kripke frame of bounded size completely determined by the number of formulas of type $(A\to B)\to C$.
 This allows us to eliminate the remaining quantifiers by simply enumerating the worlds in that Kripke model.
 
-All in all this leaves us with the following:
-
-\begin{theorem}\label{thm:prop-translation}
+Summarizing, we obtain the following result:
+
+\begin{theorem}
+\label{thm:reduction-propositional}
 	Let $\mathcal S$ be as above.
-Let $\mathcal F_\to\subseteq\mathcal S$ denote formulas $\psi$ of the form $(A_\psi\to B_\psi)\to C_\psi$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
+Let $\mathcal F_\to\subseteq\mathcal S$ denote formulas $\psi$ of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
 For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
-Obtain $\mathcal S^\#$ by including formulas as follows
+Obtain $\mathcal S^\#$ by including the following formulas:
 	\begin{itemize}
@@
 
-We then move to the first-order case.
-Here things are a bit more involved but the underlying approach is the same.
-First we again perform a Tseytin-like transformation.
+%\begin{theorem}
+%	For every predicate formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set $\mathcal S$ of formulas of the forms
+%	$$\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),$$$$\forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),
+%	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)),$$$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),$$$$ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))$$for atomic $A, B, C$ and vectors of variables $\vec z, \vec a, \vec b, \vec c$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
+The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
+%\end{theorem}
+
+We then move to the first-order case. Here the construction is a bit more involved but the underlying approach is the same.
+Again we first perform a Tseytin-like transformation.
 Here the non-classical content is encapsulated by formulas of type $(A\to B)\to C$ and $\forall xA\to B$.
-
-\begin{theorem}
-	For every predicate formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set $\mathcal S$ of formulas of the forms
-	$$\begin{matrix}
-		\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),\forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),\\
-		\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)), \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),\\ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))
-	\end{matrix}$$
-	for atomic $A, B, C$ and vectors of variables $\vec z, \vec a, \vec b, \vec c$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
-The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
-\end{theorem}
-
-We then proceed quite similarly to the propositional case.
+The details of this translation are given in Section~\ref{section:tseytin}.
+We then proceed similarly to the propositional case. 
 Note however that one difficulty arises from the fact that our (classical) domain will now contain on the one hand worlds in the Kripke frame but on the other hand also proper domain elements. 

### 2022-07-15 CSL/lipics-v2021-EIICL.tex
space reduction, updated conclusion

@@
 This entailment has most famously been established for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}.
-Further circumstances under which this entailment holds have been identified in~\cite{schwichtenberg}, however this approach cannot encapsulate the general case.\todo{can you elaborate a few words on ``further circumstances'' and ``the general case''}
+Other very restricted subsets of classical logic have been identified under which this entailment holds, e.g. in~\cite{schwichtenberg}, however this approach is not applicable in general.
 Another popular approach is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}.
-In this case we necessarily lose realizability at least of $\exists$ and $\vee$, which --- one could argue --- is why we want a program in the first place.\todo{as the curry-howard correspondence is only mentioned in passing now, we should adapt the last part of sentence}
+In this case we necessarily lose realizability at least of $\exists$ and $\vee$ in general.
 
@@
 \end{definition}
-There are many classical tautologies which are not intuitionistically valid:
-\begin{example}\label{LEMcounterexample}
-Consider the famous law of excluded middle $A\vee\neg A$.
-Recall that $\neg A := A\to \bot$.
-As a counter-model consider a Kripke structure with two nodes $u\leq w$ and $v_u(A) = 0, v_w(A) = 1$.
-Then $u\not\models A$ and $u\not\models \neg A$, i.e.
-$u\not\models A\vee\neg A$.
-\end{example}
-In classical logic a proven strategy for showing that a formula valid is to consider its negation and show that it is not satisfiable.
-This is insufficient for intuitionistic logic, e.g.
-$A\vee\neg A$ is not an intuitionistic tautology but its negation is not satisfiable either.
+There are many classical tautologies which are not intuitionistically valid, e.g. the law of excluded middle $A\vee\neg A$. On the other hand any intuitionistic theorem is a classical one.
 
@@
 	\begin{itemize}
-		\item $\mathcal M, v\not\models\bot$ and  $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$.
-		\item $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
-		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$.
-		\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
+		\item $\mathcal M, v\not\models\bot$, $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$, $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
+		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$, $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
 		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
@@
 \begin{definition}
-	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, its (Kripke) \emph{Frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$
-	\begin{itemize}
-		\item $M_u\subseteq M_w$.
-		\item $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
-		\item $R^{I_w}|_{M_u} = R^{I_u}$.
-	\end{itemize}
+	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, its (Kripke) \emph{Frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$} and $R^{I_w}|_{M_u} = R^{I_u}$.
 	Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
 	\begin{itemize}
-		\item $u, v\not\models\bot$ and $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$.
-		\item $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
-		\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$.
-		\item $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
+		\item $u, v\not\models\bot$, $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$, $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
+		\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$, $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
 		\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
@@
 
-\section{Equivalid normalization}
+\section{Normal form translation}
 \label{section:tseytin}
-
-\todo{find better name for section, as the normalization works for equisatsifiability as well as for equivalidity}
 
@@
 
-We give a transformation of structures that corresponds to the above.
-
-\begin{definition}
-	For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms of the form $P_\psi$.
-By slight abuse of notation we denote with $\vec z_\psi$ elements of the domain instead of variables.
-In the classical case have
-	$$\begin{matrix}
-		P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)\indent P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))\\P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})\\P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}\indent P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\}
-	\end{matrix}$$
-	and for intuitionistic logic and each world $u$
-	$$\begin{matrix}
-		P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)\indent P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))\\
-		 P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}\\
-		 P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}\indent P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}
-	\end{matrix}$$
-\end{definition}
-
-The above definitions are motivated by the following results which follow directly from the involved definitions and are true both for classical and intuitionistic semantics:
-
+We give a transformation of structures that corresponds to the above, i.e.
+for each structure $\mathcal M$ and formula $\varphi$ wee can give a structure $\mathcal S(\mathcal M, \varphi)$, the exact definition of which can be found in the appendix as Definition~\ref{def:transf-structure}, such that:
 \begin{lemma}
@@
 \end{definition}
-\begin{lemma}
+\begin{lemma}~\label{thm:prop-countermodel-reduction}
 	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
@@
 	\end{enumerate}
-
-\end{lemma}
-
-\begin{proof}
-
-	1. If $C^I(u)$ is true then we are done due to persistency. Otherwise $C^I(u)$ is false.
-Because $(A^I(f_\psi^I(u))\to B^I(f_\psi^I(u)))\to C^I(u)$ holds, then $A^I(f_\psi^I(u))\to B^I(f_\psi^I(u))$ must be false.
-
-	2. Let $w$ be some element with $w\geq u$.
-If $C^I(w)$ is true, we are done.
-Otherwise $C^I(w)$ is false.
-Due to persistency $C^I(u)$ is also false.
-Because $\psi$ is fulfilled at $u$, $A^I(u)\to B^I(u)$ must be false, i.e. $A^I(u)$ is true and $B^I(u)$ is false. Then $A^I(w)$ and $A^I(f^I_\psi(w))$ are also true due to persistency.
-Because $(A^I(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I(w)$ holds, we now get that  $B^I(f^I(w))$ must be false.
-But then due to persistency so is $B^I(w)$ and we have that $\psi$ is fulfilled at $w$.
-\end{proof}
-
+\end{lemma}
+
+A proof can be found in the appendix as lemma~\ref{thm:prop-countermodel-reduction2}.
 Essentially what this tells us is that for every $u$ at which $\psi$ is fulfilled, in particular each $u$ in the image of $f_\psi$, we may add the assumption that $f_\psi(w) = w$ for all $w\geq u$ without changing the validity of $\psi$.
@@
 
-An examples of this translation can be found in the appendix.
+An example of this translation can be found in the appendix as Example~\ref{ex:prop-tree-model}.
 
@@
 \end{cases}$$ Now we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula.
-This gives us the translation stated in Theorem~\ref{thm:prop-translation}.
+This gives us the translation stated in Theorem~\ref{thm:reduction-propositional}.
 \begin{theorem}
@@
 
-Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
+Fix some set $\mathcal S$ of formulas of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
 and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$.
@@
 \end{lemma}
-\begin{proof}
-Suppose $\varphi$ is not intuitionistically valid i.e.
-it $\bigwedge\mathcal S\to P$ has an intuitionistic counter-model.
-Then by the precious section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
-For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$.
-Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.
-
-	Let the $F^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified).
-If not exactly one is interpreted as true then valid$(n)$ fails, i.e.
-have $\psi_n\in\mathcal F_\to$ the only element with $F^n_{\psi_n} = 1$.
-If $F^i_{\psi_n} = 1$ for some $i < n$ then valid$(n)$ also fails, so we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions.
-Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{\psi_1\dots\psi_n}$.
-Continue this construction until $n  = |\mathcal F_\to|$.
-Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.
-
-	On the other hand suppose $\varphi^Q$ is satisfiable. We construct a counter-example to $\varphi^\#$.
-	Again we proceed iteratively.
-Interpret $A^\epsilon$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$.
-Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$ in which $F^i_\psi$ is chosen true iff $\psi = \psi_i$.
-Let $\psi_n\in\mathcal F_\to$.
-Consider some interpretation of the $A^n$ that are part of a satisfying assignment where $F^n_\psi$ is true iff $\psi = \psi_n$ and all variables quantified above are chosen as before.
-Have $A^{\psi_0\dots\psi_n} = A^n$ for each propositional variable $A$.
-From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.
-\end{proof}
+A proof can be found in the appendix as proof~\ref{lemma:QBF}.
 
@@
 \begin{align*}
-	K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \\
-				 & \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
-				 & \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge\\
-				 & \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
+	K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
+				 & \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
 \end{align*}
@@
 
-Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid iff
+Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid if and only if
 $$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
-is classically valid where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced, so let's present a proof.
-
-\begin{proof}
-	We proceed by translation of counter-examples.
-Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi^\circ$.
-As a Kripke frame $(W, \leq)$ take all $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ let $\leq$ be $\preceq^I$ restricted to $W$.
-Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A^\#(\vec x, u)$.
-It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
-
-	The other direction is a bit more involved.
-Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
-In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
-Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
-Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
-Now have
-	\begin{itemize}
-		\item $s^I = w_0$.
-		\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
-		\item $f^I(m_1\dots m_n) =\begin{cases}
-			 f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
-			 w_0, & \text{otherwise.}
-		\end{cases}$
-		\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
-			A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
-			\top, & \text{otherwise.}
-		\end{cases}$
-	\end{itemize}
-	One easily verifies that these are well-defined.
-Let us now briefly check that this indeed defines a counter-model.
-First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is.
-Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified.
-All that remains to show is $M, I\models\bigwedge\mathcal S^\circ$.
-Consider e.g.
-the case where $\psi\in\mathcal S$ is of the form
-	$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
-	We have to show that
-	$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
-	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
-	$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false.
-For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$.
-Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$.
-Note that the above is false in particular for $w\in M_0$ with $u\preceq w$.
-But this would imply
-	$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
-\end{proof}
+is classically valid where $s$ is a new constant symbol.
+This was obvious in the propositional case but here it is more nuanced.
+A proof is given in the appendix as Lemma~\ref{lemma:fo-simplification}.
 
@@
 For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$.
-As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
-$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$ with
-$$\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$$ and
-$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$$ with $$\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$$
+As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by adding formulas
+\begin{itemize}
+	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$
+	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$
+\end{itemize}
 and in each other formula in $\mathcal S^\circ$ removing the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
@@
 	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
-For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
-	For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
-\end{definition}
+For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
+	For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
+\end{definition}
+
+We get an analogous result to lemma~\ref{thm:prop-countermodel-reduction}.
 
@@
 
-\begin{proof}
-	For $\psi\in\mathcal F_\to$ the proof is analogous to the previous section.
-
-	1.
-Let $\psi\in\mathcal F_\forall$.
-If $B^I_\psi(\vec b, u)[\vec a/\vec t]$ holds then we are done due to persistency.
-Otherwise there is $a$ with $E^I(a, f^I_\psi(\vec a, u))$ such that $A^{I[a/x]}_\psi(\vec t_A, f^I_\psi(\vec a, u))$ is false.
-So $\psi$ is fulfilled at $f^I_\psi(\vec a, u)$ with $\vec a$.
-\end{proof}
-
 However Part 2 of this Lemma fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$) we have $A(x, v)$ there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
-This is why we must consider sequences that have repetitions:
+This is why we must consider sequences that have repetitions.
 \begin{wrapfigure}{r}{0.4\textwidth}
@@
 	\end{center}
+	\caption{Counter-model to $\varphi$.}
+	\label{fig1}
 \end{wrapfigure}
@@
 \begin{example}
-	Consider the $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model
-
+	Consider $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model as shown in Figure~\ref{fig1}
 	where at each node the elements that exist at it are indicated.
@@
 How adding these properties as assumption helps with proof search has to be examined.
-We have left them out of our translation for now for the sake of simplicity.
 
@@
 
-%We hope that this work will provide a foundation for a new practical approach to automated reasoning for intuitionistic logic, a problem that has not seen much progress in recent times.
-\todo{can you add a few sentences on proof translations, which is an important goal because it allows program extraction using the curry-howard correspondence?}
+As a next step we will benchmark our translation using state-of-the art automated theorem provers, in particular the Vampire theorem prover~\cite{kovacs2013first}.
+We will have to adapt the counter-model translation to a proof translation for the particular calculus, e.g.
+superposition in the case of Vampire.
+Ultimately we hope that this will also open a new path for program extraction via the Curry-Howard Correspondence.
 
@@
 
-\section{Omitted Proofs}
-
-\todo{the lemmata still use the set `T', so the notation is probably not consistent anymore}
+\section{Omitted content}
 
 \begin{lemma}\label{ap1}
-	Let $\chi$ be a formula and $\{t_1\dots t_n\} = T$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
-	\begin{itemize}
-		\item if $\mathcal M, v\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
+	Let $\chi$ be a formula and $\{z_1\dots z_n\} = Z$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
+	\begin{itemize}
+		\item if $\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
 		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
@@
 
-\begin{example}
+\begin{definition}\label{def:transf-structure}
+	For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms of the form $P_\psi$.
+By slight abuse of notation we denote with $\vec z_\psi$ elements of the domain instead of variables.
+In the classical case have
+	$$\begin{matrix}
+		P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)\\
+		P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))\\P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})\\P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}\indent P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\}
+	\end{matrix}$$
+	and for intuitionistic logic and each world $u$
+	$$\begin{matrix}
+		P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)\\
+		P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))\\
+		P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}\\
+		P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}\indent P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}
+	\end{matrix}$$
+\end{definition}
+
+\begin{lemma}~\label{thm:prop-countermodel-reduction2}
+	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
+	\begin{enumerate}
+		\item $\psi$ is fulfilled at $f_\psi^I(u)$ for all $u\in M, \psi\in\mathcal F_\to$.
+		\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\geq u$.
+	\end{enumerate}
+\end{lemma}
+
+\begin{proof}
+	1. If $C^I(u)$ is true then we are done due to persistency. Otherwise $C^I(u)$ is false.
+	Because $(A^I(f_\psi^I(u))\to B^I(f_\psi^I(u)))\to C^I(u)$ holds, then $A^I(f_\psi^I(u))\to B^I(f_\psi^I(u))$ must be false.
+
+	2. Let $w$ be some element with $w\geq u$.
+	If $C^I(w)$ is true, we are done.
+	Otherwise $C^I(w)$ is false.
+	Due to persistency $C^I(u)$ is also false.
+	Because $\psi$ is fulfilled at $u$, $A^I(u)\to B^I(u)$ must be false, i.e. $A^I(u)$ is true and $B^I(u)$ is false. Then $A^I(w)$ and $A^I(f^I_\psi(w))$ are also true due to persistency.
+	Because $(A^I(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I(w)$ holds, we now get that  $B^I(f^I(w))$ must be false.
+	But then due to persistency so is $B^I(w)$ and we have that $\psi$ is fulfilled at $w$.
+\end{proof}
+
+\begin{example}\label{ex:prop-tree-model}
 	This is an example for the model translation from Corollary~\ref{cor:prop-tree-model}. Consider
@@
 
+\begin{lemma}\label{lemma:QBF}
+	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
+\end{lemma}
+\begin{proof}
+	Suppose $\varphi$ is not intuitionistically valid i.e.
+it $\bigwedge\mathcal S\to P$ has an intuitionistic counter-model.
+Then by the precious section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
+	For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$.
+Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.
+
+	Let the $F^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified).
+If not exactly one is interpreted as true then valid$(n)$ fails, i.e.
+have $\psi_n\in\mathcal F_\to$ the only element with $F^n_{\psi_n} = 1$.
+If $F^i_{\psi_n} = 1$ for some $i < n$ then valid$(n)$ also fails, so we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions.
+Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{\psi_1\dots\psi_n}$.
+Continue this construction until $n  = |\mathcal F_\to|$.
+Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.
+
+	On the other hand suppose $\varphi^Q$ is satisfiable. We construct a counter-example to $\varphi^\#$.
+	Again we proceed iteratively.
+Interpret $A^\epsilon$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$.
+Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$ in which $F^i_\psi$ is chosen true iff $\psi = \psi_i$.
+Let $\psi_n\in\mathcal F_\to$.
+Consider some interpretation of the $A^n$ that are part of a satisfying assignment where $F^n_\psi$ is true iff $\psi = \psi_n$ and all variables quantified above are chosen as before.
+Have $A^{\psi_0\dots\psi_n} = A^n$ for each propositional variable $A$.
+From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.
+\end{proof}
+
+\begin{lemma}\label{lemma:fo-simplification}
+	$\varphi$ is intuitionistically valid if and only of $\varphi^\circ$ is classically valid.
+\end{lemma}
+
+\begin{proof}
+	We proceed by translation of counter-examples.
+Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi^\circ$.
+As a Kripke frame $(W, \leq)$ take all $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ let $\leq$ be $\preceq^I$ restricted to $W$.
+Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A^\#(\vec x, u)$.
+It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
+
+	The other direction is a bit more involved.
+Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
+In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
+Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
+Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
+	Now have
+	\begin{itemize}
+		\item $s^I = w_0$.
+		\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
+		\item $f^I(m_1\dots m_n) =\begin{cases}
+			f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
+			w_0, & \text{otherwise.}
+		\end{cases}$
+		\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
+			A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
+			\top, & \text{otherwise.}
+		\end{cases}$
+	\end{itemize}
+	One easily verifies that these are well-defined.
+Let us now briefly check that this indeed defines a counter-model.
+First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is.
+Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified.
+All that remains to show is $M, I\models\bigwedge\mathcal S^\circ$.
+Consider e.g.
+the case where $\psi\in\mathcal S$ is of the form
+	$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
+	We have to show that
+	$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
+	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
+	$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false.
+For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$.
+Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$.
+Note that the above is false in particular for $w\in M_0$ with $u\preceq w$.
+But this would imply
+	$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
+\end{proof}
+
 \end{document}

### 2022-07-15 CSL/lipics-v2021-EIICL.tex
typos

@@
 	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
-For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
-	For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say it is \textit{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
+For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
+	For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
 \end{definition}

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
some typos

@@
 The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the fact that the law of excluded middle $A\vee\neg A$ and the double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
-Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in famous Curry-Howard correspondence~\cite{howard1980formulae}.
+Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particul<ar motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in famous Curry-Howard correspondence~\cite{howard1980formulae}.
 
@@
 ILTP library website~\cite{iltp}).
-However, progress in automated theorem proving for intuitionistic logic has been slow whereas reasoners for classical logic have made tremendous progress, see e.g.
+However, progress in automated theorem proving for intuitionistic logic has been slow, whereas reasoners for classical logic have made tremendous progress, see e.g.
 the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
@@
 This approach generally leads to a much more complex search and is therefore difficult to implement efficiently.
-Finally we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, in particular for the propositional case, i.e.
+Finally, we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, in particular for the propositional case, i.e.
 SAT-solvers.
@@
 With this work we want to open a path for leveraging the progress in classical theorem proving for intuitionistic logic.
-To this end we give an embedding from intuitionistic logic into classical logic.
-That is, we give an effective procedure that returns for each formula $\varphi$ a modified formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
+To this end, we give an embedding from intuitionistic logic into classical logic.
+That is, we give an effective procedure that returns for each formula $\varphi$ a modified formula $\varphi^\#$, such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
 Interestingly, the reverse direction, the famous double-negation translation, has long been established and goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
@@
 This gives us a first idea why the reverse direction is perhaps more difficult: We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e.
-if they both occur in $\varphi$ we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$.
+if they both occur in $\varphi$, we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$.
 In particular, this necessitates the introduction of new propositional variables in our procedure, which already marks a big difference to the double negation translation.
@@
 for each intuitionistic counter-model of $\varphi$ we can effectively construct a classical counter-model of $\varphi^\#$ and vice versa.
-We note that the existence of counter-models is a key notion, which forms a proper dual to validity --- whereas the satisfiability of the negation is not necessary for invalidity (in contrast to classical logic, where it is a proper dual to validity).
+We note that the existence of counter-models is a key notion that forms a proper dual to validity --- whereas the satisfiability of the negation is not necessary for invalidity (in contrast to classical logic, where it is a proper dual to validity).
 Transforming and reducing counter-models to a normal form is also what ultimately enables our translation.
@@
 An important direction is to examine under which circumstances classical validity entails intuitionistic validity and therefore constructiveness.
-This entailment has most famously been established for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}.
-Other very restricted subsets of classical logic have been identified under which this entailment holds, e.g. in~\cite{schwichtenberg}, however this approach is not applicable in general.
-Another popular approach is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}.
-In this case we necessarily lose realizability at least of $\exists$ and $\vee$ in general.
+This entailment has most famously been established for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}, but also other settings have been studied~\cite{schwichtenberg}.
+Another important direction is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}, which however does not permit in general the realizability of at least $\exists$ and $\vee$.
 
@@
 We shall now give an overview of our main results and highlight the key arguments.
-Recall that our goal is to give an effective translation procedure that, for a given formula $\varphi$, yields a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
+Recall that our goal is to give an effective translation procedure that, for a given formula $\varphi$, yields a formula $\varphi^\#$, such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
 We start with the propositional case.
@@
 There is an effective procedure that produces a QBF $\varphi^Q$ of size $\mathcal O(|\mathcal S|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$ with $2\cdot |\mathcal F_\to|-1$ quantifier alternations such that $\varphi$ is intuitionistically valid if and only if $\varphi^Q$ is not satisfiable.
-Fixing $N\in\mathbb N$ deciding intuitionistic validity for formulas $\bigwedge \mathcal S\to P$ where $\mathcal S$ is as above and $|\mathcal F_\to| = N$ is therefore in $\Sigma_{2N-1}$.
+For fixed $N\in\mathbb N$, deciding intuitionistic validity for formulas $\bigwedge \mathcal S\to P$ where $\mathcal S$ is as above and $|\mathcal F_\to| = N$ is therefore in $\Sigma_{2N-1}$.
 \end{theorem}
@@
 There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
-The size of $\varphi^\#$ is in linear in the size of $|\varphi|$, however for each $n$-ary relation symbol in $\varphi$ there is a corresponding $n+1$-ary relation symbol in $\varphi^\#$ and furthermore $\varphi$ contains a new binary predicate $E$ as well as a number of new function symbols.
-Furthermore there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
+The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary relation symbol in $\varphi$ there is a corresponding $n+1$-ary relation symbol in $\varphi^\#$, and $\varphi$ contains a new binary predicate $E$ as well as a number of new function symbols.
+Furthermore, there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
 \end{theorem}
@@
 
-These results provide an effective framework for using classical provers to check intuitionistic validity. First experiments with the Vampire theorem prover show promise.
+These results provide an effective framework for using classical provers to check intuitionistic validity. First experiments with the Vampire theorem prover~\cite{kovacs2013first} show promise.
 
@@
 \begin{definition}
-A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds} and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
+A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds}, called the Kripke \emph{frame}, and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
 The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, where we set
@@
 \begin{definition}
-	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, its (Kripke) \emph{Frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$} and $R^{I_w}|_{M_u} = R^{I_u}$.
+A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$} and $R^{I_w}|_{M_u} = R^{I_u}$.
 	Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
@@
 	\end{itemize}
-	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$.
-A formula is valid if it is satisfied in eveery $\Sigma$-Kripke structure, it is valid for some Frame (write $W\models\varphi$) if it is satisfied in every Krike structure with that Frame.
+	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. $\varphi$ is valid if it is satisfied by every Kripke structure.
+$\varphi$ is valid for some frame (write $W\models\varphi$) if it is satisfied in every Krike structure with that frame.
 We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
@@
 \begin{remark}
-	It is well known that Kripke Frames which are trees are complete with regard to intuitionistic logic.
+	It is well known that Kripke frames which are trees are complete with regard to intuitionistic logic.
 Our contribution is giving an explicit geometry that depends only on the structure of the considered formula.
@@
 	The other direction is a bit more involved.
-Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
+Suppose we have a Kripke counter-model to $\varphi$ with frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
 In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
proof read up to QBF section

@@
 	\begin{itemize}
-		\item $A^S_T = A^H_Z = A$ for each atomic $A$.
+		\item $A^S_Z = A^H_Z = A$ for each atomic $A$.
 		\item $(\varphi\circ\psi)^X_Z = \varphi^X_Z\circ\psi^X_Z$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
 		\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$\\$(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
-		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
+		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
 		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.
@@
 \end{definition}
-
+\todo{I think with the implicit forall quanitifcation over the free variables in our semantics, now also in the skolem case the free variables should be universally quantified}
 \begin{theorem}
@@
 We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form.
-In fact our translation is often given as a first step in the Tseytin transformation after which the individual formulas are then converted to CNF, so it is well known, but we are not aware of work that uses just this step as a stand-alone.
-A similar translation for intuitionistic logic was also proposed in~\cite{statman1979intuitionistic}.
+In fact our translation is often presented as the first step in the Tseytin transformation, after which the individual formulas are then converted to CNF.
+Hence, the propositional case of our translation is well known, but we are not aware of any work that uses just this step as a stand-alone.
+A similar translation for intuitionistic propositional logic was also proposed in~\cite{statman1979intuitionistic}.
 
@@
 Let $\varphi$ be some formula.
-For each sub(semi)formula $\psi$ of $\varphi$ consider some new $n$-ary relation symbol $P_\psi$ where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$.
+For each subformula $\psi$ of $\varphi$, consider some new $n$-ary relation symbol $P_\psi$, where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$.
 We denote those variables with $\vec z_\psi$ in some arbitrary but fixed order.
@@
 	\begin{itemize}
-		\item For every atomic formula $A$\\$\mathcal S^+(A) = \{P_A(\vec z_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec z_A)\}$
-		\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$\\$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\to P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\to P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
-		\item For every implication $\varphi \to\psi$\\$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\wedge P_{\varphi}(\vec z_\varphi))\to P_{\psi}(\vec z_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec z_\varphi)\to P_{\psi}(\vec z_\psi))\to P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\psi)$
-		\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$\\$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec z_{Qx\varphi})\to QxP_{\varphi}(\vec z_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec z_{\varphi})\to P_{Qx\varphi}(\vec z_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
-	\end{itemize}
-	and have $\mathcal S(\varphi) = \mathcal S^+(\varphi)\cup\mathcal S^-(\varphi)$.
+		\item For every atomic formula $A$:\\
+$\mathcal S^+(A) = \{P_A(\vec z_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec z_A)\}$
+		\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$:\\
+$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\to P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\to P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
+		\item For every implication $\varphi \to\psi$:\\
+$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\wedge P_{\varphi}(\vec z_\varphi))\to P_{\psi}(\vec z_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec z_\varphi)\to P_{\psi}(\vec z_\psi))\to P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\psi)$
+		\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$:\\
+$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec z_{Qx\varphi})\to QxP_{\varphi}(\vec z_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec z_{\varphi})\to P_{Qx\varphi}(\vec z_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
+	\end{itemize}
+	%We further set $\mathcal S(\varphi) = \mathcal S^+(\varphi)\cup\mathcal S^-(\varphi)$.
 \end{definition}
@@
 
-We give a transformation of structures that corresponds to the above, i.e.
-for each structure $\mathcal M$ and formula $\varphi$ wee can give a structure $\mathcal S(\mathcal M, \varphi)$, the exact definition of which can be found in the appendix as Definition~\ref{def:transf-structure}, such that:
+It is straight-forward to extend structures to the additional symbols such that the formulas $S^+(\varphi)\cup\mathcal S^-(\varphi)$ are satisfied, i.e.
+for each structure $\mathcal M$ and formula $\varphi$ we can define a structure $\mathcal S(\mathcal M, \varphi)$, whose exact definition can be found in the appendix as Definition~\ref{def:transf-structure}, such that:
 \begin{lemma}
-	For every formula $\varphi$ and structure $\mathcal M$
-	$\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$
+For every formula $\varphi$ and structure $\mathcal M$, we have that	$\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$.
 \end{lemma}
@@
 \begin{lemma}
-	For every structure $\mathcal M$
-	$\mathcal M\models \varphi$ if and only if $S(\mathcal M, \varphi)\models P_\varphi$
+	For every structure $\mathcal M$,
+	$\mathcal M\models \varphi$ if and only if $S(\mathcal M, \varphi)\models P_\varphi$.
 \end{lemma}
@@
 	\begin{itemize}
-		\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)$ is.
+		\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)$ is, and
 		\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec z_\varphi)$ is.
@@
 
-We are now ready to put these two reduction together with our embedding f.
+We are now ready to combine the previously presented reductions. 
 Let $\varphi$ be some propositional formula.
@@
 	\forall k(u\preceq k\to A(k)\to (B(k)\vee C(k)))\indent \forall k(u\preceq k\to (A(k)\vee B(k))\to C(k))\\
-	\forall k(u\preceq k\to (\forall l(k\preceq l \to A(l)\to B(l))\to C(k)))
+	\forall k(u\preceq k\to (\forall l(k\preceq l \to A(l)\to B(l))\to C(k))),
 \end{matrix}$$
 with each $P(x)$ possibly being $\bot$. Applying Herbrandization we obtain a formula
-$$\varphi''' = K(\varphi')\to \bigwedge \mathcal S'\to P_\varphi(s)$$
+$$\varphi''' = K(\varphi')\to \bigwedge \mathcal S'\to P_\varphi(s),$$
 where $s$ is some new constant term and each $\psi\in\mathcal S'$ is of one of the forms
@@
 	\forall k(s\preceq k\to A(k)\to (B(k)\vee C(k)))\indent \forall k(s\preceq k\to (A(k)\vee (B(k))\to C(k)))\\
-	\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A(f_\psi(k))\to B(f_\psi(k)))\to C(k))\\
+	\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A(f_\psi(k))\to B(f_\psi(k)))\to C(k)),\\
 \end{matrix}$$
-and $f_\psi$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model then we have a counter-model $(M',I')$ where $M' = \{m\in M|s\preceq m\}$ and
-$f_\psi^{I'}(u) = f^I_\psi(u) \text{ iff $u\preceq f^{I}_\psi(u)$ and $f_\psi^{I'}(u) = u$ else.}$ that is where $s$ is a least element and the $f_\psi$ increase their argument.
+and $f_\psi$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model, then we have a counter-model $(M',I')$ where $M' = \{m\in M|s\preceq m\}$ and
+$f_\psi^{I'}(u) = f^I_\psi(u) \text{ iff $u\preceq f^{I}_\psi(u)$, and $f_\psi^{I'}(u) = u,$ else.}$ That is, where $s$ is the least element and the $f_\psi$ increase their argument.
 Therefore instead of $\varphi'''$ we may consider a formula $$\varphi^\circ = \forall k(s\preceq k)\to \bigwedge\{\forall k(k\preceq f_\psi(k))\:|\:\psi\in\mathcal F_\to\}\to K(\varphi')\to \bigwedge \mathcal S^\circ\to P_\varphi(s)$$ where each formula $\psi\in\mathcal S^\circ$ is of one of the forms
@@
 	\forall k(A(k)\to (B(k)\vee C(k)))\indent\forall k((A(k)\vee B_\psi(k))\to C(k))\\
-	\forall k((A(f_\psi(k))\to B(f_\psi(k)))\to C(k))
+	\forall k((A(f_\psi(k))\to B(f_\psi(k)))\to C(k)),
 \end{matrix}$$
@@
 
-A proof can be found in the appendix as lemma~\ref{thm:prop-countermodel-reduction2}.
-Essentially what this tells us is that for every $u$ at which $\psi$ is fulfilled, in particular each $u$ in the image of $f_\psi$, we may add the assumption that $f_\psi(w) = w$ for all $w\geq u$ without changing the validity of $\psi$.
+A proof can be found in the appendix as Lemma~\ref{thm:prop-countermodel-reduction2}.
+Essentially what this tells us is that for every $u$ at which $\psi$ is fulfilled, in particular for each $u$ in the image of $f_\psi$, we may add the assumption that $f_\psi(w) = w$ for all $w\geq u$ without changing the validity of $\psi$.
 With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.
@@
 	x, &\text{ if $\psi$ occurs in $x$,}\\
-	x, \psi&\text{ otherwise.}
-\end{cases}$$ Now we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula.
+	x\psi, &\text{ otherwise.}
+\end{cases}$$ 
+Now we can replace every $\forall$-quantifier by enumerating all the ground terms, and then replace all distinct ground instances of relations by new propositional variables yielding a CPC formula.
 This gives us the translation stated in Theorem~\ref{thm:reduction-propositional}.
-\begin{theorem}
-	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
-\end{theorem}
+%\begin{theorem}
+%	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
+%\end{theorem}
 
@@
 	Let us consider the law of excluded middle $\varphi = A\vee\neg A$. Then we get
-	$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\}$$
+	$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\},$$
 	and we know that $\varphi$ is valid iff $\bigwedge \mathcal S(\varphi)\to P_\varphi$ is valid.
@@
 	\begin{align*}
-		\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi}, (P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi}\}\cup\\ &\{A^\epsilon\to P_A^\epsilon, P_A^{\psi}\to P_A^{\psi}, (A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}
+		\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi}, (P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi}\}\cup\\ &\{A^\epsilon\to P_A^\epsilon, P_A^{\psi}\to P_A^{\psi}, (A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}.
 	\end{align*}
-	and $P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$, $P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$ defines a counter-model.
+	We now note that $P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$, $P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$ defines a counter-model.
 \end{example}

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
proof read of whole paper (without appendix) finished

@@
 We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model.
-The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
+The idea how to get a polynomially-sized formula is: Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
 The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e.
@@
 	Have $k = |\mathcal F_\to|$ and define$$\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal S}(i)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
-	for $0 < i < k$ as well as the special cases
+	for $0 < i < k$, as well as the special cases
 	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
-	for leafs and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
-	Example encodings of the above formulas would be
+	for leafs, and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
+	Example encodings of the above formulas are:
 	\begin{align*}
@@
 \begin{example}
-	With the above encodings for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
+	With the previous encoding for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
 	$$\hspace*{-.3cm}
-	\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right)
+	\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right),
 	$$
@@
 \end{lemma}
-A proof can be found in the appendix as proof~\ref{lemma:QBF}.
-
-A simple counting argument shows that for each $n < |\mathcal F_\to|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal F_\to|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal S}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal F_\to}(n)|\in \mathcal O(|\mathcal F_\to|)$ so overall we get
+A proof can be found in the appendix as Lemma~\ref{lemma:QBF}.
+
+A simple counting argument shows that for each $n < |\mathcal F_\to|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal F_\to|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal S}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal F_\to}(n)|\in \mathcal O(|\mathcal F_\to|)$ so overall we get:
 \begin{lemma}
@@
 
-First of all recall that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
+First of all recall, that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
 \[\begin{matrix}
 	\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c)))\indent \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c))\indent \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c)))\\
-	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c))\indent \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\indent\forall \vec z(\forall xA(\vec a)\to B(\vec b))\\ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))
+	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c))\indent \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\indent\forall \vec z(\forall xA(\vec a)\to B(\vec b))\\ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b)).
 \end{matrix}\]
@@
 		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
-		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}
+		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}.
 	\end{align*}
@@
 
-It is not possible to completely eliminate worlds from the domain as in the propositional case, since counter-models possibly must be infinite. We will see why that is in more detail later.
+It is not possible to completely eliminate worlds from the domain as in the propositional case, since counter-models can possibly be infinite (as we will see in more detail later).
 
@@
 
-As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
+As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$, but also $E$ has to be axiomatized. For that we use the following predicates:
 \begin{itemize}
@@
 \end{itemize}
-Then analogously to the predicate case define
+Then, analogously to the propositional case, we define:
 \begin{align*}
 	K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
-				 & \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
+				 & \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
 \end{align*}
-Example encodings could be
+Example encodings of these formulas are as follows:
 \begin{align*}
@@
 $$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
-is classically valid where $s$ is a new constant symbol.
+is classically valid, where $s$ is a new constant symbol.
 This was obvious in the propositional case but here it is more nuanced.
@@
 
-As in the predicate case we now show the existence of a canonical counter-model, if a counter-model exists.
+As in the predicate case we now show the existence of a canonical counter-model (in case such a counter-model exists).
 Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$.
@@
 For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$.
-As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by adding formulas
-\begin{itemize}
-	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$
-	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$
-\end{itemize}
-and in each other formula in $\mathcal S^\circ$ removing the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
-$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#(s)$$
+As before we may assume that $s$ is a least element and we can avoid the quantification over $w$ by considering the formulas
+\begin{itemize}
+	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$,
+	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$,
+\end{itemize}
+and by removing removing the $s\preceq u$ for the remaining formulas in $\mathcal S^\circ$.
+We denote the resulting set of formulas by $\mathcal S^\#$.
+Then, the previous formulas are equivalid to
+$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#(s).$$
+\todo{we already define $\varphi^\circ$, shouldn't this be some $\varphi^\dagger$? Also the phrase before should be: Then,$\varphi^\circ$ is equivalid to ...?}
 As in the previous section we can define a tree counter-model for this.
@@
 \begin{definition}
-	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
+	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. \todo{shouldn't this be $\varphi^\dagger$?} Define $\mathcal M_T = (M_T, I_T)$ as follows:
 	\begin{itemize}
 		\item $M_T$ are the sequences on $\{(\vec z, \psi)\:|\:\psi\in \mathcal F, \vec z\in M^n\}$.
-		\item Interpret $\preceq$ as the prefix-order.
-		\item Have $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi)$$
-		\item For propositional variables $P$ have $$P^{I_T}((\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = P^I(f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots))$$
+		\item We interpret $\preceq$ as the prefix-order.
+		\item We set $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi).$$
+		\item For propositional variables $P$, we set $$P^{I_T}((\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = P^I(f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots)).$$
 	\end{itemize}
@@
 \begin{lemma}
-	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
-\end{lemma}
-
-Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec z_1, f_{\psi_2}(\vec z_2, \dots f_{\psi_n}(\vec z_n, s)\dots))$$ is sufficient.
-In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
+	If $\mathcal M$ is a counter-model to $\varphi^\circ$ \todo{shouldn't this be $\varphi^\dagger$?} then so is $\mathcal M_T$.
+\end{lemma}
+
+Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$f_{\psi_1}(\vec z_1, f_{\psi_2}(\vec z_2, \dots f_{\psi_n}(\vec z_n, s)\dots))$$ is sufficient.
+In particular, this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
 from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
-Define
-
-$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#$$
-then we have
+We set
+
+$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#.$$
+Then, we obtain the following result:
 \begin{theorem}
@@
 \end{theorem}
-The complete translation is summarised in Theorem~\ref{fullFOtranslation}. Let us understand why a further reduction as in the propositional case is impossible.
+The complete translation is summarised in Theorem~\ref{fullFOtranslation} (stated earlier). Let us understand why a further reduction as in the propositional case is impossible.
 
@@
 
-We get an analogous result to lemma~\ref{thm:prop-countermodel-reduction}.
+We get an analogous result to Lemma~\ref{thm:prop-countermodel-reduction}:
 
@@
 
-However Part 2 of this Lemma fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$) we have $A(x, v)$ there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
-This is why we must consider sequences that have repetitions.
+However part 2 of Lemma~\ref{thm:fo-countermodel-reduction} fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$), we have $A(x, v)$, there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
+This is why we must consider sequences that have repetitions:
 \begin{wrapfigure}{r}{0.4\textwidth}
@@
 \begin{example}
-	Consider $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model as shown in Figure~\ref{fig1}
-	where at each node the elements that exist at it are indicated.
-For black ones both $A$ and $B$ hold, for red ones only $B$, for blue ones only $A$.
-Furthermore red arrows indicate $f_{\neg \forall xA(x)}$ and blue arrows $f_{\neg \forall xB(x)}$.
-Note that applying $f_{\neg \forall xA(x)}$ we must always reach the left column and $f_{\neg \forall xB(x)}$ we must always reach the right column.
-Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we can never remain stationary.
+	Consider $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model as shown in Figure~\ref{fig1},
+	where we indicate at each node the elements that exist at it.
+For black nodes, both $A$ and $B$ hold, for red ones only $B$, for blue ones only $A$.
+Furthermore, red arrows indicate the functions $f_{\neg \forall xA(x)}$ and blue arrows the functions $f_{\neg \forall xB(x)}$.
+Note that applying $f_{\neg \forall xA(x)}$ we always reach the left column and applying $f_{\neg \forall xB(x)}$ we always reach the right column.
+Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we never remain stationary.
 This is in contrast to the propositional case.
@@
 
-However Lemma~\ref{thm:fo-countermodel-reduction} still enables us to add some additional axioms which could help guide proof search.
-In particular we may assume that $f_\psi(w) = w$ if $\exists u(w = f_\psi(u))$ for $\psi\in\mathcal F_\forall$, i.e.
-$f_\psi$ is idempotent, and  $f_\psi(w) = w$ if $\exists u(w \geq f_\psi(u))$ for each $\psi\in\mathcal F_\to$, which is even stronger than idempotence.
-How adding these properties as assumption helps with proof search has to be examined.
+Let us add that Lemma~\ref{thm:fo-countermodel-reduction} still equips us with additional information that could be used to guide proof search: We could add additional axioms that restrict the interpretation of functions.
+That is, we may assume that $f_\psi(w) = w$, if $\exists u(w = f_\psi(u))$ for all $\psi\in\mathcal F_\forall$ (i.e.
+that $f_\psi$ is idempotent), and $f_\psi(w) = w$, if $\exists u(w \geq f_\psi(u))$ for all $\psi\in\mathcal F_\to$ (which is even stronger than idempotence).
+How adding such axioms impacts proof search will need to be evaluated in experiments.
 
@@
 %In particular in the first-order case it is difficult to estimate how the increased complexity of the output formula impacts proof search.
-Initial tests show promise but we plan on conducting a more thorough evaluation.
+Initial tests show promise but we plan on conducting a more thorough evaluation using state-of-the art automated theorem provers, in particular the Vampire theorem prover~\cite{kovacs2013first}.
 A key factor yet to be determined is how large the parameters $|\mathcal F_\to|$ in the propositional case and $|\mathcal F|$ in the predicate case are in practice.
@@
 
-As a next step we will benchmark our translation using state-of-the art automated theorem provers, in particular the Vampire theorem prover~\cite{kovacs2013first}.
-We will have to adapt the counter-model translation to a proof translation for the particular calculus, e.g.
+Finally, we plan to complement the counter-model translation presented in this paper by a proof translation.
+We will target the particular calculi used by state-of-the-art provers, e.g.
 superposition in the case of Vampire.

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
Incorporated Remarks, fixed Lemma 41

@@
 		\item $\psi$ is fulfilled at $f_\psi^I(u)$ for all $u\in M, \psi\in\mathcal F_\to$.
-		\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\geq u$.
+		\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\succeq u$.
 	\end{enumerate}
@@
 
-Using $E$ we then obtain $\mathcal S^\circ$ analogously to the predicate case by including formulas:
+Using $E$ we then obtain $\mathcal S'$ analogously to the predicate case by including formulas:
 
@@
 Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid if and only if
-$$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
+$$\varphi' = \text{World}(s)\to K(\varphi)\to \bigwedge S'\to P^\#(s)$$
 is classically valid, where $s$ is a new constant symbol.
@@
 \end{itemize}
-and by removing removing the $s\preceq u$ for the remaining formulas in $\mathcal S^\circ$.
+and by removing removing the $s\preceq u$ for the remaining formulas in $\mathcal S'$.
 We denote the resulting set of formulas by $\mathcal S^\#$.
@@
 $$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#(s).$$
-\todo{we already define $\varphi^\circ$, shouldn't this be some $\varphi^\dagger$? Also the phrase before should be: Then,$\varphi^\circ$ is equivalid to ...?}
 As in the previous section we can define a tree counter-model for this.
@@
 \begin{definition}
-	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. \todo{shouldn't this be $\varphi^\dagger$?} Define $\mathcal M_T = (M_T, I_T)$ as follows:
+	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
 	\begin{itemize}
@@
 \begin{lemma}
-	If $\mathcal M$ is a counter-model to $\varphi^\circ$ \todo{shouldn't this be $\varphi^\dagger$?} then so is $\mathcal M_T$.
+	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
 \end{lemma}
@@
 In particular, this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
-from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
+from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u\forall x(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
 We set
@@
 \begin{lemma}\label{lemma:fo-simplification}
-	$\varphi$ is intuitionistically valid if and only of $\varphi^\circ$ is classically valid.
+	$\varphi$ is intuitionistically valid if and only of $\varphi'$ is classically valid.
 \end{lemma}
@@
 	We proceed by translation of counter-examples.
-Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi^\circ$.
+Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi'$.
 As a Kripke frame $(W, \leq)$ take all $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ let $\leq$ be $\preceq^I$ restricted to $W$.
@@
 Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified.
-All that remains to show is $M, I\models\bigwedge\mathcal S^\circ$.
+All that remains to show is $M, I\models\bigwedge\mathcal S'$.
 Consider e.g.

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
typos proof of lemma 35

@@
 \begin{lemma}\label{ap1}
-	Let $\chi$ be a formula and $\{z_1\dots z_n\} = Z$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
-	\begin{itemize}
-		\item if $\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
+Let $\chi$ be a formula and let $\{z_1\dots z_n\} = Z$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
+	\begin{itemize}
+		\item if $M, v \models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
 		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
@@
 $\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$.
-By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v}), \mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as in the lemma.
+By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v})$ and $\mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as stated in the lemma.
 Then we can define $I_{\chi, v}$ as follows:
 
-	For every symbol $p$ that occurs in $\varphi^S$ or $\varphi^H$ but not $\psi^S, \psi^H$ have $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$.
-For every symbol $p$ that occurs in $\psi^S$ or $\psi^H$ but not $\varphi^S, \varphi^H$ have $p^{I_{\chi, v}} = p^{I_{\psi, v}}$.
+	For every symbol $p$ that occurs in $\varphi^S_Z$ or $\varphi^H_Z$ but not $\psi^S_Z, \psi^H_Z$, we set $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$.
+For every symbol $p$ that occurs in $\psi^S_Z$ or $\psi^H_Z$ but not $\varphi^S_Z, \varphi^H_Z$, we set $p^{I_{\chi, v}} = p^{I_{\psi, v}}$.
 Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.
 
-	Let $\circ=\to$.
+	Let $\circ=\ \to$.
 Suppose $\mathcal M, v\models\chi$.
-Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_T$ or $\mathcal M_\chi\models\psi^H_T$, i.e.
-$\mathcal M_\chi\models\chi^S_T$.
+Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_Z$ or $\mathcal M_\chi\models\psi^H_Z$, i.e.
+$\mathcal M_\chi\models\chi^S_Z$.
 On the other hand suppose $\mathcal M, v\not\models\chi$.
-Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_T$ and $\mathcal M_\chi, v\not\models\psi^S_T$, i.e.
-$\mathcal M_\chi\not\models\chi^H_T$.
+Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_Z$ and $\mathcal M_\chi, v\not\models\psi^S_Z$, i.e.
+$\mathcal M_\chi\not\models\chi^H_Z$.
 Analogous arguments work for $\circ\in\{\wedge, \vee\}$.
@@
 $\chi = \forall x\varphi$.
-Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise.
-Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.
+Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/z_1\dots x_n/z_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise.
+Since $Z$ contains all free variables occurring in $\chi$ this is a well-defined function.
 
 By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
-Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
+Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{Z\cup a}$ or $\varphi[a/x]^H_{Z\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
 
 Suppose $\mathcal M, v\models\chi$.
-Then for all $m\in M$ $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{M\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{M\cup\{a\}}[x/a])$, i.e.
-$\mathcal M_\chi,v\models \chi^S$.
+Then, for all $m\in M$ we have that $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{Z\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{Z\cup\{a\}}[x/a])$, i.e.
+$\mathcal M_\chi,v\models \chi^S_Z$.
 On the other hand suppose $\mathcal M, v\not\models\chi$.
-Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{T\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(t_1,\dots t_n)/x]^H_T$, i.e.
-$\mathcal M_\chi, v\not\models(\forall x\varphi)^H_T$.
+Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{Z\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(t_1,\dots t_n)/x]^H_Z$, i.e.
+$\mathcal M_\chi, v\not\models(\forall x\varphi)^H_Z$.
 
@@
 The argument runs dually to 2.
-Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise.
-Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.
+Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/z_1\dots x_n/z_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise.
+Since $Z$ contains all free variables occurring in $\chi$ this is a well-defined function.
 
 	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
-Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
-
-	Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_T$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_T$.
+Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{Z\cup a}$ or $\varphi[a/x]^H_{Z\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
+
+	Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_Z$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_Z$.
 \end{proof}

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
Definition Classical Satisfiability

--- 2022-07-16 CSL/lipics-v2021-EIICL.tex
@@
 	\end{itemize}
-	$\mathcal{M}$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$.
-A formula is valid if every $\Sigma$-structure satisfies it.
+	$\varphi$ is satisfiable if there exist $\mathcal M, v$  with $\mathcal M, v\models\varphi$.
+A $\Sigma$-structure $\mathcal M$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$, we write $\mathcal M\models\varphi$.
+A formula is valid if every $\Sigma$-structure $\mathcal M$ satisfies it.
 We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
some more typos in the proofs

--- 2022-07-16 CSL/lipics-v2021-EIICL.tex
@@
 On the other hand suppose $\mathcal M, v\not\models\chi$.
-Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{Z\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(t_1,\dots t_n)/x]^H_Z$, i.e.
+Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{Z\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(z_1,\dots z_n)/x]^H_Z$, i.e.
 $\mathcal M_\chi, v\not\models(\forall x\varphi)^H_Z$.
@@
 By slight abuse of notation we denote with $\vec z_\psi$ elements of the domain instead of variables.
-In the classical case have
+In the classical case, we set
 	$$\begin{matrix}
 		P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)\\
-		P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))\\P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})\\P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}\indent P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\}
+		P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))\\P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})\\P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}\indent P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\},
 	\end{matrix}$$
-	and for intuitionistic logic and each world $u$
+	and for intuitionistic logic and each world $u$, we set
 	$$\begin{matrix}
@@
 		P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}\\
-		P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}\indent P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}
+		P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}\indent P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}.
 	\end{matrix}$$
@@
 	%
-	Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (A\to B)\to D$, this corresponds to a transformed counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below:
+	Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (B\to A)\to D$, this corresponds to a transformed counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below:
 	\begin{center}
@@
 \begin{proof}
-	Suppose $\varphi$ is not intuitionistically valid i.e.
-it $\bigwedge\mathcal S\to P$ has an intuitionistic counter-model.
-Then by the precious section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
+Suppose $\varphi$ is not intuitionistically valid.
+Then, $\bigwedge\mathcal S\to P$ has an intuitionistic counter-model. Then by the previous section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
+We now define a QBF model iteratively.
 For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$.
-Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.
-
+Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$  in $\mathcal M$.
 Let the $F^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified).
-If not exactly one is interpreted as true then valid$(n)$ fails, i.e.
-have $\psi_n\in\mathcal F_\to$ the only element with $F^n_{\psi_n} = 1$.
-If $F^i_{\psi_n} = 1$ for some $i < n$ then valid$(n)$ also fails, so we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions.
+If not exactly one of the $F^{n}_\psi$ is interpreted as true, then valid$(n)$ fails and the remaining propositions can be chosen arbitrarily.
+If $F^i_{\psi_n} = 1$ for some $i < n$, then valid$(n)$ also fails and the remaining propositions can be chosen arbitrarily.
+So we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions. 
 Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{\psi_1\dots\psi_n}$. 
-Continue this construction until $n  = |\mathcal F_\to|$.
-Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.
+Continue this construction until $n  = |\mathcal F_\to|$. Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.
 
@@
 Interpret $A^\epsilon$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$.
-Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$ in which $F^i_\psi$ is chosen true iff $\psi = \psi_i$.
+Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$, in which $F^i_\psi$ is chosen true iff $\psi = \psi_i$.
 Let $\psi_n\in\mathcal F_\to$.

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
todo removed
appendix small corrections

--- 2022-07-16 CSL/lipics-v2021-EIICL.tex
@@
 for each structure $\mathcal M$ and formula $\varphi$ we can define a structure $\mathcal S(\mathcal M, \varphi)$, whose exact definition can be found in the appendix as Definition~\ref{def:transf-structure}, such that:
-\begin{lemma}
+\begin{lemma}\label{thm:struct1}
 For every formula $\varphi$ and structure $\mathcal M$, we have that	$\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$.
@@
 
-\begin{lemma}
+\begin{lemma}\label{thm:struct2}
 	For every structure $\mathcal M$,
@@
 \begin{lemma}\label{ap2}
-	For every structure $\mathcal M$ and $\{t_1\dots t_n\} = T$ that contains all free variables in $\chi$ and variable assignment $v$
-	\begin{itemize}
-		\item if $\mathcal M, v\models\varphi^S_T$ then $\mathcal M, v\models \varphi$
-		\item if $\mathcal M, v\not\models\varphi^H_T$ then $\mathcal M, v\not\models\varphi$.
+	For every structure $\mathcal M$ and $\{z_1\dots z_n\} = Z$ that contains all free variables in $\chi$ and variable assignment $v$
+	\begin{itemize}
+		\item if $\mathcal M, v\models\varphi^S_Z$ then $\mathcal M, v\models \varphi$
+		\item if $\mathcal M, v\not\models\varphi^H_Z$ then $\mathcal M, v\not\models\varphi$.
 	\end{itemize}
@@
 $\chi = \varphi\to\psi$.
-Suppose $\mathcal M, v\models\chi^S_T$, i.e.
-$\mathcal M, v\not\models\varphi^H_T$ or $\mathcal M, v\models\psi^S_T$.
+Suppose $\mathcal M, v\models\chi^S_Z$, i.e.
+$\mathcal M, v\not\models\varphi^H_Z$ or $\mathcal M, v\models\psi^S_Z$.
 By induction hypothesis $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e.
 $\mathcal M, v\models \chi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e.
-$\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_T$.
+On the other hand suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
+$\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_Z$.
 Again by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.
@@
 $\chi = \forall x\varphi$.
-Suppose $\mathcal M, v\models\chi^S_T$, i.e.
-for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{T\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$.
+Suppose $\mathcal M, v\models\chi^S_Z$, i.e.
+for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{Z\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$.
 Then it follows that $\mathcal M, v\models\varphi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e.
-there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{T\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e.
+On the other hand suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
+there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{Z\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e.
 $\mathcal M, v\not\models\forall x\chi$.
@@
 \end{definition}
+
+Lemmas~\ref{thm:struct1} and~\ref{thm:struct2} then follow directly by induction on the height of $\varphi$.
 

### 2022-07-16 CSL/lipics-v2021-EIICL.tex

--- 2022-07-16 CSL/lipics-v2021-EIICL.tex
@@
 \end{definition}
-\todo{I think with the implicit forall quanitifcation over the free variables in our semantics, now also in the skolem case the free variables should be universally quantified}
+
 \begin{theorem}

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
Typos, Example 34

@@
 The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the fact that the law of excluded middle $A\vee\neg A$ and the double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
-Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particul<ar motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in famous Curry-Howard correspondence~\cite{howard1980formulae}.
-
-The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmarks problems (see e.g.
+Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in the Curry-Howard correspondence~\cite{howard1980formulae}.
+
+The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmark problems (see e.g.
 the
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:Tseytin}
 For every propositional formula $\varphi$ there effectively are an atom $P$ and a set of $\mathcal S$ of formulas (over an extended set of propositions) of one of the forms\\
@@
 3) We can then argue that if there is a counter-model for the resulting formula then there is a counter-model with a certain Kripke frame of bounded size completely determined by the number of formulas of type $(A\to B)\to C$.
-This allows us to eliminate the remaining quantifiers by simply enumerating the worlds in that Kripke model.
+This allows us to eliminate the remaining quantifiers by simply enumerating the worlds in that Kripke frame.
 
@@
 \label{thm:reduction-propositional}
-	Let $\mathcal S$ be as above.
-Let $\mathcal F_\to\subseteq\mathcal S$ denote formulas $\psi$ of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
+	Let $\mathcal S$ be as in Theorem~\ref{thm:Tseytin} and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
 For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
@@
 \begin{theorem}
-	Let $\mathcal S$ be as above.
+	Let $\mathcal S$ be as in Theorem~\ref{thm:Tseytin} and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$.
 There is an effective procedure that produces a QBF $\varphi^Q$ of size $\mathcal O(|\mathcal S|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$ with $2\cdot |\mathcal F_\to|-1$ quantifier alternations such that $\varphi$ is intuitionistically valid if and only if $\varphi^Q$ is not satisfiable.
@@
 We resolve this apparent conflict by introducing a special binary predicate $E$, inspired by~\cite{iemhoff2010eskolemization}, encoding which domain elements exists at which world.
-In particular $E(x, u)$ entails that $x$ represents a domain element and $u$ a world.
 The main transformation then again proceeds in three steps:
@@
 The real difference arises in step 3):
-While we can also establish the existence of a canonic counter-model whose size only depends on formulas of type $(A\to B)\to C$ and $\forall xA\to B$, the size of this counter-model is countably infinite in general.
+While we can also establish the existence of a canonic counter-model whose frame only depends on formulas of type $(A\to B)\to C$ and $\forall xA\to B$, the size of this counter-model is countably infinite in general.
 This is not surprising since certain intuitionistically invalid formulas like the double negation shift $\forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$ don't have finite counter-models.
@@
 In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
-Analogous remarks with regard to validity and satisfiability hold as in the propositional case.
 
@@
 A fresh variable is any variable that does not occur in any of the considered formulas.
-Whenever we add function symbol we implicitly extend the signature by some not previously contained symbol.
+Whenever we add a function symbol we implicitly extend the signature by some not previously contained symbol.
 
@@
 	\end{itemize}
-	Let $K(\varphi)$ encode the Kripke semantics, i.e.
+	Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
 	$$K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))$$
@@
 		\text{\normalfont PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
-		\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w))\:|\: \text{ $A$ is a propositional variable that occurs in $\varphi$}\}
+		\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ is a propositional variable that occurs in $\varphi$}\}
 	\end{align*}
@@
 \begin{lemma}
-	$\bigwedge S^+(\varphi)\wedge P_\varphi\to\varphi$, $\bigwedge S^-(\varphi)\wedge \varphi\to P_\varphi$ hold classically and intuitionistically.
+	$\bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)\to\varphi$ and $\bigwedge S^-(\varphi)\wedge \varphi\to P_\varphi(\vec z_\varphi)$ are valid both classically and intuitionistically.
 \end{lemma}
@@
 	For every structure $\mathcal M$,
-	$\mathcal M\models \varphi$ if and only if $S(\mathcal M, \varphi)\models P_\varphi$.
-\end{lemma}
-
-Then from the previous three Lemmas we get:
+	$\mathcal M\models \varphi$ if and only if $S(\mathcal M, \varphi)\models P_\varphi(\vec z_\varphi)$.
+\end{lemma}
+
+Then from the previous three Lemmata we get:
 
@@
 	\forall k(u\preceq k\to A(k)\to (B(k)\vee C(k)))\indent \forall k(u\preceq k\to (A(k)\vee B(k))\to C(k))\\
-	\forall k(u\preceq k\to (\forall l(k\preceq l \to A(l)\to B(l))\to C(k))),
+	\forall k(u\preceq k\to (\forall l(k\preceq l \to A(l)\to B(l)))\to C(k)),
 \end{matrix}$$
@@
 \end{cases}$$
-Now we can replace every $\forall$-quantifier by enumerating all the ground terms, and then replace all distinct ground instances of relations by new propositional variables yielding a CPC formula.
+Now we can replace every $\forall$-quantifier by enumerating all the ground terms, and then replace all distinct ground instances of relations by new propositional variables yielding a propositional formula.
 This gives us the translation stated in Theorem~\ref{thm:reduction-propositional}.
@@
 	$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\},$$
-	and we know that $\varphi$ is valid iff $\bigwedge \mathcal S(\varphi)\to P_\varphi$ is valid.
+	and we know that $\varphi$ is valid iff $\bigwedge \mathcal S^-(\varphi)\to P_\varphi$ is valid.
 Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically valid, where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
@@
 We now give an analogous transformation for first order logic.
-However our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic.
+Our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic.
 However we do it nonetheless!
@@
 		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
-		\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
+		\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\psi = \forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
 		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
@@
 	\begin{align*}
-		K := &\:\bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
+		K := &\:\bigwedge\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F\}\wedge\\
 		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
@@
 	The size of the obtained formula is linear in the input.
-However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
+However $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
 \end{theorem}
@@
 		\item We set $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi).$$
-		\item For propositional variables $P$, we set $$P^{I_T}((\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = P^I(f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots)).$$
+		\item For relation symbols $R^\#$, we set $${R^\#}^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = {R^\#}^I(\vec z, f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots)).$$
 	\end{itemize}
@@
 In particular, this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
-from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u\forall x(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
-We set
+from $K(\varphi)$ and adjust DomainSubset$(u, w)$ and Persistent$(u, w)$, e.g. remove quantification over $w$ and the precondition $u\preceq w$ from $K(\varphi)$ and replace DomainSubset$(u, w)$ with
+$$\bigwedge\{\forall \vec z\forall u\forall x(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$$ and Persistent$(u, w)$ with $$\bigwedge\{\forall\vec z_1\forall\vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\psi\in\mathcal F_\to, \text{$A\in\Sigma$ is a relation symbol}\}.$$ We set
 
@@
 This is why we must consider sequences that have repetitions:
-\begin{wrapfigure}{r}{0.4\textwidth}
+
+\begin{wrapfigure}[16]{r}{0.4\textwidth}
+	\vspace{-32pt}
 	\begin{center}
@@
 			\node      (4l)   [above=of 3l] {{$\vdots$}};
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (0.north east) -- (1r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (0.north west) -- (1l.south east);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1r.north) -- (2r.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1l.north) -- (2l.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1l.north east) -- (2r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1r.north west) -- (2l.south east);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2r.north) -- (3r.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2l.north) -- (3l.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2l.north east) -- (3r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2r.north west) -- (3l.south east);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3r.north) -- (4r.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3l.north) -- (4l.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3l.north east) -- (4r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3r.north west) -- (4l.south east);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (0.north east) -- (1r.south west);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (0.north west) -- (1l.south east);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1r.north) -- (2r.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1l.north) -- (2l.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1l.north east) -- (2r.south west);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1r.north west) -- (2l.south east);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2r.north) -- (3r.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2l.north) -- (3l.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2l.north east) -- (3r.south west);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2r.north west) -- (3l.south east);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3r.north) -- (4r.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3l.north) -- (4l.south);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3l.north east) -- (4r.south west);
+			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3r.north west) -- (4l.south east);
 		\end{tikzpicture}
@@
 \end{wrapfigure}
-
 \begin{example}
 	Consider $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model as shown in Figure~\ref{fig1},
-	where we indicate at each node the elements that exist at it.
-For black nodes, both $A$ and $B$ hold, for red ones only $B$, for blue ones only $A$.
-Furthermore, red arrows indicate the functions $f_{\neg \forall xA(x)}$ and blue arrows the functions $f_{\neg \forall xB(x)}$.
-Note that applying $f_{\neg \forall xA(x)}$ we always reach the left column and applying $f_{\neg \forall xB(x)}$ we always reach the right column.
-Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we never remain stationary.
-This is in contrast to the propositional case.
+	where we indicate at each world the elements that exist at it.
+For black numbers $n$, both $A(n)$ and $B(n)$ hold at that world, for red ones only $A(n)$, for blue ones only $B(n)$.
+The arrows indicate the order.
+$\neg\forall xA(x)$ is not fulfilled at worlds in the left column and $\neg\forall xB(x)$ is not fulfilled at worlds in the right column.
+In particular applying $f_{\neg \forall xA(x)}$ we must always reach the right column and applying $f_{\neg \forall xB(x)}$ we must always reach the left column.
+Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we cannot remain stationary.
 \end{example}

### 2022-08-01 Notes/Embedding intuitionistic into classical logic/EIICL.tex
First fully working implementation of the translation (single-file only)

@@
 \usepackage{footmisc}
+\usepackage{wrapfig}
 
@@
 
+	\section{Skolemization and Herbrandization}
+
+	An important step in the embedding will be the elimination of quantifiers via Herbrandization.
+	In this process we introduce fresh variables and add additional function symbols to the signature.
+	A fresh variable is any variable that does not occur in any of the considered formulas.
+	Whenever we add a function symbol we implicitly extend the signature by some not previously contained symbol.
+
+	\begin{definition}
+		For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to $Z$ by simultaneous induction as follows:
+		\begin{itemize}
+			\item $A^S_Z = A^H_Z = A$ for each atomic $A$.
+			\item $(\varphi\circ\psi)^X_Z = \varphi^X_Z\circ\psi^X_Z$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
+			\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$\\$(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
+			\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
+			\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.
+		\end{itemize}
+		Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
+	\end{definition}
+
+	\begin{theorem}
+		\label{thm:herbrand-skolem}
+		For every formula $\varphi$
+		\begin{itemize}
+			\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
+			\item $\varphi$ and $\varphi^H$ are classically equivalid.
+		\end{itemize}
+	\end{theorem}
+	which follows from the following two Lemmata:
+	\begin{lemma}
+		Let $\chi$ be a formula and let $\{z_1\dots z_n\} = Z$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
+		\begin{itemize}
+			\item if $M, v \models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
+			\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
+		\end{itemize}
+	\end{lemma}
+
+	\begin{proof}
+		We proceed by simultaneous induction on the formula height.
+
+		For atomic $\chi$ the claims are clear.
+
+		Otherwise there are $3$ cases:
+
+		1.
+$\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$.
+By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v})$ and $\mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as stated in the lemma.
+		Then we can define $I_{\chi, v}$ as follows:
+
+		For every symbol $p$ that occurs in $\varphi^S_Z$ or $\varphi^H_Z$ but not $\psi^S_Z, \psi^H_Z$, we set $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$.
+For every symbol $p$ that occurs in $\psi^S_Z$ or $\psi^H_Z$ but not $\varphi^S_Z, \varphi^H_Z$, we set $p^{I_{\chi, v}} = p^{I_{\psi, v}}$.
+Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.
+
+		Let $\circ=\ \to$.
+Suppose $\mathcal M, v\models\chi$.
+Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_Z$ or $\mathcal M_\chi\models\psi^H_Z$, i.e.
+$\mathcal M_\chi\models\chi^S_Z$.
+On the other hand suppose $\mathcal M, v\not\models\chi$.
+Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_Z$ and $\mathcal M_\chi, v\not\models\psi^S_Z$, i.e.
+$\mathcal M_\chi\not\models\chi^H_Z$.
+Analogous arguments work for $\circ\in\{\wedge, \vee\}$.
+
+		2.
+$\chi = \forall x\varphi$.
+Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/z_1\dots x_n/z_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise.
+Since $Z$ contains all free variables occurring in $\chi$ this is a well-defined function.
+
+		By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
+Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{Z\cup a}$ or $\varphi[a/x]^H_{Z\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
+
+		Suppose $\mathcal M, v\models\chi$.
+Then, for all $m\in M$ we have that $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{Z\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{Z\cup\{a\}}[x/a])$, i.e.
+$\mathcal M_\chi,v\models \chi^S_Z$.
+On the other hand suppose $\mathcal M, v\not\models\chi$.
+Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{Z\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(z_1,\dots z_n)/x]^H_Z$, i.e.
+$\mathcal M_\chi, v\not\models(\forall x\varphi)^H_Z$.
+
+		3.
+$\chi = \exists x\varphi$.
+The argument runs dually to 2.
+Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/z_1\dots x_n/z_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise.
+Since $Z$ contains all free variables occurring in $\chi$ this is a well-defined function.
+
+		By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
+Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{Z\cup a}$ or $\varphi[a/x]^H_{Z\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
+
+		Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_Z$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_Z$.
+	\end{proof}
+
+	\begin{lemma}
+		For every structure $\mathcal M$ and $\{z_1\dots z_n\} = Z$ that contains all free variables in $\chi$ and variable assignment $v$
+		\begin{itemize}
+			\item if $\mathcal M, v\models\varphi^S_Z$ then $\mathcal M, v\models \varphi$
+			\item if $\mathcal M, v\not\models\varphi^H_Z$ then $\mathcal M, v\not\models\varphi$.
+		\end{itemize}
+	\end{lemma}
+
+	\begin{proof}
+		Again we proceed by simultaneous induction on the formula height.
+
+		For atoms the claims are clear. We distinguish 5 cases.
+
+		1.
+$\chi = \varphi\to\psi$.
+Suppose $\mathcal M, v\models\chi^S_Z$, i.e.
+$\mathcal M, v\not\models\varphi^H_Z$ or $\mathcal M, v\models\psi^S_Z$.
+By induction hypothesis $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e.
+$\mathcal M, v\models \chi$.
+On the other hand suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
+$\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_Z$.
+Again by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.
+
+		2. + 3. Conjunctions and Disjunctions are dealt with analogously.
+
+		4.
+$\chi = \forall x\varphi$.
+Suppose $\mathcal M, v\models\chi^S_Z$, i.e.
+for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{Z\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$.
+Then it follows that $\mathcal M, v\models\varphi$.
+On the other hand suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
+there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{Z\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e.
+$\mathcal M, v\not\models\forall x\chi$.
+
+		5. $\chi = \exists x\varphi$.  The argument runs dually to 4.
+	\end{proof}
+
 	\section{Proof theory}
@@
 We first present an embedding from IPC to CQC which is conceptually easier to understand, but will then reduce it to CPC.
+To that end we will first propose a Tseytin-like translation, that reduces formulas to a normal form while preserving validity:
+
+	\begin{theorem}\label{thm:Tseytin}
+		For every propositional formula $\varphi$ there effectively are an atom $P$ and a set of $\mathcal S$ of formulas (over an extended set of propositions) of one of the forms\\
+		$\text{ \hspace{1cm}} A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$\\
+		for atomic $A, B, C$, such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
+The time complexity of the transformation is linear in the input size.
+	\end{theorem}
+
+	\begin{theorem}
+		Let $\mathcal S$ be a set of formulas of the form
+		$$(A\wedge B)\to C, (A\vee B)\to C, A\to (B\wedge C), A\to(B\vee C), (A\to B)\to C$$
+		and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
+For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
+Obtain $\mathcal S^\#$ by including the following formulas:
+		\begin{itemize}
+			\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
+			\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+			\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+			\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
+		\end{itemize}
+		Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
+The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$.
+Futhermore, there is an effective procedure for translating counter-models between the sentences.
+	\end{theorem}
 
@@
 	\begin{definition}
-		Let $\varphi$ be a propositional formula. Define $\varphi^{k}$ inductively as follows:
-		\begin{itemize}
-			\item $A^{k} = A(k)$ for every propositional variable $A$.
-			\item $(\varphi\wedge\psi)^k = \varphi^k\wedge\psi^k$.
-			\item $(\varphi\vee\psi)^k = \varphi^k\vee\psi^k$.
-			\item $(\varphi\to \psi)^k = \forall u(k\preceq u\to\varphi^{u}\to\psi^{u})$ where $u$ is some new free variable.
-		\end{itemize}
-		Let $K(\varphi)$ encode that $\preceq$ is a partial order, the persistency condition for all propositional variables occurring in $\varphi$
+		Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively as follows:
+		\begin{itemize}
+			\item $A^{u} = A(u)$ for every propositional variable $A$.
+			\item $\bot^u = \bot$.
+			\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
+			\item $(\varphi\to \psi)^u = \forall w(u\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new variable.
+		\end{itemize}
+		Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
+		$$K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))$$
+		with e.g.
 		\begin{align*}
-			K(\varphi) = &\forall k(k\preceq k)\wedge \forall k\forall l(k\preceq l\to l\preceq k\to l = k) \wedge \\&\forall k\forall l\forall m(k\preceq l \to l\preceq m\to k\preceq m)\wedge\\&\bigwedge \{\forall k\forall l(k\preceq l\to A(k)\to A(l))\:|\: \text{ $A$ occurs in $\varphi$}\}
+			\text{\normalfont PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
+			\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ is a propositional variable that occurs in $\varphi$}\}
 		\end{align*}
@@
 
-	All that we have done is model Kripke frames with a finite number of propositional variables as a first-order theory and by definition we have
+	\noindent As all we have done is modelling Kripke frames as a first-order theory we directly obtain:
 
@@
 						 &= K(\varphi)\to (\forall u'( u\preceq u' \to A^{u'}\to \bot^{u'})\vee A(u))\\
-						 &= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot(u'))\vee A(u))
+						 &= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot)\vee A(u))
 		\end{align*}
@@
 
-	\subsection{Quantifiers in Classical Logic}
-
-	Of course translating a propositional formula to a first-order formula is somewhat unsatisfying.
-As a first step to obtaining a propositional formula we wish to eliminate the quantifiers.
-Since we are interested in preserving validity we use Herbrandization.
-In this process we introduce free variables and also additional function symbols to the signature.
-Whenever we speak of a new free variable what is meant is a free variable that does not occur in any of the considered formulas.
-Whenever a new function symbol is introduced we implicitly extend the signature by some not previously contained symbol.
-
-	Now usually in classical logic before removing quantifiers prenexification is performed, however since we wish to apply a similar procedure to intuitionistic logic later, where prenexification is not valid, we will do without.
-
-	\begin{definition}
-		For formulas $\varphi$ we define the Skolemization $\varphi^S_T$ and Herbrandization $\varphi^H_T$ with respect to $T$ by simultaneous induction as follows:
-		\begin{itemize}
-			\item $A^S_T = A^H_T = A$ for each atomic $A$.
-			\item $(\varphi\circ\psi)^X_T = \varphi^X_T\circ\psi^X_T$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
-			\item $(\varphi\to\psi)^S_T = \varphi^H_T\to \psi^S_T$\\$(\varphi\to\psi)^H_T = \varphi^S_T\to\psi^H_T$.
-			\item $(\forall x\varphi)^S_T = \forall x(\varphi[a/x]^S_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_T = \varphi[s(t_1,\dots,t_n)/x]^H_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$.
-			\item $(\exists x\varphi)^S_T = \varphi[s(t_1,\dots,t_n)/x]^S_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$\\$(\exists x\varphi)^H_T = \exists x(\varphi[a/x]^H_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable.
-		\end{itemize}
-		Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
-	\end{definition}
-
-	\begin{theorem}For every formula $\varphi$
-		\begin{itemize}
-			\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
-			\item $\varphi$ and $\varphi^H$ are classically equivalid.
-		\end{itemize}
-	\end{theorem}
-
-	which follows from the following two Lemmata:
-
-	\begin{lemma}
-		Let $\chi$ be a formula and $\{t_1\dots t_n\} = T$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
-		\begin{itemize}
-			\item if $\mathcal M, v\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
-			\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
-		\end{itemize}
-	\end{lemma}
-
-	\begin{proof}
-		We proceed by simultaneous induction on the formula height.
-
-		For atomic $\chi$ the claims are clear.
-
-		Otherwise there are $3$ cases:
-
-		1.
-$\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$.
-By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v}), \mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as in the lemma.
-Then we can define $I_{\chi, v}$ as follows:
-
-		For every symbol $p$ that occurs in $\varphi^S$ or $\varphi^H$ but not $\psi^S, \psi^H$ have $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$.
-For every symbol $p$ that occurs in $\psi^S$ or $\psi^H$ but not $\varphi^S, \varphi^H$ have $p^{I_{\chi, v}} = p^{I_{\psi, v}}$.
-Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.
-
-		Let $\circ=\to$.
-Suppose $\mathcal M, v\models\chi$.
-Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_T$ or $\mathcal M_\chi\models\psi^H_T$, i.e.
-$\mathcal M_\chi\models\chi^S_T$.
-On the other hand suppose $\mathcal M, v\not\models\chi$.
-Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_T$ and $\mathcal M_\chi, v\not\models\psi^S_T$, i.e.
-$\mathcal M_\chi\not\models\chi^H_T$.
-Analogous arguments work for $\circ\in\{\wedge, \vee\}$.
-
-		2.
-$\chi = \forall x\varphi$.
-Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise.
-Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.
-
-		By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
-Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
-
-		Suppose $\mathcal M, v\models\chi$.
-Then for all $m\in M$ $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{M\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{M\cup\{a\}}[x/a])$, i.e.
-$\mathcal M_\chi,v\models \chi^S$.
-On the other hand suppose $\mathcal M, v\not\models\chi$.
-Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{T\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(t_1,\dots t_n)/x]^H_T$, i.e.
-$\mathcal M_\chi, v\not\models(\forall x\varphi)^H_T$.
-
-		3.
-$\chi = \exists x\varphi$.
-The argument runs dually to 2.
-Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise.
-Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.
-
-		By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
-Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
-
-		Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_T$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_T$.
-	\end{proof}
-
-	\begin{lemma}
-		For every structure $\mathcal M$ and $\{t_1\dots t_n\} = T$ that contains all free variables in $\chi$ and variable assignment $v$
-		\begin{itemize}
-			\item if $\mathcal M, v\models\varphi^S_T$ then $\mathcal M, v\models \varphi$
-			\item if $\mathcal M, v\not\models\varphi^H_T$ then $\mathcal M, v\not\models\varphi$.
-		\end{itemize}
-	\end{lemma}
-
-	\begin{proof}
-		Again we proceed by simultaneous induction on the formula height.
-
-		For atoms the claims are clear. We distinguish 5 cases.
-
-		1.
-$\chi = \varphi\to\psi$.
-Suppose $\mathcal M, v\models\chi^S_T$, i.e.
-$\mathcal M, v\not\models\varphi^H_T$ or $\mathcal M, v\models\psi^S_T$.
-By induction hypothesis $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e.
-$\mathcal M, v\models \chi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e.
-$\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_T$.
-Again by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.
-
-		2. + 3. Conjunctions and Disjunctions are dealt with analogously.
-
-		4.
-$\chi = \forall x\varphi$.
-Suppose $\mathcal M, v\models\chi^S_T$, i.e.
-for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{T\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$.
-Then it follows that $\mathcal M, v\models\varphi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e.
-there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{T\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e.
-$\mathcal M, v\not\models\forall x\chi$.
-
-		5. $\chi = \exists x\varphi$.  The argument runs dually to 4.
-	\end{proof}
-
-	\section{Equivalid normalization}
+	\subsection{Normal form transformation}
 
 	Of course transforming a propositional formula into a predicate formula is somewhat unsatisfying.
-As a first step we want to apply Herbrandization.
+As a first step towards obtaining a propositional formula we want to apply Herbrandization.
 Now trying to apply it to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated.
 We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form.
-A similar translation was proposed in~\cite{statman1979intuitionistic}.
+In fact our translation is often presented as the first step in the Tseytin transformation, after which the individual formulas are then converted to CNF.
+	Hence, the propositional case of our translation is well known, but we are not aware of any work that uses just this step as a stand-alone.
+A similar translation for intuitionistic propositional logic was also proposed in~\cite{statman1979intuitionistic}.
 
@@
 		Let $\varphi$ be some formula.
-For each sub(semi)formula $\psi$ of $\varphi$ consider some new $n$-ary relation symbol $P_\psi$ where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$.
+For each subformula $\psi$ of $\varphi$, consider some new $n$-ary relation symbol $P_\psi$, where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$.
 We denote those variables with $\vec z_\psi$ in some arbitrary but fixed order.
@@
 		\begin{itemize}
-			\item For every atomic formula $A$\\$\mathcal S^+(A) = \{P_A(\vec z_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec z_A)\}$
-			\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$\\$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\to P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\to P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
-			\item For every implication $\varphi \to\psi$\\$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\wedge P_{\varphi}(\vec z_\varphi))\to P_{\psi}(\vec z_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec z_\varphi)\to P_{\psi}(\vec z_\psi))\to P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\psi)$
-			\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$\\$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec z_{Qx\varphi})\to QxP_{\varphi}(\vec z_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec z_{\varphi})\to P_{Qx\varphi}(\vec z_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
-		\end{itemize}
-		and have $\mathcal S(\varphi) = \mathcal S^+(\varphi)\cup\mathcal S^-(\varphi)$.
-	\end{definition}
-
-	\begin{lemma}The following are true for both classical and intuitionistic semantics:
-		\begin{itemize}
-			\item $\bigwedge S^+(\varphi)\wedge P_\varphi\models\varphi$.
-			\item $\bigwedge S^-(\varphi)\wedge \varphi\models P_\varphi$.
-		\end{itemize}
+			\item For every atomic formula $A$:\\
+			$\mathcal S^+(A) = \{P_A(\vec z_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec z_A)\}$
+			\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$:\\
+			$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\to P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\to P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
+			\item For every implication $\varphi \to\psi$:\\
+			$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\wedge P_{\varphi}(\vec z_\varphi))\to P_{\psi}(\vec z_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec z_\varphi)\to P_{\psi}(\vec z_\psi))\to P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\psi)$
+			\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$:\\
+			$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec z_{Qx\varphi})\to QxP_{\varphi}(\vec z_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec z_{\varphi})\to P_{Qx\varphi}(\vec z_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
+		\end{itemize}
+		%We further set $\mathcal S(\varphi) = \mathcal S^+(\varphi)\cup\mathcal S^-(\varphi)$.
+	\end{definition}
+
+	\begin{lemma}
+		$\bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)\to\varphi$ and $\bigwedge S^-(\varphi)\wedge \varphi\to P_\varphi(\vec z_\varphi)$ are valid both classically and intuitionistically.
 	\end{lemma}
 
-	\begin{proof}
-		The claims are verified in a straightforward manner by induction on the height of $\varphi$.
-	\end{proof}
+	It is straight-forward to extend structures to the additional symbols such that the formulas $S^+(\varphi)\cup\mathcal S^-(\varphi)$ are satisfied.
+	\begin{definition}
+		For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms of the form $P_\psi$.
+By slight abuse of notation we denote with $\vec z_\psi$ elements of the domain instead of variables.
+In the classical case, we set
+		$$\begin{matrix}
+			P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)\\
+			P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))\\P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})\\P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}\indent P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\},
+		\end{matrix}$$
+		and for intuitionistic logic and each world $u$, we set
+		$$\begin{matrix}
+			P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)\\
+			P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))\\
+			P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}\\
+			P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}\indent P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}.
+		\end{matrix}$$
+	\end{definition}
+	From this definition we directly obtain the following Lemmata by induction on the formula height.
+	\begin{lemma}\label{thm:struct1}
+		For every formula $\varphi$ and structure $\mathcal M$, we have that $\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$.
+	\end{lemma}
+
+	\begin{lemma}\label{thm:struct2}
+		For every structure $\mathcal M$,
+		$\mathcal M\models \varphi$ if and only if $S(\mathcal M, \varphi)\models P_\varphi(\vec z_\varphi)$.
+	\end{lemma}
+
+	Then from the previous three Lemmata we get:
+
+	\begin{corollary}\label{eequivalid}
+		In both intuitionistic and classical logic
+		\begin{itemize}
+			\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)$ is, and
+			\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec z_\varphi)$ is.
+		\end{itemize}
+	\end{corollary}
 
@@
 
-	We can give a transformation of structures that corresponds to the above transformation of sentences.
-
-	\begin{definition}
-		For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms of the form $P_\psi$.
-By slight abuse of notation we denote with $\vec z_\psi$ elements of the domain instead of variables.
-In the classical case have
-		$$P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)$$
-		and
-		$$P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))$$$$ P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})$$
-		$$P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}$$$$ P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\}$$
-		\\
-		and for intuitionistic logic and some world $u$
-		$$P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)$$
-		and
-		$$P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))$$$$ P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}$$
-		$$P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}$$$$ P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}$$
-		\\
-	\end{definition}
-
-	The above definitions are motivated by the following results which follow directly from the involved definitions and are true both for classical and intuitionistic semantics:
-
-	\begin{lemma}
-		For every formula $\varphi$ and structure $\mathcal M$
-		$$\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$$
-	\end{lemma}
-
-	\begin{lemma}
-		For every structure $\mathcal M$
-		$$\mathcal M\models \varphi\text{ iff }S(\mathcal M, \varphi)\models P_\varphi$$
-	\end{lemma}
-
-	Then from the previous three Lemmas we get
-
-	\begin{corollary}\label{equivalid}
-		In both intuitionistic and classical logic
-		\begin{itemize}
-			\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)$ is.
-			\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec z_\varphi)$ is.
-		\end{itemize}
-	\end{corollary}
-
-	\section{IPC to CPC}
-
-	We are now ready to put everything together.
-Recall that our goal is for a formula to give a modified formula that is valid classically iff the original formula is valid intuitionistically.
+	\subsection{IPC to simplified CQC}
+
+	We are now ready to combine the previously presented reductions.
+	Let $\varphi$ be some propositional formula.
 As a first step we perform the normalization from the last section, i.e.
-instead of $\varphi$ we consider $$\varphi' = \bigwedge \mathcal S^-(\varphi)\to P_\varphi$$
+instead of $\varphi$ we consider $$\varphi' = \bigwedge \mathcal S^-(\varphi)\to P_\varphi.$$
 
 	Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
-	$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$
-
+	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$$
 	where we can treat $A\to B$ as a special case $A\to (B\wedge B)$. We then apply the transformation to CQC, leading to a formula
-	$$\varphi'' = K(\varphi')\to\bigwedge  \mathcal S\to P_\varphi(u)$$
+	$$\varphi'' = K(\varphi')\to\bigwedge  \mathcal S\to P_\varphi(u),$$
 	where $u$ is some new free variable and each $\psi\in\mathcal S$ is of one of the forms
-	$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\wedge (B_\psi(k))\to C_\psi(k)))$$
-	$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\vee B_\psi(k))\to C_\psi(k))$$
-	$$\forall k(u\preceq k\to (\forall l(k\preceq l \to A_\psi(l)\to B_\psi(l))\to C_\psi(k)))$$
+	$$\begin{matrix}
+		\forall k(u\preceq k\to A(k)\to (B(k)\wedge C(k)))\indent  \forall k(u\preceq k\to (A(k)\wedge (B(k))\to C(k)))\\
+		\forall k(u\preceq k\to A(k)\to (B(k)\vee C(k)))\indent \forall k(u\preceq k\to (A(k)\vee B(k))\to C(k))\\
+		\forall k(u\preceq k\to (\forall l(k\preceq l \to A(l)\to B(l)))\to C(k)),
+	\end{matrix}$$
 	with each $P(x)$ possibly being $\bot$. Applying Herbrandization we obtain a formula
-	$$\varphi''' = K(\varphi')\to \bigwedge \mathcal S'\to P_\varphi(s)$$
+	$$\varphi''' = K(\varphi')\to \bigwedge \mathcal S'\to P_\varphi(s),$$
 	where $s$ is some new constant term and each $\psi\in\mathcal S'$ is of one of the forms
-	$$\hspace*{-2mm}\forall k(s\preceq k\to A_\psi(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k(s\preceq k\to (A_\psi(k)\wedge (B_\psi(k))\to C_\psi(k)))$$
-	$$\hspace*{-2mm}\forall k(s\preceq k\to A_\psi(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(s\preceq k\to (A_\psi(k)\vee (B_\psi(k))\to C_\psi(k)))$$
-	$$\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A_\psi(f_\psi(k))\to B_\psi(f_\psi(k)))\to C_\psi(k))$$
-	and $f$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model then we have a counter-model $(M',I')$ where $$M' = \{m\in M|s\preceq m\}$$ and
-	$$f_\psi^{I'}(u) = \begin{cases}
-		f^I_\psi(u)&\text{ iff $u\preceq f^{I}_\psi(u)$}\\
-		u&\text{ else}
-	\end{cases}$$ that is where $s$ is a least element and the $f_\psi$ increase their argument.
+	$$\begin{matrix}
+		\forall k(s\preceq k\to A(k)\to (B(k)\wedge C(k)))\indent \forall k(s\preceq k\to (A(k)\wedge (B(k))\to C(k)))\\
+		\forall k(s\preceq k\to A(k)\to (B(k)\vee C(k)))\indent \forall k(s\preceq k\to (A(k)\vee (B(k))\to C(k)))\\
+		\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A(f_\psi(k))\to B(f_\psi(k)))\to C(k)),\\
+	\end{matrix}$$
+	and $f_\psi$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model, then we have a counter-model $(M',I')$ where $M' = \{m\in M|s\preceq m\}$ and
+	$f_\psi^{I'}(u) = f^I_\psi(u) \text{ iff $u\preceq f^{I}_\psi(u)$, and $f_\psi^{I'}(u) = u,$ else.}$ That is, where $s$ is the least element and the $f_\psi$ increase their argument.
 Therefore instead of $\varphi'''$ we may consider a formula $$\varphi^\circ = \forall k(s\preceq k)\to \bigwedge\{\forall k(k\preceq f_\psi(k))\:|\:\psi\in\mathcal F_\to\}\to K(\varphi')\to \bigwedge \mathcal S^\circ\to P_\varphi(s)$$ where each formula $\psi\in\mathcal S^\circ$ is of one of the forms
-	$$\forall k(A_\psi(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k((A_\psi(k)\wedge B_\psi(k))\to C_\psi(k))$$
-	$$\forall k(A_\psi(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k((A_\psi(k)\vee B_\psi(k))\to C_\psi(k))$$
-	$$\forall k((A_\psi(f_\psi(k))\to B_\psi(f_\psi(k)))\to C_\psi(k))$$
+	$$\begin{matrix}
+		\forall k(A(k)\to (B(k)\wedge C(k)))\indent\forall k((A(k)\wedge B(k))\to C(k))\\
+		\forall k(A(k)\to (B(k)\vee C(k)))\indent\forall k((A(k)\vee B_\psi(k))\to C(k))\\
+		\forall k((A(f_\psi(k))\to B(f_\psi(k)))\to C(k)),
+	\end{matrix}$$
 	and $\mathcal F_\to$ denotes the set of formulas of the last kind.
 
+	\subsection{IPC to CPC}
+
 	We now argue that if a counter-model exists for $\varphi^\circ$ then a bounded counter-model exists, allowing us to return to the propositional case.
@@
 		Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
-For $\psi\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true.
+		We say $\psi\in\mathcal F_\to$ is \emph{fulfilled} at $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true.
 For $\psi\in\mathcal F_\to$ define $$g_\psi : M\to M, u\mapsto\begin{cases}
-			u&\text{ if $\psi$ is fulfilled in $u$}\\
-			f^I_\psi(u)&\text{ else}		
-		\end{cases}$$Define a $\mathcal M_T = (M_T, I_T)$ as follows:
+			u,&\text{ if $\psi$ is fulfilled in $u$,}\\
+			f^I_\psi(u),&\text{ else.}		
+		\end{cases}$$Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
 		\begin{itemize}
@@
 			\item Interpret $\preceq$ as the prefix-order.
-			\item Have $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
-				\psi_1\dots\psi_n&\text{if $\psi$ occurs in $\psi_1\dots\psi_n$}\\
-				\psi_1\dots\psi_n\psi&\text{else}			
+			\item We set $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
+				\psi_1\dots\psi_n, &\text{if $\psi$ occurs in $\psi_1\dots\psi_n$,}\\
+				\psi_1\dots\psi_n\psi, &\text{else.}			
 			\end{cases}$$
-			\item For propositional variables $P$ have $$P^{I_T}(\psi_1\dots \psi_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots))$$
-		\end{itemize}
-	\end{definition}
-	\begin{lemma}
+			\item For propositional variables $P$, we set $$P^{I_T}(\psi_1\dots \psi_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots)).$$
+		\end{itemize}
+	\end{definition}
+	\begin{lemma}~\label{thm:prop-countermodel-reduction}
 		Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
@@
 			\item $\psi$ is fulfilled at $f_\psi^I(u)$ for all $u\in M, \psi\in\mathcal F_\to$.
-			\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\geq u$.
+			\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\succeq u$.
 		\end{enumerate}
-
 	\end{lemma}
@@
 	\begin{proof}
-		1.
-If $C_\psi^I(u)$ is true then we are done due to persistency.
-Otherwise we know that $(A_\psi^I(f_\psi^I(u))\to B_\psi^I(f_\psi^I(u)))\to C_\psi^I(f_\psi^I(u))$ holds, so $A_\psi^I(f_\psi^I(u))\to B_\psi^I(f_\psi^I(u))$ must be true.
-
-		2.
-If $C_\psi^I(u)$ is true then $C_\psi^I(w)$ is also true and we are done.
-So suppose $A_\psi^I(u)\to B_\psi^I(u)$ is false, i.e.
-$A_\psi^I(u)$ true and $B_\psi^I(u)$ false.
-Then due to the persistency condition $A_\psi^I(w)$ is also true.
-Again if $C_\psi^I(w)$ is true we are done.
-Otherwise we know that $B^I(f^I_\psi(w))$ must be false, since $(A^I_\psi(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I_\psi(w)$ and $A^I_\psi(f^I_\psi(w))$ hold.
-But then due to persistency so is $B^I_\psi(w)$ and we have that $\psi$ is fulfilled at $w$.
+		1. If $C^I(u)$ is true then we are done due to persistency. Otherwise $C^I(u)$ is false.
+		Because $(A^I(f_\psi^I(u))\to B^I(f_\psi^I(u)))\to C^I(u)$ holds, then $A^I(f_\psi^I(u))\to B^I(f_\psi^I(u))$ must be false.
+
+		2. Let $w$ be some element with $w\geq u$.
+		If $C^I(w)$ is true, we are done.
+		Otherwise $C^I(w)$ is false.
+		Due to persistency $C^I(u)$ is also false.
+		Because $\psi$ is fulfilled at $u$, $A^I(u)\to B^I(u)$ must be false, i.e. $A^I(u)$ is true and $B^I(u)$ is false. Then $A^I(w)$ and $A^I(f^I_\psi(w))$ are also true due to persistency.
+		Because $(A^I(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I(w)$ holds, we now get that  $B^I(f^I(w))$ must be false.
+		But then due to persistency so is $B^I(w)$ and we have that $\psi$ is fulfilled at $w$.
 	\end{proof}
 
-	Essentially what fulfilment of $\psi$ at $u$ tells us is that having $f_\psi(w) = w$ for all $w\geq u$ works.
+	Essentially what this tells us is that for every $u$ at which $\psi$ is fulfilled, in particular for each $u$ in the image of $f_\psi$, we may add the assumption that $f_\psi(w) = w$ for all $w\geq u$ without changing the validity of $\psi$.
 	With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.
@@
 	\begin{example}
-		Consider
-		$$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model
+		Consider $$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model, depicted below,
 		\begin{center}
@@
 		\end{center}
-		\vspace*{-.3cm}where at each node the true propositions are indicated.\pagebreak
-
+		\vspace*{-.3cm}where at each node the true propositions are indicated.
+		%	
 		There is a corresponding classical counter-model $\mathcal M = (M, I)$ to $\varphi^\circ$ with $M = \{u, u_l, u_r\}$, $A^{I}(l) = D^I(l) = B^I(r) = C^I(r) = 1$ and $0$ else.
-
-		Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (A\to B)\to C$ this corresponds to a counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below
+		%
+		Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (B\to A)\to D$, this corresponds to a transformed counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below:
 		\begin{center}
@@
 
+	\begin{remark}
+		It is well known that Kripke frames which are trees are complete with regard to intuitionistic logic.
+Our contribution is giving an explicit geometry that depends only on the structure of the considered formula.
+	\end{remark}
+
 	So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
-		x&\text{ if $\psi$ occurs in $x$}\\
-		x\psi&\text{ otherwise}
-	\end{cases}$$ Now we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula.
-This gives us the following translation:
-
-	\begin{definition}
-		Suppose $\mathcal S$ is a set of formulas $\psi$ of the form
-		$$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi),(A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$
-		Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$.
-For each atom $A$ occurring in $\mathcal S$ and $k\in \mathcal F_\to$ consider a new atom $A^{k}$, which will correspond to $A^{I_F}(k)$ from the previous definition.
-Obtain $\mathcal S^\#$ by including the formulas
-		\begin{itemize}
-			\item $A^k\to A^{k\psi}$ for each $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
-			\item $A^k_\psi\to (B^k_\psi\circ C^k_\psi)$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
-			\item $(A^k_\psi\circ B^k_\psi)\to C^k_\psi$ for each $\circ\in\{\wedge,\vee\}$, $A_\psi\to (B_\psi\circ C_\psi)\in\mathcal S$, $k\in\Lambda$.
-			\item $(A^{k\psi}_\psi\to B^{k\psi}_\psi)\to C^k_\psi$ for $(A_\psi\to B_\psi)\to C_\psi\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
-		\end{itemize}
-	\end{definition}
+		x, &\text{ if $\psi$ occurs in $x$,}\\
+		x\psi, &\text{ otherwise.}
+	\end{cases}$$
+	Now we can replace every $\forall$-quantifier by enumerating all the ground terms, and then replace all distinct ground instances of relations by new propositional variables yielding a propositional formula.
+This gives us the following theorem.
 
 	\begin{theorem}
-		$$\mathcal S\models_C P\text{ iff }\mathcal S^\#\models_I P^\emptyset$$
+		Let $\mathcal S$ be a set of formulas of the form
+		$$(A\wedge B)\to C, (A\vee B)\to C, A\to (B\wedge C), A\to(B\vee C), (A\to B)\to C$$
+		and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
+For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
+Obtain $\mathcal S^\#$ by including the following formulas:
+		\begin{itemize}
+			\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
+			\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+			\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+			\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
+		\end{itemize}
+		Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
+The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$.
+Futhermore, there is an effective procedure for translating counter-models between the sentences.
 	\end{theorem}
@@
 		Let us consider the law of excluded middle $\varphi = A\vee\neg A$. Then we get
-		$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\}$$
-		and we know that $\varphi$ is valid iff $\bigwedge \mathcal S(\varphi)\to P_\varphi$ is valid.
-Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
+		$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\},$$
+		and we know that $\varphi$ is valid iff $\bigwedge \mathcal S^-(\varphi)\to P_\varphi$ is valid.
+Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically valid, where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
 		\begin{align*}
-			\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi}\}\\ &\cup\{(P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi},  A^\epsilon\to P_A^\epsilon\}\\ &\cup \{P_A^{\psi}\to P_A^{\psi}, (A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}
+			\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi}, (P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi}\}\cup\\ &\{A^\epsilon\to P_A^\epsilon, P_A^{\psi}\to P_A^{\psi}, (A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}.
 		\end{align*}
-		It is easy to construct a counter example:
-		$$P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$$
-		$$P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$$
+		We now note that $P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$, $P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$ defines a counter-model.
 	\end{example}
@@
 
-	It is well known that a polynomial time translation between QBF and IPC must exists, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}.
-However giving an explicit translation from IPC to QBF can still be useful as good solvers exists for QBF but for IPC not so much.
-The translation from the last section will serve as a useful tool for this.
-We will not explain the semantics of QBF in depth.
-Intuitively just think that the interpretation of existentially quantified variables may depend on variables quantified above them.
-
-	Fix some set $\mathcal S$ of formulas $\psi$ of the form $$A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$and$$\varphi = \bigwedge \mathcal S\to P$$
+	It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}.
+	However, giving an explicit translation from IPC to QBF still promises to be useful, allowing to leverage the progress in QBF solving.
+	The translation from the last section will serve as starting point for developing the translation in this section.
+	For lack of space, we do not define the semantics of QBF in this paper and refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
+
+	Fix some set $\mathcal S$ of formulas of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
 	and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$.
 We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model.
-The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
+The idea how to get a polynomially-sized formula is: Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
 The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e.
@@
 		Have $k = |\mathcal F_\to|$ and define$$\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal S}(i)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
-		for $0 < i < k$ as well as the special cases
+		for $0 < i < k$, as well as the special cases
 		$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
-		for leafs and
-		and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
-		Example encodings of the above formulas would be
+		for leafs, and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
+		Example encodings of the above formulas are:
 		\begin{align*}
-			\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\wedge\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
+			\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
 			\text{\normalfont Persistent}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
 			\text{\normalfont Sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
-			&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\to C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
+			&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
 			\text{\normalfont Sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
@@
 	\begin{example}
-		With the above encodings for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
+		With the previous encoding for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
 		$$\hspace*{-.3cm}
-		\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right)
+		\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right),
 		$$
@@
 	\end{lemma}
+
 	\begin{proof}
-		Suppose $\varphi$ is not intuitionistically valid, i.e.
-there exists a counter-model $\mathcal M$ for $\mathcal S^\#\to P^\epsilon$.
+		Suppose $\varphi$ is not intuitionistically valid.
+		Then, $\bigwedge\mathcal S\to P$ has an intuitionistic counter-model. Then by the previous section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
+		We now define a QBF model iteratively.
 		For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$.
-Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.
-
+Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$  in $\mathcal M$.
 		Let the $F^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified).
-If not exactly one is interpreted as true then valid$(n)$ fails, i.e.
-have $\psi_n\in\mathcal F_\to$ the only element with $F^n_{\psi_n} = 1$.
-If $F^i_{\psi_n} = 1$ for some $i < n$ then valid$(n)$ also fails, so we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions.
+If not exactly one of the $F^{n}_\psi$ is interpreted as true, then valid$(n)$ fails and the remaining propositions can be chosen arbitrarily.
+If $F^i_{\psi_n} = 1$ for some $i < n$, then valid$(n)$ also fails and the remaining propositions can be chosen arbitrarily.
+		So we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions.
 		Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{\psi_1\dots\psi_n}$.
-Continue this construction until $n  = |\mathcal F_\to|$.
-Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.
+		Continue this construction until $n  = |\mathcal F_\to|$. Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.
 
@@
 
-	Since we have not discussed QBF in detail the above argument might be a bit difficult to follow, in particular the second part, but is simple in essence. We shall demonstrate with help of a example.
-
-	\begin{example}
-		Consider the formula $\varphi = (((A\to B)\to C)\wedge((B\to A)\to C))\to C$, which is a non-intuitionistic tautology.
-Its associated QBF is $$\varphi^Q = \exists A^0B^0C^0\forall F_{\psi_1}^1F_{\psi_2}^1 \exists A^1B^1C^1\forall F_{\psi_1}^2F_{\psi_2}^2 \exists A^2B^2C^2\psi'$$where $\psi_1 := (A\to B)\to C, \psi_2 = (B\to A)\to C$ and $$\psi' = \bigwedge\left\{\begin{matrix}
-			F^1_{\psi_1}\vee F^1_{\psi_2}, \neg(F^1_{\psi_1}\wedge F^1_{\psi_2}), F^2_{\psi_1}\vee F^2_{\psi_2},\neg(F^2_{\psi_1}\wedge F^2_{\psi_2}),\\F^1_{\psi_1}\to\neg F^2_{\psi_1},F^1_{\psi_2}\to\neg F^2_{\psi_2}, A^0\to A^1,A^1\to A^2,
-			\\B^0\to B^1, B^1\to B^2, C^0\to C^1, C^1\to C^2,\neg C^0, \\
-			F^1_{\psi_1}\to (A^1\to B^1)\to C^0, F^1_{\psi_2}\to(B^1\to A^1)\to C^0\\
-			F^2_{\psi_1}\to (A^2\to B^2)\to C^1, F^2_{\psi_2}\to(B^2\to A^2)\to C^1
-		\end{matrix}\right\}$$
-		This QBF is satisfiable: Have $P^0 = 0$ for all atoms $P$, always $P^2 = 1$ for all atoms $P$ and $C^1 = 1$.
-Furthermore have $A^1 = 1$ iff $F^1_{\psi_1} = 1$ and $B^1 = 1$ iff $F^1_{\psi_2} = 1$.
-The construction of an counter-model to $\varphi^\#$ then proceeds as follows:
-		\begin{itemize}
-			\item First have $P^\epsilon = 0$ for each propositional variable $P$.
-			\item For $P^{\psi_1}$ we consider what happens when we choose $F^1_{\psi_1}$ to be true, i.e. we have $A^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
-			\item For $P^{\psi_2}$ we consider what happens when we choose $F^1_{\psi_2}$ to be true, i.e. we have $B^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
-			\item Finally for $P^{\psi_1\psi_2} = P^{\psi_2\psi_1} = 1$ for all propositional variables $P$ as $P^2$ is true regardless of choice for the $F^i_\psi$.
-		\end{itemize}
-	\end{example}
-
-	Note that the last set of existentially quantified variables is superfluous because there is only one assignment which even has the chance to falsify $\varphi^Q$, namely that which assigns $1$ to the only $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$.
-That is in our final formula we can replace $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^i$.
+	A simple counting argument shows that for each $n < |\mathcal F_\to|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal F_\to|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal S}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal F_\to}(n)|\in \mathcal O(|\mathcal F_\to|)$ so overall we get:
+	\begin{lemma}
+		The size of $\varphi^Q$ is in $\mathcal O(|\varphi|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$.
+	\end{lemma}
+
+	Note that the last set of universally quantified variables can be avoided because there is only one assignment that has the chance to falsify $\varphi^Q$, namely the assignment that assigns $1$ to the single $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$.
+Hence, in our final formula we can replace every $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^n$.
 With that we get the following:
@@
 	\begin{corollary}
-		Let $N$-INTINVALID be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid.
-Then $N$-INTINVALID is in $\Sigma_{2N-1}^P$.
-The dual problem $N$-INTVALID is in $\Pi_{2N-1}^P$.
+		Let $N$-\textbf{IntInvalid} be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid.
+Then $N$-\textbf{IntInvalid} is in $\Sigma_{2N-1}^P$.
+The dual problem $N$-\textbf{IntValid} is in $\Pi_{2N-1}^P$.
 	\end{corollary}
@@
 	We now give an analogous transformation for first order logic.
-However our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic.
+Our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic.
 However we do it nonetheless!
@@
 
-	First of all recall that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
-
-	$$\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),$$$$ \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),
-	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)),$$$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),$$$$ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))$$
-	As in the propositional case we will now encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
+	First of all recall, that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
+	\[\begin{matrix}
+		\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c)))\indent \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c))\indent \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c)))\\
+		\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c))\indent \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\indent\forall \vec z(\forall xA(\vec a)\to B(\vec b))\\ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b)).
+	\end{matrix}\]
+	The complete result that we will show is as follows:
+	\begin{theorem}\label{fullFOtranslation}
+		Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain the formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$.
+Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
+For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$.
+For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$.
+Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
+Obtain $\mathcal S^\#$ by including formulas as follows:
+		\begin{itemize}
+			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
+			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
+			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
+			\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\psi = \forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
+			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
+			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
+			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
+		\end{itemize}
+		Then $\bigwedge S\to P$ is intuitionistically iff $\exists xE(x, s)\to K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol, $\Sigma$ is the original signature and
+		\begin{align*}
+			K := &\:\bigwedge\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F\}\wedge\\
+			&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
+			&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}.
+		\end{align*}
+		The size of the obtained formula is linear in the input.
+However $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
+	\end{theorem}
+
+	It is not possible to completely eliminate worlds from the domain as in the propositional case, since counter-models can possibly be infinite (as we will see in more detail later).
+
+	\subsection{Encoding Kripke Semantics}
+
+	As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$, but also $E$ has to be axiomatized. For that we use the following predicates:
 	\begin{itemize}
@@
 	\end{itemize}
-	Then analogously to the predicate case define
+	Then, analogously to the propositional case, we define:
 	\begin{align*}
-		K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \\
-		& \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
-		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge\\
-		& \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
+		K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
+		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
 	\end{align*}
-	Example encodings could be
+	Example encodings of these formulas are as follows:
 	\begin{align*}
@@
 
-	Using $E$ we then obtain $\mathcal S^\circ$ analogously to the predicate case by including formulas:
+	Using $E$ we then obtain $\mathcal S'$ analogously to the predicate case by including formulas:
 
@@
 
-	Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid iff
-	$$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
-	is classically valid where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced, so let's present a proof.
+	Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid if and only if
+	$$\varphi' = \text{World}(s)\to K(\varphi)\to \bigwedge S'\to P^\#(s)$$
+	is classically valid, where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced.
 
@@
 		We proceed by translation of counter-examples.
-Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi^\circ$.
+Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi'$.
 As a Kripke frame $(W, \leq)$ take all $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ let $\leq$ be $\preceq^I$ restricted to $W$.
@@
 		The other direction is a bit more involved.
-Suppose we have a Kripke counter-model to $\varphi$ with Frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
+Suppose we have a Kripke counter-model to $\varphi$ with frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
 In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
 Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
-Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$ and extend $\sim$ by having $[w] = \{w\}$ for $w\in W_0$.
+Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
 		Now have
@@
 			\item $f^I(m_1\dots m_n) =\begin{cases}
-				f^{I_u}(x_1\dots x_n) &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$}\\
-				w_0 & \text{otherwise}
+				f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
+				w_0, & \text{otherwise.}
 			\end{cases}$
 			\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
-				A^{I_u}(x_1\dots x_n) &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$}\\
-				\top & \text{otherwise}
+				A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
+				\top, & \text{otherwise.}
 			\end{cases}$
@@
 
-	We proceed now as in the previous section.
+	\subsection{Reducing the encoding}
+
+	As in the predicate case we now show the existence of a canonical counter-model (in case such a counter-model exists).
 Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$.
 Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
-For each  $\psi\in\mathcal F$ consider a new function symbol $f$.
-As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
-	$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$ with
-	$$\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$$ and
-	$$\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$$ with $$\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u))))\to B^\#(\vec b, u))$$
-	and in each other formula in $\mathcal S^\circ$ remove the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
-	$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
-	As in the previous section we can define a tree-counter model for this.
-	\begin{definition}
+For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$.
+As before we may assume that $s$ is a least element and we can avoid the quantification over $w$ by considering the formulas
+	\begin{itemize}
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$,
+		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$,
+	\end{itemize}
+	and by removing removing the $s\preceq u$ for the remaining formulas in $\mathcal S'$.
+	We denote the resulting set of formulas by $\mathcal S^\#$.
+	Then, the previous formulas are equivalid to
+	$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#(s).$$
+	As in the previous section we can define a tree counter-model for this.
+
+	\begin{definition}
+		Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
+		\begin{itemize}
+			\item $M_T$ are the sequences on $\{(\vec z, \psi)\:|\:\psi\in \mathcal F, \vec z\in M^n\}$.
+			\item We interpret $\preceq$ as the prefix-order.
+			\item We set $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi).$$
+			\item For relation symbols $R^\#$, we set $${R^\#}^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = {R^\#}^I(\vec z, f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots)).$$
+		\end{itemize}
+	\end{definition}
+
+	\begin{lemma}
+		If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
+	\end{lemma}
+
+	Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$f_{\psi_1}(\vec z_1, f_{\psi_2}(\vec z_2, \dots f_{\psi_n}(\vec z_n, s)\dots))$$ is sufficient.
+	In particular, this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
+	from $K(\varphi)$ and adjust DomainSubset$(u, w)$ and Persistent$(u, w)$, e.g. remove quantification over $w$ and the precondition $u\preceq w$ from $K(\varphi)$ and replace DomainSubset$(u, w)$ with
+	$$\bigwedge\{\forall \vec z\forall u\forall x(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$$ and Persistent$(u, w)$ with $$\bigwedge\{\forall\vec z_1\forall\vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\psi\in\mathcal F_\to, \text{$A\in\Sigma$ is a relation symbol}\}.$$ We set
+
+	$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#.$$
+	Then, we obtain the following result:
+	\begin{theorem}
+		\label{thm:fo-reduction}
+		$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
+	\end{theorem}
+	The complete translation is summarised in Theorem~\ref{fullFOtranslation} (stated earlier). Let us understand why a further reduction as in the propositional case is impossible.
+
+	\begin{definition}
+		\label{def:fo-fulfilled}
 		Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
-For $\psi = \forall\vec t\psi'\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff $(A_\psi^I(\vec t_A, u)\to B_\psi^I(\vec t_B, u))[\vec a/\vec t]$ is false or $C_\psi^I(\vec t_C, u)[\vec a/\vec t]$ is true.
-		For $\psi = \forall\vec t\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff there is $a\in M_u$ such that $A^{I[a/x]}_\psi(\vec t_A)[\vec a/\vec t]$ is false or $B_\psi^I(\vec t_B)[\vec a/\vec t]$ is true.
-For $\psi\in\mathcal F$ define $$g_{\vec a, \psi} : M\to M, u\mapsto\begin{cases}
-			u&\text{ if $\psi$ is fulfilled in $u$ with $\vec a$}\\
-			f^I_\psi(\vec a, u)&\text{ else}		
-		\end{cases}$$		
-		Define a $\mathcal M_T = (M_T, I_T)$ as follows:
-		\begin{itemize}
-			\item $M_T$ are the sequences on $\{ \vec t, \psi\:|\:\psi\in \mathcal F, \vec t\in M^n\}$ without repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\to$ and without immediate repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\forall$.
-			\item Interpret $\preceq$ as the prefix-order.
-			\item Have $$f_\psi^{I_T}(x_1\dots x_n) = \begin{cases}
-				x_1\dots x_n&\text{if $x=x_n$}\\
-				x_1\dots x_nx&\text{else}			
-			\end{cases}$$
-			\item For propositional variables $P$ have $$P^{I_T}(x_1\dots x_n) = P^I(g_{x_n}(\dots(g_{x_1}(s^I))\dots))$$
-		\end{itemize}
-	\end{definition}
+For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
+		For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
+	\end{definition}
+
+	We get an analogous result to Lemma~\ref{thm:prop-countermodel-reduction}:
 
 	\begin{lemma}
+		\label{thm:fo-countermodel-reduction}
 		Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
@@
 
-	\begin{proof}
-		For $\psi\in\mathcal F_\to$ the proof is analogous to the previous section.
-
-		1.
-Let $\psi\in\mathcal F_\forall$.
-If $B^I_\psi(\vec b, u)[\vec a/\vec t]$ holds then we are done due to persistency.
-Otherwise there is $a$ with $E^I(a, f^I_\psi(\vec a, u))$ such that $A^{I[a/x]}_\psi(\vec t_A, f^I_\psi(\vec a, u))$ is false.
-So $\psi$ is fulfilled at $f^I_\psi(\vec a, u)$ with $\vec a$.
-	\end{proof}
-
-	Part 2 of this Lemma does not hold for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$) we have $A(x, v)$ there could be some term $t$ with $\neg E(t, v)$ and $w$ with $v\preceq w$ such that $E(t, w)$ and $\neg A(t, w)$.
-This is why we must consider sequences that have repetitions of such formulas:
-
-	\begin{example}
-		Consider the $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model
+	However part 2 of Lemma~\ref{thm:fo-countermodel-reduction} fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$), we have $A(x, v)$, there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
+This is why we must consider sequences that have repetitions:
+
+	\begin{wrapfigure}[16]{r}{0.4\textwidth}
+		\vspace{-32pt}
 		\begin{center}
@@
 				\node      (4l)   [above=of 3l] {{$\vdots$}};
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (0.north east) -- (1r.south west);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (0.north west) -- (1l.south east);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1r.north) -- (2r.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1l.north) -- (2l.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1l.north east) -- (2r.south west);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1r.north west) -- (2l.south east);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2r.north) -- (3r.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2l.north) -- (3l.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2l.north east) -- (3r.south west);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2r.north west) -- (3l.south east);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3r.north) -- (4r.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3l.north) -- (4l.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3l.north east) -- (4r.south west);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3r.north west) -- (4l.south east);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (0.north east) -- (1r.south west);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (0.north west) -- (1l.south east);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1r.north) -- (2r.south);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1l.north) -- (2l.south);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1l.north east) -- (2r.south west);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1r.north west) -- (2l.south east);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2r.north) -- (3r.south);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2l.north) -- (3l.south);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2l.north east) -- (3r.south west);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2r.north west) -- (3l.south east);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3r.north) -- (4r.south);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3l.north) -- (4l.south);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3l.north east) -- (4r.south west);
+				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3r.north west) -- (4l.south east);
 			\end{tikzpicture}
 		\end{center}
-		where at each node the elements that exist at it are indicated.
-For black ones both $A$ and $B$ hold, for red ones only $B$, for blue ones only $A$.
-Furthermore red arrows indicate $f_{\neg \forall xA(x)}$ and blue arrows $f_{\neg \forall xB(x)}$.
-Note that applying $f_{\neg \forall xA(x)}$ we must always reach the left column and $f_{\neg \forall xB(x)}$ we must always reach the right column.
-Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we can never remain stationary.
-This is in contrast to the propositional case.
+		\caption{Counter-model to $\varphi$.}
+		\label{fig1}
+	\end{wrapfigure}
+	\begin{example}
+		Consider $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model as shown in Figure~\ref{fig1},
+		where we indicate at each world the elements that exist at it.
+For black numbers $n$, both $A(n)$ and $B(n)$ hold at that world, for red ones only $A(n)$, for blue ones only $B(n)$.
+The arrows indicate the order.
+$\neg\forall xA(x)$ is not fulfilled at worlds in the left column and $\neg\forall xB(x)$ is not fulfilled at worlds in the right column.
+In particular applying $f_{\neg \forall xA(x)}$ we must always reach the right column and applying $f_{\neg \forall xB(x)}$ we must always reach the left column.
+Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we cannot remain stationary.
 	\end{example}
 
-	\begin{corollary}
-		If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
-	\end{corollary}
-
-	Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$$ is sufficient.
-In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PO-encoding from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$.
-Define
-
-	$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#$$
-	then we have 
-	\begin{theorem}
-		$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
-	\end{theorem}
-	The complete translation is summarised in Theorem~\ref{fullFOtranslation}.
-
-	\begin{example}
-		Consider the double negation shift (DNS)
-		$$\varphi = \forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$$
-		which is a famous non-intuitionistic tautology. We get
-		\begin{align*}
-			\mathcal S = & \{(P_{\forall x\neg\neg A(x)}\to P_{\neg\neg\forall xA(x)})\to P_\varphi, P_{\forall x\neg\neg A(x)}\to \forall xP_{\neg\neg A(x)}(x)\}\cup\\
-			& \{\forall t((P_{\neg\neg A(x)}(t)\wedge P_{\neg A(x)}(t))\to \bot), \forall t((A(t)\to \bot)\to P_{\neg A(x)})(t)\}\cup\\
-			& \{(P_{\neg\forall xA(x)}\to \bot)\to P_{\neg\neg\forall xA(x)}, (P_{\neg\forall xA(x)}\wedge P_{\forall xA(x)})\to \bot\}\cup\\
-			&\{\forall xA(x)\to P_{\forall xA(x)}\}
-		\end{align*}
-		and thus, denoting $\psi_1 = (P_{\neg\forall xA(x)}\to \bot)\to P_{\neg\neg\forall xA(x)}$ and $\psi_2 = \forall xA(x)\to P_{\forall xA(x)}$, we have
-		\begin{align*}
-			\mathcal S^\# = & \{\forall u(P_{\forall x\neg\neg A(x)}(u)\to P_{\neg\neg\forall xA(x)}(u))\to P_\varphi(u)\}\cup\\& \{\forall u(P_{\forall x\neg\neg A(x)}(u)\to \forall xP_{\neg\neg A(x)}(x, u)\}\cup\\
-			& \{\forall t, u((P_{\neg\neg A(x)}(t, u)\wedge P_{\neg A(x)}(t, u))\to \bot)\}\\& \{\forall t, u((A(t, u)\to \bot)\to P_{\neg A(x)}(t, u))\}\cup\\
-			& \{\forall u((P_{\neg\forall xA(x)}(f_{\psi_1}(u))\to \bot)\to P_{\neg\neg\forall xA(x)}(u))\}\\& \{\forall u((P_{\neg\forall xA(x)}(u)\wedge P_{\forall xA(x)}(u))\to \bot)\}\cup\\
-			&\{\forall u(\forall xA(x, f_{\psi_2}(u))\to P_{\forall xA(x)}(u))\}
-		\end{align*}
-	\end{example}
-
-	\chapter{Proof translation}
-
-	\section{Translation of counter-models}
-
-	There have been various results on when classical provability implies intuitionistic provability, most famously for $\Pi_2$-formulas in Arithmetic~\cite{Friedman_1978}, but most recently also in some other cases~\cite{schwichtenberg, practical}.
-The constructibility of a bounded counter-model as in~\ref{finitecountermodel} can be used as an inspiration to give a similar result, i.e.
-roughly when such a construction is possible then classical validity implies intuitionistic validity.
-Note that the results in this section will applicable to predicate logic already.
-
-	For this section we fix some set $\mathcal S$ of (predicate) formulas where each formula $\varphi$ is of the forms $\forall \vec x\psi$ for some $\psi$ of the form
-	$$A_\psi\to(B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi)$$$$(A_\psi\vee B_\psi)\to C_\psi, A_\psi\to B_\psi\to C_\psi, (A_\psi\to B_\psi)\to  C_\psi$$$$\forall xA_\psi\to B_\psi, A_\psi\to \forall xB_\psi, \exists xA_\psi\to B_\psi, A_\psi\to\exists xB_\psi$$
-	for some atomic relations $A_\psi, B_\psi, C_\psi$. We write $\varphi(\vec t)$ for $\psi[\vec t/\vec x]$. Note that we can interpret $A\to B$ as e.g. $A\to (B\wedge B)$.
-
-	\subsection{Dependence}
-
-	\begin{definition}
-		A \textit{dependence order} $\precsim$ (w.r.t.
-$\mathcal S$) is a pre-order on a subset of ground predicates such that for all $\varphi\in\mathcal S$, ground predicates $A, B, C$ and ground terms $\vec t$
-		\begin{itemize}
-			\item If $\varphi(\vec t) = A\to (B\wedge C)$ then $A\precsim B$ and $A\precsim C$.
-			\item If $\varphi(\vec t) = (A\wedge B)\to C$ then for each $D$ with $D\precsim B$ and $D\precsim C$ we have $D\precsim A$.
-			\item If $\varphi(\vec t) = A\to (B\vee C)$ then for each $D$ with $B\precsim D$ and $C\precsim D$ we have $A\precsim D$.
-			\item If $\varphi(\vec t) = (A\vee B)\to C$ then $A\precsim C$ and $B\precsim C$.
-			\item If $\varphi(\vec t) = A\to \forall xB$ then $A\precsim B[t/x]$ for every ground term $t$.
-			\item If $\varphi(\vec t) = \exists xA\to B$ then $A[t/x]\precsim B$ for every ground term $t$.
-			\item If $\varphi(\vec t) = A\to \exists xB$ then for each $D$ with $B[t/x]\precsim D$ for each $t$ we have $A\precsim D$.
-		\end{itemize}
-		We say that a dependence order is \textit{sufficient} for some ground predicate $P$ if for each $\varphi\in\mathcal S$, ground predicates $A, B, C$ and ground terms $\vec t$
-		\begin{itemize}
-			\item If $\varphi(\vec t) = (A\to B)\to C$ and $C\precsim P$ then $A\not\precsim P$ and $\precsim$ is sufficient for $B$.
-			\item If $\varphi(\vec t) = \forall xA\to B$ and $B\precsim P$ then $\precsim$ is sufficient for $A[t/x]$ for every ground term $t$.
-		\end{itemize}
-	\end{definition}
-
-	\begin{theorem}~\label{dependencetheorem}
-		Consider some formula $\varphi = \bigwedge\mathcal S\to P$.
-If there is a dependence order w.r.t.
-$\mathcal S$ that is sufficient for $P(s)$ and $\varphi$ is classically valid then $\varphi$ is intuitionistically valid.
-	\end{theorem}
-
-	\begin{example}
-		Consider the formula $(A\to B)\vee (B\to A)$. We know from the previous section that this is intuitionistically equivalid to $$\bigwedge\mathcal S\to P_\varphi(s)$$
-		where
-		\begin{align*}
-			\mathcal S = &\{\forall u((P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u))\}\\
-			&\cup \{\forall u((P_A(f(u))\to P_B(f(u)))\to P_{A\to B}(u))\}\\
-			&\cup \{\forall u((P_B(f(u))\to P_A(f(u)))\to P_{B\to A}(u))\}\\
-			&\cup \{\forall u(P_A(u)\to A(u)), \forall u(A(u)\to P_A(u))\}\\
-			&\cup \forall u(P_B(u)\to B(u)), \forall u(B(u)\to P_B(u))\}
-		\end{align*}
-		Consider the following dependence order
-
-		\begin{center}
-			\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
-				\node[squarednode]      (upper)                          {$P_\varphi(s), P_{A\to B}(s), P_{B\to A}(s)$};
-				\node[squarednode]      (lower)       	[below=of upper] {$P_A(f(s)), P_B(f(s)), A(f(s)), B(f(s))$};
-				\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north) -- (upper.south);
-			\end{tikzpicture}
-		\end{center}
-
-		i.e. for all $P, Q$ in the same rectangle have $P\precsim Q$ and $P\precsim Q$, and for $P$ in the lower $Q$ in the upper $P\prec Q$.
-	\end{example}
+	Let us add that Lemma~\ref{thm:fo-countermodel-reduction} still equips us with additional information that could be used to guide proof search: We could add additional axioms that restrict the interpretation of functions.
+That is, we may assume that $f_\psi(w) = w$, if $\exists u(w = f_\psi(u))$ for all $\psi\in\mathcal F_\forall$ (i.e.
+that $f_\psi$ is idempotent), and $f_\psi(w) = w$, if $\exists u(w \geq f_\psi(u))$ for all $\psi\in\mathcal F_\to$ (which is even stronger than idempotence).
 
 	\chapter{Implementation}
+
+	\chapter{Translation of proofs and counter-models}
 

### 2022-08-01 Implementation/eiicl/src/encoding.rs

@@ -1,0 +1,1322 @@
+use std::collections::HashMap;
+use tptp::visitor::Visitor;
+use tptp::{common, fof, top, TPTPIterator};
+
+#[derive(Hash, Debug)]
+pub enum PlainTerm {
+    Variable {
+        name: String,
+    },
+    Function {
+        name: String,
+        arguments: Vec<PlainTerm>,
+    },
+}
+
+impl PlainTerm {
+    fn to_tptp(&self) -> fof::Term {
+        match self {
+            PlainTerm::Variable { name } => {
+                fof::Term::Variable(common::Variable(common::UpperWord(name)))
+            }
+            PlainTerm::Function { name, arguments } => fof::Term::Function(Box::new(
+                fof::FunctionTerm::Plain(fof::PlainTerm::Function(
+                    common::Functor(common::AtomicWord::SingleQuoted(common::SingleQuoted(name))),
+                    Box::new(fof::Arguments(
+                        arguments.iter().map(|x| x.to_tptp()).collect(),
+                    )),
+                )),
+            )),
+        }
+    }
+}
+
+#[derive(Hash, Debug)]
+pub struct AtomizedSubformula {
+    name: String,
+    arguments: Vec<PlainTerm>,
+}
+
+impl AtomizedSubformula {
+    fn to_tptp(&self) -> fof::AtomicFormula {
+        fof::AtomicFormula::Plain(fof::PlainAtomicFormula(fof::PlainTerm::Function(
+            common::Functor(common::AtomicWord::SingleQuoted(common::SingleQuoted(
+                &self.name,
+            ))),
+            Box::new(fof::Arguments(
+                self.arguments.iter().map(|x| x.to_tptp()).collect(),
+            )),
+        )))
+    }
+}
+
+#[derive(Hash, Debug)]
+pub enum NormalizedFormula<'a> {
+    LeftAnd {
+        left: Vec<AtomizedSubformula>,
+        right: AtomizedSubformula,
+    },
+    RightAnd {
+        left: AtomizedSubformula,
+        right: Vec<AtomizedSubformula>,
+    },
+    LeftOr {
+        left: Vec<AtomizedSubformula>,
+        right: AtomizedSubformula,
+    },
+    RightOr {
+        left: AtomizedSubformula,
+        right: Vec<AtomizedSubformula>,
+    },
+    DoubleImplication {
+        left: (AtomizedSubformula, AtomizedSubformula),
+        right: AtomizedSubformula,
+    },
[…]

### 2022-08-01 Implementation/eiicl/src/main.rs

@@ -1,0 +1,28 @@
+mod encoding;
+
+use std::fs::File;
+use std::io::Read;
+use std::path::Path;
+
+fn example(bytes: &[u8]) {
+    let encoding = encoding::Encoding::from_tptp(bytes);
+    for annotated in encoding.to_tptp() {
+        println!("{}", annotated);
+    }
+}
+
+fn main() {
+    let path = Path::new("example.p");
+    let display = path.display();
+
+    let mut file = match File::open(&path) {
+        Err(why) => panic!("couldn't open {}: {}", display, why),
+        Ok(file) => file,
+    };
+
+    let mut s = String::new();
+    match file.read_to_string(&mut s) {
+        Err(why) => panic!("couldn't read {}: {}", display, why),
+        Ok(_) => example(s.as_bytes()),
+    }
+}

### 2022-08-01 Implementation/formula.py

@@ -1,97 +1,0 @@
-from ast import arguments
-from term import Term, Signature
-
-from functools import reduce
-from typing import Dict, Set
-
-
-class Signature:
-    def __init__(self, atoms: Dict[str, int], connectors: Dict[str, int], quantifiers: Set[str], termType: type):
-        self._d = {**atoms, **connectors}
-        self.termType = termType
-        self.atoms = set(atoms.keys()) | {c for c, a in connectors.items() if a == 0}
-        self.connectors = {c for c, a in connectors.items() if a > 0}
-        self.quantifiers = quantifiers
-        self.symbols = self.connectors | self.atoms | self.quantifiers
-        self.arity = lambda x: self._d[x]
-
-    def __repr__(self):
-        return f"""
-        Atoms: {", ".join([f"{s}/{self.termArity(s)}" for s in self.atoms])}
-        Connectors: {", ".join([f"{s}/{self.formulaArity(s)}" for s in self.connectors])}
-        Quantifiers: {", ".join([f"{s}" for s in self.quantifiers])}
-        """
-
-    def __eq__(self, other) -> bool:
-        return self._d == other._d
-
-def Formula(sig: Signature):
-    class _Formula:
-        signature = sig
-        
-        def __init__(self, name: str, arguments = [], boundVariables = []):
-            if name in _Formula.signature.quantifiers:
-                self.name = name
-                self.boundVariables = boundVariables
-                self.arguments = arguments
-            else:
-                if name in _Formula.signature.atoms | _Formula.signature.connectors:
-                    assert len(arguments) == _Formula.signature.arity(name)
-                    assert all(type(arg) == _Formula.signature.termType if name in _Formula.signature.atoms else type(arg) == type(self) for arg in arguments) 
-                    self.name = name
-                    self.arguments = arguments
-                else:
-                    assert len(arguments) == 0
-                    self.name = name
-                    self.arguments = []
-
-        def __repr__(self) -> str:
-            if self.name not in _Formula.signature.quantifiers:   
-                return self.name + (("(" + ", ".join(map(str, self.arguments)) + ")") if self.arguments else "")
-            else:
-                return self.name + (("[" + ", ".join(map(str, self.boundVariables)) + "]")if self.boundVariables else "") + (("(" + ", ".join(map(str, self.arguments)) + ")") if self.arguments else "")
-
-
-        def __eq__(self, other) -> bool:
-            return self.name == other.name and self.arguments == other.arguments
-
-        def __hash__(self) -> int:
[…]

### 2022-08-01 Implementation/proof.py

@@ -1,31 +1,0 @@
-from typing import Any, Dict
-from formula import Signature, Formula
-
-class Logic:
-    def __init__(self, sig : Signature, rules : Dict[str, Any]):
-        self.signature = Signature
-        self.rules = rules
-
-    def addAxiom(self, name, ax):
-        self.rules[name] = lambda : ax
-
-def Proof(log: Logic):
-    class _Proof:
-        logic = log
-        
-        def __init__(self, name: str, arguments = [], parameters = []):
-            self.name = name
-            self.arguments = arguments
-            self.parameters = parameters
-            self.head = self.logic.rules[name](**arguments, **parameters)
-
-        def __repr__(self) -> str:
-            return self.name + (("(" + ", ".join(map(str, self.arguments)) + ")") if self.arguments else "")
-
-        def __eq__(self, other) -> bool:
-            return self.name == other.name and self.arguments == other.arguments
-
-        def __hash__(self) -> int:
-            return hash(repr(self))
-
-    return _Proof

### 2022-08-01 Implementation/term.py

@@ -1,83 +1,0 @@
-from dataclasses import dataclass
-from functools import reduce
-from typing import Dict
-
-
-class Signature:
-    def __init__(self, d: Dict[str, int]):
-        self._d = d
-        self.constants = {c for c, a in self._d.items() if a == 0}
-        self.functions = {c for c, a in self._d.items() if a > 0}
-        self.symbols = self.constants | self.functions
-        self.arity = lambda x: self._d[x]
-
-    def __repr__(self):
-        return "[" + ", ".join([f"{s}/{self.arity(s)}" for s in self.symbols]) + "]"
-
-    def __eq__(self, other) -> bool:
-        return self._d == other._d
-
-
-def Term(sig: Signature):
-    class _Term():
-        signature = sig
-
-        def __init__(self, name: str, arguments = []):
-            assert all(type(self).signature == type(arg).signature for arg in arguments)
-            if name in _Term.signature.symbols:
-                assert len(arguments) == _Term.signature.arity(name)
-                self.name = name
-                self.arguments = arguments
-            else:
-                assert len(arguments) == 0
-                self.name = name
-                self.arguments = []
-
-        def __repr__(self) -> str:
-            return self.name + (("(" + ", ".join(map(str, self.arguments)) + ")") if self.arguments else "")
-
-        def __eq__(self, other) -> bool:
-            return self.name == other.name and self.arguments == other.arguments
-
-        def __hash__(self) -> int:
-            return hash(repr(self))
-        
-        def isVariable(self):
-            return self.name not in _Term.signature.symbols
-
-        def variables(self) -> set():
-            return (set() if self.name in _Term.signature.symbols else set([self])) | reduce(lambda x, y: x | y, map(lambda arg: arg.variables(), self.arguments), set())
-
-        def superterm(self, other):
-            return self == other or any([arg.superterm(other) for arg in self.arguments])
-
-        def subterm(self, other):
-            return other.superterm(self)
-
-        def substitute(self, d):
-            if self in d.keys():
-                return d[self]
-            else:
-                return _Term(name=self.name, arguments=list(map(lambda x: x.substitute(d), self.arguments)))
-
-        def mgu(self, other):
-            if self == other:
-                return {}
-            if other.isVariable():
-                if other not in self.variables():   
-                    return {other: self}
-                else:
-                    return None
-            if self.isVariable():
-                if self not in other.variables():
-                    return {self: other}
-                else:
-                    return None
[…]

### 2022-08-01 Implementation_old/formula.py

@@ -1,97 +1,0 @@
-from dataclasses import dataclass, field
-from typing import Sequence
-from term import Term, Variable
-
-class Formula:  
-    def variables(self):
-        vars = {}
-        for arg in self.arguments:
-            if vars:
-                vars.update(arg.variables())
-            else:
-                vars = arg.variables()
-        return vars
-    
-    def substitute(self, d):
-        if type(self) == Atom:
-            return type(self)(name=self.name, arguments=[arg.substitute(d) for arg in self.arguments])
-        if type(self) in [Conjunction, Disjunction, Implication, Equivalence, Negation]:
-            return type(self)(arguments=[arg.substitute(d) for arg in self.arguments])
-        if type(self) in [Forall, Exists]:
-            return type(self)(bound_variables=self.bound_variables, arguments=[self.arguments[0].substitute({k : d[k] for k in d.keys() if k not in self.bound_variables})])
-        else:
-            return self
-
-    def __str__(self):
-        if type(self) == Atom:
-            return self.name + '(' + ','.join(list(map(str, self.arguments))) +')'
-        if type(self) == Conjunction:
-            return '(' + ') & ('.join(list(map(str, self.arguments))) + ')'
-        if type(self) == Disjunction:
-            return '(' + ') | ('.join(list(map(str, self.arguments))) + ')'
-        if type(self) == Implication:
-            return '(' + ') => ('.join(list(map(str, self.arguments))) + ')'
-        if type(self) == Equivalence:
-            return '(' + ') <=> ('.join(list(map(str, self.arguments))) + ')'
-        if type(self) == Negation:
-            return '~(' + str(self.arguments[0]) + ')'
-        if type(self) == BTrue:
-            return 'True'
-        if type(self) == BFalse:
-            return 'False'
-        if type(self) == Forall:
-            return '! [' + ','.join(list(map(str, self.bound_variables))) + ']: ' + str(self.arguments[0])
-        if type(self) == Exists:
-            return '? [' + ','.join(list(map(str, self.bound_variables))) + ']: ' + str(self.arguments[0])
-
-    def normalize(self):
-        if type(self) == Forall:
-            return self.arguments[0].substitute({x : Variable("input" + x.name[1:], []) for x in self.bound_variables}).normalize()
-        if type(self) == Exists:
[…]

### 2022-08-01 Implementation_old/parse.py

@@ -1,118 +1,0 @@
-from parsimonious.grammar import Grammar, NodeVisitor
-from formula import Disjunction, Conjunction, Implication, Equivalence, Forall, Exists, Negation, Atom, BTrue, BFalse
-from term import Variable, FunctionApplication, Constant
-
-
-proof = Grammar(
-    """
-    proof = "\\n"? (rule "\\n"?)*
-    rule = id ". " formula " [" rulename id* "]"
-    id = ~"[0-9]+"","?
-    rulename = ~"[a-z\s]+"
-    formula = disjunction / conjunction / implication / equivalence / parensFormula / quantifiedFormula / negation / atom
-    disjunction = (quantifiedFormula / negation / atom / parensFormula) (" | " (quantifiedFormula / negation / atom / parensFormula))+
-    conjunction = (quantifiedFormula / negation / atom / parensFormula) (" & " (quantifiedFormula / negation / atom / parensFormula))+
-    implication = (quantifiedFormula / negation / atom / parensFormula) (" => " (quantifiedFormula / negation / atom / parensFormula))+
-    equivalence = (quantifiedFormula / negation / atom / parensFormula) (" <=> " (quantifiedFormula / negation / atom / parensFormula))+
-    parensFormula = "(" formula ")"
-    quantifiedFormula = forall / exists
-    forall = "! [" variable+ "] : " (parensFormula / quantifiedFormula / negation / atom)
-    exists = "? [" variable+ "] : " (parensFormula / quantifiedFormula / negation / atom)
-    negation = "~" (quantifiedFormula / atom / parensFormula)
-    atom = equality / inequality / predicate / bool
-    equality = term " = " term 
-    inequality = term " != " term
-    predicate = (name "(" term+ ")")
-    bool = "$true" / "$false"
-    term = (functionApplication / variable / constant)","?
-    functionApplication = (skolemName / name) "(" term+ ")"
-    name = ~"[a-z A-Z]([a-z A-Z 0-9]*)"
-    variable = ~"X([0-9]+)"","?
-    constant = skolemName / name
-    skolemName = ~"sK([0-9]+)"
-    """
-)
-
-class ProofVisitor(NodeVisitor):
-    def __init__(self) -> None:
-        self.result = {}
-
-    def visit_rule(self, node, visited_children):
-        id, _, formula , _, rulename, ids, _ = visited_children
-        self.result[id] = (rulename, [] if ids == '' else ids, formula)
-
[…]

### 2022-08-01 Implementation_old/proof.py

@@ -1,248 +1,0 @@
-from dataclasses import dataclass
-from formula import BFalse, BTrue, Formula, Disjunction, Negation, Atom, Forall, Exists
-from typing import Dict, Sequence
-from term import Term, Variable
-
-class Proof:
-    def parse(d):
-        head_id = max(d.keys())
-        if d[head_id][0] == 'input':
-            return Input([], None, d[head_id][2])
-        if d[head_id][0] == 'negated conjecture ':
-            return NegatedConjecture([], None, d[head_id][2])
-        if d[head_id][0] == 'rectify ':
-            return Rectify([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
-        if d[head_id][0] == 'ennf transformation ':
-            return EnnfTransformation([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
-        if d[head_id][0] == 'nnf transformation ':
-            return NnfTransformation([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
-        if d[head_id][0] == 'cnf transformation ':
-            return CnfTransformation([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
-        if d[head_id][0] == 'flattening ':
-            return Flattening([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
-        if d[head_id][0] == 'choice axiom':
-            return ChoiceAxiom([], None, d[head_id][2])
-        if d[head_id][0] == 'equality resolution ':
-            return EqualityResolution([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
-        if d[head_id][0] == 'skolemisation ':
-            return Skoemisation([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
-        if d[head_id][0] == 'superposition ':
-            return Superposition([Proof.parse({k : d[k] for k in d.keys() if k <= id}) for id in d[head_id][1]], None, d[head_id][2])
-        if d[head_id][0] == 'forward demodulation ':
[…]

### 2022-08-01 Implementation_old/term.py

@@ -1,75 +1,0 @@
-from ast import arguments
-from dataclasses import dataclass, field
-from typing import Sequence
-
-class Term:
-    def variables(self):
-        if type(self) == FunctionApplication:
-            vars = None
-            for arg in self.arguments:
-                if vars:
-                    vars.update(arg.variables())
-                else:
-                    vars = arg.variables()
-            return vars
-        if type(self) == Constant:
-            return {}
-        if type(self) == Variable:
-            return {self}
-
-    def substitute(self, d):
-        if not d:
-            return self
-        if type(self) == FunctionApplication:
-            return FunctionApplication(self.name, arguments=[arg.substitute(d) for arg in self.arguments])
-        if type(self) == Constant:
-            return Constant(self.name, [])
-        if type(self) == Variable:
-            if self in d.keys():
-                return(d[self])
-            else:
-                return(Variable(self.name, []))
-
-    def mgu(self, f):
-        if type(self) == Variable:
-            return ({self : f}, {})
-        if type(f) == Variable:
-            return ({}, {f : self})
-        if type(self) == FunctionApplication and type(f) == FunctionApplication:
-            s = [{}, {}]
-            for arg1, arg2 in zip(self.arguments, f.arguments):
-                arg1, arg2 = arg1.substitute(s[0]), arg2.substitute(s[1])
-                si = arg1.mgu(arg2)
-                for i in [0, 1]:
-                    if s[i]:
-                        s[i].update(si[i])
-                    else:
-                        s[i] = si[i]
-            return s
-        return ({}, {})
-
-    def __str__(self):
-        if type(self) in [Variable, Constant]:
-            return self.name
-        if type(self) == FunctionApplication:
-            return self.name + '(' + ','.join(list(map(str, self.arguments))) + ')'
-
-
-
-@dataclass
-class Variable(Term):
-    name : str
-    arguments : Sequence[Term] = field(default_factory = list)
-
-    def __hash__(self):
-        return hash(self.name)
-
-@dataclass
-class Constant(Term):
-    name : str
-    arguments : Sequence[Term] = field(default_factory = list)
-
-@dataclass
-class FunctionApplication(Term):
-    name : str
-    arguments : Sequence[Term]

### 2022-08-17 Notes/Embedding intuitionistic into classical logic/EIICL.tex
First Final Thesis Draft

@@
-\documentclass[a4paper,12pt]{report}
-
-\title{Embedding intuitionistic into classical logic}
-\author{Alexander Pluska}
-
+\documentclass[a4paper,11pt]{report}
+
+\usepackage{setspace}
 \usepackage{amsmath,amssymb,amsthm}
@@
 \usepackage{stackengine}
-\usepackage{bbold}
 \usepackage{footmisc}
 \usepackage{wrapfig}
+\usepackage[a4paper,top=2.5cm,bottom=3.5cm,left=2.5cm,right=2.5cm,headsep=10pt]{geometry}
+\usepackage[toc,page]{appendix}
+\usepackage{listings}
+\lstset{
+	language=bash,
+	basicstyle=\ttfamily
+}
 
@@
 
-\DeclareMathOperator*{\argmax}{arg\,max}
-
 \begin{document}
 
-	\maketitle
-
-	\begin{abstract}
-		We present an embedding of intuitionistic into classical logic, a quasi-inverse to the famous double-negation translation, i.e.
-a procedure that for each formula $\varphi$ gives a formula $\varphi_c$ such that $\varphi$ is intuitionistically valid if and only if $\varphi_c$ is classically valid.
-We consider both the propositional and predicate case.
-Since many of the more involved results for the more interesting predicate scenario involve similar arguments as in the propositional case we shall develop them simultaneously.
-	\end{abstract}
-
+	\pagenumbering{gobble}
+	\begin{flushright}
+		% https://public.univie.ac.at/en/downloads/logos/
+		\includegraphics[width=7cm]{RZ_Logo_Uni_sw}
+	\end{flushright}	
+	\vskip20mm
+	\begin{center}
+		\huge\textbf{MASTERARBEIT / MASTER'S THESIS}
+		\vskip15mm
+		\normalsize{Titel der Masterarbeit / Title of the Master's Thesis}
+
+		\LARGE\textbf{``Automated Deduction for Intuitionistic Logic\\via Embedding into Classical Logic''} % Insert title here
+		\vskip15mm
+		\normalsize{verfasst von / submitted by}
+
+		\Large\textbf{Alexander Pluska, B.Sc.} % Insert your name in this line
+		\vskip15mm
+		\normalsize{angestrebter akademischer Grad / in partial fulfillment of the requirements for the degree of}
+
+		\Large\textbf{Master of Science (M.Sc.)}
+
+		\small
+	\end{center}
+	\vskip15mm
+	\begingroup
+	\renewcommand{\arraystretch}{1.2}
+	\begin{flushleft}
+		Wien, 2022 / Vienna 2022 % Insert graduation year in this line
+		\vfill
+		\addtolength{\tabcolsep}{-6pt} 
+		\begin{table}[h!]
+			\begingroup
+			\renewcommand*{\arraystretch}{1.2}
+			%\renewcommand*{\linespacing}{1}
+			\renewcommand\baselinestretch{1}
+			\small
+			\begin{tabular}{p{7cm} p{10cm}}
+				Studienkennzahl lt. Studienblatt / \newline degree programme code as it appears on \newline the student record sheet: &A 066 821 \\
+				Studienrichtung lt. Studienblatt / \newline degree programme as it appears on \newline the student record sheet:& Mathematics \\
+				Betreut von / Supervisor: & Univ.-Prof. Matthias Aschenbrenner \\ % Insert your program/supervisor here
+				Mitbetreut von / Co-Supervisor: & Ass.-Prof. Florian Zuleger \\ 
+			\end{tabular}
+			\endgroup
+		\end{table}
+		\addtolength{\tabcolsep}{6pt} 
+	\end{flushleft}	
+	\endgroup
+
+	\newpage
+	\
+	\newpage
+
+	% Some of the code for creating ToC id adapted from the following template:
+	% https://www.sharelatex.com/templates/thesis/university-of-sussex-thesis
+
+		\
+		\vfill
+		\onehalfspacing
+		\begin{center}
+		\bfseries Abstract		
+		\end{center}
+		The famous double negation translation~\cite{glivenko1929quelques, godel1933intuitionistischen} establishes an embedding from classical into intuitionistic logic.
+Curiously the reverse direction has not been covered in literature.
+We present an effective embedding from intuitionistic into classical logic, both in the propositional and first-order case, as well as an effective embedding of intuitionistic propositional logic into quantified boolean formulas.
+
+		Furthermore we implement a practical system that takes intuitionistic first-order problems as an input and performs our transformation.
+This allows the use of classical theorem provers for establishing intuitionistic validity.
+We benchmark our implementation using the vampire theorem prover~\cite{kovacs2013first} on the ILTP problem set~\cite{iltp}.
+		%The key notion used in the proofs is the existence of counter-models - an analogon to the satisfiability of the negation in classical logic, which is not equivalent to invalidity in the intuitionistic case.
+We hope that this will allow leveraging the tremendous recent progress in automated reasoning in classical logic for checking intuitionistic validity, for which there has not been much progress in recent decades.
+
+		Finally we discuss how the generated classical proofs of the transformed problems can be translated to intuitionistic proofs of the original problem.
+All in all this establishes a novel approach to theorem proving for intuitionistic logic.
+
+	\vspace*{1cm}
+	\begin{center}
+		\bfseries Acknowledgements		
+	\end{center}
+
+	I want to thank my Matthias for accommodating my particular interests by agreeing to supervise a thesis which lies outside of his main area of research and bearing with the organizational difficulties.
+I want to thank Florian who graciously agreed to co-supervise a thesis at a different university.
+He has provided me with a fascinating topic as well as plenty of his time, expertise and thorough feedback.
+Finally I want to thank my family, in particular my parents Stefania and Cezary, who have instilled in me the joy of learning, provided all the love and help I could ask for and invariably supported me along my journey.
+
+	{\vfill\null}
+
+	\newpage
+	\
+	\singlespacing
+	\newgeometry{top=1cm}
 	\tableofcontents
+	\restoregeometry
+
+	\chapter{Introduction}
+	\onehalfspacing
+
+	\pagenumbering{arabic}
+	Constructive mathematics refers to a flavour of mathematics in which the existence of an object can only be established by explicit construction, as opposed to classical mathematics where existence can be shown implicitly, e.g.
+by assuming non-existence and deriving a contradiction.
+	The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the fact that the law of excluded middle $A\vee\neg A$ and the double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
+	Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in the Curry-Howard correspondence~\cite{howard1980formulae}.
+
+	The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmark problems (see e.g.
+the
+	ILTP library website~\cite{iltp}).
+	However, progress in automated theorem proving for intuitionistic logic has been slow, whereas reasoners for classical logic have made tremendous progress, see e.g.
+the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
+	%, e.g., the first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}
+	%Comparing benchmarks for classical and intuitionistic automated theorem proving, e.g.
+TPTP~\cite{casc} and , we see that classical provers currently have a far better success rate.
+Furthermore all the most popular first-order utilize classical logic.
+	This difference can be partially explained by fundamental differences between the logics.
+	First of all, determining intuitionistic validity is computationally harder, i.e.
+in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
+	A further advantage of classical logic is the existence of calculi that are particularly suitable for automation such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability, i.e.
+to show the validity of a formula it suffices to show the unsatisfiability of the negated formula.
+	While some (albeit more complex) normal forms also exist  for intuitionistic logic, crucially the duality between validity and satisfiability of the negation does not hold.
+	Therefore most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient, tammet1996resolution} use the naive inverse method, i.e.
+directly search for a cut-free proof by applying the rules from some proof calculus inversely, enhanced by search strategies such as focussing and polarization.
+This approach generally leads to a much more complex search and is therefore difficult to implement efficiently.
+	Finally, we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, in particular for the propositional case, i.e.
+SAT-solvers.
+
+	With this work we want to propose a novel approach for intuitionistic theorem proving that leverages the progress in classical theorem proving, that is for each formula $\varphi$:
+	\begin{itemize}
+		\item Give a formula $\varphi^\#$ that is classically valid if and only if $\varphi$ is intuitionistically valid.
+		\item Use a state-of-the-art classical prover to establish the validity/invalidity of $\varphi^\#$.
+		\item Transform the output proof/counter-model of $\varphi^\#$ to one of $\varphi$.
+	\end{itemize}
+	The most challenging part of this approach is giving the translation of $\varphi$ to $\varphi^\#$.
+	Interestingly, the reverse direction, the famous double-negation translation, has long been established and goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
+In the propositional case it is particularly simple: $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
+Intuitively the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent.
+This gives us a first idea why the reverse direction is perhaps more difficult: We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e.
+if they both occur in $\varphi$, we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$.
+In particular, this necessitates the introduction of new propositional variables in our procedure, which already marks a big difference to the double negation translation.
+	%
+	While establishing the translation of formulas we also perform an effective translation of counter-models, i.e.
+for each intuitionistic counter-model of $\varphi$ we effectively construct a classical counter-model of $\varphi^\#$ and vice versa.
+	We note that the existence of counter-models is a key notion that forms a proper dual to validity --- whereas the satisfiability of the negation is not necessary for invalidity (in contrast to classical logic, where it is a proper dual to validity).
+	Transforming and reducing counter-models to a normal form is also what ultimately enables our translation.
+Apart from a translation of counter-models we also explicitly describe how to equivalently transform classical proofs of $\varphi^\#$ to intuitionistic proofs of $\varphi$, in particular resolution proofs as given by theorem provers like vampire~\cite{Kov_cs_2013}.
+
+	As a final contribution we implement our translation in the Rust programming language.
+It takes a first-order problem in the tptp format~\cite{tptp} and outputs the translated problem in the same format.
+The code is published on github~\cite{implementation}.
+We then benchmark our implementation using the ILTP problem set~\cite{iltp}, i.e.
+we translate all problems in the set and then run the vampire theorem prover~\cite{Kov_cs_2013} on the translated problems.
+Our approach performs on par with most existing approaches for intuitionistic theorem provers, but comes short of the state-of-the art leaders.
+As there is still a lot of room for optimization this is a hopeful sign.
+
+	\chapter{Overview}
+
+	The main content is organized in chapters 3-6.
+Chapter 3 collects prerequisite definitions and serves to fix our notation and recall some important facts.
+Since this overview already contains some technical results it may be consulted for reference.
+Chapter 4 is the meat of this thesis.
+In particular it features a complete description of our embedding of intuitionistic into classical logic as well as proving its correctness.
+In chapter 5 we present our implementation and benchmark it.
+Finally in chapter 6 we explicitly describe how to translate the generated proofs and counter-models which in principle directly follows from the results of section 4.
+	We shall now for each chapter give an overview of our main results and highlight the key arguments.
+
+	\section{Embedding Intuitionistic into Classical Logic}
+
+	Recall that our goal in this chapter is to give an effective translation procedure that, for a given formula $\varphi$, yields a formula $\varphi^\#$, such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
+We start with the propositional case.
+The key arguments are the same as in the first-order case, while it is technically simpler and less cluttered.
+
+	Before our main transformation we employ a preprocessing step akin to the Tseytin transformation~\cite{tseitin1983complexity}, which is a popular pre-processing step in classical automated reasoning:
+	It gives an equisatisfiable sentence (over an extended set of propositions) in conjunctive normal form.
+	Eliminating all implications, however, is not possible in intuitionistic logic since $A\to B$ is not equivalent to $\neg A\vee B$.
+	Still we propose a similar transformation.
+	A notable feature of our transformation is that all non-classical content is encapsulated in formulas of type $(A\to B)\to C$, i.e.
+if there are no such formulas in the transformed formula then the classical validity of the transformed formula immediately implies intuitionistic validity.
+
+	\begin{theorem}\label{thm:Tseytin}
+		For every propositional formula $\varphi$ there effectively are an atom $P$ and a set of $\mathcal S$ of formulas (over an extended set of propositions) of one of the forms\\
+		$\text{ \hspace{1cm}} A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$\\
+		for atomic $A, B, C$, such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
+The time complexity of the transformation is linear in the input size.
+	\end{theorem}
+
+	The main transformation then proceeds in three steps:
+	%
+	1) We encode as a first-order sentence that the considered formula holds for every Kripke frame.
+	Since Kripke frames for propositional logic over a fixed alphabet of propositional variables form a first-order theory this step is rather straightforward.
+	%
+	2) Next we eliminate some quantifiers via Herbrandization. The only quantifiers eliminated occur in formulas derived from formulas of type $(A\to B)\to C$.
+	%
+	3) We can then argue that if there is a counter-model for the resulting formula then there is a counter-model with a certain Kripke frame of bounded size completely determined by the number of formulas of type $(A\to B)\to C$.
+	This allows us to eliminate the remaining quantifiers by simply enumerating the worlds in that Kripke frame.
+
+	Summarizing, we obtain the following result:
+
+	\begin{theorem}
+		\label{thm:reduction-propositional}
+		Let $\mathcal S$ be as in Theorem~\ref{thm:Tseytin} and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
+For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
+Obtain $\mathcal S^\#$ by including the following formulas:
+		\begin{itemize}
+			\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
+			\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+			\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
+			\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
+		\end{itemize}
+		Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
+The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$.
+Futhermore, there is an effective procedure for translating counter-models between the sentences.
+	\end{theorem}
+
+	Instead of explicitly enumerating the worlds of a counter-example of bounded size (as in the reduction stated in the above theorem), we can also do this enumeration implicitly with a quantified boolean formula (QBF).
+	We can show that this QBF is satisfiable if and only if there exists a counter-model for our original formula:
+
+	\begin{theorem}
+		Let $\mathcal S$ be as in Theorem~\ref{thm:Tseytin} and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$.
+There is an effective procedure that produces a QBF $\varphi^Q$ of size $\mathcal O(|\mathcal S|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$ with $2\cdot |\mathcal F_\to|-1$ quantifier alternations such that $\varphi$ is intuitionistically valid if and only if $\varphi^Q$ is not satisfiable.
+For fixed $N\in\mathbb N$, deciding intuitionistic validity for formulas $\bigwedge \mathcal S\to P$ where $\mathcal S$ is as above and $|\mathcal F_\to| = N$ is therefore in $\Sigma_{2N-1}$.
+	\end{theorem}
+
+	We then move to the first-order case. Here the construction is a bit more involved but the underlying approach is the same.
+	Again we first perform a Tseytin-like transformation.
+	Here the non-classical content is encapsulated by formulas of type $(A\to B)\to C$ and $\forall xA\to B$.
+	The details of this translation are given in Section~\ref{section:tseytin}.
+	We then proceed similarly to the propositional case.
+	Note however that one difficulty arises from the fact that our (classical) domain will now contain on the one hand worlds in the Kripke frame but on the other hand also proper domain elements.
+	We resolve this apparent conflict by introducing a special binary predicate $E$, inspired by~\cite{iemhoff2010eskolemization}, encoding which domain elements exists at which world.
+	The main transformation then again proceeds in three steps:
+	Step 1) and 2) are analogous to the propositional case.
+	The real difference arises in step 3):
+	While we can also establish the existence of a canonic counter-model whose frame only depends on formulas of type $(A\to B)\to C$ and $\forall xA\to B$, the size of this counter-model is countably infinite in general.
+	This is not surprising since certain intuitionistically invalid formulas like the double negation shift $\forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$ don't have finite counter-models.
+	Therefore we are not able to eliminate the $\forall$-quantifiers associated with the Kripke semantics.
+	However since our translation targets first-order logic, we believe that the introduction of these quantifiers is acceptable.
+	%
+	Summarizing, we obtain the following result:
+
+	\begin{theorem}
+		\label{thm:reduction-first-order-short}
+		There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
+The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary relation symbol in $\varphi$ there is a corresponding $n+1$-ary relation symbol in $\varphi^\#$, and $\varphi$ contains a new binary predicate $E$ as well as a number of new function symbols.
+Furthermore, there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
+	\end{theorem}
+	A more detailed account of Theorem~\ref{thm:reduction-first-order-short} is given in Theorem~\ref{fullFOtranslation}.
+
+	\section{Implementation}
+
+	In this section we describe our system that implements the translation from the previous section.
+The rust source code can be found online~\cite{implementation}.
+We then elaborate our benchmarking process which utilizes the ILTP problem set~\cite{iltp} and the vampire theorem prover~\cite{Kov_cs_2013}.
+
+	\begin{figure}[h]\label{results}
+		$$\begin{matrix}
+			&\text{AGT}&\text{ALG}&\text{COM}&\text{CSR}&\text{GEO}&\text{GRA}&\text{GRP}&\text{HAL}\\
+			\text{Proven}&5&170&0&0&58&3&4&2\\
+			\text{Disproven}&0&0&0&0&0&0&0&\\
+			\text{Timeout}&47&29&3&29&107&74&1&16
+		\end{matrix}$$
+		$$\begin{matrix}
+			&\text{KRS}&\text{LCL}&\text{MGT}&\text{MSC}&\text{NLP}&\text{NUM}&\text{PLA}&\text{PUZ}\\
+			\text{Proven}&0&0&30&1&9&1&9&2\\
+			\text{Disproven}&0&0&0&0&12&0&0&0\\
+			\text{Timeout}&5&9&128&3&69&1&237&82
+		\end{matrix}$$
+		$$\begin{matrix}
+			&\text{SET}&\text{SWC}&\text{SWV}&\text{SYN}&\text{TOP}&\text{GEJ}&\text{GPJ}&\text{SYJ}\\
+			\text{Proven}&0&5&27&1&64&65&54&1\\
+			\text{Disproven}&6&1&297&423&154&108&298&2
+		\end{matrix}$$
+		\caption{Performance on problem sections of the ILTP set.}
+	\end{figure}
+
+	Our implementation performs similar to existing dedicated provers for intuitionistic logic, but falls short of the state-of-the-art.
+Since there is still a lot of possible optimizations left for our translation this is a hopeful first sign.
+
+	\section{Translation of proofs and counter-models}
+
+	In the final section we give an explicit mechanism how to transform the proofs and counter-models for translated formula $\varphi^\#$ into intuitionistic proofs/counter-models of the original formula $\varphi$.
+The existence of these transformations is clear from the methodology of chapter~\ref{chapter:eiicl}, however they are hidden in the proofs and here we make them explicit.
+The two main results are as follows:
+	\begin{theorem}
+		There exists an effective linear time procedure that converts any classical counter-model for a translated formula $\varphi^\#$ into a intuitionistic counter-model for the original formula $\varphi$.
+	\end{theorem}
+	\begin{theorem}
+		There exists an effective linear time procedure that converts any classical proof of a translated formula $\varphi^\#$ into a intuitionistic proof for the original formula $\varphi$.
+	\end{theorem}
+
+	Given the arguments of Chapter~\ref{chapter:eiicl} the translation of counter-models is more or less just a compilation of proofs.
+On the other hand the explicit translation of proofs does require some work, since to that point all arguments were semantic and this requires a syntactic viewpoint.
 
 	\chapter{Preliminaries}
-
 	First let us recapitulate the most important definitions and some facts about them. This will serve to fix our notation as well.
@@
 
-	\section{Proof theory}
-
-	Logical truth can also be approached via syntax rather than semantics. This gives rise to proof theory which will be especially important in the third chapter.
+	\section{Proof theory}\label{proof-theory}
+
+	Logical truth can also be approached via syntax rather than semantics. This gives rise to proof theory which will be especially important in the final chapter when we establish proof translations.
 
 	\begin{definition}
-		A \textit{sequent} $\Gamma\vdash \Delta$ consists of two multisets of formulas $\Gamma, \Delta$.
-The following rules make up our classical calculus $\mathbf{GC}$, corresponding roughly to \textbf{G3c} from~\cite{basicprooftheory}:\\
+		Pairs of multisets of formulas $A = \{A_1,\dots, A_n\}, B = \{B_1,\dots, B_m\}$ form a \textit{Sequent} $A\Rightarrow B$.
+Usually we will write $A_1,\dots, A_n\Rightarrow B_1,\dots, B_m$, note however that $A$ and $B$ are \textbf{not} to be interpreted as sequences but as multisets.
+	\end{definition}
+
+	We define sequent calculi as in~\cite[p.77]{basicprooftheory}. Have following inferences\\
+
 	\begin{center}
@@
 		\RightLabel{Ax ($P$ atomic)}
-				\UnaryInfC{$P,\Gamma\vdash \Delta, P$}
+		\UnaryInfC{$P,\Gamma\Rightarrow \Delta, P$}
 		\DisplayProof&
@@
 		\RightLabel{L$\bot$}
-				\UnaryInfC{$\bot,\Gamma\vdash\Delta$}
+		\UnaryInfC{$\bot,\Gamma\Rightarrow\Delta$}
 		\DisplayProof&
 		\\&&\\
-				\AxiomC{$A, B,\Gamma\vdash\Delta$}
+		\AxiomC{$A, B,\Gamma\Rightarrow\Delta$}
 		\RightLabel{L$\wedge$}
-				\UnaryInfC{$A\wedge B, \Gamma\vdash \Delta$}
+		\UnaryInfC{$A\wedge B, \Gamma\Rightarrow \Delta$}
 		\DisplayProof&
-				\AxiomC{$\Gamma\vdash\Delta, A$}
-				\AxiomC{$\Gamma\vdash\Delta, B$}
+		\AxiomC{$\Gamma\Rightarrow\Delta, A$}
+		\AxiomC{$\Gamma\Rightarrow\Delta, B$}
 		\RightLabel{R$\wedge$}
-				\BinaryInfC{$\Gamma\vdash \Delta, A\wedge B$}
+		\BinaryInfC{$\Gamma\Rightarrow \Delta, A\wedge B$}
 		\DisplayProof&
 		\\&&\\
-				\AxiomC{$A, \Gamma\vdash\Delta$}
-				\AxiomC{$B, \Gamma\vdash\Delta$}
+		\AxiomC{$A, \Gamma\Rightarrow\Delta$}
+		\AxiomC{$B, \Gamma\Rightarrow\Delta$}
 		\RightLabel{L$\vee$}
-				\BinaryInfC{$A\vee B, \Gamma\vdash \Delta$}
+		\BinaryInfC{$A\vee B, \Gamma\Rightarrow \Delta$}
 		\DisplayProof&
-				\AxiomC{$\Gamma\vdash\Delta, A, B$}
+		\AxiomC{$\Gamma\Rightarrow\Delta, A, B$}
 		\RightLabel{R$\vee$}
-				\UnaryInfC{$\Gamma\vdash \Delta, A\vee B$}
+		\UnaryInfC{$\Gamma\Rightarrow \Delta, A\vee B$}
 		\DisplayProof&
 		\\&&\\
-				\AxiomC{$\Gamma\vdash\Delta, A$}
-				\AxiomC{$B, \Gamma\vdash\Delta$}
+		\AxiomC{$\Gamma\Rightarrow\Delta, A$}
+		\AxiomC{$B, \Gamma\Rightarrow\Delta$}
 		\RightLabel{L$\to$}
-				\BinaryInfC{$A\to B, \Gamma\vdash \Delta$}
+		\BinaryInfC{$A\to B, \Gamma\Rightarrow \Delta$}
 		\DisplayProof&
-				\AxiomC{$A,\Gamma\vdash\Delta, B$}
+		\AxiomC{$A,\Gamma\Rightarrow\Delta, B$}
 		\RightLabel{R$\to$}
-				\UnaryInfC{$\Gamma\vdash \Delta, A\to B$}
+		\UnaryInfC{$\Gamma\Rightarrow \Delta, A\to B$}
 		\DisplayProof&
 		\\&&\\
-				\AxiomC{$A[t/x], \Gamma\vdash\Delta$}
+		\AxiomC{$A[t/x], \Gamma\Rightarrow\Delta$}
 		\RightLabel{L$\forall$}
-				\UnaryInfC{$\forall xA, \Gamma\vdash \Delta$}
+		\UnaryInfC{$\forall xA, \Gamma\Rightarrow \Delta$}
 		\DisplayProof&
-				\AxiomC{$\Gamma\vdash\Delta, A[a/x]$}
+		\AxiomC{$\Gamma\Rightarrow\Delta, A[a/x]$}
 		\RightLabel{R$\forall$}
-				\UnaryInfC{$\Gamma\vdash \Delta, \forall xA$}
+		\UnaryInfC{$\Gamma\Rightarrow \Delta, \forall xA$}
 		\DisplayProof&
 		\\&&\\
-				\AxiomC{$A[a/x], \Gamma\vdash\Delta$}
+		\AxiomC{$A[a/x], \Gamma\Rightarrow\Delta$}
 		\RightLabel{L$\exists$}
-				\UnaryInfC{$\exists xA, \Gamma\vdash \Delta$}
+		\UnaryInfC{$\exists xA, \Gamma\Rightarrow \Delta$}
 		\DisplayProof&
-				\AxiomC{$\Gamma\vdash\Delta, A[t/x]$}
+		\AxiomC{$\Gamma\Rightarrow\Delta, A[t/x]$}
 		\RightLabel{R$\exists$}
-				\UnaryInfC{$\Gamma\vdash \Delta, \exists xA$}
-				\DisplayProof&
-				\\&&\\
-				\AxiomC{$t = t, \Gamma\vdash \Delta$}
-				\RightLabel{Ref}
-				\UnaryInfC{$\Gamma\vdash \Delta$}
-				\DisplayProof&
-				\AxiomC{$s = t, P[s/x], P[t/x], \Gamma\vdash \Delta$}
-				\RightLabel{Ref}
-				\UnaryInfC{$s = t, P[s/x],\Gamma\vdash \Delta$}
+		\UnaryInfC{$\Gamma\Rightarrow \Delta, \exists xA$}
 		\DisplayProof&
@@
 	\end{center}
-
 	where in R$\forall$ and L$\exists$ $a$ is a free variable not occurring in the sequent otherwise.
-		The above rules make up the calculus called \textbf{Gc}. It corresponds to \textbf{G3c}$^=$ from~\cite{basicprooftheory}.
-
-		For intuitionistic logic we remain with a multi-succedent system by adjusting R$\to$ and R$\forall$:\\
+
+	The above rules make up the calculus for classical logic called \textbf{Gc}. It corresponds to \textbf{G3c} from~\cite{basicprooftheory}.
+	Similarly we can define a calculus \textbf{Gi} for intuitionistic logic corresponding to \textbf{m-G3i}:\\
 
 		\begin{center}
-			\AxiomC{$A, \Gamma\vdash B$}
+		\begin{tabular}{lll}
+			\AxiomC{\hphantom{x}}
+			\RightLabel{Ax ($P$ atomic)}
+			\UnaryInfC{$P,\Gamma\Rightarrow \Delta, P$}
+			\DisplayProof&
+			\AxiomC{\hphantom{x}}
+			\RightLabel{L$\bot$}
+			\UnaryInfC{$\bot,\Gamma\Rightarrow\Delta$}
+			\DisplayProof&
+			\\&&\\
+			\AxiomC{$A, B,\Gamma\Rightarrow\Delta$}
+			\RightLabel{L$\wedge$}
+			\UnaryInfC{$A\wedge B, \Gamma\Rightarrow \Delta$}
+			\DisplayProof&
+			\AxiomC{$\Gamma\Rightarrow\Delta, A$}
+			\AxiomC{$\Gamma\Rightarrow\Delta, B$}
+			\RightLabel{R$\wedge$}
+			\BinaryInfC{$\Gamma\Rightarrow \Delta, A\wedge B$}
+			\DisplayProof&
+			\\&&\\
+			\AxiomC{$A, \Gamma\Rightarrow\Delta$}
+			\AxiomC{$B, \Gamma\Rightarrow\Delta$}
+			\RightLabel{L$\vee$}
+			\BinaryInfC{$A\vee B, \Gamma\Rightarrow \Delta$}
+			\DisplayProof&
+			\AxiomC{$\Gamma\Rightarrow\Delta, A, B$}
+			\RightLabel{R$\vee$}
+			\UnaryInfC{$\Gamma\Rightarrow \Delta, A\vee B$}
+			\DisplayProof&
+			\\&&\\
+			\AxiomC{$A\to B, \Gamma\Rightarrow\Delta, A$}
+			\AxiomC{$B, \Gamma\Rightarrow\Delta$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$A\to B, \Gamma\Rightarrow \Delta$}
+			\DisplayProof&
+			\AxiomC{$A,\Gamma\Rightarrow B$}
 			\RightLabel{R$\to$}
-			\UnaryInfC{$\Gamma\vdash\Delta, A\to B$}
-			\DisplayProof\hspace{2cm}
-			\AxiomC{$\Gamma\vdash A[a/x]$}
+			\UnaryInfC{$\Gamma\Rightarrow \Delta, A\to B$}
+			\DisplayProof&
+			\\&&\\
+			\AxiomC{$A[t/x], \Gamma\Rightarrow\Delta$}
+			\RightLabel{L$\forall$}
+			\UnaryInfC{$\forall xA, \Gamma\Rightarrow \Delta$}
+			\DisplayProof&
+			\AxiomC{$\Gamma\Rightarrow A[a/x]$}
 			\RightLabel{R$\forall$}
-			\UnaryInfC{$\Gamma\vdash\Delta, \forall xA$}
-			\DisplayProof
+			\UnaryInfC{$\Gamma\Rightarrow \forall xA$}
+			\DisplayProof&
+			\\&&\\
+			\AxiomC{$A[a/x], \Gamma\Rightarrow\Delta$}
+			\RightLabel{L$\exists$}
+			\UnaryInfC{$\exists xA, \Gamma\Rightarrow \Delta$}
+			\DisplayProof&
+			\AxiomC{$\Gamma\Rightarrow\Delta, A[t/x]$}
+			\RightLabel{R$\exists$}
+			\UnaryInfC{$\Gamma\Rightarrow \Delta, \exists xA$}
+			\DisplayProof&
+			\\&&\\
+		\end{tabular}
 	\end{center}
 
-		We will call the calculus containing this formulas $\mathbf{GI}$ roughly corresponding to $\mathbf{G3i}$ from~\cite{basicprooftheory}.
-Usually for intuitionistic logic a single-formula succedent system is considered, but we are interested in reducing the difference between the calculi as much as possible.
-
-		A \textit{proof} of a formula $\varphi$ in a calculus $\mathbf{G}$ is a labelled binary rooted tree in which every node is labelled with a sequent such that
-		\begin{itemize}
-			\item Every leaf is labelled in accordance to Ax or L$\bot$.
-			\item The sequent of an interior node can be derived from sequents of its child nodes according to one of the rules.
-		\end{itemize}
-		We usually present a proof directly as a sequence of applications of rules rather than in a more usual tree presentation.
-If a sequent $\mathcal S\vdash\mathcal T$ is provable in $\mathbf{GC}$ we write $\mathcal S\vdash_C\mathcal T$, if it is in $\mathbf{GI}$ we write $\mathcal S\vdash_I\mathcal T$.
+	The differences are found in L$\to$, R$\to$ and R$\forall$. Note how in particular the rules for R$\to$ and R$\forall$ allows only a single formula in the succedent.
+
+	\pagebreak
+
+	\begin{definition}
+		A \textit{Derivation} is a rooted finite labelled tree, in which each leaf is labelled with one of the two axiom rules and each other label is derived from the labels of the successor nodes in accordance to the rules above.
+All formulas not in $\Gamma$ or $\Delta$ are the \textit{active} formulas of the rule.
+The label of the root is called \textit{head}.
+A formula that is the head of a derivation is said to be \textit{derivable}.
 	\end{definition}
 
+	The following is one of the fundamental results of first-order Logic:
+
+	\begin{theorem}[Completeness theorem]
+		A sequent $\Phi\Rightarrow \varphi$ is derivable iff $\Phi\models \varphi$.
+	\end{theorem}
+
 	\begin{example}
-		The following is a proof of $(A\to B)\to A\vdash B$:
+		The following is a \textbf{Gc} proof of the law of excluded middle.\\
+
 		\begin{center}
@@
 			\RightLabel{Ax}
-			\UnaryInfC{$A\to B, A\vdash B, A$}
+			\UnaryInfC{$A\Rightarrow\bot, A$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$\Rightarrow A, A\to\bot$}
+			\RightLabel{R$\vee$}
+			\UnaryInfC{$\Rightarrow A\vee (A\to \bot)$}
+			\DisplayProof
+		\end{center}
+
+		Note how this derivation is invalid in \textbf{Gi} since R$\to$ is not applied correctly. However it is possible to prove the double negation of it in \textbf{Gi}:\\
+
+		\begin{center}
 			\AxiomC{}
 			\RightLabel{Ax}
-			\UnaryInfC{$A, B\vdash B$}
+			\UnaryInfC{$(A\vee (A\to \bot))\to\bot, A\Rightarrow A, (A\to \bot),\bot$}
+			\RightLabel{R$\vee$}
+			\UnaryInfC{$(A\vee (A\to \bot))\to\bot, A\Rightarrow A\vee (A\to \bot),\bot$}
+			\AxiomC{}
+			\RightLabel{L$\bot$}
+			\UnaryInfC{$\bot\Rightarrow$}
 			\RightLabel{L$\to$}
-			\BinaryInfC{$A\to B, A\vdash B$}
+			\BinaryInfC{$(A\vee (A\to \bot))\to\bot, A\Rightarrow\bot$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$(A\vee (A\to \bot))\to\bot\Rightarrow A, A\to \bot$}
+			\RightLabel{R$\vee$}
+			\UnaryInfC{$(A\vee (A\to \bot))\to\bot\Rightarrow A\vee (A\to \bot)$}
+			\AxiomC{}
+			\RightLabel{L$\bot$}
+			\UnaryInfC{$\bot\Rightarrow$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$(A\vee (A\to \bot))\to\bot\Rightarrow$}
 			\DisplayProof
 		\end{center}
+
 	\end{example}
 
-	\noindent We now recall some important properties of the given systems:
-
-	\begin{theorem}[subformula property]
-		Any formula that appears in a derivation is a subformula of a forumla that occurs in the root sequent.
-	\end{theorem}
-
-	We consider the following additional rules, which are not a priori part of our calculus:
-	\begin{center}
-		\begin{tabular}{lll}
-			\AxiomC{$\Gamma\vdash\Delta$}
-			\RightLabel{Lweak}
-			\UnaryInfC{$A,\Gamma\vdash \Delta$}
-			\DisplayProof&
-			\AxiomC{$\Gamma\vdash\Delta$}
-			\RightLabel{Rweak}
-			\UnaryInfC{$\Gamma\vdash \Delta, A$}
-			\DisplayProof&
-			\\&&\\
-			\AxiomC{$A, A,\Gamma\vdash\Delta$}
-			\RightLabel{Lcontr}
-			\UnaryInfC{$A, \Gamma\vdash \Delta$}
-			\DisplayProof&
-			\AxiomC{$\Gamma\vdash\Delta, A, A$}
-			\RightLabel{Rcontr}
-			\UnaryInfC{$\Gamma\vdash \Delta, A$}
-			\DisplayProof&
-			\\&&\\
-		\end{tabular}
-
-		\AxiomC{$\Gamma\vdash A, \Delta$}
-		\AxiomC{$\Gamma', A\vdash \Delta'$}
-		\RightLabel{Cut}
-		\BinaryInfC{$\Gamma, \Gamma'\vdash\Delta, \Delta'$}
-		\DisplayProof
-	\end{center}
-
-	The following central result of proof theory will be useful to us in many places:
-	\begin{theorem}[admissibility of cut (and structural rules)]
-		The 5 above rules are admissible in our calculi for classical and intuitionistic logic, i.e. adding them we are not able to prove any additional theorems.
-	\end{theorem}
-
-	This means that we may assume that all proofs are free of these rules, but in constructing new proofs we may apply them.
-We exclude them from our calculus to keep it simple for proofs and because cut destroys the subformula property.
-
-	The semantic and syntactic approach to logic are of course equivalent.
-
-	\begin{theorem}\label{completeness}[Completeness and Soundness]
-		For sets of formulas $\mathcal S,\mathcal T$ both classically and intuitionistically
-		$$\mathcal S\models\mathcal T \text{ iff }\mathcal S\vdash\mathcal T$$
-	\end{theorem}
-
-	\chapter{Embedding intuitionistic into classical logic}
+	\chapter{Embedding intuitionistic into classical logic}\label{chapter:eiicl}
 
@@
 
-	\subsection{Normal form transformation}
+	\subsection{Normal form transformation}\label{section:tseytin}
 
@@
 
-	\begin{definition}
+	\begin{definition}\label{def:tseytin}
 		Let $\varphi$ be some formula.
@@
 
-	\begin{corollary}\label{eequivalid}
+	\begin{corollary}\label{equivalid}
 		In both intuitionistic and classical logic
@@
 	\begin{example}
-		With the previous encoding for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
+		With the previous encoding for the double negation elimination $$\varphi = ((A\to \bot)\to \bot)\to A$$ we have
 		$$\hspace*{-.3cm}
@@
 
-	\section{CQC to IQC}
+	\section{CQC to IQC}\label{section:CQCIQC}
 
@@
 		\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c)))\indent \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c))\indent \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c)))\\
-		\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c))\indent \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\indent\forall \vec z(\forall xA(\vec a)\to B(\vec b))\\ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b)).
+		\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c))\indent \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\indent\forall \vec z(\forall xA(\vec a)\to B(\vec b))\\ \forall \vec z(A(\vec a)\to\forall xB(\vec b))\indent \forall \vec z(\exists xA(\vec a)\to B(\vec b))\indent\forall \vec z(A(\vec a)\to\exists xB(b)).
 	\end{matrix}\]
@@
 	\end{theorem}
+
+	\begin{example}
+		Consider the classical tautology $\varphi := \neg\forall x\neg A(x)\to \exists x A(x)$. Setting $\psi := \forall x\neg A(x)$ we have
+		\begin{align*}
+			S^\#(\varphi) = &\{\forall u((P_{\neg\psi}^\#(f_\varphi(u))\to P_{\exists xA(x)}^\#(f_\varphi(u)))\to P_\varphi^\#(u))\}\cup\\
+			 & \{\forall u(\exists x(E(x, u)\wedge A^\#(x, u))\to P_{\exists xA(x)}^\#(u)), \forall u((P_{\neg\psi}^\#(u)\wedge P^\#_\psi(u))\to \bot)\}\cup\\
+			 & \{\forall u(\forall x(E^\#(x,f_\psi(u))\to P^\#_{\neg A(x)}(x, f_\psi(u)))\to P^\#_\psi(u))\}\cup\\&\{ \forall x\forall u(E(x, u)\to (A^\#(x, f_{\neg A(x)(u)})\to \bot)\to P^\#_{\neg A(x)}(x, u))\}
+		\end{align*}
+		And as one might expect we can indeed define a counter-model for $$\exists xE(x, s)\to K\to\bigwedge S^\#\to P^\#(s),$$
+		namely take as a domain $M = \{m, v, w\}$ where $u, w$ represent worlds, and $m$ representing a single domain element.
+Have $f_\varphi^I(x) = f_\psi^I(x) = f_{\neg A(x)}^I(x) = w$ for all $x\in M$, $E^I(x, y)$ if and only if $x = m$ and $y\in\{u, v\}$ and finally $A^\#(x, y)$ iff $x = m$ and $y = w$.
+	\end{example}
 
@@
 		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
-		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
-		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
+		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
+		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec b, u)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
 	\end{itemize}
@@
 
-	However part 2 of Lemma~\ref{thm:fo-countermodel-reduction} fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$), we have $A(x, v)$, there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
-This is why we must consider sequences that have repetitions:
-
 	\begin{wrapfigure}[16]{r}{0.4\textwidth}
-		\vspace{-32pt}
 		\begin{center}
@@
 	\end{wrapfigure}
+	However part 2 of Lemma~\ref{thm:fo-countermodel-reduction} fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$), we have $A(x, v)$, there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
+This is why we must consider sequences that have repetitions:
+
 	\begin{example}
@@
 
+	In this chapter we shall give an overview over our implementation of the described translation in Rust. Rust was a good fit for this endeavour because
+	\begin{itemize}
+		\item There is a great package for parsing files in tptp format~\cite{rust-tptp}.
+		\item Static typing reduced bugs induced by the many syntactically similar expressions that require different translation.
+		\item Rust is blazingly fast.
+	\end{itemize}
+	The implementation spans over 3000 lines of code across the files 
+	\begin{center}\lstinline|normalized_formula.rs, encoding.rs, translation.rs, main.rs|.\end{center}
+	\lstinline|normalized_formula.rs| contains internal data structures representing the atomic formulas as well as normalized formulas as in $\mathcal S$ in Theorem~\ref{fullFOtranslation}.
+\lstinline|encoding.rs| defines a visitor that traverses the structure of the input file and gives for each encountered subformula the corresponding encoding as in Definition~\ref{def:tseytin}.
+\lstinline|translation.rs| handles additional included files.
+Finally \lstinline|main.rs| puts it all together to define a complete translation.
+
+	\section{Installation}
+
+	Currently the only available installation option is building the binary from source.
+The cargo rust complier~\cite{rust} is required for this purpose.
+The source can be downloaded from github~\url{https://github.com/lexpk/Proofs-as-Programs-in-Classical-Logic/tree/main/Implementation/eiicl}~\cite{implementation}.
+To build the binary go to the eiicl folder and run
+	\begin{lstlisting}
+	cargo build --bin eiicl
+	\end{lstlisting}
+	The binary can be then found under 
+	\begin{lstlisting}
+	eiicl/target/debug/eiicl
+	\end{lstlisting}
+
+	\section{Usage}
+
+	The binary is to be used from the command line. It takes one or two arguments:
+	\begin{itemize}
+		\item The location of the tptp file that is to be translated.
+		\item (Optionally) a location with additional files to be included, per default the current folder.
+	\end{itemize}
+	\begin{example}
+		Suppose the \lstinline|eiicl| binary is in the \$PATH and the current directory contains a file \lstinline|example.p| containing
+		\begin{lstlisting}
+	include('ax1.p').
+		\end{lstlisting}
+		as well as a folder \lstinline|ax| containing a single file \lstinline|ax1.p|.
+	Then running the command
+	\begin{lstlisting}
+ 	eiicl example.p axioms/
+	\end{lstlisting}
+	will output the correct translation.
+	\end{example}
+	Currently the translation does not support free variables, i.e. all free variables are expected to be explicitly quantified.
+
+	\section{Benchmarking}
+
+	\begin{figure}[h]\label{results}
+		$$\begin{matrix}
+			&\text{AGT}&\text{ALG}&\text{COM}&\text{CSR}&\text{GEO}&\text{GRA}&\text{GRP}&\text{HAL}\\
+			\text{Proven}&5&170&0&0&58&3&4&2\\
+			\text{Disproven}&0&0&0&0&0&0&0&\\
+			\text{Timeout}&47&29&3&29&107&74&1&16
+		\end{matrix}$$
+		$$\begin{matrix}
+			&\text{KRS}&\text{LCL}&\text{MGT}&\text{MSC}&\text{NLP}&\text{NUM}&\text{PLA}&\text{PUZ}\\
+			\text{Proven}&0&0&30&1&9&1&9&2\\
+			\text{Disproven}&0&0&0&0&12&0&0&0\\
+			\text{Timeout}&5&9&128&3&69&1&237&82
+		\end{matrix}$$
+		$$\begin{matrix}
+			&\text{SET}&\text{SWC}&\text{SWV}&\text{SYN}&\text{TOP}&\text{GEJ}&\text{GPJ}&\text{SYJ}\\
+			\text{Proven}&0&5&27&1&64&65&54&1\\
+			\text{Disproven}&6&1&297&423&154&108&298&2
+		\end{matrix}$$
+		\caption{Performance on problem sections of the ILTP set.}
+	\end{figure}
+
+	The program was benchmarked on the ILTP problem set~\cite{iltp}, which contains 2480 problems from the TPTP problem set~\cite{tptp}, which is commonly used to benchmark classical provers.
+The problem \lstinline|SYN007+1.014| was removed from the set as our translation mechanism would generate $>2^{14}$ formulas.
+For the purpose of the benchmark we report timeout for this problem.
+Since \lstinline|SYN007+1.014| is rather artificial this is not a big issue for us.
+For the classical proving part of our benchmark we use the Vampire theorem prover~\cite{Kov_cs_2013}, which is a state-of-the-art prover as evidenced by the CASC competition~\cite{casc}.
+Then for each problem our system transforms the described transformation and feeds the result to vampire with a time limit of 30 seconds which then reports validity, satisfiability of the negation or a timeout.
+The benchmark was run on an Intel Core i5-8400 CPU.
+
+	Out of the 2480 problems 511 were proven, 30 were refuted and 2128 resulted in a timeout, which corresponds to frequencies of $19.1\%$, $1.1\%$ and $79.7\%$ respectively.
+Out of the provers on the ILTP website this only puts it behind \lstinline|ileanCoP| which managed to resolve 690 or $27.1\%$ of the problems with a proof or refutation and ahead of  five other provers.
+A caveat is that the ILTP benchmarks were run on a slightly less powerful machine but with a timelimit of 600s, which suggests that our result might be even better.
+
+	The ILTP problem set is divided into 24 separate sections consisting of different kinds of problems, e.g.
+ALG which are primarily algebraic in nature.
+The performance on the individual sections can be seen it table~\ref{results}.
+	More detailed results for the individual problems can be found in Apendix~\ref{benchmark-results}.
+
 	\chapter{Translation of proofs and counter-models}
+
+	Finally we elaborate on how generated proofs and counter-models for $\varphi^\#$ can be explicitly translated in the first-order case.
+
+	\section{Translation of counter-models}
+
+	We will merely have to summarize results here.
+
+	\begin{definition}
+		Let $\mathcal M = (M, I)$ be a counter-model to $\varphi^\#$. Then define the Kripke structure $\mathcal K(\mathcal M) = (W, (M_w, I_w))_{u\in W})$ as follows:
+		\begin{itemize}
+			\item As the set of worlds $W$ take the term algebra over $s$ and $\{f_{\vec{z}} := u\mapsto f(\vec z, u)\:|\:f\in\mathcal F, \vec z\in M\}$.
+Let the partial order be induced by the subterm relation.
+			\item At each world $u$ define the universe $M_u := \{m\in M\:|\:E^I(m, u^I)\}$ where $f_{\vec z}(t)^I := f^I(\vec z, t^I)$.
+			\item For each function symbol $f$ have $f^{I_u} := f^I_{M_u}$.
+			\item For each relation symbol $R$ have $R^{I_u}(\vec t) := {R^\#}^I(\vec t, u^I)$
+		\end{itemize}
+	\end{definition}
+
+	\begin{theorem}
+		If $\mathcal M\not\models\varphi^\#$ then $ \mathcal K(\mathcal M)\not\models\varphi$
+	\end{theorem}
+
+	This follows directly from the proofs in section~\ref{section:CQCIQC}.
+
+	\section{Proof translation}
+
+	We want to transform Resolution proof for $\varphi^\#$ as given by vampire into a \textbf{Gi} proof for $\varphi$.
+While the results of this section are also clear in principle, there is some additional work to be done in the details.
+Since syntactic details are of importance here it might be useful to recall preliminary sections~\ref{proof-theory}.
+We will focus on how to obtain a \textbf{Gi} proof for $\varphi$ from a \textbf{Gc} proof of $\varphi^\#$, which is a much closer calculus.
+To obtain a \textbf{Gc} proof from a resolution proof one can use a standard method, e.g.
+as given in~\cite{de2005translation}.
+	Recall that $\varphi^\#$ is of the form $K\to \bigwedge\mathcal S^\#\to P^\#(s)$, which is equivalent to $K,\mathcal S^\#\Rightarrow P^\#(s)$ where each $\psi\in\mathcal S^\#$ is of one of the forms
+	\[\begin{matrix}
+		\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\wedge C^\#(\vec c, u))))\\\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\wedge B^\#(\vec b, u))\to C^\#(\vec c, u))\\ \forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\vee C^\#(\vec c, u))))\\
+		\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\vee B^\#(\vec b, u))\to C^\#(\vec c, u))\\ \forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))\\\forall \vec z\forall u(\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u)))\to B^\#(\vec b, u))\\ \forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to\forall x(E(x, u)\to B^\#(\vec b, u))))\\ \forall \vec z\forall u(\vec E(\vec z, u)\to (\exists x(E(x, u) \wedge A^\#(\vec a, u))\to B^\#(\vec b, u))\\\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to\exists x(E(x, u) \wedge B^\#(b, u))))).
+	\end{matrix}\]
+
+	Suppose now we have a \textbf{Gc} proof $\mathfrak P$ of $\varphi^\#$.
+Without loss of generality suppose that the rules introducing the left $\forall$ quantifiers are at the end of the proof, i.e.
+each atom occurring until then will have ground terms in place of $\vec t$ and $u$.
+This can be achieved by permutation of rules.
+
+	We can then split the proof into two parts $\mathfrak{A}$ and $\mathfrak{B}$, where in $\mathfrak{A}$ the left $\forall$ quantifiers have not yet been introduced.
+We can then associate with each formula $\psi$ occurring in $\mathfrak A$ a world $u(\psi)$ which is the ground term that is in the end substituted for $u$.
+We say that a rule $R$ \textit{features} a world $u$ if $u$ is a subformula of $u(\psi)$ for some active formulas $\psi$.
+
+	Due to the subformula property we can achieve by another permutation argument that each rule R$\to$ and R$\forall$ is followed by its associated L$\to$ rule, i.e.
+all rules  R$\to$ and R$\forall$ occurr as
+
+	\begin{center}
+		\AxiomC{$A^\#(\vec a, f_\psi(\vec z, w)),\Gamma\Rightarrow\Delta, B^\#(\vec b, f_\psi(\vec z, w))$}
+				\RightLabel{R$\to$}
+		\UnaryInfC{$\Gamma\Rightarrow\Delta, A^\#(\vec a, f_\psi(\vec z, w))\to B^\#(\vec b, f_\psi(\vec z, w))$}
+		\AxiomC{$C^\#(\vec c, w), \Gamma\Rightarrow\Delta$}
+		\RightLabel{L$\to$}
+		\BinaryInfC{$(A^\#(\vec a, f_\psi(\vec z, w))\to B^\#(\vec b, f_\psi(\vec z, w)))\to C^\#(\vec c, w), \Gamma\Rightarrow \Delta$}
+		\DisplayProof
+	\end{center}
+
+	\begin{center}
+		\AxiomC{$E(x, f_\psi(\vec z, w)),\Gamma\Rightarrow\Delta, A^\#(\vec b, f_\psi(\vec z, w))$}
+		\RightLabel{R$\to$}
+		\UnaryInfC{$\Gamma\Rightarrow\Delta, E(x, f_\psi(\vec z, w))\to B^\#(\vec a, f_\psi(\vec z, w))$}
+		\RightLabel{R$\forall$}
+		\UnaryInfC{$\Gamma\Rightarrow\Delta, \forall x(E(x, f_\psi(\vec z, w))\to A^\#(\vec a, f_\psi(\vec z, w)))$}
+		\AxiomC{$B^\#(\vec b, w), \Gamma\Rightarrow\Delta$}
+		\RightLabel{L$\to$}
+		\BinaryInfC{$ \forall x(E(x, f_\psi(\vec z, w))\to A^\#(\vec a, f_\psi(\vec z, w)))\to B^\#(\vec b, w), \Gamma\Rightarrow \Delta$}
+		\DisplayProof
+	\end{center}
+	for some ground term $w$.
+
+	We will now show that $\mathfrak A$ can be transformed into an intuitionistically valid proof by induction on the number of such rules.
+
+	Consider the last such rule.
+We must have $\Delta = P^\#(s)$.
+W.l.o.g.
+assume we are in the first case.
+Now consider the (possibly invalid) subproof above this rule R$\to$ consisting only of formulas that contain $f_\psi(\vec z, w)$ and rules that feature $f_\psi(\vec z, w)$.
+	\begin{itemize}
+		\item If this proof contains a valid proof of $$A^\#(\vec a, f_\psi(\vec z, w)),\Gamma\Rightarrow B^\#(\vec b, f_\psi(\vec z, w))$$ then, substituting $s$ for $f_\psi(\vec z, w)$, using the induction hypothesis we obtain an intuitionistic proof.
+But then applying the induction hypothesis also on the right branch of L$\to$ above we directly obtain an intuitionistic proof of $$(A^\#(\vec a, f_\psi(\vec z, w))\to B^\#(\vec b, f_\psi(\vec z, w)))\to C^\#(\vec c, w), \Gamma\Rightarrow \Delta.$$
+		\item  On the other hand suppose the subproof does not contain a valid proof of 
+		$$A^\#(\vec a, f_\psi(\vec z, w)),\Gamma\Rightarrow B^\#(\vec b, f_\psi(\vec z, w)).$$Then removing all formulas containing $f_\psi(\vec z, w)$ and all rules featuring it and then adding missing redundant formulas in the axioms yields a valid classical proof of $$(A^\#(\vec a, f_\psi(\vec z, w))\to B^\#(\vec b, f_\psi(\vec z, w)))\to C^\#(\vec c, w), \Gamma\Rightarrow \Delta.$$ But then, since we removed one R$\to$ formula, we also obtain an intuitionistic proof via the induction hypothesis.
+	\end{itemize}
+
+	This gives us the following result
+	\begin{theorem}
+		There is an effective linear time procedure that transforms a classical proof of $\varphi^\#$ into a intuitionistic proof of $\varphi$.
+	\end{theorem}
+	\begin{example}
+		Consider the intuitionistic theorem $$\varphi = \bigwedge\{(A\to A)\to B, (A\to B)\to B\}\to B.$$
+		which is translated to
+		\[
+			\varphi^\# = \exists xE(x, s)\to \bigwedge\left\lbrace
+			\begin{matrix}
+				\forall z\forall u(E(z, u)\to (E(z, f(u))\wedge E(z, g(u))))\\
+				\forall u(A^\#(u)\to  (A^\#(f(u))\wedge A^\#(g(u))))\\
+				\forall z\forall u(P^\#(z, u)\to (P^\#(z, f(u))\wedge P^\#(z, g(u))))\\
+				\forall u((A^\#(f(u))\to A^\#(f(u)))\to B^\#(u))\\
+				\forall u(A^\#(g(u))\to B^\#(x, g(u)))\to B^\#(u))
+			\end{matrix}\right\rbrace 
+			\to B^\#(s)
+		\]
+			For this example we shall consider the simplified but still valid alternative
+			\[
+			\varphi^\# = \bigwedge\left\lbrace
+			\begin{matrix}
+				\forall u((A^\#(f(u))\to A^\#(f(u)))\to B^\#(u))\\
+				\forall u(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))
+			\end{matrix}\right\rbrace 
+			\to B^\#(s)
+			\]
+
+			For this we obtain the following proof:
+			\begin{center}
+				\AxiomC{}
+				\RightLabel{Ax}
+				\UnaryInfC{$\substack{A^\#(g(s)),\\A^\#(f(s))}
+					\Rightarrow \substack{B^\#(g(s))),\\ B^\#(s), A^\#(f(s))}$}
+				\RightLabel{R$\to$}
+				\UnaryInfC{$A^\#(s, g(s))
+					\Rightarrow \substack{B^\#(g(s))), B^\#(s),\\A^\#(f(s))\to A^\#(f(s))}$}
+				\AxiomC{}
+				\RightLabel{Ax}
+				\UnaryInfC{$B^\#(s), A^\#(g(s))
+					\Rightarrow \substack{B^\#(g(s))),\\ B^\#(s)}$}
+				\RightLabel{L$\to$}
+				\BinaryInfC{$(A^\#(f(s))\to A^\#(f(s)))\to B^\#(s), A^\#(g(s))
+					\Rightarrow B^\#(g(s)), B^\#(s)$}
+				\RightLabel{L$\forall$}
+				\UnaryInfC{$\forall u((A^\#(f(u))\to A^\#(f(u)))\to B^\#(u)), A^\#(g(s))
+					\Rightarrow B^\#(g(s)), B^\#(s)$}
+				\RightLabel{R$\to$}
+					\UnaryInfC{$\forall u((A^\#(f(u))\to A^\#(f(u)))\to B^\#(u))
+						\Rightarrow A^\#(g(s))\to B^\#(g(s)), B^\#(s)$}
+				\AxiomC{\vdots}
+				\RightLabel{L$\to$}
+				\BinaryInfC{$\substack{\forall u((A^\#(f(u))\to A^\#(f(u)))\to B^\#(u))\\
+						A^\#(g(s))\to B^\#(g(s)))\to B^\#(s)}\Rightarrow B^\#(s)$}
+				\RightLabel{L$\forall$}
+				\UnaryInfC{$\substack{\forall u((A^\#(f(u))\to A^\#(f(u)))\to B^\#(u))\\
+						\forall u(\forall x(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))}\Rightarrow B^\#(s)$}
+					\DisplayProof
+			\end{center}
+			First we push all the L$\forall$ rules to the bottom and group L$\to$ with R$\to$/R$\forall$:
+		\begin{center}
+			\AxiomC{}
+			\RightLabel{Ax}
+			\UnaryInfC{$\substack{A^\#(g(s)),\\A^\#(f(s))}
+				\Rightarrow \substack{B^\#(g(s))),\\ B^\#(s), A^\#(f(s))}$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$A^\#(s, g(s))
+				\Rightarrow \substack{B^\#(g(s))), B^\#(s),\\A^\#(f(s))\to A^\#(f(s))}$}
+			\AxiomC{}
+			\RightLabel{Ax}
+			\UnaryInfC{$B^\#(s), A^\#(g(s))
+				\Rightarrow \substack{B^\#(g(s))),\\ B^\#(s)}$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$(A^\#(f(s))\to A^\#(f(s))\to B^\#(s)), A^\#(g(s))
+				\Rightarrow B^\#(g(s)), B^\#(s)$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$(A^\#(f(s))\to A^\#(f(s))\to B^\#(s))
+				\Rightarrow A^\#(g(s))\to B^\#(g(s)), B^\#(s)$}
+			\AxiomC{\vdots}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$\substack{(A^\#(f(s))\to A^\#(f(s))\to B^\#(s))\\
+					A^\#(g(s))\to B^\#(s, g(s)))\to B^\#(s)}\Rightarrow B^\#(s)$}
+			\RightLabel{L$\forall$}
+			\UnaryInfC{$\substack{(A^\#(f(s))\to A^\#(f(s))\to B^\#(s))\\
+					A^\#(g(s))\to B^\#(g(s)))\to B^\#(s)}\Rightarrow B^\#(s)$}
+			\RightLabel{L$\forall$}
+			\UnaryInfC{$\substack{\forall u((A^\#(f(u))\to A^\#(f(u)))\to B^\#(u))\\
+					\forall u(\forall x(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))}\Rightarrow B^\#(s)$}
+			\DisplayProof
+		\end{center}
+
+		Next we consider the subproof spanned by formulas containing $g(s)$ and rules featuring $g(s)$ above the last R$\to$ rule. But there are no rules featuring $g(s)$, i.e. we obtain an invalid proof.
+		\begin{center}
+			\AxiomC{}
+			\UnaryInfC{$A^\#(g(s))
+				\Rightarrow B^\#(g(s))$}
+			\DisplayProof
+		\end{center}
+		Therefore we remove all formulas and rules featuring $g(s)$: 
+		\begin{center}
+			\AxiomC{}
+			\RightLabel{Ax}
+			\UnaryInfC{$A^\#(g(s))
+				\Rightarrow B^\#(s), A^\#(f(s))$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$\Rightarrow B^\#(s), A^\#(f(s))\to A^\#(f(s))$}
+			\AxiomC{}
+			\RightLabel{Ax}
+			\UnaryInfC{$B^\#(s)
+				\Rightarrow B^\#(s)$}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$(A^\#(f(s))\to A^\#(f(s))\to B^\#(s))\Rightarrow B^\#(s)$}
+			\RightLabel{L$\forall$}
+			\UnaryInfC{$\forall u((A^\#(f(u))\to A^\#(f(u)))\to B^\#(u))\Rightarrow B^\#(s)$}
+			\DisplayProof
+		\end{center}
+		Then we add the superfluous formulas back in the axioms:
+		\begin{center}
+			\AxiomC{}
+			\RightLabel{Ax}
+			\UnaryInfC{$\substack{A^\#(g(s)),\\\forall u(\forall x(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))}
+				\Rightarrow B^\#(s), A^\#(f(s))$}
+			\RightLabel{R$\to$}
+			\UnaryInfC{$\forall u(\forall x(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))\Rightarrow B^\#(s), A^\#(f(s))\to A^\#(f(s))$}
+			\AxiomC{}
+			\RightLabel{Ax}
+			\UnaryInfC{\vdots}
+			\RightLabel{L$\to$}
+			\BinaryInfC{$\substack{(A^\#(f(s))\to A^\#(f(s))\to B^\#(s)),\\\forall u(\forall x(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))}\Rightarrow B^\#(s)$}
+			\RightLabel{L$\forall$}
+			\UnaryInfC{$\substack{\forall u((A^\#(f(u))\to A^\#(f(u)))\to B^\#(u))\\\forall u(\forall x(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))}\Rightarrow B^\#(s)$}
+			\DisplayProof
+		\end{center}
+	We then inductively proceed by examining the new last R$\to$ rule.
+This time however the subproof spanned by formulas and rules containing $f(s)$ is completely valid.
+Since the subproof does not contain any more R$\to$ or R$\forall$ formulas we are done.
+	\end{example}
+
+	\chapter{Conclusion}
+
+\section{Conclusion}
+
+	We have presented embeddings of intuitionistic into classical logic for the propositional and the predicate case.
+	The transformation saw an exponential blow-up parameterized by $|\mathcal F_\to|$ in the propositional case and a complexity increase reflected by an arity-increase of one for all relations and the introduction of new function symbols in the predicate case.
+	A key motivation for our work was a new approach for automated deduction for intuitionistic logic leveraging classical provers.
+	Our initial benchmarks showed decent results and proved this to be a legitimate route.
+
+	So far our translation is very straightforward and while this was conceptually very helpful we expect that better results can be achieved by more elaborate mechanisms.	We plan to establish better bounds in future work by utilizing structural properties of the input formula, in particular by constructing smaller counter-examples $\mathcal M_T$ in case certain atoms are independent from each other, and utilizing additional information about the functions in $\mathcal F$ to reduce the space of Kripke Frames that need to be considered.
+
+	On the theoretical side, we also hope to give a new translation from QBF to IPC that improves our understanding of the relationship between intuitionistic propositional logic and the polynomial hierarchy.
+
+	\appendix
+	\chapter{Zusammenfassung auf Deutsch}
+
+	\chapter{Complete Benchmark Results}\label{benchmark-results}
+	\singlespacing
+	All benchmarks that are not listed ended with timeout.\\
+
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{Time}\\
+			\text{AGT001+1.p}&\text{Proven}& 0.054 s\\
+			\text{AGT001+2.p}&\text{Proven}& 0.085 s\\
+			\text{AGT002+1.p}&\text{Proven}& 0.059 s\\
+			\text{AGT002+2.p}&\text{Proven}& 0.066 s\\
+			\text{AGT017+2.p}&\text{Proven}& 4.868 s\\
+			\text{ALG014+1.p}&\text{Proven}& 0.386 s\\
+			\text{ALG015+1.p}&\text{Proven}& 0.749 s\\
+			\text{ALG016+1.p}&\text{Proven}& 0.405 s\\
+			\text{ALG017+1.p}&\text{Proven}& 0.043 s\\
+			\text{ALG020+1.p}&\text{Proven}& 0.205 s\\
+			\text{ALG021+1.p}&\text{Proven}& 0.457 s\\
+			\text{ALG022+1.p}&\text{Proven}& 0.493 s\\
+			\text{ALG023+1.p}&\text{Proven}& 0.713 s\\
+			\text{ALG024+1.p}&\text{Proven}& 1.059 s\\
+			\text{ALG025+1.p}&\text{Proven}& 2.532 s\\
+			\text{ALG026+1.p}&\text{Proven}& 1.316 s\\
+			\text{ALG027+1.p}&\text{Proven}& 5.175 s\\
+			\text{ALG028+1.p}&\text{Proven}& 0.355 s\\
+			\text{ALG031+1.p}&\text{Proven}& 0.375 s\\
+			\text{ALG032+1.p}&\text{Proven}& 1.275 s\\
+			\text{ALG033+1.p}&\text{Proven}& 1.182 s\\
+			\text{ALG034+1.p}&\text{Proven}& 9.518 s\\
+			\text{ALG035+1.p}&\text{Proven}& 9.285 s\\
+			\text{ALG036+1.p}&\text{Proven}& 0.302 s\\
+			\text{ALG037+1.p}&\text{Proven}& 1.004 s\\
+			\text{ALG038+1.p}&\text{Proven}& 0.762 s\\
+			\text{ALG039+1.p}&\text{Proven}& 0.042 s\\
+			\text{ALG042+1.p}&\text{Proven}& 0.202 s\\
+			\text{ALG043+1.p}&\text{Proven}& 0.196 s\\
+			\text{ALG044+1.p}&\text{Proven}& 0.249 s\\
+			\text{ALG045+1.p}&\text{Proven}& 0.571 s\\
+			\text{ALG046+1.p}&\text{Proven}& 0.832 s\\
+		\end{matrix}\]
+	\end{minipage}
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{Time}\\
+			\text{ALG048+1.p}&\text{Proven}& 1.586 s\\
+			\text{ALG049+1.p}&\text{Proven}& 2.467 s\\
+			\text{ALG050+1.p}&\text{Proven}& 2.347 s\\
+			\text{ALG051+1.p}&\text{Proven}& 1.639 s\\
+			\text{ALG052+1.p}&\text{Proven}& 3.164 s\\
+			\text{ALG053+1.p}&\text{Proven}& 2.168 s\\
+			\text{ALG054+1.p}&\text{Proven}& 0.169 s\\
+			\text{ALG055+1.p}&\text{Proven}& 0.899 s\\
+			\text{ALG056+1.p}&\text{Proven}& 0.892 s\\
+			\text{ALG057+1.p}&\text{Proven}& 1.005 s\\
+			\text{ALG058+1.p}&\text{Proven}& 0.673 s\\
+			\text{ALG059+1.p}&\text{Proven}& 0.976 s\\
+			\text{ALG060+1.p}&\text{Proven}& 1.346 s\\
+			\text{ALG061+1.p}&\text{Proven}& 0.594 s\\
+			\text{ALG062+1.p}&\text{Proven}& 0.440 s\\
+			\text{ALG063+1.p}&\text{Proven}& 0.456 s\\
+			\text{ALG064+1.p}&\text{Proven}& 0.782 s\\
+			\text{ALG065+1.p}&\text{Proven}& 0.528 s\\
+			\text{ALG066+1.p}&\text{Proven}& 0.553 s\\
+			\text{ALG067+1.p}&\text{Proven}& 0.574 s\\
+			\text{ALG068+1.p}&\text{Proven}& 0.962 s\\
+			\text{ALG075+1.p}&\text{Proven}& 0.135 s\\
+			\text{ALG076+1.p}&\text{Proven}& 0.266 s\\
+			\text{ALG077+1.p}&\text{Proven}& 0.200 s\\
+			\text{ALG078+1.p}&\text{Proven}& 0.190 s\\
+			\text{ALG079+1.p}&\text{Proven}& 0.203 s\\
+			\text{ALG080+1.p}&\text{Proven}& 0.334 s\\
+			\text{ALG081+1.p}&\text{Proven}& 0.196 s\\
+			\text{ALG082+1.p}&\text{Proven}& 0.228 s\\
+			\text{ALG083+1.p}&\text{Proven}& 0.168 s\\
+			\text{ALG084+1.p}&\text{Proven}& 0.211 s\\
+			\text{ALG085+1.p}&\text{Proven}& 0.206 s\\
+		\end{matrix}\]
+	\end{minipage}
+
+	\pagebreak
+
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{Time}\\
+			\text{ALG086+1.p}&\text{Proven}& 0.273 s\\
+			\text{ALG087+1.p}&\text{Proven}& 0.305 s\\
+			\text{ALG088+1.p}&\text{Proven}& 0.195 s\\
+			\text{ALG089+1.p}&\text{Proven}& 0.293 s\\
+			\text{ALG090+1.p}&\text{Proven}& 0.270 s\\
+			\text{ALG091+1.p}&\text{Proven}& 0.466 s\\
+			\text{ALG092+1.p}&\text{Proven}& 21.681 s\\
+			\text{ALG093+1.p}&\text{Proven}& 0.551 s\\
+			\text{ALG094+1.p}&\text{Proven}& 0.280 s\\
+			\text{ALG095+1.p}&\text{Proven}& 0.358 s\\
+			\text{ALG096+1.p}&\text{Proven}& 1.859 s\\
+			\text{ALG097+1.p}&\text{Proven}& 1.531 s\\
+			\text{ALG098+1.p}&\text{Proven}& 2.604 s\\
+			\text{ALG099+1.p}&\text{Proven}& 3.372 s\\
+			\text{ALG100+1.p}&\text{Proven}& 2.517 s\\
+			\text{ALG101+1.p}&\text{Proven}& 3.101 s\\
+			\text{ALG104+1.p}&\text{Proven}& 2.526 s\\
+			\text{ALG105+1.p}&\text{Proven}& 0.765 s\\
+			\text{ALG106+1.p}&\text{Proven}& 2.808 s\\
+			\text{ALG107+1.p}&\text{Proven}& 1.168 s\\
+			\text{ALG108+1.p}&\text{Proven}& 2.324 s\\
+			\text{ALG109+1.p}&\text{Proven}& 1.958 s\\
+			\text{ALG110+1.p}&\text{Proven}& 1.049 s\\
+			\text{ALG111+1.p}&\text{Proven}& 0.960 s\\
+			\text{ALG112+1.p}&\text{Proven}& 1.348 s\\
+			\text{ALG113+1.p}&\text{Proven}& 0.846 s\\
+			\text{ALG114+1.p}&\text{Proven}& 0.889 s\\
+			\text{ALG115+1.p}&\text{Proven}& 1.004 s\\
+			\text{ALG116+1.p}&\text{Proven}& 1.253 s\\
+			\text{ALG117+1.p}&\text{Proven}& 1.121 s\\
+			\text{ALG118+1.p}&\text{Proven}& 1.133 s\\
+			\text{ALG119+1.p}&\text{Proven}& 1.539 s\\
+			\text{ALG120+1.p}&\text{Proven}& 1.905 s\\
+			\text{ALG121+1.p}&\text{Proven}& 0.962 s\\
+			\text{ALG122+1.p}&\text{Proven}& 2.674 s\\
+			\text{ALG123+1.p}&\text{Proven}& 1.099 s\\
+			\text{ALG124+1.p}&\text{Proven}& 1.390 s\\
+			\text{ALG125+1.p}&\text{Proven}& 1.243 s\\
+			\text{ALG126+1.p}&\text{Proven}& 2.903 s\\
+			\text{ALG127+1.p}&\text{Proven}& 2.391 s\\
+			\text{ALG128+1.p}&\text{Proven}& 6.782 s\\
+			\text{ALG130+1.p}&\text{Proven}& 0.548 s\\
+			\text{ALG131+1.p}&\text{Proven}& 7.478 s\\
+			\text{ALG132+1.p}&\text{Proven}& 1.040 s\\
+			\text{ALG133+1.p}&\text{Proven}& 7.531 s\\
+			\text{ALG134+1.p}&\text{Proven}& 1.663 s\\
+			\text{ALG135+1.p}&\text{Proven}& 0.914 s\\
+			\text{ALG136+1.p}&\text{Proven}& 1.122 s\\
+		\end{matrix}\]
+	\end{minipage}
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{Time}\\
+			\text{ALG137+1.p}&\text{Proven}& 1.047 s\\
+			\text{ALG138+1.p}&\text{Proven}& 0.208 s\\
+			\text{ALG139+1.p}&\text{Proven}& 0.998 s\\
+			\text{ALG140+1.p}&\text{Proven}& 1.416 s\\
+			\text{ALG141+1.p}&\text{Proven}& 0.958 s\\
+			\text{ALG142+1.p}&\text{Proven}& 0.910 s\\
+			\text{ALG143+1.p}&\text{Proven}& 1.345 s\\
+			\text{ALG145+1.p}&\text{Proven}& 1.407 s\\
+			\text{ALG146+1.p}&\text{Proven}& 1.061 s\\
+			\text{ALG147+1.p}&\text{Proven}& 6.912 s\\
+			\text{ALG148+1.p}&\text{Proven}& 1.260 s\\
+			\text{ALG149+1.p}&\text{Proven}& 1.434 s\\
+			\text{ALG150+1.p}&\text{Proven}& 1.175 s\\
+			\text{ALG151+1.p}&\text{Proven}& 29.674 s\\
+			\text{ALG152+1.p}&\text{Proven}& 3.644 s\\
+			\text{ALG153+1.p}&\text{Proven}& 26.962 s\\
+			\text{ALG157+1.p}&\text{Proven}& 2.262 s\\
+			\text{ALG158+1.p}&\text{Proven}& 4.045 s\\
+			\text{ALG159+1.p}&\text{Proven}& 14.128 s\\
+			\text{ALG161+1.p}&\text{Proven}& 1.952 s\\
+			\text{ALG162+1.p}&\text{Proven}& 2.372 s\\
+			\text{ALG164+1.p}&\text{Proven}& 1.593 s\\
+			\text{ALG165+1.p}&\text{Proven}& 1.910 s\\
+			\text{ALG166+1.p}&\text{Proven}& 1.755 s\\
+			\text{ALG167+1.p}&\text{Proven}& 1.744 s\\
+			\text{ALG168+1.p}&\text{Proven}& 0.963 s\\
+			\text{ALG169+1.p}&\text{Proven}& 0.947 s\\
+			\text{ALG170+1.p}&\text{Proven}& 0.746 s\\
+			\text{ALG171+1.p}&\text{Proven}& 0.080 s\\
+			\text{ALG172+1.p}&\text{Proven}& 0.076 s\\
+			\text{ALG173+1.p}&\text{Proven}& 0.085 s\\
+			\text{ALG174+1.p}&\text{Proven}& 0.080 s\\
+			\text{ALG175+1.p}&\text{Proven}& 0.088 s\\
+			\text{ALG180+1.p}&\text{Proven}& 0.225 s\\
+			\text{ALG181+1.p}&\text{Proven}& 0.143 s\\
+			\text{ALG182+1.p}&\text{Proven}& 0.135 s\\
+			\text{ALG183+1.p}&\text{Proven}& 0.138 s\\
+			\text{ALG184+1.p}&\text{Proven}& 0.142 s\\
+			\text{ALG185+1.p}&\text{Proven}& 0.499 s\\
+			\text{ALG186+1.p}&\text{Proven}& 0.391 s\\
+			\text{ALG187+1.p}&\text{Proven}& 0.430 s\\
+			\text{ALG188+1.p}&\text{Proven}& 0.445 s\\
+			\text{ALG189+1.p}&\text{Proven}& 0.508 s\\
+			\text{ALG190+1.p}&\text{Proven}& 2.462 s\\
+			\text{ALG191+1.p}&\text{Proven}& 2.207 s\\
+			\text{ALG192+1.p}&\text{Proven}& 0.648 s\\
+			\text{ALG193+1.p}&\text{Proven}& 0.584 s\\
+			\text{ALG194+1.p}&\text{Proven}& 28.952 s\\
+		\end{matrix}\]
+	\end{minipage}
+
+	\pagebreak
+
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{Time}\\
+			\text{ALG195+1.p}&\text{Proven}& 4.695 s\\
+			\text{ALG196+1.p}&\text{Proven}& 7.115 s\\
+			\text{ALG197+1.p}&\text{Proven}& 4.805 s\\
+			\text{ALG198+1.p}&\text{Proven}& 0.127 s\\
+			\text{ALG199+1.p}&\text{Proven}& 1.533 s\\
+			\text{ALG200+1.p}&\text{Proven}& 0.263 s\\
+			\text{ALG203+1.p}&\text{Proven}& 26.932 s\\
+			\text{ALG204+1.p}&\text{Proven}& 0.242 s\\
+			\text{ALG205+1.p}&\text{Proven}& 0.320 s\\
+			\text{ALG206+1.p}&\text{Proven}& 0.274 s\\
+			\text{ALG207+1.p}&\text{Proven}& 0.823 s\\
+			\text{ALG208+1.p}&\text{Proven}& 0.703 s\\
+			\text{ALG209+1.p}&\text{Proven}& 0.783 s\\
+			\text{ALG210+2.p}&\text{Proven}& 6.846 s\\
+			\text{ALG211+1.p}&\text{Proven}& 0.327 s\\
+			\text{GEJ002+1.p}&\text{Proven}& 7.672 s\\
+			\text{GEJ002+2.p}&\text{Proven}& 0.030 s\\
+			\text{GEJ004+3.p}&\text{Proven}& 23.293 s\\
+			\text{GEJ007+2.p}&\text{Proven}& 0.156 s\\
+			\text{GEJ007+4.p}&\text{Proven}& 0.098 s\\
+			\text{GEJ009+1.p}&\text{Proven}& 0.487 s\\
+			\text{GEJ009+2.p}&\text{Proven}& 0.054 s\\
+			\text{GEJ009+3.p}&\text{Proven}& 0.338 s\\
+			\text{GEJ009+4.p}&\text{Proven}& 0.046 s\\
+			\text{GEJ010+1.p}&\text{Proven}& 2.915 s\\
+			\text{GEJ010+2.p}&\text{Proven}& 7.168 s\\
+			\text{GEJ010+3.p}&\text{Proven}& 1.635 s\\
+			\text{GEJ038+1.p}&\text{Proven}& 0.029 s\\
+			\text{GEJ038+2.p}&\text{Proven}& 0.028 s\\
+			\text{GEJ042+1.p}&\text{Proven}& 0.057 s\\
+			\text{GEJ042+2.p}&\text{Proven}& 0.055 s\\
+			\text{GEJ043+2.p}&\text{Proven}& 17.815 s\\
+			\text{GEJ044+2.p}&\text{Proven}& 29.105 s\\
+			\text{GEJ051+1.p}&\text{Proven}& 0.030 s\\
+			\text{GEJ051+2.p}&\text{Proven}& 0.030 s\\
+			\text{GEJ060+1.p}&\text{Proven}& 0.082 s\\
+			\text{GEJ061+1.p}&\text{Proven}& 0.153 s\\
+			\text{GEJ062+1.p}&\text{Proven}& 0.036 s\\
+			\text{GEJ063+1.p}&\text{Proven}& 0.085 s\\
+			\text{GEJ064+1.p}&\text{Proven}& 0.093 s\\
+			\text{GEJ065+1.p}&\text{Proven}& 0.062 s\\
+			\text{GEJ066+1.p}&\text{Proven}& 0.078 s\\
+			\text{GEJ067+1.p}&\text{Proven}& 0.080 s\\
+			\text{GEJ068+1.p}&\text{Proven}& 0.071 s\\
+			\text{GEJ069+1.p}&\text{Proven}& 0.073 s\\
+			\text{GEJ070+1.p}&\text{Proven}& 0.111 s\\
+			\text{GEJ071+1.p}&\text{Proven}& 0.117 s\\
+			\text{GEJ072+1.p}&\text{Proven}& 0.103 s\\
+		\end{matrix}\]
+	\end{minipage}
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{Time}\\
+			\text{GEJ073+1.p}&\text{Proven}& 0.106 s\\
+			\text{GEJ074+1.p}&\text{Proven}& 0.102 s\\
+			\text{GEJ075+1.p}&\text{Proven}& 0.106 s\\
+			\text{GEJ076+1.p}&\text{Proven}& 0.091 s\\
+			\text{GEJ077+1.p}&\text{Proven}& 0.093 s\\
+			\text{GEJ078+1.p}&\text{Proven}& 0.090 s\\
+			\text{GEJ079+1.p}&\text{Proven}& 0.095 s\\
+			\text{GEJ080+1.p}&\text{Proven}& 0.120 s\\
+			\text{GEJ081+1.p}&\text{Proven}& 0.063 s\\
+			\text{GEJ082+1.p}&\text{Proven}& 0.071 s\\
+			\text{GEJ083+1.p}&\text{Proven}& 0.082 s\\
+			\text{GEJ084+1.p}&\text{Proven}& 0.088 s\\
+			\text{GEJ085+1.p}&\text{Proven}& 0.074 s\\
+			\text{GEJ086+1.p}&\text{Proven}& 0.083 s\\
+			\text{GEJ087+1.p}&\text{Proven}& 0.108 s\\
+			\text{GEJ088+1.p}&\text{Proven}& 0.108 s\\
+			\text{GEJ089+1.p}&\text{Proven}& 0.121 s\\
+			\text{GEJ090+1.p}&\text{Proven}& 0.125 s\\
+			\text{GEJ091+1.p}&\text{Proven}& 0.102 s\\
+			\text{GEJ092+1.p}&\text{Proven}& 0.110 s\\
+			\text{GEJ093+1.p}&\text{Proven}& 0.106 s\\
+			\text{GEJ094+1.p}&\text{Proven}& 0.117 s\\
+			\text{GEJ095+1.p}&\text{Proven}& 0.113 s\\
+			\text{GEJ096+1.p}&\text{Proven}& 0.119 s\\
+			\text{GEJ097+1.p}&\text{Proven}& 0.091 s\\
+			\text{GEO080+1.p}&\text{Proven}& 0.227 s\\
+			\text{GEO085+1.p}&\text{Proven}& 0.266 s\\
+			\text{GEO086+1.p}&\text{Proven}& 0.697 s\\
+			\text{GPJ001+1.p}&\text{Proven}& 13.428 s\\
+			\text{GPJ001+2.p}&\text{Proven}& 0.187 s\\
+			\text{GPJ002+1.p}&\text{Proven}& 0.067 s\\
+			\text{GPJ003+1.p}&\text{Proven}& 0.077 s\\
+			\text{GRA010+1.p}&\text{Proven}& 0.038 s\\
+			\text{GRA010+2.p}&\text{Proven}& 0.038 s\\
+			\text{KRS065+1.p}&\text{Proven}& 0.039 s\\
+			\text{KRS130+1.p}&\text{Proven}& 0.097 s\\
+			\text{KRS134+1.p}&\text{Proven}& 0.052 s\\
+			\text{KRS135+1.p}&\text{Proven}& 0.049 s\\
+			\text{KRS136+1.p}&\text{Proven}& 0.031 s\\
+			\text{KRS137+1.p}&\text{Proven}& 0.074 s\\
+			\text{KRS139+1.p}&\text{Proven}& 0.128 s\\
+			\text{KRS141+1.p}&\text{Proven}& 0.083 s\\
+			\text{KRS142+1.p}&\text{Proven}& 0.096 s\\
+			\text{KRS143+1.p}&\text{Proven}& 0.084 s\\
+			\text{KRS144+1.p}&\text{Proven}& 0.105 s\\
+			\text{KRS145+1.p}&\text{Proven}& 0.090 s\\
+			\text{KRS146+1.p}&\text{Proven}& 1.792 s\\
+			\text{KRS148+1.p}&\text{Proven}& 0.331 s\\
+		\end{matrix}\]
+	\end{minipage}
+
+	\pagebreak
+
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{Time}\\
+			\text{KRS149+1.p}&\text{Proven}& 0.493 s\\
+			\text{KRS150+1.p}&\text{Proven}& 0.247 s\\
+			\text{KRS152+1.p}&\text{Proven}& 0.090 s\\
+			\text{KRS154+1.p}&\text{Proven}& 0.788 s\\
+			\text{KRS156+1.p}&\text{Proven}& 0.539 s\\
+			\text{KRS157+1.p}&\text{Proven}& 1.134 s\\
+			\text{KRS158+1.p}&\text{Proven}& 0.184 s\\
+			\text{KRS160+1.p}&\text{Proven}& 0.102 s\\
+			\text{KRS163+1.p}&\text{Proven}& 0.065 s\\
+			\text{KRS164+1.p}&\text{Proven}& 0.072 s\\
+			\text{KRS165+1.p}&\text{Proven}& 0.045 s\\
+			\text{KRS166+1.p}&\text{Proven}& 0.047 s\\
+			\text{KRS167+1.p}&\text{Proven}& 0.170 s\\
+			\text{KRS169+1.p}&\text{Proven}& 1.030 s\\
+			\text{KRS170+1.p}&\text{Proven}& 0.048 s\\
+			\text{KRS171+1.p}&\text{Proven}& 0.055 s\\
+			\text{LCL414+1.p}&\text{Proven}& 0.880 s\\
+			\text{MGT003+1.p}&\text{Proven}& 3.148 s\\
+			\text{MGT006+1.p}&\text{Proven}& 27.800 s\\
+			\text{MGT009+1.p}&\text{Proven}& 28.410 s\\
+			\text{MGT013+1.p}&\text{Proven}& 10.306 s\\
+			\text{MGT014+1.p}&\text{Proven}& 10.485 s\\
+			\text{MGT036+3.p}&\text{Proven}& 0.087 s\\
+			\text{MGT045+1.p}&\text{Proven}& 0.077 s\\
+			\text{MGT049+1.p}&\text{Proven}& 0.059 s\\
+			\text{MGT052+1.p}&\text{Proven}& 0.086 s\\
+			\text{MSC010+1.p}&\text{Proven}& 0.038 s\\
+			\text{NLP001+1.p}&\text{Proven}& 0.237 s\\
+			\text{NLP002+1.p}&\text{Refuted}& 14.428 s\\
+			\text{NLP003+1.p}&\text{Refuted}& 14.224 s\\
+			\text{NLP004+1.p}&\text{Proven}& 4.340 s\\
+			\text{NLP007+1.p}&\text{Proven}& 4.476 s\\
+			\text{NLP009+1.p}&\text{Proven}& 4.150 s\\
+			\text{NLP011+1.p}&\text{Proven}& 4.546 s\\
+			\text{NLP104+1.p}&\text{Refuted}& 1.138 s\\
+			\text{NLP105+1.p}&\text{Refuted}& 1.216 s\\
+			\text{NLP106+1.p}&\text{Refuted}& 1.216 s\\
+			\text{NLP107+1.p}&\text{Refuted}& 1.241 s\\
+			\text{NLP108+1.p}&\text{Refuted}& 1.088 s\\
+			\text{NLP109+1.p}&\text{Refuted}& 1.178 s\\
+			\text{NLP110+1.p}&\text{Refuted}& 1.068 s\\
+			\text{NLP111+1.p}&\text{Refuted}& 1.095 s\\
+			\text{NLP112+1.p}&\text{Refuted}& 1.112 s\\
+			\text{NLP113+1.p}&\text{Refuted}& 1.380 s\\
+			\text{NLP117+1.p}&\text{Proven}& 0.304 s\\
+			\text{NLP122+1.p}&\text{Proven}& 0.302 s\\
+			\text{NLP204+1.p}&\text{Proven}& 3.293 s\\
+			\text{NLP208+1.p}&\text{Proven}& 3.487 s\\
+		\end{matrix}\]
+	\end{minipage}
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{Time}\\
+			\text{NUM304+1.p}&\text{Proven}& 2.448 s\\
+			\text{NUM333+1.p}&\text{Proven}& 27.096 s\\
+			\text{PUZ001+1.p}&\text{Proven}& 0.215 s\\
+			\text{PUZ031+1.p}&\text{Proven}& 0.252 s\\
+			\text{PUZ047+1.p}&\text{Proven}& 0.248 s\\
+			\text{PUZ060+1.p}&\text{Proven}& 0.113 s\\
+			\text{PUZ061+1.p}&\text{Proven}& 0.104 s\\
+			\text{SET002+3.p}&\text{Proven}& 0.050 s\\
+			\text{SET019+4.p}&\text{Proven}& 0.051 s\\
+			\text{SET047+1.p}&\text{Proven}& 0.656 s\\
+			\text{SET054+1.p}&\text{Proven}& 0.833 s\\
+			\text{SET055+1.p}&\text{Proven}& 0.041 s\\
+			\text{SET061+1.p}&\text{Proven}& 1.388 s\\
+			\text{SET062+3.p}&\text{Proven}& 0.127 s\\
+			\text{SET062+4.p}&\text{Proven}& 0.197 s\\
+			\text{SET063+3.p}&\text{Proven}& 6.494 s\\
+			\text{SET148+3.p}&\text{Proven}& 0.052 s\\
+			\text{SET194+3.p}&\text{Proven}& 1.440 s\\
+			\text{SET366+4.p}&\text{Proven}& 0.215 s\\
+			\text{SET574+3.p}&\text{Proven}& 0.096 s\\
+			\text{SET575+3.p}&\text{Proven}& 0.035 s\\
+			\text{SET583+3.p}&\text{Proven}& 0.036 s\\
+			\text{SET585+3.p}&\text{Proven}& 4.499 s\\
+			\text{SET589+3.p}&\text{Proven}& 0.155 s\\
+			\text{SET590+3.p}&\text{Proven}& 29.211 s\\
+			\text{SET602+3.p}&\text{Proven}& 0.494 s\\
+			\text{SET604+3.p}&\text{Proven}& 0.216 s\\
+			\text{SET618+3.p}&\text{Proven}& 17.657 s\\
+			\text{SET627+3.p}&\text{Proven}& 0.401 s\\
+			\text{SET631+3.p}&\text{Proven}& 0.306 s\\
+			\text{SET639+3.p}&\text{Proven}& 13.976 s\\
+			\text{SET658+3.p}&\text{Proven}& 0.873 s\\
+			\text{SET687+4.p}&\text{Proven}& 0.223 s\\
+			\text{SET705+4.p}&\text{Proven}& 0.240 s\\
+			\text{SWC128+1.p}&\text{Proven}& 0.086 s\\
+			\text{SWV011+1.p}&\text{Proven}& 0.053 s\\
+			\text{SWV014+1.p}&\text{Proven}& 0.281 s\\
+			\text{SWV016+1.p}&\text{Refuted}& 0.879 s\\
+			\text{SWV018+1.p}&\text{Refuted}& 0.973 s\\
+			\text{SWV022+1.p}&\text{Proven}& 0.054 s\\
+			\text{SWV023+1.p}&\text{Proven}& 0.050 s\\
+			\text{SWV043+1.p}&\text{Proven}& 0.039 s\\
+			\text{SWV045+1.p}&\text{Proven}& 0.041 s\\
+			\text{SWV047+1.p}&\text{Proven}& 0.042 s\\
+			\text{SWV054+1.p}&\text{Proven}& 0.042 s\\
+			\text{SWV057+1.p}&\text{Proven}& 0.039 s\\
+			\text{SWV058+1.p}&\text{Proven}& 0.042 s\\
+			\text{SWV059+1.p}&\text{Proven}& 0.065 s\\
+		\end{matrix}\]
+	\end{minipage}
+
+	\pagebreak
+
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{Time}\\
+			\text{SWV060+1.p}&\text{Proven}& 0.041 s\\
+			\text{SWV061+1.p}&\text{Proven}& 0.046 s\\
+			\text{SWV062+1.p}&\text{Proven}& 0.039 s\\
+			\text{SWV063+1.p}&\text{Proven}& 0.054 s\\
+			\text{SWV064+1.p}&\text{Proven}& 0.161 s\\
+			\text{SWV065+1.p}&\text{Proven}& 0.140 s\\
+			\text{SWV066+1.p}&\text{Proven}& 0.044 s\\
+			\text{SWV068+1.p}&\text{Proven}& 0.040 s\\
+			\text{SWV069+1.p}&\text{Proven}& 12.016 s\\
+			\text{SWV070+1.p}&\text{Proven}& 0.040 s\\
+			\text{SWV071+1.p}&\text{Proven}& 0.054 s\\
+			\text{SWV072+1.p}&\text{Proven}& 0.043 s\\
+			\text{SWV073+1.p}&\text{Proven}& 0.053 s\\
+			\text{SWV074+1.p}&\text{Proven}& 0.049 s\\
+			\text{SWV075+1.p}&\text{Proven}& 0.057 s\\
+			\text{SWV076+1.p}&\text{Proven}& 0.045 s\\
+			\text{SWV077+1.p}&\text{Proven}& 17.571 s\\
+			\text{SWV078+1.p}&\text{Proven}& 2.419 s\\
+			\text{SWV080+1.p}&\text{Proven}& 0.051 s\\
+			\text{SWV081+1.p}&\text{Proven}& 0.043 s\\
+			\text{SWV082+1.p}&\text{Proven}& 0.043 s\\
+			\text{SWV083+1.p}&\text{Proven}& 0.051 s\\
+			\text{SWV084+1.p}&\text{Proven}& 0.045 s\\
+			\text{SWV085+1.p}&\text{Proven}& 0.041 s\\
+			\text{SWV086+1.p}&\text{Proven}& 0.040 s\\
+			\text{SWV087+1.p}&\text{Proven}& 0.044 s\\
+			\text{SWV088+1.p}&\text{Proven}& 0.041 s\\
+			\text{SWV105+1.p}&\text{Proven}& 0.043 s\\
+			\text{SWV106+1.p}&\text{Proven}& 0.040 s\\
+			\text{SWV107+1.p}&\text{Proven}& 0.049 s\\
+			\text{SWV119+1.p}&\text{Proven}& 0.042 s\\
+			\text{SWV121+1.p}&\text{Proven}& 0.038 s\\
+			\text{SWV123+1.p}&\text{Proven}& 0.050 s\\
+			\text{SWV126+1.p}&\text{Proven}& 0.038 s\\
+			\text{SWV128+1.p}&\text{Proven}& 0.040 s\\
+			\text{SWV131+1.p}&\text{Proven}& 0.051 s\\
+			\text{SWV132+1.p}&\text{Proven}& 0.052 s\\
+			\text{SWV145+1.p}&\text{Proven}& 0.161 s\\
+			\text{SWV146+1.p}&\text{Proven}& 0.229 s\\
+			\text{SWV147+1.p}&\text{Proven}& 0.141 s\\
+			\text{SWV148+1.p}&\text{Proven}& 0.138 s\\
+			\text{SWV149+1.p}&\text{Proven}& 0.139 s\\
+			\text{SWV150+1.p}&\text{Proven}& 0.137 s\\
+			\text{SWV169+1.p}&\text{Proven}& 0.266 s\\
+			\text{SWV171+1.p}&\text{Proven}& 0.388 s\\
+			\text{SWV172+1.p}&\text{Proven}& 0.447 s\\
+			\text{SWV179+1.p}&\text{Proven}& 0.324 s\\
+			\text{SWV180+1.p}&\text{Proven}& 0.362 s\\
+		\end{matrix}\]
+	\end{minipage}
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{Time}\\
+			\text{SWV192+1.p}&\text{Proven}& 0.394 s\\
+			\text{SWV193+1.p}&\text{Proven}& 0.510 s\\
+			\text{SWV194+1.p}&\text{Proven}& 0.520 s\\
+			\text{SWV195+1.p}&\text{Proven}& 0.457 s\\
+			\text{SWV196+1.p}&\text{Proven}& 0.387 s\\
+			\text{SYJ001+1.001.p}&\text{Proven}& 0.030 s\\
+			\text{SYJ002+1.001.p}&\text{Proven}& 0.033 s\\
+			\text{SYJ004+1.001.p}&\text{Proven}& 0.052 s\\
+			\text{SYJ004+1.002.p}&\text{Proven}& 0.053 s\\
+			\text{SYJ004+1.003.p}&\text{Proven}& 0.055 s\\
+			\text{SYJ004+1.004.p}&\text{Proven}& 0.085 s\\
+			\text{SYJ004+1.005.p}&\text{Proven}& 0.078 s\\
+			\text{SYJ004+1.006.p}&\text{Proven}& 0.079 s\\
+			\text{SYJ004+1.007.p}&\text{Proven}& 0.073 s\\
+			\text{SYJ004+1.008.p}&\text{Proven}& 0.070 s\\
+			\text{SYJ004+1.009.p}&\text{Proven}& 0.096 s\\
+			\text{SYJ004+1.010.p}&\text{Proven}& 0.078 s\\
+			\text{SYJ004+1.011.p}&\text{Proven}& 0.101 s\\
+			\text{SYJ004+1.012.p}&\text{Proven}& 0.091 s\\
+			\text{SYJ004+1.013.p}&\text{Proven}& 0.104 s\\
+			\text{SYJ004+1.014.p}&\text{Proven}& 0.096 s\\
+			\text{SYJ004+1.015.p}&\text{Proven}& 0.082 s\\
+			\text{SYJ004+1.016.p}&\text{Proven}& 0.098 s\\
+			\text{SYJ004+1.017.p}&\text{Proven}& 0.092 s\\
+			\text{SYJ004+1.018.p}&\text{Proven}& 0.099 s\\
+			\text{SYJ004+1.019.p}&\text{Proven}& 0.097 s\\
+			\text{SYJ004+1.020.p}&\text{Proven}& 0.100 s\\
+			\text{SYJ013+1.p}&\text{Proven}& 0.057 s\\
+			\text{SYJ014+1.p}&\text{Proven}& 0.048 s\\
+			\text{SYJ015+1.p}&\text{Proven}& 0.050 s\\
+			\text{SYJ025+1.p}&\text{Proven}& 0.065 s\\
+			\text{SYJ026+1.p}&\text{Proven}& 0.047 s\\
+			\text{SYJ027+1.p}&\text{Proven}& 0.047 s\\
+			\text{SYJ028+1.p}&\text{Proven}& 0.068 s\\
+			\text{SYJ031+1.p}&\text{Proven}& 0.090 s\\
+			\text{SYJ033+1.p}&\text{Proven}& 0.099 s\\
+			\text{SYJ034+1.p}&\text{Proven}& 0.135 s\\
+			\text{SYJ035+1.p}&\text{Proven}& 1.537 s\\
+			\text{SYJ038+1.p}&\text{Proven}& 0.049 s\\
+			\text{SYJ039+1.p}&\text{Proven}& 0.086 s\\
+			\text{SYJ101+1.p}&\text{Proven}& 0.028 s\\
+			\text{SYJ102+1.p}&\text{Proven}& 0.032 s\\
+			\text{SYJ103+1.p}&\text{Proven}& 0.032 s\\
+			\text{SYJ104+1.p}&\text{Proven}& 0.030 s\\
+			\text{SYJ107+1.001.p}&\text{Proven}& 0.035 s\\
+			\text{SYJ107+1.002.p}&\text{Proven}& 0.047 s\\
+			\text{SYJ107+1.003.p}&\text{Proven}& 0.057 s\\
+			\text{SYJ107+1.004.p}&\text{Proven}& 0.068 s\\
+		\end{matrix}\]
+	\end{minipage}
+
+	\pagebreak
+
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{Time}\\
+			\text{SYJ108+1.p}&\text{Proven}& 0.026 s\\
+			\text{SYJ109+1.p}&\text{Proven}& 0.036 s\\
+			\text{SYJ111+1.p}&\text{Proven}& 0.049 s\\
+			\text{SYJ112+1.p}&\text{Proven}& 0.036 s\\
+			\text{SYJ114+1.p}&\text{Proven}& 0.046 s\\
+			\text{SYJ115+1.p}&\text{Proven}& 0.045 s\\
+			\text{SYJ116+1.001.p}&\text{Proven}& 0.058 s\\
+			\text{SYJ116+1.002.p}&\text{Proven}& 1.048 s\\
+			\text{SYJ117+1.p}&\text{Proven}& 0.048 s\\
+			\text{SYJ118+1.p}&\text{Proven}& 0.073 s\\
+			\text{SYJ119+1.p}&\text{Proven}& 0.052 s\\
+			\text{SYJ120+1.p}&\text{Proven}& 0.052 s\\
+			\text{SYJ121+1.p}&\text{Proven}& 0.057 s\\
+			\text{SYJ122+1.p}&\text{Proven}& 0.072 s\\
+			\text{SYJ123+1.p}&\text{Proven}& 0.059 s\\
+			\text{SYJ124+1.p}&\text{Proven}& 0.062 s\\
+			\text{SYJ202+1.002.p}&\text{Proven}& 0.056 s\\
+			\text{SYJ202+1.006.p}&\text{Proven}& 0.956 s\\
+			\text{SYJ204+1.002.p}&\text{Proven}& 0.037 s\\
+			\text{SYJ204+1.006.p}&\text{Proven}& 0.052 s\\
+			\text{SYJ204+1.010.p}&\text{Proven}& 0.073 s\\
+			\text{SYJ206+1.002.p}&\text{Proven}& 0.035 s\\
+			\text{SYJ210+1.002.p}&\text{Refuted}& 0.037 s\\
+			\text{SYJ210+1.006.p}&\text{Refuted}& 0.050 s\\
+			\text{SYJ210+1.010.p}&\text{Refuted}& 0.064 s\\
+			\text{SYN044+1.p}&\text{Proven}& 0.045 s\\
+			\text{SYN045+1.p}&\text{Proven}& 0.048 s\\
+			\text{SYN054+1.p}&\text{Refuted}& 0.089 s\\
+			\text{SYN055+1.p}&\text{Proven}& 0.061 s\\
+			\text{SYN057+1.p}&\text{Proven}& 0.104 s\\
+			\text{SYN058+1.p}&\text{Proven}& 0.096 s\\
+			\text{SYN061+1.p}&\text{Proven}& 0.057 s\\
+			\text{SYN062+1.p}&\text{Proven}& 0.052 s\\
+			\text{SYN065+1.p}&\text{Proven}& 0.066 s\\
+			\text{SYN066+1.p}&\text{Proven}& 0.086 s\\
+			\text{SYN071+1.p}&\text{Proven}& 0.059 s\\
+			\text{SYN072+1.p}&\text{Proven}& 0.089 s\\
+			\text{SYN079+1.p}&\text{Proven}& 0.056 s\\
+			\text{SYN080+1.p}&\text{Proven}& 0.218 s\\
+			\text{SYN082+1.p}&\text{Proven}& 0.101 s\\
+			\text{SYN323+1.p}&\text{Proven}& 0.066 s\\
+			\text{SYN346+1.p}&\text{Proven}& 0.061 s\\
+			\text{SYN356+1.p}&\text{Proven}& 0.076 s\\
+			\text{SYN357+1.p}&\text{Proven}& 0.041 s\\
+			\text{SYN358+1.p}&\text{Proven}& 0.048 s\\
+			\text{SYN359+1.p}&\text{Proven}& 0.060 s\\
+			\text{SYN379+1.p}&\text{Proven}& 0.062 s\\
+			\text{SYN387+1.p}&\text{Refuted}& 0.029 s\\
+		\end{matrix}\]
+	\end{minipage}
+	\begin{minipage}{0.45\textwidth}
+		\[\begin{matrix}
+			\text{Problem}&\text{Result}&\text{time}\\
+			\text{SYN390+1.p}&\text{Proven}& 0.035 s\\
+			\text{SYN401+1.p}&\text{Proven}& 0.045 s\\
+			\text{SYN402+1.p}&\text{Proven}& 0.044 s\\
+			\text{SYN404+1.p}&\text{Proven}& 0.045 s\\
+			\text{SYN405+1.p}&\text{Proven}& 0.049 s\\
+			\text{SYN406+1.p}&\text{Proven}& 0.060 s\\
+			\text{SYN410+1.p}&\text{Proven}& 0.044 s\\
+			\text{SYN490+1.p}&\text{Refuted}& 13.282 s\\
+			\text{SYN491+1.p}&\text{Refuted}& 20.030 s\\
+			\text{SYN495+1.p}&\text{Refuted}& 5.038 s\\
+			\text{SYN497+1.p}&\text{Refuted}& 25.937 s\\
+			\text{SYN516+1.p}&\text{Refuted}& 3.231 s\\
+			\text{SYN517+1.p}&\text{Refuted}& 1.989 s\\
+			\text{SYN523+1.p}&\text{Refuted}& 1.218 s\\
+			\text{SYN527+1.p}&\text{Refuted}& 8.835 s\\
+			\text{SYN532+1.p}&\text{Refuted}& 3.138 s\\
+			\text{SYN533+1.p}&\text{Refuted}& 23.206 s\\
+			\text{SYN721+1.p}&\text{Proven}& 0.078 s\\
+			\text{SYN722+1.p}&\text{Proven}& 0.105 s\\
+			\text{SYN733+1.p}&\text{Proven}& 0.073 s\\
+			\text{SYN915+1.p}&\text{Proven}& 0.034 s\\
+			\text{SYN916+1.p}&\text{Refuted}& 0.030 s\\
+			\text{SYN924+1.p}&\text{Proven}& 0.059 s\\
+			\text{SYN926+1.p}&\text{Proven}& 0.039 s\\
+			\text{SYN927+1.p}&\text{Proven}& 0.050 s\\
+			\text{SYN928+1.p}&\text{Proven}& 0.034 s\\
+			\text{SYN931+1.p}&\text{Proven}& 0.053 s\\
+			\text{SYN932+1.p}&\text{Proven}& 0.045 s\\
+			\text{SYN943+1.p}&\text{Proven}& 1.681 s\\
+			\text{SYN944+1.p}&\text{Proven}& 0.115 s\\
+			\text{SYN948+1.p}&\text{Proven}& 0.053 s\\
+			\text{SYN949+1.p}&\text{Proven}& 0.078 s\\
+			\text{SYN952+1.p}&\text{Proven}& 0.051 s\\
+			\text{SYN958+1.p}&\text{Proven}& 0.044 s\\
+			\text{SYN959+1.p}&\text{Proven}& 0.040 s\\
+			\text{SYN960+1.p}&\text{Proven}& 0.050 s\\
+			\text{SYN961+1.p}&\text{Proven}& 0.054 s\\
+			\text{SYN963+1.p}&\text{Proven}& 0.056 s\\
+			\text{SYN964+1.p}&\text{Proven}& 0.043 s\\
+			\text{SYN973+1.p}&\text{Proven}& 0.034 s\\
+			\text{SYN974+1.p}&\text{Proven}& 0.045 s\\
+			\text{SYN976+1.p}&\text{Proven}& 0.140 s\\
+			\text{SYN979+1.p}&\text{Proven}& 0.074 s\\
+			\text{SYN980+1.p}&\text{Proven}& 0.127 s\\
+			\text{TOP022+1.p}&\text{Proven}& 0.153 s\\
+		\end{matrix}\]
+	\end{minipage}
 
@@
 	\bibliography{EIICL}
-
 \end{document}

### 2022-08-17 Implementation/benchmark.py

@@ -1,0 +1,63 @@
+import subprocess
+import tqdm
+import os
+
+VAMPIRE_PATH = "/home/lambdax/.vampire/vampire"
+TRANSLATOR = "./eiicl/target/debug/eiicl"
+
+INCLUDE_ROOT = "./ILTP-v1.1.2-firstorder/"
+PROBLEM_ROOT = "./ILTP-v1.1.2-firstorder/Problems/"
+TRANSLATION_ROOT = "./ILTP-v1.1.2-firstorder-translated/"
+RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results/"
+
+total_proven = 0
+total_disproven = 0
+total_unresolved = 0
+total_total = 0
+
+if not os.path.isdir(TRANSLATION_ROOT):
+    os.mkdir(TRANSLATION_ROOT)
+if not os.path.isdir(RESULT_ROOT):
+    os.mkdir(RESULT_ROOT)
+for directory in os.listdir(PROBLEM_ROOT):
+    proven = 0
+    disproven = 0
+    unresolved = 0
+    if not os.path.isdir(os.path.join(TRANSLATION_ROOT, directory)):
+        os.mkdir(os.path.join(TRANSLATION_ROOT, directory))
+    if not os.path.isdir(os.path.join(RESULT_ROOT, directory)):
+        os.mkdir(os.path.join(RESULT_ROOT, directory))
+    for translation in tqdm.tqdm(os.listdir(os.path.join(PROBLEM_ROOT, directory)), desc=f"{directory}"):
+        problem_path = os.path.join(os.path.join(
+            PROBLEM_ROOT, directory), translation)
+        translation_path = os.path.join(os.path.join(
+            TRANSLATION_ROOT, directory), translation)
+        with open(translation_path, "w") as output_file:
+            subprocess.run([TRANSLATOR, problem_path,
+                           INCLUDE_ROOT], stdout=output_file)
+        result_path = os.path.join(os.path.join(
+            RESULT_ROOT, directory), translation)
+        with open(result_path, "w") as output_file:
+            subprocess.run([VAMPIRE_PATH, translation_path,
+                           "--time_limit", "30"], stdout=output_file)
+        with open(result_path, "r") as result:
+            result_text = result.read()
+        with open(problem_path, "r") as result:
+            problem_text = result.read()    
+        if "Refutation found." in result_text:
+            proven += 1
+            assert "Status (intuit.) : Non-Theorem" not in problem_text
[…]

### 2022-08-17 Implementation/eiicl/src/encoding.rs

@@ -1,332 +1,20 @@
+#[path = "normalized_formula.rs"]
+mod normalized_formula;
+
 use std::collections::HashMap;
+use std::ops::{Add, AddAssign};
 use tptp::visitor::Visitor;
 use tptp::{common, fof, top, TPTPIterator};
 
-#[derive(Hash, Debug)]
-pub enum PlainTerm {
-    Variable {
-        name: String,
-    },
-    Function {
-        name: String,
-        arguments: Vec<PlainTerm>,
-    },
-}
-
-impl PlainTerm {
-    fn to_tptp(&self) -> fof::Term {
-        match self {
-            PlainTerm::Variable { name } => {
-                fof::Term::Variable(common::Variable(common::UpperWord(name)))
-            }
-            PlainTerm::Function { name, arguments } => fof::Term::Function(Box::new(
-                fof::FunctionTerm::Plain(fof::PlainTerm::Function(
-                    common::Functor(common::AtomicWord::SingleQuoted(common::SingleQuoted(name))),
-                    Box::new(fof::Arguments(
-                        arguments.iter().map(|x| x.to_tptp()).collect(),
-                    )),
-                )),
-            )),
-        }
-    }
-}
-
-#[derive(Hash, Debug)]
-pub struct AtomizedSubformula {
-    name: String,
-    arguments: Vec<PlainTerm>,
-}
-
-impl AtomizedSubformula {
-    fn to_tptp(&self) -> fof::AtomicFormula {
-        fof::AtomicFormula::Plain(fof::PlainAtomicFormula(fof::PlainTerm::Function(
-            common::Functor(common::AtomicWord::SingleQuoted(common::SingleQuoted(
-                &self.name,
-            ))),
-            Box::new(fof::Arguments(
-                self.arguments.iter().map(|x| x.to_tptp()).collect(),
-            )),
-        )))
-    }
-}
-
-#[derive(Hash, Debug)]
-pub enum NormalizedFormula<'a> {
-    LeftAnd {
-        left: Vec<AtomizedSubformula>,
-        right: AtomizedSubformula,
-    },
-    RightAnd {
-        left: AtomizedSubformula,
-        right: Vec<AtomizedSubformula>,
-    },
-    LeftOr {
-        left: Vec<AtomizedSubformula>,
-        right: AtomizedSubformula,
-    },
-    RightOr {
-        left: AtomizedSubformula,
-        right: Vec<AtomizedSubformula>,
-    },
-    DoubleImplication {
[…]

### 2022-08-17 Implementation/eiicl/src/main.rs

@@ -1,28 +1,19 @@
-mod encoding;
+mod translation;
 
-use std::fs::File;
-use std::io::Read;
-use std::path::Path;
-
-fn example(bytes: &[u8]) {
-    let encoding = encoding::Encoding::from_tptp(bytes);
-    for annotated in encoding.to_tptp() {
-        println!("{}", annotated);
-    }
-}
+use std::env;
 
 fn main() {
-    let path = Path::new("example.p");
-    let display = path.display();
-
-    let mut file = match File::open(&path) {
-        Err(why) => panic!("couldn't open {}: {}", display, why),
-        Ok(file) => file,
-    };
-
-    let mut s = String::new();
-    match file.read_to_string(&mut s) {
-        Err(why) => panic!("couldn't read {}: {}", display, why),
-        Ok(_) => example(s.as_bytes()),
+    let mut args = env::args();
+    let problem : translation::Problem;
+    match args.len() {
+        1 => {
+            panic!("Please provide input file.")
+        },
+        _ => {
+            problem = translation::Problem::new(args.nth(1).unwrap(), args.nth(0));
+        },
+    }
+    for annotated in problem.translate().to_tptp() {
+        println!("{}", annotated);
     }
 }

### 2022-08-17 Implementation/eiicl/src/normalized_formula.rs

@@ -1,0 +1,642 @@
+use tptp::{common, fof};
+
+#[derive(Hash, Debug, PartialEq, Clone)]
+pub enum PlainTerm {
+    Variable {
+        name: String,
+    },
+    Function {
+        name: String,
+        arguments: Vec<PlainTerm>,
+    },
+}
+
+impl PlainTerm {
+    fn to_tptp(&self) -> fof::Term {
+        match self {
+            PlainTerm::Variable { name } => {
+                fof::Term::Variable(common::Variable(common::UpperWord(name)))
+            }
+            PlainTerm::Function { name, arguments } => fof::Term::Function(Box::new(
+                fof::FunctionTerm::Plain(fof::PlainTerm::Function(
+                    common::Functor(common::AtomicWord::SingleQuoted(common::SingleQuoted(name))),
+                    Box::new(fof::Arguments(
+                        arguments.iter().map(|x| x.to_tptp()).collect(),
+                    )),
+                )),
+            )),
+        }
+    }
+}
+
+#[derive(Hash, Debug, Clone, PartialEq)]
+pub struct AtomizedSubformula {
+    pub name: String,
+    pub arguments: Vec<PlainTerm>,
+}
+
+impl AtomizedSubformula {
+    pub fn to_tptp(&self) -> fof::AtomicFormula {
+        if self.name == "false" {
+            fof::AtomicFormula::Defined(fof::DefinedAtomicFormula::Plain(fof::DefinedPlainFormula(fof::DefinedPlainTerm::Constant(
+                common::DefinedConstant(common::DefinedFunctor(common::AtomicDefinedWord(common::DollarWord(common::LowerWord("false")))))
+            ))))
+        } else if self.name == "true" {
+            fof::AtomicFormula::Defined(fof::DefinedAtomicFormula::Plain(fof::DefinedPlainFormula(fof::DefinedPlainTerm::Constant(
+                common::DefinedConstant(common::DefinedFunctor(common::AtomicDefinedWord(common::DollarWord(common::LowerWord("true")))))
+            ))))
+        } else {
+            fof::AtomicFormula::Plain(fof::PlainAtomicFormula(fof::PlainTerm::Function(
+                common::Functor(common::AtomicWord::SingleQuoted(common::SingleQuoted(
+                    &self.name,
+                ))),
+                Box::new(fof::Arguments(
[…]

### 2022-08-17 Implementation/eiicl/src/translation.rs

@@ -1,0 +1,111 @@
+#[path = "encoding.rs"]
+mod encoding;
+
+use tptp::visitor::Visitor;
+use tptp::{top, TPTPIterator};
+
+use std::fs::File;
+use std::io::Read;
+use std::path::Path;
+
+struct Dependencies {
+    paths: Vec<String>,
+    include_root: Option<String>,
+}
+impl Dependencies {
+    fn new(input_file: String, include_root: Option<String>) -> Dependencies {
+        match include_root {
+            Some(ref dir) => {
+                if !Path::new(&dir).exists() {
+                    panic!("Path does not exists: {}", dir);
+                }
+            }
+            _ => {}
+        }
+        let mut result = Dependencies {
+            paths: vec![input_file.clone()],
+            include_root: include_root,
+        };
+        if !Path::new(&input_file).exists() {
+            panic!("File does not exists: {}", input_file);
+        }
+        let path = Path::new(&input_file);
+        let display = path.display();
+        let mut file = match File::open(&path) {
+            Err(why) => panic!("couldn't open {}: {}", display, why),
+            Ok(file) => file,
+        };
+        let mut s = String::new();
+        match file.read_to_string(&mut s) {
+            Err(why) => panic!("couldn't read {}: {}", display, why),
+            Ok(_) => {}
+        }
+        let bytes = s.as_bytes();
+        let mut parser = TPTPIterator::<()>::new(bytes);
+        for input in &mut parser {
+            let input = input.expect(&format!("syntax error in {}", input_file));
+            result.visit_tptp_input(&input);
+        }
+        return result;
+    }
+}
+
+impl<'a> Visitor<'a> for Dependencies {
+    fn visit_include(&mut self, include: &top::Include<'a>) {
+        let path = match self.include_root {
+            Some(ref dir) => dir.clone(),
+            None => String::new(),
+        } + &include.file_name.0.0.to_string();
+        self.paths.push(path.clone());
+        let mut recursive_result =
+            Dependencies::new(path, self.include_root.clone());
+        self.paths.append(&mut recursive_result.paths)
+    }
+    
+
+    fn visit_tptp_input(&mut self, input: &top::TPTPInput<'a>) {
+        match input {
+            top::TPTPInput::Annotated(_) => {}
+            top::TPTPInput::Include(include) => self.visit_include(include),
+        }
+    }
+}
+
+pub struct Problem {
+    input: Vec<Vec<u8>>,
+    dependencies: Dependencies,
+}
+
+impl Problem {
[…]

### 2022-08-17 Implementation/individual_results.py

@@ -1,0 +1,38 @@
+import os
+
+RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results/"
+
+k = 0
+i = 0
+print("\\begin{minipage}{0.45\\textwidth}\n\t\\[\\begin{matrix}\n\t\t\\text{Problem}&\\text{Result}&\\text{Time}\\\\")
+if not os.path.isdir(RESULT_ROOT):
+    os.mkdir(RESULT_ROOT)
+for directory in os.listdir(RESULT_ROOT):
+    for result in os.listdir(os.path.join(RESULT_ROOT, directory)):
+        result_path = os.path.join(os.path.join(RESULT_ROOT, directory), result)
+        with open(result_path, "r") as result:
+            result_text = result.read()
+        if "Refutation found." in result_text:
+            print(f"\t\t\\text{{{result.name.split('/')[-1]}}}&\\text{{Proven}}&{result_text.split('Time elapsed:')[1].splitlines()[0]}\\\\")
+            i += 1
+            if (k < 2 and i == 32) or (k >= 1 and i == 48):
+                print("\t\\end{matrix}\\]\n\\end{minipage}")
+                if k % 2 == 1:
+                    print("\n\\pagebreak\n")
+                k = k + 1
+                i = 0
+                print("\\begin{minipage}{0.45\\textwidth}\n\t\\[\\begin{matrix}\n\t\t\\text{Problem}&\\text{Result}&\\text{Time}\\\\")
+        if "SZS status CounterSatisfiable" in result_text:
+            print(f"\t\t\\text{{{result.name.split('/')[-1]}}}&\\text{{Refuted}}&{result_text.split('Time elapsed:')[1].splitlines()[0]}\\\\")
+            i += 1
+            if (k < 2 and i == 32) or (k >= 1 and i == 48):
+                print("\t\\end{matrix}\\]\n\\end{minipage}")
+                if k % 2 == 1:
+                    print("\n\\pagebreak\n")
+                k = k + 1
+                i = 0
+                print("\\begin{minipage}{0.45\\textwidth}\n\t\\[\\begin{matrix}\n\t\t\\text{Problem}&\\text{Result}&\\text{time}\\\\")
+        
+
+if i!=38:
+    print("\t\\end{matrix}\\]\n\\end{minipage}")

### 2022-08-17 Implementation/results.py

@@ -1,0 +1,35 @@
+import os
+
+RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results/"
+
+total_proven = 0
+total_disproven = 0
+total_unresolved = 0
+total_total = 0
+
+if not os.path.isdir(RESULT_ROOT):
+    os.mkdir(RESULT_ROOT)
+for directory in os.listdir(RESULT_ROOT):
+    proven = 0
+    disproven = 0
+    unresolved = 0
+    for translation in os.listdir(os.path.join(RESULT_ROOT, directory)):
+        result_path = os.path.join(os.path.join(
+            RESULT_ROOT, directory), translation)
+        with open(result_path, "r") as result:
+            result_text = result.read()
+        if "Refutation found." in result_text:
+            proven += 1
+        elif "SZS status CounterSatisfiable" in result_text:
+            disproven += 1
+        else:
+            unresolved += 1
+    total_proven += proven
+    total_disproven += disproven
+    total_unresolved += unresolved
+    total = proven + disproven + unresolved
+    total_total += total
+    print(
+        f"Category {directory}, Proven/Disproven/Unresolved:\t{proven}/{disproven}/{unresolved}\t{100*proven/total:.1f}%/{100*disproven/total:.1f}%/{100*unresolved/total:.1f}%")
+print("Total Result:")
+print(f"Proven/Disproven/Unresolved:\t{total_proven}/{total_disproven}/{total_unresolved}\t{100*total_proven/total_total:.1f}%/{100*total_disproven/total_total:.1f}%/{100*total_unresolved/total_total:.1f}%")

### 2022-10-05 FoSSaCS/fossacs-EIICL.tex
Format change for FoSSaCS

--- 2022-07-16 CSL/lipics-v2021-EIICL.tex
@@
-\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
-%This is a template for producing LIPIcs articles.
-%See lipics-v2021-authors-guidelines.pdf for further information.
-%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
-%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
-%for section-numbered lemmas etc., use "numberwithinsect"
-%for enabling cleveref support, use "cleveref"
-%for enabling autoref support, use "autoref"
-%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
-%for enabling thm-restate support, use "thm-restate"
-%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
-%for producing a PDF according the PDF/A standard, add "pdfa"
-
-%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
-%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-
+\documentclass[runningheads]{llncs}
+\usepackage{amsmath, amsfonts}
 \usepackage{wrapfig}
@@
 \usetikzlibrary{arrows,shapes,quotes}
-
-\usepackage{todonotes}
-
-%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
-
-\bibliographystyle{plainurl}% the mandatory bibstyle
-
-\title{Embedding Intuitionistic into Classical Logic} %TODO Please add
-
-%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
-
-\author{Alexander {Pluska}}{Faculty of Mathematics, Universtät Wien, Austria}{e11941874@student.tuwien.ac.at}{https://orcid.org/my-orcid?orcid=0000-0002-7709-3335}{}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
-Please provide at least the name of the affiliation and the country.
-The full address is optional.
-Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.
-
-\author{Florian Zuleger}{Institute of Logic and Computation, Technische Universität Wien, Austria}{florian.zuleger@tuwien.ac.at}{https://orcid.org/0000-0003-1468-8398}{}
-
-\authorrunning{A. Pluska and F. Zuleger} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
-
-\Copyright{Alexander Pluska and Florian Zuleger} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
-
-\ccsdesc[100]{\textcolor{red}{Theory of Computation Logic Constructive Mathematics}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm
-
-\keywords{Intuitionistic Logic, Automated Reasoning, Propositional Logic, First-order Logic, QBF} %TODO mandatory; please add comma-separated list of keywords
-
-\category{} %optional, e.g. invited paper
-
-\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
-%\relatedversiondetails[linktext={opt.
-text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g.
-Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional
-
-%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
-%\supplementdetails[linktext={opt.
-text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g.
-Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional
-
-%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors.
-Please enter author specific funding statements as fifth argument of the \author macro.
-
-%\nolinenumbers %uncomment to disable line numbering
-
-%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\EventEditors{John Q. Open and Joan R. Access}
-\EventNoEds{2}
-\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
-\EventShortTitle{CVIT 2016}
-\EventAcronym{CVIT}
-\EventYear{2016}
-\EventDate{December 24--27, 2016}
-\EventLocation{Little Whinging, United Kingdom}
-\EventLogo{}
-\SeriesVolume{42}
-\ArticleNo{23}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.21 of 2022/01/12
+%
+
+%
+\usepackage[T1]{fontenc}
+% T1 fonts will be used to generate the final print and online PDFs,
+% so please use T1 fonts in your manuscript whenever possible.
+% Other font encondings may result in incorrect characters.
+%
+\usepackage{graphicx}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following two lines
+% to display URLs in blue roman font according to Springer's eBook style:
+%\usepackage{color}
+%\renewcommand\UrlFont{\color{blue}\rmfamily}
+%
 \begin{document}
-
+	%
+	\title{Embedding Intuitionistic into Classical Logic}
+	%
+	%\titlerunning{Abbreviated paper title}
+	% If the paper title is too long for the running head, you can set
+	% an abbreviated paper title here
+	%
+	\author{Alexander Pluska\inst{1}\orcidID{0000-0002-7709-3335} \and
+		Florian Zuleger\inst{2,3}\orcidID{0000-0003-1468-8398}}
+	%
+	\authorrunning{A. Pluska and F. Zuleger}
+	% First names are abbreviated in the running head.
+	% If there are more than two authors, 'et al.' is used.
+	%
+	\institute{TU Wien, Vienna, Austria}
+	%
 	\maketitle
-
-%TODO mandatory: add short abstract of the document
 \begin{abstract}
-The famous double negation translation~\cite{glivenko1929quelques, godel1933intuitionistischen} establishes an embedding from classical into intuitionistic logic.
+The famous double negation translation~\cite{glivenko1929quelques,godel1933intuitionistischen} establishes an embedding from classical into intuitionistic logic.
 Curiously the reverse direction has not been covered in literature.
@@
 While some (albeit more complex) normal forms also exist  for intuitionistic logic, crucially the duality between validity and satisfiability of the negation does not hold.
-Therefore most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient, tammet1996resolution} use the naive inverse method, i.e.
+Therefore most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient,tammet1996resolution} use the naive inverse method, i.e.
 directly search for a cut-free proof by applying the rules from some proof calculus inversely, enhanced by search strategies such as focussing and polarization.
@@
 This entailment has most famously been established for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}, but also other settings have been studied~\cite{schwichtenberg}.
-Another important direction is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}, which however does not permit in general the realizability of at least $\exists$ and $\vee$.
+Another important direction is to equip classical proofs with operational semantics~\cite{Control1,Parigot1}, which however does not permit in general the realizability of at least $\exists$ and $\vee$.
 
@@
 For every propositional formula $\varphi$ there effectively are an atom $P$ and a set of $\mathcal S$ of formulas (over an extended set of propositions) of one of the forms\\
-$\text{ \hspace{1cm}} A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$\\
+$\hspace{1cm} A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$\\
 for atomic $A, B, C$, such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
@@
 
-It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}.
+It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}.
 However, giving an explicit translation from IPC to QBF still promises to be useful, allowing to leverage the progress in QBF solving.
@@
 
-\bibliography{lipics-v2021-EIICL}
+\bibliographystyle{splncs04}
+\bibliography{bibliography}
 

### 2022-10-05 Notes/Embedding intuitionistic into classical logic/EIICL.tex
Defensio

@@
 \usepackage{setspace}
+\usepackage{caption}
 \usepackage{amsmath,amssymb,amsthm}
@@
 \usepackage{wrapfig}
-\usepackage[a4paper,top=2.5cm,bottom=3.5cm,left=2.5cm,right=2.5cm,headsep=10pt]{geometry}
+\usepackage[a4paper,top=3cm,bottom=3.8cm,left=3cm,right=3cm,headsep=20pt]{geometry}
 \usepackage[toc,page]{appendix}
@@
 \newtheorem{remark}[theorem]{Remark}
-
 \begin{document}
@@
 	\pagenumbering{gobble}
+	\newgeometry{top=2cm, right=2.5cm, left=2.5cm, bottom=2.5cm}
 	\begin{flushright}
@@
 
-		\Large\textbf{Alexander Pluska, B.Sc.} % Insert your name in this line
+		\Large\textbf{Alexander Pluska, BSc} % Insert your name in this line
 		\vskip15mm
@@
 
-		\Large\textbf{Master of Science (M.Sc.)}
+		\Large\textbf{Master of Science (MSc)}
 
@@
 	\begin{flushleft}
-		Wien, 2022 / Vienna 2022 % Insert graduation year in this line
+		Wien, 2022 / Vienna, 2022 % Insert graduation year in this line
 		\vfill
@@
 			\begin{tabular}{p{7cm} p{10cm}}
-				Studienkennzahl lt. Studienblatt / \newline degree programme code as it appears on \newline the student record sheet: &A 066 821 \\
-				Studienrichtung lt. Studienblatt / \newline degree programme as it appears on \newline the student record sheet:& Mathematics \\
-				Betreut von / Supervisor: & Univ.-Prof. Matthias Aschenbrenner \\ % Insert your program/supervisor here
-				Mitbetreut von / Co-Supervisor: & Ass.-Prof. Florian Zuleger \\ 
+				Studienkennzahl lt. Studienblatt / \newline degree programme code as it appears on \newline the student record sheet: &UA 066 821 \\
+				Studienrichtung lt. Studienblatt / \newline degree programme as it appears on \newline the student record sheet:& Masterstudium Mathematik\\
+				Betreut von / Supervisor: & Univ.-Prof. Matthias Aschenbrenner, PhD
 			\end{tabular}
@@
 	\endgroup	
+	\restoregeometry
 
@@
 		The famous double negation translation~\cite{glivenko1929quelques, godel1933intuitionistischen} establishes an embedding from classical into intuitionistic logic.
-Curiously the reverse direction has not been covered in literature.
+Curiously, the reverse direction has not been covered in literature.
 We present an effective embedding from intuitionistic into classical logic, both in the propositional and first-order case, as well as an effective embedding of intuitionistic propositional logic into quantified boolean formulas.
 
-		Furthermore we implement a practical system that takes intuitionistic first-order problems as an input and performs our transformation.
-This allows the use of classical theorem provers for establishing intuitionistic validity.
-We benchmark our implementation using the vampire theorem prover~\cite{kovacs2013first} on the ILTP problem set~\cite{iltp}.
-		%The key notion used in the proofs is the existence of counter-models - an analogon to the satisfiability of the negation in classical logic, which is not equivalent to invalidity in the intuitionistic case.
-We hope that this will allow leveraging the tremendous recent progress in automated reasoning in classical logic for checking intuitionistic validity, for which there has not been much progress in recent decades.
-
-		Finally we discuss how the generated classical proofs of the transformed problems can be translated to intuitionistic proofs of the original problem.
-All in all this establishes a novel approach to theorem proving for intuitionistic logic.
+		Furthermore, we implement a system that takes intuitionistic first-order problems in the tptp file format as an input and performs our transformation.
+This allows the use of classical theorem provers for checking intuitionistic validity.
+We benchmark our implementation using the Vampire theorem prover~\cite{kovacs2013first} on the ILTP problem set~\cite{iltp}.
+
+		Finally, we discuss how the generated classical proofs of the transformed problem can be translated back into intuitionistic proofs of the original problem.
+All in all, this establishes a novel approach to theorem proving for intuitionistic logic and provides a first proof of concept.
 
@@
 
-	I want to thank my Matthias for accommodating my particular interests by agreeing to supervise a thesis which lies outside of his main area of research and bearing with the organizational difficulties.
-I want to thank Florian who graciously agreed to co-supervise a thesis at a different university.
+	I want to thank Matthias for accommodating my particular interests by agreeing to supervise a thesis which lies outside his main area of research and bearing with the organizational difficulties.
+
+	I sincerely appreciate Florian's guidance, who graciously agreed to co-supervise a thesis at a different university.
 He has provided me with a fascinating topic as well as plenty of his time, expertise and thorough feedback.
-Finally I want to thank my family, in particular my parents Stefania and Cezary, who have instilled in me the joy of learning, provided all the love and help I could ask for and invariably supported me along my journey.
+
+	Finally, I am incredibly grateful to my family, in particular my parents Stefania and Cezary, who have instilled in me the joy of learning, provided all the love and help I could ask for and invariably supported me along my journey.
 
@@
 	\singlespacing
-	\newgeometry{top=1cm}
+	\newgeometry{top=1cm, bottom=2cm}
 	\tableofcontents
@@
 
-	The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmark problems (see e.g.
+	The interest in intuitionistic logic has led to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmark problems (see e.g.
 the
@@
 the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
-	%, e.g., the first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}
-	%Comparing benchmarks for classical and intuitionistic automated theorem proving, e.g.
-TPTP~\cite{casc} and , we see that classical provers currently have a far better success rate.
-Furthermore all the most popular first-order utilize classical logic.
 	This difference can be partially explained by fundamental differences between the logics.
-	First of all, determining intuitionistic validity is computationally harder, i.e.
+	Foremost, determining intuitionistic validity is computationally harder, i.e.
 in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
@@
 	While some (albeit more complex) normal forms also exist  for intuitionistic logic, crucially the duality between validity and satisfiability of the negation does not hold.
-	Therefore most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient, tammet1996resolution} use the naive inverse method, i.e.
-directly search for a cut-free proof by applying the rules from some proof calculus inversely, enhanced by search strategies such as focussing and polarization.
+	Therefore, most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient, tammet1996resolution} use the naive inverse method, i.e.
+direct search for a cut-free proof by applying the rules from some proof calculus inversely, enhanced by search strategies such as focussing and polarization.
 This approach generally leads to a much more complex search and is therefore difficult to implement efficiently.
@@
 		\item Use a state-of-the-art classical prover to establish the validity/invalidity of $\varphi^\#$.
-		\item Transform the output proof/counter-model of $\varphi^\#$ to one of $\varphi$.
+		\item Transform the generated proof/counter-model of $\varphi^\#$ to one of $\varphi$.
 	\end{itemize}
@@
 	Interestingly, the reverse direction, the famous double-negation translation, has long been established and goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
-In the propositional case it is particularly simple: $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
-Intuitively the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent.
+In the propositional case, it is particularly simple: $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
+Intuitively, the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent.
 This gives us a first idea why the reverse direction is perhaps more difficult: We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e.
@@
 In particular, this necessitates the introduction of new propositional variables in our procedure, which already marks a big difference to the double negation translation.
-	%
+
 	While establishing the translation of formulas we also perform an effective translation of counter-models, i.e.
@@
 	Transforming and reducing counter-models to a normal form is also what ultimately enables our translation.
-Apart from a translation of counter-models we also explicitly describe how to equivalently transform classical proofs of $\varphi^\#$ to intuitionistic proofs of $\varphi$, in particular resolution proofs as given by theorem provers like vampire~\cite{Kov_cs_2013}.
-
-	As a final contribution we implement our translation in the Rust programming language.
-It takes a first-order problem in the tptp format~\cite{tptp} and outputs the translated problem in the same format.
-The code is published on github~\cite{implementation}.
+Apart from a translation of counter-models in the final chapter, we also explicitly describe how to equivalently transform classical proofs of $\varphi^\#$ to intuitionistic proofs of $\varphi$.
+
+	As a final contribution, we implement our translation in the Rust programming language.
+Our implementation transforms a first-order problem in the tptp format~\cite{tptp} and outputs the translated problem in the same format.
+The code is published on GitHub~\cite{implementation}.
 We then benchmark our implementation using the ILTP problem set~\cite{iltp}, i.e.
-we translate all problems in the set and then run the vampire theorem prover~\cite{Kov_cs_2013} on the translated problems.
-Our approach performs on par with most existing approaches for intuitionistic theorem provers, but comes short of the state-of-the art leaders.
-As there is still a lot of room for optimization this is a hopeful sign.
+we translate all problems in the set and then run the Vampire theorem prover~\cite{Kov_cs_2013} on the translated problems.
+Our approach performs on par with existing approaches for intuitionistic theorem proving, but comes short of the state-of-the art.
+As there is still a lot of room for optimization, this is a hopeful first sign.
 
@@
 
-	The main content is organized in chapters 3-6.
+	The main content is organized in chapters 3 to 6.
 Chapter 3 collects prerequisite definitions and serves to fix our notation and recall some important facts.
-Since this overview already contains some technical results it may be consulted for reference.
 Chapter 4 is the meat of this thesis.
-In particular it features a complete description of our embedding of intuitionistic into classical logic as well as proving its correctness.
-In chapter 5 we present our implementation and benchmark it.
-Finally in chapter 6 we explicitly describe how to translate the generated proofs and counter-models which in principle directly follows from the results of section 4.
+In particular, it features a complete description of our embedding of intuitionistic into classical logic as well as proving its correctness.
+In chapter 5 we present the implementation of the embedding and the results of our benchmark.
+Finally, in chapter 6 we explicitly describe how to translate the generated proofs and counter-models which in principle directly follows from the results of section 4.
 	We shall now for each chapter give an overview of our main results and highlight the key arguments.
 
-	\section{Embedding Intuitionistic into Classical Logic}
+	\section{Embedding intuitionistic into classical logic}
 
@@
 	Eliminating all implications, however, is not possible in intuitionistic logic since $A\to B$ is not equivalent to $\neg A\vee B$.
-	Still we propose a similar transformation.
+	Still, we propose a similar transformation.
 	A notable feature of our transformation is that all non-classical content is encapsulated in formulas of type $(A\to B)\to C$, i.e.
-if there are no such formulas in the transformed formula then the classical validity of the transformed formula immediately implies intuitionistic validity.
-
-	\begin{theorem}\label{thm:Tseytin}
+if there are no such formulas in the transformed formula then the classical validity of the transformed formula immediately implies intuitionistic validity and no further processing is needed.
+
+	\begin{theorem}\label{thm:Tseytin1}
 		For every propositional formula $\varphi$ there effectively are an atom $P$ and a set of $\mathcal S$ of formulas (over an extended set of propositions) of one of the forms\\
@@
 	1) We encode as a first-order sentence that the considered formula holds for every Kripke frame.
-	Since Kripke frames for propositional logic over a fixed alphabet of propositional variables form a first-order theory this step is rather straightforward.
+	Since Kripke frames for propositional logic over a fixed alphabet of propositional variables form a first-order theory, this step is rather straightforward.
 	%
@@
 	%
-	3) We can then argue that if there is a counter-model for the resulting formula then there is a counter-model with a certain Kripke frame of bounded size completely determined by the number of formulas of type $(A\to B)\to C$.
+	3) We can then argue that if there is a counter-model for the resulting formula, then there is a counter-model with a certain Kripke frame of bounded size completely determined by the number of formulas of type $(A\to B)\to C$.
 	This allows us to eliminate the remaining quantifiers by simply enumerating the worlds in that Kripke frame.
@@
 		\label{thm:reduction-propositional}
-		Let $\mathcal S$ be as in Theorem~\ref{thm:Tseytin} and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
+		Let $\mathcal S$ be as in Theorem~\ref{thm:Tseytin1} and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
 For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
@@
 			\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-			\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
+			\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occur in $k$.
 		\end{itemize}
@@
 The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$.
-Futhermore, there is an effective procedure for translating counter-models between the sentences.
+Furthermore, there is an effective procedure for translating counter-models between the sentences.
 	\end{theorem}
@@
 	Instead of explicitly enumerating the worlds of a counter-example of bounded size (as in the reduction stated in the above theorem), we can also do this enumeration implicitly with a quantified boolean formula (QBF).
-	We can show that this QBF is satisfiable if and only if there exists a counter-model for our original formula:
+	We can show that this QBF is satisfiable iff there exists a counter-model for our original formula:
 
@@
 
-	We then move to the first-order case. Here the construction is a bit more involved but the underlying approach is the same.
-	Again we first perform a Tseytin-like transformation.
+	We then move to the first-order case. Here the construction is a bit more involved, but the underlying approach is the same.
+	Again, we first perform a Tseytin-like transformation.
 	Here the non-classical content is encapsulated by formulas of type $(A\to B)\to C$ and $\forall xA\to B$.
@@
 	We then proceed similarly to the propositional case.
-	Note however that one difficulty arises from the fact that our (classical) domain will now contain on the one hand worlds in the Kripke frame but on the other hand also proper domain elements.
+	Note however that one difficulty arises from the fact that our (classical) domain will now contain on the one hand worlds in the Kripke frame, but on the other hand also proper domain elements.
 	We resolve this apparent conflict by introducing a special binary predicate $E$, inspired by~\cite{iemhoff2010eskolemization}, encoding which domain elements exists at which world.
@@
 	While we can also establish the existence of a canonic counter-model whose frame only depends on formulas of type $(A\to B)\to C$ and $\forall xA\to B$, the size of this counter-model is countably infinite in general.
-	This is not surprising since certain intuitionistically invalid formulas like the double negation shift $\forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$ don't have finite counter-models.
-	Therefore we are not able to eliminate the $\forall$-quantifiers associated with the Kripke semantics.
-	However since our translation targets first-order logic, we believe that the introduction of these quantifiers is acceptable.
-	%
+	This is not surprising, since certain intuitionistically invalid formulas like the double negation shift $\forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$ don't have finite counter-models.
+	Therefore, we are not able to eliminate the $\forall$-quantifiers associated with the Kripke semantics.
+	However, since our translation targets first-order logic, we believe that the introduction of these quantifiers is acceptable.
+
 	Summarizing, we obtain the following result:
@@
 
-	In this section we describe our system that implements the translation from the previous section.
+	In this section, we describe our system that implements the translation from the previous section.
 The rust source code can be found online~\cite{implementation}.
-We then elaborate our benchmarking process which utilizes the ILTP problem set~\cite{iltp} and the vampire theorem prover~\cite{Kov_cs_2013}.
+We then elaborate our benchmarking process, which utilizes the ILTP problem set~\cite{iltp} and the Vampire theorem prover~\cite{Kov_cs_2013}.
 
@@
 		$$\begin{matrix}
-			&\text{AGT}&\text{ALG}&\text{COM}&\text{CSR}&\text{GEO}&\text{GRA}&\text{GRP}&\text{HAL}\\
+			&\text{AGT}&\text{ALG}&\text{COM}&\text{CSR}&\text{GEJ}&\text{GEO}&\text{GPJ}&\text{GRA}\\
 			\text{Proven}&5&170&0&0&58&3&4&2\\
-			\text{Disproven}&0&0&0&0&0&0&0&\\
+			\text{Disproven}&0&0&0&0&0&0&0&0\\
 			\text{Timeout}&47&29&3&29&107&74&1&16
@@
 		$$\begin{matrix}
-			&\text{KRS}&\text{LCL}&\text{MGT}&\text{MSC}&\text{NLP}&\text{NUM}&\text{PLA}&\text{PUZ}\\
+			&\text{GRP}&\text{HAL}&\text{KRS}&\text{LCL}&\text{MGT}&\text{MSC}&\text{NLP}&\text{NUM}\\
 			\text{Proven}&0&0&30&1&9&1&9&2\\
@@
 		$$\begin{matrix}
-			&\text{SET}&\text{SWC}&\text{SWV}&\text{SYN}&\text{TOP}&\text{GEJ}&\text{GPJ}&\text{SYJ}\\
+			&\text{PLA}&\text{PUZ}&\text{SET}&\text{SWC}&\text{SWV}&\text{SYJ}&\text{SYN}&\text{TOP}\\
 			\text{Proven}&0&5&27&1&64&65&54&1\\
-			\text{Disproven}&6&1&297&423&154&108&298&2
+			\text{Disproven}&0&1&0&0&2&3&13&0\\
+			\text{Timeout}&6&1&297&422&154&108&298&2
 		\end{matrix}$$
@@
 	Our implementation performs similar to existing dedicated provers for intuitionistic logic, but falls short of the state-of-the-art.
-Since there is still a lot of possible optimizations left for our translation this is a hopeful first sign.
-
-	\section{Translation of proofs and counter-models}
-
-	In the final section we give an explicit mechanism how to transform the proofs and counter-models for translated formula $\varphi^\#$ into intuitionistic proofs/counter-models of the original formula $\varphi$.
-The existence of these transformations is clear from the methodology of chapter~\ref{chapter:eiicl}, however they are hidden in the proofs and here we make them explicit.
+Since there is still a lot of possible optimizations left for our translation, this is a hopeful first sign.
+
+	\section{Translation of counter-models and proofs}
+
+	In the final section, we give an explicit mechanism how to transform the proofs and counter-models for translated formula $\varphi^\#$ into intuitionistic proofs/counter-models of the original formula $\varphi$.
+The existence of these transformations is clear from the methodology of chapter~\ref{chapter:eiicl}, however, they are hidden in the proofs and here we make them explicit.
 The two main results are as follows:
 	\begin{theorem}
-		There exists an effective linear time procedure that converts any classical counter-model for a translated formula $\varphi^\#$ into a intuitionistic counter-model for the original formula $\varphi$.
+		There exists an effective procedure that converts any classical counter-model for a translated formula $\varphi^\#$ into an intuitionistic counter-model for the original formula $\varphi$.
 	\end{theorem}
 	\begin{theorem}
-		There exists an effective linear time procedure that converts any classical proof of a translated formula $\varphi^\#$ into a intuitionistic proof for the original formula $\varphi$.
+		There exists an effective procedure that converts any classical proof of a translated formula $\varphi^\#$ into a intuitionistic proof of the original formula $\varphi$.
 	\end{theorem}
@@
 	Given the arguments of Chapter~\ref{chapter:eiicl} the translation of counter-models is more or less just a compilation of proofs.
-On the other hand the explicit translation of proofs does require some work, since to that point all arguments were semantic and this requires a syntactic viewpoint.
+On the other hand, the explicit translation of proofs does require some work, since to that point all arguments were semantic and this requires a syntactic viewpoint.
 
 	\chapter{Preliminaries}
-	First let us recapitulate the most important definitions and some facts about them. This will serve to fix our notation as well.
+	First, let us recapitulate the most important definitions and some facts about them. This will serve to fix our notation as well.
 
@@
 
-	The analog for predicate logic is much more involved as we need to define the term language.
+	The analogon for predicate logic is much more involved as we need to define terms.
 
@@
 
-	\noindent Usually the signature we be left implicit. If explicitly stated we denote it by $$\{f_1/\text{ar}(f_1),\dots,f_n/\text{ar}(f_n), R_1/\text{ar}(R_1).\dots,R_m/\text{ar}(R_m)\}.$$
+	\noindent Usually the signature will be left implicit. If explicitly stated we denote it by $$\{f_1/\text{ar}(f_1),\dots,f_n/\text{ar}(f_n), R_1/\text{ar}(R_1).\dots,R_m/\text{ar}(R_m)\}.$$
 	Fix a countably infinite collection of free variables $a, b, c\dots$ and bound variables $x, y, z\dots$.
 	Note that the following definitions could be simplified by not distinguishing between bound and free variables.
-However this has some other drawbacks and in particular when defining a sequent calculus for first-order logic having disjoint sets of bound and free variables is convenient.
+However, this has some other drawbacks and in particular when defining a sequent calculus for first-order logic having disjoint sets of bound and free variables is convenient.
 
 	\begin{definition}
-		A \textit{term} is defined inductively via the following rules
-		\begin{itemize}
-			\item each free variable is a term
-			\item if $t_1,\dots,t_n$ are terms and $f$ an $n$-ary function symbol then $f(t_1, \dots, t_n)$ is a term.
-		\end{itemize}
-		A \textit{semiterm} is defined in the same way with the difference that bound variables are allowed in 1.
-Clearly every term is a semiterm.
-Analogously to before we can define \textit{subterms} and \textit{subsemiterms}.
+		A \textit{semiterm} is defined inductively via the following rules
+		\begin{itemize}
+			\item each variable is a term
+			\item if $t_1,\dots,t_n$ are terms and $f$ a $n$-ary function symbol then $f(t_1, \dots, t_n)$ is a term.
+		\end{itemize}
+		A \textit{term} is a semiterm in which each occurring variable is a free variable.
+Every term is a semiterm.
+Analogously to the definition of subformulas we can define \textit{subsemiterms} and \textit{subterms}.
 	\end{definition}
@@
 			\end{itemize}
-			\item A formula with no free variables is a \textit{sentence}.
+			\item A formula containing no free variables is a \textit{sentence}.
 		\end{itemize}
@@
 
-	We now give meaning to formulas in classical and intuitionistic logic.
+	Classical and intuitionistic semantics give (different) meaning to formulas.
 
@@
 	\begin{definition}
-		A \textit{valuation} $v$ is a function that assigns to each propositional variable $A$ a \textit{true value} $v(A)\in\{0, 1\}$. We extend $v$ to a function $v^*$ on formulas:
-		\begin{itemize}
-			\item $v^*(\bot) = 0$
-			\item $v^*(A) = v(A)$ for each propositional variable $A$.
-			\item $v^*(\varphi\wedge\psi) = \min(v^*(\varphi), v^*(\psi))$
-			\item $v^*(\varphi\vee\psi) = \max(v^*(\varphi), v^*(\psi))$
-			\item $v^*(\varphi\to \psi) = \max(1 - v^*(\varphi), v^*(\psi))$
-		\end{itemize}
-		$v$ is a \textit{model} for $\varphi$ if $v(\varphi) = 1$, we write $v\models\varphi$.
+		A \textit{valuation} $v$ is a function that assigns to each propositional variable $A$ a \textit{truth value} $v(A)\in\{0, 1\}$.
+We define a model relation $\models$ between valuations $v$ and formulas:
+		\begin{itemize}
+			\item $v\not\models \bot$
+			\item $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
+			\item $v\models \varphi\wedge\psi$ iff $v\models\varphi$ and $v\models\psi$.
+			\item $v\models\varphi\vee\psi$ iff $v\models\varphi$ or $v\models\psi$.
+			\item $v\models\varphi\to \psi$ iff $v\not\models\varphi$ or $v\models\psi$.
+		\end{itemize}
+		$v$ is a \textit{model} for $\varphi$ if $v\models\varphi$.
 $\varphi$ is \textit{satisfiable} if there exists a valuation which is a model.
-It is \textit{valid} if every valuation is a model and we write $\models \varphi$.
+It is \textit{valid} if every valuation is a model, and we write $\models \varphi$.
 We denote the set of valid formulas with CPC.
@@
 
-	One notable property of classical logic is that a formula $\varphi$ is valid if and only if its negation, i.e.
+	One notable property of classical logic is that a formula $\varphi$ is valid iff its negation, i.e.
 $\neg\varphi$ is not satisfiable.
-The same does not hold true for intuitionistic logic as we shall see.
+The same does not hold true for intuitionistic logic, as we shall see.
 
@@
 		A \textit{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of nodes an a family of valuations $(v_w)_{w\in W}$ such that for $u\leq w$ we have $v_u(A)\leq v_u(A)$ for every propositional variable $A$, this is called the \textit{persistency condition}.
-We extend the valuations $v_u$ slightly differently to $v_u^\#$ than before, i.e.
+In this case, we define a model relation between worlds $u$ and formulas.
+It is slightly different in the case of implications, i.e.
 have
 		\begin{itemize}
-			\item $v_u^\#(\bot) = 0$
-			\item $v_u^\#(\varphi) = v_u(A)$ if $\varphi$ is atomic.
-			\item $v_u^\#(\varphi\wedge\psi) = \min(v_u^\#(\varphi), v_u^\#(\psi))$
-			\item $v_u^\#(\varphi\vee\psi) = \max(v_u^\#(\varphi), v_u^\#(\psi))$
-			\item $v_u^\#(\varphi\to \psi) = \min\{\max(1 - v_w^\#(\varphi), v_w^\#(\psi))\:|\:w\geq u\}$
-		\end{itemize}
-		That is $v_u^\#(\varphi\to \psi) = 1$ if at every world $w\geq u$ we have $v_w^\#(\varphi) = 0$ or $v_w^\#(\psi) = 1$.
-We say that $\varphi$ is satisfied at a node $u$ if $v_u^\#(\varphi) = 1$ and write $u\models\varphi$.
-If a formula is satisfied at every node $K$ is a \textit{model} for $\varphi$ and we write $K\models \varphi$.
-$\varphi$ is satisfiable if there exists a Kripke structure which is a model.
+			\item $u\not\models \bot$
+			\item $u\models A$ iff $v_u(A) = 1$ for each propositional variable $A$.
+			\item $u\models \varphi\wedge\psi$ iff $v\models\varphi$ and $v\models\psi$.
+			\item $u\models\varphi\vee\psi$ iff $v\models\varphi$ or $v\models\psi$.
+			\item $u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $w\models\psi$.
+		\end{itemize}
+		We say that $\varphi$ is satisfied at a node $u$ if $u\models\varphi$.
+If a formula is satisfied at every node, then $K$ is a \textit{model} for $\varphi$, and we write $K\models \varphi$.
+$\varphi$ is satisfiable if there exists a Kripke structure, which is a model.
 It is \textit{valid} if every Kripke structure is a model and we write $\models \varphi$.
 We denote the set of valid formulas with IPC.
-As before we define an entailment relation $\models_I$ between sets of formulas.
+As before, we define an entailment relation $\models_I$ between sets of formulas.
 	\end{definition}
@@
 	\begin{theorem}[Glivenko's theorem]
-		A propositional formula $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
+		A propositional formula $\varphi$ is classically valid iff $\neg\neg\varphi$ is intuitionistically valid.
 	\end{theorem}
@@
 		Every classical model can be viewed as a Kripke model with a single node.
-Therefore classical satisfiability trivially implies intuitionistic satisfiability.	On the other hand in every Kripke structure that is a model every node is a classical model so intuitionistic satisfiability implies intuitionistic satisfiability.
+Therefore, classical satisfiability trivially implies intuitionistic satisfiability.	On the other hand, in every Kripke structure that is a model every node is a classical model so intuitionistic satisfiability implies intuitionistic satisfiability.
 	\end{proof}
 
-	In classical logic a proven strategy for showing that a formula valid is to consider its negation and show that it is not satisfiable.
-Due to the previous lemma this is insufficient for intuitionistic logic.
-There are formulas like $A\vee\neg A$ which are not a intuitionistic tautology while their negation is not satisfiable either.
+	In classical logic, a proven strategy for showing that a formula valid is to consider its negation and show that it is not satisfiable.
+Due to the previous lemma, this is insufficient for intuitionistic logic.
+There are formulas like $A\vee\neg A$ which are not an intuitionistic tautology, while their negation is not satisfiable either.
 This serves to motivate the consideration of equivalidity in the following chapters.
-In the context of classical logic it can be fully explained via equisatisfiability and is therefore neglected but for intuitionistic logic it is more fundamental.
-
-	Finally let us note the complexities of these various decision problems, which play a fundamental role in all of theoretical computer science.
+In the context of classical logic it can be fully explained via equisatisfiability and is therefore neglected, but for intuitionistic logic it is more fundamental.
+
+	Finally, let us note the complexities of these various decision problems, which play a fundamental role in all of theoretical computer science.
 
 	\begin{theorem}
-		Satisfiability for classical and intuitionistic propositional logic is NP-complete.
+		Satisfiability for classical and intuitionistic propositional logic is \verb+NP+-complete.
 	\end{theorem}
 
-	For classical logic this is the famous Cook-levin theorem first published in Cook's seminal paper~\cite{cook1971complexity}.
-Of course we have seen that the intuitionistic case is the same decision problem.
+	For classical logic this is the famous Cook-Levin theorem first published in Cook's seminal paper~\cite{cook1971complexity}.
+Of course, we have seen that the intuitionistic case is the same decision problem.
 
 	\begin{corollary}
-		Validity for classical logic is co-NP-complete
+		Validity for classical logic is \verb+co-NP+-complete
 	\end{corollary}
 
-	This is clear since validity is just the complementary problem to satisfiability of the negated formula.
-However in intuitionistic logic unfortunately validity is really harder than satisfiability as shown by Statman~\cite{statman1979intuitionistic}.
+	This is clear, since validity is just the complementary problem to satisfiability of the negated formula.
+However, in intuitionistic logic unfortunately validity is really harder than satisfiability as first shown by Statman~\cite{statman1979intuitionistic}.
 
 	\begin{theorem}
-		Validity for intuitionistic logic in PSPACE-complete.
+		Validity for intuitionistic logic in \verb+PSPACE+-complete.
 	\end{theorem}
 
-	One thing to note is that in contrast to predicate logic all these problems are decidable.
+	One thing to note is that in contrast to predicate logic, all these problems are decidable.
 
@@
 		\begin{itemize}
-			\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
+			\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.
 			\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
@@
 			\item $M_u\subseteq M_w$.
-			\item $f^{I_w}|_{M_u} = f^{I_u}$.
+			\item $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$.}
 			\item $R^{I_w}|_{M_u} = R^{I_u}$.
@@
 		\end{itemize}
-		We say that $\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$.
-As always a formula is satisfiable if it is satisfies by some Kripke structure and it is valid if it is satisfied in every Kripke structure.
+		We say that $\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$ and write $\mathcal K\models\varphi$.
+As always, a formula is satisfiable if it is satisfied by some Kripke structure, and it is valid if it is satisfied in every Kripke structure.
 We denote the set of valid formulas with IQC.
-As before define an entailment relation $\models_I$ between sets of formulas.
+As before, define an entailment relation $\models_I$ between sets of formulas.
 	\end{definition}
-	In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically.
+	In addition to the propositional tautologies, there are now sentences involving quantifiers, which are classically valid but not intuitionistically.
 	Consider for instance the classical tautology $\varphi$ $$\neg\forall x A(x)\to \exists x \neg A(x)$$We exhibit an intuitionistic counter model.
 Consider the Kripke structure with two nodes $u\leq w$ with $M_u = \{0\}, M_w = \{0, 1\}$ and $A^{I_u} = A^{I_w} = \{0\}$.
-Then $u, v\not\models \varphi$: First note that $w, v\not\models\forall xA(x)$ and $u, v\not\models\forall xA(x)$ for any valuation $v$ since $w, v[1, a]\not\models A(a)$, i.e.
+Then $u, v\not\models \varphi$: First note that $w, v\not\models\forall xA(x)$ and $u, v\not\models\forall xA(x)$ for any valuation $v$ since $w, v[1/a]\not\models A(a)$, i.e.
 $u, v\models\neg\forall xA(x)$ and also $w, v\models\neg\forall xA(x)$.
-Furthermore $u, v[0/a]\not\models \neg A(a)$ and $w, v[0/a]\not\models \neg A(a)$ so $u, v\not\models \exists x\neg A(x)$ for any valuation $v$ (note however that this is not true at $w$!).
-Therefore $u, v\not\models\varphi$.
-
-	Intuitively the problem is that the existential quantifier at $u$ does not see the elements at $w$, i.e.
-it must "choose" between the currently available elements.
-As in the propositional case there are well known negative translation of classical logic into intuitionistic logic going back to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit}.
+Furthermore, $u, v[0/a]\not\models \neg A(a)$ and $w, v[0/a]\not\models \neg A(a)$ so $u, v\not\models \exists x\neg A(x)$ for any valuation $v$ (note however that this is not true at $w$!).
+Therefore, $u, v\not\models\varphi$.
+
+	Intuitively, the problem is that the existential quantifier at $u$ does not ``see'' the elements at $w$, i.e.
+it must "choose" between the elements at $u$.
+As in the propositional case, there are well-known negative translations of classical logic into intuitionistic logic going back to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit}.
 
@@
 
-	But again this is not reflected in satisfiability, i.e. with an completely analogous argument to~\ref{propsat} we obtain
+	But again this is not reflected in satisfiability, i.e. with a completely analogous argument to~\ref{propsat} we obtain
 
@@
 	An important step in the embedding will be the elimination of quantifiers via Herbrandization.
-	In this process we introduce fresh variables and add additional function symbols to the signature.
+	In this process, we introduce fresh variables and add additional function symbols to the signature.
 	A fresh variable is any variable that does not occur in any of the considered formulas.
-	Whenever we add a function symbol we implicitly extend the signature by some not previously contained symbol.
+	Whenever we add a function symbol, we implicitly extend the signature by some not previously contained symbol.
 
@@
 
-		Otherwise there are $3$ cases:
+		Otherwise, there are $3$ cases:
 
@@
 For every symbol $p$ that occurs in $\psi^S_Z$ or $\psi^H_Z$ but not $\varphi^S_Z, \varphi^H_Z$, we set $p^{I_{\chi, v}} = p^{I_{\psi, v}}$.
-Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.
+Otherwise, $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.
 
@@
 $\mathcal M_\chi\models\chi^S_Z$.
-On the other hand suppose $\mathcal M, v\not\models\chi$.
+On the other hand, suppose $\mathcal M, v\not\models\chi$.
 Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_Z$ and $\mathcal M_\chi, v\not\models\psi^S_Z$, i.e.
@@
 $\mathcal M_\chi,v\models \chi^S_Z$.
-On the other hand suppose $\mathcal M, v\not\models\chi$.
+On the other hand, suppose $\mathcal M, v\not\models\chi$.
 Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{Z\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(z_1,\dots z_n)/x]^H_Z$, i.e.
@@
 	\begin{proof}
-		Again we proceed by simultaneous induction on the formula height.
+		Again, we proceed by simultaneous induction on the formula height.
 
@@
 $\mathcal M, v\not\models\varphi^H_Z$ or $\mathcal M, v\models\psi^S_Z$.
-By induction hypothesis $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e.
+By induction hypothesis, $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e.
 $\mathcal M, v\models \chi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
+On the other hand, suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
 $\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_Z$.
-Again by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.
+Again, by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.
 
@@
 Suppose $\mathcal M, v\models\chi^S_Z$, i.e.
-for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{Z\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$.
-Then it follows that $\mathcal M, v\models\varphi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
+for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{Z\cup\{a\}}$ and, by induction hypothesis, $\mathcal M, v[m/a]\models \chi[a/x]$.
+Then it follows, that $\mathcal M, v\models\varphi$.
+On the other hand, suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
 there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{Z\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e.
@@
 
-	Logical truth can also be approached via syntax rather than semantics. This gives rise to proof theory which will be especially important in the final chapter when we establish proof translations.
+	Logical truth can also be approached via syntax rather than semantics. This gives rise to proof theory, which will be especially important in the final chapter when we establish proof translations.
 
@@
 		Pairs of multisets of formulas $A = \{A_1,\dots, A_n\}, B = \{B_1,\dots, B_m\}$ form a \textit{Sequent} $A\Rightarrow B$.
-Usually we will write $A_1,\dots, A_n\Rightarrow B_1,\dots, B_m$, note however that $A$ and $B$ are \textbf{not} to be interpreted as sequences but as multisets.
+Usually, we will write $A_1,\dots, A_n\Rightarrow B_1,\dots, B_m$, note however that $A$ and $B$ are \textbf{not} to be interpreted as sequences but as multisets.
 	\end{definition}
@@
 
-	The above rules make up the calculus for classical logic called \textbf{Gc}. It corresponds to \textbf{G3c} from~\cite{basicprooftheory}.
-	Similarly we can define a calculus \textbf{Gi} for intuitionistic logic corresponding to \textbf{m-G3i}:\\
+	The above inferences make up the calculus for classical logic called \textbf{Gc}. It corresponds to \textbf{G3c$^=$} from~\cite{basicprooftheory}.
+	Similarly, we can define a calculus \textbf{Gi} for intuitionistic logic corresponding to \textbf{m-G3i$^=$}:\\
 
@@
 
-	The differences are found in L$\to$, R$\to$ and R$\forall$. Note how in particular the rules for R$\to$ and R$\forall$ allows only a single formula in the succedent.
-
-	\pagebreak
+	The differences are found in L$\to$, R$\to$ and R$\forall$. Note how in particular the rules for R$\to$ and R$\forall$ allow only a single formula in the succedent.
 
@@
 	\begin{theorem}[Completeness theorem]
-		A sequent $\Phi\Rightarrow \varphi$ is derivable iff $\Phi\models \varphi$.
+		A sequent $\Phi\Rightarrow \varphi$ is derivable if and only if $\Phi\models \varphi$.
 	\end{theorem}
 
+	A further important property of the above calculi is the subformula property.
+While it is typically used to prove soundness of propositional logic, it is also a useful tool for proving many other results.
+
+	\begin{theorem}[Subformula Property]
+		Let $\varphi$ be a formula that occurs in a derivation of $\Delta\Rightarrow\Gamma$. Then $\varphi$ is a subformula of some formula in $\Delta\cup\Gamma$.
+	\end{theorem}
+
+	A direct consequence of this result is the soundness of propositional logic:
+
+	\begin{theorem}[Soundness of propositional logic]\hfill\\
+		The empty sequent is not derivable in \textbf{Gc} and \textbf{Gi}.
+	\end{theorem}
+
 	\begin{example}
-		The following is a \textbf{Gc} proof of the law of excluded middle.\\
+		The following is a \textbf{Gc} derivation of the law of excluded middle.\\
 
@@
 
-		Note how this derivation is invalid in \textbf{Gi} since R$\to$ is not applied correctly. However it is possible to prove the double negation of it in \textbf{Gi}:\\
+		Note how this derivation is invalid in \textbf{Gi} since R$\to$ is not applied correctly. However it is possible to prove its double negation in \textbf{Gi}:
 
@@
 		\end{center}
-
 	\end{example}
@@
 The process will be divided into three steps.
-The first is a first-order encoding of intuitionistic semantics, which gives an obvious but unsatisfactory embedding from propositional intuitionistic logic.
+The first is a first-order encoding of intuitionistic semantics, which gives an obvious but unsatisfactory embedding of propositional intuitionistic logic into classical logic.
 We then discuss how some of the quantifiers can be eliminated via Herbrandization.
-As a third part we give a certain normal form that then finally allows us to realize that all quantifiers can be indeed eliminated from the translated formula and we end up with a propositional formula.
-As a quick note we also give an embedding form IPC to QBF.
-
-	For predicate formulas the process is a bit more involved in that we also need to respect the constructive content of quantifiers.
-For this we use orderization as proposed in~\cite{iemhoff2010eskolemization}.
+As a third part, we give a certain normal form that then finally allows us to realize that all quantifiers can be indeed eliminated from the translated formula, and we end up with a propositional formula.
+As a quick aside, we also give an embedding form IPC to QBF.
+
+	For predicate formulas, the process is a bit more involved in that we also need to respect the constructive content of quantifiers.
+For this, we use a special existence predicate $E$ inspired by the work in~\cite{iemhoff2010eskolemization}.
 This allows us to utilize a quite similar process as in the first section, even though it is a bit more involved.
@@
 
-	The idea is to encode the Kripke semantics of intuitionistic logic classically and is quite close to what was done in~\cite{iemhoff2010eskolemization}.
-We first present an embedding from IPC to CQC which is conceptually easier to understand, but will then reduce it to CPC.
-To that end we will first propose a Tseytin-like translation, that reduces formulas to a normal form while preserving validity:
-
-	\begin{theorem}\label{thm:Tseytin}
-		For every propositional formula $\varphi$ there effectively are an atom $P$ and a set of $\mathcal S$ of formulas (over an extended set of propositions) of one of the forms\\
-		$\text{ \hspace{1cm}} A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$\\
-		for atomic $A, B, C$, such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
-The time complexity of the transformation is linear in the input size.
-	\end{theorem}
-
-	\begin{theorem}
-		Let $\mathcal S$ be a set of formulas of the form
-		$$(A\wedge B)\to C, (A\vee B)\to C, A\to (B\wedge C), A\to(B\vee C), (A\to B)\to C$$
-		and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
-For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
-Obtain $\mathcal S^\#$ by including the following formulas:
-		\begin{itemize}
-			\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
-			\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-			\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-			\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
-		\end{itemize}
-		Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
-The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$.
-Futhermore, there is an effective procedure for translating counter-models between the sentences.
-	\end{theorem}
+	As mentioned before, we proceed by first giving an embedding into classical predicate logic, transforming that embedding and finally arguing that all the quantifiers can be eliminated, by showing that if a counter-model exists than there exists a bounded one.
 
@@
 	The most obvious approach to embedding intuitionistic into classical logic is to express intuitionistic semantics, i.e.
-Kripke frames, as a first order theory.
+Kripke frames, as a first-order theory.
 For every propositional variable $A$ consider a unary predicate $A$ of the same name where $A(u)$ expresses that $A$ is true at some world $u$.
@@
 
-	\noindent As all we have done is modelling Kripke frames as a first-order theory we directly obtain:
+	\noindent As all we have done is modelling Kripke frames as a first-order theory, we directly obtain:
 
 	\begin{lemma}
-		$\varphi$ is intuitionistically valid iff $\varphi^C$ is classically valid.
+		$\varphi$ is intuitionistically valid if and only if $\varphi^C$ is classically valid.
 	\end{lemma}
@@
 	\begin{example}
-		Consider the law of excluded middle $$\varphi = (A\to \bot)\vee A$$ which is not intuitionistically valid. Then
+		Consider the law of excluded middle $$\varphi = (A\to \bot)\vee A,$$ which is not intuitionistically valid. Then
 		\begin{align*}
@@
 		\end{align*}
-		and indeed $\varphi^{C}$ admits a counter model with domain $\{u, u'\}$ where $u \preceq u'$, $A(u) = 0, A(u') = 1$.
+		and indeed $\varphi^{C}$ admits a counter model with domain $\{u, u'\}$, where $u \preceq u'$, $A(u) = 0, A(u') = 1$.
 	\end{example}
@@
 	\begin{definition}
-		Let $\mathcal K = (W, (v_w)_{v\in W})$ be a propositional Kripke structure. Define the $\mathcal M(\mathcal K) = (M, I)$ as follows:
+		Let $\mathcal K = (W, (v_w)_{v\in W})$ be a propositional Kripke structure. Define $\mathcal M(\mathcal K) = (M, I)$ as follows:
 		\begin{itemize}
@@
 
-	\noindent By definition we then have
+	\noindent By definition, we then have
 	\begin{lemma}
@@
 
-	\indent Similarly for every $\Sigma$-structure $\mathcal M$ with $\mathcal M\models K(\varphi)$ we could construct a Kripke Frame $\mathcal K(\mathcal M)$ such that $\mathcal M\models \varphi^C$ iff $\mathcal K(\mathcal M)\models \varphi$.
+	\indent Similarly, for every $\Sigma$-structure $\mathcal M$ with $\mathcal M\models K(\varphi)$ we could construct a Kripke Frame $\mathcal K(\mathcal M)$ such that $\mathcal M\models \varphi^C$ iff $\mathcal K(\mathcal M)\models \varphi$.
 
@@
 
-	Of course transforming a propositional formula into a predicate formula is somewhat unsatisfying.
-As a first step towards obtaining a propositional formula we want to apply Herbrandization.
-Now trying to apply it to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated.
+	Of course, transforming a propositional formula into a predicate formula is somewhat unsatisfying.
+As a first step towards obtaining a propositional formula, we want to apply Herbrandization.
+Now, trying to apply it to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated.
 We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form.
-In fact our translation is often presented as the first step in the Tseytin transformation, after which the individual formulas are then converted to CNF.
+In fact, our translation is often presented as the first step in the Tseytin transformation, after which the individual formulas are then converted to CNF.
 	Hence, the propositional case of our translation is well known, but we are not aware of any work that uses just this step as a stand-alone.
@@
 		\end{itemize}
-		%We further set $\mathcal S(\varphi) = \mathcal S^+(\varphi)\cup\mathcal S^-(\varphi)$.
 	\end{definition}
+
+	The following lemma can be directly obtained by induction on the formula height:
 
@@
 		For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms of the form $P_\psi$.
-By slight abuse of notation we denote with $\vec z_\psi$ elements of the domain instead of variables.
+By slight abuse of notation, we denote with $\vec z_\psi$ elements of the domain instead of variables.
 In the classical case, we set
-		$$\begin{matrix}
-			P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)\\
-			P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))\\P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})\\P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}\indent P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\},
-		\end{matrix}$$
+		\begin{align*}
+			P_A^I(\vec z_A)&:\Leftrightarrow A^I(\vec z_A)\\
+			P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) &:\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\\ P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) &:\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))\\P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) &:\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})\\P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) &:\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}\\ P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) &:\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\},
+		\end{align*}
 		and for intuitionistic logic and each world $u$, we set
-		$$\begin{matrix}
-			P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)\\
-			P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))\\
-			P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}\\
-			P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}\indent P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}.
-		\end{matrix}$$
+		\begin{align*}
+			P_A^{I_u}(\vec z_A)&:\Leftrightarrow A^{I_u}(\vec z_A)\\
+			P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) &:\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\\ P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) &:\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))\\
+			P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) &:\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}\\
+			P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) &:\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}\\ P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) &:\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}.
+		\end{align*}
 	\end{definition}
-	From this definition we directly obtain the following Lemmata by induction on the formula height.
+	From this definition, we directly obtain the following Lemmata by induction on the formula height.
 	\begin{lemma}\label{thm:struct1}
-		For every formula $\varphi$ and structure $\mathcal M$, we have that $\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$.
+		For every formula $\varphi$, structure $\mathcal M$, we have $\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$.
 	\end{lemma}
@@
 	\begin{lemma}\label{thm:struct2}
-		For every structure $\mathcal M$,
 		$\mathcal M\models \varphi$ if and only if $S(\mathcal M, \varphi)\models P_\varphi(\vec z_\varphi)$.
@@
 	\begin{corollary}\label{equivalid}
-		In both intuitionistic and classical logic
+		In both intuitionistic and classical logic,
 		\begin{itemize}
@@
 
-	Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
+	Recall that since $\varphi$ is quantifier-free, each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
 	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$$
@@
 	\end{matrix}$$
-	with each $P(x)$ possibly being $\bot$. Applying Herbrandization we obtain a formula
+	with each $P(x)$ possibly being $\bot$. Applying Herbrandization, we obtain a formula
 	$$\varphi''' = K(\varphi')\to \bigwedge \mathcal S'\to P_\varphi(s),$$
@@
 	$f_\psi^{I'}(u) = f^I_\psi(u) \text{ iff $u\preceq f^{I}_\psi(u)$, and $f_\psi^{I'}(u) = u,$ else.}$ That is, where $s$ is the least element and the $f_\psi$ increase their argument.
-Therefore instead of $\varphi'''$ we may consider a formula $$\varphi^\circ = \forall k(s\preceq k)\to \bigwedge\{\forall k(k\preceq f_\psi(k))\:|\:\psi\in\mathcal F_\to\}\to K(\varphi')\to \bigwedge \mathcal S^\circ\to P_\varphi(s)$$ where each formula $\psi\in\mathcal S^\circ$ is of one of the forms
+Therefore, instead of $\varphi'''$ we may consider a formula $$\varphi^\circ = \forall k(s\preceq k)\to \bigwedge\{\forall k(k\preceq f_\psi(k))\:|\:\psi\in\mathcal F_\to\}\to K(\varphi')\to \bigwedge \mathcal S^\circ\to P_\varphi(s)$$ where each formula $\psi\in\mathcal S^\circ$ is of one of the forms
 	$$\begin{matrix}
@@
 
-	We now argue that if a counter-model exists for $\varphi^\circ$ then a bounded counter-model exists, allowing us to return to the propositional case.
+	We now argue that if a counter-model exists for $\varphi^\circ$, then a bounded counter-model exists, allowing us to return to the propositional case.
 
@@
 
-	Essentially what this tells us is that for every $u$ at which $\psi$ is fulfilled, in particular for each $u$ in the image of $f_\psi$, we may add the assumption that $f_\psi(w) = w$ for all $w\geq u$ without changing the validity of $\psi$.
-	With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.
-
-	\begin{corollary}
-		If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
-	\end{corollary}
-
-	\begin{example}
-		Consider $$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model, depicted below,
+	\begin{wrapfigure}[7]{r}{0.5\textwidth}
+		\vspace*{-.6cm}
 		\begin{center}
@@
 		\end{center}
-		\vspace*{-.3cm}where at each node the true propositions are indicated.
-		%	
-		There is a corresponding classical counter-model $\mathcal M = (M, I)$ to $\varphi^\circ$ with $M = \{u, u_l, u_r\}$, $A^{I}(l) = D^I(l) = B^I(r) = C^I(r) = 1$ and $0$ else.
-		%
-		Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (B\to A)\to D$, this corresponds to a transformed counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below:
+		\captionsetup{justification=centering}
+		\caption{ Counter-model $\bigwedge S\to E$.\\(Example 4.1.16.)}
+		\label{fi2}
+	\end{wrapfigure}
+
+	Essentially, what this tells us is that for every $u$ at which $\psi$ is fulfilled, in particular for each $u$ in the image of $f_\psi$, we may add the assumption that $f_\psi(w) = w$ for all $w\geq u$ without changing the validity of $\psi$.
+	With this lemma in hand, showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.
+
+	\begin{corollary}
+		If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
+	\end{corollary}
+
+	\addtocounter{theorem}{1}
+	\noindent\textbf{Example \thetheorem.} Consider $$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$
+
+	\begin{wrapfigure}[12]{r}{0.5\textwidth}
+		\vspace*{-.6cm}
 		\begin{center}
@@
 		\end{center}
-	\end{example}
+		\caption{$\mathcal M_T$}
+		\label{fi3}
+	\end{wrapfigure}
+
+	and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model depicted in~\ref{fi2}
+	where at each node the true propositions are indicated.
+
+		There is a corresponding classical counter-model $\mathcal M = (M, I)$ to $\varphi^\circ$ with $M = \{u, u_l, u_r\}$, $A^{I}(l) = D^I(l) = B^I(r) = C^I(r) = 1$ and $0$ else.
+		%
+		Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (B\to A)\to D$, this corresponds to a transformed counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ and interpretation as presented in~\ref{fi3}.\\
 
@@
 
-	So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
+	So we know that $\varphi^\circ$ is valid if and only if it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
 		x, &\text{ if $\psi$ occurs in $x$,}\\
@@
 		and we know that $\varphi$ is valid iff $\bigwedge \mathcal S^-(\varphi)\to P_\varphi$ is valid.
-Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically valid, where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
+Applying the above it is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically valid, where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
 		\begin{align*}
@@
 	The translation from the last section will serve as starting point for developing the translation in this section.
-	For lack of space, we do not define the semantics of QBF in this paper and refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
+	We do not define the semantics of QBF and refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
 
@@
 		\begin{itemize}
-			\item {\normalfont Valid}$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
+			\item {\normalfont Valid}$(n)$, which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
 if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
-			\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
-			\item {\normalfont Sat}$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
-			\item {\normalfont Sat}$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
+			\item {\normalfont Persistent}$(n)$, which encodes that the persistency condition holds.
+			\item {\normalfont Sat}$_{\mathcal S}(n)$, encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
+			\item {\normalfont Sat}$_{\mathcal F_\to}(n)$, encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
 		\end{itemize}
@@
 		Suppose $\varphi$ is not intuitionistically valid.
-		Then, $\bigwedge\mathcal S\to P$ has an intuitionistic counter-model. Then by the previous section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
+		Then, $\bigwedge\mathcal S\to P$ has an intuitionistic counter-model. Then, by the previous section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
 		We now define a QBF model iteratively.
@@
 
-		On the other hand suppose $\varphi^Q$ is satisfiable. We construct a counter-example to $\varphi^\#$.
-		Again we proceed iteratively.
+		On the other hand, suppose $\varphi^Q$ is satisfiable. We construct a counter-example to $\varphi^\#$.
+		Again, we proceed iteratively.
 Interpret $A^\epsilon$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$.
@@
 	\begin{corollary}
-		Let $N$-\textbf{IntInvalid} be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid.
-Then $N$-\textbf{IntInvalid} is in $\Sigma_{2N-1}^P$.
-The dual problem $N$-\textbf{IntValid} is in $\Pi_{2N-1}^P$.
+		Let $N$-\verb+IntInvalid+ be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid.
+Then $N$-\verb+IntInvalid+ is in $\Sigma_{2N-1}^P$.
+The dual problem $N$-\verb+IntValid+ is in $\Pi_{2N-1}^P$.
 	\end{corollary}
@@
 Our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic.
-However we do it nonetheless!
+However, we do it nonetheless!
 The domain of our classical model will feature on the one hand elements representing worlds in a Kripke frame and on the other hand the domain-elements from the Kripke model.
@@
 We fix some signature $\Sigma$ for this section.
-To encode that some $n$-ary relation $A$ holds at a world $u$ we replace each $n$-ary relation symbol $A$ with a $n+1$-ary relation symbol $A^\#$, interpreting $A^\#(\vec x, u)$ as "$A(\vec x)$ holds at u".
+To encode that some $n$-ary relation $A$ holds at a world $u$ we replace each $n$-ary relation symbol $A$ with a $n+1$-ary relation symbol $A^\#$, interpreting $A^\#(\vec x, u)$ as ``$A(\vec x)$ holds at $u$''.
 Furthermore we write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$.
@@
 	\begin{itemize}
-		\item PartialOrder($\preceq$) encoding that $\preceq$ is a partial order.
-		\item DomainSubset$(u, w)$ encoding that the domain at $u$ is a subset of that at $w$.
-		\item World($u$) encoding that $u$ represents a world.
-		\item DomainClosed($u$) which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
-		\item Persistent$(u, w)$ which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$ at $u$ we have $A(x)$ at $w$.
+		\item PartialOrder($\preceq$), encoding that $\preceq$ is a partial order.
+		\item DomainSubset$(u, w)$, encoding that the domain at $u$ is a subset of that at $w$.
+		\item World($u$), encoding that $u$ represents a world.
+		\item DomainClosed($u$), which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
+		\item Persistent$(u, w)$, which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$, at $u$ we have $A(x)$ at $w$.
 	\end{itemize}
@@
 
-	Using $E$ we then obtain $\mathcal S'$ analogously to the predicate case by including formulas:
+	Using $E$, we then obtain $\mathcal S'$ analogously to the predicate case by including formulas:
 
@@
 Suppose we have a Kripke counter-model to $\varphi$ with frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
-In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
+In particular, since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
 Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
@@
 
-	As in the predicate case we now show the existence of a canonical counter-model (in case such a counter-model exists).
+	As in the predicate case, we now show the existence of a canonical counter-model (in case such a counter-model exists).
 Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$.
@@
 	\end{theorem}
-	The complete translation is summarised in Theorem~\ref{fullFOtranslation} (stated earlier). Let us understand why a further reduction as in the propositional case is impossible.
+	The complete translation is summarised in Theorem~\ref{fullFOtranslation} (stated earlier).
+Let us understand why a further reduction as in the propositional case is impossible.
+In this case this is not such an issue, because as we started in predicate logic it is not so troublesome that we introduce some additional predicates, however it is still interesting to understand why it is required.
 
@@
 
+	However part 2 of Lemma~\ref{thm:fo-countermodel-reduction} fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$), we have $A(x, v)$, there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
+This is why we must consider sequences that have repetitions:
+	\pagebreak
 	\begin{wrapfigure}[16]{r}{0.4\textwidth}
@@
 	\end{wrapfigure}	
-	However part 2 of Lemma~\ref{thm:fo-countermodel-reduction} fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$), we have $A(x, v)$, there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
-This is why we must consider sequences that have repetitions:
-
+	\
 	\begin{example}
@@
 
-	Let us add that Lemma~\ref{thm:fo-countermodel-reduction} still equips us with additional information that could be used to guide proof search: We could add additional axioms that restrict the interpretation of functions.
+	Let us add that Lemma~\ref{thm:fo-countermodel-reduction} still equips us with additional information that can be used to guide proof search: We could add additional axioms that restrict the interpretation of functions.
 That is, we may assume that $f_\psi(w) = w$, if $\exists u(w = f_\psi(u))$ for all $\psi\in\mathcal F_\forall$ (i.e.
 that $f_\psi$ is idempotent), and $f_\psi(w) = w$, if $\exists u(w \geq f_\psi(u))$ for all $\psi\in\mathcal F_\to$ (which is even stronger than idempotence).
+And indeed in our implementation we will add the first kind of additional axiom, which can be stated without the reference to the already eliminated $\geq$.
 
@@
 
-	In this chapter we shall give an overview over our implementation of the described translation in Rust. Rust was a good fit for this endeavour because
-	\begin{itemize}
-		\item There is a great package for parsing files in tptp format~\cite{rust-tptp}.
+	In this chapter we shall give an overview over our implementation of translation defined in the previous chapter.
+We have chosen to write our system in Rust.
+Rust was a good fit for this endeavour because:
+	\begin{itemize}
+		\item There is a great package for parsing files in tptp format~\cite{rust-tptp} written by Michael Rawson that was tremendously helpful.
 		\item Static typing reduced bugs induced by the many syntactically similar expressions that require different translation.
@@
 
-	Currently the only available installation option is building the binary from source.
+	Currently, the only available installation option is building the binary from source.
 The cargo rust complier~\cite{rust} is required for this purpose.
@@
 		\item The location of the tptp file that is to be translated.
-		\item (Optionally) a location with additional files to be included, per default the current folder.
+		\item (Optionally) a file directory that contains all additional files which are to be included, per default the current directory.
 	\end{itemize}
 	\begin{example}
-		Suppose the \lstinline|eiicl| binary is in the \$PATH and the current directory contains a file \lstinline|example.p| containing
+		Suppose the \lstinline|eiicl| binary is in the \$PATH and the current directory contains a file \lstinline|example.p| containing (besides a conjecture and axioms)
 		\begin{lstlisting}
@@
 	\end{lstlisting}
-	will output the correct translation.
+	will output the correct translation according to the previous chapter.
 	\end{example}
@@
 
-	\begin{figure}[h]\label{results}
+	\begin{figure}[h]
 		$$\begin{matrix}
@@
 		\caption{Performance on problem sections of the ILTP set.}
+		\label{results-bench}
 	\end{figure}
@@
 The problem \lstinline|SYN007+1.014| was removed from the set as our translation mechanism would generate $>2^{14}$ formulas.
-For the purpose of the benchmark we report timeout for this problem.
-Since \lstinline|SYN007+1.014| is rather artificial this is not a big issue for us.
-For the classical proving part of our benchmark we use the Vampire theorem prover~\cite{Kov_cs_2013}, which is a state-of-the-art prover as evidenced by the CASC competition~\cite{casc}.
-Then for each problem our system transforms the described transformation and feeds the result to vampire with a time limit of 30 seconds which then reports validity, satisfiability of the negation or a timeout.
+For the purpose of the benchmark, we report timeout for this problem.
+Since \lstinline|SYN007+1.014| is rather artificial, this is not a big issue for us.
+For the classical proving part of our benchmark, we use the Vampire theorem prover~\cite{Kov_cs_2013}, which is a state-of-the-art prover as evidenced by the CASC competition~\cite{casc}.
+Then for each problem our system transforms the described transformation and feeds the result to Vampire with a time limit of 30 seconds which then reports validity, satisfiability of the negation or a timeout.
 The benchmark was run on an Intel Core i5-8400 CPU.
@@
 Out of the provers on the ILTP website this only puts it behind \lstinline|ileanCoP| which managed to resolve 690 or $27.1\%$ of the problems with a proof or refutation and ahead of five other provers.
-A caveat is that the ILTP benchmarks were run on a slightly less powerful machine but with a timelimit of 600s, which suggests that our result might be even better.
+A caveat is that the ILTP benchmarks were run on a slightly less powerful machine but with a timelimit of 600s, which suggests that our result could be even better when run under the same conditions.
 
 	The ILTP problem set is divided into 24 separate sections consisting of different kinds of problems, e.g.
-ALG which are primarily algebraic in nature.
-The performance on the individual sections can be seen it table~\ref{results}.
-	More detailed results for the individual problems can be found in Apendix~\ref{benchmark-results}.
-
-	\chapter{Translation of proofs and counter-models}
-
-	Finally we elaborate on how generated proofs and counter-models for $\varphi^\#$ can be explicitly translated in the first-order case.
+ALG which contains problems tjat are primarily algebraic in nature.
+The performance on the individual sections can be seen it table~\ref{results-bench}.
+	More detailed results for the individual problems can be found in Appendix~\ref{benchmark-results}.
+
+	\chapter{Translation of counter-models and proofs}
+
+	Finally, we elaborate on how generated proofs and counter-models for $\varphi^\#$ can explicitly be translated in the first-order case.
 
@@
 			\item At each world $u$ define the universe $M_u := \{m\in M\:|\:E^I(m, u^I)\}$ where $f_{\vec z}(t)^I := f^I(\vec z, t^I)$.
-			\item For each function symbol $f$ have $f^{I_u} := f^I_{M_u}$.
-			\item For each relation symbol $R$ have $R^{I_u}(\vec t) := {R^\#}^I(\vec t, u^I)$
+			\item For each function symbol $f$ have $f^{I_u} := f^I|_{M_u}$.
+			\item For each relation symbol $R$ have $R^{I_u}(\vec t) :\Leftrightarrow {R^\#}^I(\vec t, u^I)$
 		\end{itemize}
@@
 	\begin{theorem}
-		If $\mathcal M\not\models\varphi^\#$ then $ \mathcal K(\mathcal M)\not\models\varphi$
+		If $\mathcal M\not\models\varphi^\#$, then $ \mathcal K(\mathcal M)\not\models\varphi$.
 	\end{theorem}
@@
 
-	We want to transform Resolution proof for $\varphi^\#$ as given by vampire into a \textbf{Gi} proof for $\varphi$.
+	We want to transform Resolution proof for $\varphi^\#$ as given by Vampire into a \textbf{Gi} proof for $\varphi$.
 While the results of this section are also clear in principle, there is some additional work to be done in the details.
-Since syntactic details are of importance here it might be useful to recall preliminary sections~\ref{proof-theory}.
+Since syntax is of crucial importance here it might be useful to recall preliminary section~\ref{proof-theory}.
 We will focus on how to obtain a \textbf{Gi} proof for $\varphi$ from a \textbf{Gc} proof of $\varphi^\#$, which is a much closer calculus.
-To obtain a \textbf{Gc} proof from a resolution proof one can use a standard method, e.g.
-as given in~\cite{de2005translation}.
+To obtain one from a resolution proof one might first use a standard method, e.g.
+as given in~\cite{de2005translation}, to translate the resolution proof into a \textbf{Gc} proof and then obtain a \textbf{Gi} proof via our method.
+
 	Recall that $\varphi^\#$ is of the form $K\to \bigwedge\mathcal S^\#\to P^\#(s)$, which is equivalent to $K,\mathcal S^\#\Rightarrow P^\#(s)$ where each $\psi\in\mathcal S^\#$ is of one of the forms
@@
 	Suppose now we have a \textbf{Gc} proof $\mathfrak P$ of $\varphi^\#$.
-Without loss of generality suppose that the rules introducing the left $\forall$ quantifiers are at the end of the proof, i.e.
-each atom occurring until then will have ground terms in place of $\vec t$ and $u$.
+Without loss of generality, suppose that the rules introducing the left $\forall$ quantifiers are at the end of the proof, i.e.
+each atom occurring until then will have ground terms in place of $\vec z/\vec a/\vec b/\vec c$ and $u$.
 This can be achieved by permutation of rules.
@@
 
-	Due to the subformula property we can achieve by another permutation argument that each rule R$\to$ and R$\forall$ is followed by its associated L$\to$ rule, i.e.
-all rules  R$\to$ and R$\forall$ occurr as
+	Due to the subformula property, we can achieve by another permutation argument that each rule R$\to$ and R$\forall$ is followed by its associated L$\to$ rule, i.e.
+all rules R$\to$ and R$\forall$ occur as
 
@@
 		\RightLabel{L$\to$}
-		\BinaryInfC{$(A^\#(\vec a, f_\psi(\vec z, w))\to B^\#(\vec b, f_\psi(\vec z, w)))\to C^\#(\vec c, w), \Gamma\Rightarrow \Delta$}
+		\BinaryInfC{$A^\#(\vec a, f_\psi(\vec z, w))\to B^\#(\vec b, f_\psi(\vec z, w))\to C^\#(\vec c, w), \Gamma\Rightarrow \Delta$}
 		\DisplayProof
@@
 	\end{center}
-	for some ground term $w$.
-
-	We will now show that $\mathfrak A$ can be transformed into an intuitionistically valid proof by induction on the number of such rules.
-
-	Consider the last such rule.
+	for some ground term $w$. By another permutation argument we can move all these rule sequences to the bottom of the proof.
+
+	We will now show that $\mathfrak A$ can be transformed into an intuitionistically valid proof by induction on the number of such rule sequences.
+
+	Consider the last rule R$\to$.
 We must have $\Delta = P^\#(s)$.
@@
 		\item  On the other hand suppose the subproof does not contain a valid proof of 
-		$$A^\#(\vec a, f_\psi(\vec z, w)),\Gamma\Rightarrow B^\#(\vec b, f_\psi(\vec z, w)).$$Then removing all formulas containing $f_\psi(\vec z, w)$ and all rules featuring it and then adding missing redundant formulas in the axioms yields a valid classical proof of $$(A^\#(\vec a, f_\psi(\vec z, w))\to B^\#(\vec b, f_\psi(\vec z, w)))\to C^\#(\vec c, w), \Gamma\Rightarrow \Delta.$$ But then, since we removed one R$\to$ formula, we also obtain an intuitionistic proof via the induction hypothesis.
+		$$A^\#(\vec a, f_\psi(\vec z, w)),\Gamma\Rightarrow B^\#(\vec b, f_\psi(\vec z, w)).$$Then removing all formulas containing $f_\psi(\vec z, w)$ and all rules featuring it and then adding missing redundant formulas in the axioms yields a valid classical proof of $$(A^\#(\vec a, f_\psi(\vec z, w))\to B^\#(\vec b, f_\psi(\vec z, w)))\to C^\#(\vec c, w), \Gamma\Rightarrow \Delta.$$ But then, since we removed one R$\to$ formula, we obtain an intuitionistic proof of the original claim via the induction hypothesis.
 	\end{itemize}
@@
 	\begin{theorem}
-		There is an effective linear time procedure that transforms a classical proof of $\varphi^\#$ into a intuitionistic proof of $\varphi$.
+		There is an effective procedure that transforms a classical proof of $\varphi^\#$ into a intuitionistic proof of $\varphi$.
 	\end{theorem}
@@
 		\]
-			For this example we shall consider the simplified but still valid alternative
+			For this example, we shall consider the simplified but still valid alternative
 			\[
@@
 
-			For this we obtain the following proof:
+			For this, we obtain the following proof:
 			\begin{center}
@@
 			\end{center}
-			First we push all the L$\forall$ rules to the bottom and group L$\to$ with R$\to$/R$\forall$:
+			First, we push all the L$\forall$ rules to the bottom and group L$\to$ with R$\to$/R$\forall$:
 		\begin{center}
@@
 
-		Next we consider the subproof spanned by formulas containing $g(s)$ and rules featuring $g(s)$ above the last R$\to$ rule. But there are no rules featuring $g(s)$, i.e. we obtain an invalid proof.
+		Next, we examine the last R$\to$ rule and consider the subproof spanned by formulas containing $g(s)$ and rules featuring $g(s)$ above the last R$\to$ rule.
+But there are no rules featuring $g(s)$, i.e.
+we obtain an invalid proof.
 		\begin{center}
@@
 			\RightLabel{Ax}
-			\UnaryInfC{$A^\#(g(s))
+			\UnaryInfC{$A^\#(f(s))
 				\Rightarrow B^\#(s), A^\#(f(s))$}
@@
 		\end{center}
-		Then we add the superfluous formulas back in the axioms:
+		Then, we add the superfluous formulas back in the axioms:
 		\begin{center}
@@
 			\RightLabel{Ax}
-			\UnaryInfC{$\substack{A^\#(g(s)),\\\forall u(\forall x(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))}
+			\UnaryInfC{$\substack{A^\#(f(s)),\\\forall u(\forall x(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))}
 				\Rightarrow B^\#(s), A^\#(f(s))$}
@@
 		\end{center}
-	We then inductively proceed by examining the new last R$\to$ rule.
-This time however the subproof spanned by formulas and rules containing $f(s)$ is completely valid.
-Since the subproof does not contain any more R$\to$ or R$\forall$ formulas we are done.
+	We then inductively proceed by examining the new last R$\to$ rule, i.e. we consider the subproof spanned by formulas containing $f(s)$ and rules featuring $f(s)$ above the last R$\to$ rule.
+	\begin{center}
+		\AxiomC{}
+		\RightLabel{Ax}
+		\UnaryInfC{$A^\#(f(s))\Rightarrow A^\#(f(s))$}
+		\DisplayProof
+	\end{center}
+
+	This time however the subproof is completely valid.
+We then proceed inductively by transforming this subproof.
+Since it does not contain any more R$\to$ or R$\forall$ formulas we are done and can insert it back into the original proof and add the remaining formulas in the axioms, yielding an intuitionistically valid proof:\\
+	\begin{center}
+		\AxiomC{}
+		\RightLabel{Ax}
+		\UnaryInfC{$\substack{A^\#(f(s))\to A^\#(f(s))\to B^\#(s), A^\#(f(s)),\\A^\#(f(s))\to A^\#(f(s))\to B^\#(s)}\Rightarrow A^\#(f(s))$}
+		\RightLabel{R$\to$}
+		\UnaryInfC{$\substack{A^\#(f(s))\to A^\#(f(s))\to B^\#(s), A^\#(f(s)),\\\forall u(\forall x(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))}\Rightarrow B^\#(s), A^\#(f(s))\to A^\#(f(s))$}
+		\AxiomC{}
+		\RightLabel{Ax}
+		\UnaryInfC{\vdots}
+		\RightLabel{L$\to$}
+		\BinaryInfC{$\substack{A^\#(f(s))\to A^\#(f(s))\to B^\#(s),\\\forall u(\forall x(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))}\Rightarrow B^\#(s)$}
+		\RightLabel{L$\forall$}
+		\UnaryInfC{$\substack{\forall u((A^\#(f(u))\to A^\#(f(u)))\to B^\#(u))\\\forall u(\forall x(A^\#(g(u))\to B^\#(g(u)))\to B^\#(u))}\Rightarrow B^\#(s)$}
+		\DisplayProof
+	\end{center}
 	\end{example}
@@
 	\chapter{Conclusion}
-
-\section{Conclusion}
 
@@
 	A key motivation for our work was a new approach for automated deduction for intuitionistic logic leveraging classical provers.
-	Our initial benchmarks showed decent results and proved this to be a legitimate route.
-
-	So far our translation is very straightforward and while this was conceptually very helpful we expect that better results can be achieved by more elaborate mechanisms.	We plan to establish better bounds in future work by utilizing structural properties of the input formula, in particular by constructing smaller counter-examples $\mathcal M_T$ in case certain atoms are independent from each other, and utilizing additional information about the functions in $\mathcal F$ to reduce the space of Kripke Frames that need to be considered.
+
+	We have tested this approach on the ILTP problem set using the Vampire theorem provers.
+20.2\% of the benchmark problems were resolved, which is a decent result but falls short of the state-of-the-art.
+Nonetheless, this is a hopeful sign and proof that the used concept has merit.
+
+	So far, our translation is very straightforward and while this was conceptually very helpful, we expect that better results can be achieved by more elaborate mechanisms.	We plan to establish better bounds in future work by utilizing structural properties of the input formula, in particular we plan to reduce the space of Kripke Frames that need to be considered by ascertaining that certain subformulas can be examined independently of each other and utilizing additional information about the functions in $\mathcal F$.
 
@@
 	\appendix
+
+	\chapter{Deutscher Abstract}
+
+	Die berühmte Double Negation Translation stellt eine Einbettung von klassischer in intuitionistische Logik dar.
+Interessanterweise ist die umgekehrte Richtung in der Literatur noch nicht behandelt worden.
+Wir präsentieren eine effektive Einbettung von intuitionistischer in klassische Logik, sowohl im Fall der Aussagenlogik als auch im Fall der Prädikatenlogik, sowie eine effektive Einbettung von intuitionistischer Aussagenlogik in quantifizierte boolesche Formeln.
+
+	Darüber hinaus implementieren wir ein System, das intuitionistische Probleme der Prädikatenlogik im tptp-Dateiformat als Input nimmt und die Transformation durchführt.
+Dies ermöglicht die Verwendung von klassischen Beweissystemen zur Überprüfung intuitionistischer Gültigkeit.
+Wir testen unsere Implementierung mithilfe des Vampire-Beweissystems an der ILTP-Datenbank.
+
+	Schließlich diskutieren wir, wie die generierten klassischen Beweise des transformierten Problems in intuitionistische Beweise des ursprünglichen Problems zurückübersetzt werden können.
+Insgesamt wird damit ein neuartiger Ansatz zum automatisiertem Theorembeweis für intuitionistische Logik etabliert und ein erster Proof of Concept erbracht.
+
 	\chapter{Zusammenfassung auf Deutsch}
+
+	Mathematischer Konstruktivismus bezeichnet einen Ansatz, in dem die Existenz eines Objekts nur durch eine explizite Konstruktion nachgewiesen werden kann, im Gegensatz zur klassischen Mathematik, bei der die Existenz implizit nachgewiesen werden kann, z.
+B.
+durch die Annahme der Nichtexistenz und die Ableitung eines Widerspruchs.
+	Der üblicherweise mit Konstruktivismus assoziierte Formalismus ist die intuitionistische Logik, die sich von der klassischen Logik im Wesentlichen dadurch unterscheidet, dass der Satz vom ausgeschlossenen Dritten $A\vee\neg A$ und der sogennante Double Negation Shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ nicht gültig sind.
+	Neben philosophischen Überlegungen gibt es eine besondere Motivation für das Studium der konstruktiven Mathematik aus der Perspektive der Informatik, da Beweise in der intuitionistischen Logik direkt Computerprogrammen entsprechen --- wie in der Curry-Howard-Korrespondenz~\cite{howard1980formulae} ausgedrückt.
+
+	Das Interesse an intuitionistischer Logik hat zur Entwicklung von automatisierten Beweissystemen geführt.
+	Die Fortschritte beim automatisierten Beweisen für intuitionistische Logik sind jedoch langsam, während Systeme für klassische Logik enorme Fortschritte gemacht haben, siehe z.
+B.
+die Wettbewerbe TPTP~\cite{casc} und SAT~\cite{satc}.
+	Dies kann teilweise durch grundlegende Unterschiede zwischen den Logiken erklärt werden.
+	Zunächst einmal ist die Bestimmung der intuitionistischen Gültigkeit rechnerisch komplexer: In der Aussagenlogik ist intuitionistische Gültigkeit \verb+PSPACE+-vollständig~\cite{statman1979intuitionistic} während klassische Gültigkeit \verb+coNP+-vollständig~\cite{cook1971complexity} ist.
+	Ein weiterer Vorteil der klassischen Logik ist die Existenz von Kalkülen, die sich besonders gut für Automatisierung eignen, wie z.B.
+Superposition~\cite{bachmair2001resolution}, die sich auf Normalformen wie CNF und die Dualität zwischen Gültigkeit und Erfüllbarkeit stützt.
+	Zwar gibt es auch für intuitionistische Logik einige (wenn auch komplexere) Normalformen, doch ist die Dualität zwischen Gültigkeit und Erfüllbarkeit der Negation nicht gegeben.
+	Daher verwenden die meisten dedizierten intuitionistischen Beweissysteme~\cite{mclaughlin2009efficient, tammet1996resolution} eine direkte Suche nach einem schnittfreien Beweis durch die umgekehrte Anwendung der Regeln eines Beweiskalküls.
+Dieser Ansatz führt im Allgemeinen zu einer viel komplexeren Suche und ist daher nur schwer effizient zu implementieren.
+	Schließlich fügen wir hinzu, dass im Gegensatz zu intuitionistischen Beweissystemen eine enorme Menge an Arbeit in die Optimierung von Beweissystemen für klassische Logik, insbesondere in der Aussagenlogik, d.h.
+SAT-Solver, investiert wurde.
+
+	Mit dieser Arbeit wollen wir einen neuen Ansatz für das intuitionistische Theorembeweisen vorschlagen, der den Fortschritt im klassischen Theorembeweisen nutzt:
+	\begin{itemize}
+		\item Für jede Formel $\varphi$ geben wir eine Formel $\varphi^\#$ an, die genau dann klassisch gültig ist, wenn $\varphi$ intuitionistisch gültig ist.
+		\item Wir prüfen die Gültigkeit/Gültigkeit von $\varphi^\#$ mit einem klassischen Beweissystem.
+		\item Wir transformieren den erzeugte Beweis/das Gegenmodell von $\varphi^\#$ zu einem von $\varphi$.
+	\end{itemize}
+	Der schwierigste Teil dieses Ansatzes besteht darin, die Übersetzung von $\varphi$ in $\varphi^\#$ zu geben.
+	Interessanterweise ist die umgekehrte Richtung, die berühmte Double-Negation Translation, seit langem bekannt und geht der Aussagenlogik auf Glivenko~\cite{glivenko1929quelques} und der Prädikatenlogik auf G\"odel~\cite{godel1933intuitionistischen} und Gentzen~\cite{gentzen1936widerspruchsfreiheit} zurück.
+Im propositionalen Fall ist sie besonders einfach: $\varphi$ ist klassisch gültig, genau dann wenn $\neg\neg\varphi$ intuitionistisch gültig ist.
+Intuitiv kollabiert die Übersetzung für jede Teilformel $\psi$ von $\varphi$ die Wahrheitswerte von $\psi$ und $\neg\neg\psi$, die klassisch, aber nicht intuitionistisch äquivalent sind.
+Dies gibt uns eine erste Idee, warum die umgekehrte Richtung vielleicht schwieriger ist: Wir müssen die Wahrheitswerte von $\psi$ und $\neg\neg\psi$ disambiguieren, d.h.
+wenn sie beide in $\varphi$ vorkommen, müssen wir eine Möglichkeit haben, ihrem jeweiligen Gegenstücken in $\varphi^\#$ (klassisch) unterschiedliche Wahrheitswerte zuzuordnen.
+Dies erfordert insbesondere die Einführung neuer Variablen in unserem Verfahren.
+
+	Während wir die Übersetzung von Formeln etablieren, führen wir auch eine effektive Übersetzung von Gegenmodellen durch, d.
+h.
+für jedes intuitionistische Gegenmodell von $\varphi$ konstruieren wir effektiv ein klassisches Gegenmodell von $\varphi^\#$ und vice versa.
+	Wir stellen fest, dass die Existenz von Gegenmodellen ein Schlüsselkonzept ist, das ein Dual zur Gültigkeit bildet.
+	Die Transformation und Reduktion von Gegenmodellen auf eine Normalform ist auch das, was letztlich unsere Übersetzung ermöglicht.
+Neben einer Übersetzung von Gegenmodellen beschreiben wir im letzten Kapitel auch explizit, wie man klassische Beweise von $\varphi^\#$ in intuitionistische Beweise von $\varphi$ transformiert.
+
+	Als letzten Beitrag implementieren wir unsere Übersetzung in der Programmiersprache Rust.
+Unsere Implementierung transformiert ein Problem der Prädikatenlogik im tptp-Format~\cite{tptp} und gibt das übersetzte Problem in demselben Format aus.
+Der Code ist auf Github~\cite{implementation} veröffentlicht.
+Anschließend testen wir unsere Implementierung an der ILTP-Datenbank~\cite{iltp}, d.h.
+wir übersetzen alle Probleme in der Datenbank und lassen dann das Vampire-Beweissystem~\cite{Kov_cs_2013} auf den übersetzten Problemen laufen.
+Unser Ansatz kann mit bestehenden Ansätzen für das intuitionistische Theorembeweisen mithalten aber die Spitzenreiter nicht erreichen.
+Da es noch viel Raum für Optimierungen gibt, ist dies ein vielversprechendes erstes Zeichen.
+
+	\newgeometry{top=2cm, right=3cm, left=3cm, bottom=3.8cm}
 
@@
 	\begin{minipage}{0.45\textwidth}
-		\[\begin{matrix}
+		\vspace*{-1.5cm}\[\begin{matrix}
 			\text{Problem}&\text{Result}&\text{time}\\
@@
 
+	\restoregeometry
+
 	\bibliographystyle{plain}
+	\addcontentsline{toc}{chapter}{References}
 	\bibliography{EIICL}
+
+	\pagebreak
+
+	\pagenumbering{gobble}
+	\
+	\newpage
+
 \end{document}

### 2022-10-05 Presentation/presentation.tex
Defensio

\documentclass{cubeamer}

\usepackage{trimclip}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{arrows,shapes,quotes}
\usepackage{amsmath,amsfonts}

\title{Automated Deduction for Intuitionistic Logic via Embedding into Classical Logic}
\subtitle{Defensio}
\author[Alexander Pluska]{Alexander Pluska}
\date{\today} % or whatever the date you are presenting in is
\institute[University of Vienna]{University of Vienna}
% \copyrightnotice{Published by the American Institute of Aeronautics and Astronautics, Inc., with permission}

\begin{document}

	\maketitle

	\cutoc

	\section{Background}

	\begin{frame}{Automated Deduction}
			\begin{figure}[!tbp]
				\centering
				\begin{subfigure}[b]{0.35\textwidth}
					\begin{center}
						\includegraphics[width=\textwidth]{Vampire.png}
						\begin{subfigure}[b]{0.4\textwidth}
							\includegraphics[height=0.2\textheight]{Voronkov.png}
						\end{subfigure}
						\hfill
						\begin{subfigure}[b]{0.4\textwidth}
							\includegraphics[height=0.2\textheight]{Kovacs.png}
						\end{subfigure}
						\caption{Andrei Voronkov \& Laura Kovács}
					\end{center}
				\end{subfigure}
				\hfill
				\begin{subfigure}[b]{0.6\textwidth}
					\includegraphics[width=\textwidth]{CASC.png}
					\caption{CASC-28 Results}
				\end{subfigure}
			\end{figure}
	\end{frame}

	\begin{frame}{Intuitionistic Logic}
		\begin{figure}[!tbp]
			\centering
			\begin{subfigure}[][][t]{0.6\textwidth}
				\begin{itemize}
					\item Motivation:
					\begin{itemize}
						\item Mathematical Constructivism.
						\item Curry-Howard-Correspondence.
					\end{itemize}
					\item Not intuitionistically valid:
					$$\neg\neg A\to A\hspace*{1cm} \forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$$
				\end{itemize}
			\end{subfigure}
			\hfill
			\begin{subfigure}[]{0.3\textwidth}
				\centering
				\includegraphics[width=0.9\textwidth]{logicians.png}
				\caption{L. E. J. Brouwer, A. Heyting, H. B. Curry, W. A. Howard}
			\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}{Kripke Semantics}
		\begin{definition}
			A \textit{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of worlds and a family of valuations $(v_w)_{w\in W}$ such that for $u\leq w$ we have $v_u(A)\leq v_w(A)$ for every propositional variable $A$. We define a model relation between worlds $u$ and formulas as follow:
			\begin{itemize}
				\item $u\not\models \bot$
				\item $u\models A$ iff $v_u(A) = 1$ for each propositional variable $A$.
				\item $u\models \varphi\wedge\psi$ iff $v\models\varphi$ and $v\models\psi$.
				\item $u\models\varphi\vee\psi$ iff $v\models\varphi$ or $v\models\psi$.
				\item $u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $w\models\psi$.
			\end{itemize}
		\end{definition}
	\end{frame}

	\begin{frame}{Kripke Semantics - Example}
			\begin{figure}[!tbp]
			\centering
			\begin{subfigure}[][][t]{0.3\textwidth}
				\centering
					$A\vee\neg A\:\approx$ ``$A$ or never $A$''\vspace*{1cm}
					\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
						\node (center) {};
						\node[squarednode]      (left)   [left=of center] {};
						\node[squarednode]      (right)       	[right=of center] {$A$};
						\draw[-{Stealth[length=3mm, width=3mm]}] (left.east) -- (right.west);	
					\end{tikzpicture}
				\caption{Counter-model $A\vee\neg A$}
			\end{subfigure}
			\hfill
		\begin{subfigure}[][][t]{0.6\textwidth}
			\centering
			$\forall x\neg\neg A(x)\to \neg\neg\forall x A(x)\:\approx$ ``for all $x$ eventually $A(x)$ implies eventually for all $x$ $A(x)$''\vspace*{1cm}
			\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
				\node[squarednode]  (0) {\color{red}$0$};
				\node[squarednode]      (1)   [right=of 0] {\color{green}$0$, \color{red}$1$};
				\node[squarednode]      (2)       	[right=of 1] {\color{green}$0, 1$, \color{red}$2$};
				\node   (3)       	[right=of 2] {\dots};
				\draw[-{Stealth[length=3mm, width=3mm]}] (0.east) -- (1.west);	
				\draw[-{Stealth[length=3mm, width=3mm]}] (1.east) -- (2.west);
				\draw[-{Stealth[length=3mm, width=3mm]}] (2.east) -- (3.west);	
			\end{tikzpicture}
			\caption{Counter-model $\forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$}
		\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}{Automated Deduction for Intuitionistic Logic}
		\begin{figure}
			\begin{subfigure}{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{statman.png}
			\end{subfigure}
			\begin{subfigure}{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{cook.png}
			\end{subfigure}
		\begin{subfigure}[b]{0.49\textwidth}
			\centering
				\includegraphics[width=\textwidth]{iltp.png}
			\caption{ILTP benchmark}
		\end{subfigure}
		\begin{subfigure}[b]{0.49\textwidth}
			\includegraphics[width=0.9\textwidth]{CASC.png}
			\caption{CASC-28 Results}
		\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}{Existing approaches}
		\begin{figure}
			\begin{subfigure}{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{friedman.png}
			\end{subfigure}
			\begin{subfigure}{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{schwichtenberg.png}
			\end{subfigure}
		\vspace{.5cm}
			\begin{subfigure}[b]{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{griffin.png}
			\end{subfigure}
			\begin{subfigure}[b]{0.49\textwidth}
				\centering
				\includegraphics[width=0.9\textwidth]{parigot.png}
			\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}{Our approach}
		\begin{enumerate}
			\item Give an embedding of intuitionistic logic into classical logic.
			\item Leverage a classical prover to check the validity of the translated formula.
			\item Translate the generated proof/counter-example to one of the original formula.
		\end{enumerate}
	\end{frame}

	\section{The Embedding}

	\begin{frame}{Overview}
		\begin{enumerate}
			\item Tseytin-like Normalization
			\item Classical Encoding of Kripke Semantics
			\item Reduction of the Encoding
		\end{enumerate}
	\end{frame}

	\begin{frame}{Tseytin-like Normalization}
			\begin{lemma}[Tseytin-like Normalization, propositional case]
				Let $\varphi$ be a propositional formula. There is effectively a set $\mathcal S$ of formulas of the form $$(A\wedge B)\to C, A\to (B\wedge C), (A\vee B)\to C, A\to (B\vee C), (A\to B)\to C$$ and new propositional variable $P_\varphi$ such that $\varphi$ is valid if and only if $\mathcal S\to P_\varphi$ is.
			\end{lemma}	
	\end{frame}

	\begin{frame}{Example}
		For the non-intuitionistic tautology $\varphi = (A\to B)\vee (B\to A)$ we can have
		\begin{align*}
			\mathcal S &= \{\phantom{(P_{A\to B}\vee P_{B\to A})\to P_\varphi, (A\to B)\to P_{A\to B}, (B\to A)\to P_{B\to A}\}}
		\end{align*}
	\end{frame}

	\begin{frame}{Example}
		For the non-intuitionistic tautology $\varphi = (A\to B)\vee (B\to A)$ we can have
		\begin{align*}
			\mathcal S &= \{(P_{A\to B}\vee P_{B\to A})\to P_\varphi\phantom{, (A\to B)\to P_{A\to B}, (B\to A)\to P_{B\to A}\}}
		\end{align*}
	\end{frame}

	\begin{frame}{Example}
		For the non-intuitionistic tautology $\varphi = (A\to B)\vee (B\to A)$ we can have
		\begin{align*}
			\mathcal S &= \{(P_{A\to B}\vee P_{B\to A})\to P_\varphi, (A\to B)\to P_{A\to B}\phantom{, (B\to A)\to P_{B\to A}\}}
		\end{align*}
	\end{frame}

	\begin{frame}{Example}
		For the non-intuitionistic tautology $\varphi = (A\to B)\vee (B\to A)$ we can have
		\begin{align*}
			\mathcal S &= \{(P_{A\to B}\vee P_{B\to A})\to P_\varphi, (A\to B)\to P_{A\to B}, (B\to A)\to P_{B\to A}\phantom{\}}
		\end{align*}
	\end{frame}

	\begin{frame}{Example}
		For the non-intuitionistic tautology $\varphi = (A\to B)\vee (B\to A)$ we can have
		\begin{align*}
			\mathcal S &= \{(P_{A\to B}\vee P_{B\to A})\to P_\varphi, (A\to B)\to P_{A\to B}, (B\to A)\to P_{B\to A}\}
		\end{align*}
	\end{frame}

	\begin{frame}{Classical Encoding of Kripke Semantics - Propositional Case}
		Over a fixed propositional alphabet the theory of Kripke Frames is a first-order theory.
	\end{frame}

	\begin{frame}{Example}
		$$\bigwedge\{(P_{A\to B}\vee P_{B\to A})\to P_\varphi, (A\to B)\to P_{A\to B}, (B\to A)\to P_{B\to A}\} \to P_\varphi$$
		is intuitionistically valid iff
		$$\text{PO}(\preceq)\wedge\text{Persistent}\to \forall k\left(\bigwedge\left\{\substack{\forall u(k\preceq u\to (P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u)),\\ \forall u(k\preceq u\to \forall v(u\preceq v\to A(v)\to B(v))\to P_{A\to B}(u)),\\\forall u(k\preceq u\to \forall v(u\preceq v\to B(v)\to A(v))\to P_{B\to A}(u))}\right\} \to P_\varphi(k)\right)$$
		where PO$(\preceq)$ encodes that $\preceq$ is a partial order and Persistent persistency, e.g. $$\text{PO}(\preceq) = \forall u(u\preceq u)\wedge\forall u,v(u\preceq v\to v\preceq u)\wedge\forall u, v, w((u\preceq v\wedge v\preceq w)\to u\preceq w)$$
		$$\text{Persistent} = \bigwedge\{\forall u, v((u\preceq v\wedge P(u))\to P(v))\:|\:P\in\{A, B, P_{A\to B}, P_{B\to A}, P_\varphi\}\}$$
	\end{frame}

	\begin{frame}{Classical Encoding of Kripke Semantics - Predicate Case}
		We use a special predicate $E$ to encode that an element exists at a world.
		Now as in the predicate case we can encode the Kripke Semantics:
		\begin{align*}
			\varphi'= &\:\text{PartialOrder}(\preceq) \to \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w))\\
			&\to\forall u(\text{DomainClosed}(u))\to \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))\\
			&\to \forall k(\mathcal S'\to P_\varphi(k))
		\end{align*}
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		Recall that we ended with the formula
		$$\text{PO}(\preceq)\wedge\text{Persistent}\to \forall k\left(\bigwedge\left\{\substack{\forall u(k\preceq u\to (P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u)),\\ \forall u(k\preceq u\to \forall v(u\preceq v\to A(v)\to B(v))\to P_{A\to B}(u)),\\\forall u(k\preceq u\to \forall v(u\preceq v\to B(v)\to A(v))\to P_{B\to A}(u))}\right\} \to P_\varphi(k)\right).$$
		Herbrandization:
		\[\phantom{\text{PO}(\preceq)\wedge\text{Persistent}\to \bigwedge\left\{\substack{\forall u({\color{red}s}\preceq u\to (P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u)),\\ \forall u({\color{red}s}\preceq u\to (u\preceq {\color{red}f_{A\to B}(u)}\to A({\color{red}f_{A\to B}(u)})\to B({\color{red}f_{A\to B}(u)}))\to P_{A\to B}(u)),\\\forall u({\color{red}s}\preceq u\to (u\preceq  {\color{red}f_{B\to A}(u)}\to B({\color{red}f_{B\to A}(u)})\to A({\color{red}f_{B\to A}(u)}))\to P_{B\to A}(u))}\right\} \to P_\varphi({\color{red}s}).}\]
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		Recall that we ended with the formula
		$$\text{PO}(\preceq)\wedge\text{Persistent}\to \forall k\left(\bigwedge\left\{\substack{\forall u(k\preceq u\to (P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u)),\\ \forall u(k\preceq u\to \forall v(u\preceq v\to A(v)\to B(v))\to P_{A\to B}(u)),\\\forall u(k\preceq u\to \forall v(u\preceq v\to B(v)\to A(v))\to P_{B\to A}(u))}\right\} \to P_\varphi(k)\right).$$
		Herbrandization:
			\[\text{PO}(\preceq)\wedge\text{Persistent}\to \bigwedge\left\{\substack{\forall u({\color{red}s}\preceq u\to (P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u)),\\ \forall u({\color{red}s}\preceq u\to (u\preceq {\color{red}f_{A\to B}(u)}\to A({\color{red}f_{A\to B}(u)})\to B({\color{red}f_{A\to B}(u)}))\to P_{A\to B}(u)),\\\forall u({\color{red}s}\preceq u\to (u\preceq  {\color{red}f_{B\to A}(u)}\to B({\color{red}f_{B\to A}(u)})\to A({\color{red}f_{B\to A}(u)}))\to P_{B\to A}(u))}\right\} \to P_\varphi({\color{red}s}).\]
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		$$\text{PO}(\preceq)\wedge\text{Persistent}\to \bigwedge\left\{\substack{\forall u({s}\preceq u\to (P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u)),\\ \forall u({s}\preceq u\to (u\preceq {f_{A\to B}(u)}\to A({f_{A\to B}(u)})\to B({f_{A\to B}(u)}))\to P_{A\to B}(u)),\\\forall u({s}\preceq u\to (u\preceq  {f_{B\to A}(u)}\to B({f_{B\to A}(u)})\to A({f_{B\to A}(u)}))\to P_{B\to A}(u))}\right\} \to P_\varphi({s}).$$
			We may assume $s$ is minimal and $f_{A\to B}, f_{B\to A}$ are non-decreasing:
			\begin{align*}
				&\phantom{\forall k(s\preceq k)\to\forall u(u\preceq f_{A\to B}(u))\to\forall u(u\preceq f_{B\to 	A}(u))\to\text{PO}(\preceq)\wedge\text{Persistent}}\\&\phantom{\to\bigwedge\left\{\substack{\forall u(P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u)),\\ \forall u(A(f_{A\to B}(u))\to B(f_{A\to B}(u)))\to P_{A\to B}(u)),\\\forall u((B(f_{B\to A}(u))\to A(f_{B\to A}(u)))\to P_{B\to A}(u))}\right\} \to P_\varphi(s)}.
			\end{align*}
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		$$\text{PO}(\preceq)\wedge\text{Persistent}\to \bigwedge\left\{\substack{\forall u({s}\preceq u\to (P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u)),\\ \forall u({s}\preceq u\to (u\preceq {f_{A\to B}(u)}\to A({f_{A\to B}(u)})\to B({f_{A\to B}(u)}))\to P_{A\to B}(u)),\\\forall u({s}\preceq u\to (u\preceq  {f_{B\to A}(u)}\to B({f_{B\to A}(u)})\to A({f_{B\to A}(u)}))\to P_{B\to A}(u))}\right\} \to P_\varphi({s}).$$
		We may assume $s$ is minimal and $f_{A\to B}, f_{B\to A}$ are non-decreasing:
		\begin{align*}
				&\forall k(s\preceq k)\to\forall u(u\preceq f_{A\to B}(u))\to\forall u(u\preceq f_{B\to A}(u))\to\text{PO}(\preceq)\wedge\text{Persistent}\\&\to\bigwedge\left\{\substack{\forall u(P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u)),\\ \forall u(A(f_{A\to B}(u))\to B(f_{A\to B}(u)))\to P_{A\to B}(u)),\\\forall u((B(f_{B\to A}(u))\to A(f_{B\to A}(u)))\to P_{B\to A}(u))}\right\} \to P_\varphi(s).
		\end{align*}
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		It is sufficient to consider structures $\mathcal M_T = (M_T, I_T)$ where:
		\begin{itemize}
			\item $M_T$ is the set of sequences without repetition on the newly introduced function symbols ($\{\epsilon, f_{A\to B}, f_{B\to A}, f_{A\to B}f_{B\to A}, f_{B\to A}f_{A\to B}\}$ in our case).
			\item $\preceq$ is interpreted as the prefix-order.
			\item $$f_\psi^{I_T}(f_1\dots f_n) = \begin{cases}
				f_1\dots f_n, &\text{if $f_\psi$ occurs in $f_1\dots f_n$,}\\
				f_1\dots f_nf_\psi, &\text{else.}			
			\end{cases}$$
		\end{itemize}
 	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		In our example this corresponds to the following:
		\begin{figure}[width=0.9\textwidth]
			\begin{center}
				\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
					\node (center) {};
					\node[squarednode]      (right1)   [right=of center] {$f_{B\to A}$};
					\node[squarednode]      (right2)   [above=of right1] {$f_{B\to A}f_{A\to B}$};
					\node[squarednode]      (left1)    [left=of center] {$f_{A\to B}$};
					\node[squarednode]      (left2)    [above=of left1] {$f_{A\to B}f_{B\to A}$};
					\node[squarednode]      (lower)       	[below=of center] {$\epsilon$};
					\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left1.south east) node [midway, below left] {$f_{\psi_1}$};
					\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right1.south west) node [midway, below right] {$f_{\psi_2}$};
					\draw[-{Stealth[length=3mm, width=3mm]}] (left1.north) -- (left2.south) node [midway, right] {$f_{\psi_2}$};
					\draw[-{Stealth[length=3mm, width=3mm]}] (right1.north) -- (right2.south) node [midway, left] {$f_{\psi_1}$};
					\draw[-{Stealth[length=3mm, width=3mm]}] (left1.south west) to [out=225, in=135, loop, looseness=3] node [left] {$f_{\psi_1}$} (left1.north west);		
					\draw[-{Stealth[length=3mm, width=3mm]}] (right1.south east) to [out=315, in=45, loop, looseness=3] node [right] {$f_{\psi_2}$} (right1.north east);
					\draw[-{Stealth[length=3mm, width=3mm]}] (left2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (left2.north east);
					\draw[-{Stealth[length=3mm, width=3mm]}] (right2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (right2.north east);		
				\end{tikzpicture}
			\end{center}
		\end{figure}
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		And indeed there is a counter-model to $(A\to B)\vee (B\to A)$ with such a frame:
		\vspace{.5cm}
		\begin{figure}[width=0.9\textwidth]
			\begin{center}
				\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
					\node (center) {};
					\node[squarednode]      (right1)   [right=of center] {$B$};
					\node[squarednode]      (right2)   [above=of right1] {$B, A$};
					\node[squarednode]      (left1)    [left=of center] {$A$};
					\node[squarednode]      (left2)    [above=of left1] {$A, B$};
					\node[squarednode]      (lower)       	[below=of center] {$ $};
					\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left1.south east) node [midway, below left] {};
					\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right1.south west) node [midway, below right] {};
					\draw[-{Stealth[length=3mm, width=3mm]}] (left1.north) -- (left2.south) node [midway, right] {};
					\draw[-{Stealth[length=3mm, width=3mm]}] (right1.north) -- (right2.south) node [midway, left] {};
				\end{tikzpicture}
			\end{center}
		\end{figure}
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		\vspace*{-.5cm}
		\begin{align*}
			&\forall k(s\preceq k)\to\forall u(u\preceq f_{A\to B}(u))\to\forall u(u\preceq f_{B\to A}(u))\to\text{PO}(\preceq)\wedge\text{Persistent}\\&\to\bigwedge\left\{\substack{\forall u(P_{A\to B}(u)\vee P_{B\to A}(u))\to P_\varphi(u)),\\ \forall u(A(f_{A\to B}(u))\to B(f_{A\to B}(u)))\to P_{A\to B}(u)),\\\forall u((B(f_{B\to A}(u))\to A(f_{B\to A}(u)))\to P_{B\to A}(u))}\right\} \to P_\varphi(s).
		\end{align*}
		This allows us to eliminate the remaining quantifiers...
		\begin{align*}
			&\bigwedge\left\{P(u)\to P(v)\:|\:P\in\left\{\substack{A, B,\\P_{A\to B}, P_{B\to A}, P_\varphi}\right\}, (u, v)\in \left\{\substack{(\epsilon, f_{A\to B}), (\epsilon, f_{B\to A}),\\ (f_{A\to B}, f_{A\to B}f_{B\to A}), (f_{B\to A}, f_{B\to A}f_{A\to B})}\right\}\right\}\\&\to\bigwedge\left\{(P_{A\to B}(u)\to P_{B\to A}(u))\to P_\varphi(u)\:|\:u\in \{\epsilon, f_{A\to B}, f_{B\to A}, f_{A\to B}f_{B\to A}, f_{B\to A}f_{A\to B}\}\right\}\\& \to
			\bigwedge\left\{(A(v)\to B(v))\to P_{A\to B}(u)\:|\:(u, v)\in \left\{\substack{(\epsilon, f_{A\to B}), (f_{A\to B}, f_{A\to B}), (f_{A\to B}f_{B\to A}, f_{A\to B}f_{B\to A}),\\(f_{B\to A}, f_{B\to A}f_{A\to B}), (f_{B\to A}f_{A\to B}, f_{B\to A}f_{A\to B})}\right\}\right\}\\& \to
			\bigwedge\left\{(B(v)\to A(v))\to P_{B\to A}(u)\:|\:(u, v)\in \left\{\substack{(\epsilon, f_{B\to A}), (f_{B\to A}, f_{B\to A}), (f_{B\to A}f_{A\to B}, f_{B\to A}f_{A\to B}),\\(f_{A\to B}, f_{A\to B}f_{B\to A}), (f_{A\to B}f_{B\to A}, f_{A\to B}f_{B\to A})}\right\}\right\}\\&\to P_\varphi(\epsilon).
		\end{align*}
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		\begin{theorem}
			\label{thm:reduction-propositional}
			Let $\mathcal S$ be as on Slide 9 and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$. For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$. Obtain $\mathcal S^\#$ by including:
			\begin{itemize}
				\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
				\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
				\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
				\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occur in $k$.
			\end{itemize}
			Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
		\end{theorem}
	\end{frame}

	\begin{frame}{Reducing the Encoding - Predicate Case}
		\begin{itemize}
			\item Essentially works the same, it is also possible to eliminate $\preceq$ and just consider models with the term algebra of newly introduced function symbols as domain.
			\item It is \textbf{not} possible to completely eliminate quantification over worlds as some sentences only have counter-models with an infinite number of worlds.
		\end{itemize}
	\end{frame}

	\section{Implementation}

	\begin{frame}{Implementation}
		\centering
		\begin{figure}
			\includegraphics[width=\textwidth]{implementation.png}
		\end{figure}
		$\sim$3000 lines of rust code for the translation + $\sim$400 lines of python benchmarking
	\end{frame}

	\begin{frame}{Benchmark}
		\begin{figure}
				\begin{subfigure}{\textwidth}
					\scriptsize
					$$\begin{array}{*{13}c}
					&\text{AGT}&\text{ALG}&\text{COM}&\text{CSR}&\text{GEJ}&\text{GEO}&\text{GPJ}&\text{GRA}&\text{GRP}&\text{HAL}&\text{KRS}&\text{LCL}\\
					\text{Proven}&5&170&0&0&58&3&4&2&0&0&0&30\\
					\text{Disproven}&0&0&0&0&0&0&0&0&0&0&0&0\\
					\text{Timeout}&47&29&3&29&107&74&1&16&5&5&9&128
				\end{array}$$
				$$\begin{array}{*{13}c}
					&\text{MGT}&\text{MSC}&\text{NLP}&\text{NUM}&\text{PLA}&\text{PUZ}&\text{SET}&\text{SWC}&\text{SWV}&\text{SYJ}&\text{SYN}&\text{TOP}\\
					\text{Proven}&9&1&9&2&0&5&27&1&64&65&54&1\\
					\text{Disproven}&12&0&0&0&0&1&0&0&2&3&13&0\\
					\text{Timeout}&69&1&237&82&6&1&297&422&154&108&298&2
				\end{array}$$
				\end{subfigure}
				\begin{subfigure}{0.9\textwidth}
					\includegraphics[width=\textwidth]{comparison.png}
					\caption{Comparison on different problem sections}
				\end{subfigure}
		\end{figure}
	\end{frame}

	\section{Conclusion and Outlook}

\begin{frame}[standout]
	\Huge\textsc{Thank You}

	\vfill

	\LARGE\textsc{Questions?}
\end{frame}

\end{document}

### 2022-10-06 FoSSaCS/fossacs-EIICL.tex
Incorporated previous work

@@
 \usetikzlibrary{arrows,shapes,quotes}
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.21 of 2022/01/12
-%
-
-%
 \usepackage[T1]{fontenc}
@@
 	%
-	\author{Alexander Pluska\inst{1}\orcidID{0000-0002-7709-3335} \and
-		Florian Zuleger\inst{2,3}\orcidID{0000-0003-1468-8398}}
+	\author{Alexander Pluska \and Florian Zuleger}
 	%
@@
 Curiously the reverse direction has not been covered in literature.
-We present an effective embedding from intuitionistic into classical logic, both in the propositional and first-order case, as well as an effective embedding of intuitionistic propositional logic into quantified boolean formulas.
-Furthermore, in the propositional case we establish a parameter that bounds the proof complexity of intuitionistic formulas.
-As part of our construction, we also present a Tseytin-like translation that preserves intuitionistic validity and reduces formulas to a normal form.
+Based on the work in~\cite{claessen2015sat} we establish a small model property for intuitionistic propositional logic which allows us to utilize its Kripke semantics to give a simple embedding into classical propositional logic and quantified boolean formulas.
+We then transfer this procedure to first-order logic which allows us to give a embedding of intuitionistic first-order logic into classical first-order-logic.
 
@@
 
-We hope that our embeddings of classical into intuitionistic logic will allow leveraging the tremendous progress in automated reasoning in classical logic for checking intuitionistic validity, for which there has not been much progress in recent decades.
+This allows a direct application of state-of-the-art classical first-order provers to the problem of determining intuitionistic validity.
+A first benchmark shows that our approach is competitive.
 Crucially, our constructions support the transfer of counter-models to validity, which is important to model checking applications, where the output of counter-examples is an appreciated feature.
@@
 The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the fact that the law of excluded middle $A\vee\neg A$ and the double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
-Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in the Curry-Howard correspondence~\cite{howard1980formulae}.
+Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in the Curry--Howard correspondence~\cite{howard1980formulae}.
 
 The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmark problems (see e.g.
-the
-ILTP library website~\cite{iltp}).
-However, progress in automated theorem proving for intuitionistic logic has been slow, whereas reasoners for classical logic have made tremendous progress, see e.g.
-the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
+the ILTP library website~\cite{iltp}).
+However, intuitionistic provers perform poorly compared to their classical counterparts, see e.g. the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
 %, e.g., the first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}
@@
 A further advantage of classical logic is the existence of calculi that are particularly suitable for automation such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability, i.e.
-to show the validity of a formula it suffices to show the unsatisfiability of the negated formula.
-While some (albeit more complex) normal forms also exist  for intuitionistic logic, crucially the duality between validity and satisfiability of the negation does not hold.
-Therefore most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient,tammet1996resolution} use the naive inverse method, i.e.
-directly search for a cut-free proof by applying the rules from some proof calculus inversely, enhanced by search strategies such as focussing and polarization.
-This approach generally leads to a much more complex search and is therefore difficult to implement efficiently.
-Finally, we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, in particular for the propositional case, i.e.
-SAT-solvers.
-
-With this work we want to open a path for leveraging the progress in classical theorem proving for intuitionistic logic.
+to show the validity of a formula it suffices to show the unsatisfiability of the negated formula, which is insufficient in intuitionistic logic.
+The first dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient,tammet1996resolution} used the naive inverse method, i.e.
+direct search for a cut-free proof by applying the rules from some proof calculus inversely, which generally leads to a very complex search.
+More recently connection-based methods have been applied to various non-classical logics~\cite{otten2005clausal,otten2021nanocop}, including intuitionistic logic, and show promise.
+There have also been some successful attempts to study intuitionistic validity via embedding into higher-order classical logic~\cite{LEO}.
+Finally, we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic.
+
+With this work we want to examine a new but evident approach for directly leveraging the progress in classical reasoning for intuitionistic logic.
 To this end, we give an embedding from intuitionistic logic into classical logic.
@@
 
-In summary, our work suggests a new approach for obtaining a constructive proof using a classical prover, that is
+In summary, our work suggests the following approach:
 \begin{itemize}
@@
 
-\textbf{Related Work.}
-Several approaches to obtain constructive content from classical proofs have been studied in the literature.
-An important direction is to examine under which circumstances classical validity entails intuitionistic validity and therefore constructiveness.
-This entailment has most famously been established for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}, but also other settings have been studied~\cite{schwichtenberg}.
-Another important direction is to equip classical proofs with operational semantics~\cite{Control1,Parigot1}, which however does not permit in general the realizability of at least $\exists$ and $\vee$.
+\section{Preliminaries}
+
+In this section we fix notation and recall the semantics for classical and intuitionistic logic. Futhermore we recapitulate some results from~\cite{otten2005clausal}.
+
+\subsection{Propositional Semantics}
+
+\begin{definition}
+	A \emph{valuation} $v$ maps each propositional variable $A$ to a \emph{truth value} $v(A)\in\{0, 1\}$. We inductively define the model relation $\models$ between $v$ and formulas:
+	\begin{itemize}
+		\item $v\not\models \bot$ and $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
+		\item $v\models \varphi\wedge\psi$ iff $v\models\varphi$ and $v\models\psi$.
+		\item $v\models\varphi\vee\psi$ iff $v\models\varphi$ or $v\models\psi$.
+		\item $v\models\varphi\to \psi$ iff $v\not\models\varphi$ or $v\models\psi$.
+	\end{itemize}
+	A valuation $v$ is a \emph{model} for $\varphi$ if $v\models\varphi$.
+If every valuation is a model for $\varphi$ then we say $\varphi$ is \emph{valid}.
+We denote the set of valid formulas with \emph{CPC} (Classical Propositional Calculus).
+\end{definition}
+
+One notable property of classical logic is that a formula $\varphi$ is valid if and only if its negation $\neg\varphi := \varphi\to\bot$ is not satisfiable, i.e.
+there does not exist a model for it.
+The same does not hold true for intuitionistic logic as we shall see.
+
+\begin{definition}
+	A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds}, called the Kripke \emph{frame}, and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
+	The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, where we set
+	\begin{itemize}
+		\item $v_u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
+	\end{itemize}
+	We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$.
+If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$.
+A formula is valid if every Kripke structure is a model of it.
+We denote the set of valid formulas with \emph{IPC} (Intuitionistic Propositional Calculus).
+\end{definition}
+There are many classical tautologies which are not intuitionistically valid, e.g. the law of excluded middle $A\vee\neg A$. On the other hand any intuitionistic theorem is a classical one.
+
+\subsection{Predicate Semantics}
+
+We now recall the semantics of first-order logic:
+
+\begin{definition}
+	Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
+	\begin{itemize}
+		\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.
+		\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
+	\end{itemize}
+	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
+	m, &\text{if $b=a$,}\\
+	v(b), &\text{otherwise.}
+	\end{cases}$$
+	Then terms are interpreted as follows:
+	\begin{itemize}
+		\item $a^{I, v} = v(a)$ for each free variable $a$.
+		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$.
+	\end{itemize}
+	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
+	\begin{itemize}
+		\item $\mathcal M, v\not\models\bot$, $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$, $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
+		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$, $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
+		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
+		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
+		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
+	\end{itemize}
+	$\varphi$ is satisfiable if there exist $\mathcal M, v$  with $\mathcal M, v\models\varphi$.
+A $\Sigma$-structure $\mathcal M$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$, we write $\mathcal M\models\varphi$.
+A formula is valid if every $\Sigma$-structure $\mathcal M$ satisfies it.
+We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).
+\end{definition}
+
+\begin{definition}
+	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$} and $R^{I_w}|_{M_u} = R^{I_u}$.
+	Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
+	\begin{itemize}
+		\item $u, v\not\models\bot$, $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$, $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
+		\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$, $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
+		\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
+		\item $u, v\models\exists x\varphi$ iff there exists $m\in M_u$ such that $u, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
+		\item $u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
+	\end{itemize}
+	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. $\varphi$ is valid if it is satisfied by every Kripke structure.
+	$\varphi$ is valid for some frame (write $W\models\varphi$) if it is satisfied in every Krike structure with that frame.
+We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
+\end{definition}
+In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
+
+\subsection{Skolemization and Herbrandization}\label{section:herbrandiaztion}
+
+An important step in the embedding will be the elimination of quantifiers via Herbrandization.
+In this process we introduce fresh variables and add additional function symbols to the signature.
+A fresh variable is any variable that does not occur in any of the considered formulas.
+Whenever we add a function symbol we implicitly extend the signature by some not previously contained symbol.
+
+\begin{definition}
+	For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to $Z$ by simultaneous induction as follows:
+	\begin{itemize}
+		\item $A^S_Z = A^H_Z = A$ for each atomic $A$.
+		\item $(\varphi\circ\psi)^X_Z = \varphi^X_Z\circ\psi^X_Z$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
+		\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$\\$(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
+		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
+		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.
+	\end{itemize}
+	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
+\end{definition}
+
+\begin{theorem}
+	\label{thm:herbrand-skolem}
+	For every formula $\varphi$
+	\begin{itemize}
+		\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
+		\item $\varphi$ and $\varphi^H$ are classically equivalid.
+	\end{itemize}
+\end{theorem}
+
+Theorem~\ref{thm:herbrand-skolem} follows from Lemmata~\ref{ap1} and~\ref{ap2} stated in the appendix.
+
+\subsection{Intuitionistic normal forms}
+
+Finally we note the results from~\cite{otten2005clausal} which are of importance to us.
+
+\begin{lemma}\label{lemma:propositional-normal-form}
+	For every propositional formula $\varphi$ there exists an atom $P$ as well as sets of clauses $\mathcal R$ of the form
+	$$(A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)$$
+	called \emph{flat} clauses and $\mathcal X$ of the form
+	$$(A\to B)\to C$$
+	called \emph{implication} clauses such that $\varphi$ is intuitionistically equivalid to
+	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$where $A_i, B_i, A, B, C$ are atomic.
+This size of $\mathcal R$ and $\mathcal X$ is linear in the size of $\varphi$.
+In general they will contain additional propositional variables.
+\end{lemma}
+
+A similar result holds in the first order case.
+
+\begin{lemma}\label{lemma:first-order-normal-form}
+	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R$ of the form
+	$$\forall \vec x\left((A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)\right)$$
+	called \emph{flat} clauses and $\mathcal X$ of the form
+	$$\forall \vec x\left((A\to B)\to C\right)$$
+	called \emph{implication} clauses as well as $\mathcal Q$ of the form
+	$$\forall\vec x\left((\forall y A)\to B\right)$$
+	called \emph{quantification} clauses such that $\varphi$ is intuitionistically equivalid to
+	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic.
+This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$.
+In general this formula will contain new function and relation symbols.
+\end{lemma}
+
+Since we could not find a proof of this result one is given in the appendix\ref{proof:first-order-normal-form}.
 
@@
 		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
+		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occur in $k$.
 	\end{itemize}
@@
 
-\section{Preliminaries}
-
-In this section we fix notation and recall the semantics for classical and intuitionistic logic.
-
-\subsection{Propositional Semantics}
-
-\begin{definition}
-A \emph{valuation} $v$ maps each propositional variable $A$ to a \emph{truth value} $v(A)\in\{0, 1\}$. We inductively define the model relation $\models$ between $v$ and formulas:
-	\begin{itemize}
-		\item $v\not\models \bot$ and $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
-		\item $v\models \varphi\wedge\psi$ iff $v\models\varphi$ and $v\models\psi$.
-		\item $v\models\varphi\vee\psi$ iff $v\models\varphi$ or $v\models\psi$.
-		\item $v\models\varphi\to \psi$ iff $v\not\models\varphi$ or $v\models\psi$.
-	\end{itemize}
-A valuation $v$ is a \emph{model} for $\varphi$ if $v\models\varphi$.
-If every valuation is a model for $\varphi$ then we say $\varphi$ is \emph{valid}.
-We denote the set of valid formulas with \emph{CPC} (Classical Propositional Calculus).
-\end{definition}
-
-One notable property of classical logic is that a formula $\varphi$ is valid if and only if its negation $\neg\varphi := \varphi\to\bot$ is not satisfiable, i.e.
-there does not exist a model for it.
-The same does not hold true for intuitionistic logic as we shall see.
-
-\begin{definition}
-A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds}, called the Kripke \emph{frame}, and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
-The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, where we set
-	\begin{itemize}
-		\item $v_u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
-	\end{itemize}
-We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$.
-If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$.
-A formula is valid if every Kripke structure is a model of it.
-We denote the set of valid formulas with \emph{IPC} (Intuitionistic Propositional Calculus).
-\end{definition}
-There are many classical tautologies which are not intuitionistically valid, e.g. the law of excluded middle $A\vee\neg A$. On the other hand any intuitionistic theorem is a classical one.
-
-\subsection{Predicate Semantics}
-
-We now recall the semantics of first-order logic:
-
-\begin{definition}
-Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
-	\begin{itemize}
-		\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.
-		\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
-	\end{itemize}
-	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
-		m, &\text{if $b=a$,}\\
-		v(b), &\text{otherwise.}
-	\end{cases}$$
-	Then terms are interpreted as follows:
-	\begin{itemize}
-		\item $a^{I, v} = v(a)$ for each free variable $a$.
-		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$.
-	\end{itemize}
-	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
-	\begin{itemize}
-		\item $\mathcal M, v\not\models\bot$, $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$, $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
-		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$, $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
-		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
-		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
-		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
-	\end{itemize}
-	$\varphi$ is satisfiable if there exist $\mathcal M, v$  with $\mathcal M, v\models\varphi$.
-A $\Sigma$-structure $\mathcal M$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$, we write $\mathcal M\models\varphi$.
-A formula is valid if every $\Sigma$-structure $\mathcal M$ satisfies it.
-We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).
-\end{definition}
-
-\begin{definition}
-A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$} and $R^{I_w}|_{M_u} = R^{I_u}$.
-	Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
-	\begin{itemize}
-		\item $u, v\not\models\bot$, $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$, $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
-		\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$, $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
-		\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
-		\item $u, v\models\exists x\varphi$ iff there exists $m\in M_u$ such that $u, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
-		\item $u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
-	\end{itemize}
-	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. $\varphi$ is valid if it is satisfied by every Kripke structure.
-$\varphi$ is valid for some frame (write $W\models\varphi$) if it is satisfied in every Krike structure with that frame.
-We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
-\end{definition}
-In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
-
-\subsection{Skolemization and Herbrandization}
-
-An important step in the embedding will be the elimination of quantifiers via Herbrandization.
-In this process we introduce fresh variables and add additional function symbols to the signature.
-A fresh variable is any variable that does not occur in any of the considered formulas.
-Whenever we add a function symbol we implicitly extend the signature by some not previously contained symbol.
-
-\begin{definition}
-	For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to $Z$ by simultaneous induction as follows:
-	\begin{itemize}
-		\item $A^S_Z = A^H_Z = A$ for each atomic $A$.
-		\item $(\varphi\circ\psi)^X_Z = \varphi^X_Z\circ\psi^X_Z$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
-		\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$\\$(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
-		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
-		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.
-	\end{itemize}
-	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
-\end{definition}
-
-\begin{theorem}
-\label{thm:herbrand-skolem}
-For every formula $\varphi$
-	\begin{itemize}
-		\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
-		\item $\varphi$ and $\varphi^H$ are classically equivalid.
-	\end{itemize}
-\end{theorem}
-
-Theorem~\ref{thm:herbrand-skolem} follows from Lemmata~\ref{ap1} and~\ref{ap2} stated in the appendix.
-
-\section{IPC to CQC}
+\section{Bounded Intuitionistic Counter-models}
 
@@
 \begin{definition}
-	Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively as follows:
-	\begin{itemize}
+	Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively:
+	\begin{itemize}
+		\item $\bot^u = \bot$.
 		\item $A^{u} = A(u)$ for every propositional variable $A$.
-		\item $\bot^u = \bot$.
 		\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
@@
 		\text{\normalfont PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
-		\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ is a propositional variable that occurs in $\varphi$}\}
+		\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ occurs in $\varphi$}\}
 	\end{align*}
 	Then define
-	$$\varphi^{C} = K(\varphi)\to \varphi^{u}$$
-	where $u$ is some free variable.
+	$$\varphi^{C} = K(\varphi)\to \varphi^{r}$$
+	where $r$ is a new constant symbol.
 \end{definition}
 
-\noindent As all we have done is modelling Kripke frames as a first-order theory we directly obtain:
+\noindent As all we have done is modeling Kripke frames as a first-order theory we directly obtain:
 
 \begin{lemma}
-	$\varphi$ is intuitionistically valid iff $\varphi^C$ is classically valid.
-\end{lemma}
-
-\section{Normal form translation}
-\label{section:tseytin}
-
-Of course transforming a propositional formula into a predicate formula is somewhat unsatisfying.
-As a first step towards obtaining a propositional formula we want to apply Herbrandization.
-Now trying to apply it to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated.
-We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form.
-In fact our translation is often presented as the first step in the Tseytin transformation, after which the individual formulas are then converted to CNF.
-Hence, the propositional case of our translation is well known, but we are not aware of any work that uses just this step as a stand-alone.
-A similar translation for intuitionistic propositional logic was also proposed in~\cite{statman1979intuitionistic}.
+	$\varphi$ is intuitionistically valid if and only if $\varphi^C$ is classically valid.
+\end{lemma}
+
+This is somewhat unsatisfying as we end up in a much more complex logic.
+However using the normal form from Lemma~\ref{lemma:propositional-normal-form}, i.e.
+assuming that $\varphi$ is of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ for some flat clauses $\mathcal R$ and implicational clauses $\mathcal X$, further simplification is possible:
+\begin{itemize}
+	\item For flat clauses $((A_1\wedge\dots\wedge A_n)\to(B_1\vee\dots\vee B_m))^u$ is
+	$$\forall w(u\preceq w\to(A_1(u)\wedge\dots\wedge A_n(u))\to(B_1(u)\vee\dots\vee B_m(u))).$$
+	\item For implication clauses $((A\to B)\to C)^u$ is
+	$$\forall w(u\preceq w\to(\forall k(w\preceq k\to A(k)\to B(k)))\to C(w))$$
+\end{itemize}
+
+and after applying Herbrandization we end up with a formula $$\varphi^{CH} = K(\varphi)\to r\preceq s\to\left(\bigwedge\mathcal R^{(u)S}\wedge \bigwedge\mathcal X^{(u)S}\right)\to P(s)$$ where $s$ is a new constant symbol, each formula in $\mathcal R^{(u)S}$ is of the form
+$$\forall w(s\preceq w\to (A_1(u)\wedge\dots\wedge A_n(u))\to(B_1(u)\vee\dots\vee B_m(u)))$$
+and each formula $\psi\in\mathcal X^{(u)S}$ is of the form
+$$\forall w(s\preceq w\to(w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w)))\to C(w))$$
+with $f_\psi$ being a new function symbol.
+
+We now argue that if a counter-model exists for $\varphi^\circ$ then a small counter-model exists. This is ultimately what enables our translations.
 
 \begin{definition}
-Let $\varphi$ be some formula.
-For each subformula $\psi$ of $\varphi$, consider some new $n$-ary relation symbol $P_\psi$, where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$.
-We denote those variables with $\vec z_\psi$ in some arbitrary but fixed order.
-We inductively define clause sets $\mathcal S^+(\varphi)$ and $\mathcal S^-(\varphi)$ as follows:
-	\begin{itemize}
-		\item For every atomic formula $A$:\\
-$\mathcal S^+(A) = \{P_A(\vec z_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec z_A)\}$
-		\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$:\\
-$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\to P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\to P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
-		\item For every implication $\varphi \to\psi$:\\
-$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\wedge P_{\varphi}(\vec z_\varphi))\to P_{\psi}(\vec z_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec z_\varphi)\to P_{\psi}(\vec z_\psi))\to P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\psi)$
-		\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$:\\
-$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec z_{Qx\varphi})\to QxP_{\varphi}(\vec z_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec z_{\varphi})\to P_{Qx\varphi}(\vec z_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
-	\end{itemize}
-	%We further set $\mathcal S(\varphi) = \mathcal S^+(\varphi)\cup\mathcal S^-(\varphi)$.
-\end{definition}
-
-\begin{lemma}
-	$\bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)\to\varphi$ and $\bigwedge S^-(\varphi)\wedge \varphi\to P_\varphi(\vec z_\varphi)$ are valid both classically and intuitionistically.
-\end{lemma}
-
-It is straight-forward to extend structures to the additional symbols such that the formulas $S^+(\varphi)\cup\mathcal S^-(\varphi)$ are satisfied, i.e.
-for each structure $\mathcal M$ and formula $\varphi$ we can define a structure $\mathcal S(\mathcal M, \varphi)$, whose exact definition can be found in the appendix as Definition~\ref{def:transf-structure}, such that:
-\begin{lemma}\label{thm:struct1}
-For every formula $\varphi$ and structure $\mathcal M$, we have that $\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$.
-\end{lemma}
-
-\begin{lemma}\label{thm:struct2}
-	For every structure $\mathcal M$,
-	$\mathcal M\models \varphi$ if and only if $S(\mathcal M, \varphi)\models P_\varphi(\vec z_\varphi)$.
-\end{lemma}
-
-Then from the previous three Lemmata we get:
-
-\begin{corollary}\label{equivalid}
-	In both intuitionistic and classical logic
-	\begin{itemize}
-		\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)$ is, and
-		\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec z_\varphi)$ is.
-	\end{itemize}
-\end{corollary}
-
-\section{IPC to simplified CQC}
-
-We are now ready to combine the previously presented reductions.
-Let $\varphi$ be some propositional formula.
-As a first step we perform the normalization from the last section, i.e.
-instead of $\varphi$ we consider $$\varphi' = \bigwedge \mathcal S^-(\varphi)\to P_\varphi.$$
-
-Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
-$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$$
-where we can treat $A\to B$ as a special case $A\to (B\wedge B)$. We then apply the transformation to CQC, leading to a formula
-$$\varphi'' = K(\varphi')\to\bigwedge  \mathcal S\to P_\varphi(u),$$
-where $u$ is some new free variable and each $\psi\in\mathcal S$ is of one of the forms
-$$\begin{matrix}
-	\forall k(u\preceq k\to A(k)\to (B(k)\wedge C(k)))\indent  \forall k(u\preceq k\to (A(k)\wedge (B(k))\to C(k)))\\
-	\forall k(u\preceq k\to A(k)\to (B(k)\vee C(k)))\indent \forall k(u\preceq k\to (A(k)\vee B(k))\to C(k))\\
-	\forall k(u\preceq k\to (\forall l(k\preceq l \to A(l)\to B(l)))\to C(k)),
-\end{matrix}$$
-with each $P(x)$ possibly being $\bot$. Applying Herbrandization we obtain a formula
-$$\varphi''' = K(\varphi')\to \bigwedge \mathcal S'\to P_\varphi(s),$$
-where $s$ is some new constant term and each $\psi\in\mathcal S'$ is of one of the forms
-$$\begin{matrix}
-	\forall k(s\preceq k\to A(k)\to (B(k)\wedge C(k)))\indent \forall k(s\preceq k\to (A(k)\wedge (B(k))\to C(k)))\\
-	\forall k(s\preceq k\to A(k)\to (B(k)\vee C(k)))\indent \forall k(s\preceq k\to (A(k)\vee (B(k))\to C(k)))\\
-	\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A(f_\psi(k))\to B(f_\psi(k)))\to C(k)),\\
-\end{matrix}$$
-and $f_\psi$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model, then we have a counter-model $(M',I')$ where $M' = \{m\in M|s\preceq m\}$ and
-$f_\psi^{I'}(u) = f^I_\psi(u) \text{ iff $u\preceq f^{I}_\psi(u)$, and $f_\psi^{I'}(u) = u,$ else.}$ That is, where $s$ is the least element and the $f_\psi$ increase their argument.
-Therefore instead of $\varphi'''$ we may consider a formula $$\varphi^\circ = \forall k(s\preceq k)\to \bigwedge\{\forall k(k\preceq f_\psi(k))\:|\:\psi\in\mathcal F_\to\}\to K(\varphi')\to \bigwedge \mathcal S^\circ\to P_\varphi(s)$$ where each formula $\psi\in\mathcal S^\circ$ is of one of the forms
-$$\begin{matrix}
-	\forall k(A(k)\to (B(k)\wedge C(k)))\indent\forall k((A(k)\wedge B(k))\to C(k))\\
-	\forall k(A(k)\to (B(k)\vee C(k)))\indent\forall k((A(k)\vee B_\psi(k))\to C(k))\\
-	\forall k((A(f_\psi(k))\to B(f_\psi(k)))\to C(k)),
-\end{matrix}$$
-and $\mathcal F_\to$ denotes the set of formulas of the last kind.
-
-\section{IPC to CPC}
-
-We now argue that if a counter-model exists for $\varphi^\circ$ then a bounded counter-model exists, allowing us to return to the propositional case.
-
-\begin{definition}
-Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
-We say $\psi\in\mathcal F_\to$ is \emph{fulfilled} at $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true.
-For $\psi\in\mathcal F_\to$ define $$g_\psi : M\to M, u\mapsto\begin{cases}
-		u,&\text{ if $\psi$ is fulfilled in $u$,}\\
-		f^I_\psi(u),&\text{ else.}		
-	\end{cases}$$Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
-	\begin{itemize}
-		\item $M_T$ is the set of sequences without repetition on $\mathcal F_\to$.
+For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{(u)S}$. Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
+We say $(A\to B)\to C\in\mathcal X$ is \emph{fulfilled} at $m\in M$ iff $A^I(m)\to B^I(m)$ is false or $C^I(m)$ is true. For $\psi\in\mathcal X$ define $$g_\psi : M\to M, m\mapsto\begin{cases}
+		m,&\text{ if $\psi$ is fulfilled at $u$,}\\
+		f^I_{\lbrack\psi\rbrack}(m),&\text{ else.}		
+	\end{cases}$$. Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
+	\begin{itemize}
+		\item $M_T$ is the set of sequences without repetition on $\mathcal X$.
 		\item Interpret $\preceq$ as the prefix-order.
 		\item We set $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
-			\psi_1\dots\psi_n, &\text{if $\psi$ occurs in $\psi_1\dots\psi_n$,}\\
+			\psi_1\dots\psi_n, &\text{if $\psi$ occurs in $[\psi_1]\dots[\psi_n]$,}\\
 			\psi_1\dots\psi_n\psi, &\text{else.}			
 		\end{cases}$$
-		\item For propositional variables $P$, we set $$P^{I_T}(\psi_1\dots \psi_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots)).$$
+		\item For propositional variables $P$, we set $$P^{I_T}\left(\psi_1\dots \psi_n\right) = P^I\left(g_{\psi_n}\left(\dots\left(g_{\psi_1}\left(s^I\right)\right)\dots\right)\right).$$
 	\end{itemize}
@@
 \begin{lemma}~\label{thm:prop-countermodel-reduction}
-	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
+	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^{CH}$.
 	\begin{enumerate}
-		\item $\psi$ is fulfilled at $f_\psi^I(u)$ for all $u\in M, \psi\in\mathcal F_\to$.
-		\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\succeq u$.
+		\item $\psi$ is fulfilled at $f_{[\psi]}^I(u)$ for all $u\in M, \psi\in\mathcal X$.
+		\item If $\psi$ is fulfilled at some $m\in M$ then $\psi$ is fulfilled at all $n\succeq m$.
 	\end{enumerate}
@@
 
-A proof can be found in the appendix as Lemma~\ref{thm:prop-countermodel-reduction2}.
-Essentially what this tells us is that for every $u$ at which $\psi$ is fulfilled, in particular for each $u$ in the image of $f_\psi$, we may add the assumption that $f_\psi(w) = w$ for all $w\geq u$ without changing the validity of $\psi$.
+A proof can be found in the appendix as Lemma~\ref{proof:prop-countermodel-reduction}.
+Essentially what this tells us is that for every $m$ at which $\psi$ is fulfilled, in particular for each $m$ in the image of $f_\psi$, we may add the assumption that $f_\psi(n) = n$ for all $n\geq m$ without changing the validity of $\psi$.
 With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.
@@
 \begin{corollary}\label{cor:prop-tree-model}
-	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
+	If $\mathcal M$ is a counter-model to $\varphi^{CH}$ then so is $\mathcal M_T$.
 \end{corollary}
@@
 	It is well known that Kripke frames which are trees are complete with regard to intuitionistic logic.
-Our contribution is giving an explicit geometry that depends only on the structure of the considered formula.
+Our contribution is giving an explicit size and geometry that depends only on the structure of the considered formula.
 \end{remark}
 
-So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
+\begin{corollary}
+	If a formula in normal form as in Lemma~\ref{lemma:propositional-normal-form} in intuitionistically invalid then there exists a Kripke counter-model which is a rooted tree with height and degree bounded by $|\mathcal X|$.
+\end{corollary}
+
+\section{IPC to CPC}
+
+We can apply the previous results to give a direct embedding from intuitionistic into classical propositional logic.
+We know that $\varphi^{CH}$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
 	x, &\text{ if $\psi$ occurs in $x$,}\\
@@
 	\begin{align*}
-		\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi}, (P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi}\}\cup\\ &\{A^\epsilon\to P_A^\epsilon, P_A^{\psi}\to P_A^{\psi}, (A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}.
+		\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi},A^\epsilon\to P_A^\epsilon, P_A^{\psi}\to P_A^{\psi}, \}\cup\\ &\{(P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi},\}\cup\\&\{(A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}.
 	\end{align*}
-	We now note that $P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$, $P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$ defines a counter-model.
+	We now note that $P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$, $P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$ indeed defines a counter-model.
 \end{example}
@@
 It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}.
-However, giving an explicit translation from IPC to QBF still promises to be useful, allowing to leverage the progress in QBF solving.
 The translation from the last section will serve as starting point for developing the translation in this section.
@@
 
-Fix some set $\mathcal S$ of formulas of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
-and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$.
+Suppose we are given a formula in normal form $$\left(\bigwedge R\wedge\bigwedge X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}.
 We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model.
@@
 \begin{definition}
-	For every propositional variable $X$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $X^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$.
-Let $\vec X^n$ range over all propositional variables $X^n$ and $\vec F^n$ range over all $F_\psi^n$.
+	For every propositional variable $A$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $A^n$ and for every formula $\psi\in\mathcal X$ consider a new atom $X_\psi^n$.
+Let $\vec A^n$ range over all propositional variables $A^n$ and $\vec X^n$ range over all $X_\psi^n$.
 Define
 	\begin{itemize}
-		\item {\normalfont Valid}$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
-if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
+		\item {\normalfont Valid}$(n)$ which encodes that $\vec X^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
+if we interpret the $\vec X^i$ as bit-vectors, then $\vec X^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec X^i\:|\:i < n\}$.
 		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
-		\item {\normalfont Sat}$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
-		\item {\normalfont Sat}$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
-	\end{itemize}
-	Have $k = |\mathcal F_\to|$ and define$$\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal S}(i)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
+		\item {\normalfont Sat}$_{\mathcal R}(n)$ encoding that the formulas in $\mathcal R$ hold at the $n$-th world of the path.
+		\item {\normalfont Sat}$_{\mathcal X}(n)$ encoding that the formula in $\mathcal X$ that is represented by $\vec X^{n-1}$ holds.
+	\end{itemize}
+	Have $k = |\mathcal X|$ and define$$\varphi^Q_{i} = \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
 	for $0 < i < k$, as well as the special cases
-	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
-	for leafs, and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
+	$$\varphi^Q_{k} = \exists \vec A^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal R}(k)\wedge \text{\normalfont Sat}_{\mathcal X}(k)\right)$$
+	for leafs, and $$\varphi^Q = \varphi_0^Q = \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
 	Example encodings of the above formulas are:
 	\begin{align*}
-		\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
-		\text{\normalfont Persistent}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
-		\text{\normalfont Sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
-		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
-		\text{\normalfont Sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
+		\text{\normalfont Valid}(n) = &\left(\bigvee \{X^n_\psi\:|\:\psi\in\mathcal X\}\right)\wedge\left(\bigwedge \{\neg(X^n_{\psi_1}\wedge X^n_{\psi_2})\:|\:\psi_1\neq\psi_2\in\mathcal X\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{X^i_\psi\to\neg X^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal X\right\}\right)\\
+		\text{\normalfont Persistent}(n) = & \bigwedge\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\}\\
+		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal R\right\}\wedge\\
+		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal R\right\}\\
+		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\}
 	\end{align*}
@@
 	$$\hspace*{-.3cm}
-	\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right),
+	\varphi^Q = \exists A^0\forall X^0\left(\neg A^0\wedge\left(X^0\to \exists A^1\left((A^0\to A^1)\wedge(X_0\to (A^1\to \bot)\to \bot)\right)\right)\right),
 	$$
-	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $F^0$.
+	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $X^0$.
 \end{example}
@@
 \begin{lemma}
-	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
+	$\varphi$ is not intuitionistically valid if and only if $\varphi^Q$ is a satisfiable QBF.
 \end{lemma}
@@
 
-A simple counting argument shows that for each $n < |\mathcal F_\to|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal F_\to|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal S}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal F_\to}(n)|\in \mathcal O(|\mathcal F_\to|)$ so overall we get:
+A simple counting argument shows that for each $n < |\mathcal X|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal X|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal R}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal X}(n)|\in \mathcal O(|\mathcal X|)$ so overall we get:
 \begin{lemma}
-	The size of $\varphi^Q$ is in $\mathcal O(|\varphi|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$.
-\end{lemma}
-
-Note that the last set of universally quantified variables can be avoided because there is only one assignment that has the chance to falsify $\varphi^Q$, namely the assignment that assigns $1$ to the single $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$.
-Hence, in our final formula we can replace every $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^n$.
+	The size of $\varphi^Q$ is in $\mathcal O(|\varphi|\cdot|\mathcal X| + |\mathcal X|^3)$.
+\end{lemma}
+
+Note that the last set of universally quantified variables can be avoided because there is only one assignment that has the chance to falsify $\varphi^Q$, namely the assignment that assigns $1$ to the single $X_\psi^n$ such that $X_\psi^i = 0$ for all $i < n$.
+Hence, in our final formula we can replace every $X_\psi^n$ with $\bigwedge\{\neg X_\psi^i\:|\:i < n\}$ and remove that quantification over $X_\psi^n$.
 With that we get the following:
@@
 \begin{corollary}
-	Let $N$-\textbf{IntInvalid} be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid.
+	Let $N$-\textbf{IntInvalid} be the problem of deciding if a formula $(\bigwedge \mathcal R\wedge\mathcal X)\to P$ as in Lemma~\ref{lemma:propositional-normal-form} with $|\mathcal X|\leq N$ is not intuitionistically valid.
 Then $N$-\textbf{IntInvalid} is in $\Sigma_{2N-1}^P$.
@@
 	\text{DomainClosed}(u) = &\:\bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f\in\Sigma$ is a function symbol}\}\\
-	\text{Pesistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
+	\text{Persistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
 \end{align*}
@@
 superposition in the case of Vampire.
-Ultimately we hope that this will also open a new path for program extraction via the Curry-Howard Correspondence.
+Ultimately we hope that this will also open a new path for program extraction via the Curry--Howard Correspondence.
 
@@
 
-\begin{lemma}~\label{thm:prop-countermodel-reduction2}
-	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
+\begin{lemma}
+	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R$ of the form
+	$$\forall \vec x\left((A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)\right)$$
+	called \emph{flat} clauses and $\mathcal X$ of the form
+	$$\forall \vec x\left((A\to B)\to C\right)$$
+	called \emph{implication} clauses as well as $\mathcal Q$ of the form
+	$$\forall\vec x\left((\forall y A)\to B\right)$$
+	called \emph{quantification} clauses such that $\varphi$ is intuitionistically equivalid to
+	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic.
+This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$.
+In general this formula will contain new function symbols and relational symbols.
+\end{lemma}
+
+\begin{proof}\label{proof:first-order-normal-form}
+	As in~\cite{otten2005clausal} we may assume that $\varphi$ is of the form $\psi\to P$.
+Recall that the transformation works by transforming $\psi$ into a set of equivalent assumptions $\left\lbrack \psi\right\rbrack$ of the desired form.
+We extend the given procedure with rules for quantifiers as follows:
+	\begin{align*}
+		\left\lbrack(\forall x \psi)\to A\right\rbrack&:= \left\lbrack\psi[s(\vec z)/x]\to A\right\rbrack\\
+		\left\lbrack A\to (\forall x\psi)\right\rbrack&:= \{\forall x\chi[x/a]\:|\:\chi\in\left\lbrack A\to\psi[a/x]\right\rbrack\}\\
+		\left\lbrack(\exists x\psi)\to A\right\rbrack&:= \{\forall x\chi[x/a]\:|\:\chi\in\left\lbrack A\to\psi[a/x]\right\rbrack\}\\
+		A\to (\exists x\psi)&:= \left\lbrack A\to\psi[s(\vec z)/x]\right\rbrack
+	\end{align*}
+	where $s$ is a new function symbol, $a$ is a new free variable and $\vec z$ ranges over the free variables in $\psi$.
+	While the generated set of clauses is in general not equivalent to $\psi$, section~\ref{section:herbrandiaztion} guarantees that it is at least equivalid.
+\end{proof}
+
+\begin{lemma}~\label{proof:prop-countermodel-reduction}
+	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^{CH}$.
 	\begin{enumerate}
-		\item $\psi$ is fulfilled at $f_\psi^I(u)$ for all $u\in M, \psi\in\mathcal F_\to$.
-		\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\geq u$.
+		\item $\psi$ is fulfilled at $f_{[\psi]}^I(m)$ for all $m\in M, \psi\in\mathcal X$.
+		\item If $\psi$ is fulfilled at some $m\in M$ then $\psi$ is fulfilled at all $n\geq m$.
 	\end{enumerate}
@@
 \begin{proof}
-	1. If $C^I(u)$ is true then we are done due to persistency. Otherwise $C^I(u)$ is false.
-	Because $(A^I(f_\psi^I(u))\to B^I(f_\psi^I(u)))\to C^I(u)$ holds, then $A^I(f_\psi^I(u))\to B^I(f_\psi^I(u))$ must be false.
-
-	2. Let $w$ be some element with $w\geq u$.
-	If $C^I(w)$ is true, we are done.
-	Otherwise $C^I(w)$ is false.
-	Due to persistency $C^I(u)$ is also false.
-	Because $\psi$ is fulfilled at $u$, $A^I(u)\to B^I(u)$ must be false, i.e. $A^I(u)$ is true and $B^I(u)$ is false. Then $A^I(w)$ and $A^I(f^I_\psi(w))$ are also true due to persistency.
-	Because $(A^I(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I(w)$ holds, we now get that  $B^I(f^I(w))$ must be false.
-	But then due to persistency so is $B^I(w)$ and we have that $\psi$ is fulfilled at $w$.
+	1. If $C^I(m)$ is true then we are done due to persistency. Otherwise $C^I(m)$ is false.
+	Because $(A^I(f_{[\psi]}^I(m))\to B^I(f_{[\psi]}^I(m)))\to C^I(u)$ holds, then $A^I(f_{[\psi]}^I(u))\to B^I(f_{[\psi]}^I(u))$ must be false.
+
+	2. Let $n$ be some element with $n\geq m$.
+	If $C^I(n)$ is true, we are done.
+	Otherwise $C^I(n)$ is false.
+	Due to persistency $C^I(m)$ is also false.
+	Because $\psi$ is fulfilled at $m$, $A^I(m)\to B^I(m)$ must be false, i.e. $A^I(m)$ is true and $B^I(m)$ is false. Then $A^I(n)$ and $A^I(f^I_{[\psi]}(n))$ are also true due to persistency.
+	Because $(A^I(f^I_{[\psi]}(n))\to B^I(f^I_{[\psi]}(n)))\to C^I(n)$ holds, we now get that  $B^I(f^I_{[\psi]}(n))$ must be false.
+	But then due to persistency so is $B^I(n)$ and we have that $\psi$ is fulfilled at $n$.
 \end{proof}
@@
 Suppose $\varphi$ is not intuitionistically valid.
-Then, $\bigwedge\mathcal S\to P$ has an intuitionistic counter-model. Then by the previous section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
+Then, $(\bigwedge R\wedge\bigwedge X)\to P$ has an intuitionistic counter-model. By the previous section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
 We now define a QBF model iteratively.
@@
 Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$  in $\mathcal M$.
-Let the $F^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified).
-If not exactly one of the $F^{n}_\psi$ is interpreted as true, then valid$(n)$ fails and the remaining propositions can be chosen arbitrarily.
-If $F^i_{\psi_n} = 1$ for some $i < n$, then valid$(n)$ also fails and the remaining propositions can be chosen arbitrarily.
+Let the $X^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified).
+If not exactly one of the $X^{n}_\psi$ is interpreted as true, then valid$(n)$ fails and the remaining propositions can be chosen arbitrarily.
+If $X^i_{\psi_n} = 1$ for some $i < n$, then valid$(n)$ also fails and the remaining propositions can be chosen arbitrarily.
 So we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions.
 Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{\psi_1\dots\psi_n}$.
-Continue this construction until $n  = |\mathcal F_\to|$. Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.
+Continue this construction until $n  = |\mathcal X|$.
+Then from $\mathcal M$ being a counter-example to $\bigwedge\mathcal S^\#\to P^\epsilon$ it directly follows that this interpretation satisfies $\varphi^Q$.
 
@@
 Interpret $A^\epsilon$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$.
-Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$, in which $F^i_\psi$ is chosen true iff $\psi = \psi_i$.
-Let $\psi_n\in\mathcal F_\to$.
-Consider some interpretation of the $A^n$ that are part of a satisfying assignment where $F^n_\psi$ is true iff $\psi = \psi_n$ and all variables quantified above are chosen as before.
+Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$, in which $X^i_\psi$ is chosen true iff $\psi = \psi_i$.
+Let $\psi_n\in\mathcal X$.
+Consider some interpretation of the $A^n$ that are part of a satisfying assignment where $X^n_\psi$ is true iff $\psi = \psi_n$ and all variables quantified above are chosen as before.
 Have $A^{\psi_0\dots\psi_n} = A^n$ for each propositional variable $A$.

### 2022-10-07 FoSSaCS/fossacs-EIICL.tex
typos and corrected examples

@@
 \begin{lemma}\label{lemma:propositional-normal-form}
-	For every propositional formula $\varphi$ there exists an atom $P$ as well as sets of clauses $\mathcal R$ of the form
+	For every propositional formula $\varphi$ there exists an atom $P$ as well as sets of clauses $\mathcal R, \mathcal X$, with $\mathcal R$ containing  \emph{flat clauses} of the form
 	$$(A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)$$
-	called \emph{flat} clauses and $\mathcal X$ of the form
-	$$(A\to B)\to C$$
-	called \emph{implication} clauses such that $\varphi$ is intuitionistically equivalid to
+	and $\mathcal X$ containing \emph{implication clauses} of the form
+	$$(A\to B)\to C,$$
+	such that $\varphi$ is intuitionistically equivalid to
 	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$where $A_i, B_i, A, B, C$ are atomic.
@@
 \begin{lemma}\label{lemma:first-order-normal-form}
-	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R$ of the form
-	$$\forall \vec x\left((A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)\right)$$
-	called \emph{flat} clauses and $\mathcal X$ of the form
+	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R,\mathcal X, \mathcal Q$, with $\mathcal R$ containing \emph{flat clauses} of the form
+	$$\forall \vec x\left((A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)\right),$$
+	$\mathcal X$ containing \emph{implication clauses} of the form
 	$$\forall \vec x\left((A\to B)\to C\right)$$
-	called \emph{implication} clauses as well as $\mathcal Q$ of the form
-	$$\forall\vec x\left((\forall y A)\to B\right)$$
-	called \emph{quantification} clauses such that $\varphi$ is intuitionistically equivalid to
+	and $\mathcal Q$ containing \emph{quantification clauses} of the form
+	$$\forall\vec x\left((\forall y A)\to B\right),$$
+	such that $\varphi$ is intuitionistically equivalid to
 	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic.
@@
 
-Since we could not find a proof of this result one is given in the appendix\ref{proof:first-order-normal-form}.
+Since we could not find a proof of this result one is given in the appendix as Lemma~\ref{proof:first-order-normal-form}.
 
@@
 
-The most obvious approach to embedding intuitionistic into classical logic is to express intuitionistic semantics, i.e.
-Kripke frames, as a first order theory.
+The most obvious approach to embedding intuitionistic into classical logic is to examine intuitionistic semantics, e.g.
+Kripke frames, as a classical first order theory.
 For every propositional variable $A$ consider a unary predicate $A$ of the same name where $A(u)$ expresses that $A$ is true at some world $u$.
-We can then naively encode formulas as follows:
+We can then consider the following naive encoding:
 
@@
 
-\noindent As all we have done is modeling Kripke frames as a first-order theory we directly obtain:
+\noindent As all we have done is modeling Kripke frames as a first-order theory so we directly obtain:
 
@@
 
-We now argue that if a counter-model exists for $\varphi^\circ$ then a small counter-model exists. This is ultimately what enables our translations.
+We now argue that if a counter-model for $\varphi^{CH}$ exists then a small counter-model exists. This is ultimately what enables our translations.
 
@@
 		f^I_{\lbrack\psi\rbrack}(m),&\text{ else.}		
-	\end{cases}$$. Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
+	\end{cases}$$Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
 	\begin{itemize}
@@
 A proof can be found in the appendix as Lemma~\ref{proof:prop-countermodel-reduction}.
-Essentially what this tells us is that for every $m$ at which $\psi$ is fulfilled, in particular for each $m$ in the image of $f_\psi$, we may add the assumption that $f_\psi(n) = n$ for all $n\geq m$ without changing the validity of $\psi$.
-With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.
+Essentially what this tells us is that for every $m$ at which $\psi$ is fulfilled, in particular for each $m$ in the image of $f_{[\psi]}$, we may add the assumption that $f_{[\psi]}(n) = n$ for all $n\geq m$ without changing the validity of $\psi$.
+With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^{CH}$ is a straightforward check of definitions.
 
@@
 \end{cases}$$
-Now we can replace every $\forall$-quantifier by enumerating all the ground terms, and then replace all distinct ground instances of relations by new propositional variables yielding a propositional formula.
-This gives us the translation stated in Theorem~\ref{thm:reduction-propositional}.
+Now we can replace every $\forall$-quantifier in $\varphi^{CH}$ by enumerating all the ground terms, and then replace all distinct ground instances of relations by new propositional variables yielding a propositional formula.
+Note that the encoding of partial order becomes redundant with this so we drop it, but we need to keep the persistency restraints.
+This gives us the following theorem:
+
+\begin{theorem}
+	Suppose we are given a formula as in Lemma~\ref{lemma:propositional-normal-form}, i.e. of the form
+	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal S\right)\to P$$
+	with flat $\mathcal R$ and implication $\mathcal S$.
+Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$.
+For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional Variable $X^\lambda$.
+Define $R^\#$ as the union over all sets
+	$$
+		\{(A_1^\lambda \wedge\dots\wedge A_n^\lambda)\to(B_1^\lambda \vee\dots\vee B_n^\lambda)\:|\:\lambda\in\Lambda\}
+	$$
+	for $(A_1 \wedge\dots\wedge A_n)\to(B_1\vee\dots\vee B_n)\in\mathcal R$ and $\mathcal X^\#$ as the union over all sets
+	$$
+		\{({A^\lambda\psi} \to A^{\lambda\psi})\to(C^\lambda)\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\}
+	$$
+	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $P^\#$ as the union over all sets
+	$$
+		\{A^\lambda\to A^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\psi\text{ does not occur in }\lambda\}
+	$$
+	 for each propositional variable $A$ occurring in $\varphi$ encoding persistency. Then
+	$$
+		\varphi^\# := \left(\bigwedge P^\#\wedge\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\right)\to P^\epsilon
+	$$
+	is classically valid if and only if $\varphi$ is intuitionistically valid.
+\end{theorem}
 %\begin{theorem}
@@
 \begin{example}
-	Let us consider the law of excluded middle $\varphi = A\vee\neg A$. Then we get
-	$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\},$$
-	and we know that $\varphi$ is valid iff $\bigwedge \mathcal S^-(\varphi)\to P_\varphi$ is valid.
-Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically valid, where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
-	\begin{align*}
-		\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi},A^\epsilon\to P_A^\epsilon, P_A^{\psi}\to P_A^{\psi}, \}\cup\\ &\{(P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi},\}\cup\\&\{(A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}.
-	\end{align*}
-	We now note that $P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$, $P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$ indeed defines a counter-model.
+	Let us consider the law of excluded middle $\varphi = A\vee\neg A$ for which there is the equivalid normal form
+	$$((A\to P)\wedge ((A\to \bot)\to P))\to P.$$
+	Applying the above lemma this is intuitionistically valid if and only if $$((A^\epsilon\to A^\psi)\wedge(P^\epsilon\to P^\psi)\wedge(A^\epsilon\to P^\epsilon)\wedge(A^\psi\to P^\psi)\wedge ((A^\psi\to \bot)\to P^\epsilon))\to P^\epsilon$$ is classically valid where $\psi := (A\to \bot)\to P$.
+	We note that $A^\epsilon  = P^\epsilon = 0$, $A^\psi = P^{\psi}  = 1$ indeed defines a counter-model.
 \end{example}
@@
 
-Suppose we are given a formula in normal form $$\left(\bigwedge R\wedge\bigwedge X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}.
+Suppose we are given a formula $\varphi$ in normal form $$\left(\bigwedge R\wedge\bigwedge X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}.
 We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model.
 The idea how to get a polynomially-sized formula is: Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
-The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e.
+The universally quantified variables $X_\psi$ in the next definition handle the branching, i.e.
 express which path is considered.
@@
 \begin{definition}
-	For every propositional variable $A$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $A^n$ and for every formula $\psi\in\mathcal X$ consider a new atom $X_\psi^n$.
+	For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$.
 Let $\vec A^n$ range over all propositional variables $A^n$ and $\vec X^n$ range over all $X_\psi^n$.
@@
 	$$
-	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $X^0$.
+	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any choice of $X^0$.
 \end{example}
@@
 
-\begin{lemma}
+\begin{lemma}\label{proof:first-order-normal-form}
 	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R$ of the form
@@
 
-\begin{proof}\label{proof:first-order-normal-form}
+\begin{proof}
 	As in~\cite{otten2005clausal} we may assume that $\varphi$ is of the form $\psi\to P$.

### 2022-10-07 FoSSaCS/fossacs-EIICL.tex
first-order section

@@
 	For every propositional formula $\varphi$ there exists an atom $P$ as well as sets of clauses $\mathcal R, \mathcal X$, with $\mathcal R$ containing  \emph{flat clauses} of the form
-	$$(A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)$$
+	$$\bigwedge_iA_i\to\bigvee_jB_j$$
 	and $\mathcal X$ containing \emph{implication clauses} of the form
@@
 	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R,\mathcal X, \mathcal Q$, with $\mathcal R$ containing \emph{flat clauses} of the form
-	$$\forall \vec x\left((A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)\right),$$
+	$$\forall \vec x\left(\bigwedge_i A_i\to \bigvee_jB_j\right),$$
 	$\mathcal X$ containing \emph{implication clauses} of the form
@@
 \end{theorem}
-A more detailed account of Theorem~\ref{thm:reduction-first-order-short} is given in Theorem~\ref{fullFOtranslation}.
+A more detailed account of Theorem~\ref{thm:reduction-first-order-short} is given in Theorem~\ref{thm:fo-translation}.
 
@@
 \begin{itemize}
-	\item For flat clauses $((A_1\wedge\dots\wedge A_n)\to(B_1\vee\dots\vee B_m))^u$ is
-	$$\forall w(u\preceq w\to(A_1(u)\wedge\dots\wedge A_n(u))\to(B_1(u)\vee\dots\vee B_m(u))).$$
+	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^u$ is
+	$$\forall w\left(u\preceq w\to\bigwedge_i A_i(u)\to\bigvee_jB_j(u)\right).$$
 	\item For implication clauses $((A\to B)\to C)^u$ is
@@
 
-and after applying Herbrandization we end up with a formula $$\varphi^{CH} = K(\varphi)\to r\preceq s\to\left(\bigwedge\mathcal R^{(u)S}\wedge \bigwedge\mathcal X^{(u)S}\right)\to P(s)$$ where $s$ is a new constant symbol, each formula in $\mathcal R^{(u)S}$ is of the form
-$$\forall w(s\preceq w\to (A_1(u)\wedge\dots\wedge A_n(u))\to(B_1(u)\vee\dots\vee B_m(u)))$$
-and each formula $\psi\in\mathcal X^{(u)S}$ is of the form
+and after applying Herbrandization we end up with a formula $$\varphi^{CH} = K(\varphi)\to r\preceq s\to\left(\bigwedge\mathcal R^{uS}\wedge \bigwedge\mathcal X^{uS}\right)\to P(s)$$ where $s$ is a new constant symbol, each formula in $\mathcal R^{uS}$ is of the form
+$$\forall w\left(s\preceq w\to\bigwedge_iA_i(u)\to\bigvee_jB_j(u)\right)$$
+and each formula $\psi\in\mathcal X^{uS}$ is of the form
 $$\forall w(s\preceq w\to(w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w)))\to C(w))$$
@@
 \begin{definition}
-For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{(u)S}$. Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
+For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{uS}$. Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
 We say $(A\to B)\to C\in\mathcal X$ is \emph{fulfilled} at $m\in M$ iff $A^I(m)\to B^I(m)$ is false or $C^I(m)$ is true. For $\psi\in\mathcal X$ define $$g_\psi : M\to M, m\mapsto\begin{cases}
@@
 		\item $M_T$ is the set of sequences without repetition on $\mathcal X$.
+		\item $r^{I_T} = s^{I_T} = \epsilon$.
 		\item Interpret $\preceq$ as the prefix-order.
@@
 	It is well known that Kripke frames which are trees are complete with regard to intuitionistic logic.
-Our contribution is giving an explicit size and geometry that depends only on the structure of the considered formula.
+Our contribution is giving an easy to compute explicit size and geometry that depends only on the structure of the considered formula.
 \end{remark}
@@
 	$$
-		\{(A_1^\lambda \wedge\dots\wedge A_n^\lambda)\to(B_1^\lambda \vee\dots\vee B_n^\lambda)\:|\:\lambda\in\Lambda\}
-	$$
-	for $(A_1 \wedge\dots\wedge A_n)\to(B_1\vee\dots\vee B_n)\in\mathcal R$ and $\mathcal X^\#$ as the union over all sets
-	$$
-		\{({A^\lambda\psi} \to A^{\lambda\psi})\to(C^\lambda)\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\}
+		\left\{\bigwedge_iA_i^\lambda\to\bigvee_jB_j^\lambda\:|\:\lambda\in\Lambda\right\}
+	$$
+	for $\bigwedge_iA_i\to\bigvee_jB_j\in\mathcal R$ and $\mathcal X^\#$ as the union over all sets
+	$$
+		\left\{\left({A^\lambda\psi} \to A^{\lambda\psi}\right)\to(C^\lambda)\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
 	$$
@@
 	$$
-		\{A^\lambda\to A^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\psi\text{ does not occur in }\lambda\}
+		\left\{A^\lambda\to A^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\psi\text{ does not occur in }\lambda\right\}
 	$$
@@
 	\begin{align*}
-		\text{\normalfont Valid}(n) = &\left(\bigvee \{X^n_\psi\:|\:\psi\in\mathcal X\}\right)\wedge\left(\bigwedge \{\neg(X^n_{\psi_1}\wedge X^n_{\psi_2})\:|\:\psi_1\neq\psi_2\in\mathcal X\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{X^i_\psi\to\neg X^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal X\right\}\right)\\
-		\text{\normalfont Persistent}(n) = & \bigwedge\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\}\\
+		\text{\normalfont Valid}(n) = &\left(\bigvee\left\{X^n_\psi\:|\:\psi\in\mathcal X\right\}\right)\wedge\left(\bigwedge \left\{\neg(X^n_{\psi_1}\wedge X^n_{\psi_2})\:|\:\psi_1\neq\psi_2\in\mathcal X\right\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{X^i_\psi\to\neg X^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal X\right\}\right)\\
+		\text{\normalfont Persistent}(n) = & \bigwedge\left\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\right\}\\
 		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal R\right\}\wedge\\
 		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal R\right\}\\
-		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\}
+		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\left\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\right\}
 	\end{align*}
@@
 We fix some signature $\Sigma$ for this section.
-To encode that some $n$-ary relation $A$ holds at a world $u$ we replace each $n$-ary relation symbol $A$ with a $n+1$-ary relation symbol $A^\#$, interpreting $A^\#(\vec x, u)$ as "$A(\vec x)$ holds at u".
+To encode that some $n$-ary relation $A$ holds at a world $u$ we replace each $n$-ary relation symbol $A$ with a $n+1$-ary relation symbol $A$, interpreting $A(\vec x, u)$ as "$A(\vec x)$ holds at u".
 Furthermore we write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$.
-
-First of all recall, that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
-\[\begin{matrix}
-	\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c)))\indent \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c))\indent \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c)))\\
-	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c))\indent \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\indent\forall \vec z(\forall xA(\vec a)\to B(\vec b))\\ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b)).
-\end{matrix}\]
-The complete result that we will show is as follows:
-\begin{theorem}\label{fullFOtranslation}
-	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain the formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$.
-Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
-For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$.
-For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$.
-Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
-Obtain $\mathcal S^\#$ by including formulas as follows:
-	\begin{itemize}
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
-		\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\psi = \forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
-	\end{itemize}
-	Then $\bigwedge S\to P$ is intuitionistically iff $\exists xE(x, s)\to K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol, $\Sigma$ is the original signature and
-	\begin{align*}
-		K := &\:\bigwedge\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F\}\wedge\\
-		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
-		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}.
-	\end{align*}
-	The size of the obtained formula is linear in the input.
-However $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
-\end{theorem}
-
-It is not possible to completely eliminate worlds from the domain as in the propositional case, since counter-models can possibly be infinite (as we will see in more detail later).
-
-\subsection{Encoding Kripke Semantics}
-
-As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$, but also $E$ has to be axiomatized. For that we use the following predicates:
-\begin{itemize}
-	\item PartialOrder($\preceq$) encoding that $\preceq$ is a partial order.
-	\item DomainSubset$(u, w)$ encoding that the domain at $u$ is a subset of that at $w$.
-	\item World($u$) encoding that $u$ represents a world.
-	\item DomainClosed($u$) which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
-	\item Persistent$(u, w)$ which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$ at $u$ we have $A(x)$ at $w$.
-\end{itemize}
-Then, analogously to the propositional case, we define:
+With this we obtain the following encoding of Kripke Semantics:
+
+\begin{definition}
+	Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively:
+	\begin{itemize}
+		\item $\bot^u = \bot$.
+		\item $A(\vec x)^{u} = A(\vec x, u)$ for every relational symbol $A$.
+		\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
+		\item $(\varphi\to \psi)^u = \forall w(u\preceq w\to\varphi^{w}\to\psi^{w})$
+		\item $(\exists x\varphi)^u = \exists x(E(x,u)\wedge\varphi^u)$
+		\item $(\forall x\varphi)^u = \forall w(u\preceq w\to E(x, w)\to \varphi^w)$
+	\end{itemize}
+	where $w$ is some new variable. Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
 	\begin{align*}
@@
 	\end{align*}
-Example encodings of these formulas are as follows:
+	with e.g.:
 	\begin{align*}
 		\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
-	\text{DomainSubset}(u, w) = &\:\forall z(E(z, u)\to E(z, w))\\
+		\text{DomainSubset}(u, w) = &\:\forall x(E(x, u)\to E(x, w))\\
 		\text{World}(u) = &\:\exists xE(x, u)\\
-	\text{DomainClosed}(u) = &\:\bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f\in\Sigma$ is a function symbol}\}\\
-	\text{Persistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
+		\text{DomainClosed}(u) = &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
+		\text{Persistent}(u, w) = &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a relation in $\varphi$}\right\}
 	\end{align*}
-
-Using $E$ we then obtain $\mathcal S'$ analogously to the predicate case by including formulas:
-
-\begin{itemize}
-	\item $\forall \vec z\forall u(s\preceq u\to \vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
-	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
-	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to\forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$\\ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
-	\item  $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
-	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
-	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
-	\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
-\end{itemize}
-
-Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid if and only if
-$$\varphi' = \text{World}(s)\to K(\varphi)\to \bigwedge S'\to P^\#(s)$$
-is classically valid, where $s$ is a new constant symbol.
-This was obvious in the propositional case but here it is more nuanced.
-A proof is given in the appendix as Lemma~\ref{lemma:fo-simplification}.
-
-\subsection{Reducing the encoding}
-
-As in the predicate case we now show the existence of a canonical counter-model (in case such a counter-model exists).
-Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$.
-Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
-For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$.
-As before we may assume that $s$ is a least element and we can avoid the quantification over $w$ by considering the formulas
-\begin{itemize}
-	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$,
-	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$,
-\end{itemize}
-and by removing removing the $s\preceq u$ for the remaining formulas in $\mathcal S'$.
-We denote the resulting set of formulas by $\mathcal S^\#$.
-Then, the previous formulas are equivalid to
-$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#(s).$$
+	Then define
+	$$\varphi^C = \text{World}(r)\to K(\varphi)\to \varphi^r$$
+	where $r$ is a new constant symbol.
+\end{definition}
+
+\begin{lemma}\label{lemma:fo-simplification}
+	$\varphi$ is intuitionistically valid if and only if $\varphi^C$ is classically valid.
+\end{lemma}
+
+Since we are mixing worlds and proper intuitionistic domain elements in a single classical domain there are some details to be taken care of.
+A proof is given in the appendix as Lemma~\ref{proof:fo-simplification}.
+
+Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulas $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = \text{World}(s)\to K(\varphi)\to r\preceq s\to\left(\bigwedge R^{uS}\wedge \bigwedge X^{uS}\wedge\bigwedge Q^{uS}\right)\to P(s)$$
+where $s$ is a new constant symbol, each formula in $\mathcal R^{uS}$ is of the form
+$$\forall w \forall \vec x(s\preceq w\to\vec E(x, w)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)),$$
+each formula $\psi\in\mathcal X^{uS}$ is of the form
+$$
+	\resizebox{\hsize}{!}{$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to (w\preceq f_\psi(\vec x, w)\to A(\vec a, f_\psi(\vec x, w))\to B(\vec b, f_\psi(\vec x, w)))\to C(\vec c, w))$}
+$$
+and each formula $\psi\in\mathcal Q^{uS}$ is of the form
+$$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to \forall y(E(y, f_\psi(\vec x, w))\to A(\vec a, f_\psi(\vec x, w)))\to B(\vec B, w))$$
+with $f_\psi$ being a new function symbol.
+
 As in the previous section we can define a tree counter-model for this.
-
 \begin{definition}
-	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
+	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
 	\begin{itemize}
 		\item $M_T$ are the sequences on $\{(\vec z, \psi)\:|\:\psi\in \mathcal F, \vec z\in M^n\}$.
+		\item $r^{I_T} = s^{I_T} = \epsilon$.
 		\item We interpret $\preceq$ as the prefix-order.
 		\item We set $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi).$$
-		\item For relation symbols $R^\#$, we set $${R^\#}^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = {R^\#}^I(\vec z, f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots)).$$
+		\item For relation symbols $R$, we set $${R}^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = {R}^I(\vec z, f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots)).$$
 	\end{itemize}
@@
 \begin{lemma}
-	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
+	If $\mathcal M$ is a counter-model to $\varphi^{CH}$ then so is $\mathcal M_T$.
 \end{lemma}
@@
 Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$f_{\psi_1}(\vec z_1, f_{\psi_2}(\vec z_2, \dots f_{\psi_n}(\vec z_n, s)\dots))$$ is sufficient.
-In particular, this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
+In particular, this allows us to completely eliminate $\preceq$, that is we remove the PartialOrder$(\preceq)$
 from $K(\varphi)$ and adjust DomainSubset$(u, w)$ and Persistent$(u, w)$, e.g. remove quantification over $w$ and the precondition $u\preceq w$ from $K(\varphi)$ and replace DomainSubset$(u, w)$ with
-$$\bigwedge\{\forall \vec z\forall u\forall x(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$$ and Persistent$(u, w)$ with $$\bigwedge\{\forall\vec z_1\forall\vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\psi\in\mathcal F_\to, \text{$A\in\Sigma$ is a relation symbol}\}.$$ We set
-
-$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#.$$
-Then, we obtain the following result:
-\begin{theorem}
-	\label{thm:fo-reduction}
-	$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
+$$
+	\bigwedge\{\forall u\forall x\forall\vec y(E(x, u)\to E(x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\}
+$$
+and Persistent$(u, w)$ with
+$$
+	\bigwedge\{\forall\vec x\forall\vec y\forall u(A(\vec x, u)\to A(\vec x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q, \text{$A$ a relation in $\varphi$}\}.
+$$Putting all of this together yields the following translation:
+\begin{theorem}\label{thm:fo-translation}
+	Let $\mathcal R,\mathcal X,\mathcal Q$ be as above.
+For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R$.
+For every $\psi\in\mathcal \mathcal X\cup\mathcal Q$ define a new function symbol $f_\psi$.
+Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
+Obtain $\mathcal R^\#$ by including
+	$$
+		\forall u\forall \vec x(\vec E(\vec x, u)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u))
+	$$
+ 	for each flat clause 
+ 	$$
+ 		\forall \vec x(\bigwedge_i A_i(\vec a_i)\to\bigvee_j B_j(\vec b_j))\in\mathcal R,
+ 	$$
+ 	$\mathcal X$ by including
+ 	$$
+ 		\forall u\forall \vec x((A(\vec a, f_\psi(\vec z, u))\to B(\vec b, f_\psi(\vec z, u))\to C(\vec c, u)))
+ 	$$
+ 	for each implication clause 
+	$$
+		\psi = \forall \vec x((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal X,
+	$$
+	$\mathcal Q^\#$ by including
+	$$
+		\forall u\forall \vec x(\forall y(A(\vec a, f_\psi(\vec z, u)))\to B(\vec b, u))
+	$$
+	for each implication clause 
+	$$
+		\psi = \forall \vec x(\forall y(A(\vec a))\to B(\vec b))\in\mathcal X,
+	$$
+
+	Then $\varphi$ is intuitionistically if and only if
+	$$
+		\varphi^\# := \exists xE(x, s)\to K(\varphi)\to \left(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge\mathcal Q^\#\right)\to P^\#(s)
+	$$
+	is classically valid, where $s$ is a new constant symbol and
+	\begin{align*}
+		K(\varphi) := &\:\bigwedge\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F\}\wedge\\
+		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ a function in $\varphi$}\}\right)\wedge\\
+		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ a relation in $\varphi$}\}.
+	\end{align*}
+	The size of the obtained formula is linear in the input.
+However $|\mathcal X| + |\mathcal Q| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
 \end{theorem}
-The complete translation is summarised in Theorem~\ref{fullFOtranslation} (stated earlier). Let us understand why a further reduction as in the propositional case is impossible.
+
+It is not possible to completely eliminate worlds from the domain by enumerating them since there are formulas with no finite counter-examples. An analogous procedure to the propositional case fails:
 
@@
 \label{def:fo-fulfilled}
-	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
-For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
-	For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
+	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
+For $\psi = \forall\vec z\psi\in\mathcal \mathcal X$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_{[\psi]}^I(\vec a, u)\to B_{[\psi]}^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
+	For $\psi = \forall\vec z\psi\in\mathcal Q$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
 \end{definition}
@@
 	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R$ of the form
-	$$\forall \vec x\left((A_1\wedge\dots\wedge A_n)\to (B_1\vee\dots\vee B_n)\right)$$
+	$$\forall \vec x\left(\bigwedge_i A_i\to \bigvee_j B_j\right)$$
 	called \emph{flat} clauses and $\mathcal X$ of the form
@@
 
-\begin{lemma}\label{lemma:fo-simplification}
+\begin{lemma}\label{proof:fo-simplification}
 	$\varphi$ is intuitionistically valid if and only of $\varphi'$ is classically valid.

### 2022-10-10 FoSSaCS/fossacs-EIICL.tex
Update fossacs-EIICL.tex

@@
 	\begin{itemize}
-		\item $v\not\models \bot$ and $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
+		\item $v\not\models \bot$, $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
 		\item $v\models \varphi\wedge\psi$ iff $v\models\varphi$ and $v\models\psi$.
@@
 	\begin{itemize}
-		\item $v_u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
+		\item $v_u\models\varphi\to \psi$ iff for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item $\mathcal M, v\not\models\bot$, $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$, $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
-		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$, $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
+		\item $\mathcal M, v\not\models\bot$
+		\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$, $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
+		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$
+		\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
 		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
@@
 \begin{definition}
-	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$} and $R^{I_w}|_{M_u} = R^{I_u}$.
-	Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
-	\begin{itemize}
-		\item $u, v\not\models\bot$, $u, v\models R(t_1,\dots,t_n)$ iff $\mathcal M_u, v\models R(t_1,\dots,t_n)$, $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
-		\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$, $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
-		\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
-		\item $u, v\models\exists x\varphi$ iff there exists $m\in M_u$ such that $u, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
-		\item $u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
-	\end{itemize}
-	$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. $\varphi$ is valid if it is satisfied by every Kripke structure.
+	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
+	The model relation between $\mathcal M_u$, $v$ (targeting $\mathcal M_u$) and formulas $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where we set
+	\begin{itemize}
+		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
+		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
+	\end{itemize}
+	We write $u, v\models\varphi$ if $\mathcal M_u, v\models \varphi$.
+$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$.
+If such $\mathcal K$ exists then $\varphi$ is satisfiable.
+$\varphi$ is valid if it is satisfied by every Kripke structure.
 	$\varphi$ is valid for some frame (write $W\models\varphi$) if it is satisfied in every Krike structure with that frame.
@@
 	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$where $A_i, B_i, A, B, C$ are atomic.
-This size of $\mathcal R$ and $\mathcal X$ is linear in the size of $\varphi$.
+The size of $\mathcal R$ and $\mathcal X$ is linear in the size of $\varphi$.
 In general they will contain additional propositional variables.
@@
 	Then define
-	$$\varphi^{C} = K(\varphi)\to \varphi^{r}$$
-	where $r$ is a new constant symbol.
+	$$\varphi^{C} = K(\varphi)\to \forall u\varphi^{u}.$$
 \end{definition}
@@
 	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^u$ is
-	$$\forall w\left(u\preceq w\to\bigwedge_i A_i(u)\to\bigvee_jB_j(u)\right).$$
+	$$\forall w\left(u\preceq w\to\bigwedge_i A_i(w)\to\bigvee_jB_j(w)\right).$$
 	\item For implication clauses $((A\to B)\to C)^u$ is
@@
 
-and after applying Herbrandization we end up with a formula $$\varphi^{CH} = K(\varphi)\to r\preceq s\to\left(\bigwedge\mathcal R^{uS}\wedge \bigwedge\mathcal X^{uS}\right)\to P(s)$$ where $s$ is a new constant symbol, each formula in $\mathcal R^{uS}$ is of the form
-$$\forall w\left(s\preceq w\to\bigwedge_iA_i(u)\to\bigvee_jB_j(u)\right)$$
+and after applying Herbrandization we end up with a formula $$\varphi^{CH} = K(\varphi)\to\left(\bigwedge\mathcal R^{uS}\wedge \bigwedge\mathcal X^{uS}\right)\to P(s)$$ where $s$ is a new constant symbol, each formula in $\mathcal R^{uS}$ is of the form
+$$\forall w\left(s\preceq w\to\bigwedge_iA_i(w)\to\bigvee_jB_j(w)\right)$$
 and each formula $\psi\in\mathcal X^{uS}$ is of the form
@@
 		\item $M_T$ is the set of sequences without repetition on $\mathcal X$.
-		\item $r^{I_T} = s^{I_T} = \epsilon$.
+		\item $s^{I_T} = \epsilon$.
 		\item Interpret $\preceq$ as the prefix-order.
@@
 Now we can replace every $\forall$-quantifier in $\varphi^{CH}$ by enumerating all the ground terms, and then replace all distinct ground instances of relations by new propositional variables yielding a propositional formula.
-Note that the encoding of partial order becomes redundant with this so we drop it, but we need to keep the persistency restraints.
+Note that the encoding of partial order becomes redundant with this so we drop it, but we need to keep the persistency constraints.
 This gives us the following theorem:
@@
 Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$.
-For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional Variable $X^\lambda$.
+For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional Variable $A^\lambda$.
 Define $R^\#$ as the union over all sets
@@
 	$$
-		\left\{\left({A^\lambda\psi} \to A^{\lambda\psi}\right)\to(C^\lambda)\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
+		\left\{\left(A^{\lambda\psi} \to A^{\lambda\psi}\right)\to(C^\lambda)\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
 	$$
@@
 
-Suppose we are given a formula $\varphi$ in normal form $$\left(\bigwedge R\wedge\bigwedge X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}.
+Suppose we are given a formula $\varphi$ in normal form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}.
 We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model.
@@
 		\text{\normalfont Persistent}(n) = & \bigwedge\left\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\right\}\\
-		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal R\right\}\wedge\\
-		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal R\right\}\\
+		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{\bigwedge A^n_i\to\bigvee B^n_j\:|\:\bigwedge_i\mathcal R\right\}\\
 		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\left\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\right\}
@@
 \begin{corollary}
-	Let $N$-\textbf{IntInvalid} be the problem of deciding if a formula $(\bigwedge \mathcal R\wedge\mathcal X)\to P$ as in Lemma~\ref{lemma:propositional-normal-form} with $|\mathcal X|\leq N$ is not intuitionistically valid.
+	Let $N$-\textbf{IntInvalid} be the problem of deciding if a formula $(\bigwedge \mathcal R\wedge\bigwedge\mathcal X)\to P$ as in Lemma~\ref{lemma:propositional-normal-form} with $|\mathcal X|\leq N$ is not intuitionistically valid.
 Then $N$-\textbf{IntInvalid} is in $\Sigma_{2N-1}^P$.

### 2022-10-10 Implementation/benchmark_nanocop.py
added nanocop benchmark

@@ -2,56 +2,47 @@ import subprocess
 import tqdm
 import os
 
-VAMPIRE_PATH = "/home/lambdax/.vampire/vampire"
-TRANSLATOR = "./eiicl/target/debug/eiicl"
+NANO_PATH = "/home/lambdax/Implementation/provers/nanoCoP-i20/nanocopi.sh"
 
-INCLUDE_ROOT = "./ILTP-v1.1.2-firstorder/"
 PROBLEM_ROOT = "./ILTP-v1.1.2-firstorder/Problems/"
-TRANSLATION_ROOT = "./ILTP-v1.1.2-firstorder-translated/"
-RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results/"
+RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results-nanocop/"
 
 total_proven = 0
 total_disproven = 0
 total_unresolved = 0
 total_total = 0
 
-if not os.path.isdir(TRANSLATION_ROOT):
-    os.mkdir(TRANSLATION_ROOT)
 if not os.path.isdir(RESULT_ROOT):
     os.mkdir(RESULT_ROOT)
 for directory in os.listdir(PROBLEM_ROOT):
     proven = 0
     disproven = 0
     unresolved = 0
-    if not os.path.isdir(os.path.join(TRANSLATION_ROOT, directory)):
-        os.mkdir(os.path.join(TRANSLATION_ROOT, directory))
     if not os.path.isdir(os.path.join(RESULT_ROOT, directory)):
         os.mkdir(os.path.join(RESULT_ROOT, directory))
     for translation in tqdm.tqdm(os.listdir(os.path.join(PROBLEM_ROOT, directory)), desc=f"{directory}"):
         problem_path = os.path.join(os.path.join(
             PROBLEM_ROOT, directory), translation)
-        translation_path = os.path.join(os.path.join(
-            TRANSLATION_ROOT, directory), translation)
-        with open(translation_path, "w") as output_file:
-            subprocess.run([TRANSLATOR, problem_path,
-                           INCLUDE_ROOT], stdout=output_file)
         result_path = os.path.join(os.path.join(
             RESULT_ROOT, directory), translation)
         with open(result_path, "w") as output_file:
-            subprocess.run([VAMPIRE_PATH, translation_path,
-                           "--time_limit", "10"], stdout=output_file)
+            try:
+                subprocess.run([NANO_PATH, problem_path], stdout=output_file)
+            except subprocess.TimeoutExpired:
[…]

### 2022-10-10 FoSSaCS/fossacs-EIICL.tex
Update fossacs-EIICL.tex

@@
 	\end{itemize}
-	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
+	Let $$\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$$ and $$\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
 \end{definition}
@@
 	\begin{align*}
-		K(\varphi) := &\:\bigwedge\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F\}\wedge\\
+		K(\varphi) := &\:\bigwedge\left\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\right\}\wedge\\
 		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ a function in $\varphi$}\}\right)\wedge\\
-		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ a relation in $\varphi$}\}.
+		&\resizebox{.9\hsize}{!}{$\:\bigwedge\left\{\forall\vec z_1\forall \vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal X\cup\mathcal Q$, $A$ a relation in $\varphi$}\right\}$.}
 	\end{align*}
@@
 
-It is not possible to completely eliminate worlds from the domain by enumerating them since there are formulas with no finite counter-examples. An analogous procedure to the propositional case fails:
-
-\begin{definition}
-\label{def:fo-fulfilled}
-	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
-For $\psi = \forall\vec z\psi\in\mathcal \mathcal X$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_{[\psi]}^I(\vec a, u)\to B_{[\psi]}^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
-	For $\psi = \forall\vec z\psi\in\mathcal Q$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
-\end{definition}
-
-We get an analogous result to Lemma~\ref{thm:prop-countermodel-reduction}:
-
-\begin{lemma}
-\label{thm:fo-countermodel-reduction}
-	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
-	\begin{enumerate}
-		\item $\psi$ is fulfilled at $f_\psi^I(\vec a, u)$ with $\vec a$ for all $u\in M, \psi\in\mathcal F$, $\vec a\in M^n$.
-		\item If $\psi\in\mathcal F_\to$ is fulfilled at $u\in M$ with $\vec a$ then $\psi$ is fulfilled at all $w\geq u$ with $\vec a$.
-	\end{enumerate}	
-\end{lemma}
-
-However part 2 of Lemma~\ref{thm:fo-countermodel-reduction} fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$), we have $A(x, v)$, there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
-This is why we must consider sequences that have repetitions:
-
-\begin{wrapfigure}[16]{r}{0.4\textwidth}
-	\vspace{-32pt}
-	\begin{center}
-		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
-			\node[squarednode]      (0)   [] {$ $};
-			\node[squarednode]      (1r)   [above right=of 0] {{\color{blue}$1$}};
-			\node[squarednode]      (1l)   [above left=of 0] {{\color{red}$1$}};
-			\node[squarednode]      (2r)   [above=of 1r] {{1, \color{blue}$2$}};
-			\node[squarednode]      (2l)   [above=of 1l] {{1, \color{red}$2$}};
-			\node[squarednode]      (3r)   [above=of 2r] {{1, 2, \color{blue}$3$}};
-			\node[squarednode]      (3l)   [above=of 2l] {{1, 2, \color{red}$3$}};
-			\node      (4r)   [above=of 3r] {{$\vdots$}};
-			\node      (4l)   [above=of 3l] {{$\vdots$}};
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (0.north east) -- (1r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (0.north west) -- (1l.south east);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1r.north) -- (2r.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1l.north) -- (2l.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1l.north east) -- (2r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1r.north west) -- (2l.south east);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2r.north) -- (3r.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2l.north) -- (3l.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2l.north east) -- (3r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2r.north west) -- (3l.south east);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3r.north) -- (4r.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3l.north) -- (4l.south);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3l.north east) -- (4r.south west);
-			\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3r.north west) -- (4l.south east);
-		\end{tikzpicture}
-	\end{center}
-	\caption{Counter-model to $\varphi$.}
-	\label{fig1}
-\end{wrapfigure}
-\begin{example}
-	Consider $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model as shown in Figure~\ref{fig1},
-	where we indicate at each world the elements that exist at it.
-For black numbers $n$, both $A(n)$ and $B(n)$ hold at that world, for red ones only $A(n)$, for blue ones only $B(n)$.
-The arrows indicate the order.
-$\neg\forall xA(x)$ is not fulfilled at worlds in the left column and $\neg\forall xB(x)$ is not fulfilled at worlds in the right column.
-In particular applying $f_{\neg \forall xA(x)}$ we must always reach the right column and applying $f_{\neg \forall xB(x)}$ we must always reach the left column.
-Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we cannot remain stationary.
-\end{example}
-
-Let us add that Lemma~\ref{thm:fo-countermodel-reduction} still equips us with additional information that could be used to guide proof search: We could add additional axioms that restrict the interpretation of functions.
-That is, we may assume that $f_\psi(w) = w$, if $\exists u(w = f_\psi(u))$ for all $\psi\in\mathcal F_\forall$ (i.e.
-that $f_\psi$ is idempotent), and $f_\psi(w) = w$, if $\exists u(w \geq f_\psi(u))$ for all $\psi\in\mathcal F_\to$ (which is even stronger than idempotence).
-How adding such axioms impacts proof search will need to be evaluated in experiments.
+It is not possible to completely eliminate worlds from the domain by enumerating them since there are formulas with no finite counter-examples.
+An analogous procedure to the propositional case fails, however we can still retrieve additional axioms that restrict the interpretation of functions.
+That is, we may assume that $f_\psi(\vec t, w) = w$, if $\exists u(w = f_\psi(\vec t, u))$ for all $\psi\in\mathcal X\cup\mathcal Q$ (i.e.
+that $f_\psi$ parametrized by $\vec t$ is idempotent), and $f_\psi(\vec t, w) = w$, if $\exists u(w \geq f_\psi(\vec t, u))$ for all $\psi\in\mathcal X$ (which is even stronger than idempotence).
 

### 2022-10-10 Implementation/benchmark_nanocop.py
strong idempotence

@@ -27,15 +27,26 @@ for directory in os.listdir(PROBLEM_ROOT):
             RESULT_ROOT, directory), translation)
         with open(result_path, "w") as output_file:
             try:
+<<<<<<< Updated upstream
                 subprocess.run([NANO_PATH, problem_path], stdout=output_file)
             except subprocess.TimeoutExpired:
                 subprocess.run(["killall", "swipl"])
+=======
+                subprocess.run([NANO_PATH, problem_path],
+                           timeout=10,
+                           stdout=output_file)
+            except subprocess.TimeoutExpired:
+>>>>>>> Stashed changes
                 print("timeout", file=output_file)
         with open(result_path, "r") as result:
             result_text = result.read()
         with open(problem_path, "r") as result:
             problem_text = result.read()    
+<<<<<<< Updated upstream
         if "intuitionistic Theorem" in result_text:
+=======
+        if "is an intuitionistic Theorem" in result_text:
+>>>>>>> Stashed changes
             proven += 1
             assert "Status (intuit.) : Non-Theorem" not in problem_text
         elif "timeout" in result_text:

### 2022-10-10 Implementation/eiicl/src/encoding.rs

@@ -11,10 +11,13 @@ pub struct Encoding<'a> {
     axioms: Vec<normalized_formula::NormalizedFormula<'a>>,
     goal: Option<normalized_formula::AtomizedSubformula>,
     variables: HashMap<String, Vec<String>>,
-    pub f: HashMap<String, Vec<String>>,
+    pub f_implication: HashMap<String, Vec<String>>,
+    f_pred: HashMap<String, Vec<String>>,
+    pub f_forall: HashMap<String, Vec<String>>,
     constants: Vec<String>,
     function_symbols: HashMap<String, Vec<String>>,
     relation_symbols: HashMap<String, Vec<String>>,
+    new_relation_symbols: HashMap<String, Vec<String>>,
     polarity: bool,
 }
 
@@ -24,10 +27,13 @@ impl<'a> Encoding<'a> {
             axioms: vec![],
             goal: None,
             variables: HashMap::new(),
-            f: HashMap::new(),
+            f_implication: HashMap::new(),
+            f_pred: HashMap::new(),
+            f_forall: HashMap::new(),
             constants: Vec::new(),
             function_symbols: HashMap::new(),
             relation_symbols: HashMap::new(),
+            new_relation_symbols: HashMap::new(),
             polarity: true,
         }
     }
@@ -41,9 +47,179 @@ impl<'a> Encoding<'a> {
         return encoding;
     }
 
-    pub fn to_tptp(&self) -> Vec<top::FofAnnotated> {
+    pub fn to_tptp(&mut self) -> Vec<top::FofAnnotated> {
         let mut result = vec![];
-        for (f, args) in self.f.iter() {
+        for (f, args) in self.f_implication.iter() {
+            result.push(top::FofAnnotated(top::Annotated {
+                name: common::Name::AtomicWord(common::AtomicWord::Lower(common::LowerWord("a"))),
+                role: top::FormulaRole(common::LowerWord("axiom")),
+                formula: Box::new(fof::Formula(fof::LogicFormula::Binary(
+                    fof::BinaryFormula::Nonassoc(fof::BinaryNonassoc {
+                        left: Box::new(fof::UnitFormula::Unitary(fof::UnitaryFormula::Atomic(
+                            Box::new(fof::AtomicFormula::Plain(fof::PlainAtomicFormula(
+                                fof::PlainTerm::Function(
+                                    common::Functor(common::AtomicWord::Lower(common::LowerWord(
+                                        "exists",
+                                    ))),
+                                    Box::new(fof::Arguments(vec![
[…]

### 2022-10-10 Implementation/benchmark_nanocop.py
Update benchmark_nanocop.py

@@ -27,26 +27,15 @@ for directory in os.listdir(PROBLEM_ROOT):
             RESULT_ROOT, directory), translation)
         with open(result_path, "w") as output_file:
             try:
-<<<<<<< Updated upstream
                 subprocess.run([NANO_PATH, problem_path], stdout=output_file)
             except subprocess.TimeoutExpired:
                 subprocess.run(["killall", "swipl"])
-=======
-                subprocess.run([NANO_PATH, problem_path],
-                           timeout=10,
-                           stdout=output_file)
-            except subprocess.TimeoutExpired:
->>>>>>> Stashed changes
                 print("timeout", file=output_file)
         with open(result_path, "r") as result:
             result_text = result.read()
         with open(problem_path, "r") as result:
             problem_text = result.read()    
-<<<<<<< Updated upstream
-        if "intuitionistic Theorem" in result_text:
-=======
         if "is an intuitionistic Theorem" in result_text:
->>>>>>> Stashed changes
             proven += 1
             assert "Status (intuit.) : Non-Theorem" not in problem_text
         elif "timeout" in result_text:

### 2022-10-11 FoSSaCS/fossacs-EIICL.tex
updated summary, introduction

@@
 We then transfer this procedure to first-order logic which allows us to give a embedding of intuitionistic first-order logic into classical first-order-logic.
+This allows a direct application of state-of-the-art classical first-order provers to the problem of determining intuitionistic validity.
 
-%The key notion used in the proofs is the existence of counter-models - an analogon to the satisfiability of the negation in classical logic, which is not equivalent to invalidity in the intuitionistic case.
-We hope that this will allow leveraging the tremendous recent progress in automated reasoning in classical logic for checking intuitionistic validity, for which there has not been much progress in recent decades.
-
-This allows a direct application of state-of-the-art classical first-order provers to the problem of determining intuitionistic validity.
-A first benchmark shows that our approach is competitive.
+A first benchmark shows that under the right circumstances this is a valid approach.
 Crucially, our constructions support the transfer of counter-models to validity, which is important to model checking applications, where the output of counter-examples is an appreciated feature.
@@
 
-The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmark problems (see e.g.
+The interest in intuitionistic logic has lead to the development of a number of automated theorem proving systems and a collection of benchmark problems (see e.g.
 the ILTP library website~\cite{iltp}).
@@
 in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
-A further advantage of classical logic is the existence of calculi that are particularly suitable for automation such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability, i.e.
+A further advantage of classical logic is the existence of calculi that are particularly suited for automation such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability, i.e.
 to show the validity of a formula it suffices to show the unsatisfiability of the negated formula, which is insufficient in intuitionistic logic.
-The first dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient,tammet1996resolution} used the naive inverse method, i.e.
-direct search for a cut-free proof by applying the rules from some proof calculus inversely, which generally leads to a very complex search.
-More recently connection-based methods have been applied to various non-classical logics~\cite{otten2005clausal,otten2021nanocop}, including intuitionistic logic, and show promise.
+The first dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient,tammet1996resolution} used the naive inverse method, a direct search for a cut-free proof by applying the rules from some proof calculus inversely, which generally leads to a very complex search.
+More recently connection-based methods have been applied to various non-classical logics~\cite{otten2005clausal,otten2021nanocop}, including intuitionistic logic.
 There have also been some successful attempts to study intuitionistic validity via embedding into higher-order classical logic~\cite{LEO}.
@@
 
-In summary, our work suggests the following approach:
+In summary, our work establishes and then examines the following approach to theorem proving for intuitionistic logic:
 \begin{itemize}
-	\item performing our translation on the input formula $\varphi$ to obtain $\varphi^\#$,
-	\item checking the classical validity of $\varphi^\#$ with the prover resulting in a proof or counter-model,
-	\item translating the proof/counter-model of $\varphi^\#$ to a (intuitionistic) proof/counter-model of $\varphi$.
+	\item Translate the input formula $\varphi$ to $\varphi^\#$, which is classically valid if and only if $\varphi$ is intuitionistically valid.
+	\item Determine the classical validity of $\varphi^\#$ using a prover for classical logic, resulting in a proof or counter-model.
+	\item Translate the classical proof/counter-model of $\varphi^\#$ to a corresponding intuitionistic proof/counter-model of $\varphi$.
 \end{itemize}
@@
 	\begin{itemize}
-		\item $v\not\models \bot$, $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
+		\item $v\not\models \bot$
+		\item $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
 		\item $v\models \varphi\wedge\psi$ iff $v\models\varphi$ and $v\models\psi$.
@@
 	\begin{itemize}
-		\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.
-		\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
+		\item to each $n$-ary function symbol $f\in\Sigma$ a $n$-ary function $f^I: M^n\to M$.
+		\item to each $n$-ary relation symbol $R\in\Sigma$ a $n$-ary relation $R^I\subseteq M^n$.
 	\end{itemize}
@@
 		\item $a^{I, v} = v(a)$ for each free variable $a$.
-		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$.
+		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$ for each $n$-ary function symbol $f\in\Sigma$.
 	\end{itemize}
@@
 	\end{itemize}
-	$\varphi$ is satisfiable if there exist $\mathcal M, v$  with $\mathcal M, v\models\varphi$.
 	A $\Sigma$-structure $\mathcal M$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$, we write $\mathcal M\models\varphi$.
@@
 $\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$.
-If such $\mathcal K$ exists then $\varphi$ is satisfiable.
 $\varphi$ is valid if it is satisfied by every Kripke structure.
@@
 \end{definition}
-In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
+In addition to propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
 
@@
 \end{theorem}
-
-Theorem~\ref{thm:herbrand-skolem} follows from Lemmata~\ref{ap1} and~\ref{ap2} stated in the appendix.
 
@@
 		\item $\psi$ is fulfilled at $f_{[\psi]}^I(u)$ for all $u\in M, \psi\in\mathcal X$.
-		\item If $\psi$ is fulfilled at some $m\in M$ then $\psi$ is fulfilled at all $n\succeq m$.
+		\item If $\psi$ is fulfilled at some $m\in M$ then $\psi$ is fulfilled at all $n\succeq^I m$.
 	\end{enumerate}
@@
 
-A proof can be found in the appendix as Lemma~\ref{proof:prop-countermodel-reduction}.
-Essentially what this tells us is that for every $m$ at which $\psi$ is fulfilled, in particular for each $m$ in the image of $f_{[\psi]}$, we may add the assumption that $f_{[\psi]}(n) = n$ for all $n\geq m$ without changing the validity of $\psi$.
+\begin{proof}
+	1. If $C^I(m)$ is true then we are done due to persistency. Otherwise $C^I(m)$ is false.
+	Because $(A^I(f_{[\psi]}^I(m))\to B^I(f_{[\psi]}^I(m)))\to C^I(u)$ holds, then $A^I(f_{[\psi]}^I(u))\to B^I(f_{[\psi]}^I(u))$ must be false.
+
+	2. Let $n\succeq^I m$.
+	If $C^I(n)$ is true, we are done.
+	Otherwise $C^I(n)$ is false.
+	Due to persistency $C^I(m)$ is also false.
+	Because $\psi$ is fulfilled at $m$, $A^I(m)\to B^I(m)$ must be false, i.e. $A^I(m)$ is true and $B^I(m)$ is false. Then $A^I(n)$ and $A^I(f^I_{[\psi]}(n))$ are also true due to persistency.
+	Because $(A^I(f^I_{[\psi]}(n))\to B^I(f^I_{[\psi]}(n)))\to C^I(n)$ holds, we now get that  $B^I(f^I_{[\psi]}(n))$ must be false.
+	But then due to persistency so is $B^I(n)$ and we have that $\psi$ is fulfilled at $n$.\hfill$\square$
+\end{proof}
+
+Essentially what this tells us is that for every $m$ at which $\psi$ is fulfilled, in particular for each $m$ in the image of $f_{[\psi]}$, we may add the assumption that $f_{[\psi]}(n) = n$ for all $n\succeq m$ without changing the validity of $\psi$.
 With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^{CH}$ is a straightforward check of definitions.
@@
 \end{corollary}
-
-An example of this translation can be found in the appendix as Example~\ref{ex:prop-tree-model}.
 
@@
 		\text{\normalfont Persistent}(n) = & \bigwedge\left\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\right\}\\
-		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{\bigwedge A^n_i\to\bigvee B^n_j\:|\:\bigwedge_i\mathcal R\right\}\\
+		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{\bigwedge A^n_i\to\bigvee B^n_j\:|\:\bigwedge A_i\to\bigvee B_j\mathcal R\right\}\\
 		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\left\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\right\}
@@
 We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$.
-We fix some signature $\Sigma$ for this section.
 To encode that some $n$-ary relation $A$ holds at a world $u$ we replace each $n$-ary relation symbol $A$ with a $n+1$-ary relation symbol $A$, interpreting $A(\vec x, u)$ as "$A(\vec x)$ holds at u".
@@
 		\item $\bot^u = \bot$.
-		\item $A(\vec x)^{u} = A(\vec x, u)$ for every relational symbol $A$.
+		\item $A(\vec x)^{u} = A(\vec x, u)$ for every relation symbol $A$.
 		\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
@@
 	Then define
-	$$\varphi^C = \text{World}(r)\to K(\varphi)\to \varphi^r$$
-	where $r$ is a new constant symbol.
+	$$\varphi^C = K(\varphi)\to\forall u(\text{World}(u)\to \varphi^u).$$
 \end{definition}
@@
 
-Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulas $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = \text{World}(s)\to K(\varphi)\to r\preceq s\to\left(\bigwedge R^{uS}\wedge \bigwedge X^{uS}\wedge\bigwedge Q^{uS}\right)\to P(s)$$
+Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulas $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = \text{World}(s)\to K(\varphi)\to\left(\bigwedge R^{uS}\wedge \bigwedge X^{uS}\wedge\bigwedge Q^{uS}\right)\to P(s)$$
 where $s$ is a new constant symbol, each formula in $\mathcal R^{uS}$ is of the form
@@
 	Let $\mathcal R,\mathcal X,\mathcal Q$ be as above.
-For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R$.
+For every $n$-ary relation $A$ define a new $n+1$-ary relation of the same name.
 For every $\psi\in\mathcal \mathcal X\cup\mathcal Q$ define a new function symbol $f_\psi$.
@@
  	$$
- 	$\mathcal X$ by including
+ 	$\mathcal X^\#$ by including
  	$$
@@
 	$$
-	$\mathcal Q^\#$ by including
+	and $\mathcal Q^\#$ by including
 	$$
@@
 	$$
-		\psi = \forall \vec x(\forall y(A(\vec a))\to B(\vec b))\in\mathcal X,
+		\psi = \forall \vec x(\forall y(A(\vec a))\to B(\vec b))\in\mathcal X.
 	$$
-
 	Then $\varphi$ is intuitionistically if and only if
@@
 	The size of the obtained formula is linear in the input.
-However $|\mathcal X| + |\mathcal Q| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
+However $|\mathcal X| + |\mathcal Q| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation has been extended to a $n+1$-ary relation.
 \end{theorem}
@@
 
+\section{Benchmark}
+
+\begin{wrapfigure}[6]{r}{0.4\textwidth}
+		\vspace*{-1.7cm}
+		\begin{tabular}{l|c|c}
+			Embedding&&\\
+			 Embedding (idem.)&557&20.9\%\\
+			ileanCoP 1.2&&\\
+			nanoCoP-i 2.0&858&32.1\%\\\hline
+			Total&2669&100\%
+		\end{tabular}
+		\caption{Solved problems.}
+		\label{fig:a}
+\end{wrapfigure}
+
+We have benchmarked our first-order translation on version 1.1.2 of the ILTP problem set~\cite{iltp} by performing our translation and then examining the generated classical problem using Vampire~\cite{kovacs2013first}.
+For comparison the state-of-the-art intuitionistic provers ileanCoP 1.2~\cite{otten2008leancop} and nanoCoP-i 2.0~\cite{otten2021nanocop} are used.
+All benchmarks were conducted on an Intel Core i5-8400 2.80GHz machine running Linux with a time limit of 10s per problem.
+Two versions of the translation were tested, one with and one without the additional axioms encoding idempotence.
+The results can be seen in Figure \ref{fig:a}.
+In general our approach falls short of the state of the art.
+However when binning for $|\mathcal Q|$ and $|\mathcal X|$ as seen in Figure~\ref{fig:b} we observe something interesting.
+\begin{figure}
+	\centering
+	\begin{tabular}{c|c|c|c|c|c|c}
+		$|\mathcal Q|$&$|\mathcal X|$&Embedding&Embedding (idem.)&ileanCoP&nanoCoP-i 2.0&Total\\\hline
+		0&0&&157&&61&178\\
+		0&1-2&&130&&96&400\\
+		0&3-5&&17&&40&105\\
+		0&6-10&&4&&7&19\\
+		0&11-20&&0&&3&34\\
+		0&21+&&7&&1&17\\
+		1-2&0&&14&&15&17\\
+		1-2&1-2&&67&&206&270\\
+		1-2&3-5&&77&&122&312\\
+		1-2&6-10&&15&&28&64\\
+		1-2&11+&&4&&6&12\\
+		3-5&0&&14&&16&16\\
+		3-5&1-2&&11&&36&49\\
+		3-5&3-5&&15&&58&159\\
+		3-5&6-10&&16&&83&199\\
+		3-5&11+&&2&&3&38\\
+		6+&0-2&&1&&50&50\\
+		6+&3+&&3&&24&605
+	\end{tabular}
+	\caption{Solved problems binned by $|\mathcal Q|$ and $|\mathcal X|$.}
+	\label{fig:b}
+\end{figure}
+Leveraging the power of Vampire, for low $|\mathcal X|$ and $|\mathcal Q|$ the embedding performs on par or even outperforms the dedicated intuitionistic provers.
+For larger values the state of the art provers have an advantage, in particular when both parameters are combined.
+Therefore our tool is already useful if it is known that $|\mathcal X|$ and $|\mathcal Q|$ are small.
+Another hopeful sign is that the inclusion of idempotence increases performance in particular for larger $|\mathcal X|$ and $|\mathcal Q|$.
+Further improvements could be achieved by introducing more axioms based on structural properties of the formula.
+Another interesting observation is that $|\mathcal X|$ and $|\mathcal Q|$ also seem to correlate with success rate for the dedicated intuitionistic provers.
+
 \section{Conclusion}
@@
 We have presented embeddings of intuitionistic into classical logic for the propositional and the predicate case.
-The transformation saw an exponential blow-up parameterized by $|\mathcal F_\to|$ in the propositional case and a complexity increase reflected by an arity-increase of one for all relations and the introduction of new function symbols in the predicate case.
-A key motivation for our work is the potential of leveraging classical provers for showing intuitionistic validity.
-%The practical feasibility of our translation is yet to be tested.
-%In particular in the first-order case it is difficult to estimate how the increased complexity of the output formula impacts proof search.
-Initial tests show promise but we plan on conducting a more thorough evaluation using state-of-the art automated theorem provers, in particular the Vampire theorem prover~\cite{kovacs2013first}.
-A key factor yet to be determined is how large the parameters $|\mathcal F_\to|$ in the propositional case and $|\mathcal F|$ in the predicate case are in practice.
+Using Kripke semantics and model reduction we have established that the complexity overhead of intuitionistic logic over classical logic can be parametrized by $|\mathcal X|$ and $|\mathcal Q|$.
+This is directly reflected in the exponential blow-up parametrized by $|\mathcal X|$ in the propositional case and more indirectly in our benchmarks in the predicate case.
 
-The above mentioned complexity considerations are direct consequences of our straightforward constructions.
-We plan to establish better bounds in future work by utilizing structural properties of the input formula, in particular by constructing smaller counter-examples $\mathcal M_T$ in case certain atoms are independent from each other.
+A key motivation for our work has been the potential of using classical provers for intuitionistic proofs. Our tests have shown that there are cases where this is already reasonable.
+However, the current complexity considerations are direct consequences of our straightforward constructions.
+We plan to establish better bounds in future work by utilizing structural properties of the input formula, in particular by constructing smaller counter-examples $\mathcal M_T$ and using redundant axioms to limit the search.
+Our addition of idempotence axioms has demonstrated that this is a promising approach.
+
 On the theoretical side, we also hope to give a new translation from QBF to IPC that improves our understanding of the relationship between intuitionistic propositional logic and the polynomial hierarchy.
-
 Finally, we plan to complement the counter-model translation presented in this paper by a proof translation.
@@
 
-\section{Omitted content}
-
-\begin{lemma}\label{ap1}
-Let $\chi$ be a formula and let $\{z_1\dots z_n\} = Z$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
-	\begin{itemize}
-		\item if $M, v \models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
-		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
-	\end{itemize}
-\end{lemma}
-
-\begin{proof}
-	We proceed by simultaneous induction on the formula height.
-
-	For atomic $\chi$ the claims are clear.
-
-	Otherwise there are $3$ cases:
-
-	1.
-$\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$.
-By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v})$ and $\mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as stated in the lemma.
-Then we can define $I_{\chi, v}$ as follows:
-
-	For every symbol $p$ that occurs in $\varphi^S_Z$ or $\varphi^H_Z$ but not $\psi^S_Z, \psi^H_Z$, we set $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$.
-For every symbol $p$ that occurs in $\psi^S_Z$ or $\psi^H_Z$ but not $\varphi^S_Z, \varphi^H_Z$, we set $p^{I_{\chi, v}} = p^{I_{\psi, v}}$.
-Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.
-
-	Let $\circ=\ \to$.
-Suppose $\mathcal M, v\models\chi$.
-Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_Z$ or $\mathcal M_\chi\models\psi^H_Z$, i.e.
-$\mathcal M_\chi\models\chi^S_Z$.
-On the other hand suppose $\mathcal M, v\not\models\chi$.
-Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_Z$ and $\mathcal M_\chi, v\not\models\psi^S_Z$, i.e.
-$\mathcal M_\chi\not\models\chi^H_Z$.
-Analogous arguments work for $\circ\in\{\wedge, \vee\}$.
-
-	2.
-$\chi = \forall x\varphi$.
-Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/z_1\dots x_n/z_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise.
-Since $Z$ contains all free variables occurring in $\chi$ this is a well-defined function.
-
-By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
-Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{Z\cup a}$ or $\varphi[a/x]^H_{Z\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
-
-Suppose $\mathcal M, v\models\chi$.
-Then, for all $m\in M$ we have that $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{Z\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{Z\cup\{a\}}[x/a])$, i.e.
-$\mathcal M_\chi,v\models \chi^S_Z$.
-On the other hand suppose $\mathcal M, v\not\models\chi$.
-Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{Z\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(z_1,\dots z_n)/x]^H_Z$, i.e.
-$\mathcal M_\chi, v\not\models(\forall x\varphi)^H_Z$.
-
-	3.
-$\chi = \exists x\varphi$.
-The argument runs dually to 2.
-Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/z_1\dots x_n/z_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise.
-Since $Z$ contains all free variables occurring in $\chi$ this is a well-defined function.
-
-	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
-Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{Z\cup a}$ or $\varphi[a/x]^H_{Z\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
-
-	Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_Z$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_Z$.
-\end{proof}
-
-\begin{lemma}\label{ap2}
-	For every structure $\mathcal M$ and $\{z_1\dots z_n\} = Z$ that contains all free variables in $\chi$ and variable assignment $v$
-	\begin{itemize}
-		\item if $\mathcal M, v\models\varphi^S_Z$ then $\mathcal M, v\models \varphi$
-		\item if $\mathcal M, v\not\models\varphi^H_Z$ then $\mathcal M, v\not\models\varphi$.
-	\end{itemize}
-\end{lemma}
-
-\begin{proof}
-	Again we proceed by simultaneous induction on the formula height.
-
-	For atoms the claims are clear. We distinguish 5 cases.
-
-	1.
-$\chi = \varphi\to\psi$.
-Suppose $\mathcal M, v\models\chi^S_Z$, i.e.
-$\mathcal M, v\not\models\varphi^H_Z$ or $\mathcal M, v\models\psi^S_Z$.
-By induction hypothesis $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e.
-$\mathcal M, v\models \chi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
-$\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_Z$.
-Again by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.
-
-	2. + 3. Conjunctions and Disjunctions are dealt with analogously.
-
-	4.
-$\chi = \forall x\varphi$.
-Suppose $\mathcal M, v\models\chi^S_Z$, i.e.
-for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{Z\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$.
-Then it follows that $\mathcal M, v\models\varphi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
-there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{Z\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e.
-$\mathcal M, v\not\models\forall x\chi$.
-
-	5. $\chi = \exists x\varphi$.  The argument runs dually to 4.
-\end{proof}
-
-\begin{definition}\label{def:transf-structure}
-	For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms of the form $P_\psi$.
-By slight abuse of notation we denote with $\vec z_\psi$ elements of the domain instead of variables.
-In the classical case, we set
-	$$\begin{matrix}
-		P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)\\
-		P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))\\P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})\\P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}\indent P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\},
-	\end{matrix}$$
-	and for intuitionistic logic and each world $u$, we set
-	$$\begin{matrix}
-		P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)\\
-		P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))\\
-		P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}\\
-		P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}\indent P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}.
-	\end{matrix}$$
-\end{definition}
-
-Lemmas~\ref{thm:struct1} and~\ref{thm:struct2} then follow directly by induction on the height of $\varphi$.
+\section{Omitted proofs}
 
@@
 	where $s$ is a new function symbol, $a$ is a new free variable and $\vec z$ ranges over the free variables in $\psi$.
-	While the generated set of clauses is in general not equivalent to $\psi$, section~\ref{section:herbrandiaztion} guarantees that it is at least equivalid.
+	While the generated set of clauses is in general not equivalent to $\psi$, section~\ref{section:herbrandiaztion} guarantees that it is at least equivalid.\hfill$\square$
 \end{proof}
-
-\begin{lemma}~\label{proof:prop-countermodel-reduction}
-	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^{CH}$.
-	\begin{enumerate}
-		\item $\psi$ is fulfilled at $f_{[\psi]}^I(m)$ for all $m\in M, \psi\in\mathcal X$.
-		\item If $\psi$ is fulfilled at some $m\in M$ then $\psi$ is fulfilled at all $n\geq m$.
-	\end{enumerate}
-\end{lemma}
-
-\begin{proof}
-	1. If $C^I(m)$ is true then we are done due to persistency. Otherwise $C^I(m)$ is false.
-	Because $(A^I(f_{[\psi]}^I(m))\to B^I(f_{[\psi]}^I(m)))\to C^I(u)$ holds, then $A^I(f_{[\psi]}^I(u))\to B^I(f_{[\psi]}^I(u))$ must be false.
-
-	2. Let $n$ be some element with $n\geq m$.
-	If $C^I(n)$ is true, we are done.
-	Otherwise $C^I(n)$ is false.
-	Due to persistency $C^I(m)$ is also false.
-	Because $\psi$ is fulfilled at $m$, $A^I(m)\to B^I(m)$ must be false, i.e. $A^I(m)$ is true and $B^I(m)$ is false. Then $A^I(n)$ and $A^I(f^I_{[\psi]}(n))$ are also true due to persistency.
-	Because $(A^I(f^I_{[\psi]}(n))\to B^I(f^I_{[\psi]}(n)))\to C^I(n)$ holds, we now get that  $B^I(f^I_{[\psi]}(n))$ must be false.
-	But then due to persistency so is $B^I(n)$ and we have that $\psi$ is fulfilled at $n$.
-\end{proof}
-
-\begin{example}\label{ex:prop-tree-model}
-	This is an example for the model translation from Corollary~\ref{cor:prop-tree-model}. Consider
-	$$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model, depicted below,
-	\begin{center}
-		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
-			\node (center) {};
-			\node[squarednode]      (right)   [right=of center] {$B, C$};
-			\node[squarednode]      (left)    [left=of center] {$A, D$};
-			\node[squarednode]      (lower)       	[below=of center] {};
-			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left.south east);
-			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right.south west);			
-		\end{tikzpicture}
-	\end{center}
-	\vspace*{-.3cm}where at each node the true propositions are indicated.
-	%	
-	There is a corresponding classical counter-model $\mathcal M = (M, I)$ to $\varphi^\circ$ with $M = \{u, u_l, u_r\}$, $A^{I}(l) = D^I(l) = B^I(r) = C^I(r) = 1$ and $0$ else.
-	%
-	Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (B\to A)\to D$, this corresponds to a transformed counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below:
-	\begin{center}
-		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
-			\node (center) {};
-			\node[squarednode]      (right1)   [right=of center] {$B, C$};
-			\node[squarednode]      (right2)   [above=of right1] {$B, C$};
-			\node[squarednode]      (left1)    [left=of center] {$A, D$};
-			\node[squarednode]      (left2)    [above=of left1] {$A, D$};
-			\node[squarednode]      (lower)       	[below=of center] {};
-			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left1.south east) node [midway, below left] {$f_{\psi_1}$};
-			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right1.south west) node [midway, below right] {$f_{\psi_2}$};
-			\draw[-{Stealth[length=3mm, width=3mm]}] (left1.north) -- (left2.south) node [midway, right] {$f_{\psi_2}$};
-			\draw[-{Stealth[length=3mm, width=3mm]}] (right1.north) -- (right2.south) node [midway, left] {$f_{\psi_1}$};
-			\draw[-{Stealth[length=3mm, width=3mm]}] (left1.south west) to [out=225, in=135, loop, looseness=3] node [left] {$f_{\psi_1}$} (left1.north west);		
-			\draw[-{Stealth[length=3mm, width=3mm]}] (right1.south east) to [out=315, in=45, loop, looseness=3] node [right] {$f_{\psi_2}$} (right1.north east);
-			\draw[-{Stealth[length=3mm, width=3mm]}] (left2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (left2.north east);
-			\draw[-{Stealth[length=3mm, width=3mm]}] (right2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (right2.north east);		
-		\end{tikzpicture}
-	\end{center}
-\end{example}
 
@@
 Suppose $\varphi$ is not intuitionistically valid.
-Then, $(\bigwedge R\wedge\bigwedge X)\to P$ has an intuitionistic counter-model. By the previous section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
+Then, $(\bigwedge\mathcal R\wedge\bigwedge\mathcal X)\to P$ has an intuitionistic counter-model.
+By the previous section there exists a classical counter-model $\mathcal M$ for $(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#)\to P^\epsilon$.
 We now define a QBF model iteratively.
 For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$.
-Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$  in $\mathcal M$.
+Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal X$ such that $\psi_i$ is exactly the $\psi\in\mathcal X$ for which $X_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$  in $\mathcal M$.
 Let the $X^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified).
@@
 Continue this construction until $n  = |\mathcal X|$.
-Then from $\mathcal M$ being a counter-example to $\bigwedge\mathcal S^\#\to P^\epsilon$ it directly follows that this interpretation satisfies $\varphi^Q$.
+Then from $\mathcal M$ being a counter-example to $(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#)\to P^\epsilon$ it directly follows that this interpretation satisfies $\varphi^Q$.
 
@@
 Have $A^{\psi_0\dots\psi_n} = A^n$ for each propositional variable $A$.
-From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.
+From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.\hfill$\square$
 \end{proof}
@@
 \begin{lemma}\label{proof:fo-simplification}
-	$\varphi$ is intuitionistically valid if and only of $\varphi'$ is classically valid.
+	$\varphi$ is intuitionistically valid if and only of $\varphi^C$ is classically valid.
 \end{lemma}
@@
 	We proceed by translation of counter-examples.
-Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi'$.
+Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi^C$.
 As a Kripke frame $(W, \leq)$ take all $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ let $\leq$ be $\preceq^I$ restricted to $W$.
-Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A^\#(\vec x, u)$.
+Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A(\vec x, u)$.
 It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
@@
 		\item $s^I = w_0$.
-		\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
+		\item $E^I(m, w)$ iff $w\in W_0$ and $m = [x, w]$ for some $x\in M_w$.
 		\item $f^I(m_1\dots m_n) =\begin{cases}
-			f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
+			f^{I_u}(x_1\dots x_n), &\substack{\text{if there are $ u\in W_0, x_i\in M_u$}\\ \text{with $m_i= [x_i, u]$ for all $i$,}}\\
 			w_0, & \text{otherwise.}
 		\end{cases}$
-		\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
-			A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
+		\item ${A}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
+			A^{I_u}(x_1\dots x_n), &\substack{\text{if $u\in W_0$ and $\exists x_i\in M_u$}\\\text{with $m_i= [x_i, u]$ for all $i$,}}\\
 			\top, & \text{otherwise.}
@@
 Let us now briefly check that this indeed defines a counter-model.
-First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is.
-Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified.
-All that remains to show is $M, I\models\bigwedge\mathcal S'$.
+First of all there ${P}^I(w_0)$ is false, World$(w_0)$ holds and $K(\varphi)$ is also easily verified.
+All that remains to show is that (for each $v$) $M, I, v[w_0/u]\models\bigwedge\mathcal R^{u}\wedge\bigwedge\mathcal X^{u}$.
 Consider e.g.
-the case where $\psi\in\mathcal S$ is of the form
+the case where $\psi\in\mathcal X$, i.e.
+is of the form
 	$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
 	We have to show that
-	$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
-	holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
-	$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false.
-For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$.
-Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$.
-Note that the above is false in particular for $w\in M_0$ with $u\preceq w$.
-But this would imply
-	$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
+	$$M, I\models \forall \vec z(\vec E(\vec z, w_0)\to \forall w(w_0\preceq w\to A(\vec a, w)\to B(\vec b, w))\to C(\vec c, w_0))$$
+	holds. Suppose towards contradiction that it does not, i.e. there is $\vec z\in M$ such that for all $w\in M$
+	$$\vec E^I(\vec z, w_0)\to (w_0\preceq w\to {A}^I(\vec a, w)\to {B}^I(\vec b, w))\to {C}^I(\vec c, w_0)$$ is false.
+For that ${C}^I(\vec c, w_0)$ must be false and $\vec E^I(\vec z, w_0)$ must be true so we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_{w_0}$.
+But then by5 definition
+	$$((A^{I_{w_0}}(\vec a)\to B^{I_{w_0}}(\vec b))\to C^{I_{w_0}}(\vec c))[\vec z/z_1\dots z_n]$$ is false in our original Kripke counter-model, so it cannot be a counter-example of $\varphi$, which is a contradiction.
+The argument for $\psi\in\mathcal R$ is analogous.\hfill$\square$
 \end{proof}

### 2022-10-11 Implementation/benchmark-embedding-idemp.py

@@ -7,8 +7,8 @@ TRANSLATOR = "./eiicl/target/debug/eiicl"
 
 INCLUDE_ROOT = "./ILTP-v1.1.2-firstorder/"
 PROBLEM_ROOT = "./ILTP-v1.1.2-firstorder/Problems/"
-TRANSLATION_ROOT = "./ILTP-v1.1.2-firstorder-translated/"
-RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results/"
+TRANSLATION_ROOT = "./ILTP-v1.1.2-firstorder-translated-embedding-idemp/"
+RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results-embedding-idemp/"
 
 total_proven = 0
 total_disproven = 0

### 2022-10-11 Implementation/benchmark_leancop.py

@@ -1,0 +1,65 @@
+import subprocess
+import tqdm
+import os
+
+LEAN_PATH = "/home/lambdax/Implementation/provers/ileancop/ileancop.sh"
+TRANSLATOR = "/home/lambdax/Implementation/ILTP-v1.1.2-firstorder/TPTP2X/tptp2x"
+
+PROBLEM_ROOT = "/home/lambdax/Implementation/ILTP-v1.1.2-firstorder/Problems"
+TRANSLATION_ROOT = "/home/lambdax/Implementation/ILTP-v1.1.2-firstorder-translated-leancop"
+RESULT_ROOT = "/home/lambdax/Implementation/ILTP-v1.1.2-firstorder-results-leancop"
+
+total_proven = 0
+total_disproven = 0
+total_unresolved = 0
+total_total = 0
+
+if not os.path.isdir("ILTP-v1.1.2-firstorder-translated-leancop"):
+    os.mkdir("ILTP-v1.1.2-firstorder-translated-leancop")
+if not os.path.isdir("ILTP-v1.1.2-firstorder-results-leancop"):
+    os.mkdir("ILTP-v1.1.2-firstorder-results-leancop")
+for directory in os.listdir(PROBLEM_ROOT)[7:]:
+    proven = 0
+    disproven = 0
+    unresolved = 0
+    if not os.path.isdir(os.path.join(TRANSLATION_ROOT, directory)):
+        os.mkdir(os.path.join(TRANSLATION_ROOT, directory))
+    if not os.path.isdir(os.path.join(RESULT_ROOT, directory)):
+        os.mkdir(os.path.join(RESULT_ROOT, directory))
+    for problem in tqdm.tqdm(os.listdir(os.path.join(PROBLEM_ROOT, directory)), desc=f"{directory}"):
+        problem_path = os.path.join(os.path.join(
+            PROBLEM_ROOT, directory), problem)      
+        translation_path = os.path.join(os.path.join(
+            TRANSLATION_ROOT, directory), problem)
+        with open(translation_path, "w") as output_file:
+            subprocess.run([TRANSLATOR, "-f", "leancop", "-q2", "-d", "-", problem_path], stdout=output_file)
+        result_path = os.path.join(os.path.join(
+            RESULT_ROOT, directory), problem)
+        with open(result_path, "w") as output_file:
+            try:
+                subprocess.run([LEAN_PATH, translation_path], stdout=output_file, timeout=10)
+            except subprocess.TimeoutExpired:
[…]

### 2022-10-11 Implementation/benchmark_nanocop.py

@@ -27,7 +27,7 @@ for directory in os.listdir(PROBLEM_ROOT):
             RESULT_ROOT, directory), translation)
         with open(result_path, "w") as output_file:
             try:
-                subprocess.run([NANO_PATH, problem_path], stdout=output_file)
+                subprocess.run([NANO_PATH, problem_path], stdout=output_file, timeout=10)
             except subprocess.TimeoutExpired:
                 subprocess.run(["killall", "swipl"])
                 print("timeout", file=output_file)
@@ -37,12 +37,12 @@ for directory in os.listdir(PROBLEM_ROOT):
             problem_text = result.read()    
         if "is an intuitionistic Theorem" in result_text:
             proven += 1
-            assert "Status (intuit.) : Non-Theorem" not in problem_text
-        elif "timeout" in result_text:
-            unresolved += 1
-        else:
+            assert "Status (intuit.) : Non-Theorem" not in problem_text or "Problem negated" in problem_text
+        elif "intuitionistically Satisfiable" in result_text:
             disproven += 1
-            assert "Status (intuit.) : Theorem" not in problem_text
+            assert "Status (intuit.) : Theorem" not in problem_text or "Problem negated" in problem_text
+        else:
+            unresolved += 1
     total_proven += proven
     total_disproven += disproven
     total_unresolved += unresolved

### 2022-10-11 Implementation/bucketed_results.py

@@ -1,0 +1,45 @@
+import os
+
+import tqdm
+
+RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results"
+TRANSLATION_ROOT = "./ILTP-v1.1.2-firstorder-translated"
+PROVERS = ["-embedding-idemp", "-nanocop"]
+PROOF = {
+    "-embedding-idemp" : "Refutation found.",
+    "-nanocop" : "is an intuitionistic Theorem"
+}
+COUNTERMODEL = {
+    "-embedding-idemp" : "SZS status CounterSatisfiable",
+    "-nanocop" : "is an intuitionistic Non-Theorem"
+}
+
+buckets = [0, 2, 5, 10, 20, 1000]
+tally = {}
+
+for directory in os.listdir(TRANSLATION_ROOT):
+    for translation in tqdm.tqdm(os.listdir(os.path.join(TRANSLATION_ROOT, directory)), desc=f"{directory}"):
+        translation_path = os.path.join(os.path.join(
+            TRANSLATION_ROOT, directory), translation)
+        with open(translation_path, "r") as translation_text:
+            try:
+                translation_text = translation_text.read()
+                q = translation_text.count("a,axiom,'f_!")
+                x = (translation_text.count("a,axiom,'f") - q)//2
+            except MemoryError:
+                q = 100
+                x = 100
+        i = min([b for b in buckets if b >= q])
+        j = min([b for b in buckets if b >= x])
+        if not (i, j) in tally:
+            tally[(i, j)] = {p : 0 for p in PROVERS + ["total"]}
+        tally[(i, j)]["total"] += 1
+        for p in PROVERS:
+            root = RESULT_ROOT + p
+            result_path = os.path.join(os.path.join(root, directory), translation)
+            with open(result_path, "r") as result_text:
+                result_text = result_text.read()
+            if PROOF[p] in result_text or COUNTERMODEL[p] in result_text:
+                tally[(i, j)][p] += 1
+for k, v in tally.items():
+    print(f"{k}: {v}")

### 2022-10-11 Implementation/individual_results.py

@@ -1,38 +1,0 @@
-import os
-
-RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results/"
-
-k = 0
-i = 0
-print("\\begin{minipage}{0.45\\textwidth}\n\t\\[\\begin{matrix}\n\t\t\\text{Problem}&\\text{Result}&\\text{Time}\\\\")
-if not os.path.isdir(RESULT_ROOT):
-    os.mkdir(RESULT_ROOT)
-for directory in os.listdir(RESULT_ROOT):
-    for result in os.listdir(os.path.join(RESULT_ROOT, directory)):
-        result_path = os.path.join(os.path.join(RESULT_ROOT, directory), result)
-        with open(result_path, "r") as result:
-            result_text = result.read()
-        if "Refutation found." in result_text:
-            print(f"\t\t\\text{{{result.name.split('/')[-1]}}}&\\text{{Proven}}&{result_text.split('Time elapsed:')[1].splitlines()[0]}\\\\")
-            i += 1
-            if (k < 2 and i == 32) or (k >= 1 and i == 48):
-                print("\t\\end{matrix}\\]\n\\end{minipage}")
-                if k % 2 == 1:
-                    print("\n\\pagebreak\n")
-                k = k + 1
-                i = 0
-                print("\\begin{minipage}{0.45\\textwidth}\n\t\\[\\begin{matrix}\n\t\t\\text{Problem}&\\text{Result}&\\text{Time}\\\\")
-        if "SZS status CounterSatisfiable" in result_text:
-            print(f"\t\t\\text{{{result.name.split('/')[-1]}}}&\\text{{Refuted}}&{result_text.split('Time elapsed:')[1].splitlines()[0]}\\\\")
-            i += 1
-            if (k < 2 and i == 32) or (k >= 1 and i == 48):
-                print("\t\\end{matrix}\\]\n\\end{minipage}")
-                if k % 2 == 1:
-                    print("\n\\pagebreak\n")
-                k = k + 1
-                i = 0
-                print("\\begin{minipage}{0.45\\textwidth}\n\t\\[\\begin{matrix}\n\t\t\\text{Problem}&\\text{Result}&\\text{time}\\\\")
-        
-
-if i!=38:
-    print("\t\\end{matrix}\\]\n\\end{minipage}")

### 2022-10-11 Implementation/results-embedding-idemp.py

@@ -1,0 +1,35 @@
+import os
+
+RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results-embedding-idemp/"
+
+total_proven = 0
+total_disproven = 0
+total_unresolved = 0
+total_total = 0
+
+if not os.path.isdir(RESULT_ROOT):
+    os.mkdir(RESULT_ROOT)
+for directory in os.listdir(RESULT_ROOT):
+    proven = 0
+    disproven = 0
+    unresolved = 0
+    for translation in os.listdir(os.path.join(RESULT_ROOT, directory)):
+        result_path = os.path.join(os.path.join(
+            RESULT_ROOT, directory), translation)
+        with open(result_path, "r") as result:
+            result_text = result.read()
+        if "Refutation found." in result_text:
+            proven += 1
+        elif "SZS status CounterSatisfiable" in result_text:
+            disproven += 1
+        else:
+            unresolved += 1
+    total_proven += proven
+    total_disproven += disproven
+    total_unresolved += unresolved
+    total = proven + disproven + unresolved
+    total_total += total
+    print(
+        f"Category {directory}, Proven/Disproven/Unresolved:\t{proven}/{disproven}/{unresolved}\t{100*proven/total:.1f}%/{100*disproven/total:.1f}%/{100*unresolved/total:.1f}%")
+print("Total Result:")
+print(f"Proven/Disproven/Unresolved:\t{total_proven}/{total_disproven}/{total_unresolved}\t{100*total_proven/total_total:.1f}%/{100*total_disproven/total_total:.1f}%/{100*total_unresolved/total_total:.1f}%")

### 2022-10-11 Implementation/results-leancop.py

@@ -1,0 +1,36 @@
+import os
+import tqdm
+
+RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results-leancop/"
+
+total_proven = 0
+total_disproven = 0
+total_unresolved = 0
+total_total = 0
+
+if not os.path.isdir(RESULT_ROOT):
+    os.mkdir(RESULT_ROOT)
+for directory in os.listdir(RESULT_ROOT):
+    proven = 0
+    disproven = 0
+    unresolved = 0
+    for translation in tqdm.tqdm(os.listdir(os.path.join(RESULT_ROOT, directory)), desc=f"{directory}"):
+        result_path = os.path.join(os.path.join(
+            RESULT_ROOT, directory), translation)
+        with open(result_path, "r") as result:
+            result_text = result.read()
+        if "is an intuitionistic Theorem" in result_text:
+            proven += 1
+        elif "is an intuitionistic Non-Theorem" in result_text:
+            disproven += 1
+        else:
+            unresolved += 1
+    total_proven += proven
+    total_disproven += disproven
+    total_unresolved += unresolved
+    total = proven + disproven + unresolved
+    total_total += total
+    print(
+        f"Category {directory}, Proven/Disproven/Unresolved:\t{proven}/{disproven}/{unresolved}\t{100*proven/total:.1f}%/{100*disproven/total:.1f}%/{100*unresolved/total:.1f}%")
+print("Total Result:")
+print(f"Proven/Disproven/Unresolved:\t{total_proven}/{total_disproven}/{total_unresolved}\t{100*total_proven/total_total:.1f}%/{100*total_disproven/total_total:.1f}%/{100*total_unresolved/total_total:.1f}%")

### 2022-10-11 Implementation/results-nanocop.py

@@ -1,6 +1,6 @@
 import os
 
-RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results/"
+RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results-nanocop/"
 
 total_proven = 0
 total_disproven = 0
@@ -18,9 +18,9 @@ for directory in os.listdir(RESULT_ROOT):
             RESULT_ROOT, directory), translation)
         with open(result_path, "r") as result:
             result_text = result.read()
-        if "Refutation found." in result_text:
+        if "is an intuitionistic Theorem" in result_text:
             proven += 1
-        elif "SZS status CounterSatisfiable" in result_text:
+        elif "is an intuitionistic Non-Theorem" in result_text:
             disproven += 1
         else:
             unresolved += 1

### 2022-10-12 FoSSaCS/fossacs-EIICL.tex
fossacs first final draft

@@
 \usepackage{wrapfig}
-\usepackage{tikz-qtree}
-\usetikzlibrary{positioning}
-\usetikzlibrary{arrows.meta}
-\usetikzlibrary{arrows,shapes,quotes}
+\usepackage{multicol}
 \usepackage[T1]{fontenc}
@@
 We shall now give an overview of our main results and highlight the key arguments.
-Recall that our goal is to give an effective translation procedure that, for a given formula $\varphi$, yields a formula $\varphi^\#$, such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
 We start with the propositional case.
@@
 
-Before our main transformation we employ a preprocessing step akin to the Tseytin transformation~\cite{tseitin1983complexity}, which is a popular pre-processing step in classical automated reasoning:
-It gives an equisatisfiable sentence (over an extended set of propositions) in conjunctive normal form.
-Eliminating all implications, however, is not possible in intuitionistic logic since $A\to B$ is not equivalent to $\neg A\vee B$.
-Still we propose a similar transformation.
-A notable feature of our transformation is that all non-classical content is encapsulated in formulas of type $(A\to B)\to C$, i.e.
-if there are no such formulas in the transformed formula then the classical validity of the transformed formula immediately implies intuitionistic validity.
+The general idea is to use the Kripke semantics for intuitionistic logic, which are expressed in classical logic, to perform an embedding of formulas.
+That is, for each formula $\varphi$ we give a formula $\varphi^\#$ that is classically valid if and only if $\varphi$ is intuitionistically valid, establishing a dual to the double negation translation.
 
-\begin{theorem}\label{thm:Tseytin}
-For every propositional formula $\varphi$ there effectively are an atom $P$ and a set of $\mathcal S$ of formulas (over an extended set of propositions) of one of the forms\\
-$\hspace{1cm} A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$\\
-for atomic $A, B, C$, such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
-The time complexity of the transformation is linear in the input size.
-\end{theorem}
+The key property that allows us to reduce our encoding, which a priori targets first order logic, to propositional logic, is a reduction of counter-models, which is a proper dual to validity for intuitionistic logic.
+In particular, given $$\varphi = \left(\bigwedge\mathcal R\wedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form} we establish:
 
-The main transformation then proceeds in three steps:
-%
-1) We encode as a first-order sentence that the considered formula holds for every Kripke frame.
-Since Kripke frames for propositional logic over a fixed alphabet of propositional variables form a first-order theory this step is rather straightforward.
-%
-2) Next we eliminate some quantifiers via Herbrandization. The only quantifiers eliminated occur in formulas derived from formulas of type $(A\to B)\to C$.
-%
-3) We can then argue that if there is a counter-model for the resulting formula then there is a counter-model with a certain Kripke frame of bounded size completely determined by the number of formulas of type $(A\to B)\to C$.
-This allows us to eliminate the remaining quantifiers by simply enumerating the worlds in that Kripke frame.
+\begin{lemma}
+	If $\varphi$ is intuitionistically invalid then there exists a Kripke counter-model which is a rooted tree with height and degree bounded by $|\mathcal X|$.
+\end{lemma}
 
-Summarizing, we obtain the following result:
+This allows us to embed intuitionistic propositional logic into classical propositional logic.
 
-\begin{theorem}
-\label{thm:reduction-propositional}
-	Let $\mathcal S$ be as in Theorem~\ref{thm:Tseytin} and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
-For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
-Obtain $\mathcal S^\#$ by including the following formulas:
-	\begin{itemize}
-		\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
-		\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occur in $k$.
-	\end{itemize}
-Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
-The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$.
-Futhermore, there is an effective procedure for translating counter-models between the sentences.
-\end{theorem}
+\begin{lemma}
+	For each propositional formula $\varphi$ there is effectively a formula $\varphi^\#$ with $\mathcal |\varphi^\#|\in\mathcal O(|\mathcal R|\cdot\exp(|\mathcal X|\log(|\mathcal X|)))$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
+\end{lemma}
 
-Instead of explicitly enumerating the worlds of a counter-example of bounded size (as in the reduction stated in the above theorem), we can also do this enumeration implicitly with a quantified boolean formula (QBF).
-We can show that this QBF is satisfiable if and only if there exists a counter-model for our original formula:
+The exact details can be found in Theorem~\ref{thm:prop-translation}.
+The essential mechanisms is the elimination $\forall$-quantifiers by enumeration of worlds in a sufficiently general Kripke Frame.
+Instead of expressing correctness at each world of the model we can express that the worlds along some path in the tree are correct.
+The branching of paths can be expressed through quantification in a quantified boolean formula (QBF) which gives us the following:
 
-\begin{theorem}
-	Let $\mathcal S$ be as in Theorem~\ref{thm:Tseytin} and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$.
-There is an effective procedure that produces a QBF $\varphi^Q$ of size $\mathcal O(|\mathcal S|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$ with $2\cdot |\mathcal F_\to|-1$ quantifier alternations such that $\varphi$ is intuitionistically valid if and only if $\varphi^Q$ is not satisfiable.
-For fixed $N\in\mathbb N$, deciding intuitionistic validity for formulas $\bigwedge \mathcal S\to P$ where $\mathcal S$ is as above and $|\mathcal F_\to| = N$ is therefore in $\Sigma_{2N-1}$.
-\end{theorem}
+\begin{lemma}
+	For each propositional formula $\varphi$ there is effectively a QBF $\varphi^Q$ with $\mathcal |\varphi^Q|\in\mathcal O(|\mathcal R|\cdot|\mathcal X| + |\mathcal X|^3)$ such that $\varphi$ is intuitionistically invalid if and only if $\varphi^Q$ is a satisfiable QBF.
+\end{lemma}
 
-%\begin{theorem}
-%	For every predicate formula $\varphi$ there exists an effective procedure that yields an atom $P$ and a set $\mathcal S$ of formulas of the forms
-%	$$\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c))), \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c)),$$$$\forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c))),
-%	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c)),$$$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c)),\forall \vec z(\forall xA(\vec a)\to B(\vec b)),$$$$ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))$$for atomic $A, B, C$ and vectors of variables $\vec z, \vec a, \vec b, \vec c$ such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
-The size of $\mathcal S$ and time complexity of the transformation are linear in the input size.
-%\end{theorem}
-
-We then move to the first-order case. Here the construction is a bit more involved but the underlying approach is the same.
-Again we first perform a Tseytin-like transformation.
-Here the non-classical content is encapsulated by formulas of type $(A\to B)\to C$ and $\forall xA\to B$.
-The details of this translation are given in Section~\ref{section:tseytin}.
-We then proceed similarly to the propositional case.
-Note however that one difficulty arises from the fact that our (classical) domain will now contain on the one hand worlds in the Kripke frame but on the other hand also proper domain elements.
-We resolve this apparent conflict by introducing a special binary predicate $E$, inspired by~\cite{iemhoff2010eskolemization}, encoding which domain elements exists at which world.
-The main transformation then again proceeds in three steps:
-Step 1) and 2) are analogous to the propositional case.
-The real difference arises in step 3):
-While we can also establish the existence of a canonic counter-model whose frame only depends on formulas of type $(A\to B)\to C$ and $\forall xA\to B$, the size of this counter-model is countably infinite in general.
-This is not surprising since certain intuitionistically invalid formulas like the double negation shift $\forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$ don't have finite counter-models.
-Therefore we are not able to eliminate the $\forall$-quantifiers associated with the Kripke semantics.
-However since our translation targets first-order logic, we believe that the introduction of these quantifiers is acceptable.
-%
-Summarizing, we obtain the following result:
+Again, the details can be found in Section~\ref{sec:qbf}.
+Next we attempt to transport these results to the first order case.
+There are formulas for which only counter-examples with an infinite Kripke Frame exist.
+Therefore, there is no hope of completely eliminating terms representing worlds from the final embedding.
+This also means that our language necessarily becomes more complex.
+However a number of reductions is still possible.
 
@@
 \end{theorem}
-A more detailed account of Theorem~\ref{thm:reduction-first-order-short} is given in Theorem~\ref{thm:fo-translation}.
 
-These results provide an effective framework for using classical provers to check intuitionistic validity. First experiments with the Vampire theorem prover~\cite{kovacs2013first} show promise.
+A detailed account of the translation can be found in definition~\ref{def:fo-translation}.
+Finally, we benchmark our translation by translating all problems in the ILTP problem set~\cite{iltp} and applying the Vampire theorem prover~\cite{kovacs2013first} to the translated instances.
+While this approach falls short of the absolute state of the art in general, under the right circumstances it is a reasonable approach.
+Since there is still a lot of room for improvement this is a hopeful first sign.
 
@@
 \begin{corollary}
-	If a formula in normal form as in Lemma~\ref{lemma:propositional-normal-form} in intuitionistically invalid then there exists a Kripke counter-model which is a rooted tree with height and degree bounded by $|\mathcal X|$.
+	If a formula in normal form as in Lemma~\ref{lemma:propositional-normal-form} is intuitionistically invalid then there exists a Kripke counter-model which is a rooted tree with height and degree bounded by $|\mathcal X|$.
 \end{corollary}
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:prop-translation}
 	Suppose we are given a formula as in Lemma~\ref{lemma:propositional-normal-form}, i.e. of the form
@@
 
-\section{IPC to QBF}
+\section{IPC to QBF}\label{sec:qbf}
 
@@
 
-A simple counting argument shows that for each $n < |\mathcal X|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal X|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal R}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal X}(n)|\in \mathcal O(|\mathcal X|)$ so overall we get:
+A simple counting argument shows that for each $n < |\mathcal X|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal X|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal R}(n)|\in \mathcal O(|\mathcal R|)$ and $|\text{Sat}_{\mathcal X}(n)|\in \mathcal O(|\mathcal X|)$ so overall we get:
 \begin{lemma}
-	The size of $\varphi^Q$ is in $\mathcal O(|\varphi|\cdot|\mathcal X| + |\mathcal X|^3)$.
+	The size of $\varphi^Q$ is in $\mathcal O(|\mathcal R|\cdot|\mathcal X| + |\mathcal X|^3)$.
 \end{lemma}
@@
 We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$.
-To encode that some $n$-ary relation $A$ holds at a world $u$ we replace each $n$-ary relation symbol $A$ with a $n+1$-ary relation symbol $A$, interpreting $A(\vec x, u)$ as "$A(\vec x)$ holds at u".
+To encode that some $n$-ary relation $A$ holds at a world $u$ we extend each $n$-ary relation symbol to a $n+1$-ary relation symbol of the same name, interpreting $A(\vec x, u)$ as "$A(\vec x)$ holds at u".
 Furthermore we write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$.
@@
 
-\begin{definition}
+\begin{definition}\label{def:fo-translation}
 	Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively:
@@
 		\item $(\exists x\varphi)^u = \exists x(E(x,u)\wedge\varphi^u)$
-		\item $(\forall x\varphi)^u = \forall w(u\preceq w\to E(x, w)\to \varphi^w)$
+		\item $(\forall x\varphi)^u = \forall x\forall w(u\preceq w\to E(x, w)\to \varphi^w)$
 	\end{itemize}
@@
 
-Since we are mixing worlds and proper intuitionistic domain elements in a single classical domain there are some details to be taken care of.
-A proof is given in the appendix as Lemma~\ref{proof:fo-simplification}.
+\begin{proof}
+	Since we are mixing worlds and proper intuitionistic domain elements in a single classical domain there are some details to be taken care of. We proceed by translation of counter-models.
+
+	Suppose first we have a counter-model $\mathcal M = (M, I)$ to $\varphi^C$.
+As a Kripke frame $(W, \leq)$ take $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ and let $\leq$ be $\preceq^I$ restricted to $W$.
+Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A(\vec x, u)$.
+It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
+
+	The other direction is a bit more involved.
+Suppose we have a Kripke counter-model to $\varphi$ with frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
+In particular, since it is a counter-model there exists $w_0\in W$ and a variable assignment $v$ with $w_0, v\not\models\varphi$.
+Let $W_0 = \{w\in W\:|\: w_0\preceq w\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
+Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
+	Now have
+	\begin{itemize}
+		\item $\preceq^I = \leq$ when restricted to $W_0$ and $\prec$ can be interpreted as an arbitrary PO otherwise.
+		\item $E^I(m, w)$ iff $w\in W_0$ and $m = [x, w]$ for some $x\in M_w$.
+		\item $f^I(m_1\dots m_n) =\begin{cases}
+			f^{I_u}(x_1\dots x_n), &\substack{\text{if there are $ u\in W_0, x_i\in M_u$}\\ \text{with $m_i= [x_i, u]$ for all $i$,}}\\
+			w_0, & \text{otherwise.}
+		\end{cases}$
+		\item ${A}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
+			A^{I_u}(x_1\dots x_n), &\substack{\text{if $u\in W_0$ and $\exists x_i\in M_u$}\\\text{with $m_i= [x_i, u]$ for all $i$,}}\\
+			\top, & \text{otherwise.}
+		\end{cases}$
+	\end{itemize}
+	One easily verifies that these are well-defined.
+Let us now briefly check that this indeed defines a counter-model.
+Clearly World$(w_0)$ holds and $K(\varphi)$ also holds.
+All that remains to show is that $M, I, v[w_0/u]\not\models\varphi^u$.
+We proceed by induction on the height of $\varphi$.
+Consider e.g.
+the case where $\varphi$ is of the form $\forall x\psi$.
+	We have to show that
+	$$M, I, v[w_0/u]\models \forall x\forall w(u\preceq w\to E(x, w)\to \psi^w)$$
+	does not hold for some $v$.
+Now since $\varphi$ is intuitionistically invalid there exists $w_1\in W_0$ and $m\in M_{w_1}$ such that $w_1, v[m/x]\not\models\psi$.
+By induction hypothesis $M, I, v[w_1/w]\not\models\psi^w$ for some $v$.
+By definition $M, I, v[w_1/w, [m, w_1]/x]\models E(x, w)$ and $w_0\preceq^I w_1$ holds.
+Therefore indeed $$M, I, v[w_0/u, w_1/w, [m, w_1]/x]\not\models \forall w(u\preceq w\to E(x, w)\to \psi^w)$$ holds and we are done.
+The other cases are analogous.\hfill$\square$
+\end{proof}
 
@@
 	$$
-		\forall u\forall \vec x(\vec E(\vec x, u)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u))
+		\forall u\forall \vec x\left(\vec E(\vec x, u)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)\right)
 	$$
@@
  	$$
- 		\forall \vec x(\bigwedge_i A_i(\vec a_i)\to\bigvee_j B_j(\vec b_j))\in\mathcal R,
+ 		\forall \vec x\left(\bigwedge_i A_i(\vec a_i)\to\bigvee_j B_j(\vec b_j)\right)\in\mathcal R,
  	$$
@@
  	$$
- 		\forall u\forall \vec x((A(\vec a, f_\psi(\vec z, u))\to B(\vec b, f_\psi(\vec z, u))\to C(\vec c, u)))
+ 		\forall u\forall \vec x\left(A(\vec a, f_\psi(\vec z, u))\to B(\vec b, f_\psi(\vec z, u))\to C(\vec c, u)\right)
  	$$
@@
 	$$
-		\psi = \forall \vec x((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal X,
+		\psi = \forall \vec x\left(\left(A(\vec a)\to B(\vec b)\right)\to C(\vec c)\right)\in\mathcal X,
 	$$
@@
 	$$
-		\forall u\forall \vec x(\forall y(A(\vec a, f_\psi(\vec z, u)))\to B(\vec b, u))
+		\forall u\forall \vec x\left(\forall y\left(A(\vec a, f_\psi(\vec z, u))\right)\to B(\vec b, u)\right)
 	$$
@@
 	$$
-		\psi = \forall \vec x(\forall y(A(\vec a))\to B(\vec b))\in\mathcal X.
+		\psi = \forall \vec x\left(\forall y\left(A(\vec a)\right)\to B(\vec b)\right)\in\mathcal X.
 	$$
@@
 		\begin{tabular}{l|c|c}
-			Embedding&&\\
+			Embedding&571&21.4\%\\
 			 Embedding (idem.)&557&20.9\%\\
-			ileanCoP 1.2&&\\
+			ileanCoP 1.2&875&32.8\%\\
 			nanoCoP-i 2.0&858&32.1\%\\\hline
@@
 	\centering
-	\begin{tabular}{c|c|c|c|c|c|c}
-		$|\mathcal Q|$&$|\mathcal X|$&Embedding&Embedding (idem.)&ileanCoP&nanoCoP-i 2.0&Total\\\hline
-		0&0&&157&&61&178\\
-		0&1-2&&130&&96&400\\
-		0&3-5&&17&&40&105\\
-		0&6-10&&4&&7&19\\
-		0&11-20&&0&&3&34\\
-		0&21+&&7&&1&17\\
-		1-2&0&&14&&15&17\\
-		1-2&1-2&&67&&206&270\\
-		1-2&3-5&&77&&122&312\\
-		1-2&6-10&&15&&28&64\\
-		1-2&11+&&4&&6&12\\
-		3-5&0&&14&&16&16\\
-		3-5&1-2&&11&&36&49\\
-		3-5&3-5&&15&&58&159\\
-		3-5&6-10&&16&&83&199\\
-		3-5&11+&&2&&3&38\\
-		6+&0-2&&1&&50&50\\
-		6+&3+&&3&&24&605
+	\begin{tabular}{c|c|lr|lr|lr|lr|c}
+		$|\mathcal Q|$&$|\mathcal X|$&\multicolumn{2}{c}{Embedding}&\multicolumn{2}{c}{Embedding (idem.)}&\multicolumn{2}{c}{ileanCoP}&\multicolumn{2}{c}{nanoCoP-i 2.0}&Total\\\hline
+		0&0		&157&88.2\%	&157&88.2\%		&128&71.9\%		&61&38.1\%	&178\\
+		0&1-2	&119&29.75\%&130&32.5\%		&108&27.0\%		&96&24.0\%	&400\\
+		0&3-5	&16&		&17&16.2\%		&50&47.6\%		&40&28.1\%	&105\\
+		0&6-10	&16&		&4&21.1\%		&9&47.4\%		&7&36.8\%	&19\\
+		0&11-20	&2&			&0&0\%			&2&5.9\%		&3&8.8\%	&34\\
+		0&21+	&8&			&7&41.2\%		&9&52.9\%		&1&5.9\%	&17\\
+		1-2&0	&14&		&14&82.4\%		&16&94.1\%		&15&88.2\%	&17\\
+		1-2&1-2	&70&		&67&24.8\%		&211&78.1\%		&206&76.3\%	&270\\
+		1-2&3-5	&80&		&77&24.7\%		&141&45.2\%		&122&39.1\%	&312\\
+		1-2&6-10&16&		&15&23.4\%		&29&45.3\%		&28&43.8\%	&64\\
+		1-2&11+	&6&			&6&16.2\%		&10&27.0\%		&9&24.3\%	&37\\
+		3-5&0	&14&		&14&87.5\%		&16&100\%		&16&100\%	&16\\
+		3-5&1-2	&14&		&11&22.4\%		&37&75.5\%		&36&73.4\%	&49\\
+		3-5&3-5	&10&		&15&9.4\%		&72&45.3\%		&58&36.5\%	&159\\
+		3-5&6-10&18&		&16&8.0\%		&64&32.2\%		&83&41.7\%	&199\\
+		3-5&11+	&2&			&2&5.3\%		&0&0\%			&3&7.9\%	&38\\
+		6+&0-2	&1&			&1&2.0\%		&49&98.0\%		&50&100\%	&50\\
+		6+&3+	&7&			&3&	0.6\%		&24&3.4\%		&24&3.4\%	&605
 	\end{tabular}
@@
 Therefore our tool is already useful if it is known that $|\mathcal X|$ and $|\mathcal Q|$ are small.
-Another hopeful sign is that the inclusion of idempotence increases performance in particular for larger $|\mathcal X|$ and $|\mathcal Q|$.
-Further improvements could be achieved by introducing more axioms based on structural properties of the formula.
+We see that idempotence does not improve performance.
+On the other hand from a model-theoretic perspective this is quite powerful.
+We suspect that a more clever encoding of this property could improve our results.
+There are other restriction on models we could consider encoding based on structural properties of the formula, e.g.
+the relation between occurrences of different atoms.
 Another interesting observation is that $|\mathcal X|$ and $|\mathcal Q|$ also seem to correlate with success rate for the dedicated intuitionistic provers.
@@
 However, the current complexity considerations are direct consequences of our straightforward constructions.
-We plan to establish better bounds in future work by utilizing structural properties of the input formula, in particular by constructing smaller counter-examples $\mathcal M_T$ and using redundant axioms to limit the search.
-Our addition of idempotence axioms has demonstrated that this is a promising approach.
+We plan to establish better bounds in future work by utilizing structural properties of the input formula, in particular relations between different atoms.
+We hope that this will allow the construction of smaller counter-examples $\mathcal M_T$ and the use of redundant axioms to limit search in a more clever way.
 
@@
 
-\begin{lemma}\label{proof:fo-simplification}
-	$\varphi$ is intuitionistically valid if and only of $\varphi^C$ is classically valid.
-\end{lemma}
-
-\begin{proof}
-	We proceed by translation of counter-examples.
-Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi^C$.
-As a Kripke frame $(W, \leq)$ take all $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ let $\leq$ be $\preceq^I$ restricted to $W$.
-Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A(\vec x, u)$.
-It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
-
-	The other direction is a bit more involved.
-Suppose we have a Kripke counter-model to $\varphi$ with frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
-In particular, since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
-Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
-Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
-	Now have
-	\begin{itemize}
-		\item $s^I = w_0$.
-		\item $E^I(m, w)$ iff $w\in W_0$ and $m = [x, w]$ for some $x\in M_w$.
-		\item $f^I(m_1\dots m_n) =\begin{cases}
-			f^{I_u}(x_1\dots x_n), &\substack{\text{if there are $ u\in W_0, x_i\in M_u$}\\ \text{with $m_i= [x_i, u]$ for all $i$,}}\\
-			w_0, & \text{otherwise.}
-		\end{cases}$
-		\item ${A}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
-			A^{I_u}(x_1\dots x_n), &\substack{\text{if $u\in W_0$ and $\exists x_i\in M_u$}\\\text{with $m_i= [x_i, u]$ for all $i$,}}\\
-			\top, & \text{otherwise.}
-		\end{cases}$
-	\end{itemize}
-	One easily verifies that these are well-defined.
-Let us now briefly check that this indeed defines a counter-model.
-First of all there ${P}^I(w_0)$ is false, World$(w_0)$ holds and $K(\varphi)$ is also easily verified.
-All that remains to show is that (for each $v$) $M, I, v[w_0/u]\models\bigwedge\mathcal R^{u}\wedge\bigwedge\mathcal X^{u}$.
-Consider e.g.
-the case where $\psi\in\mathcal X$, i.e.
-is of the form
-	$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
-	We have to show that
-	$$M, I\models \forall \vec z(\vec E(\vec z, w_0)\to \forall w(w_0\preceq w\to A(\vec a, w)\to B(\vec b, w))\to C(\vec c, w_0))$$
-	holds. Suppose towards contradiction that it does not, i.e. there is $\vec z\in M$ such that for all $w\in M$
-	$$\vec E^I(\vec z, w_0)\to (w_0\preceq w\to {A}^I(\vec a, w)\to {B}^I(\vec b, w))\to {C}^I(\vec c, w_0)$$ is false.
-For that ${C}^I(\vec c, w_0)$ must be false and $\vec E^I(\vec z, w_0)$ must be true so we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_{w_0}$.
-But then by5 definition
-	$$((A^{I_{w_0}}(\vec a)\to B^{I_{w_0}}(\vec b))\to C^{I_{w_0}}(\vec c))[\vec z/z_1\dots z_n]$$ is false in our original Kripke counter-model, so it cannot be a counter-example of $\varphi$, which is a contradiction.
-The argument for $\psi\in\mathcal R$ is analogous.\hfill$\square$
-\end{proof}
-
 \end{document}

### 2022-10-12 Implementation/benchmark-embedding.py

@@ -1,0 +1,71 @@
+import subprocess
+import tqdm
+import os
+
+VAMPIRE_PATH = "/home/lambdax/.vampire/vampire"
+
+PROBLEM_ROOT = "./ILTP-v1.1.2-firstorder-translated-embedding-idemp/"
+TRANSLATION_ROOT = "./ILTP-v1.1.2-firstorder-translated-embedding/"
+RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results-embedding/"
+
+total_proven = 0
+total_disproven = 0
+total_unresolved = 0
+total_total = 0
+
+if not os.path.isdir(TRANSLATION_ROOT):
+    os.mkdir(TRANSLATION_ROOT)
+if not os.path.isdir(RESULT_ROOT):
+    os.mkdir(RESULT_ROOT)
+for directory in os.listdir(PROBLEM_ROOT)[-4:]:
+    proven = 0
+    disproven = 0
+    unresolved = 0
+    if not os.path.isdir(os.path.join(TRANSLATION_ROOT, directory)):
+        os.mkdir(os.path.join(TRANSLATION_ROOT, directory))
+    if not os.path.isdir(os.path.join(RESULT_ROOT, directory)):
+        os.mkdir(os.path.join(RESULT_ROOT, directory))
+    for translation in tqdm.tqdm(os.listdir(os.path.join(PROBLEM_ROOT, directory)), desc=f"{directory}"):
+        problem_path = os.path.join(os.path.join(
+            PROBLEM_ROOT, directory), translation)
+        translation_path = os.path.join(os.path.join(
+            TRANSLATION_ROOT, directory), translation)
+        result_path = os.path.join(os.path.join(
+            RESULT_ROOT, directory), translation)
+        with open(problem_path, "r") as input_file:
+            try:
+                input_text = input_file.read()
+                with open(translation_path, "w") as output_file:
+                    print(('\n').join([line for line in input_text.split("\n") if "fof(a,axiom,'f" not in line]), file = output_file)
+                with open(result_path, "w") as output_file:
+                    subprocess.run([VAMPIRE_PATH, translation_path,
+                                "--time_limit", "10"], stdout=output_file)
+            except MemoryError:
+                with open(result_path, "w") as output_file:
+                    print("Error reading file", file = output_file)
+        try:
+            with open(result_path, "r") as  input_file:
+                result_text = input_file.read()
+            with open(problem_path, "r") as input_file:
+                problem_text = input_file.read()    
+            if "Refutation found." in result_text:
[…]

### 2022-10-12 Implementation/benchmark_leancop.py

@@ -18,7 +18,7 @@ if not os.path.isdir("ILTP-v1.1.2-firstorder-translated-leancop"):
     os.mkdir("ILTP-v1.1.2-firstorder-translated-leancop")
 if not os.path.isdir("ILTP-v1.1.2-firstorder-results-leancop"):
     os.mkdir("ILTP-v1.1.2-firstorder-results-leancop")
-for directory in os.listdir(PROBLEM_ROOT)[7:]:
+for directory in os.listdir(PROBLEM_ROOT):
     proven = 0
     disproven = 0
     unresolved = 0
@@ -32,7 +32,7 @@ for directory in os.listdir(PROBLEM_ROOT)[7:]:
         translation_path = os.path.join(os.path.join(
             TRANSLATION_ROOT, directory), problem)
         with open(translation_path, "w") as output_file:
-            subprocess.run([TRANSLATOR, "-f", "leancop", "-q2", "-d", "-", problem_path], stdout=output_file)
+            subprocess.run([TRANSLATOR, "-f", "leancop", "-q2", "-t", "stdfof+add_equality", "-d", "-", problem_path], stdout=output_file)
         result_path = os.path.join(os.path.join(
             RESULT_ROOT, directory), problem)
         with open(result_path, "w") as output_file:

### 2022-10-12 Implementation/bucketed_results.py

@@ -3,18 +3,22 @@ import os
 import tqdm
 
 RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results"
-TRANSLATION_ROOT = "./ILTP-v1.1.2-firstorder-translated"
-PROVERS = ["-embedding-idemp", "-nanocop"]
+TRANSLATION_ROOT = "./ILTP-v1.1.2-firstorder-translated-embedding-idemp"
+PROVERS = ["-embedding", "-embedding-idemp", "-leancop", "-nanocop"]
 PROOF = {
+    "-embedding" : "Refutation found.",
     "-embedding-idemp" : "Refutation found.",
+    "-leancop" : "is an Intuitionistic Theorem",
     "-nanocop" : "is an intuitionistic Theorem"
 }
 COUNTERMODEL = {
+    "-embedding" : "SZS status CounterSatisfiable",
     "-embedding-idemp" : "SZS status CounterSatisfiable",
+    "-leancop" : "is an Intuitionistic Non-Theorem",
     "-nanocop" : "is an intuitionistic Non-Theorem"
 }
 
-buckets = [0, 2, 5, 10, 20, 1000]
+buckets = {0 : [0, 2, 5, 10, 20, 1000], 2 : [0, 2, 5, 10, 1000], 5 : [0, 2, 5, 10, 1000], 1000 : [2, 1000] }
 tally = {}
 
 for directory in os.listdir(TRANSLATION_ROOT):
@@ -29,17 +33,22 @@ for directory in os.listdir(TRANSLATION_ROOT):
             except MemoryError:
                 q = 100
                 x = 100
-        i = min([b for b in buckets if b >= q])
-        j = min([b for b in buckets if b >= x])
+        i = min([b for b in buckets.keys() if b >= q])
+        j = min([b for b in buckets[i] if b >= x])
         if not (i, j) in tally:
             tally[(i, j)] = {p : 0 for p in PROVERS + ["total"]}
         tally[(i, j)]["total"] += 1
         for p in PROVERS:
             root = RESULT_ROOT + p
             result_path = os.path.join(os.path.join(root, directory), translation)
-            with open(result_path, "r") as result_text:
-                result_text = result_text.read()
-            if PROOF[p] in result_text or COUNTERMODEL[p] in result_text:
-                tally[(i, j)][p] += 1
+            try:
+                with open(result_path, "r") as result_text:
+                    result_text = result_text.read()
[…]

### 2022-10-12 Implementation/results-embedding.py

@@ -1,7 +1,6 @@
 import os
-import tqdm
 
-RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results-leancop/"
+RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results-embedding/"
 
 total_proven = 0
 total_disproven = 0
@@ -14,14 +13,14 @@ for directory in os.listdir(RESULT_ROOT):
     proven = 0
     disproven = 0
     unresolved = 0
-    for result in tqdm.tqdm(os.listdir(os.path.join(RESULT_ROOT, directory)), desc=f"{directory}"):
+    for translation in os.listdir(os.path.join(RESULT_ROOT, directory)):
         result_path = os.path.join(os.path.join(
-            RESULT_ROOT, directory), result)
-        with open(result_path, "r") as input_file:
-            result_text = input_file.read()
-        if "is an Intuitionistic Theorem" in result_text:
+            RESULT_ROOT, directory), translation)
+        with open(result_path, "r") as result:
+            result_text = result.read()
+        if "Refutation found." in result_text:
             proven += 1
-        elif "is an Intuitionistic Non-Theorem" in result_text:
+        elif "SZS status CounterSatisfiable" in result_text:
             disproven += 1
         else:
             unresolved += 1

### 2022-10-13 FoSSaCS/fossacs-EIICL.tex
typos and abstract

@@
 Curiously the reverse direction has not been covered in literature.
-Based on the work in~\cite{claessen2015sat} we establish a small model property for intuitionistic propositional logic which allows us to utilize its Kripke semantics to give a simple embedding into classical propositional logic and quantified boolean formulas.
-We then transfer this procedure to first-order logic which allows us to give a embedding of intuitionistic first-order logic into classical first-order-logic.
+Based on the work in~\cite{claessen2015sat} we establish a small model property for intuitionistic propositional logic in order to utilize its Kripke semantics to give a simple embedding into classical propositional logic and quantified boolean formulas.
+We transfer there methods to the first order case and also give an embedding of intuitionistic first-order logic into classical first-order-logic.
 This allows a direct application of state-of-the-art classical first-order provers to the problem of determining intuitionistic validity.
@@
 \begin{definition}
-	A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds}, called the Kripke \emph{frame}, and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
+	A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds}, called the Kripke \emph{frame}, and an indexed family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_w(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
 	The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, where we set
 	\begin{itemize}
-		\item $v_u\models\varphi\to \psi$ iff for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
+		\item $v_u\models\varphi\to \psi$ iff for all $w\geq u$ we have $w\not\models\varphi$ or $w\models\psi$.
 	\end{itemize}
@@
 		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
-		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
-		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
+		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$.
+		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$.
 	\end{itemize}
+	 Here $a$ is a free variable that does not occur in $\varphi$.
 A $\Sigma$-structure $\mathcal M$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$, we write $\mathcal M\models\varphi$.
@@
 \begin{definition}
-	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
-	The model relation between $\mathcal M_u$, $v$ (targeting $\mathcal M_u$) and formulas $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where we set
+	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
+	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulas $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where we set
 	\begin{itemize}
 		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
-		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
+		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$.
 	\end{itemize}
@@
 \end{definition}
-In addition to propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
+In addition to propositional tautologies there are now quantified formulas which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
 
@@
 \begin{definition}
-	For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to $Z$ by simultaneous induction as follows:
+	For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to a set of variables $Z$ by simultaneous induction:
 	\begin{itemize}
@@
 		\item $(\varphi\circ\psi)^X_Z = \varphi^X_Z\circ\psi^X_Z$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
-		\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$\\$(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
-		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
-		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.
+		\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$ and $(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
+		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
+		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.
 	\end{itemize}
@@
 
-Since we could not find a proof of this result one is given in the appendix as Lemma~\ref{proof:first-order-normal-form}.
+Since we could not find a proof of this, one is given in the appendix.
 
@@
 We start with the propositional case.
-The key arguments are the same as in the first-order case, while it is technically simpler and less cluttered.
+The important insights are the same as in the first-order case, while it is technically simpler and less cluttered.
 
@@
 
-The key property that allows us to reduce our encoding, which a priori targets first order logic, to propositional logic, is a reduction of counter-models, which is a proper dual to validity for intuitionistic logic.
+The key property that allows us to reduce our encoding to propositional logic, which a priori targets first order logic, is a reduction of counter-models.
 In particular, given $$\varphi = \left(\bigwedge\mathcal R\wedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form} we establish:
@@
 
-This allows us to embed intuitionistic propositional logic into classical propositional logic.
+This allows us to embed intuitionistic propositional logic into classical propositional logic by replacing quantification over worlds by their enumeration.
 
 \begin{lemma}
-	For each propositional formula $\varphi$ there is effectively a formula $\varphi^\#$ with $\mathcal |\varphi^\#|\in\mathcal O(|\mathcal R|\cdot\exp(|\mathcal X|\log(|\mathcal X|)))$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
+	For each propositional formula $\varphi$ there is effectively a formula $\varphi^\#$ with $\mathcal |\varphi^\#|\in\mathcal O(|\mathcal R|\cdot\exp(|\mathcal X|\log|\mathcal X|))$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
 \end{lemma}
@@
 The exact details can be found in Theorem~\ref{thm:prop-translation}.
-The essential mechanisms is the elimination $\forall$-quantifiers by enumeration of worlds in a sufficiently general Kripke Frame.
-Instead of expressing correctness at each world of the model we can express that the worlds along some path in the tree are correct.
+Instead of expressing correctness at each world of the model, we can express that the worlds along some path in the tree are correct.
 The branching of paths can be expressed through quantification in a quantified boolean formula (QBF) which gives us the following:
@@
 Next we attempt to transport these results to the first order case.
-There are formulas for which only counter-examples with an infinite Kripke Frame exist.
+However, there are formulas for which only counter-examples with an infinite Kripke frame exist.
 Therefore, there is no hope of completely eliminating terms representing worlds from the final embedding.
 This also means that our language necessarily becomes more complex.
-However a number of reductions is still possible.
+Still, a number of reductions is possible.
 

### 2022-10-13 FoSSaCS/fossacs-EIICL.tex
typos

@@
 Finally, we benchmark our translation by translating all problems in the ILTP problem set~\cite{iltp} and applying the Vampire theorem prover~\cite{kovacs2013first} to the translated instances.
-While this approach falls short of the absolute state of the art in general, under the right circumstances it is a reasonable approach.
+While this method falls short of the absolute state of the art in general, under the right circumstances it is a reasonable approach.
 Since there is still a lot of room for improvement this is a hopeful first sign.
@@
 
-\noindent As all we have done is modeling Kripke frames as a first-order theory so we directly obtain:
+\noindent All we have done is model Kripke semantics as a first-order theory, so we obtain:
 
@@
 	It is well known that Kripke frames which are trees are complete with regard to intuitionistic logic.
-Our contribution is giving an easy to compute explicit size and geometry that depends only on the structure of the considered formula.
+Our contribution is giving an explicit bound to height and degree that depends only on a simple property of the considered formula.
 \end{remark}
@@
 We can apply the previous results to give a direct embedding from intuitionistic into classical propositional logic.
-We know that $\varphi^{CH}$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
+We know that $\varphi^{CH}$ is valid if and only if it is valid for structures that have as a domain the sequences without repetition over $\mathcal X$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
 	x, &\text{ if $\psi$ occurs in $x$,}\\
@@
 Now we can replace every $\forall$-quantifier in $\varphi^{CH}$ by enumerating all the ground terms, and then replace all distinct ground instances of relations by new propositional variables yielding a propositional formula.
-Note that the encoding of partial order becomes redundant with this so we drop it, but we need to keep the persistency constraints.
+Note that the encoding of the partial order becomes redundant, so we drop it.
+We need to keep the persistency constraints.
 This gives us the following theorem:
@@
 	$$
-		\left\{\left(A^{\lambda\psi} \to A^{\lambda\psi}\right)\to(C^\lambda)\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
+		\left\{\left(A^{\lambda\psi} \to A^{\lambda\psi}\right)\to C^\lambda\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
 	$$
@@
 	$$
-		\left\{A^\lambda\to A^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\psi\text{ does not occur in }\lambda\right\}
+		\left\{A^\lambda\to A^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
 	$$
-	 for each propositional variable $A$ occurring in $\varphi$ encoding persistency. Then
+	 for each propositional variable $A$ occurring in $\varphi$, encoding persistency. Then
 	$$
@@
 	\end{itemize}
-	Have $k = |\mathcal X|$ and define$$\varphi^Q_{i} = \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
-	for $0 < i < k$, as well as the special cases
-	$$\varphi^Q_{k} = \exists \vec A^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal R}(k)\wedge \text{\normalfont Sat}_{\mathcal X}(k)\right)$$
+	Define$$\varphi^Q_{i} = \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
+	for $0 < i < |\mathcal{X}|$, as well as the special cases
+	$$\varphi^Q_{|\mathcal{X}|} = \exists \vec A^{|\mathcal{X}|}\left(\text{\normalfont Persistent}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal R}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal X}(|\mathcal{X}|)\right)$$
 	for leafs, and $$\varphi^Q = \varphi_0^Q = \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.

### 2022-10-13 FoSSaCS/fossacs-EIICL.tex
Contributions + Related Work

@@
 
-A first benchmark shows that under the right circumstances this is a valid approach.
+For certain classes of benchmarks our approach can compete with state-of-the-art provers, in particular when the intuitionistic content is low.
 Crucially, our constructions support the transfer of counter-models to validity, which is important to model checking applications, where the output of counter-examples is an appreciated feature.
@@
 \end{itemize}
+
+\subsection*{Main Contributions}
+\begin{itemize}
+	\item Establishing a small model property for intuitionistic propositional logic.
+	\item Embedding intuitionistic propositional logic into classical propositional logic.
+	\item \resizebox*{0.96\hsize}{!}{Embedding intuitionistic propositional logic into quantified boolean formulas.}
+	\item Presenting a model reduction for intuitionistic first-order logic.
+	\item Embedding intuitionistic into classical first-order logic.
+	\item Benchmarking our new approach to intuitionistic theorem proving.  
+\end{itemize}
+
+\subsection*{Related work}
+
+There are various known translations between logics.
+Most relevant are the double negation translation~\cite{glivenko1929quelques,gentzen1936widerspruchsfreiheit}, the embedding of intuitionistic logic into S4~\cite{basicprooftheory} and of various non-classical logics into higher-order classical logic~\cite{LEO,wisniewski2016tptp}.
+
+Of special relevance is the reduction of intuitionistic propositional logic given in~\cite{claessen2015sat} and examined in~\cite{fiorentini2019proof}, which we further expand upon.
+
+On the other hand there have been attempts to circumvent intuitionistic logic by equipping classical proofs with operational semantics~\cite{Control1,Parigot1}, which however does not permit in general the realizability of $\exists$ and $\vee$.
+
+There are some results relating intuitionistic and classical validity, most famously for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}, but also in some other cases~\cite{schwichtenberg}.
+However, they only apply in very limited settings.
+
+Finally, of course dedicated theorem provers for intuitionistic logic exists, generally using the inverse method~\cite{mclaughlin2008imogen,pavlov2017whaleprover}, but more recently also based on connection-based calculi~\cite{otten2008leancop,otten2021nanocop}, which also work well for classical logic.
 
@@
 Finally, we benchmark our translation by translating all problems in the ILTP problem set~\cite{iltp} and applying the Vampire theorem prover~\cite{kovacs2013first} to the translated instances.
-While this method falls short of the absolute state of the art in general, under the right circumstances it is a reasonable approach.
+While this method falls short of the absolute state-of-the-art in general, for certain subsets of benchmarks it beats existing provers, in particular when the intuitionistic content is low.
 Since there is still a lot of room for improvement this is a hopeful first sign.

### 2022-10-13 FoSSaCS/fossacs-EIICL.tex
Symmetric difference

@@
 %
+\usepackage{tikz}
+\usetikzlibrary{shapes,backgrounds}
+
 \begin{document}
@@
 
-It is not possible to completely eliminate worlds from the domain by enumerating them since there are formulas with no finite counter-examples.
-An analogous procedure to the propositional case fails, however we can still retrieve additional axioms that restrict the interpretation of functions.
-That is, we may assume that $f_\psi(\vec t, w) = w$, if $\exists u(w = f_\psi(\vec t, u))$ for all $\psi\in\mathcal X\cup\mathcal Q$ (i.e.
-that $f_\psi$ parametrized by $\vec t$ is idempotent), and $f_\psi(\vec t, w) = w$, if $\exists u(w \geq f_\psi(\vec t, u))$ for all $\psi\in\mathcal X$ (which is even stronger than idempotence).
-
-\section{Benchmark}
-
-\begin{wrapfigure}[6]{r}{0.4\textwidth}
-		\vspace*{-1.7cm}
+\begin{wrapfigure}[8]{r}{0.4\textwidth}
+	\vspace*{-.5cm}
 	\begin{tabular}{l|c|c}
@@
 
+It is not possible to completely eliminate worlds from the domain by enumerating them since there are formulas with no finite counter-examples.
+An analogous procedure to the propositional case fails, however we can still retrieve additional axioms that restrict the interpretation of functions.
+That is, we may assume that $f_\psi(\vec t, w) = w$, if $\exists u(w = f_\psi(\vec t, u))$ for all $\psi\in\mathcal X\cup\mathcal Q$ (i.e.
+that $f_\psi$ parametrized by $\vec t$ is idempotent), and $f_\psi(\vec t, w) = w$, if $\exists u(w \geq f_\psi(\vec t, u))$ for all $\psi\in\mathcal X$ (which is even stronger than idempotence).
+
+\section{Benchmark}
+
+\begin{wrapfigure}[13]{r}{0.4\textwidth}
+	\vspace*{4.9cm}
+	\hspace*{1cm}
+	\begin{tikzpicture}
+		\begin{scope}[opacity=0.3, blend mode=multiply]
+			\def\firstcircle{(90:.75) circle (1.5)}
+			\def\secondcircle{(210:.75) circle (1.5)}
+			\def\thirdcircle{(330:.75) circle (1.5)}
+			\begin{tikzpicture}
+				\begin{scope}[shift={(3cm,-5cm)}, fill opacity=0.5]
+					\fill[red] \firstcircle;
+					\fill[green] \secondcircle;
+					\fill[blue] \thirdcircle;
+					\draw node {$491$};
+					\draw node at (30:1.2) {$30$};
+					\draw node at (150:1.2) {$42$};
+					\draw node at (270:1.2) {$22$};
+					\draw node at (90:1.8) {$8$};
+					\draw node at (210:1.8) {$2$};
+					\draw node at (330:1.8) {$417$};
+				\end{scope}
+			\end{tikzpicture}
+		\end{scope}
+	\end{tikzpicture}
+	\vspace*{-7.1cm}
+	\caption{Symmetric difference. Instances proven by Embedding (red), Embedding (idem.) (green), ileanCoP or nanoCoP (blue)}
+	\label{fig:c}
+\end{wrapfigure}
+
 We have benchmarked our first-order translation on version 1.1.2 of the ILTP problem set~\cite{iltp} by performing our translation and then examining the generated classical problem using Vampire~\cite{kovacs2013first}.
-For comparison the state-of-the-art intuitionistic provers ileanCoP 1.2~\cite{otten2008leancop} and nanoCoP-i 2.0~\cite{otten2021nanocop} are used.
+For comparison the state-of-the-art intuitionistic provers ileanCoP 1.2~\cite{otten2008leancop} and nanoCoP-i 2.0~\cite{otten2021nanocop} were used.
 All benchmarks were conducted on an Intel Core i5-8400 2.80GHz machine running Linux with a time limit of 10s per problem.
+The problem \textsc{SYN007+1} was removed as our normal form procedure would explode the size of instance.
+
 Two versions of the translation were tested, one with and one without the additional axioms encoding idempotence.
@@
 In general our approach falls short of the state of the art.
-However when binning for $|\mathcal Q|$ and $|\mathcal X|$ as seen in Figure~\ref{fig:b} we observe something interesting.
+However when binning for $|\mathcal Q|$ and $|\mathcal X|$ as seen in Figure~\ref{fig:b} we observe good performance for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 3$.
+Figure~\ref{fig:c} shows that we solve 52 problems not solved by the existing provers.
 \begin{figure}
@@
 		0&0		&157&88.2\%	&157&88.2\%		&128&71.9\%		&61&38.1\%	&178\\
-		0&1-2	&119&29.75\%&130&32.5\%		&108&27.0\%		&96&24.0\%	&400\\
-		0&3-5	&16&		&17&16.2\%		&50&47.6\%		&40&28.1\%	&105\\
-		0&6-10	&16&		&4&21.1\%		&9&47.4\%		&7&36.8\%	&19\\
-		0&11-20	&2&			&0&0\%			&2&5.9\%		&3&8.8\%	&34\\
-		0&21+	&8&			&7&41.2\%		&9&52.9\%		&1&5.9\%	&17\\
-		1-2&0	&14&		&14&82.4\%		&16&94.1\%		&15&88.2\%	&17\\
-		1-2&1-2	&70&		&67&24.8\%		&211&78.1\%		&206&76.3\%	&270\\
-		1-2&3-5	&80&		&77&24.7\%		&141&45.2\%		&122&39.1\%	&312\\
-		1-2&6-10&16&		&15&23.4\%		&29&45.3\%		&28&43.8\%	&64\\
-		1-2&11+	&6&			&6&16.2\%		&10&27.0\%		&9&24.3\%	&37\\
-		3-5&0	&14&		&14&87.5\%		&16&100\%		&16&100\%	&16\\
-		3-5&1-2	&14&		&11&22.4\%		&37&75.5\%		&36&73.4\%	&49\\
-		3-5&3-5	&10&		&15&9.4\%		&72&45.3\%		&58&36.5\%	&159\\
-		3-5&6-10&18&		&16&8.0\%		&64&32.2\%		&83&41.7\%	&199\\
-		3-5&11+	&2&			&2&5.3\%		&0&0\%			&3&7.9\%	&38\\
-		6+&0-2	&1&			&1&2.0\%		&49&98.0\%		&50&100\%	&50\\
-		6+&3+	&7&			&3&	0.6\%		&24&3.4\%		&24&3.4\%	&605
+		0&1-2	&119&29.8\%&130&32.5\%		&108&27.0\%		&96&24.0\%	&400\\
+		0&3-5	&16&15.2\%	&17&16.2\%		&50&47.6\%		&40&28.1\%	&105\\
+		0&6+	&27&38.6\%	&11&15.7\%		&20&28.6\%		&11&15.7\%	&70\\
+		1-2&0-2	&84&29.3\%	&81&28.2\%		&227&79.1\%		&221&77.0\%	&287\\
+		1-2&3-5	&80&25.6\%	&77&24.7\%		&141&45.2\%		&122&39.1\%	&312\\
+		1-2&6+	&22&21.8\%	&21&20.8\%		&39&38.6\%		&37&36.6\%	&101\\
+		3-5&0-5	&38&17.0\%	&40&17.9\%		&125&55.8\%		&110&49.1\%	&224\\
+		3-5&6+  &20&8.4\%	&18&7.6\%		&64&27.0\%		&86&36.3\%	&237\\
+		6+&0+	&8&1.1\%	&5&0.7\%		&73&9.7\%		&74&9.8\%	&755
 	\end{tabular}
@@
 \end{figure}
-Leveraging the power of Vampire, for low $|\mathcal X|$ and $|\mathcal Q|$ the embedding performs on par or even outperforms the dedicated intuitionistic provers.
-For larger values the state of the art provers have an advantage, in particular when both parameters are combined.
+
  Therefore our tool is already useful if it is known that $|\mathcal X|$ and $|\mathcal Q|$ are small.
 We see that idempotence does not improve performance.
-On the other hand from a model-theoretic perspective this is quite powerful.
+On the other hand from a model-theoretic perspective it is quite powerful.
 We suspect that a more clever encoding of this property could improve our results.
-There are other restriction on models we could consider encoding based on structural properties of the formula, e.g.
+There are other restrictions on models we could consider, based on structural properties of the formula, e.g.
 the relation between occurrences of different atoms.
-Another interesting observation is that $|\mathcal X|$ and $|\mathcal Q|$ also seem to correlate with success rate for the dedicated intuitionistic provers.
 

### 2022-10-13 Implementation/symmetric_difference.py

@@ -1,5 +1,12 @@
 import os
 
+from itertools import chain, combinations
+
+def powerset(iterable):
+    "list(powerset([1,2,3])) --> [(), (1,), (2,), (3,), (1,2), (1,3), (2,3), (1,2,3)]"
+    s = list(iterable)
+    return chain.from_iterable(combinations(s, r) for r in range(len(s)+1))
+
 import tqdm
 
 RESULT_ROOT = "./ILTP-v1.1.2-firstorder-results"
@@ -18,26 +25,11 @@ COUNTERMODEL = {
     "-nanocop" : "is an intuitionistic Non-Theorem"
 }
 
-buckets = {0 : [0, 2, 5, 1000], 2 : [2, 5, 1000], 5 : [5, 1000], 1000 : [1000] }
-tally = {}
+result = {tuple(provers) : 0 for provers in powerset(PROVERS)}
 
 for directory in os.listdir(TRANSLATION_ROOT):
     for translation in tqdm.tqdm(os.listdir(os.path.join(TRANSLATION_ROOT, directory)), desc=f"{directory}"):
-        translation_path = os.path.join(os.path.join(
-            TRANSLATION_ROOT, directory), translation)
-        with open(translation_path, "r") as translation_text:
-            try:
-                translation_text = translation_text.read()
-                q = translation_text.count("a,axiom,'f_!")
-                x = (translation_text.count("a,axiom,'f") - q)//2
-            except MemoryError:
-                q = 100
-                x = 100
-        i = min([b for b in buckets.keys() if b >= q])
-        j = min([b for b in buckets[i] if b >= x])
-        if not (i, j) in tally:
-            tally[(i, j)] = {p : 0 for p in PROVERS + ["total"]}
-        tally[(i, j)]["total"] += 1
+        solvedby = []
         for p in PROVERS:
             root = RESULT_ROOT + p
             result_path = os.path.join(os.path.join(root, directory), translation)
@@ -45,10 +37,10 @@ for directory in os.listdir(TRANSLATION_ROOT):
                 with open(result_path, "r") as result_text:
                     result_text = result_text.read()
                 if PROOF[p] in result_text or COUNTERMODEL[p] in result_text:
-                    tally[(i, j)][p] += 1
-            except FileNotFoundError:
+                    solvedby.append(p)
+            except:
                 pass
-for k, v in tally.items():
+        result[tuple(solvedby)] += 1
+
[…]

### 2022-10-13 FoSSaCS/fossacs-EIICL.tex
abstract

@@
 The famous double negation translation~\cite{glivenko1929quelques,godel1933intuitionistischen} establishes an embedding from classical into intuitionistic logic.
-Curiously the reverse direction has not been covered in literature.
-Based on the work in~\cite{claessen2015sat} we establish a small model property for intuitionistic propositional logic in order to utilize its Kripke semantics to give a simple embedding into classical propositional logic and quantified boolean formulas.
-We transfer there methods to the first order case and also give an embedding of intuitionistic first-order logic into classical first-order-logic.
-This allows a direct application of state-of-the-art classical first-order provers to the problem of determining intuitionistic validity.
-
-For certain classes of benchmarks our approach can compete with state-of-the-art provers, in particular when the intuitionistic content is low.
-Crucially, our constructions support the transfer of counter-models to validity, which is important to model checking applications, where the output of counter-examples is an appreciated feature.
+Curiously, the reverse direction has not been covered in literature.
+Utilizing a normal form for intuitionistic logic~\cite{claessen2015sat}, we establish a small model property for intuitionistic propositional logic.
+We use this property for a direct encoding of the Kripke semantics into classical propositional logic and quantified boolean formulas.
+Next, we transfer the developed techniques to the first order case and provide an embedding of intuitionistic first-order logic into classical first-order-logic.
+Our goal here is an encoding that facilitates the use of state-of-the-art provers for classical first-order logic for determining intuitionistic validity.
+In an experimental evaluation, we show that our approach can compete with state-of-the-art provers for certain classes of benchmarks, in particular when the intuitionistic content is low.
+We further note that our constructions support the transfer of counter-models to validity, which is an desired feature in model checking applications.
 \end{abstract}

### 2022-10-13 FoSSaCS/fossacs-EIICL.tex
asdf

@@
 the ILTP library website~\cite{iltp}).
-However, intuitionistic provers perform poorly compared to their classical counterparts, see e.g. the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
+The progress in automated reasoning for intuitionistic logic, however, has been slower than the  impressive advances in solvers for classic logics --- evidenced, e.g., by the CASC~\cite{casc} and SAT~\cite{satc} competitions.
 %, e.g., the first-order provers~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}
@@
 Furthermore all the most popular first-order utilize classical logic.
-This difference can be partially explained by fundamental differences between the logics.
-First of all, determining intuitionistic validity is computationally harder, i.e.
-in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
-A further advantage of classical logic is the existence of calculi that are particularly suited for automation such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability, i.e.
-to show the validity of a formula it suffices to show the unsatisfiability of the negated formula, which is insufficient in intuitionistic logic.
-The first dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient,tammet1996resolution} used the naive inverse method, a direct search for a cut-free proof by applying the rules from some proof calculus inversely, which generally leads to a very complex search.
-More recently connection-based methods have been applied to various non-classical logics~\cite{otten2005clausal,otten2021nanocop}, including intuitionistic logic.
+This difference can partially be explained by fundamental differences between the logics.
+First of all, determining intuitionistic validity is computationally harder, e.g., in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic}, whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
+A further advantage of classical logic is the existence of calculi that are particularly suited for automation, such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability (i.e., in order to show the validity of a formula it suffices to show the unsatisfiability of the negated formula, which is insufficient in intuitionistic logic).
+The first dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient,tammet1996resolution} used the naive inverse method, i.e., a direct search for a cut-free proof by applying the rules from some proof calculus inversely, which generally leads to a very complex search.
+More recently, connection-based methods have been applied to various non-classical logics~\cite{otten2005clausal,otten2021nanocop}, including intuitionistic logic.
 There have also been some successful attempts to study intuitionistic validity via embedding into higher-order classical logic~\cite{LEO}.
-Finally, we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic.
+We finally add that, in contrast to intuitionistic provers, a tremendous amount of work has been put into optimizing provers for classical logic.
 
-With this work we want to examine a new but evident approach for directly leveraging the progress in classical reasoning for intuitionistic logic.
+With this work we want to leverage the progress in classical reasoning for intuitionistic logic.
 To this end, we give an embedding from intuitionistic logic into classical logic. 

### 2022-10-13 FoSSaCS/fossacs-EIICL.tex
intro

@@
 \usepackage[T1]{fontenc}
+\usepackage{todonotes}
 % T1 fonts will be used to generate the final print and online PDFs,
@@
 \begin{abstract}
-The famous double negation translation~\cite{glivenko1929quelques,godel1933intuitionistischen} establishes an embedding from classical into intuitionistic logic.
+The famous double negation translation~\cite{glivenko1929quelques,godel1933intuitionistischen} establishes an embedding of classical into intuitionistic logic.
 Curiously, the reverse direction has not been covered in literature.
@@
 In an experimental evaluation, we show that our approach can compete with state-of-the-art provers for certain classes of benchmarks, in particular when the intuitionistic content is low.
-We further note that our constructions support the transfer of counter-models to validity, which is an desired feature in model checking applications.
+We further note that our constructions support the transfer of counter-models to validity, which is a desired feature in model checking applications.
 \end{abstract}
@@
 With this work we want to leverage the progress in classical reasoning for intuitionistic logic.
-To this end, we give an embedding from intuitionistic logic into classical logic. 
-That is, we give an effective procedure that returns for each formula $\varphi$ a modified formula $\varphi^\#$, such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
-Interestingly, the reverse direction, the famous double-negation translation, has long been established and goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
-In the propositional case it is particularly simple: $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
-Intuitively the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent.
-This gives us a first idea why the reverse direction is perhaps more difficult: We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e.
-if they both occur in $\varphi$, we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$.
-In particular, this necessitates the introduction of new propositional variables in our procedure, which already marks a big difference to the double negation translation.
-%
-In addition to the translation of formulas we give an effective translation of counter-models, i.e.
-for each intuitionistic counter-model of $\varphi$ we can effectively construct a classical counter-model of $\varphi^\#$ and vice versa.
-We note that the existence of counter-models is a key notion that forms a proper dual to validity --- whereas the satisfiability of the negation is not necessary for invalidity (in contrast to classical logic, where it is a proper dual to validity).
-Transforming and reducing counter-models to a normal form is also what ultimately enables our translation.
-
-In summary, our work establishes and then examines the following approach to theorem proving for intuitionistic logic:
+To this end, we propose the following methodology:
 \begin{itemize}
-	\item Translate the input formula $\varphi$ to $\varphi^\#$, which is classically valid if and only if $\varphi$ is intuitionistically valid.
+	\item Translate an input formula $\varphi$ to a formula $\varphi^\#$, which is classically valid if and only if $\varphi$ is intuitionistically valid.
 	\item Determine the classical validity of $\varphi^\#$ using a prover for classical logic, resulting in a proof or counter-model.
@@
 \end{itemize}
-
-\subsection*{Main Contributions}
+\textbf{Contributions.} We now overview the main results of this paper (which we discuss in more detail in Sections~\ref{sec:overview} and~\ref
 \begin{itemize}
-	\item Establishing a small model property for intuitionistic propositional logic.
-	\item Embedding intuitionistic propositional logic into classical propositional logic.
-	\item \resizebox*{0.96\hsize}{!}{Embedding intuitionistic propositional logic into quantified boolean formulas.}
-	\item Presenting a model reduction for intuitionistic first-order logic.
-	\item Embedding intuitionistic into classical first-order logic.
-	\item Benchmarking our new approach to intuitionistic theorem proving.
+	\item We establish a small model property for counter-models to validity in intuitionistic propositional logic.
+	\item Based on the small model property, we provide an embedding of intuitionistic propositional logic into classical propositional logic and into quantified boolean formulas, i.e., we provide an 
+	\item We discuss the reduction of counter-models in intuitionistic first-order logic and, based on this reduction, present an embedding intuitionistic into classical first-order logic.
+	\item We provide an experimental evaluation on the use of provers for classical first-order logic for determining intuitionistic validity.
+ We show that our approach can compete with state-of-the-art provers for certain classes of benchmarks, in particular when the intuitionistic content is low.
 \end{itemize}
@@
 \subsection*{Related work}
+The famous double-negation translation establishes
+an embedding of classical into intuitionistic logic, going back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
+In the propositional case the embedding is particularly simple: 
+A formula $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid. 
+Intuitively, the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent.
+For embedding intuitionistic into classical logic, we need to do the opposite:
+We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e., if they both occur in $\varphi$, we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$.
+This necessitates the introduction of new propositional variables in our procedure, which marks a big difference to the double negation translation.
 
-There are various known translations between logics.
-Most relevant are the double negation translation~\cite{glivenko1929quelques,gentzen1936widerspruchsfreiheit}, the embedding of intuitionistic logic into S4~\cite{basicprooftheory} and of various non-classical logics into higher-order classical logic~\cite{LEO,wisniewski2016tptp}.
+We further mention some translations between logics from the literature.
+Most relevant are the embedding of intuitionistic logic into S4~\cite{basicprooftheory} and of various non-classical logics into higher-order classical logic~\cite{LEO,wisniewski2016tptp} \todo{why are they relevant?}.
+Of special relevance is the reduction of intuitionistic propositional logic given in~\cite{claessen2015sat}
+\todo{reduction to what?} and examined in~\cite{fiorentini2019proof} \todo{what has been examined here?}, which we further expand upon in this paper.
 
-Of special relevance is the reduction of intuitionistic propositional logic given in~\cite{claessen2015sat} and examined in~\cite{fiorentini2019proof}, which we further expand upon.
-
+\todo{This paragraph is more about realizability, right? So remove, or connect to paper?}
 On the other hand there have been attempts to circumvent intuitionistic logic by equipping classical proofs with operational semantics~\cite{Control1,Parigot1}, which however does not permit in general the realizability of $\exists$ and $\vee$.
 
-There are some results relating intuitionistic and classical validity, most famously for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}, but also in some other cases~\cite{schwichtenberg}.
-However, they only apply in very limited settings.
+We finally mention a long-standing line of research that has investigated when classical validity implies intuitionistic validity.
+This most famously holds for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}, but also other conditions are known~\cite{schwichtenberg}. 
+In contrast, we propose a translation of an intuitionistic formula to an equi-valid classical formula.
 
-Finally, of course dedicated theorem provers for intuitionistic logic exists, generally using the inverse method~\cite{mclaughlin2008imogen,pavlov2017whaleprover}, but more recently also based on connection-based calculi~\cite{otten2008leancop,otten2021nanocop}, which also work well for classical logic.
+%Finally, of course dedicated theorem provers for intuitionistic logic exists, generally using the inverse method~\cite{mclaughlin2008imogen,pavlov2017whaleprover}, but more recently also based on connection-based calculi~\cite{otten2008leancop,otten2021nanocop}, which also work well for classical logic.
 
@@
 \section{Overview}
+\label{sec:overview}
 

### 2022-10-14 FoSSaCS/fossacs-EIICL.tex
pass over intro + experiments finished

@@
 With this work we want to leverage the progress in classical reasoning for intuitionistic logic.
-To this end, we propose the following methodology:
+To this end, we study embeddings of intuitionistic into classical logic and propose the following methodology:
 \begin{itemize}
@@
 \end{itemize}
-\textbf{Contributions.} We now overview the main results of this paper (which we discuss in more detail in Sections~\ref{sec:overview} and~\ref
+\subsection*{Contributions} 
+We now overview the main results of this paper (which we discuss in more detail in Sections~\ref{sec:overview} and~\ref{sec:evaluation}):
 \begin{itemize}
 	\item We establish a small model property for counter-models to validity in intuitionistic propositional logic.
-	\item Based on the small model property, we provide an embedding of intuitionistic propositional logic into classical propositional logic and into quantified boolean formulas, i.e., we provide an 
-	\item We discuss the reduction of counter-models in intuitionistic first-order logic and, based on this reduction, present an embedding intuitionistic into classical first-order logic.
-	\item We provide an experimental evaluation on the use of provers for classical first-order logic for determining intuitionistic validity.
+	\item Based on the small model property, we provide an embedding of intuitionistic propositional logic into classical propositional logic and into quantified boolean formulas.
+	\item We discuss the reduction of counter-models in intuitionistic first-order logic and, based on this reduction, present an embedding of intuitionistic into classical first-order logic.
+	\item We provide an experimental evaluation on the use of first-order provers for classical logic for determining intuitionistic validity.
  We show that our approach can compete with state-of-the-art provers for certain classes of benchmarks, in particular when the intuitionistic content is low.
@@
 \begin{lemma}
-	For each propositional formula $\varphi$ there is effectively a formula $\varphi^\#$ with $\mathcal |\varphi^\#|\in\mathcal O(|\mathcal R|\cdot\exp(|\mathcal X|\log|\mathcal X|))$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
+	For each propositional formula $\varphi$ there is effectively a formula $\varphi^\#$ with $\mathcal |\varphi^\#|\in\mathcal O(|\mathcal R|\cdot 2^{|\mathcal X|\log|\mathcal X|})$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
 \end{lemma}
@@
 
-A detailed account of the translation can be found in definition~\ref{def:fo-translation}.
+A detailed account of the translation can be found in Definition~\ref{def:fo-translation}.
 Finally, we benchmark our translation by translating all problems in the ILTP problem set~\cite{iltp} and applying the Vampire theorem prover~\cite{kovacs2013first} to the translated instances.
@@
 
-\section{Benchmark}
+\section{Experimental Evaluation}
+\label{sec:evaluation}
 
@@
 
-Two versions of the translation were tested, one with and one without the additional axioms encoding idempotence.
-The results can be seen in Figure \ref{fig:a}.
-In general our approach falls short of the state of the art.
-However when binning for $|\mathcal Q|$ and $|\mathcal X|$ as seen in Figure~\ref{fig:b} we observe good performance for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 3$.
-Figure~\ref{fig:c} shows that we solve 52 problems not solved by the existing provers.
+Two versions of the translation were tested, one with and one without the additional axioms encoding idempotence. The results are stated in Figure \ref{fig:a}. 
+In general, our approach falls short of the state of the art. 
+However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},
+we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$ resp. $|\mathcal X| \geq 6$.
+\todo{Note that I added $\ge 6$.}
+Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers.
 \begin{figure}
@@
 	\end{tabular}
-	\caption{Solved problems binned by $|\mathcal Q|$ and $|\mathcal X|$.}
+	\caption{Solved problems grouped by $|\mathcal Q|$ and $|\mathcal X|$.}
 	\label{fig:b}
@@
 
- Therefore our tool is already useful if it is known that $|\mathcal X|$ and $|\mathcal Q|$ are small.
-We see that idempotence does not improve performance.
-On the other hand from a model-theoretic perspective it is quite powerful.
+We conclude that our approach already has merits when $|\mathcal X|$ and $|\mathcal Q|$ are small. 
+We see that encoding idempotence does help to solve some additional benchmarks, but in general does not improve performance.
+However, idempotence is a quite powerful concept from a model-theoretic perspective.
 We suspect that a more clever encoding of this property could improve our results.

### 2022-10-14 FoSSaCS/fossacs-EIICL.tex
incoprporated notes

@@
 We further mention some translations between logics from the literature.
-Most relevant are the embedding of intuitionistic logic into S4~\cite{basicprooftheory} and of various non-classical logics into higher-order classical logic~\cite{LEO,wisniewski2016tptp} \todo{why are they relevant?}.
-Of special relevance is the reduction of intuitionistic propositional logic given in~\cite{claessen2015sat}
-\todo{reduction to what?} and examined in~\cite{fiorentini2019proof} \todo{what has been examined here?}, which we further expand upon in this paper.
-
-\todo{This paragraph is more about realizability, right? So remove, or connect to paper?}
-On the other hand there have been attempts to circumvent intuitionistic logic by equipping classical proofs with operational semantics~\cite{Control1,Parigot1}, which however does not permit in general the realizability of $\exists$ and $\vee$.
+Most relevant are the embedding of intuitionistic logic into S4~\cite{basicprooftheory} and of various non-classical logics into higher-order classical logic~\cite{LEO,wisniewski2016tptp}.
+Of special relevance are the normalization procedures for intuitionistic propositional logic given in~\cite{claessen2015sat}
+and examined from a proof-theoretic perspective in~\cite{fiorentini2019proof}, which we further expand upon in this paper.
 
@@
 we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$ resp. $|\mathcal X| \geq 6$.
-\todo{Note that I added $\ge 6$.}
-Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers.
+Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers.\pagebreak
 \begin{figure}
@@
 \end{figure}
+\vspace*{-.5cm}
 

### 2022-10-14 FoSSaCS/fossacs-EIICL.tex
Update fossacs-EIICL.tex

@@
 	\begin{tikzpicture}
-		\begin{scope}[opacity=0.3, blend mode=multiply]
+		\begin{scope}[opacity=0.1, blend mode=multiply]
 			\def\firstcircle{(90:.75) circle (1.5)}
@@
 	\vspace*{-7.1cm}
-	\caption{Symmetric difference. Instances proven by Embedding (red), Embedding (idem.) (green), ileanCoP or nanoCoP (blue)}
+	\caption{Instances proven by combinations of Embedding (red), Embedding (idem.) (green), ileanCoP or nanoCoP (blue)}
 	\label{fig:c}
@@
 we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$ resp. $|\mathcal X| \geq 6$.
-Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers.\pagebreak
+Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers.
+Notably, while there is great overlap in solved instances, ileanCoP and nanoCoP also have a significant symmetric difference.\pagebreak
 \begin{figure}
@@
 
-We conclude that our approach already has merits when $|\mathcal X|$ and $|\mathcal Q|$ are small. 
+We conclude that our approach already has merits when $|\mathcal Q| = 0$ and $|\mathcal X|$ is small. 
 We see that encoding idempotence does help to solve some additional benchmarks, but in general does not improve performance.

### 2023-03-06 FoSSaCS/rebuttal.txt
LPAR

Thank you for your insightful reviews. We will take the feedback into account for preparing the final version. We now address the points raised by the reviewers:

R1, "erroneous translations":

We agree that the translation is incomplete as given. However, this issue is a minor mistake in the write-up and can be easily be fixed (moreover, our implementation already handles the observed issue correctly): That the translation is incomplete is due to the fact that definition 9 handles universally quantified variables differently than free variables (which are not translated at all). This incompleteness arises from the fact that the translation is only intended to be used on sentences. However, the inductive nature of definition 9 necessitates that φᶜ be defined even when not all variables are quantified. With this in mind we can easily fix the problem by changing definition 9 analogously to definition 5, i.e. replacing all free variables with bound ones and universally quantifying over them before applying the transformation. We note that our tool already introduces the necessary universal quantifiers during the normalization, and so is not affected by the observed issue. We further note that the results of our tool are validated against the known results of the ILTP set, which increases the confidence in the correctness of our translation (and indeed allowed us to fix bugs during implementation).

R1, "lack of clarity and rigour":

We believe that the level of technical detail of our paper is on par with related papers from the literature. We judged that, besides taking up precious space, the inclusion/repetition of standard definitions would occlude the contributions of our paper. Such omissions are quite usual, even in papers where syntax is relevant, e.g. [1, 2, 3]. In particular notions such as atomic formulas or ground expressions are rarely ever defined.  While our definitions of semantics suggest the underlying languages we concede that they should be made explicit.

 [1] Otten, Jens. "Clausal connection-based theorem proving in intuitionistic first-order logic." International Conference on Automated Reasoning with Analytic Tableaux and Related Methods. Springer, Berlin, Heidelberg, 2005.

 [2] Parigot, Michel. "λμ-calculus: an algorithmic interpretation of classical natural deduction." International Conference on Logic for Programming Artificial Intelligence and Reasoning. Springer, Berlin, Heidelberg, 1992.

 [3] Baaz, Matthias and Iemhoff, Rosalie. "The Skolemization of existential quantifiers in intuitionistic logic." Annals of Pure and Applied Logic 142.1-3 (2006): 269-295.

We are thankful for the points raised by the reviewer and we will do our best to integrate them, which will improve the quality of our paper.

We address a few selected points by the reviewer:

- We agree that a definition of the formal languages should be noted and we will give either an exact definition or reference.

- As correctly observed, the paper was written with distinct sets of free and bound variables in mind, with letters a, b usually indicating free variables and u, w, x, y bound ones (see e.g. definition 5). We agree that the distinction between free and bound variables should be noted.

- We cannot fully understand the criticism of Definition 7:
 * Cᴵ(m) is true iff m ∈ Cᴵ, similarly Aᴵ(m) → Bᴵ(m) is false iff m ∈ Aᴵ and not m ∈ Bᴵ. While one could be more deliberate in their terminology, this is not so unusual.
 * The notation [ψ] was introduced precisely so that f^I_{[ψ]} would be defined for ψ ∈ X, since f_ψ is only defined for ψ ∈ X^{uS} in Lemma 6.
 * s is also introduced in Lemma 6.
 * Definition 7 defines for every counter-model M of ψ^{CH} a modified model M_T. Corollary 1 is in line with this.

- We will incorporate the suggestions by the reviewer to improve clarity, e.g. requiring certain sets to be finite.

R2, "use of QBF solvers for the propositional case": We agree that it would be interesting to test the performance of our translation for the propositional case. However the interest in intuitionistic propositional logic is lower than for the first-order case, where applications are more evident, e.g. program extraction. Therefore we focused our evaluation efforts on the first-order case.

R3, "However, it isn't clear whether these adjustments would indeed close the gap.": Indeed, the results do not demonstrate that our approach has the potential to outperform other algorithms on the general problem. However, there is sufficient evidence that on certain subsets it can, namely at the very least for |Q| = 0 and |X| < 3 (referring to Fig. 3). Note in particular that our tool is quite efficient in applying a classical solver if |Q| = |X| = 0. It could be useful as part of a portfolio solver even at this point.

### 2023-03-06 LPAR/LPAR-EIICL.tex
LPAR

--- 2022-10-14 FoSSaCS/fossacs-EIICL.tex
@@
-\documentclass[runningheads]{llncs}
-\usepackage{amsmath, amsfonts}
+\documentclass{easychair}
+\usepackage{amsmath, amsfonts, amsthm}
 \usepackage{wrapfig}
@@
 \usepackage{todonotes}
-% T1 fonts will be used to generate the final print and online PDFs,
-% so please use T1 fonts in your manuscript whenever possible.
-% Other font encondings may result in incorrect characters.
-%
-\usepackage{graphicx}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following two lines
-% to display URLs in blue roman font according to Springer's eBook style:
-%\usepackage{color}
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
 %
@@
 \usetikzlibrary{shapes,backgrounds}
+
+\theoremstyle{definition}
+\newtheorem{theorem}{Theorem}[section]
+\theoremstyle{definition}
+\newtheorem{corollary}[theorem]{Corollary}
+\theoremstyle{definition}
+\newtheorem{lemma}[theorem]{Lemma}
+\theoremstyle{definition}
+\newtheorem{proposition}[theorem]{Proposition}
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\theoremstyle{definition}
+\newtheorem{example}[theorem]{Example}
+\theoremstyle{definition}
+\newtheorem{remark}[theorem]{Remark}
 
@@
 	\authorrunning{A. Pluska and F. Zuleger}
+	\titlerunning{Embedding Intuitionistic into Classical Logic}
 	% First names are abbreviated in the running head.
@@
 
-Constructive mathematics refers to a flavour of mathematics in which the existence of an object can only be established by explicit construction, as opposed to classical mathematics where existence can be shown implicitly, e.g.
+Constructive mathematics refers to a flavor of mathematics in which the existence of an object can only be established by explicit construction, as opposed to classical mathematics where existence can be shown implicitly, e.g.
 by assuming non-existence and deriving a contradiction.
@@
 A further advantage of classical logic is the existence of calculi that are particularly suited for automation, such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability (i.e., in order to show the validity of a formula it suffices to show the unsatisfiability of the negated formula, which is insufficient in intuitionistic logic).
-The first dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient,tammet1996resolution} used the naive inverse method, i.e., a direct search for a cut-free proof by applying the rules from some proof calculus inversely, which generally leads to a very complex search.
+The first dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient,tammet1996resolution} used the naïve inverse method, i.e., a direct search for a cut-free proof by applying the rules from some proof calculus inversely, which generally leads to a very complex search.
 More recently, connection-based methods have been applied to various non-classical logics~\cite{otten2005clausal,otten2021nanocop}, including intuitionistic logic.
@@
 
-In this section we fix notation and recall the semantics for classical and intuitionistic logic. Futhermore we recapitulate some results from~\cite{otten2005clausal}.
+In this section we fix notation and recall the semantics for classical and intuitionistic logic. Furthermore we recapitulate some results from~\cite{otten2005clausal}.
+
+\subsection{Syntax}
+
+For propositional logic we assume an ambient set of propositional variables. A propositional formula is defined inductively as follows:
+\begin{itemize}
+	\item The propositional constant $\bot$ is a formula.
+	\item Each propositional variable is a formula.
+	\item Given formulas $\varphi$ and $\psi$ and a connector $\circ\in\{\wedge, \vee,\to\}$ there is a formula $\varphi\circ\psi$.
+\end{itemize}
+
+For predicate logic we assume an ambient signature $\Sigma$, i.e., a set of function and predicate symbols with their respective arities, and sets of free and bound variables.
+Generally free variables will be denoted will letters at the beginning of the alphabet while bound variables will be denoted with letters at the end of the alphabet.
+$\nu$ will usually be a placeholder that can be instantiated with free or bound variables.
+We can define terms $t$ and substitutions $t[\nu/b]$ by simultaneous induction as follows:
+\begin{itemize}
+	\item Each free variable $a$ is a term and $a[\nu/b] := \begin{cases}
+		\nu & \text{if } a = b \\
+		a & \text{otherwise}
+	\end{cases}$.
+	\item For each $n$-ary function symbol $f$ and terms $t_1, \ldots, t_n$ we have that $f(t_1, \ldots, t_n)$ is a term and $f(t_1, \ldots, t_n)[\nu/b] := f(t_1[\nu/b], \ldots, t_n[\nu/b])$.
+\end{itemize}
+We can then define formulas $\varphi$ and substitutions $\varphi[\nu/b]$ by simultaneous induction as follows:
+\begin{itemize}
+	\item The constant $\bot$ is a formula and $\bot[\nu/b] := \bot$.
+	\item For each $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$ we have that $P(t_1, \ldots, t_n)$ is a formula and $P(t_1,\ldots, t_n)[\nu/b] := P(t_1[\nu/b], \ldots, t_n[\nu/b])$.
+	\item Given formulas $\varphi$ and $\psi$ and a connector $\circ\in \{\wedge,\vee,\to\}$ there is a formula $\varphi\circ\psi$ and $(\varphi\circ\psi)[\nu/b] := \varphi[\nu/b]\circ\psi[\nu/b]$.
+	\item Given a formula $\varphi$, a free variable $a$, a bound variable $x$ not occurring in $\varphi$ and a quantifier $Q\in\{\exists,\forall\}$ there is a formula $Qx(\varphi[x/a])$ and $Qx(\varphi[x/a])[\nu/b] := Qx(\varphi[x/a][\nu/b])$.
+\end{itemize}
 
@@
 		\item to each $n$-ary function symbol $f\in\Sigma$ a $n$-ary function $f^I: M^n\to M$.
-		\item to each $n$-ary relation symbol $R\in\Sigma$ a $n$-ary relation $R^I\subseteq M^n$.
+		\item to each $n$-ary predicate symbol $R\in\Sigma$ a $n$-ary predicate $R^I\subseteq M^n$.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
-		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$.
+		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $\mathcal M_w, v\not\models\varphi$ or $\mathcal M_w, v\models\psi$.
+		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $\mathcal M_w, v[m/a]\models\varphi[a/x]$.
 	\end{itemize}
@@
 $\varphi$ is valid if it is satisfied by every Kripke structure.
-	$\varphi$ is valid for some frame (write $W\models\varphi$) if it is satisfied in every Krike structure with that frame.
+	$\varphi$ is valid for some frame (write $W\models\varphi$) if it is satisfied in every Kripke structure with that frame.
 We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
@@
 
-\begin{definition}
-	For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to a set of variables $Z$ by simultaneous induction:
+\begin{definition}\label{definition:skolemization}
+	For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to a set of free variables $Z$ by simultaneous induction:
 	\begin{itemize}
@@
 \begin{lemma}\label{lemma:first-order-normal-form}
-	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R,\mathcal X, \mathcal Q$, with $\mathcal R$ containing \emph{flat clauses} of the form
+	For every predicate formula $\varphi$ there exists a nullary predicate symbol $P$ as well as sets of clauses $\mathcal R,\mathcal X, \mathcal Q$, with $\mathcal R$ containing \emph{flat clauses} of the form
 	$$\forall \vec x\left(\bigwedge_i A_i\to \bigvee_jB_j\right),$$
@@
 This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$.
-In general this formula will contain new function and relation symbols.
+In general this formula will contain new function and predicate symbols.
 \end{lemma}
@@
 	There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
-The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary relation symbol in $\varphi$ there is a corresponding $n+1$-ary relation symbol in $\varphi^\#$, and $\varphi$ contains a new binary predicate $E$ as well as a number of new function symbols.
+The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary predicate symbol in $\varphi$ there is a corresponding $n+1$-ary predicate symbol in $\varphi^\#$, and $\varphi$ contains a new binary predicate $E$ as well as a number of new function symbols.
 Furthermore, there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
@@
 For every propositional variable $A$ consider a unary predicate $A$ of the same name where $A(u)$ expresses that $A$ is true at some world $u$.
-We can then consider the following naive encoding:
+We can then consider the following naïve encoding:
 
 \begin{definition}
-	Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively:
+	Let $\varphi$ be a propositional formula. Define $\varphi^{\nu}$ inductively:
 	\begin{itemize}
-		\item $\bot^u = \bot$.
-		\item $A^{u} = A(u)$ for every propositional variable $A$.
-		\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
-		\item $(\varphi\to \psi)^u = \forall w(u\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new variable.
+		\item $\bot^\nu = \bot$.
+		\item $A^{\nu} = A(\nu)$ for every propositional variable $A$.
+		\item $(\varphi\circ\psi)^\nu = \varphi^\nu\circ\psi^\nu$ for $\circ\in\{\wedge, \vee\}$.
+		\item $(\varphi\to \psi)^\nu = \forall w(\nu\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new bound variable.
 	\end{itemize}
@@
 	Then define
-	$$\varphi^{C} = K(\varphi)\to \forall u\varphi^{u}.$$
+	$$\varphi^{C} = K(\varphi)\to \varphi^{b}$$
+	where $b$ is a new free variable.
 \end{definition}
@@
 \begin{itemize}
-	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^u$ is
-	$$\forall w\left(u\preceq w\to\bigwedge_i A_i(w)\to\bigvee_jB_j(w)\right).$$
-	\item For implication clauses $((A\to B)\to C)^u$ is
-	$$\forall w(u\preceq w\to(\forall k(w\preceq k\to A(k)\to B(k)))\to C(w))$$
+	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^\nu$ is
+	$$\forall w\left(\nu\preceq w\to\bigwedge_i A_i(w)\to\bigvee_jB_j(w)\right).$$
+	\item For implication clauses $((A\to B)\to C)^\nu$ is
+	$$\forall w(\nu\preceq w\to(\forall k(w\preceq k\to A(k)\to B(k)))\to C(w))$$
 \end{itemize}
 
-and after applying Herbrandization we end up with a formula $$\varphi^{CH} = K(\varphi)\to\left(\bigwedge\mathcal R^{uS}\wedge \bigwedge\mathcal X^{uS}\right)\to P(s)$$ where $s$ is a new constant symbol, each formula in $\mathcal R^{uS}$ is of the form
+and after applying Herbrandization we end up with a formula $$\varphi^{CH} = K(\varphi)\to\left(\bigwedge\mathcal R^{b S}\wedge \bigwedge\mathcal X^{b S}\right)\to P(s)$$ where $s$ is a new constant symbol, each formula in $\mathcal R^{b S}$ is of the form
 $$\forall w\left(s\preceq w\to\bigwedge_iA_i(w)\to\bigvee_jB_j(w)\right)$$
-and each formula $\psi\in\mathcal X^{uS}$ is of the form
+and each formula $\psi\in\mathcal X^{b S}$ is of the form
 $$\forall w(s\preceq w\to(w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w)))\to C(w))$$
@@
 \begin{definition}
-For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{uS}$. Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
+For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{b S}$. Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
 We say $(A\to B)\to C\in\mathcal X$ is \emph{fulfilled} at $m\in M$ iff $A^I(m)\to B^I(m)$ is false or $C^I(m)$ is true. For $\psi\in\mathcal X$ define $$g_\psi : M\to M, m\mapsto\begin{cases}
-		m,&\text{ if $\psi$ is fulfilled at $u$,}\\
+		m,&\text{ if $\psi$ is fulfilled at $m$,}\\
 		f^I_{\lbrack\psi\rbrack}(m),&\text{ else.}		
@@
 	$$\varphi^Q_{|\mathcal{X}|} = \exists \vec A^{|\mathcal{X}|}\left(\text{\normalfont Persistent}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal R}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal X}(|\mathcal{X}|)\right)$$
-	for leafs, and $$\varphi^Q = \varphi_0^Q = \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
+	for leaves, and $$\varphi^Q = \varphi_0^Q = \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
 	Example encodings of the above formulas are:
@@
 We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$.
-To encode that some $n$-ary relation $A$ holds at a world $u$ we extend each $n$-ary relation symbol to a $n+1$-ary relation symbol of the same name, interpreting $A(\vec x, u)$ as "$A(\vec x)$ holds at u".
+To encode that some $n$-ary predicate $A$ holds at a world $u$ we extend each $n$-ary predicate symbol to a $n+1$-ary predicate symbol of the same name, interpreting $A(\vec x, u)$ as "$A(\vec x)$ holds at u".
 Furthermore we write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$.
@@
 \begin{definition}\label{def:fo-translation}
-	Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively:
+	Let $\varphi$ be a predicate formula. Define $\varphi^{\nu}$ inductively:
 	\begin{itemize}
-		\item $\bot^u = \bot$.
-		\item $A(\vec x)^{u} = A(\vec x, u)$ for every relation symbol $A$.
-		\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
-		\item $(\varphi\to \psi)^u = \forall w(u\preceq w\to\varphi^{w}\to\psi^{w})$
-		\item $(\exists x\varphi)^u = \exists x(E(x,u)\wedge\varphi^u)$
-		\item $(\forall x\varphi)^u = \forall x\forall w(u\preceq w\to E(x, w)\to \varphi^w)$
+		\item $\bot^\nu = \bot$.
+		\item $A(\vec t)^{\nu} = A(\vec t, \nu)$ for every predicate symbol $A$.
+		\item $(\varphi\circ\psi)^\nu = \varphi^\nu\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
+		\item $(\varphi\to \psi)^\nu = \forall w(\nu\preceq w\to\varphi^{w}\to\psi^{w})$.
+		\item $(\exists x\varphi)^\nu = \exists x(E(x,\nu)\wedge\varphi[a/x]^\nu[x/a])$ where $a$ is a new free variable.
+		\item $(\forall x\varphi)^\nu = \forall w(i\preceq w\to \forall x(E(x, w)\to \varphi[a/x]^w[x/a]))$ where $a$ is a new free variable and $w$ is a new bound variable.
 	\end{itemize}
-	where $w$ is some new variable. Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
+	where $w$ is some new bound variable. Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
 	\begin{align*}
@@
 		\text{DomainClosed}(u) = &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
-		\text{Persistent}(u, w) = &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a relation in $\varphi$}\right\}
+		\text{Persistent}(u, w) = &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a predicate in $\varphi$}\right\}
 	\end{align*}
 	Then define
-	$$\varphi^C = K(\varphi)\to\forall u(\text{World}(u)\to \varphi^u).$$
+	$$\varphi^C = K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b$$where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
 \end{definition}
@@
 	\begin{itemize}
-		\item $\preceq^I = \leq$ when restricted to $W_0$ and $\prec$ can be interpreted as an arbitrary PO otherwise.
+		\item $\preceq^I = \leq\cup \{(m, m): m\in M\setminus W_0\}$ as a set, i.e. $\preceq$ corresponds to $\leq$ on $W_0$ and is trivial elsewhere.
 		\item $E^I(m, w)$ iff $w\in W_0$ and $m = [x, w]$ for some $x\in M_w$.
@@
 			A^{I_u}(x_1\dots x_n), &\substack{\text{if $u\in W_0$ and $\exists x_i\in M_u$}\\\text{with $m_i= [x_i, u]$ for all $i$,}}\\
-			\top, & \text{otherwise.}
+			\bot, & \text{otherwise.}
 		\end{cases}$
 	\end{itemize}
-	One easily verifies that these are well-defined.
-Let us now briefly check that this indeed defines a counter-model.
-Clearly World$(w_0)$ holds and $K(\varphi)$ also holds.
-All that remains to show is that $M, I, v[w_0/u]\not\models\varphi^u$.
-We proceed by induction on the height of $\varphi$.
-Consider e.g.
-the case where $\varphi$ is of the form $\forall x\psi$.
-	We have to show that
-	$$M, I, v[w_0/u]\models \forall x\forall w(u\preceq w\to E(x, w)\to \psi^w)$$
-	does not hold for some $v$.
-Now since $\varphi$ is intuitionistically invalid there exists $w_1\in W_0$ and $m\in M_{w_1}$ such that $w_1, v[m/x]\not\models\psi$.
-By induction hypothesis $M, I, v[w_1/w]\not\models\psi^w$ for some $v$.
-By definition $M, I, v[w_1/w, [m, w_1]/x]\models E(x, w)$ and $w_0\preceq^I w_1$ holds.
-Therefore indeed $$M, I, v[w_0/u, w_1/w, [m, w_1]/x]\not\models \forall w(u\preceq w\to E(x, w)\to \psi^w)$$ holds and we are done.
-The other cases are analogous.\hfill$\square$
+	One easily verifies that these are well-defined. We now show by induction on the height of the formula the following:
+
+	\emph{Claim.} For any formula $\psi$ in which the free variable $b$ doesn't occur, world $w_k\in W_0$ and variable assignments $v_c, v_i$ such that $v_c(a) = [v_k(a), w_k]$ for all free variables $a$ occurring in $\psi$ it holds that \[M, I, v_c[w_k/b]\models \psi^b\text{ if and only if $w_k, v_i\models \psi$.}\]
+
+	For the base case, i.e. if $\psi = P(t_1,\dots, t_n)$ for some $n$-ary predicate symbol $P$, the claim follows directly from the definitions.
+
+	For the induction step consider e.g. the case where $\psi$ is of the form $\forall x \psi'$.
+	Suppose first that $w_k, v_i\not\models \forall x\psi'$, i.e.
+there exists $w_l\geq w_k$ and $m_l\in w_l$ such that $w_l, v_i[m_l/a]\not\models \psi'[a/x]$.
+Then by the induction hypothesis $M, I, v_c[[v_i(a), w_l],w_l/c]\not\models \psi'[a/x]^c$.
+Furthermore by definition $M, I, v_c[w_l/c, w_k/b]\models b\preceq c$ and $M, I, v_c[v_c[[v_i(a), w_l],w_l/c]\models E(a, c)$.
+Hence $M, I, v_c[w_k/b]\not\models \psi^b$.
+Suppose on the other hand that $w_k, v_i\not\models\forall x\psi'$, i.e.
+for all $w_l\geq w_k$ and $m_l\in w_l$ it holds that $w_l, v_i[m_l/a]\models \psi'[a/x]$.
+Then by the induction hypothesis $M, I, v_c[[m_l, w_l]/A,w_l/c]\models \psi'[a/x]^c$ for all $w_l\geq w_k$ and $m_l\in w_l$.
+It can be shown that for every other assignment of $c$ and $a$ either $b\preceq c$ or $E(a, c)$ is violated.
+Hence $M, I, v_c[w_k/b]\models \psi^b$.
+For other forms of $\psi$ one proceeds in similar fashion.
+This completes the proof of the claim.
+
+	Recall that $w_0, v\not\models\varphi$.
+From the claim it follows that \[M, I, v'[[v(a_0), w_0]/a_0\dots[v(a_n), w_0]/a_0, w_0/b]\not\models \varphi^b\]where $a_0\dots a_n$ are the free variables occurring in $\varphi$ and $v'$ is any variable assignment.
+Furthermore a straightforward check of definitions yields \[M, I, v'[[v(a_0), w_0]/a_0\dots[v(a_n), w_0]/a_0, w_0/b]\models K(\varphi)\wedge \text{World}(b)\wedge\vec E(\vec a, b).\]Therefore $M, I\not\models \varphi^C$ and we are done.
+
 \end{proof}
 
-Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulas $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = \text{World}(s)\to K(\varphi)\to\left(\bigwedge R^{uS}\wedge \bigwedge X^{uS}\wedge\bigwedge Q^{uS}\right)\to P(s)$$
-where $s$ is a new constant symbol, each formula in $\mathcal R^{uS}$ is of the form
+Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulas $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = \text{World}(s)\to K(\varphi)\to\left(\bigwedge R^{bS}\wedge \bigwedge X^{bS}\wedge\bigwedge Q^{bS}\right)\to P(s)$$
+where $s$ is a new constant symbol, each formula in $\mathcal R^{bS}$ is of the form
 $$\forall w \forall \vec x(s\preceq w\to\vec E(x, w)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)),$$
-each formula $\psi\in\mathcal X^{uS}$ is of the form
+each formula $\psi\in\mathcal X^{bS}$ is of the form
 $$
@@
 $$
-and each formula $\psi\in\mathcal Q^{uS}$ is of the form
+and each formula $\psi\in\mathcal Q^{bS}$ is of the form
 $$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to \forall y(E(y, f_\psi(\vec x, w))\to A(\vec a, f_\psi(\vec x, w)))\to B(\vec B, w))$$
@@
 	\begin{itemize}
-		\item $M_T$ are the sequences on $\{(\vec z, \psi)\:|\:\psi\in \mathcal F, \vec z\in M^n\}$.
+		\item $M_T$ are the sequences on $\{(\vec m, \psi)\:|\:\psi\in \mathcal F, \vec m\in M^n\}$.
 		\item $r^{I_T} = s^{I_T} = \epsilon$.
 		\item We interpret $\preceq$ as the prefix-order.
-		\item We set $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi).$$
-		\item For relation symbols $R$, we set $${R}^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = {R}^I(\vec z, f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots)).$$
+		\item We set $$f_\psi^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)(\vec m, \psi).$$
+		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, f_{\psi_n}(\vec m_n, \dots(f_{\psi_1}(\vec m_1, s^I))\dots)).$$
 	\end{itemize}
@@
 
-Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$f_{\psi_1}(\vec z_1, f_{\psi_2}(\vec z_2, \dots f_{\psi_n}(\vec z_n, s)\dots))$$ is sufficient.
+Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$f_{\psi_1}(\vec m_1, f_{\psi_2}(\vec m_2, \dots f_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
 In particular, this allows us to completely eliminate $\preceq$, that is we remove the PartialOrder$(\preceq)$
@@
 $$
-	\bigwedge\{\forall\vec x\forall\vec y\forall u(A(\vec x, u)\to A(\vec x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q, \text{$A$ a relation in $\varphi$}\}.
+	\bigwedge\{\forall\vec x\forall\vec y\forall u(A(\vec x, u)\to A(\vec x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q, \text{$A$ a predicate in $\varphi$}\}.
 $$Putting all of this together yields the following translation:
@@
 	Let $\mathcal R,\mathcal X,\mathcal Q$ be as above.
-For every $n$-ary relation $A$ define a new $n+1$-ary relation of the same name.
+For every $n$-ary predicate $A$ define a new $n+1$-ary predicate of the same name.
 For every $\psi\in\mathcal \mathcal X\cup\mathcal Q$ define a new function symbol $f_\psi$.
-Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
+Consider a new binary predicate symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
 Obtain $\mathcal R^\#$ by including
@@
  	$$
- 		\forall u\forall \vec x\left(A(\vec a, f_\psi(\vec z, u))\to B(\vec b, f_\psi(\vec z, u))\to C(\vec c, u)\right)
+ 		\forall u\forall \vec x\left(A(\vec a, f_\psi(\vec x, u))\to B(\vec b, f_\psi(\vec x, u))\to C(\vec c, u)\right)
  	$$
@@
 	$$
-		\forall u\forall \vec x\left(\forall y\left(A(\vec a, f_\psi(\vec z, u))\right)\to B(\vec b, u)\right)
+		\forall u\forall \vec x\left(\forall y\left(A(\vec a, f_\psi(\vec x, u))\right)\to B(\vec b, u)\right)
 	$$
@@
 	$$
-	Then $\varphi$ is intuitionistically if and only if
+	Then $\varphi$ is intuitionistically valid if and only if
 	$$
@@
 		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ a function in $\varphi$}\}\right)\wedge\\
-		&\resizebox{.9\hsize}{!}{$\:\bigwedge\left\{\forall\vec z_1\forall \vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal X\cup\mathcal Q$, $A$ a relation in $\varphi$}\right\}$.}
+		&\resizebox{.9\hsize}{!}{$\:\bigwedge\left\{\forall\vec z_1\forall \vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal X\cup\mathcal Q$, $A$ a predicate in $\varphi$}\right\}$.}
 	\end{align*}
 	The size of the obtained formula is linear in the input.
-However $|\mathcal X| + |\mathcal Q| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation has been extended to a $n+1$-ary relation.
+However $|\mathcal X| + |\mathcal Q| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary predicate has been extended to a $n+1$-ary predicate.
 \end{theorem}
@@
 Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers.
-Notably, while there is great overlap in solved instances, ileanCoP and nanoCoP also have a significant symmetric difference.\pagebreak
+Notably, while there is big overlap in solved instances, ileanCoP and nanoCoP also have a significant symmetric difference.\pagebreak
 \begin{figure}
@@
 
-\bibliographystyle{splncs04}
+\bibliographystyle{plain}
 \bibliography{bibliography}
@@
 \begin{lemma}\label{proof:first-order-normal-form}
-	For every predicate formula $\varphi$ there exists a nullary relation symol $P$ as well as sets of clauses $\mathcal R$ of the form
+	For every predicate formula $\varphi$ there exists a nullary relation symbol $P$ as well as sets of clauses $\mathcal R$ of the form
 	$$\forall \vec x\left(\bigwedge_i A_i\to \bigvee_j B_j\right)$$
@@
 This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$.
-In general this formula will contain new function symbols and relational symbols.
+In general this formula will contain new function symbols and predicate symbols.
 \end{lemma}

### 2023-03-06 Lighning Talk/presentation.tex
LPAR

\documentclass{cubeamer}

\usepackage{trimclip}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{arrows,shapes,quotes}
\usepackage{amsmath,amsfonts}

\title{Intuitionistic Proofs from Classical Provers}
\subtitle{A new approach to program synthesis}
\author[Alexander Pluska]{Alexander Pluska}
\date{\today} % or whatever the date you are presenting in is
\institute[TU Wien]{TU Wien}
% \copyrightnotice{Published by the American Institute of Aeronautics and Astronautics, Inc., with permission}

\begin{document}

	\maketitle

	\begin{frame}{Motivation}		
		\begin{figure}[!tbp]
			\begin{subfigure}[t]{0.48\textwidth}
				\noindent We have a prover...
				\begin{center}
					\includegraphics[width=0.8\textwidth]{Vampire.png}
					\includegraphics[width=\textwidth]{CASC.png}
					\caption{CASC-28 Results}
				\end{center}
				\vfill
			\end{subfigure}
			\begin{subfigure}[t]{0.48\textwidth}
				\hfill
			\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}{Motivation}		
			\begin{figure}[!tbp]
				\begin{subfigure}[t]{0.48\textwidth}
					\noindent We have a prover...

					\begin{center}
						\includegraphics[width=0.8\textwidth]{Vampire.png}
						\includegraphics[width=\textwidth]{CASC.png}
						\caption{CASC-28 Results}
					\end{center}
					\vfill
				\end{subfigure}
				\begin{subfigure}[t]{0.48\textwidth}
					\hfill...and proofs are programs.
					\begin{center}
						\includegraphics[width=\textwidth]{proofs_as_programs.png}
					\end{center}
					\vfill
				\end{subfigure}
			\end{figure}
	\end{frame}

	\begin{frame}{Motivation}		
		\begin{figure}[!tbp]
			\begin{subfigure}[t]{0.48\textwidth}
				\noindent We have a {\color{red}classical} prover...
				\begin{center}
					\includegraphics[width=0.8\textwidth]{Vampire.png}
					\includegraphics[width=\textwidth]{CASC.png}
					\caption{CASC-28 Results}
				\end{center}
				\vfill
			\end{subfigure}
			\begin{subfigure}[t]{0.48\textwidth}
				\hfill...and {\color{red}intuitionistic} proofs are programs.
				\begin{center}
					\includegraphics[width=\textwidth]{proofs_as_programs.png}
				\end{center}
				\vfill
			\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}{Kripke Semantics}
		\begin{figure}
			\centering
			\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
				\node[squarednode]      (0)   [] {$ $};
				\node[squarednode]      (1r)   [above right=of 0] {{\color{green}$1$}};
				\node[squarednode]      (1l)   [below right=of 0] {{\color{red}$1$}};
				\node[squarednode]      (2r)   [right=of 1r] {{1, \color{green}$2$}};
				\node[squarednode]      (2l)   [right=of 1l] {{1, \color{red}$2$}};
				\node[squarednode]      (3r)   [right=of 2r] {{1, 2, \color{green}$3$}};
				\node[squarednode]      (3l)   [right=of 2l] {{1, 2, \color{red}$3$}};
				\node      (4r)   [right=of 3r] {{$\hdots$}};
				\node      (4l)   [right=of 3l] {{$\hdots$}};
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (0.north east) -- (1r.south west);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (0.south east) -- (1l.north west);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1r.east) -- (2r.west);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1l.east) -- (2l.west);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1l.north) -- (2r.south);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1r.south) -- (2l.north);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2r.east) -- (3r.west);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2l.east) -- (3l.west);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2l.north) -- (3r.south);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2r.south) -- (3l.north);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3r.east) -- (4r.west);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3l.east) -- (4l.west);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3l.north) -- (4r.south);
				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3r.south) -- (4l.north);
			\end{tikzpicture}
			\vfill
			\caption{Kripke counter-model to $\neg(\neg \forall xA(x)\vee\neg \forall xB(x))$.}
		\end{figure}
	\end{frame}

	\begin{frame}{Encoding Kripke Semantics - Propositional Case}
		\begin{theorem}
			\label{thm:reduction-propositional}
			Let $\mathcal S$ be a set of flat and implicational clauses, $\mathcal F_\to\subseteq\mathcal S$ denote the subset implicational ones and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$. For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$. Obtain $\mathcal S^\#$ by including:
			\begin{itemize}
				\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
				\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
				\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
				\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occur in $k$.
			\end{itemize}
			Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
		\end{theorem}
	\end{frame}

	\begin{frame}{Reducing the Encoding - Predicate Case}
		\begin{itemize}
			\item Essentially the same, but...
		\end{itemize}
	\end{frame}

	\begin{frame}{Implementation}
		\centering
		\begin{figure}
			\includegraphics[width=\textwidth]{implementation.png}
		\end{figure}
		$\sim$3000 lines of rust code for the translation + $\sim$400 lines of python benchmarking
	\end{frame}

	\begin{frame}{Benchmark}

		\begin{figure}
			\begin{subfigure}[8]{0.4\textwidth}
				\def\firstcircle{(90:.75) circle (1.5)}
				\def\secondcircle{(210:.75) circle (1.5)}
				\def\thirdcircle{(330:.75) circle (1.5)}
				\begin{tikzpicture}
					\begin{scope}[shift={(3cm,-5cm)}, opacity=0.5, blend mode=multiply]
						\fill[red] \firstcircle;
						\fill[green] \secondcircle;
						\fill[blue] \thirdcircle;
						\draw node {$491$};
						\draw node at (30:1.2) {$30$};
						\draw node at (150:1.2) {$42$};
						\draw node at (270:1.2) {$22$};
						\draw node at (90:1.8) {$8$};
						\draw node at (210:1.8) {$2$};
						\draw node at (330:1.8) {$417$};
					\end{scope}
				\end{tikzpicture}
			\end{subfigure}
			\begin{subfigure}[8]{0.4\textwidth}
				\vspace*{-.5cm}
				\begin{tabular}{l|c|c}
					Embedding&571&21.4\%\\
					Embedding (idem.)&557&20.9\%\\
					ileanCoP 1.2&875&32.8\%\\
					nanoCoP-i 2.0&858&32.1\%\\\hline
					Total&2669&100\%
				\end{tabular}
			\end{subfigure}
			\vfill
			\caption{Number of theorems proven via our Embedding (red) with extra axioms (green) and a combination of existing state-of-the-art provers (blue) on the ILTP benchmark set.}
		\end{figure}
	\end{frame}

\begin{frame}[standout]
	\Huge\textsc{Thank You}
\end{frame}

\end{document}

### 2023-03-06 Master's Thesis Summary/summary.tex
LPAR

--- 2022-08-01 Notes/Embedding intuitionistic into classical logic/EIICL.tex
@@
-\documentclass[a4paper,12pt]{report}
+\documentclass{article}
+\usepackage[english]{babel}
+\usepackage{setspace}
 
-\title{Embedding intuitionistic into classical logic}
+\title{Master's Thesis Summary}
 \author{Alexander Pluska}
-
-\usepackage{amsmath,amssymb,amsthm}
-\usepackage{ wasysym }
-\usepackage{ stmaryrd }
-\usepackage{ mathpartir }
-\usepackage{bussproofs}
-\usepackage{enumerate}
-\usepackage{tikz-qtree}
-\usetikzlibrary{positioning}
-\usetikzlibrary{arrows.meta}
-\usetikzlibrary{arrows,shapes,quotes}
-\usepackage{stackengine}
-\usepackage{bbold}
-\usepackage{footmisc}
-\usepackage{wrapfig}
-
-\usepackage{hyperref}% http://ctan.org/pkg/hyperref
-\usepackage{cleveref}% http://ctan.org/pkg/cleveref
-
-\theoremstyle{definition}
-\newtheorem{theorem}{Theorem}[section]
-\theoremstyle{definition}
-\newtheorem{corollary}[theorem]{Corollary}
-\theoremstyle{definition}
-\newtheorem{lemma}[theorem]{Lemma}
-\theoremstyle{definition}
-\newtheorem{proposition}[theorem]{Proposition}
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\theoremstyle{definition}
-\newtheorem{example}[theorem]{Example}
-\theoremstyle{definition}
-\newtheorem{remark}[theorem]{Remark}
-
-\DeclareMathOperator*{\argmax}{arg\,max}
-
 \begin{document}
@@
 
-	\begin{abstract}
-		We present an embedding of intuitionistic into classical logic, a quasi-inverse to the famous double-negation translation, i.e.
-a procedure that for each formula $\varphi$ gives a formula $\varphi_c$ such that $\varphi$ is intuitionistically valid if and only if $\varphi_c$ is classically valid.
-We consider both the propositional and predicate case.
-Since many of the more involved results for the more interesting predicate scenario involve similar arguments as in the propositional case we shall develop them simultaneously.
-	\end{abstract}
+	Constructive mathematics refers to a flavour of mathematics in which the existence of an object can only be established by explicit construction, as opposed to classical mathematics where existence can be shown implicitly, e.g.
+by assuming non-existence and deriving a contradiction.
+	The formalism usually associated with constructive mathematics is intuitionistic logic, which essentially differentiates itself from classical logic by the fact that the law of excluded middle $A\vee\neg A$ and the double negation shift $\forall x\neg\neg P(x)\to\neg\neg\forall xP(x)$ are not valid.
+	Besides philosophical considerations, most prominently advocated by Brouwer~\cite{brouwer1907over} and Bishop~\cite{bishop1967foundations}, there is a particular motivation for studying constructive mathematics from the perspective of computer science in that proofs in  intuitionistic logic directly correspond to computer programs --- as expressed in the Curry-Howard correspondence~\cite{howard1980formulae}.
 
-	\tableofcontents
+	The interest in intuitionistic logic has led to the development of a number of automated theorem proving systems for propositional as well as for predicate logic and a collection of benchmark problems (see e.g.
+the
+	ILTP library website~\cite{iltp}).
+	However, progress in automated theorem proving for intuitionistic logic has been slow, whereas reasoners for classical logic have made tremendous progress, see e.g.
+the TPTP~\cite{casc} and SAT~\cite{satc} competitions.
+	This difference can be partially explained by fundamental differences between the logics.
+	Foremost, determining intuitionistic validity is computationally harder, i.e.
+in the propositional case intuitionistic validity is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} whereas classical validity is \verb+coNP+-complete~\cite{cook1971complexity}.
+	A further advantage of classical logic is the existence of calculi that are particularly suitable for automation such as superposition~\cite{bachmair2001resolution}, which rely on the existence of convenient normal forms such as CNF, and the duality between validity and satisfiability, i.e.
+to show the validity of a formula it suffices to show the unsatisfiability of the negated formula.
+	While some (albeit more complex) normal forms also exist  for intuitionistic logic, crucially the duality between validity and satisfiability of the negation does not hold.
+	Therefore, most dedicated intuitionistic theorem provers~\cite{mclaughlin2009efficient, tammet1996resolution} use the naive inverse method, i.e.
+direct search for a cut-free proof by applying the rules from some proof calculus inversely, enhanced by search strategies such as focussing and polarization.
+This approach generally leads to a much more complex search and is therefore difficult to implement efficiently.
+	Finally, we add that in contrast to intuitionistic provers a tremendous amount of work has been put into optimizing provers for classical logic, in particular for the propositional case, i.e.
+SAT-solvers.
 
-	\chapter{Preliminaries}
+	With this work we want to propose a novel approach for intuitionistic theorem proving that leverages the progress in classical theorem proving, that is for each formula $\varphi$:
+	\begin{itemize}
+		\item Give a formula $\varphi^\#$ that is classically valid if and only if $\varphi$ is intuitionistically valid.
+		\item Use a state-of-the-art classical prover to establish the validity/invalidity of $\varphi^\#$.
+		\item Transform the generated proof/counter-model of $\varphi^\#$ to one of $\varphi$.
+	\end{itemize}
+	The most challenging part of this approach is giving the translation of $\varphi$ to $\varphi^\#$.
+	Interestingly, the reverse direction, the famous double-negation translation, has long been established and goes back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
+In the propositional case, it is particularly simple: $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
+Intuitively, the translation collapses for each subformula $\psi$ of $\varphi$ the truth values of $\psi$ and $\neg\neg\psi$, which are classically but not intuitionistically equivalent.
+This gives us a first idea why the reverse direction is perhaps more difficult: We need to expand the truth values of $\psi$ and $\neg\neg\psi$, i.e.
+if they both occur in $\varphi$, we must have a way to (classically) assign different truth values to their respective counterparts in $\varphi^\#$.
+In particular, this necessitates the introduction of new propositional variables in our procedure, which already marks a big difference to the double negation translation.
 
-	First let us recapitulate the most important definitions and some facts about them. This will serve to fix our notation as well.
+	While establishing the translation of formulas we also perform an effective translation of counter-models, i.e.
+for each intuitionistic counter-model of $\varphi$ we effectively construct a classical counter-model of $\varphi^\#$ and vice versa.
+	We note that the existence of counter-models is a key notion that forms a proper dual to validity --- whereas the satisfiability of the negation is not necessary for invalidity (in contrast to classical logic, where it is a proper dual to validity).
+	Transforming and reducing counter-models to a normal form is also what ultimately enables our translation.
+Apart from a translation of counter-models in the final chapter, we also explicitly describe how to equivalently transform classical proofs of $\varphi^\#$ to intuitionistic proofs of $\varphi$.
 
-	\section{Syntax}
+	As a final contribution, we implement our translation in the Rust programming language.
+Our implementation transforms a first-order problem in the tptp format~\cite{tptp} and outputs the translated problem in the same format.
+The code is published on GitHub~\cite{implementation}.
+We then benchmark our implementation using the ILTP problem set~\cite{iltp}, i.e.
+we translate all problems in the set and then run the Vampire theorem prover~\cite{Kov_cs_2013} on the translated problems.
+Our approach performs on par with existing approaches for intuitionistic theorem proving, but comes short of the state-of-the art.
+As there is still a lot of room for optimization, this is a hopeful first sign.
 
-	\subsection{Propositional logic}
-
-	Let us fix some countably infinite set of propositional variables $A, B, C\dots$.
-
-	\begin{definition}
-		A \textit{formula} and \textit{subformula} is defined inductively via the following rules:
-		\begin{itemize}
-			\item every propositional variable as well as a special symbol $\bot$ is a formula, such a formula is called \textit{atomic} and $A,B,C,\dots$ and $\bot$ are called \textit{atoms}.
-			\item if $\varphi$ and $\psi$ are formulas then $\varphi\circ\psi$ is a formula for $\circ\in\{\wedge,\vee,\to\}$.
-			\item every formula is a subformula of itself.
-			\item for a formula $\chi = \varphi\circ\psi$ every subformula of $\varphi$ and $\psi$ is also a subformula of $\chi$.
-		\end{itemize}
-	We write $\neg \varphi$ for $\varphi\to \bot$.
-	\end{definition}
-
-	\subsection{Predicate logic}
-
-	The analog for predicate logic is much more involved as we need to define the term language.
-
-	\begin{definition}
-		A \textit{signature} consists of
-		\begin{itemize}
-			\item a finite set $S_f$ of function symbols $f_1, \dots, f_n$
-			\item a finite set $S_R$ of relation symbols $R_1,\dots, R_m$
-			\item  a function ar$: S_f\cup S_R\to \mathbb{N}$ assigning to each symbol its arity (possibly 0).
-		\end{itemize}
-	\end{definition}
-
-	\noindent Usually the signature we be left implicit. If explicitly stated we denote it by $$\{f_1/\text{ar}(f_1),\dots,f_n/\text{ar}(f_n), R_1/\text{ar}(R_1).\dots,R_m/\text{ar}(R_m)\}.$$
-	Fix a countably infinite collection of free variables $a, b, c\dots$ and bound variables $x, y, z\dots$.
-	Note that the following definitions could be simplified by not distinguishing between bound and free variables.
-However this has some other drawbacks and in particular when defining a sequent calculus for first-order logic having disjoint sets of bound and free variables is convenient.
-
-	\begin{definition}
-		A \textit{term} is defined inductively via the following rules
-		\begin{itemize}
-			\item each free variable is a term
-			\item if $t_1,\dots,t_n$ are terms and $f$ an $n$-ary function symbol then $f(t_1, \dots, t_n)$ is a term.
-		\end{itemize}
-		A \textit{semiterm} is defined in the same way with the difference that bound variables are allowed in 1.
-Clearly every term is a semiterm.
-Analogously to before we can define \textit{subterms} and \textit{subsemiterms}.
-	\end{definition}
-
-	\begin{definition}
-		A \textit{Substitution} is a function $\sigma$ from free variables to semiterms. For a semiterm $t$ we define $t\sigma$ as follows:
-		\begin{itemize}
-			\item $x\sigma = x$ if $x$ is bound.
-			\item $a\sigma = \sigma(a)$ if $a$ is free.
-			\item $f(t_1,\dots,t_n)\sigma = f(t_1\sigma,\dots,t_n\sigma)$.
-		\end{itemize}
-	\end{definition}
-
-	\begin{definition}
-		A \textit{formula} is defined inductively via the following rules
-		\begin{itemize}
-			\item $\bot$ is a formula.
-			\item if $t_1,\dots,t_n$ are terms and $R$ and an $n$-ary relation symbol then $R(t_1, \dots, t_n)$ is a formula.
-			\item if $s, t$ are terms then $s = t$ is a formula.
-		\end{itemize}	
-		Formulas formed by the above rules are called \textit{atomic}.
-		\begin{itemize}
-			\item If $\varphi$ and $\psi$ are formulas then $\varphi\circ\psi$ is a formula for $\circ\in\{\wedge,\vee,\to\}$.
-			\item If $\varphi$ is a formula and $a$ is a free variable occurring in $\varphi$ and $x$ is a bound variable not occurring in $\varphi$ then $Qx(\varphi[x/a])$ is a formula for $Q\in\{\exists,\forall\}$ where the \textit{substitution} $\varphi\sigma$ is defined inductively via
-			\begin{itemize}
-				\item $\bot\sigma = \bot$
-				\item $R(t_1,\dots,t_n)\sigma = R(t_1\sigma,\dots, t_n\sigma)$
-				\item $(s = t)\sigma = (s\sigma = t\sigma)$
-				\item $(\varphi\circ\psi)\sigma = \varphi\sigma\circ\psi\sigma$ for $\circ\in\{\wedge,\vee,\to\}$.
-				\item $(Qy\varphi)\sigma = Qy(\varphi\sigma)$ for $Q\in\{\exists,\forall\}$.
-			\end{itemize}
-			\item A formula with no free variables is a \textit{sentence}.
-		\end{itemize}
-	\end{definition}
-
-	\section{Semantics}
-
-	We now give meaning to formulas in classical and intuitionistic logic.
-
-	\subsection{Propositional Logic}
-	\subsubsection{Classical Semantics}
-
-	\begin{definition}
-		A \textit{valuation} $v$ is a function that assigns to each propositional variable $A$ a \textit{true value} $v(A)\in\{0, 1\}$. We extend $v$ to a function $v^*$ on formulas:
-		\begin{itemize}
-			\item $v^*(\bot) = 0$
-			\item $v^*(A) = v(A)$ for each propositional variable $A$.
-			\item $v^*(\varphi\wedge\psi) = \min(v^*(\varphi), v^*(\psi))$
-			\item $v^*(\varphi\vee\psi) = \max(v^*(\varphi), v^*(\psi))$
-			\item $v^*(\varphi\to \psi) = \max(1 - v^*(\varphi), v^*(\psi))$
-		\end{itemize}
-		$v$ is a \textit{model} for $\varphi$ if $v(\varphi) = 1$, we write $v\models\varphi$.
-$\varphi$ is \textit{satisfiable} if there exists a valuation which is a model.
-It is \textit{valid} if every valuation is a model and we write $\models \varphi$.
-We denote the set of valid formulas with CPC.
-We can furthermore define an entailment relation between sets of clauses, that is $\mathcal S\models_C \mathcal T$ if for every valuation $v$ with $v\models\bigwedge S$ we have $v\models\bigvee \mathcal T$.
-	\end{definition}
-
-	One notable property of classical logic is that a formula $\varphi$ is valid if and only if its negation, i.e.
-$\neg\varphi$ is not satisfiable.
-The same does not hold true for intuitionistic logic as we shall see.
-
-	\subsubsection{Intuitionistic Semantics}
-
-	\begin{definition}
-		A \textit{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of nodes an a family of valuations $(v_w)_{w\in W}$ such that for $u\leq w$ we have $v_u(A)\leq v_u(A)$ for every propositional variable $A$, this is called the \textit{persistency condition}.
-We extend the valuations $v_u$ slightly differently to $v_u^\#$ than before, i.e.
-have
-		\begin{itemize}
-			\item $v_u^\#(\bot) = 0$
-			\item $v_u^\#(\varphi) = v_u(A)$ if $\varphi$ is atomic.
-			\item $v_u^\#(\varphi\wedge\psi) = \min(v_u^\#(\varphi), v_u^\#(\psi))$
-			\item $v_u^\#(\varphi\vee\psi) = \max(v_u^\#(\varphi), v_u^\#(\psi))$
-			\item $v_u^\#(\varphi\to \psi) = \min\{\max(1 - v_w^\#(\varphi), v_w^\#(\psi))\:|\:w\geq u\}$
-		\end{itemize}
-		That is $v_u^\#(\varphi\to \psi) = 1$ if at every world $w\geq u$ we have $v_w^\#(\varphi) = 0$ or $v_w^\#(\psi) = 1$.
-We say that $\varphi$ is satisfied at a node $u$ if $v_u^\#(\varphi) = 1$ and write $u\models\varphi$.
-If a formula is satisfied at every node $K$ is a \textit{model} for $\varphi$ and we write $K\models \varphi$.
-$\varphi$ is satisfiable if there exists a Kripke structure which is a model.
-It is \textit{valid} if every Kripke structure is a model and we write $\models \varphi$.
-We denote the set of valid formulas with IPC.
-As before we define an entailment relation $\models_I$ between sets of formulas.
-	\end{definition}
-	There are many classical tautologies which are not intuitionistically valid
-	\begin{example}\label{LEMcounterexample}
-		Consider the famous law of excluded middle $A\vee\neg A$.
-As a counter model consider a Kripke structure with two nodes $u\leq w$ and $v_u(A) = 0, v_w(A) = 1$.
-Then $u\not\models A$ and $u\not\models \neg A$, i.e.
-$u\not\models A\vee\neg A$.
-	\end{example}
-	The following is a famous and important result due to Glivenko~\cite{glivenko1929}
-
-	\begin{theorem}[Glivenko's theorem]
-		A propositional formula $\varphi$ is classically valid if and only if $\neg\neg\varphi$ is intuitionistically valid.
-	\end{theorem}
-	\noindent Note however that this difference between classical and intuitionistic logic is not reflected in satisfiability, only in validity:
-	\begin{lemma}\label{propsat}
-		A propositional formula $\varphi$ is intuitionistically satisfiable if and only if it is classically satisfiable.
-	\end{lemma}
-
-	\begin{proof}
-		Every classical model can be viewed as a Kripke model with a single node.
-Therefore classical satisfiability trivially implies intuitionistic satisfiability.	On the other hand in every Kripke structure that is a model every node is a classical model so intuitionistic satisfiability implies intuitionistic satisfiability.
-	\end{proof}
-
-	In classical logic a proven strategy for showing that a formula valid is to consider its negation and show that it is not satisfiable.
-Due to the previous lemma this is insufficient for intuitionistic logic.
-There are formulas like $A\vee\neg A$ which are not a intuitionistic tautology while their negation is not satisfiable either.
-This serves to motivate the consideration of equivalidity in the following chapters.
-In the context of classical logic it can be fully explained via equisatisfiability and is therefore neglected but for intuitionistic logic it is more fundamental.
-
-	Finally let us note the complexities of these various decision problems, which play a fundamental role in all of theoretical computer science.
-
-	\begin{theorem}
-		Satisfiability for classical and intuitionistic propositional logic is NP-complete.
-	\end{theorem}
-
-	For classical logic this is the famous Cook-levin theorem first published in Cook's seminal paper~\cite{cook1971complexity}.
-Of course we have seen that the intuitionistic case is the same decision problem.
-
-	\begin{corollary}
-		Validity for classical logic is co-NP-complete
-	\end{corollary}
-
-	This is clear since validity is just the complementary problem to satisfiability of the negated formula.
-However in intuitionistic logic unfortunately validity is really harder than satisfiability as shown by Statman~\cite{statman1979intuitionistic}.
-
-	\begin{theorem}
-		Validity for intuitionistic logic in PSPACE-complete.
-	\end{theorem}
-
-	One thing to note is that in contrast to predicate logic all these problems are decidable.
-
-	\subsection{Predicate Logic}
-
-	We now give account of the more involved semantics of first-order logic
-
-	\subsubsection{Classical Semantics}
-
-	\begin{definition}
-		Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the domain of $\mathcal{M}$, and a function $I$ that assigns
-		\begin{itemize}
-			\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
-			\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
-		\end{itemize}
-		A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
-			m&\text{if $b=a$}\\
-			v(b)&\text{otherwise}
-		\end{cases}$$
-		Then terms are interpreted as follows:
-		\begin{itemize}
-			\item $a^{I, v} = v(a)$ for each free variable $a$.
-			\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$.
-		\end{itemize}
-		We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
-		\begin{itemize}
-			\item $\mathcal M, v\not\models\bot$.
-			\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$.
-			\item $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
-			\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$.
-			\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
-			\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
-			\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
-			\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
-		\end{itemize}
-		$\varphi$ is satisfiable if for some $\mathcal{M}, v$ we have  $\mathcal M, v\models\varphi$.
-It is valid if for all  $\mathcal M, v$ we have  $\mathcal M, v\models\varphi$.
-We write $\mathcal{M}\models\varphi$ if for every $v$ we have  $\mathcal M, v\models\varphi$.
-We denote the set of valid formulas with CQC.
-As before define an entailment relation $\models_C$ between sets of formulas.
-	\end{definition}
-
-	\subsubsection{Intuitionistic Semantics}
-
-	\begin{definition}
-		A $\Sigma$-\textit{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$ and a family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$
-		\begin{itemize}
-			\item $M_u\subseteq M_w$.
-			\item $f^{I_w}|_{M_u} = f^{I_u}$.
-			\item $R^{I_w}|_{M_u} = R^{I_u}$.
-		\end{itemize}
-		Then we define a model relation between worlds $u\in W$ and variable assignments $v$ (targeting $M_u$) and formulas $\varphi$ as follows:
-		\begin{itemize}
-			\item $u, v\not\models\bot$
-			\item $u, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I_u, v},\dots,t_n^{I_u, v})\in R^{I_u}$.
-			\item $u, v\models s = t$ iff $s^{I_u, v} = t^{I_u, v}$.
-			\item $u, v\models \varphi\wedge \psi$ iff $u, v\models\varphi$ and $u, v\models\psi$.
-			\item $u, v\models \varphi\vee\psi$ iff $u, v\models\varphi$ or $u, v\models\psi$.
-			\item $u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $w, v\not\models\varphi$ or $w, v\models\psi$.
-			\item $u, v\models\exists x\varphi$ iff there exists $m\in M_u$ such that $u, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
-			\item $u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $w, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
-		\end{itemize}
-		We say that $\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$.
-As always a formula is satisfiable if it is satisfies by some Kripke structure and it is valid if it is satisfied in every Kripke structure.
-We denote the set of valid formulas with IQC.
-As before define an entailment relation $\models_I$ between sets of formulas.
-	\end{definition}
-	In addition to the propositional tautologies there are now sentences involving quantifiers which are classically valid but not intuitionistically.
-	Consider for instance the classical tautology $\varphi$ $$\neg\forall x A(x)\to \exists x \neg A(x)$$We exhibit an intuitionistic counter model.
-Consider the Kripke structure with two nodes $u\leq w$ with $M_u = \{0\}, M_w = \{0, 1\}$ and $A^{I_u} = A^{I_w} = \{0\}$.
-Then $u, v\not\models \varphi$: First note that $w, v\not\models\forall xA(x)$ and $u, v\not\models\forall xA(x)$ for any valuation $v$ since $w, v[1, a]\not\models A(a)$, i.e.
-$u, v\models\neg\forall xA(x)$ and also $w, v\models\neg\forall xA(x)$.
-Furthermore $u, v[0/a]\not\models \neg A(a)$ and $w, v[0/a]\not\models \neg A(a)$ so $u, v\not\models \exists x\neg A(x)$ for any valuation $v$ (note however that this is not true at $w$!).
-Therefore $u, v\not\models\varphi$.
-
-	Intuitively the problem is that the existential quantifier at $u$ does not see the elements at $w$, i.e.
-it must "choose" between the currently available elements.
-As in the propositional case there are well known negative translation of classical logic into intuitionistic logic going back to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit}.
-
-	\begin{definition}
-		For a predicate formula $\varphi$ define inductively $\varphi^N$ as follows:
-		\begin{itemize}
-			\item $\varphi^N = \varphi$ if $\varphi$ is atomic
-			\item $(\varphi\wedge\psi)^N = \varphi^N\wedge\psi^N$
-			\item $(\varphi\vee\psi)^N = \neg(\neg\varphi^N\wedge\neg\varphi^N)$
-			\item $(\varphi\to\psi)^N = \varphi^N\to\psi^N$
-			\item $(\forall x\varphi)^N = \forall x\varphi^N$
-			\item $(\exists x\varphi)^N = \neg\forall x\neg\varphi^N$
-		\end{itemize}
-	\end{definition}
-	\begin{theorem}
-		A $\Sigma$-formula $\varphi$ is classically valid if and only if $\varphi^N$ is intuitionistically valid.
-	\end{theorem}
-
-	But again this is not reflected in satisfiability, i.e. with an completely analogous argument to~\ref{propsat} we obtain
-
-	\begin{lemma}
-		Any $\Sigma$-formula $\varphi$ is classically satisfiable if and only if it is intuitionistically satisfiable.
-	\end{lemma}
-
-	\section{Skolemization and Herbrandization}
-
-	An important step in the embedding will be the elimination of quantifiers via Herbrandization.
-	In this process we introduce fresh variables and add additional function symbols to the signature.
-	A fresh variable is any variable that does not occur in any of the considered formulas.
-	Whenever we add a function symbol we implicitly extend the signature by some not previously contained symbol.
-
-	\begin{definition}
-		For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to $Z$ by simultaneous induction as follows:
-		\begin{itemize}
-			\item $A^S_Z = A^H_Z = A$ for each atomic $A$.
-			\item $(\varphi\circ\psi)^X_Z = \varphi^X_Z\circ\psi^X_Z$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
-			\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$\\$(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
-			\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
-			\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.
-		\end{itemize}
-		Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
-	\end{definition}
-
-	\begin{theorem}
-		\label{thm:herbrand-skolem}
-		For every formula $\varphi$
-		\begin{itemize}
-			\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
-			\item $\varphi$ and $\varphi^H$ are classically equivalid.
-		\end{itemize}
-	\end{theorem}
-	which follows from the following two Lemmata:
-	\begin{lemma}
-		Let $\chi$ be a formula and let $\{z_1\dots z_n\} = Z$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
-		\begin{itemize}
-			\item if $M, v \models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
-			\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
-		\end{itemize}
-	\end{lemma}
-
-	\begin{proof}
-		We proceed by simultaneous induction on the formula height.
-
-		For atomic $\chi$ the claims are clear.
-
-		Otherwise there are $3$ cases:
-
-		1.
-$\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$.
-By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v})$ and $\mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as stated in the lemma.
-		Then we can define $I_{\chi, v}$ as follows:
-
-		For every symbol $p$ that occurs in $\varphi^S_Z$ or $\varphi^H_Z$ but not $\psi^S_Z, \psi^H_Z$, we set $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$.
-For every symbol $p$ that occurs in $\psi^S_Z$ or $\psi^H_Z$ but not $\varphi^S_Z, \varphi^H_Z$, we set $p^{I_{\chi, v}} = p^{I_{\psi, v}}$.
-Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.
-
-		Let $\circ=\ \to$.
-Suppose $\mathcal M, v\models\chi$.
-Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_Z$ or $\mathcal M_\chi\models\psi^H_Z$, i.e.
-$\mathcal M_\chi\models\chi^S_Z$.
-On the other hand suppose $\mathcal M, v\not\models\chi$.
-Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_Z$ and $\mathcal M_\chi, v\not\models\psi^S_Z$, i.e.
-$\mathcal M_\chi\not\models\chi^H_Z$.
-Analogous arguments work for $\circ\in\{\wedge, \vee\}$.
-
-		2.
-$\chi = \forall x\varphi$.
-Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/z_1\dots x_n/z_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise.
-Since $Z$ contains all free variables occurring in $\chi$ this is a well-defined function.
-
-		By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
-Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{Z\cup a}$ or $\varphi[a/x]^H_{Z\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
-
-		Suppose $\mathcal M, v\models\chi$.
-Then, for all $m\in M$ we have that $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{Z\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{Z\cup\{a\}}[x/a])$, i.e.
-$\mathcal M_\chi,v\models \chi^S_Z$.
-On the other hand suppose $\mathcal M, v\not\models\chi$.
-Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{Z\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(z_1,\dots z_n)/x]^H_Z$, i.e.
-$\mathcal M_\chi, v\not\models(\forall x\varphi)^H_Z$.
-
-		3.
-$\chi = \exists x\varphi$.
-The argument runs dually to 2.
-Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/z_1\dots x_n/z_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise.
-Since $Z$ contains all free variables occurring in $\chi$ this is a well-defined function.
-
-		By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma.
-Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{Z\cup a}$ or $\varphi[a/x]^H_{Z\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.
-
-		Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_Z$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_Z$.
-	\end{proof}
-
-	\begin{lemma}
-		For every structure $\mathcal M$ and $\{z_1\dots z_n\} = Z$ that contains all free variables in $\chi$ and variable assignment $v$
-		\begin{itemize}
-			\item if $\mathcal M, v\models\varphi^S_Z$ then $\mathcal M, v\models \varphi$
-			\item if $\mathcal M, v\not\models\varphi^H_Z$ then $\mathcal M, v\not\models\varphi$.
-		\end{itemize}
-	\end{lemma}
-
-	\begin{proof}
-		Again we proceed by simultaneous induction on the formula height.
-
-		For atoms the claims are clear. We distinguish 5 cases.
-
-		1.
-$\chi = \varphi\to\psi$.
-Suppose $\mathcal M, v\models\chi^S_Z$, i.e.
-$\mathcal M, v\not\models\varphi^H_Z$ or $\mathcal M, v\models\psi^S_Z$.
-By induction hypothesis $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e.
-$\mathcal M, v\models \chi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
-$\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_Z$.
-Again by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.
-
-		2. + 3. Conjunctions and Disjunctions are dealt with analogously.
-
-		4.
-$\chi = \forall x\varphi$.
-Suppose $\mathcal M, v\models\chi^S_Z$, i.e.
-for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{Z\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$.
-Then it follows that $\mathcal M, v\models\varphi$.
-On the other hand suppose $\mathcal M, v\not\models\chi^H_Z$, i.e.
-there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{Z\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e.
-$\mathcal M, v\not\models\forall x\chi$.
-
-		5. $\chi = \exists x\varphi$.  The argument runs dually to 4.
-	\end{proof}
-
-	\section{Proof theory}
-
-	Logical truth can also be approached via syntax rather than semantics. This gives rise to proof theory which will be especially important in the third chapter.
-
-	\begin{definition}
-		A \textit{sequent} $\Gamma\vdash \Delta$ consists of two multisets of formulas $\Gamma, \Delta$.
-The following rules make up our classical calculus $\mathbf{GC}$, corresponding roughly to \textbf{G3c} from~\cite{basicprooftheory}:\\
-		\begin{center}
-			\begin{tabular}{lll}
-				\AxiomC{\hphantom{x}}
-				\RightLabel{Ax ($P$ atomic)}
-				\UnaryInfC{$P,\Gamma\vdash \Delta, P$}
-				\DisplayProof&
-				\AxiomC{\hphantom{x}}
-				\RightLabel{L$\bot$}
-				\UnaryInfC{$\bot,\Gamma\vdash\Delta$}
-				\DisplayProof&
-				\\&&\\
-				\AxiomC{$A, B,\Gamma\vdash\Delta$}
-				\RightLabel{L$\wedge$}
-				\UnaryInfC{$A\wedge B, \Gamma\vdash \Delta$}
-				\DisplayProof&
-				\AxiomC{$\Gamma\vdash\Delta, A$}
-				\AxiomC{$\Gamma\vdash\Delta, B$}
-				\RightLabel{R$\wedge$}
-				\BinaryInfC{$\Gamma\vdash \Delta, A\wedge B$}
-				\DisplayProof&
-				\\&&\\
-				\AxiomC{$A, \Gamma\vdash\Delta$}
-				\AxiomC{$B, \Gamma\vdash\Delta$}
-				\RightLabel{L$\vee$}
-				\BinaryInfC{$A\vee B, \Gamma\vdash \Delta$}
-				\DisplayProof&
-				\AxiomC{$\Gamma\vdash\Delta, A, B$}
-				\RightLabel{R$\vee$}
-				\UnaryInfC{$\Gamma\vdash \Delta, A\vee B$}
-				\DisplayProof&
-				\\&&\\
-				\AxiomC{$\Gamma\vdash\Delta, A$}
-				\AxiomC{$B, \Gamma\vdash\Delta$}
-				\RightLabel{L$\to$}
-				\BinaryInfC{$A\to B, \Gamma\vdash \Delta$}
-				\DisplayProof&
-				\AxiomC{$A,\Gamma\vdash\Delta, B$}
-				\RightLabel{R$\to$}
-				\UnaryInfC{$\Gamma\vdash \Delta, A\to B$}
-				\DisplayProof&
-				\\&&\\
-				\AxiomC{$A[t/x], \Gamma\vdash\Delta$}
-				\RightLabel{L$\forall$}
-				\UnaryInfC{$\forall xA, \Gamma\vdash \Delta$}
-				\DisplayProof&
-				\AxiomC{$\Gamma\vdash\Delta, A[a/x]$}
-				\RightLabel{R$\forall$}
-				\UnaryInfC{$\Gamma\vdash \Delta, \forall xA$}
-				\DisplayProof&
-				\\&&\\
-				\AxiomC{$A[a/x], \Gamma\vdash\Delta$}
-				\RightLabel{L$\exists$}
-				\UnaryInfC{$\exists xA, \Gamma\vdash \Delta$}
-				\DisplayProof&
-				\AxiomC{$\Gamma\vdash\Delta, A[t/x]$}
-				\RightLabel{R$\exists$}
-				\UnaryInfC{$\Gamma\vdash \Delta, \exists xA$}
-				\DisplayProof&
-				\\&&\\
-				\AxiomC{$t = t, \Gamma\vdash \Delta$}
-				\RightLabel{Ref}
-				\UnaryInfC{$\Gamma\vdash \Delta$}
-				\DisplayProof&
-				\AxiomC{$s = t, P[s/x], P[t/x], \Gamma\vdash \Delta$}
-				\RightLabel{Ref}
-				\UnaryInfC{$s = t, P[s/x],\Gamma\vdash \Delta$}
-				\DisplayProof&
-				\\&&\\
-			\end{tabular}
-		\end{center}
-
-		where in R$\forall$ and L$\exists$ $a$ is a free variable not occurring in the sequent otherwise.
-		The above rules make up the calculus called \textbf{Gc}. It corresponds to \textbf{G3c}$^=$ from~\cite{basicprooftheory}.
-
-		For intuitionistic logic we remain with a multi-succedent system by adjusting R$\to$ and R$\forall$:\\
-
-		\begin{center}
-			\AxiomC{$A, \Gamma\vdash B$}
-			\RightLabel{R$\to$}
-			\UnaryInfC{$\Gamma\vdash\Delta, A\to B$}
-			\DisplayProof\hspace{2cm}
-			\AxiomC{$\Gamma\vdash A[a/x]$}
-			\RightLabel{R$\forall$}
-			\UnaryInfC{$\Gamma\vdash\Delta, \forall xA$}
-			\DisplayProof
-		\end{center}
-
-		We will call the calculus containing this formulas $\mathbf{GI}$ roughly corresponding to $\mathbf{G3i}$ from~\cite{basicprooftheory}.
-Usually for intuitionistic logic a single-formula succedent system is considered, but we are interested in reducing the difference between the calculi as much as possible.
-
-		A \textit{proof} of a formula $\varphi$ in a calculus $\mathbf{G}$ is a labelled binary rooted tree in which every node is labelled with a sequent such that
-		\begin{itemize}
-			\item Every leaf is labelled in accordance to Ax or L$\bot$.
-			\item The sequent of an interior node can be derived from sequents of its child nodes according to one of the rules.
-		\end{itemize}
-		We usually present a proof directly as a sequence of applications of rules rather than in a more usual tree presentation.
-If a sequent $\mathcal S\vdash\mathcal T$ is provable in $\mathbf{GC}$ we write $\mathcal S\vdash_C\mathcal T$, if it is in $\mathbf{GI}$ we write $\mathcal S\vdash_I\mathcal T$.
-	\end{definition}
-
-	\begin{example}
-		The following is a proof of $(A\to B)\to A\vdash B$:
-		\begin{center}
-			\AxiomC{}
-			\RightLabel{Ax}
-			\UnaryInfC{$A\to B, A\vdash B, A$}
-			\AxiomC{}
-			\RightLabel{Ax}
-			\UnaryInfC{$A, B\vdash B$}
-			\RightLabel{L$\to$}
-			\BinaryInfC{$A\to B, A\vdash B$}
-			\DisplayProof
-		\end{center}
-	\end{example}
-
-	\noindent We now recall some important properties of the given systems:
-
-	\begin{theorem}[subformula property]
-		Any formula that appears in a derivation is a subformula of a forumla that occurs in the root sequent.
-	\end{theorem}
-
-	We consider the following additional rules, which are not a priori part of our calculus:
-	\begin{center}
-		\begin{tabular}{lll}
-			\AxiomC{$\Gamma\vdash\Delta$}
-			\RightLabel{Lweak}
-			\UnaryInfC{$A,\Gamma\vdash \Delta$}
-			\DisplayProof&
-			\AxiomC{$\Gamma\vdash\Delta$}
-			\RightLabel{Rweak}
-			\UnaryInfC{$\Gamma\vdash \Delta, A$}
-			\DisplayProof&
-			\\&&\\
-			\AxiomC{$A, A,\Gamma\vdash\Delta$}
-			\RightLabel{Lcontr}
-			\UnaryInfC{$A, \Gamma\vdash \Delta$}
-			\DisplayProof&
-			\AxiomC{$\Gamma\vdash\Delta, A, A$}
-			\RightLabel{Rcontr}
-			\UnaryInfC{$\Gamma\vdash \Delta, A$}
-			\DisplayProof&
-			\\&&\\
-		\end{tabular}
-
-		\AxiomC{$\Gamma\vdash A, \Delta$}
-		\AxiomC{$\Gamma', A\vdash \Delta'$}
-		\RightLabel{Cut}
-		\BinaryInfC{$\Gamma, \Gamma'\vdash\Delta, \Delta'$}
-		\DisplayProof
-	\end{center}
-
-	The following central result of proof theory will be useful to us in many places:
-	\begin{theorem}[admissibility of cut (and structural rules)]
-		The 5 above rules are admissible in our calculi for classical and intuitionistic logic, i.e. adding them we are not able to prove any additional theorems.
-	\end{theorem}
-
-	This means that we may assume that all proofs are free of these rules, but in constructing new proofs we may apply them.
-We exclude them from our calculus to keep it simple for proofs and because cut destroys the subformula property.
-
-	The semantic and syntactic approach to logic are of course equivalent.
-
-	\begin{theorem}\label{completeness}[Completeness and Soundness]
-		For sets of formulas $\mathcal S,\mathcal T$ both classically and intuitionistically
-		$$\mathcal S\models\mathcal T \text{ iff }\mathcal S\vdash\mathcal T$$
-	\end{theorem}
-
-	\chapter{Embedding intuitionistic into classical logic}
-
-	The goal of this section is to give a complete embedding of intuitionistic into classical logic.
-We shall start with the simpler case of propositional formulas.
-The process will be divided into three steps.
-The first is a first-order encoding of intuitionistic semantics, which gives an obvious but unsatisfactory embedding from propositional intuitionistic logic.
-We then discuss how some of the quantifiers can be eliminated via Herbrandization.
-As a third part we give a certain normal form that then finally allows us to realize that all quantifiers can be indeed eliminated from the translated formula and we end up with a propositional formula.
-As a quick note we also give an embedding form IPC to QBF.
-
-	For predicate formulas the process is a bit more involved in that we also need to respect the constructive content of quantifiers.
-For this we use orderization as proposed in~\cite{iemhoff2010eskolemization}.
-This allows us to utilize a quite similar process as in the first section, even though it is a bit more involved.
-
-	\section{Propositional logic}
-
-	The idea is to encode the Kripke semantics of intuitionistic logic classically and is quite close to what was done in~\cite{iemhoff2010eskolemization}.
-We first present an embedding from IPC to CQC which is conceptually easier to understand, but will then reduce it to CPC.
-To that end we will first propose a Tseytin-like translation, that reduces formulas to a normal form while preserving validity:
-
-	\begin{theorem}\label{thm:Tseytin}
-		For every propositional formula $\varphi$ there effectively are an atom $P$ and a set of $\mathcal S$ of formulas (over an extended set of propositions) of one of the forms\\
-		$\text{ \hspace{1cm}} A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$\\
-		for atomic $A, B, C$, such that $\varphi$ is intuitionistically valid if and only if $\bigwedge\mathcal S\to P$ is intuitionistically valid.
-The time complexity of the transformation is linear in the input size.
-	\end{theorem}
-
-	\begin{theorem}
-		Let $\mathcal S$ be a set of formulas of the form
-		$$(A\wedge B)\to C, (A\vee B)\to C, A\to (B\wedge C), A\to(B\vee C), (A\to B)\to C$$
-		and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
-For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
-Obtain $\mathcal S^\#$ by including the following formulas:
-		\begin{itemize}
-			\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
-			\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-			\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-			\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
-		\end{itemize}
-		Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
-The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$.
-Futhermore, there is an effective procedure for translating counter-models between the sentences.
-	\end{theorem}
-
-	\subsection{IPC to CQC}
-
-	The most obvious approach to embedding intuitionistic into classical logic is to express intuitionistic semantics, i.e.
-Kripke frames, as a first order theory.
-For every propositional variable $A$ consider a unary predicate $A$ of the same name where $A(u)$ expresses that $A$ is true at some world $u$.
-We can then naively encode formulas as follows:
-
-	\begin{definition}
-		Let $\varphi$ be a propositional formula. Define $\varphi^{u}$ inductively as follows:
-		\begin{itemize}
-			\item $A^{u} = A(u)$ for every propositional variable $A$.
-			\item $\bot^u = \bot$.
-			\item $(\varphi\circ\psi)^u = \varphi^u\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
-			\item $(\varphi\to \psi)^u = \forall w(u\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new variable.
-		\end{itemize}
-		Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
-		$$K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))$$
-		with e.g.
-		\begin{align*}
-			\text{\normalfont PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
-			\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ is a propositional variable that occurs in $\varphi$}\}
-		\end{align*}
-		Then define
-		$$\varphi^{C} = K(\varphi)\to \varphi^{u}$$
-		where $u$ is some free variable.
-	\end{definition}
-
-	\noindent As all we have done is modelling Kripke frames as a first-order theory we directly obtain:
-
-	\begin{lemma}
-		$\varphi$ is intuitionistically valid iff $\varphi^C$ is classically valid.
-	\end{lemma}
-
-	\begin{example}
-		Consider the law of excluded middle $$\varphi = (A\to \bot)\vee A$$ which is not intuitionistically valid. Then
-		\begin{align*}
-			\varphi^{C} &= K(\varphi)\to ((A\to \bot)\vee A)^u\\
-						 &= K(\varphi)\to (A\to \bot)^u\vee A^u\\
-						 &= K(\varphi)\to (\forall u'( u\preceq u' \to A^{u'}\to \bot^{u'})\vee A(u))\\
-						 &= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot)\vee A(u))
-		\end{align*}
-		and indeed $\varphi^{C}$ admits a counter model with domain $\{u, u'\}$ where $u \preceq u'$, $A(u) = 0, A(u') = 1$.
-	\end{example}
-
-	Note how this counter model corresponds exactly to the intuitionistic counter model presented in example~\ref{LEMcounterexample}. We can make this correspondence explicit.
-
-	\begin{definition}
-		Let $\mathcal K = (W, (v_w)_{v\in W})$ be a propositional Kripke structure. Define the $\mathcal M(\mathcal K) = (M, I)$ as follows:
-		\begin{itemize}
-			\item As a domain take $M = W$.
-			\item Have $u\in A^{I}$ iff $u, v_u\models A$.
-			\item Interpret $\preceq$ as the partial order on $W$.
-		\end{itemize}
-	\end{definition}
-
-	\noindent By definition we then have
-	\begin{lemma}
-		$\mathcal K\models \varphi$ iff $\mathcal M(\mathcal K)\models \varphi^{C}$.
-	\end{lemma}
-
-	\indent Similarly for every $\Sigma$-structure $\mathcal M$ with $\mathcal M\models K(\varphi)$ we could construct a Kripke Frame $\mathcal K(\mathcal M)$ such that $\mathcal M\models \varphi^C$ iff $\mathcal K(\mathcal M)\models \varphi$.
-
-	\subsection{Normal form transformation}
-
-	Of course transforming a propositional formula into a predicate formula is somewhat unsatisfying.
-As a first step towards obtaining a propositional formula we want to apply Herbrandization.
-Now trying to apply it to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated.
-We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form.
-In fact our translation is often presented as the first step in the Tseytin transformation, after which the individual formulas are then converted to CNF.
-	Hence, the propositional case of our translation is well known, but we are not aware of any work that uses just this step as a stand-alone.
-A similar translation for intuitionistic propositional logic was also proposed in~\cite{statman1979intuitionistic}.
-
-	\begin{definition}
-		Let $\varphi$ be some formula.
-For each subformula $\psi$ of $\varphi$, consider some new $n$-ary relation symbol $P_\psi$, where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$.
-We denote those variables with $\vec z_\psi$ in some arbitrary but fixed order.
-We inductively define clause sets $\mathcal S^+(\varphi)$ and $\mathcal S^-(\varphi)$ as follows:
-		\begin{itemize}
-			\item For every atomic formula $A$:\\
-			$\mathcal S^+(A) = \{P_A(\vec z_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec z_A)\}$
-			\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$:\\
-			$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\to P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec z_\varphi)\circ P_{\psi}(\vec z_\psi)\to P_{\varphi\circ\psi}(\vec z_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
-			\item For every implication $\varphi \to\psi$:\\
-			$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\wedge P_{\varphi}(\vec z_\varphi))\to P_{\psi}(\vec z_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec z_\varphi)\to P_{\psi}(\vec z_\psi))\to P_{\varphi\to\psi}(\vec z_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\psi)$
-			\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$:\\
-			$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec z_{Qx\varphi})\to QxP_{\varphi}(\vec z_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec z_{\varphi})\to P_{Qx\varphi}(\vec z_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
-		\end{itemize}
-		%We further set $\mathcal S(\varphi) = \mathcal S^+(\varphi)\cup\mathcal S^-(\varphi)$.
-	\end{definition}
-
-	\begin{lemma}
-		$\bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)\to\varphi$ and $\bigwedge S^-(\varphi)\wedge \varphi\to P_\varphi(\vec z_\varphi)$ are valid both classically and intuitionistically.
-	\end{lemma}
-
-	It is straight-forward to extend structures to the additional symbols such that the formulas $S^+(\varphi)\cup\mathcal S^-(\varphi)$ are satisfied.
-	\begin{definition}
-		For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms of the form $P_\psi$.
-By slight abuse of notation we denote with $\vec z_\psi$ elements of the domain instead of variables.
-In the classical case, we set
-		$$\begin{matrix}
-			P_A^I(\vec z_A):\Leftrightarrow A^I(\vec z_A)\\
-			P_{\varphi\wedge\psi}^I(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\wedge P_{\psi}^I(\vec z_\psi)\indent P_{\varphi\vee\psi}^I(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^I(\vec z_\varphi)\vee P_{\psi}^I((\vec z_\psi))\\P_{\varphi\to\psi}^I(\vec z_{\varphi\to\psi}) :\Leftrightarrow (\neg P_{\varphi}^I(\vec z_{\varphi}))\vee P_{\psi}^I(\vec z_{\psi})\\P_{\forall x\varphi}^I(\vec z_{\forall x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for all $x\in M$}\}\indent P_{\exists x\varphi}^I(\vec z_{\exists x\varphi}) :\Leftrightarrow \{\vec z\:|\:\:P_{\varphi}^I(\vec z_\varphi) \text{ for some $x\in M$}\},
-		\end{matrix}$$
-		and for intuitionistic logic and each world $u$, we set
-		$$\begin{matrix}
-			P_A^{I_u}(\vec z_A):\Leftrightarrow A^{I_u}(\vec z_A)\\
-			P_{\varphi\wedge\psi}^{I_u}(\vec z_{\varphi\wedge\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\wedge P_{\psi}^{I_u}(\vec z_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec z_{\varphi\vee\psi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi)\vee P_{\psi}^{I_u}((\vec z_\psi))\\
-			P_{\varphi\to\psi}^{I_u}(\vec z_{\varphi\to\psi}) :\Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec z_{\varphi}))\vee P_{\psi}^{I_w}(\vec z_{\psi})\text{ for all $w\geq u$}\\
-			P_{\forall x\varphi}^{I_u}(\vec z_{\forall x\varphi}) :\Leftrightarrow P_{\varphi}^{I_w}(\vec z_\varphi) \text{ for all $w\geq u$, $x\in M_w$}\indent P_{\exists x\varphi}^{I_u}(\vec z_{\exists x\varphi}) :\Leftrightarrow P_{\varphi}^{I_u}(\vec z_\varphi) \text{ for some $x\in M_u$}.
-		\end{matrix}$$
-	\end{definition}
-	From this definition we directly obtain the following Lemmata by induction on the formula height.
-	\begin{lemma}\label{thm:struct1}
-		For every formula $\varphi$ and structure $\mathcal M$, we have that $\mathcal S(\mathcal M, \varphi)\models\mathcal S^+(\varphi)\cup S^-(\varphi)$.
-	\end{lemma}
-
-	\begin{lemma}\label{thm:struct2}
-		For every structure $\mathcal M$,
-		$\mathcal M\models \varphi$ if and only if $S(\mathcal M, \varphi)\models P_\varphi(\vec z_\varphi)$.
-	\end{lemma}
-
-	Then from the previous three Lemmata we get:
-
-	\begin{corollary}\label{eequivalid}
-		In both intuitionistic and classical logic
-		\begin{itemize}
-			\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec z_\varphi)$ is, and
-			\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec z_\varphi)$ is.
-		\end{itemize}
-	\end{corollary}
-
-	\begin{example}
-		Consider $\varphi = \neg\forall x A(x)\to \exists x\neg A(x)$. Then
-		\begin{align*}
-			\mathcal S^+(\varphi) = &\{(P_\varphi\wedge P_{\neg\forall x A(x)})\to P_{\exists x\neg A(x)}, (P_{\forall xA(x)}\to\bot)\to P_{\neg \forall xA(x)}\}\\&\cup\{\forall xP_{A(x)}(x)\to P_{\forall xA(x)}, \forall x(P_{A(x)}(x)\to A(x))\}\\&\cup\{P_{\exists x\neg A(x)}\to \exists xP_{\neg A(x)}(x), \forall x(P_{\neg A(x)}(x)\to P_{A(x)}(x)\to \bot)\}\\&\cup\{\forall x(A(x)\to P_{A(a)}(x))\}\\
-			\mathcal S^-(\varphi) = &\{(P_{\neg\forall x A(x)}\to P_{\exists x\neg A(x)})\to P_\varphi,  (P_{\neg \forall xA(x)}\wedge P_{\forall xA(x)})\to\bot\}\\&\cup\{P_{\forall xA(x)}\to \forall xP_{A(x)}(x), \forall x(A(x)\to P_{A(x)}(x))\}\\&\cup\{\exists xP_{\neg A(x)}(x)\to P_{\exists x\neg A(x)}, \forall x((P_{A(x)}(x)\to \bot)\to P_{\neg A(x)}(x))\}\\&\cup\{\forall x(P_{A(a)}(x)\to A(x))\}
-		\end{align*}
-	\end{example}
-
-	\subsection{IPC to simplified CQC}
-
-	We are now ready to combine the previously presented reductions.
-	Let $\varphi$ be some propositional formula.
-As a first step we perform the normalization from the last section, i.e.
-instead of $\varphi$ we consider $$\varphi' = \bigwedge \mathcal S^-(\varphi)\to P_\varphi.$$
-
-	Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
-	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C,$$
-	where we can treat $A\to B$ as a special case $A\to (B\wedge B)$. We then apply the transformation to CQC, leading to a formula
-	$$\varphi'' = K(\varphi')\to\bigwedge  \mathcal S\to P_\varphi(u),$$
-	where $u$ is some new free variable and each $\psi\in\mathcal S$ is of one of the forms
-	$$\begin{matrix}
-		\forall k(u\preceq k\to A(k)\to (B(k)\wedge C(k)))\indent  \forall k(u\preceq k\to (A(k)\wedge (B(k))\to C(k)))\\
-		\forall k(u\preceq k\to A(k)\to (B(k)\vee C(k)))\indent \forall k(u\preceq k\to (A(k)\vee B(k))\to C(k))\\
-		\forall k(u\preceq k\to (\forall l(k\preceq l \to A(l)\to B(l)))\to C(k)),
-	\end{matrix}$$
-	with each $P(x)$ possibly being $\bot$. Applying Herbrandization we obtain a formula
-	$$\varphi''' = K(\varphi')\to \bigwedge \mathcal S'\to P_\varphi(s),$$
-	where $s$ is some new constant term and each $\psi\in\mathcal S'$ is of one of the forms
-	$$\begin{matrix}
-		\forall k(s\preceq k\to A(k)\to (B(k)\wedge C(k)))\indent \forall k(s\preceq k\to (A(k)\wedge (B(k))\to C(k)))\\
-		\forall k(s\preceq k\to A(k)\to (B(k)\vee C(k)))\indent \forall k(s\preceq k\to (A(k)\vee (B(k))\to C(k)))\\
-		\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A(f_\psi(k))\to B(f_\psi(k)))\to C(k)),\\
-	\end{matrix}$$
-	and $f_\psi$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model, then we have a counter-model $(M',I')$ where $M' = \{m\in M|s\preceq m\}$ and
-	$f_\psi^{I'}(u) = f^I_\psi(u) \text{ iff $u\preceq f^{I}_\psi(u)$, and $f_\psi^{I'}(u) = u,$ else.}$ That is, where $s$ is the least element and the $f_\psi$ increase their argument.
-Therefore instead of $\varphi'''$ we may consider a formula $$\varphi^\circ = \forall k(s\preceq k)\to \bigwedge\{\forall k(k\preceq f_\psi(k))\:|\:\psi\in\mathcal F_\to\}\to K(\varphi')\to \bigwedge \mathcal S^\circ\to P_\varphi(s)$$ where each formula $\psi\in\mathcal S^\circ$ is of one of the forms
-	$$\begin{matrix}
-		\forall k(A(k)\to (B(k)\wedge C(k)))\indent\forall k((A(k)\wedge B(k))\to C(k))\\
-		\forall k(A(k)\to (B(k)\vee C(k)))\indent\forall k((A(k)\vee B_\psi(k))\to C(k))\\
-		\forall k((A(f_\psi(k))\to B(f_\psi(k)))\to C(k)),
-	\end{matrix}$$
-	and $\mathcal F_\to$ denotes the set of formulas of the last kind.
-
-	\subsection{IPC to CPC}
-
-	We now argue that if a counter-model exists for $\varphi^\circ$ then a bounded counter-model exists, allowing us to return to the propositional case.
-
-	\begin{definition}
-		Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
-		We say $\psi\in\mathcal F_\to$ is \emph{fulfilled} at $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true.
-For $\psi\in\mathcal F_\to$ define $$g_\psi : M\to M, u\mapsto\begin{cases}
-			u,&\text{ if $\psi$ is fulfilled in $u$,}\\
-			f^I_\psi(u),&\text{ else.}		
-		\end{cases}$$Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
-		\begin{itemize}
-			\item $M_T$ is the set of sequences without repetition on $\mathcal F_\to$.
-			\item Interpret $\preceq$ as the prefix-order.
-			\item We set $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
-				\psi_1\dots\psi_n, &\text{if $\psi$ occurs in $\psi_1\dots\psi_n$,}\\
-				\psi_1\dots\psi_n\psi, &\text{else.}			
-			\end{cases}$$
-			\item For propositional variables $P$, we set $$P^{I_T}(\psi_1\dots \psi_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots)).$$
-		\end{itemize}
-	\end{definition}
-	\begin{lemma}~\label{thm:prop-countermodel-reduction}
-		Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
-		\begin{enumerate}
-			\item $\psi$ is fulfilled at $f_\psi^I(u)$ for all $u\in M, \psi\in\mathcal F_\to$.
-			\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\succeq u$.
-		\end{enumerate}
-	\end{lemma}
-
-	\begin{proof}
-		1. If $C^I(u)$ is true then we are done due to persistency. Otherwise $C^I(u)$ is false.
-		Because $(A^I(f_\psi^I(u))\to B^I(f_\psi^I(u)))\to C^I(u)$ holds, then $A^I(f_\psi^I(u))\to B^I(f_\psi^I(u))$ must be false.
-
-		2. Let $w$ be some element with $w\geq u$.
-		If $C^I(w)$ is true, we are done.
-		Otherwise $C^I(w)$ is false.
-		Due to persistency $C^I(u)$ is also false.
-		Because $\psi$ is fulfilled at $u$, $A^I(u)\to B^I(u)$ must be false, i.e. $A^I(u)$ is true and $B^I(u)$ is false. Then $A^I(w)$ and $A^I(f^I_\psi(w))$ are also true due to persistency.
-		Because $(A^I(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I(w)$ holds, we now get that  $B^I(f^I(w))$ must be false.
-		But then due to persistency so is $B^I(w)$ and we have that $\psi$ is fulfilled at $w$.
-	\end{proof}
-
-	Essentially what this tells us is that for every $u$ at which $\psi$ is fulfilled, in particular for each $u$ in the image of $f_\psi$, we may add the assumption that $f_\psi(w) = w$ for all $w\geq u$ without changing the validity of $\psi$.
-	With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.
-
-	\begin{corollary}
-		If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
-	\end{corollary}
-
-	\begin{example}
-		Consider $$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model, depicted below,
-		\begin{center}
-			\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
-				\node (center) {};
-				\node[squarednode]      (right)   [right=of center] {$B, C$};
-				\node[squarednode]      (left)    [left=of center] {$A, D$};
-				\node[squarednode]      (lower)       	[below=of center] {};
-				\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left.south east);
-				\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right.south west);			
-			\end{tikzpicture}
-		\end{center}
-		\vspace*{-.3cm}where at each node the true propositions are indicated.
-		%	
-		There is a corresponding classical counter-model $\mathcal M = (M, I)$ to $\varphi^\circ$ with $M = \{u, u_l, u_r\}$, $A^{I}(l) = D^I(l) = B^I(r) = C^I(r) = 1$ and $0$ else.
-		%
-		Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (B\to A)\to D$, this corresponds to a transformed counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below:
-		\begin{center}
-			\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
-				\node (center) {};
-				\node[squarednode]      (right1)   [right=of center] {$B, C$};
-				\node[squarednode]      (right2)   [above=of right1] {$B, C$};
-				\node[squarednode]      (left1)    [left=of center] {$A, D$};
-				\node[squarednode]      (left2)    [above=of left1] {$A, D$};
-				\node[squarednode]      (lower)       	[below=of center] {};
-				\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left1.south east) node [midway, below left] {$f_{\psi_1}$};
-				\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right1.south west) node [midway, below right] {$f_{\psi_2}$};
-				\draw[-{Stealth[length=3mm, width=3mm]}] (left1.north) -- (left2.south) node [midway, right] {$f_{\psi_2}$};
-				\draw[-{Stealth[length=3mm, width=3mm]}] (right1.north) -- (right2.south) node [midway, left] {$f_{\psi_1}$};
-				\draw[-{Stealth[length=3mm, width=3mm]}] (left1.south west) to [out=225, in=135, loop, looseness=3] node [left] {$f_{\psi_1}$} (left1.north west);		
-				\draw[-{Stealth[length=3mm, width=3mm]}] (right1.south east) to [out=315, in=45, loop, looseness=3] node [right] {$f_{\psi_2}$} (right1.north east);
-				\draw[-{Stealth[length=3mm, width=3mm]}] (left2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (left2.north east);
-				\draw[-{Stealth[length=3mm, width=3mm]}] (right2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (right2.north east);		
-			\end{tikzpicture}
-		\end{center}
-	\end{example}
-
-	\begin{remark}
-		It is well known that Kripke frames which are trees are complete with regard to intuitionistic logic.
-Our contribution is giving an explicit geometry that depends only on the structure of the considered formula.
-	\end{remark}
-
-	So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
-		x, &\text{ if $\psi$ occurs in $x$,}\\
-		x\psi, &\text{ otherwise.}
-	\end{cases}$$
-	Now we can replace every $\forall$-quantifier by enumerating all the ground terms, and then replace all distinct ground instances of relations by new propositional variables yielding a propositional formula.
-This gives us the following theorem.
-
-	\begin{theorem}
-		Let $\mathcal S$ be a set of formulas of the form
-		$$(A\wedge B)\to C, (A\vee B)\to C, A\to (B\wedge C), A\to(B\vee C), (A\to B)\to C$$
-		and $\mathcal F_\to\subseteq\mathcal S$ denote the subset of formulas of the form $(A\to B)\to C$ and $\Lambda$ denote the set of sequences without repetition over $\mathcal F_\to$.
-For each atom $A$ and $k\in\Lambda$ consider a new atom $A^k$.
-Obtain $\mathcal S^\#$ by including the following formulas:
-		\begin{itemize}
-			\item $A^k\to A^{k\psi}$ for each atom $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
-			\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-			\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
-			\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
-		\end{itemize}
-		Then, $\bigwedge S\to P$ is intuitionistically valid iff $\bigwedge S^\#\to P^\epsilon$ is classically valid, where $\epsilon$ denotes the empty sequence.
-The size of $\mathcal S^\#$ is in $\mathcal O(|\mathcal S|\cdot2^{|\mathcal F_\to|\cdot\log(|\mathcal F_\to|)})$.
-Futhermore, there is an effective procedure for translating counter-models between the sentences.
-	\end{theorem}
-
-	\begin{example}
-		Let us consider the law of excluded middle $\varphi = A\vee\neg A$. Then we get
-		$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\},$$
-		and we know that $\varphi$ is valid iff $\bigwedge \mathcal S^-(\varphi)\to P_\varphi$ is valid.
-Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically valid, where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
-		\begin{align*}
-			\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi}, (P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi}\}\cup\\ &\{A^\epsilon\to P_A^\epsilon, P_A^{\psi}\to P_A^{\psi}, (A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}.
-		\end{align*}
-		We now note that $P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$, $P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$ defines a counter-model.
-	\end{example}
-
-	\section{IPC to QBF}
-
-	It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}.
-	However, giving an explicit translation from IPC to QBF still promises to be useful, allowing to leverage the progress in QBF solving.
-	The translation from the last section will serve as starting point for developing the translation in this section.
-	For lack of space, we do not define the semantics of QBF in this paper and refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
-
-	Fix some set $\mathcal S$ of formulas of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
-	and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$.
-We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model.
-The idea how to get a polynomially-sized formula is: Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
-The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e.
-express which path is considered.
-
-	\begin{definition}
-		For every propositional variable $X$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $X^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$.
-Let $\vec X^n$ range over all propositional variables $X^n$ and $\vec F^n$ range over all $F_\psi^n$.
-Define
-		\begin{itemize}
-			\item {\normalfont Valid}$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
-if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
-			\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
-			\item {\normalfont Sat}$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
-			\item {\normalfont Sat}$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
-		\end{itemize}
-		Have $k = |\mathcal F_\to|$ and define$$\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal S}(i)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
-		for $0 < i < k$, as well as the special cases
-		$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
-		for leafs, and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
-		Example encodings of the above formulas are:
-		\begin{align*}
-			\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
-			\text{\normalfont Persistent}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
-			\text{\normalfont Sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
-			&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
-			\text{\normalfont Sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
-		\end{align*}
-	\end{definition}
-
-	\begin{example}
-		With the previous encoding for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
-		$$\hspace*{-.3cm}
-		\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right),
-		$$
-		which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $F^0$.
-	\end{example}
-
-	\begin{lemma}
-		$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
-	\end{lemma}
-
-	\begin{proof}
-		Suppose $\varphi$ is not intuitionistically valid.
-		Then, $\bigwedge\mathcal S\to P$ has an intuitionistic counter-model. Then by the previous section there exists a classical counter-model $\mathcal M$ for $\bigwedge\mathcal S^\#\to P^\epsilon$.
-		We now define a QBF model iteratively.
-		For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$.
-Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$  in $\mathcal M$.
-		Let the $F^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified).
-If not exactly one of the $F^{n}_\psi$ is interpreted as true, then valid$(n)$ fails and the remaining propositions can be chosen arbitrarily.
-If $F^i_{\psi_n} = 1$ for some $i < n$, then valid$(n)$ also fails and the remaining propositions can be chosen arbitrarily.
-		So we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions.
-		Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{\psi_1\dots\psi_n}$.
-		Continue this construction until $n  = |\mathcal F_\to|$. Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.
-
-		On the other hand suppose $\varphi^Q$ is satisfiable. We construct a counter-example to $\varphi^\#$.
-		Again we proceed iteratively.
-Interpret $A^\epsilon$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$.
-Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$, in which $F^i_\psi$ is chosen true iff $\psi = \psi_i$.
-Let $\psi_n\in\mathcal F_\to$.
-Consider some interpretation of the $A^n$ that are part of a satisfying assignment where $F^n_\psi$ is true iff $\psi = \psi_n$ and all variables quantified above are chosen as before.
-Have $A^{\psi_0\dots\psi_n} = A^n$ for each propositional variable $A$.
-From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.
-	\end{proof}
-
-	A simple counting argument shows that for each $n < |\mathcal F_\to|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal F_\to|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal S}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal F_\to}(n)|\in \mathcal O(|\mathcal F_\to|)$ so overall we get:
-	\begin{lemma}
-		The size of $\varphi^Q$ is in $\mathcal O(|\varphi|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$.
-	\end{lemma}
-
-	Note that the last set of universally quantified variables can be avoided because there is only one assignment that has the chance to falsify $\varphi^Q$, namely the assignment that assigns $1$ to the single $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$.
-Hence, in our final formula we can replace every $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^n$.
-With that we get the following:
-
-	\begin{corollary}
-		Let $N$-\textbf{IntInvalid} be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid.
-Then $N$-\textbf{IntInvalid} is in $\Sigma_{2N-1}^P$.
-The dual problem $N$-\textbf{IntValid} is in $\Pi_{2N-1}^P$.
-	\end{corollary}
-
-	\section{CQC to IQC}
-
-	We now give an analogous transformation for first order logic.
-Our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic.
-However we do it nonetheless!
-The domain of our classical model will feature on the one hand elements representing worlds in a Kripke frame and on the other hand the domain-elements from the Kripke model.
-We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$.
-We fix some signature $\Sigma$ for this section.
-To encode that some $n$-ary relation $A$ holds at a world $u$ we replace each $n$-ary relation symbol $A$ with a $n+1$-ary relation symbol $A^\#$, interpreting $A^\#(\vec x, u)$ as "$A(\vec x)$ holds at u".
-Furthermore we write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$.
-
-	First of all recall, that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
-	\[\begin{matrix}
-		\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c)))\indent \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c))\indent \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c)))\\
-		\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c))\indent \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\indent\forall \vec z(\forall xA(\vec a)\to B(\vec b))\\ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b)).
-	\end{matrix}\]
-	The complete result that we will show is as follows:
-	\begin{theorem}\label{fullFOtranslation}
-		Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain the formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$.
-Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
-For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$.
-For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$.
-Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
-Obtain $\mathcal S^\#$ by including formulas as follows:
-		\begin{itemize}
-			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
-			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
-			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
-			\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\psi = \forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
-			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
-			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
-			\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
-		\end{itemize}
-		Then $\bigwedge S\to P$ is intuitionistically iff $\exists xE(x, s)\to K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol, $\Sigma$ is the original signature and
-		\begin{align*}
-			K := &\:\bigwedge\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F\}\wedge\\
-			&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
-			&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}.
-		\end{align*}
-		The size of the obtained formula is linear in the input.
-However $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
-	\end{theorem}
-
-	It is not possible to completely eliminate worlds from the domain as in the propositional case, since counter-models can possibly be infinite (as we will see in more detail later).
-
-	\subsection{Encoding Kripke Semantics}
-
-	As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$, but also $E$ has to be axiomatized. For that we use the following predicates:
-	\begin{itemize}
-		\item PartialOrder($\preceq$) encoding that $\preceq$ is a partial order.
-		\item DomainSubset$(u, w)$ encoding that the domain at $u$ is a subset of that at $w$.
-		\item World($u$) encoding that $u$ represents a world.
-		\item DomainClosed($u$) which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
-		\item Persistent$(u, w)$ which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$ at $u$ we have $A(x)$ at $w$.
-	\end{itemize}
-	Then, analogously to the propositional case, we define:
-	\begin{align*}
-		K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
-		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
-	\end{align*}
-	Example encodings of these formulas are as follows:
-	\begin{align*}
-		\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
-		\text{DomainSubset}(u, w) = &\:\forall z(E(z, u)\to E(z, w))\\
-		\text{World}(u) = &\:\exists xE(x, u)\\
-		\text{DomainClosed}(u) = &\:\bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f\in\Sigma$ is a function symbol}\}\\
-		\text{Pesistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
-	\end{align*}
-
-	Using $E$ we then obtain $\mathcal S'$ analogously to the predicate case by including formulas:
-
-	\begin{itemize}
-		\item $\forall \vec z\forall u(s\preceq u\to \vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
-		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
-		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to\forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$\\ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
-		\item  $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
-		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
-		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
-		\item $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
-	\end{itemize}
-
-	Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid if and only if
-	$$\varphi' = \text{World}(s)\to K(\varphi)\to \bigwedge S'\to P^\#(s)$$
-	is classically valid, where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced.
-
-	\begin{proof}
-		We proceed by translation of counter-examples.
-Suppose first we have a counter-example $\mathcal M = (M, I)$ to $\varphi'$.
-As a Kripke frame $(W, \leq)$ take all $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ let $\leq$ be $\preceq^I$ restricted to $W$.
-Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A^\#(\vec x, u)$.
-It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
-
-		The other direction is a bit more involved.
-Suppose we have a Kripke counter-model to $\varphi$ with frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
-In particular since it is a counter-model there exists $w_0\in W$ with $w_0\not\models\varphi$.
-Let $W_0 = \{w\in W\:|\: w\geq w_0\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
-Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
-		Now have
-		\begin{itemize}
-			\item $s^I = w_0$.
-			\item $E^I(m, w)$ iff $w\in W_0$ and $m \sim [x, w]$ for some $x\in M_w$.
-			\item $f^I(m_1\dots m_n) =\begin{cases}
-				f^{I_u}(x_1\dots x_n), &\text{if there are $ u\in W_0, x_i\in M_u$ with $m_i\sim [x_i, u]$ for all $i$,}\\
-				w_0, & \text{otherwise.}
-			\end{cases}$
-			\item ${A^\#}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
-				A^{I_u}(x_1\dots x_n), &\text{if $u\in W_0$ and $\exists x_i\in M_u$ w. $m_i\sim [x_i, u]$ for all $i$,}\\
-				\top, & \text{otherwise.}
-			\end{cases}$
-		\end{itemize}
-		One easily verifies that these are well-defined.
-Let us now briefly check that this indeed defines a counter-model.
-First of all ${P^\#}^I(s^I)$ is false since $P^{I_{w_0}}$ is.
-Clearly World$(w_0)$ holds and $K(\varphi)$ is also easily verified.
-All that remains to show is $M, I\models\bigwedge\mathcal S'$.
-Consider e.g.
-the case where $\psi\in\mathcal S$ is of the form
-		$$ \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$$
-		We have to show that
-		$$M, I\models \forall \vec z\forall u(\vec E(t, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$$
-		holds. Suppose towards contradiction that it does not, i.e. there are $u, \vec z\in M$ such that for all $w\in M$
-		$$\vec E^I(\vec z, u)\to (u\preceq w\to {A^\#}^I(\vec a, w)\to {B^\#}^I(\vec b, w))\to {C^\#}^I(\vec c, u)$$ is false.
-For that ${C^\#}^I(\vec c, u)$ must be false and in particular $u\in W_0$.
-Futhermore $\vec E^I(\vec z, u)$ must be true and we can write $\vec z = [z_1, u]\dots[z_n, u]$ for some $z_i\in W_u$.
-Note that the above is false in particular for $w\in M_0$ with $u\preceq w$.
-But this would imply
-		$$((A^{I_u}(\vec a)\to B^{I_u}(\vec b))\to C^{I_u}(\vec c))[\vec z/z_1\dots z_n]$$in our original Kripke counter-model. The other cases are analogous.
-	\end{proof}
-
-	\subsection{Reducing the encoding}
-
-	As in the predicate case we now show the existence of a canonical counter-model (in case such a counter-model exists).
-Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$.
-Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$.
-For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$.
-As before we may assume that $s$ is a least element and we can avoid the quantification over $w$ by considering the formulas
-	\begin{itemize}
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$,
-		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$,
-	\end{itemize}
-	and by removing removing the $s\preceq u$ for the remaining formulas in $\mathcal S'$.
-	We denote the resulting set of formulas by $\mathcal S^\#$.
-	Then, the previous formulas are equivalid to
-	$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#(s).$$
-	As in the previous section we can define a tree counter-model for this.
-
-	\begin{definition}
-		Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
-		\begin{itemize}
-			\item $M_T$ are the sequences on $\{(\vec z, \psi)\:|\:\psi\in \mathcal F, \vec z\in M^n\}$.
-			\item We interpret $\preceq$ as the prefix-order.
-			\item We set $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi).$$
-			\item For relation symbols $R^\#$, we set $${R^\#}^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = {R^\#}^I(\vec z, f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots)).$$
-		\end{itemize}
-	\end{definition}
-
-	\begin{lemma}
-		If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
-	\end{lemma}
-
-	Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$f_{\psi_1}(\vec z_1, f_{\psi_2}(\vec z_2, \dots f_{\psi_n}(\vec z_n, s)\dots))$$ is sufficient.
-	In particular, this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
-	from $K(\varphi)$ and adjust DomainSubset$(u, w)$ and Persistent$(u, w)$, e.g. remove quantification over $w$ and the precondition $u\preceq w$ from $K(\varphi)$ and replace DomainSubset$(u, w)$ with
-	$$\bigwedge\{\forall \vec z\forall u\forall x(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$$ and Persistent$(u, w)$ with $$\bigwedge\{\forall\vec z_1\forall\vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\psi\in\mathcal F_\to, \text{$A\in\Sigma$ is a relation symbol}\}.$$ We set
-
-	$$\varphi^{\#} = K\to \bigwedge S^\#\to P^\#.$$
-	Then, we obtain the following result:
-	\begin{theorem}
-		\label{thm:fo-reduction}
-		$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
-	\end{theorem}
-	The complete translation is summarised in Theorem~\ref{fullFOtranslation} (stated earlier). Let us understand why a further reduction as in the propositional case is impossible.
-
-	\begin{definition}
-		\label{def:fo-fulfilled}
-		Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$.
-For $\psi = \forall\vec z\psi'\in\mathcal F_\to$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff $(A_\psi^I(\vec a, u)\to B_\psi^I(\vec b, u))[\vec m/\vec z]$ is false or $C_\psi^I(\vec c, u)[\vec m/\vec z]$ is true.
-		For $\psi = \forall\vec z\psi'\in\mathcal F_\forall$ we say it is \emph{fulfilled} at $u\in M$ with $\vec m\in M^n$ iff there is $m\in M$ with $E^{I}(m, u)$ such that $A^{I[m/x]}_\psi(\vec a)[\vec m/\vec z]$ is false or $B_\psi^I(\vec b)[\vec m/\vec z]$ is true.
-	\end{definition}
-
-	We get an analogous result to Lemma~\ref{thm:prop-countermodel-reduction}:
-
-	\begin{lemma}
-		\label{thm:fo-countermodel-reduction}
-		Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
-		\begin{enumerate}
-			\item $\psi$ is fulfilled at $f_\psi^I(\vec a, u)$ with $\vec a$ for all $u\in M, \psi\in\mathcal F$, $\vec a\in M^n$.
-			\item If $\psi\in\mathcal F_\to$ is fulfilled at $u\in M$ with $\vec a$ then $\psi$ is fulfilled at all $w\geq u$ with $\vec a$.
-		\end{enumerate}	
-	\end{lemma}
-
-	However part 2 of Lemma~\ref{thm:fo-countermodel-reduction} fails for $\psi\in\mathcal F_\forall$, because even if for all $x$ with $E(x, v$), we have $A(x, v)$, there could be some term $y$ with $\neg E(y, v)$ and $w$ with $v\preceq w$ such that $E(y, w)$ and $\neg A(y, w)$.
-This is why we must consider sequences that have repetitions:
-
-	\begin{wrapfigure}[16]{r}{0.4\textwidth}
-		\vspace{-32pt}
-		\begin{center}
-			\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
-				\node[squarednode]      (0)   [] {$ $};
-				\node[squarednode]      (1r)   [above right=of 0] {{\color{blue}$1$}};
-				\node[squarednode]      (1l)   [above left=of 0] {{\color{red}$1$}};
-				\node[squarednode]      (2r)   [above=of 1r] {{1, \color{blue}$2$}};
-				\node[squarednode]      (2l)   [above=of 1l] {{1, \color{red}$2$}};
-				\node[squarednode]      (3r)   [above=of 2r] {{1, 2, \color{blue}$3$}};
-				\node[squarednode]      (3l)   [above=of 2l] {{1, 2, \color{red}$3$}};
-				\node      (4r)   [above=of 3r] {{$\vdots$}};
-				\node      (4l)   [above=of 3l] {{$\vdots$}};
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (0.north east) -- (1r.south west);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (0.north west) -- (1l.south east);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1r.north) -- (2r.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1l.north) -- (2l.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1l.north east) -- (2r.south west);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (1r.north west) -- (2l.south east);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2r.north) -- (3r.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2l.north) -- (3l.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2l.north east) -- (3r.south west);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (2r.north west) -- (3l.south east);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3r.north) -- (4r.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3l.north) -- (4l.south);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3l.north east) -- (4r.south west);
-				\draw[-{Stealth[length=3mm, width=3mm]}, color=black] (3r.north west) -- (4l.south east);
-			\end{tikzpicture}
-		\end{center}
-		\caption{Counter-model to $\varphi$.}
-		\label{fig1}
-	\end{wrapfigure}
-	\begin{example}
-		Consider $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model as shown in Figure~\ref{fig1},
-		where we indicate at each world the elements that exist at it.
-For black numbers $n$, both $A(n)$ and $B(n)$ hold at that world, for red ones only $A(n)$, for blue ones only $B(n)$.
-The arrows indicate the order.
-$\neg\forall xA(x)$ is not fulfilled at worlds in the left column and $\neg\forall xB(x)$ is not fulfilled at worlds in the right column.
-In particular applying $f_{\neg \forall xA(x)}$ we must always reach the right column and applying $f_{\neg \forall xB(x)}$ we must always reach the left column.
-Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we cannot remain stationary.
-	\end{example}
-
-	Let us add that Lemma~\ref{thm:fo-countermodel-reduction} still equips us with additional information that could be used to guide proof search: We could add additional axioms that restrict the interpretation of functions.
-That is, we may assume that $f_\psi(w) = w$, if $\exists u(w = f_\psi(u))$ for all $\psi\in\mathcal F_\forall$ (i.e.
-that $f_\psi$ is idempotent), and $f_\psi(w) = w$, if $\exists u(w \geq f_\psi(u))$ for all $\psi\in\mathcal F_\to$ (which is even stronger than idempotence).
-
-	\chapter{Implementation}
-
-	\chapter{Translation of proofs and counter-models}
-
+	\begingroup
+	\setstretch{0.8}
 	\bibliographystyle{plain}
 	\bibliography{EIICL}
-
+	\endgroup
 \end{document}

### 2023-03-07 LPAR/LPAR-EIICL.tex
formatting

@@
 \newtheorem{remark}[theorem]{Remark}
+
+\let\OLDitemize\itemize
+\renewcommand\itemize{\OLDitemize\addtolength{\itemsep}{-5pt}}
 
@@
 \begin{itemize}
-	\item Translate an input formula $\varphi$ to a formula $\varphi^\#$, which is classically valid if and only if $\varphi$ is intuitionistically valid.
-	\item Determine the classical validity of $\varphi^\#$ using a prover for classical logic, resulting in a proof or counter-model.
-	\item Translate the classical proof/counter-model of $\varphi^\#$ to a corresponding intuitionistic proof/counter-model of $\varphi$.
+	\item Give for each $\varphi$ a formula $\varphi^\#$, which is classically valid iff $\varphi$ is intuitionistically valid.
+	\item Determine the classical validity of $\varphi^\#$ using a prover for classical logic.
+	\item Translate the classical proof (resp. counter-model) of $\varphi^\#$ given by the prover to a corresponding intuitionistic proof (resp. counter-model) of $\varphi$.
 \end{itemize}
 \subsection*{Contributions} 
-We now overview the main results of this paper (which we discuss in more detail in Sections~\ref{sec:overview} and~\ref{sec:evaluation}):
+The following main contributions will be discussed in more detail in Sections~\ref{sec:overview} and~\ref{sec:evaluation}:
 \begin{itemize}
@@
 We can define terms $t$ and substitutions $t[\nu/b]$ by simultaneous induction as follows:
+\vspace*{-.3cm}
 \begin{itemize}
@@
 	Let $\varphi$ be a propositional formula. Define $\varphi^{\nu}$ inductively:
+	\vspace*{-.1cm}
 	\begin{itemize}
@@
 	Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
-	$$K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))$$
+	\vspace*{-.1cm}
+	\[K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))\]
+	\vspace*{-.4cm}
 	with e.g.
+	\vspace*{-.2cm}
 	\begin{align*}
@@
 	Then define
-	$$\varphi^{C} = K(\varphi)\to \varphi^{b}$$
+	$\varphi^{C} = K(\varphi)\to \varphi^{b}$
 	where $b$ is a new free variable.
@@
 	Because $(A^I(f^I_{[\psi]}(n))\to B^I(f^I_{[\psi]}(n)))\to C^I(n)$ holds, we now get that  $B^I(f^I_{[\psi]}(n))$ must be false.
-	But then due to persistency so is $B^I(n)$ and we have that $\psi$ is fulfilled at $n$.\hfill$\square$
+	But then due to persistency so is $B^I(n)$ and we have that $\psi$ is fulfilled at $n$.
 \end{proof}
@@
 
-\begin{wrapfigure}[8]{r}{0.4\textwidth}
-	\vspace*{-.5cm}
+It is not possible to completely eliminate worlds from the domain by enumerating them since there are formulas with no finite counter-examples.
+An analogous procedure to the propositional case fails, however we can still retrieve additional axioms that restrict the interpretation of functions.
+That is, we may assume that $f_\psi(\vec t, w) = w$, if $\exists u(w = f_\psi(\vec t, u))$ for all $\psi\in\mathcal X\cup\mathcal Q$ (i.e.
+that $f_\psi$ parametrized by $\vec t$ is idempotent), and $f_\psi(\vec t, w) = w$, if $\exists u(w \geq f_\psi(\vec t, u))$ for all $\psi\in\mathcal X$ (which is even stronger than idempotence).
+
+\section{Experimental Evaluation}
+\label{sec:evaluation}
+
+\begin{wrapfigure}[4]{r}{0.4\textwidth}
+	\vspace*{-2cm}
 	\begin{tabular}{l|c|c}
@@
 
-It is not possible to completely eliminate worlds from the domain by enumerating them since there are formulas with no finite counter-examples.
-An analogous procedure to the propositional case fails, however we can still retrieve additional axioms that restrict the interpretation of functions.
-That is, we may assume that $f_\psi(\vec t, w) = w$, if $\exists u(w = f_\psi(\vec t, u))$ for all $\psi\in\mathcal X\cup\mathcal Q$ (i.e.
-that $f_\psi$ parametrized by $\vec t$ is idempotent), and $f_\psi(\vec t, w) = w$, if $\exists u(w \geq f_\psi(\vec t, u))$ for all $\psi\in\mathcal X$ (which is even stronger than idempotence).
+We have benchmarked our first-order translation on version 1.1.2 of the ILTP problem set~\cite{iltp} by performing our translation and then examining the generated classical problem using Vampire~\cite{kovacs2013first}.
+For comparison the state-of-the-art intuitionistic provers ileanCoP 1.2~\cite{otten2008leancop} and nanoCoP-i 2.0~\cite{otten2021nanocop} were used.
+All benchmarks were conducted on an Intel Core i5-8400 2.80GHz machine running Linux with a time limit of 10s per problem.
+The problem \textsc{SYN007+1} was removed as our normal form procedure would explode the size of instance.
 
-\section{Experimental Evaluation}
-\label{sec:evaluation}
-
-\begin{wrapfigure}[13]{r}{0.4\textwidth}
-	\vspace*{4.9cm}
-	\hspace*{1cm}
+\begin{wrapfigure}[16]{r}{0.4\textwidth}
+	\vspace{6cm}
+	\hspace*{1.5cm}
 	\begin{tikzpicture}
@@
 	\end{tikzpicture}
-	\vspace*{-7.1cm}
+	\vspace*{-7.5cm}
 	\caption{Instances proven by combinations of Embedding (red), Embedding (idem.) (green), ileanCoP or nanoCoP (blue)}
@@
 \end{wrapfigure}
-
-We have benchmarked our first-order translation on version 1.1.2 of the ILTP problem set~\cite{iltp} by performing our translation and then examining the generated classical problem using Vampire~\cite{kovacs2013first}.
-For comparison the state-of-the-art intuitionistic provers ileanCoP 1.2~\cite{otten2008leancop} and nanoCoP-i 2.0~\cite{otten2021nanocop} were used.
-All benchmarks were conducted on an Intel Core i5-8400 2.80GHz machine running Linux with a time limit of 10s per problem.
-The problem \textsc{SYN007+1} was removed as our normal form procedure would explode the size of instance.
 
@@
 Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers.
-Notably, while there is big overlap in solved instances, ileanCoP and nanoCoP also have a significant symmetric difference.\pagebreak
+Notably, while there is big overlap in solved instances, ileanCoP and nanoCoP also have a significant symmetric difference.
 \begin{figure}
@@
 \end{figure}
-\vspace*{-.5cm}
 
@@
 Have $A^{\psi_0\dots\psi_n} = A^n$ for each propositional variable $A$.
-From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.\hfill$\square$
+From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.
 \end{proof}

### 2023-03-07 LPAR/LPAR-EIICL.tex
Syntax redefined, Proof 7.2 to appendix.

@@
 
-For propositional logic we assume an ambient set of propositional variables. A propositional formula is defined inductively as follows:
-\begin{itemize}
-	\item The propositional constant $\bot$ is a formula.
-	\item Each propositional variable is a formula.
-	\item Given formulas $\varphi$ and $\psi$ and a connector $\circ\in\{\wedge, \vee,\to\}$ there is a formula $\varphi\circ\psi$.
-\end{itemize}
+For propositional logic we assume an ambient set of propositional variables.
+As usual a formula is either $\bot$, a propositional variable or, given formulas $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$.
 
-For predicate logic we assume an ambient signature $\Sigma$, i.e., a set of function and predicate symbols with their respective arities, and sets of free and bound variables.
-Generally free variables will be denoted will letters at the beginning of the alphabet while bound variables will be denoted with letters at the end of the alphabet.
-$\nu$ will usually be a placeholder that can be instantiated with free or bound variables.
-We can define terms $t$ and substitutions $t[\nu/b]$ by simultaneous induction as follows:
-\vspace*{-.3cm}
-\begin{itemize}
-	\item Each free variable $a$ is a term and $a[\nu/b] := \begin{cases}
-		\nu & \text{if } a = b \\
-		a & \text{otherwise}
-	\end{cases}$.
-	\item For each $n$-ary function symbol $f$ and terms $t_1, \ldots, t_n$ we have that $f(t_1, \ldots, t_n)$ is a term and $f(t_1, \ldots, t_n)[\nu/b] := f(t_1[\nu/b], \ldots, t_n[\nu/b])$.
-\end{itemize}
-We can then define formulas $\varphi$ and substitutions $\varphi[\nu/b]$ by simultaneous induction as follows:
-\begin{itemize}
-	\item The constant $\bot$ is a formula and $\bot[\nu/b] := \bot$.
-	\item For each $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$ we have that $P(t_1, \ldots, t_n)$ is a formula and $P(t_1,\ldots, t_n)[\nu/b] := P(t_1[\nu/b], \ldots, t_n[\nu/b])$.
-	\item Given formulas $\varphi$ and $\psi$ and a connector $\circ\in \{\wedge,\vee,\to\}$ there is a formula $\varphi\circ\psi$ and $(\varphi\circ\psi)[\nu/b] := \varphi[\nu/b]\circ\psi[\nu/b]$.
-	\item Given a formula $\varphi$, a free variable $a$, a bound variable $x$ not occurring in $\varphi$ and a quantifier $Q\in\{\exists,\forall\}$ there is a formula $Qx(\varphi[x/a])$ and $Qx(\varphi[x/a])[\nu/b] := Qx(\varphi[x/a][\nu/b])$.
-\end{itemize}
+For predicate logic we assume an ambient signature $\Sigma$, i.e., a set of function and predicate symbols with their respective arities, and a set of variables.
+As usual we can inductively define terms and formuas.
+A term is either a variable or, given an $n$-ary function symbol $f$ and terms $t_1, \ldots, t_n$, $f(t_1, \ldots, t_n)$.
+A formula is either $\bot$, or, given a $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$, $P(t_1, \ldots, t_n)$, or, given formulas $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$, or, given a formula $\varphi$ and a variable $x$, $\exists x\varphi$ or $\forall x\varphi$.
+Free and bound variables are defined as usual and we assume that they comprise distinct sets, such that bound variables are not affected by substitutions of free variables.
+Generally we will denote free variables with letters at the beginning of the alphabet and bound variables with letters at the end of the alphabet.
 
@@
 	 Here $a$ is a free variable that does not occur in $\varphi$.
-A $\Sigma$-structure $\mathcal M$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$, we write $\mathcal M\models\varphi$.
-A formula is valid if every $\Sigma$-structure $\mathcal M$ satisfies it.
+$\varphi$ is satisfiable if $\mathcal M, v\models\varphi$ for some $\mathcal M, v$.
+It is valid if $\mathcal M, v\models\varphi$ for every $\mathcal M, v$.
 We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).
@@
 $\varphi$ is valid if it is satisfied by every Kripke structure.
-	$\varphi$ is valid for some frame (write $W\models\varphi$) if it is satisfied in every Kripke structure with that frame.
 	We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
@@
 
-\begin{proof}
-	Since we are mixing worlds and proper intuitionistic domain elements in a single classical domain there are some details to be taken care of. We proceed by translation of counter-models.
-
-	Suppose first we have a counter-model $\mathcal M = (M, I)$ to $\varphi^C$.
-As a Kripke frame $(W, \leq)$ take $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ and let $\leq$ be $\preceq^I$ restricted to $W$.
-Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A(\vec x, u)$.
-It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
-
-	The other direction is a bit more involved.
-Suppose we have a Kripke counter-model to $\varphi$ with frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
-In particular, since it is a counter-model there exists $w_0\in W$ and a variable assignment $v$ with $w_0, v\not\models\varphi$.
-Let $W_0 = \{w\in W\:|\: w_0\preceq w\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
-Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
-	Now have
-	\begin{itemize}
-		\item $\preceq^I = \leq\cup \{(m, m): m\in M\setminus W_0\}$ as a set, i.e. $\preceq$ corresponds to $\leq$ on $W_0$ and is trivial elsewhere.
-		\item $E^I(m, w)$ iff $w\in W_0$ and $m = [x, w]$ for some $x\in M_w$.
-		\item $f^I(m_1\dots m_n) =\begin{cases}
-			f^{I_u}(x_1\dots x_n), &\substack{\text{if there are $ u\in W_0, x_i\in M_u$}\\ \text{with $m_i= [x_i, u]$ for all $i$,}}\\
-			w_0, & \text{otherwise.}
-		\end{cases}$
-		\item ${A}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
-			A^{I_u}(x_1\dots x_n), &\substack{\text{if $u\in W_0$ and $\exists x_i\in M_u$}\\\text{with $m_i= [x_i, u]$ for all $i$,}}\\
-			\bot, & \text{otherwise.}
-		\end{cases}$
-	\end{itemize}
-	One easily verifies that these are well-defined. We now show by induction on the height of the formula the following:
-
-	\emph{Claim.} For any formula $\psi$ in which the free variable $b$ doesn't occur, world $w_k\in W_0$ and variable assignments $v_c, v_i$ such that $v_c(a) = [v_k(a), w_k]$ for all free variables $a$ occurring in $\psi$ it holds that \[M, I, v_c[w_k/b]\models \psi^b\text{ if and only if $w_k, v_i\models \psi$.}\]
-
-	For the base case, i.e. if $\psi = P(t_1,\dots, t_n)$ for some $n$-ary predicate symbol $P$, the claim follows directly from the definitions.
-
-	For the induction step consider e.g. the case where $\psi$ is of the form $\forall x \psi'$.
-	Suppose first that $w_k, v_i\not\models \forall x\psi'$, i.e.
-there exists $w_l\geq w_k$ and $m_l\in w_l$ such that $w_l, v_i[m_l/a]\not\models \psi'[a/x]$.
-Then by the induction hypothesis $M, I, v_c[[v_i(a), w_l],w_l/c]\not\models \psi'[a/x]^c$.
-Furthermore by definition $M, I, v_c[w_l/c, w_k/b]\models b\preceq c$ and $M, I, v_c[v_c[[v_i(a), w_l],w_l/c]\models E(a, c)$.
-Hence $M, I, v_c[w_k/b]\not\models \psi^b$.
-Suppose on the other hand that $w_k, v_i\not\models\forall x\psi'$, i.e.
-for all $w_l\geq w_k$ and $m_l\in w_l$ it holds that $w_l, v_i[m_l/a]\models \psi'[a/x]$.
-Then by the induction hypothesis $M, I, v_c[[m_l, w_l]/A,w_l/c]\models \psi'[a/x]^c$ for all $w_l\geq w_k$ and $m_l\in w_l$.
-It can be shown that for every other assignment of $c$ and $a$ either $b\preceq c$ or $E(a, c)$ is violated.
-Hence $M, I, v_c[w_k/b]\models \psi^b$.
-For other forms of $\psi$ one proceeds in similar fashion.
-This completes the proof of the claim.
-
-	Recall that $w_0, v\not\models\varphi$.
-From the claim it follows that \[M, I, v'[[v(a_0), w_0]/a_0\dots[v(a_n), w_0]/a_0, w_0/b]\not\models \varphi^b\]where $a_0\dots a_n$ are the free variables occurring in $\varphi$ and $v'$ is any variable assignment.
-Furthermore a straightforward check of definitions yields \[M, I, v'[[v(a_0), w_0]/a_0\dots[v(a_n), w_0]/a_0, w_0/b]\models K(\varphi)\wedge \text{World}(b)\wedge\vec E(\vec a, b).\]Therefore $M, I\not\models \varphi^C$ and we are done.
-
-\end{proof}
+A proof can be found in the appendix~\ref*{proof:fo-simplification}.
 
@@
 
+\begin{lemma}\label{proof:fo-simplification}
+	$\varphi$ is intuitionistically valid if and only if $\varphi^C$ is classically valid.
+\end{lemma}
+
+\begin{proof}
+	Since we are mixing worlds and proper intuitionistic domain elements in a single classical domain there are some details to be taken care of. We proceed by translation of counter-models.
+
+	Suppose first we have a counter-model $\mathcal M = (M, I)$ to $\varphi^C$.
+As a Kripke frame $(W, \leq)$ take $ W = \{m\in M\:|\:\text{ World}^I(m)\}$ and let $\leq$ be $\preceq^I$ restricted to $W$.
+Then let $M_u = \{m\in M\:|\: E(m, u)\}$ and let $f^{I_u}$ be $f^I$ restricted to $M_u$ and $A^{I_u}(\vec x): \Leftrightarrow A(\vec x, u)$.
+It is then a straightforward check of definitions that this defines a Kripke counter-model to $\varphi$.
+
+	The other direction is a bit more involved.
+Suppose we have a Kripke counter-model to $\varphi$ with frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
+In particular, since it is a counter-model there exists $w_0\in W$ and a variable assignment $v$ with $w_0, v\not\models\varphi$.
+Let $W_0 = \{w\in W\:|\: w_0\preceq w\}$ and define an equivalence relation $\sim$ on $\{(x, u)\:|\:u\in W_0, x\in M_u\}$ via $(x, u)\sim (y, w)$ iff $x = y$ and there exists $v\in W_0$ comparable with both $u, w$ such that $x\in v$ and denote the equivalence class of $(x, u)$ with $[x, u]$.
+Let $M = W_0\cup \{[x, u]\:|\:u\in W_0, x\in M_u\}$.
+	Now have
+	\begin{itemize}
+		\item $\preceq^I = \leq\cup \{(m, m): m\in M\setminus W_0\}$ as a set, i.e. $\preceq$ corresponds to $\leq$ on $W_0$ and is trivial elsewhere.
+		\item $E^I(m, w)$ iff $w\in W_0$ and $m = [x, w]$ for some $x\in M_w$.
+		\item $f^I(m_1\dots m_n) =\begin{cases}
+			f^{I_u}(x_1\dots x_n), &\substack{\text{if there are $ u\in W_0, x_i\in M_u$}\\ \text{with $m_i= [x_i, u]$ for all $i$,}}\\
+			w_0, & \text{otherwise.}
+		\end{cases}$
+		\item ${A}^I(m_1\dots m_n, u) \Leftrightarrow\begin{cases}
+			A^{I_u}(x_1\dots x_n), &\substack{\text{if $u\in W_0$ and $\exists x_i\in M_u$}\\\text{with $m_i= [x_i, u]$ for all $i$,}}\\
+			\bot, & \text{otherwise.}
+		\end{cases}$
+	\end{itemize}
+	One easily verifies that these are well-defined. We now show by induction on the height of the formula the following:
+
+	\emph{Claim.} For any formula $\psi$ in which the free variable $b$ doesn't occur, world $w_k\in W_0$ and variable assignments $v_c, v_i$ such that $v_c(a) = [v_k(a), w_k]$ for all free variables $a$ occurring in $\psi$ it holds that \[M, I, v_c[w_k/b]\models \psi^b\text{ if and only if $w_k, v_i\models \psi$.}\]
+
+	For the base case, i.e. if $\psi = P(t_1,\dots, t_n)$ for some $n$-ary predicate symbol $P$, the claim follows directly from the definitions.
+
+	For the induction step consider e.g. the case where $\psi$ is of the form $\forall x \psi'$.
+	Suppose first that $w_k, v_i\not\models \forall x\psi'$, i.e.
+there exists $w_l\geq w_k$ and $m_l\in w_l$ such that $w_l, v_i[m_l/a]\not\models \psi'[a/x]$.
+Then by the induction hypothesis $M, I, v_c[[v_i(a), w_l],w_l/c]\not\models \psi'[a/x]^c$.
+Furthermore by definition $M, I, v_c[w_l/c, w_k/b]\models b\preceq c$ and $M, I, v_c[v_c[[v_i(a), w_l],w_l/c]\models E(a, c)$.
+Hence $M, I, v_c[w_k/b]\not\models \psi^b$.
+Suppose on the other hand that $w_k, v_i\not\models\forall x\psi'$, i.e.
+for all $w_l\geq w_k$ and $m_l\in w_l$ it holds that $w_l, v_i[m_l/a]\models \psi'[a/x]$.
+Then by the induction hypothesis $M, I, v_c[[m_l, w_l]/A,w_l/c]\models \psi'[a/x]^c$ for all $w_l\geq w_k$ and $m_l\in w_l$.
+It can be shown that for every other assignment of $c$ and $a$ either $b\preceq c$ or $E(a, c)$ is violated.
+Hence $M, I, v_c[w_k/b]\models \psi^b$.
+For other forms of $\psi$ one proceeds in similar fashion.
+This completes the proof of the claim.
+
+	Recall that $w_0, v\not\models\varphi$.
+From the claim it follows that \[M, I, v'[[v(a_0), w_0]/a_0\dots[v(a_n), w_0]/a_0, w_0/b]\not\models \varphi^b\]where $a_0\dots a_n$ are the free variables occurring in $\varphi$ and $v'$ is any variable assignment.
+Furthermore a straightforward check of definitions yields \[M, I, v'[[v(a_0), w_0]/a_0\dots[v(a_n), w_0]/a_0, w_0/b]\models K(\varphi)\wedge \text{World}(b)\wedge\vec E(\vec a, b).\]Therefore $M, I\not\models \varphi^C$ and we are done.
+
+\end{proof}
+
 \end{document}

### 2023-03-09 LPAR/LPAR-EIICL.tex
formatting

@@
 For predicate logic we assume an ambient signature $\Sigma$, i.e., a set of function and predicate symbols with their respective arities, and a set of variables.
-As usual we can inductively define terms and formuas.
 A term is either a variable or, given an $n$-ary function symbol $f$ and terms $t_1, \ldots, t_n$, $f(t_1, \ldots, t_n)$.
@@
 Free and bound variables are defined as usual and we assume that they comprise distinct sets, such that bound variables are not affected by substitutions of free variables.
-Generally we will denote free variables with letters at the beginning of the alphabet and bound variables with letters at the end of the alphabet.
+Generally letters at the beginning of the alphabet will denote free variables and letters at the end of the alphabet bound variables.
 
@@
 In this process we introduce fresh variables and add additional function symbols to the signature.
-A fresh variable is any variable that does not occur in any of the considered formulas.
-Whenever we add a function symbol we implicitly extend the signature by some not previously contained symbol.
+A fresh variable is any variable that does not occur in any considered formula.
+Whenever we add a function symbol we implicitly extend the signature by a not previously contained symbol.
 
@@
 	\end{itemize}
-	Let $$\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$$ and $$\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
+	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
 \end{definition}
@@
 
-A similar result holds in the first order case.
-
 \begin{lemma}\label{lemma:first-order-normal-form}
@@
 
-Since we could not find a proof of this, one is given in the appendix.
+Since we could not find a proof of this, one is given in the appendix~\ref{proof:first-order-normal-form}.
 
@@
 
-The general idea is to use the Kripke semantics for intuitionistic logic, which are expressed in classical logic, to perform an embedding of formulas.
-That is, for each formula $\varphi$ we give a formula $\varphi^\#$ that is classically valid if and only if $\varphi$ is intuitionistically valid, establishing a dual to the double negation translation.
+The idea is to use the Kripke semantics for intuitionistic logic, expressed in classical logic, to perform an embedding of formulas.
+That is, for each $\varphi$ we give a $\varphi^\#$ that is classically valid iff $\varphi$ is intuitionistically valid, establishing a dual to the double negation translation.
 
@@
 	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^\nu$ is
-	$$\forall w\left(\nu\preceq w\to\bigwedge_i A_i(w)\to\bigvee_jB_j(w)\right).$$
+	$\forall w\left(\nu\preceq w\to\bigwedge_i A_i(w)\to\bigvee_jB_j(w)\right).$
 	\item For implication clauses $((A\to B)\to C)^\nu$ is
@@
 with $f_\psi$ being a new function symbol.
-
 We now argue that if a counter-model for $\varphi^{CH}$ exists then a small counter-model exists. This is ultimately what enables our translations.
@@
 \begin{corollary}
-	If a formula in normal form as in Lemma~\ref{lemma:propositional-normal-form} is intuitionistically invalid then there exists a Kripke counter-model which is a rooted tree with height and degree bounded by $|\mathcal X|$.
+	If a formula in normal form as in Lemma~\ref{lemma:propositional-normal-form} is intuitionistically invalid then there exists a Kripke counter-model that is a rooted tree of height and degree bounded by $|\mathcal X|$.
 \end{corollary}
@@
 To encode that some $n$-ary predicate $A$ holds at a world $u$ we extend each $n$-ary predicate symbol to a $n+1$-ary predicate symbol of the same name, interpreting $A(\vec x, u)$ as "$A(\vec x)$ holds at u".
-Furthermore we write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$.
+Write $\vec E(\vec x, u)$ for $\bigwedge\{E(x_i, u)\:|\:x_i\in \vec x\}$.
 With this we obtain the following encoding of Kripke Semantics:
@@
 	\end{align*}
-	with e.g.:
+	with e.g.:\vspace*{-.76cm}
 	\begin{align*}
@@
 	Then define
-	$$\varphi^C = K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b$$where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
+	$\varphi^C = K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b$ where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
 \end{definition}
@@
 		\item We interpret $\preceq$ as the prefix-order.
-		\item We set $$f_\psi^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)(\vec m, \psi).$$
+		\item We set $f_\psi^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)(\vec m, \psi).$
 		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, f_{\psi_n}(\vec m_n, \dots(f_{\psi_1}(\vec m_1, s^I))\dots)).$$
@@
  	for each flat clause
- 	$$
+ 	$
  		\forall \vec x\left(\bigwedge_i A_i(\vec a_i)\to\bigvee_j B_j(\vec b_j)\right)\in\mathcal R,
- 	$$
+ 	$
  	$\mathcal X^\#$ by including
@@
  	for each implication clause
-	$$
+	$
 		\psi = \forall \vec x\left(\left(A(\vec a)\to B(\vec b)\right)\to C(\vec c)\right)\in\mathcal X,
-	$$
+	$
 	and $\mathcal Q^\#$ by including
@@
 	for each implication clause
-	$$
+	$
 		\psi = \forall \vec x\left(\forall y\left(A(\vec a)\right)\to B(\vec b)\right)\in\mathcal X.
-	$$
+	$
 	Then $\varphi$ is intuitionistically valid if and only if
@@
 That is, we may assume that $f_\psi(\vec t, w) = w$, if $\exists u(w = f_\psi(\vec t, u))$ for all $\psi\in\mathcal X\cup\mathcal Q$ (i.e.
-that $f_\psi$ parametrized by $\vec t$ is idempotent), and $f_\psi(\vec t, w) = w$, if $\exists u(w \geq f_\psi(\vec t, u))$ for all $\psi\in\mathcal X$ (which is even stronger than idempotence).
+that $f_\psi$ parametrized by $\vec t$ is idempotent), and $f_\psi(\vec t, w) = w$, if $\exists u(w \geq f_\psi(\vec t, u))$ for all $\psi\in\mathcal X$.
 
@@
 
-\begin{wrapfigure}[4]{r}{0.4\textwidth}
-	\vspace*{-2cm}
+\begin{wrapfigure}[5]{r}{0.4\textwidth}
+	\vspace*{-1.6cm}
 	\begin{tabular}{l|c|c}
@@
 
-\begin{wrapfigure}[16]{r}{0.4\textwidth}
-	\vspace{6cm}
+\begin{wrapfigure}[14]{r}{0.4\textwidth}
+	\vspace{6.4cm}
 	\hspace*{1.5cm}
@@
 	\end{tikzpicture}
-	\vspace*{-7.5cm}
+	\vspace*{-7.6cm}
 	\caption{Instances proven by combinations of Embedding (red), Embedding (idem.) (green), ileanCoP or nanoCoP (blue)}
@@
 
+\vspace*{-.5cm}
 Two versions of the translation were tested, one with and one without the additional axioms encoding idempotence. The results are stated in Figure \ref{fig:a}. 
@@
 		0&1-2	&119&29.8\%&130&32.5\%		&108&27.0\%		&96&24.0\%	&400\\
-		0&3-5	&16&15.2\%	&17&16.2\%		&50&47.6\%		&40&28.1\%	&105\\
-		0&6+	&27&38.6\%	&11&15.7\%		&20&28.6\%		&11&15.7\%	&70\\
+		0&3+	&33&18.9\%	&38&21.7\%		&70&40.0\%		&51&29.1\%	&175\\
 		1-2&0-2	&84&29.3\%	&81&28.2\%		&227&79.1\%		&221&77.0\%	&287\\
-		1-2&3-5	&80&25.6\%	&77&24.7\%		&141&45.2\%		&122&39.1\%	&312\\
-		1-2&6+	&22&21.8\%	&21&20.8\%		&39&38.6\%		&37&36.6\%	&101\\
-		3-5&0-5	&38&17.0\%	&40&17.9\%		&125&55.8\%		&110&49.1\%	&224\\
-		3-5&6+  &20&8.4\%	&18&7.6\%		&64&27.0\%		&86&36.3\%	&237\\
-		6+&0+	&8&1.1\%	&5&0.7\%		&73&9.7\%		&74&9.8\%	&755
+		1-2&3+	&102&24.7\%	&98&27.3\%		&180&43.6\%		&157&39.1\%	&413\\
+		3+&0+	&66&5.4\%	&63&5.2\%		&262&20.7\%		&270&22.2\%	&1216\\
 	\end{tabular}
@@
 However, idempotence is a quite powerful concept from a model-theoretic perspective.
-We suspect that a more clever encoding of this property could improve our results.
+We suspect that a more clever encoding of this property could improve results.
 There are other restrictions on models we could consider, based on structural properties of the formula, e.g.
@@
 
-We have presented embeddings of intuitionistic into classical logic for the propositional and the predicate case.
-Using Kripke semantics and model reduction we have established that the complexity overhead of intuitionistic logic over classical logic can be parametrized by $|\mathcal X|$ and $|\mathcal Q|$.
+We have presented embeddings of intuitionistic into classical.
+Using Kripke semantics and model reduction we have established a complexity overhead of intuitionistic logic over classical logic parametrized by $|\mathcal X|$ and $|\mathcal Q|$.
 This is directly reflected in the exponential blow-up parametrized by $|\mathcal X|$ in the propositional case and more indirectly in our benchmarks in the predicate case.
@@
 However, the current complexity considerations are direct consequences of our straightforward constructions.
-We plan to establish better bounds in future work by utilizing structural properties of the input formula, in particular relations between different atoms.
+We plan to establish better bounds in future work by utilizing structural properties of the input formula.
 We hope that this will allow the construction of smaller counter-examples $\mathcal M_T$ and the use of redundant axioms to limit search in a more clever way.

### 2023-03-09 LPAR/LPAR-EIICL.tex
Fixed small errors

@@
 	\begin{enumerate}
-		\item $\psi$ is fulfilled at $f_{[\psi]}^I(u)$ for all $u\in M, \psi\in\mathcal X$.
+		\item $\psi$ is fulfilled at $f_{[\psi]}^I(m)$ for all $m\in M, \psi\in\mathcal X$.
 		\item If $\psi$ is fulfilled at some $m\in M$ then $\psi$ is fulfilled at all $n\succeq^I m$.
@@
 
-Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulas $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = \text{World}(s)\to K(\varphi)\to\left(\bigwedge R^{bS}\wedge \bigwedge X^{bS}\wedge\bigwedge Q^{bS}\right)\to P(s)$$
-where $s$ is a new constant symbol, each formula in $\mathcal R^{bS}$ is of the form
+Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulas $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge R^{bS}\wedge \bigwedge X^{bS}\wedge\bigwedge Q^{bS}\right)\to P(s)$$
+where $\vec r, s$ are new constant symbols, each formula in $\mathcal R^{bS}$ is of the form
 $$\forall w \forall \vec x(s\preceq w\to\vec E(x, w)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)),$$
@@
 \begin{definition}
-	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
+	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
+W.l.o.g.
+suppose $\{m\in M\:|\:\text{World}(m)\}$ has a minimal element, say $w_0$.
+Define $\mathcal M_T = (M_T, I_T)$ as follows:
 	\begin{itemize}
-		\item $M_T$ are the sequences on $\{(\vec m, \psi)\:|\:\psi\in \mathcal F, \vec m\in M^n\}$.
-		\item $r^{I_T} = s^{I_T} = \epsilon$.
-		\item We interpret $\preceq$ as the prefix-order.
-		\item We set $f_\psi^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)(\vec m, \psi).$
-		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, f_{\psi_n}(\vec m_n, \dots(f_{\psi_1}(\vec m_1, s^I))\dots)).$$
+		\item $M_T$ is the union of $M$ and sequences $(\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)$ such that $\vec E^I(\vec m_1, w_0)$ and $\vec E^I(\vec m_i, f_{\psi_i}(\vec m_i, f_{\psi_{i-1}}^I(\vec m_{i-1}, \dots f_{\psi_1}^I(\vec m_1, w_0)\dots)))$ for $i > 0$ hold.
+		\item $s^{I_T} = w_0$, $\vec r^{I_T} = \vec r^I$.
+		\item We interpret $\preceq$ trivially on $M$ and as the prefix-order elsewhere.
+		\item Have $E^{I_T}(m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n))$ iff $E^I(m, f_{\psi_n}^I(\vec m_n, \dots f_{\psi_1}^I(\vec m_1, w_0)\dots))$.
+		\item We set $f_\psi^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)(\vec m, \psi)$ and arbitrary else.
+		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, f_{\psi_n}^I(\vec m_n, \dots(f_{\psi_1}^I(\vec m_1, s^I))\dots))$$ and arbitrary otherwise, for function symbols $f^{I_T}(\vec m) = f^I(\vec m)$ and arbitrary else.
 	\end{itemize}
@@
 	$$
-		\varphi^\# := \exists xE(x, s)\to K(\varphi)\to \left(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge\mathcal Q^\#\right)\to P^\#(s)
+		\varphi^\# := K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge\mathcal Q^\#\right)\to P^\#(s)
 	$$
-	is classically valid, where $s$ is a new constant symbol and
+	is classically valid, where $s, \vec r$ are new constants introduced by Herbrandization of free variables,
 	\begin{align*}

### 2023-03-09 LPAR/LPAR-EIICL.tex
comments

--- 2023-03-09 LPAR/LPAR-EIICL.tex
@@
 
-In this section we fix notation and recall the semantics for classical and intuitionistic logic. Furthermore we recapitulate some results from~\cite{otten2005clausal}.
+In this section we fix notation and recall the semantics for classical and intuitionistic logic. 
+Furthermore, we recapitulate some results from~\cite{otten2005clausal}.
 
@@
 		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$.
-		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$.
+		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$.\todo{Hier und im Rest vom paper bin nicht so glücklich über diese Substitution von durch a..}
 	\end{itemize}
@@
 A fresh variable is any variable that does not occur in any considered formula.
-Whenever we add a function symbol we implicitly extend the signature by a not previously contained symbol.
+Whenever we add a function symbol we implicitly extend the signature by a fresh symbol.
 
@@
 		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
-		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.
+		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$
+ where $a$ is a new free variable.
 	\end{itemize}
@@
 
-Finally we note the results from~\cite{otten2005clausal} which are of importance to us.
+Finally we recall a result from~\cite{otten2005clausal}, on which we build in this paper.
 
@@
 The size of $\mathcal R$ and $\mathcal X$ is linear in the size of $\varphi$.
-In general they will contain additional propositional variables.
+In general these formulae will contain additional propositional variables.
 \end{lemma}
@@
 This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$.
-In general this formula will contain new function and predicate symbols.
+In general these formulae will contain new function and predicate symbols.
 \end{lemma}
 
-Since we could not find a proof of this, one is given in the appendix~\ref{proof:first-order-normal-form}.
+Since we could not find a proof for the first-order case, we provide one in the appendix~\ref{proof:first-order-normal-form}.
 
@@
 We shall now give an overview of our main results and highlight the key arguments. 
-We start with the propositional case.
-The important insights are the same as in the first-order case, while it is technically simpler and less cluttered.
+We start with the propositional case, as the important insights are the same as in the first-order case, while it is technically simpler and less cluttered.
 
-The idea is to use the Kripke semantics for intuitionistic logic, expressed in classical logic, to perform an embedding of formulas.
+The first step is to express the Kripke semantics of intuitionistic logic in classical first-order logic.
 That is, for each $\varphi$ we give a $\varphi^\#$ that is classically valid iff $\varphi$ is intuitionistically valid, establishing a dual to the double negation translation.
 
-The key property that allows us to reduce our encoding to propositional logic, which a priori targets first order logic, is a reduction of counter-models.
+Then, the key to reducing the first-order logic encoding to a propositional logic encoding, is a small model property that supports the reduction of counter-models.
 In particular, given $$\varphi = \left(\bigwedge\mathcal R\wedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form} we establish:
@@
 
-This allows us to embed intuitionistic propositional logic into classical propositional logic by replacing quantification over worlds by their enumeration.
+This allows us to embed intuitionistic propositional logic into classical propositional logic by replacing the quantification over all possible worlds by an explicit enumeration:
 
@@
 
-The exact details can be found in Theorem~\ref{thm:prop-translation}.
-Instead of expressing correctness at each world of the model, we can express that the worlds along some path in the tree are correct.
-The branching of paths can be expressed through quantification in a quantified boolean formula (QBF) which gives us the following:
+The details can be found in Section~\ref{sec:ipc-to-cpc}.
+Building on the same idea, we can also give a quantified boolean formula (QBF) of polynomial size, where the branches of the tree-model can be enumerated implicitly using (alternating) quantifiers.
+We obtain the following result:
 
 \begin{lemma}
-	For each propositional formula $\varphi$ there is effectively a QBF $\varphi^Q$ with $\mathcal |\varphi^Q|\in\mathcal O(|\mathcal R|\cdot|\mathcal X| + |\mathcal X|^3)$ such that $\varphi$ is intuitionistically invalid if and only if $\varphi^Q$ is a satisfiable QBF.
+	For each propositional formula $\varphi$ there is effectively a QBF $\varphi^Q$ \todo{Ich würde die Formeln $\varphi^Q$ hier und in Section 6 aus Symmetriegründen auch $\varphi^\#$ nennen.} with $\mathcal |\varphi^Q|\in\mathcal O(|\mathcal R|\cdot|\mathcal X| + |\mathcal X|^3)$ such that $\varphi$ is intuitionistically invalid if and only if $\varphi^Q$ is a satisfiable QBF.
 \end{lemma}
@@
 This also means that our language necessarily becomes more complex.
-Still, a number of reductions is possible.
+Still, a number of reductions are possible.
 
@@
 \begin{definition}
-	Let $\varphi$ be a propositional formula. Define $\varphi^{\nu}$ inductively:
+	Let $\varphi$ be a propositional formula.\todo{genau genommen müsstest du noch sagen, dass $\nu$ eine freie Variable ist.
+Im letzten Fall der Definition benützen wir dann aber w gleichzeitig als freie und gebundene Variable}
+ Define $\varphi^{\nu}$ inductively:
 	\vspace*{-.1cm}
@@
 \section{IPC to CPC}
+\label{sec:ipc-to-cpc}
 
@@
 For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional Variable $A^\lambda$.
-Define $R^\#$ as the union over all sets
+Define $\mathcal R^\#$ as the union over all sets
 	$$
@@
 	$$
-	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $P^\#$ as the union over all sets
+	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $P^\#$ \todo{Hier ist nur P statt mathcal P, was aus Symmetriegründen vielleicht hilfreich wäre, aber P wird auch schon für die finale Präposition verwendet, also wäre ein anderer Buchstabe wohl besser} as the union over all sets
 	$$
@@
 It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}.
-The translation from the last section will serve as starting point for developing the translation in this section.
-For lack of space, we do not define the semantics of QBF in this paper and refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
+\todo{Wollen wir noch etwas sagen wie ``However the previous PSPACE result does not support the transfer of countermodels that we develop in this section''?}
+For lack of space, we do not give the semantics of QBF in this paper and instead refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
+The translation from the last section will serve as starting point for this section.
 
 Suppose we are given a formula $\varphi$ in normal form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}.
-We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model.
-The idea how to get a polynomially-sized formula is: Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
-The universally quantified variables $X_\psi$ in the next definition handle the branching, i.e.
-express which path is considered.
+We now given an intuition on how to obtain a QBF of polynomial size that expresses that $\varphi$ has an intuitionistic counter-model:
+Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
+The universally quantified variables $X_\psi$ in the next definition handle the branching, i.e. express which path is considered.
 
 \begin{definition}
-	For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$.
+For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ (representing the distance of a node from the root on a path) consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$.
 Let $\vec A^n$ range over all propositional variables $A^n$ and $\vec X^n$ range over all $X_\psi^n$. 
@@
 		\item {\normalfont Valid}$(n)$ which encodes that $\vec X^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
-if we interpret the $\vec X^i$ as bit-vectors, then $\vec X^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec X^i\:|\:i < n\}$.
+if we interpret the $\vec X^i$ as bit-vectors, then $\vec X^n$ has exactly one bit set to $1$, indicating the formula $\psi$ at the $n$-th position of the sequence, and it is at a position that is $0$ in all $\vec X^i$ with $i < n$ (i.e., it did not occur in the sequence previously).
 		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
@@
 	Let $\varphi$ be a predicate formula. Define $\varphi^{\nu}$ inductively:
+\todo{genau genommen müsstest du noch sagen, dass $\nu$ eine freie Variable ist. Im letzten Fall der Definition benützen wir dann aber w gleichzeitig als freie und gebundene Variable}
 	\begin{itemize}
@@
 This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$.
-In general this formula will contain new function symbols and predicate symbols.
+In general these formulae will contain new function symbols and predicate symbols.
 \end{lemma}

### 2023-03-09 LPAR/LPAR-EIICL.tex
comments appendix

@@
 	1. If $C^I(m)$ is true then we are done due to persistency. Otherwise $C^I(m)$ is false.
-	Because $(A^I(f_{[\psi]}^I(m))\to B^I(f_{[\psi]}^I(m)))\to C^I(u)$ holds, then $A^I(f_{[\psi]}^I(u))\to B^I(f_{[\psi]}^I(u))$ must be false.
+	Because $(A^I(f_{[\psi]}^I(m))\to B^I(f_{[\psi]}^I(m)))\to C^I(m)$ holds, then $A^I(f_{[\psi]}^I(m))\to B^I(f_{[\psi]}^I(m))$ must be false.
 
@@
 \begin{definition}
-For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ (representing the distance of a node from the root on a path) consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$.
+For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ (representing the $n$-th position of a node on a path) consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$.
 Let $\vec A^n$ range over all propositional variables $A^n$ and $\vec X^n$ range over all $X_\psi^n$.
@@
 		\text{\normalfont Persistent}(n) = & \bigwedge\left\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\right\}\\
-		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{\bigwedge A^n_i\to\bigvee B^n_j\:|\:\bigwedge A_i\to\bigvee B_j\mathcal R\right\}\\
+		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{\bigwedge A^n_i\to\bigvee B^n_j\:|\:\bigwedge A_i\to\bigvee B_j \in \mathcal R\right\}\\
 		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\left\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\right\}
@@
 Then, $(\bigwedge\mathcal R\wedge\bigwedge\mathcal X)\to P$ has an intuitionistic counter-model.
-By the previous section there exists a classical counter-model $\mathcal M$ for $(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#)\to P^\epsilon$.
+By the previous section\todo{unklar was mit "previous section" gemeint ist, und was für ein model du genau meinst.} there exists a classical counter-model $\mathcal M$ for $(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#)\to P^\epsilon$.
 We now define a QBF model iteratively.
@@
 
-	The other direction is a bit more involved.
-Suppose we have a Kripke counter-model to $\varphi$ with frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$.
+	The other direction is a bit more involved. Suppose we have a Kripke counter-model to $\varphi$ with frame $(W, \preceq)$ and family of $\Sigma$-structures $(M_w, I_w)_{w\in W}$. 
 In particular, since it is a counter-model there exists $w_0\in W$ and a variable assignment $v$ with $w_0, v\not\models\varphi$.
@@
 
-	\emph{Claim.} For any formula $\psi$ in which the free variable $b$ doesn't occur, world $w_k\in W_0$ and variable assignments $v_c, v_i$ such that $v_c(a) = [v_k(a), w_k]$ for all free variables $a$ occurring in $\psi$ it holds that \[M, I, v_c[w_k/b]\models \psi^b\text{ if and only if $w_k, v_i\models \psi$.}\]
+	\emph{Claim.} For any formula $\psi$ in which the free variable $b$ doesn't occur, world $w_k\in W_0$ and variable assignments $v_c, v_i$ such that $v_c(a) = [v_k(a), w_k]$ \todo{das sollte $v_i$ statt $v_k$ sein, oder} for all free variables $a$ occurring in $\psi$ it holds that
+\todo{Hier und weiter und sind immer M,I genannt, aber eigentlich haben wir Notation für $\mathcal M,v \models$ definiert}
+\[M, I, v_c[w_k/b]\models \psi^b\text{ if and only if $w_k, v_i\models \psi$.}\]
 
@@
 there exists $w_l\geq w_k$ and $m_l\in w_l$ such that $w_l, v_i[m_l/a]\not\models \psi'[a/x]$.
-Then by the induction hypothesis $M, I, v_c[[v_i(a), w_l],w_l/c]\not\models \psi'[a/x]^c$.
-Furthermore by definition $M, I, v_c[w_l/c, w_k/b]\models b\preceq c$ and $M, I, v_c[v_c[[v_i(a), w_l],w_l/c]\models E(a, c)$.
+Then by the induction hypothesis $M, I, v_c[[v_i(a), w_l],w_l/c]\not\models \psi'[a/x]^c$\todo{hier fehlt die Ersetzung für $[v_i(a), w_l]$.
+Fehlt außerdem nicht irgendein Bezug zu $m_l$?}.
+Furthermore by definition $M, I, v_c[w_l/c, w_k/b]\models b\preceq c$ and $M, I, v_c[v_c[[v_i(a), w_l],w_l/c]\models E(a, c)$\todo{hier fehlt die Ersetzung für $[v_i(a), w_l]$.
+Außerdem ist hier wirklich $v_c[v_c$ gemeint?}.
 Hence $M, I, v_c[w_k/b]\not\models \psi^b$.
@@
 for all $w_l\geq w_k$ and $m_l\in w_l$ it holds that $w_l, v_i[m_l/a]\models \psi'[a/x]$.
-Then by the induction hypothesis $M, I, v_c[[m_l, w_l]/A,w_l/c]\models \psi'[a/x]^c$ for all $w_l\geq w_k$ and $m_l\in w_l$.
+Then by the induction hypothesis $M, I, v_c[[m_l, w_l]/A,w_l/c]\models \psi'[a/x]^c$\todo{Hier wohl a statt A} for all $w_l\geq w_k$ and $m_l\in w_l$.
 It can be shown that for every other assignment of $c$ and $a$ either $b\preceq c$ or $E(a, c)$ is violated.
@@
 Furthermore a straightforward check of definitions yields \[M, I, v'[[v(a_0), w_0]/a_0\dots[v(a_n), w_0]/a_0, w_0/b]\models K(\varphi)\wedge \text{World}(b)\wedge\vec E(\vec a, b).\]Therefore $M, I\not\models \varphi^C$ and we are done.
-
+\todo{Da ich hier nicht durchgekommen bin, kannst du den ganzen Beweis nochmal checken :p ?}
 \end{proof}

### 2023-03-10 LPAR/LPAR-EIICL.tex
Lemma 4.4

--- 2023-03-09 LPAR/LPAR-EIICL.tex
@@
 
-In this section we fix notation and recall the semantics for classical and intuitionistic logic. 
-Furthermore, we recapitulate some results from~\cite{otten2005clausal}.
+In this section we fix notation and recall the semantics for classical and intuitionistic logic. Furthermore we recapitulate some results from~\cite{otten2005clausal}.
 
@@
 A formula is either $\bot$, or, given a $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$, $P(t_1, \ldots, t_n)$, or, given formulas $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$, or, given a formula $\varphi$ and a variable $x$, $\exists x\varphi$ or $\forall x\varphi$.
-Free and bound variables are defined as usual and we assume that they comprise distinct sets, such that bound variables are not affected by substitutions of free variables.
-Generally letters at the beginning of the alphabet will denote free variables and letters at the end of the alphabet bound variables.
+Free and bound variables are defined as usual and there is a Barendregt's convention in place, i.e.
+we implicitly assume that they comprise distinct sets, such that bound variables are not affected by substitutions of free variables.
 
@@
 		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$.
-		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$.\todo{Hier und im Rest vom paper bin nicht so glücklich über diese Substitution von durch a..}
+		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$.
 	\end{itemize}
@@
 A fresh variable is any variable that does not occur in any considered formula.
-Whenever we add a function symbol we implicitly extend the signature by a fresh symbol.
+Whenever we add a function symbol we implicitly extend the signature by a not previously contained symbol.
 
@@
 		\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$ and $(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
-		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
-		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$
- where $a$ is a new free variable.
+		\item $(\forall x\varphi)^S_Z = \forall x(\varphi^S_{Z\cup\{x\}})$.\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
+		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.\\$(\exists x\varphi)^H_Z = \exists x(\varphi^H_{Z\cup\{x\}})$.
 	\end{itemize}
@@
 
-Finally we recall a result from~\cite{otten2005clausal}, on which we build in this paper.
+Finally we note the results from~\cite{otten2005clausal} which are of importance to us.
 
@@
 The size of $\mathcal R$ and $\mathcal X$ is linear in the size of $\varphi$.
-In general these formulae will contain additional propositional variables.
+In general they will contain additional propositional variables.
 \end{lemma}
@@
 This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$.
-In general these formulae will contain new function and predicate symbols.
+In general this formula will contain new function and predicate symbols.
 \end{lemma}
 
-Since we could not find a proof for the first-order case, we provide one in the appendix~\ref{proof:first-order-normal-form}.
+Since we could not find a proof of this, one is given in the appendix~\ref{proof:first-order-normal-form}.
 
@@
 We shall now give an overview of our main results and highlight the key arguments.
-We start with the propositional case, as the important insights are the same as in the first-order case, while it is technically simpler and less cluttered.
+We start with the propositional case.
+The important insights are the same as in the first-order case, while it is technically simpler and less cluttered.
 
-The first step is to express the Kripke semantics of intuitionistic logic in classical first-order logic.
+The idea is to use the Kripke semantics for intuitionistic logic, expressed in classical logic, to perform an embedding of formulas.
 That is, for each $\varphi$ we give a $\varphi^\#$ that is classically valid iff $\varphi$ is intuitionistically valid, establishing a dual to the double negation translation.
 
-Then, the key to reducing the first-order logic encoding to a propositional logic encoding, is a small model property that supports the reduction of counter-models.
+The key property that allows us to reduce our encoding to propositional logic, which a priori targets first order logic, is a reduction of counter-models.
 In particular, given $$\varphi = \left(\bigwedge\mathcal R\wedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form} we establish:
@@
 
-This allows us to embed intuitionistic propositional logic into classical propositional logic by replacing the quantification over all possible worlds by an explicit enumeration:
+This allows us to embed intuitionistic propositional logic into classical propositional logic by replacing quantification over worlds by their enumeration.
 
@@
 
-The details can be found in Section~\ref{sec:ipc-to-cpc}.
-Building on the same idea, we can also give a quantified boolean formula (QBF) of polynomial size, where the branches of the tree-model can be enumerated implicitly using (alternating) quantifiers.
-We obtain the following result:
+The exact details can be found in Theorem~\ref{thm:prop-translation}.
+Instead of expressing correctness at each world of the model, we can express that the worlds along some path in the tree are correct.
+The branching of paths can be expressed through quantification in a quantified boolean formula (QBF) which gives us the following:
 
 \begin{lemma}
-	For each propositional formula $\varphi$ there is effectively a QBF $\varphi^Q$ \todo{Ich würde die Formeln $\varphi^Q$ hier und in Section 6 aus Symmetriegründen auch $\varphi^\#$ nennen.} with $\mathcal |\varphi^Q|\in\mathcal O(|\mathcal R|\cdot|\mathcal X| + |\mathcal X|^3)$ such that $\varphi$ is intuitionistically invalid if and only if $\varphi^Q$ is a satisfiable QBF.
+	For each propositional formula $\varphi$ there is effectively a QBF $\varphi^Q$ with $\mathcal |\varphi^Q|\in\mathcal O(|\mathcal R|\cdot|\mathcal X| + |\mathcal X|^3)$ such that $\varphi$ is intuitionistically invalid if and only if $\varphi^Q$ is a satisfiable QBF.
 \end{lemma}
@@
 This also means that our language necessarily becomes more complex.
-Still, a number of reductions are possible.
+Still, a number of reductions is possible.
 
@@
 \begin{definition}
-	Let $\varphi$ be a propositional formula.\todo{genau genommen müsstest du noch sagen, dass $\nu$ eine freie Variable ist.
-Im letzten Fall der Definition benützen wir dann aber w gleichzeitig als freie und gebundene Variable}
- Define $\varphi^{\nu}$ inductively:
+	Let $\varphi$ be a propositional formula. Define $\varphi^{\nu}$ inductively:
 	\vspace*{-.1cm}
@@
 	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^\nu$ is
-	$\forall w\left(\nu\preceq w\to\bigwedge_i A_i(w)\to\bigvee_jB_j(w)\right).$
+	$$\forall w\left(\nu\preceq w\to\bigwedge_i A_i(w)\to\bigvee_jB_j(w)\right).$$
 	\item For implication clauses $((A\to B)\to C)^\nu$ is
@@
 with $f_\psi$ being a new function symbol.
-We now argue that if a counter-model for $\varphi^{CH}$ exists then a small counter-model exists. This is ultimately what enables our translations.
 
 \begin{definition}
-For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{b S}$. Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
+For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{b S}$.
+Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$ such that $m\preceq f_\psi^I(m)$ for each $m\in M$, $\psi\in \mathcal X$.
 We say $(A\to B)\to C\in\mathcal X$ is \emph{fulfilled} at $m\in M$ iff $A^I(m)\to B^I(m)$ is false or $C^I(m)$ is true. For $\psi\in\mathcal X$ define $$g_\psi : M\to M, m\mapsto\begin{cases}
-		m,&\text{ if $\psi$ is fulfilled at $m$,}\\
+		m,&\text{ if $\psi$ is fulfilled at $m$ or $m\not\preceq f^I_{[\psi]}(m)$,}\\
 		f^I_{\lbrack\psi\rbrack}(m),&\text{ else.}		
@@
 \end{definition}
-\begin{lemma}~\label{thm:prop-countermodel-reduction}
-	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^{CH}$.
-	\begin{enumerate}
-		\item $\psi$ is fulfilled at $f_{[\psi]}^I(m)$ for all $m\in M, \psi\in\mathcal X$.
-		\item If $\psi$ is fulfilled at some $m\in M$ then $\psi$ is fulfilled at all $n\succeq^I m$.
-	\end{enumerate}
+
+\begin{lemma}
+	Suppose $(M, T)$ is a counter-model to $\varphi^{CH}$, then so is $(M_T, I_T)$.
 \end{lemma}
-
 \begin{proof}
-	1. If $C^I(m)$ is true then we are done due to persistency. Otherwise $C^I(m)$ is false.
-	Because $(A^I(f_{[\psi]}^I(m))\to B^I(f_{[\psi]}^I(m)))\to C^I(u)$ holds, then $A^I(f_{[\psi]}^I(u))\to B^I(f_{[\psi]}^I(u))$ must be false.
-
-	2. Let $n\succeq^I m$.
-	If $C^I(n)$ is true, we are done.
-	Otherwise $C^I(n)$ is false.
-	Due to persistency $C^I(m)$ is also false.
-	Because $\psi$ is fulfilled at $m$, $A^I(m)\to B^I(m)$ must be false, i.e. $A^I(m)$ is true and $B^I(m)$ is false. Then $A^I(n)$ and $A^I(f^I_{[\psi]}(n))$ are also true due to persistency.
-	Because $(A^I(f^I_{[\psi]}(n))\to B^I(f^I_{[\psi]}(n)))\to C^I(n)$ holds, we now get that  $B^I(f^I_{[\psi]}(n))$ must be false.
-	But then due to persistency so is $B^I(n)$ and we have that $\psi$ is fulfilled at $n$.
+	It follows directly from the definitions that $M_T, I_T\not\models P(s)$ and $M_T, I_T\models K(\varphi)$.
+It is easy to show that $M_T, I_T\models \psi$ for $\psi\in\mathcal R^{bS}$ from persistency.
+Let $\psi\in\mathcal X^{bS}$, we show $M_T, I_T[\psi_1\dots\psi_n/w] \models s\preceq w\to (w\preceq f_\psi(w)\to A(f_\psi(w)\to B(f_\psi(w)))\to C(w))$ for all $\psi_1\dots\psi_n\in M_T$.
+Let $\psi_1\dots\psi_n\in M_T$ and $m = g_{\psi_n}(\dots(g_{\psi_1}(s^I))$.
+If $M, I[m/w]\models C(w)$ we are done.
+Otherwise $M, I[m/w]\not\models s\preceq w$ or $M, I[m/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w)) \to B(f_\psi(w))$.
+The former is impossible since $g_\psi$ is increasing for all $\psi$.
+For the latter to hold we must have $M, I[m/w]\models w\preceq f_\psi(w)$, $M, I[m/w]\models A(f_\psi(w))$ and $M, I[m/w]\not\models B(f_\psi(w))$.
+If $\psi$ is fulfilled at $m$ we have $g_\psi(m) = m$, otherwise $g_\psi(m) = f_{[\psi]}^I(m)$.
+In any case $M, I'[\psi_1\dots\psi_n/w]\models A(f_\psi(w))$ and $M_T, I_T[\psi_1\dots\psi_n/w]\not\models B(f_\psi(w))$.
+It follows that $M_T, I_T[\psi_1\dots\psi_n/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w))$ and therefore $M_T, I_T\models \psi$.
 \end{proof}
-
-Essentially what this tells us is that for every $m$ at which $\psi$ is fulfilled, in particular for each $m$ in the image of $f_{[\psi]}$, we may add the assumption that $f_{[\psi]}(n) = n$ for all $n\succeq m$ without changing the validity of $\psi$.
-With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^{CH}$ is a straightforward check of definitions.
 
@@
 \section{IPC to CPC}
-\label{sec:ipc-to-cpc}
 
@@
 For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional Variable $A^\lambda$.
-Define $\mathcal R^\#$ as the union over all sets
+Define $R^\#$ as the union over all sets
 	$$
@@
 	$$
-	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $P^\#$ \todo{Hier ist nur P statt mathcal P, was aus Symmetriegründen vielleicht hilfreich wäre, aber P wird auch schon für die finale Präposition verwendet, also wäre ein anderer Buchstabe wohl besser} as the union over all sets
+	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $P^\#$ as the union over all sets
 	$$
@@
 It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}.
-\todo{Wollen wir noch etwas sagen wie ``However the previous PSPACE result does not support the transfer of countermodels that we develop in this section''?}
-For lack of space, we do not give the semantics of QBF in this paper and instead refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
-The translation from the last section will serve as starting point for this section.
+The translation from the last section will serve as starting point for developing the translation in this section.
+For lack of space, we do not define the semantics of QBF in this paper and refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
 
 Suppose we are given a formula $\varphi$ in normal form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}.
-We now given an intuition on how to obtain a QBF of polynomial size that expresses that $\varphi$ has an intuitionistic counter-model:
-Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
-The universally quantified variables $X_\psi$ in the next definition handle the branching, i.e. express which path is considered.
+We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model.
+The idea how to get a polynomially-sized formula is: Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
+The universally quantified variables $X_\psi$ in the next definition handle the branching, i.e.
+express which path is considered.
 
 \begin{definition}
-For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ (representing the distance of a node from the root on a path) consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$.
+	For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$.
 Let $\vec A^n$ range over all propositional variables $A^n$ and $\vec X^n$ range over all $X_\psi^n$.
@@
 		\item {\normalfont Valid}$(n)$ which encodes that $\vec X^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
-if we interpret the $\vec X^i$ as bit-vectors, then $\vec X^n$ has exactly one bit set to $1$, indicating the formula $\psi$ at the $n$-th position of the sequence, and it is at a position that is $0$ in all $\vec X^i$ with $i < n$ (i.e., it did not occur in the sequence previously).
+if we interpret the $\vec X^i$ as bit-vectors, then $\vec X^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec X^i\:|\:i < n\}$.
 		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
@@
 	Let $\varphi$ be a predicate formula. Define $\varphi^{\nu}$ inductively:
-\todo{genau genommen müsstest du noch sagen, dass $\nu$ eine freie Variable ist. Im letzten Fall der Definition benützen wir dann aber w gleichzeitig als freie und gebundene Variable}
 	\begin{itemize}
@@
 		\item $(\varphi\to \psi)^\nu = \forall w(\nu\preceq w\to\varphi^{w}\to\psi^{w})$.
-		\item $(\exists x\varphi)^\nu = \exists x(E(x,\nu)\wedge\varphi[a/x]^\nu[x/a])$ where $a$ is a new free variable.
-		\item $(\forall x\varphi)^\nu = \forall w(i\preceq w\to \forall x(E(x, w)\to \varphi[a/x]^w[x/a]))$ where $a$ is a new free variable and $w$ is a new bound variable.
+		\item $(\exists x\varphi)^\nu = \exists x(E(x,\nu)\wedge\varphi^\nu)$.
+		\item $(\forall x\varphi)^\nu = \forall w(i\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
 	\end{itemize}
@@
 
-\begin{wrapfigure}[14]{r}{0.4\textwidth}
-	\vspace{6.4cm}
+\begin{wrapfigure}[15]{r}{0.4\textwidth}
+	\vspace{6.6cm}
 	\hspace*{1.5cm}
@@
 
-\vspace*{-.5cm}
 Two versions of the translation were tested, one with and one without the additional axioms encoding idempotence. The results are stated in Figure \ref{fig:a}. 
@@
 This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$.
-In general these formulae will contain new function symbols and predicate symbols.
+In general this formula will contain new function symbols and predicate symbols.
 \end{lemma}

### 2023-03-10 LPAR/LPAR-EIICL.tex

--- 2023-03-09 LPAR/LPAR-EIICL.tex
@@
 
-In this section we fix notation and recall the semantics for classical and intuitionistic logic.
-Furthermore, we recapitulate some results from~\cite{otten2005clausal}.
+In this section we fix notation and recall the semantics for classical and intuitionistic logic. Furthermore we recapitulate some results from~\cite{otten2005clausal}.
 
@@
 		\item $(\varphi\to\psi)^S_Z = \varphi^H_Z\to \psi^S_Z$ and $(\varphi\to\psi)^H_Z = \varphi^S_Z\to\psi^H_Z$.
+		\item $(\forall x\varphi)^S_Z = \forall x(\varphi^S_{Z\cup\{x\}})$.\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
+		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.\\$(\exists x\varphi)^H_Z = \exists x(\varphi^H_{Z\cup\{x\}})$.
+=======
 		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
@@
  where $a$ is a new free variable.
+>>>>>>> c8d6f4a7fa194cc08089d36f91f2f60b706bd539
 	\end{itemize}
@@
 \end{definition}
-\begin{lemma}~\label{thm:prop-countermodel-reduction}
-	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^{CH}$.
-	\begin{enumerate}
-		\item $\psi$ is fulfilled at $f_{[\psi]}^I(m)$ for all $m\in M, \psi\in\mathcal X$.
-		\item If $\psi$ is fulfilled at some $m\in M$ then $\psi$ is fulfilled at all $n\succeq^I m$.
-	\end{enumerate}
+
+\begin{lemma}
+	Suppose $(M, T)$ is a counter-model to $\varphi^{CH}$, then so is $(M_T, I_T)$.
 \end{lemma}
-
 \begin{proof}
+	It follows directly from the definitions that $M_T, I_T\not\models P(s)$ and $M_T, I_T\models K(\varphi)$.
+It is easy to show that $M_T, I_T\models \psi$ for $\psi\in\mathcal R^{bS}$ from persistency.
+Let $\psi\in\mathcal X^{bS}$, we show $M_T, I_T[\psi_1\dots\psi_n/w] \models s\preceq w\to (w\preceq f_\psi(w)\to A(f_\psi(w)\to B(f_\psi(w)))\to C(w))$ for all $\psi_1\dots\psi_n\in M_T$.
+Let $\psi_1\dots\psi_n\in M_T$ and $m = g_{\psi_n}(\dots(g_{\psi_1}(s^I))$.
+If $M, I[m/w]\models C(w)$ we are done.
+Otherwise $M, I[m/w]\not\models s\preceq w$ or $M, I[m/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w)) \to B(f_\psi(w))$.
+The former is impossible since $g_\psi$ is increasing for all $\psi$.
+For the latter to hold we must have $M, I[m/w]\models w\preceq f_\psi(w)$, $M, I[m/w]\models A(f_\psi(w))$ and $M, I[m/w]\not\models B(f_\psi(w))$.
+If $\psi$ is fulfilled at $m$ we have $g_\psi(m) = m$, otherwise $g_\psi(m) = f_{[\psi]}^I(m)$.
+In any case $M, I'[\psi_1\dots\psi_n/w]\models A(f_\psi(w))$ and $M_T, I_T[\psi_1\dots\psi_n/w]\not\models B(f_\psi(w))$.
+It follows that $M_T, I_T[\psi_1\dots\psi_n/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w))$ and therefore $M_T, I_T\models \psi$.
+=======
 	1. If $C^I(m)$ is true then we are done due to persistency. Otherwise $C^I(m)$ is false.
@@
 	But then due to persistency so is $B^I(n)$ and we have that $\psi$ is fulfilled at $n$.
+>>>>>>> c8d6f4a7fa194cc08089d36f91f2f60b706bd539
 \end{proof}
-
-Essentially what this tells us is that for every $m$ at which $\psi$ is fulfilled, in particular for each $m$ in the image of $f_{[\psi]}$, we may add the assumption that $f_{[\psi]}(n) = n$ for all $n\succeq m$ without changing the validity of $\psi$.
-With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^{CH}$ is a straightforward check of definitions.
 
@@
 
+Two versions of the translation were tested, one with and one without the additional axioms encoding idempotence. The results are stated in Figure \ref{fig:a}. 
+In general, our approach falls short of the state of the art. 
+=======
 \vspace*{-.5cm}
@@
 In general, our approach falls short of the state of the art.
+>>>>>>> c8d6f4a7fa194cc08089d36f91f2f60b706bd539
 However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},

### 2023-03-10 LPAR/LPAR-EIICL.tex
Adressed Remarks

@@
 		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.\\$(\exists x\varphi)^H_Z = \exists x(\varphi^H_{Z\cup\{x\}})$.
-=======
-		\item $(\forall x\varphi)^S_Z = \forall x(\varphi[a/x]^S_{Z\cup\{a\}}[x/a])$ where $a$ is a new free variable.\\$(\forall x\varphi)^H_Z = \varphi[s(z_1,\dots,z_n)/x]^H_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.
-		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.\\$(\exists x\varphi)^H_Z = \exists x(\varphi[a/x]^H_{Z\cup\{a\}}[x/a])$
- where $a$ is a new free variable.
->>>>>>> c8d6f4a7fa194cc08089d36f91f2f60b706bd539
 	\end{itemize}
@@
 \begin{lemma}
-	For each propositional formula $\varphi$ there is effectively a QBF $\varphi^Q$ \todo{Ich würde die Formeln $\varphi^Q$ hier und in Section 6 aus Symmetriegründen auch $\varphi^\#$ nennen.} with $\mathcal |\varphi^Q|\in\mathcal O(|\mathcal R|\cdot|\mathcal X| + |\mathcal X|^3)$ such that $\varphi$ is intuitionistically invalid if and only if $\varphi^Q$ is a satisfiable QBF.
+	For each propositional formula $\varphi$ there is effectively a QBF $\varphi^Q$ with $\mathcal |\varphi^Q|\in\mathcal O(|\mathcal R|\cdot|\mathcal X| + |\mathcal X|^3)$ such that $\varphi$ is intuitionistically invalid if and only if $\varphi^Q$ is a satisfiable QBF.
 \end{lemma}
@@
 \begin{lemma}
-	Suppose $(M, T)$ is a counter-model to $\varphi^{CH}$, then so is $(M_T, I_T)$.
+	Suppose $\mathcal M = (M, T)$ is a counter-model to $\varphi^{CH}$, then so is $\mathcal M_T = (M_T, I_T)$.
 \end{lemma}
 \begin{proof}
-	It follows directly from the definitions that $M_T, I_T\not\models P(s)$ and $M_T, I_T\models K(\varphi)$.
-It is easy to show that $M_T, I_T\models \psi$ for $\psi\in\mathcal R^{bS}$ from persistency.
-Let $\psi\in\mathcal X^{bS}$, we show $M_T, I_T[\psi_1\dots\psi_n/w] \models s\preceq w\to (w\preceq f_\psi(w)\to A(f_\psi(w)\to B(f_\psi(w)))\to C(w))$ for all $\psi_1\dots\psi_n\in M_T$.
+	Let $v$ be a variable assignment.
+It follows directly from the definitions that $\mathcal M_T, v\not\models P(s)$ and $\mathcal M_T, v\models K(\varphi)$.
+It is easy to show that $\mathcal M_T, v\models \psi$ for $\psi\in\mathcal R^{bS}$ from persistency.
+Let $\psi\in\mathcal X^{bS}$, we show $\mathcal M_T, v[\psi_1\dots\psi_n/w] \models s\preceq w\to (w\preceq f_\psi(w)\to A(f_\psi(w)\to B(f_\psi(w)))\to C(w))$ for all $\psi_1\dots\psi_n\in M_T$.
 Let $\psi_1\dots\psi_n\in M_T$ and $m = g_{\psi_n}(\dots(g_{\psi_1}(s^I))$.
-If $M, I[m/w]\models C(w)$ we are done.
-Otherwise $M, I[m/w]\not\models s\preceq w$ or $M, I[m/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w)) \to B(f_\psi(w))$.
+If $\mathcal M, v[m/w]\models C(w)$ we are done.
+Otherwise $\mathcal M, v[m/w]\not\models s\preceq w$ or $\mathcal M, v[m/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w)) \to B(f_\psi(w))$.
 The former is impossible since $g_\psi$ is increasing for all $\psi$.
-For the latter to hold we must have $M, I[m/w]\models w\preceq f_\psi(w)$, $M, I[m/w]\models A(f_\psi(w))$ and $M, I[m/w]\not\models B(f_\psi(w))$.
+For the latter to hold we must have $\mathcal M, v[m/w]\models w\preceq f_\psi(w)$, $\mathcal M, v[m/w]\models A(f_\psi(w))$ and $\mathcal M, v[m/w]\not\models B(f_\psi(w))$.
 If $\psi$ is fulfilled at $m$ we have $g_\psi(m) = m$, otherwise $g_\psi(m) = f_{[\psi]}^I(m)$.
-In any case $M, I'[\psi_1\dots\psi_n/w]\models A(f_\psi(w))$ and $M_T, I_T[\psi_1\dots\psi_n/w]\not\models B(f_\psi(w))$.
-It follows that $M_T, I_T[\psi_1\dots\psi_n/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w))$ and therefore $M_T, I_T\models \psi$.
-=======
-	1. If $C^I(m)$ is true then we are done due to persistency. Otherwise $C^I(m)$ is false.
-	Because $(A^I(f_{[\psi]}^I(m))\to B^I(f_{[\psi]}^I(m)))\to C^I(m)$ holds, then $A^I(f_{[\psi]}^I(m))\to B^I(f_{[\psi]}^I(m))$ must be false.
-
-	2. Let $n\succeq^I m$.
-	If $C^I(n)$ is true, we are done.
-	Otherwise $C^I(n)$ is false.
-	Due to persistency $C^I(m)$ is also false.
-	Because $\psi$ is fulfilled at $m$, $A^I(m)\to B^I(m)$ must be false, i.e. $A^I(m)$ is true and $B^I(m)$ is false. Then $A^I(n)$ and $A^I(f^I_{[\psi]}(n))$ are also true due to persistency.
-	Because $(A^I(f^I_{[\psi]}(n))\to B^I(f^I_{[\psi]}(n)))\to C^I(n)$ holds, we now get that  $B^I(f^I_{[\psi]}(n))$ must be false.
-	But then due to persistency so is $B^I(n)$ and we have that $\psi$ is fulfilled at $n$.
->>>>>>> c8d6f4a7fa194cc08089d36f91f2f60b706bd539
+In any case $\mathcal M_T, v[\psi_1\dots\psi_n/w]\models A(f_\psi(w))$ and $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models B(f_\psi(w))$.
+It follows that $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w))$ and therefore $\mathcal M_T, v\models \psi$.
 \end{proof}
@@
 	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal S\right)\to P$$
-	with flat $\mathcal R$ and implication $\mathcal S$.
+	with flat $\mathcal R$ and implicational $\mathcal S$.
 Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$.
@@
 	$$
-	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $P^\#$ \todo{Hier ist nur P statt mathcal P, was aus Symmetriegründen vielleicht hilfreich wäre, aber P wird auch schon für die finale Präposition verwendet, also wäre ein anderer Buchstabe wohl besser} as the union over all sets
+	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $\mathcal P^\#$ as the union over all sets
 	$$
@@
 	$$
-		\varphi^\# := \left(\bigwedge P^\#\wedge\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\right)\to P^\epsilon
+		\varphi^\# := \left(\bigwedge \mathcal P^\#\wedge\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\right)\to P^\epsilon
 	$$
@@
 In general, our approach falls short of the state of the art. 
-=======
-\vspace*{-.5cm}
-Two versions of the translation were tested, one with and one without the additional axioms encoding idempotence. The results are stated in Figure \ref{fig:a}.
-In general, our approach falls short of the state of the art.
->>>>>>> c8d6f4a7fa194cc08089d36f91f2f60b706bd539
 However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},
@@
 Then, $(\bigwedge\mathcal R\wedge\bigwedge\mathcal X)\to P$ has an intuitionistic counter-model.
-By the previous section\todo{unklar was mit "previous section" gemeint ist, und was für ein model du genau meinst.} there exists a classical counter-model $\mathcal M$ for $(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#)\to P^\epsilon$.
+By theorem~\ref{thm:prop-translation} there exists a classical counter-model $\mathcal M$ for $(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#)\to P^\epsilon$.
 We now define a QBF model iteratively.

### 2023-03-10 LPAR/LPAR-EIICL.tex
formulas -> formulae

@@
 Utilizing a normal form for intuitionistic logic~\cite{claessen2015sat}, we establish a small model property for intuitionistic propositional logic.
-We use this property for a direct encoding of the Kripke semantics into classical propositional logic and quantified boolean formulas.
+We use this property for a direct encoding of the Kripke semantics into classical propositional logic and quantified boolean formulae.
 Next, we transfer the developed techniques to the first order case and provide an embedding of intuitionistic first-order logic into classical first-order-logic.
@@
 	\item We establish a small model property for counter-models to validity in intuitionistic propositional logic.
-	\item Based on the small model property, we provide an embedding of intuitionistic propositional logic into classical propositional logic and into quantified boolean formulas.
+	\item Based on the small model property, we provide an embedding of intuitionistic propositional logic into classical propositional logic and into quantified boolean formulae.
 	\item We discuss the reduction of counter-models in intuitionistic first-order logic and, based on this reduction, present an embedding of intuitionistic into classical first-order logic.
@@
 We finally mention a long-standing line of research that has investigated when classical validity implies intuitionistic validity.
-This most famously holds for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}, but also other conditions are known~\cite{schwichtenberg}.
+This most famously holds for $\Pi_2$-formulae in Peano Arithmetic~\cite{friedman1978classically}, but also other conditions are known~\cite{schwichtenberg}.
 In contrast, we propose a translation of an intuitionistic formula to an equi-valid classical formula.
@@
 For propositional logic we assume an ambient set of propositional variables.
-As usual a formula is either $\bot$, a propositional variable or, given formulas $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$.
+As usual a formula is either $\bot$, a propositional variable or, given formulae $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$.
 
@@
 A term is either a variable or, given an $n$-ary function symbol $f$ and terms $t_1, \ldots, t_n$, $f(t_1, \ldots, t_n)$.
-A formula is either $\bot$, or, given a $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$, $P(t_1, \ldots, t_n)$, or, given formulas $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$, or, given a formula $\varphi$ and a variable $x$, $\exists x\varphi$ or $\forall x\varphi$.
+A formula is either $\bot$, or, given a $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$, $P(t_1, \ldots, t_n)$, or, given formulae $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$, or, given a formula $\varphi$ and a variable $x$, $\exists x\varphi$ or $\forall x\varphi$.
 Free and bound variables are defined as usual and there is a Barendregt's convention in place, i.e.
@@
 \begin{definition}
-	A \emph{valuation} $v$ maps each propositional variable $A$ to a \emph{truth value} $v(A)\in\{0, 1\}$. We inductively define the model relation $\models$ between $v$ and formulas:
+	A \emph{valuation} $v$ maps each propositional variable $A$ to a \emph{truth value} $v(A)\in\{0, 1\}$. We inductively define the model relation $\models$ between $v$ and formulae:
 	\begin{itemize}
@@
 If every valuation is a model for $\varphi$ then we say $\varphi$ is \emph{valid}.
-We denote the set of valid formulas with \emph{CPC} (Classical Propositional Calculus).
+We denote the set of valid formulae with \emph{CPC} (Classical Propositional Calculus).
 \end{definition}
@@
 	A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds}, called the Kripke \emph{frame}, and an indexed family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_w(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
-	The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, where we set
+	The model relation between the $v_u$ and formulae $\varphi$ is defined as before, except in the case of implications, where we set
 	\begin{itemize}
@@
 A formula is valid if every Kripke structure is a model of it.
-We denote the set of valid formulas with \emph{IPC} (Intuitionistic Propositional Calculus).
+We denote the set of valid formulae with \emph{IPC} (Intuitionistic Propositional Calculus).
 \end{definition}
@@
 	\end{itemize}
-	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
+	We define a model relation between pairs $\mathcal M, v$ and formulae $\varphi$ as follows
 	\begin{itemize}
@@
 It is valid if $\mathcal M, v\models\varphi$ for every $\mathcal M, v$.
-We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).
+We denote the set of valid formulae with \emph{CQC} (Classical Quantified Calculus).
 \end{definition}
@@
 	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
-	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulas $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where we set
+	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulae $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where we set
 	\begin{itemize}
@@
 $\varphi$ is valid if it is satisfied by every Kripke structure.
-	We denote the set of valid formulas with \emph{IQC} (Intuitionistic Quantified Calculus).
+	We denote the set of valid formulae with \emph{IQC} (Intuitionistic Quantified Calculus).
 \end{definition}
-In addition to propositional tautologies there are now quantified formulas which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
+In addition to propositional tautologies there are now quantified formulae which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.
 
@@
 \begin{definition}\label{definition:skolemization}
-	For formulas $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to a set of free variables $Z$ by simultaneous induction:
+	For formulae $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to a set of free variables $Z$ by simultaneous induction:
 	\begin{itemize}
@@
 Next we attempt to transport these results to the first order case.
-However, there are formulas for which only counter-examples with an infinite Kripke frame exist.
+However, there are formulae for which only counter-examples with an infinite Kripke frame exist.
 Therefore, there is no hope of completely eliminating terms representing worlds from the final embedding.
@@
 		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
-		\item {\normalfont Sat}$_{\mathcal R}(n)$ encoding that the formulas in $\mathcal R$ hold at the $n$-th world of the path.
+		\item {\normalfont Sat}$_{\mathcal R}(n)$ encoding that the formulae in $\mathcal R$ hold at the $n$-th world of the path.
 		\item {\normalfont Sat}$_{\mathcal X}(n)$ encoding that the formula in $\mathcal X$ that is represented by $\vec X^{n-1}$ holds.
@@
 	for leaves, and $$\varphi^Q = \varphi_0^Q = \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
-	Example encodings of the above formulas are:
+	Example encodings of the above formulae are:
 	\begin{align*}
@@
 
-Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulas $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge R^{bS}\wedge \bigwedge X^{bS}\wedge\bigwedge Q^{bS}\right)\to P(s)$$
+Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulae $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge R^{bS}\wedge \bigwedge X^{bS}\wedge\bigwedge Q^{bS}\right)\to P(s)$$
 where $\vec r, s$ are new constant symbols, each formula in $\mathcal R^{bS}$ is of the form
@@
 
-It is not possible to completely eliminate worlds from the domain by enumerating them since there are formulas with no finite counter-examples.
+It is not possible to completely eliminate worlds from the domain by enumerating them since there are formulae with no finite counter-examples.
 An analogous procedure to the propositional case fails, however we can still retrieve additional axioms that restrict the interpretation of functions.

### 2023-03-12 LPAR/LPAR-EIICL.tex
graphics
Appendix A.3

@@
 
+Two versions of the translation were tested, one with and one without the additional axioms encoding idempotence. The results are stated in Figure \ref{fig:a}.
+
+In general, our approach falls short of the state of the art. 
+However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},
+we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$ resp. $|\mathcal X| \geq 6$.
+
+\begin{figure}
+	\centering
+	\begin{tabular}{c|c|lr|lr|lr|lr|c}
+		$|\mathcal Q|$&$|\mathcal X|$&\multicolumn{2}{c}{Embedding}&\multicolumn{2}{c}{Embedding (idem.)}&\multicolumn{2}{c}{ileanCoP}&\multicolumn{2}{c}{nanoCoP-i 2.0}&Total\\\hline
+		0&0		&157&88.2\%	&157&88.2\%		&128&71.9\%		&61&38.1\%	&178\\
+		0&1-2	&119&29.8\%&130&32.5\%		&108&27.0\%		&96&24.0\%	&400\\
+		0&3+	&33&18.9\%	&38&21.7\%		&70&40.0\%		&51&29.1\%	&175\\
+		1-2&0-2	&84&29.3\%	&81&28.2\%		&227&79.1\%		&221&77.0\%	&287\\
+		1-2&3+	&102&24.7\%	&98&27.3\%		&180&43.6\%		&157&39.1\%	&413\\
+		3+&0+	&66&5.4\%	&63&5.2\%		&262&20.7\%		&270&22.2\%	&1216\\
+	\end{tabular}
+	\caption{Solved problems grouped by $|\mathcal Q|$ and $|\mathcal X|$.}
+	\label{fig:b}
+\end{figure}
+
 \begin{wrapfigure}[15]{r}{0.4\textwidth}
@@
 
-Two versions of the translation were tested, one with and one without the additional axioms encoding idempotence. The results are stated in Figure \ref{fig:a}. 
-In general, our approach falls short of the state of the art. 
-However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},
-we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$ resp. $|\mathcal X| \geq 6$.
 Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers.
 Notably, while there is big overlap in solved instances, ileanCoP and nanoCoP also have a significant symmetric difference.
-\begin{figure}
-	\centering
-	\begin{tabular}{c|c|lr|lr|lr|lr|c}
-		$|\mathcal Q|$&$|\mathcal X|$&\multicolumn{2}{c}{Embedding}&\multicolumn{2}{c}{Embedding (idem.)}&\multicolumn{2}{c}{ileanCoP}&\multicolumn{2}{c}{nanoCoP-i 2.0}&Total\\\hline
-		0&0		&157&88.2\%	&157&88.2\%		&128&71.9\%		&61&38.1\%	&178\\
-		0&1-2	&119&29.8\%&130&32.5\%		&108&27.0\%		&96&24.0\%	&400\\
-		0&3+	&33&18.9\%	&38&21.7\%		&70&40.0\%		&51&29.1\%	&175\\
-		1-2&0-2	&84&29.3\%	&81&28.2\%		&227&79.1\%		&221&77.0\%	&287\\
-		1-2&3+	&102&24.7\%	&98&27.3\%		&180&43.6\%		&157&39.1\%	&413\\
-		3+&0+	&66&5.4\%	&63&5.2\%		&262&20.7\%		&270&22.2\%	&1216\\
-	\end{tabular}
-	\caption{Solved problems grouped by $|\mathcal Q|$ and $|\mathcal X|$.}
-	\label{fig:b}
-\end{figure}
 
@@
 	\begin{itemize}
-		\item $\preceq^I = \leq\cup \{(m, m): m\in M\setminus W_0\}$ as a set, i.e. $\preceq$ corresponds to $\leq$ on $W_0$ and is trivial elsewhere.
+		\item $\preceq^I = \leq\cup \{(m, m): m\in M\setminus W_0\}$, i.e. $\preceq$ corresponds to $\leq$ on $W_0$ and is trivial elsewhere.
 		\item $E^I(m, w)$ iff $w\in W_0$ and $m = [x, w]$ for some $x\in M_w$.
@@
 			A^{I_u}(x_1\dots x_n), &\substack{\text{if $u\in W_0$ and $\exists x_i\in M_u$}\\\text{with $m_i= [x_i, u]$ for all $i$,}}\\
-			\bot, & \text{otherwise.}
+			\bot, & \text{otherwise}
 		\end{cases}$
 	\end{itemize}
-	One easily verifies that these are well-defined. We now show by induction on the height of the formula the following:
+	and $\mathcal M = (M, I)$. One easily verifies that these are well-defined. We now show by induction on the height of the formula the following:
 
-	\emph{Claim.} For any formula $\psi$ in which the free variable $b$ doesn't occur, world $w_k\in W_0$ and variable assignments $v_c, v_i$ such that $v_c(a) = [v_k(a), w_k]$ \todo{das sollte $v_i$ statt $v_k$ sein, oder} for all free variables $a$ occurring in $\psi$ it holds that
-\todo{Hier und weiter und sind immer M,I genannt, aber eigentlich haben wir Notation für $\mathcal M,v \models$ definiert}
-\[M, I, v_c[w_k/b]\models \psi^b\text{ if and only if $w_k, v_i\models \psi$.}\]
+	\emph{Claim.} For any formula $\psi$ in which the free variable $b$ doesn't occur, world $w_k\in W_0$ and variable assignments $v_c, v_i$ such that $v_c(a) = [v_i(a), w_k]$ for all free variables $a$ occurring in $\psi$ it holds that
+\[\mathcal M, v_c[w_k/b]\models \psi^b\text{ if and only if $w_k, v_i\models \psi$.}\]
 
-	For the base case, i.e. if $\psi = P(t_1,\dots, t_n)$ for some $n$-ary predicate symbol $P$, the claim follows directly from the definitions.
+	For the base case, i.e. if $\psi = \bot$ or $\psi = P(t_1,\dots, t_n)$ for some $n$-ary predicate symbol $P$, the claim follows directly from the definitions.
 
-	For the induction step consider e.g. the case where $\psi$ is of the form $\forall x \psi'$.
+	For the induction step we must consider the cases where $\psi$ is of the form $\psi_0\wedge \psi$, $\psi_0\vee \psi$, $\psi_0\to \psi$, $\exists x\psi'$, $\forall x \psi'$, there former two are straightforward and are left out.
+
+	\emph{Case 1:} $\psi = \psi_0\to \psi_1$.\\
+	Suppose first that $w_k, v_i\not\models \psi_0\to\psi_1$.
+Then there exists $w_l\geq w_k$ such that $w_l, v_i\models \psi_0$ and $w_l, v_i\not\models\psi_1$.
+Then by induction hypothesis $\mathcal M, v_c[w_l/b]\models \psi_0^b$ and $\mathcal M, v_c[w_l/b]\models \psi_1^b$, i.e.
+$\mathcal M, v_c[w_l/b]\models \psi$.
+Suppose on the other hand that $w_l, v_i\models \psi_0\to\psi_1$.
+Then for all $w_k\geq w_l$ we have $w_k\models \psi_0\to \psi_1$, i.e.
+$w_k, v_i\not\models\psi_0$ or $w_k, v_i\models\psi_1$.
+Then by induction hypothesis $\mathcal M, v_c[w_k/b]\not\models\psi_0$ or $\mathcal M, v_c[w_k/b]\models\psi_1$, i.e.
+$\mathcal M, v_c[w_k/b]\models \psi_0\to\psi_1$.
+
+	\emph{Case 2:} $\psi = \exists x\psi'$.\\
+	Suppose first that $w_k, v_i\not\models\exists x\psi'$, i.e.
+for all $m_k\in M_{w_k}$ we have $w_k, v_i[m_k/x]\not\models \psi'$ and by induction hypothesis $\mathcal M, v_i[[m_k, w_k]/x, w_k/b]\not\models \psi'^b$.
+Every other substitution for $x$ violates $E(x, b)$.
+Therefore $\mathcal M, v_i[w_k/b]\models \forall x(\neg E(x, b)\vee \neg \psi')$, i.e.
+$\mathcal M, v_i[w_k/b]\not\models \psi^b$.
+Suppose on the other hand that $w_k, v_i\models \exists x\psi'$ then there exists $m_k\in M_{w_k}$ such that $w_k, v_i[m_k/x]\models\psi'$.
+Then by induction hypothesis $\mathcal M, v_i[[m_k, w_k]/x, w_k/b]\models \psi'^b$.
+By definition $\mathcal M, v_i[[m_k, w_k]/x, w_k/b]\models E(x, b)$ holds.
+Therefore $\mathcal M, v_i[w_k/b]\models \exists x(E(x, b)\wedge \psi')$, i.e.
+$\mathcal M, v_i[w_k/b]\models \psi^b$.
+
+	\emph{Case 3:} $\psi = \forall x\psi'$.\\
 	Suppose first that $w_k, v_i\not\models \forall x\psi'$, i.e.
-there exists $w_l\geq w_k$ and $m_l\in w_l$ such that $w_l, v_i[m_l/a]\not\models \psi'[a/x]$.
-Then by the induction hypothesis $M, I, v_c[[v_i(a), w_l],w_l/c]\not\models \psi'[a/x]^c$\todo{hier fehlt die Ersetzung für $[v_i(a), w_l]$.
-Fehlt außerdem nicht irgendein Bezug zu $m_l$?}.
-Furthermore by definition $M, I, v_c[w_l/c, w_k/b]\models b\preceq c$ and $M, I, v_c[v_c[[v_i(a), w_l],w_l/c]\models E(a, c)$\todo{hier fehlt die Ersetzung für $[v_i(a), w_l]$.
-Außerdem ist hier wirklich $v_c[v_c$ gemeint?}.
-Hence $M, I, v_c[w_k/b]\not\models \psi^b$.
-Suppose on the other hand that $w_k, v_i\not\models\forall x\psi'$, i.e.
-for all $w_l\geq w_k$ and $m_l\in w_l$ it holds that $w_l, v_i[m_l/a]\models \psi'[a/x]$.
-Then by the induction hypothesis $M, I, v_c[[m_l, w_l]/A,w_l/c]\models \psi'[a/x]^c$\todo{Hier wohl a statt A} for all $w_l\geq w_k$ and $m_l\in w_l$.
-It can be shown that for every other assignment of $c$ and $a$ either $b\preceq c$ or $E(a, c)$ is violated.
-Hence $M, I, v_c[w_k/b]\models \psi^b$.
-For other forms of $\psi$ one proceeds in similar fashion.
-This completes the proof of the claim.
+there exists $w_l\geq w_k$ and $m_l\in M_{w_l}$ such that $w_l, v_i[m_l/x]\not\models \psi'$.
+Then by the induction hypothesis $\mathcal M, v_c[[m_l, w_l]/x,w_l/c]\not\models \psi'^c$.
+Furthermore by definition $\mathcal M, v_c[w_l/c, w_k/b]\models b\preceq c$ and $\mathcal M, v_c[[m_l, w_l]/x,w_l/c]\models E(x, c)$.
+Hence $\mathcal M, v_c[w_k/b]\not\models \psi^b$.
+Suppose on the other hand that $w_k, v_i\models\forall x\psi'$, i.e.
+for all $w_l\geq w_k$ and $m_l\in M_{w_l}$ it holds that $w_l, v_i[m_l/x]\models \psi'$.
+Then by the induction hypothesis $\mathcal M, v_c[[m_l, w_l]/x,w_l/c]\models \psi'^c$ for all $w_l\geq w_k$ and $m_l\in w_l$.
+On the other hand for every other assignment of $c$ and $x$ either $b\preceq c$ or $E(x, c)$ is violated.
+Hence $\mathcal M, v_c[w_k/b]\models \psi^b$.
 
+	This concludes our proof of the claim.
+
+	Now to show that $\mathcal M$ is a counter-model to $\varphi^C$:
 	Recall that $w_0, v\not\models\varphi$.
-From the claim it follows that \[M, I, v'[[v(a_0), w_0]/a_0\dots[v(a_n), w_0]/a_0, w_0/b]\not\models \varphi^b\]where $a_0\dots a_n$ are the free variables occurring in $\varphi$ and $v'$ is any variable assignment.
-Furthermore a straightforward check of definitions yields \[M, I, v'[[v(a_0), w_0]/a_0\dots[v(a_n), w_0]/a_0, w_0/b]\models K(\varphi)\wedge \text{World}(b)\wedge\vec E(\vec a, b).\]Therefore $M, I\not\models \varphi^C$ and we are done.
-\todo{Da ich hier nicht durchgekommen bin, kannst du den ganzen Beweis nochmal checken :p ?}
+From the claim it follows that \[\mathcal M, v'[[v(a_0), w_0]/a_0\dots[v(a_n), w_0]/a_0, w_0/b]\not\models \varphi^b\]where $a_0\dots a_n$ are the free variables occurring in $\varphi$ and $v'$ is any variable assignment.
+Furthermore a straightforward check of definitions yields \[\mathcal M, v'[[v(a_0), w_0]/a_0\dots[v(a_n), w_0]/a_0, w_0/b]\models K(\varphi)\wedge \text{World}(b)\wedge\vec E(\vec a, b).\]Therefore $\mathcal M, v'\not\models \varphi^C$ and we are done.
 \end{proof}

### 2023-03-15 LPAR/LPAR-EIICL.tex
typo
small fixes

@@
 Free and bound variables are defined as usual and there is a Barendregt's convention in place, i.e.
-we implicitly assume that they comprise distinct sets, such that bound variables are not affected by substitutions of free variables.
+we implicitly assume that they comprise distinct sets, such that bound variables are not affected by substitutions of free variables in any formula.
 
@@
 		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
-		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$.
-		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$.
+		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/x]\models\varphi$.
+		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/x]\models\varphi$.
 	\end{itemize}
-	 Here $a$ is a free variable that does not occur in $\varphi$.
 	$\varphi$ is satisfiable if $\mathcal M, v\models\varphi$ for some $\mathcal M, v$.
@@
 \begin{definition}
-For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{b S}$.
-Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$ such that $m\preceq f_\psi^I(m)$ for each $m\in M$, $\psi\in \mathcal X$.
+For $\psi\in\mathcal X$ let $[\psi]$ denote the corresponding clause in $\mathcal X^{b S}$. Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
 We say $(A\to B)\to C\in\mathcal X$ is \emph{fulfilled} at $m\in M$ iff $A^I(m)\to B^I(m)$ is false or $C^I(m)$ is true. For $\psi\in\mathcal X$ define $$g_\psi : M\to M, m\mapsto\begin{cases}
@@
 \end{proof}
-
-\begin{corollary}\label{cor:prop-tree-model}
-	If $\mathcal M$ is a counter-model to $\varphi^{CH}$ then so is $\mathcal M_T$.
-\end{corollary}
 
@@
 Then there exists $w_l\geq w_k$ such that $w_l, v_i\models \psi_0$ and $w_l, v_i\not\models\psi_1$.
-Then by induction hypothesis $\mathcal M, v_c[w_l/b]\models \psi_0^b$ and $\mathcal M, v_c[w_l/b]\models \psi_1^b$, i.e.
-$\mathcal M, v_c[w_l/b]\models \psi$.
+Then by induction hypothesis $\mathcal M, v_c[w_l/b]\models \psi_0^b$ and $\mathcal M, v_c[w_l/b]\not\models \psi_1^b$, i.e.
+$\mathcal M, v_c[w_l/b]\not\models \psi$.
 Suppose on the other hand that $w_l, v_i\models \psi_0\to\psi_1$.

### 2023-03-18 LPAR/LPAR-EIICL.tex
small fixes

@@
 		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $\mathcal M_w, v\not\models\varphi$ or $\mathcal M_w, v\models\psi$.
-		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $\mathcal M_w, v[m/a]\models\varphi[a/x]$.
+		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $\mathcal M_w, v[m/x]\models\varphi$.
 	\end{itemize}
@@
 \begin{definition}
-	Let $\varphi$ be a propositional formula.\todo{genau genommen müsstest du noch sagen, dass $\nu$ eine freie Variable ist.
-Im letzten Fall der Definition benützen wir dann aber w gleichzeitig als freie und gebundene Variable}
- Define $\varphi^{\nu}$ inductively:
+	Let $\varphi$ be a propositional formula.
+ Define $\varphi^{b}$ inductively:
 	\vspace*{-.1cm}
 	\begin{itemize}
-		\item $\bot^\nu = \bot$.
-		\item $A^{\nu} = A(\nu)$ for every propositional variable $A$.
-		\item $(\varphi\circ\psi)^\nu = \varphi^\nu\circ\psi^\nu$ for $\circ\in\{\wedge, \vee\}$.
-		\item $(\varphi\to \psi)^\nu = \forall w(\nu\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new bound variable.
+		\item $\bot^b = \bot$.
+		\item $A^{b} = A(b)$ for every propositional variable $A$.
+		\item $(\varphi\circ\psi)^b = \varphi^b\circ\psi^b$ for $\circ\in\{\wedge, \vee\}$.
+		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new bound variable.
 	\end{itemize}
@@
 \begin{itemize}
-	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^\nu$ is
-	$$\forall w\left(\nu\preceq w\to\bigwedge_i A_i(w)\to\bigvee_jB_j(w)\right).$$
-	\item For implication clauses $((A\to B)\to C)^\nu$ is
-	$$\forall w(\nu\preceq w\to(\forall k(w\preceq k\to A(k)\to B(k)))\to C(w))$$
+	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^b$ is
+	$$\forall w\left(b\preceq w\to\bigwedge_i A_i(w)\to\bigvee_jB_j(w)\right).$$
+	\item For implication clauses $((A\to B)\to C)^b$ is
+	$$\forall w(b\preceq w\to(\forall k(w\preceq k\to A(k)\to B(k)))\to C(w))$$
 \end{itemize}
@@
 \end{lemma}
-\begin{proof}
-	Let $v$ be a variable assignment.
-It follows directly from the definitions that $\mathcal M_T, v\not\models P(s)$ and $\mathcal M_T, v\models K(\varphi)$.
-It is easy to show that $\mathcal M_T, v\models \psi$ for $\psi\in\mathcal R^{bS}$ from persistency.
-Let $\psi\in\mathcal X^{bS}$, we show $\mathcal M_T, v[\psi_1\dots\psi_n/w] \models s\preceq w\to (w\preceq f_\psi(w)\to A(f_\psi(w)\to B(f_\psi(w)))\to C(w))$ for all $\psi_1\dots\psi_n\in M_T$.
-Let $\psi_1\dots\psi_n\in M_T$ and $m = g_{\psi_n}(\dots(g_{\psi_1}(s^I))$.
-If $\mathcal M, v[m/w]\models C(w)$ we are done.
-Otherwise $\mathcal M, v[m/w]\not\models s\preceq w$ or $\mathcal M, v[m/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w)) \to B(f_\psi(w))$.
-The former is impossible since $g_\psi$ is increasing for all $\psi$.
-For the latter to hold we must have $\mathcal M, v[m/w]\models w\preceq f_\psi(w)$, $\mathcal M, v[m/w]\models A(f_\psi(w))$ and $\mathcal M, v[m/w]\not\models B(f_\psi(w))$.
-If $\psi$ is fulfilled at $m$ we have $g_\psi(m) = m$, otherwise $g_\psi(m) = f_{[\psi]}^I(m)$.
-In any case $\mathcal M_T, v[\psi_1\dots\psi_n/w]\models A(f_\psi(w))$ and $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models B(f_\psi(w))$.
-It follows that $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w))$ and therefore $\mathcal M_T, v\models \psi$.
-\end{proof}
+
+A proof can be found in Appendix~\ref{proof:prop-simplification}.
 
@@
 
-\begin{corollary}
-	If a formula in normal form as in Lemma~\ref{lemma:propositional-normal-form} is intuitionistically invalid then there exists a Kripke counter-model that is a rooted tree of height and degree bounded by $|\mathcal X|$.
+\begin{corollary}[Small model property]
+	If a formula in normal form as in Lemma~\ref{lemma:propositional-normal-form} is intuitionistically invalid then there exists a Kripke counter-model which is a rooted tree of height and degree bounded by $|\mathcal X|$.
 \end{corollary}
@@
 
-It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}.
-\todo{Wollen wir noch etwas sagen wie ``However the previous PSPACE result does not support the transfer of countermodels that we develop in this section''?}
+It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}, but we make it explicit.
+In particular, our translation enables a direct translation of counter-models.
 For lack of space, we do not give the semantics of QBF in this paper and instead refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
@@
 \begin{definition}\label{def:fo-translation}
-	Let $\varphi$ be a predicate formula. Define $\varphi^{\nu}$ inductively:
-\todo{genau genommen müsstest du noch sagen, dass $\nu$ eine freie Variable ist. Im letzten Fall der Definition benützen wir dann aber w gleichzeitig als freie und gebundene Variable}
+	Let $\varphi$ be a predicate formula. Define $\varphi^{b}$ inductively:
+\todo{genau genommen müsstest du noch sagen, dass $b$ eine freie Variable ist. Im letzten Fall der Definition benützen wir dann aber w gleichzeitig als freie und gebundene Variable}
 	\begin{itemize}
-		\item $\bot^\nu = \bot$.
-		\item $A(\vec t)^{\nu} = A(\vec t, \nu)$ for every predicate symbol $A$.
-		\item $(\varphi\circ\psi)^\nu = \varphi^\nu\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
-		\item $(\varphi\to \psi)^\nu = \forall w(\nu\preceq w\to\varphi^{w}\to\psi^{w})$.
-		\item $(\exists x\varphi)^\nu = \exists x(E(x,\nu)\wedge\varphi^\nu)$.
-		\item $(\forall x\varphi)^\nu = \forall w(i\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
+		\item $\bot^b = \bot$.
+		\item $A(\vec t)^{b} = A(\vec t, b)$ for every predicate symbol $A$.
+		\item $(\varphi\circ\psi)^b = \varphi^b\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
+		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$.
+		\item $(\exists x\varphi)^b = \exists x(E(x,b)\wedge\varphi^b)$.
+		\item $(\forall x\varphi)^b = \forall w(i\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
 	\end{itemize}
@@
 		\left\lbrack(\forall x \psi)\to A\right\rbrack&:= \left\lbrack\psi[s(\vec z)/x]\to A\right\rbrack\\
-		\left\lbrack A\to (\forall x\psi)\right\rbrack&:= \{\forall x\chi[x/a]\:|\:\chi\in\left\lbrack A\to\psi[a/x]\right\rbrack\}\\
-		\left\lbrack(\exists x\psi)\to A\right\rbrack&:= \{\forall x\chi[x/a]\:|\:\chi\in\left\lbrack A\to\psi[a/x]\right\rbrack\}\\
+		\left\lbrack A\to (\forall x\psi)\right\rbrack&:= \{\forall x\chi\:|\:\chi\in\left\lbrack A\to\psi\right\rbrack\}\\
+		\left\lbrack(\exists x\psi)\to A\right\rbrack&:= \{\forall x\chi\:|\:\chi\in\left\lbrack A\to\psi\right\rbrack\}\\
 		A\to (\exists x\psi)&:= \left\lbrack A\to\psi[s(\vec z)/x]\right\rbrack
@@
 	where $s$ is a new function symbol, $a$ is a new free variable and $\vec z$ ranges over the free variables in $\psi$.
-	While the generated set of clauses is in general not equivalent to $\psi$, section~\ref{section:herbrandiaztion} guarantees that it is at least equivalid.\hfill$\square$
+	While the generated set of clauses is in general not equivalent to $\psi$, section~\ref{section:herbrandiaztion} guarantees that it is at least equivalid.
+\end{proof}
+
+\begin{lemma}\label{proof:prop-simplification}
+	Suppose $\mathcal M=(M, T)$ is a counter-model to $\varphi^{CH}$, then so is $\mathcal M_T = (M_T, I_T)$.
+\end{lemma}
+\begin{proof}
+	Let $v$ be a variable assignment.
+It follows directly from the definitions that $\mathcal M_T, v\not\models P(s)$ and $\mathcal M_T, v\models K(\varphi)$.
+It is easy to show that $\mathcal M_T, v\models \psi$ for $\psi\in\mathcal R^{bS}$ from persistency.
+Let $\psi\in\mathcal X^{bS}$, we show $\mathcal M_T, v[\psi_1\dots\psi_n/w] \models s\preceq w\to (w\preceq f_\psi(w)\to A(f_\psi(w)\to B(f_\psi(w)))\to C(w))$ for all $\psi_1\dots\psi_n\in M_T$.
+Let $\psi_1\dots\psi_n\in M_T$ and $m = g_{\psi_n}(\dots(g_{\psi_1}(s^I))$.
+If $\mathcal M, v[m/w]\models C(w)$ we are done.
+Otherwise $\mathcal M, v[m/w]\not\models s\preceq w$ or $\mathcal M, v[m/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w)) \to B(f_\psi(w))$.
+The former is impossible since $g_\psi$ is increasing for all $\psi$.
+For the latter to hold we must have $\mathcal M, v[m/w]\models w\preceq f_\psi(w)$, $\mathcal M, v[m/w]\models A(f_\psi(w))$ and $\mathcal M, v[m/w]\not\models B(f_\psi(w))$.
+If $\psi$ is fulfilled at $m$ we have $g_\psi(m) = m$, otherwise $g_\psi(m) = f_{[\psi]}^I(m)$.
+In any case $\mathcal M_T, v[\psi_1\dots\psi_n/w]\models A(f_\psi(w))$ and $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models B(f_\psi(w))$.
+It follows that $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w))$ and therefore $\mathcal M_T, v\models \psi$.
 \end{proof}

### 2023-03-18 LPAR/LPAR-EIICL.tex
Complexity of CPC translation

@@
 
-\begin{lemma}
-	For each propositional formula $\varphi$ there is effectively a formula $\varphi^\#$ with $\mathcal |\varphi^\#|\in\mathcal O(|\mathcal R|\cdot 2^{|\mathcal X|\log|\mathcal X|})$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
-\end{lemma}
-
 The details can be found in Section~\ref{sec:ipc-to-cpc}.
@@
 	$$
-		\left\{\left(A^{\lambda\psi} \to A^{\lambda\psi}\right)\to C^\lambda\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
+		\left\{\left(A^{\lambda\psi} \to B^{\lambda\psi}\right)\to C^\lambda\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
 	$$
@@
 	$$
-		\left\{A^\lambda\to A^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
+		\left\{A^\lambda\to B^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
 	$$
@@
 	$$
-		\varphi^\# := \left(\bigwedge \mathcal P^\#\wedge\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\right)\to P^\epsilon
+		\varphi^\# := \left(\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge \mathcal P^\#\right)\to P^\epsilon
 	$$
@@
 \end{example}
+
+Since $|\Lambda|\in O(2^{|\mathcal X|\log |\mathcal X|})$ a simple counting argument shows that $|\mathcal P^\#|\in\mathcal O(|\varphi|\cdot 2^{|\mathcal X|\log |\mathcal X|})$, $|\mathcal R^\#|\in\mathcal O(|\mathcal R|\cdot2^{|\mathcal X|\log |\mathcal X|})$ and $|\mathcal X|\in \mathcal O(|\mathcal X|\cdot 2^{|\mathcal X|\log|\mathcal X|})$.
+This gives us a bound on the size of $\varphi^#$.
+
+\begin{lemma}
+	The size of $\varphi^\#$ is in $\mathcal O(|\varphi|\cdot 2^{|\mathcal X|\log|\mathcal X|})$.
+\end{corollary}
 

### 2023-03-18 LPAR/LPAR-EIICL.tex
Proof A.2

@@
 Utilizing a normal form for intuitionistic logic~\cite{claessen2015sat}, we establish a small model property for intuitionistic propositional logic.
-We use this property for a direct encoding of the Kripke semantics into classical propositional logic and quantified boolean formulae.
+We use this property for a direct encoding of the Kripke semantics into classical propositional logic and quantified Boolean formulae.
 Next, we transfer the developed techniques to the first order case and provide an embedding of intuitionistic first-order logic into classical first-order-logic.
@@
 	\item We establish a small model property for counter-models to validity in intuitionistic propositional logic.
-	\item Based on the small model property, we provide an embedding of intuitionistic propositional logic into classical propositional logic and into quantified boolean formulae.
+	\item Based on the small model property, we provide an embedding of intuitionistic propositional logic into classical propositional logic and into quantified Boolean formulae.
 	\item We discuss the reduction of counter-models in intuitionistic first-order logic and, based on this reduction, present an embedding of intuitionistic into classical first-order logic.
@@
 The famous double-negation translation establishes
-an embedding of classical into intuitionistic logic, going back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to G\"odel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
+an embedding of classical into intuitionistic logic, going back to Glivenko~\cite{glivenko1929quelques} in the propositional case, and to Gödel~\cite{godel1933intuitionistischen} and Gentzen~\cite{gentzen1936widerspruchsfreiheit} in the first-order case.
 In the propositional case the embedding is particularly simple:
@@
 This most famously holds for $\Pi_2$-formulae in Peano Arithmetic~\cite{friedman1978classically}, but also other conditions are known~\cite{schwichtenberg}.
-In contrast, we propose a translation of an intuitionistic formula to an equi-valid classical formula.
+In contrast, we propose a translation of an intuitionistic formula to an equivalid classical formula.
 
@@
 The details can be found in Section~\ref{sec:ipc-to-cpc}.
-Building on the same idea, we can also give a quantified boolean formula (QBF) of polynomial size, where the branches of the tree-model can be enumerated implicitly using (alternating) quantifiers.
+Building on the same idea, we can also give a quantified Boolean formula (QBF) of polynomial size, where the branches of the tree-model can be enumerated implicitly using (alternating) quantifiers.
 We obtain the following result:
@@
  Define $\varphi^{b}$ inductively:
-	\vspace*{-.1cm}
 	\begin{itemize}
@@
 	Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
-	\vspace*{-.1cm}
-	\[K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))\]
-	\vspace*{-.4cm}
+	\[K(\varphi) := \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))\]
 	with e.g.
-	\vspace*{-.2cm}
 	\begin{align*}
-		\text{\normalfont PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
-		\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ occurs in $\varphi$}\}
+		\text{\normalfont PartialOrder}(\preceq) :=&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
+		\text{\normalfont Persistent}(u, w):=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ occurs in $\varphi$}\}
 	\end{align*}
 	Then define
-	$\varphi^{C} = K(\varphi)\to \varphi^{b}$
+	\[\varphi^{C} := K(\varphi)\to \varphi^{b}\]
 	where $b$ is a new free variable.
@@
 Since $|\Lambda|\in O(2^{|\mathcal X|\log |\mathcal X|})$ a simple counting argument shows that $|\mathcal P^\#|\in\mathcal O(|\varphi|\cdot 2^{|\mathcal X|\log |\mathcal X|})$, $|\mathcal R^\#|\in\mathcal O(|\mathcal R|\cdot2^{|\mathcal X|\log |\mathcal X|})$ and $|\mathcal X|\in \mathcal O(|\mathcal X|\cdot 2^{|\mathcal X|\log|\mathcal X|})$.
-This gives us a bound on the size of $\varphi^#$.
+This gives us a bound on the size of $\varphi^\#$.
 
@@
 	The size of $\varphi^\#$ is in $\mathcal O(|\varphi|\cdot 2^{|\mathcal X|\log|\mathcal X|})$.
-\end{corollary}
+\end{lemma}
 
@@
 	\end{itemize}
-	Define$$\varphi^Q_{i} = \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
+	Define$$\varphi^Q_{i} := \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
 	for $0 < i < |\mathcal{X}|$, as well as the special cases
-	$$\varphi^Q_{|\mathcal{X}|} = \exists \vec A^{|\mathcal{X}|}\left(\text{\normalfont Persistent}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal R}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal X}(|\mathcal{X}|)\right)$$
-	for leaves, and $$\varphi^Q = \varphi_0^Q = \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
+	$$\varphi^Q_{|\mathcal{X}|} := \exists \vec A^{|\mathcal{X}|}\left(\text{\normalfont Persistent}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal R}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal X}(|\mathcal{X}|)\right)$$
+	for leaves, and $$\varphi^Q = \varphi_0^Q := \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
 	Example encodings of the above formulae are:
 	\begin{align*}
-		\text{\normalfont Valid}(n) = &\left(\bigvee\left\{X^n_\psi\:|\:\psi\in\mathcal X\right\}\right)\wedge\left(\bigwedge \left\{\neg(X^n_{\psi_1}\wedge X^n_{\psi_2})\:|\:\psi_1\neq\psi_2\in\mathcal X\right\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{X^i_\psi\to\neg X^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal X\right\}\right)\\
-		\text{\normalfont Persistent}(n) = & \bigwedge\left\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\right\}\\
-		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{\bigwedge A^n_i\to\bigvee B^n_j\:|\:\bigwedge A_i\to\bigvee B_j \in \mathcal R\right\}\\
-		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\left\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\right\}
+		\text{\normalfont Valid}(n) := &\left(\bigvee\left\{X^n_\psi\:|\:\psi\in\mathcal X\right\}\right)\wedge\left(\bigwedge \left\{\neg(X^n_{\psi_1}\wedge X^n_{\psi_2})\:|\:\psi_1\neq\psi_2\in\mathcal X\right\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{X^i_\psi\to\neg X^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal X\right\}\right)\\
+		\text{\normalfont Persistent}(n) := & \bigwedge\left\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\right\}\\
+		\text{\normalfont Sat}_{\mathcal R}(n) := &\bigwedge\left\{\bigwedge A^n_i\to\bigvee B^n_j\:|\:\bigwedge A_i\to\bigvee B_j \in \mathcal R\right\}\\
+		\text{\normalfont Sat}_{\mathcal X}(n) := & \bigwedge\left\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\right\}
 	\end{align*}
@@
 	Let $\varphi$ be a predicate formula. Define $\varphi^{b}$ inductively:
-\todo{genau genommen müsstest du noch sagen, dass $b$ eine freie Variable ist. Im letzten Fall der Definition benützen wir dann aber w gleichzeitig als freie und gebundene Variable}
 	\begin{itemize}
@@
 	\begin{align*}
-		K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
+		K(\varphi) := \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
 		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
 	\end{align*}
-	with e.g.:\vspace*{-.76cm}
+	with e.g.:
 	\begin{align*}
-		\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
-		\text{DomainSubset}(u, w) = &\:\forall x(E(x, u)\to E(x, w))\\
-		\text{World}(u) = &\:\exists xE(x, u)\\
-		\text{DomainClosed}(u) = &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
-		\text{Persistent}(u, w) = &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a predicate in $\varphi$}\right\}
+		\text{PartialOrder}(\preceq) := &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
+		\text{DomainSubset}(u, w) := &\:\forall x(E(x, u)\to E(x, w))\\
+		\text{World}(u) := &\:\exists xE(x, u)\\
+		\text{DomainClosed}(u) := &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
+		\text{Persistent}(u, w) := &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a predicate in $\varphi$}\right\}
 	\end{align*}
 	Then define
-	$\varphi^C = K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b$ where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
+	\[\varphi^C:= K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b\] where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
 \end{definition}
@@
 \begin{definition}
-	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$.
-W.l.o.g.
-suppose $\{m\in M\:|\:\text{World}(m)\}$ has a minimal element, say $w_0$.
-Define $\mathcal M_T = (M_T, I_T)$ as follows:
+	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^{CH}$. Similarly to the propositional case for $\psi\in\mathcal X$ define \[g_\psi :M^n\to M, (\vec z, m) \mapsto
+	\begin{cases} 
+	m,&\text{if }m\not\preceq f_\psi(\vec z, m)\\
+	f_{[\psi]}^I(\vec z, m)&\text{otherwise}
+	\end{cases}
+	\]where $[\psi]$ denotes the corresponding element of $\mathcal X^{bS}$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
 	\begin{itemize}
-		\item $M_T$ is the union of $M$ and sequences $(\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)$ such that $\vec E^I(\vec m_1, w_0)$ and $\vec E^I(\vec m_i, f_{\psi_i}(\vec m_i, f_{\psi_{i-1}}^I(\vec m_{i-1}, \dots f_{\psi_1}^I(\vec m_1, w_0)\dots)))$ for $i > 0$ hold.
-		\item $s^{I_T} = w_0$, $\vec r^{I_T} = \vec r^I$.
+		\item $M_T$ is the union of $\{m\in M\:|\:\text{World}^I(m)\wedge s^I\preceq^I m\}$ and sequences $(\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)$ such that $\vec E^I(\vec m_1, w_0)$ and $\vec E^I(\vec m_i, g_{\psi_i}(\vec m_i, g_{\psi_{i-1}}(\vec m_{i-1}, \dots g_{\psi_1}(\vec m_1, w_0)\dots)))$ for $i > 0$ hold.
+		\item $s^{I_T} = s^I$, $\vec r^{I_T} = \vec r^I$.
 		\item We interpret $\preceq$ trivially on $M$ and as the prefix-order elsewhere.
-		\item Have $E^{I_T}(m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n))$ iff $E^I(m, f_{\psi_n}^I(\vec m_n, \dots f_{\psi_1}^I(\vec m_1, w_0)\dots))$.
+		\item Have $E^{I_T}(m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n))$ iff $E^I(m, g_{\psi_n}(\vec m_n, \dots g_{\psi_1}(\vec m_1, w_0)\dots))$.
 		\item We set $f_\psi^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)(\vec m, \psi)$ and arbitrary else.
-		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, f_{\psi_n}^I(\vec m_n, \dots(f_{\psi_1}^I(\vec m_1, s^I))\dots))$$ and arbitrary otherwise, for function symbols $f^{I_T}(\vec m) = f^I(\vec m)$ and arbitrary else.
+		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, g_{\psi_n}(\vec m_n, \dots(g_{\psi_1}(\vec m_1, s^I))\dots))$$ and arbitrary otherwise, for function symbols $f^{I_T}(\vec m) = f^I(\vec m)$ and arbitrary else.
 	\end{itemize}
@@
 
-Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$f_{\psi_1}(\vec m_1, f_{\psi_2}(\vec m_2, \dots f_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
+Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$g_{\psi_1}(\vec m_1, g_{\psi_2}(\vec m_2, \dots g_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
 In particular, this allows us to completely eliminate $\preceq$, that is we remove the PartialOrder$(\preceq)$
@@
 	Let $\mathcal R,\mathcal X,\mathcal Q$ be as above.
-For every $n$-ary predicate $A$ define a new $n+1$-ary predicate of the same name.
+For each $n$-ary predicate $A$ define a new $n+1$-ary predicate of the same name.
 For every $\psi\in\mathcal \mathcal X\cup\mathcal Q$ define a new function symbol $f_\psi$.
-Consider a new binary predicate symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
+Consider a new binary predicate $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$.
 Obtain $\mathcal R^\#$ by including
@@
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
@@
 
+\newpage
+
 \section{Omitted proofs}
@@
 \begin{proof}
-	Let $v$ be a variable assignment.
-It follows directly from the definitions that $\mathcal M_T, v\not\models P(s)$ and $\mathcal M_T, v\models K(\varphi)$.
-It is easy to show that $\mathcal M_T, v\models \psi$ for $\psi\in\mathcal R^{bS}$ from persistency.
-Let $\psi\in\mathcal X^{bS}$, we show $\mathcal M_T, v[\psi_1\dots\psi_n/w] \models s\preceq w\to (w\preceq f_\psi(w)\to A(f_\psi(w)\to B(f_\psi(w)))\to C(w))$ for all $\psi_1\dots\psi_n\in M_T$.
-Let $\psi_1\dots\psi_n\in M_T$ and $m = g_{\psi_n}(\dots(g_{\psi_1}(s^I))$.
-If $\mathcal M, v[m/w]\models C(w)$ we are done.
+	Let $v$ be any variable assignment.
+We show that $\mathcal M_T, v\not\models \varphi$.
+It follows directly from the definitions and the same facts for $\mathcal M$ that $\mathcal M_T, v\not\models P(s)$ and $\mathcal M_T, v\models K(\varphi)$.
+It remains to show that $\mathcal M_T, v\models\bigwedge \mathcal R^{bS}$ and $\mathcal M_T, v\models \bigwedge\mathcal X^{bS}$.
+
+	Let \[\forall w\phi = \forall w\left(s\preceq w\to\bigwedge_i A_i(w)\to \bigvee_j B_j(w)\right)\in\mathcal R^{bS}\] and $\psi_1\dots\psi_n\in M_T$.
+Since $\mathcal M$ is a counter-model to $\varphi$ we know that $\mathcal M, v\models\forall w\phi$ and thus $\mathcal M, v[g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots)/w]\models \phi$.
+Then by definition $\mathcal M_T, v[\psi_1\dots\psi_n/w]\models\phi$.
+This shows that $\mathcal M_T, v\models\forall w\phi$.
+
+	On the other hand consider \[\forall w\phi = \forall w(s\preceq w\to(w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w)))\to C(w))\in\mathcal X^{bS}\] and $\psi_1\dots\psi_n\in M_T$.
+Let $m = g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots)$.
+If $\mathcal M, v[m/w]\models C(w)$ it follows that $\mathcal M_T, v[\psi_1\dots\psi_m]\models C(w)$ and we are done.
 Otherwise $\mathcal M, v[m/w]\not\models s\preceq w$ or $\mathcal M, v[m/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w)) \to B(f_\psi(w))$.
@@
 For the latter to hold we must have $\mathcal M, v[m/w]\models w\preceq f_\psi(w)$, $\mathcal M, v[m/w]\models A(f_\psi(w))$ and $\mathcal M, v[m/w]\not\models B(f_\psi(w))$.
-If $\psi$ is fulfilled at $m$ we have $g_\psi(m) = m$, otherwise $g_\psi(m) = f_{[\psi]}^I(m)$.
-In any case $\mathcal M_T, v[\psi_1\dots\psi_n/w]\models A(f_\psi(w))$ and $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models B(f_\psi(w))$.
-It follows that $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w))$ and therefore $\mathcal M_T, v\models \psi$.
+
+	There are now 2 cases.
+Suppose $\psi$ is fulfilled at $m$.
+Then we have $\mathcal M, v[m/w]\not\models A(w)\to B(w)$ and $g_\psi(m) = m$ and therefore $\mathcal M, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f\psi(w))$, i.e.
+$\mathcal M, v[\psi_1\dots\psi_n/w]\models \phi$.
+
+	On the other hand suppose $\psi$ is not fulfilled at $m$.
+Then $\mathcal M, v[m/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$ and $g_\psi(m) = f^I_\psi(m)$.
+In this case it directly follows that $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$ and thus $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$, i.e.
+$\mathcal M_T, v[\psi_1\dots\psi_n/w]\models\phi$.
+
+	In any case $\mathcal M_T, v\models\forall w\phi$.
 \end{proof}

### 2023-03-19 LPAR/LPAR-EIICL.tex
Formatting
Added newpage
typos

@@
 Suppose we are given a formula $\varphi$ in normal form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}.
-We now given an intuition on how to obtain a QBF of polynomial size that expresses that $\varphi$ has an intuitionistic counter-model:
+We now give an intuition on how to obtain a QBF of polynomial size that expresses that $\varphi$ has an intuitionistic counter-model:
 Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
@@
 
+\newpage
+
 \bibliographystyle{plain}
@@
 
+\newpage
+
 \appendix
-
-\newpage
 
@@
 Let $m = g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots)$.
-If $\mathcal M, v[m/w]\models C(w)$ it follows that $\mathcal M_T, v[\psi_1\dots\psi_m]\models C(w)$ and we are done.
+If $\mathcal M, v[m/w]\models C(w)$ it follows directly that $\mathcal M_T, v[\psi_1\dots\psi_m]\models C(w)$ and we are done.
 Otherwise $\mathcal M, v[m/w]\not\models s\preceq w$ or $\mathcal M, v[m/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w)) \to B(f_\psi(w))$.
 The former is impossible since $g_\psi$ is increasing for all $\psi$.
-For the latter to hold we must have $\mathcal M, v[m/w]\models w\preceq f_\psi(w)$, $\mathcal M, v[m/w]\models A(f_\psi(w))$ and $\mathcal M, v[m/w]\not\models B(f_\psi(w))$.
+For the latter to hold we must have $\mathcal M, v[m/w]\models w\preceq f_\psi(w)$ as well as $\mathcal M, v[m/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$.
+Since by definition $\mathcal M_T, v[\psi_1\dots\psi_m/w]\models s\preceq w\wedge w\preceq f_\psi(w)$ we must show $\mathcal M_T, v[\psi_1\dots\psi_m/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$.
 
@@
 Suppose $\psi$ is fulfilled at $m$.
-Then we have $\mathcal M, v[m/w]\not\models A(w)\to B(w)$ and $g_\psi(m) = m$ and therefore $\mathcal M, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f\psi(w))$, i.e.
-$\mathcal M, v[\psi_1\dots\psi_n/w]\models \phi$.
+Then we have $\mathcal M, v[m/w]\not\models A(w)\to B(w)$ and $g_\psi(m) = m$ and therefore $\mathcal M, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$, i.e.
+indeed $\mathcal M, v[\psi_1\dots\psi_n/w]\models\phi$.
 
-	On the other hand suppose $\psi$ is not fulfilled at $m$.
-Then $\mathcal M, v[m/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$ and $g_\psi(m) = f^I_\psi(m)$.
-In this case it directly follows that $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$ and thus $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$, i.e.
-$\mathcal M_T, v[\psi_1\dots\psi_n/w]\models\phi$.
+	On the other hand suppose $\psi$ is not fulfilled at $m$, i.e.
+$g_\psi(m) = f^I_\psi(m)$.
+Since $\mathcal M, v[m/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$ it directly follows that $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$ and thus $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$, i.e.
+indeed $\mathcal M_T, v[\psi_1\dots\psi_n/w]\models\phi$.
 

### 2023-03-19 LPAR/LPAR-EIICL.tex
conclusion
LPAR-fianl

@@
 A term is either a variable or, given an $n$-ary function symbol $f$ and terms $t_1, \ldots, t_n$, $f(t_1, \ldots, t_n)$.
-A formula is either $\bot$, or, given a $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$, $P(t_1, \ldots, t_n)$, or, given formulae $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$, or, given a formula $\varphi$ and a variable $x$, $\exists x\varphi$ or $\forall x\varphi$.
+A formula is either $\bot$ or, given a $n$-ary predicate symbol $P$ and terms $t_1, \ldots, t_n$, $P(t_1, \ldots, t_n)$ or, given formulae $\varphi$ and $\psi$ and $\circ\in\{\wedge, \vee,\to\}$, $\varphi\circ\psi$ or, given a formula $\varphi$ and a variable $x$, $\exists x\varphi$ or $\forall x\varphi$.
 Free and bound variables are defined as usual and there is a Barendregt's convention in place, i.e.
@@
 	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
-	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulae $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where we set
+	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulae $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where
 	\begin{itemize}
@@
 
-Then, the key to reducing the first-order logic encoding to a propositional logic encoding, is a small model property that supports the reduction of counter-models.
+Then, the key to reducing the first-order logic encoding to a propositional logic encoding is a small model property that supports the reduction of counter-models.
 In particular, given $$\varphi = \left(\bigwedge\mathcal R\wedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form} we establish:
@@
 Since $|\Lambda|\in O(2^{|\mathcal X|\log |\mathcal X|})$ a simple counting argument shows that $|\mathcal P^\#|\in\mathcal O(|\varphi|\cdot 2^{|\mathcal X|\log |\mathcal X|})$, $|\mathcal R^\#|\in\mathcal O(|\mathcal R|\cdot2^{|\mathcal X|\log |\mathcal X|})$ and $|\mathcal X|\in \mathcal O(|\mathcal X|\cdot 2^{|\mathcal X|\log|\mathcal X|})$.
-This gives us a bound on the size of $\varphi^\#$.
+This gives a bound on the size of $\varphi^\#$.
 
@@
 		\item $(\varphi\circ\psi)^b = \varphi^b\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
-		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$.
+		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is a new bound variable.
 		\item $(\exists x\varphi)^b = \exists x(E(x,b)\wedge\varphi^b)$.
@@
 	\end{itemize}
-	where $w$ is some new bound variable. Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
+	Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
 	\begin{align*}
@@
 	\begin{cases} 
-	m,&\text{if }m\not\preceq f_\psi(\vec z, m)\\
+	m,&\text{if }m\not\preceq f_{[\psi]}^I(\vec z, m)\\
 	f_{[\psi]}^I(\vec z, m)&\text{otherwise}
@@
 	\begin{itemize}
-		\item $M_T$ is the union of $\{m\in M\:|\:\text{World}^I(m)\wedge s^I\preceq^I m\}$ and sequences $(\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)$ such that $\vec E^I(\vec m_1, w_0)$ and $\vec E^I(\vec m_i, g_{\psi_i}(\vec m_i, g_{\psi_{i-1}}(\vec m_{i-1}, \dots g_{\psi_1}(\vec m_1, w_0)\dots)))$ for $i > 0$ hold.
+		\item $M_T$ is the union of $\{m\in M\:|\:\neg\text{World}^I(m)\}$ and sequences $(\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)$ such that $\vec E^I(\vec m_1, s^I)$ and $\vec E^I(\vec m_i, g_{\psi_{i-1}}(\vec m_{i-1}, g_{\psi_{i-2}}(\dots g_{\psi_1}(\vec m_1, s^I)\dots)))$ for $i > 0$ hold.
 		\item $s^{I_T} = s^I$, $\vec r^{I_T} = \vec r^I$.
-		\item We interpret $\preceq$ trivially on $M$ and as the prefix-order elsewhere.
-		\item Have $E^{I_T}(m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n))$ iff $E^I(m, g_{\psi_n}(\vec m_n, \dots g_{\psi_1}(\vec m_1, w_0)\dots))$.
+		\item We interpret $\preceq$ as the prefix-order on the set of sequences and trivially elsewhere.
+		\item Have $E^{I_T}(m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n))$ iff $E^I(m, g_{\psi_n}(\vec m_n, \dots g_{\psi_1}(\vec m_1, s^I)\dots))$.
 		\item We set $f_\psi^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)(\vec m, \psi)$ and arbitrary else.
-		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, g_{\psi_n}(\vec m_n, \dots(g_{\psi_1}(\vec m_1, s^I))\dots))$$ and arbitrary otherwise, for function symbols $f^{I_T}(\vec m) = f^I(\vec m)$ and arbitrary else.
+		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, g_{\psi_n}(\vec m_n, \dots(g_{\psi_1}(\vec m_1, s^I))\dots))$$ and arbitrarily elsewhere, for function symbols $f$ we set $f^{I_T}(\vec m) = f^I(\vec m)$ so long as $\vec m\in \{m\in M\:|\:\neg\text{World}^I(m)\}^n$ and arbitrarily elsewhere.
 	\end{itemize}
@@
 
-We have presented embeddings of intuitionistic into classical.
+We have presented embeddings of intuitionistic into classical logic for the propositional and the predicate case.
 Using Kripke semantics and model reduction we have established a complexity overhead of intuitionistic logic over classical logic parametrized by $|\mathcal X|$ and $|\mathcal Q|$.

### 2023-05-03 LPAR/LPAR-EIICL.tex
changes from reviews

@@
 \usepackage[T1]{fontenc}
-\usepackage{todonotes}
-%
 \usepackage{tikz}
 \usetikzlibrary{shapes,backgrounds}
+
+%used for d
+\usepackage{todonotes}
+\usepackage{xcolor}
+\usepackage[normalem]{ulem}
+\newcommand{\red}[1]{{\color{red} #1}}
 
@@
 \begin{definition}
-	A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds}, called the Kripke \emph{frame}, and an indexed family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_w(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
+	A \emph{Kripke structure}  \sout{$\mathcal K = (W, (v_w)_{w\in W})$} \red{$\mathcal K = (\mathbf W, (v_w)_{w\in W})$} consists of a partially ordered set $\mathbf W = (W, \leq)$ of \emph{worlds}, called the Kripke \emph{frame}, and an indexed family of valuations $(v_w)_{w\in W}$ such that \sout{$v_u(A)\leq v_w(A)$} \red{$v_u(A) = 1$ entails $v_w(A) = 1$} for all worlds $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
 	The model relation between the $v_u$ and formulae $\varphi$ is defined as before, except in the case of implications, where we set
@@
 \begin{definition}
-	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
+	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set \sout{$W$} \red{$(W, \leq)$}, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
 	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulae $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where
@@
 	We write $u, v\models\varphi$ if $\mathcal M_u, v\models \varphi$.
-$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$.
+$\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u\in W$ and variable assignment $v$.
 $\varphi$ is valid if it is satisfied by every Kripke structure.
@@
 
-Finally we recall a result from~\cite{otten2005clausal}, on which we build in this paper.
+Finally we recall a result from~\cite{claessen2015sat}, on which we build in this paper.
 
@@
 	There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
-The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary predicate symbol in $\varphi$ there is a corresponding $n+1$-ary predicate symbol in $\varphi^\#$, and $\varphi$ contains a new binary predicate $E$ as well as a number of new function symbols.
+The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary predicate symbol in $\varphi$ there is a corresponding $n+1$-ary predicate symbol in $\varphi^\#$, and \sout{$\varphi$}\red{$\varphi^\#$} contains a new binary predicate $E$ as well as a number of new function symbols.
 Furthermore, there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
@@
 Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$.
-For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional Variable $A^\lambda$.
+For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional \sout{Variable}\red{variable} $A^\lambda$.
 Define $\mathcal R^\#$ as the union over all sets
@@
 For lack of space, we do not give the semantics of QBF in this paper and instead refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
-The translation from the last section will serve as starting point for this section.
+The translation from the last section will serve as \red{a} starting point for this section.
 
@@
 		\item $A(\vec t)^{b} = A(\vec t, b)$ for every predicate symbol $A$.
-		\item $(\varphi\circ\psi)^b = \varphi^b\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
+		\item $(\varphi\circ\psi)^b = \varphi^b\circ\sout{\psi^u}\red{\psi^b}$ for $\circ\in\{\wedge, \vee\}$.
 		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is a new bound variable.
 		\item $(\exists x\varphi)^b = \exists x(E(x,b)\wedge\varphi^b)$.
-		\item $(\forall x\varphi)^b = \forall w(i\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
+		\item $(\forall x\varphi)^b = \forall w(\sout{i}\red{b}\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
 	\end{itemize}
@@
 and each formula $\psi\in\mathcal Q^{bS}$ is of the form
-$$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to \forall y(E(y, f_\psi(\vec x, w))\to A(\vec a, f_\psi(\vec x, w)))\to B(\vec B, w))$$
+$$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to \forall y(E(y, f_\psi(\vec x, w))\to A(\vec a, f_\psi(\vec x, w)))\to B(\vec{\red{b}}, w))$$
 with $f_\psi$ being a new function symbol.
@@
 
-Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$g_{\psi_1}(\vec m_1, g_{\psi_2}(\vec m_2, \dots g_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
+Even if we \sout{don't}\red{don't} have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$g_{\psi_1}(\vec m_1, g_{\psi_2}(\vec m_2, \dots g_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
 In particular, this allows us to completely eliminate $\preceq$, that is we remove the PartialOrder$(\preceq)$
@@
 All benchmarks were conducted on an Intel Core i5-8400 2.80GHz machine running Linux with a time limit of 10s per problem.
-The problem \textsc{SYN007+1} was removed as our normal form procedure would explode the size of instance.
+The problem \textsc{SYN007+1} was removed as our normal form procedure would explode the size of \red{the} instance.
 
@@
 However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},
-we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$ resp. $|\mathcal X| \geq 6$.
+we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$\sout{resp. $|\mathcal X| \geq 6$}.
 

### 2023-05-10 LPAR/LPAR-EIICL.tex
small adjustments

@@
 \usetikzlibrary{shapes,backgrounds}
-
-%used for d
-\usepackage{todonotes}
-\usepackage{xcolor}
-\usepackage[normalem]{ulem}
-\newcommand{\red}[1]{{\color{red} #1}}
 
@@
 \begin{definition}
-	A \emph{Kripke structure}  \sout{$\mathcal K = (W, (v_w)_{w\in W})$} \red{$\mathcal K = (\mathbf W, (v_w)_{w\in W})$} consists of a partially ordered set $\mathbf W = (W, \leq)$ of \emph{worlds}, called the Kripke \emph{frame}, and an indexed family of valuations $(v_w)_{w\in W}$ such that \sout{$v_u(A)\leq v_w(A)$} \red{$v_u(A) = 1$ entails $v_w(A) = 1$} for all worlds $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
+	A \emph{Kripke structure}  $\mathcal K = (\mathbf W, (v_w)_{w\in W})$ consists of a partially ordered set $\mathbf W = (W, \leq)$ of \emph{worlds}, called the Kripke \emph{frame}, and an indexed family of valuations $(v_w)_{w\in W}$ such that $v_u(A) = 1$ entails $v_w(A) = 1$ for all worlds $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
 	The model relation between the $v_u$ and formulae $\varphi$ is defined as before, except in the case of implications, where we set
@@
 \begin{definition}
-	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set \sout{$W$} \red{$(W, \leq)$}, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
+	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $(W, \leq)$, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
 	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulae $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where
@@
 	There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid.
-The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary predicate symbol in $\varphi$ there is a corresponding $n+1$-ary predicate symbol in $\varphi^\#$, and \sout{$\varphi$}\red{$\varphi^\#$} contains a new binary predicate $E$ as well as a number of new function symbols.
+The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary predicate symbol in $\varphi$ there is a corresponding $n+1$-ary predicate symbol in $\varphi^\#$, and $\varphi^\#$ contains a new binary predicate $E$ as well as a number of new function symbols.
 Furthermore, there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
@@
 A proof can be found in Appendix~\ref{proof:prop-simplification}.
-
 \begin{remark}
@@
 Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$.
-For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional \sout{Variable}\red{variable} $A^\lambda$.
+For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional variable $A^\lambda$.
 Define $\mathcal R^\#$ as the union over all sets
@@
 
+Note that it is straightforward to obtain an intuitionistic counter-model to $\varphi$ with frame $\Lambda$ from a classical counter-model to $\varphi^\#$.
+
 Since $|\Lambda|\in O(2^{|\mathcal X|\log |\mathcal X|})$ a simple counting argument shows that $|\mathcal P^\#|\in\mathcal O(|\varphi|\cdot 2^{|\mathcal X|\log |\mathcal X|})$, $|\mathcal R^\#|\in\mathcal O(|\mathcal R|\cdot2^{|\mathcal X|\log |\mathcal X|})$ and $|\mathcal X|\in \mathcal O(|\mathcal X|\cdot 2^{|\mathcal X|\log|\mathcal X|})$.
@@
 For lack of space, we do not give the semantics of QBF in this paper and instead refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
-The translation from the last section will serve as \red{a} starting point for this section.
+The translation from the last section will serve as a starting point for this section.
 
@@
 		\item $A(\vec t)^{b} = A(\vec t, b)$ for every predicate symbol $A$.
-		\item $(\varphi\circ\psi)^b = \varphi^b\circ\sout{\psi^u}\red{\psi^b}$ for $\circ\in\{\wedge, \vee\}$.
+		\item $(\varphi\circ\psi)^b = \varphi^b\circ\psi^b$ for $\circ\in\{\wedge, \vee\}$.
 		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is a new bound variable.
 		\item $(\exists x\varphi)^b = \exists x(E(x,b)\wedge\varphi^b)$.
-		\item $(\forall x\varphi)^b = \forall w(\sout{i}\red{b}\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
+		\item $(\forall x\varphi)^b = \forall w(b\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
 	\end{itemize}
@@
 and each formula $\psi\in\mathcal Q^{bS}$ is of the form
-$$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to \forall y(E(y, f_\psi(\vec x, w))\to A(\vec a, f_\psi(\vec x, w)))\to B(\vec{\red{b}}, w))$$
+$$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to \forall y(E(y, f_\psi(\vec x, w))\to A(\vec a, f_\psi(\vec x, w)))\to B(\vec{b}, w))$$
 with $f_\psi$ being a new function symbol.
@@
 
-Even if we \sout{don't}\red{don't} have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$g_{\psi_1}(\vec m_1, g_{\psi_2}(\vec m_2, \dots g_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
+Even if we do not have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$g_{\psi_1}(\vec m_1, g_{\psi_2}(\vec m_2, \dots g_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
 In particular, this allows us to completely eliminate $\preceq$, that is we remove the PartialOrder$(\preceq)$
@@
 All benchmarks were conducted on an Intel Core i5-8400 2.80GHz machine running Linux with a time limit of 10s per problem.
-The problem \textsc{SYN007+1} was removed as our normal form procedure would explode the size of \red{the} instance.
+The problem \textsc{SYN007+1} was removed as our normal form procedure would explode the size of the instance.
 
@@
 However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},
-we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$\sout{resp. $|\mathcal X| \geq 6$}.
+we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$.
+Also, there were some instances with high $|\mathcal Q|$ and low $|\mathcal X|$ that were solved by our approach but not the previous solvers.
 
@@
 Figure~\ref{fig:c} shows that our approach solved 52 problems not solved by the state-of-the-art provers.
-Notably, while there is big overlap in solved instances, ileanCoP and nanoCoP also have a significant symmetric difference.
+Note that ileanCoP and nanoCop were combined for presentation purposes but they also have a significant symmetric difference, e.g.
+of the 417 problem not solved by our approach there were 68 solved only by ileanCoP and 55 solved only by nanoCoP.
 
@@
 However, idempotence is a quite powerful concept from a model-theoretic perspective.
-We suspect that a more clever encoding of this property could improve results.
+A more clever encoding of this property could improve results.
 There are other restrictions on models we could consider, based on structural properties of the formula, e.g.

### 2023-05-14 LPAR/LPAR-EIICL.tex
appendix presentation

@@
 
-\let\OLDitemize\itemize
-\renewcommand\itemize{\OLDitemize\addtolength{\itemsep}{-5pt}}
+\newcommand\slimitemize{\itemize\addtolength{\itemsep}{-5pt}}
 
@@
 To this end, we study embeddings of intuitionistic into classical logic and propose the following methodology:
-\begin{itemize}
+\begin{slimitemize}
 	\item Give for each $\varphi$ a formula $\varphi^\#$, which is classically valid iff $\varphi$ is intuitionistically valid.
@@
 	\item Translate the classical proof (resp. counter-model) of $\varphi^\#$ given by the prover to a corresponding intuitionistic proof (resp. counter-model) of $\varphi$.
-\end{itemize}
+\end{slimitemize}
 \subsection*{Contributions}
 The following main contributions will be discussed in more detail in Sections~\ref{sec:overview} and~\ref{sec:evaluation}:
-\begin{itemize}
+\begin{slimitemize}
 	\item We establish a small model property for counter-models to validity in intuitionistic propositional logic.
@@
  We show that our approach can compete with state-of-the-art provers for certain classes of benchmarks, in particular when the intuitionistic content is low.
-\end{itemize}
+\end{slimitemize}
 
@@
 	A \emph{valuation} $v$ maps each propositional variable $A$ to a \emph{truth value} $v(A)\in\{0, 1\}$. We inductively define the model relation $\models$ between $v$ and formulae:
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $v\not\models \bot$
@@
 		\item $v\models\varphi\to \psi$ iff $v\not\models\varphi$ or $v\models\psi$.
-	\end{itemize}
+	\end{slimitemize}
 	A valuation $v$ is a \emph{model} for $\varphi$ if $v\models\varphi$.
@@
 	The model relation between the $v_u$ and formulae $\varphi$ is defined as before, except in the case of implications, where we set
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $v_u\models\varphi\to \psi$ iff for all $w\geq u$ we have $w\not\models\varphi$ or $w\models\psi$.
-	\end{itemize}
+	\end{slimitemize}
 	We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$.
@@
 	Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
-	\begin{itemize}
+	\begin{slimitemize}
 		\item to each $n$-ary function symbol $f\in\Sigma$ a $n$-ary function $f^I: M^n\to M$.
 		\item to each $n$-ary predicate symbol $R\in\Sigma$ a $n$-ary predicate $R^I\subseteq M^n$.
-	\end{itemize}
+	\end{slimitemize}
 	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
@@
 	Then terms are interpreted as follows:
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $a^{I, v} = v(a)$ for each free variable $a$.
 		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$ for each $n$-ary function symbol $f\in\Sigma$.
-	\end{itemize}
+	\end{slimitemize}
 	We define a model relation between pairs $\mathcal M, v$ and formulae $\varphi$ as follows
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $\mathcal M, v\not\models\bot$
@@
 		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/x]\models\varphi$.
-	\end{itemize}
+	\end{slimitemize}
 	$\varphi$ is satisfiable if $\mathcal M, v\models\varphi$ for some $\mathcal M, v$.
@@
 	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulae $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $\mathcal M_w, v\not\models\varphi$ or $\mathcal M_w, v\models\psi$.
 		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $\mathcal M_w, v[m/x]\models\varphi$.
-	\end{itemize}
+	\end{slimitemize}
 	We write $u, v\models\varphi$ if $\mathcal M_u, v\models \varphi$.
@@
 	For formulae $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to a set of free variables $Z$ by simultaneous induction:
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $A^S_Z = A^H_Z = A$ for each atomic $A$.
@@
 		\item $(\exists x\varphi)^S_Z = \varphi[s(z_1,\dots,z_n)/x]^S_Z$ where $s$ is a new function, $\{z_1\dots z_n\} = Z$.\\$(\exists x\varphi)^H_Z = \exists x(\varphi^H_{Z\cup\{x\}})$.
-	\end{itemize}
+	\end{slimitemize}
 	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\forall x_1\dots\forall x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
@@
 	For every formula $\varphi$
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
 		\item $\varphi$ and $\varphi^H$ are classically equivalid.
-	\end{itemize}
+	\end{slimitemize}
 \end{theorem}
@@
  Define $\varphi^{b}$ inductively:
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $\bot^b = \bot$.
@@
 		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new bound variable.
-	\end{itemize}
+	\end{slimitemize}
 	Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
@@
 assuming that $\varphi$ is of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ for some flat clauses $\mathcal R$ and implicational clauses $\mathcal X$, further simplification is possible:
-\begin{itemize}
+\begin{slimitemize}
 	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^b$ is
@@
 	$$\forall w(b\preceq w\to(\forall k(w\preceq k\to A(k)\to B(k)))\to C(w))$$
-\end{itemize}
+\end{slimitemize}
 
@@
 	\end{cases}$$Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $M_T$ is the set of sequences without repetition on $\mathcal X$.
@@
 		\item For propositional variables $P$, we set $$P^{I_T}\left(\psi_1\dots \psi_n\right) = P^I\left(g_{\psi_n}\left(\dots\left(g_{\psi_1}\left(s^I\right)\right)\dots\right)\right).$$
-	\end{itemize}
+	\end{slimitemize}
 \end{definition}
@@
 Define
-	\begin{itemize}
+	\begin{slimitemize}
 		\item {\normalfont Valid}$(n)$ which encodes that $\vec X^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
@@
 		\item {\normalfont Sat}$_{\mathcal X}(n)$ encoding that the formula in $\mathcal X$ that is represented by $\vec X^{n-1}$ holds.
-	\end{itemize}
+	\end{slimitemize}
 	Define$$\varphi^Q_{i} := \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
@@
 	Let $\varphi$ be a predicate formula. Define $\varphi^{b}$ inductively:
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $\bot^b = \bot$.
@@
 		\item $(\forall x\varphi)^b = \forall w(b\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
-	\end{itemize}
+	\end{slimitemize}
 	Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
@@
 	\]where $[\psi]$ denotes the corresponding element of $\mathcal X^{bS}$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $M_T$ is the union of $\{m\in M\:|\:\neg\text{World}^I(m)\}$ and sequences $(\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)$ such that $\vec E^I(\vec m_1, s^I)$ and $\vec E^I(\vec m_i, g_{\psi_{i-1}}(\vec m_{i-1}, g_{\psi_{i-2}}(\dots g_{\psi_1}(\vec m_1, s^I)\dots)))$ for $i > 0$ hold.
@@
 		\item For predicate symbols $R$, we set $${R}^{I_T}(\vec m, (\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)) = {R}^I(\vec m, g_{\psi_n}(\vec m_n, \dots(g_{\psi_1}(\vec m_1, s^I))\dots))$$ and arbitrarily elsewhere, for function symbols $f$ we set $f^{I_T}(\vec m) = f^I(\vec m)$ so long as $\vec m\in \{m\in M\:|\:\neg\text{World}^I(m)\}^n$ and arbitrarily elsewhere.
-	\end{itemize}
+	\end{slimitemize}
 \end{definition}
@@
 \begin{lemma}\label{proof:first-order-normal-form}
-	For every predicate formula $\varphi$ there exists a nullary relation symbol $P$ as well as sets of clauses $\mathcal R$ of the form
-	$$\forall \vec x\left(\bigwedge_i A_i\to \bigvee_j B_j\right)$$
-	called \emph{flat} clauses and $\mathcal X$ of the form
-	$$\forall \vec x\left((A\to B)\to C\right)$$
-	called \emph{implication} clauses as well as $\mathcal Q$ of the form
-	$$\forall\vec x\left((\forall y A)\to B\right)$$
-	called \emph{quantification} clauses such that $\varphi$ is intuitionistically equivalid to
+		For every predicate formula $\varphi$ there exists a nullary predicate symbol $P$ as well as sets of clauses $\mathcal R,\mathcal X, \mathcal Q$, with $\mathcal R$ containing \emph{flat clauses} of the form
+		$$\forall \vec x\left(\bigwedge_i A_i\to \bigvee_jB_j\right),$$
+		$\mathcal X$ containing \emph{implication clauses} of the form
+		$\forall \vec x\left((A\to B)\to C\right)$
+		and $\mathcal Q$ containing \emph{quantification clauses} of the form
+		$\forall\vec x\left((\forall y A)\to B\right),$
+		such that $\varphi$ is intuitionistically equivalid to
 		$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic.
 This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$.
-In general these formulae will contain new function symbols and predicate symbols.
+In general these formulae will contain new function and predicate symbols.
 \end{lemma}
@@
 	Now have
-	\begin{itemize}
+	\begin{slimitemize}
 		\item $\preceq^I = \leq\cup \{(m, m): m\in M\setminus W_0\}$, i.e. $\preceq$ corresponds to $\leq$ on $W_0$ and is trivial elsewhere.
@@
 		\end{cases}$
-	\end{itemize}
+	\end{slimitemize}
 	and $\mathcal M = (M, I)$. One easily verifies that these are well-defined. We now show by induction on the height of the formula the following:

### 2023-05-15 LPAR/LPAR-EIICL.tex
replace slimitemize

--- 2023-05-10 LPAR/LPAR-EIICL.tex
@@
 \newtheorem{remark}[theorem]{Remark}
-
-\let\OLDitemize\itemize
-\renewcommand\itemize{\OLDitemize\addtolength{\itemsep}{-5pt}}
 
@@
 To this end, we study embeddings of intuitionistic into classical logic and propose the following methodology:
-\begin{itemize}
+\begin{itemize}\addtolength{\itemsep}{-5pt}
 	\item Give for each $\varphi$ a formula $\varphi^\#$, which is classically valid iff $\varphi$ is intuitionistically valid.
@@
 The following main contributions will be discussed in more detail in Sections~\ref{sec:overview} and~\ref{sec:evaluation}:
-\begin{itemize}
+\begin{itemize}\addtolength{\itemsep}{-5pt}
 	\item We establish a small model property for counter-models to validity in intuitionistic propositional logic.
@@
 	A \emph{valuation} $v$ maps each propositional variable $A$ to a \emph{truth value} $v(A)\in\{0, 1\}$. We inductively define the model relation $\models$ between $v$ and formulae:
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $v\not\models \bot$
@@
 	The model relation between the $v_u$ and formulae $\varphi$ is defined as before, except in the case of implications, where we set
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $v_u\models\varphi\to \psi$ iff for all $w\geq u$ we have $w\not\models\varphi$ or $w\models\psi$.
@@
 	Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item to each $n$-ary function symbol $f\in\Sigma$ a $n$-ary function $f^I: M^n\to M$.
@@
 	Then terms are interpreted as follows:
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $a^{I, v} = v(a)$ for each free variable $a$.
@@
 	We define a model relation between pairs $\mathcal M, v$ and formulae $\varphi$ as follows
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $\mathcal M, v\not\models\bot$
@@
 	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulae $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $\mathcal M_w, v\not\models\varphi$ or $\mathcal M_w, v\models\psi$.
@@
 	For formulae $\varphi$ we define the Skolemization $\varphi^S_Z$ and Herbrandization $\varphi^H_Z$ with respect to a set of free variables $Z$ by simultaneous induction:
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $A^S_Z = A^H_Z = A$ for each atomic $A$.
@@
 	For every formula $\varphi$
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
@@
  Define $\varphi^{b}$ inductively:
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $\bot^b = \bot$.
@@
 assuming that $\varphi$ is of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ for some flat clauses $\mathcal R$ and implicational clauses $\mathcal X$, further simplification is possible:
-\begin{itemize}
+\begin{itemize}\addtolength{\itemsep}{-5pt}
 	\item For flat clauses $(\bigwedge A_i\to\bigvee B_j)^b$ is
@@
 	\end{cases}$$Define a model $\mathcal M_T = (M_T, I_T)$ as follows:
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $M_T$ is the set of sequences without repetition on $\mathcal X$.
@@
 Define
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item {\normalfont Valid}$(n)$ which encodes that $\vec X^n_\psi$ represents a valid next element of a sequence without repetition, i.e.
@@
 	Let $\varphi$ be a predicate formula. Define $\varphi^{b}$ inductively:
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $\bot^b = \bot$.
@@
 	\]where $[\psi]$ denotes the corresponding element of $\mathcal X^{bS}$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $M_T$ is the union of $\{m\in M\:|\:\neg\text{World}^I(m)\}$ and sequences $(\vec m_1, \psi_1)\dots (\vec m_n, \psi_n)$ such that $\vec E^I(\vec m_1, s^I)$ and $\vec E^I(\vec m_i, g_{\psi_{i-1}}(\vec m_{i-1}, g_{\psi_{i-2}}(\dots g_{\psi_1}(\vec m_1, s^I)\dots)))$ for $i > 0$ hold.
@@
 
-\begin{wrapfigure}[5]{r}{0.4\textwidth}
+\begin{wrapfigure}[5]{r}{0.41\textwidth}
 	\vspace*{-1.6cm}
@@
 
-\begin{wrapfigure}[15]{r}{0.4\textwidth}
+\begin{wrapfigure}[15]{r}{0.41\textwidth}
 	\vspace{6.6cm}
@@
 	Now have
-	\begin{itemize}
+	\begin{itemize}\addtolength{\itemsep}{-5pt}
 		\item $\preceq^I = \leq\cup \{(m, m): m\in M\setminus W_0\}$, i.e. $\preceq$ corresponds to $\leq$ on $W_0$ and is trivial elsewhere.

### 2023-06-13 Presentation LPAR/presentation.tex
LPAR Presentation

\documentclass{cubeamer}

\usepackage{trimclip}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{tikz}

\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{arrows,shapes,backgrounds}
\usepackage{amsmath,amsfonts}

\newcommand{\btVFill}{\vskip0pt plus 1filll}

\title{Embedding Intuitionistic into Classical Logic}
\subtitle{LPAR-24}
\author[A. Pluska, F. Zuleger]{Alexander Pluska, Florian Zuleger}
\date{\today} % or whatever the date you are presenting in is
\institute[TU Vienna]{Vienna University of Technology}
% \copyrightnotice{Published by the American Institute of Aeronautics and Astronautics, Inc., with permission}

\begin{document}

	\maketitle

	\cutoc

	\section{Background}

	\begin{frame}{Languages}
		Propositional Logic:
		\begin{itemize}
			\item Propositional variables $A, B, C, \dots$ and bottom $\bot$.
			\item Connectives $\wedge, \vee, \to$.
		\end{itemize}
		\[((A\to\bot)\vee B)\wedge C\]
		First-order Logic:
		\begin{itemize}
			\item Ambient Signature $\Sigma$.
			\item Inductively defined terms from variables and function symbols.
			\item Atoms from predicate symbols, equalities between terms and $\bot$.
			\item Connectives $\wedge, \vee, \to$ and quantifiers $\forall, \exists$.
		\end{itemize}
		\[\forall x (P(x)\to (f(x) = c\wedge \exists y Q(f(x), g(y))))\]
	\end{frame}

	\begin{frame}{Intuitionistic Logic}
		\begin{figure}[!tbp]
			\centering
			\begin{subfigure}[][][t]{0.6\textwidth}
				\begin{itemize}
					\item Motivation:
					\begin{itemize}
						\item Mathematical Constructivism.
						\item Curry-Howard-Correspondence.
					\end{itemize}
					\item Not intuitionistically valid:
					$$A\vee\neg A\hspace*{1cm} \forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$$
					\item Classical Logic $\longrightarrow$ Intuitionisic logic:\[ \text{Double Negation Translation}\]
					\item Intuitionistic Logic $\longrightarrow$ Classical Logic\[\text ?\]
				\end{itemize}
			\end{subfigure}
			\hfill
			\begin{subfigure}[]{0.3\textwidth}
				\centering
				\includegraphics[width=0.9\textwidth]{logicians.png}
				\caption{L. E. J. Brouwer, A. Heyting, H. B. Curry, W. A. Howard}
			\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}{Existing approaches}
		\begin{figure}
			\begin{subfigure}{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{friedman.png}
			\end{subfigure}
			\begin{subfigure}{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{schwichtenberg.png}
			\end{subfigure}
		\vspace{.5cm}
			\begin{subfigure}[b]{0.49\textwidth}
				\centering
				\includegraphics[width=\textwidth]{griffin.png}
			\end{subfigure}
			\begin{subfigure}[b]{0.49\textwidth}
				\centering
				\includegraphics[width=0.9\textwidth]{parigot.png}
			\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}{Kripke Semantics}
		\centering
		\vspace*{1cm}
		\begin{tabular}{c|c}
			Classical Logic&Intuitionistic Logic\\
			\hline
			Single valuation $v$.&Partially ordered set $W$ of worlds and family of\\&valuations $(v_u)_{u\in W}$ satisfying \emph{persistency}.\\
			\hline
			$v\models\varphi$&$u\models\varphi$ for each $u\in W$.	
		\end{tabular}
		\vspace*{.5cm}

		For atoms: $u\models A$ iff $v_u\models A$.\\
		Inductive extensions to formulas generally as in the first order case, e.g. \[u\models \varphi\wedge\psi \text{ iff } u\models\varphi \text{ and } u\models\psi.\]\vspace*{1.5cm}
		Difference: $u\models\varphi\to\psi$ iff \textbf{for all} $\mathbf{w}\geq \mathbf{u}$ we have $w\models\varphi$ or $w\models\psi$.
	\end{frame}

	\begin{frame}{Kripke Semantics - Example}
		\begin{figure}[!tbp]
		\centering
		\begin{subfigure}[][][t]{0.3\textwidth}
			\centering
				$A\vee\neg A\:\approx$ ``$A$ or never $A$''\vspace*{1cm}
				\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
					\node (center) {};
					\node[squarednode]      (left)   [left=of center] {};
					\node[squarednode]      (right)       	[right=of center] {$A$};
					\draw[-{Stealth[length=3mm, width=3mm]}] (left.east) -- (right.west);	
				\end{tikzpicture}
			\caption{Counter-model $A\vee\neg A$ $A\vee(A\to\bot)$}
		\end{subfigure}
			\hfill
		\begin{subfigure}[][][t]{0.6\textwidth}
			\centering
			$\forall x\neg\neg A(x)\to \neg\neg\forall x A(x)\:\approx$ ``for all $x$ eventually $A(x)$ implies eventually for all $x$ $A(x)$''\vspace*{1cm}
			\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
				\node[squarednode]  (0) {\color{red}$0$};
				\node[squarednode]      (1)   [right=of 0] {\color{green}$0$, \color{red}$1$};
				\node[squarednode]      (2)       	[right=of 1] {\color{green}$0, 1$, \color{red}$2$};
				\node   (3)       	[right=of 2] {\dots};
				\draw[-{Stealth[length=3mm, width=3mm]}] (0.east) -- (1.west);	
				\draw[-{Stealth[length=3mm, width=3mm]}] (1.east) -- (2.west);
				\draw[-{Stealth[length=3mm, width=3mm]}] (2.east) -- (3.west);	
			\end{tikzpicture}
			\caption{Counter-model $\forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$}
		\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}{A second motivation: Automated Deduction for Intuitionistic Logic}
		\begin{figure}
		\begin{subfigure}[b]{\textwidth}
			\centering
				\includegraphics[width=\textwidth]{iltp.png}
			\caption{ILTP benchmark}
		\end{subfigure}
		\begin{subfigure}[b]{\textwidth}
			\includegraphics[width=\textwidth]{CASC.png}
			\caption{CASC-28 Results}
		\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}{Not quite CNF}
		\begin{lemma}[Intuitionistic normal form, propositional case]
	For every propositional formula $\varphi$ there exists an atom $P$ as well as sets of clauses $\mathcal R, \mathcal X$, with $\mathcal R$ containing  \emph{flat clauses} of the form
	$$\bigwedge_iA_i\to\bigvee_jB_j$$
	and $\mathcal X$ containing \emph{implication clauses} of the form
	$(A\to B)\to C,$
	such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$where $A_i, B_i, A, B, C$ are atomic.
			\end{lemma}	
	\end{frame}

	\begin{frame}{Not quite CNF - Example}
		The formulas \[A\vee\neg A\] and \[((A\to P)\wedge(\neg A\to P))\to P\] are intuitionistically equivalid.
	\end{frame}

	\begin{frame}{Not quite CNF - First-order case}
		In addition to sets of flat clauses $\mathcal R$ of the form
	$$\forall\vec x\left(\bigwedge A_i\to\bigvee B_j\right)$$ and implication clauses $\mathcal X$ of the form
	$$\forall\vec x((A\to B)\to C)$$ there is a set of \emph{quantification clauses} $\mathcal Q$ of the form $$\forall \vec x(\forall \vec yQ\to P).$$
	\end{frame}

	\section{The Embedding}

	\begin{frame}{Overview}
		\begin{enumerate}
			\item Encoding of Kripke Semantics
			\item Reduction of the Encoding
		\end{enumerate}
	\end{frame}

	\begin{frame}{Classical Encoding of Kripke Semantics - Propositional Case}
		\begin{lemma}
			Let $\varphi$ be a propositional formula.
		 	We can inductively define $\varphi^{b}$ 
			and have $K(\varphi)$ encode the theory of Kripke structures, i.e.
			\[K(\varphi) := \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w)),\]
			such that
			\[\varphi^{C} := K(\varphi)\to \varphi^{b}\]
		    is classically valid if and only if $\varphi$ is intuitionistically valid.
		\end{lemma}
		In this process propositional variables are replaced by unary predicates, where $M, [u/x]\models A(x)$ can be interpreted as $u\models A$.
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		\vspace*{.5cm}
		\begin{lemma}
			Let $\Lambda$ be the set of sequences without repetition over $\mathcal X$. A countermodel to $\varphi^C$ exists if and only if a countermodel with domain $\Lambda$ exists in which $\preceq$ is interpreted as the prefix ordering.
		\end{lemma}
		\btVFill
		Recall: Intuitionistic normal from $\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$ with $\mathcal R$ containing  \emph{flat clauses} $\bigwedge A_i\to \bigvee B_j$ and $\mathcal X$ containing \emph{implication clauses} $(A\to B)\to C$.
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		\vspace*{.5cm}
		\begin{lemma}
			Let $\Lambda$ be the set of sequences without repetition over $\mathcal X$. A countermodel to $\varphi^C$ exists if and only if a countermodel exists with domain $\Lambda$ exists and in which $\preceq$ is interpreted as the prefix ordering.
		\end{lemma}
		Example, $((A\to P)\wedge(\neg A\to P))\to P$:
		\begin{center}
		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
			\node (center) {};
			\node[squarednode]      (left)   [left=of center, label=$\epsilon$] {};
			\node[squarednode]      (right)       	[right=of center, label=$\neg A\to P$] {};
			\draw[-{Stealth[length=3mm, width=3mm]}] (left.east) -- (right.west);	
		\end{tikzpicture}
		\end{center}
		\btVFill
		Recall: Intuitionistic normal from $\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$ with $\mathcal R$ containing  \emph{flat clauses} $\bigwedge A_i\to \bigvee B_j$ and $\mathcal X$ containing \emph{implication clauses} $(A\to B)\to C$.
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		\vspace*{.5cm}
		\begin{lemma}
			Let $\Lambda$ be the set of sequences without repetition over $\mathcal X$. A countermodel to $\varphi^C$ exists if and only if a countermodel exists with domain $\Lambda$ exists and in which $\preceq$ is interpreted as the prefix ordering.
		\end{lemma}
		Example, $((A\to P)\wedge(\neg A\to P))\to P$:
		\begin{center}
		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
			\node (center) {};
			\node[squarednode]      (left)   [left=of center, label=$\epsilon$] {};
			\node[squarednode]      (right)       	[right=of center, label=$\neg A\to P$] {$A$};
			\draw[-{Stealth[length=3mm, width=3mm]}] (left.east) -- (right.west);	
		\end{tikzpicture}
		\end{center}
		\btVFill
		Recall: Intuitionistic normal from $\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$ with $\mathcal R$ containing  \emph{flat clauses} $\bigwedge A_i\to \bigvee B_j$ and $\mathcal X$ containing \emph{implication clauses} $(A\to B)\to C$.
	\end{frame}

	\begin{frame}{Reducing the Encoding - Propositional Case}
		\vspace*{.5cm}
		\begin{lemma}
			Let $\Lambda$ be the set of sequences without repetition over $\mathcal X$. A countermodel to $\varphi^C$ exists if and only if a countermodel exists with domain $\Lambda$ exists and in which $\preceq$ is interpreted as the prefix ordering.
		\end{lemma}
		\begin{corollary}[Small Model Property]
			If $\varphi$ is intuitionistically invalid then there exists a Kripke counter-model which is a rooted tree with height and degree bounded by $|\mathcal X|$.
		\end{corollary}
		\btVFill
		Recall: Intuitionistic normal from $\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$ with $\mathcal R$ containing  \emph{flat clauses} $\bigwedge A_i\to \bigvee B_j$ and $\mathcal X$ containing \emph{implication clauses} $(A\to B)\to C$.
	\end{frame}

	\begin{frame}{Propositional Encoding}
		Knowing the shape of potential countermodels we can essentially replace universal quantification by enumeration. This gives us an embedding into propositional logic.
		\begin{itemize}
			\item For each atom $A$ and $\lambda\in\Lambda$ consider a new atom $A^\lambda$.
			\item $\bigwedge A_i\to \bigvee B_j$ is encoded as $\bigwedge \{\bigwedge A_i^\lambda\to \bigvee B_j^\lambda\mid \lambda\in\Lambda\}$.
			\item $\psi = (A\to B)\to C$ is encoded as $\bigwedge\{(A^{\lambda\psi}\to B^{\lambda\psi})\to C^\lambda\mid \lambda\in\Lambda, \psi\notin\lambda\}$.
			\item $P$ is encoded as $P^\epsilon$.
		\end{itemize}
		Size of the new formula is in $\mathcal O(|\varphi|\cdot 2^{|\mathcal X|})$.
		\btVFill
		Recall: Intuitionistic normal from $\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$ with $\mathcal R$ containing  \emph{flat clauses} $\bigwedge A_i\to \bigvee B_j$ and $\mathcal X$ containing \emph{implication clauses} $(A\to B)\to C$.
	\end{frame}

	\begin{frame}{Aside - Encoding to QBF}
		\vspace*{2cm}

		\begin{lemma}
			For each propositional formula $\varphi$ there is effectively a QBF $\varphi^Q$ with $\mathcal |\varphi^Q|\in\mathcal O(|\mathcal R|\cdot|\mathcal X| + |\mathcal X|^3)$ such that $\varphi$ is intuitionistically invalid if and only if $\varphi^Q$ is a satisfiable QBF.
		\end{lemma}
		\btVFill
		Recall: Intuitionistic normal from $\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$ with $\mathcal R$ containing  \emph{flat clauses} $\bigwedge A_i\to \bigvee B_j$ and $\mathcal X$ containing \emph{implication clauses} $(A\to B)\to C$.
	\end{frame}

	\begin{frame}{Classical Encoding of Kripke Semantics - First-order Case}
		\begin{itemize}
			\item Classical models must contain elements both representing worlds in the Kripke frame and elements in the domain of the models at each world.
			\item 
			We use a special predicate $E$ to encode that an element exists at a world.
		\end{itemize}
		As in the propositional case we can define $\varphi^b$ and encode the Kripke Semantics
		\begin{align*}
			K(\varphi)= &\:\text{PartialOrder}(\preceq) \to \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w))\\
			&\to\forall u(\text{DomainClosed}(u))\to \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
		\end{align*}
		such that
		\[\varphi^C:= K(\varphi)\to E(sk_0, b)\to \vec E(\vec a, b)\to \varphi^b\]
		is classically valid if and only if $\varphi$ is intuitionistically valid.
	\end{frame}

	\begin{frame}{Reducing the Encoding - First-order Case}
		\begin{itemize}
			\item We show that most model reduction techniques from the propositional case carry over to the first-order case.
			\item We are able to completely eliminate $\preceq$.
			\item It is \textbf{not} possible to completely eliminate quantification over worlds as some sentences only have counter-models with an infinite number of worlds.
			\item Redundant clauses can encode further model reductions.
			\item No exponential blowup \textbf{but} new symbols.
		\end{itemize}
	\end{frame}

	\section{Benchmark}

	\begin{frame}{Banchmark}
		\begin{itemize}
			\item Translation written in Rust.
			\item Classical backend: Vampire.
			\item ILTP library, 2670 problems.
			\item Problem \textsc{SYN007+1} removed.
			\item Intuitionistic SOTA provers for comparison:
			\begin{itemize}
				\item ileanCoP 1.2
				\item nanoCoP-i 2.0
			\end{itemize}
		\end{itemize}
	\end{frame}

	\begin{frame}{Benchmark}
		\begin{figure}
			\begin{subfigure}{0.39\textwidth}			
				\centering
					\begin{tabular}{l|c|c}
						Embedding&571&21.4\%\\
						Emb. (idem.)&557&20.9\%\\
						ileanCoP 1.2&875&32.8\%\\
						nanoCoP-i 2.0&858&32.1\%\\\hline
						Total&2669&100\%
					\end{tabular}
					\vspace{1.2cm}
					\caption{Solved problems by each prover.}
			\end{subfigure}	
			\begin{subfigure}{0.6\textwidth}
				\centering
				\includegraphics[width=0.66\textwidth]{circleplot.png}
				\caption{Instances proven by combinations of Embedding (red), Embedding (idem.) (green), ileanCoP or nanoCoP (blue)}
			\end{subfigure}
		\end{figure}
	\end{frame}

	\begin{frame}{Benchmark}
		\begin{figure}
			\begin{subfigure}{\textwidth}			
				\centering
					\begin{tabular}{c|c|lr|lr|lr|lr|c}
		$|\mathcal Q|$&$|\mathcal X|$&\multicolumn{2}{c}{Embedding}&\multicolumn{2}{c}{Emb. (idem.)}&\multicolumn{2}{c}{ileanCoP}&\multicolumn{2}{c}{nanoCoP-i 2.0}&Total\\\hline
		0&0		&157&88.2\%	&157&88.2\%		&128&71.9\%		&61&38.1\%	&178\\
		0&1-2	&119&29.8\%&130&32.5\%		&108&27.0\%		&96&24.0\%	&400\\
		0&3+	&33&18.9\%	&38&21.7\%		&70&40.0\%		&51&29.1\%	&175\\
		1-2&0-2	&84&29.3\%	&81&28.2\%		&227&79.1\%		&221&77.0\%	&287\\
		1-2&3+	&102&24.7\%	&98&27.3\%		&180&43.6\%		&157&39.1\%	&413\\
		3+&0+	&66&5.4\%	&63&5.2\%		&262&20.7\%		&270&22.2\%	&1216\\
					\end{tabular}
					\caption{Solved problems grouped by $|\mathcal Q|$ and $|\mathcal X|$.}
				\end{subfigure}
		\end{figure}
	\end{frame}

	\section{Summary and Future Work}

	\begin{frame}{Summary}
		Our main contributions:
		\begin{itemize}
			\item Use model reductions to establish a small model property for intuitionistic propositional logic.
			\item Based on this, give an embedding of intuitionistic propositional logic into classical propositional logic and an embedding into QBF.
			\item Utilize analogous model reductions to give a reduced embedding of intuitionistic first-order logic into classical first-order logic.
			\item Evaluate the practical utility of this embedding for automated theorem proving.
		\end{itemize}
	\end{frame}

	\begin{frame}{Future Work}
		\begin{itemize}
			\item Implication/quantification clauses explode search space.
			\item Usually we don't need (all of) them.
			\item Potential Solution: Overapproximation, e.g. replace \[\forall\vec x\forall u((A(\vec a, f(u))\to B(\vec b, f(u)))\to C(\vec c, u))\] with \[\forall\vec x\forall u(B(\vec b,u)\to  C(\vec c, u)).\]
			\item Add them back incrementally as required.
		\end{itemize}
	\end{frame}

\begin{frame}[standout]
	\Huge\textsc{Thanks for your attention!}
\end{frame}

\end{document}
